### 2018-07-17 EMNLP/divergence.tex
add EMNLP paper

%
% File emnlp2018.tex
%
%% Based on the style files for EMNLP 2018, which were
%% Based on the style files for ACL 2018, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{emnlp2018}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
%%% YOUR PACKAGES BELOW THIS LINE %%%
\usepackage[small,bf]{caption} % added MLF 20171211
\usepackage{multicol}
\usepackage{adjustbox}
\usepackage{graphicx}
\usepackage{array,multirow}

%\aclfinalcopy % Uncomment this line for the final submission

\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}
\newcommand\confname{EMNLP 2018}
\newcommand\conforg{SIGDAT}

\title{Fixing Translation Divergences in Parallel Corpora for Neural MT}
%\title{A Recurrent Network for Parallel Corpora Divergence Analysis}

\author{%Minh Pham, \ Fran\c cois Yvon\\
  First author\\
  Affiliation / Address line 1\\
  Affiliation / Address line 2\\
  Affiliation / Address line 3\\
  {\tt email@domain}  
  \And
  %Josep Crego, \ Jean Senellart\\
  Second author\\
  Affiliation / Address line 1\\
  Affiliation / Address line 2\\
  Affiliation / Address line 3\\
  {\tt email@domain}}

\date{}

\begin{document}
\maketitle
\begin{abstract}

%http://www.airccj.org/CSCP/vol4/csit42503.pdf

Corpus-based approaches to machine translation rely on the availability of parallel corpora.
%and their performance is highly linked to the quality of the parallel data. 
According to the process followed to compile a parallel corpus, it may contain multiple parallel sentences that are often not as parallel as one might assume. 
%Examples are easily found when corpora is created from non-parallel texts or included by noisy document/sentence alignment tools.
This paper describes an unsupervised method for detecting translation divergences in parallel sentences. 
%We use a deep neural network to predict word alignment scores in parallel sentences. 
We present a neural network to predict sentence similarity that minimises a loss function based on word alignments.
We show that accurate predictions are obtained allowing divergent sentences to be filtered out. 
Furthermore, word similarity scores predicted by the network are used to identify and fix some divergences guiding to more parallel segments.
%Misaligned words are then identified allowing divergent parallel sentences to be filtered out and in some cases making sentences parallel by removing such words. 
We evaluate the presented method on an English-French and an English-German machine translation tasks. 
Results show that neural MT systems trained on the filtered/corrected corpus outperform the MT systems trained on the original data.
%This method can be used on any parallel corpus without any manual annotation.

\end{abstract}

\section{Introduction}

Parallel sentence pairs are the only necessary resource to build Machine Translation (MT) systems. 
In the case of Neural MT, a large neural network is trained through maximising translation performance on a given parallel corpus. 
Therefore, the quality of an MT engine is heavily dependent upon the amount and quality of parallel sentences. 
Unfortunately, parallel texts are scarce resources. 
There are relatively few language pairs for which parallel corpora of reasonable sizes are available, and even for those pairs, the corpora come mostly from few domains. 
To alleviate the lack of parallel data several approaches have been developed in the last years. 
They range from methods using non-parallel, or comparable data ~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,W17-2509,P17-3003} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714} using in all cases automated alignment/translation engines that are prone to the introduction of noise in the resulting parallel sentences. 
Mismatches on parallel sentences extracted from translated texts are also reported ~\cite{tiedemann2011bitext,XU16.310}. 
This problem is mostly ignored in machine translation, where parallel sentences are considered to convey the exact same meaning, and is particularly important for neural MT engines as suggested by~\cite{chen2016adaptation}.

Table~\ref{tab:examples} illustrates some examples of English-French parallel sentences which are not completely semantically equivalent. Examples are extracted from the OpenSubtitles corpus~\cite{LisonTiedemann2016}. 
Divergences are outlined using bold letters. 
Additional segments are included on either side of the parallel sentences (first and second rows). 
Some translations may be completely uncorrelated (third row) or inaccurate (fourth row).
Note that divergent translations are due to many different reasons ~\cite{C14-1055}, the study of which is beyond the scope of this paper.

\begin{table}[ht]
\small
\center
\begin{tabular}{ c|l }
  \hline  
  \texttt{en} & \it{What do you feel}\bf{, Spock}\it{?} \\
  \texttt{fr} & \it{Que ressentez-vous?} \\
  \texttt{gl} & {\small \it{What do you feel?}} \\
  \hline
  \texttt{en} & \it{How much do you get paid?} \\
  \texttt{fr} & \it{T'es pay\'e combien} \bf{de l'heure}\it{?} \\
  \texttt{gl} & {\small \it{How much do you get paid per hour?}} \\
  \hline  
  \texttt{en} &  \bf{That seems a lot.} \\
  \texttt{fr} & \bf{40 livres?} \\
  \texttt{gl} & {\small \it{40 pounds?}} \\
  \hline  
  \texttt{en} & \it{I brought you} \bf{french fries}\it{!} \\
  \texttt{fr} & \it{Je t'ai rapport\'e des} \bf{saucisses}\it{!} \\
  \texttt{gl} & {\small \it{I brought you sausage!}} \\
  \hline
\end{tabular}
\caption[Table caption text]{Examples of semantically divergent parallel sentences. English (\texttt{en}), French (\texttt{fr}) and gloss of French (\texttt{gl}). }
\label{tab:examples}
\end{table}

In this work, we present a neural divergence classifier aimed at detecting words on either side of parallel sentences for which the corresponding meaning is not present on the translated counterpart. 
We evaluate the classifier on an English-French and an English-German translation tasks showing that translation accuracy can be improved by filtering out divergent sentence pairs.
In addition, we  show that some divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.

The remainder of this paper is structured as follows. Section~\ref{related} overviews related work. 
We describe in detail the core of the neural divergence classifier and the correction algorithm in Section~\ref{sec:divergence}. 
Section~\ref{experiments} details the experiments conducted to measure the ability of the presented classifier to identify and fix divergent sentences.
Section~\ref{sec:results} evaluates results. 
Finally, conclusions are drawn in Section~\ref{conclusions} while further work is outlined in Section~\ref{further}.
All the code used in this paper as well as a human annotated test set are freely available\footnote{\texttt{https://github.com/anonymised}}.

\section{Related Work}
\label{related}

Attempts to measure the impact of translation divergences in MT systems have focused on the introduction of noise in sentence alignments ~\cite{goute2012}, showing that statistical MT systems are highly tolerant to noise, and that performance only degrades seriously at very high noise levels. 
In constrast, neural MT engines seem to be more sensitive~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.

Efforts have been devoted to characterise the degree of semantic equivalence between two snippets of text in the same or different languages~\cite{conf/semeval/AgirreBCDGMRW16}, a workshop devoted to an objective similar to our work. 
In~\cite{Mueller:2016:SRA:3016100.3016291}, a monolingual sentence similarity network is proposed, making use of a simple LSTM layer to compute sentence representations. 
The authors show that a simple SVM classifier can be built on top of the sentence representations to achieve state-of-the-art results in a semantic entailment classification task. 
With the same objective, the system presented in~\cite{N16-1108} uses multiple convolutional layers and models pairwise word interactions. %Our work differs from the previous as we build a different network and use it to measure similarity of sentences in different languages.

Our work is inspired by~\cite{W17-3209} where the authors train a cross-lingual divergence detector using word alignments and sentence length features to train a linear SVM classifier. 
Their work shows that an NMT system trained only on non-divergent sentences yields slightly higher translation quality scores and requires clearly less training time. 
The same authors have recently updated their work in~\cite{DBLP:journals/corr/abs-1803-11112}. 
The objective is the same as the neural network presented in~\cite{N16-1108} and their network further outperforms their previous work. 
Our work differs from the previous as we make use of a network with different topology. 
We model sentence similarity by means of optimising a loss function based on word alignments. 
Furthermore, the network predicts word similarity scores that we further use to correct divergent sentences.

\section{Neural Divergence Classifier}
\label{sec:divergence}

The architecture of our model is inspired by the work on Word Alignment in~\cite{W16-2207}. Figure~\ref{network} illustrates the model.

\begin{figure}[h]
\center
    \includegraphics[width=0.8\linewidth]{network}
    \caption{Illustration of the model.} 
    \label{network}
\end{figure}

%\subsection{Divergence Network}

In the following, we consider a source-target sentence pair $(s,t)$ with $s=(s_1,...,s_I)$ and $t=(t_1,...,t_J)$. 
The model is composed of 2 Bi-directional LSTM subnetworks, $net_s$ and $net_t$, which respectively encode source and target sentences. 
Since both $net_s$ and $net_t$ take the same form we describe only the source architecture.

The source-sentence Bi-LSTM network outputs forward and backward hidden states, $\overrightarrow{h}^{src}_i$ and $\overleftarrow{h}^{src}_i$, which are then concatenated into a single vector encoding the $i^{th}$ word of the source sentence, 
$h^{src}_i = [ \overrightarrow{h}^{src}_i ; \overleftarrow{h}^{src}_i ]$.
In addition, the last forward/backward hidden states (outlined using dark grey in Figure~\ref{network}) are also concatenated into a single vector  to represent whole sentences 
$h_{src} = [ \overrightarrow{h}^{src}_I ; \overleftarrow{h}^{src}_1 ]$.
At this point, a measure of similarity between sentence pairs can be obtained by cosine similarity:
\begin{equation}
    sim(h_{src}, h_{tgt}) = \frac{h_{src} \cdotp h_{tgt}}{||h_{src}|| * ||h_{tgt}||}
    \label{cosine}
\end{equation}
%\noindent where two vectors (embeddings) with the same orientation have a cosine similarity of $1$, while two vectors with opposed orientation have a similarity of $-1$, independent of their magnitude.

Our model can be optimised following two different objective functions:

\begin{description}
    \item[wemb]

Maximise the word alignment scores between words of both sentences using aggregation functions that summarise the alignment scores for each source/target word. 
Similar to~\cite{W16-2207} alignment scores $s(i,j)$ are given by the dot-product $S(i,j) = h_i^{src} \cdotp h_j^{tgt}$, while aggregation functions are defined as:   
\begin{equation}
\begin{split}
    aggr_s(i,S) = \frac{1}{r} \ log \left( \displaystyle \sum_{j=1}^{J} e^{r * S(i,j)}\right) \\
    aggr_t(j,S) = \frac{1}{r} \ log \left( \displaystyle \sum_{i=1}^{I} e^{r * S(i,j)}\right)
\end{split}
\label{aggregation}
\end{equation}

The loss function is defined as:
%\vspace{-8mm}
\begin{equation}
\begin{split}
\mathcal{L}(src,tgt) = \ \ \ \ & \\
    \sum_{i=1}^I log(1+e&^{s_{aggr}(i,tgt) * sign_i}) \ \ +\\
 + \sum_{j=1}^J log(1+e&^{s_{aggr}(src,j) * sign_j})
\end{split}
\label{loss_wemb}
\end{equation}
%\vspace{-8mm}

    \item[semb] Maximise the similarity score of  both sentence embeddings as computed by Equation~\ref{cosine}. With loss function defined as: 
%\vspace{-5mm}
\begin{equation}
\mathcal{L}(src,tgt) = log(1+e^{sim(src,tgt) * sign}) % where $sign(src,tgt) =1
\label{loss_semb}
\end{equation}

\end{description}

\subsection{Training with Negative Examples}
\label{negative}

Training is performed by minimising Equation~\ref{loss_wemb}, for which examples with annotations for source $sign_i$ and target $sign_j$ words are needed.
As positive examples we use {\bf (P)}aired sentences of a parallel corpus. All words in paired sentences are labelled as parallel ($-1$). %, $sign_i=-1$ and $sign_j=-1$. 
As negative examples we use random {\bf (U)}npaired sentences. 
In this case, all words are labelled as divergent ($+1$). %, $sign_i=+1$ and $sign_j=+1$. 
Since negative pairs may be very easy to classify and we want our network to detect less obvious divergences, we further create negative examples following: % the next two methods:

We {\bf (R)}eplace random sequences of words by a sequence of words with same part-of-speeches. 
The rationale behind this method is to keep the new sentences as grammatical as possible. 
%Otherwise the network can learn to detect non-grammatical sentences to predict divergence.
Words that are not replaced are considered parallel ($-1$) while those replaced are assigned the divergent label ($+1$). 
Words aligned to any replaced word are also assigned the divergent label ($+1$). For instance, given the original sentence pair:

\begin{table}[h]
\small
\center
\begin{tabular}{ll}
$src$: & { \small \texttt{What do you feel ?}} \\
$tgt$: & { \small \texttt{Que ressentez-vous ?}} \\
\end{tabular}
\end{table}

We may replace '\texttt{you feel}', with tags '\texttt{PRP VB}', by another with same tags (i.e. '\texttt{we want}'):

\begin{table}[h]
\small
\center
\begin{tabular}{ll}
$src$: & { \small \texttt{What do {\bf we \ want} ?}} \\
$sign_i$: & { \small \texttt{-1 \ \  -1 {\bf +1\ \ +1} \ \  -1}} \\
$tgt$: & { \small \texttt{Que {\bf ressentez-vous} ?}} \\
$sign_j$: & { \small \texttt{-1\ \ {\bf +1}\ \ \ \ \ \ \ \ \ -1}} \\
\end{tabular}
\end{table}

%Divergent words are shown in bold. 
%Note that new sentences are not assured to be grammatical after replacing sequences with the same part-of-speech.
Word alignments are used to identify '\texttt{ressentez-vous}' as divergent, since it is aligned to '\texttt{you feel}'.

Motivated by sentence segmentation errors observed in many corpora, we also build negative examples by {\bf (I)}nserting a second sentence following/preceding an %at the beginning (or end) of the source (or target) pair. 
original sentence. Only inserted words are considered divergent ($+1$).
Following with our example, we may add the sentence '\texttt{Not .}' at the end of the original source sentence:

\begin{table}[h]
\small
\center
\begin{tabular}{ll}
$src$: & { \small \texttt{What do you want ? {\bf Not \ .}}} \\
$sign_i$: & { \small \texttt{-1 \ \  -1 -1 \ -1  \ \ -1 {\bf +1\ \ \  +1}}} \\
$tgt$: & { \small \texttt{Que ressentez-vous ?}} \\
$sign_j$: & { \small \texttt{-1\ \ -1\ \ \ \ \ \ \ \ \ \ \ \ \ -1}} \\
\end{tabular}
\end{table}

Replace and insert methods are applied on either side of the sentence pairs.
To avoid that negative examples are easily predicted by using the difference in length of sentences, we restrict negative examples to have a difference in length not exceeding a ratio of $2.0$ ($3.0$ for shortest sentences).

The loss function of Equation~\ref{loss_semb} uses a single label for each sentence pair. 
Positive examples are labelled $sign=-1$ while any of the previous negative examples are labelled $sign=+1$.

\subsection{Divergence Correction}
\label{correction}

We observed in our training corpora that many divergent sentences follow a common pattern, consisting of adding some extra leading/trailing words. % (as in the first and second examples of table~\ref{tab:examples}). 
Accordingly, we implemented a simple algorithm that discards sequences of leading/trailing words on both sides. 
Hence considering as parallel $s_u^v$ and $t_x^y$. 
To find the optimal source ($u, v$) and target ($x, y$) indexes that enclose the parallel segments within the original sentence, we implement:
\begin{equation*}
\underset{u, v, x, y}{\arg\max} \Big \{      \underset{u \le I \le v}{\sum} \underset{x \le j \le y}{\max} \{ S(i,j) \}    \Big \}
\end{equation*}

The $\mathcal{N}$-best sequences following the previous function ($s_u^v$, $t_x^y$) are considered as valid corrections, but only the highest ranked according to their similarity score is used as replacement for the original $(s_1^I, t_1^J)$.
Short sentences are not considered. This is, $v - u > \tau$ and $y - x > \tau $. 
Figure~\ref{matrix} (left) shows an example of an alignment matrix $S(i,j)$ for a given sentence pair. 
An acceptable correction is: \textit{Que ressentez-vous ? $\Leftrightarrow$ What do you feel ?}. 
Hence, with indexes $u=1$, $v=5$, $x=1$ and $y=3$.

\section{Experiments}
\label{experiments}

%In this section we report on the experiments conducted to evaluate the proposed methods. %We begin with details of the corpora employed.

\subsection{Corpora}
\label{corpora}

We filter out divergences found in the English-French OpenSubtitles corpus~\cite{LisonTiedemann2016}, which consists of a collection of movie and TV subtitles. We also use the English-German Paracrawl\footnote{http://paracrawl.eu/} corpus. 
Both corpora present many potential divergences.
To evaluate English-French performance we use the En-Fr Microsoft Spoken Language Translation corpus, created from actual conversations over Skype~\cite{mslt-corpus-iwslt-2016-release}. 
English-German performance is evaluated on the publicly available newstest-2017 En-De test set, corresponding to news stories selected from online sources~\cite{W17-4717}.
In order to better assess the quality of our classifier when facing different types of word divergences we collected and annotated at the word level $500$ sentences from the original OpenSubtitles corpus, containing:
$200$ paired sentences;
$100$ unpaired sentences;
$100$ sentences with replace examples; and
$100$ sentences with insert examples as detailed in Section~\ref{negative}.

%For training, our systems use a random selection ($3$ million sentence pairs) of the OpenSubtitles corpus~\cite{LisonTiedemann2016} which consists of a collection of movie and TV subtitles. %, aligned at the sentence level after a number of automatic preprocessing steps. 
%The corpus presents many potential divergences as outlined in Table~\ref{tab:examples}.

%We evaluated the suitability of the filtering systems on the En-Fr Microsoft Spoken Language Translation task ({\bf MSLT}), created from actual conversations~\cite{mslt-corpus-iwslt-2016-release}. We also tested on a second domain, using the publicly available newstest-2013 En-Fr test set ({\bf NEWS}), corresponding to news stories selected from online sources~\cite{WMT:2013}.

\subsection{Neural Divergence}
\label{divergence}

All data is preprocessed with an in-house toolkit that performs minimal tokenisation. %, basically splitting-off punctuation.
After tokenisation, each out-of-vocabulary word is mapped to a special UNK token.
Vocabularies consist of the $50,000$ more frequent words.
%Our divergence classifier is described in Section~\ref{divergence}. 
Word embeddings are initialised using \texttt{fastText}\footnote{https://github.com/facebookresearch/fastText}. %, further aligned by means of \texttt{MUSE}\footnote{https://github.com/facebookresearch/MUSE} following the unsupervised method detailed in~\cite{lample2018word}. 
Size of embeddings is $E_s=E_t=256$ cells. 
Both Bi-LSTM use $256$-dimensional hidden representations ($E=512$).
We use $r=1.0$. 
Network optimization is done using the SGD method along with gradient clipping~\cite{Pascanu:2013:DTR:3042817.3043083}. 
For each epoch we randomly select $1$ million sentence pairs that we place in batches of $32$ examples.  
We run $10$ epochs and start decaying at each epoch by $0.8$ when the loss on validation set increases. 
Two versions of the network are evaluated. 
The first optimises the loss shown by equation~\ref{loss_wemb}, the second optimises the loss shown by equation~\ref{loss_semb}. 
Divergence is always computed following equation~\ref{cosine}. 
For divergence correction, we used $\mathcal{N}=20$ and $\tau=3$.

\subsection{Neural Translation}
\label{translation}

In addition to the basic tokenisation detailed in Section~\ref{divergence} we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30,000$ merge operations learned from both English and French data.
Neural MT systems are based on the open-source project \texttt{OpenNMT}\footnote{http://opennmt.net}. We use a bidirectional RNN encoder with $4$ LSTM layers with each containing $1,000$ cells. 
Word embeddings have a size of $300$ cells. We set the dropout probability to $0.3$. Batch size is set to $64$. The maximum length of both source and target sentences is set to $80$. % and we limit the vocabulary size to $50K$ words for both source and target languages.
The default optimiser is SGD with the starting learning rate of $1.0$. We start to decay the learning rate from epoch $10$ or when we detect increasing perplexity as compared to the previous epoch on a validation set. We stop training after $20$ epochs.

\section{Results}
\label{sec:results}

%We evaluate first the ability of our similarity classifier to predict different types of divergencies. 
%We use the test set manually annotated for that purpose and train our model on the OpenSubtitles corpus.

Table~\ref{results_puri} shows accuracies obtained by our model when trained over different combinations of negative examples.
We use the test set with manual annotations of word divergences.
In training, the same number of examples are always generated for each type of example (P, U, R and I). 
A word is considered divergent if it has a negative aggregation score, Equation~\ref{aggregation}.

\begin{table}[h]
\small
\center
\begin{tabular}{crccccc}
\hline
\multicolumn{2}{l}{\bf Accuracy} & \multicolumn{5}{c}{Test examples} \\
 &  & \texttt{P} & \texttt{U} & \texttt{R} & \texttt{I} & \texttt{PURI} \\
 \hline
\parbox[t]{0mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Train examples}}} &  \texttt{PU}     & \bf 0.996 & \bf 0.994 & 0.671 & 0.673 & 0.874 \\
 &  \texttt{PR}     & \bf 0.995 &      0.033 & \bf 0.951 &      0.689 & 0.746 \\
 &  \texttt{PI}       & \bf 0.998 &      0.071 &      0.697 & \bf 0.725 & 0.705 \\
 &  \texttt{PUR}  & \bf 0.994 & \bf 0.989 & \bf 0.919 &      0.710 & 0.932 \\
 &  \texttt{PUI}    & \bf 0.995 & \bf 0.996 &      0.662 & \bf 0.769 & 0.887 \\
 &  \texttt{PRI}    & \bf 0.991 &      0.161 & \bf 0.924 & \bf 0.719 & 0.768 \\
 &  \texttt{PURI} & \bf 0.995 & \bf 0.980 & \bf 0.916 & \bf 0.788 & \bf 0.942 \\
\hline
\end{tabular}
\caption{Word divergence accuracies according to different type of examples used in train/test.}% for networks trained with different combinations of negative examples. P, U, R and I stand respectively for pair, unpair, replace and insert.}
\label{results_puri}
\end{table}

As it can be seen, non-divergent words in parallel and unpaired sentences (columns \texttt{P} and \texttt{U}) are easy to identify, as far as the model has seen these types of examples in training. 
Regarding columns \texttt{R} and \texttt{I}, accuracies are lower since these sentences contain a mix of divergent and non-divergent words. 
%Hence, word divergence is more difficult to predict. 
Again, models that were trained with the corresponding examples obtain the highest accuracies (outlined in bold letters).
Column \texttt{PURI} shows results over the entire test set, mixing all type of examples. 
The best accuracy is obtained by the system trained on all type of examples.

\begin{figure}[h]
\centering
    \includegraphics[width=0.35\linewidth]{feel}    
    \includegraphics[width=0.35\linewidth]{feel_pu}    
\caption{Sentence pair with similarity scores produced by our model when trained over PU examples (right) and over PURI examples (left). Aggregation scores (Equation~\ref{aggregation}) are shown next to words. Matrices contain alignment scores. Sentence similarities (Equation~\ref{cosine})  shown below matrices.}
\label{matrix}
\end{figure}

Figure~\ref{matrix} illustrates the output of our network when trained using PU examples (right) and PURI examples (left).  
The former (right) fails to predict some word divergences, most likely because in training it never saw sentences mixing divergent and non-divergent words in the same example. 
%Furthermore, the latter (right) correctly assigns a lower similarity score to the sentence pair as both sentences do not convey the exact same meaning.

Table~\ref{results} shows BLEU results for our neural MT engine trained over different data configurations:
The entire\footnote{The original Paracrawl corpus contains more than $100$ million sentences. We reduced its size to $22.2$ millions using standard filtering techniques.} data (\texttt{all}); 
Most similar pairs after optimising equation~\ref{loss_semb} (\texttt{semb}); 
Most similar pairs after optimising equation~\ref{loss_wemb} (\texttt{wemb}); 
Finally, we apply the correction algorithm detailed in Section~\ref{correction} (\texttt{wemb+fix}).
Columns Ref and Fix indicate the number of original and corrected sentences (in millions) considered to train NMT.

\begin{table}[h!]
\small
\center
\begin{tabular}{lccl}
\hline
\bf Data & \bf Ref (M) & \bf Fix (M) & \bf Test (BLEU) \\ %MSLT & \bf NEWS \\
\hline
\multicolumn{3}{c}{\scriptsize{OpenSubtitles English-French}} \\
\texttt{all}                  & 27.2 & - & 42.18 \\
\texttt{semb}            & 18.0 & - & 41.95\\
\texttt{wemb}           & 15.5 & - & 43.12  ($+0.94$)\\
%\texttt{wemb}           & 21.5 & - & 42.56 \\
\texttt{wemb+fix}   & 15.5 & 2.5 & \\
\texttt{wemb}           & 18.0 & - & 43.19  ($+1.01$)\\
%\texttt{wemb+fix}   & 15.5 & 5.0 & \\
\hline
\multicolumn{3}{c}{\scriptsize{Paracrawl English-German}} \\
%\texttt{all}                  & 100.0 & 12.56 \\ 
\texttt{all}                  & 22.2 & - & 19.27 \\ 
\texttt{semb}            & 15.0 & - & \\
\texttt{wemb}           & 15.0 & - & 21.52 ($+2.25$)\\
%\texttt{wemb}           & 12.5 &   \\
%\texttt{wemb+fix}   & 12.5+2.5 & \\
\hline
\end{tabular}
\caption{BLEU scores obtained by neural MT using different subsets of the OpenSubtitles and Paracrawl corpora.}
\label{results}
\end{table}

Results for \texttt{wemb} clearly outperform the baseline \texttt{all} ($1^{st}$ rows) by about $+1.0$ and $+2.25$ BLEU ($3^{rd}$ rows). 
In contrast, \texttt{semb} does not improve over the baseline ($2^{nd}$ rows).
Regarding OpenSubtitles, an additional gain of $+X.X$ is found when fixing $2.5$ million sentences ($4^{th}$ row). 
The same $2.5$ million sentences did not show any improvement when added in their original form ($5^{th}$ row).

At this moment experiments showing correction results regarding Paracrawl  are not yet finished.

\section{Conclusions}
\label{conclusions}

We presented an unsupervised method based on deep neural networks for detecting translation divergences in parallel sentence pairs. 
Our model also predicts misaligned words that can then be filtered out allowing for reusing some divergent sentences.
We evaluated our model on two neural machine translation tasks, showing improvements when compared to training over the entire data set. 
%The method can be used on any parallel corpus without any manual annotation.

\section{Further Work}
\label{further}

%We plan to further evaluate divergence classification under different data noise levels. %and on additional language pairs. 
%as well as to measure the impact of using less noisy data when learning the similarity model
%We also would like to evaluate the impact of using less noisy (already filtered) data to learn the similarity model. 
%We would also like to use our model to predict other types of divergences 
We plan to use our model to predict sentence embeddings over monolingual corpora allowing to collect parallel pairs through vector similarity measures.
%We also plan to study the use of a neural network to predict divergent words, as a replacement for the algorithm detailed in Section~\ref{correction}.

%\section*{Acknowledgements}
%We are grateful to Jo\"el Legrand for his fruitful comments when building the neural divergence classifier.

\bibliography{biblio}
\bibliographystyle{acl_natbib_nourl}

\end{document}

### 2018-07-17 WMT/divergence.tex
add WMT paper

%
% File emnlp2018.tex
%
%% Based on the style files for EMNLP 2018, which were
%% Based on the style files for ACL 2018, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{emnlp2018}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
%%% YOUR PACKAGES BELOW THIS LINE %%%
\usepackage[small,bf]{caption} % added MLF 20171211
\usepackage{multicol}
\usepackage{adjustbox}
\usepackage{graphicx}
\usepackage{array,multirow}
%\aclfinalcopy % Uncomment this line for the final submission

\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}
\newcommand\confname{EMNLP 2018}
\newcommand\conforg{SIGDAT}

\title{Crosslingual Sentence Embeddings for Neural Machine Translation}
%\title{A Recurrent Network for Parallel Corpora Divergence Analysis}

\author{%Minh Pham, \ Fran\c cois Yvon\\
  First author\\
  Affiliation / Address line 1\\
  Affiliation / Address line 2\\
  Affiliation / Address line 3\\
  {\tt email@domain}  
  \And
  %Josep Crego, \ Jean Senellart\\
  Second author\\
  Affiliation / Address line 1\\
  Affiliation / Address line 2\\
  Affiliation / Address line 3\\
  {\tt email@domain}}

\date{}

\begin{document}
\maketitle
\begin{abstract}

%http://www.airccj.org/CSCP/vol4/csit42503.pdf

Corpus-based approaches to machine translation rely on the availability and quality of parallel corpora.
In the case of neural machine translation, a large neural network is trained to maximise the translation performance on a given parallel corpus. 
%This paper describes an unsupervised method for building cross lingual sentence embeddings, which can then be used to compute semantic similarity between sentences in different languages.
This paper describes an unsupervised method for building cross lingual sentence embeddings that optimises word alignments. Embeddings can then be used to compute semantic similarity between sentences of different languages.
We report competitive performance on two different tasks. 
Firstly, sentence embeddings are used to detect translation divergencies in parallel sentence pairs, hence allowing for filtering out noisy pairs.
Then, we use sentence embeddings to discover parallel sentences on comparable corpora.
We evaluate the presented model for different language pairs and data resources.
The model can be used on any parallel corpus without any manual annotation.

%According to the process followed to compile a parallel corpus, it may contain multiple parallel sentences that are often not as parallel as one might assume. Examples are easily found when corpora is created from non-parallel texts or included by noisy document/sentence alignment tools.
%This paper describes an unsupervised method for detecting translation divergencies in parallel sentence pairs. We use a deep neural network to predict word alignment scores in parallel sentences. Misaligned words are then identified allowing divergent parallel sentences to be filtered out and in some cases making sentences parallel by removing such words. 
%We evaluate the presented method on a machine translation task. Results show that a neural MT system trained on the filtered/corrected sentences outperforms the neural MT system trained on the original sentence pairs. This method can be used on any parallel corpus without any manual annotation.

\end{abstract}

\section{Introduction}

Parallel sentence pairs are the only necessary resource to build Machine Translation (MT) systems.
In the case of neural MT, a large neural network is trained through maximising translation performance on a given parallel corpus. 
Therefore, the quality of an MT engine is heavily dependent upon the amount and quality of parallel sentences. 
Note that recent work on neural MT remove completely the need of parallel data following unsupervised methods~\cite{lample2018word,artetxe2018iclr} obtaining accuracy improvements over word-by-word statistical MT systems.

Unfortunately, parallel texts are scarce resources. 
There are relatively few language pairs for which parallel corpora of reasonable sizes are available, and even for those pairs, the corpora come mostly from few domains. 
To alleviate the lack of parallel data several approaches have been developed in the last years.
They range from methods using non-parallel, or comparable data ~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,W17-2509,P17-3003} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714} using in all cases automated alignment/translation engines that are prone to the introduction of noise in the resulting parallel sentences. 
Mismatches on parallel sentences extracted from translated texts are also reported ~\cite{tiedemann2011bitext,XU16.310}. 
This problem is mostly ignored in machine translation, where parallel sentences are considered to convey the exact same meaning, and is particularly important for neural MT engines as suggested by~\cite{chen2016adaptation}.

Table~\ref{tab:examples} illustrates some examples of English-French parallel sentences which are not completely semantically equivalent. Divergencies are outlined using bold letters. Examples extracted from the OpenSubtitles corpus ~\cite{LisonTiedemann2016}.

\begin{table}[ht]
%\begin{adjustbox}{width=0.48\textwidth}
\small
\center
\begin{tabular}{ c|l }
  %\hline
  \hline  
  \texttt{en} & \it{What do you feel}\bf{, Spock}\it{?} \\
  \texttt{fr} & \it{Que ressentez-vous?} \\
  \texttt{gl} & {\small \it{What do you feel?}} \\
  \hline
  \texttt{en} & \it{How much do you get paid?} \\
  \texttt{fr} & \it{T'es pay\'e combien} \bf{de l'heure}\it{?} \\
  \texttt{gl} & {\small \it{How much do you get paid per hour?}} \\
  \hline  
  \texttt{en} &  \bf{That seems a lot.} \\
  \texttt{fr} & \bf{40 livres?} \\
  \texttt{gl} & {\small \it{40 pounds?}} \\
  \hline  
  \texttt{en} & \it{I brought you} \bf{french fries}\it{!} \\
  \texttt{fr} & \it{Je t'ai rapport\'e des} \bf{saucisses}\it{!} \\
  \texttt{gl} & {\small \it{I brought you sausage!}} \\
  \hline
  %\hline  
\end{tabular}
%\end{adjustbox}
\caption[Table caption text]{Examples of semantically divergent parallel sentences. English (\texttt{en}), French (\texttt{fr}) and gloss of French (\texttt{gl}).}
\label{tab:examples}
\end{table}

Different types of translation divergencies can be found in a parallel corpus:
Additional segments are included on either side of the parallel sentences (first and second rows) most likely due to errors in sentence segmentation tools;
Some translations may be completely uncorrelated (third row);
Inaccurate translations are also included (fourth row). 
Note that divergent translations can be due to many different reasons ~\cite{C14-1055}, the study of which is beyond the scope of this paper.

In this work, we present an unsupervised method for building cross lingual sentence embeddings in the form of continuous vectors. 
Embeddings are then used to measure semantic equivalence of their corresponding sentences.
We evaluate the method on two different tasks:
First, we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation task.
Then, we also show competitive results when identifying parallel sentences in comparable corpora using the French-English data made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}.

%In addition, we  show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.

The remainder of this paper is structured as follows. 
Section~\ref{related} overviews related work. 
We describe in detail the core of the neural similarity classifier in Section~\ref{sec:similarity}. 
We report experiments with the presented model on two different tasks in Section~\ref{experiments}.
Section~\ref{sec:results} evaluates results. 
Finally, conclusions are drawn in Section~\ref{conclusions} and further work is outlined in Section~\ref{further}.
All the code used in this paper is freely available\footnote{https://github.com/anonymised}.

\section{Related Work}
\label{related}

\subsection{Translation Divergencies}

Prior work on translation divergencies focused on reflecting that translation of one language into another results in very different forms ~\cite{J94-4004} according to several categories (e.g. lexical, structural, categorical).

Attempts to measure the impact of translation divergencies in MT systems have focused on the introduction of noise in sentence alignments ~\cite{goute2012}, showing that statistical MT systems are highly tolerant to noise, and that performance only degrades seriously at very high noise levels. In constrast, neural MT engines seem to be more sensitive~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.

Efforts have been devoted to characterising the degree of semantic equivalence between two snippets of text in the same or different languages~\cite{conf/semeval/AgirreBCDGMRW16}, a workshop devoted to an objective similar to our work. In~\cite{Mueller:2016:SRA:3016100.3016291}, a monolingual sentence similarity network is proposed, making use of a simple LSTM layer to compute sentence representations. The authors show that a simple SVM classifier can be built on top of the sentence representations to achieve state-of-the-art results in a semantic entailment classification task. With the same objective, the system presented in~\cite{N16-1108} uses multiple convolutional layers and models pairwise word interactions. %Our work differs from the previous as we build a different network and use it to measure similarity of sentences in different languages.

Our work is inspired by~\cite{W17-3209} where the authors train a cross-lingual divergence detector using word alignments and sentence length features to train a linear SVM classifier. Their work shows that an NMT system trained only on non-divergent sentences yields slightly higher translation quality scores and requires clearly less training time. The same authors have recently updated their work in~\cite{DBLP:journals/corr/abs-1803-11112}. The objective is the same as the neural network presented in~\cite{N16-1108} and their network further outperforms their previous work. Our work differs from the previous as we make use of a network with different topology. We model sentence similarity by means of optimising word alignment and in addition, we propose a simple method that employs alignment scores to not only detect but correct divergent sentences.

\subsection{Extracting Parallel Sentences}

As previously outlined, parallel texts are scarce resources as well as the only necessary resource to build Machine Translation systems. In contrast, there is an increasing amount of comparable corpora available on the Internet, a potential solution to alleviate the data scarcity problem.

Several approaches exist to extract parallel sentences from comparable corpora. 
In~\cite{J05-4003} is presented a maximum entropy classifier that can reliably determine whether a pair of sentences are translations of each other. 
A similar approach is proposed by~\cite{E09-1003}, where a SMT system is used to translate the source language side of a comparable corpus to find candidate sentences on the target side. 
Word error rate and translation error rate measures are used to decide when a candidate is finally considered parallel. 
More recently,~\cite{W17-2508} present a method based on seed lexical translations, simple set expansion operations and the Jaccard similarity coefficient. The method obtains state-of-the-art results in the task of parallel sentence extraction from comparable corpora.

Closely related to the method proposed in this paper, different works employ continuous vector sentence representations as a preceding step to compute sentence similarity. 
In~\cite{W15-1521}, word representations from bilingual word embeddings are used to build a convolutional network that predicts wether a sentence pair is aligned or not.
~\cite{W17-2509} propose a recurrent network to estimate sentence embeddings followed by dense layers that estimates the conditional probability distribution that two sentences are parallel.

These approaches are different from ours FINISH

This strong interest in comparable corpora was further boosted after the apparition of the BUCC workshop series on Building and Using Comparable Corpora\footnote{https://comparable.limsi.fr/bucc2018/} which helped to provide a task definition, datasets and evaluation methods to assess the state of the art.

\section{Neural Similarity Classifier}
\label{sec:similarity}

In this section we describe the architecture of our model, very much inspired by the work on Word Alignment in~\cite{W16-2207}. Figure~\ref{network} illustrates the network.

In the following, we consider a source-target sentence pair $(s,t)$ with $s=(s_1,...,s_I)$ and $t=(t_1,...,t_J)$. The model is composed of 2 Bi-directional LSTM subnetworks, $net_s$ and $net_t$, which respectively encode source and target sentences. Since both $net_s$ and $net_t$ take the same form we describe only the source architecture.

The source-sentence Bi-LSTM network outputs forward and backward hidden states, $\overrightarrow{h}^{src}_i$ and $\overleftarrow{h}^{src}_i$, which are then concatenated into a single vector encoding the $i^{th}$ word of the source sentence, 
$h^{src}_i = [ \overrightarrow{h}^{src}_i ; \overleftarrow{h}^{src}_i ]$.
%$h^{src}_i = \left( \begin{array}{cc} \overrightarrow{h}_i \\ \overleftarrow{h}_i \\ \end{array} \right)$.

In addition, the last forward/backward hidden states (outlined using dark grey in Figure~\ref{network}) are also concatenated into a single vector  to represent whole sentences 
$h_{src} = [ \overrightarrow{h}^{src}_I ; \overleftarrow{h}^{src}_1 ]$.
%$h_{src} = \left( \begin{array}{cc} \overrightarrow{h}^{src}_I \\ \overleftarrow{h}^{src}_1 \\ \end{array} \right)$.
At this point a measure of similarity between sentence pairs can be obtained by cosine similarity:

\begin{equation}
    sim(h_{src}, h_{tgt}) = \frac{h_{src} \cdotp h_{tgt}}{||h_{src}|| * ||h_{tgt}||}
    \label{cosine}
\end{equation}
\noindent where two vectors (embeddings) with the same orientation have a cosine similarity of $1$, while two vectors with opposed orientation have a similarity of $-1$, independent of their magnitude.

\begin{figure}[h]
\center
    \includegraphics[width=1.0\linewidth]{network}
    \caption{Illustration of the model. The network is composed of source and target word embedding lookup tables ($LT_s$ and $LT_t$) and two identical subnetworks ($net_s$ and $net_t$) that compute in context representations of source ($s_i$) and target words ($t_j$).} 
    \label{network}
\end{figure}
%The similarity function outputs the level of semantic similarity between source and target sentences.

%Figure~\ref{net_lstm} illustrates the first component of our network. In order to optimise our model we follow the functions depicted in Figure~\ref{net_align}.

Similar to ~\cite{W16-2207} our model extracts context information from source and target sentences and then computes simple dot-products to estimate word alignments. The objective function is computed at the level of words. 
To enable unsupervised training, we use an aggregation operation that summarizes the alignment scores for a given target word. A soft-margin objective increases scores for true target words while decreasing scores for target words that are not present.
The aggregation function combines the scores of all source (or target) words for a particular target (or source) word and promotes source words which are likely to be aligned with a given target word according to the knowledge the model has learned so far.

Alignment scores $S(i,j)$ are given by the dot-product $S(i,j) = h_i^{src} \cdotp h_j^{tgt}$, while aggregation functions are defined as:

\begin{equation}
\begin{split}
    aggr_s(i,S) = \frac{1}{r} \ log \left( \displaystyle \sum_{j=1}^{J} e^{r * S(i,j)}\right) \\
    aggr_t(j,S) = \frac{1}{r} \ log \left( \displaystyle \sum_{i=1}^{I} e^{r * S(i,j)}\right)
\end{split}
\label{aggregation}
\end{equation}

%Hence, rather than sentence similarity our model optimises word alignments.
The loss function is defined as:

%\vspace{-8mm}
\begin{equation}
\begin{split}
\mathcal{L}(src,tgt) = & \\
    \sum_{i=1}^I log&\left(1+e^{aggr_{s}(i,S) * \mathcal{Y}_i^{src}}\right) +\\
 + \sum_{j=1}^J log&\left(1+e^{aggr_{t}(j,S) * \mathcal{Y}_j^{tgt}}\right)
\end{split}
\label{loss_wemb}
\end{equation}
\noindent where $\mathcal{Y}_i^{src}$ (or $\mathcal{Y}_j^{tgt}$) is a vector with reference labels containing $-1$ when the $i^{th}$ source word (or $j^{th}$ target word) is present in the translated target (or source) sentence, and $+1$ for divergent (unpaired) words.

%\begin{figure}[h]
%\center
%    \includegraphics[width=0.9\linewidth]{net_align}
%    \caption{Illustration of the model.}
%    \label{net_align}
%\end{figure}

%\vspace{-8mm}

For comparison purposes we also implemented the model introduced in \cite{W17-2509}. The network employs the same initial Bi-LSTM layers outlined in Figure~\ref{network} which encode source and target sentences into $h_{src}$ and $h_{tgt}$ vectors. The remaining of the network is illustrated by Figure~\ref{net_sentence}, where sentence pair matching information is captured by using their element-wise product and absolute element-wise difference. Sentence similarities are finally estimated by feeding the matching vectors into a fully connected layer.

\begin{equation}
\begin{split}
h^{(1)} =  &\ \ h_{src} * h_{tgt} \\
h^{(2)} =  &\ \ |h_{src} - h_{tgt}| \\
h^{(3)} = &\ \ tanh(W[h^{(1)};h^{(2)}] + b^{(3)}) \\
h_{sim} = &\ \ Wh^{(3)} + b^{sim}
\end{split}
\end{equation}

This approach is different from our model since it uses a single end-to-end model to estimate the conditional probability distribution that two sentences are parallel.

\begin{figure}[h]
\center
    \includegraphics[width=0.8\linewidth]{net_sentence}
    \caption{Illustration of the model presented in~\cite{W17-2509}. Sentence embeddings (vectors in dark grey) are computed by the same Bi-LSTM layers shown in Figure~\ref{network}.} %The network is composed of source and target word embedding lookup tables ($LT_s$ and $LT_t$) and two identical subnetworks ($net_s$ and $net_t$) that compute in context representations of source ($s_i$) and target words ($t_j$). The similarity function outputs the level of semantic similarity between source and target sentences.}
    \label{net_sentence}
\end{figure}

For this second network the loss function is defined as: 
%\vspace{-5mm}
\begin{equation}
\mathcal{L}(src,tgt) = log(1+e^{h_{sim} * \mathcal{Y})}) % where $sign(src,tgt) =1
\label{loss_semb}
\end{equation}
\noindent where $\mathcal{Y}$ is the reference label containing $-1$ when the given sentence pair is non-divergent (parallel) and $+1$ for divergent (unpair) examples.

Note that different from the network described in the original work, our implementation does not estimates a probability distribution but FINISH.

%For each source(target) word, we consider the aggregated matching score over the target(source) sentence:
%$$s_{aggr}(i,tgt) = Aggr_{j=1}^{J} s(i,j)$$
%$$s_{aggr}(src,j) = Aggr_{i=1}^{I} s(i,j)$$

%Where $Aggr$ is LSE aggregation operator used in~\cite{W16-2207}. Given the sign of each word noted $sign(w)$, that indicates whether the word $w$ is matched to the other side (by convention, $1$ indicates non-matching, $-1$ indicates matching), the loss function is defined as follows:

%In the second option, we directly optimize the score of similarity between two sentences by defining the loss function as follows:

%$$L(src,tgt) = log(1+e^{h_{src} . h_{tgt} * sign(src,tgt)})$$ where $sign(src,tgt) =1 $ if two sentences are not parallel and $sign(src,tgt) =-1 $ otherwise.

\subsection{Training with Negative Examples}
\label{training}

Training is performed by minimising Equation~\ref{loss_wemb}, for which examples with annotations for source $\mathcal{Y}_i^{src}$ and target $\mathcal{Y}_j^{tgt}$ words are needed.

As positive examples we use {\bf paired} sentences of a parallel corpus. All words of the pair of sentences are labelled as parallel, $\mathcal{Y}_i^{src}=-1$ and $\mathcal{Y}_j^{tgt}=-1$. 
As negative examples we use random {\bf unpaired} sentences. In this case, all words are labelled as divergent, $\mathcal{Y}_i^{src}=+1$ and $\mathcal{Y}_j^{tgt}=+1$. 
Since negative examples may be very easy to classify and we want our network to detect less obvious divergencies, we further create negative examples following the next two methods:

We {\bf replace} random sequences of words on either side of the sentence pair by a sequence of words with the same part-of-speeches. 
The rationale behind this method is to keep the new sentences as grammatical as possible. 
Otherwise the network can learn to detect non-grammatical sentences to predict divergence.
Words that are not replaced are considered parallel ($-1$) while those replaced are assigned the divergent label ($+1$). 
Words aligned to some replaced words are also assigned the divergent label ($+1$). For instance, given the original sentence pair:

\begin{table}[h]
\begin{tabular}{ll}
src: & { \small \texttt{What do you feel ?}} \\
tgt: & { \small \texttt{Que ressentez-vous ?}} \\
\end{tabular}
\end{table}

We may replace '\texttt{you feel}', with part-of-speech tags '\texttt{PRP VB}', by another sequence with same tags (i.e. '\texttt{we want}'):

\begin{table}[h]
\begin{tabular}{ll}
src: & { \small \texttt{What do {\bf we \ want} ?}} \\
$\mathcal{Y}^{src}$: & { \small \texttt{-1 \ \  -1 {\bf +1\ \ +1} \ \  -1}} \\
tgt: & { \small \texttt{Que {\bf voulez-vous} ?}} \\
$\mathcal{Y}^{tgt}$: & { \small \texttt{-1\ \ {\bf +1}\ \ \ \ \ \ \ -1}} \\
\end{tabular}
\end{table}

Divergent words are shown in bold. 
Note that the new sentences are not assured to be grammatical after replacing sequences with the same part-of-speech.
%Note that part-of-speeches and word alignments are needed to build this negative example. 
We use word alignments to identify as divergent the sequence '\texttt{ressentez-vous}' since it is aligned to the replaced sequence '\texttt{you feel}'.

Finally, motivated by sentence segmentation errors observed in many corpora, we also build negative examples by {\bf inserting} a second sentence at the beginning (or end) of the source (or target) sentence pair. 
Words in the original sentence pair are assigned the parallel label ($-1$) while the new words inserted are considered divergent ($+1$).
Given the original sentence pair previously shown, we may want to build the next negative example by inserting the sentence '\texttt{Not .}' at the end of the original source sentence:

\begin{table}[h]
\begin{tabular}{ll}
src: & { \small \texttt{What do you want ? {\bf Not \ .}}} \\
$\mathcal{Y}^{src}$: & { \small \texttt{-1 \ \  -1 -1 \ -1  \ \ -1 {\bf +1\ \ \  +1}}} \\
tgt: & { \small \texttt{Que ressentez-vous ?}} \\
$\mathcal{Y}^{tgt}$: & { \small \texttt{-1\ \ -1\ \ \ \ \ \ \ \ \ \ \ \ \ -1}} \\
\end{tabular}
\end{table}

In order to avoid that negative examples are easily predicted just by looking at the difference in length of training sentences we constraint all negative examples to have a difference in length not exceeding $2.0$. Very short sentences, up to $4$ words, are accepted if the length ratio does not exceeds $3.0$.

The network outlined in Figure~\ref{net_sentence} uses a single label for each sentence pair. 
Hence, positive examples are labelled $\mathcal{Y}=-1$ while any of the previously described negative examples are labelled $\mathcal{Y}=+1$.

\subsection{Fixing Translation Divergencies}
\label{correction}

We observed in our training corpus that many divergent sentences follow a common pattern, consisting of containing some extra leading/trailing words. See the first and second examples of table~\ref{tab:examples}. 
Accordingly, we implemented an algorithm that discards sequences of leading/trailing words making use of alignment scores $S(i,j)$. 
Hence considering as parallel $s_[^]$ and $t_\langle^\rangle$. To find the optimal source ($[, ]$) and target ($\langle, \rangle$) indexes that enclose the corrected segment within the original sentence, we implement:
%\vspace{-4mm}
\begin{equation*}
\begin{split}
%\underset{[, ], \{, \}}{\arg\max}&\{\mathcal{P}([,],\{,\}) - \mathcal{Q}([,],\{,\})\} \\
\underset{[, ], \langle, \rangle}{\arg\max}& \Big \{      \underset{[ \le i \le ]}{\sum} \underset{\langle \le j \le \rangle}{\max} \{ S(i,j) \}  \Big \} % -   \underset{[ > i, i > ]}{\sum} \underset{\langle > j, j > \rangle}{\max} \{ s(i,j) \}   \Big \}
\end{split}
\end{equation*}
%\noindent where $\mathcal{P}$ and $\mathcal{Q}$ are defined:
%\vspace{-4mm}
%\begin{equation*}
%\begin{split}
%\mathcal{P}&([,],\{,\}) = \underset{[ \le i \le ]}{\sum} \underset{\{ \le j \le \}}{\max} \{ s(i,j) \} \\%+ \underset{\{ \le j \le \}}{\sum} \underset{[ \le i \le ]}{\max} \{ s(i,j) \}  \} \\
%\mathcal{Q}&([,],\{,\}) = \underset{[ > i, i > ]}{\sum} \underset{\{ > j, j > \}}{\max} \{ s(i,j) \} %+ \underset{\{ > j < \}}{\sum} \underset{[ > i < ]}{\max} \{ s(i,j) \}  \} \\
%\end{split}
%\end{equation*}

The $\mathcal{N}$-best sequences following the previous function ($s_[^]$, $t_\langle^\rangle$) are considered as valid corrections, but only the highest ranked according to their similarity score is used as replacement for the original $(s_1^I, t_1^J)$. % if it has a higher similarity score: $sim(s_[^], t_\{^\}) > sim(s_1^I, t_1^J)$. 
Short sentences are not considered. This is, $] - [ > \tau$ and $\langle - \rangle > \tau$.

Figure~\ref{matrix} (left) shows an example of an alignment matrix $s(i,j)$ for a given sentence pair. A possible correction is: \texttt{What do you feel ? $\Leftrightarrow$ Que ressentez-vous ?}. Hence, with indexes $[=1$, $]=5$, $\langle=1$ and $\rangle=3$.

%\begin{figure*}
%\centering
%\begin{multicols}{2}
%    \includegraphics[width=1.1\linewidth]{learn}    
%    \includegraphics[width=1.1\linewidth]{summer}    
%    \includegraphics[width=1.1\linewidth]{swim} 
%    \includegraphics[width=1.4\linewidth]{awesome}
%\end{multicols}
%\begin{multicols}{3}
%    \includegraphics[width=\linewidth]{sandwich}
%    \includegraphics[width=\linewidth]{sandwich}
%    \includegraphics[width=\linewidth]{peu}
%\end{multicols}
%\caption{Examples of neural alignment output by the matrix $S_{i,j}$ in Figure~\ref{network}. Only positive scores, indicating alignment between words $s_i$ and $t_j$, are printed. Different levels of grey are used to highlight the alignment.}
%\end{figure*}

\section{Experiments}
\label{experiments}

In this section we report on the experiments conducted to evaluate the proposed network. We begin with details of the corpora employed.

\subsection{Corpora}
\label{corpora}

Different corpora are used in this work to evaluate our similarity classifier.

We filter out divergencies found in the English-French OpenSubtitles corpus~\cite{LisonTiedemann2016}, which consists of a collection of movie and TV subtitles, aligned at the sentence level after a number of automatic preprocessing steps. 
The corpus presents many potential divergencies as outlined in Table~\ref{tab:examples}. 
To evaluate performance we used the En-Fr Microsoft Spoken Language Translation task, created from actual conversations over Skype~\cite{mslt-corpus-iwslt-2016-release}. 
%We also tested on a second domain, using the publicly available newstest-2013 En-Fr test set, corresponding to news stories selected from online sources~\cite{WMT:2013}.
In order to better assess the quality of our classifier when facing different types of word divergencies we collected from the original corpus and annotated at the word level $500$ sentences containing:
200 paired sentences;
100 unpaired sentences;
100 sentences with replace examples; and
100 sentences with insert examples as detailed in Section~\ref{training}.

We also use the Paracrawl corpus, which consists of a very noisy $1$ billion word (English token count) German-English corpus crawled from the web as part of the Paracrawl project\footnote{http://paracrawl.eu/}.
Performance is evaluated on the publicly available newstest-2017 En-De test set, corresponding to news stories selected from online sources~\cite{W17-4717}.

To evaluate the ability of our sentence embeddings to identify parallel sentences in comparable corpora we followed the data conditions made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}. 
For training we used the French-English European Parliament corpus from WMT15\footnote{http://www.statmt.org/wmt15/translation-task.html}.
For testing we used the test set made available by the task organisers, consisting of about $300,000$ monolingual sentences per language, French and English, that contains near $10,000$ parallel sentences.
Test data come from Wikipedia\footnote{http://ftp.acc.umu.se/mirror/wikimedia.org/dumps} and News Commentary\footnote{http://www.casmacat.eu/corpus/news-commentary.html}.

\subsection{Training}

All data used in this paper is preprocessed with an in-house toolkit that performs minimal tokenisation, basically splitting-off punctuation.

\subsubsection{Neural Similarity Classifier}
\label{divergence}

After tokenisation, each out-of-vocabulary word is mapped to a special UNK token.
Our similarity classifier is described in Section~\ref{sec:similarity}. 
Word embeddings ($LT_s$ and $LT_t$) are initialised using \texttt{fastText}\footnote{https://github.com/facebookresearch/fastText}, further aligned by means of \texttt{MUSE}\footnote{https://github.com/facebookresearch/MUSE} following the unsupervised method detailed in~\cite{lample2018word}. 
Size of embeddings is $E_s=E_t=256$ cells. 
Both Bi-LSTM use 256-dimensional hidden representations ($E=512$). 
Optimization of the parameters is done using the stochastic gradient descent method along with gradient clipping (rescaling gradients whose norm exceeds a threshold) to avoid the exploding gradients problem~\cite{Pascanu:2013:DTR:3042817.3043083}. 
For each epoch we randomly select $1$ million sentence pairs that we place in batches of $32$ examples.  
Negative examples are created following the strategies detailed in~\ref{training}.% ({\bf P}aired, {\bf U}unpaired, {\bf R}eplace and {\bf I}nsert).
We run $10$ epochs and start decaying at each epoch by $0.8$ when score on validation set increases.

Two networks are evaluated. 
The first is the work presented in this paper, it optimises the loss shown by equation~\ref{loss_wemb}, and uses Equation~\ref{cosine} to compute similarity scores (\texttt{wemb}).
The second implements the work in~\cite{W17-2509}, it optimises the loss shown by Equation~\ref{loss_semb} which in an end-to-end fashion predicts wether two sentences are parallel (\texttt{semb}).
To fix translation divergencies we used $\mathcal{N}=20$ and $\tau=3$.

\subsubsection{Neural MT}
\label{translation}

In addition to the basic tokenisation we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30,000$ merge operations learned from both English and French data.

We build our NMT systems based on the open-source project \texttt{OpenNMT}\footnote{http://opennmt.net}. We use a bidirectional RNN encoder with $4$ LSTM layers with each containing $1,000$ cells. Word embeddings have a size of $300$ cells. We set the dropout probability to $0.3$. Batch size is set to $64$. The maximum length of both source and target sentences is set to $80$. % and we limit the vocabulary size to $50K$ words for both source and target languages.
The default optimiser is SGD with the starting learning rate of $1.0$. We start to decay the learning rate from epoch $10$ or when we detect increasing perplexity as compared to the previous epoch on a validation set. We stop training after $20$ epochs.

\section{Results}
\label{sec:results}

We evaluate first the ability of our similarity classifier to predict different types of divergencies. 
We use the test set manually annotated for that purpose and train our model on the OpenSubtitles corpus.
The same number of paired examples (P) is used than for any other class (U, R or I). %This is, in the case of the model trained using with all types of examples, the same number of examples of each class is used.

\begin{table}[h]
\small
\center
\begin{tabular}{crccccc}
\hline
\multicolumn{2}{l}{\bf Accuracy} & \multicolumn{5}{c}{Test examples} \\
 &  & \texttt{P} & \texttt{U} & \texttt{R} & \texttt{I} & \texttt{PURI} \\
 \hline
\parbox[t]{0mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Train examples}}} &  \texttt{PU}     & \bf 0.996 & \bf 0.994 & 0.671 & 0.673 & 0.874 \\
 &  \texttt{PR}     & \bf 0.995 &      0.033 & \bf 0.951 &      0.689 & 0.746 \\
 &  \texttt{PI}       & \bf 0.998 &      0.071 &      0.697 & \bf 0.725 & 0.705 \\
 &  \texttt{PUR}  & \bf 0.994 & \bf 0.989 & \bf 0.919 &      0.710 & 0.932 \\
 &  \texttt{PUI}    & \bf 0.995 & \bf 0.996 &      0.662 & \bf 0.769 & 0.887 \\
 &  \texttt{PRI}    & \bf 0.991 &      0.161 & \bf 0.924 & \bf 0.719 & 0.768 \\
 &  \texttt{PURI} & \bf 0.995 & \bf 0.980 & \bf 0.916 & \bf 0.788 & \bf 0.942 \\
\hline
\end{tabular}
\caption{Word divergence accuracies for networks trained with different combinations of negative examples. P, U, R and I stand respectively for pair, unpair, replace and insert.}
\label{results_puri}
\end{table}

Table~\ref{results_puri} shows accuracies obtained by our model when trained over different combinations of negative examples on the task of predicting wether words are divergent. 
We consider divergent those words for which the similarity score predicted by our model, Equation~\ref{cosine}, is negative.

As it can be seen, non-divergent words in parallel sentences (column \texttt{P}) are easy to identify. 
%In all cases the accuracy reaches $100\%$. 
Divergent words in unpaired sentences (column \texttt{U}) are also easy to identify as far as the model has seen unpaired examples in training. 
The accuracy drops dramatically when the model is not used to see unpaired sentences (rows  \texttt{PR}, \texttt{PI} and \texttt{PRI}).

Regarding columns \texttt{R} and \texttt{I}, accuracies are lower since these sentences contain a mix of divergent and non-divergent words. 
Hence, word divergence is more difficult to predict. 
Again, models that were trained with the corresponding examples obtain the highest accuracies (outlined in bold letters).
Column \texttt{PURI} shows accuracy results over the entire test set. This is, mixing all type of examples. 
The best accuracy is obtained by the system trained on all type of negative examples.

%The test set was annotated after looking at the entire train set and looking for these type of examples. 
%Existance and distribution of types of

\begin{figure}[h]
\centering
    \includegraphics[width=0.4\linewidth]{feel}    
    \includegraphics[width=0.4\linewidth]{feel_pu}    
\caption{Example of neural alignment matrices and divergence scores obtained with our model when trained over pair and unpair examples (right), and over all types of examples (left) as detailed in Section~\ref{training}. Aggregation scores of Equation~\ref{aggregation} are shown next to source and target words. Matrices contain word alignment scores. Sentence similarity scores computed by Equation~\ref{cosine} are shown below each matrix.}
\label{matrix}
\end{figure}

Figure~\ref{matrix} shows an example of word and sentence similarity predictions when our network is trained using all types of negative examples (left); and trained using only paired and unpaired sentences (right). 
As it can be seen, the network trained using examples where all words are either divergent or non-divergent (right), fails to correctly predict the class for words in sentences with mixed divergent/non-divergent words. In contrast, the model trained with all types of examples (left) correctly predicts the right class for each word. Furthermore, it correctly assigns a lower similarity score to the sentence pair, as both sentences do not convey the exact same meaning.

\subsection{Divergent Sentences Identification Task}

Next, we evaluate the ability of our neural similarity classifier to detect divergent sentence pairs. 
Since we lack of a gold standard test set to directly evaluate sentence similarity, 
we measure the accuracy of neural MT systems trained over the original corpus and after filtering divergent sentences. 
Notice that some divergent sentences may not be completely useless to train a neural MT system. 
Consider for instance the example of Figure~\ref{matrix}. 
Despite not conveying the exact same meaning it still contains useful information when training a neural MT engine. 
It is not the case of the example in the third row of Table~\ref{tab:examples}.

\begin{table}[h]
\small
\center
\begin{tabular}{lcc}
\hline
\bf $sim(h_{src},h_{tgt})$ & \bf Size (M) & \bf Test (BLEU)\\%& \bf NEWS \\
\hline
$(-\infty,+\infty)$   & 27.2 & 42.18 \\ %& 26.60  \\
$[0.000,+\infty)$   & 24.0 & 42.68 \\ %& 26.87  \\
$[0.003,+\infty)$   & 21.5 & 42.56 \\ %&  \\
$[0.076,+\infty)$   & 18.0 & \bf 43.19 \\ %& 26.38 \\
$[0.100,+\infty)$   & 15.5 & \\%&  \\
\hline
\end{tabular}
\caption{BLEU scores obtained by a neural MT network when trained over the OpenSubtitles corpus filtered using different similarity thresholds.}
\label{results_wemb}
\end{table}

Table~\ref{results_wemb} shows BLEU~\cite{P02-1040} scores of the same neural MT network when trained over different samples of the OpenSubtitles corpus.
Samples result from filtering sentences of the original corpus using different similarity thresholds.
As it can be seen, the best performing training sample is obtained when using sentence pairs with similarity score in the range $[0.076, +\infty)$. 
This is, training with the most similar $18$ million sentences of the original data set.

%Table~\ref{results_nmt} shows BLEU scores for the same neural MT system trained with different samples of the OpenSubtitles and Paracrawl corpus.
For each of the corpora, OpenSubtitles and Paracrawl, Table~\ref{results_nmt} shows BLEU scores of the neural MT system trained over the entire data set (\texttt{All}), and over the $18$ (and $15$) million sentences with highest similarity score as predicted by \texttt{semb} and \texttt{wemb} classifiers. REPHRASE

\begin{table}[h]
\small
\center
\begin{tabular}{lcc}
\hline
\bf Data & \bf Size (M) & \bf Test (BLEU) \\ %MSLT & \bf NEWS \\
\hline
\multicolumn{3}{c}{\scriptsize{OpenSubtitles English-French}} \\
\texttt{All}                    & 27.2 & 42.18 \\%& 26.60 \\ 
\texttt{semb}               & 18.0 & \\%&  \\
\texttt{wemb}              & 18.0 & 43.19 \\%& 26.38 \\
\hline
\multicolumn{3}{c}{\scriptsize{Paracrawl English-German}} \\
\texttt{All}                    & 100 & 12.56 \\ 
\texttt{base}                & 22.0 & 19.27 \\ 
\texttt{semb}               & 15.0 &  \\
\texttt{wemb}              & 15.0 &   \\
\hline
\end{tabular}
\caption{.}
\label{results_nmt}
\end{table}

%clean.min1.max99.maxw40.fert6-5-3.data.tok.en.lid.filtered
In the case of the Paracrawl corpus, we applied first a \texttt{base} filtering that reduced the amount of sentences to $22$ millions. 
It consists of limiting the sentence size to $100$ words. 
We also limit the length ratio between source/target sentences to $6$, and we employed an in-house language Id classifier that helped us to detect sentence pairs not being English-German.

As it can be seen FINISH

\subsection{Parallel Sentences Identification Task}

\begin{table}[h]
\small
\center
\begin{tabular}{lrccccc}
\hline
\bf Model & $N=1$ & $10$ & $50$ & $100$ & $500$ & $1,000$ \\
\hline
\texttt{MUSE}           & & & & & & \\
\texttt{semb}            & & & & & & \\
\texttt{wemb}           & ssaling & & & & & \\
\hline
\end{tabular}
\caption{Number of parallel sentences found on the $N$-best cosine similarity filter (FAISS) according to the sentence embeddings produced by different networks.}
\label{results_embeddings}
\end{table}

The Cartesian product is costly for semb a filtering preprocess is necessary. Not for our model since we use faiss.

\begin{table}[h]
\small
\center
\begin{tabular}{lcccc}
\hline
\bf Model  & $\lambda$ & Precision & Recall & F-measure \\
\hline
\texttt{MUSE}             & & & & \\
\texttt{wemb}             & & & & \\
\texttt{semb}              & & & & \\
\hline
\end{tabular}
\caption{Precission, Recall and F-measure for the three evaluated models on the BUCC task.}
\label{results_similarity}
\end{table}

\section{Conclusions}
\label{conclusions}

We presented an unsupervised method based on deep neural networks for detecting translation divergencies in parallel sentence pairs. Our model also predicts misaligned words that can then be filtered out allowing for reusing some divergent sentences.
We evaluated our model on a neural machine translation task showing that it outperforms a baseline system trained on the entire (unfiltered) data set. The method can be used on any parallel corpus without any manual annotation.

\section{Further Work}
\label{further}

We plan to further evaluate divergence classification under larger data size conditions and different noise levels and on additional language pairs. 
%In this work we used a (noisy) parallel corpora to learn a divergence model. We also would like to evaluate the impact of using using less noisy (already filtered) data to learn the model. 
We also plan to study the use of a neural network to predict divergent words, as a replacement for the algorithm detailed in Section~\ref{correction}.

%\section*{Acknowledgements}
%We are grateful to Jo\"el Legrand for his fruitful comments when building the neural divergence classifier.

\bibliography{biblio}
\bibliographystyle{acl_natbib_nourl}

\end{document}

### 2018-07-25 WMT/divergence.tex
fix the abstract

@@
 %This paper describes an unsupervised method for building cross lingual sentence embeddings, which can then be used to compute semantic similarity between sentences in different languages.
-This paper describes an unsupervised method for building cross lingual sentence embeddings that optimises word alignments.
-Embeddings can then be used to compute semantic similarity between sentences of different languages.
+This paper describes an unsupervised method for building cross-lingual
+sentence embeddings that optimise word alignments. Such embeddings can then be used to compute semantic similarity between sentences in different languages.
 We report competitive performance on two different tasks. 
-Firstly, sentence embeddings are used to detect translation divergencies in parallel sentence pairs, hence allowing for filtering out noisy pairs.
-Then, we use sentence embeddings to discover parallel sentences on comparable corpora.
-We evaluate the presented model for different language pairs and data resources.
-The model can be used on any parallel corpus without any manual annotation.
+Sentence embeddings are first used to detect translation divergences
+in parallel sentence pairs, therby helping to filter out noisy pairs.
+Then, sentence embeddings help to discover parallel sentences in comparable corpora.
+We evaluate our model on multiple language pairs and data resources.
+This model can be used on any parallel corpus without any manual annotation.
 

### 2018-07-25 WMT/divergence.tex
fix introduction

@@
 In the case of neural MT, a large neural network is trained through maximising translation performance on a given parallel corpus. 
-Therefore, the quality of an MT engine is heavily dependent upon the amount and quality of parallel sentences. 
-Note that recent work on neural MT remove completely the need of parallel data following unsupervised methods~\cite{lample2018word,artetxe2018iclr} obtaining accuracy improvements over word-by-word statistical MT systems.
+Therefore, the quality of an MT engine is heavily dependent upon the
+amount and quality of the training parallel sentences.\footnote{Note
+  that some recent work on neural MT
+  \cite{lample2018word,artetxe2018iclr} completely dispenses with the
+  need of parallel data, using unsupervised methods to obtain performance
+  improvements over word-by-word statistical MT systems. However, these systems
+  still lag far behind systems trained in a supervised fashion, as
+  considered in this work.}
 
 Unfortunately, parallel texts are scarce resources. 
-There are relatively few language pairs for which parallel corpora of reasonable sizes are available, and even for those pairs, the corpora come mostly from few domains. 
-To alleviate the lack of parallel data several approaches have been developed in the last years.
-They range from methods using non-parallel, or comparable data ~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,W17-2509,P17-3003} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714} using in all cases automated alignment/translation engines that are prone to the introduction of noise in the resulting parallel sentences.
-Mismatches on parallel sentences extracted from translated texts are also reported ~\cite{tiedemann2011bitext,XU16.310}. 
-This problem is mostly ignored in machine translation, where parallel sentences are considered to convey the exact same meaning, and is particularly important for neural MT engines as suggested by~\cite{chen2016adaptation}.
+There are relatively few language pairs for which parallel corpora of
+reasonable sizes are available, and even for those pairs, the corpora
+are only representative of a small number of domains or genres. 
+To mitigate the lack of parallel data, several approaches have been developed in the last years.
+They range from methods using non-parallel, or comparable data ~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,W17-2509,P17-3003} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714} using in all cases automated alignment/translation engines that are prone to introducing noise in the resulting parallel sentences.
+Mismatches in parallel sentences extracted from translated texts are also reported ~\cite{tiedemann2011bitext,XU16.310}. 
+This problem is mostly ignored in MT, where parallel sentences are considered to convey the exact same meaning, but seems particularly important for neural MT engines, as noted in~\citet{chen2016adaptation}.
 
-Table~\ref{tab:examples} illustrates some examples of English-French parallel sentences which are not completely semantically equivalent.
-Divergencies are outlined using bold letters.
-Examples extracted from the OpenSubtitles corpus ~\cite{LisonTiedemann2016}.
+Table~\ref{tab:examples} gives some examples of English-French
+parallel sentences that are not completely semantically equivalent, 
+extracted from the OpenSubtitles corpus ~\cite{LisonTiedemann2016}. 
+Divergences are outlined using bold letters.
 
@@
 
-Different types of translation divergencies can be found in a parallel corpus:
+Different types of translation divergences exist in a parallel corpus:
 Additional segments are included on either side of the parallel sentences (first and second rows) most likely due to errors in sentence segmentation tools;
 Some translations may be completely uncorrelated (third row);
-Inaccurate translations are also included (fourth row). 
+Inaccurate translations also exist (fourth row). 
 Note that divergent translations can be due to many different reasons ~\cite{C14-1055}, the study of which is beyond the scope of this paper.
 
-In this work, we present an unsupervised method for building cross lingual sentence embeddings in the form of continuous vectors. 
-Embeddings are then used to measure semantic equivalence of their corresponding sentences.
-We evaluate the method on two different tasks:
-First, we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation task.
-Then, we also show competitive results when identifying parallel sentences in comparable corpora using the French-English data made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}.
+In this work, we present an unsupervised method for building cross-lingual sentence embeddings in the form of continuous vectors. 
+These embeddings are then used to measure semantic equivalence of their corresponding sentences.
+We evaluate our method on two different tasks:
+First, we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
+Then, we also report competitive results when identifying parallel sentences in comparable corpora, using the French-English data made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}.
 

### 2018-07-26 WMT/divergence.tex
delete bucc experiments

@@
 amount and quality of the training parallel sentences.\footnote{Note
-  that some recent work on neural MT
+  that  recent work on neural MT
   \cite{lample2018word,artetxe2018iclr} completely dispenses with the
@@
 To mitigate the lack of parallel data, several approaches have been developed in the last years.
-They range from methods using non-parallel, or comparable data ~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,W17-2509,P17-3003} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714} using in all cases automated alignment/translation engines that are prone to introducing noise in the resulting parallel sentences.
+They range from methods using non-parallel, or comparable data ~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,W17-2509,P17-3003,P18-2037} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714} using in all cases automated alignment/translation engines that are prone to introducing noise in the resulting parallel sentences.
 Mismatches in parallel sentences extracted from translated texts are also reported ~\cite{tiedemann2011bitext,XU16.310}. 
@@
 These embeddings are then used to measure semantic equivalence of their corresponding sentences.
-We evaluate our method on two different tasks:
-First, we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
-Then, we also report competitive results when identifying parallel sentences in comparable corpora, using the French-English data made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}.
-
-%In addition, we  show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.
+%We evaluate our method on two different tasks:
+To evaluate our method we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
+%Then, we also report competitive results when identifying parallel sentences in comparable corpora, using the French-English data made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}.
+Then, we  show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.
 
@@
 We describe in detail the core of the neural similarity classifier in Section~\ref{sec:similarity}. 
-We report experiments with the presented model on two different tasks in Section~\ref{experiments}.
+We report experiments with the presented model in Section~\ref{experiments}.
 Section~\ref{sec:results} evaluates results. 
@@
 
-\subsection{Translation Divergencies}
+%\subsection{Translation Divergencies}
 
@@
 Attempts to measure the impact of translation divergencies in MT systems have focused on the introduction of noise in sentence alignments ~\cite{goute2012}, showing that statistical MT systems are highly tolerant to noise, and that performance only degrades seriously at very high noise levels.
-In constrast, neural MT engines seem to be more sensitive~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.
+In contrast, neural MT engines seem to be more sensitive~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.
 
@@
 
-\subsection{Extracting Parallel Sentences}
+%\subsection{Extracting Parallel Sentences}
 
-As previously outlined, parallel texts are scarce resources as well as the only necessary resource to build Machine Translation systems.
+%As previously outlined, parallel texts are scarce resources as well as the only necessary resource to build Machine Translation systems.
 In contrast, there is an increasing amount of comparable corpora available on the Internet, a potential solution to alleviate the data scarcity problem.
 
-Several approaches exist to extract parallel sentences from comparable corpora. 
-In~\cite{J05-4003} is presented a maximum entropy classifier that can reliably determine whether a pair of sentences are translations of each other. 
-A similar approach is proposed by~\cite{E09-1003}, where a SMT system is used to translate the source language side of a comparable corpus to find candidate sentences on the target side. 
-Word error rate and translation error rate measures are used to decide when a candidate is finally considered parallel. 
-More recently,~\cite{W17-2508} present a method based on seed lexical translations, simple set expansion operations and the Jaccard similarity coefficient.
+%Several approaches exist to extract parallel sentences from comparable corpora. 
+%In~\cite{J05-4003} is presented a maximum entropy classifier that can reliably determine whether a pair of sentences are translations of each other. 
+%A similar approach is proposed by~\cite{E09-1003}, where a SMT system is used to translate the source language side of a comparable corpus to find candidate sentences on the target side. 
+%Word error rate and translation error rate measures are used to decide when a candidate is finally considered parallel. 
+%More recently,~\cite{W17-2508} present a method based on seed lexical translations, simple set expansion operations and the Jaccard similarity coefficient.
 The method obtains state-of-the-art results in the task of parallel sentence extraction from comparable corpora.
 
-Closely related to the method proposed in this paper, different works employ continuous vector sentence representations as a preceding step to compute sentence similarity. 
-In~\cite{W15-1521}, word representations from bilingual word embeddings are used to build a convolutional network that predicts wether a sentence pair is aligned or not.
-~\cite{W17-2509} propose a recurrent network to estimate sentence embeddings followed by dense layers that estimates the conditional probability distribution that two sentences are parallel.
+%Closely related to the method proposed in this paper, different works employ continuous vector sentence representations as a preceding step to compute sentence similarity. 
+%In~\cite{W15-1521}, word representations from bilingual word embeddings are used to build a convolutional network that predicts wether a sentence pair is aligned or not.
+%~\cite{W17-2509} propose a recurrent network to estimate sentence embeddings followed by dense layers that estimates the conditional probability distribution that two sentences are parallel.
 
-These approaches are different from ours FINISH
+%These approaches are different from ours FINISH
 
-This strong interest in comparable corpora was further boosted after the apparition of the BUCC workshop series on Building and Using Comparable Corpora\footnote{https://comparable.limsi.fr/bucc2018/} which helped to provide a task definition, datasets and evaluation methods to assess the state of the art.
+%This strong interest in comparable corpora was further boosted after the apparition of the BUCC workshop series on Building and Using Comparable Corpora\footnote{https://comparable.limsi.fr/bucc2018/} which helped to provide a task definition, datasets and evaluation methods to assess the state of the art.
 
@@
 
-For comparison purposes we also implemented the model introduced in \cite{W17-2509}.
+%For comparison purposes we also implemented the model introduced in \cite{W17-2509}.
 The network employs the same initial Bi-LSTM layers outlined in Figure~\ref{network} which encode source and target sentences into $h_{src}$ and $h_{tgt}$ vectors.
@@
 
-\begin{equation}
-\begin{split}
-h^{(1)} =  &\ \ h_{src} * h_{tgt} \\
-h^{(2)} =  &\ \ |h_{src} - h_{tgt}| \\
-h^{(3)} = &\ \ tanh(W[h^{(1)};h^{(2)}] + b^{(3)}) \\
-h_{sim} = &\ \ Wh^{(3)} + b^{sim}
-\end{split}
-\end{equation}
+%\begin{equation}
+%\begin{split}
+%h^{(1)} =  &\ \ h_{src} * h_{tgt} \\
+%h^{(2)} =  &\ \ |h_{src} - h_{tgt}| \\
+%h^{(3)} = &\ \ tanh(W[h^{(1)};h^{(2)}] + b^{(3)}) \\
+%h_{sim} = &\ \ Wh^{(3)} + b^{sim}
+%\end{split}
+%\end{equation}
 
-This approach is different from our model since it uses a single end-to-end model to estimate the conditional probability distribution that two sentences are parallel.
+%This approach is different from our model since it uses a single end-to-end model to estimate the conditional probability distribution that two sentences are parallel.
 
-\begin{figure}[h]
-\center
-    \includegraphics[width=0.8\linewidth]{net_sentence}
-    \caption{Illustration of the model presented in~\cite{W17-2509}.
+%\begin{figure}[h]
+%\center
+%    \includegraphics[width=0.8\linewidth]{net_sentence}
+%    \caption{Illustration of the model presented in~\cite{W17-2509}.
 Sentence embeddings (vectors in dark grey) are computed by the same Bi-LSTM layers shown in Figure~\ref{network}.} %The network is composed of source and target word embedding lookup tables ($LT_s$ and $LT_t$) and two identical subnetworks ($net_s$ and $net_t$) that compute in context representations of source ($s_i$) and target words ($t_j$).
 The similarity function outputs the level of semantic similarity between source and target sentences.}
-    \label{net_sentence}
-\end{figure}
+%    \label{net_sentence}
+%\end{figure}
 
-For this second network the loss function is defined as: 
+%For this second network the loss function is defined as: 
 %\vspace{-5mm}
-\begin{equation}
-\mathcal{L}(src,tgt) = log(1+e^{h_{sim} * \mathcal{Y})}) % where $sign(src,tgt) =1
-\label{loss_semb}
-\end{equation}
-\noindent where $\mathcal{Y}$ is the reference label containing $-1$ when the given sentence pair is non-divergent (parallel) and $+1$ for divergent (unpair) examples.
+%\begin{equation}
+%\mathcal{L}(src,tgt) = log(1+e^{h_{sim} * \mathcal{Y})}) % where $sign(src,tgt) =1
+%\label{loss_semb}
+%\end{equation}
+%\noindent where $\mathcal{Y}$ is the reference label containing $-1$ when the given sentence pair is non-divergent (parallel) and $+1$ for divergent (unpair) examples.
 
-Note that different from the network described in the original work, our implementation does not estimates a probability distribution but FINISH.
+%Note that different from the network described in the original work, our implementation does not estimates a probability distribution but FINISH.
 
@@
 
-The network outlined in Figure~\ref{net_sentence} uses a single label for each sentence pair. 
-Hence, positive examples are labelled $\mathcal{Y}=-1$ while any of the previously described negative examples are labelled $\mathcal{Y}=+1$.
+%The network outlined in Figure~\ref{net_sentence} uses a single label for each sentence pair. 
+%Hence, positive examples are labelled $\mathcal{Y}=-1$ while any of the previously described negative examples are labelled $\mathcal{Y}=+1$.
 
@@
 
-To evaluate the ability of our sentence embeddings to identify parallel sentences in comparable corpora we followed the data conditions made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}.
-For training we used the French-English European Parliament corpus from WMT15\footnote{http://www.statmt.org/wmt15/translation-task.html}.
-For testing we used the test set made available by the task organisers, consisting of about $300,000$ monolingual sentences per language, French and English, that contains near $10,000$ parallel sentences.
-Test data come from Wikipedia\footnote{http://ftp.acc.umu.se/mirror/wikimedia.org/dumps} and News Commentary\footnote{http://www.casmacat.eu/corpus/news-commentary.html}.
+%To evaluate the ability of our sentence embeddings to identify parallel sentences in comparable corpora we followed the data conditions made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}.
+%For training we used the French-English European Parliament corpus from WMT15\footnote{http://www.statmt.org/wmt15/translation-task.html}.
+%For testing we used the test set made available by the task organisers, consisting of about $300,000$ monolingual sentences per language, French and English, that contains near $10,000$ parallel sentences.
+%Test data come from Wikipedia\footnote{http://ftp.acc.umu.se/mirror/wikimedia.org/dumps} and News Commentary\footnote{http://www.casmacat.eu/corpus/news-commentary.html}.
 

### 2018-07-26 WMT/divergence.tex
section 2

--- 2018-07-25 WMT/divergence.tex
@@
 amount and quality of the training parallel sentences.\footnote{Note
-  that some recent work on neural MT
+  that recent works on neural MT
   \cite{lample2018word,artetxe2018iclr} completely dispenses with the
@@
   improvements over word-by-word statistical MT systems. However, these systems
-  still lag far behind systems trained in a supervised fashion, as
-  considered in this work.}
+  lag far behind systems trained in a supervised fashion, as
+  considered here.}
 
@@
 
-Different types of translation divergences exist in a parallel corpus:
-Additional segments are included on either side of the parallel sentences (first and second rows) most likely due to errors in sentence segmentation tools;
+Different types of translation divergences exist in a parallel corpus.
+Additional segments are included on either side of the parallel
+sentences (first and second rows in Table~\ref{tab:examples}), most likely due to errors in sentence segmentation tools;
 Some translations may be completely uncorrelated (third row);
 Inaccurate translations also exist (fourth row). 
-Note that divergent translations can be due to many different reasons ~\cite{C14-1055}, the study of which is beyond the scope of this paper.
+Note that divergent translations can be due to multiple reasons~\cite{C14-1055}, the study of which is beyond the scope of this paper.
 
@@
 These embeddings are then used to measure semantic equivalence of their corresponding sentences.
-We evaluate our method on two different tasks:
-First, we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
-Then, we also report competitive results when identifying parallel sentences in comparable corpora, using the French-English data made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}.
+We evaluate our method on two different tasks.
+We first show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
+Then, we report competitive results when identifying parallel sentences in comparable corpora, using the French-English data made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}.
 
@@
 
-The remainder of this paper is structured as follows. 
-Section~\ref{related} overviews related work. 
-We describe in detail the core of the neural similarity classifier in Section~\ref{sec:similarity}. 
-We report experiments with the presented model on two different tasks in Section~\ref{experiments}.
-Section~\ref{sec:results} evaluates results. 
+The remainder of this paper is structured as follows:
+Section~\ref{related} overviews related work;
+we describe in detail the core of the neural similarity classifier in Section~\ref{sec:similarity}. 
+We report experiments with the presented model on two different tasks
+in Section~\ref{experiments}, and further discuss our results in Section~\ref{sec:results}. 
 Finally, conclusions are drawn in Section~\ref{conclusions} and further work is outlined in Section~\ref{further}.
@@
 
-\subsection{Translation Divergencies}
+\subsection{Translation Divergences}
 
-Prior work on translation divergencies focused on reflecting that translation of one language into another results in very different forms ~\cite{J94-4004} according to several categories (e.g.
+Prior work on translation divergences focused on reflecting that translation of one language into another results in very different forms ~\cite{J94-4004} according to several categories (e.g.
 lexical, structural, categorical).
 
-Attempts to measure the impact of translation divergencies in MT systems have focused on the introduction of noise in sentence alignments ~\cite{goute2012}, showing that statistical MT systems are highly tolerant to noise, and that performance only degrades seriously at very high noise levels.
-In constrast, neural MT engines seem to be more sensitive~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.
+Attempts to measure the impact of translation divergencies in MT
+systems are based on the introduction of noise in sentence
+alignments~\cite{goute2012}, showing that statistical MT systems are
+highly tolerant to noise, and that performance only degrades seriously
+at very high noise levels. In constrast, neural MT engines seem to be
+more sensitive to noise~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.
 
-Efforts have been devoted to characterising the degree of semantic equivalence between two snippets of text in the same or different languages~\cite{conf/semeval/AgirreBCDGMRW16}, a workshop devoted to an objective similar to our work.
+Efforts have also been devoted to characterising the degree of
+semantic equivalence between two snippets of text in the same or in
+different languages (see notably~\cite{conf/semeval/AgirreBCDGMRW16}, a workshop devoted to an objective similar to our work).
 In~\cite{Mueller:2016:SRA:3016100.3016291}, a monolingual sentence similarity network is proposed, making use of a simple LSTM layer to compute sentence representations.
 The authors show that a simple SVM classifier can be built on top of the sentence representations to achieve state-of-the-art results in a semantic entailment classification task.
-With the same objective, the system presented in~\cite{N16-1108} uses multiple convolutional layers and models pairwise word interactions.
+Aiming similar goals, the system presented in~\cite{N16-1108} uses multiple convolutional layers and models pairwise word interactions.
 %Our work differs from the previous as we build a different network and use it to measure similarity of sentences in different languages.
 
-Our work is inspired by~\cite{W17-3209} where the authors train a cross-lingual divergence detector using word alignments and sentence length features to train a linear SVM classifier.
-Their work shows that an NMT system trained only on non-divergent sentences yields slightly higher translation quality scores and requires clearly less training time.
-The same authors have recently updated their work in~\cite{DBLP:journals/corr/abs-1803-11112}.
-The objective is the same as the neural network presented in~\cite{N16-1108} and their network further outperforms their previous work.
-Our work differs from the previous as we make use of a network with different topology.
-We model sentence similarity by means of optimising word alignment and in addition, we propose a simple method that employs alignment scores to not only detect but correct divergent sentences.
+Our work is mostly inspired by~\citet{W17-3209}, who train a
+cross-lingual divergence detector using word alignments and sentence
+length features to train a linear SVM classifier. Their work shows
+that an NMT system trained only on non-divergent sentences yields
+slightly higher translation quality scores and also requires less
+training time. The same authors have recently updated their work
+in~\cite{DBLP:journals/corr/abs-1803-11112}. Using an objective
+similar to the neural network presented in~\cite{N16-1108}, they were
+able to improve their previous results. Our work differs from these
+prior art in several ways, as we make use of a network with different
+topology and we model sentence similarity by means of optimising word
+alignment. In addition, we propose a simple method that employs
+alignment scores to not only detect but also correct divergent sentences.
 
@@
 
-As previously outlined, parallel texts are scarce resources as well as the only necessary resource to build Machine Translation systems.
-In contrast, there is an increasing amount of comparable corpora available on the Internet, a potential solution to alleviate the data scarcity problem.
+As previously outlined, Machine Translation systems depend on the
+availability of parallel texts, which are scarce resources.
+In contrast, there is an vast amount of comparable corpora available on the Internet, a potential solution to mitigate the data scarcity problem.
 
-Several approaches exist to extract parallel sentences from comparable corpora. 
-In~\cite{J05-4003} is presented a maximum entropy classifier that can reliably determine whether a pair of sentences are translations of each other. 
-A similar approach is proposed by~\cite{E09-1003}, where a SMT system is used to translate the source language side of a comparable corpus to find candidate sentences on the target side. 
-Word error rate and translation error rate measures are used to decide when a candidate is finally considered parallel. 
-More recently,~\cite{W17-2508} present a method based on seed lexical translations, simple set expansion operations and the Jaccard similarity coefficient.
-The method obtains state-of-the-art results in the task of parallel sentence extraction from comparable corpora.
+Several approaches have been proposed to extract parallel sentences from comparable corpora. 
+In~\cite{J05-4003} is presented a maximum entropy classifier that can
+reliably assess whether a pair of sentences are mutual translations. 
+A similar approach is proposed by~\cite{E09-1003}, where a SMT system
+is used to translate the source language side of a comparable corpus
+to find candidate sentences on the target side. Word error rate and
+translation error rate measures are used to decide whether a candidate
+pair is finally deemed parallel. 
+More recently,~\cite{W17-2508} present a method based on seed lexical
+translations, simple set expansion operations and the Jaccard
+similarity coefficient. The method obtains state-of-the-art results in
+the task of mining parallel sentence extraction from comparable corpora.
 
-Closely related to the method proposed in this paper, different works employ continuous vector sentence representations as a preceding step to compute sentence similarity. 
-In~\cite{W15-1521}, word representations from bilingual word embeddings are used to build a convolutional network that predicts wether a sentence pair is aligned or not.
-~\cite{W17-2509} propose a recurrent network to estimate sentence embeddings followed by dense layers that estimates the conditional probability distribution that two sentences are parallel.
+Several works employ continuous vector sentence representations as a
+preceding step to compute sentence similarity, an approach we also
+adopt in this paper. In~\cite{W15-1521}, word representations from
+bilingual word embeddings are used to build a convolutional network
+that predicts wether a sentence pair is parallel or not.
+\citet{W17-2509} propose a recurrent network to estimate sentence embeddings followed by dense layers that estimates the conditional probability distribution that two sentences are parallel.
 
-These approaches are different from ours FINISH
+% These approaches are different from ours FINISH
 
-This strong interest in comparable corpora was further boosted after the apparition of the BUCC workshop series on Building and Using Comparable Corpora\footnote{https://comparable.limsi.fr/bucc2018/} which helped to provide a task definition, datasets and evaluation methods to assess the state of the art.
+This strong interest in comparable corpora was further boosted after
+the apparition of the BUCC \footnote{Building and Using Comparable
+  Corpora  - https://comparable.limsi.fr/bucc2018/} evaluation
+campaign which helped to provide a task definition, datasets and
+evaluation methods to improve the state-of-the-art.
 
@@
 
-In this section we describe the architecture of our model, very much inspired by the work on Word Alignment in~\cite{W16-2207}. Figure~\ref{network} illustrates the network.
+In this section, we describe the architecture of our model, drawing
+inspiration from the work of~\citet{W16-2207} on word alignment.
 
-In the following, we consider a source-target sentence pair $(s,t)$ with $s=(s_1,...,s_I)$ and $t=(t_1,...,t_J)$.
-The model is composed of 2 Bi-directional LSTM subnetworks, $net_s$ and $net_t$, which respectively encode source and target sentences.
-Since both $net_s$ and $net_t$ take the same form we describe only the source architecture.
+In the following, we consider a source-target sentence pair $(s,t)$
+with $s=(s_1,...,s_I)$ and $t=(t_1,...,t_J)$. The model is composed of
+2 Bi-directional LSTM subnetworks, $net_s$ and $net_t$, which
+respectively encode source and target sentences (see Figure~\ref{network}). Since both $net_s$ and $net_t$ take the same form, we describe only the source architecture.
 
-The source-sentence Bi-LSTM network outputs forward and backward hidden states, $\overrightarrow{h}^{src}_i$ and $\overleftarrow{h}^{src}_i$, which are then concatenated into a single vector encoding the $i^{th}$ word of the source sentence,
+The source-sentence Bi-LSTM network outputs forward and backward hidden states, $\overrightarrow{h}^{src}_i$ and $\overleftarrow{h}^{src}_i$, which are concatenated into a single vector encoding the $i^{th}$ word in the source sentence,
 $h^{src}_i = [ \overrightarrow{h}^{src}_i ; \overleftarrow{h}^{src}_i ]$.
@@
 
-In addition, the last forward/backward hidden states (outlined using dark grey in Figure~\ref{network}) are also concatenated into a single vector  to represent whole sentences 
+In addition, the last forward/backward hidden states (outlined in dark grey in Figure~\ref{network}) are also concatenated into a single vector to represent whole sentences 
 $h_{src} = [ \overrightarrow{h}^{src}_I ; \overleftarrow{h}^{src}_1 ]$.
 %$h_{src} = \left( \begin{array}{cc} \overrightarrow{h}^{src}_I \\ \overleftarrow{h}^{src}_1 \\ \end{array} \right)$.
-At this point a measure of similarity between sentence pairs can be obtained by cosine similarity:
+At this point, a measure of similarity between sentence pairs can be
+obtained by by computing the cosine between these representations:
 
 \begin{equation}
-    sim(h_{src}, h_{tgt}) = \frac{h_{src} \cdotp h_{tgt}}{||h_{src}|| * ||h_{tgt}||}
+    sim(h_{src}, h_{tgt}) = \frac{h_{src} \cdotp h_{tgt}}{||h_{src}|| * ||h_{tgt}||},
     \label{cosine}

### 2018-07-26 WMT/divergence.tex

--- 2018-07-26 WMT/divergence.tex
@@
 amount and quality of the training parallel sentences.\footnote{Note
-  that  recent work on neural MT
+  that recent works on neural MT
   \cite{lample2018word,artetxe2018iclr} completely dispenses with the
@@
 These embeddings are then used to measure semantic equivalence of their corresponding sentences.
+
 %We evaluate our method on two different tasks:
-To evaluate our method we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
+To evaluate the benefits of our method, we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
 %Then, we also report competitive results when identifying parallel sentences in comparable corpora, using the French-English data made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}.
-Then, we  show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.
+We also show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.
 
@@
 Section~\ref{sec:results} evaluates results.
+
 Finally, conclusions are drawn in Section~\ref{conclusions} and further work is outlined in Section~\ref{further}.
@@
 
-%\subsection{Translation Divergencies}
+\subsection{Translation Divergences}
 
-Prior work on translation divergencies focused on reflecting that translation of one language into another results in very different forms ~\cite{J94-4004} according to several categories (e.g.
+Prior work on translation divergences focused on reflecting that translation of one language into another results in very different forms ~\cite{J94-4004} according to several categories (e.g.
 lexical, structural, categorical).
 
-Attempts to measure the impact of translation divergencies in MT systems have focused on the introduction of noise in sentence alignments ~\cite{goute2012}, showing that statistical MT systems are highly tolerant to noise, and that performance only degrades seriously at very high noise levels.
-In contrast, neural MT engines seem to be more sensitive~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.
+Attempts to measure the impact of translation divergencies in MT
+systems are based on the introduction of noise in sentence
+alignments~\cite{goute2012}, showing that statistical MT systems are
+highly tolerant to noise, and that performance only degrades seriously
+at very high noise levels. In constrast, neural MT engines seem to be
+more sensitive to noise~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.
 
-Efforts have been devoted to characterising the degree of semantic equivalence between two snippets of text in the same or different languages~\cite{conf/semeval/AgirreBCDGMRW16}, a workshop devoted to an objective similar to our work.
+Efforts have also been devoted to characterising the degree of
+semantic equivalence between two snippets of text in the same or in
+different languages (see notably~\cite{conf/semeval/AgirreBCDGMRW16}, a workshop devoted to an objective similar to our work).
 In~\cite{Mueller:2016:SRA:3016100.3016291}, a monolingual sentence similarity network is proposed, making use of a simple LSTM layer to compute sentence representations.
 The authors show that a simple SVM classifier can be built on top of the sentence representations to achieve state-of-the-art results in a semantic entailment classification task.
-With the same objective, the system presented in~\cite{N16-1108} uses multiple convolutional layers and models pairwise word interactions.
+Aiming similar goals, the system presented in~\cite{N16-1108} uses multiple convolutional layers and models pairwise word interactions.
 %Our work differs from the previous as we build a different network and use it to measure similarity of sentences in different languages.
 
-Our work is inspired by~\cite{W17-3209} where the authors train a cross-lingual divergence detector using word alignments and sentence length features to train a linear SVM classifier. 
-Their work shows that an NMT system trained only on non-divergent sentences yields slightly higher translation quality scores and requires clearly less training time. 
-The same authors have recently updated their work in~\cite{DBLP:journals/corr/abs-1803-11112}. 
-The objective is the same as the neural network presented in~\cite{N16-1108} and their network further outperforms their previous work. 
-Our work differs from the previous as we make use of a network with different topology. 
-We model sentence similarity by means of optimising word alignment and in addition, we propose a simple method that employs alignment scores to not only detect but correct divergent sentences.
+Our work is mostly inspired by~\citet{W17-3209}, who train a
+cross-lingual divergence detector using word alignments and sentence
+length features to train a linear SVM classifier. Their work shows
+that an NMT system trained only on non-divergent sentences yields
+slightly higher translation quality scores and also requires less
+training time. The same authors have recently updated their work
+in~\cite{DBLP:journals/corr/abs-1803-11112}. Using an objective
+similar to the neural network presented in~\cite{N16-1108}, they were
+able to improve their previous results. Our work differs from these
+prior art in several ways, as we make use of a network with different
+topology and we model sentence similarity by means of optimising word
+alignment. In addition, we propose a simple method that employs
+alignment scores to not only detect but also correct divergent sentences.
 
-%\subsection{Extracting Parallel Sentences}
+% As previously outlined, Machine Translation systems depend on the
+% availability of parallel texts, which are scarce resources.
+In contrast, there is an vast amount of comparable corpora available on the Internet, a potential solution to mitigate the data scarcity problem.
 
-%As previously outlined, parallel texts are scarce resources as well as the only necessary resource to build Machine Translation systems.
-In contrast, there is an increasing amount of comparable corpora available on the Internet, a potential solution to alleviate the data scarcity problem.
+% Several approaches have been proposed to extract parallel sentences from comparable corpora. 
+% In~\cite{J05-4003} is presented a maximum entropy classifier that can
+% reliably assess whether a pair of sentences are mutual translations. 
+% A similar approach is proposed by~\cite{E09-1003}, where a SMT system
+% is used to translate the source language side of a comparable corpus
+% to find candidate sentences on the target side. Word error rate and
+% translation error rate measures are used to decide whether a candidate
+% pair is finally deemed parallel. 
+% More recently,~\cite{W17-2508} present a method based on seed lexical
+% translations, simple set expansion operations and the Jaccard
+% similarity coefficient. The method obtains state-of-the-art results in
+% the task of mining parallel sentence extraction from comparable corpora.
 
-%Several approaches exist to extract parallel sentences from comparable corpora. 
-%In~\cite{J05-4003} is presented a maximum entropy classifier that can reliably determine whether a pair of sentences are translations of each other. 
-%A similar approach is proposed by~\cite{E09-1003}, where a SMT system is used to translate the source language side of a comparable corpus to find candidate sentences on the target side. 
-%Word error rate and translation error rate measures are used to decide when a candidate is finally considered parallel. 
-%More recently,~\cite{W17-2508} present a method based on seed lexical translations, simple set expansion operations and the Jaccard similarity coefficient.
-The method obtains state-of-the-art results in the task of parallel sentence extraction from comparable corpora.
+% Several works employ continuous vector sentence representations as a
+% preceding step to compute sentence similarity, an approach we also
+% adopt in this paper. In~\cite{W15-1521}, word representations from
+% bilingual word embeddings are used to build a convolutional network
+% that predicts wether a sentence pair is parallel or not.
+\citet{W17-2509} propose a recurrent network to estimate sentence embeddings followed by dense layers that estimates the conditional probability distribution that two sentences are parallel.
 
-%Closely related to the method proposed in this paper, different works employ continuous vector sentence representations as a preceding step to compute sentence similarity. 
-%In~\cite{W15-1521}, word representations from bilingual word embeddings are used to build a convolutional network that predicts wether a sentence pair is aligned or not.
-%~\cite{W17-2509} propose a recurrent network to estimate sentence embeddings followed by dense layers that estimates the conditional probability distribution that two sentences are parallel.
+% % These approaches are different from ours FINISH
 
-%These approaches are different from ours FINISH
-
-%This strong interest in comparable corpora was further boosted after the apparition of the BUCC workshop series on Building and Using Comparable Corpora\footnote{https://comparable.limsi.fr/bucc2018/} which helped to provide a task definition, datasets and evaluation methods to assess the state of the art.
+% This strong interest in comparable corpora was further boosted after
+% the apparition of the BUCC \footnote{Building and Using Comparable
+%   Corpora  - https://comparable.limsi.fr/bucc2018/} evaluation
+% campaign which helped to provide a task definition, datasets and
+% evaluation methods to improve the state-of-the-art.
 

### 2018-07-26 EMNLP/divergence.tex

@@
 
-We {\bf (R)}eplace random sequences of words by a sequence of words with same part-of-speeches. 
+We {\bf (R)}eplace random sequences of words by a sequence of words with same parts-of-speech. 
 The rationale behind this method is to keep the new sentences as grammatical as possible. 
@@
 In addition to the basic tokenisation detailed in Section~\ref{divergence} we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30,000$ merge operations learned from both English and French data.
-Neural MT systems are based on the open-source project \texttt{OpenNMT}\footnote{http://opennmt.net}. We use a bidirectional RNN encoder with $4$ LSTM layers with each containing $1,000$ cells. 
-Word embeddings have a size of $300$ cells.
-We set the dropout probability to $0.3$.
-Batch size is set to $64$.
-The maximum length of both source and target sentences is set to $80$.
-% and we limit the vocabulary size to $50K$ words for both source and target languages.
-The default optimiser is SGD with the starting learning rate of $1.0$.
-We start to decay the learning rate from epoch $10$ or when we detect increasing perplexity as compared to the previous epoch on a validation set.
-We stop training after $20$ epochs.
+Neural MT systems are based on the open-source project \texttt{OpenNMT}\footnote{http://opennmt.net}.
+We use a Transformer model with same configuration in paper \cite{Transformer2017}, i.e, both encoder and decoder have 6 layers; Multi-head attention is performed over 8 heads.
+Hidden layer's size is 512.
+The inner layer of feedforward network is of size $2048$.
+Word embeddings have a size of $512$ cells.
+We set the dropout probability to $0.1$.
+Batch size is set to $3072$.
+The maximum length of both source and target sentences is set to $80$ and we limit the vocabulary size to $50K$ words for both source and target languages.
+
+The optimiser is Adam avec learning rate that was introduced in paper \cite{Transformer2017} with $\beta_1 = 0.9$, $\beta_2 = 0.98$, $\eta = 10^{-9}$, $warmup\_steps = 4000$.
+We stop training after 30 epochs.
 
@@
 %\texttt{wemb}           & 21.5 & - & 42.56 \\
-\texttt{wemb+fix}   & 15.5 & 2.5 & \\
+\texttt{wemb+fix}   & 15.5 & 2.5 & 44.19 \\
 \texttt{wemb}           & 18.0 & - & 43.19  ($+1.01$)\\
@@
 \texttt{all}                  & 22.2 & - & 19.27 \\ 
-\texttt{semb}            & 15.0 & - & \\
+\texttt{semb}            & 15.0 & - & 20.23\\
 \texttt{wemb}           & 15.0 & - & 21.52 ($+2.25$)\\
-%\texttt{wemb}           & 12.5 &   \\
-%\texttt{wemb+fix}   & 12.5+2.5 & \\
+\texttt{wemb}           & 17.5 & - & 21.97   \\
+\texttt{wemb+fix}   & 15 & 2.5 & 22.42\\
 \hline

### 2018-07-26 WMT/divergence.tex

--- 2018-07-26 WMT/divergence.tex
@@
 
-\title{Crosslingual Sentence Embeddings for Neural Machine Translation}
-%\title{A Recurrent Network for Parallel Corpora Divergence Analysis}
+%\title{Crosslingual Sentence Embeddings for Neural Machine Translation}
+\title{A Recurrent Network for Parallel Corpora Similarity Analysis}
 
@@
 In the case of neural machine translation, a large neural network is trained to maximise the translation performance on a given parallel corpus. 
-%This paper describes an unsupervised method for building cross lingual sentence embeddings, which can then be used to compute semantic similarity between sentences in different languages.
-This paper describes an unsupervised method for building cross-lingual
-sentence embeddings that optimise word alignments. Such embeddings can then be used to compute semantic similarity between sentences in different languages.
-We report competitive performance on two different tasks. 
-Sentence embeddings are first used to detect translation divergences
-in parallel sentence pairs, therby helping to filter out noisy pairs.
-Then, sentence embeddings help to discover parallel sentences in comparable corpora.
+This paper describes an unsupervised method for building cross-lingual sentence embeddings that optimise word alignments. 
+Such embeddings can then be used to compute semantic similarity between sentences in different languages.
+We first use sentence embeddings to detect translation divergences in parallel sentence pairs, thereby helping to filter out noisy pairs.
+Then, word similarity scores predicted by the network are used to identify and fix some divergences guiding to collect additional parallel segments.
 We evaluate our model on multiple language pairs and data resources.
@@
 In the case of neural MT, a large neural network is trained through maximising translation performance on a given parallel corpus. 
-Therefore, the quality of an MT engine is heavily dependent upon the
-amount and quality of the training parallel sentences.\footnote{Note
-  that  recent work on neural MT
-  \cite{lample2018word,artetxe2018iclr} completely dispenses with the
-  need of parallel data, using unsupervised methods to obtain performance
-  improvements over word-by-word statistical MT systems. However, these systems
-  still lag far behind systems trained in a supervised fashion, as
-  considered in this work.}
+Therefore, the quality of an MT engine is heavily dependent upon the amount and quality of the training parallel sentences.\footnote{Note that  recent work on neural MT~\cite{lample2018word,artetxe2018iclr} completely dispenses with the need of parallel data, using unsupervised methods to obtain performance improvements over word-by-word statistical MT systems.
+However, these systems still lag far behind systems trained in a supervised fashion, as considered in this work.}
 
@@
 These embeddings are then used to measure semantic equivalence of their corresponding sentences.
-%We evaluate our method on two different tasks:
 To evaluate our method we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
-%Then, we also report competitive results when identifying parallel sentences in comparable corpora, using the French-English data made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}.
 Then, we  show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.
@@
 
-%\subsection{Translation Divergencies}
-
 Prior work on translation divergencies focused on reflecting that translation of one language into another results in very different forms ~\cite{J94-4004} according to several categories (e.g.
@@
 With the same objective, the system presented in~\cite{N16-1108} uses multiple convolutional layers and models pairwise word interactions. 
-%Our work differs from the previous as we build a different network and use it to measure similarity of sentences in different languages.
-
+Recently, in~\cite{P18-2037} is presented a network able to learn a multilingual sentence embedding space that is successfully used to filter noisy parallel data and to mine for bi-texts in large monolingual collections.
 Our work is inspired by~\cite{W17-3209} where the authors train a cross-lingual divergence detector using word alignments and sentence length features to train a linear SVM classifier. 
@@
 The objective is the same as the neural network presented in~\cite{N16-1108} and their network further outperforms their previous work.
+
 Our work differs from the previous as we make use of a network with different topology. 
@@
 
-%\subsection{Extracting Parallel Sentences}
-
 %As previously outlined, parallel texts are scarce resources as well as the only necessary resource to build Machine Translation systems.
 In contrast, there is an increasing amount of comparable corpora available on the Internet, a potential solution to alleviate the data scarcity problem.
-
 %Several approaches exist to extract parallel sentences from comparable corpora. 
@@
 The method obtains state-of-the-art results in the task of parallel sentence extraction from comparable corpora.
-
 %Closely related to the method proposed in this paper, different works employ continuous vector sentence representations as a preceding step to compute sentence similarity. 
@@
 %~\cite{W17-2509} propose a recurrent network to estimate sentence embeddings followed by dense layers that estimates the conditional probability distribution that two sentences are parallel.
-
 %These approaches are different from ours FINISH
-
 %This strong interest in comparable corpora was further boosted after the apparition of the BUCC workshop series on Building and Using Comparable Corpora\footnote{https://comparable.limsi.fr/bucc2018/} which helped to provide a task definition, datasets and evaluation methods to assess the state of the art.
@@
 The aggregation function combines the scores of all source (or target) words for a particular target (or source) word and promotes source words which are likely to be aligned with a given target word according to the knowledge the model has learned so far.
-
 Alignment scores $S(i,j)$ are given by the dot-product $S(i,j) = h_i^{src} \cdotp h_j^{tgt}$, while aggregation functions are defined as:
@@
 \end{equation}
-\noindent where $\mathcal{Y}_i^{src}$ (or $\mathcal{Y}_j^{tgt}$) is a vector with reference labels containing $-1$ when the $i^{th}$ source word (or $j^{th}$ target word) is present in the translated target (or source) sentence, and $+1$ for divergent (unpaired) words.
+\noindent where $\mathcal{Y}_i^{src}$ and $\mathcal{Y}_j^{tgt}$ are vectors with reference labels containing $-1$ when the word is present in the translated sentence, and $+1$ for divergent (unpaired) words.
 
@@
 Training is performed by minimising Equation~\ref{loss_wemb}, for which examples with annotations for source $\mathcal{Y}_i^{src}$ and target $\mathcal{Y}_j^{tgt}$ words are needed.
+Different kind of examples are used to train our network:
+\begin{itemize}
 
-As positive examples we use {\bf paired} sentences of a parallel corpus. All words of the pair of sentences are labelled as parallel, $\mathcal{Y}_i^{src}=-1$ and $\mathcal{Y}_j^{tgt}=-1$. 
-As negative examples we use random {\bf unpaired} sentences. In this case, all words are labelled as divergent, $\mathcal{Y}_i^{src}=+1$ and $\mathcal{Y}_j^{tgt}=+1$. 
+\item As positive examples we use paired sentences of the parallel corpus.
+In this case, all words of the pair of sentences are labelled as parallel, $\mathcal{Y}_i^{src}=-1$ and $\mathcal{Y}_j^{tgt}=-1$.
+
+\item As negative examples we use random unpaired sentences. In this case, all words are labelled as divergent, $\mathcal{Y}_i^{src}=+1$ and $\mathcal{Y}_j^{tgt}=+1$. 
 Since negative examples may be very easy to classify and we want our network to detect less obvious divergencies, we further create negative examples following the next two methods:
 
-We {\bf replace} random sequences of words on either side of the sentence pair by a sequence of words with the same part-of-speeches. 
+\item We replace random sequences of words on either side of the sentence pair by a sequence of words with the same part-of-speeches. 
 The rationale behind this method is to keep the new sentences as grammatical as possible. 
-Otherwise the network can learn to detect non-grammatical sentences to predict divergence.
+Otherwise, to predict divergence the network can learn to detect non-grammatical sentences.
 Words that are not replaced are considered parallel ($-1$) while those replaced are assigned the divergent label ($+1$). 
@@
 \begin{table}[h]
+\center
 \begin{tabular}{ll}
@@
 $\mathcal{Y}^{src}$: & { \small \texttt{-1 \ \  -1 {\bf +1\ \ +1} \ \  -1}} \\
-tgt: & { \small \texttt{Que {\bf voulez-vous} ?}} \\
-$\mathcal{Y}^{tgt}$: & { \small \texttt{-1\ \ {\bf +1}\ \ \ \ \ \ \ -1}} \\
+tgt: & { \small \texttt{Que {\bf ressentez-vous} ?}} \\
+$\mathcal{Y}^{tgt}$: & { \small \texttt{-1\ \ {\bf +1}\ \ \ \ \ \ \ \ \ -1}} \\
 \end{tabular}
@@
 Note that the new sentences are not assured to be grammatical after replacing sequences with the same part-of-speech.
-%Note that part-of-speeches and word alignments are needed to build this negative example. 
-We use word alignments to identify as divergent the sequence '\texttt{ressentez-vous}' since it is aligned to the replaced sequence '\texttt{you feel}'.
+We use word alignments to identify as divergent the sequence '\texttt{ressentez-vous}' since it is aligned to '\texttt{you feel}' in the original sentence pair.
 
-Finally, motivated by sentence segmentation errors observed in many corpora, we also build negative examples by {\bf inserting} a second sentence at the beginning (or end) of the source (or target) sentence pair.
+\item Finally, motivated by sentence segmentation errors observed in many corpora, we also build negative examples by inserting a second sentence at the beginning (or end) of the source (or target) sentence pair.
 Words in the original sentence pair are assigned the parallel label ($-1$) while the new words inserted are considered divergent ($+1$).
@@
 
+\end{itemize}
+
 In order to avoid that negative examples are easily predicted just by looking at the difference in length of training sentences we constraint all negative examples to have a difference in length not exceeding $2.0$.
-Very short sentences, up to $4$ words, are accepted if the length ratio does not exceeds $3.0$.
-
-%The network outlined in Figure~\ref{net_sentence} uses a single label for each sentence pair. 
-%Hence, positive examples are labelled $\mathcal{Y}=-1$ while any of the previously described negative examples are labelled $\mathcal{Y}=+1$.
+Very short sentences, of up to $4$ words, are accepted if the length ratio does not exceeds $3.0$.
 
@@
 
-We observed in our training corpus that many divergent sentences follow a common pattern, consisting of containing some extra leading/trailing words.
-See the first and second examples of table~\ref{tab:examples}.
-Accordingly, we implemented an algorithm that discards sequences of leading/trailing words making use of alignment scores $S(i,j)$. 
-Hence considering as parallel $s_[^]$ and $t_\langle^\rangle$.
-To find the optimal source ($[, ]$) and target ($\langle, \rangle$) indexes that enclose the corrected segment within the original sentence, we implement:
-%\vspace{-4mm}
+We observed in our training corpora that many divergent sentences follow a common pattern, consisting of adding some extra leading/trailing words.
+As in the first and second examples of table~\ref{tab:examples}.
+Accordingly, we implemented a simple algorithm that discards sequences of leading/trailing words on both sides of the source and/or target sentences. 
+Hence considering as parallel $s_u^v$ and $t_x^y$. 
+To find the optimal source ($u, v$) and target ($x, y$) indices that enclose the parallel segments within the original sentence, we implement:
 \begin{equation*}
-\begin{split}
-%\underset{[, ], \{, \}}{\arg\max}&\{\mathcal{P}([,],\{,\}) - \mathcal{Q}([,],\{,\})\} \\
-\underset{[, ], \langle, \rangle}{\arg\max}& \Big \{      \underset{[ \le i \le ]}{\sum} \underset{\langle \le j \le \rangle}{\max} \{ S(i,j) \}  \Big \} % -   \underset{[ > i, i > ]}{\sum} \underset{\langle > j, j > \rangle}{\max} \{ s(i,j) \}   \Big \}
-\end{split}
+\underset{u, v, x, y}{\arg\max} \Big \{      \underset{u \le I \le v}{\sum} \underset{x \le j \le y}{\max} \{ S(i,j) \}    \Big \}
 \end{equation*}
-%\noindent where $\mathcal{P}$ and $\mathcal{Q}$ are defined:
-%\vspace{-4mm}
-%\begin{equation*}
-%\begin{split}
-%\mathcal{P}&([,],\{,\}) = \underset{[ \le i \le ]}{\sum} \underset{\{ \le j \le \}}{\max} \{ s(i,j) \} \\%+ \underset{\{ \le j \le \}}{\sum} \underset{[ \le i \le ]}{\max} \{ s(i,j) \}  \} \\
-%\mathcal{Q}&([,],\{,\}) = \underset{[ > i, i > ]}{\sum} \underset{\{ > j, j > \}}{\max} \{ s(i,j) \} %+ \underset{\{ > j < \}}{\sum} \underset{[ > i < ]}{\max} \{ s(i,j) \}  \} \\
-%\end{split}
-%\end{equation*}
 
-The $\mathcal{N}$-best sequences following the previous function ($s_[^]$, $t_\langle^\rangle$) are considered as valid corrections, but only the highest ranked according to their similarity score is used as replacement for the original $(s_1^I, t_1^J)$.
-% if it has a higher similarity score: $sim(s_[^], t_\{^\}) > sim(s_1^I, t_1^J)$.
-Short sentences are not considered. This is, $] - [ > \tau$ and $\langle - \rangle > \tau$.
-
-Figure~\ref{matrix} (left) shows an example of an alignment matrix $s(i,j)$ for a given sentence pair.
-A possible correction is: \texttt{What do you feel ?
-$\Leftrightarrow$ Que ressentez-vous ?}.
-Hence, with indexes $[=1$, $]=5$, $\langle=1$ and $\rangle=3$.
-
-%\begin{figure*}
-%\centering
-%\begin{multicols}{2}
-%    \includegraphics[width=1.1\linewidth]{learn}    
-%    \includegraphics[width=1.1\linewidth]{summer}    
-%    \includegraphics[width=1.1\linewidth]{swim} 
-%    \includegraphics[width=1.4\linewidth]{awesome}
-%\end{multicols}
-%\begin{multicols}{3}
-%    \includegraphics[width=\linewidth]{sandwich}
-%    \includegraphics[width=\linewidth]{sandwich}
-%    \includegraphics[width=\linewidth]{peu}
-%\end{multicols}
-%\caption{Examples of neural alignment output by the matrix $S_{i,j}$ in Figure~\ref{network}.
-Only positive scores, indicating alignment between words $s_i$ and $t_j$, are printed.
-Different levels of grey are used to highlight the alignment.}
-%\end{figure*}
+The $\mathcal{N}$-best sequences following the previous function ($s_u^v$, $t_x^y$) are considered as valid corrections, but only the highest ranked according to its similarity score is used as replacement for the original $(s_1^I, t_1^J)$.
+Short sentences are not considered. This is, $v - u > \tau$ and $y - x > \tau $. 
+Figure~\ref{matrix} (left) shows an example of an alignment matrix $S(i,j)$ for a given sentence pair. 
+An acceptable correction is: \textit{Que ressentez-vous ? $\Leftrightarrow$ What do you feel ?}. 
+Hence, with indices $u=1$, $v=5$, $x=1$ and $y=3$.
 
@@
 Different corpora are used in this work to evaluate our similarity classifier.
-
 We filter out divergencies found in the English-French OpenSubtitles corpus~\cite{LisonTiedemann2016}, which consists of a collection of movie and TV subtitles, aligned at the sentence level after a number of automatic preprocessing steps.
@@
 %We also tested on a second domain, using the publicly available newstest-2013 En-Fr test set, corresponding to news stories selected from online sources~\cite{WMT:2013}.
-In order to better assess the quality of our classifier when facing different types of word divergencies we collected from the original corpus and annotated at the word level $500$ sentences containing:
+In order to better assess the quality of our classifier when facing different types of word divergencies we collected from the original OpenSubtitles corpus and annotated at the word level $500$ sentences containing:
 200 paired sentences;
@@
 
-All data used in this paper is preprocessed with an in-house toolkit that performs minimal tokenisation, basically splitting-off punctuation.
+All data is preprocessed with \texttt{OpenNMT}\footnote{http://opennmt.net}, performing minimal tokenisation, basically splitting-off punctuation.
 
@@
 
-After tokenisation, each out-of-vocabulary word is mapped to a special UNK token.
+After tokenisation, the $50,000$ most frequent words of each language are used as vocabulary.
+Each out-of-vocabulary word is mapped to a special UNK token.
 Our similarity classifier is described in Section~\ref{sec:similarity}. 
@@
 Both Bi-LSTM use 256-dimensional hidden representations ($E=512$). 
+We use $r=1.0$. 
 Optimization of the parameters is done using the stochastic gradient descent method along with gradient clipping (rescaling gradients whose norm exceeds a threshold) to avoid the exploding gradients problem~\cite{Pascanu:2013:DTR:3042817.3043083}.
 For each epoch we randomly select $1$ million sentence pairs that we place in batches of $32$ examples.  
-Negative examples are created following the strategies detailed in~\ref{training}.% ({\bf P}aired, {\bf U}unpaired, {\bf R}eplace and {\bf I}nsert).
+Negative examples are created following the strategies detailed in~\ref{training}.
+Word alignments and English part-of-speeches used to build negative examples were performed by \texttt{fast\_align}\footnote{https://github.com/clab/fast\_align} and \texttt{FreeLing}\footnote{https://github.com/TALP-UPC/FreeLing.git} respectively.
 We run $10$ epochs and start decaying at each epoch by $0.8$ when score on validation set increases. 
-
-Two networks are evaluated. 
-The first is the work presented in this paper, it optimises the loss shown by equation~\ref{loss_wemb}, and uses Equation~\ref{cosine} to compute similarity scores (\texttt{wemb}).
-The second implements the work in~\cite{W17-2509}, it optimises the loss shown by Equation~\ref{loss_semb} which in an end-to-end fashion predicts wether two sentences are parallel (\texttt{semb}).
-To fix translation divergencies we used $\mathcal{N}=20$ and $\tau=3$.
+Divergence is always computed following equation~\ref{cosine}. 
+For divergence correction, we use $\mathcal{N}=20$ and $\tau=3$.
 
@@
 
-In addition to the basic tokenisation we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30,000$ merge operations learned from both English and French data.
+In addition to the basic tokenisation detailed in Section~\ref{divergence} we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30000$ merge operations learned joined from both language sides.
+Neural MT systems are based on the open-source project  \texttt{OpenNMT}. 
+We use a Transformer model with same configuration than in paper~\cite{vaswani2017attention}, i.e, both encoder and decoder have $6$ layers; Multi-head attention is performed over $8$ heads. 
+Hidden layer’s size is $512$. 
+The inner layer of feed forward network is of size $2048$. 
+Word embeddings have $512$ cells. We set the dropout probability to $0.1$. 
+Batch size is set to $3072$.
+%The maximum length of both source and target sentences is set to $80$ and we limit the vocabulary size to $50,000$ words for both source and target languages. 
+The optimiser is Lazy Adam with $\beta_1 = 0.9$, $\beta_2 = 0.98$, $\epsilon = 10^{-9}$, $warmup\_steps = 4000$. We stop training after $30$ epochs.
 
-We build our NMT systems based on the open-source project \texttt{OpenNMT}\footnote{http://opennmt.net}.
+%In addition to the basic tokenisation we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30,000$ merge operations learned from both English and French data.
+%We build our NMT systems based on the open-source project \texttt{OpenNMT}.
 We use a bidirectional RNN encoder with $4$ LSTM layers with each containing $1,000$ cells.
@@
 % and we limit the vocabulary size to $50K$ words for both source and target languages.
-The default optimiser is SGD with the starting learning rate of $1.0$.
+%The default optimiser is SGD with the starting learning rate of $1.0$.
 We start to decay the learning rate from epoch $10$ or when we detect increasing perplexity as compared to the previous epoch on a validation set.
@@
 
-We evaluate first the ability of our similarity classifier to predict different types of divergencies. 
+We evaluate first the ability of our similarity classifier to predict different types of divergencies at the level of words. 
 We use the test set manually annotated for that purpose and train our model on the OpenSubtitles corpus.
-The same number of paired examples (P) is used than for any other class (U, R or I).
-%This is, in the case of the model trained using with all types of examples, the same number of examples of each class is used.
+In training, when different types of examples are used to train the network, the same number of examples are always considered for each type ({\texttt P}aired, {\texttt U}npaired, {\texttt R}eplace and {\texttt I}nsert).
+A word is considered divergent when associated to a negative aggregation score (see Equation~\ref{aggregation}).
+Table~\ref{results_puri} shows accuracies obtained by our model when trained over different combinations of negative examples.% on the task of predicting wether words are divergent. 
+We consider as divergent those words for which the similarity score predicted by our model, Equation~\ref{cosine}, is negative.
 
@@
 \end{tabular}
-\caption{Word divergence accuracies for networks trained with different combinations of negative examples. P, U, R and I stand respectively for pair, unpair, replace and insert.}
+\caption{Word divergence accuracies according to different type of examples used in train/test. P, U, R and I stand respectively for pair, unpair, replace and insert.}
 \label{results_puri}
@@
 
-Table~\ref{results_puri} shows accuracies obtained by our model when trained over different combinations of negative examples on the task of predicting wether words are divergent. 
-We consider divergent those words for which the similarity score predicted by our model, Equation~\ref{cosine}, is negative.
-
 As it can be seen, non-divergent words in parallel sentences (column \texttt{P}) are easy to identify. 
-%In all cases the accuracy reaches $100\%$. 
-Divergent words in unpaired sentences (column \texttt{U}) are also easy to identify as far as the model has seen unpaired examples in training. 
-The accuracy drops dramatically when the model is not used to see unpaired sentences (rows  \texttt{PR}, \texttt{PI} and \texttt{PRI}).
-
+In all cases the accuracy reaches $100\%$. 
+Divergent words in unpaired sentences (column \texttt{U}) are also easy to identify as far as the model has seen these types of examples in training. 
+However, the accuracy drops dramatically when the model is not trained with unpaired sentences (rows  \texttt{PR}, \texttt{PI} and \texttt{PRI}).
 Regarding columns \texttt{R} and \texttt{I}, accuracies are lower since these sentences contain a mix of divergent and non-divergent words. 
-Hence, word divergence is more difficult to predict. 
-Again, models that were trained with the corresponding examples obtain the highest accuracies (outlined in bold letters).
+%Hence, word divergence is more difficult to predict. 
+Again, models that were trained with the corresponding examples (\texttt{R} and \texttt{I}) obtain the highest accuracies (outlined in bold letters).
 Column \texttt{PURI} shows accuracy results over the entire test set. This is, mixing all type of examples. 
-The best accuracy is obtained by the system trained on all type of negative examples.
+As expected, the best accuracy is also obtained by the system trained on all type of  examples.
 
@@
     \includegraphics[width=0.4\linewidth]{feel_pu}    
-\caption{Example of neural alignment matrices and divergence scores obtained with our model when trained over pair and unpair examples (right), and over all types of examples (left) as detailed in Section~\ref{training}.
-Aggregation scores of Equation~\ref{aggregation} are shown next to source and target words.
-Matrices contain word alignment scores.
-Sentence similarity scores computed by Equation~\ref{cosine} are shown below each matrix.}
+\caption{Sentence pair with similarity scores produced by our model when trained over \texttt{PU} examples (right) and over \texttt{PURI} examples (left).
+Aggregation scores (Equation~\ref{aggregation}) are shown next to words.
+Matrices contain alignment scores.
+Sentence similarities (Equation~\ref{cosine})  shown below matrices.}
 \label{matrix}
@@
 
-Figure~\ref{matrix} shows an example of word and sentence similarity predictions when our network is trained using all types of negative examples (left); and trained using only paired and unpaired sentences (right).
-As it can be seen, the network trained using examples where all words are either divergent or non-divergent (right), fails to correctly predict the class for words in sentences with mixed divergent/non-divergent words.
-In contrast, the model trained with all types of examples (left) correctly predicts the right class for each word.
-Furthermore, it correctly assigns a lower similarity score to the sentence pair, as both sentences do not convey the exact same meaning.
-
-\subsection{Divergent Sentences Identification Task}
+Figure~\ref{matrix} illustrates the output of our network when trained using \texttt{PU} examples (right) and \texttt{PURI} examples (left).  
+The former (right) fails to predict some word divergences, most likely because in training it never saw sentences mixing divergent and non-divergent words in the same example. 
+Furthermore, the network trained over \texttt{PURI} examples (left) correctly assigns a lower similarity score to the sentence pair as both sentences do not convey the exact same meaning.
 
@@
 Consider for instance the example of Figure~\ref{matrix}. 
-Despite not conveying the exact same meaning it still contains useful information when training a neural MT engine. 
+Despite not conveying the exact same meaning it still contains useful information to train a neural MT engine. 
 It is not the case of the example in the third row of Table~\ref{tab:examples}.
+
+Table~\ref{results_wemb} shows BLEU~\cite{P02-1040} scores of the same neural MT network when trained over different samples of the OpenSubtitles corpus.
+Samples result from filtering sentences of the original corpus using different similarity thresholds.
 
@@
 
-Table~\ref{results_wemb} shows BLEU~\cite{P02-1040} scores of the same neural MT network when trained over different samples of the OpenSubtitles corpus.
-Samples result from filtering sentences of the original corpus using different similarity thresholds.
 As it can be seen, the best performing training sample is obtained when using sentence pairs with similarity score in the range $[0.076, +\infty)$. 
 This is, training with the most similar $18$ million sentences of the original data set. 
+Note that despite assigned a positive similarity score, sentence pairs in the range $[0, 0.076]$ (almost $10$ million sentences) does not seem to contribute when training a neural MT network. 
+On the contrary, when added as training data the MT accuracy is reduced.
 
-%Table~\ref{results_nmt} shows BLEU scores for the same neural MT system trained with different samples of the OpenSubtitles and Paracrawl corpus.
-For each of the corpora, OpenSubtitles and Paracrawl, Table~\ref{results_nmt} shows BLEU scores of the neural MT system trained over the entire data set (\texttt{All}), and over the $18$ (and $15$) million sentences with highest similarity score as predicted by \texttt{semb} and \texttt{wemb} classifiers.
-REPHRASE
+Table~\ref{results} shows BLEU results for our neural MT engine trained over different data configurations:
+The entire\footnote{The original Paracrawl corpus contains more than $100$ million sentences.
+We reduced its size to $22.2$ millions using standard filtering techniques.} data sets (\texttt{all}) of both corpora;
+Most similar pairs after as predicted by our network (\texttt{sim}); 
+Finally, we apply the correction algorithm detailed in Section~\ref{correction} (\texttt{sim+fix}).
+Columns Ref and Fix indicate the number of original and corrected sentences (in millions) considered to train the neural MT system.
 
-\begin{table}[h]
+\begin{table}[h!]
 \small
 \center
-\begin{tabular}{lcc}
+\begin{tabular}{lccl}
 \hline
-\bf Data & \bf Size (M) & \bf Test (BLEU) \\ %MSLT & \bf NEWS \\
+\bf Data & \bf Ref (M) & \bf Fix (M) & \bf Test (BLEU) \\ %MSLT & \bf NEWS \\
 \hline
 \multicolumn{3}{c}{\scriptsize{OpenSubtitles English-French}} \\
-\texttt{All}                    & 27.2 & 42.18 \\%& 26.60 \\ 
-\texttt{semb}               & 18.0 & \\%&  \\
-\texttt{wemb}              & 18.0 & 43.19 \\%& 26.38 \\
+\texttt{all}                   & 27.2 & - & 42.18 \\
+%\texttt{semb}             & 18.0 & - & \\
+\texttt{sim}            & 15.5 & - & 43.12  ($+0.94$)\\
+%\texttt{sim}           & 21.5 & - & 42.56 \\
+\texttt{sim}           & 18.0 & - & 43.19  ($+1.01$)\\
+\texttt{sim+fix}     & 15.5 & 2.5 & \bf 44.19 ($+2.01$)\\
+%\texttt{sim+fix}   & 15.5 & 5.0 & \\
 \hline
 \multicolumn{3}{c}{\scriptsize{Paracrawl English-German}} \\
-\texttt{All}                    & 100 & 12.56 \\ 
-\texttt{base}                & 22.0 & 19.27 \\ 
-\texttt{semb}               & 15.0 &  \\
-\texttt{wemb}              & 15.0 &   \\
+%\texttt{all}                  & 100.0 & 12.56 \\ 
+\texttt{all}                  & 22.2 & - & 19.27 \\ 
+%\texttt{semb}            & 15.0 & - & \\
+\texttt{sim}           & 15.0 & - & 21.52 ($+2.25$)\\
+\texttt{sim}           & 17.5 & - & 21.97 ($+2.70$)\\
+\texttt{sim+fix}   & 15.0 & 2.5 & \bf 22.42 ($+3.15$) \\
 \hline
 \end{tabular}
-\caption{.}
-\label{results_nmt}
+\caption{BLEU scores obtained by neural MT using different subsets of the OpenSubtitles and Paracrawl corpora.}
+\label{results}
 \end{table}
 
-%clean.min1.max99.maxw40.fert6-5-3.data.tok.en.lid.filtered
-In the case of the Paracrawl corpus, we applied first a \texttt{base} filtering that reduced the amount of sentences to $22$ millions. 
-It consists of limiting the sentence size to $100$ words. 
-We also limit the length ratio between source/target sentences to $6$, and we employed an in-house language Id classifier that helped us to detect sentence pairs not being English-German.
+Results obtained after filtering sentence pairs by our network, \texttt{sim}, clearly outperform the baseline \texttt{all} by $+0.94$ and $+2.25$ BLEU respectively.
 
-As it can be seen FINISH
+Regarding OpenSubtitles, when fixing $2.5$ million sentences ($4^{th}$ row) the accuracy is further boosted to $+2.01$, while the same sentence pairs does not show any improvement when added in their original form ($3^{rd}$ row).
+%results are further improved $+2.01$ when fixing $2.5$ million sentences ($4^{th}$ row). 
+%The same $2.5$ million sentences did not show any improvement when added in their original form
 
-\subsection{Parallel Sentences Identification Task}
-
-\begin{table}[h]
-\small
-\center
-\begin{tabular}{lrccccc}
-\hline
-\bf Model & $N=1$ & $10$ & $50$ & $100$ & $500$ & $1,000$ \\
-\hline
-\texttt{MUSE}           & & & & & & \\
-\texttt{semb}            & & & & & & \\
-\texttt{wemb}           & ssaling & & & & & \\
-\hline
-\end{tabular}
-\caption{Number of parallel sentences found on the $N$-best cosine similarity filter (FAISS) according to the sentence embeddings produced by different networks.}
-\label{results_embeddings}
-\end{table}
-
-The Cartesian product is costly for semb a filtering preprocess is necessary. Not for our model since we use faiss.
-
-\begin{table}[h]
-\small
-\center
-\begin{tabular}{lcccc}
-\hline
-\bf Model  & $\lambda$ & Precision & Recall & F-measure \\
-\hline
-\texttt{MUSE}             & & & & \\
-\texttt{wemb}             & & & & \\
-\texttt{semb}              & & & & \\
-\hline
-\end{tabular}
-\caption{Precission, Recall and F-measure for the three evaluated models on the BUCC task.}
-\label{results_similarity}
-\end{table}
+Similar results are obtained over the Paracrawl corpus.
+Results after fixing $2.5$ million sentences ($4^{th} row$) outperform those obtained by the same same sentence pairs when used by the neural MT system in their original form ($3^{rd} row$).
 
@@
 
-We presented an unsupervised method based on deep neural networks for detecting translation divergencies in parallel sentence pairs.
-Our model also predicts misaligned words that can then be filtered out allowing for reusing some divergent sentences.
-We evaluated our model on a neural machine translation task showing that it outperforms a baseline system trained on the entire (unfiltered) data set.
-The method can be used on any parallel corpus without any manual annotation.
+We presented an unsupervised method based on deep neural networks for detecting translation divergences in parallel sentence pairs. 
+The model also predicts misaligned words that can then be filtered out allowing for reusing some divergent sentences. 
+We evaluated our model on two neural machine translation tasks, showing significant improvements when compared to training over the entire data set. 
+%The method can be used on any parallel corpus without any manual annotation.
 
@@
 
-We plan to further evaluate divergence classification under larger data size conditions and different noise levels and on additional language pairs. 
-%In this work we used a (noisy) parallel corpora to learn a divergence model. We also would like to evaluate the impact of using using less noisy (already filtered) data to learn the model. 
-We also plan to study the use of a neural network to predict divergent words, as a replacement for the algorithm detailed in Section~\ref{correction}.
+%We plan to further evaluate divergence classification under different data noise levels. %and on additional language pairs. 
+%as well as to measure the impact of using less noisy data when learning the similarity model
+%We also would like to evaluate the impact of using less noisy (already filtered) data to learn the similarity model. 
+%We would also like to use our model to predict other types of divergences 
+We plan to use our model to predict sentence embeddings over monolingual corpora, allowing to collect parallel pairs through vector similarity measures.
+%We also plan to further evaluate the correction algorithm.
+In addition, we would like to measure the performance of our model after applying subword tokenisation,
+%when working with a reduced vocabulary set, as a result of applying any subword tokenisation,
+as well as using multiple LSTM layers, which is well known to capture hierarchical structure in the context of MT.
 

### 2018-07-26 WMT/divergence.tex
more changes in section 3

@@
 
-In this section we describe the architecture of our model, very much inspired by the work on Word Alignment in~\cite{W16-2207}. Figure~\ref{network} illustrates the network.
+In this section we describe the architecture of our model, inspired by the work on word alignment of~\citet{W16-2207}.
 
-In the following, we consider a source-target sentence pair $(s,t)$ with $s=(s_1,...,s_I)$ and $t=(t_1,...,t_J)$.
-The model is composed of 2 Bi-directional LSTM subnetworks, $net_s$ and $net_t$, which respectively encode source and target sentences.
-Since both $net_s$ and $net_t$ take the same form we describe only the source architecture.
+In the following, we consider a source-target sentence pair $(s,t)$
+with $s=(s_1,...,s_I)$ and $t=(t_1,...,t_J)$. The model is composed of
+2 Bi-directional LSTM subnetworks (see Figure~\ref{network}), $net_s$ and $net_t$, which respectively encode source and target sentences.
+Since both $net_s$ and $net_t$ take the same form, we only describe the source architecture.
 
@@
 %$h_{src} = \left( \begin{array}{cc} \overrightarrow{h}^{src}_I \\ \overleftarrow{h}^{src}_1 \\ \end{array} \right)$.
-At this point a measure of similarity between sentence pairs can be obtained by cosine similarity:
-
+At this point, a measure of similarity between sentence pairs is the
+cosine of their represention vector: 
 \begin{equation}
-    sim(h_{src}, h_{tgt}) = \frac{h_{src} \cdotp h_{tgt}}{||h_{src}|| * ||h_{tgt}||}
+    sim(h_{src}, h_{tgt}) = \frac{h_{src} \cdotp h_{tgt}}{||h_{src}|| * ||h_{tgt}||},
     \label{cosine}
@@
     \includegraphics[width=1.0\linewidth]{network}
-    \caption{Illustration of the model.
+    \caption{Description of our model.
 The network is composed of source and target word embedding lookup tables ($LT_s$ and $LT_t$) and two identical subnetworks ($net_s$ and $net_t$) that compute in context representations of source ($s_i$) and target words ($t_j$).}
@@
 The objective function is computed at the level of words.
-To enable unsupervised training, we use an aggregation operation that summarizes the alignment scores for a given target word. 
-A soft-margin objective increases scores for true target words while decreasing scores for target words that are not present.
-The aggregation function combines the scores of all source (or target) words for a particular target (or source) word and promotes source words which are likely to be aligned with a given target word according to the knowledge the model has learned so far.
+To enable unsupervised training, we use an aggregation operation
+($aggr()$) that summarizes the alignment scores for any given target word. 
+A soft-margin objective increases scores for true target words, while decreasing scores for target words that are not present.
+The aggregation function combines the scores of all source (resp.\ target) words for a particular target (resp.\ source) word and promotes source words which are likely to be aligned with a given target word according to the knowledge the model has learned so far.
 Alignment scores $S(i,j)$ are given by the dot-product $S(i,j) = h_i^{src} \cdotp h_j^{tgt}$, while aggregation functions are defined as:
@@
 
-Training is performed by minimising Equation~\ref{loss_wemb}, for which examples with annotations for source $\mathcal{Y}_i^{src}$ and target $\mathcal{Y}_j^{tgt}$ words are needed.
+Training is performed by minimising the loss in Equation~\eqref{loss_wemb}, for which examples with annotations for source $\mathcal{Y}_i^{src}$ and target $\mathcal{Y}_j^{tgt}$ words are needed.
 Different kind of examples are used to train our network:
@@
 
-\item As positive examples we use paired sentences of the parallel corpus.
-In this case, all words of the pair of sentences are labelled as parallel, $\mathcal{Y}_i^{src}=-1$ and $\mathcal{Y}_j^{tgt}=-1$.
+\item As positive examples, we use paired sentences of the parallel corpus.
+In this case, all words in the pair of sentences are labelled as parallel, $\mathcal{Y}_i^{src}=-1$ and $\mathcal{Y}_j^{tgt}=-1$.
 
-\item As negative examples we use random unpaired sentences. In this case, all words are labelled as divergent, $\mathcal{Y}_i^{src}=+1$ and $\mathcal{Y}_j^{tgt}=+1$. 
+\item As negative examples, we use random unpaired sentences. In this case, all words are labelled as divergent, $\mathcal{Y}_i^{src}=+1$ and $\mathcal{Y}_j^{tgt}=+1$. 
 Since negative examples may be very easy to classify and we want our network to detect less obvious divergencies, we further create negative examples following the next two methods:
@@
 We use a Transformer model with same configuration than in paper~\cite{vaswani2017attention}, i.e, both encoder and decoder have $6$ layers; Multi-head attention is performed over $8$ heads. 
-Hidden layer’s size is $512$. 
+Hidden layer's size is $512$. 
 The inner layer of feed forward network is of size $2048$. 

### 2018-07-26 WMT/divergence.tex
relecture complete - reste quelques doutes - look for FY

@@
 Section~\ref{sec:results} evaluates results. 
-Finally, conclusions are drawn in Section~\ref{conclusions} and further work is outlined in Section~\ref{further}.
-All the code used in this paper is freely available\footnote{https://github.com/anonymised}.
+Finally, we present some conclusions and discuss future work in
+Section~\ref{conclusions+further}.
+All the code used in this paper is freely
+available\footnote{https://github.com/anonymised}.
 
@@
 
-Training is performed by minimising the loss in Equation~\eqref{loss_wemb}, for which examples with annotations for source $\mathcal{Y}_i^{src}$ and target $\mathcal{Y}_j^{tgt}$ words are needed.
+Training is performed by minimising the loss in
+Equation~\eqref{loss_wemb}, for which annotated examples of
+source $\mathcal{Y}_i^{src}$ and target $\mathcal{Y}_j^{tgt}$ words
+are needed. 
 Different kind of examples are used to train our network:
@@
 
-\item As negative examples, we use random unpaired sentences. In this case, all words are labelled as divergent, $\mathcal{Y}_i^{src}=+1$ and $\mathcal{Y}_j^{tgt}=+1$. 
-Since negative examples may be very easy to classify and we want our network to detect less obvious divergencies, we further create negative examples following the next two methods:
+\item As negative examples, we use random unpaired sentences. In this
+  case, all words are labelled as divergent, $\mathcal{Y}_i^{src}=+1$
+  and $\mathcal{Y}_j^{tgt}=+1$.
+
+Since negative examples may be very easy to classify and we want our
+network to detect less obvious divergences, we further create negative
+examples according to the next two methods:
 
@@
 The rationale behind this method is to keep the new sentences as grammatical as possible. 
-Otherwise, to predict divergence the network can learn to detect non-grammatical sentences.
-Words that are not replaced are considered parallel ($-1$) while those replaced are assigned the divergent label ($+1$). 
+Otherwise, to predict divergence the network could learn to detect non-grammatical sentences.
+Words that are not replaced are considered parallel ($-1$), whereas those replaced are assigned the divergent label ($+1$). 
 Words aligned to some replaced words are also assigned the divergent label ($+1$). For instance, given the original sentence pair:
@@
 
-We may replace '\texttt{you feel}', with part-of-speech tags '\texttt{PRP VB}', by another sequence with same tags (i.e. '\texttt{we want}'):
+We may replace '\texttt{you feel}', with part-of-speech tags
+'\texttt{PRP VB}', by another sequence with same tags (i.e. '\texttt{we want}'):
 
@@
 
-In order to avoid that negative examples are easily predicted just by looking at the difference in length of training sentences we constraint all negative examples to have a difference in length not exceeding $2.0$.
-Very short sentences, of up to $4$ words, are accepted if the length ratio does not exceeds $3.0$.
+In order to avoid that negative examples are easily predicted just by looking at the difference in length of training sentences we constraint all negative examples to have a difference in length not exceeding $2$.
+Very short sentences, of up to $4$ words, are accepted if the length ratio does not exceeds $3$.
 
-\subsection{Fixing Translation Divergencies}
+\subsection{Fixing Translation Divergences}
 \label{correction}
@@
 We observed in our training corpora that many divergent sentences follow a common pattern, consisting of adding some extra leading/trailing words.
-As in the first and second examples of table~\ref{tab:examples}.
-Accordingly, we implemented a simple algorithm that discards sequences of leading/trailing words on both sides of the source and/or target sentences. 
-Hence considering as parallel $s_u^v$ and $t_x^y$. 
-To find the optimal source ($u, v$) and target ($x, y$) indices that enclose the parallel segments within the original sentence, we implement:
+This is for instance the case of the first and second examples in Table~\ref{tab:examples}.
+Accordingly, we implemented a simple algorithm that discards sequences
+of leading/trailing words on both sides of the source and/or target
+sentences, and considers that only two substrings $s_u^v$ and $t_x^y$
+of the original sentences are actually parallel.
+To find the optimal source ($u, v$) and target ($x, y$) indices that
+define the parallel segments in the original sentence, we solve:
 \begin{equation*}
-\underset{u, v, x, y}{\arg\max} \Big \{      \underset{u \le I \le v}{\sum} \underset{x \le j \le y}{\max} \{ S(i,j) \}    \Big \}
+\underset{u, v, x, y}{\arg\max} \Big \{      \underset{u \le I \le v}{\sum} \underset{x \le j \le y}{\max} \{ S(i,j) \}    \Big \}.
 \end{equation*}
@@
 The $\mathcal{N}$-best sequences following the previous function ($s_u^v$, $t_x^y$) are considered as valid corrections, but only the highest ranked according to its similarity score is used as replacement for the original $(s_1^I, t_1^J)$.
-Short sentences are not considered. This is, $v - u > \tau$ and $y - x > \tau $. 
-Figure~\ref{matrix} (left) shows an example of an alignment matrix $S(i,j)$ for a given sentence pair. 
-An acceptable correction is: \textit{Que ressentez-vous ? $\Leftrightarrow$ What do you feel ?}. 
-Hence, with indices $u=1$, $v=5$, $x=1$ and $y=3$.
+Short sentences are not considered. This is, we enforce $v - u > \tau$
+and $y - x > \tau $, with $\tau$ a predefined parameter. 
+Figure~\ref{matrix} (left) displays an example of an alignment matrix $S(i,j)$ for a given sentence pair. 
+An acceptable correction is: \textit{Que ressentez-vous ?
+  $\Leftrightarrow$ What do you feel ?}, corresponding to the 4-uples
+of indices $u=1$, $v=5$, $x=1$ and $y=3$.
 
@@
 
-In this section we report on the experiments conducted to evaluate the proposed network. We begin with details of the corpora employed.
+In this section we report on the experiments conducted to evaluate our
+architecture. We begin with details regarding our training and test corpora.
 
@@
 
-Different corpora are used in this work to evaluate our similarity classifier.
+Multiple corpora are used in this work to evaluate our similarity classifier.
 We filter out divergencies found in the English-French OpenSubtitles corpus~\cite{LisonTiedemann2016}, which consists of a collection of movie and TV subtitles, aligned at the sentence level after a number of automatic preprocessing steps.
-The corpus presents many potential divergencies as outlined in Table~\ref{tab:examples}. 
-To evaluate performance we used the En-Fr Microsoft Spoken Language Translation task, created from actual conversations over Skype~\cite{mslt-corpus-iwslt-2016-release}. 
+The corpus presents many potential divergences as illustrated in Table~\ref{tab:examples}.
+
 %We also tested on a second domain, using the publicly available newstest-2013 En-Fr test set, corresponding to news stories selected from online sources~\cite{WMT:2013}.
-In order to better assess the quality of our classifier when facing different types of word divergencies we collected from the original OpenSubtitles corpus and annotated at the word level $500$ sentences containing:
+In order to better assess the quality of our classifier when facing different types of word divergences, we selected from the original OpenSubtitles corpus and annotated at the word level $500$ sentences containing:
 200 paired sentences;
@@
 
-We also use the Paracrawl corpus, which consists of a very noisy $1$ billion word (English token count) German-English corpus crawled from the web as part of the Paracrawl project\footnote{http://paracrawl.eu/}.
-Performance is evaluated on the publicly available newstest-2017 En-De test set, corresponding to news stories selected from online sources~\cite{W17-4717}.
+As an indirect evaluation of our filtering method, we also consider
+two translation tasks.
+Again using OpenSubtitles for training our system, we test on the En-Fr Microsoft Spoken Language Translation task, created from actual conversations over Skype~\cite{mslt-corpus-iwslt-2016-release}.
+
+A second translation task uses the Paracrawl corpus, which consists of
+a very noisy $1$ billion word (English token count) German-English
+corpus crawled from the web as part of the Paracrawl
+project\footnote{http://paracrawl.eu/}.
+Translation performance is evaluated on the publicly available newstest-2017 En-De test set, corresponding to news stories selected from online sources~\cite{W17-4717}.
 
@@
 
-After tokenisation, the $50,000$ most frequent words of each language are used as vocabulary.
+After tokenisation, the $50,000$ most frequent words of each language
+define the vocabulary.
 Each out-of-vocabulary word is mapped to a special UNK token.
-Our similarity classifier is described in Section~\ref{sec:similarity}. 
+% Our similarity classifier is described in Section~\ref{sec:similarity}. 
 Word embeddings ($LT_s$ and $LT_t$) are initialised using \texttt{fastText}\footnote{https://github.com/facebookresearch/fastText}, further aligned by means of \texttt{MUSE}\footnote{https://github.com/facebookresearch/MUSE} following the unsupervised method detailed in~\cite{lample2018word}.
-Size of embeddings is $E_s=E_t=256$ cells. 
-Both Bi-LSTM use 256-dimensional hidden representations ($E=512$). 
-We use $r=1.0$. 
-Optimization of the parameters is done using the stochastic gradient descent method along with gradient clipping (rescaling gradients whose norm exceeds a threshold) to avoid the exploding gradients problem~\cite{Pascanu:2013:DTR:3042817.3043083}.
-For each epoch we randomly select $1$ million sentence pairs that we place in batches of $32$ examples.  
+The size of embeddings is $E_s=E_t=256$ cells. 
+Both Bi-LSTMs use 256-dimensional hidden representations ($E=512$); and
+we also set $r=1.0$ in Equation~\eqref{aggregation}.
+Optimization of the parameters is done using stochastic gradient descent method along with gradient clipping (rescaling gradients whose norm exceeds a threshold) to avoid the exploding gradients problem~\cite{Pascanu:2013:DTR:3042817.3043083}.
+For each epoch, we randomly select $1$ million sentence pairs that we
+place in batches of $32$ examples. We run $10$ epochs and start decaying at each epoch by $0.8$ when the
+score on a validation set increases.
+
 Negative examples are created following the strategies detailed in~\ref{training}.
-Word alignments and English part-of-speeches used to build negative examples were performed by \texttt{fast\_align}\footnote{https://github.com/clab/fast\_align} and \texttt{FreeLing}\footnote{https://github.com/TALP-UPC/FreeLing.git} respectively.
-We run $10$ epochs and start decaying at each epoch by $0.8$ when score on validation set increases. 
-Divergence is always computed following equation~\ref{cosine}. 
+Word alignments and English part-of-speeches used to build negative
+examples were computed respectively by \texttt{fast\_align}\footnote{https://github.com/clab/fast\_align} and \texttt{FreeLing}\footnote{https://github.com/TALP-UPC/FreeLing.git}.
+Divergence is always computed according to Equation~\eqref{cosine}. 
 For divergence correction, we use $\mathcal{N}=20$ and $\tau=3$.
@@
 
-In addition to the basic tokenisation detailed in Section~\ref{divergence} we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30000$ merge operations learned joined from both language sides.
+In addition to the basic tokenisation detailed in Section~\ref{divergence}, we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30000$ merge operations learned jointly from both language sides.
 Neural MT systems are based on the open-source project  \texttt{OpenNMT}. 
-We use a Transformer model with same configuration than in paper~\cite{vaswani2017attention}, i.e, both encoder and decoder have $6$ layers; Multi-head attention is performed over $8$ heads. 
-Hidden layer's size is $512$. 
-The inner layer of feed forward network is of size $2048$. 
-Word embeddings have $512$ cells. We set the dropout probability to $0.1$. 
-Batch size is set to $3072$.
+We use a Transformer model with configuration similar to the
+description in~\cite{vaswani2017attention}: both encoder and decoder
+have $6$ layers; Multi-head attention is performed over $8$
+heads. Additional parameters are set follows:
+the hidden layer's size is $512$; the inner layer of the feed forward network is of size $2048$;
+word embeddings have $512$ cells. We alsoe set the dropout probability
+to $0.1$, and the batch size to $3072$.
 %The maximum length of both source and target sentences is set to $80$ and we limit the vocabulary size to $50,000$ words for both source and target languages. 
@@
 
-We evaluate first the ability of our similarity classifier to predict different types of divergencies at the level of words. 
-We use the test set manually annotated for that purpose and train our model on the OpenSubtitles corpus.
+We first evaluate the ability of our similarity classifier to predict different types of divergences at the level of words. 
+We use the manually annotated test set for that purpose and train our model on the OpenSubtitles corpus.
 In training, when different types of examples are used to train the network, the same number of examples are always considered for each type ({\texttt P}aired, {\texttt U}npaired, {\texttt R}eplace and {\texttt I}nsert).
-A word is considered divergent when associated to a negative aggregation score (see Equation~\ref{aggregation}).
-Table~\ref{results_puri} shows accuracies obtained by our model when trained over different combinations of negative examples.% on the task of predicting wether words are divergent. 
-We consider as divergent those words for which the similarity score predicted by our model, Equation~\ref{cosine}, is negative.
+A word is considered divergent when associated to a negative aggregation score (see Equation~\eqref{aggregation}).
+Table~\ref{results_puri} reports the accuracies obtained by our model
+when trained over different combinations of negative examples.% on the
+                                % task of predicting wether words are
+                                % divergent. 
+% FY - Is it not contradictory with previous sentence ? Do you mean
+% divergent *sentences*
+We consider as divergent, words for which the similarity score predicted by our model, Equation~\eqref{cosine}, is negative.
 
@@
 Divergent words in unpaired sentences (column \texttt{U}) are also easy to identify as far as the model has seen these types of examples in training. 
-However, the accuracy drops dramatically when the model is not trained with unpaired sentences (rows  \texttt{PR}, \texttt{PI} and \texttt{PRI}).
+However, the accuracy dramatically drops when the model is not trained with unpaired sentences (rows  \texttt{PR}, \texttt{PI} and \texttt{PRI}).
 Regarding columns \texttt{R} and \texttt{I}, accuracies are lower since these sentences contain a mix of divergent and non-divergent words. 
@@
 Again, models that were trained with the corresponding examples (\texttt{R} and \texttt{I}) obtain the highest accuracies (outlined in bold letters).
-Column \texttt{PURI} shows accuracy results over the entire test set. This is, mixing all type of examples. 
-As expected, the best accuracy is also obtained by the system trained on all type of  examples.
+Column \texttt{PURI} shows accuracy results over the entire test set,
+that is, when mixing all type of examples. As expected, the best
+accuracy is also obtained by the system trained on all type of  examples.
 
@@
     \includegraphics[width=0.4\linewidth]{feel_pu}    
-\caption{Sentence pair with similarity scores produced by our model when trained over \texttt{PU} examples (right) and over \texttt{PURI} examples (left).
-Aggregation scores (Equation~\ref{aggregation}) are shown next to words.
-Matrices contain alignment scores.
-Sentence similarities (Equation~\ref{cosine})  shown below matrices.}
+\caption{Sentence pair with similarity scores produced by our model
+  when trained over \texttt{PU} examples (right) and over
+  \texttt{PURI} examples (left). Aggregation scores
+  (Equation~\eqref{aggregation}) are shown next to words. Matrices
+  contain alignment scores. Sentence similarities
+  (Equation~\eqref{cosine})  are at the bottom.}
 \label{matrix}
@@
 The former (right) fails to predict some word divergences, most likely because in training it never saw sentences mixing divergent and non-divergent words in the same example. 
-Furthermore, the network trained over \texttt{PURI} examples (left) correctly assigns a lower similarity score to the sentence pair as both sentences do not convey the exact same meaning.
+Furthermore, the network trained over \texttt{PURI} examples (left) correctly assigns a low similarity score to the sentence pair, since both sentences do not convey the exact same meaning.
 
@@
 
-Table~\ref{results_wemb} shows BLEU~\cite{P02-1040} scores of the same neural MT network when trained over different samples of the OpenSubtitles corpus.
+Table~\ref{results_wemb} reports BLEU~\cite{P02-1040} scores of the same neural MT network when trained over different samples of the OpenSubtitles corpus.
 Samples result from filtering sentences of the original corpus using different similarity thresholds.
@@
 
-As it can be seen, the best performing training sample is obtained when using sentence pairs with similarity score in the range $[0.076, +\infty)$. 
-This is, training with the most similar $18$ million sentences of the original data set. 
-Note that despite assigned a positive similarity score, sentence pairs in the range $[0, 0.076]$ (almost $10$ million sentences) does not seem to contribute when training a neural MT network. 
-On the contrary, when added as training data the MT accuracy is reduced.
-
-Table~\ref{results} shows BLEU results for our neural MT engine trained over different data configurations:
-The entire\footnote{The original Paracrawl corpus contains more than $100$ million sentences.
+As it can be seen in Table~\ref{results_wemb}, the best performing
+training sample is obtained when using sentence pairs with similarity
+score in the range $[0.076, +\infty)$, which means training with the 
+most similar $18$ million sentences in the original data set. 
+Note that despite assigned a positive similarity score, sentence pairs in the range $[0, 0.076]$ (almost $10$ million sentences) does not seem to contribute when training a neural MT system. 
+On the contrary, when they are added as training data, the MT accuracy is reduced.
+% FY - cite standard filtering techniques
+Table~\ref{results} reports BLEU results for our neural MT engine trained over three data configurations.
+when using the entire\footnote{The original Paracrawl corpus contains more than $100$ million sentences.
 We reduced its size to $22.2$ millions using standard filtering techniques.} data sets (\texttt{all}) of both corpora;
-Most similar pairs after as predicted by our network (\texttt{sim}); 
-Finally, we apply the correction algorithm detailed in Section~\ref{correction} (\texttt{sim+fix}).
-Columns Ref and Fix indicate the number of original and corrected sentences (in millions) considered to train the neural MT system.
+when using similar pairs after as predicted by our network (\texttt{sim}); 
+after also applying the correction algorithm of Section~\ref{correction} (\texttt{sim+fix}).
+Columns Ref and Fix indicate the number of original and corrected
+sentences (in millions) used to train the neural MT system.
 
@@
 
-Regarding OpenSubtitles, when fixing $2.5$ million sentences ($4^{th}$ row) the accuracy is further boosted to $+2.01$, while the same sentence pairs does not show any improvement when added in their original form ($3^{rd}$ row).
+Regarding OpenSubtitles, when fixing $2.5$ million sentences ($4^{th}$ row) the accuracy is further boosted to $+2.01$, while the same sentence pairs do not show any improvement when added in their original form ($3^{rd}$ row).
 %results are further improved $+2.01$ when fixing $2.5$ million sentences ($4^{th}$ row). 
@@
 
-Similar results are obtained over the Paracrawl corpus.
+Similar results are obtained with the Paracrawl corpus.
 Results after fixing $2.5$ million sentences ($4^{th} row$) outperform those obtained by the same same sentence pairs when used by the neural MT system in their original form ($3^{rd} row$).
 
-\section{Conclusions}
-\label{conclusions}
+\section{Conclusions and future work}
+\label{conclusions+further}
 
 We presented an unsupervised method based on deep neural networks for detecting translation divergences in parallel sentence pairs. 
-The model also predicts misaligned words that can then be filtered out allowing for reusing some divergent sentences. 
-We evaluated our model on two neural machine translation tasks, showing significant improvements when compared to training over the entire data set. 
+The model also predicts misaligned words that can then be filtered
+out, thus enabling us to reuse some divergent sentences. 
+We evaluated our model on two neural machine translation tasks, 
+reporting significant improvements when compared to training over the entire data set. 
 %The method can be used on any parallel corpus without any manual annotation.
-
-\section{Further Work}
-\label{further}
 
@@
 %We also would like to evaluate the impact of using less noisy (already filtered) data to learn the similarity model. 
-%We would also like to use our model to predict other types of divergences 
-We plan to use our model to predict sentence embeddings over monolingual corpora, allowing to collect parallel pairs through vector similarity measures.
+%We would also like to use our model to predict other types of
+%divergences 
+% FY - this should be on comparable corpora ?
+We plan to use our model to predict sentence embeddings in monolingual corpora, allowing to collect parallel pairs through vector similarity measures.
 %We also plan to further evaluate the correction algorithm.

### 2018-07-27 WMT/divergence.tex
score osub
rewrite

--- 2018-07-26 WMT/divergence.tex
@@
 %\title{Crosslingual Sentence Embeddings for Neural Machine Translation}
-\title{A Recurrent Network for Parallel Corpora Similarity Analysis}
+%\title{A Recurrent Network for Parallel Corpora Similarity Analysis}
+\title{Detecting and Fixing Translation Divergences in Noisy Parallel Corpora}
 
-\author{%Minh Pham, \ Fran\c cois Yvon\\
-  First author\\
-  Affiliation / Address line 1\\
-  Affiliation / Address line 2\\
-  Affiliation / Address line 3\\
-  {\tt email@domain}  
-  \And
-  %Josep Crego, \ Jean Senellart\\
-  Second author\\
-  Affiliation / Address line 1\\
-  Affiliation / Address line 2\\
-  Affiliation / Address line 3\\
-  {\tt email@domain}}
+\author{MinhQuang Pham$^{\dag\ddag}$, \ \ Josep Crego$^\dag$,\ \ Jean Senellart$^\dag$ ,\ \ Fran\c cois Yvon$^\ddag$\\
+  $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France\\
+  {\tt firstname.lastname@systrangroup.com}\\
+  $^\ddag$LIMSI-CNRS / Universit\'e Paris-Saclay 91405 Orsay, France\\
+  {\tt firstname.lastname@limsi.fr}}
 
@@
 
-In this work, we present an unsupervised method for building cross-lingual sentence embeddings in the form of continuous vectors. 
-These embeddings are then used to measure semantic equivalence of their corresponding sentences.
+In this work, we present an unsupervised method for building cross-lingual sentence embeddings based on modelling word similarity. % in the form of continuous vectors. 
+The architecture of our network allows us to distinguish between different types of divergences common to many bi-texts.
+The resulting sentence embeddings are then used to measure semantic equivalence of the corresponding sentences.
 To evaluate our method we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
@@
 The authors show that a simple SVM classifier can be built on top of the sentence representations to achieve state-of-the-art results in a semantic entailment classification task.
-With the same objective, the system presented in~\cite{N16-1108} uses multiple convolutional layers and models pairwise word interactions. 
-Recently, in~\cite{P18-2037} is presented a network able to learn a multilingual sentence embedding space that is successfully used to filter noisy parallel data and to mine for bi-texts in large monolingual collections.
-Our work is inspired by~\cite{W17-3209} where the authors train a cross-lingual divergence detector using word alignments and sentence length features to train a linear SVM classifier. 
+
+Similar to our work, in ~\cite{W17-3209} the authors train a cross-lingual divergence detector using word alignments and sentence length features to train a linear SVM classifier. 
 Their work shows that an NMT system trained only on non-divergent sentences yields slightly higher translation quality scores and requires clearly less training time. 
 The same authors have recently updated their work in~\cite{DBLP:journals/corr/abs-1803-11112}. 
-The objective is the same as the neural network presented in~\cite{N16-1108} and their network further outperforms their previous work.
+They use the system presented in ~\cite{N16-1108} that employs multiple convolutional layers and models pairwise word interactions. 
+The updated network further outperforms their previous work.
 
-Our work differs from the previous as we make use of a network with different topology. 
-We model sentence similarity by means of optimising word alignment and in addition, we propose a simple method that employs alignment scores to not only detect but correct divergent sentences.
+Recently, in~\cite{P18-2037} is presented a network able to learn a joint multilingual sentence embedding space, with the goal that the distance in that space reflects their semantic difference, independently of the language.
+The network learns sentence embeddings by max-pooling over several Bi-LSTM outputs. 
+The system is successfully used to filter noisy parallel data and to mine for bi-texts in large monolingual collections.
+
+As in the previous works we also use a recurrent network to build sentence embeddings but we optimise word similarity rather than sentence similarity. 
+Thus, enabling fine grained divergence prediction at the level of words.
+In addition, we propose a simple method that employs word similarity scores to not only detect but to correct divergent sentences, which allows to reuse additional sentence pairs in the context of neural machine translation.
 
@@
 Since both $net_s$ and $net_t$ take the same form we describe only the source architecture.
-
 The source-sentence Bi-LSTM network outputs forward and backward hidden states, $\overrightarrow{h}^{src}_i$ and $\overleftarrow{h}^{src}_i$, which are then concatenated into a single vector encoding the $i^{th}$ word of the source sentence,
 $h^{src}_i = [ \overrightarrow{h}^{src}_i ; \overleftarrow{h}^{src}_i ]$.
-%$h^{src}_i = \left( \begin{array}{cc} \overrightarrow{h}_i \\ \overleftarrow{h}_i \\ \end{array} \right)$.
 
@@
 $h_{src} = [ \overrightarrow{h}^{src}_I ; \overleftarrow{h}^{src}_1 ]$.
-%$h_{src} = \left( \begin{array}{cc} \overrightarrow{h}^{src}_I \\ \overleftarrow{h}^{src}_1 \\ \end{array} \right)$.
-At this point a measure of similarity between sentence pairs can be obtained by cosine similarity:
 
+\begin{figure}[h!]
+\center
+    \includegraphics[width=1.0\linewidth]{network}
+    \caption{Illustration of the model.
+The network is composed of source and target word embedding lookup tables ($LT_s$ and $LT_t$) and two identical subnetworks ($net_s$ and $net_t$) that compute in context representations of source ($s_i$) and target words ($t_j$).}
+    \label{network}
+\end{figure}
+
+At this point a measure of similarity between sentences can be obtained by cosine similarity: 
 \begin{equation}
@@
 
-\begin{figure}[h]
-\center
-    \includegraphics[width=1.0\linewidth]{network}
-    \caption{Illustration of the model.
-The network is composed of source and target word embedding lookup tables ($LT_s$ and $LT_t$) and two identical subnetworks ($net_s$ and $net_t$) that compute in context representations of source ($s_i$) and target words ($t_j$).}
-    \label{network}
-\end{figure}
 %The similarity function outputs the level of semantic similarity between source and target sentences.
@@
 
-\item As positive examples we use paired sentences of the parallel corpus.
-In this case, all words of the pair of sentences are labelled as parallel, $\mathcal{Y}_i^{src}=-1$ and $\mathcal{Y}_j^{tgt}=-1$.
+\item As positive examples we use paired sentences of the parallel corpus. In this case, all words in both sentences are labelled as parallel, $\mathcal{Y}_i^{src}=-1$ and $\mathcal{Y}_j^{tgt}=-1$.
 
@@
 Very short sentences, of up to $4$ words, are accepted if the length ratio does not exceeds $3.0$.
+Note that the network can be trained with any combination of examples ({\texttt P}aired, {\texttt U}npaired, {\texttt R}eplace and {\texttt I}nsert).
+In all cases, the same number of examples is used for each type.
 
@@
 Figure~\ref{matrix} (left) shows an example of an alignment matrix $S(i,j)$ for a given sentence pair. 
-An acceptable correction is: \textit{Que ressentez-vous ? $\Leftrightarrow$ What do you feel ?}. 
-Hence, with indices $u=1$, $v=5$, $x=1$ and $y=3$.
+An acceptable correction is:
+
+\begin{table}[h]
+\center
+\begin{tabular}{c}
+	\textit{Que ressentez-vous ? \ \ $\Leftrightarrow$\ \ \ What do you feel ?} \\
+\end{tabular}
+\end{table}
+
+\noindent Hence, with optimal indices being: $u=1$, $v=5$, $x=1$ and $y=3$.
 
@@
 %We also tested on a second domain, using the publicly available newstest-2013 En-Fr test set, corresponding to news stories selected from online sources~\cite{WMT:2013}.
-In order to better assess the quality of our classifier when facing different types of word divergencies we collected from the original OpenSubtitles corpus and annotated at the word level $500$ sentences containing:
-200 paired sentences;
-100 unpaired sentences;
-100 sentences with replace examples; and
-100 sentences with insert examples as detailed in Section~\ref{training}.
+
+In order to better assess the quality of our classifier when facing different types of word divergences we collected from the original OpenSubtitles corpus $500$ sentences containing different types of examples:
+\begin{itemize}
+\item $200$ paired sentences,
+\item $100$ unpaired sentences,
+\item $100$ sentences with replace examples and
+%\item $100$ sentences with replace examples; 
+\item $100$ sentences with insert examples.
+\end{itemize}
+
+Details of each type of example are given in Section~\ref{training}.
+Sentence pairs have divergence annotations at the level of words.
 
@@
 We use the test set manually annotated for that purpose and train our model on the OpenSubtitles corpus.
-In training, when different types of examples are used to train the network, the same number of examples are always considered for each type ({\texttt P}aired, {\texttt U}npaired, {\texttt R}eplace and {\texttt I}nsert).
 A word is considered divergent when associated to a negative aggregation score (see Equation~\ref{aggregation}).
 Table~\ref{results_puri} shows accuracies obtained by our model when trained over different combinations of negative examples.% on the task of predicting wether words are divergent. 
-We consider as divergent those words for which the similarity score predicted by our model, Equation~\ref{cosine}, is negative.
+%We consider as divergent those words for which the similarity score predicted by our model, Equation~\ref{cosine}, is negative.
 
@@
 We presented an unsupervised method based on deep neural networks for detecting translation divergences in parallel sentence pairs. 
-The model also predicts misaligned words that can then be filtered out allowing for reusing some divergent sentences. 
+Our model optimises word alignments. Hence, allowing for fine grained divergence prediction at the level of words. 
+Misaligned/divergent words can then be filtered out allowing for reusing some divergent sentences. 
 We evaluated our model on two neural machine translation tasks, showing significant improvements when compared to training over the entire data set. 
-%The method can be used on any parallel corpus without any manual annotation.
+The method can be used on any parallel corpus without any manual annotation.
 
@@
 %We would also like to use our model to predict other types of divergences 
-We plan to use our model to predict sentence embeddings over monolingual corpora, allowing to collect parallel pairs through vector similarity measures.
+We plan to use our model to predict sentence embeddings over comparable corpora, allowing to collect parallel pairs through vector similarity measures.
 %We also plan to further evaluate the correction algorithm.

### 2018-07-27 WMT/divergence.tex
copy

@@
 In the case of neural machine translation, a large neural network is trained to maximise the translation performance on a given parallel corpus. 
-This paper describes an unsupervised method for building cross-lingual sentence embeddings that optimise word alignments. 
-Such embeddings can then be used to compute semantic similarity between sentences in different languages.
-We first use sentence embeddings to detect translation divergences in parallel sentence pairs, thereby helping to filter out noisy pairs.
-Then, word similarity scores predicted by the network are used to identify and fix some divergences guiding to collect additional parallel segments.
-We evaluate our model on multiple language pairs and data resources.
-This model can be used on any parallel corpus without any manual annotation.
+Such resource is not naturally existing, and because of the process necessary to compile a parallel corpus, it may contain multiple sentence pairs that are often not as parallel as one might assume.
+%%
+This paper describes an unsupervised method for detecting translation divergences in parallel sentences.
+We present a neural network to predict sentence similarity that minimises a loss function based on word divergence.
+We show that accurate predictions are obtained allowing divergent sentences to be filtered out.
+Furthermore, word similarity scores predicted by the network are used to identify and fix some divergences guiding to collect additional parallel segments.
+We evaluate the presented method on  English-French and  English-German machine translation tasks.
+Results show that neural MT systems trained on the filtered/corrected corpus outperform the MT systems trained on the original data.
+%%
+%This paper describes an unsupervised method for building cross-lingual sentence embeddings that optimise word alignments. 
+%Such embeddings can then be used to compute semantic similarity between sentences in different languages.
+%We first use sentence embeddings to detect translation divergences in parallel sentence pairs, thereby helping to filter out noisy pairs.
+%Then, word similarity scores predicted by the network are used to identify and fix some divergences guiding to collect additional parallel segments.
+%We evaluate our model on multiple language pairs and data resources.
+%This model can be used on any parallel corpus without any manual annotation.
 
@@
 This problem is mostly ignored in MT, where parallel sentences are considered to convey the exact same meaning, but seems particularly important for neural MT engines, as noted in~\citet{chen2016adaptation}.
-
-Table~\ref{tab:examples} gives some examples of English-French
-parallel sentences that are not completely semantically equivalent, 
-extracted from the OpenSubtitles corpus ~\cite{LisonTiedemann2016}. 
-Divergences are outlined using bold letters.
 
@@
 
+Table~\ref{tab:examples} gives some examples of English-French
+parallel sentences that are not completely semantically equivalent, 
+extracted from the OpenSubtitles corpus ~\cite{LisonTiedemann2016}. 
+Divergences are outlined using bold letters. 
 Different types of translation divergences exist in a parallel corpus:
@@
 In this work, we present an unsupervised method for building cross-lingual sentence embeddings based on modelling word similarity. % in the form of continuous vectors. 
-The architecture of our network allows us to distinguish between different types of divergences common to many bi-texts.
+The architecture of our network allows us to distinguish between different types of divergences which are common to many existing bi-texts.
 The resulting sentence embeddings are then used to measure semantic equivalence of the corresponding sentences.
@@
 
-%As previously outlined, parallel texts are scarce resources as well as the only necessary resource to build Machine Translation systems.
-In contrast, there is an increasing amount of comparable corpora available on the Internet, a potential solution to alleviate the data scarcity problem.
-%Several approaches exist to extract parallel sentences from comparable corpora. 
-%In~\cite{J05-4003} is presented a maximum entropy classifier that can reliably determine whether a pair of sentences are translations of each other. 
-%A similar approach is proposed by~\cite{E09-1003}, where a SMT system is used to translate the source language side of a comparable corpus to find candidate sentences on the target side. 
-%Word error rate and translation error rate measures are used to decide when a candidate is finally considered parallel. 
-%More recently,~\cite{W17-2508} present a method based on seed lexical translations, simple set expansion operations and the Jaccard similarity coefficient.
-The method obtains state-of-the-art results in the task of parallel sentence extraction from comparable corpora.
-%Closely related to the method proposed in this paper, different works employ continuous vector sentence representations as a preceding step to compute sentence similarity. 
-%In~\cite{W15-1521}, word representations from bilingual word embeddings are used to build a convolutional network that predicts wether a sentence pair is aligned or not.
-%~\cite{W17-2509} propose a recurrent network to estimate sentence embeddings followed by dense layers that estimates the conditional probability distribution that two sentences are parallel.
-%These approaches are different from ours FINISH
-%This strong interest in comparable corpora was further boosted after the apparition of the BUCC workshop series on Building and Using Comparable Corpora\footnote{https://comparable.limsi.fr/bucc2018/} which helped to provide a task definition, datasets and evaluation methods to assess the state of the art.
-
 \section{Neural Similarity Classifier}
@@
 In this section we describe the architecture of our model, very much inspired by the work on Word Alignment in~\cite{W16-2207}. Figure~\ref{network} illustrates the network.
-
 In the following, we consider a source-target sentence pair $(s,t)$ with $s=(s_1,...,s_I)$ and $t=(t_1,...,t_J)$.
-The model is composed of 2 Bi-directional LSTM subnetworks, $net_s$ and $net_t$, which respectively encode source and target sentences.
-Since both $net_s$ and $net_t$ take the same form we describe only the source architecture.
-The source-sentence Bi-LSTM network outputs forward and backward hidden states, $\overrightarrow{h}^{src}_i$ and $\overleftarrow{h}^{src}_i$, which are then concatenated into a single vector encoding the $i^{th}$ word of the source sentence,
-$h^{src}_i = [ \overrightarrow{h}^{src}_i ; \overleftarrow{h}^{src}_i ]$.
-
-In addition, the last forward/backward hidden states (outlined using dark grey in Figure~\ref{network}) are also concatenated into a single vector  to represent whole sentences 
-$h_{src} = [ \overrightarrow{h}^{src}_I ; \overleftarrow{h}^{src}_1 ]$.
 
@@
 
+The model is composed of 2 Bi-directional LSTM subnetworks, $net_s$ and $net_t$, which respectively encode source and target sentences.
+Since both $net_s$ and $net_t$ take the same form we describe only the source architecture.
+The source-sentence Bi-LSTM network outputs forward and backward hidden states, $\overrightarrow{h}^{src}_i$ and $\overleftarrow{h}^{src}_i$, which are then concatenated into a single vector encoding the $i^{th}$ word of the source sentence,
+$h^{src}_i = [ \overrightarrow{h}^{src}_i ; \overleftarrow{h}^{src}_i ]$.
+In addition, the last forward/backward hidden states (outlined using dark grey in Figure~\ref{network}) are also concatenated into a single vector  to represent whole sentences 
+$h_{src} = [ \overrightarrow{h}^{src}_I ; \overleftarrow{h}^{src}_1 ]$.
 At this point a measure of similarity between sentences can be obtained by cosine similarity: 
@@
 
+In this section we evaluate the performance of our neural similarity classifier and correction algorithm detailed in Section~\ref{sec:similarity}.
+
+\subsection{Divergence Prediction}
+
 We evaluate first the ability of our similarity classifier to predict different types of divergencies at the level of words. 
@@
 
-%The test set was annotated after looking at the entire train set and looking for these type of examples. 
-%Existance and distribution of types of
+Figure~\ref{matrix} illustrates the output of our network when trained using \texttt{PU} examples (right) and \texttt{PURI} examples (left).  
+The former (right) fails to predict some word divergences, most likely because in training it never saw sentences mixing divergent and non-divergent words in the same example. 
+Furthermore, the network trained over \texttt{PURI} examples (left) correctly assigns a lower similarity score to the sentence pair as both sentences do not convey the exact same meaning.
 
@@
 
-Figure~\ref{matrix} illustrates the output of our network when trained using \texttt{PU} examples (right) and \texttt{PURI} examples (left).  
-The former (right) fails to predict some word divergences, most likely because in training it never saw sentences mixing divergent and non-divergent words in the same example. 
-Furthermore, the network trained over \texttt{PURI} examples (left) correctly assigns a lower similarity score to the sentence pair as both sentences do not convey the exact same meaning.
+\subsection{Neural Machine Translation}
 
@@
 
-Table~\ref{results_wemb} shows BLEU~\cite{P02-1040} scores of the same neural MT network when trained over different samples of the OpenSubtitles corpus.
-Samples result from filtering sentences of the original corpus using different similarity thresholds.
+Table~\ref{results_wemb} shows BLEU~\cite{P02-1040} scores of the same neural MT network when trained over different subsets of the OpenSubtitles data set.
+Subsets result from filtering sentences of the original corpus using different similarity thresholds as computed by Equation~\ref{cosine} of our similarity classifier.
+As it can be seen, the best performing subset is obtained when using sentence pairs with similarity score in the range $[0.100, +\infty)$. 
+This is, training with the most similar $15.5$ million sentences of the original data set.
 
@@
 $[0.003,+\infty)$   & 21.5 & 42.56 \\ %&  \\
-$[0.076,+\infty)$   & 18.0 & \bf 43.19 \\ %& 26.38 \\
-$[0.100,+\infty)$   & 15.5 & \\%&  \\
+$[0.076,+\infty)$   & 18.0 & 43.19 \\ %& 26.38 \\
+$[0.100,+\infty)$   & 15.5 & 43.12\\%&  \\
 \hline
@@
 
-As it can be seen, the best performing training sample is obtained when using sentence pairs with similarity score in the range $[0.076, +\infty)$. 
-This is, training with the most similar $18$ million sentences of the original data set. 
-Note that despite assigned a positive similarity score, sentence pairs in the range $[0, 0.076]$ (almost $10$ million sentences) does not seem to contribute when training a neural MT network. 
-On the contrary, when added as training data the MT accuracy is reduced.
+Note that sentences in the range $[0.076, 0.100]$ ($2.5$ million sentences) does not contribute to increase the accuracy of the resulting neural MT system.
+%The BLEU score of the system that considers these sentences in training remains unchanged.
+Furthermore, sentences in the range $[0, 0.076]$ (about $6$ million sentences) despite assigned a positive similarity score they degrade the overall performance of the network.
 
-Table~\ref{results} shows BLEU results for our neural MT engine trained over different data configurations:
-The entire\footnote{The original Paracrawl corpus contains more than $100$ million sentences.
-We reduced its size to $22.2$ millions using standard filtering techniques.} data sets (\texttt{all}) of both corpora;
-Most similar pairs after as predicted by our network (\texttt{sim}); 
-Finally, we apply the correction algorithm detailed in Section~\ref{correction} (\texttt{sim+fix}).
+%Note that despite assigned a positive similarity score, sentence pairs in the range $[0, 0.076]$ (aobut $6$ million sentences) does not seem to contribute when training a neural MT network. 
+%On the contrary, when added as training data the MT accuracy is reduced.
+
+Finally, table~\ref{results} shows BLEU results for our neural MT engine trained over different data configurations:
+\begin{itemize}
+\item The entire\footnote{The original Paracrawl corpus contains more than $100$ million sentences.
+We reduced its size to $22.2$ millions using standard filtering techniques.} data sets (\texttt{all}) of both corpora,
+\item The most similar pairs as predicted by our network (\texttt{sim}),
+\item After applying the correction algorithm detailed in Section~\ref{correction} (\texttt{sim+fix}).
+\end{itemize}
 Columns Ref and Fix indicate the number of original and corrected sentences (in millions) considered to train the neural MT system.
@@
 
-Results obtained after filtering sentence pairs by our network, \texttt{sim}, clearly outperform the baseline \texttt{all} by $+0.94$ and $+2.25$ BLEU respectively.
+Results obtained after filtering sentence pairs by our network (\texttt{sim}) clearly outperform the baseline (\texttt{all}) by $+0.94$ and $+2.25$ BLEU respectively.
 
@@
 Similar results are obtained over the Paracrawl corpus.
-Results after fixing $2.5$ million sentences ($4^{th} row$) outperform those obtained by the same same sentence pairs when used by the neural MT system in their original form ($3^{rd} row$).
+Results after fixing $2.5$ million sentences ($4^{th} row$) outperform those obtained by the same sentence pairs when used by the neural MT system in their original form ($3^{rd} row$).
 

### 2018-08-22 EMNLP/divergence.tex
start final version by JMC

--- 2018-07-17 EMNLP/divergence.tex
@@
 
-%\aclfinalcopy % Uncomment this line for the final submission
+\aclfinalcopy % Uncomment this line for the final submission
 
@@
 
-\author{%Minh Pham, \ Fran\c cois Yvon\\
-  First author\\
-  Affiliation / Address line 1\\
-  Affiliation / Address line 2\\
-  Affiliation / Address line 3\\
-  {\tt email@domain}  
-  \And
-  %Josep Crego, \ Jean Senellart\\
-  Second author\\
-  Affiliation / Address line 1\\
-  Affiliation / Address line 2\\
-  Affiliation / Address line 3\\
-  {\tt email@domain}}
+\author{MinhQuang Pham$^{\dag\ddag}$, \ \ Josep Crego$^\dag$,\ \ Jean Senellart$^\dag$ ,\ \ Fran\c cois Yvon$^\ddag$\\
+  $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France\\
+  {\tt firstname.lastname@systrangroup.com}\\
+  $^\ddag$LIMSI-CNRS / Universit\'e Paris-Saclay 91405 Orsay, France\\
+  {\tt firstname.lastname@limsi.fr}}
 
@@
 
-Corpus-based approaches to machine translation rely on the availability of parallel corpora.
-%and their performance is highly linked to the quality of the parallel data. 
-According to the process followed to compile a parallel corpus, it may contain multiple parallel sentences that are often not as parallel as one might assume. 
-%Examples are easily found when corpora is created from non-parallel texts or included by noisy document/sentence alignment tools.
+Corpus-based approaches to machine translation rely on the availability and quality of parallel corpora.
+%In the case of neural machine translation, a large neural network is trained to maximise the translation performance on a given parallel corpus. 
+Such resource is not naturally existing, and because of the process necessary to compile a parallel corpus, it may contain multiple sentence pairs that are often not as parallel as one might assume.
 This paper describes an unsupervised method for detecting translation divergences in parallel sentences.
-%We use a deep neural network to predict word alignment scores in parallel sentences. 
-We present a neural network to predict sentence similarity that minimises a loss function based on word alignments.
+We present a neural network to predict sentence similarity that minimises a loss function based on word divergence.
 We show that accurate predictions are obtained allowing divergent sentences to be filtered out.
-Furthermore, word similarity scores predicted by the network are used to identify and fix some divergences guiding to more parallel segments.
-%Misaligned words are then identified allowing divergent parallel sentences to be filtered out and in some cases making sentences parallel by removing such words. 
-We evaluate the presented method on an English-French and an English-German machine translation tasks. 
+Furthermore, word similarity scores predicted by the network are used to identify and fix some divergences guiding to collect additional parallel segments.
+We evaluate the presented method on  English-French and  English-German machine translation tasks.
 Results show that neural MT systems trained on the filtered/corrected corpus outperform the MT systems trained on the original data.
-%This method can be used on any parallel corpus without any manual annotation.
 
@@
 In the case of Neural MT, a large neural network is trained through maximising translation performance on a given parallel corpus. 
-Therefore, the quality of an MT engine is heavily dependent upon the amount and quality of parallel sentences. 
+Therefore, the quality of an MT engine is heavily dependent upon the amount but also the quality of parallel sentences.\footnote{Note that  recent work on neural MT~\cite{lample2018word,artetxe2018iclr} completely dispenses with the need of parallel data, using unsupervised methods to obtain performance improvements over word-by-word statistical MT systems.
+However, these systems still lag far behind systems trained in a supervised fashion, as considered in this work.}
+
 Unfortunately, parallel texts are scarce resources. 
-There are relatively few language pairs for which parallel corpora of reasonable sizes are available, and even for those pairs, the corpora come mostly from few domains. 
-To alleviate the lack of parallel data several approaches have been developed in the last years. 
-They range from methods using non-parallel, or comparable data ~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,W17-2509,P17-3003} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714} using in all cases automated alignment/translation engines that are prone to the introduction of noise in the resulting parallel sentences.
-Mismatches on parallel sentences extracted from translated texts are also reported ~\cite{tiedemann2011bitext,XU16.310}. 
+There are relatively few language pairs for which parallel corpora of large sizes are available, and even for those pairs, the corpora come mostly from few restricted domains. 
+To alleviate the lack of parallel data, several approaches have been developed in the last years. 
+They range from methods using non-parallel, or comparable data 
+~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,GregoireL18,P17-3003,P18-2037} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714} using in all cases automated alignment/translation engines that are prone to the introduction of noise in the resulting parallel sentences.
+Mismatches in parallel sentences extracted from translated texts are also reported ~\cite{tiedemann2011bitext,XU16.310}. 
 This problem is mostly ignored in machine translation, where parallel sentences are considered to convey the exact same meaning, and is particularly important for neural MT engines as suggested by~\cite{chen2016adaptation}.
-
-Table~\ref{tab:examples} illustrates some examples of English-French parallel sentences which are not completely semantically equivalent.
-Examples are extracted from the OpenSubtitles corpus~\cite{LisonTiedemann2016}.
-Divergences are outlined using bold letters. 
-Additional segments are included on either side of the parallel sentences (first and second rows). 
-Some translations may be completely uncorrelated (third row) or inaccurate (fourth row).
-Note that divergent translations are due to many different reasons ~\cite{C14-1055}, the study of which is beyond the scope of this paper.
 
@@
 
-In this work, we present a neural divergence classifier aimed at detecting words on either side of parallel sentences for which the corresponding meaning is not present on the translated counterpart. 
-We evaluate the classifier on an English-French and an English-German translation tasks showing that translation accuracy can be improved by filtering out divergent sentence pairs.
-In addition, we  show that some divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.
+Table~\ref{tab:examples} gives some examples of English-French parallel sentences that are not completely semantically equivalent, extracted from the OpenSubtitles corpus ~\cite{LisonTiedemann2016}. 
+Divergences are outlined using bold letters. 
+Different types of translation divergences exist in a parallel corpus:
+Additional segments are included on either side of the parallel sentences (first and second rows) most likely due to errors in sentence segmentation tools;
+Some translations may be completely uncorrelated (third row);
+Inaccurate translations also exist (fourth row). 
+Note that divergent translations can be due to many different reasons ~\cite{C14-1055}, the study of which is beyond the scope of this paper.
 
-The remainder of this paper is structured as follows. Section~\ref{related} overviews related work. 
-We describe in detail the core of the neural divergence classifier and the correction algorithm in Section~\ref{sec:divergence}. 
-Section~\ref{experiments} details the experiments conducted to measure the ability of the presented classifier to identify and fix divergent sentences.
+In this work, we present an unsupervised method for building cross-lingual sentence embeddings based on modelling word similarity. % in the form of continuous vectors. 
+The architecture of our network allows us to distinguish between different types of divergences which are common to many existing bi-texts.
+The resulting sentence embeddings are then used to measure semantic equivalence of the corresponding sentences.
+To evaluate our method we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
+Then, we  show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.
+
+The remainder of this paper is structured as follows. 
+Section~\ref{related} overviews related work. 
+We describe in detail the core of the neural divergence classifier in Section~\ref{sec:divergence}. 
+We report experiments with the presented model in Section~\ref{experiments}.
 Section~\ref{sec:results} evaluates results. 
-Finally, conclusions are drawn in Section~\ref{conclusions} while further work is outlined in Section~\ref{further}.
-All the code used in this paper as well as a human annotated test set are freely available\footnote{\texttt{https://github.com/anonymised}}.
+Finally, conclusions are drawn in Section~\ref{conclusions} and further work is outlined in Section~\ref{further}.
+All the code used in this paper is freely available\footnote{https://github.com/jmcrego/similarity}.
 
@@
 
-Attempts to measure the impact of translation divergences in MT systems have focused on the introduction of noise in sentence alignments ~\cite{goute2012}, showing that statistical MT systems are highly tolerant to noise, and that performance only degrades seriously at very high noise levels.
-In constrast, neural MT engines seem to be more sensitive~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.
+Attempts to measure the impact of translation divergences in MT systems have focused on the introduction of noise in sentence alignments~\cite{goute2012}, showing that statistical MT systems are highly tolerant to noise, and that performance only degrades seriously at very high noise levels.
+In contrast, neural MT engines seem to be more sensitive~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.
 
-Efforts have been devoted to characterise the degree of semantic equivalence between two snippets of text in the same or different languages~\cite{conf/semeval/AgirreBCDGMRW16}, a workshop devoted to an objective similar to our work.
+Efforts have been devoted to characterising the degree of semantic equivalence between two snippets of text in the same or different languages~\cite{conf/semeval/AgirreBCDGMRW16}, a workshop devoted to an objective similar to our work.
 In~\cite{Mueller:2016:SRA:3016100.3016291}, a monolingual sentence similarity network is proposed, making use of a simple LSTM layer to compute sentence representations. 
@@
 Their work shows that an NMT system trained only on non-divergent sentences yields slightly higher translation quality scores and requires clearly less training time. 
-The same authors have recently updated their work in~\cite{DBLP:journals/corr/abs-1803-11112}. 
+The same authors have recently updated their work in~\cite{N18-1136}. 
 The objective is the same as the neural network presented in~\cite{N16-1108} and their network further outperforms their previous work. 
-Our work differs from the previous as we make use of a network with different topology. 
+Our work differs from the previous as we make use of a network with a different topology. 
 We model sentence similarity by means of optimising a loss function based on word alignments. 
@@
 
-Our model can be optimised following two different objective functions:
-
-\begin{description}
-    \item[wemb]
-
-Maximise the word alignment scores between words of both sentences using aggregation functions that summarise the alignment scores for each source/target word. 
-Similar to~\cite{W16-2207} alignment scores $s(i,j)$ are given by the dot-product $S(i,j) = h_i^{src} \cdotp h_j^{tgt}$, while aggregation functions are defined as:   
+Our model is optimised through maximising word alignment scores between words of both sentences using aggregation functions that summarise the alignment scores for each source/target word. 
+Similar to~\cite{W16-2207} alignment scores $S(i,j)$ are given by the dot-product $S(i,j) = h_i^{src} \cdotp h_j^{tgt}$, and aggregation functions are defined:   
 \begin{equation}
@@
 \mathcal{L}(src,tgt) = \ \ \ \ & \\
-    \sum_{i=1}^I log(1+e&^{s_{aggr}(i,tgt) * sign_i}) \ \ +\\
- + \sum_{j=1}^J log(1+e&^{s_{aggr}(src,j) * sign_j})
+    \sum_{i=1}^I log(1+e&^{aggr_s(i,S) * sign_i}) \ \ +\\
+ + \sum_{j=1}^J log(1+e&^{aggr_t(j,S) * sign_j})
 \end{split}
@@
 
-    \item[semb] Maximise the similarity score of  both sentence embeddings as computed by Equation~\ref{cosine}. With loss function defined as: 
-%\vspace{-5mm}
-\begin{equation}
-\mathcal{L}(src,tgt) = log(1+e^{sim(src,tgt) * sign}) % where $sign(src,tgt) =1
-\label{loss_semb}
-\end{equation}
-
-\end{description}
-
 \subsection{Training with Negative Examples}
@@
 Training is performed by minimising Equation~\ref{loss_wemb}, for which examples with annotations for source $sign_i$ and target $sign_j$ words are needed.
-As positive examples we use {\bf (P)}aired sentences of a parallel corpus. All words in paired sentences are labelled as parallel ($-1$). %, $sign_i=-1$ and $sign_j=-1$. 
-As negative examples we use random {\bf (U)}npaired sentences. 
+Different kind of examples are used to train our network:
+\begin{itemize}
+
+\item As positive examples we use paired sentences of a parallel corpus. All words in paired sentences are labelled as parallel ($-1$). %, $sign_i=-1$ and $sign_j=-1$. 
+\item As negative examples we use random unpaired sentences. 
 In this case, all words are labelled as divergent ($+1$). %, $sign_i=+1$ and $sign_j=+1$. 
-Since negative pairs may be very easy to classify and we want our network to detect less obvious divergences, we further create negative examples following: % the next two methods:
+Since negative pairs may be very easy to classify and we want our network to detect less obvious divergences, we further create negative examples following the next two methods:
 
-We {\bf (R)}eplace random sequences of words by a sequence of words with same part-of-speeches. 
+\item We replace random sequences of words on either side of the sentence pair by a sequence of words with the same part-of-speeches. 
 The rationale behind this method is to keep the new sentences as grammatical as possible. 
-%Otherwise the network can learn to detect non-grammatical sentences to predict divergence.
+Otherwise, to predict divergence the network can learn to detect non-grammatical sentences.
 Words that are not replaced are considered parallel ($-1$) while those replaced are assigned the divergent label ($+1$). 
-Words aligned to any replaced word are also assigned the divergent label ($+1$). For instance, given the original sentence pair:
+Words aligned to some replaced words are also assigned the divergent label ($+1$). For instance, given the original sentence pair:
 
 \begin{table}[h]
-\small
 \center
 \begin{tabular}{ll}
-$src$: & { \small \texttt{What do you feel ?}} \\
-$tgt$: & { \small \texttt{Que ressentez-vous ?}} \\
+src: & { \small \texttt{What do you feel ?}} \\
+tgt: & { \small \texttt{Que ressentez-vous ?}} \\
 \end{tabular}
@@
 
-We may replace '\texttt{you feel}', with tags '\texttt{PRP VB}', by another with same tags (i.e. '\texttt{we want}'):
+We may replace '\texttt{you feel}', with part-of-speech tags '\texttt{PRP VB}', by another sequence with same tags (i.e. '\texttt{we want}'):
 
 \begin{table}[h]
-\small
 \center
 \begin{tabular}{ll}
-$src$: & { \small \texttt{What do {\bf we \ want} ?}} \\
-$sign_i$: & { \small \texttt{-1 \ \  -1 {\bf +1\ \ +1} \ \  -1}} \\
-$tgt$: & { \small \texttt{Que {\bf ressentez-vous} ?}} \\
-$sign_j$: & { \small \texttt{-1\ \ {\bf +1}\ \ \ \ \ \ \ \ \ -1}} \\
+src: & { \small \texttt{What do {\bf we \ want} ?}} \\
+$\mathcal{Y}^{src}$: & { \small \texttt{-1 \ \  -1 {\bf +1\ \ +1} \ \  -1}} \\
+tgt: & { \small \texttt{Que {\bf ressentez-vous} ?}} \\
+$\mathcal{Y}^{tgt}$: & { \small \texttt{-1\ \ {\bf +1}\ \ \ \ \ \ \ \ \ -1}} \\
 \end{tabular}
@@
 
-%Divergent words are shown in bold. 
-%Note that new sentences are not assured to be grammatical after replacing sequences with the same part-of-speech.
-Word alignments are used to identify '\texttt{ressentez-vous}' as divergent, since it is aligned to '\texttt{you feel}'.
+Divergent words are shown in bold. 
+Note that the new sentences are not assured to be grammatical after replacing sequences with the same part-of-speech.
+We use word alignments to identify as divergent the sequence '\texttt{ressentez-vous}' since it is aligned to '\texttt{you feel}' in the original sentence pair.
 
-Motivated by sentence segmentation errors observed in many corpora, we also build negative examples by {\bf (I)}nserting a second sentence following/preceding an %at the beginning (or end) of the source (or target) pair.
-original sentence. Only inserted words are considered divergent ($+1$).
-Following with our example, we may add the sentence '\texttt{Not .}' at the end of the original source sentence:
+\item Finally, motivated by sentence segmentation errors observed in many corpora, we also build negative examples by inserting a second sentence at the beginning (or end) of the source (or target) sentence pair.
+Words in the original sentence pair are assigned the parallel label ($-1$) while the new words inserted are considered divergent ($+1$).
+Given the original sentence pair previously shown, we may want to build the next negative example by inserting the sentence '\texttt{Not .}' at the end of the original source sentence:
 
 \begin{table}[h]
-\small
 \center
 \begin{tabular}{ll}
-$src$: & { \small \texttt{What do you want ? {\bf Not \ .}}} \\
-$sign_i$: & { \small \texttt{-1 \ \  -1 -1 \ -1  \ \ -1 {\bf +1\ \ \  +1}}} \\
-$tgt$: & { \small \texttt{Que ressentez-vous ?}} \\
-$sign_j$: & { \small \texttt{-1\ \ -1\ \ \ \ \ \ \ \ \ \ \ \ \ -1}} \\
+src: & { \small \texttt{What do you want ? {\bf Not \ .}}} \\
+$\mathcal{Y}^{src}$: & { \small \texttt{-1 \ \  -1 -1 \ -1  \ \ -1 {\bf +1\ \ \  +1}}} \\
+tgt: & { \small \texttt{Que ressentez-vous ?}} \\
+$\mathcal{Y}^{tgt}$: & { \small \texttt{-1\ \ -1\ \ \ \ \ \ \ \ \ \ \ \ \ -1}} \\
 \end{tabular}
@@
 
+\end{itemize}
+
 Replace and insert methods are applied on either side of the sentence pairs.
 To avoid that negative examples are easily predicted by using the difference in length of sentences, we restrict negative examples to have a difference in length not exceeding a ratio of $2.0$ ($3.0$ for shortest sentences).
-
-The loss function of Equation~\ref{loss_semb} uses a single label for each sentence pair. 
-Positive examples are labelled $sign=-1$ while any of the previous negative examples are labelled $sign=+1$.
 
@@
 Hence considering as parallel $s_u^v$ and $t_x^y$. 
-To find the optimal source ($u, v$) and target ($x, y$) indexes that enclose the parallel segments within the original sentence, we implement:
+To find optimal source ($u, v$) and target ($x, y$) indices that enclose parallel segments within the original sentence, we implement:
 \begin{equation*}
@@
 
-The $\mathcal{N}$-best sequences following the previous function ($s_u^v$, $t_x^y$) are considered as valid corrections, but only the highest ranked according to their similarity score is used as replacement for the original $(s_1^I, t_1^J)$.
+The $\mathcal{N}$-best sequences following the previous function ($s_u^v$, $t_x^y$) are considered as valid corrections, but only the highest ranked according to its similarity score is used as replacement for the original $(s_1^I, t_1^J)$.
 Short sentences are not considered. This is, $v - u > \tau$ and $y - x > \tau $. 
@@
 An acceptable correction is: \textit{Que ressentez-vous ? $\Leftrightarrow$ What do you feel ?}. 
-Hence, with indexes $u=1$, $v=5$, $x=1$ and $y=3$.
+Hence, with indices $u=1$, $v=5$, $x=1$ and $y=3$.
 
@@
 We filter out divergences found in the English-French OpenSubtitles corpus~\cite{LisonTiedemann2016}, which consists of a collection of movie and TV subtitles. 
-We also use the English-German Paracrawl\footnote{http://paracrawl.eu/} corpus.
+We also use the very noisy English-German Paracrawl\footnote{http://paracrawl.eu/} corpus. 
 Both corpora present many potential divergences.
 To evaluate English-French performance we use the En-Fr Microsoft Spoken Language Translation corpus, created from actual conversations over Skype~\cite{mslt-corpus-iwslt-2016-release}. 
-English-German performance is evaluated on the publicly available newstest-2017 En-De test set, corresponding to news stories selected from online sources~\cite{W17-4717}.
-In order to better assess the quality of our classifier when facing different types of word divergences we collected and annotated at the word level $500$ sentences from the original OpenSubtitles corpus, containing:
+English-German performance is evaluated on the publicly available newstest-2017 En-De test set~\cite{W17-4717}, corresponding to news stories selected from online sources.
+
+In order to better assess the quality of our classifier when facing different word divergences we collected from the original OpenSubtitles corpus $500$ sentences containing different types of examples:
 $200$ paired sentences;
@@
 $100$ sentences with insert examples as detailed in Section~\ref{negative}.
-
-%For training, our systems use a random selection ($3$ million sentence pairs) of the OpenSubtitles corpus~\cite{LisonTiedemann2016} which consists of a collection of movie and TV subtitles.
-%, aligned at the sentence level after a number of automatic preprocessing steps.
-%The corpus presents many potential divergences as outlined in Table~\ref{tab:examples}.
-
-%We evaluated the suitability of the filtering systems on the En-Fr Microsoft Spoken Language Translation task ({\bf MSLT}), created from actual conversations~\cite{mslt-corpus-iwslt-2016-release}.
-We also tested on a second domain, using the publicly available newstest-2013 En-Fr test set ({\bf NEWS}), corresponding to news stories selected from online sources~\cite{WMT:2013}.
+All data is preprocessed with \texttt{OpenNMT}\footnote{http://opennmt.net}, performing minimal tokenisation. %, basically splitting-off punctuation.
 
@@
 
-All data is preprocessed with an in-house toolkit that performs minimal tokenisation. %, basically splitting-off punctuation.
 After tokenisation, each out-of-vocabulary word is mapped to a special UNK token.
 Vocabularies consist of the $50,000$ more frequent words.
-%Our divergence classifier is described in Section~\ref{divergence}. 
-Word embeddings are initialised using \texttt{fastText}\footnote{https://github.com/facebookresearch/fastText}.
-%, further aligned by means of \texttt{MUSE}\footnote{https://github.com/facebookresearch/MUSE} following the unsupervised method detailed in~\cite{lample2018word}.
+Word embeddings are initialised using \texttt{fastText}\footnote{https://github.com/facebookresearch/fastText}, further aligned by means of \texttt{MUSE}\footnote{https://github.com/facebookresearch/MUSE} following the unsupervised method detailed in~\cite{lample2018word}.
 Size of embeddings is $E_s=E_t=256$ cells. 
@@
 We run $10$ epochs and start decaying at each epoch by $0.8$ when the loss on validation set increases. 
-Two versions of the network are evaluated. 
-The first optimises the loss shown by equation~\ref{loss_wemb}, the second optimises the loss shown by equation~\ref{loss_semb}. 
-Divergence is always computed following equation~\ref{cosine}. 
-For divergence correction, we used $\mathcal{N}=20$ and $\tau=3$.
+Divergence is computed following equation~\ref{cosine}. 
+For divergence correction, we use $\mathcal{N}=20$ and $\tau=3$.
+The same number of examples are always generated for each type of example ({\texttt P}aired, {\texttt U}npaired, {\texttt R}eplace and {\texttt I}nsert). 
+Alignments needed for {\texttt R}eplace and {\texttt I}nsert methods are performed using \texttt{fast\_align}\footnote{https://github.com/clab/fast\_align}.
 
@@
 
-In addition to the basic tokenisation detailed in Section~\ref{divergence} we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30,000$ merge operations learned from both English and French data.
-Neural MT systems are based on the open-source project \texttt{OpenNMT}\footnote{http://opennmt.net}. We use a bidirectional RNN encoder with $4$ LSTM layers with each containing $1,000$ cells. 
-Word embeddings have a size of $300$ cells.
-We set the dropout probability to $0.3$.
-Batch size is set to $64$.
-The maximum length of both source and target sentences is set to $80$.
-% and we limit the vocabulary size to $50K$ words for both source and target languages.
-The default optimiser is SGD with the starting learning rate of $1.0$.
-We start to decay the learning rate from epoch $10$ or when we detect increasing perplexity as compared to the previous epoch on a validation set.
-We stop training after $20$ epochs.
+In addition to the basic tokenisation detailed in Section~\ref{divergence} we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30000$ merge operations learned by joining both language sides.
+Neural MT systems are based on the open-source project  \texttt{OpenNMT}. 
+We use a Transformer model with same configuration than in paper~\cite{vaswani2017attention}, i.e, both encoder and decoder have $6$ layers; Multi-head attention is performed over $8$ heads. 
+Hidden layer size is $512$. 
+The inner layer of feed forward network is of size $2048$. 
+Word embeddings have $512$ cells. We set the dropout probability to $0.1$. 
+Batch size is set to $3072$.
+%The maximum length of both source and target sentences is set to $80$ and we limit the vocabulary size to $50,000$ words for both source and target languages. 
+The optimiser is Lazy Adam with $\beta_1 = 0.9$, $\beta_2 = 0.98$, $\epsilon = 10^{-9}$, $warmup\_steps = 4000$. We stop training after $30$ epochs.
 
@@
 
-%We evaluate first the ability of our similarity classifier to predict different types of divergencies. 
-%We use the test set manually annotated for that purpose and train our model on the OpenSubtitles corpus.
+We evaluate first the ability of our divergence classifier to predict different types of divergences at the level of words. 
+We use the test set manually annotated for that purpose and train our model on the OpenSubtitles corpus.
+A word is considered divergent when associated to a negative aggregation score (see Equation~\ref{aggregation}).
+Table~\ref{results_puri} shows accuracies obtained by our model when trained over different combinations of negative examples.
 
-Table~\ref{results_puri} shows accuracies obtained by our model when trained over different combinations of negative examples.
-We use the test set with manual annotations of word divergences.
-In training, the same number of examples are always generated for each type of example (P, U, R and I). 
-A word is considered divergent if it has a negative aggregation score, Equation~\ref{aggregation}.
+As it can be seen, non-divergent words in parallel and unpaired sentences (columns \texttt{P} and \texttt{U}) are easy to identify, as far as the model has seen these types of examples in training. 
+However, the accuracy drops dramatically when the model is not trained with unpaired sentences (rows  \texttt{PR}, \texttt{PI} and \texttt{PRI}).
+Regarding columns \texttt{R} and \texttt{I}, accuracies are lower since these sentences contain a mix of divergent and non-divergent words.
 
@@
 \end{tabular}
-\caption{Word divergence accuracies according to different type of examples used in train/test.}% for networks trained with different combinations of negative examples.
-P, U, R and I stand respectively for pair, unpair, replace and insert.}
+\caption{Word divergence accuracies according to different type of examples used in train/test.}
 \label{results_puri}
@@
 
-As it can be seen, non-divergent words in parallel and unpaired sentences (columns \texttt{P} and \texttt{U}) are easy to identify, as far as the model has seen these types of examples in training. 
-Regarding columns \texttt{R} and \texttt{I}, accuracies are lower since these sentences contain a mix of divergent and non-divergent words. 
-%Hence, word divergence is more difficult to predict. 
-Again, models that were trained with the corresponding examples obtain the highest accuracies (outlined in bold letters).
+Again, models that were trained with the corresponding examples (\texttt{R} and \texttt{I}) obtain the highest accuracies (outlined in bold letters).
 Column \texttt{PURI} shows results over the entire test set, mixing all type of examples. 
-The best accuracy is obtained by the system trained on all type of examples.
+As expected, the best accuracy is also obtained by the system trained on all type of  examples.
 
+Figure~\ref{matrix} illustrates the output of our network when trained using \texttt{PU} examples (right) and \texttt{PURI} examples (left).  
+The former (right) fails to predict some word divergences, most likely because in training it never saw sentences mixing divergent and non-divergent words in the same example.
+
+%\vspace{-3mm}
 \begin{figure}[h]
@@
     \includegraphics[width=0.35\linewidth]{feel_pu}    
-\caption{Sentence pair with similarity scores produced by our model when trained over PU examples (right) and over PURI examples (left).
+\caption{Sentence pair with similarity scores produced by our model when trained over \texttt{PU} examples (right) and over \texttt{PURI} examples (left).
 Aggregation scores (Equation~\ref{aggregation}) are shown next to words.
@@
 
-Figure~\ref{matrix} illustrates the output of our network when trained using PU examples (right) and PURI examples (left).  
-The former (right) fails to predict some word divergences, most likely because in training it never saw sentences mixing divergent and non-divergent words in the same example. 
-%Furthermore, the latter (right) correctly assigns a lower similarity score to the sentence pair as both sentences do not convey the exact same meaning.
+Furthermore, the network trained over \texttt{PURI} examples (left) correctly assigns a lower similarity score to the sentence pair as both sentences do not convey the exact same meaning.
 
-Table~\ref{results} shows BLEU results for our neural MT engine trained over different data configurations:
-The entire\footnote{The original Paracrawl corpus contains more than $100$ million sentences. We reduced its size to $22.2$ millions using standard filtering techniques.} data (\texttt{all}); 
-Most similar pairs after optimising equation~\ref{loss_semb} (\texttt{semb}); 
-Most similar pairs after optimising equation~\ref{loss_wemb} (\texttt{wemb}); 
-Finally, we apply the correction algorithm detailed in Section~\ref{correction} (\texttt{wemb+fix}).
-Columns Ref and Fix indicate the number of original and corrected sentences (in millions) considered to train NMT.
+Finally, table~\ref{results} shows BLEU results for our neural MT engine trained over different data configurations:
+The entire\footnote{The original Paracrawl corpus contains more than $100$ million sentences.
+We reduced its size to $22.2$ millions using standard filtering techniques.} data sets (\texttt{all}) of both corpora;
+Most similar pairs after optimising equation~\ref{loss_wemb} (\texttt{sim}); 
+Finally, we apply the correction algorithm detailed in Section~\ref{correction} (\texttt{sim+fix}).
+Columns Ref and Fix indicate the number of original and corrected sentences (in millions) considered to train our NMT system.
 
@@
 \texttt{all}                   & 27.2 & - & 42.18 \\
-\texttt{semb}            & 18.0 & - & 41.95\\
-\texttt{wemb}           & 15.5 & - & 43.12  ($+0.94$)\\
-%\texttt{wemb}           & 21.5 & - & 42.56 \\
-\texttt{wemb+fix}   & 15.5 & 2.5 & \\
-\texttt{wemb}           & 18.0 & - & 43.19  ($+1.01$)\\
-%\texttt{wemb+fix}   & 15.5 & 5.0 & \\
+%\texttt{semb}             & 18.0 & - & \\
+\texttt{sim}            & 15.5 & - & 43.12  ($+0.94$)\\
+%\texttt{sim}           & 21.5 & - & 42.56 \\
+\texttt{sim}           & 18.0 & - & 43.19  ($+1.01$)\\
+\texttt{sim+fix}     & 15.5 & 2.5 & \bf 44.19 ($+2.01$)\\
+%\texttt{sim+fix}   & 15.5 & 5.0 & \\
 \hline
@@
 \texttt{all}                  & 22.2 & - & 19.27 \\ 
-\texttt{semb}            & 15.0 & - & \\
-\texttt{wemb}           & 15.0 & - & 21.52 ($+2.25$)\\
-%\texttt{wemb}           & 12.5 &   \\
-%\texttt{wemb+fix}   & 12.5+2.5 & \\
+%\texttt{semb}            & 15.0 & - & \\
+\texttt{sim}           & 15.0 & - & 21.52 ($+2.25$)\\
+\texttt{sim}           & 17.5 & - & 21.97 ($+2.70$)\\
+\texttt{sim+fix}   & 15.0 & 2.5 & \bf 22.42 ($+3.15$) \\
 \hline
@@
 
-Results for \texttt{wemb} clearly outperform the baseline \texttt{all} ($1^{st}$ rows) by about $+1.0$ and $+2.25$ BLEU ($3^{rd}$ rows). 
-In contrast, \texttt{semb} does not improve over the baseline ($2^{nd}$ rows).
-Regarding OpenSubtitles, an additional gain of $+X.X$ is found when fixing $2.5$ million sentences ($4^{th}$ row). 
-The same $2.5$ million sentences did not show any improvement when added in their original form ($5^{th}$ row).
-
-At this moment experiments showing correction results regarding Paracrawl  are not yet finished.
+Results obtained after filtering sentence pairs by our network (\texttt{sim}) clearly outperform the baseline (\texttt{all}) by $+0.94$ and $+2.25$ BLEU respectively.
+Regarding OpenSubtitles, when fixing $2.5$ million sentences ($4^{th}$ row) the accuracy is further boosted to $+2.01$, while the same sentence pairs does not show any improvement when added in their original form ($3^{rd}$ row).
+Similar results are obtained over the Paracrawl corpus.
+Results after fixing $2.5$ million sentences ($4^{th} row$) outperform those obtained by the same sentence pairs when used by the neural MT system in their original form ($3^{rd} row$).
 
@@
 We presented an unsupervised method based on deep neural networks for detecting translation divergences in parallel sentence pairs. 
-Our model also predicts misaligned words that can then be filtered out allowing for reusing some divergent sentences.
-We evaluated our model on two neural machine translation tasks, showing improvements when compared to training over the entire data set. 
+Our model optimises word alignments. Hence, allowing for fine grained divergence prediction at the level of words. 
+Misaligned/divergent words can then be filtered out allowing for reusing some divergent sentences. 
+We evaluated our model on two neural machine translation tasks, showing significant improvements when compared to training over the entire data set. 
 %The method can be used on any parallel corpus without any manual annotation.
@@
 %We would also like to use our model to predict other types of divergences 
-We plan to use our model to predict sentence embeddings over monolingual corpora allowing to collect parallel pairs through vector similarity measures.
-%We also plan to study the use of a neural network to predict divergent words, as a replacement for the algorithm detailed in Section~\ref{correction}.
+We plan to use our model to predict sentence embeddings over monolingual corpora, allowing to collect parallel pairs through vector similarity measures.
+%We also plan to further evaluate the correction algorithm.
+In addition, we would like to measure the performance of our model after applying subword tokenisation,
+%when working with a reduced vocabulary set, as a result of applying any subword tokenisation,
+as well as using multiple LSTM layers, which is well known to capture hierarchical structure in the context of MT.
 

### 2018-08-22 EMNLP/divergence.tex
Restent deux lignes en trop

@@
   {\tt firstname.lastname@systrangroup.com}\\
-  $^\ddag$LIMSI-CNRS / Universit\'e Paris-Saclay 91405 Orsay, France\\
+  $^\ddag$LIMSI, CNRS,  Universit\'e Paris-Saclay 91405 Orsay, France\\
   {\tt firstname.lastname@limsi.fr}}
@@
 \maketitle
+
 \begin{abstract}
-
 %http://www.airccj.org/CSCP/vol4/csit42503.pdf
@@
 %In the case of neural machine translation, a large neural network is trained to maximise the translation performance on a given parallel corpus. 
-Such resource is not naturally existing, and because of the process necessary to compile a parallel corpus, it may contain multiple sentence pairs that are often not as parallel as one might assume.
-This paper describes an unsupervised method for detecting translation divergences in parallel sentences.
-We present a neural network to predict sentence similarity that minimises a loss function based on word divergence.
-We show that accurate predictions are obtained allowing divergent sentences to be filtered out.
-Furthermore, word similarity scores predicted by the network are used to identify and fix some divergences guiding to collect additional parallel segments.
-We evaluate the presented method on  English-French and  English-German machine translation tasks.
-Results show that neural MT systems trained on the filtered/corrected corpus outperform the MT systems trained on the original data.
+Such resources are scarce, and because of the processes involved in their preparation, they may contain sentence pairs that are not as parallel as one would expect.
+This paper describes an unsupervised method for detecting such translation divergences in parallel sentences. We rely on a  neural network trained
+to predict sentence similarity that minimises a loss function based on word divergence. Its predictions are sufficiently accurate to effectively filtered out divergent translations.
+Furthermore, word similarity scores predicted by the network are used to identify and fix some partial divergences, yielding additional parallel segments.
+We evaluate these methods for  English-French and  English-German machine translation tasks: 
+neural MT systems trained on the filtered/corrected corpora outperform the MT systems trained on the original data.
 
@@
 Parallel sentence pairs are the only necessary resource to build Machine Translation (MT) systems. 
-In the case of Neural MT, a large neural network is trained through maximising translation performance on a given parallel corpus. 
-Therefore, the quality of an MT engine is heavily dependent upon the amount but also the quality of parallel sentences.\footnote{Note that  recent work on neural MT~\cite{lample2018word,artetxe2018iclr} completely dispenses with the need of parallel data, using unsupervised methods to obtain performance improvements over word-by-word statistical MT systems.
-However, these systems still lag far behind systems trained in a supervised fashion, as considered in this work.}
+In the case of neural MT, a large neural network is trained through maximising a proxy of translation performance on a parallel corpus. 
+Therefore, the quality of MT engines is heavily dependent on the amount but also the quality of available parallel sentences.\footnote{Recent work on neural MT~\cite{lample2018word,artetxe2018iclr} completely dispenses with parallel data, using unsupervised methods to obtain performance improvements over word-by-word statistical MT.
+These systems however lag far behind supervised systems, as considered in this work.}
 
-Unfortunately, parallel texts are scarce resources. 
-There are relatively few language pairs for which parallel corpora of large sizes are available, and even for those pairs, the corpora come mostly from few restricted domains. 
-To alleviate the lack of parallel data, several approaches have been developed in the last years. 
+Parallel texts are unfortunately, scarce resources:
+There are relatively few language pairs for which parallel corpora of large sizes exist, and even for those pairs, available corpora only concern few restricted domains. 
+To alleviate the lack of parallel data, several approaches have been developed over the years. 
 They range from methods using non-parallel, or comparable data 
-~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,GregoireL18,P17-3003,P18-2037} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714} using in all cases automated alignment/translation engines that are prone to the introduction of noise in the resulting parallel sentences.
+~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,GregoireL18,P17-3003,P18-2037} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714}, using automated alignment/translation engines that are prone to the introduction of noise in the resulting parallel sentences.
 Mismatches in parallel sentences extracted from translated texts are also reported ~\cite{tiedemann2011bitext,XU16.310}. 
-This problem is mostly ignored in machine translation, where parallel sentences are considered to convey the exact same meaning, and is particularly important for neural MT engines as suggested by~\cite{chen2016adaptation}.
+This problem is mostly ignored in MT, where parallel sentences are considered to convey the exact same meaning; yet it seems particularly important for neural MT engines~\cite{chen2016adaptation}.
 
@@
 \end{tabular}
-\caption[Table caption text]{Examples of semantically divergent parallel sentences. English (\texttt{en}), French (\texttt{fr}) and gloss of French (\texttt{gl}). }
+\caption[Table caption text]{Examples of semantically divergent parallel sentences. English (\texttt{en}), French (\texttt{fr}) and gloss of French (\texttt{gl}). Divergences are in bold letters. }
 \label{tab:examples}
@@
 Table~\ref{tab:examples} gives some examples of English-French parallel sentences that are not completely semantically equivalent, extracted from the OpenSubtitles corpus ~\cite{LisonTiedemann2016}.
-Divergences are outlined using bold letters. 
-Different types of translation divergences exist in a parallel corpus:
-Additional segments are included on either side of the parallel sentences (first and second rows) most likely due to errors in sentence segmentation tools;
+
+Multiples types of translation divergences are found in parallel corpora:
+Additional segments are included on either side of the parallel sentences (first and second rows) most likely due to errors in sentence segmentation;
 Some translations may be completely uncorrelated (third row);
 Inaccurate translations also exist (fourth row). 
-Note that divergent translations can be due to many different reasons ~\cite{C14-1055}, the study of which is beyond the scope of this paper.
+Note that divergent translations can be due various reasons~\cite{C14-1055}, the study of which is beyond the scope of this paper.
 
-In this work, we present an unsupervised method for building cross-lingual sentence embeddings based on modelling word similarity. % in the form of continuous vectors. 
-The architecture of our network allows us to distinguish between different types of divergences which are common to many existing bi-texts.
-The resulting sentence embeddings are then used to measure semantic equivalence of the corresponding sentences.
-To evaluate our method we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
-Then, we  show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.
+In this work, we present an unsupervised method for building cross-lingual sentence embeddings based on modelling word similarity,  % in the form of continuous vectors. 
+relying on a neural architecture (see \textsection~\ref{sec:divergence}) that is able to identify several types of common cross-lingual divergences. % which are common to many existing bi-texts.
+The resulting embeddings are then used to measure semantic equivalence between sentences.
+To evaluate our method, we show in \textsection~\ref{experiments} that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
+We also show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.
+All the code used in this paper is freely available.\footnote{https://github.com/jmcrego/similarity}
 
-The remainder of this paper is structured as follows. 
-Section~\ref{related} overviews related work. 
-We describe in detail the core of the neural divergence classifier in Section~\ref{sec:divergence}. 
-We report experiments with the presented model in Section~\ref{experiments}.
-Section~\ref{sec:results} evaluates results. 
-Finally, conclusions are drawn in Section~\ref{conclusions} and further work is outlined in Section~\ref{further}.
-All the code used in this paper is freely available\footnote{https://github.com/jmcrego/similarity}.
+% Useless for a short
+% The remainder of this paper is structured as follows. 
+% Section~\ref{related} overviews related work. 
+% We describe in detail the core of the neural divergence classifier in Section. 
+% We report experiments with the presented model in Section
+% Section~\ref{sec:results} evaluates results. 
+% Finally, conclusions are drawn in Section~\ref{conclusions} and further work is outlined in Section~\ref{further}.
 
@@
 
-Attempts to measure the impact of translation divergences in MT systems have focused on the introduction of noise in sentence alignments~\cite{goute2012}, showing that statistical MT systems are highly tolerant to noise, and that performance only degrades seriously at very high noise levels.
-In contrast, neural MT engines seem to be more sensitive~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.
+Attempts to measure the impact of translation divergences in MT have focused on the introduction of noise in sentence alignments~\cite{goute2012}, showing that statistical MT is highly robust to noise, and that performance only degrades seriously at very high noise levels.
+In contrast, neural MTs seem to be more sensitive to noise~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.
 
-Efforts have been devoted to characterising the degree of semantic equivalence between two snippets of text in the same or different languages~\cite{conf/semeval/AgirreBCDGMRW16}, a workshop devoted to an objective similar to our work.
+Efforts devoted to characterising the degree of semantic equivalence between two snippets of texts in the same or different languages are presented~\cite{conf/semeval/AgirreBCDGMRW16}.
+% a workshop devoted to an objective similar to our work.
 In~\cite{Mueller:2016:SRA:3016100.3016291}, a monolingual sentence similarity network is proposed, making use of a simple LSTM layer to compute sentence representations. 
-The authors show that a simple SVM classifier can be built on top of the sentence representations to achieve state-of-the-art results in a semantic entailment classification task. 
-With the same objective, the system presented in~\cite{N16-1108} uses multiple convolutional layers and models pairwise word interactions.
+The authors show that a simple SVM classifier exploiting such sentence representations achieves state-of-the-art results in a textual entailment task. 
+With the same objective, the system of~\citet{N16-1108} uses multiple convolutional layers and models pairwise word interactions.
 %Our work differs from the previous as we build a different network and use it to measure similarity of sentences in different languages.
 
-Our work is inspired by~\cite{W17-3209} where the authors train a cross-lingual divergence detector using word alignments and sentence length features to train a linear SVM classifier. 
-Their work shows that an NMT system trained only on non-divergent sentences yields slightly higher translation quality scores and requires clearly less training time. 
-The same authors have recently updated their work in~\cite{N18-1136}. 
-The objective is the same as the neural network presented in~\cite{N16-1108} and their network further outperforms their previous work. 
-Our work differs from the previous as we make use of a network with a different topology. 
+Our work is inspired by~\citet{W17-3209}, who train a SVM-based cross-lingual divergence detector using word alignments and sentence length features. 
+Their work shows that an NMT system trained only on non-divergent sentences yields slightly better translation scores, while requiring less training time. 
+A follow-up study by the same authors~\cite{N18-1136} achieves even better results, using the neural architecture of \citet{N16-1108}.
+Our work differs from theirs as we make use of a network with a different, arguably simpler, topology. 
 We model sentence similarity by means of optimising a loss function based on word alignments. 
@@
 
-The architecture of our model is inspired by the work on Word Alignment in~\cite{W16-2207}. Figure~\ref{network} illustrates the model.
-
+The architecture of our network is inspired by the work on word alignment of~\citet{W16-2207}, using however contextual, rather than fixed, word embeddings (see Figure~\ref{network}).
 \begin{figure}[h]
@@
 
-In the following, we consider a source-target sentence pair $(s,t)$ with $s=(s_1,...,s_I)$ and $t=(t_1,...,t_J)$. 
-The model is composed of 2 Bi-directional LSTM subnetworks, $net_s$ and $net_t$, which respectively encode source and target sentences. 
-Since both $net_s$ and $net_t$ take the same form we describe only the source architecture.
-
-The source-sentence Bi-LSTM network outputs forward and backward hidden states, $\overrightarrow{h}^{src}_i$ and $\overleftarrow{h}^{src}_i$, which are then concatenated into a single vector encoding the $i^{th}$ word of the source sentence,
-$h^{src}_i = [ \overrightarrow{h}^{src}_i ; \overleftarrow{h}^{src}_i ]$.
-In addition, the last forward/backward hidden states (outlined using dark grey in Figure~\ref{network}) are also concatenated into a single vector  to represent whole sentences 
+It computes the similarity of any source-target sentence pair $(s,t)$, where $s=(s_1,...,s_I)$ and $t=(t_1,...,t_J)$. 
+The model is composed of 2~bi-directional LSTM subnetworks, $net_s$ and $net_t$, which respectively encode source and target sentences. 
+Since both $net_s$ and $net_t$ take the same form, we only describe the former network: 
+it outputs forward and backward hidden states, $\overrightarrow{h}^{src}_i$ and $\overleftarrow{h}^{src}_i$, which are then concatenated into a vector encoding the $i^{th}$ source word as $h^{src}_i = [ \overrightarrow{h}^{src}_i ; \overleftarrow{h}^{src}_i ]$.
+In addition, the last forward/backward hidden states (in dark grey on Figure~\ref{network}) are also concatenated  to represent whole sentences 
 $h_{src} = [ \overrightarrow{h}^{src}_I ; \overleftarrow{h}^{src}_1 ]$.
-At this point, a measure of similarity between sentence pairs can be obtained by cosine similarity:
+The similarity between sentence pairs can then be obtained using eg.\ the cosine similarity:
 \begin{equation}
@@
 
-Our model is optimised through maximising word alignment scores between words of both sentences using aggregation functions that summarise the alignment scores for each source/target word. 
-Similar to~\cite{W16-2207} alignment scores $S(i,j)$ are given by the dot-product $S(i,j) = h_i^{src} \cdotp h_j^{tgt}$, and aggregation functions are defined:   
+Our model is trained to maximize word alignment scores between words in both sentences, using aggregation functions that summarise the alignment scores for each source/target word. 
+Similar to~\cite{W16-2207}, alignment scores $S(i,j)$ are given by the dot-product $S(i,j) = h_i^{src} \cdotp h_j^{tgt}$, further aggregated as follows:
 \begin{equation}
@@
 
-The loss function is defined as:
+The training loss function is then defined as:
 %\vspace{-8mm}
@@
 
-Training is performed by minimising Equation~\ref{loss_wemb}, for which examples with annotations for source $sign_i$ and target $sign_j$ words are needed.
-Different kind of examples are used to train our network:
-\begin{itemize}
+Training is performed by minimizing Eq.~\eqref{loss_wemb}, for which annotated examples of source ($\operatorname{sign}_i$) and target ($\operatorname{sign_j}$) words are needed.
+As positive examples, we use paired sentences of a parallel corpus; all words in such sentences are labelled as parallel ($ \forall i, j, \operatorname{sign}_i=\operatorname{sign}_j=-1$).
+%, $sign_i=-1$ and $sign_j=-1$.
+We consider three types of negative instances: the basic case uses random unpaired sentences; in this case, all words are labelled as divergent ($ \forall i, j, \operatorname{sign}_i=\operatorname{sign}_j=+1.$).
+%, $sign_i=+1$ and $sign_j=+1$.
+Since negative pairs may be very easy to classify and we want our network to detect less obvious divergences, we further create more difficult negative examples as follows.
 
-\item As positive examples we use paired sentences of a parallel corpus. All words in paired sentences are labelled as parallel ($-1$). %, $sign_i=-1$ and $sign_j=-1$. 
-\item As negative examples we use random unpaired sentences. 
-In this case, all words are labelled as divergent ($+1$). %, $sign_i=+1$ and $sign_j=+1$. 
-Since negative pairs may be very easy to classify and we want our network to detect less obvious divergences, we further create negative examples following the next two methods:
-
-\item We replace random sequences of words on either side of the sentence pair by a sequence of words with the same part-of-speeches. 
-The rationale behind this method is to keep the new sentences as grammatical as possible. 
-Otherwise, to predict divergence the network can learn to detect non-grammatical sentences.
-Words that are not replaced are considered parallel ($-1$) while those replaced are assigned the divergent label ($+1$). 
-Words aligned to some replaced words are also assigned the divergent label ($+1$). For instance, given the original sentence pair:
+We first replace random sequences of words in source or target by a sequence of words with the same part-of-speeches.\footnote{The rationale is to try to keep the generated sentences as grammatical as possible;
+Otherwise, the network could learn to flag non-grammatical sentences as non-parallel.}
+ Words that are not replaced are deemed parallel ($\operatorname{sign}_i=-1$) while those replaced are annotated as $\operatorname{sign}_i=+1$. 
+Words aligned to some replaced words are also assigned the divergent label ($\operatorname{sign}_i=+1$). For instance, given the original sentence pair:
 
@@
 tgt: & { \small \texttt{Que ressentez-vous ?}} \\
-\end{tabular}
+\end{tabular},
 \end{table}
 
-We may replace '\texttt{you feel}', with part-of-speech tags '\texttt{PRP VB}', by another sequence with same tags (i.e. '\texttt{we want}'):
-
+\noindent we may replace '\texttt{you feel}', with part-of-speech tags '\texttt{PRP VB}', by another sequence with same tags (i.e.
+'\texttt{we want}'), yielding a new negative instance (divergent words are in bold):
 \begin{table}[h]
@@
 
-Divergent words are shown in bold. 
-Note that the new sentences are not assured to be grammatical after replacing sequences with the same part-of-speech.
-We use word alignments to identify as divergent the sequence '\texttt{ressentez-vous}' since it is aligned to '\texttt{you feel}' in the original sentence pair.
-
-\item Finally, motivated by sentence segmentation errors observed in many corpora, we also build negative examples by inserting a second sentence at the beginning (or end) of the source (or target) sentence pair.
-Words in the original sentence pair are assigned the parallel label ($-1$) while the new words inserted are considered divergent ($+1$).
-Given the original sentence pair previously shown, we may want to build the next negative example by inserting the sentence '\texttt{Not .}' at the end of the original source sentence:
+% Note that the new sentences are not assured to be grammatical after replacing sequences with the same part-of-speech.
+Note that we need word alignments to identify as divergent the sequence '\texttt{ressentez-vous}', which was aligned to '\texttt{you feel}' in the original sentence.
+Finally, motivated by sentence segmentation errors observed in many corpora, we also build negative examples by inserting a sentence at the beginning (or end) of the source (or target) sentence. 
+Words in the original sentence pair are annotated $\operatorname{sign}_i=-1$, while the new words inserted are considered divergent ($\operatorname{sign}_i=+1$).
+Given the same sentence pair as above, a negative example is created by inserting the sentence '\texttt{Not .}' at the end of the original source:
 
@@
 
-\end{itemize}
-
-Replace and insert methods are applied on either side of the sentence pairs.
-To avoid that negative examples are easily predicted by using the difference in length of sentences, we restrict negative examples to have a difference in length not exceeding a ratio of $2.0$ ($3.0$ for shortest sentences).
+% Replace and insert methods are applied on either side of the sentence pairs. 
+To finally avoid the generation of easy negative sentence pairs having a large difference in sentence length, we restrict negative examples to have a length ratio $< 2.0$ ($3.0$ for shortest sentences).
 
@@
 
-We observed in our training corpora that many divergent sentences follow a common pattern, consisting of adding some extra leading/trailing words.
+Our training corpora contains many divergent sentences that follow a common pattern, consisting of adding some extra leading/trailing words.
 % (as in the first and second examples of table~\ref{tab:examples}).
 Accordingly, we implemented a simple algorithm that discards sequences of leading/trailing words on both sides. 
-Hence considering as parallel $s_u^v$ and $t_x^y$. 
-To find optimal source ($u, v$) and target ($x, y$) indices that enclose parallel segments within the original sentence, we implement:
+% Hence considering as parallel $s_u^v$ and $t_x^y$. 
+To find optimal source ($u, v$) and target ($x, y$) indices that enclose parallel segments within the original sentence, we compute:
 \begin{equation*}
@@
 \end{equation*}
-
-The $\mathcal{N}$-best sequences following the previous function ($s_u^v$, $t_x^y$) are considered as valid corrections, but only the highest ranked according to its similarity score is used as replacement for the original $(s_1^I, t_1^J)$.
-Short sentences are not considered. This is, $v - u > \tau$ and $y - x > \tau $. 
-Figure~\ref{matrix} (left) shows an example of an alignment matrix $S(i,j)$ for a given sentence pair. 
+The $\mathcal{N}$-best sequences ($s_u^v$, $t_x^y$) are considered as likely corrections, in which we use the one having the highest similarity score to replace the original $(s_1^I, t_1^J)$.
+Note that short sentences are not considered and we enforce $v - u > \tau$ and $y - x > \tau $. 
+Figure~\ref{matrix} (left) displays an example of an alignment matrix $S(i,j)$.
 An acceptable correction is: \textit{Que ressentez-vous ? $\Leftrightarrow$ What do you feel ?}. 
-Hence, with indices $u=1$, $v=5$, $x=1$ and $y=3$.
+corresponding to $u=1$, $v=5$, $x=1$ and $y=3$.
 
@@
 
-We filter out divergences found in the English-French OpenSubtitles corpus~\cite{LisonTiedemann2016}, which consists of a collection of movie and TV subtitles. 
-We also use the very noisy English-German Paracrawl\footnote{http://paracrawl.eu/} corpus. 
-Both corpora present many potential divergences.
-To evaluate English-French performance we use the En-Fr Microsoft Spoken Language Translation corpus, created from actual conversations over Skype~\cite{mslt-corpus-iwslt-2016-release}. 
-English-German performance is evaluated on the publicly available newstest-2017 En-De test set~\cite{W17-4717}, corresponding to news stories selected from online sources.
+We filter out divergences from the English-French OpenSubtitles corpus~\cite{LisonTiedemann2016}, which consists of a collection of movie and TV subtitles. 
+We also use the very noisy English-German Paracrawl\footnote{http://paracrawl.eu/} corpus. Both corpora present many potential divergences.
+To evaluate English-French translation performance, we use the En-Fr Microsoft Spoken Language Translation corpus, created from actual Skype conversations~\cite{mslt-corpus-iwslt-2016-release}. 
+English-German performance is evaluated on the publicly available Newstest-2017~\cite{W17-4717}, corresponding to news stories selected from online sources.
 
-In order to better assess the quality of our classifier when facing different word divergences we collected from the original OpenSubtitles corpus $500$ sentences containing different types of examples:
+In order to better assess the quality of our classifier when facing different word divergences, we also collected from the original OpenSubtitles corpus $500$ sentences containing different types of examples:
 $200$ paired sentences;
@@
 $100$ sentences with replace examples; and
-$100$ sentences with insert examples as detailed in Section~\ref{negative}.
+$100$ sentences with insert examples (see \textsection~\ref{negative}).
 All data is preprocessed with \texttt{OpenNMT}\footnote{http://opennmt.net}, performing minimal tokenisation. %, basically splitting-off punctuation.
+After tokenisation, each out-of-vocabulary word is mapped to a special UNK token, assuming a vocabulary containing the $50,000$ more frequent words.
 
@@
 
-After tokenisation, each out-of-vocabulary word is mapped to a special UNK token.
-Vocabularies consist of the $50,000$ more frequent words.
-Word embeddings are initialised using \texttt{fastText}\footnote{https://github.com/facebookresearch/fastText}, further aligned by means of \texttt{MUSE}\footnote{https://github.com/facebookresearch/MUSE} following the unsupervised method detailed in~\cite{lample2018word}.
-Size of embeddings is $E_s=E_t=256$ cells. 
-Both Bi-LSTM use $256$-dimensional hidden representations ($E=512$).
-We use $r=1.0$. 
-Network optimization is done using the SGD method along with gradient clipping~\cite{Pascanu:2013:DTR:3042817.3043083}. 
-For each epoch we randomly select $1$ million sentence pairs that we place in batches of $32$ examples.  
+Word embeddings of $E_s=E_t=256$ cells are initialised using \texttt{fastText},\footnote{https://github.com/facebookresearch/fastText} further aligned by means of \texttt{MUSE}\footnote{https://github.com/facebookresearch/MUSE} following the unsupervised method of~\citet{lample2018word}.
+Both bi-LSTMs use $256$-dimensional hidden representations ($E=512$).
+Network optimization is done using SGD with gradient clipping~\cite{Pascanu:2013:DTR:3042817.3043083}. 
+For each epoch, we randomly select $1$ million sentence pairs that we place in batches of $32$ examples.  
 We run $10$ epochs and start decaying at each epoch by $0.8$ when the loss on validation set increases. 
-Divergence is computed following equation~\ref{cosine}. 
+Divergence is computed as in equation~\eqref{cosine} and setting $r=1.0$ ; 
 For divergence correction, we use $\mathcal{N}=20$ and $\tau=3$.
@@
 
-In addition to the basic tokenisation detailed in Section~\ref{divergence} we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30000$ merge operations learned by joining both language sides.
-Neural MT systems are based on the open-source project  \texttt{OpenNMT}. 
-We use a Transformer model with same configuration than in paper~\cite{vaswani2017attention}, i.e, both encoder and decoder have $6$ layers; Multi-head attention is performed over $8$ heads. 
-Hidden layer size is $512$. 
-The inner layer of feed forward network is of size $2048$. 
-Word embeddings have $512$ cells. We set the dropout probability to $0.1$. 
-Batch size is set to $3072$.
+In addition to the basic tokenisation detailed above, we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30000$ merge operations learned by joining both language sides.
+Neural systems are based on the open-source project  \texttt{OpenNMT};  using a Transformer model similar to the model of~\citet{vaswani2017attention}: both encoder and decoder have $6$ layers; Multi-head attention is performed over $8$ head;
+the hidden layer size is $512$; and the inner layer of feed forward network is of size $2048$. 
+Word embeddings have $512$ cells. We set the dropout probability to $0.1$ and the batch size to $3072$.
 %The maximum length of both source and target sentences is set to $80$ and we limit the vocabulary size to $50,000$ words for both source and target languages. 
-The optimiser is Lazy Adam with $\beta_1 = 0.9$, $\beta_2 = 0.98$, $\epsilon = 10^{-9}$, $warmup\_steps = 4000$. We stop training after $30$ epochs.
+The optimiser is Lazy Adam with $\beta_1 = 0.9$, $\beta_2 = 0.98$, $\epsilon = 10^{-9}$, $warmup\_steps = 4000$. Training stops after $30$ epochs.
 
@@
 
-We evaluate first the ability of our divergence classifier to predict different types of divergences at the level of words. 
+We first evaluate the ability of our divergence classifier to predict different types of divergences at the level of words. 
 We use the test set manually annotated for that purpose and train our model on the OpenSubtitles corpus.
-A word is considered divergent when associated to a negative aggregation score (see Equation~\ref{aggregation}).
-Table~\ref{results_puri} shows accuracies obtained by our model when trained over different combinations of negative examples.
-
-As it can be seen, non-divergent words in parallel and unpaired sentences (columns \texttt{P} and \texttt{U}) are easy to identify, as far as the model has seen these types of examples in training. 
+A word is considered divergent when associated to a negative aggregation score (see Equation~\eqref{aggregation}).
+Accuracies obtained for various combinations of negative examples, where we see that
+non-divergent words in parallel and unpaired sentences (columns \texttt{P} and \texttt{U}) are easy to spot, as long as the model has seen these types of examples in training. 
 However, the accuracy drops dramatically when the model is not trained with unpaired sentences (rows  \texttt{PR}, \texttt{PI} and \texttt{PRI}).
@@
 
-Again, models that were trained with the corresponding examples (\texttt{R} and \texttt{I}) obtain the highest accuracies (outlined in bold letters).
-Column \texttt{PURI} shows results over the entire test set, mixing all type of examples. 
-As expected, the best accuracy is also obtained by the system trained on all type of  examples.
+Models that were trained with the matching examples (\texttt{R} and \texttt{I}) obtain the highest accuracies (in bold letters).
+Column \texttt{PURI} gives results for the complete test set, mixing all type of examples. 
+As expected, the best accuracy is also obtained when training on all types of  examples.
 
 Figure~\ref{matrix} illustrates the output of our network when trained using \texttt{PU} examples (right) and \texttt{PURI} examples (left).  
-The former (right) fails to predict some word divergences, most likely because in training it never saw sentences mixing divergent and non-divergent words in the same example.
+The former (right) fails to predict some divergences, most likely because its training set does not contain sentences mixing divergent and non-divergent words. 
+Furthermore, the network trained with \texttt{PURI} examples correctly assigns a lower similarity score to this pair, as both sentences do not convey the exact same meaning.
 
@@
 
-Furthermore, the network trained over \texttt{PURI} examples (left) correctly assigns a lower similarity score to the sentence pair as both sentences do not convey the exact same meaning.
-
-Finally, table~\ref{results} shows BLEU results for our neural MT engine trained over different data configurations:
-The entire\footnote{The original Paracrawl corpus contains more than $100$ million sentences.
-We reduced its size to $22.2$ millions using standard filtering techniques.} data sets (\texttt{all}) of both corpora;
-Most similar pairs after optimising equation~\ref{loss_wemb} (\texttt{sim}); 
-Finally, we apply the correction algorithm detailed in Section~\ref{correction} (\texttt{sim+fix}).
-Columns Ref and Fix indicate the number of original and corrected sentences (in millions) considered to train our NMT system.
+Finally, table~\ref{results} gives BLEU results when training our MT with varying data configurations:
+The entire\footnote{Paracrawl contains more than $100$M sentences. We reduced its size to $22.2$M using standard filtering techniques.} data sets (\texttt{all}) of both corpora; 
+The most similar pairs after optimizing equation~\eqref{loss_wemb} (\texttt{sim}); 
+After applying the correction algorithm of Section~\ref{correction} (\texttt{sim+fix}).
+Columns Ref and Fix indicate the number of original and corrected sentences (in millions) used in training.
 
@@
 
-Results obtained after filtering sentence pairs by our network (\texttt{sim}) clearly outperform the baseline (\texttt{all}) by $+0.94$ and $+2.25$ BLEU respectively.
-Regarding OpenSubtitles, when fixing $2.5$ million sentences ($4^{th}$ row) the accuracy is further boosted to $+2.01$, while the same sentence pairs does not show any improvement when added in their original form ($3^{rd}$ row).
-Similar results are obtained over the Paracrawl corpus.
-Results after fixing $2.5$ million sentences ($4^{th} row$) outperform those obtained by the same sentence pairs when used by the neural MT system in their original form ($3^{rd} row$).
+Results obtained after filtering sentence pairs (\texttt{sim}) clearly outperform the baseline (\texttt{all}) by $+0.94$ and $+2.25$ BLEU respectively.
+Regarding OpenSubtitles, when fixing $2.5$M sentences ($4^{th}$ row) the accuracy is further boosted to $+2.01$, whereas the same sentence pairs do not show any improvement when added in their original form ($3^{rd}$ row).
+Similar results are obtained for the Paracrawl corpus. Results after fixing $2.5$M sentences ($4^{th} row$) outperform those obtained with their original form ($3^{rd} row$).
 
-\section{Conclusions}
-\label{conclusions}
+\section{Conclusions and outlook}
+\label{conclusions+further}
 
@@
 %The method can be used on any parallel corpus without any manual annotation.
-
-\section{Further Work}
-\label{further}
 
@@
 %when working with a reduced vocabulary set, as a result of applying any subword tokenisation,
-as well as using multiple LSTM layers, which is well known to capture hierarchical structure in the context of MT.
+as well as using multiple LSTM layers, a technique well known to capture hierarchical structure in the context of MT.
 
-%\section*{Acknowledgements}
-%We are grateful to Jo\"el Legrand for his fruitful comments when building the neural divergence classifier.
+\section*{Acknowledgements}
+We are grateful to J. Legrand for his fruitful comments regarding the neural divergence classifier.
 

### 2018-08-23 2018-EMNLP/divergence.tex
Follow up remaming

--- 2018-08-22 EMNLP/divergence.tex
@@
 
-Corpus-based approaches to machine translation rely on the availability and quality of parallel corpora.
+Corpus-based approaches to machine translation rely on the availability of clean parallel corpora.
 %In the case of neural machine translation, a large neural network is trained to maximise the translation performance on a given parallel corpus. 
-Such resources are scarce, and because of the processes involved in their preparation, they may contain sentence pairs that are not as parallel as one would expect.
-This paper describes an unsupervised method for detecting such translation divergences in parallel sentences. We rely on a  neural network trained
-to predict sentence similarity that minimises a loss function based on word divergence. Its predictions are sufficiently accurate to effectively filtered out divergent translations.
-Furthermore, word similarity scores predicted by the network are used to identify and fix some partial divergences, yielding additional parallel segments.
-We evaluate these methods for  English-French and  English-German machine translation tasks: 
-neural MT systems trained on the filtered/corrected corpora outperform the MT systems trained on the original data.
+Such resources are scarce, and because of the automatic processes involved in their preparation, they are often noisy. % may contain sentence pairs that are not as parallel as one would expect.
+This paper describes an unsupervised method for detecting translation divergences in parallel sentences.
+We rely on a neural network that computes cross-lingual sentence similarity scores, which are then used to effectively filter out divergent translations.
+Furthermore, similarity scores predicted by the network are used to identify and fix some partial divergences, yielding additional parallel segments.
+We evaluate these methods for English-French and English-German machine translation tasks,  and show that using filtered/corrected corpora actually improves MT performance.
 
@@
 In this work, we present an unsupervised method for building cross-lingual sentence embeddings based on modelling word similarity,  % in the form of continuous vectors. 
-relying on a neural architecture (see \textsection~\ref{sec:divergence}) that is able to identify several types of common cross-lingual divergences. % which are common to many existing bi-texts.
+relying on a neural architecture (see \textsection~\ref{sec:divergence}) that is able to identify several types of common cross-lingual divergences. 
+% which are common to many existing bi-texts.
 The resulting embeddings are then used to measure semantic equivalence between sentences.
 To evaluate our method, we show in \textsection~\ref{experiments} that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
-We also show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.
+We also show that in some cases, divergent sentences can be fixed by removing divergent segments, further increasing translation quality.
 All the code used in this paper is freely available.\footnote{https://github.com/jmcrego/similarity}
@@
 \end{equation*}
-The $\mathcal{N}$-best sequences ($s_u^v$, $t_x^y$) are considered as likely corrections, in which we use the one having the highest similarity score to replace the original $(s_1^I, t_1^J)$.
+The $\mathcal{N}$-best sequences ($s_u^v$, $t_x^y$) are considered as likely corrections, in which we select the one having the highest sentence-level similarity to replace the original $(s_1^I, t_1^J)$.
 Note that short sentences are not considered and we enforce $v - u > \tau$ and $y - x > \tau $. 
@@
     \includegraphics[width=0.35\linewidth]{feel_pu}    
-\caption{Sentence pair with similarity scores produced by our model when trained over \texttt{PU} examples (right) and over \texttt{PURI} examples (left).
-Aggregation scores (Equation~\ref{aggregation}) are shown next to words.
+\caption{Sentence pair with similarity scores produced by our model when trained with \texttt{PU} examples (right) and over \texttt{PURI} examples (left).
+Aggregation scores (Eq.~\eqref{aggregation}) are shown next to words.
 Matrices contain alignment scores.
-Sentence similarities (Equation~\ref{cosine})  shown below matrices.}
+Sentence similarities (Eq.~\eqref{cosine}) are below matrices.}
 \label{matrix}
@@
 
-Finally, table~\ref{results} gives BLEU results when training our MT with varying data configurations:
-The entire\footnote{Paracrawl contains more than $100$M sentences. We reduced its size to $22.2$M using standard filtering techniques.} data sets (\texttt{all}) of both corpora; 
-The most similar pairs after optimizing equation~\eqref{loss_wemb} (\texttt{sim}); 
-After applying the correction algorithm of Section~\ref{correction} (\texttt{sim+fix}).
+Finally, BLEU scores obtained with varying training data configurations are in Table~\ref{results}:
+The entire\footnote{Paracrawl contains more than $100$M sentences. We reduced its size to $22.2$M using standard filtering techniques.} data sets (\texttt{all}); 
+The most similar pairs after optimizing Eq.~\eqref{loss_wemb} (\texttt{sim}); 
+After applying the correction algorithm of \textsection~\ref{correction} (\texttt{sim+fix}).
 Columns Ref and Fix indicate the number of original and corrected sentences (in millions) used in training.
@@
 
-We presented an unsupervised method based on deep neural networks for detecting translation divergences in parallel sentence pairs. 
-Our model optimises word alignments. Hence, allowing for fine grained divergence prediction at the level of words. 
-Misaligned/divergent words can then be filtered out allowing for reusing some divergent sentences. 
-We evaluated our model on two neural machine translation tasks, showing significant improvements when compared to training over the entire data set. 
+We presented an unsupervised method based on deep neural networks for detecting translation divergences in parallel corpora. 
+Our model optimizes word alignments, and computes a fine grained divergence prediction at the level of words. 
+Misaligned/divergent words can then be filtered out, yielding larger and better training sets. 
+Our experiments on two machine translation tasks show significant improvements in comparison to training with the entire data set. 
 %The method can be used on any parallel corpus without any manual annotation.

### 2018-11-07 2019/naacl/naaclhlt2019.tex
Sparse embedding

--- 2018-07-27 WMT/divergence.tex
@@
 %
-% File emnlp2018.tex
+% File naacl2019.tex
 %
-%% Based on the style files for EMNLP 2018, which were
-%% Based on the style files for ACL 2018, which were
+%% Based on the style files for ACL 2018 and NAACL 2018, which were
 %% Based on the style files for ACL-2015, with some improvements
@@
 \documentclass[11pt,a4paper]{article}
-\usepackage[hyperref]{emnlp2018}
+\usepackage[hyperref]{naaclhlt2019}
 \usepackage{times}
 \usepackage{latexsym}
+
 \usepackage{url}
-\usepackage{amsmath}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-%%% YOUR PACKAGES BELOW THIS LINE %%%
-\usepackage[small,bf]{caption} % added MLF 20171211
-\usepackage{multicol}
-\usepackage{adjustbox}
-\usepackage{graphicx}
-\usepackage{array,multirow}
+
 %\aclfinalcopy % Uncomment this line for the final submission
+%\def\aclpaperid{***} %  Enter the acl Paper ID here
 
-\setlength\titlebox{5cm}
+%\setlength\titlebox{5cm}
 % You can expand the titlebox if you need extra space
@@
 \newcommand\BibTeX{B{\sc ib}\TeX}
-\newcommand\confname{EMNLP 2018}
-\newcommand\conforg{SIGDAT}
 
-%\title{Crosslingual Sentence Embeddings for Neural Machine Translation}
-%\title{A Recurrent Network for Parallel Corpora Similarity Analysis}
-\title{Detecting and Fixing Translation Divergences in Noisy Parallel Corpora}
+\title{Le titre}
 
-\author{MinhQuang Pham$^{\dag\ddag}$, \ \ Josep Crego$^\dag$,\ \ Jean Senellart$^\dag$ ,\ \ Fran\c cois Yvon$^\ddag$\\
-  $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France\\
-  {\tt firstname.lastname@systrangroup.com}\\
-  $^\ddag$LIMSI-CNRS / Universit\'e Paris-Saclay 91405 Orsay, France\\
-  {\tt firstname.lastname@limsi.fr}}
+\author{Minh Quang Pham\\
+  Affiliation / Address line 1 \\
+  Affiliation / Address line 2 \\
+  Affiliation / Address line 3 \\
+  {\tt email@domain} \\\And
+  Second Author \\
+  Affiliation / Address line 1 \\
+  Affiliation / Address line 2 \\
+  Affiliation / Address line 3 \\
+  {\tt email@domain} \\}
 
@@
 \begin{abstract}
-
-%http://www.airccj.org/CSCP/vol4/csit42503.pdf
-
-Corpus-based approaches to machine translation rely on the availability and quality of parallel corpora.
-In the case of neural machine translation, a large neural network is trained to maximise the translation performance on a given parallel corpus. 
-Such resource is not naturally existing, and because of the process necessary to compile a parallel corpus, it may contain multiple sentence pairs that are often not as parallel as one might assume.
-%%
-This paper describes an unsupervised method for detecting translation divergences in parallel sentences.
-We present a neural network to predict sentence similarity that minimises a loss function based on word divergence.
-We show that accurate predictions are obtained allowing divergent sentences to be filtered out.
-Furthermore, word similarity scores predicted by the network are used to identify and fix some divergences guiding to collect additional parallel segments.
-We evaluate the presented method on  English-French and  English-German machine translation tasks.
-Results show that neural MT systems trained on the filtered/corrected corpus outperform the MT systems trained on the original data.
-%%
-%This paper describes an unsupervised method for building cross-lingual sentence embeddings that optimise word alignments. 
-%Such embeddings can then be used to compute semantic similarity between sentences in different languages.
-%We first use sentence embeddings to detect translation divergences in parallel sentence pairs, thereby helping to filter out noisy pairs.
-%Then, word similarity scores predicted by the network are used to identify and fix some divergences guiding to collect additional parallel segments.
-%We evaluate our model on multiple language pairs and data resources.
-%This model can be used on any parallel corpus without any manual annotation.
-
-%According to the process followed to compile a parallel corpus, it may contain multiple parallel sentences that are often not as parallel as one might assume.
-Examples are easily found when corpora is created from non-parallel texts or included by noisy document/sentence alignment tools.
-%This paper describes an unsupervised method for detecting translation divergencies in parallel sentence pairs.
-We use a deep neural network to predict word alignment scores in parallel sentences.
-Misaligned words are then identified allowing divergent parallel sentences to be filtered out and in some cases making sentences parallel by removing such words.
-%We evaluate the presented method on a machine translation task.
-Results show that a neural MT system trained on the filtered/corrected sentences outperforms the neural MT system trained on the original sentence pairs.
-This method can be used on any parallel corpus without any manual annotation.
-
 \end{abstract}
@@
 
-Parallel sentence pairs are the only necessary resource to build Machine Translation (MT) systems.
-In the case of neural MT, a large neural network is trained through maximising translation performance on a given parallel corpus. 
-Therefore, the quality of an MT engine is heavily dependent upon the amount and quality of the training parallel sentences.\footnote{Note that  recent work on neural MT~\cite{lample2018word,artetxe2018iclr} completely dispenses with the need of parallel data, using unsupervised methods to obtain performance improvements over word-by-word statistical MT systems.
-However, these systems still lag far behind systems trained in a supervised fashion, as considered in this work.}
+\section{Related Work}
 
-Unfortunately, parallel texts are scarce resources. 
-There are relatively few language pairs for which parallel corpora of
-reasonable sizes are available, and even for those pairs, the corpora
-are only representative of a small number of domains or genres. 
-To mitigate the lack of parallel data, several approaches have been developed in the last years.
-They range from methods using non-parallel, or comparable data ~\cite{Zhao:2002:APS:844380.844785,W04-3208,J05-4003,W17-2509,P17-3003,P18-2037} to techniques that produce synthetic parallel data from monolingual corpora ~\cite{P16-1009,W17-4714} using in all cases automated alignment/translation engines that are prone to introducing noise in the resulting parallel sentences.
-Mismatches in parallel sentences extracted from translated texts are also reported ~\cite{tiedemann2011bitext,XU16.310}. 
-This problem is mostly ignored in MT, where parallel sentences are considered to convey the exact same meaning, but seems particularly important for neural MT engines, as noted in~\citet{chen2016adaptation}.
-
-\begin{table}[ht]
-%\begin{adjustbox}{width=0.48\textwidth}
-\small
-\center
-\begin{tabular}{ c|l }
-  %\hline
-  \hline  
-  \texttt{en} & \it{What do you feel}\bf{, Spock}\it{?} \\
-  \texttt{fr} & \it{Que ressentez-vous?} \\
-  \texttt{gl} & {\small \it{What do you feel?}} \\
-  \hline
-  \texttt{en} & \it{How much do you get paid?} \\
-  \texttt{fr} & \it{T'es pay\'e combien} \bf{de l'heure}\it{?} \\
-  \texttt{gl} & {\small \it{How much do you get paid per hour?}} \\
-  \hline  
-  \texttt{en} &  \bf{That seems a lot.} \\
-  \texttt{fr} & \bf{40 livres?} \\
-  \texttt{gl} & {\small \it{40 pounds?}} \\
-  \hline  
-  \texttt{en} & \it{I brought you} \bf{french fries}\it{!} \\
-  \texttt{fr} & \it{Je t'ai rapport\'e des} \bf{saucisses}\it{!} \\
-  \texttt{gl} & {\small \it{I brought you sausage!}} \\
-  \hline
-  %\hline  
-\end{tabular}
-%\end{adjustbox}
-\caption[Table caption text]{Examples of semantically divergent parallel sentences. English (\texttt{en}), French (\texttt{fr}) and gloss of French (\texttt{gl}).}
-\label{tab:examples}
-\end{table}
-
-Table~\ref{tab:examples} gives some examples of English-French
-parallel sentences that are not completely semantically equivalent, 
-extracted from the OpenSubtitles corpus ~\cite{LisonTiedemann2016}. 
-Divergences are outlined using bold letters. 
-Different types of translation divergences exist in a parallel corpus:
-Additional segments are included on either side of the parallel sentences (first and second rows) most likely due to errors in sentence segmentation tools;
-Some translations may be completely uncorrelated (third row);
-Inaccurate translations also exist (fourth row). 
-Note that divergent translations can be due to many different reasons ~\cite{C14-1055}, the study of which is beyond the scope of this paper.
-
-In this work, we present an unsupervised method for building cross-lingual sentence embeddings based on modelling word similarity. % in the form of continuous vectors. 
-The architecture of our network allows us to distinguish between different types of divergences which are common to many existing bi-texts.
-The resulting sentence embeddings are then used to measure semantic equivalence of the corresponding sentences.
-To evaluate our method we show that translation accuracy can be improved after filtering out divergent sentence pairs in an English-to-French and an English-to-German translation tasks.
-Then, we  show that in some cases, divergent sentences can be fixed by removing divergent words, further boosting translation accuracy.
-
-The remainder of this paper is structured as follows. 
-Section~\ref{related} overviews related work. 
-We describe in detail the core of the neural similarity classifier in Section~\ref{sec:similarity}. 
-We report experiments with the presented model in Section~\ref{experiments}.
-Section~\ref{sec:results} evaluates results. 
-Finally, conclusions are drawn in Section~\ref{conclusions} and further work is outlined in Section~\ref{further}.
-All the code used in this paper is freely available\footnote{https://github.com/anonymised}.
-
-\section{Related Work}
-\label{related}
-
-Prior work on translation divergencies focused on reflecting that translation of one language into another results in very different forms ~\cite{J94-4004} according to several categories (e.g.
-lexical, structural, categorical).
-
-Attempts to measure the impact of translation divergencies in MT systems have focused on the introduction of noise in sentence alignments ~\cite{goute2012}, showing that statistical MT systems are highly tolerant to noise, and that performance only degrades seriously at very high noise levels.
-In contrast, neural MT engines seem to be more sensitive~\cite{chen2016adaptation}, as they tend to assign high probabilities to rare events~\cite{Hassan2018AchievingHP}.
-
-Efforts have been devoted to characterising the degree of semantic equivalence between two snippets of text in the same or different languages~\cite{conf/semeval/AgirreBCDGMRW16}, a workshop devoted to an objective similar to our work.
-In~\cite{Mueller:2016:SRA:3016100.3016291}, a monolingual sentence similarity network is proposed, making use of a simple LSTM layer to compute sentence representations. 
-The authors show that a simple SVM classifier can be built on top of the sentence representations to achieve state-of-the-art results in a semantic entailment classification task.
-
-Similar to our work, in ~\cite{W17-3209} the authors train a cross-lingual divergence detector using word alignments and sentence length features to train a linear SVM classifier. 
-Their work shows that an NMT system trained only on non-divergent sentences yields slightly higher translation quality scores and requires clearly less training time. 
-The same authors have recently updated their work in~\cite{DBLP:journals/corr/abs-1803-11112}. 
-They use the system presented in ~\cite{N16-1108} that employs multiple convolutional layers and models pairwise word interactions. 
-The updated network further outperforms their previous work.
-
-Recently, in~\cite{P18-2037} is presented a network able to learn a joint multilingual sentence embedding space, with the goal that the distance in that space reflects their semantic difference, independently of the language.
-The network learns sentence embeddings by max-pooling over several Bi-LSTM outputs. 
-The system is successfully used to filter noisy parallel data and to mine for bi-texts in large monolingual collections.
-
-As in the previous works we also use a recurrent network to build sentence embeddings but we optimise word similarity rather than sentence similarity. 
-Thus, enabling fine grained divergence prediction at the level of words.
-In addition, we propose a simple method that employs word similarity scores to not only detect but to correct divergent sentences, which allows to reuse additional sentence pairs in the context of neural machine translation.
-
-\section{Neural Similarity Classifier}
-\label{sec:similarity}
-
-In this section we describe the architecture of our model, very much inspired by the work on Word Alignment in~\cite{W16-2207}. Figure~\ref{network} illustrates the network.
-In the following, we consider a source-target sentence pair $(s,t)$ with $s=(s_1,...,s_I)$ and $t=(t_1,...,t_J)$.
-
-\begin{figure}[h!]
-\center
-    \includegraphics[width=1.0\linewidth]{network}
-    \caption{Illustration of the model.
-The network is composed of source and target word embedding lookup tables ($LT_s$ and $LT_t$) and two identical subnetworks ($net_s$ and $net_t$) that compute in context representations of source ($s_i$) and target words ($t_j$).}
-    \label{network}
-\end{figure}
-
-The model is composed of 2 Bi-directional LSTM subnetworks, $net_s$ and $net_t$, which respectively encode source and target sentences.
-Since both $net_s$ and $net_t$ take the same form we describe only the source architecture.
-The source-sentence Bi-LSTM network outputs forward and backward hidden states, $\overrightarrow{h}^{src}_i$ and $\overleftarrow{h}^{src}_i$, which are then concatenated into a single vector encoding the $i^{th}$ word of the source sentence,
-$h^{src}_i = [ \overrightarrow{h}^{src}_i ; \overleftarrow{h}^{src}_i ]$.
-In addition, the last forward/backward hidden states (outlined using dark grey in Figure~\ref{network}) are also concatenated into a single vector  to represent whole sentences 
-$h_{src} = [ \overrightarrow{h}^{src}_I ; \overleftarrow{h}^{src}_1 ]$.
-At this point a measure of similarity between sentences can be obtained by cosine similarity: 
-\begin{equation}
-    sim(h_{src}, h_{tgt}) = \frac{h_{src} \cdotp h_{tgt}}{||h_{src}|| * ||h_{tgt}||}
-    \label{cosine}
-\end{equation}
-\noindent where two vectors (embeddings) with the same orientation have a cosine similarity of $1$, while two vectors with opposed orientation have a similarity of $-1$, independent of their magnitude.
-
-%The similarity function outputs the level of semantic similarity between source and target sentences.
-
-%Figure~\ref{net_lstm} illustrates the first component of our network. In order to optimise our model we follow the functions depicted in Figure~\ref{net_align}.
-
-Similar to ~\cite{W16-2207} our model extracts context information from source and target sentences and then computes simple dot-products to estimate word alignments. 
-The objective function is computed at the level of words. 
-To enable unsupervised training, we use an aggregation operation that summarizes the alignment scores for a given target word. 
-A soft-margin objective increases scores for true target words while decreasing scores for target words that are not present.
-The aggregation function combines the scores of all source (or target) words for a particular target (or source) word and promotes source words which are likely to be aligned with a given target word according to the knowledge the model has learned so far.
-Alignment scores $S(i,j)$ are given by the dot-product $S(i,j) = h_i^{src} \cdotp h_j^{tgt}$, while aggregation functions are defined as:
-
-\begin{equation}
-\begin{split}
-    aggr_s(i,S) = \frac{1}{r} \ log \left( \displaystyle \sum_{j=1}^{J} e^{r * S(i,j)}\right) \\
-    aggr_t(j,S) = \frac{1}{r} \ log \left( \displaystyle \sum_{i=1}^{I} e^{r * S(i,j)}\right)
-\end{split}
-\label{aggregation}
-\end{equation}
-
-%Hence, rather than sentence similarity our model optimises word alignments.
-The loss function is defined as:
-
-%\vspace{-8mm}
-\begin{equation}
-\begin{split}
-\mathcal{L}(src,tgt) = & \\
-    \sum_{i=1}^I log&\left(1+e^{aggr_{s}(i,S) * \mathcal{Y}_i^{src}}\right) +\\
- + \sum_{j=1}^J log&\left(1+e^{aggr_{t}(j,S) * \mathcal{Y}_j^{tgt}}\right)
-\end{split}
-\label{loss_wemb}
-\end{equation}
-\noindent where $\mathcal{Y}_i^{src}$ and $\mathcal{Y}_j^{tgt}$ are vectors with reference labels containing $-1$ when the word is present in the translated sentence, and $+1$ for divergent (unpaired) words.
-
-%\begin{figure}[h]
-%\center
-%    \includegraphics[width=0.9\linewidth]{net_align}
-%    \caption{Illustration of the model.}
-%    \label{net_align}
-%\end{figure}
-
-%\vspace{-8mm}
-
-%For comparison purposes we also implemented the model introduced in \cite{W17-2509}.
-The network employs the same initial Bi-LSTM layers outlined in Figure~\ref{network} which encode source and target sentences into $h_{src}$ and $h_{tgt}$ vectors.
-The remaining of the network is illustrated by Figure~\ref{net_sentence}, where sentence pair matching information is captured by using their element-wise product and absolute element-wise difference.
-Sentence similarities are finally estimated by feeding the matching vectors into a fully connected layer.
-
-%\begin{equation}
-%\begin{split}
-%h^{(1)} =  &\ \ h_{src} * h_{tgt} \\
-%h^{(2)} =  &\ \ |h_{src} - h_{tgt}| \\
-%h^{(3)} = &\ \ tanh(W[h^{(1)};h^{(2)}] + b^{(3)}) \\
-%h_{sim} = &\ \ Wh^{(3)} + b^{sim}
-%\end{split}
-%\end{equation}
-
-%This approach is different from our model since it uses a single end-to-end model to estimate the conditional probability distribution that two sentences are parallel.
-
-%\begin{figure}[h]
-%\center
-%    \includegraphics[width=0.8\linewidth]{net_sentence}
-%    \caption{Illustration of the model presented in~\cite{W17-2509}.
-Sentence embeddings (vectors in dark grey) are computed by the same Bi-LSTM layers shown in Figure~\ref{network}.} %The network is composed of source and target word embedding lookup tables ($LT_s$ and $LT_t$) and two identical subnetworks ($net_s$ and $net_t$) that compute in context representations of source ($s_i$) and target words ($t_j$).
-The similarity function outputs the level of semantic similarity between source and target sentences.}
-%    \label{net_sentence}
-%\end{figure}
-
-%For this second network the loss function is defined as: 
-%\vspace{-5mm}
-%\begin{equation}
-%\mathcal{L}(src,tgt) = log(1+e^{h_{sim} * \mathcal{Y})}) % where $sign(src,tgt) =1
-%\label{loss_semb}
-%\end{equation}
-%\noindent where $\mathcal{Y}$ is the reference label containing $-1$ when the given sentence pair is non-divergent (parallel) and $+1$ for divergent (unpair) examples.
-
-%Note that different from the network described in the original work, our implementation does not estimates a probability distribution but FINISH.
-
-%For each source(target) word, we consider the aggregated matching score over the target(source) sentence:
-%$$s_{aggr}(i,tgt) = Aggr_{j=1}^{J} s(i,j)$$
-%$$s_{aggr}(src,j) = Aggr_{i=1}^{I} s(i,j)$$
-
-%Where $Aggr$ is LSE aggregation operator used in~\cite{W16-2207}.
-Given the sign of each word noted $sign(w)$, that indicates whether the word $w$ is matched to the other side (by convention, $1$ indicates non-matching, $-1$ indicates matching), the loss function is defined as follows:
-
-%In the second option, we directly optimize the score of similarity between two sentences by defining the loss function as follows:
-
-%$$L(src,tgt) = log(1+e^{h_{src} . h_{tgt} * sign(src,tgt)})$$ where $sign(src,tgt) =1 $ if two sentences are not parallel and $sign(src,tgt) =-1 $ otherwise.
-
-\subsection{Training with Negative Examples}
-\label{training}
-
-Training is performed by minimising Equation~\ref{loss_wemb}, for which examples with annotations for source $\mathcal{Y}_i^{src}$ and target $\mathcal{Y}_j^{tgt}$ words are needed.
-Different kind of examples are used to train our network:
-\begin{itemize}
-
-\item As positive examples we use paired sentences of the parallel corpus. In this case, all words in both sentences are labelled as parallel, $\mathcal{Y}_i^{src}=-1$ and $\mathcal{Y}_j^{tgt}=-1$.
-
-\item As negative examples we use random unpaired sentences. In this case, all words are labelled as divergent, $\mathcal{Y}_i^{src}=+1$ and $\mathcal{Y}_j^{tgt}=+1$. 
-Since negative examples may be very easy to classify and we want our network to detect less obvious divergencies, we further create negative examples following the next two methods:
-
-\item We replace random sequences of words on either side of the sentence pair by a sequence of words with the same part-of-speeches. 
-The rationale behind this method is to keep the new sentences as grammatical as possible. 
-Otherwise, to predict divergence the network can learn to detect non-grammatical sentences.
-Words that are not replaced are considered parallel ($-1$) while those replaced are assigned the divergent label ($+1$). 
-Words aligned to some replaced words are also assigned the divergent label ($+1$). For instance, given the original sentence pair:
-
-\begin{table}[h]
-\center
-\begin{tabular}{ll}
-src: & { \small \texttt{What do you feel ?}} \\
-tgt: & { \small \texttt{Que ressentez-vous ?}} \\
-\end{tabular}
-\end{table}
-
-We may replace '\texttt{you feel}', with part-of-speech tags '\texttt{PRP VB}', by another sequence with same tags (i.e. '\texttt{we want}'):
-
-\begin{table}[h]
-\center
-\begin{tabular}{ll}
-src: & { \small \texttt{What do {\bf we \ want} ?}} \\
-$\mathcal{Y}^{src}$: & { \small \texttt{-1 \ \  -1 {\bf +1\ \ +1} \ \  -1}} \\
-tgt: & { \small \texttt{Que {\bf ressentez-vous} ?}} \\
-$\mathcal{Y}^{tgt}$: & { \small \texttt{-1\ \ {\bf +1}\ \ \ \ \ \ \ \ \ -1}} \\
-\end{tabular}
-\end{table}
-
-Divergent words are shown in bold. 
-Note that the new sentences are not assured to be grammatical after replacing sequences with the same part-of-speech.
-We use word alignments to identify as divergent the sequence '\texttt{ressentez-vous}' since it is aligned to '\texttt{you feel}' in the original sentence pair.
-
-\item Finally, motivated by sentence segmentation errors observed in many corpora, we also build negative examples by inserting a second sentence at the beginning (or end) of the source (or target) sentence pair.
-Words in the original sentence pair are assigned the parallel label ($-1$) while the new words inserted are considered divergent ($+1$).
-Given the original sentence pair previously shown, we may want to build the next negative example by inserting the sentence '\texttt{Not .}' at the end of the original source sentence:
-
-\begin{table}[h]
-\center
-\begin{tabular}{ll}
-src: & { \small \texttt{What do you want ? {\bf Not \ .}}} \\
-$\mathcal{Y}^{src}$: & { \small \texttt{-1 \ \  -1 -1 \ -1  \ \ -1 {\bf +1\ \ \  +1}}} \\
-tgt: & { \small \texttt{Que ressentez-vous ?}} \\
-$\mathcal{Y}^{tgt}$: & { \small \texttt{-1\ \ -1\ \ \ \ \ \ \ \ \ \ \ \ \ -1}} \\
-\end{tabular}
-\end{table}
-
-\end{itemize}
-
-In order to avoid that negative examples are easily predicted just by looking at the difference in length of training sentences we constraint all negative examples to have a difference in length not exceeding $2.0$.
-Very short sentences, of up to $4$ words, are accepted if the length ratio does not exceeds $3.0$.
-Note that the network can be trained with any combination of examples ({\texttt P}aired, {\texttt U}npaired, {\texttt R}eplace and {\texttt I}nsert).
-In all cases, the same number of examples is used for each type.
-
-\subsection{Fixing Translation Divergencies}
-\label{correction}
-
-We observed in our training corpora that many divergent sentences follow a common pattern, consisting of adding some extra leading/trailing words.
-As in the first and second examples of table~\ref{tab:examples}.
-Accordingly, we implemented a simple algorithm that discards sequences of leading/trailing words on both sides of the source and/or target sentences. 
-Hence considering as parallel $s_u^v$ and $t_x^y$. 
-To find the optimal source ($u, v$) and target ($x, y$) indices that enclose the parallel segments within the original sentence, we implement:
-\begin{equation*}
-\underset{u, v, x, y}{\arg\max} \Big \{      \underset{u \le I \le v}{\sum} \underset{x \le j \le y}{\max} \{ S(i,j) \}    \Big \}
-\end{equation*}
-
-The $\mathcal{N}$-best sequences following the previous function ($s_u^v$, $t_x^y$) are considered as valid corrections, but only the highest ranked according to its similarity score is used as replacement for the original $(s_1^I, t_1^J)$.
-Short sentences are not considered. This is, $v - u > \tau$ and $y - x > \tau $. 
-Figure~\ref{matrix} (left) shows an example of an alignment matrix $S(i,j)$ for a given sentence pair. 
-An acceptable correction is:
-
-\begin{table}[h]
-\center
-\begin{tabular}{c}
-	\textit{Que ressentez-vous ? \ \ $\Leftrightarrow$\ \ \ What do you feel ?} \\
-\end{tabular}
-\end{table}
-
-\noindent Hence, with optimal indices being: $u=1$, $v=5$, $x=1$ and $y=3$.
+\section{The Model}
 
 \section{Experiments}
-\label{experiments}
-
-In this section we report on the experiments conducted to evaluate the proposed network. We begin with details of the corpora employed.
 
 \subsection{Corpora}
-\label{corpora}
 
-Different corpora are used in this work to evaluate our similarity classifier.
-We filter out divergencies found in the English-French OpenSubtitles corpus~\cite{LisonTiedemann2016}, which consists of a collection of movie and TV subtitles, aligned at the sentence level after a number of automatic preprocessing steps.
-The corpus presents many potential divergencies as outlined in Table~\ref{tab:examples}. 
-To evaluate performance we used the En-Fr Microsoft Spoken Language Translation task, created from actual conversations over Skype~\cite{mslt-corpus-iwslt-2016-release}. 
-%We also tested on a second domain, using the publicly available newstest-2013 En-Fr test set, corresponding to news stories selected from online sources~\cite{WMT:2013}.
+\subsection{NMT engine}
 
-In order to better assess the quality of our classifier when facing different types of word divergences we collected from the original OpenSubtitles corpus $500$ sentences containing different types of examples:
-\begin{itemize}
-\item $200$ paired sentences,
-\item $100$ unpaired sentences,
-\item $100$ sentences with replace examples and
-%\item $100$ sentences with replace examples; 
-\item $100$ sentences with insert examples.
-\end{itemize}
-
-Details of each type of example are given in Section~\ref{training}.
-Sentence pairs have divergence annotations at the level of words.
-
-We also use the Paracrawl corpus, which consists of a very noisy $1$ billion word (English token count) German-English corpus crawled from the web as part of the Paracrawl project\footnote{http://paracrawl.eu/}.
-Performance is evaluated on the publicly available newstest-2017 En-De test set, corresponding to news stories selected from online sources~\cite{W17-4717}.
-
-%To evaluate the ability of our sentence embeddings to identify parallel sentences in comparable corpora we followed the data conditions made available for the BUCC shared task~\cite{ZWEIGENBAUM18.12}.
-%For training we used the French-English European Parliament corpus from WMT15\footnote{http://www.statmt.org/wmt15/translation-task.html}.
-%For testing we used the test set made available by the task organisers, consisting of about $300,000$ monolingual sentences per language, French and English, that contains near $10,000$ parallel sentences.
-%Test data come from Wikipedia\footnote{http://ftp.acc.umu.se/mirror/wikimedia.org/dumps} and News Commentary\footnote{http://www.casmacat.eu/corpus/news-commentary.html}.
-
-\subsection{Training}
-
-All data is preprocessed with \texttt{OpenNMT}\footnote{http://opennmt.net}, performing minimal tokenisation, basically splitting-off punctuation.
-
-\subsubsection{Neural Similarity Classifier}
-\label{divergence}
-
-After tokenisation, the $50,000$ most frequent words of each language are used as vocabulary.
-Each out-of-vocabulary word is mapped to a special UNK token.
-Our similarity classifier is described in Section~\ref{sec:similarity}. 
-Word embeddings ($LT_s$ and $LT_t$) are initialised using \texttt{fastText}\footnote{https://github.com/facebookresearch/fastText}, further aligned by means of \texttt{MUSE}\footnote{https://github.com/facebookresearch/MUSE} following the unsupervised method detailed in~\cite{lample2018word}.
-Size of embeddings is $E_s=E_t=256$ cells. 
-Both Bi-LSTM use 256-dimensional hidden representations ($E=512$). 
-We use $r=1.0$. 
-Optimization of the parameters is done using the stochastic gradient descent method along with gradient clipping (rescaling gradients whose norm exceeds a threshold) to avoid the exploding gradients problem~\cite{Pascanu:2013:DTR:3042817.3043083}.
-For each epoch we randomly select $1$ million sentence pairs that we place in batches of $32$ examples.  
-Negative examples are created following the strategies detailed in~\ref{training}.
-Word alignments and English part-of-speeches used to build negative examples were performed by \texttt{fast\_align}\footnote{https://github.com/clab/fast\_align} and \texttt{FreeLing}\footnote{https://github.com/TALP-UPC/FreeLing.git} respectively.
-We run $10$ epochs and start decaying at each epoch by $0.8$ when score on validation set increases. 
-Divergence is always computed following equation~\ref{cosine}. 
-For divergence correction, we use $\mathcal{N}=20$ and $\tau=3$.
-
-\subsubsection{Neural MT}
-\label{translation}
-
-In addition to the basic tokenisation detailed in Section~\ref{divergence} we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30000$ merge operations learned joined from both language sides.
-Neural MT systems are based on the open-source project  \texttt{OpenNMT}. 
-We use a Transformer model with same configuration than in paper~\cite{vaswani2017attention}, i.e, both encoder and decoder have $6$ layers; Multi-head attention is performed over $8$ heads. 
-Hidden layer’s size is $512$. 
-The inner layer of feed forward network is of size $2048$. 
-Word embeddings have $512$ cells. We set the dropout probability to $0.1$. 
-Batch size is set to $3072$.
-%The maximum length of both source and target sentences is set to $80$ and we limit the vocabulary size to $50,000$ words for both source and target languages. 
-The optimiser is Lazy Adam with $\beta_1 = 0.9$, $\beta_2 = 0.98$, $\epsilon = 10^{-9}$, $warmup\_steps = 4000$. We stop training after $30$ epochs.
-
-%In addition to the basic tokenisation we perform Byte-Pair Encoding~\cite{Sennrich2016} with $30,000$ merge operations learned from both English and French data.
-%We build our NMT systems based on the open-source project \texttt{OpenNMT}.
-We use a bidirectional RNN encoder with $4$ LSTM layers with each containing $1,000$ cells.
-Word embeddings have a size of $300$ cells.
-We set the dropout probability to $0.3$.
-Batch size is set to $64$.
-The maximum length of both source and target sentences is set to $80$.
-% and we limit the vocabulary size to $50K$ words for both source and target languages.
-%The default optimiser is SGD with the starting learning rate of $1.0$.
-We start to decay the learning rate from epoch $10$ or when we detect increasing perplexity as compared to the previous epoch on a validation set.
-We stop training after $20$ epochs.
-
-\section{Results}
-\label{sec:results}
-
-In this section we evaluate the performance of our neural similarity classifier and correction algorithm detailed in Section~\ref{sec:similarity}.
-
-\subsection{Divergence Prediction}
-
-We evaluate first the ability of our similarity classifier to predict different types of divergencies at the level of words. 
-We use the test set manually annotated for that purpose and train our model on the OpenSubtitles corpus.
-A word is considered divergent when associated to a negative aggregation score (see Equation~\ref{aggregation}).
-Table~\ref{results_puri} shows accuracies obtained by our model when trained over different combinations of negative examples.% on the task of predicting wether words are divergent. 
-%We consider as divergent those words for which the similarity score predicted by our model, Equation~\ref{cosine}, is negative.
-
-\begin{table}[h]
-\small
-\center
-\begin{tabular}{crccccc}
-\hline
-\multicolumn{2}{l}{\bf Accuracy} & \multicolumn{5}{c}{Test examples} \\
- &  & \texttt{P} & \texttt{U} & \texttt{R} & \texttt{I} & \texttt{PURI} \\
- \hline
-\parbox[t]{0mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Train examples}}} &  \texttt{PU}     & \bf 0.996 & \bf 0.994 & 0.671 & 0.673 & 0.874 \\
- &  \texttt{PR}     & \bf 0.995 &      0.033 & \bf 0.951 &      0.689 & 0.746 \\
- &  \texttt{PI}       & \bf 0.998 &      0.071 &      0.697 & \bf 0.725 & 0.705 \\
- &  \texttt{PUR}  & \bf 0.994 & \bf 0.989 & \bf 0.919 &      0.710 & 0.932 \\
- &  \texttt{PUI}    & \bf 0.995 & \bf 0.996 &      0.662 & \bf 0.769 & 0.887 \\
- &  \texttt{PRI}    & \bf 0.991 &      0.161 & \bf 0.924 & \bf 0.719 & 0.768 \\
- &  \texttt{PURI} & \bf 0.995 & \bf 0.980 & \bf 0.916 & \bf 0.788 & \bf 0.942 \\
-\hline
-\end{tabular}
-\caption{Word divergence accuracies according to different type of examples used in train/test. P, U, R and I stand respectively for pair, unpair, replace and insert.}
-\label{results_puri}
-\end{table}
-
-As it can be seen, non-divergent words in parallel sentences (column \texttt{P}) are easy to identify. 
-In all cases the accuracy reaches $100\%$. 
-Divergent words in unpaired sentences (column \texttt{U}) are also easy to identify as far as the model has seen these types of examples in training. 
-However, the accuracy drops dramatically when the model is not trained with unpaired sentences (rows  \texttt{PR}, \texttt{PI} and \texttt{PRI}).
-Regarding columns \texttt{R} and \texttt{I}, accuracies are lower since these sentences contain a mix of divergent and non-divergent words. 
-%Hence, word divergence is more difficult to predict. 
-Again, models that were trained with the corresponding examples (\texttt{R} and \texttt{I}) obtain the highest accuracies (outlined in bold letters).
-Column \texttt{PURI} shows accuracy results over the entire test set. This is, mixing all type of examples. 
-As expected, the best accuracy is also obtained by the system trained on all type of  examples.
-
-Figure~\ref{matrix} illustrates the output of our network when trained using \texttt{PU} examples (right) and \texttt{PURI} examples (left).  
-The former (right) fails to predict some word divergences, most likely because in training it never saw sentences mixing divergent and non-divergent words in the same example. 
-Furthermore, the network trained over \texttt{PURI} examples (left) correctly assigns a lower similarity score to the sentence pair as both sentences do not convey the exact same meaning.
-
-\begin{figure}[h]
-\centering
-    \includegraphics[width=0.4\linewidth]{feel}    
-    \includegraphics[width=0.4\linewidth]{feel_pu}    
-\caption{Sentence pair with similarity scores produced by our model when trained over \texttt{PU} examples (right) and over \texttt{PURI} examples (left).
-Aggregation scores (Equation~\ref{aggregation}) are shown next to words.
-Matrices contain alignment scores.
-Sentence similarities (Equation~\ref{cosine})  shown below matrices.}
-\label{matrix}
-\end{figure}
-
-\subsection{Neural Machine Translation}
-
-Next, we evaluate the ability of our neural similarity classifier to detect divergent sentence pairs. 
-Since we lack of a gold standard test set to directly evaluate sentence similarity, 
-we measure the accuracy of neural MT systems trained over the original corpus and after filtering divergent sentences. 
-Notice that some divergent sentences may not be completely useless to train a neural MT system. 
-Consider for instance the example of Figure~\ref{matrix}. 
-Despite not conveying the exact same meaning it still contains useful information to train a neural MT engine. 
-It is not the case of the example in the third row of Table~\ref{tab:examples}.
-
-Table~\ref{results_wemb} shows BLEU~\cite{P02-1040} scores of the same neural MT network when trained over different subsets of the OpenSubtitles data set.
-Subsets result from filtering sentences of the original corpus using different similarity thresholds as computed by Equation~\ref{cosine} of our similarity classifier.
-As it can be seen, the best performing subset is obtained when using sentence pairs with similarity score in the range $[0.100, +\infty)$. 
-This is, training with the most similar $15.5$ million sentences of the original data set.
-
-\begin{table}[h]
-\small
-\center
-\begin{tabular}{lcc}
-\hline
-\bf $sim(h_{src},h_{tgt})$ & \bf Size (M) & \bf Test (BLEU)\\%& \bf NEWS \\
-\hline
-$(-\infty,+\infty)$   & 27.2 & 42.18 \\ %& 26.60  \\
-$[0.000,+\infty)$   & 24.0 & 42.68 \\ %& 26.87  \\
-$[0.003,+\infty)$   & 21.5 & 42.56 \\ %&  \\
-$[0.076,+\infty)$   & 18.0 & 43.19 \\ %& 26.38 \\
-$[0.100,+\infty)$   & 15.5 & 43.12\\%&  \\
-\hline
-\end{tabular}
-\caption{BLEU scores obtained by a neural MT network when trained over the OpenSubtitles corpus filtered using different similarity thresholds.}
-\label{results_wemb}
-\end{table}
-
-Note that sentences in the range $[0.076, 0.100]$ ($2.5$ million sentences) does not contribute to increase the accuracy of the resulting neural MT system.
-%The BLEU score of the system that considers these sentences in training remains unchanged.
-Furthermore, sentences in the range $[0, 0.076]$ (about $6$ million sentences) despite assigned a positive similarity score they degrade the overall performance of the network.
-
-%Note that despite assigned a positive similarity score, sentence pairs in the range $[0, 0.076]$ (aobut $6$ million sentences) does not seem to contribute when training a neural MT network. 
-%On the contrary, when added as training data the MT accuracy is reduced.
-
-Finally, table~\ref{results} shows BLEU results for our neural MT engine trained over different data configurations:
-\begin{itemize}
-\item The entire\footnote{The original Paracrawl corpus contains more than $100$ million sentences.
-We reduced its size to $22.2$ millions using standard filtering techniques.} data sets (\texttt{all}) of both corpora,
-\item The most similar pairs as predicted by our network (\texttt{sim}),
-\item After applying the correction algorithm detailed in Section~\ref{correction} (\texttt{sim+fix}).
-\end{itemize}
-Columns Ref and Fix indicate the number of original and corrected sentences (in millions) considered to train the neural MT system.
-
-\begin{table}[h!]
-\small
-\center
-\begin{tabular}{lccl}
-\hline
-\bf Data & \bf Ref (M) & \bf Fix (M) & \bf Test (BLEU) \\ %MSLT & \bf NEWS \\
-\hline
-\multicolumn{3}{c}{\scriptsize{OpenSubtitles English-French}} \\
-\texttt{all}                   & 27.2 & - & 42.18 \\
-%\texttt{semb}             & 18.0 & - & \\
-\texttt{sim}            & 15.5 & - & 43.12  ($+0.94$)\\
-%\texttt{sim}           & 21.5 & - & 42.56 \\
-\texttt{sim}           & 18.0 & - & 43.19  ($+1.01$)\\
-\texttt{sim+fix}     & 15.5 & 2.5 & \bf 44.19 ($+2.01$)\\
-%\texttt{sim+fix}   & 15.5 & 5.0 & \\
-\hline
-\multicolumn{3}{c}{\scriptsize{Paracrawl English-German}} \\
-%\texttt{all}                  & 100.0 & 12.56 \\ 
-\texttt{all}                  & 22.2 & - & 19.27 \\ 
-%\texttt{semb}            & 15.0 & - & \\
-\texttt{sim}           & 15.0 & - & 21.52 ($+2.25$)\\
-\texttt{sim}           & 17.5 & - & 21.97 ($+2.70$)\\
-\texttt{sim+fix}   & 15.0 & 2.5 & \bf 22.42 ($+3.15$) \\
-\hline
-\end{tabular}
-\caption{BLEU scores obtained by neural MT using different subsets of the OpenSubtitles and Paracrawl corpora.}
-\label{results}
-\end{table}
-
-Results obtained after filtering sentence pairs by our network (\texttt{sim}) clearly outperform the baseline (\texttt{all}) by $+0.94$ and $+2.25$ BLEU respectively.
-
-Regarding OpenSubtitles, when fixing $2.5$ million sentences ($4^{th}$ row) the accuracy is further boosted to $+2.01$, while the same sentence pairs does not show any improvement when added in their original form ($3^{rd}$ row).
-%results are further improved $+2.01$ when fixing $2.5$ million sentences ($4^{th}$ row). 
-%The same $2.5$ million sentences did not show any improvement when added in their original form
-
-Similar results are obtained over the Paracrawl corpus.
-Results after fixing $2.5$ million sentences ($4^{th} row$) outperform those obtained by the same sentence pairs when used by the neural MT system in their original form ($3^{rd} row$).
+\subsection{Results}
 
 \section{Conclusions}
-\label{conclusions}
 
-We presented an unsupervised method based on deep neural networks for detecting translation divergences in parallel sentence pairs. 
-Our model optimises word alignments. Hence, allowing for fine grained divergence prediction at the level of words. 
-Misaligned/divergent words can then be filtered out allowing for reusing some divergent sentences. 
-We evaluated our model on two neural machine translation tasks, showing significant improvements when compared to training over the entire data set. 
-The method can be used on any parallel corpus without any manual annotation.
+\section*{Acknowledgments}
 
-\section{Further Work}
-\label{further}
-
-%We plan to further evaluate divergence classification under different data noise levels. %and on additional language pairs. 
-%as well as to measure the impact of using less noisy data when learning the similarity model
-%We also would like to evaluate the impact of using less noisy (already filtered) data to learn the similarity model. 
-%We would also like to use our model to predict other types of divergences 
-We plan to use our model to predict sentence embeddings over comparable corpora, allowing to collect parallel pairs through vector similarity measures.
-%We also plan to further evaluate the correction algorithm.
-In addition, we would like to measure the performance of our model after applying subword tokenisation,
-%when working with a reduced vocabulary set, as a result of applying any subword tokenisation,
-as well as using multiple LSTM layers, which is well known to capture hierarchical structure in the context of MT.
-
-%\section*{Acknowledgements}
-%We are grateful to Jo\"el Legrand for his fruitful comments when building the neural divergence classifier.
-
-\bibliography{biblio}
-\bibliographystyle{acl_natbib_nourl}
+\bibliography{naaclhlt2019}
+\bibliographystyle{acl_natbib}
 

### 2018-11-07 2019/naacl/naaclhlt2019.tex
add bib

@@
 
-\title{Le titre}
+\title{Sparse Word Embedding for Multi-domain Machine Translation}
 
@@
 \section{Related Work}
+Multi-domain Machine Translation has largely interested NLP community by its promising application in industry and its relation to fundamental problem in theory.
+Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}.
+Data centric methods aim to augment in-domain training data by using data selection techniques such as
+The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041}.
+\cite{D18-1041} create in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
+$h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively.
+While introducing new features in the architecture, \cite{D18-1041} introduces also new objective which is sum of word-level weighted MT objective and objectives of Domain-classifier in source side and target side and Adversarial Domain-classifier in source side.
+The author have very good approach to well seperate Domain-shared information flow and Domain-specify information flow and determine their contributions to the inference that mitigates the catastrophic interference happens in the network during forward step.
 
 \section{The Model}
+There is not any changes in the architecture of Neural Machine Translation network except the construction of word embedding.
 

### 2018-11-08 2019/naacl/naaclhlt2019.tex
continue writing

@@
 Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}.
-Data centric methods aim to augment in-domain training data by using data selection techniques such as
+Data centric methods have goal to collect related-domain data from existing in-domain data by using different techniques such as scoring by Language Model \cite{P10-2041}; \cite{D11-1033}; \cite{P13-2119} or by generating pseudo parallel data (\cite{P03-1010}; \cite{C16-1295}; \cite{D14-1023}).
+On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation, which can be either the training objective (\cite{Luong2015SNMT}; \cite{P16-1009}; \cite{D17-1155}; \cite{W17-3205}; \cite{D17-1156}), the NMT architecture (\cite{R17-1049}; \cite{gulcehre2016monolingual}; \cite{W17-4712}) or the decoding algorithm (\cite{gulcehre2016monolingual}; \cite{I17-2004}).
+Beside these works, we could consider
 The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041}.

### 2018-11-12 2019/naacl/naaclhlt2019.tex
naacl

@@
 \usepackage{latexsym}
-
+\usepackage{graphicx}
+\usepackage{multirow}
 \usepackage{url}
-
+\usepackage{float} 
 %\aclfinalcopy % Uncomment this line for the final submission
@@
 \begin{abstract}
+In this paper, we introduce an application of an old technique \cite{P07-1033} in Multi-Domain Neural Machine Translation.
+By defining general information encoding region and domain-specific encoding region in word embedding vector, we are able to  seperate domain-related information flows in forward and backward propagation during training
+mitigate the catastrophic interference happens between the inference of different domains in Neural Machine Translation system, i.e to improve the performance of the system for every domain in which it was trained.
 \end{abstract}
@@
 \section{Introduction}
+Despite fast improvement thanks to new architectures \cite{NIPS2017_7181} \cite{bahdanau2014neural} \cite{D13-1176} \cite{Sutskever2014Sequence}, Machine Translation still struggles with data scarcity.
+Domain adaptation, or it's special case Multi-domain, is the situation where in-domain data for example banking is rare while out-domain data such as news are plenty, one aims to leverage out-domain data to improve quality of Machine Translation system which is perfectible while being trained with small in-domain dataset.
+The problem is that we have not been able to apply Domain-adaptation to several domains at once, i.e to build one model which is best for several domains.
+It is obvious to realize that the bias to in-domain dataset makes model worse in translating other domains.
+This phenomenon can be implied by well-known problem in Machine Learning, Catastrophic Interference \cite{Michael1989Catastrophic}.
+In Multi-domain Machine Translation, one have to deal with polysemies whose meaning can be totally different in different domains, for example, the word "chair" means an household tool in "if you are taking a medicine which may cause low blood pressure when rising from a chair or bed " (extracted from medical corpora EMEA \cite{Tiedemann2009RANLP5}) while the word "chair" means to preside over a meeting in following sentence "The President of the ECB or , in his absence , the Vice-President , chairs the meetings of the Governing Council , the Executive Board and the General Council of the ECB.
+" (extracted from European Central Bank corpora \cite{Tiedemann2009RANLP5}).
+Using the same word embedding for polysemy obstructs the model from differentiating the meaning of a polysemy in one domain from it's other meaning in the other domain.
 
@@
 There is not any changes in the architecture of Neural Machine Translation network except the construction of word embedding. 
-
+\begin{figure}[h]
+\center
+    \includegraphics[width=0.8\linewidth]{Sparse1}
+    \caption{Illustration of the model.} 
+    \label{network}
+\end{figure}
 \section{Experiments}
@@
 \subsection{Corpora}
+To evaluate the performance of proposed architecture, we use 4 corpora for training: European Medicines Agency; European Central Bank, News-commentary and European Parliament \cite{Tiedemann2009RANLP5} corresponding to 4 domains: medical, banking, news and administrative  respectively and one non-domain corpora Common-crawl  which was introduced in Shared Task: Machine Translation of News the Third Conference On Machine Translation 2018(WMT2018).
+We use Khresmoi-dev, a test set used in the Biomedical Translation Task (WMT2018), newstest2008, and test2006 as validation set for domains: medical, news administrative; except for banking domain, we use a subset of size 2000 extracted from original corpora and excluded from training set.
+To test the model, we use Khresmoi-test; newstest2009 and test2007 for domains: medical, news administrative; except for banking domain, we use a subset of size 2000 extracted from original corpora and excluded from training set.
+The information of used corpora is presented in table \ref{table:Corpora}
+\begin{table}[H]
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{ |l|l|l|l|l|l| }
+\hline
+Task & Corpora & Train & Dev & Test & Vocabulary \\ \hline
+\multirow{5}{*}{$English \rightarrow French$} & European Medicines Agency & 1092568 & 500 & 1000 & \\
+ & European Central Bank & 195960 & 2000 & 2000 &\\
+ & News-commentary & 258432 & 2051 & 3027 & \\
+ & European Parliament & 2007723 & 2000 & 2000 & \\
+ & Common-crawl & 3244152 &  &  & \\ \hline
+\multirow{5}{*}{$English \rightarrow German$} & European Medicines Agency & 1108752 & 500 & 1000 & \\
+ & European Central Bank & 113174 & 2000 & 2000 &\\
+ & News-commentary & 270769 & 2051 & 2525 & \\
+ & European Parliament & 1920209 & 2000 & 2000 & \\ 
+ & Common-crawl & 2399123 &  &  & \\ \hline
+\hline
+\end{tabular}
+}
+\caption{Corpora}
+\label{table:Corpora}
+\end{table}
 

### 2018-11-13 2019/naacl/naaclhlt2019.tex
naacl

@@
 \section{Related Work}
-Multi-domain Machine Translation has largely interested NLP community by its promising application in industry and its relation to fundamental problem in theory.
+Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
 Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}.
-Data centric methods have goal to collect related-domain data from existing in-domain data by using different techniques such as scoring by Language Model \cite{P10-2041}; \cite{D11-1033}; \cite{P13-2119} or by generating pseudo parallel data (\cite{P03-1010}; \cite{C16-1295}; \cite{D14-1023}).
-On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation, which can be either the training objective (\cite{Luong2015SNMT}; \cite{P16-1009}; \cite{D17-1155}; \cite{W17-3205}; \cite{D17-1156}), the NMT architecture (\cite{R17-1049}; \cite{gulcehre2016monolingual}; \cite{W17-4712}) or the decoding algorithm (\cite{gulcehre2016monolingual}; \cite{I17-2004}).
-Beside these works, we could consider
+Data centric methods have goal to collect related-domain data from existing in-domain data by using different techniques such as scoring relatedness by Language Model (\cite{P10-2041}; \cite{D11-1033}; \cite{P13-2119} or by metric on the space of sentence embedding \cite{P17-2089} or generating pseudo parallel data \cite{P03-1010}; \cite{C16-1295}; \cite{D14-1023}.
+On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation.
+The novelty can be either trainining objective \cite{Luong2015SNMT}; \cite{P16-1009}; \cite{D17-1155}; \cite{W17-3205}; \cite{D17-1156}; \cite{C18-1269} or architecture \cite{R17-1049}; \cite{gulcehre2016monolingual}; \cite{W17-4712}, \cite{Biao2017CARENMT}; \cite{N18-2080}; \cite{W18-6313}; \cite{C16-1170}; \cite{P18-2050} or the decoding algorithm \cite{gulcehre2016monolingual}; \cite{I17-2004}.
+Beside these works, we could also consider "out-of-domain" works such as sequence labeling tasks \cite{P07-1033}; learning multiple visual domains \cite{NIPS2017_6654}.
 The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041}.

### 2019-04-08 2019/emnlp/sparse.tex
wmt,emnlp

--- 2018-11-13 2019/naacl/naaclhlt2019.tex
@@
 \usepackage{times}
+\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
 \usepackage{latexsym}
@@
 \usepackage{float} 
+\usepackage{tabularx}
+\newtheorem{lemma}{Lemma}[section]
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{proposition}{Proposition}[section]
+\newtheorem{definition}{Definition}[section]
+\DeclareMathOperator*{\argmax}{argmax}
 %\aclfinalcopy % Uncomment this line for the final submission
@@
 % in the camera-ready version and ask you to change it back.
-
-\newcommand\BibTeX{B{\sc ib}\TeX}
 
@@
 In this paper, we introduce an application of an old technique \cite{P07-1033} in Multi-Domain Neural Machine Translation.
-By defining general information encoding region and domain-specific encoding region in word embedding vector, we are able to  seperate domain-related information flows in forward and backward propagation during training
+By defining general information encoding region and domain-specific encoding region in word embedding vector, we are able to  seperate domain-related information flows in forward and backward propagation during training thus to
 mitigate the catastrophic interference happens between the inference of different domains in Neural Machine Translation system, i.e to improve the performance of the system for every domain in which it was trained.
@@
 On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation.
-The novelty can be either trainining objective \cite{Luong2015SNMT}; \cite{P16-1009}; \cite{D17-1155}; \cite{W17-3205}; \cite{D17-1156}; \cite{C18-1269} or architecture \cite{R17-1049}; \cite{gulcehre2016monolingual}; \cite{W17-4712}, \cite{Biao2017CARENMT}; \cite{N18-2080}; \cite{W18-6313}; \cite{C16-1170}; \cite{P18-2050} or the decoding algorithm \cite{gulcehre2016monolingual}; \cite{I17-2004}.
+The novelty can be either trainining objective; for example \cite{Luong2015SNMT}; \cite{P16-1009}; \cite{D17-1155}; \cite{W17-3205}; \cite{D17-1156}; \cite{C18-1269} or architecture \cite{R17-1049}; \cite{gulcehre2016monolingual}; \cite{W17-4712}, \cite{Biao2017CARENMT}; \cite{N18-2080}; \cite{W18-6313}; \cite{C16-1170}; \cite{P18-2050} or the decoding algorithm \cite{gulcehre2016monolingual}; \cite{I17-2004}.
 Beside these works, we could also consider "out-of-domain" works such as sequence labeling tasks \cite{P07-1033}; learning multiple visual domains \cite{NIPS2017_6654}.
@@
 \section{The Model}
-There is not any changes in the architecture of Neural Machine Translation network except the construction of word embedding. 
+\begin{definition}{Domain}
+\label{def:1}
+Domain is a hypothetical probability that generates a set of corpora. We will denote D(C) is domain of set of corpora C.
+\end{definition}
+
+In machine translation, the loss function for training is
+\begin{equation}
+\begin{split}
+E_{(x,y) \sim p_{rel}}[-log(p_{\theta}(y|x))] &= \\
+= E_{x \sim p_{rel}(x)}[KL(p_{rel}(y|x) & \mid p_{\theta}(y|x))] + \\
++ E_{x \sim p_{rel}} &[H(p_{rel}(y|x))]
+\end{split}
+\end{equation}.
+Where $p_{\theta}(y|x)) = p(y|x,\theta)$ or $p_{\theta}(y|x)) = p(y,\theta | x)$
+ In this quantity, domain is exactly $p_{rel}(x,y)$.
+However, in the inference or translation, the most important goal is to predict target sentence given source sentence, i.e, to well approximate the conditional probability $p_{rel}(y|x)$.
+However, this relation could be very different if consider 2 domains generating 2 corpora of different topic.
+Indeed, in financial document, to translate "security" into "titre financier" is more likely than to translate it into "securit\'e" but in IT document.
+In reality, to translate correctly with respect to the topic of sentence is very important.
+We formalize the multi-domain problem for machine translation in following equation
+\begin{equation}
+\begin{split}
+E_{(x,y) \sim D(C_{i})}[-log(p_{\theta^*}(y|x))] & = \\
+= \argmax_{\theta} E_{(x,y) \sim D(C_{i})} [-log & (p_{\theta}(y|x))] \; \forall i \in [1..d]
+\end{split}
+\end{equation}
+A trivial solution for this equation is $\theta^* = (\theta^*_{0},...,\theta^*_{d})$ and $p_{(\theta_{0},...,\theta_d)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times C(D_i) \rightarrow [0,1]: f_{\theta_i}(y,x) = p_{(\theta_{0},...,\theta_d)}(y|x,i)$ where $\Omega_i$ is set of all possible values of $\theta_i$.
+\\
+The model above is not different than ensemble of $d$ models specialized for one domain.
+However, we could mitigate this seperation by using shared paramaters denoted $\theta_s$ then reformulate the model as following: \\
+$\theta^* = (\theta^*_{0},...,\theta^*_{d}, \theta_s^*)$ and $p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times \Omega_s \times C(D_i) \rightarrow [0,1]: f_{\theta_i, \theta_s}(y,x) = p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ where $\Omega_i$ is set of all possible values of $\theta_i$ and $\Omega_s$ is set of all possible values of $\theta_s$.
+\\
+However, there is a problem that the couple $\theta_s^*$ and $\theta_i^*$ have to be optimal $\forall i \in [1..d]$.\\
+Our first approach to multi-domain machine translation is to build such trivial solution in calibrating positions and sizes of domain paramters $\theta_i$ and shared parameters $\theta_s$ inside one architecture which is harly achieved.
+Despite that fact, we could consider $\theta_i$ is additional factor used for fine-tuning the model to the domain $D(C_{i})$, i.e, lowering the perplexity with respect to true distribution of domain $E_{(x,y) \sim D(C_{i})}[-log(p_{\theta}(y|x,i))]$.
 \begin{figure}[h]
@@
 \multirow{5}{*}{$English \rightarrow French$} & European Medicines Agency & 1092568 & 500 & 1000 & \\
- & European Central Bank & 195960 & 2000 & 2000 &\\
+ & European Central Bank & 191960 & 2000 & 2000 &\\
  & News-commentary & 258432 & 2051 & 3027 & \\
@@
 \multirow{5}{*}{$English \rightarrow German$} & European Medicines Agency & 1108752 & 500 & 1000 & \\
- & European Central Bank & 113174 & 2000 & 2000 &\\
+ & European Central Bank & 109174 & 2000 & 2000 &\\
  & News-commentary & 270769 & 2051 & 2525 & \\
@@
 \subsection{Results}
+\begin{figure}[h]
+\caption{BLEU score in language pair English - French}
+\begin{center}
+ \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
+ \hline
+ Models & EMEA test & EPPS test & ECB test & khresmoi & test 2007 & newstest 2009 & IWSLT 2010 \\ [0.5ex] 
+ \hline\hline
+ Mixed & & & & & & & \\
+ \hline
+ EMEA-finetuned & & & & & & & \\
+ \hline
+ EPPS-finetuned & & & & & & & \\
+ \hline
+ ECB-finetuned & & & & & & & \\
+ \hline
+ Sparse & & & & & & & \\
+ \hline
+ Domain Control & & & & & & & \\
+ \hline
+\end{tabularx}}
+\end{center}
+\label{tab:1}
+\end{figure}
 
+\begin{figure}[h]
+\caption{BLEU score in language pair English - German}
+\begin{center}
+ \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
+ \hline
+ Models & EMEA test & EPPS test & ECB test & khresmoi & test 2007 & newstest 2009 & IWSLT 2010 \\ [0.5ex] 
+ \hline\hline
+ Mixed & & & & & & & \\
+ \hline
+ EMEA-finetuned & & & & & & & \\
+ \hline
+ EPPS-finetuned & & & & & & & \\
+ \hline
+ ECB-finetuned & & & & & & & \\
+ \hline
+ Sparse & & & & & & & \\
+ \hline
+ Domain Control & & & & & & & \\
+ \hline
+\end{tabularx}}
+\end{center}
+\label{tab:2}
+\end{figure}
 \section{Conclusions}
@@
 \bibliography{naaclhlt2019}
-\bibliographystyle{acl_natbib}
+\bibliographystyle{plainnat}
 

### 2019-04-08 2019/emnlp/sparse.tex
2019/emnlp/sparse.tex

@@
 \DeclareMathOperator*{\argmax}{argmax}
+\DeclareMathOperator*{\argmin}{argmin}
 %\aclfinalcopy % Uncomment this line for the final submission
@@
 \begin{abstract}
+Finetuning a generic neural machine translation model to a certain domain usually degrade its performance on the other domains, thus it's very difficult to build a unique model which is best for all.
+Indeed, Catastrophic Forgetting, well-known problem in Machine Learning, states that standard backprogation neural network suffers a lot while there is a shift in training data distribution.
+However, Finetuning is very costly and unefficient in the industry therefore one best model is always desired.
 In this paper, we introduce an application of an old technique \cite{P07-1033} in Multi-Domain Neural Machine Translation.
 By defining general information encoding region and domain-specific encoding region in word embedding vector, we are able to seperate domain-related information flows in forward and backward propagation during training thus to
-mitigate the catastrophic interference happens between the inference of different domains in Neural Machine Translation system, i.e to improve the performance of the system for every domain in which it was trained.
+mitigate the catastrophic interference happens during the training over different domains.
+Empirically, we achieved a unique model that has comparable score to finetuned models for every domains.
+Furthermore, our method is simple and could be applied for any architecture.
 \end{abstract}
@@
 \section{Introduction}
-Despite fast improvement thanks to new architectures \cite{NIPS2017_7181} \cite{bahdanau2014neural} \cite{D13-1176} \cite{Sutskever2014Sequence}, Machine Translation still struggles with data scarcity.
-Domain adaptation, or it's special case Multi-domain, is the situation where in-domain data for example banking is rare while out-domain data such as news are plenty, one aims to leverage out-domain data to improve quality of Machine Translation system which is perfectible while being trained with small in-domain dataset.
-The problem is that we have not been able to apply Domain-adaptation to several domains at once, i.e to build one model which is best for several domains.
-It is obvious to realize that the bias to in-domain dataset makes model worse in translating other domains.
+Despite fast developpemnt thanks to new architectures \cite{NIPS2017_7181} \cite{bahdanau2014neural} \cite{D13-1176} \cite{Sutskever2014Sequence}, Machine Translation still struggles with data scarcity.
+Domain adaptation, or it's special case Multi-domain, is the situation where in-domain data (such as banking) is rare while out-domain data such as news are plenty, one aims to leverage out-domain data to improve quality of Machine Translation system which is perfectible while being trained with only the in-domain dataset.
+Traditionally, ones train a generic model using a very large corpora then use in-domain dataset to finetune the model.
+It is ubiquitous that the bias to the in-domain dataset used in finetuning makes model worse in translating other domains, thus that we have not been able to finetune a model to several domains at once.
 This phenomenon can be implied by well-known problem in Machine Learning, Catastrophic Interference \cite{Michael1989Catastrophic}.
-In Multi-domain Machine Translation, one have to deal with polysemies whose meaning can be totally different in different domains, for example, the word "chair" means an household tool in "if you are taking a medicine which may cause low blood pressure when rising from a chair or bed " (extracted from medical corpora EMEA \cite{Tiedemann2009RANLP5}) while the word "chair" means to preside over a meeting in following sentence "The President of the ECB or , in his absence , the Vice-President , chairs the meetings of the Governing Council , the Executive Board and the General Council of the ECB.
+When there is a shift in the distribution of training data, standard backpropagation neural network gradually lose information learned from other domains while learning only target domain.\\
+In Multi-domain Machine Translation, ones have to deal with polysemies whose meaning can be totally different in different domains, for example, the word "chair" means an household tool in "if you are taking a medicine which may cause low blood pressure when rising from a chair or bed " (extracted from medical corpora EMEA \cite{Tiedemann2009RANLP5}) while the word "chair" means to preside over a meeting in following sentence "The President of the ECB or , in his absence , the Vice-President , chairs the meetings of the Governing Council , the Executive Board and the General Council of the ECB.
 " (extracted from European Central Bank corpora \cite{Tiedemann2009RANLP5}).
-Using the same word embedding for polysemy obstructs the model from differentiating the meaning of a polysemy in one domain from it's other meaning in the other domain.
+Using context information is not sufficient to differ these polysemies.
+Therefore using the same word embedding for polysemies obstructs the model from differentiating the meaning of a polysemy in one domain from it's other meaning in the other domain.
+\\
+\begin{table}[h]
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{|l|l|l|}
+\hline
+English Words & French translation & Domain \\ \hline
+security & titre financier & Financial \\
+security & s\'ecuti\'e & IT\\
+joint & articulaire & medical\\
+joint & conjoint & banking\\
+chair & chaise & general \\
+chair & pr\'esider, pr\'esident & banking \\
+\hline
+\end{tabular}
+}
+\caption{Polysemies}
+\label{table:polysemy}
+\end{table}
 
+\begin{table}[h]
+\resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||}
+\hline
+English sentence & French translation & Domain \\ 
+\hline
+"Flu-like" symptoms such as headaches, joint pains, feelings of weakness, dizziness, and tiredness may occur, especially at the start of treatment.
+& Des symptômes grippaux tels que céphalées, douleurs articulaires, sensation de faiblesse, vertige et asthénie, peuvent survenir, en particulier en début de traitement.
+& medical \\
+\hline
+\end{tabularx}
+}
+\caption{Polysemies}
+\label{table:polysemy}
+\end{table}
+
+In this work, we present a new design for word embedding that is composed of different regions which are activate and deactivate corresponding to domains.
+To evaluate, we use several test sets representing each domains and two other general test sets.
 \section{Related Work}
@@
 Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}.
-Data centric methods have goal to collect related-domain data from existing in-domain data by using different techniques such as scoring relatedness by Language Model (\cite{P10-2041}; \cite{D11-1033}; \cite{P13-2119} or by metric on the space of sentence embedding \cite{P17-2089} or generating pseudo parallel data \cite{P03-1010}; \cite{C16-1295}; \cite{D14-1023}.
+Data centric methods have goal to collect related-domain data from existing in-domain data by using different techniques such as scoring relatedness by Language Model \cite{P10-2041}; \cite{D11-1033}; \cite{P13-2119} or by metric on the space of sentence embedding \cite{P17-2089} or generating pseudo parallel data \cite{P03-1010}; \cite{C16-1295}; \cite{D14-1023}.
 On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation.
@@
 
-\section{The Model}
+\section{Proposed method}
+\subsection{Problem statement}
 \begin{definition}{Domain}
@@
 \begin{split}
-E_{(x,y) \sim D(C_{i})}[-log(p_{\theta^*}(y|x))] & = \\
-= \argmax_{\theta} E_{(x,y) \sim D(C_{i})} [-log & (p_{\theta}(y|x))] \; \forall i \in [1..d]
+\theta^* = \argmin_{\theta} E_{(x,y) \sim D(C_{i})} [-log & (p_{\theta}(y|x,i))] \; \forall i \in [1..d]
 \end{split}
@@
 However, we could mitigate this seperation by using shared paramaters denoted $\theta_s$ then reformulate the model as following: \\
-$\theta^* = (\theta^*_{0},...,\theta^*_{d}, \theta_s^*)$ and $p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times \Omega_s \times C(D_i) \rightarrow [0,1]: f_{\theta_i, \theta_s}(y,x) = p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ where $\Omega_i$ is set of all possible values of $\theta_i$ and $\Omega_s$ is set of all possible values of $\theta_s$.
-\\
+$\theta^* = (\theta^*_{0},...,\theta^*_{d}, \theta_s^*)$ and $p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times \Omega_s \times C(D_i) \rightarrow [0,1]: f_{\theta_i, \theta_s}(y,x) = p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ \\ $p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x)$ is independent of $\theta_{j}$ $\forall j$ i.e, $\exists \bar{f}:\Omega_s \times C \rightarrow [0,1]: \bar{f}_{\theta_s}(y,x) = p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x)$
+where $\Omega_i$ is set of all possible values of $\theta_i$ and $\Omega_s$ is set of all possible values of $\theta_s$. \\
 However, there is a problem that the couple $\theta_s^*$ and $\theta_i^*$ have to be optimal $\forall i \in [1..d]$.\\
-Our first approach to multi-domain machine translation is to build such trivial solution in calibrating positions and sizes of domain paramters $\theta_i$ and shared parameters $\theta_s$ inside one architecture which is harly achieved.
-Despite that fact, we could consider $\theta_i$ is additional factor used for fine-tuning the model to the domain $D(C_{i})$, i.e, lowering the perplexity with respect to true distribution of domain $E_{(x,y) \sim D(C_{i})}[-log(p_{\theta}(y|x,i))]$.
+Despite the fact that $\theta^* = (\theta^*_{0},...,\theta^*_{d}, \theta_s^*)$ is hardly achieved by backpropagation, we could attempt to optimize $\theta_s$:
+\begin{equation}
+\theta^*_{s} = \displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))]
+\end{equation} 
+then optimize each $\theta_i$ (as additional factor used for fine-tuning the model to the domain $D(C_{i})$) the perplexity with respect to true distribution of domain
+\begin{equation}
+\theta^*_{i} = \displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
+\end{equation} 
+Our first approach to multi-domain machine translation is to build such trivial solution in calibrating positions and sizes of domain paramters $\theta_i$ and shared parameters $\theta_s$ inside one architecture.
+\subsection{Sparse word embedding}
+This section explains a new design for word embedding used in multi-domain model.
+As mentioned above, we separate parameter set into $D+1$ parameter sets correspond to D domains and one shared paramater set.
+Our first attempt is to split standard vector word embedding into $D+1$ regions named $\theta_i$ for $i \in [1,..,D]$ and $\theta_s$.
 \begin{figure}[h]
@@
     \includegraphics[width=0.8\linewidth]{Sparse1}
-    \caption{Illustration of the model.} 
+    \caption{Sparse word embedding.
+The red region contains activate units used only when presenting its correpondence domain, the other green regions are deactivate, the bleu generic region is always activate}
     \label{network}
 \end{figure}
+The figure \ref{network} presents our proposed design.
+During backprogation, the rotation matrix receive vanished gradient at regions corresponding to deactivate regions in word embedding.
+Those regions are also vanished in forward step, thus do not interfere the training on the domain to which they does not correspond.\\
+Denote $p_{t} = P * w_{t}$ where $w_{t}$ is word embedding of token $x_{t}$  is projection of word embedding (Transformer \cite{NIPS2017_7181}, Recurrent \cite{bahdanau2014neural})
+Assume $x_{t}$ is in domain i, thus, $w_j(t) = 0 \forall j \neq i$. Denote $P_j$ are rows of projection matrix $P$ corresponding to deactivate region $w_j$.
+\begin{equation}
+\begin{split}
+\frac{\partial L(y_t,x_t)}{\partial P_j} &=\frac{\partial L}{\partial p}(p_t) * \frac{\partial p_t}{\partial P_j} \\
+								&= \frac{\partial L}{\partial p}(p_t) * w_j(t)\\
+									&=0
+\end{split} 
+\forall j \neq i
+\end{equation}
+\subsection{Detailed implementation}
+The complex structure of State-of-the-art architecture Transformer \cite{NIPS2017_7181} and popular recurrent network \cite{bahdanau2014neural} makes the implementation of sparse embedding non-obvious.
+Indeed, Transformer involve positional embedding and residual connections which are not homogeneous for domain-regions.
+Therefore, in our first implementation, we used an dense layer to fuse domain-region and generic region before applying positional embedding and residual connection.
+This implementation limits the finetuning effect at only word level.\\
+Furthermore, because word embeddings are involved in three parts of standard encoder-decoder structure, which are input of encoder, input of decoder and projection before softmax, we have 7 possibilities to apply sparse embedding.
+However, due to high memory cost, we only want to apply sparse embedding in encoder.
 \section{Experiments}
-
-\subsection{Corpora}
+\subsection{Domains and data}
 To evaluate the performance of proposed architecture, we use 4 corpora for training: European Medicines Agency; European Central Bank, News-commentary and European Parliament \cite{Tiedemann2009RANLP5} corresponding to 4 domains: medical, banking, news and administrative  respectively and one non-domain corpora Common-crawl  which was introduced in Shared Task: Machine Translation of News the Third Conference On Machine Translation 2018(WMT2018).
@@
 The information of used corpora is presented in table \ref{table:Corpora}
+
 \begin{table}[H]
@@
 \subsection{Results}
-\begin{figure}[h]
-\caption{BLEU score in language pair English - French}
+\begin{figure}[H]
 \begin{center}
@@
  \hline
- Models & EMEA test & EPPS test & ECB test & khresmoi & test 2007 & newstest 2009 & IWSLT 2010 \\ [0.5ex] 
+ Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT 2010 \\ [0.5ex] 
  \hline\hline
- Mixed & & & & & & & \\
+ Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 & 25.7 \\
  \hline
@@
  \hline
- Sparse & & & & & & & \\
+ Sparse(1) & 70.45 & 38.23 & 54.69 & 34.66 & 31.24 & 21.36 & 26.02 \\
  \hline
- Domain Control & & & & & & & \\
+ Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & 26.02 \\
+ \hline
+ Sparse(3) &  &  &  &  &  &  & \\
+ \hline
+ Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 & & \\
  \hline
@@
 \end{center}
+\caption{bpe-detokenized BLEU score in language pair English - French.
+Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT.
+Sparse(2) uses generic tag for all of test sets.
+Sparse(3) uses tag predicted by pretrained domain classififer}
 \label{tab:1}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \caption{BLEU score in language pair English - German}
@@
  \hline
- Models & EMEA test & EPPS test & ECB test & khresmoi & test 2007 & newstest 2009 & IWSLT 2010 \\ [0.5ex] 
+ Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT 2010 \\ [0.5ex] 
  \hline\hline

### 2019-04-09 2019/emnlp/sparse.tex
new

@@
 \subsection{Domains and data}
-To evaluate the performance of proposed architecture, we use 4 corpora for training: European Medicines Agency; European Central Bank, News-commentary and European Parliament \cite{Tiedemann2009RANLP5} corresponding to 4 domains: medical, banking, news and administrative  respectively and one non-domain corpora Common-crawl  which was introduced in Shared Task: Machine Translation of News the Third Conference On Machine Translation 2018(WMT2018).
-We use Khresmoi-dev, a test set used in the Biomedical Translation Task (WMT2018), newstest2008, and test2006 as validation set for domains: medical, news administrative; except for banking domain, we use a subset of size 2000 extracted from original corpora and excluded from training set.
-To test the model, we use Khresmoi-test; newstest2009 and test2007 for domains: medical, news administrative; except for banking domain, we use a subset of size 2000 extracted from original corpora and excluded from training set.
-The information of used corpora is presented in table \ref{table:Corpora}
+To evaluate the performance of proposed architecture, we design experiments in which we want to train an NMT for 3 domains such as medical, administrative documents and business briefly denoted by EMEA, EPPS, ECB according to training corpora name respectively.
+We use 3 corpora for training: European Medicines Agency; European Central Bank and European Parliament \cite{Tiedemann2009RANLP5}.
+We randomly split those 3 corpora into training, validation and test set with size described in table \ref{table:Corpora}.
+We use validation sets for choosing the best model according to the average BLEU score on 3 validation sets.
+We also realize that the test sets origined from same corpora as training sets are likely easy for trained model, thus can not provide fair evaluation.
+Therefore, we we use Khresmoi-test; newstest2009 and test2007 for domains: EMEA, EPPS.
+Unfortunately, we could not find any official test sets for ECB domain.
+We also want to evaluate the generalization of trained model.
+The model is also required to achieve comparable performance to generic model.
+To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
 
@@
 Task & Corpora & Train & Dev & Test & Vocabulary \\ \hline
-\multirow{5}{*}{$English \rightarrow French$} & European Medicines Agency & 1092568 & 500 & 1000 & \\
- & European Central Bank & 191960 & 2000 & 2000 &\\
- & News-commentary & 258432 & 2051 & 3027 & \\
- & European Parliament & 2007723 & 2000 & 2000 & \\
- & Common-crawl & 3244152 &  &  & \\ \hline
-\multirow{5}{*}{$English \rightarrow German$} & European Medicines Agency & 1108752 & 500 & 1000 & \\
- & European Central Bank & 109174 & 2000 & 2000 &\\
- & News-commentary & 270769 & 2051 & 2525 & \\
- & European Parliament & 1920209 & 2000 & 2000 & \\ 
- & Common-crawl & 2399123 &  &  & \\ \hline
+\multirow{5}{*}{$English \rightarrow French$} & European Medicines Agency & 1090568 & 1000 & 1000 & English: 30164, French: 30398\\
+ & European Central Bank & 191960 & 1000 & 1000 &\\
+ & European Parliament & 2005723 & 1000 & 1000 & \\ \hline
+\multirow{5}{*}{$English \rightarrow German$} & European Medicines Agency & 1108752 & 1000 & 1000 & English: 30159, German: 30698\\
+ & European Central Bank & 111174 & 1000 & 1000 &\\
+ & European Parliament & 1920209 & 1000 & 1000 & \\ \hline
 \hline
@@
 \end{table}
-
-\subsection{NMT engine}
-
+\subsection{NMT model configuration}
+We evaluate our method on 2 popular architecture: Transformer of and Attentional Birectional Long Short-Term Memmory (LSTM).\\
+More precisely, for Transformer we use the same setting as described in \cite{NIPS2017_7181}.For Attentional Birectional Long Short-Term Memmory, in order to fairly compare to the work of \cite{D18-1041}, we use standard configuration with: 1 layer bidirectional LSTM with hidden size 1024 in encoder; 1 layer attentional unidirectional LSTM with same hidden size in decoder; and word embedding of size 512.
+The code use Tensorflow and OpenNMT \cite{P17-4012}.\\
+For sparse embedding implementation, we split vector embedding of size 512 into 4 regions $[8,8,8,488]$ corresponding to EMEA, EPPS, ECB domain and generic region.
+If sentence comes from domain $i$, Domain regions $j,\forall j \neq i$ will be masked as 0.
+We then use a dense layer of size 512 to fuse active domain region and generic region.
 \subsection{Results}
@@
  \hline
- Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT 2010 \\ [0.5ex] 
+ Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT test 2010 \\ [0.5ex] 
  \hline\hline
@@
  \hline
- EMEA-finetuned & & & & & & & \\
+ EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.10 \\
  \hline
- EPPS-finetuned & & & & & & & \\
+ EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 & 11.1 \\
  \hline
- ECB-finetuned & & & & & & & \\
+ ECB-finetuned & 9.2 & 18.39 & 55.45 & 12.13 & 16.12 & 8.71 & 7.19 \\
  \hline
@@
 \begin{figure}[H]
-\caption{BLEU score in language pair English - German}
 \begin{center}
@@
 \end{center}
+\caption{BLEU score for Transformer in language pair English - German. The notation is same as \ref{tab:1}}
 \label{tab:2}
 \end{figure}
+
+\begin{figure}[H]
+\begin{center}
+ \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
+ \hline
+ Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT test 2010 \\ [0.5ex] 
+ \hline\hline
+ Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 & 25.7 \\
+ \hline
+ EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.10 \\
+ \hline
+ EPPS-finetuned & & & & & & & \\
+ \hline
+ ECB-finetuned & & & & & & & \\
+ \hline
+ Sparse(1) & 70.45 & 38.23 & 54.69 & 34.66 & 31.24 & 21.36 & 26.02 \\
+ \hline
+ Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & 26.02 \\
+ \hline
+ Sparse(3) &  &  &  &  &  &  & \\
+ \hline
+ Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 & & \\
+ \hline
+ WDCMT (conditional GRU) & & & & & & & \\
+ \hline
+\end{tabularx}}
+\end{center}
+\caption{bpe-detokenized BLEU score for Recurrent network architecture on language pair English-French.
+Notations as same as \ref{tab:1} except that other models from WDCMT of \cite{D18-1041} use Bidirectional-LTSM}
+\label{tab:3}
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+ \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
+ \hline
+ Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT test 2010 \\ [0.5ex] 
+ \hline\hline
+ Mixed & & & & & & & \\
+ \hline
+ EMEA-finetuned & & & & & & & \\
+ \hline
+ EPPS-finetuned & & & & & & & \\
+ \hline
+ ECB-finetuned & & & & & & & \\
+ \hline
+ Sparse & & & & & & & \\
+ \hline
+ Domain Control & & & & & & & \\
+ \hline
+ WDCMT (conditional GRU) & & & & & & & \\
+ \hline 
+\end{tabularx}}
+\end{center}
+\caption{BLEU score for Recurrent network on language pair English - German. The notation is same as \ref{tab:3}}
+\label{tab:4}
+\end{figure}
+
 \section{Conclusions}

### 2019-04-15 2019/emnlp/sparse.tex
new

@@
 \DeclareMathOperator*{\argmin}{argmin}
+\usepackage{algorithm}
+\usepackage{algorithmic}
+\floatname{algorithm}{Procedure}
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
 %\aclfinalcopy % Uncomment this line for the final submission
@@
 More precisely, for Transformer we use the same setting as described in \cite{NIPS2017_7181}.For Attentional Birectional Long Short-Term Memmory, in order to fairly compare to the work of \cite{D18-1041}, we use standard configuration with: 1 layer bidirectional LSTM with hidden size 1024 in encoder; 1 layer attentional unidirectional LSTM with same hidden size in decoder; and word embedding of size 512.
-The code use Tensorflow and OpenNMT \cite{P17-4012}.\\
+The implementation can be found in OpenNMT-tf \cite{P17-4012}.\\
 For sparse embedding implementation, we split vector embedding of size 512 into 4 regions $[8,8,8,488]$ corresponding to EMEA, EPPS, ECB domain and generic region.
@@
 We then use a dense layer of size 512 to fuse active domain region and generic region.
+
+\begin{algorithm}[H]
+\caption{Multi-domain Training}
+\label{algo:1}
+\begin{algorithmic}[1]
+\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for d domains}
+\REPEAT 
+\STATE{Randomly pick $i \in [1,..,d]$ w.r.t multinomial distribution $[\frac{|C_i|}{\displaystyle{\mathop{\sum}_{i\in [1,..,d]}|C_i|}}]$.}
+\STATE{Randomly pick a batch of sentences from corpora $C_i$.}
+\STATE{Using only generic activate embedding and domain activate embedding for this batch of sentences, i.e, we will have 2 batches of word embedding.}
+\STATE{Pass the batch of word embedding to a dense layer to fuse regions.}
+\STATE{Apply backpropagation as usual.}
+\UNTIL{convergence}
+\end{algorithmic}
+\end{algorithm}
+
+Each iteration of algorithm \ref{algo:1}, the batch is doubled because it contains 2 types of word embedding for each sentence (using only generic region; using generic region and active domain region).
+The reason to do this is to minimize simultaneously both losses \\
+$\theta^*_{s} = \displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))]$ \\ 
+$ \theta^*_{i} =\displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
+$ 
+We admit that this is not the optimal way. However, this procdedure make the training very simple.\\
+We also take in account the problem of overfitting that could occur for small corpora.
+Indeed, real data are not always balanced for every domain.
+For instance, in our experiments, the size of corpora are very different where EPPS is largest with approximately 2 millions sentences, EMEA contains about 1 million sentences and ECB is the smallest with only 200 000 sentences for English French and 100 000 for English German.
+If we train each corpora with same number of batches, ECB will be trained too oftenly, that easily leads to overfitting.
+Therefore, we perform backpropagation on corpora with frequencies according to sizes of corpora.
 \subsection{Results}

### 2019-04-18 2019/emnlp/sparse.tex
new

@@
 \usepackage{tabularx}
+\usepackage{xcolor}
 \newtheorem{lemma}{Lemma}[section]
@@
 security & s\'ecuti\'e & IT\\
+\hline
 joint & articulaire & medical\\
 joint & conjoint & banking\\
+\hline
 chair & chaise & general \\
@@
 \hline
-English sentence & French translation & Domain \\ 
+English sentence & French translation & Corpora \\ 
 \hline
-"Flu-like" symptoms such as headaches, joint pains, feelings of weakness, dizziness, and tiredness may occur, especially at the start of treatment.
-& Des symptômes grippaux tels que céphalées, douleurs articulaires, sensation de faiblesse, vertige et asthénie, peuvent survenir, en particulier en début de traitement.
-& medical \\
+"Flu-like" symptoms such as headaches, \colorbox{red}{joint} pains, feelings of weakness, dizziness, and tiredness may occur, especially at the start of treatment.
+& Des symptômes grippaux tels que céphalées, douleurs \colorbox{red}{articulaires}, sensation de faiblesse, vertige et asthénie, peuvent survenir, en particulier en début de traitement.
+& EMEA \\
+\hline
+In my opinion, Mr President, this \colorbox{red}{joint} debate fully bears this out. & Cette discussion \colorbox{red}{commune} en est, Monsieur le Président, la preuve vivante. & EPPS \\
+\hline
+The increasing number of central banks using the same IT platform for their own purposes has provided unique opportunities to undertake \colorbox{red}{joint} projects , exploit synergies and better manage the relationship with the vendor.
+& Le nombre croissant de banques centrales utilisant la même plate-forme informatique pour leurs besoins propres a offert des occasions uniques de lancer des projets \colorbox{red}{communs}, d'exploiter des synergies et de mieux gérer les relations avec le fournisseur.
+& ECB \\
 \hline
@@
 }
-\caption{Polysemies}
+\caption{Translation}
 \label{table:polysemy}
@@
 Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
-Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}.
-Data centric methods have goal to collect related-domain data from existing in-domain data by using different techniques such as scoring relatedness by Language Model \cite{P10-2041}; \cite{D11-1033}; \cite{P13-2119} or by metric on the space of sentence embedding \cite{P17-2089} or generating pseudo parallel data \cite{P03-1010}; \cite{C16-1295}; \cite{D14-1023}.
-On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation.
-The novelty can be either trainining objective; for example \cite{Luong2015SNMT}; \cite{P16-1009}; \cite{D17-1155}; \cite{W17-3205}; \cite{D17-1156}; \cite{C18-1269} or architecture \cite{R17-1049}; \cite{gulcehre2016monolingual}; \cite{W17-4712}, \cite{Biao2017CARENMT}; \cite{N18-2080}; \cite{W18-6313}; \cite{C16-1170}; \cite{P18-2050} or the decoding algorithm \cite{gulcehre2016monolingual}; \cite{I17-2004}.
-Beside these works, we could also consider "out-of-domain" works such as sequence labeling tasks \cite{P07-1033}; learning multiple visual domains \cite{NIPS2017_6654}.
+Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
+The methods in this category can be listed as following:
+\begin{itemize}
+	\item score relatedness of sentences to the target domain by Language Model \cite{P10-2041}; \cite{D11-1033}; \cite{P13-2119}.
+	\item generating pseudo parallel data \cite{P03-1010}; \cite{C16-1295}; \cite{D14-1023}
+\end{itemize}
+On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features
+\begin{itemize}
+	\item Trainining objective \cite{Luong2015SNMT}; \cite{P16-1009}; \cite{D17-1155}; \cite{W17-3205}; \cite{D17-1156}; \cite{C18-1269}
+	\item Architecture \cite{R17-1049}; \cite{gulcehre2016monolingual}; \cite{W17-4712}, \cite{Biao2017CARENMT}; \cite{N18-2080}; \cite{W18-6313}; \cite{C16-1170}; \cite{P18-2050}.
+	\item The decoding algorithm \cite{gulcehre2016monolingual}. \cite{I17-2004}
+\end{itemize} 
+Beside these works, we could also consider works in other tasks such as sequence labeling tasks \cite{P07-1033}; learning multiple visual domains \cite{NIPS2017_6654}.
 The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041}.
@@
 $h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively.
-While introducing new features in the architecture, \cite{D18-1041} introduces also new objective which is sum of word-level weighted MT objective and objectives of Domain-classifier in source side and target side and Adversarial Domain-classifier in source side.
-The author have very good approach to well seperate Domain-shared information flow and Domain-specify information flow and determine their contributions to the inference that mitigates the catastrophic interference happens in the network during forward step.
-
+While introducing new features in the architecture, \cite{D18-1041} also introduces new loss function which is sum of word-level weighted MT loss and losses of Domain-classifier in source side and target side; Adversarial Domain-classifier in source side.
+The authors have very good approach to seperate Domain-shared information flow and Domain-specify information flow and to determine their contributions to the inference thus mitigate the catastrophic interference occuring in training with backpropagation.
+\cite{D18-1039} has other idea to circumvent the catastrophic interference.
+They supposed that parameters of each domain specialized model could be generated from a universal parameter-generator.
+By introducing domain embedding (or Language embedding in the original paper), they could conduct domain related error in backpropagation toward domain embedding thus mitigates catastrophic interference between domains during training.\\
+Compared to \cite{D18-1039}, our sparse embedding is a bit similar to the idea.
+However, our method gives much more free degree for domain embedding since we give each word a domain region while the author uses only one vector domain embedding for each parameter set(encoder, decoder).\\
+Compared to the work of \cite{D18-1041}, our approach is much simpler and
+more generic since sparse embedding can be implemented for every Neural architecture while Domain-specific gate and Domain-shared gate are applied only for Recurrent architectures.
+However, in the inference, the model of \cite{D18-1041} can infer the domain the closest to input sentence that will be used in special gates while our method need to know in which domain the inference is executed.
+To circumvent this disadvantage, we could use a domain classifier or simply use only generic region.
+For more detail, the reader could refer to the section \ref{sec:sparse}
 \section{Proposed method}
@@
 \subsection{Sparse word embedding}
+\label{sec:sparse}
 This section explains a new design for word embedding used in multi-domain model.
@@
 Those regions are also vanished in forward step, thus do not interfere the training on the domain to which they does not correspond.\\
-Denote $p_{t} = P * w_{t}$ where $w_{t}$ is word embedding of token $x_{t}$  is projection of word embedding (Transformer \cite{NIPS2017_7181}, Recurrent \cite{bahdanau2014neural})
+Denote \\
+$p_{t} = P * w_{t} = P_g \omega_g(t) + \displaystyle{\mathop{\sum}_{i \in [1,..,d]} P_i * \omega_i(t)}$\\
+where $w_{t}$ is word embedding of token $x_{t}$  is projection of word embedding (Transformer \cite{NIPS2017_7181}, Recurrent \cite{bahdanau2014neural})
 Assume $x_{t}$ is in domain i, thus, $w_j(t) = 0 \forall j \neq i$. Denote $P_j$ are rows of projection matrix $P$ corresponding to deactivate region $w_j$.
@@
 Furthermore, because word embeddings are involved in three parts of standard encoder-decoder structure, which are input of encoder, input of decoder and projection before softmax, we have 7 possibilities to apply sparse embedding.
-However, due to high memory cost, we only want to apply sparse embedding in encoder.
+However, due to high memory cost, we only want to apply sparse embedding in source side.
 \section{Experiments}
@@
 The reason to do this is to minimize simultaneously both losses \\
-$\theta^*_{s} = \displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))]$ \\ 
-$ \theta^*_{i} =\displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
+$\theta^*_{s}=\displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))]$ \\ 
+$ \theta^*_{i}=\displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
 $ 
@@
 \end{center}
-\caption{BLEU score for Transformer in language pair English - German. The notation is same as \ref{tab:1}}
+\caption{BLEU score for Transformer in language pair English - German. The notation is same as figure \ref{tab:1}}
 \label{tab:2}
@@
  \hline\hline
- Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 & 25.7 \\
+ Mixed & 58.55 & 22.86 & 35.11 & 13.97 & 19.59 & 7.39 & 9.47 \\
  \hline
- EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.10 \\
+ EMEA-finetuned & & & & & & & \\
  \hline
@@
  \hline
- Sparse(1) & 70.45 & 38.23 & 54.69 & 34.66 & 31.24 & 21.36 & 26.02 \\
+ Sparse(1) & 58.45 & 23.22 & 35.98 & 13.66 & 19.5 & 7.63 & 10.39 \\
  \hline
- Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & 26.02 \\
+ Sparse(2) & 56.18 & 23.05 & 35.16 & 15.95 & 20.08 & 7.63 & 10.39 \\
  \hline
@@
  \hline
- Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 & & \\
+ Domain Control & 58.55 & 22.62 & 35.15 & 11.93 & 19.34 & & \\
  \hline
@@
 \caption{bpe-detokenized BLEU score for Recurrent network architecture on language pair English-French.
-Notations as same as \ref{tab:1} except that other models from WDCMT of \cite{D18-1041} use Bidirectional-LTSM}
+Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{D18-1041} use Bidirectional-LTSM}
 \label{tab:3}
@@
 \end{center}
-\caption{BLEU score for Recurrent network on language pair English - German. The notation is same as \ref{tab:3}}
+\caption{BLEU score for Recurrent network on language pair English - German. The notation is same as figure \ref{tab:3}}
 \label{tab:4}
 \end{figure}
+\subsection{Word embedding analysis}
 

### 2019-04-22 2019/emnlp/sparse.tex
new
some comments to be discussed

@@
 \usepackage{algorithmic}
+\usepackage[draft]{todo}
 \floatname{algorithm}{Procedure}
@@
 %\def\aclpaperid{***} %  Enter the acl Paper ID here
+
+\newcommand{\fyTodo}[1]{\Todo[FY>]{\textcolor{orange}{#1}}}
+\newcommand{\fyDone}[1]{\done[FY]\Todo[FY>]{\textcolor{orange}{#1}}}
 
@@
 \maketitle
+
+\fyTodo{Too long, selfcontained, noref, rewrite}
 \begin{abstract}
-Finetuning a generic neural machine translation model to a certain domain usually degrade its performance on the other domains, thus it's very difficult to build a unique model which is best for all.
+Fine-tuning a generic neural machine translation model to a certain domain usually degrade its performance on the other domains, thus it's very difficult to build a unique model which is best for all.
 Indeed, Catastrophic Forgetting, well-known problem in Machine Learning, states that standard backprogation neural network suffers a lot while there is a shift in training data distribution.
@@
 \section{Introduction}
-Despite fast developpemnt thanks to new architectures \cite{NIPS2017_7181} \cite{bahdanau2014neural} \cite{D13-1176} \cite{Sutskever2014Sequence}, Machine Translation still struggles with data scarcity.
+\fyTodo{Merge refs and order chronologically} \fyTodo{Homogeneize cite keys, avoid arxiv}
+Despite fast developpement thanks to new architectures \cite{NIPS2017_7181} \cite{bahdanau2014neural} \cite{D13-1176} \cite{Sutskever2014Sequence}, Machine Translation (MT) still struggles with data scarcity.
 Domain adaptation, or it's special case Multi-domain, is the situation where in-domain data (such as banking) is rare while out-domain data such as news are plenty, one aims to leverage out-domain data to improve quality of Machine Translation system which is perfectible while being trained with only the in-domain dataset.
-Traditionally, ones train a generic model using a very large corpora then use in-domain dataset to finetune the model.
-It is ubiquitous that the bias to the in-domain dataset used in finetuning makes model worse in translating other domains, thus that we have not been able to finetune a model to several domains at once.
+Traditionally, ones train a generic model using a very large corpora then use in-domain dataset to fine-tune the model.
+It is ubiquitous that the bias to the in-domain dataset used in fine-tuning makes model worse in translating other domains, thus that we have not been able to fine-tune a model to several domains at once.
 This phenomenon can be implied by well-known problem in Machine Learning, Catastrophic Interference \cite{Michael1989Catastrophic}.
-When there is a shift in the distribution of training data, standard backpropagation neural network gradually lose information learned from other domains while learning only target domain.\\
+When there is a shift in the distribution of training data, standard backpropagation neural network\fyTodo{Generic ?
+backprop or NN problem ?} gradually lose information learned from other domains while learning only target domain.
+
 In Multi-domain Machine Translation, ones have to deal with polysemies whose meaning can be totally different in different domains, for example, the word "chair" means an household tool in "if you are taking a medicine which may cause low blood pressure when rising from a chair or bed " (extracted from medical corpora EMEA \cite{Tiedemann2009RANLP5}) while the word "chair" means to preside over a meeting in following sentence "The President of the ECB or , in his absence , the Vice-President , chairs the meetings of the Governing Council , the Executive Board and the General Council of the ECB.
@@
 Therefore using the same word embedding for polysemies obstructs the model from differentiating the meaning of a polysemy in one domain from it's other meaning in the other domain.
-\\
 \begin{table}[h]
@@
 \end{table}
-
+\fyTodo{Typical plan: the problems, existing solutions and their limits, our idea and contributions}
 In this work, we present a new design for word embedding that is composed of different regions which are activate and deactivate corresponding to domains.
 To evaluate, we use several test sets representing each domains and two other general test sets.
-\section{Related Work}
+\fyTodo{More precise: two architectures, two languages pairs, multiple domains, also improves generic}
+
+\section{Related Work \label{sec:related_work}}
+\fyTodo{Add standard labels to sections}
+\fyTodo{Related work goes last}
 Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
@@
 The methods in this category can be listed as following:
+\fyTodo{Organize refs}
 \begin{itemize}
@@
 The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041}.
-\cite{D18-1041} create in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
+\citet{D18-1041} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
 $h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively.
@@
 They supposed that parameters of each domain specialized model could be generated from a universal parameter-generator.
-By introducing domain embedding (or Language embedding in the original paper), they could conduct domain related error in backpropagation toward domain embedding thus mitigates catastrophic interference between domains during training.\\
+By introducing domain embedding (or Language embedding in the original paper), they could conduct domain related error in backpropagation toward domain embedding thus mitigates catastrophic interference between domains during training.
+
 Compared to \cite{D18-1039}, our sparse embedding is a bit similar to the idea.
-However, our method gives much more free degree for domain embedding since we give each word a domain region while the author uses only one vector domain embedding for each parameter set(encoder, decoder).\\
+However, our method gives much more free degree for domain embedding since we give each word a domain region while the author uses only one vector domain embedding for each parameter set(encoder, decoder).
+
 Compared to the work of \cite{D18-1041}, our approach is much simpler and
-more generic since sparse embedding can be implemented for every Neural architecture while Domain-specific gate and Domain-shared gate are applied only for Recurrent architectures.
-However, in the inference, the model of \cite{D18-1041} can infer the domain the closest to input sentence that will be used in special gates while our method need to know in which domain the inference is executed.
-To circumvent this disadvantage, we could use a domain classifier or simply use only generic region.
-For more detail, the reader could refer to the section \ref{sec:sparse}
-\section{Proposed method}
+more generic, since sparse embedding can be implemented for every neural architecture while domain-specific gate and domain-shared gate can only be applied for recurrent architectures.
+However, during inference, the model of \cite{D18-1041} can infer the domain that is the closest to the input sentence and that will be used in special gates; while our method needs to know in which domain the inference is executed.
+To circumvent this disadvantage, we could use a domain classifier or simply use only the generic region.
+For more detail, the reader could refer to the Section \ref{sec:sparse_embeddings}.
+
+\section{Sparse embeddings \label{sec:sparse_embeddings}}
+\fyTodo{Use meaningful titles}
+
 \subsection{Problem statement}
+\fyTodo{Use math for notation, label equations and such}
 \begin{definition}{Domain}
-\label{def:1}
-Domain is a hypothetical probability that generates a set of corpora. We will denote D(C) is domain of set of corpora C.
+\label{def:domain}
+A domain is a hypothetical probability distribution that generates a set of corpora. We will denote $D(C)$ the domain of the set of corpora $C$.
 \end{definition}
 
-In machine translation, the loss function for training is
+\fyTodo{We need more: the source, the target, etc}
+In machine translation, the loss function for training is:
 \begin{equation}
@@
 \end{equation}.
+
 Where $p_{\theta}(y|x)) = p(y|x,\theta)$ or $p_{\theta}(y|x)) = p(y,\theta | x)$
@@
 However, in the inference or translation, the most important goal is to predict target sentence given source sentence, i.e, to well approximate the conditional probability $p_{rel}(y|x)$.
+\fyTodo{Transition is abrupt - this is more a motivating than a modeling aspect}
 However, this relation could be very different if consider 2 domains generating 2 corpora of different topic.
@@
     \includegraphics[width=0.8\linewidth]{Sparse1}
-    \caption{Sparse word embedding.
-The red region contains activate units used only when presenting its correpondence domain, the other green regions are deactivate, the bleu generic region is always activate}
-    \label{network}
+    \caption{Sparse word embeddings.
+The red region contains units that are activated only when an input from the corresponding domain is presented; the other (green) regions are deactivated; the bleu (generic) region is always activated.}
+    \label{fig:network}
   \end{figure}
-The figure \ref{network} presents our proposed design.
-During backprogation, the rotation matrix receive vanished gradient at regions corresponding to deactivate regions in word embedding.
-Those regions are also vanished in forward step, thus do not interfere the training on the domain to which they does not correspond.\\
-Denote \\
+\fyTodo{Explain vanishing ?}
+Figure~\ref{fig:network} presents our proposed design.
+During backprogation, the \Todo{rotation}{Why ?} matrix receives vanishing gradient at regions corresponding to deactivated regions in the word embedding.
+Those regions are also vanished in forward step, thus do not interfere the training on the domain to which they do not correspond.
+
+Denote
 $p_{t} = P * w_{t} = P_g \omega_g(t) + \displaystyle{\mathop{\sum}_{i \in [1,..,d]} P_i * \omega_i(t)}$\\
-where $w_{t}$ is word embedding of token $x_{t}$  is projection of word embedding (Transformer \cite{NIPS2017_7181}, Recurrent \cite{bahdanau2014neural})
-Assume $x_{t}$ is in domain i, thus, $w_j(t) = 0 \forall j \neq i$. Denote $P_j$ are rows of projection matrix $P$ corresponding to deactivate region $w_j$.
+where $w_{t}$ is word embedding of token $x_{t}$  is projection of word embedding (Transformer \cite{NIPS2017_7181}, recurrent \cite{bahdanau2014neural})
+Assume $x_{t}$ belongs to domain $i$, thus $w_j(t) = 0 \forall j \neq i$. Denote $P_j$ the rows of the projection matrix $P$ corresponding to the deactivated region $w_j$.
 \begin{equation}
@@
 Furthermore, because word embeddings are involved in three parts of standard encoder-decoder structure, which are input of encoder, input of decoder and projection before softmax, we have 7 possibilities to apply sparse embedding.
-However, due to high memory cost, we only want to apply sparse embedding in source side.
+However, due to high memory cost, we only want to apply sparse embedding on the source side.
+\fyTodo{Explain this.
+Future work ?}
+
 \section{Experiments}
@@
 The model is also required to achieve comparable performance to generic model.
-To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
+To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.\fyTodo{Clarify / simplify text}
 
 \begin{table}[H]
+\fyTodo{Too small ! Use abbrevs (En, Fr, ECB, EMEA, EPPS)}
   \resizebox{\columnwidth}{!}{
@@
 Task & Corpora & Train & Dev & Test & Vocabulary \\ \hline
-\multirow{5}{*}{$English \rightarrow French$} & European Medicines Agency & 1090568 & 1000 & 1000 & English: 30164, French: 30398\\
+\multirow{5}{*}{$En \rightarrow Fr$} & European Medicines Agency & 1090568 & 1000 & 1000 & English: 30164, French: 30398\\
  & European Central Bank & 191960 & 1000 & 1000 &\\
  & European Parliament & 2005723 & 1000 & 1000 & \\ \hline
-\multirow{5}{*}{$English \rightarrow German$} & European Medicines Agency & 1108752 & 1000 & 1000 & English: 30159, German: 30698\\
+\multirow{5}{*}{$En \rightarrow Ge$} & European Medicines Agency & 1108752 & 1000 & 1000 & English: 30159, German: 30698\\
  & European Central Bank & 111174 & 1000 & 1000 &\\
@@
 }
-\caption{Corpora}
+\caption{Corpora }
 \label{table:Corpora}
@@
 \subsection{NMT model configuration}
-We evaluate our method on 2 popular architecture: Transformer of and Attentional Birectional Long Short-Term Memmory (LSTM).\\
-More precisely, for Transformer we use the same setting as described in \cite{NIPS2017_7181}.For Attentional Birectional Long Short-Term Memmory, in order to fairly compare to the work of \cite{D18-1041}, we use standard configuration with: 1 layer bidirectional LSTM with hidden size 1024 in encoder; 1 layer attentional unidirectional LSTM with same hidden size in decoder; and word embedding of size 512.
-The implementation can be found in OpenNMT-tf \cite{P17-4012}.\\
-For sparse embedding implementation, we split vector embedding of size 512 into 4 regions $[8,8,8,488]$ corresponding to EMEA, EPPS, ECB domain and generic region.
-If sentence comes from domain $i$, Domain regions $j,\forall j \neq i$ will be masked as 0.
-We then use a dense layer of size 512 to fuse active domain region and generic region.
+We evaluate our method on two state-of-the-art encoder-decoder architectures: a recurrent architecture based on attentional birectional Long Short-Term Memory (LSTM) cells and the Transformer architecture of \citet{NIPS2017_7181}.
+More precisely, for the Transformer we use the same settings as described in \cite{NIPS2017_7181}.
+For the recurrent architecture, in order to fairly compare to the work of \citet{D18-1041}, we use a standard configuration with one layer bidirectional LSTM with a encoder hidden size of 1024; one attentional unidirectional LSTM layer with same hidden size in decoder; and word embeddings of size 512.
+The implementation can be found in OpenNMT-tf \cite{P17-4012}.
+\fyTodo{Language pairs, BPE number, optimizer}
+\fyTodo{Motivate the split}
+For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions $[8,8,8,488]$, corresponding respectively to EMEA, EPPS, ECB domain and generic region.
+If sentence comes from domain $i$, regions for domains $j,\forall j \neq i$ will be masked as 0.
+We then use a dense layer of size 512 to fuse the active domain region and the generic region.
 
@@
 \STATE{Randomly pick a batch of sentences from corpora $C_i$.}
-\STATE{Using only generic activate embedding and domain activate embedding for this batch of sentences, i.e, we will have 2 batches of word embedding.}
+\STATE{Using only generic activate embedding and domain activate embedding for this batch of sentences, i.e, we will have 2 batches of word embedding.\fyTodo{why 2 }}
 \STATE{Pass the batch of word embedding to a dense layer to fuse regions.}
@@
 \begin{center}
- \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
+ \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
  \hline
- Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT test 2010 \\ [0.5ex] 
+ Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
  \hline\hline
- Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 & 25.7 \\
+ Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 &  & 25.7 \\
  \hline
- EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.10 \\
+ EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 &  & 11.10 \\
  \hline
- EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 & 11.1 \\
+ EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 &  & 11.1 \\
  \hline
- ECB-finetuned & 9.2 & 18.39 & 55.45 & 12.13 & 16.12 & 8.71 & 7.19 \\
+ ECB-finetuned & 9.2 & 18.39 & 55.45 & 12.13 & 16.12 & 8.71 &  & 7.19 \\
  \hline
- Sparse(1) & 70.45 & 38.23 & 54.69 & 34.66 & 31.24 & 21.36 & 26.02 \\
+ Sparse(1) & 70.45 & 38.23 & 54.69 & 34.66 & 31.24 & 21.36 & 28.12 & 26.02 \\
  \hline
- Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & 26.02 \\
+ Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & & 26.02 \\
  \hline
- Sparse(3) &  &  &  &  &  &  & \\
+ Sparse(3) &  &  &  &  &  &  &  & \\
  \hline
- Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 & & \\
+ Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 &  & & \\
  \hline
@@
 \begin{center}
- \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
+ \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
  \hline
- Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT test 2010 \\ [0.5ex] 
+ Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
  \hline\hline
- Mixed & & & & & & & \\
+ Mixed & 64.57 & 26.47 & 68.67 & 21.77 & 23.50 & 13.9 & 10.10 & \\
  \hline
- EMEA-finetuned & & & & & & & \\
+ EMEA-finetuned & & & & & & & & \\
  \hline
- EPPS-finetuned & & & & & & & \\
+ EPPS-finetuned & & & & & & & & \\
  \hline
- ECB-finetuned & & & & & & & \\
+ ECB-finetuned & & & & & & & & \\
  \hline
- Sparse & & & & & & & \\
+ Sparse & & & & & & & & \\
  \hline
- Domain Control & & & & & & & \\
+ Domain Control & & & & & & & & \\
  \hline
@@
  \hline\hline
- Mixed & 58.55 & 22.86 & 35.11 & 13.97 & 19.59 & 7.39 & 9.47 \\
+ Mixed & 64.57 & 26.47 & 68.67 & & & & \\
  \hline
@@
  \hline
- Sparse(1) & 58.45 & 23.22 & 35.98 & 13.66 & 19.5 & 7.63 & 10.39 \\
+ Sparse & & & & & & & \\
  \hline
- Sparse(2) & 56.18 & 23.05 & 35.16 & 15.95 & 20.08 & 7.63 & 10.39 \\
- \hline
- Sparse(3) &  &  &  &  &  &  & \\
- \hline
- Domain Control & 58.55 & 22.62 & 35.15 & 11.93 & 19.34 & & \\
- \hline
- WDCMT (conditional GRU) & & & & & & & \\
+ Domain Control & & & & & & & \\
  \hline

### 2019-04-23 2019/emnlp/sparse.tex
new

@@
 \fyTodo{Merge refs and order chronologically} \fyTodo{Homogeneize cite keys, avoid arxiv}
-Despite fast developpement thanks to new architectures \cite{NIPS2017_7181} \cite{bahdanau2014neural} \cite{D13-1176} \cite{Sutskever2014Sequence}, Machine Translation (MT) still struggles with data scarcity.
+Despite fast developpement thanks to new architectures \cite{D13-1176, Sutskever2014Sequence, bahdanau2014neural, NIPS2017_7181}, Machine Translation (MT) still struggles with data scarcity.
 Domain adaptation, or it's special case Multi-domain, is the situation where in-domain data (such as banking) is rare while out-domain data such as news are plenty, one aims to leverage out-domain data to improve quality of Machine Translation system which is perfectible while being trained with only the in-domain dataset.
@@
 In this work, we present a new design for word embedding that is composed of different regions which are activate and deactivate corresponding to domains. 
-To evaluate, we use several test sets representing each domains and two other general test sets.
+\\ To evaluate, we use 3 corpora of 3 different domains: Medical, Banking and Administration and 2 different architectures: Transformer and Attentional recurrent network.
+The test sets are selectioned from shared task each domains and two other general test sets such as news commentary and IWSLT.
+The reader could find more details in Experimetal section \ref{subsec:data}
 \fyTodo{More precise: two architectures, two languages pairs, multiple domains, also improves generic}
@@
 \begin{itemize}
-	\item score relatedness of sentences to the target domain by Language Model \cite{P10-2041}; \cite{D11-1033}; \cite{P13-2119}.
-	\item generating pseudo parallel data \cite{P03-1010}; \cite{C16-1295}; \cite{D14-1023}
+	\item score relatedness of sentences to the target domain by Language Model \cite{P10-2041,D11-1033,P13-2119}.
+	\item generating pseudo parallel data \cite{P03-1010,C16-1295,D14-1023}
 \end{itemize}
@@
 \begin{itemize}
-	\item Trainining objective \cite{Luong2015SNMT}; \cite{P16-1009}; \cite{D17-1155}; \cite{W17-3205}; \cite{D17-1156}; \cite{C18-1269}
-	\item Architecture \cite{R17-1049}; \cite{gulcehre2016monolingual}; \cite{W17-4712}, \cite{Biao2017CARENMT}; \cite{N18-2080}; \cite{W18-6313}; \cite{C16-1170}; \cite{P18-2050}.
-	\item The decoding algorithm \cite{gulcehre2016monolingual}. \cite{I17-2004}
+	\item Trainining objective \cite{Luong2015SNMT,P16-1009,D17-1155,W17-3205,D17-1156,C18-1269}
+	\item Architecture \cite{gulcehre2016monolingual,C16-1170,R17-1049,W17-4712,Biao2017CARENMT,N18-2080,W18-6313,P18-2050}.
+	\item The decoding algorithm \cite{gulcehre2016monolingual,I17-2004}
 \end{itemize} 
 Beside these works, we could also consider works in other tasks such as sequence labeling tasks \cite{P07-1033}; learning multiple visual domains \cite{NIPS2017_6654}.
-The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041}.
-\citet{D18-1041} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
+The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041,D18-1041} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
 $h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively.
@@
 \label{def:domain}
-A domain is a hypothetical probability distribution that generates a set of corpora. We will denote $D(C)$ the domain of the set of corpora $C$.
+A domain is probabilistic distribution of translation $(x,y)$ in a given corpora. Given corpora $C$, we denote $D(C)$ the domain of $C$.
 \end{definition}
-
 \fyTodo{We need more: the source, the target, etc}
@@
 
-Where $p_{\theta}(y|x)) = p(y|x,\theta)$ or $p_{\theta}(y|x)) = p(y,\theta | x)$
-In this quantity, domain is exactly $p_{rel}(x,y)$.
-However, in the inference or translation, the most important goal is to predict target sentence given source sentence, i.e, to well approximate the conditional probability $p_{rel}(y|x)$.
-\fyTodo{Transition is abrupt - this is more a motivating than a modeling aspect}
-However, this relation could be very different if consider 2 domains generating 2 corpora of different topic.
-Indeed, in financial document, to translate "security" into "titre financier" is more likely than to translate it into "securit\'e" but in IT document.
-In reality, to translate correctly with respect to the topic of sentence is very important.
-We formalize the multi-domain problem for machine translation in following equation
+Where $p_{\theta}(y|x) = p(y|x,\theta)$ or $p_{\theta}(y|x) = p(y,\theta | x)$ \\
+In this equation, the domain is $p_{rel}(x,y)$.
+One aims to approximate to model to target domain $p_{rel}$.
+The more the domain of training data close to the domain of test sets, the higher score it could achieve.
+\\
+In the training of a generic model, the $p_{rel}$ is oftenly a general domain, noted $p_{gen}$ which is defined by a set of several corpora which belongs to different topics (e.g medical, banking, adminstration).
+Therefore the generic model is not usually good enough for translating by topic in which one is more interested.
+In order to specialize the model, one usually finetunes generic model by continue the traing from $\theta^*$, which was optimized on general domain, with corpora of the interested topic.
+However, there is always a dramatical shift from generic domain $p_{gen}$ to specialization domain $p_{spec}$, that usually leads to catastrophic forgetting in neural network using backpropagation \cite{Michael1989Catastrophic}.
+The finetuned model has lower performance in translating sentences, which are more related to other topic, than generic model.
+Indeed, the translation $(x,y)$ could be very different in different topic.
+For instance, in financial document, to translate "security" into "titre financier" is more likely than to translate it into "securit\'e" but in IT document.
+Therefore, the shift from general domain containing these two corpora to one of them is not negligible.
+\\
+In order to conduct the model translating by a certain topic, we might use conditional probability conditioned by the given topic $p(y|x,d)$.
+Given a set of training corpora, we assume that there are a limited d engendering topic, i.e, there are d corpora $C_i$ $i \in [1,..,d]$ representing d topical domains $D(C_i)$ respectively.
+We are then interested in approximating $p(y,x|i)$ for $i \in [1,..,d]$ and use $p(y|x,i)$ for the inference in topic $i$.
+From now, we formalize the multi-domain problem for machine translation in following equation
 \begin{equation}
@@
     \caption{Sparse word embeddings.
-The red region contains units that are activated only when an input from the corresponding domain is presented; the other (green) regions are deactivated; the bleu (generic) region is always activated.}
+The red region contains units that are activated only when an input from the corresponding domain is presented; the other (green) regions are deactivated; the blue (generic) region is always activated.}
     \label{fig:network}
@@
 \section{Experiments}
-\subsection{Domains and data}
+\subsection{Domains and data \label{subsec:data}}
 To evaluate the performance of proposed architecture, we design experiments in which we want to train an NMT for 3 domains such as medical, administrative documents and business briefly denoted by EMEA, EPPS, ECB according to training corpora name respectively.
@@
 \STATE{Randomly pick a batch of sentences from corpora $C_i$.}
-\STATE{Using only generic activate embedding and domain activate embedding for this batch of sentences, i.e, we will have 2 batches of word embedding.\fyTodo{why 2 }}
+\STATE{Activating only generic activate embedding to create generic embedding batch, denoted $W_g$}
+\STATE{Pass the batch of word embedding to a dense layer to fuse regions}
+\STATE{Computing gradient of $\theta_s$, $\frac{\partial L}{\partial \theta_s}$ using generic batch $W_g$}
+\STATE{Activating corresponding domain region and generic region to create domain embedding batch}
 \STATE{Pass the batch of word embedding to a dense layer to fuse regions.}
-\STATE{Apply backpropagation as usual.}
+\STATE{Computing gradient of domain parameters $\theta_i$, $\frac{\partial L}{\partial \theta_i}$ using domain batch }
+\STATE{Apply backprogation using computed gradients $\frac{\partial L}{\partial \theta_s}$ and $\frac{\partial L}{\partial \theta_i}$}
 \UNTIL{convergence}
@@
  \hline\hline
- Mixed & 64.57 & 26.47 & 68.67 & 21.77 & 23.50 & 13.9 & 10.10 & \\
+ Mixed & 64.57 & 26.47 & 68.67 & 21.77 & 23.50 & 13.9 & 17.04 & \\
  \hline
@@
  \hline
- Sparse & & & & & & & & \\
+ Sparse(1) & 70.55 & 26.17 & 68.07 & 21.50 & 23.30 & & & \\
  \hline
@@
  \hline\hline
- Mixed & 64.57 & 26.47 & 68.67 & & & & \\
+ Mixed & & & & & & & \\
  \hline
@@
  \hline
+ WDCMT (conditional GRU) & & & & & & & \\
+ \hline 
 \end{tabularx}}
@@
 \caption{bpe-detokenized BLEU score for Recurrent network architecture on language pair English-French.
-Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{D18-1041} use Bidirectional-LTSM}
+Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{D18-1041} use Bidirectional-GRU}
 \label{tab:3}

### 2019-04-23 2019/emnlp/sparse.tex
new

@@
 
-Where $p_{\theta}(y|x) = p(y|x,\theta)$ or $p_{\theta}(y|x) = p(y,\theta | x)$ \\
+Where $p_{\theta}(y|x) = p(y|x,\theta)$ \\
 In this equation, the domain is $p_{rel}(x,y)$.
 One aims to approximate to model to target domain $p_{rel}$.
-The more the domain of training data close to the domain of test sets, the higher score it could achieve.
-\\
+The more the domain of training data close to the domain of test set, the better performance it could achieve over test set.\\
 In the training of a generic model, the $p_{rel}$ is oftenly a general domain, noted $p_{gen}$ which is defined by a set of several corpora which belongs to different topics (e.g medical, banking, adminstration).
@@
  \hline
- Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & & 26.02 \\
+ Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & 28.12 & 26.02 \\
  \hline
@@
  \hline\hline
- Mixed & 64.57 & 26.47 & 68.67 & 21.77 & 23.50 & 13.9 & 17.04 & \\
+ Mixed & 64.57 & 26.47 & 68.67 & 21.77 & 23.50 & 13.9 & 17.04 & 18.85 \\
  \hline
@@
  \hline
- Sparse(1) & 70.55 & 26.17 & 68.07 & 21.50 & 23.30 & & & \\
+ Sparse(1) & 70.55 & 26.17 & 68.07 & 21.50 & 23.30 & 14.05 & 16.64 & 19.04\\
+ \hline
+ Sparse(2) & 70.67 & 26.24 & 66.97 & & & 14.05 & 16.64 & 19.04 \\
  \hline
@@
 \begin{center}
- \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
+ \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
  \hline
- Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT test 2010 \\ [0.5ex] 
+ Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
  \hline\hline
- Mixed & & & & & & & \\
+ Mixed & & & & & & & & \\
  \hline
- EMEA-finetuned & & & & & & & \\
+ EMEA-finetuned & & & & & & & & \\
  \hline
- EPPS-finetuned & & & & & & & \\
+ EPPS-finetuned & & & & & & & & \\
  \hline
- ECB-finetuned & & & & & & & \\
+ ECB-finetuned & & & & & & & & \\
  \hline
- Sparse & & & & & & & \\
+ Sparse & & & & & & & & \\
  \hline
- Domain Control & & & & & & & \\
+ Domain Control & & & & & & & & \\
  \hline
- WDCMT (conditional GRU) & & & & & & & \\
+ WDCMT (conditional GRU) & 68.76 & 35.71 & 52.75 & 32.03 & 29.81 & 19.58 & 24.63 & 23.28 \\
  \hline 
@@
 \caption{bpe-detokenized BLEU score for Recurrent network architecture on language pair English-French.
-Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{D18-1041} use Bidirectional-GRU}
+Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{D18-1041} use Bidirectional-LSTM}
 \label{tab:3}

### 2019-04-23 2019/emnlp/sparse.tex
more todos

@@
 \fyTodo{Related work goes last}
+\fyTodo{Compare also to Peng 2017}
+
 Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
@@
 \end{table}
+\subsection{Baselines \label{ssec:baselines}}
+\fyTodo{Describe baselines}
+
 \subsection{NMT model configuration}
@@
 Indeed, real data are not always balanced for every domain.
-For instance, in our experiments, the size of corpora are very different where EPPS is largest with approximately 2 millions sentences, EMEA contains about 1 million sentences and ECB is the smallest with only 200 000 sentences for English French and 100 000 for English German.
+For instance, in our experiments, the size of corpora are very different where EPPS is largest with approximately 2 millions sentences, EMEA contains about 1 million sentences and ECB is the smallest with only 200 000 sentences for English French and 100 000 for English-German.
 If we train each corpora with same number of batches, ECB will be trained too oftenly, that easily leads to overfitting.
@@
 \end{center}
-\caption{bpe-detokenized BLEU score in language pair English - French.
+\caption{BPE-detokenized BLEU score in language pair English - French.
 Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT.
@@
 \end{center}
-\caption{BLEU score for Recurrent network on language pair English - German. The notation is same as figure \ref{tab:3}}
+\caption{BLEU score for Recurrent network on language pair English - German. Notations are the same as for figure~\ref{tab:3}}
 \label{tab:4}
@@
 \section{Conclusions}
+\fyTodo{natural continuations: go beyond words; learn the projection matrix at the word level; other ?}
 

### 2019-04-24 2019/emnlp/sparse.tex
more comments / todos

@@
 
-\newcommand{\fyTodo}[1]{\Todo[FY>]{\textcolor{orange}{#1}}}
-\newcommand{\fyDone}[1]{\done[FY]\Todo[FY>]{\textcolor{orange}{#1}}}
+\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
 
@@
 
-\author{Minh Quang Pham\\
+\author{First Author \\% Minh Quang Pham\\
   Affiliation / Address line 1 \\
@@
 
+\fyTodo{s/naacl/EMNLP/g}
 \fyTodo{Too long, selfcontained, noref, rewrite}
@@
 \section{Introduction}
-\fyTodo{Merge refs and order chronologically} \fyTodo{Homogeneize cite keys, avoid arxiv}
+\fyTodo{Merge refs and order chronologically}
+\fyTodo{Homogeneize cite keys, avoid arxiv}
+
 Despite fast developpement thanks to new architectures \cite{D13-1176, Sutskever2014Sequence, bahdanau2014neural, NIPS2017_7181}, Machine Translation (MT) still struggles with data scarcity.
@@
 
-In Multi-domain Machine Translation, ones have to deal with polysemies whose meaning can be totally different in different domains, for example, the word "chair" means an household tool in "if you are taking a medicine which may cause low blood pressure when rising from a chair or bed " (extracted from medical corpora EMEA \cite{Tiedemann2009RANLP5}) while the word "chair" means to preside over a meeting in following sentence "The President of the ECB or , in his absence , the Vice-President , chairs the meetings of the Governing Council , the Executive Board and the General Council of the ECB.
+In \fyTodostar{Specify DA}{Multi-domain Machine Translation}, ones have to deal with polysemies whose meaning can be totally different in different domains, for example, the word "chair" means an household tool in "if you are taking a medicine which may cause low blood pressure when rising from a chair or bed " (extracted from medical corpora EMEA \cite{Tiedemann2009RANLP5}) while the word "chair" means to preside over a meeting in following sentence "The President of the ECB or , in his absence , the Vice-President , chairs the meetings of the Governing Council , the Executive Board and the General Council of the ECB.
 " (extracted from European Central Bank corpora \cite{Tiedemann2009RANLP5}).
@@
 \section{Sparse embeddings \label{sec:sparse_embeddings}}
-\fyTodo{Use meaningful titles}
+\fyDone{Use meaningful titles throughout}
 
@@
 \end{equation}
-\subsection{Detailed implementation}
+\subsection{Implementation Details}
 The complex structure of State-of-the-art architecture Transformer \cite{NIPS2017_7181} and popular recurrent network \cite{bahdanau2014neural} makes the implementation of sparse embedding non-obvious.
@@
 
-\section{Experiments}
-\subsection{Domains and data \label{subsec:data}}
+\section{Experiments \label{sec:experiments}}
+
+\subsection{Domains and data \label{ssec:data}}
 To evaluate the performance of proposed architecture, we design experiments in which we want to train an NMT for 3 domains such as medical, administrative documents and business briefly denoted by EMEA, EPPS, ECB according to training corpora name respectively.
@@
 $ 
-We admit that this is not the optimal way. However, this procdedure make the training very simple.\\
+We admit that this is not the optimal way. However, this procedure makes training very simple.\fyTodo{Explain technicalities - future work ?}
+
 We also take in account the problem of overfitting that could occur for small corpora.
@@
 Therefore, we perform backpropagation on corpora with frequencies according to sizes of corpora.
+\fyTodo{Refs on this ?}
+
 \subsection{Results}
@@
 \section*{Acknowledgments}
-
+\fyTodo{Homegenize refs - urls, addresses, etc}
 \bibliography{naaclhlt2019}

### 2019-04-24 2019/emnlp/sparse.tex
following our discussion

@@
 
-\section{Introduction}
-\fyTodo{Merge refs and order chronologically}
+\section{Introduction \label{sec:introduction}}
+\fyDone{Merge refs and order chronologically}
 \fyTodo{Homogeneize cite keys, avoid arxiv}
@@
 
-In \fyTodostar{Specify DA}{Multi-domain Machine Translation}, ones have to deal with polysemies whose meaning can be totally different in different domains, for example, the word "chair" means an household tool in "if you are taking a medicine which may cause low blood pressure when rising from a chair or bed " (extracted from medical corpora EMEA \cite{Tiedemann2009RANLP5}) while the word "chair" means to preside over a meeting in following sentence "The President of the ECB or , in his absence , the Vice-President , chairs the meetings of the Governing Council , the Executive Board and the General Council of the ECB.
+In \fyTodostar{Specify DA}{Multi-domain Machine Translation}, ones have to deal with polysemies whose meaning can be totally different in different domains, for example, the word "chair" means an household tool in "if you are taking a medicine which may cause low blood pressure when rising from a chair or bed " (extracted from medical corpus EMEA \cite{Tiedemann2009RANLP5}) while the word "chair" means to preside over a meeting in following sentence "The President of the ECB or , in his absence , the Vice-President , chairs the meetings of the Governing Council , the Executive Board and the General Council of the ECB.
 " (extracted from European Central Bank corpora \cite{Tiedemann2009RANLP5}).
@@
 \hline
-English sentence & French translation & Corpora \\ 
+English sentence & French translation & Corpus \\ 
 \hline
@@
 \fyTodo{More precise: two architectures, two languages pairs, multiple domains, also improves generic}
-
-\section{Related Work \label{sec:related_work}}
-\fyTodo{Add standard labels to sections}
-\fyTodo{Related work goes last}
-\fyTodo{Compare also to Peng 2017}
-
-Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
-Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
-The methods in this category can be listed as following:
-\fyTodo{Organize refs}
-\begin{itemize}
-	\item score relatedness of sentences to the target domain by Language Model \cite{P10-2041,D11-1033,P13-2119}.
-	\item generating pseudo parallel data \cite{P03-1010,C16-1295,D14-1023}
-\end{itemize}
-On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features
-\begin{itemize}
-	\item Trainining objective \cite{Luong2015SNMT,P16-1009,D17-1155,W17-3205,D17-1156,C18-1269}
-	\item Architecture \cite{gulcehre2016monolingual,C16-1170,R17-1049,W17-4712,Biao2017CARENMT,N18-2080,W18-6313,P18-2050}.
-	\item The decoding algorithm \cite{gulcehre2016monolingual,I17-2004}
-\end{itemize} 
-Beside these works, we could also consider works in other tasks such as sequence labeling tasks \cite{P07-1033}; learning multiple visual domains \cite{NIPS2017_6654}.
-The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041,D18-1041} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
-$h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively.
-While introducing new features in the architecture, \cite{D18-1041} also introduces new loss function which is sum of word-level weighted MT loss and losses of Domain-classifier in source side and target side; Adversarial Domain-classifier in source side.
-The authors have very good approach to seperate Domain-shared information flow and Domain-specify information flow and to determine their contributions to the inference thus mitigate the catastrophic interference occuring in training with backpropagation.
-\cite{D18-1039} has other idea to circumvent the catastrophic interference.
-They supposed that parameters of each domain specialized model could be generated from a universal parameter-generator.
-By introducing domain embedding (or Language embedding in the original paper), they could conduct domain related error in backpropagation toward domain embedding thus mitigates catastrophic interference between domains during training.
-
-Compared to \cite{D18-1039}, our sparse embedding is a bit similar to the idea.
-However, our method gives much more free degree for domain embedding since we give each word a domain region while the author uses only one vector domain embedding for each parameter set(encoder, decoder).
-
-Compared to the work of \cite{D18-1041}, our approach is much simpler and
-more generic, since sparse embedding can be implemented for every neural architecture while domain-specific gate and domain-shared gate can only be applied for recurrent architectures.
-However, during inference, the model of \cite{D18-1041} can infer the domain that is the closest to the input sentence and that will be used in special gates; while our method needs to know in which domain the inference is executed.
-To circumvent this disadvantage, we could use a domain classifier or simply use only the generic region.
-For more detail, the reader could refer to the Section \ref{sec:sparse_embeddings}.
 
@@
 Figure~\ref{fig:network} presents our proposed design.
-During backprogation, the \Todo{rotation}{Why ?} matrix receives vanishing gradient at regions corresponding to deactivated regions in the word embedding.
+During backprogation, the \fyTodo{fix notation}{projection matrix $P$} receives vanishing gradient at regions corresponding to deactivated regions in the word embedding.
 Those regions are also vanished in forward step, thus do not interfere the training on the domain to which they do not correspond.
@@
 \fyTodo{Language pairs, BPE number, optimizer}
-\fyTodo{Motivate the split}
+\fyTodo{Motivate the split - discuss experimentally embedding size}
 For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions $[8,8,8,488]$, corresponding respectively to EMEA, EPPS, ECB domain and generic region.
@@
 Therefore, we perform backpropagation on corpora with frequencies according to sizes of corpora.
-\fyTodo{Refs on this ?}
+\fyTodo{Refs on this ?
+or contrast ?}
 
 \subsection{Results}
-\begin{figure}[H]
+\begin{table}[H]
 \begin{center}
@@
 \end{center}
-\caption{BPE-detokenized BLEU score in language pair English - French.
+\caption{BPE-detokenized BLEU score for English-French with the Transformer architecture.
 Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT.
 Sparse(2) uses generic tag for all of test sets.
-Sparse(3) uses tag predicted by pretrained domain classififer}
+Sparse(3) uses tag predicted by a pre-trained domain classififer}
 \label{tab:1}
-\end{figure}
+\end{table}
 
-\begin{figure}[H]
+\begin{table}[H]
 \begin{center}
@@
 \end{center}
-\caption{BLEU score for Transformer in language pair English - German. The notation is same as figure \ref{tab:1}}
+\caption{BLEU score for Transformer in language pair English - German. Notations are same as Table~\ref{tab:1}}
 \label{tab:2}
-\end{figure}
+\end{table}
 
-\begin{figure}[H]
+\begin{table}[H]
 \begin{center}
@@
 \label{tab:3}
-\end{figure}
+\end{table}
 
-\begin{figure}[H]
+\begin{table}[H]
 \begin{center}
@@
 \end{center}
-\caption{BLEU score for Recurrent network on language pair English - German. Notations are the same as for figure~\ref{tab:3}}
+\caption{BLEU scores for RNN architecture on English-German. Notations are the same as for figure~\ref{tab:3}}
 \label{tab:4}
-\end{figure}
+\end{table}
+
 \subsection{Word embedding analysis}
+
+\section{Related Work \label{sec:related_work}}
+\fyDone{Add standard labels to sections}
+\fyDone{Related work goes last}
+\fyTodo{Compare also to Peng 2017}
+
+Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
+Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
+The methods in this category can be listed as following:
+\fyTodo{Organize refs}
+\begin{itemize}
+	\item score relatedness of sentences to the target domain by Language Model \cite{P10-2041,D11-1033,P13-2119}.
+	\item generating pseudo parallel data \cite{P03-1010,C16-1295,D14-1023}
+\end{itemize}
+On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features
+\begin{itemize}
+	\item Trainining objective \cite{Luong2015SNMT,P16-1009,D17-1155,W17-3205,D17-1156,C18-1269}
+	\item Architecture \cite{gulcehre2016monolingual,C16-1170,R17-1049,W17-4712,Biao2017CARENMT,N18-2080,W18-6313,P18-2050}.
+	\item The decoding algorithm \cite{gulcehre2016monolingual,I17-2004}
+\end{itemize} 
+Beside these works, we could also consider works in other tasks such as sequence labeling tasks \cite{P07-1033}; learning multiple visual domains \cite{NIPS2017_6654}.
+The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041,D18-1041} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
+$h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively.
+While introducing new features in the architecture, \cite{D18-1041} also introduces new loss function which is sum of word-level weighted MT loss and losses of Domain-classifier in source side and target side; Adversarial Domain-classifier in source side.
+The authors have very good approach to seperate Domain-shared information flow and Domain-specify information flow and to determine their contributions to the inference thus mitigate the catastrophic interference occuring in training with backpropagation.
+\cite{D18-1039} has other idea to circumvent the catastrophic interference.
+They supposed that parameters of each domain specialized model could be generated from a universal parameter-generator.
+By introducing domain embedding (or Language embedding in the original paper), they could conduct domain related error in backpropagation toward domain embedding thus mitigates catastrophic interference between domains during training.
+
+Compared to \cite{D18-1039}, our sparse embedding is a bit similar to the idea.
+However, our method gives much more free degree for domain embedding since we give each word a domain region while the author uses only one vector domain embedding for each parameter set(encoder, decoder).
+
+Compared to the work of \cite{D18-1041}, our approach is much simpler and
+more generic, since sparse embedding can be implemented for every neural architecture while domain-specific gate and domain-shared gate can only be applied for recurrent architectures.
+However, during inference, the model of \cite{D18-1041} can infer the domain that is the closest to the input sentence and that will be used in special gates; while our method needs to know in which domain the inference is executed.
+To circumvent this disadvantage, we could use a domain classifier or simply use only the generic region.
+For more detail, the reader could refer to the Section \ref{sec:sparse_embeddings}.
 

### 2019-05-01 2019/emnlp/sparse.tex
work on the introduction

@@
 
-\title{Sparse Word Embedding for Multi-domain Machine Translation}
+\title{Sparse Word Embeddings for Multi-domain Machine Translation}
 
@@
 \fyTodo{More precise: two architectures, two languages pairs, multiple domains, also improves generic}
+
+\section{Introduction \label{sec:introduction}}
+% The problem
+Owing to the development of flexible and powerful architectures based on neural networks \cite{bahdanau2014neural;Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date.
+One issue which repeatedly pops up with neural MT is the need to perform training with large parallel corpus, typically consisting of millions of parallel sentences as close as possible to the application domain.
+A lot of the recent research effort has focused on building MT systems in low data conditions for instance building multi-lingual \cite{} MTs or even building systems without any data \cite{}.
+
+Another important scenario for industrial MT systems is to adapt an neural system trained using a parallel data in one  domain to the peculiarity of other domains.
+Domain Adaption (DA) in MT is an ancient issue \cite{Foster07amixture,Axelrod11domain}, which comes in various guise, and for which a number of solutions has been studied \cite{Chu18asurvey}.
+The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain.
+We study here a different scenario, multi-domain adaptation \cite{Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
+
+Multi-domain adaptation is conceptually close to multi-lingual MT\cite{}, or more generally to multi-task learning \cite{}, and can be approached in a number of ways.
 

### 2019-05-01 2019/emnlp/sparse.tex
end alternate take for the intro

@@
 Owing to the development of flexible and powerful architectures based on neural networks \cite{bahdanau2014neural;Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date.
-One issue which repeatedly pops up with neural MT is the need to perform training with large parallel corpus, typically consisting of millions of parallel sentences as close as possible to the application domain.
-A lot of the recent research effort has focused on building MT systems in low data conditions for instance building multi-lingual \cite{} MTs or even building systems without any data \cite{}.
+One issue which repeatedly pops up with neural MT is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain.
+A lot of the recent research effort has thus focused on developping MT systems in low data conditions, for instance building multi-lingual MTs enabling zero-shot translation\cite{} or even building systems without any data \cite{}.
 
 Another important scenario for industrial MT systems is to adapt an neural system trained using a parallel data in one  domain to the peculiarity of other domains.
-Domain Adaption (DA) in MT is an ancient issue \cite{Foster07amixture,Axelrod11domain}, which comes in various guise, and for which a number of solutions has been studied \cite{Chu18asurvey}.
+Domain Adaption (DA) in MT is an ancient issue \cite{Foster07amixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied \cite{Chu18asurvey}.
 The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain.
 We study here a different scenario, multi-domain adaptation \cite{Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
+This allows us to be both data-efficient (all data is used to train all domains) and compuationally efficient (we only need to train one domain).
+In our setting, we assume that the domain of test documents is known.
 
 Multi-domain adaptation is conceptually close to multi-lingual MT\cite{}, or more generally to multi-task learning \cite{}, and can be approached in a number of ways.
+We adapt here the adaptation framework of \cite{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains.
+Our main hypothesis is that most of the differences between domains takes place at the lexical level, due to massive cross-domain polysemy.
+By contrast, the deeper layer, which arguably model more abstract phenomena, should be made shareable.
+To this end, we make sure to design word embeddings in such that they will contain a generic as well as a domain-specific region.
+We perform experiments with three domains, two architectures and two language pairs and find that our technique can effectively build multi-domain NMT, outperforming several baselines, including domain-specific MT engines trained through fine-tuning.
+Our contribution is thus twofold:
+\begin{itemize}
+\item methods to adapt the idea of \cite{Daume07frustratingly} for two NMT architectures;
+\item experimental results that demonstrate the effectiveness of this technique;
+\end{itemize}
+\fyTodo{can we train in random order ? can we get away with catastrophic forgetting ?}
+\fyTodo{how to analyze the embeddings ? how can we test or claim ?}
 

### 2019-05-02 2019/emnlp/sparse.tex
new

@@
 \documentclass[11pt,a4paper]{article}
-\usepackage[hyperref]{naaclhlt2019}
+\usepackage[hyperref]{emnlp-ijcnlp-2019}
 \usepackage{times}
+\usepackage{latexsym}
+
+\usepackage{url}
+
+%\aclfinalcopy % Uncomment this line for the final submission
+
+%\setlength\titlebox{5cm}
+% You can expand the titlebox if you need extra space
+% to show all the authors. Please do not make the titlebox
+% smaller than 5cm (the original size); we will check this
+% in the camera-ready version and ask you to change it back.
+
+\newcommand\BibTeX{B{\sc ib}\TeX}
+\newcommand\confname{EMNLP-IJCNLP 2019}
+\newcommand\conforg{SIGDAT}
 \usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
-\usepackage{latexsym}
 \usepackage{graphicx}
@@
 
-\begin{table}[H]
+\begin{figure}[H]
 \begin{center}
@@
  \hline\hline
- Mixed & & & & & & & & \\
+ Mixed & 65.42 & 34.7 & 51.38 & 29.27 & 28.95 & 17.82 & 22.21 & 21.63 \\
+ \hline
+ EMEA-finetuned & & & & & & & & \\
+ \hline
+ EPPS-finetuned & & & & & & & & \\
+ \hline
+ ECB-finetuned & & & & & & & & \\
+ \hline
+ Sparse & 68.51 & 35.75 & 51.19 & 28.88 & 28.92 & 17.89 & 22.73 & 21.08\\
+ \hline
+ Domain Control & & & & & & & & \\
+ \hline
+ WDCMT (conditional GRU) & 68.76 & 35.71 & 52.75 & 32.03 & 29.81 & 19.58 & 24.63 & 23.28 \\
+ \hline 
+\end{tabularx}}
+\end{center}
+\caption{bpe-detokenized BLEU score for Recurrent network architecture on language pair English-French.
+Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{D18-1041} use Bidirectional-LSTM}
+\label{tab:3}
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+ \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
+ \hline
+ Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
+ \hline\hline
+ Mixed & 62.63 & 23.91 & 66.95 & 16.92 & 21.29 & 10.9 & & 14.19 \\
  \hline
@@
  \hline
- WDCMT (conditional GRU) & 68.76 & 35.71 & 52.75 & 32.03 & 29.81 & 19.58 & 24.63 & 23.28 \\
+ WDCMT (conditional GRU) & & & & & & & & \\
  \hline 
@@
 \end{center}
-\caption{bpe-detokenized BLEU score for Recurrent network architecture on language pair English-French.
-Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{D18-1041} use Bidirectional-LSTM}
-\label{tab:3}
-\end{table}
-
-\begin{table}[H]
-\begin{center}
- \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
- \hline
- Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT test 2010 \\ [0.5ex] 
- \hline\hline
- Mixed & & & & & & & \\
- \hline
- EMEA-finetuned & & & & & & & \\
- \hline
- EPPS-finetuned & & & & & & & \\
- \hline
- ECB-finetuned & & & & & & & \\
- \hline
- Sparse & & & & & & & \\
- \hline
- Domain Control & & & & & & & \\
- \hline
- WDCMT (conditional GRU) & & & & & & & \\
- \hline 
-\end{tabularx}}
-\end{center}
-\caption{BLEU scores for RNN architecture on English-German. Notations are the same as for figure~\ref{tab:3}}
+\caption{BLEU score for Recurrent network on language pair English - German. The notation is same as figure \ref{tab:3}}
 \label{tab:4}
-\end{table}
+\end{figure}
 
@@
 \fyTodo{Homegenize refs - urls, addresses, etc}
-\bibliography{naaclhlt2019}
-\bibliographystyle{plainnat}
-
+\bibliography{emnlp-ijcnlp-2019}
+\bibliographystyle{acl_natbib}
 \end{document}

### 2019-05-02 2019/emnlp/sparse.tex

@@
 However, Finetuning is very costly and unefficient in the industry therefore one best model is always desired.
-In this paper, we introduce an application of an old technique \cite{P07-1033} in Multi-Domain Neural Machine Translation.
+In this paper, we introduce an application of an old technique \cite{Daume07frustratingly} in Multi-Domain Neural Machine Translation.
 By defining general information encoding region and domain-specific encoding region in word embedding vector, we are able to seperate domain-related information flows in forward and backward propagation during training thus to
@@
 
-Despite fast developpement thanks to new architectures \cite{D13-1176, Sutskever2014Sequence, bahdanau2014neural, NIPS2017_7181}, Machine Translation (MT) still struggles with data scarcity.
+Despite fast developpement thanks to new architectures \cite{Kalchbrenner13recurrent, Sutskever2014Sequence, bahdanau2014neural, Vaswani17attention}, Machine Translation (MT) still struggles with data scarcity.
 Domain adaptation, or it's special case Multi-domain, is the situation where in-domain data (such as banking) is rare while out-domain data such as news are plenty, one aims to leverage out-domain data to improve quality of Machine Translation system which is perfectible while being trained with only the in-domain dataset.
@@
 It is ubiquitous that the bias to the in-domain dataset used in fine-tuning makes model worse in translating other domains, thus that we have not been able to fine-tune a model to several domains at once.
-This phenomenon can be implied by well-known problem in Machine Learning, Catastrophic Interference \cite{Michael1989Catastrophic}.
+This phenomenon can be implied by well-known problem in Machine Learning, Catastrophic Interference \cite{McCloskey1989Catastrophic}.
 When there is a shift in the distribution of training data, standard backpropagation neural network\fyTodo{Generic ?
@@
 % The problem
-Owing to the development of flexible and powerful architectures based on neural networks \cite{bahdanau2014neural;Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date.
+Owing to the development of flexible and powerful architectures based on neural networks \cite{bahdanau2014neural,Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date.
 One issue which repeatedly pops up with neural MT is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain.
@@
 Another important scenario for industrial MT systems is to adapt an neural system trained using a parallel data in one  domain to the peculiarity of other domains.
-Domain Adaption (DA) in MT is an ancient issue \cite{Foster07amixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied \cite{Chu18asurvey}.
+Domain Adaption (DA) in MT is an ancient issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied \cite{Chu18asurvey}.
 The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain.
@@
 In order to specialize the model, one usually finetunes generic model by continue the traing from $\theta^*$, which was optimized on general domain, with corpora of the interested topic.
-However, there is always a dramatical shift from generic domain $p_{gen}$ to specialization domain $p_{spec}$, that usually leads to catastrophic forgetting in neural network using backpropagation \cite{Michael1989Catastrophic}.
+However, there is always a dramatical shift from generic domain $p_{gen}$ to specialization domain $p_{spec}$, that usually leads to catastrophic forgetting in neural network using backpropagation \cite{McCloskey1989Catastrophic}.
 The finetuned model has lower performance in translating sentences, which are more related to other topic, than generic model.
@@
 $p_{t} = P * w_{t} = P_g \omega_g(t) + \displaystyle{\mathop{\sum}_{i \in [1,..,d]} P_i * \omega_i(t)}$\\
-where $w_{t}$ is word embedding of token $x_{t}$  is projection of word embedding (Transformer \cite{NIPS2017_7181}, recurrent \cite{bahdanau2014neural})
+where $w_{t}$ is word embedding of token $x_{t}$  is projection of word embedding (Transformer \cite{Vaswani17attention}, recurrent \cite{bahdanau2014neural})
 Assume $x_{t}$ belongs to domain $i$, thus $w_j(t) = 0 \forall j \neq i$. Denote $P_j$ the rows of the projection matrix $P$ corresponding to the deactivated region $w_j$.
@@
 \subsection{Implementation Details}
-The complex structure of State-of-the-art architecture Transformer \cite{NIPS2017_7181} and popular recurrent network \cite{bahdanau2014neural} makes the implementation of sparse embedding non-obvious.
+The complex structure of State-of-the-art architecture Transformer \cite{Vaswani17attention} and popular recurrent network \cite{bahdanau2014neural} makes the implementation of sparse embedding non-obvious.
 Indeed, Transformer involve positional embedding and residual connections which are not homogeneous for domain-regions.
@@
 \subsection{NMT model configuration}
-We evaluate our method on two state-of-the-art encoder-decoder architectures: a recurrent architecture based on attentional birectional Long Short-Term Memory (LSTM) cells and the Transformer architecture of \citet{NIPS2017_7181}.
-More precisely, for the Transformer we use the same settings as described in \cite{NIPS2017_7181}.
-For the recurrent architecture, in order to fairly compare to the work of \citet{D18-1041}, we use a standard configuration with one layer bidirectional LSTM with a encoder hidden size of 1024; one attentional unidirectional LSTM layer with same hidden size in decoder; and word embeddings of size 512.
+We evaluate our method on two state-of-the-art encoder-decoder architectures: a recurrent architecture based on attentional birectional Long Short-Term Memory (LSTM) cells and the Transformer architecture of \citet{Vaswani17attention}.
+More precisely, for the Transformer we use the same settings as described in \cite{Vaswani17attention}.
+For the recurrent architecture, in order to fairly compare to the work of \citet{Zeng18multidomain}, we use a standard configuration with one layer bidirectional LSTM with a encoder hidden size of 1024; one attentional unidirectional LSTM layer with same hidden size in decoder; and word embeddings of size 512.
 The implementation can be found in OpenNMT-tf \cite{P17-4012}.
@@
 \caption{bpe-detokenized BLEU score for Recurrent network architecture on language pair English-French.
-Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{D18-1041} use Bidirectional-LSTM}
+Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{Zeng18multidomain} use Bidirectional-LSTM}
 \label{tab:3}
@@
 Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
-Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
+Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{Chu18asurvey}; \cite{P17-2061}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
 The methods in this category can be listed as following:
@@
 \begin{itemize}
-	\item Trainining objective \cite{Luong2015SNMT,P16-1009,D17-1155,W17-3205,D17-1156,C18-1269}
-	\item Architecture \cite{gulcehre2016monolingual,C16-1170,R17-1049,W17-4712,Biao2017CARENMT,N18-2080,W18-6313,P18-2050}.
-	\item The decoding algorithm \cite{gulcehre2016monolingual,I17-2004}
+	\item Trainining objective \cite{Luong2015SNMT,Senrich2016Mono,Wang2017Instance,Chen2017Cost,Miceli2017Regularize,Zhang18sentence}
+	\item Architecture \cite{gulcehre2016monolingual,Zhang16topicinformed,Kobus17domaincontrol,Britz2017mixing,Biao2017CARENMT,Britz2017mixing,Thompson18freezing,Michel2018extreme}.
+	\item The decoding algorithm \cite{gulcehre2016monolingual,Khayrallah2017lattice}
 \end{itemize} 
-Beside these works, we could also consider works in other tasks such as sequence labeling tasks \cite{P07-1033}; learning multiple visual domains \cite{NIPS2017_6654}.
-The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041,D18-1041} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
+Beside these works, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.
+The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{Platanios18contextual}; \cite{Zeng18multidomain} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
 $h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively.
-While introducing new features in the architecture, \cite{D18-1041} also introduces new loss function which is sum of word-level weighted MT loss and losses of Domain-classifier in source side and target side; Adversarial Domain-classifier in source side.
+While introducing new features in the architecture, \cite{Zeng18multidomain} also introduces new loss function which is sum of word-level weighted MT loss and losses of Domain-classifier in source side and target side; Adversarial Domain-classifier in source side.
 The authors have very good approach to seperate Domain-shared information flow and Domain-specify information flow and to determine their contributions to the inference thus mitigate the catastrophic interference occuring in training with backpropagation.
-\cite{D18-1039} has other idea to circumvent the catastrophic interference.
+\cite{Platanios18contextual} has other idea to circumvent the catastrophic interference.
 They supposed that parameters of each domain specialized model could be generated from a universal parameter-generator.
@@
 
-Compared to \cite{D18-1039}, our sparse embedding is a bit similar to the idea.
+Compared to \cite{Platanios18contextual}, our sparse embedding is a bit similar to the idea.
 However, our method gives much more free degree for domain embedding since we give each word a domain region while the author uses only one vector domain embedding for each parameter set(encoder, decoder).
 
-Compared to the work of \cite{D18-1041}, our approach is much simpler and
+Compared to the work of \cite{Zeng18multidomain}, our approach is much simpler and
 more generic, since sparse embedding can be implemented for every neural architecture while domain-specific gate and domain-shared gate can only be applied for recurrent architectures.
-However, during inference, the model of \cite{D18-1041} can infer the domain that is the closest to the input sentence and that will be used in special gates; while our method needs to know in which domain the inference is executed.
+However, during inference, the model of \cite{Zeng18multidomain} can infer the domain that is the closest to the input sentence and that will be used in special gates; while our method needs to know in which domain the inference is executed.
 To circumvent this disadvantage, we could use a domain classifier or simply use only the generic region.

### 2019-05-02 2019/emnlp/sparse.tex
new
more refs, start revising soa

@@
 
-\fyTodo{s/naacl/EMNLP/g}
+\fyDone{s/naacl/EMNLP/g}
 \fyTodo{Too long, selfcontained, noref, rewrite}
@@
 Empirically, we achieved a unique model that has comparable score to finetuned models for every domains.
-Furthermore, our method is simple and could be applied for any architecture.
+Furthermore, our method is simple and could be applied to any architecture.
 \end{abstract}
@@
 % The problem
-Owing to the development of flexible and powerful architectures based on neural networks \cite{bahdanau2014neural,Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date.
+Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,bahdanau2014neural,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date.
 One issue which repeatedly pops up with neural MT is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain.
-A lot of the recent research effort has thus focused on developping MT systems in low data conditions, for instance building multi-lingual MTs enabling zero-shot translation\cite{} or even building systems without any data \cite{}.
+A lot of the recent research effort has thus focused on developping MT systems in low data conditions, for instance building multi-lingual MTs enabling zero-shot translation\cite{Firat16multiway,Johnson17google} or even building systems without any data \cite{Artetxe18unsupervised,Lample18unsupervised}.
 
 Another important scenario for industrial MT systems is to adapt an neural system trained using a parallel data in one  domain to the peculiarity of other domains.
-Domain Adaption (DA) in MT is an ancient issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied \cite{Chu18asurvey}.
+Domain Adaption (DA) in MT is an old issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied (see the recent survey of \cite{Chu18asurvey} for neural MT).
 The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain.
-We study here a different scenario, multi-domain adaptation \cite{Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
-This allows us to be both data-efficient (all data is used to train all domains) and compuationally efficient (we only need to train one domain).
-In our setting, we assume that the domain of test documents is known.
+We study here a different scenario, multi-domain adaptation \cite{}Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
+This allows us to be both data efficient (all data is used to train all domains) and compuationally efficient (we only need to train one domain).
+In our setting, we assume that the domain of test documents is known in advance.
 
-Multi-domain adaptation is conceptually close to multi-lingual MT\cite{}, or more generally to multi-task learning \cite{}, and can be approached in a number of ways.
-We adapt here the adaptation framework of \cite{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains.
-Our main hypothesis is that most of the differences between domains takes place at the lexical level, due to massive cross-domain polysemy.
-By contrast, the deeper layer, which arguably model more abstract phenomena, should be made shareable.
+Multi-domain adaptation is conceptually close to multi-lingual MT, or more generally to multi-task learning \cite{Caruana97multitask}, and can be approached in a number of ways.
+We adapt here the feature-expansion framework of \cite{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains.
+Our main hypothesis is that most of the differences between domains take place at the lexical level, due to massive cross-domain polysemy.
+By contrast, the deeper layers, which arguably model more abstract linguistic phenomena, should be made shareable across domains.
 To this end, we make sure to design word embeddings in such that they will contain a generic as well as a domain-specific region.
-We perform experiments with three domains, two architectures and two language pairs and find that our technique can effectively build multi-domain NMT, outperforming several baselines, including domain-specific MT engines trained through fine-tuning.
+We perform experiments with three domains, two architectures and two language pairs and find that our technique can effectively build multi-domain NMT, outperforming several baselines.
 Our contribution is thus twofold:
 \begin{itemize}
-\item methods to adapt the idea of \cite{Daume07frustratingly} for two NMT architectures;
+\item methods to adapt the ideas of \cite{Daume07frustratingly} for two NMT architectures;
 \item experimental results that demonstrate the effectiveness of this technique;
@@
 \end{itemize}
+
 Beside these works, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.
+
+The multi-domain scenario has been studied in a number of recent work.
+\citet{Farajian17multidomain} learn a generic system, and propose to dynamically adapt the network weights using a small sample of training data that resembles the test data; their main contribution is to propose a method to relate the amount of adaptation of the network parameters to the similarity between the adaptation sample and the test sentence: the higher the similarity, the more agressive the adaptation.
+
+Similarly to our approach, \cite{Zeng18multidomain} tries to separate on the encoder side domain-specific representation from generic representations in two different sub-networks, where shared versus domain-specific representations automatically emerge from two adversarial networks.
+The decoder side can thus attend separately to these two representations to generate its output.
+In this approach, the shared and domain-specific part kept separated in the deeper layers of the network, whereas we try to localize the differences between domain to take place at the lexical level.
+
+creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
+$h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively.
+While introducing new features in the architecture, \cite{Zeng18multidomain} also introduces new loss function which is sum of word-level weighted MT loss and losses of Domain-classifier in source side and target side; Adversarial Domain-classifier in source side.
+The authors have very good approach to seperate Domain-shared information flow and Domain-specify information flow and to determine their contributions to the inference thus mitigate the catastrophic interference occuring in training with backpropagation.
+
 The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{Platanios18contextual}; \cite{Zeng18multidomain} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
@@
 \section{Conclusions}
-\fyTodo{natural continuations: go beyond words; learn the projection matrix at the word level; other ?}
+\fyTodo{natural continuations: bayesian version ? linear combination of specialized engines ? go beyond words; learn the projection matrix at the word level; other ?}
 

### 2019-05-06 2019/emnlp/sparse.tex
new

@@
 A lot of the recent research effort has thus focused on developping MT systems in low data conditions, for instance building multi-lingual MTs enabling zero-shot translation\cite{Firat16multiway,Johnson17google} or even building systems without any data \cite{Artetxe18unsupervised,Lample18unsupervised}.
-
 Another important scenario for industrial MT systems is to adapt an neural system trained using a parallel data in one  domain to the peculiarity of other domains.
@@
 The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain.
-We study here a different scenario, multi-domain adaptation \cite{}Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
+We study here a different scenario, multi-domain adaptation \cite{Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
 This allows us to be both data efficient (all data is used to train all domains) and compuationally efficient (we only need to train one domain).
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}[h]
 \caption{Multi-domain Training}
@@
  \hline\hline
- Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 &  & 25.7 \\
+ Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 & 28.56 & 25.7 \\
  \hline
- EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 &  & 11.10 \\
+ EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.64 & 11.10 \\
  \hline
@@
  \hline
- Sparse & 68.51 & 35.75 & 51.19 & 28.88 & 28.92 & 17.89 & 22.73 & 21.08\\
+ Sparse (1) & 68.51 & 35.75 & 51.19 & 28.88 & 28.92 & 17.89 & 22.73 & 21.08 \\
+ \hline
+ Sparse (2) & & & & & & & & \\
  \hline
@@
  \hline\hline
- Mixed & 62.63 & 23.91 & 66.95 & 16.92 & 21.29 & 10.9 & & 14.19 \\
+ Mixed & 57.37 & 23.1 & 63.54 & 17.39 & 20.65 & 10.69 & 11.51 & 14.29 \\
  \hline
- EMEA-finetuned & & & & & & & & \\
+ EMEA-finetuned & 65.64 & 12.36 & 15.93 & 15.38 & 11.1 & 5.54 & 6.22 & 7.6 \\
  \hline
@@
  \hline
- Sparse & & & & & & & & \\
+ Sparse (1) & 63.43 & 22.66 & 64.4 & 16.99 & 20.63 & 10.89 & 11.45 & 14.26 \\
+ \hline
+ Sparse (2) & 63.27 & 22.62 & 63.7 & 17.26 & 20.69 & 10.89 & 11.45 & 14.26 \\
  \hline

### 2019-05-06 2019/emnlp/sparse.tex
new bib

@@
 For the recurrent architecture, in order to fairly compare to the work of \citet{Zeng18multidomain}, we use a standard configuration with one layer bidirectional LSTM with a encoder hidden size of 1024; one attentional unidirectional LSTM layer with same hidden size in decoder; and word embeddings of size 512.
-The implementation can be found in OpenNMT-tf \cite{P17-4012}.
+The implementation can be found in OpenNMT-tf \cite{Klein2017OpenNMT}.
 \fyTodo{Language pairs, BPE number, optimizer}
@@
 Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
-Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{Chu18asurvey}; \cite{P17-2061}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
+Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{Chu18asurvey}; \cite{Chu2017comparison}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
 The methods in this category can be listed as following:
@@
 \begin{itemize}
-\item score relatedness of sentences to the target domain by Language Model \cite{P10-2041,D11-1033,P13-2119}.
-\item generating pseudo parallel data \cite{P03-1010,C16-1295,D14-1023}
+\item score relatedness of sentences to the target domain by Language Model \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}.
+\item generating pseudo parallel data \cite{Utiyama2003measure,Wang2016connecting,Wang2014neural}
 \end{itemize}

### 2019-05-10 2019/emnlp/sparse.tex
new
rewrite table 5

@@
 \section{Introduction \label{sec:introduction}}
-\fyDone{Merge refs and order chronologically}
-\fyTodo{Homogeneize cite keys, avoid arxiv}
-
-Despite fast developpement thanks to new architectures \cite{Kalchbrenner13recurrent, Sutskever2014Sequence, bahdanau2014neural, Vaswani17attention}, Machine Translation (MT) still struggles with data scarcity.
-Domain adaptation, or it's special case Multi-domain, is the situation where in-domain data (such as banking) is rare while out-domain data such as news are plenty, one aims to leverage out-domain data to improve quality of Machine Translation system which is perfectible while being trained with only the in-domain dataset.
-Traditionally, ones train a generic model using a very large corpora then use in-domain dataset to fine-tune the model.
-It is ubiquitous that the bias to the in-domain dataset used in fine-tuning makes model worse in translating other domains, thus that we have not been able to fine-tune a model to several domains at once.
-This phenomenon can be implied by well-known problem in Machine Learning, Catastrophic Interference \cite{McCloskey1989Catastrophic}.
-When there is a shift in the distribution of training data, standard backpropagation neural network\fyTodo{Generic ?
-backprop or NN problem ?} gradually lose information learned from other domains while learning only target domain.
-
-In \fyTodostar{Specify DA}{Multi-domain Machine Translation}, ones have to deal with polysemies whose meaning can be totally different in different domains, for example, the word "chair" means an household tool in "if you are taking a medicine which may cause low blood pressure when rising from a chair or bed " (extracted from medical corpus EMEA \cite{Tiedemann2009RANLP5}) while the word "chair" means to preside over a meeting in following sentence "The President of the ECB or , in his absence , the Vice-President , chairs the meetings of the Governing Council , the Executive Board and the General Council of the ECB.
-" (extracted from European Central Bank corpora \cite{Tiedemann2009RANLP5}).
-Using context information is not sufficient to differ these polysemies.
-Therefore using the same word embedding for polysemies obstructs the model from differentiating the meaning of a polysemy in one domain from it's other meaning in the other domain.
-\begin{table}[h]
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{|l|l|l|}
-\hline
-English Words & French translation & Domain \\ \hline
-security & titre financier & Financial \\
-security & s\'ecuti\'e & IT\\
-\hline
-joint & articulaire & medical\\
-joint & conjoint & banking\\
-\hline
-chair & chaise & general \\
-chair & pr\'esider, pr\'esident & banking \\
-\hline
-\end{tabular}
-}
-\caption{Polysemies}
-\label{table:polysemy}
-\end{table}
-
-\begin{table}[h]
-\resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||}
-\hline
-English sentence & French translation & Corpus \\ 
-\hline
-"Flu-like" symptoms such as headaches, \colorbox{red}{joint} pains, feelings of weakness, dizziness, and tiredness may occur, especially at the start of treatment.
-& Des symptômes grippaux tels que céphalées, douleurs \colorbox{red}{articulaires}, sensation de faiblesse, vertige et asthénie, peuvent survenir, en particulier en début de traitement.
-& EMEA \\
-\hline
-In my opinion, Mr President, this \colorbox{red}{joint} debate fully bears this out. & Cette discussion \colorbox{red}{commune} en est, Monsieur le Président, la preuve vivante. & EPPS \\
-\hline
-The increasing number of central banks using the same IT platform for their own purposes has provided unique opportunities to undertake \colorbox{red}{joint} projects , exploit synergies and better manage the relationship with the vendor.
-& Le nombre croissant de banques centrales utilisant la même plate-forme informatique pour leurs besoins propres a offert des occasions uniques de lancer des projets \colorbox{red}{communs}, d'exploiter des synergies et de mieux gérer les relations avec le fournisseur.
-& ECB \\
-\hline
-\end{tabularx}
-}
-\caption{Translation}
-\label{table:polysemy}
-\end{table}
-\fyTodo{Typical plan: the problems, existing solutions and their limits, our idea and contributions}
-In this work, we present a new design for word embedding that is composed of different regions which are activate and deactivate corresponding to domains. 
-\\ To evaluate, we use 3 corpora of 3 different domains: Medical, Banking and Administration and 2 different architectures: Transformer and Attentional recurrent network.
-The test sets are selectioned from shared task each domains and two other general test sets such as news commentary and IWSLT.
-The reader could find more details in Experimetal section \ref{subsec:data}
-\fyTodo{More precise: two architectures, two languages pairs, multiple domains, also improves generic}
-
-\section{Introduction \label{sec:introduction}}
 % The problem
@@
 In machine translation, the loss function for training is:
-\begin{equation}
+\begin{equation}\label{eq:1}
 \begin{split}
@@
 In this equation, the domain is $p_{rel}(x,y)$.
-One aims to approximate to model to target domain $p_{rel}$.
-The more the domain of training data close to the domain of test set, the better performance it could achieve over test set.\\
-In the training of a generic model, the $p_{rel}$ is oftenly a general domain, noted $p_{gen}$ which is defined by a set of several corpora which belongs to different topics (e.g medical, banking, adminstration).
-Therefore the generic model is not usually good enough for translating by topic in which one is more interested.
-In order to specialize the model, one usually finetunes generic model by continue the traing from $\theta^*$, which was optimized on general domain, with corpora of the interested topic.
-However, there is always a dramatical shift from generic domain $p_{gen}$ to specialization domain $p_{spec}$, that usually leads to catastrophic forgetting in neural network using backpropagation \cite{McCloskey1989Catastrophic}.
-The finetuned model has lower performance in translating sentences, which are more related to other topic, than generic model.
+In minimizing given objective \ref{eq:1}, one aims to approximate the objective domain $p_{rel}$.
+In inference, provided the model was trained on certain training data, the more likely test set is generated from domain of training data, the better performance.\\
+In general, a generic model is trained first with a mix of several corpora which belongs to different topics (e.g medical, banking, adminstration) whose the domain $p_{rel}$ is considered as a general domain, noted $p_{gen}$.
+Being trained with a general mixed corpora, the generic model is not usually good enough for translating by topic.
+In order to improve topical translation quality, specializing generic model is required.
+One usually finetunes the generic model by continue the training from initial value $\theta^*$, which was optimized on general domain, using corpora of interested topic.
+However, there is always a dramatical shift from generic domain $p_{gen}$ to specialization domain $p_{spec}$.
 Indeed, the translation $(x,y)$ could be very different in different topic.
@@
 Therefore, the shift from general domain containing these two corpora to one of them is not negligible.
-\\
-In order to conduct the model translating by a certain topic, we might use conditional probability conditioned by the given topic $p(y|x,d)$.
+This shift usually leads to catastrophic forgetting in neural network using backpropagation \cite{McCloskey1989Catastrophic}.
+In effect, the finetuned model usually shows lower performance in translating sentences, which are less related to the finetuning topic, than generic model.
+In order to conduct the model translating topic by topic, we might refer to conditional probability conditioned by topic $p(y|x,d)$.
 Given a set of training corpora, we assume that there are a limited d engendering topic, i.e, there are d corpora $C_i$ $i \in [1,..,d]$ representing d topical domains $D(C_i)$ respectively.
@@
 From now, we formalize the multi-domain problem for machine translation in following equation
-\begin{equation}
+\begin{equation} \label{eq:2}
 \begin{split}
-\theta^* = \argmin_{\theta} E_{(x,y) \sim D(C_{i})} [-log & (p_{\theta}(y|x,i))] \; \forall i \in [1..d]
+\theta^* = \argmin_{\theta} \displaystyle{\mathop{\sum}_{i \in [1..d]}} E_{(x,y) \sim D(C_{i})}[-log(p_{\theta}(y|x,i))]
 \end{split}
 \end{equation}
-A trivial solution for this equation is $\theta^* = (\theta^*_{0},...,\theta^*_{d})$ and $p_{(\theta_{0},...,\theta_d)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times C(D_i) \rightarrow [0,1]: f_{\theta_i}(y,x) = p_{(\theta_{0},...,\theta_d)}(y|x,i)$ where $\Omega_i$ is set of all possible values of $\theta_i$.
+It is obvious that
+\begin{equation} \label{eq:3}
+\begin{split}
+\argmin_{\theta} \displaystyle{\mathop{\sum}_{i \in [1..d]}} E_{(x,y) \sim D(C_{i})} &[-log(p_{\theta}(y|x,i))] \\
+\geq \displaystyle{\mathop{\sum}_{i \in [1..d]}} \argmin_{\theta} E_{(x,y) \sim D(C_{i})} &[-log(p_{\theta}(y|x,i))]
+\end{split}
+\end{equation}
+A trivial solution for the equation \ref{eq:2} is $\theta^* = (\theta^*_{0},...,\theta^*_{d})$ and $p_{(\theta_{0},...,\theta_d)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times C(D_i) \rightarrow [0,1]: f_{\theta_i}(y,x) = p_{(\theta_{0},...,\theta_d)}(y|x,i)$ where $\Omega_i$ is set of all possible values of $\theta_i$.
 \\
-The model above is not different than ensemble of $d$ models specialized for one domain.
-However, we could mitigate this seperation by using shared paramaters denoted $\theta_s$ then reformulate the model as following: \\
+The trivial solution is not rather than ensemble of $d$ models specialized for one domain.
+However, we could reduce the seperation by using shared paramaters denoted $\theta_s$ then reformulate the model as following: \\
 $\theta^* = (\theta^*_{0},...,\theta^*_{d}, \theta_s^*)$ and $p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times \Omega_s \times C(D_i) \rightarrow [0,1]: f_{\theta_i, \theta_s}(y,x) = p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ \\ $p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x)$ is independent of $\theta_{j}$ $\forall j$ i.e, $\exists \bar{f}:\Omega_s \times C \rightarrow [0,1]: \bar{f}_{\theta_s}(y,x) = p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x)$
-where $\Omega_i$ is set of all possible values of $\theta_i$ and $\Omega_s$ is set of all possible values of $\theta_s$. \\
-However, there is a problem that the couple $\theta_s^*$ and $\theta_i^*$ have to be optimal $\forall i \in [1..d]$.\\
+where $\Omega_i$ is set of all possible values of $\theta_i$ and $\Omega_s$ is set of all possible values of $\theta_s$. The optimization problem will become
+\begin{equation}
+\displaystyle{\mathop{\sum}_{i \in [1..d]}} \argmin_{\theta_{0},...,\theta_{d}, \theta_s} E_{(x,y) \sim D(C_{i})} [-log(p_{\theta_s,\theta_i}(y|x,i))]
+\end{equation}
+However, the couple $\theta_s^*$ and $\theta_i^*$ are hardly optimal $\forall i \in [1..d]$.\\
 Despite the fact that $\theta^* = (\theta^*_{0},...,\theta^*_{d}, \theta_s^*)$ is hardly achieved by backpropagation, we could attempt to optimize $\theta_s$:
@@
 \end{equation}
+
 then optimize each $\theta_i$ (as additional factor used for fine-tuning the model to the domain $D(C_{i})$) the perplexity with respect to true distribution of domain
@@
 \subsection{Domains and data \label{ssec:data}}
-To evaluate the performance of proposed architecture, we design experiments in which we want to train an NMT for 3 domains such as medical, administrative documents and business briefly denoted by EMEA, EPPS, ECB according to training corpora name respectively.
+In our experiments, there are 3 domains such as medical, administrative documents and business briefly denoted by EMEA, EPPS, ECB according to training corpora name respectively and 2 language pairs: English-French, English-German.
 We use 3 corpora for training: European Medicines Agency; European Central Bank and European Parliament \cite{Tiedemann2009RANLP5}.
@@
 The model is also required to achieve comparable performance to generic model.
-To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.\fyTodo{Clarify / simplify text}
-
-\begin{table}[H]
-\fyTodo{Too small ! Use abbrevs (En, Fr, ECB, EMEA, EPPS)}
-  \resizebox{\columnwidth}{!}{
+To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
+\fyTodo{Clarify / simplify text}
+\begin{table*}
   \begin{tabular}{ |l|l|l|l|l|l| }
@@
 Task & Corpora & Train & Dev & Test & Vocabulary \\ \hline
-\multirow{5}{*}{$En \rightarrow Fr$} & European Medicines Agency & 1090568 & 1000 & 1000 & English: 30164, French: 30398\\
- & European Central Bank & 191960 & 1000 & 1000 &\\
- & European Parliament & 2005723 & 1000 & 1000 & \\ \hline
-\multirow{5}{*}{$En \rightarrow Ge$} & European Medicines Agency & 1108752 & 1000 & 1000 & English: 30159, German: 30698\\
- & European Central Bank & 111174 & 1000 & 1000 &\\
- & European Parliament & 1920209 & 1000 & 1000 & \\ \hline
+\multirow{5}{*}{$En \rightarrow Fr$} & EMEA & 1090568 & 1000 & 1000 & English: 30164, French: 30398\\
+ & ECB & 191960 & 1000 & 1000 &\\
+ & EPPS & 2005723 & 1000 & 1000 & \\ \hline
+\multirow{5}{*}{$En \rightarrow Ge$} & EMEA & 1108752 & 1000 & 1000 & English: 30159, German: 30698\\
+ & ECB & 111174 & 1000 & 1000 &\\
+ & EPPS & 1920209 & 1000 & 1000 & \\ \hline
 \hline
 \end{tabular}
-}
 \caption{Corpora }
 \label{table:Corpora}
-\end{table}
+\end{table*}
+To reduce size of vocabulary, we apply Byte-Pair Encoding with 30000 merge operations, then select approximately 30000 frequent words to build vocabulary for each language. 
 \subsection{Baselines \label{ssec:baselines}}
-\fyTodo{Describe baselines}
-
-\subsection{NMT model configuration}
-We evaluate our method on two state-of-the-art encoder-decoder architectures: a recurrent architecture based on attentional birectional Long Short-Term Memory (LSTM) cells and the Transformer architecture of \citet{Vaswani17attention}.
-More precisely, for the Transformer we use the same settings as described in \cite{Vaswani17attention}.
-For the recurrent architecture, in order to fairly compare to the work of \citet{Zeng18multidomain}, we use a standard configuration with one layer bidirectional LSTM with a encoder hidden size of 1024; one attentional unidirectional LSTM layer with same hidden size in decoder; and word embeddings of size 512.
-The implementation can be found in OpenNMT-tf \cite{Klein2017OpenNMT}.
-\fyTodo{Language pairs, BPE number, optimizer}
+To validate our findings, we compared sparse embedding model with standard models using State-of-the-Art Transformer \cite{Vaswani17attention} architecture and Attention Recurrent Neural Network architecture.
+Our baselines consist of generic models trained with mixed corpora (denoted mixed in the table of scores); models finetuned to each domain using 15000 iterations over in-domain data (denoted $+ft_{EMEA}$, $+ft_{EPPS}$, $+ft_{ECB}$ in the table of scores); models using Domain features of \cite{Kobus17domaincontrol} (noted DC in the table of scores) and multi-domain model proposed by \cite{Zeng18multidomain} (denoted WDCMT in the table of scores).
+\\
+For all of models, we set size of word embedding equal to 512; size of hidden layers equal to 1024 for Recurrent Neural Network and 512 for Transformer. 
+\begin{itemize}
+	\item Transformer: size of word embedding: 512; size of hidden layers: 512; size of inner feed forward layer: 2048; Multi-head: 8; number of layers: 6.
+	\item Attention Recurrent Neural Network: Bidirection LSTM encoder, unidirectional LSTM decoder, Luong Attention Mechanism, size of word embedding: 512; size of hidden layers: 1024; number of layers: 1
+	\item DC: size of domain embedding:2 ; same settings for both architecture: Transformer and Attention Recurrent Neural Network.
+	\item WDCMT: Bidirection Gated recurrent units (denoted GRU) encoder, unidirectional conditional GRU decoder, size of word embedding: 512; size of hidden layers: 1024; size of domain layers: 300; number of layers: 1
+\end{itemize}
+The implementation can be found in OpenNMT-tf\cite{Klein2017OpenNMT} and \url{https://github.com/DeepLearnXMU/WDCNMT}
+\subsection{Sparse NMT model configuration}
 \fyTodo{Motivate the split - discuss experimentally embedding size}
-For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions $[8,8,8,488]$, corresponding respectively to EMEA, EPPS, ECB domain and generic region.
-If sentence comes from domain $i$, regions for domains $j,\forall j \neq i$ will be masked as 0.
+For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions with size $[8,8,8,488]$ for instance, the choice will be dicussed in \ref{secc:region_size}.
+If sentence comes from domain $i$, regions for domains $j,\forall j \neq i$ will be masked as 0 while generic region is always activated.
 We then use a dense layer of size 512 to fuse the active domain region and the generic region.
@@
 
-Each iteration of algorithm \ref{algo:1}, the batch is doubled because it contains 2 types of word embedding for each sentence (using only generic region; using generic region and active domain region).
-The reason to do this is to minimize simultaneously both losses \\
+Each iteration of algorithm \ref{algo:1}, the batch contains 2 types of word embedding for each sentence: using only generic region; using generic region and present domain region.
+The reason to do this is to minimize simultaneously both losses: \\
 $\theta^*_{s}=\displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))]$ \\ 
@@
 $ 
-We admit that this is not the optimal way. However, this procedure makes training very simple.\fyTodo{Explain technicalities - future work ?}
+We admit that this is not the optimal way. However, this procedure makes training very simple. \fyTodo{Explain technicalities - future work ?}
 
-We also take in account the problem of overfitting that could occur for small corpora.
+We also take into account the problem of overfitting that could occur for small corpora.
 Indeed, real data are not always balanced for every domain.
@@
 If we train each corpora with same number of batches, ECB will be trained too oftenly, that easily leads to overfitting.
-Therefore, we perform backpropagation on corpora with frequencies according to sizes of corpora.
+Therefore, we select batches from each corpora with frequencies proportional to its size.
 \fyTodo{Refs on this ?
@@
 \subsection{Results}
-\begin{table}[H]
+\begin{table*}
 \begin{center}
- \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
+ \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
  \hline
@@
  \hline
- ECB-finetuned & 9.2 & 18.39 & 55.45 & 12.13 & 16.12 & 8.71 &  & 7.19 \\
+ ECB-finetuned & 26.93 & 27.09 & 56.52 & 22.04 & 22.92 & 12.94 & 17.51 & 13.99 \\
  \hline
- Sparse(1) & 70.45 & 38.23 & 54.69 & 34.66 & 31.24 & 21.36 & 28.12 & 26.02 \\
+ Sparse(1) & 74.26 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.70 \\
  \hline
- Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & 28.12 & 26.02 \\
+ Sparse(2) & 73.97 & 37.81 & 53.67 & 34.17 & 30.93 & 20.97 & 28.11 & 25.7 \\
  \hline
@@
  \hline
-\end{tabularx}}
+\end{tabularx}
 \end{center}
+
 \caption{BPE-detokenized BLEU score for English-French with the Transformer architecture.
@@
 \label{tab:1}
-\end{table}
+\end{table*}
 
-\begin{table}[H]
+\begin{table*}
 \begin{center}
- \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
+ \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
  \hline
@@
  \hline
- EMEA-finetuned & & & & & & & & \\
+ EMEA-finetuned & 68.35 & 17.02 & 32.87 & 21.15 & 15.94 & 8.92 & 11.21 & 13.49 \\
  \hline
- EPPS-finetuned & & & & & & & & \\
+ EPPS-finetuned & 36.19 & 26.29 & 40.71 & 19.67 & 23.69 & 14.08 & 16.47 & 19.24 \\
  \hline
- ECB-finetuned & & & & & & & & \\
+ ECB-finetuned & 24.72 & 18.36 & 74.05 & 11.31 & 16.64 & 9.11 & 10.8 & 11.0 \\
  \hline
- Sparse(1) & 70.55 & 26.17 & 68.07 & 21.50 & 23.30 & 14.05 & 16.64 & 19.04\\
+ Sparse(1) & 70.9 & 26.30 & 68.9 & 21.42 & 23.73 & 13.99 & 16.57 & 18.82 \\
  \hline
- Sparse(2) & 70.67 & 26.24 & 66.97 & & & 14.05 & 16.64 & 19.04 \\
+ Sparse(2) & 70.86 & 26.14 & 68.59 & 21.95 & 23.79 & 13.99 & 16.57 & 18.82 \\
  \hline
- Domain Control & & & & & & & & \\
+ Sparse (3) &  &  &  &  &  &  &  & \\
  \hline
-\end{tabularx}}
+ Domain Control & 63.48 & 26.27 & 66.95 & 21.45 & 23.95 & & & \\
+ \hline
+\end{tabularx}
 \end{center}
@@
 \label{tab:2}
-\end{table}
+\end{table*}
 
-\begin{figure}[H]
+\begin{table*}
 \begin{center}
- \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
+ \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
  \hline
@@
  \hline
- EMEA-finetuned  & & & & & & & & \\
+ EMEA-finetuned  & 72.06 & 18.62 & 16.78 & 25.88 & 13.38 & 9.26 & 10.94 & 5.73 \\
  \hline
- EPPS-finetuned & & & & & & & & \\
+ EPPS-finetuned & 35.47 & 34.61 & 39.56 & 27.11 & 28.8 & 17.3 & 21.85 & 20.61 \\
  \hline
- ECB-finetuned & & & & & & & & \\
+ ECB-finetuned & 21.93 & 22.6 & 51.53 & 14.74 & 20.23 & 9.74 & 12.10 & 10.47 \\
  \hline
- Sparse (1) & 68.51 & 35.75 & 51.19 & 28.88 & 28.92 & 17.89 & 22.73 & 21.08 \\
+ Sparse (1) & 71.73 & 35.21 & 50.91 & 29.62 & 29.05 & 17.92 & 23.00 & 20.96 \\
  \hline
- Sparse (2) & & & & & & & & \\
+ Sparse (2) & 70.59 & 35.22 & 49.68 & 29.51 & 29.32 & 17.92 & 23.00 & 20.96 \\
  \hline
- Domain Control & & & & & & & & \\
+ Sparse (3) &  &  &  &  &  &  &  & \\
+ \hline
+ Domain Control & 68.26 & 35.13 & 50.09 & 27.86 & 29.34 & & & \\
  \hline
@@
  \hline 
-\end{tabularx}}
+\end{tabularx}
 \end{center}
@@
 \label{tab:3}
-\end{figure}
+\end{table*}
 
-\begin{figure}[H]
+\begin{table*}
 \begin{center}
- \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
+ \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
  \hline
@@
  \hline
- EPPS-finetuned & & & & & & & & \\
+ EPPS-finetuned & 24.9 & 22.98 & 26.26 & 13.55 & 20.95 & 10.58 & 11.33 & 14.32 \\
  \hline
@@
  \hline
- Domain Control & & & & & & & & \\
+ Sparse (3) &  &  &  &  &  &  &  & \\
+ \hline
+ Domain Control & 62.53 & 23.74 & 65.71 & 16.88 & 21.85 & & & \\
  \hline
@@
 \label{tab:4}
-\end{figure}
+\end{table*}
 
+\section{Discussion\label{sec:Discussion}}
+\subsection{Region's size \label{secc:region_size}}
 \subsection{Word embedding analysis}

### 2019-05-13 2019/emnlp/sparse.tex
add
new

@@
 \end{equation} 
-Our first approach to multi-domain machine translation is to build such trivial solution in calibrating positions and sizes of domain paramters $\theta_i$ and shared parameters $\theta_s$ inside one architecture.
+Our first approach to multi-domain machine translation is to build such trivial solution by assigning different regions of word embedding vector to set of domain paramters $\theta_i$ and set of shared parameters $\theta_s$ while keep sharing other parameters between domains.
 \subsection{Sparse word embedding}
@@
 Figure~\ref{fig:network} presents our proposed design.
-During backprogation, the \fyTodo{fix notation}{projection matrix $P$} receives vanishing gradient at regions corresponding to deactivated regions in the word embedding.
-Those regions are also vanished in forward step, thus do not interfere the training on the domain to which they do not correspond.
+During backprogation, the \fyTodo{fix notation}{projection matrix $M$} receives gradient 0 at regions corresponding to deactivated regions in the word embedding.
+Those regions are also masked in forward step, thus do not interfere the training on the domain to which they are not assigned.
 
-Denote
-$p_{t} = P * w_{t} = P_g \omega_g(t) + \displaystyle{\mathop{\sum}_{i \in [1,..,d]} P_i * \omega_i(t)}$\\
-where $w_{t}$ is word embedding of token $x_{t}$  is projection of word embedding (Transformer \cite{Vaswani17attention}, recurrent \cite{bahdanau2014neural})
-Assume $x_{t}$ belongs to domain $i$, thus $w_j(t) = 0 \forall j \neq i$. Denote $P_j$ the rows of the projection matrix $P$ corresponding to the deactivated region $w_j$.
+Let $\tilde{x}_{t} = M * e_{t} = M_g e_g(t) + \displaystyle{\mathop{\sum}_{i \in [1,..,d]} M_i * e_i(t)}$ be the projection of word embedding where $e_{t}$ is word embedding of token $x_{t}$, M is projection matrix, $M_j \forall j \in [1...d]$ is row block corresponding to domain region $j$ (Transformer \cite{Vaswani17attention}, recurrent \cite{bahdanau2014neural})\\
+Assume the batch belongs to domain $i$, then $e_j(t) = 0$ $\forall j \neq i$,$\forall t$. 
 \begin{equation}
 \begin{split}
-\frac{\partial L(y_t,x_t)}{\partial P_j} &=\frac{\partial L}{\partial p}(p_t) * \frac{\partial p_t}{\partial P_j} \\
-								&= \frac{\partial L}{\partial p}(p_t) * w_j(t)\\
+\frac{\partial L(x,y)}{\partial M_j} &= \sum_{t}\frac{\partial L}{\partial \tilde{x}_t}(x_t) * \frac{\partial \tilde{x}_t}{\partial M_j} \\
+								&= \sum_{t} \frac{\partial L}{\partial \tilde{x}_t}(x_t) * e_j(t)\\
 									&=0
@@
 \end{equation}
-\subsection{Implementation Details}
+
 The complex structure of State-of-the-art architecture Transformer \cite{Vaswani17attention} and popular recurrent network \cite{bahdanau2014neural} makes the implementation of sparse embedding non-obvious.
@@
 \end{table*}
-To reduce size of vocabulary, we apply Byte-Pair Encoding with 30000 merge operations, then select approximately 30000 frequent words to build vocabulary for each language. 
+To reduce size of vocabulary, we apply Byte-Pair Encoding with 30000 merge operations, then select approximately 30000 the most frequent words to build vocabulary for each language. 
 \subsection{Baselines \label{ssec:baselines}}
 To validate our findings, we compared sparse embedding model with standard models using State-of-the-Art Transformer \cite{Vaswani17attention} architecture and Attention Recurrent Neural Network architecture.
-Our baselines consist of generic models trained with mixed corpora (denoted mixed in the table of scores); models finetuned to each domain using 15000 iterations over in-domain data (denoted $+ft_{EMEA}$, $+ft_{EPPS}$, $+ft_{ECB}$ in the table of scores); models using Domain features of \cite{Kobus17domaincontrol} (noted DC in the table of scores) and multi-domain model proposed by \cite{Zeng18multidomain} (denoted WDCMT in the table of scores).
+Our baselines consist of generic models trained with mixed corpora (denoted mixed in the table of scores); models finetuned to each domain using (10000, 15000, 5000) iterations over in-domain data (denoted $+ft_{EMEA}$, $+ft_{EPPS}$, $+ft_{ECB}$ in the table of scores); models using Domain features of \cite{Kobus17domaincontrol} (noted DC in the table of scores) and multi-domain model proposed by \cite{Zeng18multidomain} (denoted WDCMT in the table of scores).
 \\
@@
 \end{itemize}
+To train NMT systems, we use Adam($\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$) for Attention Recurrent Neural Network; and Adam with $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} ($warmup\_steps=4000$) for Transformer.
 The implementation can be found in OpenNMT-tf\cite{Klein2017OpenNMT} and \url{https://github.com/DeepLearnXMU/WDCNMT}
@@
 \fyTodo{Motivate the split - discuss experimentally embedding size}
-For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions with size $[8,8,8,488]$ for instance, the choice will be dicussed in \ref{secc:region_size}.
-If sentence comes from domain $i$, regions for domains $j,\forall j \neq i$ will be masked as 0 while generic region is always activated.
+For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions: 3 domain specific and 1 generic, with size $[8,8,8,488]$, the choice will be dicussed in \ref{secc:region_size}.
+If sentence comes from domain $i$, regions for domains $j \neq i$ will be masked as 0 while generic region is always activated.
 We then use a dense layer of size 512 to fuse the active domain region and the generic region.
@@
 
-Each iteration of algorithm \ref{algo:1}, the batch contains 2 types of word embedding for each sentence: using only generic region; using generic region and present domain region.
-The reason to do this is to minimize simultaneously both losses: \\
+In the training, we use same optimization algorithm as the baselines.
+Each iteration of algorithm \ref{algo:1}, we use 2 batches: generic batch using only generic region; and domain batch using generic region and domain region.
+The reason to do that is to minimize simultaneously both losses: \\
 $\theta^*_{s}=\displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))]$ \\ 
@@
 $
-We admit that this is not the optimal way. However, this procedure makes training very simple. \fyTodo{Explain technicalities - future work ?}
 
@@
  \hline
- EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.64 & 11.10 \\
+ EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.64 & 11.1 \\
  \hline
- EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 &  & 11.1 \\
+ EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 & 28.33 & 11.1 \\
  \hline
@@
  \hline
- Sparse(1) & 74.26 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.70 \\
+ Sparse(1) & 74.26 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.7 \\
  \hline
@@
  \hline
- Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 &  & & \\
+ Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 &  20.1 & 26.52 & 24.81\\
  \hline
@@
  \hline
- Domain Control & 63.48 & 26.27 & 66.95 & 21.45 & 23.95 & & & \\
+ Domain Control & 63.48 & 26.27 & 66.95 & 21.45 & 23.95 & 13.94 & 16.45 & 18.78  \\
  \hline
@@
  \hline
- Domain Control & 68.26 & 35.13 & 50.09 & 27.86 & 29.34 & & & \\
+ Domain Control & 68.26 & 35.13 & 50.09 & 27.86 & 29.34 & 16.25 & 20.95 & 21.16 \\
  \hline
@@
  \hline
- ECB-finetuned & & & & & & & & \\
+ ECB-finetuned & 41.8 & 15.97 & 71.07 & 10.77 & 14.67 & 7.06 & 7.66 & 8.07 \\
  \hline
@@
  \hline
- Domain Control & 62.53 & 23.74 & 65.71 & 16.88 & 21.85 & & & \\
+ Domain Control & 62.53 & 23.74 & 65.71 & 16.88 & 21.85 & 9.24 & 11.15 & 14.97 \\
  \hline
- WDCMT (conditional GRU) & & & & & & & & \\
+ WDCMT (conditional GRU) & 64.05 & 22.82 & 67 & 15.94 & 20.57 & 10.15 & 11.45 & 13.66 \\
  \hline 
@@
 \section{Discussion\label{sec:Discussion}}
+\subsection{Sparse's performance}
+The 4 table show remarquable improvement in EMEA translation which is comparable with finetuned model. Sparse model 
 \subsection{Region's size \label{secc:region_size}}

### 2019-05-13 2019/emnlp/sparse.tex
rewrote the abstract, and some part in the experiments

@@
 \fyTodo{Too long, selfcontained, noref, rewrite}
-\begin{abstract}
+\begin{abstract}help train
+  Supervised machine translation is well suited to the situation where the training and testing data are from the same domain.
+When this is not the case, \emph{domain adaptation} techniques are needed to ensure that the knowledge of the source, out-of-domain data, generalizes properly to the target, in domain, texts.
+We study a related setting, multi-domain adaptation, where the number of domains is potentially large, and where adapting separately to each woud waste training resources.
+Our proposal transposes to neural machine translation the feature expansion technique of (Daum\'e, 2007), and isolate domain-agnostic from domain-specific lexical representations, while sharing the rest of the architecture across domains.
+Our experiments use two standard architecture and two language pairs: they show that our approach, while remaining simple and computationally unexpensive, outperforms several strong baselines, and delivers a multi-domain system that can sucessfully process heterogeneous input sentences.
+
+  THIS IS THE OLD ABSTRACT 
   Fine-tuning a generic neural machine translation model to a certain domain usually degrade its performance on the other domains, thus it's very difficult to build a unique model which is best for all.
@@
 Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,bahdanau2014neural,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date.
-One issue which repeatedly pops up with neural MT is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain.
-A lot of the recent research effort has thus focused on developping MT systems in low data conditions, for instance building multi-lingual MTs enabling zero-shot translation\cite{Firat16multiway,Johnson17google} or even building systems without any data \cite{Artetxe18unsupervised,Lample18unsupervised}.
+One issue which repeatedly pops up with MT, be it neural or statistical, is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain.
+A lot of the recent research effort has thus focused on developping MT systems in low data conditions, for instance building multilingual MTs, thereby enabling zero-shot translation \cite{Firat16multiway,Johnson17google} or even building systems without any parallel data \cite{Artetxe18unsupervised,Lample18unsupervised}.
 Another important scenario for industrial MT systems is to adapt an neural system trained using a parallel data in one  domain to the peculiarity of other domains.
@@
 
-Multi-domain adaptation is conceptually close to multi-lingual MT, or more generally to multi-task learning \cite{Caruana97multitask}, and can be approached in a number of ways.
-We adapt here the feature-expansion framework of \cite{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains.
+Multi-domain adaptation is conceptually close to multilingual MT, or more generally to multi-task learning \cite{Caruana97multitask}, and can be approached in a number of ways.
+We adapt here the of \cite{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains.
 Our main hypothesis is that most of the differences between domains take place at the lexical level, due to massive cross-domain polysemy.
@@
 \center
-    \includegraphics[width=0.8\linewidth]{Sparse1}
+    \includegraphics[width=\linewidth]{embeddings}
     \caption{Sparse word embeddings.
@@
 \subsection{Domains and data \label{ssec:data}}
-In our experiments, there are 3 domains such as medical, administrative documents and business briefly denoted by EMEA, EPPS, ECB according to training corpora name respectively and 2 language pairs: English-French, English-German.
-We use 3 corpora for training: European Medicines Agency; European Central Bank and European Parliament \cite{Tiedemann2009RANLP5}.
-We randomly split those 3 corpora into training, validation and test set with size described in table \ref{table:Corpora}.
-We use validation sets for choosing the best model according to the average BLEU score on 3 validation sets.
-We also realize that the test sets origined from same corpora as training sets are likely easy for trained model, thus can not provide fair evaluation.
-Therefore, we we use Khresmoi-test; newstest2009 and test2007 for domains: EMEA, EPPS.
-Unfortunately, we could not find any official test sets for ECB domain.
-We also want to evaluate the generalization of trained model.
-The model is also required to achieve comparable performance to generic model.
-To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
-\fyTodo{Clarify / simplify text}
-\begin{table*}
-  \begin{tabular}{ |l|l|l|l|l|l| }
+We perform our experiments with two language pairs (English-French, English-German )and data originating from three domains, corresponding to texts from three European institutions: from the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}.
+We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{table:Corpora}).
+The validation sets are used to chose the best model according to the average BLEU score.
+Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other, related, domains: the official Khresmoi-test \fyTodo{REF needed}, which is close to EMEA, and News test 2009 \cite{WMT2009}.
+This enables us to evaluate the loss in performance when the test set is from a different domain.
+% The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
+\fyTodo{Check which corpus are useful}
+\begin{table}
+  \centering
+  \begin{tabular}{ *{4}{|r|}}
     \hline
-Task & Corpora & Train & Dev & Test & Vocabulary \\ \hline
-\multirow{5}{*}{$En \rightarrow Fr$} & EMEA & 1090568 & 1000 & 1000 & English: 30164, French: 30398\\
- & ECB & 191960 & 1000 & 1000 &\\
- & EPPS & 2005723 & 1000 & 1000 & \\ \hline
-\multirow{5}{*}{$En \rightarrow Ge$} & EMEA & 1108752 & 1000 & 1000 & English: 30159, German: 30698\\
- & ECB & 111174 & 1000 & 1000 &\\
- & EPPS & 1920209 & 1000 & 1000 & \\ \hline
+    Corpus & Train & Dev & Test \\ \hline
+    \multicolumn{4}{|l|}{$En \rightarrow Fr$ }\\
+    \multicolumn{4}{|l|}{Vocab: En 30,165, Fr: 30,398}\\
     \hline
+    EMEA  & 1,09M & 1000 & 1000 \\
+    ECB    & 0,19M & 1000 & 1000     \\
+    EPPS   & 2,0M  & 1000 & 1000  \\ \hline \hline
+    \multicolumn{4}{|l|}{$En \rightarrow Ge$}\\
+    \multicolumn{4}{|l|}{ Vocab: En 30,159, Ge: 30,698}\\ \hline
+    EMEA  & 1,11M & 1000 & 1000 \\
+    ECB     &  0,11M & 1000 & 1000  \\
+    EPPS   & 1,92M & 1000 & 1000 \\ \hline
 \end{tabular}
 \caption{Corpora }
-\label{table:Corpora}
-\end{table*}
+\label{table:Corpora from three domains}
+\end{table}
+
 To reduce size of vocabulary, we apply Byte-Pair Encoding with 30000 merge operations, then select approximately 30000 the most frequent words to build vocabulary for each language. 

### 2019-05-13 2019/emnlp/sparse.tex
fix a glitch in abstract

@@
 \fyDone{s/naacl/EMNLP/g}
-\fyTodo{Too long, selfcontained, noref, rewrite}
-\begin{abstract}help train
-  Supervised machine translation is well suited to the situation where the training and testing data are from the same domain.
-When this is not the case, \emph{domain adaptation} techniques are needed to ensure that the knowledge of the source, out-of-domain data, generalizes properly to the target, in domain, texts.
-We study a related setting, multi-domain adaptation, where the number of domains is potentially large, and where adapting separately to each woud waste training resources.
-Our proposal transposes to neural machine translation the feature expansion technique of (Daum\'e, 2007), and isolate domain-agnostic from domain-specific lexical representations, while sharing the rest of the architecture across domains.
-Our experiments use two standard architecture and two language pairs: they show that our approach, while remaining simple and computationally unexpensive, outperforms several strong baselines, and delivers a multi-domain system that can sucessfully process heterogeneous input sentences.
+\fyDone{Too long, selfcontained, noref, rewrite}
+\begin{abstract}
+  Supervised machine translation is well suited to the case where the train and test data are from the same domain.
+When this is not the case, \emph{domain adaptation} techniques help ensure that the knowledge of the out-of-domain data generalizes properly to target domain texts.
+We study a related setting, multi-domain adaptation, where the number of domains is potentially large, and where adapting separately to each domain woud waste training resources.
+Our proposal transposes to neural machine translation the feature expansion technique of (Daum\'e, 2007), and isolates domain-agnostic from domain-specific lexical representations, while sharing the rest of the network across domains.
+Our experiments use two architectures and two language pairs: they show that our approach, while remaining simple and computationally unexpensive, outperforms several strong baselines, and delivers a multi-domain system that can sucessfully process heterogeneous input sentences.
 

### 2019-05-13 2019/emnlp/sparse.tex
new

--- 2019-05-13 2019/emnlp/sparse.tex
@@
 \begin{algorithmic}[1]
-\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for d domains}
+\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for d domains, Optimization algorithm Opt}
 \REPEAT 
@@
 \STATE{Randomly pick a batch of sentences from corpora $C_i$.}
-\STATE{Activating only generic activate embedding to create generic embedding batch, denoted $W_g$}
-\STATE{Pass the batch of word embedding to a dense layer to fuse regions}
-\STATE{Computing gradient of $\theta_s$, $\frac{\partial L}{\partial \theta_s}$ using generic batch $W_g$}
+\STATE{Activating only generic region to create generic embedding batch, denoted $W_g$}
+\STATE{Pass the generic embedding batch to a dense layer.}
+\STATE{Computing gradient of $\theta_s$, $\frac{\partial L}{\partial \theta_s}$ using generic embedding batch $W_g$}
 \STATE{Activating corresponding domain region and generic region to create domain embedding batch}
-\STATE{Pass the batch of word embedding to a dense layer to fuse regions.}
-\STATE{Computing gradient of domain parameters $\theta_i$, $\frac{\partial L}{\partial \theta_i}$ using domain batch }
-\STATE{Apply backprogation using computed gradients $\frac{\partial L}{\partial \theta_s}$ and $\frac{\partial L}{\partial \theta_i}$}
+\STATE{Pass the batch of domain embedding to a dense layer}
+\STATE{Computing gradient of domain parameters $\theta_i$, $\frac{\partial L}{\partial \theta_i}$ using domain batch}
+\STATE{Apply learning rules using computed gradients: 
+$$ \theta_s = \theta_s - Opt(\frac{\partial L}{\partial \theta_s}(W_g))$$ 
+$$ \theta_i = \theta_i - Opt(\frac{\partial L}{\partial \theta_i}(W_d))$$}
 \UNTIL{convergence}
@@
  \hline
- Sparse(1) & 74.26 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.7 \\
+ Sparse (1) & 74.26 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.7 \\
  \hline
- Sparse(2) & 73.97 & 37.81 & 53.67 & 34.17 & 30.93 & 20.97 & 28.11 & 25.7 \\
+ Sparse (2) & 73.97 & 37.81 & 53.67 & 34.17 & 30.93 & 20.97 & 28.11 & 25.7 \\
  \hline
- Sparse(3) &  &  &  &  &  &  &  & \\
+ Sparse (3) &  &  &  &  &  &  &  & \\
  \hline
@@
 \subsection{Sparse's performance}
-The 4 table show remarquable improvement in EMEA translation which is comparable with finetuned model. Sparse model 
+First, we analyze the performance of models on inner test sets which are created by randomly splited from same corpora as training set.
+The 4 table show remarquable improvement 6 points BLEU in average in EMEA translation which is comparable with finetuned model.
+In all most the cases, Sparse model has slightly higher score on EPPS test and ECB test than mixed and finetuned models.
+Sparse model also achieve comparable scores with Domain Control and WDCMT in average.
+In considering the score of finetuned model on out-of-domain test sets, we observed a fast catastrophic forgetting while the out-of-domain scores drop dramatically compared to mixed model.
+Taking into account the average score on 3 inner test sets, Sparse achieved far better score than all finetuned models.
+\\
+Secondly, we compare performance of models on IWSLT test 2010 and newstest 2014 which are more general than inner test sets.
+On these test sets, mixed, Sparse has approximately same score in Transformer and are a little better than Domain Control while in Recurrent Net Work architecture, WDCMT is better than mixed, Sparse and Domain Control for English -French pair.
+We also noted a quick fall of performance in finetuned model.
+
 \subsection{Region's size \label{secc:region_size}}

### 2019-05-14 2019/emnlp/sparse.tex
Fix the abstract and intro, more refs, fix experiments

@@
 \begin{abstract}
-  Supervised machine translation is well suited to the case where the train and test data are from the same domain.
-When this is not the case, \emph{domain adaptation} techniques help ensure that the knowledge of the out-of-domain data generalizes properly to target domain texts.
-We study a related setting, multi-domain adaptation, where the number of domains is potentially large, and where adapting separately to each domain woud waste training resources.
-Our proposal transposes to neural machine translation the feature expansion technique of (Daum\'e, 2007), and isolates domain-agnostic from domain-specific lexical representations, while sharing the rest of the network across domains.
-Our experiments use two architectures and two language pairs: they show that our approach, while remaining simple and computationally unexpensive, outperforms several strong baselines, and delivers a multi-domain system that can sucessfully process heterogeneous input sentences.
+  Supervised machine translation works well when the train and test data are sampled from the same distribution.
+When this is not the case, \emph{adaptation} techniques help ensure that the knowledge learned from out-of-domain texts generalizes to target domain sentences.
+We study here a related setting, \emph{multi-domain adaptation}, where the number of domains is potentially large and adapting separately to each domain woud waste training resources.
+Our proposal transposes to neural machine translation the feature expansion technique of (Daum\'e III, 2007): it isolates domain-agnostic from domain-specific lexical representations, while sharing the rest of the network across domains.
+Our experiments use two neural architectures and two language pairs: they show that our approach, while remaining simple and computationally unexpensive, outperforms several strong baselines and delivers a multi-domain system that can sucessfully translate heterogeneous texts.
 
-  THIS IS THE OLD ABSTRACT 
-  Fine-tuning a generic neural machine translation model to a certain domain usually degrade its performance on the other domains, thus it's very difficult to build a unique model which is best for all.
+%   \fyTodo{Remove this}
+%   THIS IS THE OLD ABSTRACT 
+%   Fine-tuning a generic neural machine translation model to a certain domain usually degrade its performance on the other domains, thus it's very difficult to build a unique model which is best for all.
 Indeed, Catastrophic Forgetting, well-known problem in Machine Learning, states that standard backprogation neural network suffers a lot while there is a shift in training data distribution.
@@
 By defining general information encoding region and domain-specific encoding region in word embedding vector, we are able to seperate domain-related information flows in forward and backward propagation during training thus to
-mitigate the catastrophic interference happens during the training over different domains.
+% mitigate the catastrophic interference happens during the training over different domains.
 Empirically, we achieved a unique model that has comparable score to finetuned models for every domains.
@@
 % The problem
-Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,bahdanau2014neural,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date.
+Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,bahdanau2014neural,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progresses over the past years, and constitute to date the standard for most production engines.
 One issue which repeatedly pops up with MT, be it neural or statistical, is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain.
-A lot of the recent research effort has thus focused on developping MT systems in low data conditions, for instance building multilingual MTs, thereby enabling zero-shot translation \cite{Firat16multiway,Johnson17google} or even building systems without any parallel data \cite{Artetxe18unsupervised,Lample18unsupervised}.
-Another important scenario for industrial MT systems is to adapt an neural system trained using a parallel data in one  domain to the peculiarity of other domains.
+A lot of the recent research effort has thus focused on developping MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Johnson17google}; or even building systems without any parallel data \cite{Artetxe18unsupervised,Lample18unsupervised}.
+Another important scenario for industrial MT systems is to adapt a neural system trained using parallel data in one domain to the peculiarity of other domains.
 Domain Adaption (DA) in MT is an old issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied (see the recent survey of \cite{Chu18asurvey} for neural MT).
 The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain.
-We study here a different scenario, multi-domain adaptation \cite{Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
+We study here a different scenario, multi-domain adaptation \cite{Sennrich13multidomain,Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
 This allows us to be both data efficient (all data is used to train all domains) and compuationally efficient (we only need to train one domain).
@@
 Multi-domain adaptation is conceptually close to multilingual MT, or more generally to multi-task learning \cite{Caruana97multitask}, and can be approached in a number of ways.
-We adapt here the of \cite{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains.
-Our main hypothesis is that most of the differences between domains take place at the lexical level, due to massive cross-domain polysemy.
-By contrast, the deeper layers, which arguably model more abstract linguistic phenomena, should be made shareable across domains.
+We adapt here theideas of \cite{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains.
+Our main hypothesis is that most of the differences between domains take place at the lexical level, due to cross-domain polysemy.
+By contrast, the deeper layers, which arguably model more abstract linguistic phenomena, are made shareable across domains.
 To this end, we make sure to design word embeddings in such that they will contain a generic as well as a domain-specific region.
-We perform experiments with three domains, two architectures and two language pairs and find that our technique can effectively build multi-domain NMT, outperforming several baselines.
+We perform experiments with three domains, two neural architectures and two language pairs and find that our technique can effectively build multi-domain NMT, outperforming several baselines.
 Our contribution is thus twofold:
@@
 \subsection{Domains and data \label{ssec:data}}
-We perform our experiments with two language pairs (English-French, English-German )and data originating from three domains, corresponding to texts from three European institutions: from the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}.
+
+We experiment with two language pairs (English-French, English-German )and data originating from three domains, corresponding to texts from three European institutions: from the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}.
 We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{table:Corpora}).
@@
 
-To reduce size of vocabulary, we apply Byte-Pair Encoding with 30000 merge operations, then select approximately 30000 the most frequent words to build vocabulary for each language. 
+To reduce the vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16 }with 30,000 merge operations, then select approximately 30000 the most frequent words to build vocabulary for each language.\fyTodo{I need explanations here}
+
 \subsection{Baselines \label{ssec:baselines}}
-To validate our findings, we compared sparse embedding model with standard models using State-of-the-Art Transformer \cite{Vaswani17attention} architecture and Attention Recurrent Neural Network architecture.
-Our baselines consist of generic models trained with mixed corpora (denoted mixed in the table of scores); models finetuned to each domain using (10000, 15000, 5000) iterations over in-domain data (denoted $+ft_{EMEA}$, $+ft_{EPPS}$, $+ft_{ECB}$ in the table of scores); models using Domain features of \cite{Kobus17domaincontrol} (noted DC in the table of scores) and multi-domain model proposed by \cite{Zeng18multidomain} (denoted WDCMT in the table of scores).
-\\
-For all of models, we set size of word embedding equal to 512; size of hidden layers equal to 1024 for Recurrent Neural Network and 512 for Transformer. 
+To validate our findings, we compared sparse embedding models with standard models using both attentional Recurrent Neural Networks (RNNs) \cite{Bahdanau15learning} and the recent Transformer architecture of \citet{Vaswani17attention}.
+Our baselines consist of generic models trained with a simple concatenation of all corpora (denoted \texttt{mixed} in the score tables below); models finetuned separately to each domain using (10000, 15000, 5000) iterations over in-domain data (denoted $+ft_{EMEA}$, $+ft_{EPPS}$, $+ft_{ECB}$ below); models using domain tags as in \cite{Kobus17domaincontrol} (noted \texttt{DC} below) and as well as the multi-domain model of \citet{Zeng18multidomain} (denoted \texttt{WDCMT} in the tables below).
+
+For all models, we set the size of word embeddings equal to 512; the size of hidden layers is equal to 1024 for RNNs and 512 for Transformer.
+\fyTodo{Improve this}
 \begin{itemize}
@@
 \end{itemize}
-To train NMT systems, we use Adam($\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$) for Attention Recurrent Neural Network; and Adam with $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} ($warmup\_steps=4000$) for Transformer.
-The implementation can be found in OpenNMT-tf\cite{Klein2017OpenNMT} and \url{https://github.com/DeepLearnXMU/WDCNMT}
+To train NMT systems, we use Adam($\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$) for RNNs; and Adam with $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} ($warmup\_steps=4000$) for Transformer.
+Our implementation is from OpenNMT-tf\cite{Klein2017OpenNMT}, except for WDCMT where we use the authors implementation\footnote{\url{https://github.com/DeepLearnXMU/WDCNMT}.}.
+
 \subsection{Sparse NMT model configuration}

### 2019-05-15 2019/emnlp/sparse.tex
ligth change

@@
   Supervised machine translation works well when the train and test data are sampled from the same distribution. 
-When this is not the case, \emph{adaptation} techniques help ensure that the knowledge learned from out-of-domain texts generalizes to target domain sentences.
-We study here a related setting, \emph{multi-domain adaptation}, where the number of domains is potentially large and adapting separately to each domain woud waste training resources.
+  When this is not the case, \emph{adaptation} techniques help ensure that the knowledge learned from out-of-domain texts generalises to target domain sentences. 
+  We study here a related setting, \emph{multi-domain adaptation}, where the number of domains is potentially large and adapting separately to each domain would waste training resources. 
   Our proposal transposes to neural machine translation the feature expansion technique of (Daum\'e III, 2007): it isolates domain-agnostic from domain-specific lexical representations, while sharing the rest of the network across domains.
-Our experiments use two neural architectures and two language pairs: they show that our approach, while remaining simple and computationally unexpensive, outperforms several strong baselines and delivers a multi-domain system that can sucessfully translate heterogeneous texts.
+  Our experiments use two neural architectures and two language pairs: they show that our approach, while remaining simple and computationally unexpensive, outperforms several strong baselines and delivers a multi-domain system that can successfully translate heterogeneous texts.
 

### 2019-05-15 2019/emnlp/sparse.tex
ligth change

@@
   Our proposal transposes to neural machine translation the feature expansion technique of (Daum\'e III, 2007): it isolates domain-agnostic from domain-specific lexical representations, while sharing the rest of the network across domains.
-  Our experiments use two neural architectures and two language pairs: they show that our approach, while remaining simple and computationally unexpensive, outperforms several strong baselines and delivers a multi-domain system that can successfully translate heterogeneous texts.
+  Our experiments use two neural architectures and two language pairs: they show that our approach, while remaining simple and computationally inexpensive, outperforms several strong baselines and delivers a multi-domain system that can successfully translate heterogeneous texts.
 

### 2019-05-15 2019/emnlp/sparse.tex
More refs + more related work

@@
 \title{Sparse Word Embeddings for Multi-domain Machine Translation}
+\title{Lexicalized Domain Embeddings for Multi-domain Machine Translation}
 
@@
 
-To reduce the vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16 }with 30,000 merge operations, then select approximately 30000 the most frequent words to build vocabulary for each language.\fyTodo{I need explanations here}
+To reduce the vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations, then select approximately 30,000 the most frequent words to build vocabulary for each language.\fyTodo{I need explanations here}
 
@@
 \end{itemize}
-On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features
+On the other hand, model-centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features
 \begin{itemize}
@@
 
-The multi-domain scenario has been studied in a number of recent work.
-\citet{Farajian17multidomain} learn a generic system, and propose to dynamically adapt the network weights using a small sample of training data that resembles the test data; their main contribution is to propose a method to relate the amount of adaptation of the network parameters to the similarity between the adaptation sample and the test sentence: the higher the similarity, the more agressive the adaptation.
+%% HERE
+
+The multi-domain scenario has been studied in a number of recent works.
+\citet{Farajian17multidomain} learn a generic system and propose to dynamically adapt the network weights using a small sample of training data that resembles the test data, an idea already explored in \cite{Sennrich13multidomain}.
+Their main contribution is to propose a method to relate the amount of adaptation of the network parameters to the similarity between the adaptation sample and the test sentence: the higher the similarity, the more agressive the adaptation.
+
+By analogy with multilingual NMT, \citet{Kobus17domaincontrol} and \citet{Chu18multilingual} separately propose to extend the representation of the source text with a domain tag (or its embedding).
+Our model also modifies input representations, but allows for each source word to have its own domain representation, thereby improving training of the shared parts of the network.\fyTodo{Talk about target embeddings somewhere.}
 
@@
 The decoder side can thus attend separately to these two representations to generate its output.
-In this approach, the shared and domain-specific part kept separated in the deeper layers of the network, whereas we try to localize the differences between domain to take place at the lexical level.
+In this approach, the shared and domain-specific part kept separated in the deeper layers of the network, whereas we try to localize the differences between domains at the lexical level, based on a much cheaper computational architecture.
 

### 2019-05-15 2019/emnlp/sparse.tex
region size experiments, polysemy experiment

@@
 % The problem
-Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,bahdanau2014neural,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progresses over the past years, and constitute to date the standard for most production engines.
+Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,Bahdanau15learning,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progresses over the past years, and constitute to date the standard for most production engines.
 One issue which repeatedly pops up with MT, be it neural or statistical, is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain.
@@
 
-Let $\tilde{x}_{t} = M * e_{t} = M_g e_g(t) + \displaystyle{\mathop{\sum}_{i \in [1,..,d]} M_i * e_i(t)}$ be the projection of word embedding where $e_{t}$ is word embedding of token $x_{t}$, M is projection matrix, $M_j \forall j \in [1...d]$ is row block corresponding to domain region $j$ (Transformer \cite{Vaswani17attention}, recurrent \cite{bahdanau2014neural})\\
+Let $\tilde{x}_{t} = M * e_{t} = M_g e_g(t) + \displaystyle{\mathop{\sum}_{i \in [1,..,d]} M_i * e_i(t)}$ be the projection of word embedding where $e_{t}$ is word embedding of token $x_{t}$, M is projection matrix, $M_j \forall j \in [1...d]$ is row block corresponding to domain region $j$ (Transformer \cite{Vaswani17attention}, recurrent \cite{Bahdanau15learning})\\
 Assume the batch belongs to domain $i$, then $e_j(t) = 0$ $\forall j \neq i$,$\forall t$. 
@@
 
-The complex structure of State-of-the-art architecture Transformer \cite{Vaswani17attention} and popular recurrent network \cite{bahdanau2014neural} makes the implementation of sparse embedding non-obvious.
+The complex structure of State-of-the-art architecture Transformer \cite{Vaswani17attention} and popular recurrent network \cite{Bahdanau15learning} makes the implementation of sparse embedding non-obvious.
 Indeed, Transformer involve positional embedding and residual connections which are not homogeneous for domain-regions.
@@
 We experiment with two language pairs (English-French, English-German )and data originating from three domains, corresponding to texts from three European institutions: from the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}.
-We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{table:Corpora}).
+We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).
 The validation sets are used to chose the best model according to the average BLEU score.
-Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other, related, domains: the official Khresmoi-test \fyTodo{REF needed}, which is close to EMEA, and News test 2009 \cite{WMT2009}.
+Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other, related, domains: the official Khresmoi-test \fyTodo{REF needed}, which is close to EMEA, and News test 2014 \fyTodo{REF needed}.
 This enables us to evaluate the loss in performance when the test set is from a different domain.
@@
 \end{tabular}
-\caption{Corpora }
-\label{table:Corpora from three domains}
+\caption{Corpora}
+\label{tab:Corpora}
 \end{table}
@@
 \subsection{Region's size \label{secc:region_size}}
+We also test different setting of region size in word embedding.
+However, we limit our research in 4 settings: $[4,4,4,496]$, $[8,8,8,488]$, $[16,16,16,464]$, $[32,32,32,32,416]$.
+We test those settings with the same condition as above.
+The scores is showed in table \ref{tab:5}
+\begin{table*}
+\begin{center}
+ \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X | X ||} 
+ \hline
+ Models & EMEA test & EMEA excluded & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
+ \hline\hline 
+ Sparse - 4 & 74.65 & 49.61 & 37.42 & 54.49 & 33.68 & 31.01 & 20.71 & 28.03 & 25.28 \\
+ \hline
+ Sparse - 8 & 74.26 & 49.9 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.7 \\
+ \hline
+ Sparse - 16 & 74.15 & 49.1 & 37.78 & 54.56 & 33.4 & 30.93 & 20.73 & 28.1 & 25.33 \\
+ \hline
+ Sparse - 32 & 75.1 & 48.61 & 37.64 & 54.29 & 33.2 & 30.94 & 21.07 & 27.99 & 25.44 \\
+ \hline
+\end{tabularx}
+\end{center}
+\caption{BPE-detokenized BLEU score for English-French with the Transformer architecture}
+\label{tab:5}
+\end{table*}
+
 \subsection{Word embedding analysis}
+We analyze neigborhood in word embedding space of polysemies.
+
+\begin{table}
+  \centering
+  \begin{tabular}{*{4}{|r|}}
+    \hline
+    Domain & Word & Neighbor & Cosinus distance \\ \hline
+    \multirow{5}{*}{EMEA} & \multirow{5}{*}{joint} 
+    & Joint & 0.371 \\  
+    &  & jointly & 0.545 \\
+    &  & artic￭ & 0.605 \\
+    &  & joints & 0.607 \\
+	&  & collective & 0.609 \\
+	\hline
+	\multirow{5}{*}{EPPS} & \multirow{5}{*}{joint} 
+	& Joint & 0.365 \\
+	& & jointly & 0.497 \\
+	& & join￭ & 0.606 \\
+	& & common & 0.607 \\
+	& & collective & 0.626 \\
+	& & artic￭ & 0.633\\
+	\hline
+	\multirow{5}{*}{ECB} & \multirow{5}{*}{joint}
+	& Joint & 0.423 \\
+	& & jointly & 0.467 \\
+	& & common & 0.601 \\
+	& & together & 0.632 \\
+	& & artic￭ & 0.663 \\
+	& & coordinated & 0.668 \\
+\end{tabular}
+\caption{Corpora}
+\label{tab:Polysemy}
+\end{table}
 

### 2019-05-15 2019/emnlp/sparse.tex
new references

@@
 The validation sets are used to chose the best model according to the average BLEU score.
-Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other, related, domains: the official Khresmoi-test \fyTodo{REF needed}, which is close to EMEA, and News test 2014 \fyTodo{REF needed}.
+Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other related domains: the official Khresmoi-test \cite{Khresmoi17test}, which is close to EMEA, and News test 2014 \cite{Bojar14findings}.
 This enables us to evaluate the loss in performance when the test set is from a different domain.

### 2019-05-16 2019/emnlp/sparse.tex
add / remove comments

--- 2019-05-15 2019/emnlp/sparse.tex
@@
 
-We experiment with two language pairs (English-French, English-German )and data originating from three domains, corresponding to texts from three European institutions: from the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}.
+We experiment with two language pairs (English-French, English-German) and data originating from three domains, corresponding to texts from three European institutions: from the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}.
 We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).
 The validation sets are used to chose the best model according to the average BLEU score.
-Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other, related, domains: the official Khresmoi-test \fyTodo{REF needed}, which is close to EMEA, and News test 2014 \fyTodo{REF needed}.
+Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other related domains: the official Khresmoi-test \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, IWSLT 2010 (Ted Talks) \cite{}.
 This enables us to evaluate the loss in performance when the test set is from a different domain.
 % The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
-\fyTodo{Check which corpus are useful}
+\fyDone{Check which corpus are useful}
 \begin{table}
@@
 
-To reduce the vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations, then select approximately 30,000 the most frequent words to build vocabulary for each language.\fyTodo{I need explanations here}
+To reduce the vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations separately for each language. \fyDone{I need explanations here}
 
@@
 To train NMT systems, we use Adam($\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$) for RNNs; and Adam with $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} ($warmup\_steps=4000$) for Transformer.
-Our implementation is from OpenNMT-tf\cite{Klein2017OpenNMT}, except for WDCMT where we use the authors implementation\footnote{\url{https://github.com/DeepLearnXMU/WDCNMT}.}.
+Our implementation is from OpenNMT-tf\cite{Klein2017OpenNMT}, except for WDCMT where we use the authors implementation.\footnote{\url{https://github.com/DeepLearnXMU/WDCNMT}.}
 
-\subsection{Sparse NMT model configuration}
+\subsection{Sparse NMT model configuration}\fyTodo{Change sparse.}
 \fyTodo{Motivate the split - discuss experimentally embedding size}
@@
 \subsection{Results}
+Discussion 1:
+\begin{itemize}
+\item transformer better than RNN
+\item Fine tuning is the best strategy - but subject to catastrophic
+\item Sparse better than joint - better than domain control
+\item Across th board / language \& architecture - always good for EMEA , less good for EPPS, varying for ECB
+\item better ? than domain control ? WMCD ?
+\end{itemize}
+
+Discussion 2:
+\begin{itemize}
+\item Sparse 1 > Sparse 2
+\item out of domain better than Mixed
+\item Sparse 3 ??
+\item ajouter Sparse 4
+\end{itemize}
+
 \begin{table*}
@@
  \hline
- Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
+ Models & EMEA & EPPS & ECB & Khresmoi & test 2007 (EPPS domain) & newstest 2009 & Newstest 2014 & IWSLT 2010 \\ [0.5ex] 
  \hline\hline
@@
 Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT.
-Sparse(2) uses generic tag for all of test sets.
-Sparse(3) uses tag predicted by a pre-trained domain classififer}
+Sparse (2) uses generic tag for all of test sets.
+Sparse (3) uses tag predicted by a pre-trained domain classififer}
 \label{tab:1}
@@
 \subsection{Region's size \label{secc:region_size}}
+
 We also test different setting of region size in word embedding.
@@
 We test those settings with the same condition as above.
-The scores is showed in table \ref{tab:5}
+The scores are in Table~\ref{tab:embedding-size}.\fyTodo{Comment, homegenize captions.}
+
+Results are very close so we pick 8.\fyTodo{Experiment with 64}.
+
 \begin{table*}
@@
     &  & jointly & 0.545 \\
-    &  & artic￭ & 0.605 \\
+    &  & artic- & 0.605 \\
     &  & joints & 0.607 \\
@@
 	& & jointly & 0.497 \\
-	& & join￭ & 0.606 \\
+	& & join- & 0.606 \\
 	& & common & 0.607 \\
 	& & collective & 0.626 \\
-	& & artic￭ & 0.633\\
+	& & artic- & 0.633\\
 	\hline
@@
 	& & together & 0.632 \\
-	& & artic￭ & 0.663 \\
+	& & artic- & 0.663 \\
 	& & coordinated & 0.668 \\

### 2019-05-16 2019/emnlp/sparse.tex
new scores

@@
 \title{Sparse Word Embeddings for Multi-domain Machine Translation}
-\title{Lexicalized Domain Embeddings for Multi-domain Machine Translation}
+\title{Lexicalized Domain Representation for Neural Machine Translation}
 
@@
 \begin{center}
- \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
+ \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X | X ||} 
  \hline
- Models & EMEA & EPPS & ECB & Khresmoi & test 2007 (EPPS domain) & newstest 2009 & Newstest 2014 & IWSLT 2010 \\ [0.5ex] 
+ Models & EMEA & EMEA non-included & EPPS & ECB & Khresmoi & test 2007 (EPPS domain) & newstest 2009 & Newstest 2014 & IWSLT 2010 \\ [0.5ex] 
  \hline\hline
- Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 & 28.56 & 25.7 \\
+ Mixed & 67.69 & 47.6 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 & 28.56 & 25.7 \\
  \hline
- EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.64 & 11.1 \\
+ EMEA-finetuned & 76.77 & 49.43 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.64 & 11.1 \\
  \hline
- EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 & 28.33 & 11.1 \\
+ EPPS-finetuned & 20.86 & & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 & 28.33 & 11.1 \\
  \hline
- ECB-finetuned & 26.93 & 27.09 & 56.52 & 22.04 & 22.92 & 12.94 & 17.51 & 13.99 \\
+ ECB-finetuned & 26.93 & & 27.09 & 56.52 & 22.04 & 22.92 & 12.94 & 17.51 & 13.99 \\
  \hline
- Sparse (1) & 74.26 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.7 \\
+ Sparse (1) & 74.26 & 49.9 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.7 \\
  \hline
- Sparse (2) & 73.97 & 37.81 & 53.67 & 34.17 & 30.93 & 20.97 & 28.11 & 25.7 \\
+ Sparse (2) & 73.97 & 49.54 & 37.81 & 53.67 & 34.17 & 30.93 & 20.97 & 28.11 & 25.7 \\
  \hline
- Sparse (3) &  &  &  &  &  &  &  & \\
+ Sparse (3) &  &  &  & &  &  &  &  & \\
  \hline
- Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 &  20.1 & 26.52 & 24.81\\
+ Sparse (4) & 72.95 & 49.78 & 37.62 & 53.35 &  &  &  &  & \\
+ \hline
+ Domain Control & 67.87 & 45.42 & 37.31 & 54.14 & 33.63 & 31.09 &  20.1 & 26.52 & 24.81\\
  \hline
@@
 \begin{center}
- \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
+ \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X | X ||} 
  \hline
- Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
+ Models & EMEA test & EMEA non-included & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
  \hline\hline
- Mixed & 64.57 & 26.47 & 68.67 & 21.77 & 23.50 & 13.9 & 17.04 & 18.85 \\
+ Mixed & 64.57 & 42.99 & 26.47 & 68.67 & 21.77 & 23.50 & 13.9 & 17.04 & 18.85 \\
  \hline
- EMEA-finetuned & 68.35 & 17.02 & 32.87 & 21.15 & 15.94 & 8.92 & 11.21 & 13.49 \\
+ EMEA-finetuned & 68.35 & & 17.02 & 32.87 & 21.15 & 15.94 & 8.92 & 11.21 & 13.49 \\
  \hline
- EPPS-finetuned & 36.19 & 26.29 & 40.71 & 19.67 & 23.69 & 14.08 & 16.47 & 19.24 \\
+ EPPS-finetuned & 36.19 & & 26.29 & 40.71 & 19.67 & 23.69 & 14.08 & 16.47 & 19.24 \\
  \hline
- ECB-finetuned & 24.72 & 18.36 & 74.05 & 11.31 & 16.64 & 9.11 & 10.8 & 11.0 \\
+ ECB-finetuned & 24.72 & & 18.36 & 74.05 & 11.31 & 16.64 & 9.11 & 10.8 & 11.0 \\
  \hline
- Sparse(1) & 70.9 & 26.30 & 68.9 & 21.42 & 23.73 & 13.99 & 16.57 & 18.82 \\
+ Sparse(1) & 70.9 & 46.12 & 26.30 & 68.9 & 21.42 & 23.73 & 13.99 & 16.57 & 18.82 \\
  \hline
- Sparse(2) & 70.86 & 26.14 & 68.59 & 21.95 & 23.79 & 13.99 & 16.57 & 18.82 \\
+ Sparse(2) & 70.86 & 45.6 & 26.14 & 68.59 & 21.95 & 23.79 & 13.99 & 16.57 & 18.82 \\
  \hline
- Sparse (3) &  &  &  &  &  &  &  & \\
+ Sparse (3) &  &  &  &  &  &  &  &  & \\
  \hline
- Domain Control & 63.48 & 26.27 & 66.95 & 21.45 & 23.95 & 13.94 & 16.45 & 18.78  \\
+ Sparse (4) & 69.51 & 43.5 & 26.31 & 66.86 &  &  &  &  & \\
+ \hline
+ Domain Control & 63.48 & 42.98 & 26.27 & 66.95 & 21.45 & 23.95 & 13.94 & 16.45 & 18.78  \\
  \hline
@@
 \begin{center}
- \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
+ \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X | X ||} 
  \hline
- Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
+ Models & EMEA test & EMEA non-included & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
  \hline\hline
- Mixed & 65.42 & 34.7 & 51.38 & 29.27 & 28.95 & 17.82 & 22.21 & 21.63 \\
+ Mixed & 65.42 & 45.11 & 34.7 & 51.38 & 29.27 & 28.95 & 17.82 & 22.21 & 21.63 \\
  \hline
- EMEA-finetuned  & 72.06 & 18.62 & 16.78 & 25.88 & 13.38 & 9.26 & 10.94 & 5.73 \\
+ EMEA-finetuned  & 72.06 & & 18.62 & 16.78 & 25.88 & 13.38 & 9.26 & 10.94 & 5.73 \\
  \hline
- EPPS-finetuned & 35.47 & 34.61 & 39.56 & 27.11 & 28.8 & 17.3 & 21.85 & 20.61 \\
+ EPPS-finetuned & 35.47 & & 34.61 & 39.56 & 27.11 & 28.8 & 17.3 & 21.85 & 20.61 \\
  \hline
- ECB-finetuned & 21.93 & 22.6 & 51.53 & 14.74 & 20.23 & 9.74 & 12.10 & 10.47 \\
+ ECB-finetuned & 21.93 & & 22.6 & 51.53 & 14.74 & 20.23 & 9.74 & 12.10 & 10.47 \\
  \hline
- Sparse (1) & 71.73 & 35.21 & 50.91 & 29.62 & 29.05 & 17.92 & 23.00 & 20.96 \\
+ Sparse (1) & 71.73 & 46.3 & 35.21 & 50.91 & 29.62 & 29.05 & 17.92 & 23.00 & 20.96 \\
  \hline
- Sparse (2) & 70.59 & 35.22 & 49.68 & 29.51 & 29.32 & 17.92 & 23.00 & 20.96 \\
+ Sparse (2) & 70.59 & 46.37 & 35.22 & 49.68 & 29.51 & 29.32 & 17.92 & 23.00 & 20.96 \\
  \hline
- Sparse (3) &  &  &  &  &  &  &  & \\
+ Sparse (3) &  &  &  &  & &  &  &  & \\
  \hline
- Domain Control & 68.26 & 35.13 & 50.09 & 27.86 & 29.34 & 16.25 & 20.95 & 21.16 \\
+ Sparse (4) & 62.1 & 43.29 & 34.17 & 48.79 &  &  &  &  & \\
  \hline
- WDCMT (conditional GRU) & 68.76 & 35.71 & 52.75 & 32.03 & 29.81 & 19.58 & 24.63 & 23.28 \\
+ Domain Control & 68.26 & 43.76 & 35.13 & 50.09 & 27.86 & 29.34 & 16.25 & 20.95 & 21.16 \\
+ \hline
+ WDCMT (conditional GRU) & 68.76 & 45.29 & 35.71 & 52.75 & 32.03 & 29.81 & 19.58 & 24.63 & 23.28 \\
  \hline 
@@
 \begin{center}
- \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
+ \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X | X ||} 
  \hline
- Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
+ Models & EMEA test & EMEA non-included & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
  \hline\hline
- Mixed & 57.37 & 23.1 & 63.54 & 17.39 & 20.65 & 10.69 & 11.51 & 14.29 \\
+ Mixed & 57.37 & 37.94 & 23.1 & 63.54 & 17.39 & 20.65 & 10.69 & 11.51 & 14.29 \\
  \hline
- EMEA-finetuned & 65.64 & 12.36 & 15.93 & 15.38 & 11.1 & 5.54 & 6.22 & 7.6 \\
+ EMEA-finetuned & 65.64 & & 12.36 & 15.93 & 15.38 & 11.1 & 5.54 & 6.22 & 7.6 \\
  \hline
- EPPS-finetuned & 24.9 & 22.98 & 26.26 & 13.55 & 20.95 & 10.58 & 11.33 & 14.32 \\
+ EPPS-finetuned & 24.9 & & 22.98 & 26.26 & 13.55 & 20.95 & 10.58 & 11.33 & 14.32 \\
  \hline
- ECB-finetuned & 41.8 & 15.97 & 71.07 & 10.77 & 14.67 & 7.06 & 7.66 & 8.07 \\
+ ECB-finetuned & 41.8 & & 15.97 & 71.07 & 10.77 & 14.67 & 7.06 & 7.66 & 8.07 \\
  \hline
- Sparse (1) & 63.43 & 22.66 & 64.4 & 16.99 & 20.63 & 10.89 & 11.45 & 14.26 \\
+ Sparse (1) & 63.43 & 40.04 & 22.66 & 64.4 & 16.99 & 20.63 & 10.89 & 11.45 & 14.26 \\
  \hline
- Sparse (2) & 63.27 & 22.62 & 63.7 & 17.26 & 20.69 & 10.89 & 11.45 & 14.26 \\
+ Sparse (2) & 63.27 & 39.75 & 22.62 & 63.7 & 17.26 & 20.69 & 10.89 & 11.45 & 14.26 \\
  \hline
- Sparse (3) &  &  &  &  &  &  &  & \\
+ Sparse (3) &  &  &  &  & &  &  &  & \\
  \hline
- Domain Control & 62.53 & 23.74 & 65.71 & 16.88 & 21.85 & 9.24 & 11.15 & 14.97 \\
+ Sparse (4) & 56.84 & 37.05 & 22.06 & 61.66 &  &  &  &  & \\
  \hline
- WDCMT (conditional GRU) & 64.05 & 22.82 & 67 & 15.94 & 20.57 & 10.15 & 11.45 & 13.66 \\
+ Domain Control & 62.53 & 39.25 & 23.74 & 65.71 & 16.88 & 21.85 & 9.24 & 11.15 & 14.97 \\
+ \hline
+ WDCMT (conditional GRU) & 64.05 & 38.54 & 22.82 & 67 & 15.94 & 20.57 & 10.15 & 11.45 & 13.66 \\
  \hline 

### 2019-05-16 2019/emnlp/sparse.tex
title
small error
revised section 3

@@
 
-\title{Sparse Word Embeddings for Multi-domain Machine Translation}
 \title{Lexicalized Domain Representation for Neural Machine Translation}
@@
 The validation sets are used to chose the best model according to the average BLEU score.
-Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other related domains: the official Khresmoi-test \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, IWSLT 2010 (Ted Talks) \cite{}.
-This enables us to evaluate the loss in performance when the test set is from a different domain.
+Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
+This enables us to evaluate the loss in performance when the test set is from a domain not seen in training.
 % The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
@@
 \end{tabular}
-\caption{Corpora}
+\caption{Train and test corpora}
 \label{tab:Corpora}
 \end{table}
-
-To reduce the vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations separately for each language. \fyDone{I need explanations here}
+To reduce the number of lexical units and make our systems open-vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16BPE} separately for each language with 30,000 merge operations.
+\fyDone{I need explanations here}
 
@@
 To validate our findings, we compared sparse embedding models with standard models using both attentional Recurrent Neural Networks (RNNs) \cite{Bahdanau15learning} and the recent Transformer architecture of \citet{Vaswani17attention}.
-Our baselines consist of generic models trained with a simple concatenation of all corpora (denoted \texttt{mixed} in the score tables below); models finetuned separately to each domain using (10000, 15000, 5000) iterations over in-domain data (denoted $+ft_{EMEA}$, $+ft_{EPPS}$, $+ft_{ECB}$ below); models using domain tags as in \cite{Kobus17domaincontrol} (noted \texttt{DC} below) and as well as the multi-domain model of \citet{Zeng18multidomain} (denoted \texttt{WDCMT} in the tables below).
+Our baselines consist of generic models trained with a simple concatenation of all corpora (denoted \texttt{mixed} in the score tables below); models finetuned separately to each domain using (10000, 15000, 5000) iterations over in-domain data (denoted \texttt{+ft(EMEA)}, \texttt{+ft(EPPS)}, \texttt{+ft(ECB)} below); models using domain tags as in \cite{Kobus17domaincontrol} (noted \texttt{DC} below); and finally the multi-domain model of \citet{Zeng18multidomain} (denoted \texttt{WDCMT} below).\fyTodo{Change names in tables}
 
 For all models, we set the size of word embeddings equal to 512; the size of hidden layers is equal to 1024 for RNNs and 512 for Transformer.
-\fyTodo{Improve this}
+Other important configuration details that were kept fixed throughout are as follows:
+\fyDone{Improve this}
 \begin{itemize}
-\item Transformer: size of word embedding: 512; size of hidden layers: 512; size of inner feed forward layer: 2048; Multi-head: 8; number of layers: 6.
-\item Attention Recurrent Neural Network: Bidirection LSTM encoder, unidirectional LSTM decoder, Luong Attention Mechanism, size of word embedding: 512; size of hidden layers: 1024; number of layers: 1
-\item DC: size of domain embedding:2 ; same settings for both architecture: Transformer and Attention Recurrent Neural Network.
-\item WDCMT: Bidirection Gated recurrent units (denoted GRU) encoder, unidirectional conditional GRU decoder, size of word embedding: 512; size of hidden layers: 1024; size of domain layers: 300; number of layers: 1
+\item Our Transformer models use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 components; 
+%  size of word embedding: 512; size of hidden layers: 512; size of inner feed forward layer: 2048; Multi-head: 8; number of layers: 6.
+\item Our RNN models use 1 layer on both sides: a bidirectional LSTM encoder and a unidirectional LST decoder with attention;
+%  Attention Recurrent Neural Network: Bidirection LSTM encoder, unidirectional LSTM decoder, Luong Attention Mechanism, size of word embedding: 512; size of hidden layers: 1024; number of layers: 1
+\item The domain control systems are exactly as their baseline counterpart (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.\fyTodo{Check This!}
+  % C: size of domain embedding:2 ; same settings for both architecture: Transformer and Attention Recurrent Neural Network.
+\item The \texttt{WDCMT} system also has one layer on the encoder side (a bidirectional Gated recurrent units (GRU)), as well as one layer on the decoder side (a unidirectional conditional GRU); the dimension of ``domain'' layers is 300.
 \end{itemize}
-To train NMT systems, we use Adam($\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$) for RNNs; and Adam with $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} ($warmup\_steps=4000$) for Transformer.
-Our implementation is from OpenNMT-tf\cite{Klein2017OpenNMT}, except for WDCMT where we use the authors implementation.\footnote{\url{https://github.com/DeepLearnXMU/WDCNMT}.}
+To train NMT systems, we use Adam, with parameters $\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$ for RNNs; with parameters $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} ($warmup\_steps=4000$) for Transformer.
+Our implementation is from OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein2017OpenNMT}, except for WDCMT where we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyTodo{batch size}
 
@@
 \fyTodo{Motivate the split - discuss experimentally embedding size}
-For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions: 3 domain specific and 1 generic, with size $[8,8,8,488]$, the choice will be dicussed in \ref{secc:region_size}.
-If sentence comes from domain $i$, regions for domains $j \neq i$ will be masked as 0 while generic region is always activated.
-We then use a dense layer of size 512 to fuse the active domain region and the generic region.
+For the implementation of lexicalized domain embeddings, we split the embedding vector (of size 512) into four regions: 3 are domain specific and 1 is generic, with respective sizes $[8,8,8,488]$, a choice that will be dicussed in Section~\ref{secc:region_size}.
+If a sentence originates from domain $i$, the domain specific regions for all domains $j \neq i$ will be zeroed out; while the other regions are activated.
+We then use a dense layer of size~512 to fuse the region for the active domain and the ``generic'' region.
+Training is formalized in algorithm~\ref{alg:multidomain}.\fyTodo{Notation B for batch size}
 
@@
 \caption{Multi-domain Training}
-\label{algo:1}
+\label{alg:multidomain}
 \begin{algorithmic}[1]
-\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for d domains, Optimization algorithm Opt}
+\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for $d$ domains, Optimization algorithm Opt}
 \REPEAT 
 \STATE{Randomly pick $i \in [1,..,d]$ w.r.t multinomial distribution $[\frac{|C_i|}{\displaystyle{\mathop{\sum}_{i\in [1,..,d]}|C_i|}}]$.}
-\STATE{Randomly pick a batch of sentences from corpora $C_i$.}
-\STATE{Activating only generic region to create generic embedding batch, denoted $W_g$}
+\STATE{Randomly pick $B$ sentences from $C_i$.}
+\STATE{Activate only generic region to create generic embedding batch, denoted $W_g$.}
 \STATE{Pass the generic embedding batch to a dense layer.}
-\STATE{Computing gradient of $\theta_s$, $\frac{\partial L}{\partial \theta_s}$ using generic embedding batch $W_g$}
-\STATE{Activating corresponding domain region and generic region to create domain embedding batch}
-\STATE{Pass the batch of domain embedding to a dense layer}
-\STATE{Computing gradient of domain parameters $\theta_i$, $\frac{\partial L}{\partial \theta_i}$ using domain batch}
+\STATE{Compute gradient of $\theta_s$, $\frac{\partial L}{\partial \theta_s}$ using $W_g$.}
+\STATE{Activate domain region and generic region to create domain embedding batch $W_i$}
+\STATE{Pass $W_i$ to a dense layer}
+\STATE{Compute gradient of domain-specific parameters $\theta_i$, $\frac{\partial L}{\partial \theta_i}$ using $W_d$.}
 \STATE{Apply learning rules using computed gradients: 
 $$ \theta_s = \theta_s - Opt(\frac{\partial L}{\partial \theta_s}(W_g))$$ 
-$$ \theta_i = \theta_i - Opt(\frac{\partial L}{\partial \theta_i}(W_d))$$}
+$$ \theta_i = \theta_i - Opt(\frac{\partial L}{\partial \theta_i}(W_i))$$}
 \UNTIL{convergence}
@@
 
-In the training, we use same optimization algorithm as the baselines.
-Each iteration of algorithm \ref{algo:1}, we use 2 batches: generic batch using only generic region; and domain batch using generic region and domain region.
-The reason to do that is to minimize simultaneously both losses: \\
-$\theta^*_{s}=\displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))]$ \\ 
-$ \theta^*_{i}=\displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
-$
+In training, we use same optimization algorithm as for the baselines.
+Note that each iteration of algorithm \ref{alg:multidomain} uses 2 batches: generic batch using only generic region; and domain batch using generic region and domain region.
+The reason to do that is to minimize simultaneously two losses:
+\begin{align*}
+\theta^*_{s}=&\displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))] \\ 
+ \theta^*_{i}=&\displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
+\end{align*}
 
@@
  \hline
- Models & EMEA test & EMEA excluded & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
+ Models & EMEA test & EMEA non-included & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
  \hline\hline 

### 2019-05-17 2019/emnlp/sparse.tex
title
transformer table

@@
 
-\title{Lexicalized Domain Representation for Neural Machine Translation}
+\title{Lexicalized Domain Representations for Neural Machine Translation}
 
@@
 
-\begin{table*}
+\begin{table*}[!h]
 \begin{center}
- \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X | X ||} 
+\scalebox{0.81}{
+\begin{tabular}{|l|ccc|ccc||c|}
 \hline
- Models & EMEA & EMEA non-included & EPPS & ECB & Khresmoi & test 2007 (EPPS domain) & newstest 2009 & Newstest 2014 & IWSLT 2010 \\ [0.5ex] 
- \hline\hline
- Mixed & 67.69 & 47.6 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 & 28.56 & 25.7 \\
+Model & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & Avg. \\
 \hline
- EMEA-finetuned & 76.77 & 49.43 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.64 & 11.1 \\
+\multicolumn{8}{l}{English$\rightarrow$French} \\
 \hline
- EPPS-finetuned & 20.86 & & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 & 28.33 & 11.1 \\
+$Mixed$ & 67.69 $_{47.60}$ & 37.50 & 53.49 & 34.86 & 28.56 & 25.70 & avg \\
 \hline
- ECB-finetuned & 26.93 & & 27.09 & 56.52 & 22.04 & 22.92 & 12.94 & 17.51 & 13.99 \\
+$FT_{EMEA}$ & 76.77 $_{49.43}$ & 17.16 & 11.99 & 29.58 & 11.64 & 11.10 & avg \\
+$FT_{EPPS}$ & 20.86 $$ & 37.04 & 24.53 & 31.86 & 28.33 & 11.10 & avg \\
+$FT_{ECB}$   & 26.93 & 27.09 & 56.52 & 22.04 & 17.51 & 13.99 & avg \\
 \hline
- Sparse (1) & 74.26 & 49.9 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.7 \\
+$DC$ & 67.87 $_{45.42}$ & 37.31 & 54.14 & 33.63 & 26.52 & 24.81 & avg\\
 \hline
- Sparse (2) & 73.97 & 49.54 & 37.81 & 53.67 & 34.17 & 30.93 & 20.97 & 28.11 & 25.7 \\
+$LDR_{oracle}$   & 74.26 $_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 & avg \\
+$LDR_{generic}$ & 73.97 $_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & avg \\
+$LDR_{rnn}$ & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg \\
+$LDR_{wrong}$ & 72.95 $_{49.78}$ & 37.62 & 53.35 & khreshmoi & newstest & iwslt & avg\\
 \hline
- Sparse (3) &  &  &  & &  &  &  &  & \\
+\multicolumn{8}{l}{English$\rightarrow$German} \\
 \hline
- Sparse (4) & 72.95 & 49.78 & 37.62 & 53.35 &  &  &  &  & \\
+$Mixed$ & 64.57 $_{42.99}$ & 26.47 & 68.67 & 21.77 & 17.04 & 18.85 & avg \\
 \hline
- Domain Control & 67.87 & 45.42 & 37.31 & 54.14 & 33.63 & 31.09 &  20.1 & 26.52 & 24.81\\
+$FT_{EMEA}$ & 68.35 $$ & 17.02 & 32.87 & 21.15 & 11.21 & 13.49 & avg\\
+$FT_{EPPS}$ & 36.19 & 26.29 & 40.71 & 19.67 & 16.47 & 19.24 & avg \\
+$FT_{ECB}$   & 24.72 & 18.36 & 74.05 & 11.31  & 10.8 & 11.00 & avg \\
 \hline
-\end{tabularx}
+$DC$ & 63.48 $_{42.98}$ & 26.27 & 66.95 & 21.45 & 16.45 & 18.78 & avg \\
+\hline
+$LDR_{oracle}$ & 70.9 $_{46.12}$ & 26.30 & 68.9 & 21.42 & 16.57 & 18.82 & avg\\
+$LDR_{generic}$ & 70.86 $_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & avg\\
+$LDR_{rnn}$ & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg \\
+$LDR_{wrong}$ & 69.51 $_{43.50}$ & 26.31 & 66.86 & khreshmoi & newstest & iwslt & avg \\
+\hline
+\end{tabular}
+} %scalebox
 \end{center}
-
-\caption{BPE-detokenized BLEU score for English-French with the Transformer architecture.
-Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT.
-Sparse (2) uses generic tag for all of test sets.
-Sparse (3) uses tag predicted by a pre-trained domain classififer}
-\label{tab:1}
+\caption{BPE-detokenized BLEU scores for the Transformer architecture over test sets.}
 \end{table*}

### 2019-05-17 2019/emnlp/sparse.tex
transformer table
rnn table

@@
 \begin{center}
-\scalebox{0.81}{
+\scalebox{1.0}{
 \begin{tabular}{|l|ccc|ccc||c|}
 \hline
-Model & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & Avg. \\
+Model & EMEA & EPPS & ECB & khreshmoi & newstest2014 & iwslt2010 & Avg. \\
 \hline
@@
 
-\begin{table*}
+\begin{table*}[!h]
 \begin{center}
- \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X | X ||} 
+\scalebox{1.0}{
+\begin{tabular}{|l|ccc|ccc||c|}
 \hline
- Models & EMEA test & EMEA non-included & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
- \hline\hline
- Mixed & 64.57 & 42.99 & 26.47 & 68.67 & 21.77 & 23.50 & 13.9 & 17.04 & 18.85 \\
+Model & EMEA & EPPS & ECB & khreshmoi & newstest2014 & iwslt2010 & Avg. \\
 \hline
- EMEA-finetuned & 68.35 & & 17.02 & 32.87 & 21.15 & 15.94 & 8.92 & 11.21 & 13.49 \\
+\multicolumn{8}{l}{English$\rightarrow$French} \\
 \hline
- EPPS-finetuned & 36.19 & & 26.29 & 40.71 & 19.67 & 23.69 & 14.08 & 16.47 & 19.24 \\
+$Mixed$               & 65.42 $_{45.11}$ & 34.7 & 51.38 & 29.27 & 22.21 & 21.63 & avg \\
 \hline
- ECB-finetuned & 24.72 & & 18.36 & 74.05 & 11.31 & 16.64 & 9.11 & 10.8 & 11.0 \\
+$FT_{EMEA}$     & 72.06 $$ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & avg \\
+$FT_{EPPS}$      & 35.47 $$ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & avg \\
+$FT_{ECB}$        & 21.93 $$ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & avg \\
 \hline
- Sparse(1) & 70.9 & 46.12 & 26.30 & 68.9 & 21.42 & 23.73 & 13.99 & 16.57 & 18.82 \\
+$DC$                   & 68.26 $_{43.76}$ & 35.13 & 50.09 & 27.86 & 20.95 & 21.16 & avg \\
  \hline
- Sparse(2) & 70.86 & 45.6 & 26.14 & 68.59 & 21.95 & 23.79 & 13.99 & 16.57 & 18.82 \\
+$WDCMT$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & avg \\
 \hline
- Sparse (3) &  &  &  &  &  &  &  &  & \\
+$LDR_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & avg\\
+$LDR_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & avg\\
+$LDR_{rnn}$        & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg\\
+$LDR_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 & khreshmoi & newstest & iwslt & avg \\
 \hline
- Sparse (4) & 69.51 & 43.5 & 26.31 & 66.86 &  &  &  &  & \\
+\multicolumn{8}{l}{English$\rightarrow$German} \\
 \hline
- Domain Control & 63.48 & 42.98 & 26.27 & 66.95 & 21.45 & 23.95 & 13.94 & 16.45 & 18.78  \\
+$Mixed$               & 57.37 $_{37.94}$ & 23.1 & 63.54 & 17.39 & 11.51 & 14.29 & avg\\
 \hline
-\end{tabularx}
+$FT_{EMEA}$      & 65.64 $$ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & avg\\
+$FT_{EPPS}$       & 24.90 $$ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & avg \\
+$FT_{ECB}$         & 41.8 $$ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & avg\\
+\hline
+$DC$                    & 62.53 $_{39.25}$ & 23.74 & 65.71 & 16.88 & 11.15 & 14.97 & avg\\
+\hline
+WDCMT               & 64.05 $_{38.54}$ & 22.82 & 67.00 & 15.94 & 11.45 & 13.66 & avg\\
+\hline
+$LDR_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & avg\\
+$LDR_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & avg\\
+$LDR_{rnn}$        & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg\\
+$LDR_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & khreshmoi & newstest & iwslt & avg \\
+\hline
+\end{tabular}
+} %scalebox
 \end{center}
-\caption{BLEU score for Transformer in language pair English - German. Notations are same as Table~\ref{tab:1}}
-\label{tab:2}
-\end{table*}
-
-\begin{table*}
-\begin{center}
- \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X | X ||} 
- \hline
- Models & EMEA test & EMEA non-included & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
- \hline\hline
- Mixed & 65.42 & 45.11 & 34.7 & 51.38 & 29.27 & 28.95 & 17.82 & 22.21 & 21.63 \\
- \hline
- EMEA-finetuned  & 72.06 & & 18.62 & 16.78 & 25.88 & 13.38 & 9.26 & 10.94 & 5.73 \\
- \hline
- EPPS-finetuned & 35.47 & & 34.61 & 39.56 & 27.11 & 28.8 & 17.3 & 21.85 & 20.61 \\
- \hline
- ECB-finetuned & 21.93 & & 22.6 & 51.53 & 14.74 & 20.23 & 9.74 & 12.10 & 10.47 \\
- \hline
- Sparse (1) & 71.73 & 46.3 & 35.21 & 50.91 & 29.62 & 29.05 & 17.92 & 23.00 & 20.96 \\
- \hline
- Sparse (2) & 70.59 & 46.37 & 35.22 & 49.68 & 29.51 & 29.32 & 17.92 & 23.00 & 20.96 \\
- \hline
- Sparse (3) &  &  &  &  & &  &  &  & \\
- \hline
- Sparse (4) & 62.1 & 43.29 & 34.17 & 48.79 &  &  &  &  & \\
- \hline
- Domain Control & 68.26 & 43.76 & 35.13 & 50.09 & 27.86 & 29.34 & 16.25 & 20.95 & 21.16 \\
- \hline
- WDCMT (conditional GRU) & 68.76 & 45.29 & 35.71 & 52.75 & 32.03 & 29.81 & 19.58 & 24.63 & 23.28 \\
- \hline 
-\end{tabularx}
-\end{center}
-\caption{bpe-detokenized BLEU score for Recurrent network architecture on language pair English-French.
-Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{Zeng18multidomain} use Bidirectional-LSTM}
-\label{tab:3}
-\end{table*}
-
-\begin{table*}
-\begin{center}
- \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X | X ||} 
- \hline
- Models & EMEA test & EMEA non-included & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
- \hline\hline
- Mixed & 57.37 & 37.94 & 23.1 & 63.54 & 17.39 & 20.65 & 10.69 & 11.51 & 14.29 \\
- \hline
- EMEA-finetuned & 65.64 & & 12.36 & 15.93 & 15.38 & 11.1 & 5.54 & 6.22 & 7.6 \\
- \hline
- EPPS-finetuned & 24.9 & & 22.98 & 26.26 & 13.55 & 20.95 & 10.58 & 11.33 & 14.32 \\
- \hline
- ECB-finetuned & 41.8 & & 15.97 & 71.07 & 10.77 & 14.67 & 7.06 & 7.66 & 8.07 \\
- \hline
- Sparse (1) & 63.43 & 40.04 & 22.66 & 64.4 & 16.99 & 20.63 & 10.89 & 11.45 & 14.26 \\
- \hline
- Sparse (2) & 63.27 & 39.75 & 22.62 & 63.7 & 17.26 & 20.69 & 10.89 & 11.45 & 14.26 \\
- \hline
- Sparse (3) &  &  &  &  & &  &  &  & \\
- \hline
- Sparse (4) & 56.84 & 37.05 & 22.06 & 61.66 &  &  &  &  & \\
- \hline
- Domain Control & 62.53 & 39.25 & 23.74 & 65.71 & 16.88 & 21.85 & 9.24 & 11.15 & 14.97 \\
- \hline
- WDCMT (conditional GRU) & 64.05 & 38.54 & 22.82 & 67 & 15.94 & 20.57 & 10.15 & 11.45 & 13.66 \\
- \hline 
-\end{tabularx}
-\end{center}
-\caption{BLEU score for Recurrent network on language pair English - German. The notation is same as figure \ref{tab:3}}
-\label{tab:4}
+\caption{BPE-detokenized BLEU scores for the RNN architecture over test sets.}
 \end{table*}

### 2019-05-17 2019/emnlp/sparse.tex
rnn table

@@
 
-\begin{table*}
+\begin{table*}[!h]
 \begin{center}
- \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X | X ||} 
+\scalebox{1.0}{
+\begin{tabular}{|l|ccc|ccc|}
 \hline
- Models & EMEA test & EMEA non-included & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
- \hline\hline 
- Sparse - 4 & 74.65 & 49.61 & 37.42 & 54.49 & 33.68 & 31.01 & 20.71 & 28.03 & 25.28 \\
+$LDR$ & EMEA & EPPS & ECB & khreshmoi & newstest2014 & iwslt2010 \\
 \hline
- Sparse - 8 & 74.26 & 49.9 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.7 \\
+\multicolumn{7}{l}{English$\rightarrow$French} \\
 \hline
- Sparse - 16 & 74.15 & 49.1 & 37.78 & 54.56 & 33.4 & 30.93 & 20.73 & 28.1 & 25.33 \\
+ size=4   & 74.65 $_{49.61}$ & 37.42 & 54.49 & 33.68 & 28.03 & 25.28 \\
+ size=8   & 74.26 $_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 \\
+ size=16 & 74.15 $_{49.10}$ & 37.78 & 54.56 & 33.40 & 28.10 & 25.33 \\
+ size=32 & 75.10 $_{48.61}$ & 37.64 & 54.29 & 33.20 & 27.99 & 25.44 \\
 \hline
- Sparse - 32 & 75.1 & 48.61 & 37.64 & 54.29 & 33.2 & 30.94 & 21.07 & 27.99 & 25.44 \\
- \hline
-\end{tabularx}
+\end{tabular}
+} %scalebox
 \end{center}
-\caption{BPE-detokenized BLEU score for English-French with the Transformer architecture}
-\label{tab:5}
+\caption{BPE-detokenized BLEU scores for the RNN architecture over test sets.}
 \end{table*}

### 2019-05-18 2019/emnlp/sparse.tex
more bib and related work

@@
 We experiment with two language pairs (English-French, English-German) and data originating from three domains, corresponding to texts from three European institutions: from the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}.
-We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).
-The validation sets are used to chose the best model according to the average BLEU score.
+We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).\footnote{The EMEA dataset as distributed on the OPUS site contains many duplicates.
+We therefore report two numbers: the first is comparable to what has been published on earlier studies (eg.\ \fyTodo{Missing ref}), the second one is obtained by making the test set entirely disjoint from the train.} The validation sets are used to chose the best model according to the average BLEU score.
 Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
@@
     \hline
-    EMEA  & 1,09M & 1000 & 1000 \\
+    EMEA  & 1,09M & 1000 & 1000 [300]\\
     ECB    & 0,19M & 1000 & 1000     \\
@@
 Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
-Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{Chu18asurvey}; \cite{Chu2017comparison}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
+Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{;.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
 The methods in this category can be listed as following:
@@
 %% HERE
+Domain adaptation (DA)is a vexing problem in NLP that correspond to a wide range of practical situations and data scenarios (eg.\ supervised vs.
+unsupervised adaptation), and  has been thoroughly  studied from a number of perspectives, ranging from theoretical analysis to more applied work, and for which many solutions have been proposed.
+The literature of DA for Machine Translation reflects this diversity, and is typically organized in data-based approaches vs model-based approaches \cite{Chu2017comparison,Chu18asurvey}.
+The standard scenario is using (mostly) out-of-domain training data while testing on a low-resource domain.
+In this setting, data-based approaches aim to bias the distribution of the train (out-of-domain) data towards matching that of the target domain, using data selection or instance weighting techniques  \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}, or generating adapted pseudo-parallel data through back-translation  \cite{Utiyama2003measure,Wang2014neural,Wang2016connecting}.
+Model-centring approaches ...
 
-The multi-domain scenario has been studied in a number of recent works.
-\citet{Farajian17multidomain} learn a generic system and propose to dynamically adapt the network weights using a small sample of training data that resembles the test data, an idea already explored in \cite{Sennrich13multidomain}.
-Their main contribution is to propose a method to relate the amount of adaptation of the network parameters to the similarity between the adaptation sample and the test sentence: the higher the similarity, the more agressive the adaptation.
+Our scenario is a bit different as we aim to train a unique system that will work well for several domains.
+This corresponds to a practical scenario in the industry, where one would like to maintain one single multi-domain engine, trained on all the available (heteregeneous) sources of data.
+Our primary source of inspiration is the proposal of \citet{Daume07frustratingly} who proposes to use several copies of the same features (one ``generic'' shared across domains, and one for each domain), letting the training adjust their respective weights.
+This work has been reanalyzed in a Bayesian framework in \cite{Finkel09hierarchical}, and revisited notably in \cite{Chang10necessity}.
+The recent proposal of \cite{ZengYang15multidomain}
+
+The multi-domain translation scenario has been studied in a number of recent works.
+\citet{Farajian17multidomain} learn a generic system and propose to dynamically adapt the network weights in an unsupervised manner using a small sample of training data that resembles the test data, an idea already explored in \cite{Sennrich13multidomain}.
+Their main contribution is to propose an a method to relate the amount of adaptation of the network parameters to the similarity between the adaptation sample and the test sentence: the higher the similarity, the more agressive the adaptation.
 

### 2019-05-19 2019/emnlp/sparse.tex
more related works and refs

@@
 One issue which repeatedly pops up with MT, be it neural or statistical, is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain.
-A lot of the recent research effort has thus focused on developping MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Johnson17google}; or even building systems without any parallel data \cite{Artetxe18unsupervised,Lample18unsupervised}.
+A lot of the recent research effort has thus focused on developping MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Ha16towards,Johnson17google}; or even building systems without any parallel data \cite{Artetxe18unsupervised,Lample18unsupervised}.
 Another important scenario for industrial MT systems is to adapt a neural system trained using parallel data in one domain to the peculiarity of other domains.
@@
 
-Beside these works, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.
+Beside these works\fyTdodo{Works= oeuvres d'art}, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.
 
 %% HERE
-Domain adaptation (DA)is a vexing problem in NLP that correspond to a wide range of practical situations and data scenarios (eg.\ supervised vs.
+Domain adaptation (DA) is a vexing problem in NLP, which appears in a wide range of practical situations and data scenarios (eg.\ supervised vs.
 unsupervised adaptation), and  has been thoroughly  studied from a number of perspectives, ranging from theoretical analysis to more applied work, and for which many solutions have been proposed.
-The literature of DA for Machine Translation reflects this diversity, and is typically organized in data-based approaches vs model-based approaches \cite{Chu2017comparison,Chu18asurvey}.
-The standard scenario is using (mostly) out-of-domain training data while testing on a low-resource domain.
-In this setting, data-based approaches aim to bias the distribution of the train (out-of-domain) data towards matching that of the target domain, using data selection or instance weighting techniques  \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}, or generating adapted pseudo-parallel data through back-translation  \cite{Utiyama2003measure,Wang2014neural,Wang2016connecting}.
-Model-centring approaches ...
+The literature of DA for Machine Translation reflects this diversity and typically distinguishes data-based approaches from model-based approaches \cite{Chu2017comparison,Chu18asurvey}.
+
+The most common adaptation scenario is using (mostly) out-of-domain training data while testing on a low-resource domain.
+In this setting, data-based approaches aim to bias the distribution of the train (out-of-domain) data towards matching that of the target domain, using data selection techniques  \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}, or generating adapted pseudo-parallel data through back-translation  \cite{Utiyama2003measure,Wang2014neural,Wang2016connecting,Sennrich2016Mono}.
+Model-centric approaches build domain-adapted models by combining (eg.\ with mixture weights) multiple data sources or multiple systems \cite{Foster07mixture,Wang17instanceweighting}, or by biasing the training objective towards the desired domain using in-domain adaptation data \cite{Luong2015SNMT,Freitag2016FastDA,Chen17costweighting}.
+Another approach worth mentioning integrates domain information (for instance a domain language model) in the decoding algorithm \cite{Gulcehre2016monolingual}.
 
@@
 This work has been reanalyzed in a Bayesian framework in \cite{Finkel09hierarchical}, and revisited notably in \cite{Chang10necessity}.
-The recent proposal of \cite{ZengYang15multidomain}
+Following up on \cite{Yang15multidomain}, the recent proposals of \cite{Peng17multitask} apply the same idea with neural architectures, using domain specific masking to zero out the parameters modeling domains irrelevant to the current input sentence.
+Compared to our work, these techniques are used in deep layers of the network and applied to sequence labelling tasks.
 
-The multi-domain translation scenario has been studied in a number of recent works.
+The multi-domain scenario in NMT has been studied in a number of recent works.
 \citet{Farajian17multidomain} learn a generic system and propose to dynamically adapt the network weights in an unsupervised manner using a small sample of training data that resembles the test data, an idea already explored in \cite{Sennrich13multidomain}.
@@
 
-By analogy with multilingual NMT, \citet{Kobus17domaincontrol} and \citet{Chu18multilingual} separately propose to extend the representation of the source text with a domain tag (or its embedding).
-Our model also modifies input representations, but allows for each source word to have its own domain representation, thereby improving training of the shared parts of the network.\fyTodo{Talk about target embeddings somewhere.}
+By analogy with the proposal of \citet{Johnson17google} for multilingual NMT, \citet{Kobus17domaincontrol} and \citet{Chu18multilingual} separately propose to extend the representation of the source text with a domain tag.
+Our model also modifies input representations, but allows each source word to have its own domain-specific representation, thereby improving training of the shared parts of the network.\fyTodo{Talk about target embeddings somewhere.}
 
-Similarly to our approach, \cite{Zeng18multidomain} tries to separate on the encoder side domain-specific representation from generic representations in two different sub-networks, where shared versus domain-specific representations automatically emerge from two adversarial networks.
+Similarly to our approach, \cite{Zeng18multidomain} attempts to separate on the encoder side domain-specific representation from generic representations in two different sub-networks, where generic versus domain-specific representations automatically emerge from two adversarial networks.
 The decoder side can thus attend separately to these two representations to generate its output.
 In this approach, the shared and domain-specific part kept separated in the deeper layers of the network, whereas we try to localize the differences between domains at the lexical level, based on a much cheaper computational architecture.
+Another trait of this proposal is the ability to automatically infer domain information for test sentences; as we have shown, our architecture can also effectively accomodate sentences lacking domain information.
 
-creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
+% creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
 $h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively.
@@
 
-The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{Platanios18contextual}; \cite{Zeng18multidomain} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
+% The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{Platanios18contextual}; \cite{Zeng18multidomain} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
 $h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively.
 While introducing new features in the architecture, \cite{Zeng18multidomain} also introduces new loss function which is sum of word-level weighted MT loss and losses of Domain-classifier in source side and target side; Adversarial Domain-classifier in source side.
-The authors have very good approach to seperate Domain-shared information flow and Domain-specify information flow and to determine their contributions to the inference thus mitigate the catastrophic interference occuring in training with backpropagation.
-\cite{Platanios18contextual} has other idea to circumvent the catastrophic interference.
+The authors have very good approach to seperate Domain-shared information flow and Domain-specify information flow and to determine their contributions to the inference thus mitigate the catastrophic interference occuring in training with backpropagation
+
+% Platanios is multilingual !!
+% \cite{Platanios18contextual} has other idea to circumvent the catastrophic interference.
 They supposed that parameters of each domain specialized model could be generated from a universal parameter-generator.
-By introducing domain embedding (or Language embedding in the original paper), they could conduct domain related error in backpropagation toward domain embedding thus mitigates catastrophic interference between domains during training.
+By introducing domain embedding (or language embedding in the original paper), they could conduct domain related error in backpropagation toward domain embedding thus mitigates catastrophic interference between domains during training.
 
-Compared to \cite{Platanios18contextual}, our sparse embedding is a bit similar to the idea.
+% Compared to \cite{Platanios18contextual}, our sparse embedding is a bit similar to the idea.
 However, our method gives much more free degree for domain embedding since we give each word a domain region while the author uses only one vector domain embedding for each parameter set(encoder, decoder).
 
-Compared to the work of \cite{Zeng18multidomain}, our approach is much simpler and
-more generic, since sparse embedding can be implemented for every neural architecture while domain-specific gate and domain-shared gate can only be applied for recurrent architectures.
+% Compared to the work of \cite{Zeng18multidomain}, our approach is much simpler and
+% more generic, since sparse embedding can be implemented for every neural architecture while domain-specific gate and domain-shared gate can only be applied for recurrent architectures.
 However, during inference, the model of \cite{Zeng18multidomain} can infer the domain that is the closest to the input sentence and that will be used in special gates; while our method needs to know in which domain the inference is executed.

### 2019-05-20 2019/emnlp/sparse.tex
resize image, more scores

@@
 Our first attempt is to split standard vector word embedding into $D+1$ regions named $\theta_i$ for $i \in [1,..,D]$ and $\theta_s$.
-\begin{figure}[h]
+\begin{figure}
 \center
-    \includegraphics[width=\linewidth]{embeddings}
+    \includegraphics[width=0.5\linewidth]{embeddings}
     \caption{Sparse word embeddings.
@@
 \hline
-$Mixed$ & 67.69 $_{47.60}$ & 37.50 & 53.49 & 34.86 & 28.56 & 25.70 & avg \\
+$Mixed$ & 67.69$_{47.60}$ & 37.50 & 53.49 & 34.86 & 28.56 & 25.70 & avg \\
 \hline
-$FT_{EMEA}$ & 76.77 $_{49.43}$ & 17.16 & 11.99 & 29.58 & 11.64 & 11.10 & avg \\
-$FT_{EPPS}$ & 20.86 $$ & 37.04 & 24.53 & 31.86 & 28.33 & 11.10 & avg \\
+$FT_{EMEA}$ & 76.77$_{49.43}$ & 17.16 & 11.99 & 29.58 & 11.64 & 11.10 & avg \\
+$FT_{EPPS}$ & 20.86 & 37.04 & 24.53 & 31.86 & 28.33 & 11.10 & avg \\
 $FT_{ECB}$   & 26.93 & 27.09 & 56.52 & 22.04 & 17.51 & 13.99 & avg \\
 \hline
-$DC$ & 67.87 $_{45.42}$ & 37.31 & 54.14 & 33.63 & 26.52 & 24.81 & avg\\
+$DC$ & 67.87$_{45.42}$ & 37.31 & 54.14 & 33.63 & 26.52 & 24.81 & avg\\
 \hline
-$LDR_{oracle}$   & 74.26 $_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 & avg \\
-$LDR_{generic}$ & 73.97 $_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & avg \\
+$LDR_{oracle}$   & 74.26$_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 & avg \\
+$LDR^{v2}_{oracle}$   & 74.95$_{49.38}$ & 37.35 & 55.91 & 33.57 & 27.59 & 25.27 & avg \\
+$LDR_{generic}$ & 73.97$_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & avg \\
 $LDR_{rnn}$ & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg \\
-$LDR_{wrong}$ & 72.95 $_{49.78}$ & 37.62 & 53.35 & khreshmoi & newstest & iwslt & avg\\
+$LDR_{wrong}$ & 72.95$_{49.78}$ & 37.62 & 53.35 & khreshmoi & newstest & iwslt & avg\\
 \hline
-\multicolumn{8}{l}{English$\rightarrow$German} \\
+\multicolumn{8}{l}{English $\rightarrow$German} \\
 \hline
-$Mixed$ & 64.57 $_{42.99}$ & 26.47 & 68.67 & 21.77 & 17.04 & 18.85 & avg \\
+$Mixed$ & 64.57$_{42.99}$ & 26.47 & 68.67 & 21.77 & 17.04 & 18.85 & avg \\
 \hline
-$FT_{EMEA}$ & 68.35 $$ & 17.02 & 32.87 & 21.15 & 11.21 & 13.49 & avg\\
+$FT_{EMEA}$ & 68.35 & 17.02 & 32.87 & 21.15 & 11.21 & 13.49 & avg\\
 $FT_{EPPS}$ & 36.19 & 26.29 & 40.71 & 19.67 & 16.47 & 19.24 & avg \\
@@
 \hline
-$DC$ & 63.48 $_{42.98}$ & 26.27 & 66.95 & 21.45 & 16.45 & 18.78 & avg \\
+$DC$ & 63.48$_{42.98}$ & 26.27 & 66.95 & 21.45 & 16.45 & 18.78 & avg \\
 \hline
-$LDR_{oracle}$ & 70.9 $_{46.12}$ & 26.30 & 68.9 & 21.42 & 16.57 & 18.82 & avg\\
-$LDR_{generic}$ & 70.86 $_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & avg\\
+$LDR_{oracle}$ & 70.9$_{46.12}$ & 26.30 & 68.9 & 21.42 & 16.57 & 18.82 & avg \\
+$LDR^{v2}_{oracle}$ & 71.31$_{45.23}$ & 25.98 & 73.74 & 21.6 & 16.37 & 19.2 & avg \\
+$LDR_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & avg\\
 $LDR_{rnn}$ & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg \\
-$LDR_{wrong}$ & 69.51 $_{43.50}$ & 26.31 & 66.86 & khreshmoi & newstest & iwslt & avg \\
+$LDR_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & khreshmoi & newstest & iwslt & avg \\
 \hline
@@
 \hline
-$Mixed$               & 65.42 $_{45.11}$ & 34.7 & 51.38 & 29.27 & 22.21 & 21.63 & avg \\
+$Mixed$               & 65.42$_{45.11}$ & 34.7 & 51.38 & 29.27 & 22.21 & 21.63 & avg \\
 \hline
-$FT_{EMEA}$     & 72.06 $$ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & avg \\
-$FT_{EPPS}$      & 35.47 $$ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & avg \\
-$FT_{ECB}$        & 21.93 $$ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & avg \\
+$FT_{EMEA}$     & 72.06 $ $ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & avg \\
+$FT_{EPPS}$      & 35.47 $ $ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & avg \\
+$FT_{ECB}$        & 21.93 $ $ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & avg \\
 \hline
@@
 \hline
-$FT_{EMEA}$      & 65.64 $$ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & avg\\
-$FT_{EPPS}$       & 24.90 $$ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & avg \\
-$FT_{ECB}$         & 41.8 $$ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & avg\\
+$FT_{EMEA}$      & 65.64 $ $ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & avg\\
+$FT_{EPPS}$       & 24.90 $ $ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & avg \\
+$FT_{ECB}$         & 41.8 $ $ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & avg\\
 \hline
@@
  size=32 & 75.10 $_{48.61}$ & 37.64 & 54.29 & 33.20 & 27.99 & 25.44 \\
+ size=64 & 74.5 $_{50.17}$ & 37.27 & 54.5 & 33.26 & 27.95 & 25.65 \\
 \hline
@@
 Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
-Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{;.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
+Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{Chu18asurvey}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
 The methods in this category can be listed as following:
 \fyTodo{Organize refs}
+
 \begin{itemize}
@@
 On the other hand, model-centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features
+
 \begin{itemize}
-\item Trainining objective \cite{Luong2015SNMT,Senrich2016Mono,Wang2017Instance,Chen2017Cost,Miceli2017Regularize,Zhang18sentence}
-\item Architecture \cite{gulcehre2016monolingual,Zhang16topicinformed,Kobus17domaincontrol,Britz2017mixing,Biao2017CARENMT,Britz2017mixing,Thompson18freezing,Michel2018extreme}.
-\item The decoding algorithm \cite{gulcehre2016monolingual,Khayrallah2017lattice}
+\item Trainining objective\cite{Luong2015SNMT,Senrich2016Mono,Wang2017Instance,Chen2017Cost,Miceli2017Regularize,Zhang18sentence}
+\item Architecture %\cite{gulcehre2016monolingual,Zhang16topicinformed,Kobus17domaincontrol,Britz2017mixing,Biao2017CARENMT,Britz2017mixing,Thompson18freezing,Michel2018extreme}.
+\item The decoding algorithm %\cite{gulcehre2016monolingual,Khayrallah2017lattice}
 \end{itemize}
 
-Beside these works\fyTdodo{Works= oeuvres d'art}, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.
+%Beside these works \fyTdodo{Works= oeuvres d'art}, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.
 

### 2019-05-20 2019/emnlp/sparse.tex
small errors

@@
 \end{itemize}
-On the other hand, model-centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features
-
+On the other hand, model-centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features:
 \begin{itemize}
-\item Trainining objective\cite{Luong2015SNMT,Senrich2016Mono,Wang2017Instance,Chen2017Cost,Miceli2017Regularize,Zhang18sentence}
-\item Architecture %\cite{gulcehre2016monolingual,Zhang16topicinformed,Kobus17domaincontrol,Britz2017mixing,Biao2017CARENMT,Britz2017mixing,Thompson18freezing,Michel2018extreme}.
-\item The decoding algorithm %\cite{gulcehre2016monolingual,Khayrallah2017lattice}
+\item Trainining objective\cite{Luong2015SNMT,Sennrich2016Mono,Wang17instance,Chen17costweighting,Miceli2017Regularize,Zhang18sentence}
+\item Architecture \cite{gulcehre2016monolingual,Zhang16topicinformed,Kobus17domaincontrol,Britz2017mixing,Biao2017CARENMT,Britz2017mixing,Thompson18freezing,Michel2018extreme}.
+\item The decoding algorithm \cite{gulcehre2016monolingual,Khayrallah2017lattice}
 \end{itemize}
-
-%Beside these works \fyTdodo{Works= oeuvres d'art}, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.
+Beside these works \fyTodo{Works= oeuvres d'art}, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.
 
@@
 In this setting, data-based approaches aim to bias the distribution of the train (out-of-domain) data towards matching that of the target domain, using data selection techniques  \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}, or generating adapted pseudo-parallel data through back-translation  \cite{Utiyama2003measure,Wang2014neural,Wang2016connecting,Sennrich2016Mono}.
-Model-centric approaches build domain-adapted models by combining (eg.\ with mixture weights) multiple data sources or multiple systems \cite{Foster07mixture,Wang17instanceweighting}, or by biasing the training objective towards the desired domain using in-domain adaptation data \cite{Luong2015SNMT,Freitag2016FastDA,Chen17costweighting}.
+Model-centric approaches build domain-adapted models by combining (eg.\ with mixture weights) multiple data sources or multiple systems \cite{Foster07mixture,Wang17instance}, or by biasing the training objective towards the desired domain using in-domain adaptation data \cite{Luong2015SNMT,Freitag2016FastDA,Chen17costweighting}.
 Another approach worth mentioning integrates domain information (for instance a domain language model) in the decoding algorithm \cite{Gulcehre2016monolingual}.
@@
 This work has been reanalyzed in a Bayesian framework in \cite{Finkel09hierarchical}, and revisited notably in \cite{Chang10necessity}.
-Following up on \cite{Yang15multidomain}, the recent proposals of \cite{Peng17multitask} apply the same idea with neural architectures, using domain specific masking to zero out the parameters modeling domains irrelevant to the current input sentence.
+Following up on \cite{Yang15unified}, the recent proposals of \cite{Peng17multitask} apply the same idea with neural architectures, using domain specific masking to zero out the parameters modeling domains irrelevant to the current input sentence.
 Compared to our work, these techniques are used in deep layers of the network and applied to sequence labelling tasks.

### 2019-05-20 2019/emnlp/sparse.tex
discuss most results, fix a couple of problems with the bib

@@
 \subsection{Results}
-Discussion 1:
-\begin{itemize}
-\item transformer better than RNN
-\item Fine tuning is the best strategy - but subject to catastrophic
-\item Sparse better than joint - better than domain control
-\item Across th board / language \& architecture - always good for EMEA , less good for EPPS, varying for ECB
-\item better ? than domain control ? WMCD ?
-\end{itemize}
+% Discussion 1:
+% \begin{itemize}
+% \item transformer better than RNN
+% \item Fine tuning is the best strategy - but subject to catastrophic
+% \item Sparse better than joint - better than domain control
+% \item Across th board / language \& architecture - always good for EMEA , less good for EPPS, varying for ECB
+% \item better ? than domain control ? WMCD ?
+% \end{itemize}
 
-Discussion 2:
-\begin{itemize}
-\item Sparse 1 > Sparse 2
-\item out of domain better than Mixed
-\item Sparse 3 ??
-\item ajouter Sparse 4
-\end{itemize}
+% Discussion 2:
+% \begin{itemize}
+% \item Sparse 1 > Sparse 2
+% \item out of domain better than Mixed
+% \item Sparse 3 ??
+% \item ajouter Sparse 4
+% \end{itemize}
+
+Our main results are summarized respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems. \fyTodo{Check missing numbers in Table}.
+
+We first concentrate on the left part of these tables, where we analyze the performance for the three domains of interest.
+A first series of observations is that Transformer is consistently better than RNN, and that fine-tuning on a domain-specific corpus, when applicable, is the best way to optimize the performance on that domain, albeit at the cost of a marked (even sometimes catastrophic, eg.\ for the EMEA-tuned Transformer system) decrease in performance for the other domains.
+
+For these test sets, our proposal (\texttt{LDR}$_{oracle}$) is consistently better that the mix-domain strategy, with gains than range from very large (for EMEA and ECB) to unsignificant (for EPPS in most conditions).
+This means that our architecture is somehow able to compensate for the data unbalance, and to raise the performance of the multi-domain system close to the best (fine-tuned) system in each domain.
+We even observe rare cases where the \texttt{LDR}$_{oracle}$ system outperforms fine-tuning (eg.\ Transformer en:ge).
+\texttt{LDR}$_{oracle}$ is also better than Domain Control in three conditions out of four, \texttt{DC} being seemingly a better choice for the RNN than for the Transformer architecture.
+The comparison with the WCDMT model is less favourable for our RNN systems, which are arguably much simpler.\fyTodo{More on this.}
+As expected, ignoring the domain label yields a drop in performance for the two fallback strategies considered in this study.
+The first, \texttt{LDR}$_{generic}$, only use the generic embeddings.
+The second, \texttt{LDR}$_{pred}$\fyTodo{Fix this}, resorts to automatically predicted domain labels.\fyTodo{Explain how} Note that this decrease is however quite limited\fyTodo{How much on average ?}, showing that our architecture is quite robust to noisy labels.
+Even in the worst case scenario where all domain tags are intently wrong (\texttt{LDR}$_{wrong}$), we see that the generic part still ensures a satisfying level of performance.\fyTodo{Comment V2 from Minh}
+
+The right part of Tables~\ref{tab:results-trsf} and \ref{tab:results-rnn} report performance on test sets for new domains not seen in training.
+Khresmoi texts, which are extracted from summaries of medical article are assumed to from the same domain as EMEA data; for the other two test sets, we assume no domain and only use the generic embeddings.
+Here, we see that our best system \texttt{LDR}$_{oracle}$ is slightly lagging behind the \texttt{Mixed} baselines (the difference ranges from unsignificant to one Bleu point), which can be attributed to larger useful (generic) embedding in the \texttt{Mixed} setting.
+Here again, we see that using the sole generic part makes the system robust with respect to unknown or misleading domain tags.
+We finally outline that our approach outperforms DC for these test sets (were we have to predict a tag)\fyTodo{Fix this}, and even outperforms WCDMT, which has no built in mechanism to handle data from unknown domains.
 
@@
 \hline
-Model & EMEA & EPPS & ECB & khreshmoi & newstest2014 & iwslt2010 & Avg. \\
+Model & EMEA & EPPS & ECB & Khresmoi & NewsTest2014 & IWSLT2010 & Avg. \\
 \hline
+\multicolumn{8}{l}{} \\[-9pt]  
 \multicolumn{8}{l}{English$\rightarrow$French} \\
@@
   \hline
+\multicolumn{8}{l}{} \\[-9pt]
 \multicolumn{8}{l}{English $\rightarrow$German} \\
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for the Transformer architecture over test sets.}
+\caption{BPE-detokenized BLEU scores for the Transformer architecture \label{tab:results-trsf}}
 \end{table*}
@@
 \hline
-Model & EMEA & EPPS & ECB & khreshmoi & newstest2014 & iwslt2010 & Avg. \\
+Model & EMEA & EPPS & ECB & Khresmoi & NewsTest2014 & IWSLT2010 & Avg. \\
 \hline
+\multicolumn{8}{l}{} \\[-9pt]
 \multicolumn{8}{l}{English$\rightarrow$French} \\
@@
 \hline
+\multicolumn{8}{l}{} \\[-9pt]
 \multicolumn{8}{l}{English$\rightarrow$German} \\
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for the RNN architecture over test sets.}
+\caption{BPE-detokenized BLEU scores for the RNN architecture \label{tab:results-rnn}}
 \end{table*}
+% \clearfloat
 
 \section{Discussion\label{sec:Discussion}}
-\subsection{Sparse's performance}
-First, we analyze the performance of models on inner test sets which are created by randomly splited from same corpora as training set.
+% \subsection{Sparse's performance}
+% First, we analyze the performance of models on inner test sets which are created by randomly splited from same corpora as training set.
 The 4 table show remarquable improvement 6 points BLEU in average in EMEA translation which is comparable with finetuned model.
@@
 \\
-Secondly, we compare performance of models on IWSLT test 2010 and newstest 2014 which are more general than inner test sets.
+% Secondly, we compare performance of models on IWSLT test 2010 and newstest 2014 which are more general than inner test sets.
 On these test sets, mixed, Sparse has approximately same score in Transformer and are a little better than Domain Control while in Recurrent Net Work architecture, WDCMT is better than mixed, Sparse and Domain Control for English -French pair.
@@
 
-\subsection{Region's size \label{secc:region_size}}
+\subsection{Balancing generic and domain-specific representations\label{secc:region_size}}
 
-We also test different setting of region size in word embedding.
-However, we limit our research in 4 settings: $[4,4,4,496]$, $[8,8,8,488]$, $[16,16,16,464]$, $[32,32,32,32,416]$.
-We test those settings with the same condition as above.
-The scores are in Table~\ref{tab:embedding-size}.\fyTodo{Comment, homegenize captions.}
-
-Results are very close so we pick 8.\fyTodo{Experiment with 64}.
-
+An important practical question concerns the balance between the generic and the domain-specific part of the embedding.
+In the limit where the domain specific part is very small, we should recover the performance of the \texttt{Mixed} system; conversely, we expect to see a less effective sharing of data across domains by increasing the domain-specific regions.
+Table~\ref{tab:embedding-size} reports the result of  a series of experiments for the Transformer\fyTodo{Check This} architecture with varying domain-specific sizes allocating between 4 and 64 cells for domain-specific information, and the complement to 512 for the generic part.
+The differences being overall quite small \fyTodo{Comments - is this due to prior ?}, we decided to allocate $8$ cells for the domain specific part.
+\fyTodo{Easily accomodate more domains}
+\fyTodo{Given the ways embeddings are computed, why not add more domains, and test robustness agains data presentation order ?}
 \begin{table*}[!h]
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for the RNN architecture over test sets.}
+\caption{BPE-detokenized BLEU scores for the Transformer architecture for varying domain-specific embedding sizes}
 \end{table*}
@@
 
-Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
+% Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
 Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{Chu18asurvey}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
 The methods in this category can be listed as following:
-\fyTodo{Organize refs}
+% \fyTodo{Organize refs}
 
-\begin{itemize}
-\item score relatedness of sentences to the target domain by Language Model \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}.
-\item generating pseudo parallel data \cite{Utiyama2003measure,Wang2016connecting,Wang2014neural}
-\end{itemize}
-On the other hand, model-centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features:
-\begin{itemize}
-\item Trainining objective\cite{Luong2015SNMT,Sennrich2016Mono,Wang17instance,Chen17costweighting,Miceli2017Regularize,Zhang18sentence}
-\item Architecture \cite{gulcehre2016monolingual,Zhang16topicinformed,Kobus17domaincontrol,Britz2017mixing,Biao2017CARENMT,Britz2017mixing,Thompson18freezing,Michel2018extreme}.
-\item The decoding algorithm \cite{gulcehre2016monolingual,Khayrallah2017lattice}
-\end{itemize}
-Beside these works \fyTodo{Works= oeuvres d'art}, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.
+% \begin{itemize}
+% \item score relatedness of sentences to the target domain by Language Model \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}.
+% \item generating pseudo parallel data \cite{Utiyama2003measure,Wang2016connecting,Wang2014neural}
+% \end{itemize}
+% On the other hand, model-centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features:
+% \begin{itemize}
+% \item Trainining objective\cite{Luong2015SNMT,Sennrich2016Mono,Wang17instance,Chen17costweighting,Miceli2017Regularize,Zhang18sentence}
+% \item Architecture \cite{Gulcehre2016monolingual,Zhang16topicinformed,Kobus17domaincontrol,Britz2017mixing,Biao2017CARENMT,Britz2017mixing,Thompson18freezing,Michel2018extreme}.
+% \item The decoding algorithm \cite{Gulcehre2016monolingual,Khayrallah2017lattice}
+% \end{itemize}
+% Beside these works \fyTodo{Works= oeuvres d'art}, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.
 
@@
 
-The most common adaptation scenario is using (mostly) out-of-domain training data while testing on a low-resource domain.
-In this setting, data-based approaches aim to bias the distribution of the train (out-of-domain) data towards matching that of the target domain, using data selection techniques  \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}, or generating adapted pseudo-parallel data through back-translation  \cite{Utiyama2003measure,Wang2014neural,Wang2016connecting,Sennrich2016Mono}.
-Model-centric approaches build domain-adapted models by combining (eg.\ with mixture weights) multiple data sources or multiple systems \cite{Foster07mixture,Wang17instance}, or by biasing the training objective towards the desired domain using in-domain adaptation data \cite{Luong2015SNMT,Freitag2016FastDA,Chen17costweighting}.
+The most common adaptation scenario uses (mostly) out-of-domain data in training, while testing on a low-resource in-domain set of texts.
+In this setting, \emph{data-based approaches} aim to bias the distribution of the train (out-of-domain) data towards matching that of the target domain, using data selection techniques \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}, or generating adapted pseudo-parallel data through back-translation  \cite{Utiyama2003measure,Wang2014neural,Wang2016connecting,Sennrich2016Mono}.
+\emph{Model-centric approaches} build domain-adapted models by combining (eg.\ with mixture weights) multiple data sources or multiple systems \cite{Foster07mixture,Wang17instance}, or by biasing the training objective towards the desired domain using in-domain adaptation data \cite{Luong2015SNMT,Freitag2016FastDA,Chen17costweighting}.
 Another approach worth mentioning integrates domain information (for instance a domain language model) in the decoding algorithm \cite{Gulcehre2016monolingual}.

### 2019-05-21 2019/emnlp/sparse.tex
start the analysis of embeddings

@@
 
-\subsection{Word embedding analysis}
-We analyze neigborhood in word embedding space of polysemies.
+\subsection{Word embedding analysis \label{ssec:word_embeddings}}
+One of our main assumption is that the difference between domains can be confined at the lexical level, warranting our decision to specialize lexical representation for each domain, while the remaining part of the network is shared across domains.
+Linguistically, this assumption relates to the ``one domain, one meaning'' assumption, which express the fact that in many cases, polysemy correspond to variation in use across domain.
+In its weaker form, it allows us to assume that all occurrences of a given form in a given domain correspond to the same sense and share the same embedding; the same form occuring in different domain is allowed to have one distinct embedding per domain, which may help capture polysemy and lexical ambiguity in translation.
+
+To check the validity of this hypothesis, we performed the following analysis of embeddings learned with the multi-domain system\fyTodo{Language and system} For each unit\footnote{In this study, we work with BPE units, not words, meaning that in many cases we observe the variation of use of word parts.
+As we work with a large inventory, many of these units correspond to actual words, and we focus on these in our comments.} in our dictionary, we compute the $k$ nearest neighbours for each domain, \fyTodo{As follows} yielding three lists of nearest neighbours.
+The size of their intersection should then be a sign of a variation of use across domains; conversely, a stable set of neighbours across domains should reflect the stability of word use.
+
+- impairment - mostly renal / hepatic impairment - translates as insuffisance in EMEA, quite different in ECB
+- convenience - mostly credit card convenience in ECB (), mostly for convenience (par commodité in EMEA).
+- turnover in ECB= chiffre d'affaire; EMEA - we observe collocations such as ``bone turnover'' (remodelage or renouvellement osseux') and more generally of cell renewall.
+
+Conversely, same sense
+
+This only works for sufficiently frequent words.
 

### 2019-05-21 2019/emnlp/sparse.tex
new score

@@
 $LDR_{generic}$ & 73.97$_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & avg \\
-$LDR_{rnn}$ & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg \\
+$LDR_{rnn}$ & 74.29 $_{49.84}$ & 37.73 & 54.01 & 33.34 & 28.32 & 26 & avg \\
 $LDR_{wrong}$ & 72.95$_{49.78}$ & 37.62 & 53.35 & khreshmoi & newstest & iwslt & avg\\
@@
 $LDR_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & avg\\
-$LDR_{rnn}$ & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg \\
+$LDR_{rnn}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & 21.45 & 16.61 & 19.07 & avg \\
 $LDR_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & khreshmoi & newstest & iwslt & avg \\
@@
 $LDR_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & avg\\
-$LDR_{rnn}$        & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg\\
+$LDR_{rnn}$        & 72.76 $_{46.35}$ & 35.1 & 50.38 & 30.02 & 23.28 & 21.32 & avg\\
 $LDR_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 & khreshmoi & newstest & iwslt & avg \\
@@
 $LDR_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & avg\\
+$LDR^{v2}_{oracle}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & avg\\
 $LDR_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & avg\\
-$LDR_{rnn}$        & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg\\
+$LDR_{rnn}$        & 63.17 $_{39.92}$ & 22.51 & 64 & 15.94 & 11.32 & 14.27 & avg\\
 $LDR_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & khreshmoi & newstest & iwslt & avg \\

### 2019-05-21 2019/emnlp/sparse.tex
more changes in results

@@
 The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain.
-We study here a different scenario, multi-domain adaptation \cite{Sennrich13multidomain,Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
+We study here a different scenario, \emph{multi-domain adaptation} \cite{Sennrich13multidomain,Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
 This allows us to be both data efficient (all data is used to train all domains) and compuationally efficient (we only need to train one domain).
-In our setting, we assume that the domain of test documents is known in advance.
+In our setting, we assume that the domain of test documents is known.
 
@@
 \item experimental results that demonstrate the effectiveness of this technique;
+\item analysis of embeddings
 \end{itemize}
@@
 
-\section{Sparse embeddings \label{sec:sparse_embeddings}}
+\section{Lexicalize domain representations\label{sec:lexicalized_embeddings}}
 \fyDone{Use meaningful titles throughout}
 
-\subsection{Problem statement}
+\subsection{Problem statement \label{ssec:statement}}
 \fyTodo{Use math for notation, label equations and such}
@@
 Our first approach to multi-domain machine translation is to build such trivial solution by assigning different regions of word embedding vector to set of domain paramters $\theta_i$ and set of shared parameters $\theta_s$ while keep sharing other parameters between domains.
-\subsection{Sparse word embedding}
-\label{sec:sparse}
-This section explains a new design for word embedding used in multi-domain model.
+
+\subsection{Lexicalized domain embeddings \label{ssec:lde}}
+
+This section explains a new design for word embeddings used in a multi-domain model.
 As mentioned above, we separate parameter set into $D+1$ parameter sets correspond to D domains and one shared paramater set.
-Our first attempt is to split standard vector word embedding into $D+1$ regions named $\theta_i$ for $i \in [1,..,D]$ and $\theta_s$.
+Our first attempt is to split standard vector word embedding into $D+1$ regions named $\theta_i$ for $i \in [1,..,D]$ and $\theta_s$.\fyTodo{nullified not vanishing}
 \begin{figure}
 \center
-    \includegraphics[width=0.5\linewidth]{embeddings}
-    \caption{Sparse word embeddings.
-The red region contains units that are activated only when an input from the corresponding domain is presented; the other (green) regions are deactivated; the blue (generic) region is always activated.}
+    \includegraphics[width=0.52\textwidth]{embeddings}
+    \caption{Lexicalized domain embeddings.
+When processing a sample from domain $2$, we only activate the corresponding parameter region ($\theta_2$) in the input embeddings; the remaining domain-specific parts are zeroed out, and do not receive any update.
+The generic part is always active, and is updated irrespective of the input domain.}
     \label{fig:network}
@@
 We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).\footnote{The EMEA dataset as distributed on the OPUS site contains many duplicates.
-We therefore report two numbers: the first is comparable to what has been published on earlier studies (eg.\ \fyTodo{Missing ref}), the second one is obtained by making the test set entirely disjoint from the train.} The validation sets are used to chose the best model according to the average BLEU score.
+We therefore report two numbers: the first is comparable to what has been published on earlier studies (eg.\ \cite{Zeng18multidomain}), the second one is obtained by making the test set entirely disjoint from the train.} The validation sets are used to chose the best model according to the average BLEU score.
 Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
@@
 \subsection{Baselines \label{ssec:baselines}}
-To validate our findings, we compared sparse embedding models with standard models using both attentional Recurrent Neural Networks (RNNs) \cite{Bahdanau15learning} and the recent Transformer architecture of \citet{Vaswani17attention}.
+To validate our findings, we compared lexicalized domain embedding models with standard models using both attentional Recurrent Neural Networks (RNNs) \cite{Bahdanau15learning} and the recent Transformer architecture of \citet{Vaswani17attention}.
 Our baselines consist of generic models trained with a simple concatenation of all corpora (denoted \texttt{mixed} in the score tables below); models finetuned separately to each domain using (10000, 15000, 5000) iterations over in-domain data (denoted \texttt{+ft(EMEA)}, \texttt{+ft(EPPS)}, \texttt{+ft(ECB)} below); models using domain tags as in \cite{Kobus17domaincontrol} (noted \texttt{DC} below); and finally the multi-domain model of \citet{Zeng18multidomain} (denoted \texttt{WDCMT} below).\fyTodo{Change names in tables}
@@
 %  Attention Recurrent Neural Network: Bidirection LSTM encoder, unidirectional LSTM decoder, Luong Attention Mechanism, size of word embedding: 512; size of hidden layers: 1024; number of layers: 1
-\item The domain control systems are exactly as their baseline counterpart (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.\fyTodo{Check This!}
+\item The domain control systems are exactly as their baseline counterpart (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.\fyDone{Check This!}
   % C: size of domain embedding:2 ; same settings for both architecture: Transformer and Attention Recurrent Neural Network.
-\item The \texttt{WDCMT} system also has one layer on the encoder side (a bidirectional Gated recurrent units (GRU)), as well as one layer on the decoder side (a unidirectional conditional GRU); the dimension of ``domain'' layers is 300.
+\item The \texttt{WDCMT} system also has one layer on the encoder side (a bidirectional Gated recurrent units (GRU) layer), as well as one layer on the decoder side (a unidirectional conditional GRU layer); the dimension of ``domain'' layers is~300.
 \end{itemize}
-To train NMT systems, we use Adam, with parameters $\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$ for RNNs; with parameters $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} ($warmup\_steps=4000$) for Transformer.
-Our implementation is from OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein2017OpenNMT}, except for WDCMT where we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyTodo{batch size}
+To train NMT systems, we use Adam, with parameters $\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$ for RNNs; with parameters $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} for Transformer (using $warmup\_steps=4000$).
+For all systems, we use a batch size of~128.
+Our implementation is from OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein2017OpenNMT}, except for WDCMT where we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyDone{batch size}
 
-\subsection{Sparse NMT model configuration}\fyTodo{Change sparse.}
+\subsection{Implementing LDR}\fyTodo{Check acronym.}
 \fyTodo{Motivate the split - discuss experimentally embedding size}
-For the implementation of lexicalized domain embeddings, we split the embedding vector (of size 512) into four regions: 3 are domain specific and 1 is generic, with respective sizes $[8,8,8,488]$, a choice that will be dicussed in Section~\ref{secc:region_size}.
-If a sentence originates from domain $i$, the domain specific regions for all domains $j \neq i$ will be zeroed out; while the other regions are activated.
+For the implementation of lexicalized domain representations, we split the embedding vector (of size 512) into four regions: 3 are domain specific and 1 is generic, with respective sizes $[8,8,8,488]$, a choice that will be dicussed in Section~\ref{secc:region_size}.
+If a sentence originates from domain $i$, the domain specific regions for all domains $j \neq i$ will be zeroed out; while the other regions are activated (cf.
+Figure~\ref{fig:network}).
 We then use a dense layer of size~512 to fuse the region for the active domain and the ``generic'' region.
@@
 \begin{algorithmic}[1]
-\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for $d$ domains, Optimization algorithm Opt}
+\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for $d$ domains, Batch size $B$, Optimization algorithm $\operatorname{Opt}$}
 \REPEAT 
-\STATE{Randomly pick $i \in [1,..,d]$ w.r.t multinomial distribution $[\frac{|C_i|}{\displaystyle{\mathop{\sum}_{i\in [1,..,d]}|C_i|}}]$.}
+\STATE{Randomly pick $i \in [1,..,d]$ w.r.t the multinomial distribution $[\frac{|C_i|}{\displaystyle{\mathop{\sum}_{i\in [1,..,d]}|C_i|}}]$.}
 \STATE{Randomly pick $B$ sentences from $C_i$.}
@@
 \STATE{Apply learning rules using computed gradients: 
-$$ \theta_s = \theta_s - Opt(\frac{\partial L}{\partial \theta_s}(W_g))$$ 
-$$ \theta_i = \theta_i - Opt(\frac{\partial L}{\partial \theta_i}(W_i))$$}
+$$ \theta_s = \theta_s - \operatorname{Opt}(\frac{\partial L}{\partial \theta_s}(W_g))$$ 
+$$ \theta_i = \theta_i - \operatorname{Opt} (\frac{\partial L}{\partial \theta_i}(W_i))$$}
 \UNTIL{convergence}
@@
 In training, we use same optimization algorithm as for the baselines.
-Note that each iteration of algorithm \ref{alg:multidomain} uses 2 batches: generic batch using only generic region; and domain batch using generic region and domain region.
+Note that each iteration of algorithm \ref{alg:multidomain} uses 2~batches: a ``generic'' batch updating only the generic region; and ``domain-specific'' batch updating just the domain-specific parameters.
 The reason to do that is to minimize simultaneously two losses:
@@
 
-We also take into account the problem of overfitting that could occur for small corpora.
+% We also take into account the problem of overfitting that could occur for small corpora.
 Indeed, real data are not always balanced for every domain.
@@
 Therefore, we select batches from each corpora with frequencies proportional to its size.
-\fyTodo{Refs on this ?
-or contrast ?}
+The batch selection procedure (step~2 of algorithm~\ref{alg:multidomain}) ensures that each domain is represented according to its frequency in the training data.
+In our experiments, this means that sentences from the Europarl domain will be selected more frequently that the two other domains.
+In an attempt to build a more balanced training procedure, we have also experimented with an alternative selection procedure, where $i$ is selected according to the multinomial distribution $[\frac{\sqrt{|C_i|}}{\displaystyle{\mathop{\sum}_{i\in [1,..,d]}\sqrt{|C_i|}}}]$.
+\fyTodo{Refs on this ? or contrast ?}
 
-\subsection{Results}
+\subsection{Results \label{ssec:results}}
 % Discussion 1:
@@
 
-Our main results are summarized respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems. \fyTodo{Check missing numbers in Table}.
+Our main results are summarized respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems. \fyTodo{Footnote for missing numbers}.
 
 We first concentrate on the left part of these tables, where we analyze the performance for the three domains of interest.
-A first series of observations is that Transformer is consistently better than RNN, and that fine-tuning on a domain-specific corpus, when applicable, is the best way to optimize the performance on that domain, albeit at the cost of a marked (even sometimes catastrophic, eg.\ for the EMEA-tuned Transformer system) decrease in performance for the other domains.
+A first series of observations is that Transformer is consistently better than RNN, and that fine-tuning on a domain-specific corpus, when applicable, is almost the best way to optimize the performance on that domain.\footnote{This is not so clear for the EPPS, where fine-tuning does not seem to help.}  Note that fine-tuning will however yield a marked (even sometimes catastrophic, eg.\ for the EMEA-tuned Transformer system) decrease in performance for the other domains.
 
 For these test sets, our proposal (\texttt{LDR}$_{oracle}$) is consistently better that the mix-domain strategy, with gains than range from very large (for EMEA and ECB) to unsignificant (for EPPS in most conditions).
-This means that our architecture is somehow able to compensate for the data unbalance, and to raise the performance of the multi-domain system close to the best (fine-tuned) system in each domain.
-We even observe rare cases where the \texttt{LDR}$_{oracle}$ system outperforms fine-tuning (eg.\ Transformer en:ge).
+This means that our architecture is somehow able to compensate for the data unbalance and to raise the performance of the multi-domain system close to the best (fine-tuned) system in each domain.
+We even observe rare cases where the \texttt{LDR}$_{oracle}$ system outperforms fine-tuning (eg.\ Transformer en:ge in the EMEA domain).
 \texttt{LDR}$_{oracle}$ is also better than Domain Control in three conditions out of four, \texttt{DC} being seemingly a better choice for the RNN than for the Transformer architecture.
-The comparison with the WCDMT model is less favourable for our RNN systems, which are arguably much simpler.\fyTodo{More on this.}
+The comparison with the WDCMT model is less favourable for our RNN systems, which are arguably much simpler.\footnote{The direct comparison with WDCMT is difficult as the implementation differ in many ways: a different framework, different cell types (LSTMs vs.
+conditional GRUs), etc.} \fyDone{More on this.}
 As expected, ignoring the domain label yields a drop in performance for the two fallback strategies considered in this study.

### 2019-05-21 2019/emnlp/sparse.tex
More review in section 4

@@
 We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).\footnote{The EMEA dataset as distributed on the OPUS site contains many duplicates.
-We therefore report two numbers: the first is comparable to what has been published on earlier studies (eg.\ \cite{Zeng18multidomain}), the second one is obtained by making the test set entirely disjoint from the train.} The validation sets are used to chose the best model according to the average BLEU score.
-Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
+We therefore report two numbers: the first is comparable to what has been published on earlier studies (eg.\ \cite{Zeng18multidomain}), the second one is obtained by making the test set entirely disjoint from the train.} The validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.
+Using internal test sets corresponds to somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from three other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
 This enables us to evaluate the loss in performance when the test set is from a domain not seen in training.
@@
 To train NMT systems, we use Adam, with parameters $\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$ for RNNs; with parameters $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} for Transformer (using $warmup\_steps=4000$).
-For all systems, we use a batch size of~128.
+For all systems, we use a batch size of~128 and a dropout rate of 0.1 for all layers.
 Our implementation is from OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein2017OpenNMT}, except for WDCMT where we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyDone{batch size}
@@
 We then use a dense layer of size~512 to fuse the region for the active domain and the ``generic'' region.
-Training is formalized in algorithm~\ref{alg:multidomain}.\fyTodo{Notation B for batch size}
+Training is formalized in algorithm~\ref{alg:multidomain}.
 
@@
 Note that each iteration of algorithm \ref{alg:multidomain} uses 2~batches: a ``generic'' batch updating only the generic region; and ``domain-specific'' batch updating just the domain-specific parameters.
-The reason to do that is to minimize simultaneously two losses:
+This is because we wish to simultaneously minimize two losses:
 \begin{align*}
@@
 
-Our main results are summarized respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems. \fyTodo{Footnote for missing numbers}.
+Our main results are summarized respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems, where we report BLEU scores \citecomputed after detokenization.
+\fyTodo{Footnote for missing numbers}.
 
@@
 \texttt{LDR}$_{oracle}$ is also better than Domain Control in three conditions out of four, \texttt{DC} being seemingly a better choice for the RNN than for the Transformer architecture.
-The comparison with the WDCMT model is less favourable for our RNN systems, which are arguably much simpler.\footnote{The direct comparison with WDCMT is difficult as the implementation differ in many ways: a different framework, different cell types (LSTMs vs.
+The comparison with the \texttt{WDCMT} model is less favourable for our RNN systems, which are arguably much simpler.\footnote{The direct comparison with WDCMT is difficult as the implementation differs in many ways from our RNN: a different framework, different cell types (LSTMs vs.
 conditional GRUs), etc.} \fyDone{More on this.}
 As expected, ignoring the domain label yields a drop in performance for the two fallback strategies considered in this study.
-The first, \texttt{LDR}$_{generic}$, only use the generic embeddings.
-The second, \texttt{LDR}$_{pred}$\fyTodo{Fix this}, resorts to automatically predicted domain labels.\fyTodo{Explain how} Note that this decrease is however quite limited\fyTodo{How much on average ?}, showing that our architecture is quite robust to noisy labels.
-Even in the worst case scenario where all domain tags are intently wrong (\texttt{LDR}$_{wrong}$), we see that the generic part still ensures a satisfying level of performance.\fyTodo{Comment V2 from Minh}
+The first, \texttt{LDR}$_{generic}$, only uses the generic part of the embeddings.
+The second, \texttt{LDR}$_{pred}$, resorts to automatically predicted domain labels\footnote{Our domain classifier uses a bi-LSTM RNN encoder, followed by a simple softmax layer.
+Its precision on a development set are higher than 95\%.}
+\fyTodo{Explain how} Note that this decrease is however quite limited (and hardly significant for \texttt{LDR}$_{pred}$), showing that our architecture is quite robust to noisy labels.
+Even in the worst case scenario where all domain tags are intently wrong (\texttt{LDR}$_{wrong}$), we see that the generic part still ensures a satisfying level of performance.
+A last contrast is with \texttt{LDR}$_{oracle}^{0.5}$ where we change the distribution of training sentence to increase the number of ECB samples: as a result, we see a small decrease for EMEA and EPPS, and a large boost for ECB.
+This shows that our technique can be used in conjunction to other well known strategies for performing domain adaptation.
 
-The right part of Tables~\ref{tab:results-trsf} and \ref{tab:results-rnn} report performance on test sets for new domains not seen in training.
-Khresmoi texts, which are extracted from summaries of medical article are assumed to from the same domain as EMEA data; for the other two test sets, we assume no domain and only use the generic embeddings.
-Here, we see that our best system \texttt{LDR}$_{oracle}$ is slightly lagging behind the \texttt{Mixed} baselines (the difference ranges from unsignificant to one Bleu point), which can be attributed to larger useful (generic) embedding in the \texttt{Mixed} setting.
+% \fyTodo{Comment V2 from Minh}
+
+The right parts of Tables~\ref{tab:results-trsf} and \ref{tab:results-rnn} report performance on test sets for new domains not seen in training.
+Khresmoi texts, which are extracted from summaries of medical articles are assumed to be from the same domain as EMEA data; for the other two test sets, we assume no domain and only use the generic part of the embeddings.
+Here, we see that our best system \texttt{LDR}$_{oracle}$ is slightly lagging behind the \texttt{Mixed} baselines (the difference ranges from unsignificant to one Bleu point), which can be attributed to larger useful (generic) embeddings in the \texttt{Mixed} setting.
 Here again, we see that using the sole generic part makes the system robust with respect to unknown or misleading domain tags.
-We finally outline that our approach outperforms DC for these test sets (were we have to predict a tag)\fyTodo{Fix this}, and even outperforms WCDMT, which has no built in mechanism to handle data from unknown domains.
+We finally outline that our approach outperforms DC for these test sets (were we have to commit to a predicted domain tag), and even outperforms WDCMT (for German-English).
+This might be due to the fact that WDCMT has no built-in mechanism to ignore, as we do in the \texttt{LDR}$_{generic}$ systems, the possibly misleading predicted domain information.
 
@@
 $LDR_{generic}$ & 73.97$_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & avg \\
-$LDR_{rnn}$ & 74.29 $_{49.84}$ & 37.73 & 54.01 & 33.34 & 28.32 & 26 & avg \\
-$LDR_{wrong}$ & 72.95$_{49.78}$ & 37.62 & 53.35 & khreshmoi & newstest & iwslt & avg\\
+$LDR_{pred}$ & 74.29 $_{49.84}$ & 37.73 & 54.01 & 33.34 & 28.32 & 26 & avg \\
+$LDR_{wrong}$ & 72.95$_{49.78}$ & 37.62 & 53.35 & - & - & - & avg\\
   \hline
@@
 $LDR_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & avg\\
-$LDR_{rnn}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & 21.45 & 16.61 & 19.07 & avg \\
-$LDR_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & khreshmoi & newstest & iwslt & avg \\
+$LDR_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & 21.45 & 16.61 & 19.07 & avg \\
+$LDR_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & - & - & - & avg \\
 \hline
@@
 $LDR_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & avg\\
-$LDR_{rnn}$        & 72.76 $_{46.35}$ & 35.1 & 50.38 & 30.02 & 23.28 & 21.32 & avg\\
-$LDR_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 & khreshmoi & newstest & iwslt & avg \\
+$LDR_{pred}$        & 72.76 $_{46.35}$ & 35.1 & 50.38 & 30.02 & 23.28 & 21.32 & avg\\
+$LDR_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 &  - & - & - & avg \\
 \hline
@@
 $LDR_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & avg\\
-$LDR_{rnn}$        & 63.17 $_{39.92}$ & 22.51 & 64 & 15.94 & 11.32 & 14.27 & avg\\
-$LDR_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & khreshmoi & newstest & iwslt & avg \\
+$LDR_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64 & 15.94 & 11.32 & 14.27 & avg\\
+$LDR_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & - & - & - & avg \\
 \hline
@@
 
-\section{Discussion\label{sec:Discussion}}
+\section{Complementary experiments\label{sec:Discussion}}
 % \subsection{Sparse's performance}
@@
 
-An important practical question concerns the balance between the generic and the domain-specific part of the embedding.
+An important practical question concerns the balance between the generic and the domain-specific part of the embeddings.
 In the limit where the domain specific part is very small, we should recover the performance of the \texttt{Mixed} system; conversely, we expect to see a less effective sharing of data across domains by increasing the domain-specific regions.
-Table~\ref{tab:embedding-size} reports the result of  a series of experiments for the Transformer\fyTodo{Check This} architecture with varying domain-specific sizes allocating between 4 and 64 cells for domain-specific information, and the complement to 512 for the generic part.
-The differences being overall quite small \fyTodo{Comments - is this due to prior ?}, we decided to allocate $8$ cells for the domain specific part.
-\fyTodo{Easily accomodate more domains}
+Table~\ref{tab:embedding-size} reports the result of  a series of experiments for the Transformer architecture (English-French) with varying domain-specific sizes allocating between 4 and 64 cells for domain-specific information, and the complement to 512 for the generic part.
+The differences are overall quite small in our experimental setting, where the training data is relatively limited and does not require to use a large embedding size.
+We therefore decided to allocate $8$~cells for the domain specific part.
+This suggests that we could easily accomodate more domains with the same architecture, and even reserve some regions to handle supplementary incoming training data.
+% \fyTodo{Easily accomodate more domains}
 \fyTodo{Given the ways embeddings are computed, why not add more domains, and test robustness agains data presentation order ?}
+
 \begin{table*}[!h]
@@
 \hline
-$LDR$ & EMEA & EPPS & ECB & khreshmoi & newstest2014 & iwslt2010 \\
+$LDR$ & EMEA & EPPS & ECB & Khresmoi & NewsTest2014 & IWSLT2010 \\
 \hline
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for the Transformer architecture for varying domain-specific embedding sizes}
+\caption{BPE-detokenized BLEU scores for the Transformer architecture for varying domain-specific embedding sizes \label{tab:embedding-size}}
 \end{table*}
 
-\subsection{Word embedding analysis \label{ssec:word_embeddings}}
-One of our main assumption is that the difference between domains can be confined at the lexical level, warranting our decision to specialize lexical representation for each domain, while the remaining part of the network is shared across domains.
-Linguistically, this assumption relates to the ``one domain, one meaning'' assumption, which express the fact that in many cases, polysemy correspond to variation in use across domain.
-In its weaker form, it allows us to assume that all occurrences of a given form in a given domain correspond to the same sense and share the same embedding; the same form occuring in different domain is allowed to have one distinct embedding per domain, which may help capture polysemy and lexical ambiguity in translation.
+\subsection{Analysis of Word Embeddings \label{ssec:word_embeddings}}
+One of our main assumptions is that the difference between domains can be confined at the lexical level, warranting our decision to specialize lexical representations for each domain, while the remaining part of the network is shared across domains.
+Linguistically, this assumption relates to the ``one domain, one sense'' assumption \fyTodo{Refs}, which expresses the fact that in many cases, polysemy corresponds to variation in use across domain.
+In its weaker form, it allows us to assume that all occurrences of a given form in a given domain correspond to the same sense and share the same embedding; the same form occuring in different domains is allowed to have one distinct embedding per domain, which may help capture polysemy and lexical ambiguity in translation.
 
-To check the validity of this hypothesis, we performed the following analysis of embeddings learned with the multi-domain system\fyTodo{Language and system} For each unit\footnote{In this study, we work with BPE units, not words, meaning that in many cases we observe the variation of use of word parts.
-As we work with a large inventory, many of these units correspond to actual words, and we focus on these in our comments.} in our dictionary, we compute the $k$ nearest neighbours for each domain, \fyTodo{As follows} yielding three lists of nearest neighbours.
-The size of their intersection should then be a sign of a variation of use across domains; conversely, a stable set of neighbours across domains should reflect the stability of word use.
+To check the validity of this hypothesis, we performed the following analysis of embeddings learned with the multi-domain Transformer system for English:French.
+For each unit\footnote{In this study, we work with BPE units, not words, meaning that in many cases we observe the variation of use of word parts.
+As we work with a large inventory, many of these units correspond to actual words, and we focus on these in our comments.} in our dictionary, we compute the $k$ nearest neighbours for each domain, \fyTodo{As follows} yielding $k$ lists of nearest neighbours.
+The size of their intersection should then be a sign of a variation of use across domains; conversely, an identical set of neighbours across domains should reflect the stability of word use.
 

### 2019-05-21 2019/emnlp/sparse.tex
More review in section 4

@@
     \multicolumn{4}{|l|}{$En \rightarrow Fr$ }\\
-    \multicolumn{4}{|l|}{Vocab: En 30,165, Fr: 30,398}\\
+    \multicolumn{4}{|l|}{Vocab\. En: 30,165, Fr: 30,398}\\
     \hline
@@
     \multicolumn{4}{|l|}{$En \rightarrow Ge$}\\
-    \multicolumn{4}{|l|}{ Vocab: En 30,159, Ge: 30,698}\\ \hline
+    \multicolumn{4}{|l|}{ Vocab.\ En:30,159, De: 30,698}\\ \hline
     EMEA  & 1,11M & 1000 & 1000 \\
@@
 For each unit\footnote{In this study, we work with BPE units, not words, meaning that in many cases we observe the variation of use of word parts.
-As we work with a large inventory, many of these units correspond to actual words, and we focus on these in our comments.} in our dictionary, we compute the $k$ nearest neighbours for each domain, \fyTodo{As follows} yielding $k$ lists of nearest neighbours.
+As we work with a large inventory, many of these units correspond to actual words, and we focus on these in our comments.
+We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $d$ is $e(v,i) = M_ge_g(u) + M_de_i(v)$ (cf.\ equation~\eqref{}).\fyTodo{Equation}.
+This process yields $d$ lists of $k$ nearest neighbours.
 The size of their intersection should then be a sign of a variation of use across domains; conversely, an identical set of neighbours across domains should reflect the stability of word use.
+Table~\ref{tab:embeddings} list the 10 units with the larger intersection, and the 10 units with the smaller intersection.
+
+\begin{table}
+  \centering
+  \begin{tabularx}{1.0\linewidth}{c}
+    blab
+  \end{tabularx}
+  \caption{Analyzing the variation of embeddings across domains.}
+  \label{tab:embeddings}
+\end{table}
 

### 2019-05-21 2019/emnlp/sparse.tex
analysis of word embeddings

@@
 
-Our main results are summarized respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems, where we report BLEU scores \citecomputed after detokenization.
+Our main results are summarized respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems, where we report BLEU scores \cite{Papineni02bleu} computed after detokenization.
 \fyTodo{Footnote for missing numbers}.
@@
 The size of their intersection should then be a sign of a variation of use across domains; conversely, an identical set of neighbours across domains should reflect the stability of word use.
-Table~\ref{tab:embeddings} list the 10 units with the larger intersection, and the 10 units with the smaller intersection.
+Table~\ref{tab:embeddings} list the 10 units with the larger intersection, and the 10 units with the smaller intersection (we use $k=10$ and $d=3$)
 
@@
 
-- impairment - mostly renal / hepatic impairment - translates as insuffisance in EMEA, quite different in ECB
-- convenience - mostly credit card convenience in ECB (), mostly for convenience (par commodité in EMEA).
-- turnover in ECB= chiffre d'affaire; EMEA - we observe collocations such as ``bone turnover'' (remodelage or renouvellement osseux') and more generally of cell renewall.
+Let us first consider the complete words in the left column of Table~\ref{tab:embeddings}.
+The case of \emph{impairment} is pretty clear, collocating mostly in EMEA in terms such as ``hepatic impairment'' or ``renal impairment'', and translating into French as \textsl{insuffisance}; in ECB, the collocates of ``impairment'' are quite different, and impairment often occurs in terms such as ``cost subject to impairments'' (French: \emph{coût soumis à des réductions de valeur}).
+Likewise, ``convenience'' seems to have its general meaning (``for convenience'') in EMEA, but appears in ECB in the specific context of ``convenience credit card'' (French \textsl{carte de crédit à remboursement différé}).
+We finally see the same phenomena with ``turnover'', which is consistently translated with its economic meaning (French \textsl{chiffre d'affaire}) in ECB and EPPS, but whose collocates in EMEA ("bone turnover", ``hepatic turnover'') are associated with the idea of the cell renewall process, yielding translations such as \textsl{remodelage osseux} in French.
+Subword units can be analysed in the same ways: ``ums'', for instance, appears in words such as ``gums'', ``serums'', ``vacuums''  in EMEA; in ECB, ``ums'' is mostly the suffix of ``maximums'', ``minimums'', or ``premiums''; EPPS finally contains a more diverse set of ``-ums'' ending words (``stadium'', ``forum'', equilibrium'', etc).
 
-Conversely, same sense
+Let us now consider the list of putative monosemic words (on the right part of Table~\ref{tab:embeddings}), ie.\ words for which the nearest neighbors are the same in all domains, contains mostly words for which we do not expect much variation in translation: adverbs (`` ``)
 
-This only works for sufficiently frequent words.
-
-\begin{table}
-  \centering
-  \begin{tabular}{*{4}{|r|}}
-    \hline
-    Domain & Word & Neighbor & Cosinus distance \\ \hline
-    \multirow{5}{*}{EMEA} & \multirow{5}{*}{joint} 
-    & Joint & 0.371 \\  
-    &  & jointly & 0.545 \\
-    &  & artic- & 0.605 \\
-    &  & joints & 0.607 \\
-	&  & collective & 0.609 \\
-	\hline
-	\multirow{5}{*}{EPPS} & \multirow{5}{*}{joint} 
-	& Joint & 0.365 \\
-	& & jointly & 0.497 \\
-	& & join- & 0.606 \\
-	& & common & 0.607 \\
-	& & collective & 0.626 \\
-	& & artic- & 0.633\\
-	\hline
-	\multirow{5}{*}{ECB} & \multirow{5}{*}{joint}
-	& Joint & 0.423 \\
-	& & jointly & 0.467 \\
-	& & common & 0.601 \\
-	& & together & 0.632 \\
-	& & artic- & 0.663 \\
-	& & coordinated & 0.668 \\
-\end{tabular}
-\caption{Corpora}
-\label{tab:Polysemy}
-\end{table}
+% \begin{table}
+%   \centering
+%   \begin{tabular}{*{4}{|r|}}
+%     \hline
+%     Domain & Word & Neighbor & Cosinus distance \\ \hline
+%     \multirow{5}{*}{EMEA} & \multirow{5}{*}{joint} 
+%     & Joint & 0.371 \\  
+%     &  & jointly & 0.545 \\
+%     &  & artic- & 0.605 \\
+%     &  & joints & 0.607 \\
+% 	&  & collective & 0.609 \\
+% 	\hline
+% 	\multirow{5}{*}{EPPS} & \multirow{5}{*}{joint} 
+% 	& Joint & 0.365 \\
+% 	& & jointly & 0.497 \\
+% 	& & join- & 0.606 \\
+% 	& & common & 0.607 \\
+% 	& & collective & 0.626 \\
+% 	& & artic- & 0.633\\
+% 	\hline
+% 	\multirow{5}{*}{ECB} & \multirow{5}{*}{joint}
+% 	& Joint & 0.423 \\
+% 	& & jointly & 0.467 \\
+% 	& & common & 0.601 \\
+% 	& & together & 0.632 \\
+% 	& & artic- & 0.663 \\
+% 	& & coordinated & 0.668 \\
+% \end{tabular}
+% \caption{Corpora}
+% \label{tab:Polysemy}
+% \end{table}
 

### 2019-05-21 2019/emnlp/sparse.tex
conclusion is there

@@
 
-For these test sets, our proposal (\texttt{LDR}$_{oracle}$) is consistently better that the mix-domain strategy, with gains than range from very large (for EMEA and ECB) to unsignificant (for EPPS in most conditions).
+For these test sets, our proposal (\texttt{LDR}$_{oracle}$) is consistently better that the mix-domain strategy, with gains that range from very large (for EMEA and ECB) to unsignificant (for EPPS in most conditions).
 This means that our architecture is somehow able to compensate for the data unbalance and to raise the performance of the multi-domain system close to the best (fine-tuned) system in each domain.
@@
 The size of their intersection should then be a sign of a variation of use across domains; conversely, an identical set of neighbours across domains should reflect the stability of word use.
-Table~\ref{tab:embeddings} list the 10 units with the larger intersection, and the 10 units with the smaller intersection (we use $k=10$ and $d=3$)
+Table~\ref{tab:embeddings} list the 10 units with the smaller (respectively larger) intersection (we use $k=10$ and $d=3$).
 
@@
   \centering
-  \begin{tabularx}{1.0\linewidth}{c}
-    blab
+  \begin{tabularx}{1.0\linewidth}{c|c}
+    Polysemic ``words'' & Monosemic ``words'' \\ \hline
+     ases (0) &               obtain (10) \\       
+     impairment (1) &     virtually (10) \\    
+     convenience (1) &    represent (10) \\    
+     oring (1) &              safety (10) \\       
+     ums (1) &               defence (10) \\      
+     turnover (1) &         coordinated (10) \\  
+     occurrence (1) &     handling (10) \\     
+     tent (2) &               July (10) \\         
+     ture (2) &               previous (10) \\     
+     mation (2) &           better (10) \\ \hline
   \end{tabularx}
-  \caption{Analyzing the variation of embeddings across domains.}
+  \caption{Analyzing the variation of embeddings across domains. Each word or subword is listed with the size of the intersection (between 0 and 10).}
   \label{tab:embeddings}
@@
 
-Let us first consider the complete words in the left column of Table~\ref{tab:embeddings}.
-The case of \emph{impairment} is pretty clear, collocating mostly in EMEA in terms such as ``hepatic impairment'' or ``renal impairment'', and translating into French as \textsl{insuffisance}; in ECB, the collocates of ``impairment'' are quite different, and impairment often occurs in terms such as ``cost subject to impairments'' (French: \emph{coût soumis à des réductions de valeur}).
-Likewise, ``convenience'' seems to have its general meaning (``for convenience'') in EMEA, but appears in ECB in the specific context of ``convenience credit card'' (French \textsl{carte de crédit à remboursement différé}).
+Let us first consider the full words in the left column of Table~\ref{tab:embeddings}.
+The case of \emph{impairment} is pretty clear, occuring in EMEA mostly in terms such as ``hepatic impairment'' or ``renal impairment'', and translating into French as \textsl{insuffisance}.
+In ECB, its collocate are quite different, and impairment often occurs in terms such as ``cost subject to impairments'' (French: \emph{co\^ut soumis \`a des r\'eductions de valeur}).
+Likewise, ``convenience'' seems to have its general meaning (``for convenience'') in EMEA, but appears in ECB in the specific context of ``convenience credit card'' (French \textsl{carte de cr\' edit \`a remboursement diff\'er\'e}).
 We finally see the same phenomena with ``turnover'', which is consistently translated with its economic meaning (French \textsl{chiffre d'affaire}) in ECB and EPPS, but whose collocates in EMEA ("bone turnover", ``hepatic turnover'') are associated with the idea of the cell renewall process, yielding translations such as \textsl{remodelage osseux} in French.
@@
 
-Let us now consider the list of putative monosemic words (on the right part of Table~\ref{tab:embeddings}), ie.\ words for which the nearest neighbors are the same in all domains, contains mostly words for which we do not expect much variation in translation: adverbs (`` ``)
+Let us now consider the list of putative monosemic words (on the right part of Table~\ref{tab:embeddings}), ie.\ words for which the nearest neighbors are the same in all domains, contains mostly words for which we do not expect much variation in translation: adjective (``previous'', ``better''), adverbs (``virtually``), generic verbs (``handling'', ``coordinated''), etc; and the list of words having a constant set of neighbors also contains prepositions (``at'', ``in''), auxiliary (``been'') etc.
 
@@
 \fyDone{Related work goes last}
-\fyTodo{Compare also to Peng 2017}
+\fyDone{Compare also to Peng 2017}
 
@@
 \section{Conclusions}
-\fyTodo{natural continuations: bayesian version ? linear combination of specialized engines ? go beyond words; learn the projection matrix at the word level; other ?}
+In this paper, we have presented a new technique for multi-domain machine translation, adapting to two common NMT architectures the ``frustratingly easy'' idea of \cite{Daume07frustratingly}.
+Our experiments have shown that for both architecture and for two language pairs, our multi-domain models improve over several baselines of the literature, and that it is robust to noise in domain labels.
+In our future work, we intend to develop this work so as to make the architecture more self-configurable and able to adapt the size of the domain-specific regions depending upon the actual variation of use across domains; we also would like to find ways to make the architecture able to integrate new domains in a dynamic fashion, as this is an important requirement in industrial systems.
+
+% \fyTodo{natural continuations: bayesian version ? linear combination of specialized engines ? go beyond words; learn the projection matrix at the word level; other ?}
 

### 2019-05-21 2019/emnlp/sparse.tex
revise the table, write conclusion

@@
 We study here a different scenario, \emph{multi-domain adaptation} \cite{Sennrich13multidomain,Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
-This allows us to be both data efficient (all data is used to train all domains) and compuationally efficient (we only need to train one domain).
+This allows us to be both data efficient (all data is used to train all domains) and computationally efficient (we only train one system).
 In our setting, we assume that the domain of test documents is known.
@@
 We perform experiments with three domains, two neural architectures and two language pairs and find that our technique can effectively build multi-domain NMT, outperforming several baselines.
-Our contribution is thus twofold:
+Our contribution is thus as follows
 \begin{itemize}
-\item methods to adapt the ideas of \cite{Daume07frustratingly} for two NMT architectures;
-\item experimental results that demonstrate the effectiveness of this technique;
-\item analysis of embeddings
+\item we develop methods to adapt the ideas of \cite{Daume07frustratingly} for two NMT architectures;
+\item we provide experimental evidence that suggest the effectiveness of this technique;
+\item our analysis of embeddings seems to confort our assumption that their variation across domain actually reflects a variation of senses.
 \end{itemize}
-\fyTodo{can we train in random order ? can we get away with catastrophic forgetting ?}
-\fyTodo{how to analyze the embeddings ? how can we test or claim ?}
+\fyDone{can we train in random order ? can we get away with catastrophic forgetting ?}
+\fyDone{how to analyze the embeddings ? how can we test or claim ?}
 
@@
     Corpus & Train & Dev & Test \\ \hline
-    \multicolumn{4}{|l|}{$En \rightarrow Fr$ }\\
-    \multicolumn{4}{|l|}{Vocab\. En: 30,165, Fr: 30,398}\\
+    \multicolumn{4}{|l|}{En $\rightarrow$ Fr }\\
+    \multicolumn{4}{|l|}{Vocabulary size - En: 30,165, Fr: 30,398}\\
     \hline
-    EMEA  & 1,09M & 1000 & 1000 [300]\\
+    EMEA  & 1,09M & 1000 & 1000$_{300}$\\
     ECB    & 0,19M & 1000 & 1000     \\
     EPPS   & 2,0M  & 1000 & 1000  \\ \hline \hline
-    \multicolumn{4}{|l|}{$En \rightarrow Ge$}\\
-    \multicolumn{4}{|l|}{ Vocab.\ En:30,159, De: 30,698}\\ \hline
-    EMEA  & 1,11M & 1000 & 1000 \\
+    \multicolumn{4}{|l|}{En $\rightarrow$ De}\\
+    \multicolumn{4}{|l|}{Vocabulary size -  En:30,159, De: 30,698}\\ \hline
+    EMEA  & 1,11M & 1000 & 1000$_{300}$ \\
     ECB     &  0,11M & 1000 & 1000  \\
@@
 \hline
-$FT_{EMEA}$ & 76.77$_{49.43}$ & 17.16 & 11.99 & 29.58 & 11.64 & 11.10 & avg \\
-$FT_{EPPS}$ & 20.86 & 37.04 & 24.53 & 31.86 & 28.33 & 11.10 & avg \\
-$FT_{ECB}$   & 26.93 & 27.09 & 56.52 & 22.04 & 17.51 & 13.99 & avg \\
+$\mathtt{FT} (EMEA)$ & 76.77$_{49.43}$ & 17.16 & 11.99 & 29.58 & 11.64 & 11.10 & avg \\
+$\mathtt{FT} (EPPS)$  & 20.86 & 37.04 & 24.53 & 31.86 & 28.33 & 11.10 & avg \\
+$\mathtt{FT} (ECB)$   & 26.93 & 27.09 & 56.52 & 22.04 & 17.51 & 13.99 & avg \\
 \hline
@@
 \hline
-$FT_{EMEA}$ & 68.35 & 17.02 & 32.87 & 21.15 & 11.21 & 13.49 & avg\\
-$FT_{EPPS}$ & 36.19 & 26.29 & 40.71 & 19.67 & 16.47 & 19.24 & avg \\
-$FT_{ECB}$   & 24.72 & 18.36 & 74.05 & 11.31  & 10.8 & 11.00 & avg \\
+$\mathtt{FT} (EMEA)$ & 68.35$_{42.97}$ & 17.02 & 32.87 & 21.15 & 11.21 & 13.49 & avg\\
+$\mathtt{FT} (EPPS)$ & 36.19 & 26.29 & 40.71 & 19.67 & 16.47 & 19.24 & avg \\
+$\mathtt{FT} (ECB)$  & 24.72 & 18.36 & 74.05 & 11.31  & 10.8 & 11.00 & avg \\
 \hline
@@
 \hline
-$FT_{EMEA}$     & 72.06 $ $ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & avg \\
-$FT_{EPPS}$      & 35.47 $ $ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & avg \\
-$FT_{ECB}$        & 21.93 $ $ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & avg \\
+$\mathtt{FT} (EMEA)$  & 72.06 $ $ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & avg \\
+$\mathtt{FT} (EPPS)$    & 35.47 $ $ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & avg \\
+$\mathtt{FT} (ECB)$     & 21.93 $ $ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & avg \\
 \hline
@@
 \hline
-$FT_{EMEA}$      & 65.64 $ $ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & avg\\
-$FT_{EPPS}$       & 24.90 $ $ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & avg \\
-$FT_{ECB}$         & 41.8 $ $ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & avg\\
+$\mathtt{FT} (EMEA)$  & 65.64$_{44.71}$ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & avg\\
+$\mathtt{FT} (EPPS)$   & 24.90 $ $ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & avg \\
+$\mathtt{FT} (ECB)$    & 41.8 $ $ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & avg\\
 \hline

### 2019-05-21 2019/emnlp/sparse.tex
add number, shorter introduction

@@
 Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,Bahdanau15learning,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progresses over the past years, and constitute to date the standard for most production engines.
-One issue which repeatedly pops up with MT, be it neural or statistical, is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain.
+The development of MT systems, be they neural or statistical, require very large parallel corpora consisting of millions of sentence pairs, a resource that only exist in very few application domains and language pairs.
 A lot of the recent research effort has thus focused on developping MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Ha16towards,Johnson17google}; or even building systems without any parallel data \cite{Artetxe18unsupervised,Lample18unsupervised}.
-Another important scenario for industrial MT systems is to adapt a neural system trained using parallel data in one domain to the peculiarity of other domains.
+Another important scenario for industrial MT is to adapt a neural system trained using parallel data in one domain to the peculiarity of other domains.
 Domain Adaption (DA) in MT is an old issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied (see the recent survey of \cite{Chu18asurvey} for neural MT).
@@
 Multi-domain adaptation is conceptually close to multilingual MT, or more generally to multi-task learning \cite{Caruana97multitask}, and can be approached in a number of ways.
-We adapt here theideas of \cite{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains.
+We adapt here ideas of \citet{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains.
 Our main hypothesis is that most of the differences between domains take place at the lexical level, due to cross-domain polysemy.
 By contrast, the deeper layers, which arguably model more abstract linguistic phenomena, are made shareable across domains.
-To this end, we make sure to design word embeddings in such that they will contain a generic as well as a domain-specific region.
+To this end, we design word embeddings made of a generic as well as several domain-specific regions.
 We perform experiments with three domains, two neural architectures and two language pairs and find that our technique can effectively build multi-domain NMT, outperforming several baselines.
-Our contribution is thus as follows
+Our contribution is thus as follows:
 \begin{itemize}
-\item we develop methods to adapt the ideas of \cite{Daume07frustratingly} for two NMT architectures;
+\item we implement the ideas of \cite{Daume07frustratingly} for two NMT architectures;
 \item we provide experimental evidence that suggest the effectiveness of this technique;
@@
 To validate our findings, we compared lexicalized domain embedding models with standard models using both attentional Recurrent Neural Networks (RNNs) \cite{Bahdanau15learning} and the recent Transformer architecture of \citet{Vaswani17attention}.
-Our baselines consist of generic models trained with a simple concatenation of all corpora (denoted \texttt{mixed} in the score tables below); models finetuned separately to each domain using (10000, 15000, 5000) iterations over in-domain data (denoted \texttt{+ft(EMEA)}, \texttt{+ft(EPPS)}, \texttt{+ft(ECB)} below); models using domain tags as in \cite{Kobus17domaincontrol} (noted \texttt{DC} below); and finally the multi-domain model of \citet{Zeng18multidomain} (denoted \texttt{WDCMT} below).\fyTodo{Change names in tables}
-
-For all models, we set the size of word embeddings equal to 512; the size of hidden layers is equal to 1024 for RNNs and 512 for Transformer.
-Other important configuration details that were kept fixed throughout are as follows:
+Our baselines consist of generic models trained with a simple concatenation of all corpora (denoted \texttt{mMixed} in the tables below); models tuned separately on each domain using respectively (10000, 15000, 5000) iterations over in-domain data (denoted \texttt{+ft(EMEA)}, \texttt{+ft(EPPS)}, \texttt{+ft(ECB)} below); models using domain tags as in \cite{Kobus17domaincontrol} (\texttt{DC} below); and finally the multi-domain model of \citet{Zeng18multidomain} (\texttt{WDCMT} below).\fyDone{Change names in tables}
+For all models, we set the embeddings size equal to 512; the size of hidden layers is equal to 1024 for RNNs and 512 for Transformer. Other important configuration details are as follows:
 \fyDone{Improve this}
 \begin{itemize}
-\item Our Transformer models use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 components; 
+\item Transformer models use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells; 
 %  size of word embedding: 512; size of hidden layers: 512; size of inner feed forward layer: 2048; Multi-head: 8; number of layers: 6.
-\item Our RNN models use 1 layer on both sides: a bidirectional LSTM encoder and a unidirectional LST decoder with attention;
+\item RNN models use 1 layer on both sides: a bidirectional LSTM encoder and a unidirectional LSTM decoder with attention;
 %  Attention Recurrent Neural Network: Bidirection LSTM encoder, unidirectional LSTM decoder, Luong Attention Mechanism, size of word embedding: 512; size of hidden layers: 1024; number of layers: 1
-\item The domain control systems are exactly as their baseline counterpart (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.\fyDone{Check This!}
+\item The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.\fyDone{Check This!}
   % C: size of domain embedding:2 ; same settings for both architecture: Transformer and Attention Recurrent Neural Network.
-\item The \texttt{WDCMT} system also has one layer on the encoder side (a bidirectional Gated recurrent units (GRU) layer), as well as one layer on the decoder side (a unidirectional conditional GRU layer); the dimension of ``domain'' layers is~300.
+\item The \texttt{WDCMT} system has one bidirectional Gated recurrent units (GRU) layer on the encoder side and one unidirectional conditional GRU layer on the decoder side; the dimension of ``domain'' layers is~300.
 \end{itemize}
@@
 \subsection{Implementing LDR}\fyTodo{Check acronym.}
-\fyTodo{Motivate the split - discuss experimentally embedding size}
+\fyDone{Motivate the split - discuss experimentally embedding size}
 For the implementation of lexicalized domain representations, we split the embedding vector (of size 512) into four regions: 3 are domain specific and 1 is generic, with respective sizes $[8,8,8,488]$, a choice that will be dicussed in Section~\ref{secc:region_size}.
@@
 \texttt{LDR}$_{oracle}$ is also better than Domain Control in three conditions out of four, \texttt{DC} being seemingly a better choice for the RNN than for the Transformer architecture.
-The comparison with the \texttt{WDCMT} model is less favourable for our RNN systems, which are arguably much simpler.\footnote{The direct comparison with WDCMT is difficult as the implementation differs in many ways from our RNN: a different framework, different cell types (LSTMs vs.
-conditional GRUs), etc.} \fyDone{More on this.}
+The comparison with the \texttt{WDCMT} model is less favourable for our RNN systems, which are arguably much simpler.\footnote{The direct comparison with WDCMT is difficult as the implementation differs in many ways from our RNN: a different framework, different cell types,
+  % (LSTMs vs. conditional GRUs),
+  etc.} \fyDone{More on this.}
 As expected, ignoring the domain label yields a drop in performance for the two fallback strategies considered in this study.
@@
 \hline
-$Mixed$ & 67.69$_{47.60}$ & 37.50 & 53.49 & 34.86 & 28.56 & 25.70 & avg \\
+$Mixed$ & 67.69$_{47.60}$ & 37.50 & 53.49 & 34.86 & 28.56 & 25.70 & 41.3\\
 \hline
-$\mathtt{FT} (EMEA)$ & 76.77$_{49.43}$ & 17.16 & 11.99 & 29.58 & 11.64 & 11.10 & avg \\
-$\mathtt{FT} (EPPS)$  & 20.86 & 37.04 & 24.53 & 31.86 & 28.33 & 11.10 & avg \\
-$\mathtt{FT} (ECB)$   & 26.93 & 27.09 & 56.52 & 22.04 & 17.51 & 13.99 & avg \\
+$\mathtt{FT} (EMEA)$ & 76.77$_{49.43}$ & 17.16 & 11.99 & 29.58 & 11.64 & 11.10 & 26.37 \\
+$\mathtt{FT} (EPPS)$  & 20.86 & 37.04 & 24.53 & 31.86 & 28.33 & 11.10 & 25.62 \\
+$\mathtt{FT} (ECB)$   & 26.93 & 27.09 & 56.52 & 22.04 & 17.51 & 13.99 &  27.35\\
 \hline
-$DC$ & 67.87$_{45.42}$ & 37.31 & 54.14 & 33.63 & 26.52 & 24.81 & avg\\
+$DC$ & 67.87$_{45.42}$ & 37.31 & 54.14 & 33.63 & 26.52 & 24.81 & 40.71\\
 \hline
-$LDR_{oracle}$   & 74.26$_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 & avg \\
-$LDR^{v2}_{oracle}$   & 74.95$_{49.38}$ & 37.35 & 55.91 & 33.57 & 27.59 & 25.27 & avg \\
-$LDR_{generic}$ & 73.97$_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & avg \\
-$LDR_{pred}$ & 74.29 $_{49.84}$ & 37.73 & 54.01 & 33.34 & 28.32 & 26 & avg \\
-$LDR_{wrong}$ & 72.95$_{49.78}$ & 37.62 & 53.35 & - & - & - & avg\\
+$LDR_{oracle}$   & 74.26$_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 & 42.28\\
+$LDR^{v2}_{oracle}$   & 74.95$_{49.38}$ & 37.35 & 55.91 & 33.57 & 27.59 & 25.27 & 42.44\\
+$LDR_{generic}$ & 73.97$_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & 42.24 \\
+$LDR_{pred}$ & 74.29 $_{49.84}$ & 37.73 & 54.01 & 33.34 & 28.32 & 26 & 42.28\\
+$LDR_{wrong}$ & 72.95$_{49.78}$ & 37.62 & 53.35 & - & - & - & - \\
   \hline
@@
 \hline
-$Mixed$ & 64.57$_{42.99}$ & 26.47 & 68.67 & 21.77 & 17.04 & 18.85 & avg \\
+$Mixed$ & 64.57$_{42.99}$ & 26.47 & 68.67 & 21.77 & 17.04 & 18.85 &36.23 \\
 \hline
-$\mathtt{FT} (EMEA)$ & 68.35$_{42.97}$ & 17.02 & 32.87 & 21.15 & 11.21 & 13.49 & avg\\
-$\mathtt{FT} (EPPS)$ & 36.19 & 26.29 & 40.71 & 19.67 & 16.47 & 19.24 & avg \\
-$\mathtt{FT} (ECB)$  & 24.72 & 18.36 & 74.05 & 11.31  & 10.8 & 11.00 & avg \\
+$\mathtt{FT} (EMEA)$ & 68.35$_{42.97}$ & 17.02 & 32.87 & 21.15 & 11.21 & 13.49 & 27.35\\
+$\mathtt{FT} (EPPS)$ & 36.19 & 26.29 & 40.71 & 19.67 & 16.47 & 19.24 & 31.42\\
+$\mathtt{FT} (ECB)$  & 24.72 & 18.36 & 74.05 & 11.31  & 10.8 & 11.00 & 25\\
 \hline
-$DC$ & 63.48$_{42.98}$ & 26.27 & 66.95 & 21.45 & 16.45 & 18.78 & avg \\
+$DC$ & 63.48$_{42.98}$ & 26.27 & 66.95 & 21.45 & 16.45 & 18.78 & 35.56\\
 \hline
-$LDR_{oracle}$ & 70.9$_{46.12}$ & 26.30 & 68.9 & 21.42 & 16.57 & 18.82 & avg \\
-$LDR^{v2}_{oracle}$ & 71.31$_{45.23}$ & 25.98 & 73.74 & 21.6 & 16.37 & 19.2 & avg \\
-$LDR_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & avg\\
-$LDR_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & 21.45 & 16.61 & 19.07 & avg \\
-$LDR_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & - & - & - & avg \\
+$LDR_{oracle}$ & 70.9$_{46.12}$ & 26.30 & 68.9 & 21.42 & 16.57 & 18.82 & 37.15\\
+$LDR^{v2}_{oracle}$ & 71.31$_{45.23}$ & 25.98 & 73.74 & 21.6 & 16.37 & 19.2 & 38.03\\
+$LDR_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & 37.14\\
+$LDR_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & 21.45 & 16.61 & 19.07 & 32.2 \\
+$LDR_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & - & - & - & - \\
 \hline
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for the Transformer architecture \label{tab:results-trsf}}
+\caption{BPE-detokenized BLEU scores for the Transformer systems \label{tab:results-trsf}}
 \end{table*}
@@
 \hline
-$Mixed$               & 65.42$_{45.11}$ & 34.7 & 51.38 & 29.27 & 22.21 & 21.63 & avg \\
+$Mixed$               & 65.42$_{45.11}$ & 34.7 & 51.38 & 29.27 & 22.21 & 21.63 &37.43 \\
 \hline
-$\mathtt{FT} (EMEA)$  & 72.06 $ $ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & avg \\
-$\mathtt{FT} (EPPS)$    & 35.47 $ $ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & avg \\
-$\mathtt{FT} (ECB)$     & 21.93 $ $ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & avg \\
+$\mathtt{FT} (EMEA)$  & 72.06$_{47.33}$ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & 20 \\
+$\mathtt{FT} (EPPS)$    & 35.47 $ $ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & 29.87 \\
+$\mathtt{FT} (ECB)$     & 21.93 $ $ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & 22.23 \\
 \hline
-$DC$                   & 68.26 $_{43.76}$ & 35.13 & 50.09 & 27.86 & 20.95 & 21.16 & avg \\
+$DC$                   & 68.26 $_{43.76}$ & 35.13 & 50.09 & 27.86 & 20.95 & 21.16 & 37.24\\
  \hline
-$WDCMT$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & avg \\
+$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & 38.97 \\
 \hline
-$LDR_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & avg\\
-$LDR_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & avg\\
-$LDR_{pred}$        & 72.76 $_{46.35}$ & 35.1 & 50.38 & 30.02 & 23.28 & 21.32 & avg\\
-$LDR_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 &  - & - & - & avg \\
+$LDR_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & 38.57\\
+$LDR^{v2}_{oracle}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & 32.6\\
+$LDR_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & 38.16\\
+$LDR_{pred}$        & 72.76 $_{46.35}$ & 35.1 & 50.38 & 30.02 & 23.28 & 21.32 & 38.81\\
+$LDR_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 &  - & - & - & - \\
 \hline
@@
 \hline
-$Mixed$               & 57.37 $_{37.94}$ & 23.1 & 63.54 & 17.39 & 11.51 & 14.29 & avg\\
+$Mixed$               & 57.37 $_{37.94}$ & 23.1 & 63.54 & 17.39 & 11.51 & 14.29 & 31.2\\
 \hline
-$\mathtt{FT} (EMEA)$  & 65.64$_{44.71}$ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & avg\\
-$\mathtt{FT} (EPPS)$   & 24.90 $ $ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & avg \\
-$\mathtt{FT} (ECB)$    & 41.8 $ $ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & avg\\
+$\mathtt{FT} (EMEA)$  & 65.64$_{44.71}$ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & 20.53\\
+$\mathtt{FT} (EPPS)$   & 24.90 $ $ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & 18.89 \\
+$\mathtt{FT} (ECB)$    & 41.8 $ $ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & 25.89\\
 \hline
-$DC$                    & 62.53 $_{39.25}$ & 23.74 & 65.71 & 16.88 & 11.15 & 14.97 & avg\\
+$DC$                      & 62.53 $_{39.25}$ & 23.74 & 65.71 & 16.88 & 11.15 & 14.97 & 32.50\\
 \hline
-WDCMT               & 64.05 $_{38.54}$ & 22.82 & 67.00 & 15.94 & 11.45 & 13.66 & avg\\
+$\mathtt{WDCMT}$ & 64.05 $_{38.54}$ & 22.82 & 67.00 & 15.94 & 11.45 & 13.66 & 32.49\\
 \hline
-$LDR_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & avg\\
-$LDR^{v2}_{oracle}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & avg\\
-$LDR_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & avg\\
-$LDR_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64 & 15.94 & 11.32 & 14.27 & avg\\
+$LDR_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & 32.2\\
+$LDR^{v2}_{oracle}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & 32.6\\
+$LDR_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & 32.1\\
+$LDR_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64 & 15.94 & 11.32 & 14.27 & 31.87\\
 $LDR_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & - & - & - & avg \\
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for the RNN architecture \label{tab:results-rnn}}
+\caption{BPE-detokenized BLEU scores for the RNN systems\label{tab:results-rnn}}
 \end{table*}
@@
 
-\section*{Acknowledgments}
-\fyTodo{Homegenize refs - urls, addresses, etc}
+%\section*{Acknowledgments}
+% \fyTodo{Homegenize refs - urls, addresses, etc}
 \bibliography{emnlp-ijcnlp-2019}

### 2019-05-21 2019/emnlp/sparse.tex
rewrite algorithm

@@
 Multi-domain adaptation is conceptually close to multilingual MT, or more generally to multi-task learning \cite{Caruana97multitask}, and can be approached in a number of ways.
-We adapt here ideas of \citet{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains.
-Our main hypothesis is that most of the differences between domains take place at the lexical level, due to cross-domain polysemy.
+We adapt here ideas of \citet{Daume07frustratingly} to neural MT.
+% in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains.
+Our main hypothesis is that domains mostly differ at the lexical level, due to cross-domain polysemy, which motivates domain specific embeddings.
 By contrast, the deeper layers, which arguably model more abstract linguistic phenomena, are made shareable across domains.
-To this end, we design word embeddings made of a generic as well as several domain-specific regions.
-We perform experiments with three domains, two neural architectures and two language pairs and find that our technique can effectively build multi-domain NMT, outperforming several baselines.
+To this end, we design word embeddings containing a generic and several domain-specific regions.
+We experiment with three domains, two neural architectures and two language pairs and find that our technique yields effective multi-domain NMTs, outperforming several baselines.
 Our contribution is thus as follows:
@@
 
-\section{Lexicalize domain representations\label{sec:lexicalized_embeddings}}
+\section{Lexicalized domain representations\label{sec:lexicalized_embeddings}}
 \fyDone{Use meaningful titles throughout}
@@
 \begin{algorithmic}[1]
-\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for $d$ domains, Batch size $B$, Optimization algorithm $\operatorname{Opt}$}
+\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for $d$ domains, Batch size $B$}%, Optimization algorithm $\operatorname{Opt}$}
 \REPEAT 
-\STATE{Randomly pick $i \in [1,..,d]$ w.r.t the multinomial distribution $[\frac{|C_i|}{\displaystyle{\mathop{\sum}_{i\in [1,..,d]}|C_i|}}]$.}
+\STATE{Randomly pick $i \in [1,..,d]$ w.r.t the multinomial distribution $[\frac{|C_i|}{\sum_{i\in [1,..,d]}|C_i|}]$.}
 \STATE{Randomly pick $B$ sentences from $C_i$.}
-\STATE{Activate only generic region to create generic embedding batch, denoted $W_g$.}
-\STATE{Pass the generic embedding batch to a dense layer.}
+\STATE{Activate only generic region to create generic batch, denoted $W_g$.}
+% \STATE{Pass the generic embedding batch to a dense layer.}
 \STATE{Compute gradient of $\theta_s$, $\frac{\partial L}{\partial \theta_s}$ using $W_g$.}
-\STATE{Activate domain region and generic region to create domain embedding batch $W_i$}
-\STATE{Pass $W_i$ to a dense layer}
+\STATE{Activate domain-specific and generic regions to create domain-specific batch $W_i$}
+% \STATE{Pass $W_i$ to a dense layer}
 \STATE{Compute gradient of domain-specific parameters $\theta_i$, $\frac{\partial L}{\partial \theta_i}$ using $W_d$.}
-\STATE{Apply learning rules using computed gradients: 
-$$ \theta_s = \theta_s - \operatorname{Opt}(\frac{\partial L}{\partial \theta_s}(W_g))$$ 
-$$ \theta_i = \theta_i - \operatorname{Opt} (\frac{\partial L}{\partial \theta_i}(W_i))$$}
+\STATE{Update parameters $\theta_s$ using $\frac{\partial L}{\partial \theta_s}(W_g)$ and $\theta_i$ using $\frac{\partial L}{\partial \theta_i}(W_i)$}
+% $$ \theta_s = \theta_s - \operatorname{Opt}(\frac{\partial L}{\partial \theta_s}(W_g))$$ 
+% $$ \theta_i = \theta_i - \operatorname{Opt} (\frac{\partial L}{\partial \theta_i}(W_i))$$}
 \UNTIL{convergence}
@@
 In our experiments, this means that sentences from the Europarl domain will be selected more frequently that the two other domains.
-In an attempt to build a more balanced training procedure, we have also experimented with an alternative selection procedure, where $i$ is selected according to the multinomial distribution $[\frac{\sqrt{|C_i|}}{\displaystyle{\mathop{\sum}_{i\in [1,..,d]}\sqrt{|C_i|}}}]$.
-\fyTodo{Refs on this ? or contrast ?}
+In an attempt to build a more balanced training sample, we also consider an alternative selection procedure, where $i$ is selected according to the distribution $[\frac{\sqrt{|C_i|}}{\sum_{i\in [1,..,d]}\sqrt{|C_i|}}]$.
+% \fyTodo{Refs on this ? or contrast ?}
 
@@
 Its precision on a development set are higher than 95\%.}
-\fyTodo{Explain how} Note that this decrease is however quite limited (and hardly significant for \texttt{LDR}$_{pred}$), showing that our architecture is quite robust to noisy labels.
+\fyDone{Explain how} Note that this decrease is however quite limited (and hardly significant for \texttt{LDR}$_{pred}$), showing that our architecture is quite robust to noisy labels.
 Even in the worst case scenario where all domain tags are intently wrong (\texttt{LDR}$_{wrong}$), we see that the generic part still ensures a satisfying level of performance.
@@
 By analogy with the proposal of \citet{Johnson17google} for multilingual NMT, \citet{Kobus17domaincontrol} and \citet{Chu18multilingual} separately propose to extend the representation of the source text with a domain tag.
-Our model also modifies input representations, but allows each source word to have its own domain-specific representation, thereby improving training of the shared parts of the network.\fyTodo{Talk about target embeddings somewhere.}
+Our model also modifies input representations, but allows each source word to have a domain-specific representation, thereby improving training of the shared parts of the network.\fyDone{Talk about target embeddings somewhere.}
 

### 2019-05-21 2019/emnlp/sparse.tex
intro

@@
 The development of MT systems, be they neural or statistical, require very large parallel corpora consisting of millions of sentence pairs, a resource that only exist in very few application domains and language pairs.
-A lot of the recent research effort has thus focused on developping MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Ha16towards,Johnson17google}; or even building systems without any parallel data \cite{Artetxe18unsupervised,Lample18unsupervised}.
+A lot of the recent research effort has thus focused on developping MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Ha16towards,Johnson17google}; %or even building systems without any parallel data \cite{Artetxe18unsupervised,Lample18unsupervised}.
 Another important scenario for industrial MT is to adapt a neural system trained using parallel data in one domain to the peculiarity of other domains.
-Domain Adaption (DA) in MT is an old issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied (see the recent survey of \cite{Chu18asurvey} for neural MT).
-The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain.
+Domain Adaption (DA) in MT is an old issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied.
+See the recent survey of \cite{Chu18asurvey} for neural MT.
+The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tune the parameters of a system trained on a large amount of texts in a source domain.
 We study here a different scenario, \emph{multi-domain adaptation} \cite{Sennrich13multidomain,Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
 This allows us to be both data efficient (all data is used to train all domains) and computationally efficient (we only train one system).
-In our setting, we assume that the domain of test documents is known.
+%In our setting, we assume that the domain of test documents is known.
 
@@
 Our contribution is thus as follows:
-\begin{itemize}
-\item we implement the ideas of \cite{Daume07frustratingly} for two NMT architectures;
-\item we provide experimental evidence that suggest the effectiveness of this technique;
-\item our analysis of embeddings seems to confort our assumption that their variation across domain actually reflects a variation of senses.
-\end{itemize}
+%\begin{itemize}
+%\item 
+we implement the ideas of \cite{Daume07frustratingly} for two NMT architectures;
+%\item 
+we provide experimental evidence that suggest the effectiveness of this technique;
+%\item 
+our analysis of embeddings seems to confort our assumption that their variation across domain actually reflects a variation of senses.
+%\end{itemize}
 \fyDone{can we train in random order ? can we get away with catastrophic forgetting ?}
@@
 \center
-    \includegraphics[width=0.52\textwidth]{embeddings}
+    \includegraphics[width=0.48\textwidth]{embeddings}
     \caption{Lexicalized domain embeddings.

### 2019-05-21 2019/emnlp/sparse.tex
sec3

@@
 
-We experiment with two language pairs (English-French, English-German) and data originating from three domains, corresponding to texts from three European institutions: from the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}.
+We experiment with two language pairs (English-French, English-German) and data originating from three domains, corresponding to texts from three European institutions: the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}.
 We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).\footnote{The EMEA dataset as distributed on the OPUS site contains many duplicates.
@@
 To validate our findings, we compared lexicalized domain embedding models with standard models using both attentional Recurrent Neural Networks (RNNs) \cite{Bahdanau15learning} and the recent Transformer architecture of \citet{Vaswani17attention}.
-Our baselines consist of generic models trained with a simple concatenation of all corpora (denoted \texttt{mMixed} in the tables below); models tuned separately on each domain using respectively (10000, 15000, 5000) iterations over in-domain data (denoted \texttt{+ft(EMEA)}, \texttt{+ft(EPPS)}, \texttt{+ft(ECB)} below); models using domain tags as in \cite{Kobus17domaincontrol} (\texttt{DC} below); and finally the multi-domain model of \citet{Zeng18multidomain} (\texttt{WDCMT} below).\fyDone{Change names in tables}
+Our baselines consist of generic models trained with a simple concatenation of all corpora (denoted \texttt{Mixed} in the tables below); models tuned separately on each domain using respectively (10000, 15000, 5000) iterations over in-domain data (denoted \texttt{ft(EMEA)}, \texttt{ft(EPPS)}, \texttt{ft(ECB)} below); models using domain tags as in \cite{Kobus17domaincontrol} (\texttt{DC} below); and finally the multi-domain model of \citet{Zeng18multidomain} (\texttt{WDCMT} below).\fyDone{Change names in tables}
 For all models, we set the embeddings size equal to 512; the size of hidden layers is equal to 1024 for RNNs and 512 for Transformer. Other important configuration details are as follows:
 \fyDone{Improve this}
-\begin{itemize}
-\item Transformer models use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells; 
+%\begin{itemize}
+%\item 
+Transformer models use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells; 
 %  size of word embedding: 512; size of hidden layers: 512; size of inner feed forward layer: 2048; Multi-head: 8; number of layers: 6.
-\item RNN models use 1 layer on both sides: a bidirectional LSTM encoder and a unidirectional LSTM decoder with attention;
+%\item 
+RNN models use 1 layer on both sides: a bidirectional LSTM encoder and a unidirectional LSTM decoder with attention;
 %  Attention Recurrent Neural Network: Bidirection LSTM encoder, unidirectional LSTM decoder, Luong Attention Mechanism, size of word embedding: 512; size of hidden layers: 1024; number of layers: 1
-\item The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.\fyDone{Check This!}
+%\item 
+The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.\fyDone{Check This!}
   % C: size of domain embedding:2 ; same settings for both architecture: Transformer and Attention Recurrent Neural Network.
-\item The \texttt{WDCMT} system has one bidirectional Gated recurrent units (GRU) layer on the encoder side and one unidirectional conditional GRU layer on the decoder side; the dimension of ``domain'' layers is~300.
-\end{itemize}
-To train NMT systems, we use Adam, with parameters $\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$ for RNNs; with parameters $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} for Transformer (using $warmup\_steps=4000$).
-For all systems, we use a batch size of~128 and a dropout rate of 0.1 for all layers.
-Our implementation is from OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein2017OpenNMT}, except for WDCMT where we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyDone{batch size}
+%\item 
+The \texttt{WDCMT} system has one bidirectional Gated recurrent units (GRU) layer on the encoder side and one unidirectional conditional GRU layer on the decoder side; the dimension of ``domain'' layers is~300.
+%\end{itemize}
+
+To train NMT systems, we use Adam, with parameters $\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$ for RNNs; with parameters $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} for Transformer ($warmup\_steps=4000$).
+In all cases, we use a batch size of~128 and a dropout rate of 0.1 for all layers.
+All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein2017OpenNMT} except for WDCMT for which we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyDone{batch size}
 
 \subsection{Implementing LDR}\fyTodo{Check acronym.}
+
 \fyDone{Motivate the split - discuss experimentally embedding size}
-For the implementation of lexicalized domain representations, we split the embedding vector (of size 512) into four regions: 3 are domain specific and 1 is generic, with respective sizes $[8,8,8,488]$, a choice that will be dicussed in Section~\ref{secc:region_size}.
-If a sentence originates from domain $i$, the domain specific regions for all domains $j \neq i$ will be zeroed out; while the other regions are activated (cf.
+In order to implement lexicalised domain representations (LDR), we split the embedding vector into four regions: 3 are domain specific and 1 is generic, with sizes $[8,8,8,488]$ respectively.
+%, a choice that will be discussed in Section~\ref{secc:region_size}.
+If a sentence originates from domain $i$, the domain specific regions for all domains $j \neq i$ will be zeroed out; while the rest of regions are activated (cf.
 Figure~\ref{fig:network}).
 We then use a dense layer of size~512 to fuse the region for the active domain and the ``generic'' region.
-Training is formalized in algorithm~\ref{alg:multidomain}.
+Training is formalised in algorithm~\ref{alg:multidomain}.
 
@@
 
-In training, we use same optimization algorithm as for the baselines.
-Note that each iteration of algorithm \ref{alg:multidomain} uses 2~batches: a ``generic'' batch updating only the generic region; and ``domain-specific'' batch updating just the domain-specific parameters.
-This is because we wish to simultaneously minimize two losses:
+%In training, we use same optimization algorithm as for the baselines. 
+%The same optimisation algorithm is always used.
+Note that each iteration of algorithm \ref{alg:multidomain} uses 2~batches: a ``generic'' batch updating only the generic region; and a ``domain-specific'' batch updating just the domain-specific parameters.
+Thus, we minimise two losses:
+%This is because we wish to simultaneously minimize two losses:
 \begin{align*}
@@
 Therefore, we select batches from each corpora with frequencies proportional to its size.
-The batch selection procedure (step~2 of algorithm~\ref{alg:multidomain}) ensures that each domain is represented according to its frequency in the training data.
+The batch selection procedure (step~2 of algorithm~\ref{alg:multidomain}) ensures that the number of examples seen in training of each domain follows the distribution of examples seen in the training data.
+%each domain is represented according to its frequency in the training data. 
 In our experiments, this means that sentences from the Europarl domain will be selected more frequently that the two other domains.
-In an attempt to build a more balanced training sample, we also consider an alternative selection procedure, where $i$ is selected according to the distribution $[\frac{\sqrt{|C_i|}}{\sum_{i\in [1,..,d]}\sqrt{|C_i|}}]$.
+In an attempt to build a more balanced training sample, we also consider an alternative selection procedure, where $i$ is selected according to  distribution $[\frac{\sqrt{|C_i|}}{\sum_{i\in [1,..,d]}\sqrt{|C_i|}}]$.
 % \fyTodo{Refs on this ? or contrast ?}
@@
 This means that our architecture is somehow able to compensate for the data unbalance and to raise the performance of the multi-domain system close to the best (fine-tuned) system in each domain.
-We even observe rare cases where the \texttt{LDR}$_{oracle}$ system outperforms fine-tuning (eg.\ Transformer en:ge in the EMEA domain).
+We even observe rare cases where the \texttt{LDR}$_{oracle}$ system outperforms fine-tuning (eg.\ Transformer en:de in the EMEA domain).
 \texttt{LDR}$_{oracle}$ is also better than Domain Control in three conditions out of four, \texttt{DC} being seemingly a better choice for the RNN than for the Transformer architecture.

### 2019-05-21 2019/emnlp/sparse.tex
start fixing part 2

--- 2019-05-21 2019/emnlp/sparse.tex
@@
 
-\subsection{Problem statement \label{ssec:statement}}
+\subsection{Multi-domain machine translation \label{ssec:statement}}
+
+Multi-domain machine translation is formalized as follows: we assume that our observations correspond to  domain-tagged sentence pairs $[(x,y),i]$, with $x$ in the source language, $y$ in the target language and $i$ a domain tag in $[1\dots d]$.
+We further assume a two-stage sampling process: first select a domain $i$ according to $p(i)$, then select a sentence pair according to a domain specific distribution $D_i$.
+Our objective is to find a set of parameters $\{\theta_1 \dots \theta_d \} \in \mathbb{R}^D \times \mathbb{R}^D \dots \times \mathbb{R}^D$ minimizing:
+\begin{equation} \label{eq:loss}
+\begin{split}
+\sum_{i \in [1..d]} p(i) E_{(x,y) \sim D_{i}} [-log(p_{\theta}(y|x,i))]
+\end{split}
+\end{equation}
+The training data for domain $i$ is denoted $C_i$.
+
+A straightforward solution is to process each domain separately, computing the value $\theta_i^*$ that minimizes the empirical loss on the training corpus $C_i$.
+This strategy is only effective if we have sufficient training data for each domain; when this is not the case, some estimates $\theta_i^*$ may be far from their optimal value.
+\fyDone{Risk}.
+The alternative we consider here constraints each parameter $\theta_i$ to be made of two parts: one ($\theta_s \in R^{D_g}$) is shared across all domains, while the second part $\theta_i \in \mathbb{R}^{D_i}$ is only used in domain $i$.
+The parameter set is now much more constrained, yet we expect that tying parameters across domains will yield better estimates for $\theta_s$ due to a larger training corpus.
+In this setting, the optimization program defined by equation~\eqref{eq:loss} can no longer be performed separately for each training corpus.
+
+%%% 
 \fyTodo{Use math for notation, label equations and such}
@@
 $LDR_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & 38.57\\
-$LDR^{v2}_{oracle}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & 32.6\\
+$LDR^{v2}_{oracle}$   & 71.7$_{46.41}$ & 34.24& 52.37 & 29.21 & 22.7& 20.69 & 38.48\\
 $LDR_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & 38.16\\

### 2019-05-21 2019/emnlp/sparse.tex
sec3

--- 2019-05-21 2019/emnlp/sparse.tex
@@
 
-To check the validity of this hypothesis, we performed the following analysis of embeddings learned with the multi-domain Transformer system for English:French.
+To check the validity of this hypothesis, we performed the following analysis of embeddings learned with the multi-domain Transformer system for en:fr.
 For each unit\footnote{In this study, we work with BPE units, not words, meaning that in many cases we observe the variation of use of word parts.

### 2019-05-21 2019/emnlp/sparse.tex
sec3

@@
 
-To check the validity of this hypothesis, we performed the following analysis of embeddings learned with the multi-domain Transformer system for en:fr.
+To check the validity of this hypothesis, we performed the following analysis of embeddings learned with the multi-domain Transformer system for English:French.
 For each unit\footnote{In this study, we work with BPE units, not words, meaning that in many cases we observe the variation of use of word parts.

### 2019-05-21 2019/emnlp/sparse.tex
final

@@
 This strategy is only effective if we have sufficient training data for each domain; when this is not the case, some estimates $\theta_i^*$ may be far from their optimal value.
-\fyDone{Risk}.
-The alternative we consider here constraints each parameter $\theta_i$ to be made of two parts: one ($\theta_s \in R^{D_g}$) is shared across all domains, while the second part $\theta_i \in \mathbb{R}^{D_i}$ is only used in domain $i$.
+\fyDone{Risk}
+The alternative we consider here constraints each parameter $\theta_i$ to be made of two parts: $\theta_i = [\theta_s; \theta'_i]$ ($\theta_s \in R^{D_g}$) is shared across all domains, while the second part $\theta'_i \in \mathbb{R}^{D_i}$ is only used in domain $i$.
 The parameter set is now much more constrained, yet we expect that tying parameters across domains will yield better estimates for $\theta_s$ due to a larger training corpus.
@@
 
-%%% 
-\fyTodo{Use math for notation, label equations and such}
-\begin{definition}{Domain}
-\label{def:domain}
-A domain is probabilistic distribution of translation $(x,y)$ in a given corpora. Given corpora $C$, we denote $D(C)$ the domain of $C$.
-\end{definition}
-\fyTodo{We need more: the source, the target, etc}
-In machine translation, the loss function for training is:
-\begin{equation}\label{eq:1}
-\begin{split}
-E_{(x,y) \sim p_{rel}}[-log(p_{\theta}(y|x))] &= \\
-= E_{x \sim p_{rel}(x)}[KL(p_{rel}(y|x) & \mid p_{\theta}(y|x))] + \\
-+ E_{x \sim p_{rel}} &[H(p_{rel}(y|x))]
-\end{split}
-\end{equation}.
+\subsection{Lexicalized domain embeddings \label{ssec:lde}}
 
-Where $p_{\theta}(y|x) = p(y|x,\theta)$ \\
-In this equation, the domain is $p_{rel}(x,y)$.
+To actually implement this idea for NMT, we need to define the subset of parameters that will be common to all domain.
+We make the assumption that domain specificities can be confined to the lexical level, and define $\theta_s$ to contain all the network parameters, except for a subpart of the word embeddings.
+For each word $v$, the embedding vector $e(v)$ is thus decomposed as $e(v) = [e_g(v); e_1(v); \dots; e_d(v)]$, where $e_g(v)$ stores the generic lexical embedding, while $e_i(v)$ stores the subpart that is specific to domain $i$.
+In our NMT architectures, the actual embedding layer composes these vectors linearly to generate the word embedding for domain $k$ according to:
+\begin{align}
+  \tilde{e}_k(v) =& M_g e_g(v) + \sum_{i \in [1,..,d]} M_i * e_i(v) * \delta(i=k) \nonumber \\
+   & = M [e_g(v); e_1'(v,k) \dots; e_d'(v,k)], \label{eq:embedding}
+\end{align}
+where $\delta()$ is the Kronecker function, $M$ is the matrix made of blocks $M_g, M_1 \dots, M_d$, and $e_i'(v,k)$ is the masked embedding: $e_i'(v,k)= e_i(v) * \delta(i=k)$.
+Making sure that the actual embedding do not contain any zero is important for the Transformer model, since the lexical representations are then summed to the positional encoding, which would undo the effect of domain masking, and propagate a gradient even to regions that should not be modified.
+With our design, we make sure that during backprogation, the matrix $M$ receives gradient $0$ at regions corresponding to deactivated regions in the word embedding.
+Those regions are also masked in forward step, thus do not interfere the training on the domain to which they are not assigned (see Figure~\ref{fig:network}).
+
+\begin{figure}
+  \center
+  \includegraphics[width=0.48\textwidth]{embeddings}
+  \caption{Lexicalized domain embeddings.
+When processing a sample from domain $2$, we only activate the corresponding parameter region ($\theta_2$) in the input embeddings; the remaining domain-specific parts are zeroed out, and do not receive any update.
+The generic part is always active, and is updated irrespective of the input domain.}
+  \label{fig:network}
+\end{figure}
+
+Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers the computation in equation~\ref{eq:embedding}.
+In our experiment, we consider both the attentional RNN architectures of \citet{Bahdanau15learning} and for Transformer architectures of \citet{Vaswani17attention}.
+
+% Let $\tilde{x}_{t} = M * e_{t} = M_g e_g(t) + \displaystyle{\mathop{\sum}_{i \in [1,..,d]} M_i * e_i(t)}$ be the projection of word embedding where $e_{t}$ is word embedding of token $x_{t}$, M is projection matrix, $M_j \forall j \in [1...d]$ is row block corresponding to domain region $j$ (Transformer \cite{Vaswani17attention}, recurrent \cite{Bahdanau15learning})\\
+% Assume the batch belongs to domain $i$, then $e_j(t) = 0$ $\forall j \neq i$,$\forall t$. 
+% \begin{equation}
+% \begin{split}
+% \frac{\partial L(x,y)}{\partial M_j} &= \sum_{t}\frac{\partial L}{\partial \tilde{x}_t}(x_t) * \frac{\partial \tilde{x}_t}{\partial M_j} \\
+% 								&= \sum_{t} \frac{\partial L}{\partial \tilde{x}_t}(x_t) * e_j(t)\\
+% 									&=0
+% \end{split} 
+% \forall j \neq i
+% \end{equation}
+
+% The complex structure of State-of-the-art architecture Transformer \cite{Vaswani17attention} and popular recurrent network \cite{Bahdanau15learning} makes the implementation of sparse embedding non-obvious.
+Indeed, Transformer involve positional embedding and residual connections which are not homogeneous for domain-regions.
+Therefore, in our first implementation, we used an dense layer to fuse domain-region and generic region before applying positional embedding and residual connection.
+This implementation limits the finetuning effect at only word level.\\
+% Furthermore, because word embeddings are involved in three parts of standard encoder-decoder structure, which are input of encoder, input of decoder and projection before softmax, we have 7 possibilities to apply sparse embedding.
+However, due to high memory cost, we only want to apply sparse embedding on the source side.
+\fyTodo{Explain this.
+Future work ?}
+
+% %%% 
+% \fyTodo{Use math for notation, label equations and such}
+% \begin{definition}{Domain}
+% \label{def:domain}
+% A domain is probabilistic distribution of translation $(x,y)$ in a given corpora. Given corpora $C$, we denote $D(C)$ the domain of $C$.
+% \end{definition}
+% \fyTodo{We need more: the source, the target, etc}
+% In machine translation, the loss function for training is:
+% \begin{equation}\label{eq:1}
+% \begin{split}
+% E_{(x,y) \sim p_{rel}}[-log(p_{\theta}(y|x))] &= \\
+% = E_{x \sim p_{rel}(x)}[KL(p_{rel}(y|x) & \mid p_{\theta}(y|x))] + \\
+% + E_{x \sim p_{rel}} &[H(p_{rel}(y|x))]
+% \end{split}
+% \end{equation}.
+
+% Where $p_{\theta}(y|x) = p(y|x,\theta)$ \\
+% In this equation, the domain is $p_{rel}(x,y)$.
 In minimizing given objective \ref{eq:1}, one aims to approximate the objective domain $p_{rel}$.
 In inference, provided the model was trained on certain training data, the more likely test set is generated from domain of training data, the better performance.\\
-In general, a generic model is trained first with a mix of several corpora which belongs to different topics (e.g medical, banking, adminstration) whose the domain $p_{rel}$ is considered as a general domain, noted $p_{gen}$.
+% In general, a generic model is trained first with a mix of several corpora which belongs to different topics (e.g medical, banking, adminstration) whose the domain $p_{rel}$ is considered as a general domain, noted $p_{gen}$.
 Being trained with a general mixed corpora, the generic model is not usually good enough for translating by topic.
@@
 In effect, the finetuned model usually shows lower performance in translating sentences, which are less related to the finetuning topic, than generic model.
-In order to conduct the model translating topic by topic, we might refer to conditional probability conditioned by topic $p(y|x,d)$.
+% In order to conduct the model translating topic by topic, we might refer to conditional probability conditioned by topic $p(y|x,d)$.
 Given a set of training corpora, we assume that there are a limited d engendering topic, i.e, there are d corpora $C_i$ $i \in [1,..,d]$ representing d topical domains $D(C_i)$ respectively.
@@
 From now, we formalize the multi-domain problem for machine translation in following equation
-\begin{equation} \label{eq:2}
-\begin{split}
-\theta^* = \argmin_{\theta} \displaystyle{\mathop{\sum}_{i \in [1..d]}} E_{(x,y) \sim D(C_{i})}[-log(p_{\theta}(y|x,i))]
-\end{split}
-\end{equation}
-It is obvious that
-\begin{equation} \label{eq:3}
-\begin{split}
-\argmin_{\theta} \displaystyle{\mathop{\sum}_{i \in [1..d]}} E_{(x,y) \sim D(C_{i})} &[-log(p_{\theta}(y|x,i))] \\
-\geq \displaystyle{\mathop{\sum}_{i \in [1..d]}} \argmin_{\theta} E_{(x,y) \sim D(C_{i})} &[-log(p_{\theta}(y|x,i))]
-\end{split}
-\end{equation}
-A trivial solution for the equation \ref{eq:2} is $\theta^* = (\theta^*_{0},...,\theta^*_{d})$ and $p_{(\theta_{0},...,\theta_d)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times C(D_i) \rightarrow [0,1]: f_{\theta_i}(y,x) = p_{(\theta_{0},...,\theta_d)}(y|x,i)$ where $\Omega_i$ is set of all possible values of $\theta_i$.
+% \begin{equation} \label{eq:2}
+% \begin{split}
+% \theta^* = \argmin_{\theta} \displaystyle{\mathop{\sum}_{i \in [1..d]}} E_{(x,y) \sim D(C_{i})}[-log(p_{\theta}(y|x,i))]
+% \end{split}
+% \end{equation}
+% It is obvious that
+% \begin{equation} \label{eq:3}
+% \begin{split}
+% \argmin_{\theta} \displaystyle{\mathop{\sum}_{i \in [1..d]}} E_{(x,y) \sim D(C_{i})} &[-log(p_{\theta}(y|x,i))] \\
+% \geq \displaystyle{\mathop{\sum}_{i \in [1..d]}} \argmin_{\theta} E_{(x,y) \sim D(C_{i})} &[-log(p_{\theta}(y|x,i))]
+% \end{split}
+% \end{equation}
+% A trivial solution for the equation \ref{eq:2} is $\theta^* = (\theta^*_{0},...,\theta^*_{d})$ and $p_{(\theta_{0},...,\theta_d)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times C(D_i) \rightarrow [0,1]: f_{\theta_i}(y,x) = p_{(\theta_{0},...,\theta_d)}(y|x,i)$ where $\Omega_i$ is set of all possible values of $\theta_i$.
 \\
-The trivial solution is not rather than ensemble of $d$ models specialized for one domain.
+% The trivial solution is not rather than ensemble of $d$ models specialized for one domain.
 However, we could reduce the seperation by using shared paramaters denoted $\theta_s$ then reformulate the model as following: \\
-$\theta^* = (\theta^*_{0},...,\theta^*_{d}, \theta_s^*)$ and $p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times \Omega_s \times C(D_i) \rightarrow [0,1]: f_{\theta_i, \theta_s}(y,x) = p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ \\ $p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x)$ is independent of $\theta_{j}$ $\forall j$ i.e, $\exists \bar{f}:\Omega_s \times C \rightarrow [0,1]: \bar{f}_{\theta_s}(y,x) = p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x)$
-where $\Omega_i$ is set of all possible values of $\theta_i$ and $\Omega_s$ is set of all possible values of $\theta_s$. The optimization problem will become
-\begin{equation}
-\displaystyle{\mathop{\sum}_{i \in [1..d]}} \argmin_{\theta_{0},...,\theta_{d}, \theta_s} E_{(x,y) \sim D(C_{i})} [-log(p_{\theta_s,\theta_i}(y|x,i))]
-\end{equation}
-However, the couple $\theta_s^*$ and $\theta_i^*$ are hardly optimal $\forall i \in [1..d]$.\\
-Despite the fact that $\theta^* = (\theta^*_{0},...,\theta^*_{d}, \theta_s^*)$ is hardly achieved by backpropagation, we could attempt to optimize $\theta_s$:
-\begin{equation}
-\theta^*_{s} = \displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))]
-\end{equation}
+% $\theta^* = (\theta^*_{0},...,\theta^*_{d}, \theta_s^*)$ and $p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times \Omega_s \times C(D_i) \rightarrow [0,1]: f_{\theta_i, \theta_s}(y,x) = p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ \\ $p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x)$ is independent of $\theta_{j}$ $\forall j$ i.e, $\exists \bar{f}:\Omega_s \times C \rightarrow [0,1]: \bar{f}_{\theta_s}(y,x) = p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x)$
+% where $\Omega_i$ is set of all possible values of $\theta_i$ and $\Omega_s$ is set of all possible values of $\theta_s$. The optimization problem will become
+% \begin{equation}
+% \displaystyle{\mathop{\sum}_{i \in [1..d]}} \argmin_{\theta_{0},...,\theta_{d}, \theta_s} E_{(x,y) \sim D(C_{i})} [-log(p_{\theta_s,\theta_i}(y|x,i))]
+% \end{equation}
+% However, the couple $\theta_s^*$ and $\theta_i^*$ are hardly optimal $\forall i \in [1..d]$.\\
+% Despite the fact that $\theta^* = (\theta^*_{0},...,\theta^*_{d}, \theta_s^*)$ is hardly achieved by backpropagation, we could attempt to optimize $\theta_s$:
+% \begin{equation}
+% \theta^*_{s} = \displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))]
+% \end{equation}
 
-then optimize each $\theta_i$ (as additional factor used for fine-tuning the model to the domain $D(C_{i})$) the perplexity with respect to true distribution of domain
-\begin{equation}
-\theta^*_{i} = \displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
-\end{equation} 
-Our first approach to multi-domain machine translation is to build such trivial solution by assigning different regions of word embedding vector to set of domain paramters $\theta_i$ and set of shared parameters $\theta_s$ while keep sharing other parameters between domains.
-
-\subsection{Lexicalized domain embeddings \label{ssec:lde}}
-
-This section explains a new design for word embeddings used in a multi-domain model.
-As mentioned above, we separate parameter set into $D+1$ parameter sets correspond to D domains and one shared paramater set.
-Our first attempt is to split standard vector word embedding into $D+1$ regions named $\theta_i$ for $i \in [1,..,D]$ and $\theta_s$.\fyTodo{nullified not vanishing}
-\begin{figure}
-\center
-    \includegraphics[width=0.48\textwidth]{embeddings}
-    \caption{Lexicalized domain embeddings.
-When processing a sample from domain $2$, we only activate the corresponding parameter region ($\theta_2$) in the input embeddings; the remaining domain-specific parts are zeroed out, and do not receive any update.
-The generic part is always active, and is updated irrespective of the input domain.}
-    \label{fig:network}
-  \end{figure}
-\fyTodo{Explain vanishing ?}
-Figure~\ref{fig:network} presents our proposed design.
-During backprogation, the \fyTodo{fix notation}{projection matrix $M$} receives gradient 0 at regions corresponding to deactivated regions in the word embedding.
-Those regions are also masked in forward step, thus do not interfere the training on the domain to which they are not assigned.
-
-Let $\tilde{x}_{t} = M * e_{t} = M_g e_g(t) + \displaystyle{\mathop{\sum}_{i \in [1,..,d]} M_i * e_i(t)}$ be the projection of word embedding where $e_{t}$ is word embedding of token $x_{t}$, M is projection matrix, $M_j \forall j \in [1...d]$ is row block corresponding to domain region $j$ (Transformer \cite{Vaswani17attention}, recurrent \cite{Bahdanau15learning})\\
-Assume the batch belongs to domain $i$, then $e_j(t) = 0$ $\forall j \neq i$,$\forall t$. 
-\begin{equation}
-\begin{split}
-\frac{\partial L(x,y)}{\partial M_j} &= \sum_{t}\frac{\partial L}{\partial \tilde{x}_t}(x_t) * \frac{\partial \tilde{x}_t}{\partial M_j} \\
-								&= \sum_{t} \frac{\partial L}{\partial \tilde{x}_t}(x_t) * e_j(t)\\
-									&=0
-\end{split} 
-\forall j \neq i
-\end{equation}
-
-The complex structure of State-of-the-art architecture Transformer \cite{Vaswani17attention} and popular recurrent network \cite{Bahdanau15learning} makes the implementation of sparse embedding non-obvious.
-Indeed, Transformer involve positional embedding and residual connections which are not homogeneous for domain-regions.
-Therefore, in our first implementation, we used an dense layer to fuse domain-region and generic region before applying positional embedding and residual connection.
-This implementation limits the finetuning effect at only word level.\\
-Furthermore, because word embeddings are involved in three parts of standard encoder-decoder structure, which are input of encoder, input of decoder and projection before softmax, we have 7 possibilities to apply sparse embedding.
-However, due to high memory cost, we only want to apply sparse embedding on the source side.
-\fyTodo{Explain this.
-Future work ?}
+% then optimize each $\theta_i$ (as additional factor used for fine-tuning the model to the domain $D(C_{i})$) the perplexity with respect to true distribution of domain
+% \begin{equation}
+% \theta^*_{i} = \displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
+% \end{equation} 
+% Our first approach to multi-domain machine translation is to build such trivial solution by assigning different regions of word embedding vector to set of domain paramters $\theta_i$ and set of shared parameters $\theta_s$ while keep sharing other parameters between domains.
 
@@
 
-\subsection{Implementing LDR}\fyTodo{Check acronym.}
+\subsection{Implementing Lexicalized Domain Representations}\fyDone{Check acronym.}
 
 \fyDone{Motivate the split - discuss experimentally embedding size}
-In order to implement lexicalised domain representations (LDR), we split the embedding vector into four regions: 3 are domain specific and 1 is generic, with sizes $[8,8,8,488]$ respectively.
+In order to implement lexicalised domain representations (henceforth LDR), we split the embedding vector into four regions: 3 are domain specific and 1 is generic, with sizes $[8,8,8,488]$ respectively.
 %, a choice that will be discussed in Section~\ref{secc:region_size}.
@@
 
-Our main results are summarized respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems, where we report BLEU scores \cite{Papineni02bleu} computed after detokenization.
-\fyTodo{Footnote for missing numbers}.
+Our main results are summarized respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems, where we report BLEU scores \cite{Papineni02bleu} computed after detokenization.\footnote{As explained above, we consistently report two numbers when testing with EMEA, except for the fine-tuning scenarios when tuning on ECB and EPPS.} % \fyTodo{Footnote for missing numbers}.
 
@@
 One of our main assumptions is that the difference between domains can be confined at the lexical level, warranting our decision to specialize lexical representations for each domain, while the remaining part of the network is shared across domains.
-Linguistically, this assumption relates to the ``one domain, one sense'' assumption \fyTodo{Refs}, which expresses the fact that in many cases, polysemy corresponds to variation in use across domain.
-In its weaker form, it allows us to assume that all occurrences of a given form in a given domain correspond to the same sense and share the same embedding; the same form occuring in different domains is allowed to have one distinct embedding per domain, which may help capture polysemy and lexical ambiguity in translation.
+Linguistically, this assumption relates to the classical ``one discourse, one sense'' \cite{Yarowsky95onediscourse}, and corresponds to the fact that in many cases, polysemy corresponds to variation of use across domain.
+In its weaker form, it allows us to assume that all occurrences of a given form in a given domain correspond to the same sense and share the same representation; the same form occuring in different domains is allowed to have one distinct embedding per domain, which may help capture polysemy and lexical ambiguity in translation.
 

### 2019-05-21 2019/emnlp/sparse.tex
remove comment and equation

@@
 Note that each iteration of algorithm \ref{alg:multidomain} uses 2~batches: a ``generic'' batch updating only the generic region; and a ``domain-specific'' batch updating just the domain-specific parameters.
-Thus, we minimise two losses:
+% Thus, we minimise two losses:
 %This is because we wish to simultaneously minimize two losses:
-\begin{align*}
-\theta^*_{s}=&\displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))] \\ 
- \theta^*_{i}=&\displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
-\end{align*}
+% \begin{align*}
+% \theta^*_{s}=&\displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))] \\ 
+%  \theta^*_{i}=&\displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
+% \end{align*}
 
@@
 % \fyTodo{Easily accomodate more domains}
-\fyTodo{Given the ways embeddings are computed, why not add more domains, and test robustness agains data presentation order ?}
+\fyToDone{Given the ways embeddings are computed, why not add more domains, and test robustness agains data presentation order ?}
 
@@
 One of our main assumptions is that the difference between domains can be confined at the lexical level, warranting our decision to specialize lexical representations for each domain, while the remaining part of the network is shared across domains.
-Linguistically, this assumption relates to the classical ``one discourse, one sense'' \cite{Yarowsky95onediscourse}, and corresponds to the fact that in many cases, polysemy corresponds to variation of use across domain.
+Linguistically, this assumption relates to the classical ``one sense per collocation'' \cite{Yarowsky93onesense} and corresponds to the fact that in many cases, polysemy corresponds to variation of use across domain.
 In its weaker form, it allows us to assume that all occurrences of a given form in a given domain correspond to the same sense and share the same representation; the same form occuring in different domains is allowed to have one distinct embedding per domain, which may help capture polysemy and lexical ambiguity in translation.

### 2019-05-21 2019/emnlp/sparse.tex
checked all

@@
 Even in the worst case scenario where all domain tags are intently wrong (\texttt{LDR}$_{wrong}$), we see that the generic part still ensures a satisfying level of performance.
-A last contrast is with \texttt{LDR}$_{oracle}^{0.5}$ where we change the distribution of training sentence to increase the number of ECB samples: as a result, we see a small decrease for EMEA and EPPS, and a large boost for ECB.
+A last contrast is with \texttt{LDR}$_{oracle}^{0.5}$ where we change the distribution of training sentences to decrease the weight of EPPS data and increase the number of ECB samples.
+As a result, we see a small decrease for EMEA and EPPS, and a large boost for ECB.
 This shows that our technique can be used in conjunction to other well known strategies for performing domain adaptation.
@@
 \hline
-$DC$ & 67.87$_{45.42}$ & 37.31 & 54.14 & 33.63 & 26.52 & 24.81 & 40.71\\
+$\mathtt{DC}$ & 67.87$_{45.42}$ & 37.31 & 54.14 & 33.63 & 26.52 & 24.81 & 40.71\\
 \hline
-$LDR_{oracle}$   & 74.26$_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 & 42.28\\
-$LDR^{v2}_{oracle}$   & 74.95$_{49.38}$ & 37.35 & 55.91 & 33.57 & 27.59 & 25.27 & 42.44\\
-$LDR_{generic}$ & 73.97$_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & 42.24 \\
-$LDR_{pred}$ & 74.29 $_{49.84}$ & 37.73 & 54.01 & 33.34 & 28.32 & 26 & 42.28\\
-$LDR_{wrong}$ & 72.95$_{49.78}$ & 37.62 & 53.35 & - & - & - & - \\
+$\mathtt{LDR}_{oracle}$   & 74.26$_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 & 42.28\\
+$\mathtt{LDR}_{oracle}^{0.5}$   & 74.95$_{49.38}$ & 37.35 & 55.91 & 33.57 & 27.59 & 25.27 & 42.44\\
+$\mathtt{LDR}_{generic}$ & 73.97$_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & 42.24 \\
+$\mathtt{LDR}_{pred}$ & 74.29 $_{49.84}$ & 37.73 & 54.01 & 33.34 & 28.32 & 26 & 42.28\\
+$\mathtt{LDR}_{wrong}$ & 72.95$_{49.78}$ & 37.62 & 53.35 & - & - & - & - \\
   \hline
@@
 \hline
-$DC$ & 63.48$_{42.98}$ & 26.27 & 66.95 & 21.45 & 16.45 & 18.78 & 35.56\\
+$\mathtt{DC}$ & 63.48$_{42.98}$ & 26.27 & 66.95 & 21.45 & 16.45 & 18.78 & 35.56\\
 \hline
-$LDR_{oracle}$ & 70.9$_{46.12}$ & 26.30 & 68.9 & 21.42 & 16.57 & 18.82 & 37.15\\
-$LDR^{v2}_{oracle}$ & 71.31$_{45.23}$ & 25.98 & 73.74 & 21.6 & 16.37 & 19.2 & 38.03\\
-$LDR_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & 37.14\\
-$LDR_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & 21.45 & 16.61 & 19.07 & 32.2 \\
-$LDR_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & - & - & - & - \\
+$\mathtt{LDR}_{oracle}$ & 70.9$_{46.12}$ & 26.30 & 68.9 & 21.42 & 16.57 & 18.82 & 37.15\\
+$\mathtt{LDR}_{oracle}^{0.5}$ & 71.31$_{45.23}$ & 25.98 & 73.74 & 21.6 & 16.37 & 19.2 & 38.03\\
+$\mathtt{LDR}_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & 37.14\\
+$\mathtt{LDR}_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & 21.45 & 16.61 & 19.07 & 32.2 \\
+$\mathtt{LDR}_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & - & - & - & - \\
 \hline
@@
 \hline
-$DC$                   & 68.26 $_{43.76}$ & 35.13 & 50.09 & 27.86 & 20.95 & 21.16 & 37.24\\
+$\mathtt{DC}$                   & 68.26 $_{43.76}$ & 35.13 & 50.09 & 27.86 & 20.95 & 21.16 & 37.24\\
  \hline
@@
 \hline
-$LDR_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & 38.57\\
-$LDR^{v2}_{oracle}$   & 71.7$_{46.41}$ & 34.24& 52.37 & 29.21 & 22.7& 20.69 & 38.48\\
-$LDR_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & 38.16\\
-$LDR_{pred}$        & 72.76 $_{46.35}$ & 35.1 & 50.38 & 30.02 & 23.28 & 21.32 & 38.81\\
-$LDR_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 &  - & - & - & - \\
+$\mathtt{LDR}_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & 38.57\\
+$\mathtt{LDR}_{oracle}^{0.5}$   & 71.7$_{46.41}$ & 34.24& 52.37 & 29.21 & 22.7& 20.69 & 38.48\\
+$\mathtt{LDR}_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & 38.16\\
+$\mathtt{LDR}_{pred}$        & 72.76 $_{46.35}$ & 35.1 & 50.38 & 30.02 & 23.28 & 21.32 & 38.81\\
+$\mathtt{LDR}_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 &  - & - & - & - \\
 \hline
@@
 \hline
-$DC$                      & 62.53 $_{39.25}$ & 23.74 & 65.71 & 16.88 & 11.15 & 14.97 & 32.50\\
+$\mathtt{DC}$                      & 62.53 $_{39.25}$ & 23.74 & 65.71 & 16.88 & 11.15 & 14.97 & 32.50\\
 \hline
@@
 \hline
-$LDR_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & 32.2\\
-$LDR^{v2}_{oracle}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & 32.6\\
-$LDR_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & 32.1\\
-$LDR_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64 & 15.94 & 11.32 & 14.27 & 31.87\\
-$LDR_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & - & - & - & avg \\
+$\mathtt{LDR}_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & 32.2\\
+$\mathtt{LDR}_{oracle}^{0.5}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & 32.6\\
+$\mathtt{LDR}_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & 32.1\\
+$\mathtt{LDR}_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64 & 15.94 & 11.32 & 14.27 & 31.87\\
+$\mathtt{LDR}_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & - & - & - & - \\
 \hline
@@
 % \fyTodo{Easily accomodate more domains}
-\fyToDone{Given the ways embeddings are computed, why not add more domains, and test robustness agains data presentation order ?}
+\fyDone{Given the ways embeddings are computed, why not add more domains, and test robustness agains data presentation order ?}
 

### 2019-05-21 2019/emnlp/sparse.tex
equation missing reference

@@
 As we work with a large inventory, many of these units correspond to actual words, and we focus on these in our comments.
-We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $d$ is $e(v,i) = M_ge_g(u) + M_de_i(v)$ (cf.\ equation~\eqref{}).\fyTodo{Equation}.
+We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $d$ is $e(v,i) = M_ge_g(u) + M_de_i(v)$ (cf.\ equation~\eqref{eq:embedding}).
 This process yields $d$ lists of $k$ nearest neighbours.

### 2019-05-21 2019/emnlp/sparse.tex
correct domain d to domain i line 471

@@
 As we work with a large inventory, many of these units correspond to actual words, and we focus on these in our comments.
-We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $d$ is $e(v,i) = M_ge_g(u) + M_de_i(v)$ (cf.\ equation~\eqref{eq:embedding}).
+We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $i$ is $e(v,i) = M_ge_g(u) + M_de_i(v)$ (cf.\ equation~\eqref{eq:embedding}).
 This process yields $d$ lists of $k$ nearest neighbours.

### 2019-06-14 2019/emnlp/sparse.tex
one typo

@@
 In this paper, we have presented a new technique for multi-domain machine translation, adapting to two common NMT architectures the ``frustratingly easy'' idea of \cite{Daume07frustratingly}.
-Our experiments have shown that for both architecture and for two language pairs, our multi-domain models improve over several baselines of the literature, and that it is robust to noise in domain labels.
+Our experiments have shown that for both architectures and for two language pairs, our multi-domain models improve over several baselines of the literature, and that it is robust to noise in domain labels.
 In our future work, we intend to develop this work so as to make the architecture more self-configurable and able to adapt the size of the domain-specific regions depending upon the actual variation of use across domains; we also would like to find ways to make the architecture able to integrate new domains in a dynamic fashion, as this is an important requirement in industrial systems.

### 2019-06-19 2019/emnlp/sparse.tex
correction ligne 471

--- 2019-05-21 2019/emnlp/sparse.tex
@@
 As we work with a large inventory, many of these units correspond to actual words, and we focus on these in our comments.
-We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $i$ is $e(v,i) = M_ge_g(u) + M_de_i(v)$ (cf.\ equation~\eqref{eq:embedding}).
+We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $i$ is $e(v,i) = M_ge_g(v) + M_ie_i(v)$ (cf.\ equation~\eqref{eq:embedding}).
 This process yields $d$ lists of $k$ nearest neighbours.

### 2019-06-28 2019/emnlp/sparse.tex
fix some typos

@@
 %\item 
-we implement the ideas of \cite{Daume07frustratingly} for two NMT architectures;
+we adapt the ideas of \cite{Daume07frustratingly} for two NMT architectures;
 %\item 
@@
 %\item 
-our analysis of embeddings seems to confort our assumption that their variation across domain actually reflects a variation of senses.
+our analysis of embeddings conforts our assumption that their variation across domain actually reflects a variation of senses.
 %\end{itemize}
@@
 The alternative we consider here constraints each parameter $\theta_i$ to be made of two parts: $\theta_i = [\theta_s; \theta'_i]$ ($\theta_s \in R^{D_g}$) is shared across all domains, while the second part $\theta'_i \in \mathbb{R}^{D_i}$ is only used in domain $i$.
-The parameter set is now much more constrained, yet we expect that tying parameters across domains will yield better estimates for $\theta_s$ due to a larger training corpus.
+The parameter set is now much more constrained, yet we expect that partly tying parameters across domains will yield better estimates for $\theta_s$ due to a larger training corpus.
 In this setting, the optimization program defined by equation~\eqref{eq:loss} can no longer be performed separately for each training corpus.
@@
 Making sure that the actual embedding do not contain any zero is important for the Transformer model, since the lexical representations are then summed to the positional encoding, which would undo the effect of domain masking, and propagate a gradient even to regions that should not be modified.
-With our design, we make sure that during backprogation, the matrix $M$ receives gradient $0$ at regions corresponding to deactivated regions in the word embedding.
+With our design, we make sure that during backprogation, the matrix $M$ receives gradient $0$ at regions corresponding to deactivated regions in the word embeddings.
 Those regions are also masked in forward step, thus do not interfere the training on the domain to which they are not assigned (see Figure~\ref{fig:network}).
@@
 
-Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers the computation in equation~\ref{eq:embedding}.
-In our experiment, we consider both the attentional RNN architectures of \citet{Bahdanau15learning} and for Transformer architectures of \citet{Vaswani17attention}.
+Our architecture is then readily compatible with any NMT framework, where we simply replace standard embedding layers the computation in equation~\eqref{eq:embedding}.
+In our experiments, we consider both the attentional RNN architectures of \citet{Bahdanau15learning} and for Transformer architectures of \citet{Vaswani17attention}.
 
@@
 \STATE{Randomly pick $B$ sentences from $C_i$.}
-\STATE{Activate only generic region to create generic batch, denoted $W_g$.}
+\STATE{Activate only the generic region to create generic batch, denoted $W_g$.}
 % \STATE{Pass the generic embedding batch to a dense layer.}

### 2019-07-02 2019/emnlp/sparse.tex
gru cond results

--- 2019-06-19 2019/emnlp/sparse.tex
@@
 
-We experiment with two language pairs (English-French, English-German) and data originating from three domains, corresponding to texts from three European institutions: the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}.
+We experiment with two language pairs (English-French, English-German) and data originating from three domains, corresponding to texts from three European institutions: the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann09news}.
 We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).\footnote{The EMEA dataset as distributed on the OPUS site contains many duplicates.
@@
 In all cases, we use a batch size of~128 and a dropout rate of 0.1 for all layers.
-All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein2017OpenNMT} except for WDCMT for which we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyDone{batch size}
+All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt} except for WDCMT for which we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyDone{batch size}
 
@@
  \hline
-$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & 38.97 \\
+$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & 38.97 (39.52) \\
 \hline
 $\mathtt{LDR}_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & 38.57\\
+$\mathtt{LDR_{condgru}}_{pred}$   & 71.7 $_{46.21}$ & 35.09 & 51.22 & 31.75 & 24.25 & 22.63 & 39.44\\
 $\mathtt{LDR}_{oracle}^{0.5}$   & 71.7$_{46.41}$ & 34.24& 52.37 & 29.21 & 22.7& 20.69 & 38.48\\
@@
 As we work with a large inventory, many of these units correspond to actual words, and we focus on these in our comments.
-We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $i$ is $e(v,i) = M_ge_g(v) + M_ie_i(v)$ (cf.\ equation~\eqref{eq:embedding}).
+We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $i$ is $e(v,i) = M_ge_g(u) + M_de_i(v)$ (cf.\ equation~\eqref{eq:embedding}).
 This process yields $d$ lists of $k$ nearest neighbours.
@@
 The most common adaptation scenario uses (mostly) out-of-domain data in training, while testing on a low-resource in-domain set of texts.
-In this setting, \emph{data-based approaches} aim to bias the distribution of the train (out-of-domain) data towards matching that of the target domain, using data selection techniques \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}, or generating adapted pseudo-parallel data through back-translation  \cite{Utiyama2003measure,Wang2014neural,Wang2016connecting,Sennrich2016Mono}.
-\emph{Model-centric approaches} build domain-adapted models by combining (eg.\ with mixture weights) multiple data sources or multiple systems \cite{Foster07mixture,Wang17instance}, or by biasing the training objective towards the desired domain using in-domain adaptation data \cite{Luong2015SNMT,Freitag2016FastDA,Chen17costweighting}.
-Another approach worth mentioning integrates domain information (for instance a domain language model) in the decoding algorithm \cite{Gulcehre2016monolingual}.
+In this setting, \emph{data-based approaches} aim to bias the distribution of the train (out-of-domain) data towards matching that of the target domain, using data selection techniques \cite{Moore10selection,Axelrod11domain,Duh13selection}, or generating adapted pseudo-parallel data through back-translation  \cite{Utiyama03measure,Wang14neural,Wang16connecting,Sennrich16improving}.
+\emph{Model-centric approaches} build domain-adapted models by combining (eg.\ with mixture weights) multiple data sources or multiple systems \cite{Foster07mixture,Wang17instance}, or by biasing the training objective towards the desired domain using in-domain adaptation data \cite{Luong15stanford,Freitag16fast,Chen17cost}.
+Another approach worth mentioning integrates domain information (for instance a domain language model) in the decoding algorithm \cite{Gulcehre16monolingual}.
 

### 2019-07-10 2019/emnlp/sparse.tex
our answers

@@
 %\item 
-we implement the ideas of \cite{Daume07frustratingly} for two NMT architectures;
+we adapt and implement the ideas of \cite{Daume07frustratingly} for two NMT architectures;
 %\item 
-we provide experimental evidence that suggest the effectiveness of this technique;
+we provide experimental evidence that show the effectiveness of this technique;
 %\item 
-our analysis of embeddings seems to confort our assumption that their variation across domain actually reflects a variation of senses.
+our analysis of embeddings conforts our assumption that their variation across domains actually reflects a variation of senses.
 %\end{itemize}
@@
 \end{align}
-where $\delta()$ is the Kronecker function, $M$ is the matrix made of blocks $M_g, M_1 \dots, M_d$, and $e_i'(v,k)$ is the masked embedding: $e_i'(v,k)= e_i(v) * \delta(i=k)$.
+where $\delta()$ is the indicator function, $M$ is the matrix made of blocks $M_g, M_1 \dots, M_d$, and $e_i'(v,k)$ is the masked embedding: $e_i'(v,k)= e_i(v) * \delta(i=k)$.
 Making sure that the actual embedding do not contain any zero is important for the Transformer model, since the lexical representations are then summed to the positional encoding, which would undo the effect of domain masking, and propagate a gradient even to regions that should not be modified.
@@
 
-For these test sets, our proposal (\texttt{LDR}$_{oracle}$) is consistently better that the mix-domain strategy, with gains that range from very large (for EMEA and ECB) to unsignificant (for EPPS in most conditions).
+For these test sets, our proposal (\texttt{LDR}$_{oracle}$) is consistently better than the mix-domain strategy, with gains that range from very large (for EMEA and ECB) to unsignificant (for EPPS in most conditions).
 This means that our architecture is somehow able to compensate for the data unbalance and to raise the performance of the multi-domain system close to the best (fine-tuned) system in each domain.
@@
 The first, \texttt{LDR}$_{generic}$, only uses the generic part of the embeddings.
-The second, \texttt{LDR}$_{pred}$, resorts to automatically predicted domain labels\footnote{Our domain classifier uses a bi-LSTM RNN encoder, followed by a simple softmax layer.
+The second, \texttt{LDR}$_{pred}$, resorts to automatically predicted domain labels.\footnote{Our domain classifier uses a bi-LSTM RNN encoder, followed by a simple softmax layer.
 Its precision on a development set are higher than 95\%.}
-\fyDone{Explain how} Note that this decrease is however quite limited (and hardly significant for \texttt{LDR}$_{pred}$), showing that our architecture is quite robust to noisy labels.
-Even in the worst case scenario where all domain tags are intently wrong (\texttt{LDR}$_{wrong}$), we see that the generic part still ensures a satisfying level of performance.
+% \fyDone{Explain how}
+Note that this decrease is however quite limited (and hardly significant for \texttt{LDR}$_{pred}$), showing that our architecture is quite robust to noisy labels.
+Even in the worst case scenario where all domain tags are intentionally wrong (\texttt{LDR}$_{wrong}$), we see that the generic part still ensures a satisfying level of performance.
 A last contrast is with \texttt{LDR}$_{oracle}^{0.5}$ where we change the distribution of training sentences to decrease the weight of EPPS data and increase the number of ECB samples.
@@
 In this setting, \emph{data-based approaches} aim to bias the distribution of the train (out-of-domain) data towards matching that of the target domain, using data selection techniques \cite{Moore10selection,Axelrod11domain,Duh13selection}, or generating adapted pseudo-parallel data through back-translation  \cite{Utiyama03measure,Wang14neural,Wang16connecting,Sennrich16improving}.
-\emph{Model-centric approaches} build domain-adapted models by combining (eg.\ with mixture weights) multiple data sources or multiple systems \cite{Foster07mixture,Wang17instance}, or by biasing the training objective towards the desired domain using in-domain adaptation data \cite{Luong15stanford,Freitag16fast,Chen17cost}.
-Another approach worth mentioning integrates domain information (for instance a domain language model) in the decoding algorithm \cite{Gulcehre16monolingual}.
+\emph{Model-centric approaches} build domain-adapted models by combining (eg.\ with mixture weights) multiple data sources or multiple systems \cite{Foster07mixture,Wang17instance}, or by biasing the training objective towards the desired domain using in-domain adaptation data \cite{Luong15stanford,Freitag16fast,Chen17costweighting}.
+Another approach worth mentioning integrates domain information (for instance a domain language model) in the decoding algorithm \cite{Gulcehre17onintegrating}.
 

### 2019-07-11 2019/iwslt/iwslt2018-submission.tex
iwslt18 submission

--- 2019-07-10 2019/emnlp/sparse.tex
@@
-%
-% File naacl2019.tex
-%
-%% Based on the style files for ACL 2018 and NAACL 2018, which were
-%% Based on the style files for ACL-2015, with some improvements
-%%  taken from the NAACL-2016 style
-%% Based on the style files for ACL-2014, which were, in turn,
-%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
-%% EACL-2009, IJCNLP-2008...
-%% Based on the style files for EACL 2006 by 
-%%e.agirre@ehu.es or Sergi.Balari@uab.es
-%% and that of ACL 08 by Joakim Nivre and Noah Smith
-
-\documentclass[11pt,a4paper]{article}
-\usepackage[hyperref]{emnlp-ijcnlp-2019}
-\usepackage{times}
-\usepackage{latexsym}
-
-\usepackage{url}
-
-%\aclfinalcopy % Uncomment this line for the final submission
-
-%\setlength\titlebox{5cm}
-% You can expand the titlebox if you need extra space
-% to show all the authors. Please do not make the titlebox
-% smaller than 5cm (the original size); we will check this
-% in the camera-ready version and ask you to change it back.
-
-\newcommand\BibTeX{B{\sc ib}\TeX}
-\newcommand\confname{EMNLP-IJCNLP 2019}
-\newcommand\conforg{SIGDAT}
+\documentclass[a4paper]{article}
+\usepackage{iwslt18,amssymb,amsmath,epsfig}
+\setcounter{page}{1}
+\sloppy		% better line breaks
+%\ninept
+%SM below a registered trademark definition
+\def\reg{{\rm\ooalign{\hfil
+     \raise.07ex\hbox{\scriptsize R}\hfil\crcr\mathhexbox20D}}}
 \usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
@@
 
-%\setlength\titlebox{5cm}
-% You can expand the titlebox if you need extra space
-% to show all the authors. Please do not make the titlebox
-% smaller than 5cm (the original size); we will check this
-% in the camera-ready version and ask you to change it back.
+%% \newcommand{\reg}{\textsuperscript{\textcircled{\textsc r}}}
 
@@
 
-\author{First Author \\% Minh Quang Pham\\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain} \\\And
-  Second Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain} \\}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Please make sure to keep technical paper submissions anonymous  !
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\name{Firstname Lastname}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% If multiple authors, uncomment and edit the lines shown below.       %%
+%% Note that each line must be emphasized {\em } by itself.             %%
+%% (by Stephen Martucci, author of spconf.sty).                         %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \makeatletter
+% \def\name#1{\gdef\@name{#1\\}}
+% \makeatother
+% \name{{\em Firstname1 Lastname1, Firstname2 Lastname2, Firstname3 Lastname3,}\\
+%      {\em Firstname4 Lastname4}}
+%%%%%%%%%%%%%%% End of required multiple authors changes %%%%%%%%%%%%%%%%%
 
-\date{}
-
+\address{ Affiliation / Address line 1 \\
+{\small \tt email@domain}
+}
+%
 \begin{document}
 \maketitle
-
-\fyDone{s/naacl/EMNLP/g}
-\fyDone{Too long, selfcontained, noref, rewrite}
+%
 \begin{abstract}
@@
   Our experiments use two neural architectures and two language pairs: they show that our approach, while remaining simple and computationally inexpensive, outperforms several strong baselines and delivers a multi-domain system that can successfully translate heterogeneous texts.
-
-%   \fyTodo{Remove this}
-%   THIS IS THE OLD ABSTRACT 
-%   Fine-tuning a generic neural machine translation model to a certain domain usually degrade its performance on the other domains, thus it's very difficult to build a unique model which is best for all.
-Indeed, Catastrophic Forgetting, well-known problem in Machine Learning, states that standard backprogation neural network suffers a lot while there is a shift in training data distribution.
-However, Finetuning is very costly and unefficient in the industry therefore one best model is always desired.
-In this paper, we introduce an application of an old technique \cite{Daume07frustratingly} in Multi-Domain Neural Machine Translation.
-By defining general information encoding region and domain-specific encoding region in word embedding vector, we are able to seperate domain-related information flows in forward and backward propagation during training thus to
-% mitigate the catastrophic interference happens during the training over different domains.
-Empirically, we achieved a unique model that has comparable score to finetuned models for every domains.
-Furthermore, our method is simple and could be applied to any architecture.
 \end{abstract}
@@
 \section{Introduction \label{sec:introduction}}
-% The problem
 Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,Bahdanau15learning,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progresses over the past years, and constitute to date the standard for most production engines.
 The development of MT systems, be they neural or statistical, require very large parallel corpora consisting of millions of sentence pairs, a resource that only exist in very few application domains and language pairs.
-A lot of the recent research effort has thus focused on developping MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Ha16towards,Johnson17google}; %or even building systems without any parallel data \cite{Artetxe18unsupervised,Lample18unsupervised}.
+A lot of the recent research effort has thus focused on developping MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Ha16towards,Johnson17google};
 Another important scenario for industrial MT is to adapt a neural system trained using parallel data in one domain to the peculiarity of other domains.
@@
 This allows us to be both data efficient (all data is used to train all domains) and computationally efficient (we only train one system).
-%In our setting, we assume that the domain of test documents is known.
-
 Multi-domain adaptation is conceptually close to multilingual MT, or more generally to multi-task learning \cite{Caruana97multitask}, and can be approached in a number of ways.
-We adapt here ideas of \citet{Daume07frustratingly} to neural MT.
-% in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains.
+We adapt here ideas of \cite{Daume07frustratingly} to neural MT.
 Our main hypothesis is that domains mostly differ at the lexical level, due to cross-domain polysemy, which motivates domain specific embeddings.
@@
 Our contribution is thus as follows:
-%\begin{itemize}
-%\item 
 we adapt and implement the ideas of \cite{Daume07frustratingly} for two NMT architectures;
-%\item 
 we provide experimental evidence that show the effectiveness of this technique;
-%\item 
 our analysis of embeddings conforts our assumption that their variation across domains actually reflects a variation of senses.
-%\end{itemize}
-\fyDone{can we train in random order ? can we get away with catastrophic forgetting ?}
-\fyDone{how to analyze the embeddings ? how can we test or claim ?}
-
 \section{Lexicalized domain representations\label{sec:lexicalized_embeddings}}
-\fyDone{Use meaningful titles throughout}
 
@@
 Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers the computation in equation~\ref{eq:embedding}.
-In our experiment, we consider both the attentional RNN architectures of \citet{Bahdanau15learning} and for Transformer architectures of \citet{Vaswani17attention}.
+In our experiment, we consider both the attentional RNN architectures of \cite{Bahdanau15learning} and for Transformer architectures of \cite{Vaswani17attention}.
 
@@
 \subsection{Baselines \label{ssec:baselines}}
-To validate our findings, we compared lexicalized domain embedding models with standard models using both attentional Recurrent Neural Networks (RNNs) \cite{Bahdanau15learning} and the recent Transformer architecture of \citet{Vaswani17attention}.
-Our baselines consist of generic models trained with a simple concatenation of all corpora (denoted \texttt{Mixed} in the tables below); models tuned separately on each domain using respectively (10000, 15000, 5000) iterations over in-domain data (denoted \texttt{ft(EMEA)}, \texttt{ft(EPPS)}, \texttt{ft(ECB)} below); models using domain tags as in \cite{Kobus17domaincontrol} (\texttt{DC} below); and finally the multi-domain model of \citet{Zeng18multidomain} (\texttt{WDCMT} below).\fyDone{Change names in tables}
+To validate our findings, we compared lexicalized domain embedding models with standard models using both attentional Recurrent Neural Networks (RNNs) \cite{Bahdanau15learning} and the recent Transformer architecture of \cite{Vaswani17attention}.
+Our baselines consist of generic models trained with a simple concatenation of all corpora (denoted \texttt{Mixed} in the tables below); models tuned separately on each domain using respectively (10000, 15000, 5000) iterations over in-domain data (denoted \texttt{ft(EMEA)}, \texttt{ft(EPPS)}, \texttt{ft(ECB)} below); models using domain tags as in \cite{Kobus17domaincontrol} (\texttt{DC} below); and finally the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT} below).\fyDone{Change names in tables}
 For all models, we set the embeddings size equal to 512; the size of hidden layers is equal to 1024 for RNNs and 512 for Transformer. Other important configuration details are as follows:
-\fyDone{Improve this}
-%\begin{itemize}
-%\item 
 Transformer models use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells;  
-%  size of word embedding: 512; size of hidden layers: 512; size of inner feed forward layer: 2048; Multi-head: 8; number of layers: 6.
-%\item 
 RNN models use 1 layer on both sides: a bidirectional LSTM encoder and a unidirectional LSTM decoder with attention;
-%  Attention Recurrent Neural Network: Bidirection LSTM encoder, unidirectional LSTM decoder, Luong Attention Mechanism, size of word embedding: 512; size of hidden layers: 1024; number of layers: 1
-%\item 
-The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.\fyDone{Check This!}
-  % C: size of domain embedding:2 ; same settings for both architecture: Transformer and Attention Recurrent Neural Network.
-%\item 
+The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
 The \texttt{WDCMT} system has one bidirectional Gated recurrent units (GRU) layer on the encoder side and one unidirectional conditional GRU layer on the decoder side; the dimension of ``domain'' layers is~300.
-%\end{itemize}
 
@@
 
-%In training, we use same optimization algorithm as for the baselines. 
-%The same optimisation algorithm is always used.
 Note that each iteration of algorithm \ref{alg:multidomain} uses 2~batches: a ``generic'' batch updating only the generic region; and a ``domain-specific'' batch updating just the domain-specific parameters.
-% Thus, we minimise two losses:
-%This is because we wish to simultaneously minimize two losses:
-% \begin{align*}
-% \theta^*_{s}=&\displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))] \\ 
-%  \theta^*_{i}=&\displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
-% \end{align*}
 
-% We also take into account the problem of overfitting that could occur for small corpora.
-Indeed, real data are not always balanced for every domain.
-For instance, in our experiments, the size of corpora are very different where EPPS is largest with approximately 2 millions sentences, EMEA contains about 1 million sentences and ECB is the smallest with only 200 000 sentences for English French and 100 000 for English-German.
-If we train each corpora with same number of batches, ECB will be trained too oftenly, that easily leads to overfitting.
-Therefore, we select batches from each corpora with frequencies proportional to its size.
 The batch selection procedure (step~2 of algorithm~\ref{alg:multidomain}) ensures that the number of examples seen in training of each domain follows the distribution of examples seen in the training data.
@@
 In an attempt to build a more balanced training sample, we also consider an alternative selection procedure, where $i$ is selected according to  distribution $[\frac{\sqrt{|C_i|}}{\sum_{i\in [1,..,d]}\sqrt{|C_i|}}]$.
-% \fyTodo{Refs on this ? or contrast ?}
 
 \subsection{Results \label{ssec:results}}
-% Discussion 1:
-% \begin{itemize}
-% \item transformer better than RNN
-% \item Fine tuning is the best strategy - but subject to catastrophic
-% \item Sparse better than joint - better than domain control
-% \item Across th board / language \& architecture - always good for EMEA , less good for EPPS, varying for ECB
-% \item better ? than domain control ? WMCD ?
-% \end{itemize}
-
-% Discussion 2:
-% \begin{itemize}
-% \item Sparse 1 > Sparse 2
-% \item out of domain better than Mixed
-% \item Sparse 3 ??
-% \item ajouter Sparse 4
-% \end{itemize}
 
@@
 $\mathtt{LDR}_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & 37.14\\
-$\mathtt{LDR}_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & 21.45 & 16.61 & 19.07 & 32.2 \\
+$\mathtt{LDR}_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & 21.45 & 16.61 & 19.07 & 37.23 \\
 $\mathtt{LDR}_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & - & - & - & - \\
@@
  \hline
-$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & 38.97 (39.52) \\
+$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & 39.52 \\
 \hline
@@
 
-% \begin{table}
-%   \centering
-%   \begin{tabular}{*{4}{|r|}}
-%     \hline
-%     Domain & Word & Neighbor & Cosinus distance \\ \hline
-%     \multirow{5}{*}{EMEA} & \multirow{5}{*}{joint} 
-%     & Joint & 0.371 \\  
-%     &  & jointly & 0.545 \\
-%     &  & artic- & 0.605 \\
-%     &  & joints & 0.607 \\
-% 	&  & collective & 0.609 \\
-% 	\hline
-% 	\multirow{5}{*}{EPPS} & \multirow{5}{*}{joint} 
-% 	& Joint & 0.365 \\
-% 	& & jointly & 0.497 \\
-% 	& & join- & 0.606 \\
-% 	& & common & 0.607 \\
-% 	& & collective & 0.626 \\
-% 	& & artic- & 0.633\\
-% 	\hline
-% 	\multirow{5}{*}{ECB} & \multirow{5}{*}{joint}
-% 	& Joint & 0.423 \\
-% 	& & jointly & 0.467 \\
-% 	& & common & 0.601 \\
-% 	& & together & 0.632 \\
-% 	& & artic- & 0.663 \\
-% 	& & coordinated & 0.668 \\
-% \end{tabular}
-% \caption{Corpora}
-% \label{tab:Polysemy}
-% \end{table}
+\section{Related Work \label{sec:related_work}}
 
-\section{Related Work \label{sec:related_work}}
-\fyDone{Add standard labels to sections}
-\fyDone{Related work goes last}
-\fyDone{Compare also to Peng 2017}
-
-% Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory.
-Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{Chu18asurvey}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora.
-The methods in this category can be listed as following:
-% \fyTodo{Organize refs}
-
-% \begin{itemize}
-% \item score relatedness of sentences to the target domain by Language Model \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}.
-% \item generating pseudo parallel data  \cite{Utiyama2003measure,Wang2016connecting,Wang2014neural}
-% \end{itemize}
-% On the other hand, model-centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features:
-% \begin{itemize}
-% \item Trainining objective\cite{Luong2015SNMT,Sennrich2016Mono,Wang17instance,Chen17costweighting,Miceli2017Regularize,Zhang18sentence}
-% \item Architecture \cite{Gulcehre2016monolingual,Zhang16topicinformed,Kobus17domaincontrol,Britz2017mixing,Biao2017CARENMT,Britz2017mixing,Thompson18freezing,Michel2018extreme}.
-% \item The decoding algorithm \cite{Gulcehre2016monolingual,Khayrallah2017lattice}
-% \end{itemize}
-% Beside these works \fyTodo{Works= oeuvres d'art}, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.
-
-%% HERE
 Domain adaptation (DA) is a vexing problem in NLP, which appears in a wide range of practical situations and data scenarios (eg.\ supervised vs.
@@
 This corresponds to a practical scenario in the industry, where one would like to maintain one single multi-domain engine, trained on all the available (heteregeneous) sources of data.
-Our primary source of inspiration is the proposal of \citet{Daume07frustratingly} who proposes to use several copies of the same features (one ``generic'' shared across domains, and one for each domain), letting the training adjust their respective weights.
+Our primary source of inspiration is the proposal of \cite{Daume07frustratingly} who proposes to use several copies of the same features (one ``generic'' shared across domains, and one for each domain), letting the training adjust their respective weights.
 This work has been reanalyzed in a Bayesian framework in \cite{Finkel09hierarchical}, and revisited notably in \cite{Chang10necessity}.
@@
 The multi-domain scenario in NMT has been studied in a number of recent works.
-\citet{Farajian17multidomain} learn a generic system and propose to dynamically adapt the network weights in an unsupervised manner using a small sample of training data that resembles the test data, an idea already explored in \cite{Sennrich13multidomain}.
+\cite{Farajian17multidomain} learn a generic system and propose to dynamically adapt the network weights in an unsupervised manner using a small sample of training data that resembles the test data, an idea already explored in \cite{Sennrich13multidomain}.
 Their main contribution is to propose an a method to relate the amount of adaptation of the network parameters to the similarity between the adaptation sample and the test sentence: the higher the similarity, the more agressive the adaptation.
 
-By analogy with the proposal of \citet{Johnson17google} for multilingual NMT, \citet{Kobus17domaincontrol} and \citet{Chu18multilingual} separately propose to extend the representation of the source text with a domain tag.
+By analogy with the proposal of \cite{Johnson17google} for multilingual NMT, \cite{Kobus17domaincontrol} and \cite{Chu18multilingual} separately propose to extend the representation of the source text with a domain tag.
 Our model also modifies input representations, but allows each source word to have a domain-specific representation, thereby improving training of the shared parts of the network.\fyDone{Talk about target embeddings somewhere.}
@@
 
-% creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
-$h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively.
-While introducing new features in the architecture, \cite{Zeng18multidomain} also introduces new loss function which is sum of word-level weighted MT loss and losses of Domain-classifier in source side and target side; Adversarial Domain-classifier in source side.
-The authors have very good approach to seperate Domain-shared information flow and Domain-specify information flow and to determine their contributions to the inference thus mitigate the catastrophic interference occuring in training with backpropagation.
-
-% The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{Platanios18contextual}; \cite{Zeng18multidomain} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$.
-$h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively.
-While introducing new features in the architecture, \cite{Zeng18multidomain} also introduces new loss function which is sum of word-level weighted MT loss and losses of Domain-classifier in source side and target side; Adversarial Domain-classifier in source side.
-The authors have very good approach to seperate Domain-shared information flow and Domain-specify information flow and to determine their contributions to the inference thus mitigate the catastrophic interference occuring in training with backpropagation
-
-% Platanios is multilingual !!
-% \cite{Platanios18contextual} has other idea to circumvent the catastrophic interference.
-They supposed that parameters of each domain specialized model could be generated from a universal parameter-generator.
-By introducing domain embedding (or language embedding in the original paper), they could conduct domain related error in backpropagation toward domain embedding thus mitigates catastrophic interference between domains during training.
-
-% Compared to \cite{Platanios18contextual}, our sparse embedding is a bit similar to the idea.
-However, our method gives much more free degree for domain embedding since we give each word a domain region while the author uses only one vector domain embedding for each parameter set(encoder, decoder).
-
-% Compared to the work of \cite{Zeng18multidomain}, our approach is much simpler and
-% more generic, since sparse embedding can be implemented for every neural architecture while domain-specific gate and domain-shared gate can only be applied for recurrent architectures.
-However, during inference, the model of \cite{Zeng18multidomain} can infer the domain that is the closest to the input sentence and that will be used in special gates; while our method needs to know in which domain the inference is executed.
-To circumvent this disadvantage, we could use a domain classifier or simply use only the generic region.
-For more detail, the reader could refer to the Section \ref{sec:sparse_embeddings}.
-
 \section{Conclusions}
@@
 
-% \fyTodo{natural continuations: bayesian version ? linear combination of specialized engines ? go beyond words; learn the projection matrix at the word level; other ?}
+\section{Acknowledgements}
+The IWSLT 2015 organizing committee would like to thank the
+organizing committees of INTERSPEECH 2004 for their
+help and for kindly providing the template files.
 
-%\section*{Acknowledgments}
-% \fyTodo{Homegenize refs - urls, addresses, etc}
-\bibliography{emnlp-ijcnlp-2019}
-\bibliographystyle{acl_natbib}
+%
+\bibliography{iwslt18}
+\bibliographystyle{IEEEtran}
 \end{document}

### 2019-08-26 2019/iwslt/iwslt2018-v2.tex
remove last test sets

--- 2019-07-11 2019/iwslt/iwslt2018-submission.tex
@@
 We experiment with two language pairs (English-French, English-German) and data originating from three domains, corresponding to texts from three European institutions: the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann09news}.
-We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).\footnote{The EMEA dataset as distributed on the OPUS site contains many duplicates.
-We therefore report two numbers: the first is comparable to what has been published on earlier studies (eg.\ \cite{Zeng18multidomain}), the second one is obtained by making the test set entirely disjoint from the train.} The validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.
-Using internal test sets corresponds to somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from three other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
+We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).
+%\footnote{The EMEA dataset as distributed on the OPUS site contains many duplicates.
+We therefore report two numbers: the first is comparable to what has been published on earlier studies (eg.\ \cite{Zeng18multidomain}), the second one is obtained by making the test set entirely disjoint from the train.}
+Development sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.
+Using internal test sets corresponds to somewhat ideal conditions.
+%for contrast experiments, we also use supplementary test sets from three other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
 This enables us to evaluate the loss in performance when the test set is from a domain not seen in training.
@@
   \centering
-  \begin{tabular}{ *{4}{|r|}}
+  \begin{tabular}{ *{4}{llll}}
     \hline
     Corpus & Train & Dev & Test \\ \hline
-    \multicolumn{4}{|l|}{En $\rightarrow$ Fr }\\
-    \multicolumn{4}{|l|}{Vocabulary size - En: 30,165, Fr: 30,398}\\
+    \multicolumn{4}{l}{En $\rightarrow$ Fr }\\
+%    \multicolumn{4}{|l|}{Vocab - En: 30,165, Fr: 30,398}\\
     \hline
-    EMEA  & 1,09M & 1000 & 1000$_{300}$\\
-    ECB    & 0,19M & 1000 & 1000     \\
-    EPPS   & 2,0M  & 1000 & 1000  \\ \hline \hline
-    \multicolumn{4}{|l|}{En $\rightarrow$ De}\\
-    \multicolumn{4}{|l|}{Vocabulary size -  En:30,159, De: 30,698}\\ \hline
-    EMEA  & 1,11M & 1000 & 1000$_{300}$ \\
-    ECB     &  0,11M & 1000 & 1000  \\
-    EPPS   & 1,92M & 1000 & 1000 \\ \hline
+    EMEA  & 1,09M & 1,000 & 1,000\\
+    ECB    & 0,19M & 1,000 & 1,000     \\
+    EPPS  & 2,01M  & 1,000 & 1,000  \\ \hline 
+    \multicolumn{4}{l}{En $\rightarrow$ De}\\
+%    \multicolumn{4}{|l|}{Vocab -  En: 30,159, De: 30,698}\\ \hline
+    \hline
+    EMEA  & 1,11M & 1,000 & 1,000 \\
+    ECB     &  0,11M & 1,000 & 1,000  \\
+    EPPS   & 1,92M & 1,000 & 1,000 \\ \hline
 \end{tabular}
-\caption{Train and test corpora}
+\caption{Train, Development and test corpora.}
 \label{tab:Corpora}
@@
 To validate our findings, we compared lexicalized domain embedding models with standard models using both attentional Recurrent Neural Networks (RNNs) \cite{Bahdanau15learning} and the recent Transformer architecture of \cite{Vaswani17attention}.
-Our baselines consist of generic models trained with a simple concatenation of all corpora (denoted \texttt{Mixed} in the tables below); models tuned separately on each domain using respectively (10000, 15000, 5000) iterations over in-domain data (denoted \texttt{ft(EMEA)}, \texttt{ft(EPPS)}, \texttt{ft(ECB)} below); models using domain tags as in \cite{Kobus17domaincontrol} (\texttt{DC} below); and finally the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT} below).\fyDone{Change names in tables}
+Our baselines consist of:
+\begin{itemize}
+\item generic models trained with a simple concatenation of all corpora (denoted \texttt{Mixed} in the tables below); 
+\item models tuned separately on each domain using respectively (10,000, 15,000, 5,000) iterations over in-domain data (denoted \texttt{FT} below); 
+\item models using domain tags as in \cite{Kobus17domaincontrol} (\texttt{DC} below); 
+\item and finally the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT} below).\fyDone{Change names in tables}
+\end{itemize}
 For all models, we set the embeddings size equal to 512; the size of hidden layers is equal to 1024 for RNNs and 512 for Transformer. Other important configuration details are as follows:
@@
 In all cases, we use a batch size of~128 and a dropout rate of 0.1 for all layers.
-All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt} except for WDCMT for which we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyDone{batch size}
+All systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt} except for WDCMT for which we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyDone{batch size}
 
-\subsection{Implementing Lexicalized Domain Representations}\fyDone{Check acronym.}
+\subsection{Lexicalized Domain Representations}\fyDone{Check acronym.}
 
 \fyDone{Motivate the split - discuss experimentally embedding size}
-In order to implement lexicalised domain representations (henceforth LDR), we split the embedding vector into four regions: 3 are domain specific and 1 is generic, with sizes $[8,8,8,488]$ respectively.
+In order to implement lexicalised domain representations (henceforth \texttt{LDR}), we split the embedding vector into four regions: 3 are domain specific and 1 is generic, with sizes $[8,8,8,488]$ respectively.
 %, a choice that will be discussed in Section~\ref{secc:region_size}.
@@
 
-The right parts of Tables~\ref{tab:results-trsf} and \ref{tab:results-rnn} report performance on test sets for new domains not seen in training.
+%The right parts of Tables~\ref{tab:results-trsf} and \ref{tab:results-rnn} report performance on test sets for new domains not seen in training.
 Khresmoi texts, which are extracted from summaries of medical articles are assumed to be from the same domain as EMEA data; for the other two test sets, we assume no domain and only use the generic part of the embeddings.
@@
 
-\begin{table*}[!h]
+\begin{table}[!h]
 \begin{center}
 \scalebox{1.0}{
-\begin{tabular}{|l|ccc|ccc||c|}
+\begin{tabular}{|l|ccc|c|}
 \hline
-Model & EMEA & EPPS & ECB & Khresmoi & NewsTest2014 & IWSLT2010 & Avg. \\
+Model & EMEA & EPPS & ECB & Avg. \\ %Khresmoi & NewsTest2014 & IWSLT2010 & Avg. \\
 \hline
-\multicolumn{8}{l}{} \\[-9pt]  
-\multicolumn{8}{l}{English$\rightarrow$French} \\
+%\multicolumn{5}{l}{} %\\[-9pt]  
+\multicolumn{5}{l}{English$\rightarrow$French} \\
 \hline
-$Mixed$ & 67.69$_{47.60}$ & 37.50 & 53.49 & 34.86 & 28.56 & 25.70 & 41.3\\
+$\mathtt{Mixed}$ & 67.69 & 37.50 & 53.49 & 52.89\\
+$\mathtt{FT}$ & 76.77 & 37.04 & 56.52 & 56.77\\
+$\mathtt{DC}$ & 67.87 & 37.31 & 54.14 & 53.10\\
 \hline
-$\mathtt{FT} (EMEA)$ & 76.77$_{49.43}$ & 17.16 & 11.99 & 29.58 & 11.64 & 11.10 & 26.37 \\
-$\mathtt{FT} (EPPS)$  & 20.86 & 37.04 & 24.53 & 31.86 & 28.33 & 11.10 & 25.62 \\
-$\mathtt{FT} (ECB)$   & 26.93 & 27.09 & 56.52 & 22.04 & 17.51 & 13.99 &  27.35\\
+$\mathtt{LDR}_{oracle}$   & 74.26 & 37.67 & 54.07 & 55.33\\
+$\mathtt{LDR}_{oracle}^{0.5}$   & 74.95 & 37.35 & 55.91 & 56.07\\
+$\mathtt{LDR}_{generic}$ & 73.97 & 37.81 & 53.67 & 55.15\\
+$\mathtt{LDR}_{pred}$ & 74.29 & 37.73 & 54.01 & 55.34\\
+$\mathtt{LDR}_{wrong}$ & 72.95 & 37.62 & 53.35 & 54.64\\
   \hline
-$\mathtt{DC}$ & 67.87$_{45.42}$ & 37.31 & 54.14 & 33.63 & 26.52 & 24.81 & 40.71\\
+%\multicolumn{5}{l}{} %\\[-9pt]
+\multicolumn{5}{l}{English $\rightarrow$German} \\
 \hline
-$\mathtt{LDR}_{oracle}$   & 74.26$_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 & 42.28\\
-$\mathtt{LDR}_{oracle}^{0.5}$   & 74.95$_{49.38}$ & 37.35 & 55.91 & 33.57 & 27.59 & 25.27 & 42.44\\
-$\mathtt{LDR}_{generic}$ & 73.97$_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & 42.24 \\
-$\mathtt{LDR}_{pred}$ & 74.29 $_{49.84}$ & 37.73 & 54.01 & 33.34 & 28.32 & 26 & 42.28\\
-$\mathtt{LDR}_{wrong}$ & 72.95$_{49.78}$ & 37.62 & 53.35 & - & - & - & - \\
+$\mathtt{Mixed}$ & 64.57 & 26.47 & 68.67 & 53.23\\
+$\mathtt{FT}$ & 68.35 & 26.29 & 74.05 & 56.23\\
+$\mathtt{DC}$ & 63.48 & 26.27 & 66.95 & 52.23\\
 \hline
-\multicolumn{8}{l}{} \\[-9pt]
-\multicolumn{8}{l}{English $\rightarrow$German} \\
-\hline
-$Mixed$ & 64.57$_{42.99}$ & 26.47 & 68.67 & 21.77 & 17.04 & 18.85 &36.23 \\
-\hline
-$\mathtt{FT} (EMEA)$ & 68.35$_{42.97}$ & 17.02 & 32.87 & 21.15 & 11.21 & 13.49 & 27.35\\
-$\mathtt{FT} (EPPS)$ & 36.19 & 26.29 & 40.71 & 19.67 & 16.47 & 19.24 & 31.42\\
-$\mathtt{FT} (ECB)$  & 24.72 & 18.36 & 74.05 & 11.31  & 10.8 & 11.00 & 25\\
-\hline
-$\mathtt{DC}$ & 63.48$_{42.98}$ & 26.27 & 66.95 & 21.45 & 16.45 & 18.78 & 35.56\\
-\hline
-$\mathtt{LDR}_{oracle}$ & 70.9$_{46.12}$ & 26.30 & 68.9 & 21.42 & 16.57 & 18.82 & 37.15\\
-$\mathtt{LDR}_{oracle}^{0.5}$ & 71.31$_{45.23}$ & 25.98 & 73.74 & 21.6 & 16.37 & 19.2 & 38.03\\
-$\mathtt{LDR}_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & 37.14\\
-$\mathtt{LDR}_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & 21.45 & 16.61 & 19.07 & 37.23 \\
-$\mathtt{LDR}_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & - & - & - & - \\
+$\mathtt{LDR}_{oracle}$ & 70.90 & 26.30 & 68.90 & 55.36\\
+$\mathtt{LDR}_{oracle}^{0.5}$ & 71.31 & 25.98 & 73.74 & 57.01\\
+$\mathtt{LDR}_{generic}$ & 70.86 & 26.14 & 68.59 & 55.19\\
+$\mathtt{LDR}_{pred}$ & 70.89 & 26.53 & 68.63 & 55.35\\
+$\mathtt{LDR}_{wrong}$ & 69.51 & 26.31 & 66.86 & 54.22\\
 \hline
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for the Transformer systems \label{tab:results-trsf}}
-\end{table*}
+\caption{BPE-detokenized BLEU scores (Transformer) \label{tab:results-trsf}}
+\end{table}
 
-\begin{table*}[!h]
+\begin{table}[!h]
 \begin{center}
 \scalebox{1.0}{
-\begin{tabular}{|l|ccc|ccc||c|}
+\begin{tabular}{|l|ccc|c|}
 \hline
-Model & EMEA & EPPS & ECB & Khresmoi & NewsTest2014 & IWSLT2010 & Avg. \\
+Model & EMEA & EPPS & ECB & Avg. \\
 \hline
-\multicolumn{8}{l}{} \\[-9pt]
-\multicolumn{8}{l}{English$\rightarrow$French} \\
+%\multicolumn{5}{l}{} %\\[-9pt]
+\multicolumn{5}{l}{English$\rightarrow$French} \\
 \hline
-$Mixed$               & 65.42$_{45.11}$ & 34.7 & 51.38 & 29.27 & 22.21 & 21.63 &37.43 \\
+$\mathtt{Mixed}$               & 65.42 & 34.7 & 51.38 & \\
+$\mathtt{FT}$                    & 72.06 & 34.61 & 51.53 & \\
+$\mathtt{DC}$                    & 68.26 & 35.13 & 50.09 & \\
+$\mathtt{WDCMT}$           & 68.76 & 35.71 & 52.75 &\\
 \hline
-$\mathtt{FT} (EMEA)$  & 72.06$_{47.33}$ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & 20 \\
-$\mathtt{FT} (EPPS)$    & 35.47 $ $ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & 29.87 \\
-$\mathtt{FT} (ECB)$     & 21.93 $ $ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & 22.23 \\
+$\mathtt{LDR}_{oracle}$    & 71.73 & 35.21 & 50.91 & \\
+$\mathtt{LDR_{condgru}}_{pred}$   & 71.70 & 35.09 & 51.22 & \\
+$\mathtt{LDR}_{oracle}^{0.5}$   & 71.70 & 34.24& 52.37 & \\
+$\mathtt{LDR}_{generic}$  & 70.59 & 35.22 & 49.68 & \\
+$\mathtt{LDR}_{pred}$       & 72.76 & 35.10 & 50.38 & \\
+$\mathtt{LDR}_{wrong}$    & 62.10 & 34.17 & 48.79 &  \\
 \hline
-$\mathtt{DC}$                   & 68.26 $_{43.76}$ & 35.13 & 50.09 & 27.86 & 20.95 & 21.16 & 37.24\\
+%\multicolumn{5}{l}{} %\\[-9pt]
+\multicolumn{5}{l}{English$\rightarrow$German} \\
 \hline
-$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & 39.52 \\
+$\mathtt{Mixed}$                & 57.37 & 23.1 & 63.54 & \\
 \hline
-$\mathtt{LDR}_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & 38.57\\
-$\mathtt{LDR_{condgru}}_{pred}$   & 71.7 $_{46.21}$ & 35.09 & 51.22 & 31.75 & 24.25 & 22.63 & 39.44\\
-$\mathtt{LDR}_{oracle}^{0.5}$   & 71.7$_{46.41}$ & 34.24& 52.37 & 29.21 & 22.7& 20.69 & 38.48\\
-$\mathtt{LDR}_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & 38.16\\
-$\mathtt{LDR}_{pred}$        & 72.76 $_{46.35}$ & 35.1 & 50.38 & 30.02 & 23.28 & 21.32 & 38.81\\
-$\mathtt{LDR}_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 &  - & - & - & - \\
+$\mathtt{FT} (EMEA)$        & 65.64 & 12.36 & 15.93 & \\
+$\mathtt{FT} (EPPS)$        & 24.90 & 22.98 & 26.26 & \\
+$\mathtt{FT} (ECB)$          & 41.80  & 15.97 & 71.07 & \\
+$\mathtt{DC}$                     & 62.53 & 23.74 & 65.71 & \\
 \hline
-\multicolumn{8}{l}{} \\[-9pt]
-\multicolumn{8}{l}{English$\rightarrow$German} \\
+$\mathtt{WDCMT}$            & 64.05 & 22.82 & 67.00 & \\
 \hline
-$Mixed$               & 57.37 $_{37.94}$ & 23.1 & 63.54 & 17.39 & 11.51 & 14.29 & 31.2\\
-\hline
-$\mathtt{FT} (EMEA)$  & 65.64$_{44.71}$ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & 20.53\\
-$\mathtt{FT} (EPPS)$   & 24.90 $ $ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & 18.89 \\
-$\mathtt{FT} (ECB)$    & 41.8 $ $ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & 25.89\\
-\hline
-$\mathtt{DC}$                      & 62.53 $_{39.25}$ & 23.74 & 65.71 & 16.88 & 11.15 & 14.97 & 32.50\\
-\hline
-$\mathtt{WDCMT}$ & 64.05 $_{38.54}$ & 22.82 & 67.00 & 15.94 & 11.45 & 13.66 & 32.49\\
-\hline
-$\mathtt{LDR}_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & 32.2\\
-$\mathtt{LDR}_{oracle}^{0.5}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & 32.6\\
-$\mathtt{LDR}_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & 32.1\\
-$\mathtt{LDR}_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64 & 15.94 & 11.32 & 14.27 & 31.87\\
-$\mathtt{LDR}_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & - & - & - & - \\
+$\mathtt{LDR}_{oracle}$    & 63.43 & 22.66 & 64.40 & \\
+$\mathtt{LDR}_{oracle}^{0.5}$   & 63.27 & 21.83 & 69.55 & \\
+$\mathtt{LDR}_{generic}$  & 63.27 & 22.62 & 63.70 & \\
+$\mathtt{LDR}_{pred}$       & 63.17 & 22.51 & 64.00 & \\
+$\mathtt{LDR}_{wrong}$    & 56.84 & 22.06 & 61.66 & \\
 \hline
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for the RNN systems\label{tab:results-rnn}}
-\end{table*}
+\caption{BPE-detokenized BLEU scores (RNN)\label{tab:results-rnn}}
+\end{table}
 % \clearfloat
@@
 
-\begin{table*}[!h]
+\begin{table}[!h]
 \begin{center}
 \scalebox{1.0}{
-\begin{tabular}{|l|ccc|ccc|}
+\begin{tabular}{|l|ccc|}
 \hline
-$LDR$ & EMEA & EPPS & ECB & Khresmoi & NewsTest2014 & IWSLT2010 \\
+$LDR$ & EMEA & EPPS & ECB \\
 \hline
-\multicolumn{7}{l}{English$\rightarrow$French} \\
+\multicolumn{4}{l}{English$\rightarrow$French} \\
 \hline
- size=4   & 74.65 $_{49.61}$ & 37.42 & 54.49 & 33.68 & 28.03 & 25.28 \\
- size=8   & 74.26 $_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 \\
- size=16 & 74.15 $_{49.10}$ & 37.78 & 54.56 & 33.40 & 28.10 & 25.33 \\
- size=32 & 75.10 $_{48.61}$ & 37.64 & 54.29 & 33.20 & 27.99 & 25.44 \\
- size=64 & 74.5 $_{50.17}$ & 37.27 & 54.5 & 33.26 & 27.95 & 25.65 \\
+ size=4   & 74.65 & 37.42 & 54.49 \\
+ size=8   & 74.26 & 37.67 & 54.07 \\
+ size=16 & 74.15 & 37.78 & 54.56 \\
+ size=32 & 75.10 & 37.64 & 54.29 \\
+ size=64 & 74.50 & 37.27 & 54.50 \\
 \hline
@@
 \caption{BPE-detokenized BLEU scores for the Transformer architecture for varying domain-specific embedding sizes \label{tab:embedding-size}}
-\end{table*}
+\end{table}
 

### 2019-08-26 2019/iwslt/submission.tex
changes

--- 2019-08-26 2019/iwslt/iwslt2018-v2.tex
@@
 \documentclass[a4paper]{article}
-\usepackage{iwslt18,amssymb,amsmath,epsfig}
+\usepackage{iwslt19,amssymb,amsmath,epsfig}
 \setcounter{page}{1}
@@
 \usepackage[draft]{todo}
+
 \floatname{algorithm}{Procedure}
@@
 \newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{orange}{#1}}}
 
@@
 we provide experimental evidence that show the effectiveness of this technique;
-our analysis of embeddings conforts our assumption that their variation across domains actually reflects a variation of senses.
+\jcTodo{focus on that the analysis of embeddings is an original technique.}
+our analysis of embeddings comforts our assumption that their variation across domains actually reflects a variation of senses.
+
 \section{Lexicalized domain representations\label{sec:lexicalized_embeddings}}
@@
 
-Multi-domain machine translation is formalized as follows: we assume that our observations correspond to  domain-tagged sentence pairs $[(x,y),i]$, with $x$ in the source language, $y$ in the target language and $i$ a domain tag in $[1\dots d]$.
+Multi-domain machine translation is formalized as follows: we assume observations that correspond to  domain-tagged sentence pairs $[(x,y),i]$, with $x$ in the source language, $y$ in the target language and $i$ a domain tag in $[1\dots d]$.
 We further assume a two-stage sampling process: first select a domain $i$ according to $p(i)$, then select a sentence pair according to a domain specific distribution $D_i$.
-Our objective is to find a set of parameters $\{\theta_1 \dots \theta_d \} \in \mathbb{R}^D \times \mathbb{R}^D \dots \times \mathbb{R}^D$ minimizing:
+Our objective is to find a tuple of parameters $\{\theta_1 \dots \theta_d \} \in \mathbb{R}^D \times \mathbb{R}^D \dots \times \mathbb{R}^D$ minimizing:
 \begin{equation} \label{eq:loss}
 \begin{split}
-\sum_{i \in [1..d]} p(i) E_{(x,y) \sim D_{i}} [-log(p_{\theta}(y|x,i))]
+\sum_{i \in [1..d]} p(i) E_{(x,y) \sim D_{i}} [-log(p_{\theta_i}(y|x,i))]
 \end{split}
@@
 
-A straightforward solution is to process each domain separately, computing the value $\theta_i^*$ that minimizes the empirical loss on the training corpus $C_i$.
+A straightforward solution is to process each domain separately, computing the value $\theta_i^*$ that minimizes the empirical loss on $C_i$.
 This strategy is only effective if we have sufficient training data for each domain; when this is not the case, some estimates $\theta_i^*$ may be far from their optimal value.
@@
 
-To actually implement this idea for NMT, we need to define the subset of parameters that will be common to all domain.
+To actually implement this idea for NMT, we need to define the subset of parameters that will be common to all domains. 
 We make the assumption that domain specificities can be confined to the lexical level, and define $\theta_s$ to contain all the network parameters, except for a subpart of the word embeddings. 
@@
 In our NMT architectures, the actual embedding layer composes these vectors linearly to generate the word embedding for domain $k$ according to:
+
 \begin{align}
@@
 where $\delta()$ is the indicator function, $M$ is the matrix made of blocks $M_g, M_1 \dots, M_d$, and $e_i'(v,k)$ is the masked embedding: $e_i'(v,k)= e_i(v) * \delta(i=k)$.
+
 Making sure that the actual embedding do not contain any zero is important for the Transformer model, since the lexical representations are then summed to the positional encoding, which would undo the effect of domain masking, and propagate a gradient even to regions that should not be modified.
@@
 
-\begin{figure}
+\begin{figure}[h]
   \center
@@
 
-Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers the computation in equation~\ref{eq:embedding}.
-In our experiment, we consider both the attentional RNN architectures of \cite{Bahdanau15learning} and for Transformer architectures of \cite{Vaswani17attention}.
+Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers the computation in equation~\eqref{eq:embedding}.
+In our experiments, we consider both the attentional RNN architecture of \cite{Bahdanau15learning} and for Transformer architecture of \cite{Vaswani17attention}.
 
@@
 
-We experiment with two language pairs (English-French, English-German) and data originating from three domains, corresponding to texts from three European institutions: the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann09news}.
+We experiment with two language pairs (English-French, English-German) and data originated from three domains, corresponding to texts from three European institutions: 
+the European Parlement (EPPS) \cite{Koehn05europarl}, 
+the European Medicines Agency (EMEA), 
+the European Central Bank (ECB) \cite{Tiedemann09news},
+In addition, for English-French we also used IT-domain corpora obtained from the OPUS web site\footnote{\url{http://opus.nlpl.eu}} corresponding to KDE, Ubuntu, GNOME and PHP datasets (IT).
+
 We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).
-%\footnote{The EMEA dataset as distributed on the OPUS site contains many duplicates.
-We therefore report two numbers: the first is comparable to what has been published on earlier studies (eg.\ \cite{Zeng18multidomain}), the second one is obtained by making the test set entirely disjoint from the train.}
-Development sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.
+Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.
 Using internal test sets corresponds to somewhat ideal conditions.
+
 %for contrast experiments, we also use supplementary test sets from three other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
@@
 % The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
+
 \fyDone{Check which corpus are useful}
-\begin{table}
+\begin{table}[h]
   \centering
-  \begin{tabular}{ *{4}{llll}}
+  \begin{tabular}{ |llll|} %*{4}{|r|}}
     \hline
-    Corpus & Train & Dev & Test \\ \hline
-    \multicolumn{4}{l}{En $\rightarrow$ Fr }\\
-%    \multicolumn{4}{|l|}{Vocab - En: 30,165, Fr: 30,398}\\
+    Corpus & Train & Valid & Test \\ 
     \hline
-    EMEA  & 1,09M & 1,000 & 1,000\\
-    ECB    & 0,19M & 1,000 & 1,000     \\
-    EPPS  & 2,01M  & 1,000 & 1,000  \\ \hline 
-    \multicolumn{4}{l}{En $\rightarrow$ De}\\
-%    \multicolumn{4}{|l|}{Vocab -  En: 30,159, De: 30,698}\\ \hline
+    \multicolumn{4}{l}{English $\rightarrow$ French }\\
+    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
     \hline
-    EMEA  & 1,11M & 1,000 & 1,000 \\
-    ECB     &  0,11M & 1,000 & 1,000  \\
-    EPPS   & 1,92M & 1,000 & 1,000 \\ \hline
+    EMEA  & 1.09M & 1,000 & 1,000$_{300}$\\
+    ECB    & 0.19M & 1,000 & 1,000     \\
+    EPPS   & 2.01M  & 1,000 & 1,000  \\
+    IT         & 0.54M  & 1,000 & 1,000 \\  
+    \hline
+    \multicolumn{4}{l}{English $\rightarrow$ German}\\
+    %\multicolumn{4}{|l|}{Vocab size -  En:30,159, De: 30,698}\\ 
+    \hline
+    EMEA  & 1.11M & 1,000 & 1,000$_{300}$ \\
+    ECB     &  0.11M & 1,000 & 1,000  \\
+    EPPS   & 1.92M & 1,000 & 1,000 \\ 
+    \hline
 \end{tabular}
-\caption{Train, Development and test corpora.}
+\caption{Corpora statistics.}
 \label{tab:Corpora}
 \end{table}
+
+\jcTodo{I delete vocabs since they make no sense if computed after BPE}
+
+Note that the EMEA dataset as distributed on the OPUS site contains multiple sentence duplicates. 
+We therefore report two numbers: the first is comparable to what has been published on earlier studies (eg.\ \cite{Zeng18multidomain}), the second one is obtained by making the test set entirely disjoint from the training set (700 duplicated sentences are discarded).
+
 To reduce the number of lexical units and make our systems open-vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16BPE} separately for each language with 30,000 merge operations.
@@
 \begin{itemize}
-\item generic models trained with a simple concatenation of all corpora (denoted \texttt{Mixed} in the tables below); 
-\item models tuned separately on each domain using respectively (10,000, 15,000, 5,000) iterations over in-domain data (denoted \texttt{FT} below); 
-\item models using domain tags as in \cite{Kobus17domaincontrol} (\texttt{DC} below); 
-\item and finally the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT} below).\fyDone{Change names in tables}
+\item generic models trained with a simple concatenation of all corpora (\texttt{Mixed}); 
+\item models tuned separately on each domain using respectively (10000, 15000, 5000) iterations over in-domain data (\texttt{ft(EMEA)}, \texttt{ft(EPPS)}, \texttt{ft(ECB)}); 
+\item models using domain tags as in \cite{Kobus17domaincontrol} (\texttt{DC}); 
 \end{itemize}
-For all models, we set the embeddings size equal to 512; the size of hidden layers is equal to 1024 for RNNs and 512 for Transformer. Other important configuration details are as follows:
-Transformer models use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells;  
-RNN models use 1 layer on both sides: a bidirectional LSTM encoder and a unidirectional LSTM decoder with attention;
+
+For all models, we set the embeddings size equal to 512; the size of hidden layers is equal to 1024 for RNNs and 512 for Transformer. 
+Other important configuration details are as follows:
+Transformer models use multi-head attention with 8 heads in each of the 6 layers; 
+the inner feedforward layer contains 2048 cells;  
+RNN models use 1 layer on both sides: 
+a bidirectional LSTM encoder and a unidirectional LSTM decoder with attention;
 The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
 The \texttt{WDCMT} system has one bidirectional Gated recurrent units (GRU) layer on the encoder side and one unidirectional conditional GRU layer on the decoder side; the dimension of ``domain'' layers is~300.
-
+%
 To train NMT systems, we use Adam, with parameters $\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$ for RNNs; with parameters $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} for Transformer ($warmup\_steps=4000$).
 In all cases, we use a batch size of~128 and a dropout rate of 0.1 for all layers.
-All systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt} except for WDCMT for which we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyDone{batch size}
+All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt} except for WDCMT for which we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyDone{batch size}
 
-\subsection{Lexicalized Domain Representations}\fyDone{Check acronym.}
+\subsection{Implementing Lexicalized Domain Representations}\fyDone{Check acronym.}
 
 \fyDone{Motivate the split - discuss experimentally embedding size}
-In order to implement lexicalised domain representations (henceforth \texttt{LDR}), we split the embedding vector into four regions: 3 are domain specific and 1 is generic, with sizes $[8,8,8,488]$ respectively.
-%, a choice that will be discussed in Section~\ref{secc:region_size}.
-If a sentence originates from domain $i$, the domain specific regions for all domains $j \neq i$ will be zeroed out; while the rest of regions are activated (cf.
-Figure~\ref{fig:network}).
-We then use a dense layer of size~512 to fuse the region for the active domain and the ``generic'' region.
-Training is formalised in algorithm~\ref{alg:multidomain}.
+In order to implement lexicalised domain representations (henceforth \texttt{LDR}), we split the embedding vector into four regions: 
+3 are domain specific and 1 is generic, with sizes $[8,8,8,488]$ respectively. %, a choice that will be discussed in Section~\ref{secc:region_size}. 
+If a sentence originates from domain $i$, the domain specific regions for all domains $j \neq i$ will be zeroed out; while the other regions are activated (cf. Figure~\ref{fig:network}). 
+We then use a dense layer of size~512 to fuse the region for the active domain and the ``generic'' region. Training is formalised in algorithm~\ref{alg:multidomain}.
+%
+Note that each iteration of algorithm \ref{alg:multidomain} uses 2~batches: a ``generic'' batch updating only the generic region; and a ``domain-specific'' batch updating just the domain-specific parameters.
 
@@
 
-Note that each iteration of algorithm \ref{alg:multidomain} uses 2~batches: a ``generic'' batch updating only the generic region; and a ``domain-specific'' batch updating just the domain-specific parameters.
-
-The batch selection procedure (step~2 of algorithm~\ref{alg:multidomain}) ensures that the number of examples seen in training of each domain follows the distribution of examples seen in the training data.
+The batch selection procedure (step~2 of algorithm~\ref{alg:multidomain}) ensures that the number of examples of each domain used in training  follows the distribution of the training data.
 %each domain is represented according to its frequency in the training data. 
@@
 
-Our main results are summarized respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems, where we report BLEU scores \cite{Papineni02bleu} computed after detokenization.\footnote{As explained above, we consistently report two numbers when testing with EMEA, except for the fine-tuning scenarios when tuning on ECB and EPPS.} % \fyTodo{Footnote for missing numbers}.
+Results are summarised respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems, where we report BLEU~\cite{Papineni02bleu} scores computed after detokenization.\footnote{As explained above, we  report two numbers when testing with EMEA, except for the fine-tuning scenarios when tuning on ECB and EPPS.} % \fyTodo{Footnote for missing numbers}.
 
-We first concentrate on the left part of these tables, where we analyze the performance for the three domains of interest.
-A first series of observations is that Transformer is consistently better than RNN, and that fine-tuning on a domain-specific corpus, when applicable, is almost the best way to optimize the performance on that domain.\footnote{This is not so clear for the EPPS, where fine-tuning does not seem to help.}  Note that fine-tuning will however yield a marked (even sometimes catastrophic, eg.\ for the EMEA-tuned Transformer system) decrease in performance for the other domains.
+%We first concentrate on the left part of these tables, where we analyze the performance for the three domains of interest. 
+First, we observe that Transformer is consistently better than RNN, and that fine-tuning on a domain-specific corpus, when applicable, is almost the best way to optimise the performance on that domain.\footnote{This is not so clear for EPPS, where fine-tuning does not seem to help.}
+Note that fine-tuning will however yield a marked (even sometimes catastrophic, eg.\ for the EMEA-tuned Transformer system) decrease in performance for the other domains.
 
-For these test sets, our proposal (\texttt{LDR}$_{oracle}$) is consistently better than the mix-domain strategy, with gains that range from very large (for EMEA and ECB) to unsignificant (for EPPS in most conditions).
+The presented model (\texttt{LDR}$_{oracle}$) is consistently better than the \texttt{Mixed} strategy, with gains that range from very large (for EMEA and ECB) to unsignificant (for EPPS in most conditions).
 This means that our architecture is somehow able to compensate for the data unbalance and to raise the performance of the multi-domain system close to the best (fine-tuned) system in each domain. 
@@
 \texttt{LDR}$_{oracle}$ is also better than Domain Control in three conditions out of four, \texttt{DC} being seemingly a better choice for the RNN than for the Transformer architecture. 
-The comparison with the \texttt{WDCMT} model is less favourable for our RNN systems, which are arguably much simpler.\footnote{The direct comparison with WDCMT is difficult as the implementation differs in many ways from our RNN: a different framework, different cell types,
-  % (LSTMs vs. conditional GRUs),
-  etc.} \fyDone{More on this.}
 As expected, ignoring the domain label yields a drop in performance for the two fallback strategies considered in this study. 
@@
 The second, \texttt{LDR}$_{pred}$, resorts to automatically predicted domain labels.\footnote{Our domain classifier uses a bi-LSTM RNN encoder, followed by a simple softmax layer.
-Its precision on a development set are higher than 95\%.}
+Its precision on a development set exceeds 95\%.}
 % \fyDone{Explain how}
@@
 %The right parts of Tables~\ref{tab:results-trsf} and \ref{tab:results-rnn} report performance on test sets for new domains not seen in training. 
-Khresmoi texts, which are extracted from summaries of medical articles are assumed to be from the same domain as EMEA data; for the other two test sets, we assume no domain and only use the generic part of the embeddings.
-Here, we see that our best system \texttt{LDR}$_{oracle}$ is slightly lagging behind the \texttt{Mixed} baselines (the difference ranges from unsignificant to one Bleu point), which can be attributed to larger useful (generic) embeddings in the \texttt{Mixed} setting.
-Here again, we see that using the sole generic part makes the system robust with respect to unknown or misleading domain tags.
-We finally outline that our approach outperforms DC for these test sets (were we have to commit to a predicted domain tag), and even outperforms WDCMT (for German-English).
-This might be due to the fact that WDCMT has no built-in mechanism to ignore, as we do in the \texttt{LDR}$_{generic}$ systems, the possibly misleading predicted domain information.
+%Khresmoi texts, which are extracted from summaries of medical articles are assumed to be from the same domain as EMEA data; 
+%for the other two test sets, we assume no domain and only use the generic part of the embeddings. 
+%Here, we see that our best system \texttt{LDR}$_{oracle}$ is slightly lagging behind the \texttt{Mixed} baselines (the difference ranges from unsignificant to one Bleu point), which can be attributed to larger useful (generic) embeddings in the \texttt{Mixed} setting.
+%Here again, we see that using the sole generic part makes the system robust with respect to unknown or misleading domain tags. 
+%We finally outline that our approach outperforms DC for these test sets (were we have to commit to a predicted domain tag), and even outperforms WDCMT (for German-English). 
+%This might be due to the fact that WDCMT has no built-in mechanism to ignore, as we do in the \texttt{LDR}$_{generic}$ systems, the possibly misleading predicted domain information.
+\jcTodo{compute averages}
 
@@
 \hline
-Model & EMEA & EPPS & ECB & Avg. \\ %Khresmoi & NewsTest2014 & IWSLT2010 & Avg. \\
+Model & EMEA & EPPS & ECB & Avg. \\
 \hline
-%\multicolumn{5}{l}{} %\\[-9pt]  
+%\multicolumn{8}{l}{} \\[-9pt]  
 \multicolumn{5}{l}{English$\rightarrow$French} \\
 \hline
-$\mathtt{Mixed}$ & 67.69 & 37.50 & 53.49 & 52.89\\
-$\mathtt{FT}$ & 76.77 & 37.04 & 56.52 & 56.77\\
-$\mathtt{DC}$ & 67.87 & 37.31 & 54.14 & 53.10\\
+$\mathtt{Mixed}$          & 67.69$_{47.60}$ & 37.50 & 53.49 & \\
 \hline
-$\mathtt{LDR}_{oracle}$   & 74.26 & 37.67 & 54.07 & 55.33\\
-$\mathtt{LDR}_{oracle}^{0.5}$   & 74.95 & 37.35 & 55.91 & 56.07\\
-$\mathtt{LDR}_{generic}$ & 73.97 & 37.81 & 53.67 & 55.15\\
-$\mathtt{LDR}_{pred}$ & 74.29 & 37.73 & 54.01 & 55.34\\
-$\mathtt{LDR}_{wrong}$ & 72.95 & 37.62 & 53.35 & 54.64\\
+$\mathtt{FT} (EMEA)$ & 76.77$_{49.43}$ & 17.16 & 11.99 &\\
+$\mathtt{FT} (EPPS)$  & 20.86 & 37.04 & 24.53 &  \\
+$\mathtt{FT} (ECB)$    & 26.93 & 27.09 & 56.52 & \\
 \hline
-%\multicolumn{5}{l}{} %\\[-9pt]
+$\mathtt{DC}$              & 67.87$_{45.42}$ & 37.31 & 54.14 & \\
+\hline
+$\mathtt{LDR}_{oracle}$            & 74.26$_{49.90}$ & 37.67 & 54.07 & \\
+$\mathtt{LDR}_{oracle}^{0.5}$   & 74.95$_{49.38}$ & 37.35 & 55.91 & \\
+$\mathtt{LDR}_{generic}$          & 73.97$_{49.54}$ & 37.81 & 53.67 & \\
+$\mathtt{LDR}_{pred}$               & 74.29 $_{49.84}$ & 37.73 & 54.01 & \\
+$\mathtt{LDR}_{wrong}$            & 72.95$_{49.78}$ & 37.62 & 53.35 & \\
+  \hline
+%\multicolumn{8}{l}{} \\[-9pt]
 \multicolumn{5}{l}{English $\rightarrow$German} \\
 \hline
-$\mathtt{Mixed}$ & 64.57 & 26.47 & 68.67 & 53.23\\
-$\mathtt{FT}$ & 68.35 & 26.29 & 74.05 & 56.23\\
-$\mathtt{DC}$ & 63.48 & 26.27 & 66.95 & 52.23\\
+$\mathtt{Mixed}$         & 64.57$_{42.99}$ & 26.47 & 68.67 & \\
 \hline
-$\mathtt{LDR}_{oracle}$ & 70.90 & 26.30 & 68.90 & 55.36\\
-$\mathtt{LDR}_{oracle}^{0.5}$ & 71.31 & 25.98 & 73.74 & 57.01\\
-$\mathtt{LDR}_{generic}$ & 70.86 & 26.14 & 68.59 & 55.19\\
-$\mathtt{LDR}_{pred}$ & 70.89 & 26.53 & 68.63 & 55.35\\
-$\mathtt{LDR}_{wrong}$ & 69.51 & 26.31 & 66.86 & 54.22\\
+$\mathtt{FT} (EMEA)$ & 68.35$_{42.97}$ & 17.02 & 32.87 & \\
+$\mathtt{FT} (EPPS)$  & 36.19 & 26.29 & 40.71 & \\
+$\mathtt{FT} (ECB)$   & 24.72 & 18.36 & 74.05 & \\
+\hline
+$\mathtt{DC}$ & 63.48$_{42.98}$ & 26.27 & 66.95 & \\
+\hline
+$\mathtt{LDR}_{oracle}$ & 70.90$_{46.12}$ & 26.30 & 68.90 & \\
+$\mathtt{LDR}_{oracle}^{0.5}$ & 71.31$_{45.23}$ & 25.98 & 73.74 & \\
+$\mathtt{LDR}_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & \\
+$\mathtt{LDR}_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & \\
+$\mathtt{LDR}_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & \\
 \hline
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores (Transformer) \label{tab:results-trsf}}
+\caption{BPE-detokenized BLEU scores for  Transformer systems \label{tab:results-trsf}}
 \end{table}
@@
 \hline
-%\multicolumn{5}{l}{} %\\[-9pt]
+%\multicolumn{8}{l}{} \\[-9pt]
 \multicolumn{5}{l}{English$\rightarrow$French} \\
 \hline
-$\mathtt{Mixed}$               & 65.42 & 34.7 & 51.38 & \\
-$\mathtt{FT}$                    & 72.06 & 34.61 & 51.53 & \\
-$\mathtt{DC}$                    & 68.26 & 35.13 & 50.09 & \\
-$\mathtt{WDCMT}$           & 68.76 & 35.71 & 52.75 &\\
+$Mixed$                         & 65.42$_{45.11}$ & 34.70 & 51.38 & \\
 \hline
-$\mathtt{LDR}_{oracle}$    & 71.73 & 35.21 & 50.91 & \\
-$\mathtt{LDR_{condgru}}_{pred}$   & 71.70 & 35.09 & 51.22 & \\
-$\mathtt{LDR}_{oracle}^{0.5}$   & 71.70 & 34.24& 52.37 & \\
-$\mathtt{LDR}_{generic}$  & 70.59 & 35.22 & 49.68 & \\
-$\mathtt{LDR}_{pred}$       & 72.76 & 35.10 & 50.38 & \\
-$\mathtt{LDR}_{wrong}$    & 62.10 & 34.17 & 48.79 &  \\
+$\mathtt{FT} (EMEA)$   & 72.06$_{47.33}$ & 18.62 & 16.78 & \\
+$\mathtt{FT} (EPPS)$    & 35.47 $ $ & 34.61 & 39.56 & \\
+$\mathtt{FT} (ECB)$      & 21.93 $ $ & 22.60 & 51.53 & \\
 \hline
-%\multicolumn{5}{l}{} %\\[-9pt]
+$\mathtt{DC}$                            & 68.26 $_{43.76}$ & 35.13 & 50.09 & \\
+% \hline
+%$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & \\
+\hline
+$\mathtt{LDR}_{oracle}$            & 71.73 $_{46.30}$ & 35.21 & 50.91 & \\
+%$\mathtt{LDR_{condgru}}_{pred}$   & 71.7 $_{46.21}$ & 35.09 & 51.22 & \\
+$\mathtt{LDR}_{oracle}^{0.5}$   & 71.70$_{46.41}$ & 34.24& 52.37 & \\
+$\mathtt{LDR}_{generic}$          & 70.59 $_{46.37}$ & 35.22 & 49.68 & \\
+$\mathtt{LDR}_{pred}$               & 72.76 $_{46.35}$ & 35.10 & 50.38 & \\
+$\mathtt{LDR}_{wrong}$            & 62.10 $_{43.29}$ & 34.17 & 48.79 & \\
+\hline
+%\multicolumn{8}{l}{} \\[-9pt]
 \multicolumn{5}{l}{English$\rightarrow$German} \\
 \hline
-$\mathtt{Mixed}$                & 57.37 & 23.1 & 63.54 & \\
+$Mixed$                        & 57.37 $_{37.94}$ & 23.10 & 63.54 & \\
 \hline
-$\mathtt{FT} (EMEA)$        & 65.64 & 12.36 & 15.93 & \\
-$\mathtt{FT} (EPPS)$        & 24.90 & 22.98 & 26.26 & \\
-$\mathtt{FT} (ECB)$          & 41.80  & 15.97 & 71.07 & \\
-$\mathtt{DC}$                     & 62.53 & 23.74 & 65.71 & \\
+$\mathtt{FT} (EMEA)$  & 65.64$_{44.71}$ & 12.36 & 15.93 & \\
+$\mathtt{FT} (EPPS)$   & 24.90 $ $ & 22.98 & 26.26 & \\
+$\mathtt{FT} (ECB)$     & 41.80 $ $ & 15.97 & 71.07 & \\
 \hline
-$\mathtt{WDCMT}$            & 64.05 & 22.82 & 67.00 & \\
+$\mathtt{DC}$                & 62.53 $_{39.25}$ & 23.74 & 65.71 & \\
+%\hline
+%$\mathtt{WDCMT}$ & 64.05 $_{38.54}$ & 22.82 & 67.00 & \\
 \hline
-$\mathtt{LDR}_{oracle}$    & 63.43 & 22.66 & 64.40 & \\
-$\mathtt{LDR}_{oracle}^{0.5}$   & 63.27 & 21.83 & 69.55 & \\
-$\mathtt{LDR}_{generic}$  & 63.27 & 22.62 & 63.70 & \\
-$\mathtt{LDR}_{pred}$       & 63.17 & 22.51 & 64.00 & \\
-$\mathtt{LDR}_{wrong}$    & 56.84 & 22.06 & 61.66 & \\
+$\mathtt{LDR}_{oracle}$     & 63.43 $_{40.04}$ & 22.66 & 64.40 & \\
+$\mathtt{LDR}_{oracle}^{0.5}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & \\
+$\mathtt{LDR}_{generic}$   & 63.27 $_{39.75}$ & 22.62 & 63.70 & \\
+$\mathtt{LDR}_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64.00 & \\
+$\mathtt{LDR}_{wrong}$     & 56.84 $_{37.05}$ & 22.06 & 61.66 & \\
 \hline
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores (RNN)\label{tab:results-rnn}}
+\caption{BPE-detokenized BLEU scores for RNN systems\label{tab:results-rnn}}
 \end{table}
@@
 
+\jcTodo{introduce WDCMT experiments}
+Finally the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT}).
+The comparison with the \texttt{WDCMT} model is less favourable for our RNN systems, which are arguably much simpler.
+The direct comparison with WDCMT is difficult as the implementation differs in many ways from our RNN: a different framework, different cell types,  etc.
+  % (LSTMs vs. conditional GRUs),
+
+\begin{table}[!h]
+\begin{center}
+\scalebox{1.0}{
+\begin{tabular}{|l|ccc|c|}
+\hline
+Model & EMEA & EPPS & ECB & Avg. \\
+\hline
+%\multicolumn{8}{l}{} \\[-9pt]
+\multicolumn{5}{l}{English$\rightarrow$French} \\
+\hline
+$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & \\
+$\mathtt{LDR_{condgru}}_{pred}$   & 71.70 $_{46.21}$ & 35.09 & 51.22 & \\
+\hline
+\end{tabular}
+} %scalebox
+\end{center}
+\caption{BPE-detokenized BLEU scores for RNN systems\label{tab:results-rnn-wdcmt}}
+\end{table}
+
 \section{Complementary experiments\label{sec:Discussion}}
+
 % \subsection{Sparse's performance}
@@
 \hline
- size=4   & 74.65 & 37.42 & 54.49 \\
- size=8   & 74.26 & 37.67 & 54.07 \\
- size=16 & 74.15 & 37.78 & 54.56 \\
- size=32 & 75.10 & 37.64 & 54.29 \\
- size=64 & 74.50 & 37.27 & 54.50 \\
+ size=4   & 74.65 $_{49.61}$ & 37.42 & 54.49 \\
+ size=8   & 74.26 $_{49.90}$ & 37.67 & 54.07 \\
+ size=16 & 74.15 $_{49.10}$ & 37.78 & 54.56 \\
+ size=32 & 75.10 $_{48.61}$ & 37.64 & 54.29 \\
+ size=64 & 74.50 $_{50.17}$ & 37.27 & 54.50 \\
 \hline
@@
 
+\subsection{Additional domain experiment \label{ssec:additiona_domain}}
+
 \subsection{Analysis of Word Embeddings \label{ssec:word_embeddings}}
@@
 
-To check the validity of this hypothesis, we performed the following analysis of embeddings learned with the multi-domain Transformer system for English:French.
+To check this hypothesis, we performed the following analysis of embeddings learned with the multi-domain Transformer system for English:French.
 For each unit\footnote{In this study, we work with BPE units, not words, meaning that in many cases we observe the variation of use of word parts.
@@
   \centering
-  \begin{tabularx}{1.0\linewidth}{c|c}
+%  \begin{tabularx}{1.0\linewidth}{c|c}
+  \begin{tabular}{c|c}
     Polysemic ``words'' & Monosemic ``words'' \\ \hline
@@
      mation (2) &           better (10) \\ \hline
-  \end{tabularx}
+  \end{tabular}
   \caption{Analyzing the variation of embeddings across domains. Each word or subword is listed with the size of the intersection (between 0 and 10).}
@@
 %
-\bibliography{iwslt18}
+\bibliography{iwslt19}
 \bibliographystyle{IEEEtran}

### 2019-08-26 2019/iwslt/submission.tex
additional domain scenario

@@
 \scalebox{1.0}{
-\begin{tabular}{|l|ccc|}
+\begin{tabular}{|l|ccc|c|}
 \hline
-$LDR$ & EMEA & EPPS & ECB \\
+$LDR$ & EMEA & EPPS & ECB & Avg. \\
 \hline
-\multicolumn{4}{l}{English$\rightarrow$French} \\
+\multicolumn{5}{l}{English$\rightarrow$French} \\
 \hline
- size=4   & 74.65 $_{49.61}$ & 37.42 & 54.49 \\
- size=8   & 74.26 $_{49.90}$ & 37.67 & 54.07 \\
- size=16 & 74.15 $_{49.10}$ & 37.78 & 54.56 \\
- size=32 & 75.10 $_{48.61}$ & 37.64 & 54.29 \\
- size=64 & 74.50 $_{50.17}$ & 37.27 & 54.50 \\
+ size=4   & 74.65 $_{49.61}$ & 37.42 & 54.49 & \\
+ size=8   & 74.26 $_{49.90}$ & 37.67 & 54.07 & \\
+ size=16 & 74.15 $_{49.10}$ & 37.78 & 54.56 & \\
+ size=32 & 75.10 $_{48.61}$ & 37.64 & 54.29 & \\
+ size=64 & 74.50 $_{50.17}$ & 37.27 & 54.50 & \\
 \hline
@@
 
-\subsection{Additional domain experiment \label{ssec:additiona_domain}}
+\subsection{Additional Domain Scenario \label{ssec:additiona_domain}}
+
+We evaluate now the ability of our model to cope with new unseen domains, a very common scenario for industrial MT.
+We simulate the apparition of the IT domain (see Section~\ref{sec:data}) once the $\mathtt{LDR}_{oracle}$ model is built for EMEA, EPPS and ECB domains.
+Note that to accommodate a new domain the network must reserve the new IT region from the beginning (never updated).
+Hence, word embeddings contain 32 cells (8 x 4 domain regions) and 480 cells for the generic region.
+Equivalently to our previous $\mathtt{LDR}_{oracle}$ model, 200,000 iterations are performed using EMEA, EPPS and ECB data sets.
+Then, 100,000 additional iterations are performed using  EMEA, EPPS, ECB and IT data sets ($\mathtt{LDR}_{oracle}\ cont$).
+Experiments are only reported for the English$\rightarrow$French language pair.
+For comparison purposes we also continue training our $\mathtt{Mixed}$ model applying the same number of iterations using the same four datasets ($\mathtt{Mixed}\ cont$).
+BLEU results are shown in Table~\ref{tab:add}.
+
+\begin{table}[!h]
+\begin{center}
+\scalebox{0.9}{
+\begin{tabular}{|l|cccc|c|}
+\hline
+Model & EMEA & EPPS & ECB & IT & Avg.\\
+\hline
+\multicolumn{5}{l}{English$\rightarrow$French} \\
+\hline
+$\mathtt{Mixed}$                     & 67.69$_{47.60}$ & 37.50 & 53.49 & 13.91 & \\
+$\mathtt{Mixed}\ cont$             & 66.49$_{45.79}$ & 37.59 & 55.07 & 51.78 & \\
+\hline
+$\mathtt{LDR}_{oracle}$          & 74.26$_{49.90}$ & 37.67 & 54.07 & 13.40 & \\
+$\mathtt{LDR}_{oracle}\ cont$  & 76.17$_{49.71}$ & 37.48 & 55.12 & 55.24 & \\
+\hline
+\end{tabular}
+} %scalebox
+\end{center}
+\caption{BPE-detokenized BLEU scores for the Transformer architecture for varying domain-specific embedding sizes \label{tab:add}}
+\end{table}
+
+Huge gap in IT test sets thanks to using in-domain data.
+Light difference for EPPS test sets (as before).
+Both $\mathtt{Mixed}$ and $\mathtt{LDR}_{oracle}$ systems show similar gains on ECB when further trained with IT data.
+Only $\mathtt{LDR}_{oracle}$ shows gains on EMEA when further training (only on the large test set).
+\jcTodo{rewrite this part}
 
 \subsection{Analysis of Word Embeddings \label{ssec:word_embeddings}}
-One of our main assumptions is that the difference between domains can be confined at the lexical level, warranting our decision to specialize lexical representations for each domain, while the remaining part of the network is shared across domains.
+One of our main assumptions is that the difference between domains can be confined at the lexical level, warranting our decision to specialise lexical representations for each domain, while the remaining part of the network is shared across domains.
 Linguistically, this assumption relates to the classical ``one sense per collocation'' \cite{Yarowsky93onesense} and corresponds to the fact that in many cases, polysemy corresponds to variation of use across domain.
-In its weaker form, it allows us to assume that all occurrences of a given form in a given domain correspond to the same sense and share the same representation; the same form occuring in different domains is allowed to have one distinct embedding per domain, which may help capture polysemy and lexical ambiguity in translation.
+In its weaker form, it allows us to assume that all occurrences of a given form in a given domain correspond to the same sense and share the same representation; the same form occurring in different domains is allowed to have one distinct embedding per domain, which may help capture polysemy and lexical ambiguity in translation.
 
@@
 The decoder side can thus attend separately to these two representations to generate its output. 
-In this approach, the shared and domain-specific part kept separated in the deeper layers of the network, whereas we try to localize the differences between domains at the lexical level, based on a much cheaper computational architecture.
-Another trait of this proposal is the ability to automatically infer domain information for test sentences; as we have shown, our architecture can also effectively accomodate sentences lacking domain information.
+In this approach, the shared and domain-specific part kept separated in the deeper layers of the network, whereas we try to localise the differences between domains at the lexical level, based on a much cheaper computational architecture.
+Another trait of this proposal is the ability to automatically infer domain information for test sentences; 
+as we have shown, our architecture can also effectively accommodate sentences lacking domain information.
 
@@
 Our experiments have shown that for both architecture and for two language pairs, our multi-domain models improve over several baselines of the literature, and that it is robust to noise in domain labels.
-In our future work, we intend to develop this work so as to make the architecture more self-configurable and able to adapt the size of the domain-specific regions depending upon the actual variation of use across domains; we also would like to find ways to make the architecture able to integrate new domains in a dynamic fashion, as this is an important requirement in industrial systems.
+In our future work, we intend to develop this work so as to make the architecture more self-configurable and able to adapt the size of the domain-specific regions depending upon the actual variation of use across domains;
+we also would like to find additional ways to make the architecture able to integrate new domains in a dynamic fashion, as this is an important requirement in industrial systems.
 
-\section{Acknowledgements}
-The IWSLT 2015 organizing committee would like to thank the
-organizing committees of INTERSPEECH 2004 for their
-help and for kindly providing the template files.
+%\section{Acknowledgements}
+%The IWSLT 2015 organizing committee would like to thank the
+%organizing committees of INTERSPEECH 2004 for their
+%help and for kindly providing the template files.
 

### 2019-08-26 2019/iwslt/submission.tex
additional domain scenario

@@
 we provide experimental evidence that show the effectiveness of this technique;
-\jcTodo{focus on that the analysis of embeddings is an original technique.}
+\jcTodo{highlight the analysis of embeddings as an original technique.}
 our analysis of embeddings comforts our assumption that their variation across domains actually reflects a variation of senses.
+\jcTodo{introduce the additional domain scenario experiment.}
 
@@
 Those regions are also masked in forward step, thus do not interfere the training on the domain to which they are not assigned (see Figure~\ref{fig:network}).
+%
+Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers the computation in equation~\eqref{eq:embedding}.
+In our experiments, we consider both the attentional RNN architecture of \cite{Bahdanau15learning} and for Transformer architecture of \cite{Vaswani17attention}.
 
@@
 \end{figure}
-
-Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers the computation in equation~\eqref{eq:embedding}.
-In our experiments, we consider both the attentional RNN architecture of \cite{Bahdanau15learning} and for Transformer architecture of \cite{Vaswani17attention}.
 
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for  Transformer systems \label{tab:results-trsf}}
+\caption{BLEU scores for  Transformer systems \label{tab:results-trsf}}
 \end{table}
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for RNN systems\label{tab:results-rnn}}
+\caption{BLEU scores for RNN systems\label{tab:results-rnn}}
 \end{table}
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for RNN systems\label{tab:results-rnn-wdcmt}}
+\caption{BLEU scores for RNN systems\label{tab:results-rnn-wdcmt}}
 \end{table}
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for the Transformer architecture for varying domain-specific embedding sizes \label{tab:embedding-size}}
+\caption{BLEU scores for the Transformer architecture for varying domain-specific embedding sizes \label{tab:embedding-size}}
 \end{table}
@@
 We evaluate now the ability of our model to cope with new unseen domains, a very common scenario for industrial MT.
-We simulate the apparition of the IT domain (see Section~\ref{sec:data}) once the $\mathtt{LDR}_{oracle}$ model is built for EMEA, EPPS and ECB domains.
-Note that to accommodate a new domain the network must reserve the new IT region from the beginning (never updated).
+We simulate the apparition of the IT domain (see Section~\ref{ssec:data}) once the $\mathtt{LDR}_{oracle}$ model is built for EMEA, EPPS and ECB domains.
+Note that to accommodate a new domain the network must reserve the new IT region from the beginning.
 Hence, word embeddings contain 32 cells (8 x 4 domain regions) and 480 cells for the generic region.
-Equivalently to our previous $\mathtt{LDR}_{oracle}$ model, 200,000 iterations are performed using EMEA, EPPS and ECB data sets.
-Then, 100,000 additional iterations are performed using  EMEA, EPPS, ECB and IT data sets ($\mathtt{LDR}_{oracle}\ cont$).
+Equivalently to our previous $\mathtt{LDR}_{oracle}$ model, 200,000 iterations are performed using EMEA, EPPS and ECB data sets (the IT region is never updated).
+Then, 100,000 additional iterations are performed using  EMEA, EPPS, ECB and IT data sets ($\mathtt{LDR}_{oracle}^*$).
 Experiments are only reported for the English$\rightarrow$French language pair.
-For comparison purposes we also continue training our $\mathtt{Mixed}$ model applying the same number of iterations using the same four datasets ($\mathtt{Mixed}\ cont$).
+For comparison purposes we also continue training our $\mathtt{Mixed}$ model applying the same number of iterations and using the same four datasets ($\mathtt{Mixed}^*$).
 BLEU results are shown in Table~\ref{tab:add}.
@@
 $\mathtt{Mixed}$                     & 67.69$_{47.60}$ & 37.50 & 53.49 & 13.91 & \\
-$\mathtt{Mixed}\ cont$             & 66.49$_{45.79}$ & 37.59 & 55.07 & 51.78 & \\
+$\mathtt{Mixed}^*$             & 66.49$_{45.79}$ & 37.59 & 55.07 & 51.78 & \\
 \hline
 $\mathtt{LDR}_{oracle}$          & 74.26$_{49.90}$ & 37.67 & 54.07 & 13.40 & \\
-$\mathtt{LDR}_{oracle}\ cont$  & 76.17$_{49.71}$ & 37.48 & 55.12 & 55.24 & \\
+$\mathtt{LDR}_{oracle}^*$  & 76.17$_{49.71}$ & 37.48 & 55.12 & 55.24 & \\
 \hline
@@
 \end{center}
-\caption{BPE-detokenized BLEU scores for the Transformer architecture for varying domain-specific embedding sizes \label{tab:add}}
+\caption{BLEU scores for the Transformer architecture for varying domain-specific embedding sizes \label{tab:add}}
 \end{table}
 
-Huge gap in IT test sets thanks to using in-domain data.
-Light difference for EPPS test sets (as before).
-Both $\mathtt{Mixed}$ and $\mathtt{LDR}_{oracle}$ systems show similar gains on ECB when further trained with IT data.
-Only $\mathtt{LDR}_{oracle}$ shows gains on EMEA when further training (only on the large test set).
+As expected, a huge gap is observed over the IT test sets when using in-domain data.
+Note that $\mathtt{LDR}_{oracle}^*$ clearly outperforms $\mathtt{Mixed}^*$ by a large margin.
+As previously, all models show very little differences on the EPPS test set.
+Both models show similar results over the ECB test set when further trained with IT data.
+Only our $\mathtt{LDR}_{oracle}$ model shows gains on the EMEA test set when further trained (only on the large test set).
+Overall better results for our LDR when further trained.
 \jcTodo{rewrite this part}

### 2019-08-27 2019/iwslt/submission.tex
additional domain scenario

@@
 \section{Introduction \label{sec:introduction}}
-Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,Bahdanau15learning,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progresses over the past years, and constitute to date the standard for most production engines.
+Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,Bahdanau15learning,Ghering17convolutional,Vaswani17attention}, 
+Machine Translation (MT) has made significant progresses over the past years, and constitute to date the standard for most production engines. 
 The development of MT systems, be they neural or statistical, require very large parallel corpora consisting of millions of sentence pairs, a resource that only exist in very few application domains and language pairs.
-A lot of the recent research effort has thus focused on developping MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Ha16towards,Johnson17google};
+A lot of the recent research effort has thus focused on developing MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Ha16towards,Johnson17google};
 Another important scenario for industrial MT is to adapt a neural system trained using parallel data in one domain to the peculiarity of other domains.
+
 Domain Adaption (DA) in MT is an old issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied. 
@@
 Multi-domain adaptation is conceptually close to multilingual MT, or more generally to multi-task learning \cite{Caruana97multitask}, and can be approached in a number of ways.
+
 We adapt here ideas of \cite{Daume07frustratingly} to neural MT.
@@
 we provide experimental evidence that show the effectiveness of this technique;
-\jcTodo{highlight the analysis of embeddings as an original technique.}
-our analysis of embeddings comforts our assumption that their variation across domains actually reflects a variation of senses.
-\jcTodo{introduce the additional domain scenario experiment.}
+we evaluate the ability of our network to efficiently accommodate new domains;
+and we introduce a novel technique to analyse word polysemy using word embeddings which comforts our assumption that their variation across domains actually reflects a variation of senses.
 
@@
 
-Making sure that the actual embedding do not contain any zero is important for the Transformer model, since the lexical representations are then summed to the positional encoding, which would undo the effect of domain masking, and propagate a gradient even to regions that should not be modified.
-With our design, we make sure that during backprogation, the matrix $M$ receives gradient $0$ at regions corresponding to deactivated regions in the word embedding. 
-Those regions are also masked in forward step, thus do not interfere the training on the domain to which they are not assigned (see Figure~\ref{fig:network}).
-%
-Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers the computation in equation~\eqref{eq:embedding}.
-In our experiments, we consider both the attentional RNN architecture of \cite{Bahdanau15learning} and for Transformer architecture of \cite{Vaswani17attention}.
-
 \begin{figure}[h]
@@
 \end{figure}
+
+Making sure that the actual embedding do not contain any zero is important for the Transformer model, since the lexical representations are then summed to the positional encoding, which would undo the effect of domain masking, and propagate a gradient even to regions that should not be modified.
+With our design, we make sure that during backprogation, the matrix $M$ receives gradient $0$ at regions corresponding to deactivated regions in the word embedding. 
+Those regions are also masked in forward step, thus do not interfere the training on the domain to which they are not assigned (see Figure~\ref{fig:network}).
+%
+Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers the computation in equation~\eqref{eq:embedding}.
+In our experiments, we consider both the attentional RNN architecture of \cite{Bahdanau15learning} and for Transformer architecture of \cite{Vaswani17attention}.
 
@@
 In addition, for English-French we also used IT-domain corpora obtained from the OPUS web site\footnote{\url{http://opus.nlpl.eu}} corresponding to KDE, Ubuntu, GNOME and PHP datasets (IT).
-
+%
 We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).
 Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.
-Using internal test sets corresponds to somewhat ideal conditions.
+%Using internal test sets corresponds to somewhat ideal conditions.
 
@@
 \end{table}
-
-\jcTodo{I delete vocabs since they make no sense if computed after BPE}
 
@@
 %This might be due to the fact that WDCMT has no built-in mechanism to ignore, as we do in the \texttt{LDR}$_{generic}$ systems, the possibly misleading predicted domain information.
-\jcTodo{compute averages}
+\jcTodo{compute averages on all tables}
 
@@
 
-\jcTodo{introduce WDCMT experiments}
-Finally the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT}).
-The comparison with the \texttt{WDCMT} model is less favourable for our RNN systems, which are arguably much simpler.
-The direct comparison with WDCMT is difficult as the implementation differs in many ways from our RNN: a different framework, different cell types,  etc.
-  % (LSTMs vs. conditional GRUs),
+We also compare our presented network with the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT}) for English$\rightarrow$French translation.
+\jcTodo{(explain why conditional GRU are used for the comparison)},
+Results are shown by Table~\ref{tab:results-rnn-wdcmt}. 
+The comparison is less favourable for our RNN systems, which are arguably much simpler.\footnote{The direct comparison with \texttt{WDCMT} is difficult as the implementation differs in many ways from our RNN: framework, cell types,  etc.}
 
@@
 \end{center}
-\caption{BLEU scores for RNN systems\label{tab:results-rnn-wdcmt}}
+\caption{BLEU scores when comparing \texttt{LDR} to \texttt{WDCMT}.\label{tab:results-rnn-wdcmt}}
 \end{table}
@@
 \hline
-$LDR$ & EMEA & EPPS & ECB & Avg. \\
+$\mathtt{LDR}_{oracle}$ & EMEA & EPPS & ECB & Avg. \\
 \hline
@@
 
-We evaluate now the ability of our model to cope with new unseen domains, a very common scenario for industrial MT.
-We simulate the apparition of the IT domain (see Section~\ref{ssec:data}) once the $\mathtt{LDR}_{oracle}$ model is built for EMEA, EPPS and ECB domains.
-Note that to accommodate a new domain the network must reserve the new IT region from the beginning.
-Hence, word embeddings contain 32 cells (8 x 4 domain regions) and 480 cells for the generic region.
+We evaluate now the ability of our model to integrate new domains, a very common scenario for industrial MT.
+We simulate the apparition of the IT domain once the $\mathtt{LDR}_{oracle}$ model is built for EMEA, EPPS and ECB domains.
+Note that to accommodate a new domain the network must be designed with a region reserved for the new domain.
+Hence, for this experiment word embeddings contain 480 cells for the generic region and 32 cells for domain specific regions (8 cells x 4 regions).
 Equivalently to our previous $\mathtt{LDR}_{oracle}$ model, 200,000 iterations are performed using EMEA, EPPS and ECB data sets (the IT region is never updated).
-Then, 100,000 additional iterations are performed using  EMEA, EPPS, ECB and IT data sets ($\mathtt{LDR}_{oracle}^*$).
-Experiments are only reported for the English$\rightarrow$French language pair.
+Then, 100,000 additional iterations are performed using  EMEA, EPPS, ECB and IT datasets ($\mathtt{LDR}_{oracle}^*$).
+Experiments are reported for the English$\rightarrow$French language pair.
 For comparison purposes we also continue training our $\mathtt{Mixed}$ model applying the same number of iterations and using the same four datasets ($\mathtt{Mixed}^*$).
@@
 \end{center}
-\caption{BLEU scores for the Transformer architecture for varying domain-specific embedding sizes \label{tab:add}}
+\caption{BLEU scores for the Transformer architecture when including IT as additional domain \label{tab:add}}
 \end{table}
 
-As expected, a huge gap is observed over the IT test sets when using in-domain data.
-Note that $\mathtt{LDR}_{oracle}^*$ clearly outperforms $\mathtt{Mixed}^*$ by a large margin.
-As previously, all models show very little differences on the EPPS test set.
-Both models show similar results over the ECB test set when further trained with IT data.
-Only our $\mathtt{LDR}_{oracle}$ model shows gains on the EMEA test set when further trained (only on the large test set).
-Overall better results for our LDR when further trained.
-\jcTodo{rewrite this part}
+As expected, a huge gap is observed over the IT test set when learning includes in-domain data.
+Note that $\mathtt{LDR}_{oracle}^*$ clearly outperforms $\mathtt{Mixed}^*$ by a large margin over the IT test set.
+%
+Considering the other test sets: both models similarly enhance their performance on the ECB test set;
+$\mathtt{LDR}_{oracle}^*$ shows gains over the EMEA test set while a loss of quality is observed on $\mathtt{Mixed}^*$;
+using IT data does not impact the quality of translations of the EPPS test set of any of the models.
+Overall better results are obtained by our $\mathtt{LDR}_{oracle}$ model when trained including the additional domain dataset.
+New domains are easily integrated in the model if the corresponding embedding regions are reserved in advance.
 

### 2019-08-27 2019/iwslt/submission.tex
additional domain scenario

@@
 \item generic models trained with a simple concatenation of all corpora (\texttt{Mixed}); 
-\item models tuned separately on each domain using respectively (10000, 15000, 5000) iterations over in-domain data (\texttt{ft(EMEA)}, \texttt{ft(EPPS)}, \texttt{ft(ECB)}); 
+\item models tuned separately on each domain using respectively (10000, 15000, 5000) iterations over in-domain data (\texttt{ft$_{EMEA}$}, \texttt{ft$_{EPPS}$}, \texttt{ft$_{ECB}$}); 
 \item models using domain tags as in \cite{Kobus17domaincontrol} (\texttt{DC}); 
@@
 \hline
-$\mathtt{FT} (EMEA)$ & 76.77$_{49.43}$ & 17.16 & 11.99 &\\
-$\mathtt{FT} (EPPS)$  & 20.86 & 37.04 & 24.53 &  \\
-$\mathtt{FT} (ECB)$    & 26.93 & 27.09 & 56.52 & \\
+$\mathtt{FT}_{EMEA}$ & 76.77$_{49.43}$ & 17.16 & 11.99 &\\
+$\mathtt{FT}_{EPPS}$  & 20.86 & 37.04 & 24.53 &  \\
+$\mathtt{FT}_{ECB}$    & 26.93 & 27.09 & 56.52 & \\
 \hline
@@
 \hline
-$\mathtt{FT} (EMEA)$ & 68.35$_{42.97}$ & 17.02 & 32.87 & \\
-$\mathtt{FT} (EPPS)$  & 36.19 & 26.29 & 40.71 & \\
-$\mathtt{FT} (ECB)$   & 24.72 & 18.36 & 74.05 & \\
+$\mathtt{FT}_{EMEA}$ & 68.35$_{42.97}$ & 17.02 & 32.87 & \\
+$\mathtt{FT}_{EPPS}$  & 36.19 & 26.29 & 40.71 & \\
+$\mathtt{FT}_{ECB}$   & 24.72 & 18.36 & 74.05 & \\
 \hline
@@
 \hline
-$\mathtt{FT} (EMEA)$   & 72.06$_{47.33}$ & 18.62 & 16.78 & \\
-$\mathtt{FT} (EPPS)$    & 35.47 $ $ & 34.61 & 39.56 & \\
-$\mathtt{FT} (ECB)$      & 21.93 $ $ & 22.60 & 51.53 & \\
+$\mathtt{FT}_{EMEA}$   & 72.06$_{47.33}$ & 18.62 & 16.78 & \\
+$\mathtt{FT}_{EPPS}$    & 35.47 $ $ & 34.61 & 39.56 & \\
+$\mathtt{FT}_{ECB}$      & 21.93 $ $ & 22.60 & 51.53 & \\
 \hline
@@
 \hline
-$\mathtt{FT} (EMEA)$  & 65.64$_{44.71}$ & 12.36 & 15.93 & \\
-$\mathtt{FT} (EPPS)$   & 24.90 $ $ & 22.98 & 26.26 & \\
-$\mathtt{FT} (ECB)$     & 41.80 $ $ & 15.97 & 71.07 & \\
+$\mathtt{FT}_{EMEA}$  & 65.64$_{44.71}$ & 12.36 & 15.93 & \\
+$\mathtt{FT}_{EPPS}$   & 24.90 $ $ & 22.98 & 26.26 & \\
+$\mathtt{FT}_{ECB}$     & 41.80 $ $ & 15.97 & 71.07 & \\
 \hline
@@
 We therefore decided to allocate $8$~cells for the domain specific part.
-This suggests that we could easily accomodate more domains with the same architecture, and even reserve some regions to handle supplementary incoming training data.
+This suggests that we could easily accommodate more domains with the same architecture, and even reserve some regions to handle supplementary incoming training data.
 % \fyTodo{Easily accomodate more domains}
@@
 Overall better results are obtained by our $\mathtt{LDR}_{oracle}$ model when trained including the additional domain dataset.
-New domains are easily integrated in the model if the corresponding embedding regions are reserved in advance.
+Thus, confirming the ability of the model to easily integrate new domains as supplementary training data becomes available.
+%New domains are easily integrated in the model if the corresponding embedding regions are reserved in advance.
 

### 2019-08-27 2019/iwslt/submission.tex
additional domain scenario

@@
 \texttt{LDR}$_{oracle}$ is also better than Domain Control in three conditions out of four, \texttt{DC} being seemingly a better choice for the RNN than for the Transformer architecture. 
-As expected, ignoring the domain label yields a drop in performance for the two fallback strategies considered in this study. 
-The first, \texttt{LDR}$_{generic}$, only uses the generic part of the embeddings. 
-The second, \texttt{LDR}$_{pred}$, resorts to automatically predicted domain labels.\footnote{Our domain classifier uses a bi-LSTM RNN encoder, followed by a simple softmax layer.
+As expected, ignoring the true domain label yields a light drop in performance. %for the two fallback strategies considered in this study. 
+%The first, \texttt{LDR}$_{generic}$, only uses the generic part of the embeddings. 
+%The second, 
+\texttt{LDR}$_{pred}$ employs automatically predicted domain labels.\footnote{Our domain classifier uses a bi-LSTM RNN encoder, followed by a simple softmax layer. 
 Its precision on a development set exceeds 95\%.}
 % \fyDone{Explain how}
-Note that this decrease is however quite limited (and hardly significant for \texttt{LDR}$_{pred}$), showing that our architecture is quite robust to noisy labels.
+Note that this decrease is however hardly significant, showing that our architecture is quite robust to noisy labels. 
 Even in the worst case scenario where all domain tags are intentionally wrong (\texttt{LDR}$_{wrong}$), we see that the generic part still ensures a satisfying level of performance. 
@@
 $\mathtt{LDR}_{oracle}^{0.5}$   & 74.95$_{49.38}$ & 37.35 & 55.91 & \\
-$\mathtt{LDR}_{generic}$          & 73.97$_{49.54}$ & 37.81 & 53.67 & \\
+%$\mathtt{LDR}_{generic}$          & 73.97$_{49.54}$ & 37.81 & 53.67 & \\
 $\mathtt{LDR}_{pred}$               & 74.29 $_{49.84}$ & 37.73 & 54.01 & \\
@@
 %\multicolumn{8}{l}{} \\[-9pt]
-\multicolumn{5}{l}{English $\rightarrow$German} \\
+\multicolumn{5}{l}{English$\rightarrow$German} \\
 \hline
@@
 $\mathtt{LDR}_{oracle}^{0.5}$ & 71.31$_{45.23}$ & 25.98 & 73.74 & \\
-$\mathtt{LDR}_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & \\
+%$\mathtt{LDR}_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & \\
 $\mathtt{LDR}_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & \\
@@
 $\mathtt{LDR}_{oracle}^{0.5}$   & 71.70$_{46.41}$ & 34.24& 52.37 & \\
-$\mathtt{LDR}_{generic}$          & 70.59 $_{46.37}$ & 35.22 & 49.68 & \\
+%$\mathtt{LDR}_{generic}$          & 70.59 $_{46.37}$ & 35.22 & 49.68 & \\
 $\mathtt{LDR}_{pred}$               & 72.76 $_{46.35}$ & 35.10 & 50.38 & \\
@@
 $\mathtt{LDR}_{oracle}^{0.5}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & \\
-$\mathtt{LDR}_{generic}$   & 63.27 $_{39.75}$ & 22.62 & 63.70 & \\
+%$\mathtt{LDR}_{generic}$   & 63.27 $_{39.75}$ & 22.62 & 63.70 & \\
 $\mathtt{LDR}_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64.00 & \\
@@
 We also compare our presented network with the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT}) for English$\rightarrow$French translation.
-\jcTodo{(explain why conditional GRU are used for the comparison)},
+However, the comparison with \texttt{WDCMT} is difficult as the implementation differs in many ways from our RNN: framework, cell types,  etc.
+Thus, in order to compare closer architectures we replace our model to use a single GRU layer in the encoder and a decoder with a single conditional GRU layer ($\mathtt{LDR}_{pred}^{condgru}$).
 Results are shown by Table~\ref{tab:results-rnn-wdcmt}.
-The comparison is less favourable for our RNN systems, which are arguably much simpler.\footnote{The direct comparison with \texttt{WDCMT} is difficult as the implementation differs in many ways from our RNN: framework, cell types,  etc.}
+The comparison is less favourable for our RNN system, which is arguably much simpler.
 
@@
 \hline
+$\mathtt{LDR}_{pred}^{condgru}$  & 71.70 $_{46.21}$ & 35.09 & 51.22 & \\
 $\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & \\
-$\mathtt{LDR_{condgru}}_{pred}$   & 71.70 $_{46.21}$ & 35.09 & 51.22 & \\
 \hline
@@
 \end{center}
-\caption{BLEU scores when comparing \texttt{LDR} to \texttt{WDCMT}.\label{tab:results-rnn-wdcmt}}
+\caption{BLEU scores for RNN systems. Comparison between a version of $\texttt{LDR}_{pred}$ using conditional GRUs and \texttt{WDCMT}.\label{tab:results-rnn-wdcmt}}
 \end{table}
@@
 Overall better results are obtained by our $\mathtt{LDR}_{oracle}$ model when trained including the additional domain dataset.
-Thus, confirming the ability of the model to easily integrate new domains as supplementary training data becomes available.
+Thus, confirming the ability of the model to easily integrate new domains. % as supplementary training data becomes available.
 %New domains are easily integrated in the model if the corresponding embedding regions are reserved in advance.

### 2019-08-27 2019/iwslt/submission.tex
additional domain scenario

@@
 \scalebox{1.0}{
-\begin{tabular}{|l|ccc|c|}
+\begin{tabular}{|l|lcc|c|}
 \hline
@@
 %$\mathtt{LDR}_{generic}$          & 73.97$_{49.54}$ & 37.81 & 53.67 & \\
-$\mathtt{LDR}_{pred}$               & 74.29 $_{49.84}$ & 37.73 & 54.01 & \\
+$\mathtt{LDR}_{pred}$               & 74.29$_{49.84}$ & 37.73 & 54.01 & \\
 $\mathtt{LDR}_{wrong}$            & 72.95$_{49.78}$ & 37.62 & 53.35 & \\
@@
 $\mathtt{FT}_{EMEA}$   & 72.06$_{47.33}$ & 18.62 & 16.78 & \\
-$\mathtt{FT}_{EPPS}$    & 35.47 $ $ & 34.61 & 39.56 & \\
-$\mathtt{FT}_{ECB}$      & 21.93 $ $ & 22.60 & 51.53 & \\
+$\mathtt{FT}_{EPPS}$    & 35.47$ $ & 34.61 & 39.56 & \\
+$\mathtt{FT}_{ECB}$      & 21.93$ $ & 22.60 & 51.53 & \\
 \hline
-$\mathtt{DC}$                            & 68.26 $_{43.76}$ & 35.13 & 50.09 & \\
+$\mathtt{DC}$                            & 68.26$_{43.76}$ & 35.13 & 50.09 & \\
 % \hline
-%$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & \\
+%$\mathtt{WDCMT}$           & 68.76$_{45.29}$ & 35.71 & 52.75 & \\
 \hline
-$\mathtt{LDR}_{oracle}$            & 71.73 $_{46.30}$ & 35.21 & 50.91 & \\
-%$\mathtt{LDR_{condgru}}_{pred}$   & 71.7 $_{46.21}$ & 35.09 & 51.22 & \\
+$\mathtt{LDR}_{oracle}$            & 71.73$_{46.30}$ & 35.21 & 50.91 & \\
+%$\mathtt{LDR_{condgru}}_{pred}$   & 71.7$_{46.21}$ & 35.09 & 51.22 & \\
 $\mathtt{LDR}_{oracle}^{0.5}$   & 71.70$_{46.41}$ & 34.24& 52.37 & \\
-%$\mathtt{LDR}_{generic}$          & 70.59 $_{46.37}$ & 35.22 & 49.68 & \\
-$\mathtt{LDR}_{pred}$               & 72.76 $_{46.35}$ & 35.10 & 50.38 & \\
-$\mathtt{LDR}_{wrong}$            & 62.10 $_{43.29}$ & 34.17 & 48.79 & \\
+%$\mathtt{LDR}_{generic}$          & 70.59$_{46.37}$ & 35.22 & 49.68 & \\
+$\mathtt{LDR}_{pred}$               & 72.76$_{46.35}$ & 35.10 & 50.38 & \\
+$\mathtt{LDR}_{wrong}$            & 62.10$_{43.29}$ & 34.17 & 48.79 & \\
 \hline
@@
 \hline
-$Mixed$                        & 57.37 $_{37.94}$ & 23.10 & 63.54 & \\
+$Mixed$                        & 57.37$_{37.94}$ & 23.10 & 63.54 & \\
 \hline
 $\mathtt{FT}_{EMEA}$  & 65.64$_{44.71}$ & 12.36 & 15.93 & \\
-$\mathtt{FT}_{EPPS}$   & 24.90 $ $ & 22.98 & 26.26 & \\
-$\mathtt{FT}_{ECB}$     & 41.80 $ $ & 15.97 & 71.07 & \\
+$\mathtt{FT}_{EPPS}$   & 24.90$ $ & 22.98 & 26.26 & \\
+$\mathtt{FT}_{ECB}$     & 41.80$ $ & 15.97 & 71.07 & \\
 \hline
-$\mathtt{DC}$                & 62.53 $_{39.25}$ & 23.74 & 65.71 & \\
+$\mathtt{DC}$                & 62.53$_{39.25}$ & 23.74 & 65.71 & \\
 %\hline
-%$\mathtt{WDCMT}$ & 64.05 $_{38.54}$ & 22.82 & 67.00 & \\
+%$\mathtt{WDCMT}$ & 64.05$_{38.54}$ & 22.82 & 67.00 & \\
 \hline
-$\mathtt{LDR}_{oracle}$     & 63.43 $_{40.04}$ & 22.66 & 64.40 & \\
-$\mathtt{LDR}_{oracle}^{0.5}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & \\
+$\mathtt{LDR}_{oracle}$     & 63.43$_{40.04}$ & 22.66 & 64.40 & \\
+$\mathtt{LDR}_{oracle}^{0.5}$   & 63.27$_{38.16}$ & 21.83 & 69.55 & \\
 %$\mathtt{LDR}_{generic}$   & 63.27 $_{39.75}$ & 22.62 & 63.70 & \\
-$\mathtt{LDR}_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64.00 & \\
-$\mathtt{LDR}_{wrong}$     & 56.84 $_{37.05}$ & 22.06 & 61.66 & \\
+$\mathtt{LDR}_{pred}$        & 63.17$_{39.92}$ & 22.51 & 64.00 & \\
+$\mathtt{LDR}_{wrong}$     & 56.84$_{37.05}$ & 22.06 & 61.66 & \\
 \hline
@@
 \hline
-$\mathtt{LDR}_{pred}^{condgru}$  & 71.70 $_{46.21}$ & 35.09 & 51.22 & \\
-$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & \\
+$\mathtt{LDR}_{pred}^{condgru}$  & 71.70$_{46.21}$ & 35.09 & 51.22 & \\
+$\mathtt{WDCMT}$           & 68.76$_{45.29}$ & 35.71 & 52.75 & \\
 \hline
@@
 
-\subsection{Balancing generic and domain-specific representations\label{secc:region_size}}
+\subsection{Balancing generic and domain representations\label{secc:region_size}}
 
 An important practical question concerns the balance between the generic and the domain-specific part of the embeddings. 
-In the limit where the domain specific part is very small, we should recover the performance of the \texttt{Mixed} system; conversely, we expect to see a less effective sharing of data across domains by increasing the domain-specific regions.
+In the limit where the domain specific part is very small, we should recover the performance of the \texttt{Mixed} system; 
+conversely, we expect to see a less effective sharing of data across domains by increasing the domain-specific regions. 
 Table~\ref{tab:embedding-size} reports the result of  a series of experiments for the Transformer architecture (English-French) with varying domain-specific sizes allocating between 4 and 64 cells for domain-specific information, and the complement to 512 for the generic part.
@@
 \hline
- size=4   & 74.65 $_{49.61}$ & 37.42 & 54.49 & \\
- size=8   & 74.26 $_{49.90}$ & 37.67 & 54.07 & \\
- size=16 & 74.15 $_{49.10}$ & 37.78 & 54.56 & \\
- size=32 & 75.10 $_{48.61}$ & 37.64 & 54.29 & \\
- size=64 & 74.50 $_{50.17}$ & 37.27 & 54.50 & \\
+ size=4   & 74.65$_{49.61}$ & 37.42 & 54.49 & \\
+ size=8   & 74.26$_{49.90}$ & 37.67 & 54.07 & \\
+ size=16 & 74.15$_{49.10}$ & 37.78 & 54.56 & \\
+ size=32 & 75.10$_{48.61}$ & 37.64 & 54.29 & \\
+ size=64 & 74.50$_{50.17}$ & 37.27 & 54.50 & \\
 \hline
@@
 As expected, a huge gap is observed over the IT test set when learning includes in-domain data.
-Note that $\mathtt{LDR}_{oracle}^*$ clearly outperforms $\mathtt{Mixed}^*$ by a large margin over the IT test set.
+Note that on this test set $\mathtt{LDR}_{oracle}^*$ clearly outperforms $\mathtt{Mixed}^*$ by a large margin.
 %

### 2019-08-27 2019/iwslt/submission.tex
additional domain scenario

@@
 \name{Firstname Lastname}
+%\name{MinhQuang Pham$^{\dag\ddag}$, \ Josep Crego$^\dag$, \ Fran\c cois Yvon$^\ddag$, \ Jean Senellart$^\dag$}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\address{ Affiliation / Address line 1 \\
-{\small \tt email@domain}
-}
+\address{ Affiliation / Address line 1 \\ Affiliation / Address line 2 \\ Affiliation / Address line 3 \\ {\small \tt email@domain}}
+%\address{ $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France \\ \small \tt firstname.lastname@systrangroup.com \\ $^\ddag$LIMSI, CNRS,  Universit\'e Paris-Saclay 91405 Orsay, France \\ {\small \tt firstname.lastname@limsi.fr}}
+
 %
@@
 \end{center}
-\caption{BLEU scores for RNN systems. Comparison between a version of $\texttt{LDR}_{pred}$ using conditional GRUs and \texttt{WDCMT}.\label{tab:results-rnn-wdcmt}}
+\caption{BLEU scores for RNN systems. Comparison between \texttt{WDCMT} and $\texttt{LDR}_{pred}$ built using conditional GRUs.\label{tab:results-rnn-wdcmt}}
 \end{table}
@@
 We therefore decided to allocate $8$~cells for the domain specific part. 
-This suggests that we could easily accommodate more domains with the same architecture, and even reserve some regions to handle supplementary incoming training data.  
+This suggests that we could easily accommodate more domains with the same architecture, and even reserve some regions to handle supplementary data. %incoming training data.  
 % \fyTodo{Easily accomodate more domains}
@@
 
-\begin{table}
+\begin{table}[h]
   \centering
@@
 \section{Related Work \label{sec:related_work}}
-
 Domain adaptation (DA) is a vexing problem in NLP, which appears in a wide range of practical situations and data scenarios (eg.\ supervised vs.

### 2019-08-27 2019/iwslt/submission.tex
typo

@@
 the European Central Bank (ECB) \cite{Tiedemann09news},
-In addition, for English-French we also used IT-domain corpora obtained from the OPUS web site\footnote{\url{http://opus.nlpl.eu}} corresponding to KDE, Ubuntu, GNOME and PHP datasets (IT).
+In addition, for English-French we also use IT-domain corpora obtained from the OPUS web site\footnote{\url{http://opus.nlpl.eu}} corresponding to KDE, Ubuntu, GNOME and PHP datasets (IT).
 %

### 2019-08-27 2019/iwslt/submission.tex
typo

@@
 The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-The \texttt{WDCMT} system has one bidirectional Gated recurrent units (GRU) layer on the encoder side and one unidirectional conditional GRU layer on the decoder side; the dimension of ``domain'' layers is~300.
  %
@@
 In all cases, we use a batch size of~128 and a dropout rate of 0.1 for all layers.
-All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt} except for WDCMT for which we use the authors' implementation.\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}}\fyDone{batch size}
+All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}
 
@@
 We also compare our presented network with the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT}) for English$\rightarrow$French translation.
-However, the comparison with \texttt{WDCMT} is difficult as the implementation differs in many ways from our RNN: framework, cell types,  etc.
-Thus, in order to compare closer architectures we replace our model to use a single GRU layer in the encoder and a decoder with a single conditional GRU layer ($\mathtt{LDR}_{pred}^{condgru}$).
+We use the author's network implementation\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}} that is composed of one bidirectional Gated recurrent units (GRU) layer on the encoder side; and one unidirectional conditional GRU layer on the decoder side; the dimension of ``domain'' layers is~300.
+%
+Since the comparison is difficult as both networks differ in many ways: framework, cell types, \textit{etc.} we rearrange the architecture of our model to make it closer to that of the \texttt{WDCMT} network.
+This is, we build our model with a single GRU layer in the encoder and a single conditional GRU layer in the decoder ($\mathtt{LDR}_{pred}^{condgru}$).
+%However, the comparison with \texttt{WDCMT} is difficult as the implementation differs in many ways from our RNN: framework, cell types,  etc.
+%Thus, in order to compare closer architectures we replace our model to use a single GRU layer in the encoder and a decoder with a single conditional GRU layer ($\mathtt{LDR}_{pred}^{condgru}$).
+%
 Results are shown by Table~\ref{tab:results-rnn-wdcmt}.
@@
 \hline
+$\mathtt{LDR}_{pred}$               & 72.76$_{46.35}$ & 35.10 & 50.38 & \\
 $\mathtt{LDR}_{pred}^{condgru}$  & 71.70$_{46.21}$ & 35.09 & 51.22 & \\
@@
 Another approach worth mentioning integrates domain information (for instance a domain language model) in the decoding algorithm \cite{Gulcehre17onintegrating}.
-
+%
 Our scenario is a bit different as we aim to train a unique system that will work well for several domains.
@@
 Following up on \cite{Yang15unified}, the recent proposals of \cite{Peng17multitask} apply the same idea with neural architectures, using domain specific masking to zero out the parameters modeling domains irrelevant to the current input sentence.
-Compared to our work, these techniques are used in deep layers of the network and applied to sequence labelling tasks.
+Compared to our work, these techniques are used in deep network layers and applied to sequence labelling tasks.
 
@@
 Their main contribution is to propose an a method to relate the amount of adaptation of the network parameters to the similarity between the adaptation sample and the test sentence: the higher the similarity, the more agressive the adaptation.
-
+%
 By analogy with the proposal of \cite{Johnson17google} for multilingual NMT, \cite{Kobus17domaincontrol} and \cite{Chu18multilingual} separately propose to extend the representation of the source text with a domain tag.

### 2019-08-27 2019/iwslt/submission.tex
Reviewed by FY

@@
 We experiment with three domains, two neural architectures and two language pairs and find that our technique yields effective multi-domain NMTs, outperforming several baselines. 
-Our contribution is thus as follows:
+Our contributions are thus as follows:
 we adapt and implement the ideas of \cite{Daume07frustratingly} for two NMT architectures;
 we provide experimental evidence that show the effectiveness of this technique;
-we evaluate the ability of our network to efficiently accommodate new domains;
-and we introduce a novel technique to analyse word polysemy using word embeddings which comforts our assumption that their variation across domains actually reflects a variation of senses.
+we evaluate the ability of our network to dynamically accommodate new domains;
+and we introduce a novel technique to analyse word polysemy using word embeddings, which comforts the assumption that their variation across domains actually reflects a variation of senses.
 
@@
 
-Multi-domain machine translation is formalized as follows: we assume observations that correspond to  domain-tagged sentence pairs $[(x,y),i]$, with $x$ in the source language, $y$ in the target language and $i$ a domain tag in $[1\dots d]$.
+Multi-domain machine translation is formalized as follows: we assume observations taking the form of  domain-tagged sentence pairs $[(x,y),i]$, with $x$ in the source language, $y$ in the target language and $i$ a domain tag in $[1\dots d]$.
 We further assume a two-stage sampling process: first select a domain $i$ according to $p(i)$, then select a sentence pair according to a domain specific distribution $D_i$.
@@
 
-To actually implement this idea for NMT, we need to define the subset of parameters that will be common to all domains. 
-We make the assumption that domain specificities can be confined to the lexical level, and define $\theta_s$ to contain all the network parameters, except for a subpart of the word embeddings. 
+To actually implement this idea for NMT, we need to define the subset of parameters that will be shared across domains. 
+Our hypothesis is that domain specificities can be confined to the lexical level and define $\theta_s$ to contain all the network parameters except for a subpart of the word embeddings. 
 For each word $v$, the embedding vector $e(v)$ is thus decomposed as $e(v) = [e_g(v); e_1(v); \dots; e_d(v)]$, where $e_g(v)$ stores the generic lexical embedding, while $e_i(v)$ stores the subpart that is specific to domain $i$.
@@
 When processing a sample from domain $2$, we only activate the corresponding parameter region ($\theta_2$) in the input embeddings; the remaining domain-specific parts are zeroed out, and do not receive any update.
-The generic part is always active, and is updated irrespective of the input domain.}
+The generic part is always active and is updated irrespective of the input domain.}
   \label{fig:network}
@@
 \subsection{Baselines \label{ssec:baselines}}
-To validate our findings, we compared lexicalized domain embedding models with standard models using both attentional Recurrent Neural Networks (RNNs) \cite{Bahdanau15learning} and the recent Transformer architecture of \cite{Vaswani17attention}.
+To validate our findings, we compared lexicalized domain embedding models with standard models using both attentional Recurrent Neural Networks (RNNs) \cite{Bahdanau15learning} and the Transformer architecture of \cite{Vaswani17attention}.
 Our baselines consist of:
@@
 \item generic models trained with a simple concatenation of all corpora (\texttt{Mixed});
-\item models tuned separately on each domain using respectively (10000, 15000, 5000) iterations over in-domain data (\texttt{ft$_{EMEA}$}, \texttt{ft$_{EPPS}$}, \texttt{ft$_{ECB}$}); 
+\item models tuned separately on each domain for respectively (10000, 15000, 5000) iterations using in-domain data (\texttt{ft$_{EMEA}$}, \texttt{ft$_{EPPS}$}, \texttt{ft$_{ECB}$}); 
 \item models using domain tags as in \cite{Kobus17domaincontrol} (\texttt{DC}); 
@@
 
-Results are summarised respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems, where we report BLEU~\cite{Papineni02bleu} scores computed after detokenization.\footnote{As explained above, we  report two numbers when testing with EMEA, except for the fine-tuning scenarios when tuning on ECB and EPPS.} % \fyTodo{Footnote for missing numbers}.
+Results are summarized respectively in Table~\ref{tab:results-trsf} for the Transformer systems and Table~\ref{tab:results-rnn} for the RNN systems, where we report BLEU~\cite{Papineni02bleu} scores computed after detokenization.\footnote{As explained above, we  report two numbers when testing with EMEA, except for the fine-tuning scenarios when tuning on ECB and EPPS.} % \fyTodo{Footnote for missing numbers}.
 
 %We first concentrate on the left part of these tables, where we analyze the performance for the three domains of interest. 
-First, we observe that Transformer is consistently better than RNN, and that fine-tuning on a domain-specific corpus, when applicable, is almost the best way to optimise the performance on that domain.\footnote{This is not so clear for EPPS, where fine-tuning does not seem to help.}
+First, we observe that Transformer is consistently better than RNNs and that fine-tuning on a domain-specific corpus, when applicable, is almost the best way to optimize the performance on that domain.\footnote{This is not so clear for EPPS, where fine-tuning does not seem to help.}
 Note that fine-tuning will however yield a marked (even sometimes catastrophic, eg.\ for the EMEA-tuned Transformer system) decrease in performance for the other domains.
 
-The presented model (\texttt{LDR}$_{oracle}$) is consistently better than the \texttt{Mixed} strategy, with gains that range from very large (for EMEA and ECB) to unsignificant (for EPPS in most conditions).
+Our approach (\texttt{LDR}$_{oracle}$) is consistently better than the \texttt{Mixed} strategy, with gains that range from very large (for EMEA and ECB) to unsignificant (for EPPS in most conditions).
 This means that our architecture is somehow able to compensate for the data unbalance and to raise the performance of the multi-domain system close to the best (fine-tuned) system in each domain. 
@@
 
-We also compare our presented network with the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT}) for English$\rightarrow$French translation.
-We use the author's network implementation\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}} that is composed of one bidirectional Gated recurrent units (GRU) layer on the encoder side; and one unidirectional conditional GRU layer on the decoder side; the dimension of ``domain'' layers is~300.
+We also compare our architecture with the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT}) for the pair English$\rightarrow$French.
+We use the author's implementation\footnote{\noindent\url{http://github.com/DeepLearnXMU/WDCNMT}} that is composed of one bidirectional Gated recurrent units (GRU) layer on the encoder side; and one unidirectional conditional GRU layer on the decoder side; the dimension of ``domain'' layers is~300.
 %
-Since the comparison is difficult as both networks differ in many ways: framework, cell types, \textit{etc.} we rearrange the architecture of our model to make it closer to that of the \texttt{WDCMT} network.
-This is, we build our model with a single GRU layer in the encoder and a single conditional GRU layer in the decoder ($\mathtt{LDR}_{pred}^{condgru}$).
+The direct comparison with our RNN is difficult, as both networks differ in many ways: framework, cell types, \textit{etc.}.
+The results in Table~\ref{tab:results-rnn-wdcmt} therefore use a variant of our model that makes it more similar to that of the \texttt{WDCMT} network.
+In particular, this variant also uses a single GRU layer in the encoder and a single conditional GRU layer in the decoder ($\mathtt{LDR}_{pred}^{condgru}$).
+As can be seen in this table, our model is on average comparable to \texttt{WDCMT}, while using a much simpler design.
 %However, the comparison with \texttt{WDCMT} is difficult as the implementation differs in many ways from our RNN: framework, cell types,  etc.
@@
 %
-Results are shown by Table~\ref{tab:results-rnn-wdcmt}.
-The comparison is less favourable for our RNN system, which is arguably much simpler.
+% Results are shown by Table~\ref{tab:results-rnn-wdcmt}.
+% The comparison is less favourable for our RNN system, which is arguably much simpler.
 
@@
 
-We evaluate now the ability of our model to integrate new domains, a very common scenario for industrial MT.
-We simulate the apparition of the IT domain once the $\mathtt{LDR}_{oracle}$ model is built for EMEA, EPPS and ECB domains.
-Note that to accommodate a new domain the network must be designed with a region reserved for the new domain.
-Hence, for this experiment word embeddings contain 480 cells for the generic region and 32 cells for domain specific regions (8 cells x 4 regions).
-Equivalently to our previous $\mathtt{LDR}_{oracle}$ model, 200,000 iterations are performed using EMEA, EPPS and ECB data sets (the IT region is never updated).
-Then, 100,000 additional iterations are performed using  EMEA, EPPS, ECB and IT datasets ($\mathtt{LDR}_{oracle}^*$).
-Experiments are reported for the English$\rightarrow$French language pair.
-For comparison purposes we also continue training our $\mathtt{Mixed}$ model applying the same number of iterations and using the same four datasets ($\mathtt{Mixed}^*$).
-BLEU results are shown in Table~\ref{tab:add}.
+We now evaluate the ability of our model to integrate new domains, a very common scenario for industrial MT.
+In this setting, we consider that we have a model ($\mathtt{LDR}_{oracle}$) trained as before for EMEA, EPPS and ECB during 200,000 iterations, which needs to accomodate new training data in the IT domain.
+Assuming that we have reserved extra empty embedding cells\footnote{For this experiment, word embeddings contain 480 cells for the generic region and 32 cells for domain specific regions (8 cells x 4 regions).} for this new domain, we resume training with 4 domains during 100,000 additional iterations, yielding a updated model  $\mathtt{LDR}_{oracle}^*$.
+Experiments are reported for the English$\rightarrow$French language pair in Table~\ref{tab:add}, where for comparison purposes we also report numbers obtained with continued training with the $\mathtt{Mixed}$ model, training for the same number of iterations and using the same four datasets ($\mathtt{Mixed}^*$).
 
@@
 
-As expected, a huge gap is observed over the IT test set when learning includes in-domain data.
-Note that on this test set $\mathtt{LDR}_{oracle}^*$ clearly outperforms $\mathtt{Mixed}^*$ by a large margin.
+As expected, a huge improvement in performance is observed for the IT test set when learning includes in-domain data for both models, with$\mathtt{LDR}_{oracle}^*$ outperforming $\mathtt{Mixed}^*$ by a wide margin.
 %
-Considering the other test sets: both models similarly enhance their performance on the ECB test set;
-$\mathtt{LDR}_{oracle}^*$ shows gains over the EMEA test set while a loss of quality is observed on $\mathtt{Mixed}^*$;
-using IT data does not impact the quality of translations of the EPPS test set of any of the models.
-Overall better results are obtained by our $\mathtt{LDR}_{oracle}$ model when trained including the additional domain dataset.
-Thus, confirming the ability of the model to easily integrate new domains. % as supplementary training data becomes available.
+It is interesting to see that this additional data has also a positive impact on other test sets: both models similarly increase their performance for the ECB domain, and $\mathtt{LDR}_{oracle}^*$ additionally improves the results for the EMEA test, which is not the case for $\mathtt{Mixed}^*$;
+finally, using IT data does not impact the quality of translations for the EPPS domain of any of the models.
+Overall better results are obtained by our $\mathtt{LDR}_{oracle}$ model when trained including the additional domain dataset, demonstrating the ability of the model to seamlessly integrate new domains.
+% as supplementary training data becomes available.
 %New domains are easily integrated in the model if the corresponding embedding regions are reserved in advance.
@@
 \section{Conclusions}
-In this paper, we have presented a new technique for multi-domain machine translation, adapting to two common NMT architectures the ``frustratingly easy'' idea of \cite{Daume07frustratingly}. 
-Our experiments have shown that for both architecture and for two language pairs, our multi-domain models improve over several baselines of the literature, and that it is robust to noise in domain labels.
-In our future work, we intend to develop this work so as to make the architecture more self-configurable and able to adapt the size of the domain-specific regions depending upon the actual variation of use across domains;
-we also would like to find additional ways to make the architecture able to integrate new domains in a dynamic fashion, as this is an important requirement in industrial systems.
+In this paper, we have presented a new technique for multi-domain machine translation,  adapting the ``frustratingly easy'' idea of \cite{Daume07frustratingly} to two standard NMT architectures. 
+Our experiments have shown that for both architectures and for two language pairs, our multi-domain models improve over several baselines of the literature and that it is robust to noise in domain labels.
+It is noticeable that these results are obtained without impacting the architecture or training complexity
+, making our approach a simple baseline for further studies in multi-domain translation
+.
+We have also shown that our approach can dynamically handle new domains; and that the domain-specific embeddings often reflect differences of senses.
+In our future work, we intend to develop these ideas so as to make the architecture more self-configurable and able to adapt the size of the domain-specific regions depending upon the actual variation of use across domains; we also would like to find additional ways to make the architecture able to integrate an arbitrary number of new domains in a dynamic fashion, as this is an important requirement in industrial systems.
 

### 2019-08-28 2019/iwslt/submission.tex
couple of late fixes

@@
   \end{tabular}
-  \caption{Analyzing the variation of embeddings across domains. Each word or subword is listed with the size of the intersection (between 0 and 10).}
+  \caption{Analyzing the variation of embeddings across domains. For each word or subword we also report the size of the intersection (between 0 and 10).}
   \label{tab:embeddings}
@@
 
-Let us now consider the list of putative monosemic words (on the right part of Table~\ref{tab:embeddings}), ie.\ words for which the nearest neighbors are the same in all domains, contains mostly words for which we do not expect much variation in translation: adjective (``previous'', ``better''), adverbs (``virtually``), generic verbs (``handling'', ``coordinated''), etc; and the list of words having a constant set of neighbors also contains prepositions (``at'', ``in''), auxiliary (``been'') etc.
+Let us now consider the list of putative monosemic words (on the right part of Table~\ref{tab:embeddings}), ie.\ words for which the nearest neighbors are the same in all domains.
+This list contains mostly words for which we do not expect much variation in translation: adjectives (``previous'', ``better''), adverbs (``virtually``), generic verbs (``handling'', ``coordinated'').
+Further down this list, we will also find prepositions (``at'', ``in''), auxiliary (``been'') etc.
 

### 2019-08-28 2019/iwslt/submission.tex
light

--- 2019-08-27 2019/iwslt/submission.tex
@@
 It is noticeable that these results are obtained without impacting the architecture or training complexity
-, making our approach a simple baseline for further studies in multi-domain translation
-.
+, making our approach a simple baseline for further studies in multi-domain translation.
 We have also shown that our approach can dynamically handle new domains; and that the domain-specific embeddings often reflect differences of senses.

### 2019-09-02 2019/iwslt/submission.tex
added averages using large set for emea
boldfaced best scores of each table
derniere relecture: 6 pages

@@
 \title{Lexicalized Domain Representations for Neural Machine Translation}
+\title{Generic and Specialized Word Embeddings for Multi-Domain Machine Translation}
 
@@
   Supervised machine translation works well when the train and test data are sampled from the same distribution.
-  When this is not the case, \emph{adaptation} techniques help ensure that the knowledge learned from out-of-domain texts generalises to target domain sentences.
+  When this is not the case, \emph{adaptation} techniques help ensure that the knowledge learned from out-of-domain texts generalises to in-domain sentences.
   We study here a related setting, \emph{multi-domain adaptation}, where the number of domains is potentially large and adapting separately to each domain would waste training resources.
-  Our proposal transposes to neural machine translation the feature expansion technique of (Daum\'e III, 2007): it isolates domain-agnostic from domain-specific lexical representations, while sharing the rest of the network across domains.
-  Our experiments use two neural architectures and two language pairs: they show that our approach, while remaining simple and computationally inexpensive, outperforms several strong baselines and delivers a multi-domain system that can successfully translate heterogeneous texts.
+  Our proposal transposes to neural machine translation the feature expansion technique of (Daum\'e III, 2007): it isolates domain-agnostic from domain-specific lexical representations, while sharing the most of the network across domains.
+  Our experiments use two architectures and two language pairs: they show that our approach, while simple and computationally inexpensive, outperforms several strong baselines and delivers a multi-domain system that successfully translates texts from diverse sources.
 \end{abstract}
@@
 Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,Bahdanau15learning,Ghering17convolutional,Vaswani17attention}, 
-Machine Translation (MT) has made significant progresses over the past years, and constitute to date the standard for most production engines. 
+Machine Translation (MT) has made significant progresses over the past years and constitute to date the standard for most production engines. 
 The development of MT systems, be they neural or statistical, require very large parallel corpora consisting of millions of sentence pairs, a resource that only exist in very few application domains and language pairs.
-A lot of the recent research effort has thus focused on developing MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Ha16towards,Johnson17google};
+A lot of the recent research effort has thus focused on developing MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Ha16towards,Johnson17google}.
 Another important scenario for industrial MT is to adapt a neural system trained using parallel data in one domain to the peculiarity of other domains.
 
-Domain Adaption (DA) in MT is an old issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied. 
+Domain Adaption (DA) in MT is an old issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises and for which a number of solutions have been studied. 
 See the recent survey of \cite{Chu18asurvey} for neural MT. 
-The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tune the parameters of a system trained on a large amount of texts in a source domain.
+The typical setting is \emph{supervised adaptation}, where a (small) amount of data in the target domain of interest is used to fine-tune the parameters of a system trained on a large amount of texts in a source domain.
 We study here a different scenario, \emph{multi-domain adaptation} \cite{Sennrich13multidomain,Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.
 This allows us to be both data efficient (all data is used to train all domains) and computationally efficient (we only train one system).
-Multi-domain adaptation is conceptually close to multilingual MT, or more generally to multi-task learning \cite{Caruana97multitask}, and can be approached in a number of ways.
+Multi-domain adaptation is conceptually close to multilingual MT, or more generally to multi-task learning \cite{Caruana97multitask} and can be approached in a number of ways.
 
@@
 To this end, we design word embeddings containing a generic and several domain-specific regions. 
-We experiment with three domains, two neural architectures and two language pairs and find that our technique yields effective multi-domain NMTs, outperforming several baselines. 
+We experiment with four domains, two neural architectures and two language pairs and find that our technique yields effective multi-domain NMTs, outperforming several baselines. 
 Our contributions are thus as follows:
@@
 we provide experimental evidence that show the effectiveness of this technique;
-we evaluate the ability of our network to dynamically accommodate new domains;
-and we introduce a novel technique to analyse word polysemy using word embeddings, which comforts the assumption that their variation across domains actually reflects a variation of senses.
+we evaluate the ability of our networks to dynamically accommodate new domains;
+and we introduce a new technique to analyze word polysemy using embeddings, which comforts the assumption that their variation across domains actually reflects a variation of senses.
 
@@
 We further assume a two-stage sampling process: first select a domain $i$ according to $p(i)$, then select a sentence pair according to a domain specific distribution $D_i$.
-Our objective is to find a tuple of parameters $\{\theta_1 \dots \theta_d \} \in \mathbb{R}^D \times \mathbb{R}^D \dots \times \mathbb{R}^D$ minimizing:
+Our objective is to find a tuple of parameters $\{\theta_1 \dots \theta_d \} \in \mathbb{R}^D \times \dots \times \mathbb{R}^D$ minimizing:
 \begin{equation} \label{eq:loss}
 \begin{split}
-\sum_{i \in [1..d]} p(i) E_{(x,y) \sim D_{i}} [-log(p_{\theta_i}(y|x,i))]
+\sum_{i \in [1..d]} p(i) E_{(x,y) \sim D_{i}} [-log(p_{\theta_i}(y|x,i))] \text{.}
 \end{split}
@@
 \fyDone{Risk}
-The alternative we consider here constraints each parameter $\theta_i$ to be made of two parts: $\theta_i = [\theta_s; \theta'_i]$ ($\theta_s \in R^{D_g}$) is shared across all domains, while the second part $\theta'_i \in \mathbb{R}^{D_i}$ is only used in domain $i$.
+The alternative we consider here constraints each parameter $\theta_i$ to be made of two parts: $\theta_i = [\theta_s; \theta'_i]$.
+$\theta_s \in R^{D_g}$ is shared across all domains, while the second part $\theta'_i \in \mathbb{R}^{D_i}$ is only used in domain $i$.
 The parameter set is now much more constrained, yet we expect that tying parameters across domains will yield better estimates for $\theta_s$ due to a larger training corpus.
@@
 To actually implement this idea for NMT, we need to define the subset of parameters that will be shared across domains. 
-Our hypothesis is that domain specificities can be confined to the lexical level and define $\theta_s$ to contain all the network parameters except for a subpart of the word embeddings. 
+Our hypothesis is that domain specificities can be confined to the lexical level and we define $\theta_s$ to contain all the network parameters except for a subpart of the word embeddings. 
 For each word $v$, the embedding vector $e(v)$ is thus decomposed as $e(v) = [e_g(v); e_1(v); \dots; e_d(v)]$, where $e_g(v)$ stores the generic lexical embedding, while $e_i(v)$ stores the subpart that is specific to domain $i$.
@@
   \caption{Lexicalized domain embeddings.
-When processing a sample from domain $2$, we only activate the corresponding parameter region ($\theta_2$) in the input embeddings; the remaining domain-specific parts are zeroed out, and do not receive any update.
+When processing a sample from domain $2$, we only activate the corresponding parameter region ($\theta_2$) in the input embeddings; the remaining domain-specific parts are zeroed out and do not receive any update.
 The generic part is always active and is updated irrespective of the input domain.}
@@
 
-Making sure that the actual embedding do not contain any zero is important for the Transformer model, since the lexical representations are then summed to the positional encoding, which would undo the effect of domain masking, and propagate a gradient even to regions that should not be modified.
+Making sure that the actual embedding do not contain any zero is important for the Transformer model, since the lexical representations are then added to the positional encoding, which would undo the effect of domain masking, and propagate a gradient even to regions that should not be modified.
 With our design, we make sure that during backprogation, the matrix $M$ receives gradient $0$ at regions corresponding to deactivated regions in the word embedding. 
-Those regions are also masked in forward step, thus do not interfere the training on the domain to which they are not assigned (see Figure~\ref{fig:network}).
+Those regions are also masked in forward step, thus do not interfere the training on the domains to which they are not assigned (see Figure~\ref{fig:network}).
 %
 Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers the computation in equation~\eqref{eq:embedding}.
-In our experiments, we consider both the attentional RNN architecture of \cite{Bahdanau15learning} and for Transformer architecture of \cite{Vaswani17attention}.
+In our experiments, we consider both the attentional RNN architecture of \cite{Bahdanau15learning} and the Transformer architecture of \cite{Vaswani17attention}.
 
@@
 
-We experiment with two language pairs (English-French, English-German) and data originated from three domains, corresponding to texts from three European institutions: 
+We experiment with two language pairs (English-French, English-German) and data originating from three domains, corresponding to texts from three European institutions: 
 the European Parlement (EPPS) \cite{Koehn05europarl}, 
@@
     \hline
-    EMEA  & 1.09M & 1,000 & 1,000$_{300}$\\
+    EMEA  & 1.09M & 1,000 & 1,000$_{(300)}$\\
     ECB    & 0.19M & 1,000 & 1,000     \\
@@
     \hline
-    EMEA  & 1.11M & 1,000 & 1,000$_{300}$ \\
+    EMEA  & 1.11M & 1,000 & 1,000$_{(300)}$ \\
     ECB     &  0.11M & 1,000 & 1,000  \\
@@
 
-Note that the EMEA dataset as distributed on the OPUS site contains multiple sentence duplicates. 
-We therefore report two numbers: the first is comparable to what has been published on earlier studies (eg.\ \cite{Zeng18multidomain}), the second one is obtained by making the test set entirely disjoint from the training set (700 duplicated sentences are discarded).
+Note that the EMEA dataset distributed on the OPUS site contains multiple sentence duplicates. 
+We therefore report below two numbers as $S_{(T)}$: the first ($S$) is comparable to what has been published on earlier studies (eg.\ \cite{Zeng18multidomain}), the second one ($T$) is obtained by making the test entirely disjoint from the training (700~duplicated sentences are discarded).
 
@@
 RNN models use 1 layer on both sides: 
-a bidirectional LSTM encoder and a unidirectional LSTM decoder with attention;
+a bidirectional LSTM encoder and a unidirectional LSTM decoder with attention.
 The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
@@
 In all cases, we use a batch size of~128 and a dropout rate of 0.1 for all layers.
-All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}
+All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
 
@@
 3 are domain specific and 1 is generic, with sizes $[8,8,8,488]$ respectively. %, a choice that will be discussed in Section~\ref{secc:region_size}. 
-If a sentence originates from domain $i$, the domain specific regions for all domains $j \neq i$ will be zeroed out; while the other regions are activated (cf. Figure~\ref{fig:network}). 
+If a sentence originates from domain $i$, the domain specific regions for all domains $j \neq i$ will be zeroed out while the other regions are activated (cf. Figure~\ref{fig:network}). 
 We then use a dense layer of size~512 to fuse the region for the active domain and the ``generic'' region. Training is formalised in algorithm~\ref{alg:multidomain}.
@@
 In our experiments, this means that sentences from the Europarl domain will be selected more frequently that the two other domains.
-In an attempt to build a more balanced training sample, we also consider an alternative selection procedure, where $i$ is selected according to  distribution $[\frac{\sqrt{|C_i|}}{\sum_{i\in [1,..,d]}\sqrt{|C_i|}}]$.
-
+We also consider a more balanced sampling procedure, where $i$ is selected according to  distribution $[\frac{\sqrt{|C_i|}}{\sum_{i\in [1,..,d]}\sqrt{|C_i|}}]$.
+The corresponding results are reported as \texttt{LDR}$^{0.5}$.
 \subsection{Results \label{ssec:results}}
@@
 First, we observe that Transformer is consistently better than RNNs and that fine-tuning on a domain-specific corpus, when applicable, is almost the best way to optimize the performance on that domain.\footnote{This is not so clear for EPPS, where fine-tuning does not seem to help.}
-Note that fine-tuning will however yield a marked (even sometimes catastrophic, eg.\ for the EMEA-tuned Transformer system) decrease in performance for the other domains.
+Note that fine-tuning however yields a marked (even sometimes catastrophic, eg.\ for the EMEA-tuned Transformer system) decrease in performance for the other domains.
 
@@
 \texttt{LDR}$_{oracle}$ is also better than Domain Control in three conditions out of four, \texttt{DC} being seemingly a better choice for the RNN than for the Transformer architecture. 
-As expected, ignoring the true domain label yields a light drop in performance. %for the two fallback strategies considered in this study. 
+As expected, ignoring the true domain label yields a light drop in performance: this is reflected in the results of %for the two fallback strategies considered in this study. 
 %The first, \texttt{LDR}$_{generic}$, only uses the generic part of the embeddings. 
 %The second, 
-\texttt{LDR}$_{pred}$ employs automatically predicted domain labels.\footnote{Our domain classifier uses a bi-LSTM RNN encoder, followed by a simple softmax layer. 
+\texttt{LDR}$_{pred}$, which relies on automatically predicted domain labels.\footnote{Our domain classifier uses a bi-LSTM RNN encoder, followed by a simple softmax layer. 
 Its precision on a development set exceeds 95\%.}
@@
 %This might be due to the fact that WDCMT has no built-in mechanism to ignore, as we do in the \texttt{LDR}$_{generic}$ systems, the possibly misleading predicted domain information.
-\jcTodo{compute averages on all tables}
 
@@
 \hline
-$\mathtt{Mixed}$          & 67.69$_{47.60}$ & 37.50 & 53.49 & \\
+$\mathtt{Mixed}$          & 67.69$_{47.60}$ & 37.50 & 53.49 & 52.89\\
 \hline
-$\mathtt{FT}_{EMEA}$ & 76.77$_{49.43}$ & 17.16 & 11.99 &\\
-$\mathtt{FT}_{EPPS}$  & 20.86 & 37.04 & 24.53 &  \\
-$\mathtt{FT}_{ECB}$    & 26.93 & 27.09 & 56.52 & \\
+$\mathtt{FT}_{EMEA}$ & 76.77$_{49.43}$ & 17.16 & 11.99 & 35.30\\
+$\mathtt{FT}_{EPPS}$  & 20.86 & 37.04 & 24.53 & 27.47\\
+$\mathtt{FT}_{ECB}$    & 26.93 & 27.09 & \bf 56.52 & 36.84\\
 \hline
-$\mathtt{DC}$              & 67.87$_{45.42}$ & 37.31 & 54.14 & \\
+$\mathtt{DC}$              & 67.87$_{45.42}$ & 37.31 & 54.14 & 53.10\\
 \hline
-$\mathtt{LDR}_{oracle}$            & 74.26$_{49.90}$ & 37.67 & 54.07 & \\
-$\mathtt{LDR}_{oracle}^{0.5}$   & 74.95$_{49.38}$ & 37.35 & 55.91 & \\
+$\mathtt{LDR}_{oracle}$            & 74.26$_{\bf 49.90}$ & 37.67 & 54.07 & 55.33\\
+$\mathtt{LDR}_{oracle}^{0.5}$   & \bf 74.95$_{49.38}$ & 37.35 & 55.91 & \bf 56.07\\
 %$\mathtt{LDR}_{generic}$          & 73.97$_{49.54}$ & 37.81 & 53.67 & \\
-$\mathtt{LDR}_{pred}$               & 74.29$_{49.84}$ & 37.73 & 54.01 & \\
-$\mathtt{LDR}_{wrong}$            & 72.95$_{49.78}$ & 37.62 & 53.35 & \\
+$\mathtt{LDR}_{pred}$               & 74.29$_{49.84}$ & \bf 37.73 & 54.01 & 55.34\\
+$\mathtt{LDR}_{wrong}$            & 72.95$_{49.78}$ & 37.62 & 53.35 & 54.64\\
   \hline
@@
 \hline
-$\mathtt{Mixed}$         & 64.57$_{42.99}$ & 26.47 & 68.67 & \\
+$\mathtt{Mixed}$         & 64.57$_{42.99}$ & 26.47 & 68.67 & 53.23\\
 \hline
-$\mathtt{FT}_{EMEA}$ & 68.35$_{42.97}$ & 17.02 & 32.87 & \\
-$\mathtt{FT}_{EPPS}$  & 36.19 & 26.29 & 40.71 & \\
-$\mathtt{FT}_{ECB}$   & 24.72 & 18.36 & 74.05 & \\
+$\mathtt{FT}_{EMEA}$ & 68.35$_{42.97}$ & 17.02 & 32.87 & 39.41\\
+$\mathtt{FT}_{EPPS}$  & 36.19 & 26.29 & 40.71 & 34.39\\
+$\mathtt{FT}_{ECB}$   & 24.72 & 18.36 & \bf 74.05 & 39.04\\
 \hline
-$\mathtt{DC}$ & 63.48$_{42.98}$ & 26.27 & 66.95 & \\
+$\mathtt{DC}$ & 63.48$_{42.98}$ & 26.27 & 66.95 & 52.23\\
 \hline
-$\mathtt{LDR}_{oracle}$ & 70.90$_{46.12}$ & 26.30 & 68.90 & \\
-$\mathtt{LDR}_{oracle}^{0.5}$ & 71.31$_{45.23}$ & 25.98 & 73.74 & \\
+$\mathtt{LDR}_{oracle}$ & 70.90$_{\bf 46.12}$ & 26.30 & 68.90 & 55.36\\
+$\mathtt{LDR}_{oracle}^{0.5}$ & \bf 71.31$_{45.23}$ & 25.98 & 73.74 & \bf 57.01\\
 %$\mathtt{LDR}_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & \\
-$\mathtt{LDR}_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & \\
-$\mathtt{LDR}_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & \\
+$\mathtt{LDR}_{pred}$ & 70.89$_{\bf 46.12}$ & \bf 26.53 & 68.63 & 55.35\\
+$\mathtt{LDR}_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & 54.22\\
 \hline
@@
 \hline
-$Mixed$                         & 65.42$_{45.11}$ & 34.70 & 51.38 & \\
+$Mixed$                         & 65.42$_{45.11}$ & 34.70 & 51.38 & 50.50\\
 \hline
-$\mathtt{FT}_{EMEA}$   & 72.06$_{47.33}$ & 18.62 & 16.78 & \\
-$\mathtt{FT}_{EPPS}$    & 35.47$ $ & 34.61 & 39.56 & \\
-$\mathtt{FT}_{ECB}$      & 21.93$ $ & 22.60 & 51.53 & \\
+$\mathtt{FT}_{EMEA}$   & 72.06$_{\bf 47.33}$ & 18.62 & 16.78 & 35.82\\
+$\mathtt{FT}_{EPPS}$    & 35.47$ $ & 34.61 & 39.56 & 36.55\\
+$\mathtt{FT}_{ECB}$      & 21.93$ $ & 22.60 & 51.53 & 32.02\\
 \hline
-$\mathtt{DC}$                            & 68.26$_{43.76}$ & 35.13 & 50.09 & \\
+$\mathtt{DC}$                            & 68.26$_{43.76}$ & 35.13 & 50.09 & 51.16\\
 % \hline
@@
 \hline
-$\mathtt{LDR}_{oracle}$            & 71.73$_{46.30}$ & 35.21 & 50.91 & \\
+$\mathtt{LDR}_{oracle}$            & 71.73$_{46.30}$ & \bf 35.21 & 50.91 & 52.62\\
 %$\mathtt{LDR_{condgru}}_{pred}$   & 71.7$_{46.21}$ & 35.09 & 51.22 & \\
-$\mathtt{LDR}_{oracle}^{0.5}$   & 71.70$_{46.41}$ & 34.24& 52.37 & \\
+$\mathtt{LDR}_{oracle}^{0.5}$   & 71.70$_{46.41}$ & 34.24& \bf 52.37 & \bf 52.77\\
 %$\mathtt{LDR}_{generic}$          & 70.59$_{46.37}$ & 35.22 & 49.68 & \\
-$\mathtt{LDR}_{pred}$               & 72.76$_{46.35}$ & 35.10 & 50.38 & \\
-$\mathtt{LDR}_{wrong}$            & 62.10$_{43.29}$ & 34.17 & 48.79 & \\
+$\mathtt{LDR}_{pred}$               &\bf 72.76$_{46.35}$ & 35.10 & 50.38 & 52.75\\
+$\mathtt{LDR}_{wrong}$            & 62.10$_{43.29}$ & 34.17 & 48.79 & 48.35\\
 \hline
@@
 \hline
-$Mixed$                        & 57.37$_{37.94}$ & 23.10 & 63.54 & \\
+$Mixed$                        & 57.37$_{37.94}$ & 23.10 & 63.54 & 48.00\\
 \hline
-$\mathtt{FT}_{EMEA}$  & 65.64$_{44.71}$ & 12.36 & 15.93 & \\
-$\mathtt{FT}_{EPPS}$   & 24.90$ $ & 22.98 & 26.26 & \\
-$\mathtt{FT}_{ECB}$     & 41.80$ $ & 15.97 & 71.07 & \\
+$\mathtt{FT}_{EMEA}$  & 65.64$_{\bf 44.71}$ & 12.36 & 15.93 & 31.31\\
+$\mathtt{FT}_{EPPS}$   & 24.90$ $ & 22.98 & 26.26 & 24.71\\
+$\mathtt{FT}_{ECB}$     & 41.80$ $ & 15.97 & 71.07 & 42.95\\
 \hline
-$\mathtt{DC}$                & 62.53$_{39.25}$ & 23.74 & 65.71 & \\
+$\mathtt{DC}$                & 62.53$_{39.25}$ & \bf 23.74 & 65.71 & 50.66\\
 %\hline
@@
 \hline
-$\mathtt{LDR}_{oracle}$     & 63.43$_{40.04}$ & 22.66 & 64.40 & \\
-$\mathtt{LDR}_{oracle}^{0.5}$   & 63.27$_{38.16}$ & 21.83 & 69.55 & \\
+$\mathtt{LDR}_{oracle}$     & \bf 63.43$_{40.04}$ & 22.66 & 64.40 & 50.16\\
+$\mathtt{LDR}_{oracle}^{0.5}$   & 63.27$_{38.16}$ & 21.83 & \bf 69.55 & \bf 51.55\\
 %$\mathtt{LDR}_{generic}$   & 63.27 $_{39.75}$ & 22.62 & 63.70 & \\
-$\mathtt{LDR}_{pred}$        & 63.17$_{39.92}$ & 22.51 & 64.00 & \\
-$\mathtt{LDR}_{wrong}$     & 56.84$_{37.05}$ & 22.06 & 61.66 & \\
+$\mathtt{LDR}_{pred}$        & 63.17$_{39.92}$ & 22.51 & 64.00 & 49.89\\
+$\mathtt{LDR}_{wrong}$     & 56.84$_{37.05}$ & 22.06 & 61.66 & 46.85\\
 \hline
@@
 %
-The direct comparison with our RNN is difficult, as both networks differ in many ways: framework, cell types, \textit{etc.}.
-The results in Table~\ref{tab:results-rnn-wdcmt} therefore use a variant of our model that makes it more similar to that of the \texttt{WDCMT} network.
+The direct comparison with our RNN is difficult, as both networks differ in many ways: framework, cell types, \textit{etc}.
+Results in Table~\ref{tab:results-rnn-wdcmt} therefore use a variant of our model that makes it more similar to the \texttt{WDCMT} network.
 In particular, this variant also uses a single GRU layer in the encoder and a single conditional GRU layer in the decoder ($\mathtt{LDR}_{pred}^{condgru}$).
@@
 \hline
-$\mathtt{LDR}_{pred}$               & 72.76$_{46.35}$ & 35.10 & 50.38 & \\
-$\mathtt{LDR}_{pred}^{condgru}$  & 71.70$_{46.21}$ & 35.09 & 51.22 & \\
-$\mathtt{WDCMT}$                  & 68.76$_{45.29}$ & 35.71 & 52.75 & \\
+$\mathtt{LDR}_{pred}$                   & \bf 72.76$_{46.35}$ & 35.10 & 50.38 & \bf 52.75\\
+$\mathtt{LDR}_{pred}^{condgru}$  & 71.70$_{46.21}$ & 35.09 & 51.22 & 52.67\\
+$\mathtt{WDCMT}$                        & 68.76$_{45.29}$ & \bf 35.71 & \bf 52.75 & 52.40\\
 \hline
@@
 We therefore decided to allocate $8$~cells for the domain specific part. 
-This suggests that we could easily accommodate more domains with the same architecture, and even reserve some regions to handle supplementary data. %incoming training data.  
+This suggests that we could easily accommodate more domains with the same architecture and even reserve some regions to handle supplementary data (see below). %incoming training data.  
 % \fyTodo{Easily accomodate more domains}
@@
 \hline
- size=4   & 74.65$_{49.61}$ & 37.42 & 54.49 & \\
- size=8   & 74.26$_{49.90}$ & 37.67 & 54.07 & \\
- size=16 & 74.15$_{49.10}$ & 37.78 & 54.56 & \\
- size=32 & 75.10$_{48.61}$ & 37.64 & 54.29 & \\
- size=64 & 74.50$_{50.17}$ & 37.27 & 54.50 & \\
+ size=4   & 74.65$_{49.61}$ & 37.42 & 54.49 & 55.52\\
+ size=8   & 74.26$_{49.90}$ & 37.67 & 54.07 & 55.33\\
+ size=16 & 74.15$_{49.10}$ & \bf 37.78 & \bf 54.56 & 55.50\\
+ size=32 & \bf 75.10$_{48.61}$ & 37.64 & 54.29 & \bf 55.68\\
+ size=64 & 74.50$_{\bf 50.17}$ & 37.27 & 54.50 & 55.42\\
 \hline
@@
 
-\subsection{Additional Domain Scenario \label{ssec:additiona_domain}}
+\subsection{Additional Domain Scenario \label{ssec:additional_domain}}
 
 We now evaluate the ability of our model to integrate new domains, a very common scenario for industrial MT.
-In this setting, we consider that we have a model ($\mathtt{LDR}_{oracle}$) trained as before for EMEA, EPPS and ECB during 200,000 iterations, which needs to accomodate new training data in the IT domain.
-Assuming that we have reserved extra empty embedding cells\footnote{For this experiment, word embeddings contain 480 cells for the generic region and 32 cells for domain specific regions (8 cells x 4 regions).} for this new domain, we resume training with 4 domains during 100,000 additional iterations, yielding a updated model  $\mathtt{LDR}_{oracle}^*$.
-Experiments are reported for the English$\rightarrow$French language pair in Table~\ref{tab:add}, where for comparison purposes we also report numbers obtained with continued training with the $\mathtt{Mixed}$ model, training for the same number of iterations and using the same four datasets ($\mathtt{Mixed}^*$).
+In this setting, we consider that we have a model ($\mathtt{LDR}_{oracle}$) trained as before for EMEA, EPPS and ECB during 200,000 iterations, which needs to process new training data from the IT domain.
+Assuming that we have reserved extra empty embedding cells\footnote{For this experiment, word embeddings contain 480 cells for the generic region and 32 cells for domain specific regions (8 cells x 4 regions).} for this domain, we resume training with 4 domains during 100,000 additional iterations, yielding a updated model  $\mathtt{LDR}_{oracle}^*$.
+Results for the English$\rightarrow$French language pair are in Table~\ref{tab:add}, where for comparison purposes we also report numbers obtained with continued training with the $\mathtt{Mixed}$ model, training for the same number of iterations and using the same four datasets ($\mathtt{Mixed}^*$).
 
@@
 \hline
-$\mathtt{Mixed}$                & 67.69$_{47.60}$ & 37.50 & 53.49 & 13.91 & \\
-$\mathtt{Mixed}^*$             & 66.49$_{45.79}$ & 37.59 & 55.07 & 51.78 & \\
+$\mathtt{Mixed}$                & 67.69$_{47.60}$ & 37.50 & 53.49 & 13.91 & 43.15\\
+$\mathtt{Mixed}^*$             & 66.49$_{45.79}$ & 37.59 & 55.07 & 51.78 & 52.73\\
 \hline
-$\mathtt{LDR}_{oracle}$     & 74.26$_{49.90}$ & 37.67 & 54.07 & 13.40 & \\
-$\mathtt{LDR}_{oracle}^*$  & 76.17$_{49.71}$ & 37.48 & 55.12 & 55.24 & \\
+$\mathtt{LDR}_{oracle}$     & 74.26$_{\bf 49.90}$ & \bf 37.67 & 54.07 & 13.40 & 44.85\\
+$\mathtt{LDR}_{oracle}^*$  & \bf 76.17$_{49.71}$ & 37.48 & \bf 55.12 & \bf 55.24 & \bf 56.00\\
 \hline
@@
 
-As expected, a huge improvement in performance is observed for the IT test set when learning includes in-domain data for both models, with$\mathtt{LDR}_{oracle}^*$ outperforming $\mathtt{Mixed}^*$ by a wide margin.
+As expected, a huge improvement in performance is observed for the IT test set when learning includes in-domain data for both models, with $\mathtt{LDR}_{oracle}^*$ outperforming $\mathtt{Mixed}^*$ by a wide margin.
 %
@@
 finally, using IT data does not impact the quality of translations for the EPPS domain of any of the models.
-Overall better results are obtained by our $\mathtt{LDR}_{oracle}$ model when trained including the additional domain dataset, demonstrating the ability of the model to seamlessly integrate new domains.
+Overall better results are obtained by our $\mathtt{LDR}_{oracle}^*$ model trained with data from an additional source, demonstrating the ability of our architecture to seamlessly integrate new domains.
 % as supplementary training data becomes available.
@@
 To check this hypothesis, we performed the following analysis of embeddings learned with the multi-domain Transformer system for English:French.
-For each unit\footnote{In this study, we work with BPE units, not words, meaning that in many cases we observe the variation of use of word parts.
-As we work with a large inventory, many of these units correspond to actual words, and we focus on these in our comments.
+For each unit\footnote{In this study, we work with BPE units meaning that in many cases we observe the variation of use of \emph{word parts}.
+As we work with a large inventory, many of these units still correspond to actual words and we focus on these in our comments.
 We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $i$ is $e(v,i) = M_ge_g(u) + M_de_i(v)$ (cf.\ equation~\eqref{eq:embedding}).
 This process yields $d$ lists of $k$ nearest neighbours.
-The size of their intersection should then be a sign of a variation of use across domains; conversely, an identical set of neighbours across domains should reflect the stability of word use.
+A small intersection should then be a sign of a variation of use across domains; conversely, an near-identical set of neighbours across domains should reflect the stability of word use.
 Table~\ref{tab:embeddings} list the 10 units with the smaller (respectively larger) intersection (we use $k=10$ and $d=3$).
@@
 The case of \emph{impairment} is pretty clear, occuring in EMEA mostly in terms such as ``hepatic impairment'' or ``renal impairment'', and translating into French as \textsl{insuffisance}.
-In ECB, its collocate are quite different, and impairment often occurs in terms such as ``cost subject to impairments'' (French: \emph{co\^ut soumis \`a des r\'eductions de valeur}).
+In ECB, its collocates are quite different and impairment often occurs in terms such as ``cost subject to impairments'' (French: \emph{co\^ut soumis \`a des r\'eductions de valeur}).
 Likewise, ``convenience'' seems to have its general meaning (``for convenience'') in EMEA, but appears in ECB in the specific context of ``convenience credit card'' (French \textsl{carte de cr\' edit \`a remboursement diff\'er\'e}).
@@
 The decoder side can thus attend separately to these two representations to generate its output. 
-In this approach, the shared and domain-specific part kept separated in the deeper layers of the network, whereas we try to localise the differences between domains at the lexical level, based on a much cheaper computational architecture.
+In this approach, the shared and domain-specific parts are kept separated in the deeper layers of the network, whereas we try to localise the differences between domains at the lexical level, based on a much cheaper computational architecture.
 Another trait of this proposal is the ability to automatically infer domain information for test sentences; 
@@
 It is noticeable that these results are obtained without impacting the architecture or training complexity
-, making our approach a simple baseline for further studies in multi-domain translation.
+, making our approach an effective baseline for further studies in multi-domain translation.
 We have also shown that our approach can dynamically handle new domains; and that the domain-specific embeddings often reflect differences of senses.
@@
 \end{document}
+
+64.57	26.47	68.67
+68.35	17.02	32.87
+36.19	26.29	40.71
+24.72	18.36	74.05
+63.48	26.27	66.95
+70.90	26.30	68.90
+71.31	25.98	73.74
+70.89	26.53	68.63
+69.51	26.31	66.86

### 2019-10-11 2019/iwslt/submission.tex
CR

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\name{Firstname Lastname}
-%\name{MinhQuang Pham$^{\dag\ddag}$, \ Josep Crego$^\dag$, \ Fran\c cois Yvon$^\ddag$, \ Jean Senellart$^\dag$}
+%\name{Firstname Lastname}
+\name{MinhQuang Pham$^{\dag\ddag}$, \ Josep Crego$^\dag$, \ Fran\c cois Yvon$^\ddag$, \ Jean Senellart$^\dag$}
 
@@
 
-\address{ Affiliation / Address line 1 \\ Affiliation / Address line 2 \\ Affiliation / Address line 3 \\ {\small \tt email@domain}}
-%\address{ $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France \\ \small \tt firstname.lastname@systrangroup.com \\ $^\ddag$LIMSI, CNRS,  Universit\'e Paris-Saclay 91405 Orsay, France \\ {\small \tt firstname.lastname@limsi.fr}}
+%\address{ Affiliation / Address line 1 \\ Affiliation / Address line 2 \\ Affiliation / Address line 3 \\ {\small \tt email@domain}}
+\address{ $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France \\ \small \tt firstname.lastname@systrangroup.com \\ $^\ddag$LIMSI, CNRS,  Universit\'e Paris-Saclay 91405 Orsay, France \\ {\small \tt firstname.lastname@limsi.fr}}
 
@@
 \begin{align}
-  \tilde{e}_k(v) =& M_g e_g(v) + \sum_{i \in [1,..,d]} M_i * e_i(v) * \delta(i=k) \nonumber \\
+  \tilde{e}_k(v) =& M_g e_g(v) + \sum_{i \in [1,..,d]} M_i \times e_i(v) \times \delta(i=k) \nonumber \\
    & = M [e_g(v); e_1'(v,k) \dots; e_d'(v,k)], \label{eq:embedding}
@@
 %
-Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers the computation in equation~\eqref{eq:embedding}.
-In our experiments, we consider both the attentional RNN architecture of \cite{Bahdanau15learning} and the Transformer architecture of \cite{Vaswani17attention}.
+Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers 
+by the embeddings defined
+in equation~\eqref{eq:embedding}. In our experiments, we consider both the attentional RNN architecture of \cite{Bahdanau15learning} and the Transformer architecture of \cite{Vaswani17attention}.
 
@@
 In this setting, we consider that we have a model ($\mathtt{LDR}_{oracle}$) trained as before for EMEA, EPPS and ECB during 200,000 iterations, which needs to process new training data from the IT domain.
-Assuming that we have reserved extra empty embedding cells\footnote{For this experiment, word embeddings contain 480 cells for the generic region and 32 cells for domain specific regions (8 cells x 4 regions).} for this domain, we resume training with 4 domains during 100,000 additional iterations, yielding a updated model  $\mathtt{LDR}_{oracle}^*$.
+Assuming that we have reserved extra empty embedding cells\footnote{For this experiment, word embeddings contain 480 cells for the generic region and 32 cells for domain specific regions (8 cells x 4 regions).} for this domain, we resume training with 4 domains during 100,000 additional iterations, yielding an updated model  $\mathtt{LDR}_{oracle}^*$.
 Results for the English$\rightarrow$French language pair are in Table~\ref{tab:add}, where for comparison purposes we also report numbers obtained with continued training with the $\mathtt{Mixed}$ model, training for the same number of iterations and using the same four datasets ($\mathtt{Mixed}^*$).
@@
 finally, using IT data does not impact the quality of translations for the EPPS domain of any of the models.
-Overall better results are obtained by our $\mathtt{LDR}_{oracle}^*$ model trained with data from an additional source, demonstrating the ability of our architecture to seamlessly integrate new domains.
+Overall better results are obtained by our $\mathtt{LDR}_{oracle}^*$ model trained with data from an additional source, demonstrating the ability of our architecture to seamlessly integrate a new domain.
 % as supplementary training data becomes available.
@@
 As we work with a large inventory, many of these units still correspond to actual words and we focus on these in our comments.
-We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $i$ is $e(v,i) = M_ge_g(u) + M_de_i(v)$ (cf.\ equation~\eqref{eq:embedding}).
+We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $i$ is $e(v,i) = M_ge_g(u) + M_ie_i(v)$ (cf.\ equation~\eqref{eq:embedding}).
 This process yields $d$ lists of $k$ nearest neighbours.
@@
 \cite{Farajian17multidomain} learn a generic system and propose to dynamically adapt the network weights in an unsupervised manner using a small sample of training data that resembles the test data, an idea already explored in \cite{Sennrich13multidomain}.
-Their main contribution is to propose an a method to relate the amount of adaptation of the network parameters to the similarity between the adaptation sample and the test sentence: the higher the similarity, the more agressive the adaptation.
+Their main contribution is to propose a method to relate the amount of adaptation of the network parameters to the similarity between the adaptation sample and the test sentence: the higher the similarity, the more agressive the adaptation.
 %

### 2020-03-04 2020/coling2020/multidomain.tex
start writing things

%
% File coling2020.tex
%
% Contact: feiliu@cs.ucf.edu & liang.huang.sh@gmail.com
%% Based on the style files for COLING-2018, which were, in turn,
%% Based on the style files for COLING-2016, which were, in turn,
%% Based on the style files for COLING-2014, which were, in turn,
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn, 
%% based on the style files for ACL-2010, which were, in turn, 
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{coling2020}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage[draft]{todo}
%\setlength\titlebox{5cm}
%\colingfinalcopy % Uncomment this line for the final submission

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\title{Revisiting Multi-Domain Machine Translation}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
Abstract goes here
\end{abstract}

\section{Introduction}
\label{intro}

%
% The following footnote without marker is needed for the camera-ready
% version of the paper.
% Comment out the instructions (first text) and uncomment the 8 lines
% under "final paper" for your variant of English.
% 
\blfootnote{
    %
    % for review submission
    %
    \hspace{-0.65cm}  % space normally used by the marker
    Place licence statement here for the camera-ready version. See
    Section~\ref{licence} of the instructions for preparing a
    manuscript.
    %
    % % final paper: en-uk version 
    %
    % \hspace{-0.65cm}  % space normally used by the marker
    % This work is licensed under a Creative Commons 
    % Attribution 4.0 International Licence.
    % Licence details:
    % \url{http://creativecommons.org/licenses/by/4.0/}.
    % 
    % % final paper: en-us version 
    %
    % \hspace{-0.65cm}  % space normally used by the marker
    % This work is licensed under a Creative Commons 
    % Attribution 4.0 International License.
    % License details:
    % \url{http://creativecommons.org/licenses/by/4.0/}.
}

Statistical Machine Translation, whether probabilistic or neural, rests on well understood machine learning principles. Given a training sample of paired source-target sentences $(\src,trg)$ drawn from an unknown underlying distribution $\mathcal{D}$, a model parameterized (here a translation function $f_{\theta}$) by $\theta$ is trained by minimizing the empirical expectation of a loss function $\ell(f_\theta(\src), \trg)$. This approach ensures that the expectation of the translation loss will remain low when translating new (test) sentences \emph{drawn from the same distribution}.

Owing to the great variability of language data, this ideal situation is rarely met in practice, warranting the study of an alternative scenario, where the test distribution $\mathcal{D'}$ differs from $\mathcal{D}$. In this setting, \emph{domain adaptation} (DA) methods are required. DA has a long history in Machine Learning in general \cite{} and in NLP in particular \cite{Daume,Blitzer}, and various techniques exist to handle both the situation where a (small) training sample drawn from $\mathcal{D'}$ is available, or where only samples of the source (or from the target) domain are given (see \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
% In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the

A related problem is multi-domain (MD) translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol} when \emph{a single system} has to be trained  and tested with data from multiple domains. As pointed out eg.\ by \newcite{Dredze08online}, this setting borrows both from DA, but also from multi-task learning \cite{}.  If the intuitions behind MD may seem simple, the requirements of a MDMT system are rarely spelled out, which makes their evaluation difficult. For instance, should a simple average BLEU score be reported, as in \cite{}, or a domain-weighted average? How about domains not seen in training or domain labelling errors? Should they be taken into account?

A related question: Can a MD learning improve over DA ?

In this paper, we try to answer these questions and to spell out in a more precise fashion the requirements that a MDMT system should meet, and to design adapted metrics aimed to evaluate how well these expectations are met. We then use these metrics to compare several recent proposals from the state-of-the-art in a typical MDMT scenario.

\section{Requirements of multi-domain MT \label{sec:requirements}}

\section{Experimental settings \label{sec:experiments}}
\subsection{Data}
\subsection{MD systems}

\section{Results and discussion \label{sec:results}}

\section{Conclusion and outlook \label{sec:conclusion}}

% --------------
% \section*{Acknowledgements}

% The acknowledgements should go immediately before the references.  Do
% not number the acknowledgements section. Do not include this section
% when submitting your paper for review.

% include your own bib file like this:
\bibliographystyle{coling}
\bibliography{multidomain}

\end{document}

### 2020-03-11 2020/coling2020/multidomain.tex
more thoughts, more refs

@@
 
-Statistical Machine Translation, whether probabilistic or neural, rests on well understood machine learning principles.
-Given a training sample of paired source-target sentences $(\src,trg)$ drawn from an unknown underlying distribution $\mathcal{D}$, a model parameterized (here a translation function $f_{\theta}$) by $\theta$ is trained by minimizing the empirical expectation of a loss function $\ell(f_\theta(\src), \trg)$.
+Data-based Machine Translation (MT), whether statistical or neural, rests on well understood machine learning principles.
+Given a training sample of paired source-target sentences $(\src,trg)$ drawn from an underlying distribution $\mathcal{D}$, a model parameterized by $\theta$ (here a translation function $f_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(f_\theta(\src, \trg))$.
 This approach ensures that the expectation of the translation loss will remain low when translating new (test) sentences \emph{drawn from the same distribution}.
@@
 In this setting, \emph{domain adaptation} (DA) methods are required.
-DA has a long history in Machine Learning in general \cite{} and in NLP in particular \cite{Daume,Blitzer}, and various techniques exist to handle both the situation where a (small) training sample drawn from $\mathcal{D'}$ is available, or where only samples of the source (or from the target) domain are given (see \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
+DA has a long history in Machine Learning in general \cite{} and in NLP in particular \cite{Daume,Blitzer}, and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available, or where only samples of source-side (or target-side) sentences are available (see \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
 % In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
 
-A related problem is multi-domain (MD) translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol} when \emph{a single system} has to be trained  and tested with data from multiple domains.
+A seemingly related problem is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multi-domain,Pham19generic} when \emph{a single system} has to be trained  and tested with data from multiple domains.
 As pointed out eg.\ by \newcite{Dredze08online}, this setting borrows both from DA, but also from multi-task learning \cite{}.
-If the intuitions behind MD may seem simple, the requirements of a MDMT system are rarely spelled out, which makes their evaluation difficult.
-For instance, should a simple average BLEU score be reported, as in \cite{}, or a domain-weighted average?
+If the intuitions behind MD may seem simple, the exact specifications of MDMTs system are rarely spelled out, which makes their evaluation difficult.
+For instance, should a simple score (eg.\ BLEU) averaged across domain be reported, as in \cite{Farajian17multidomain}, or a domain-weighted average as in \cite{}?
 How about domains not seen in training or domain labelling errors?
-Should they be taken into account?
+How should they be taken into account in the evaluation?
 
@@
 
-In this paper, we try to answer these questions and to spell out in a more precise fashion the requirements that a MDMT system should meet, and to design adapted metrics aimed to evaluate how well these expectations are met.
-We then use these metrics to compare several recent proposals from the state-of-the-art in a typical MDMT scenario.
+In this paper, we try to answer these questions and to spell out in a more precise fashion the requirements that a MDMT system should meet.
+Our main contribution are (a) ;  (b) adapted metrics aimed to evaluate how well these expectations are met; (c) a thorough reanalysis of recent approaches from the SOTA based on these new metrics.
+
+% From Kobus et al
+% Our goal is toallow a model built from a diverse set of trainingdata  to  produce  in-domain  translations.
+This is, to extend the coverage of generic NMT models to specific domains, with their specialized terminology and style, without lowering translation quality on more generic data.
 
 \section{Requirements of multi-domain MT \label{sec:requirements}}
+
+\section{Evaluation issues \label{sec:evaluation}}
+
+A general consideration is that evaluation should report our best estimates of the test loss, computed on a sample that is representative of the anticipated test conditions.
+Three scenarios need be distinguished:
+\begin{enumerate}
+\item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
+\item all domains are equally likely and important in testing (the prior over domains is uniform), which warrants the computation of per domain scores, and of unweighted averages; note that in this situation, the optimal training regime is to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data;
+\item the test distribution is neither equal to the training set, nor uniform; again if this distribution is known in advance an optimal domain mix should be used in training to compute the baseline scores, rather than using the empirical distribution of all the training data;
+\end{itemize}
 
@@
 \section{Results and discussion \label{sec:results}}
+
+\section{Related work \label{sec:related}} 
+\cite{Farajian17multidomain} performs on-line adaptation at the sentence-level - processing a test sentence triggers the selection of a dedicated (small) set of fine-tuning instances assumed to be similar to the input text; using these, the generic NMT system is then tuned for some epochs, before delivering its translation.
+No clear notion of a domain, so it would pass all our tests, presumably.
+Also improves upon systems fine tuned individually, sometimes by a wide margin.
+Test distribution is different from the training distribution (for domains).
+
+The authors of \cite{Kobus17domaincontrol} propose to add extra information, either in the form of an extra (initial) domain-token, or in the form of additional domain-feature associated to each word.
+This approach is extended in \cite{Pham19generic} where source word embeddings are likely to vary across domains: this is achieved by spliting the embedding vector into a generic and a domain-specific part (one for each possible domain).
+
+The motivation of \cite{Zeng18multidomain} are to leverage as much as possible the available information in the training data.
+The main assumption is that translation examples can be useful not only for test sentences of the same domains, but also for other domains.
+In this approach, the upper layers the MT aimed to sort out domain specific information on the one hand, and domain-agnostic information on the other hand, using auxiliary tasks, hoping that the domain-agnostic part will be useful across the board.
+These two sets of representations are then input to a conventional attentional sequence to sequence models; scores are reported separately per domain, suggesting that all domains are equally important at test time.
+
+The motivations of \cite{Tars18multidomain} conjoins both (a) they argue that fine-tuning is brittle, owing to the reduced size of the tuning data, and can impair generalization; (b) they also argue that having one domain-agnostic system is a practical solution in many contexts
+\begin{quote}
+  In-domain fine-tuning has two main shortcomings: it depends on the availability of sufficientamounts of in-domain data in order to avoid over-fitting and it results in degraded performance for all other domains.
+The latter means that for translating multiple domains one has to run an individual NMT system for each domain.
+\end{quote}
+In their experiments, they consider that all domains are equally important and report individual scores for each known domain; further experiments are performed with automatically induced domain tags, which show that working with automatically induced tags (at the sentence level) is a viable alternative to using corpus labels as domain tags.
 

### 2020-03-16 2020/coling2020/multidomain.tex
more thoughts

@@
 % in the camera-ready version and ask you to change it back.
+
+\newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
+\newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
 
@@
 
-A seemingly related problem is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multi-domain,Pham19generic} when \emph{a single system} has to be trained  and tested with data from multiple domains.
-As pointed out eg.\ by \newcite{Dredze08online}, this setting borrows both from DA, but also from multi-task learning \cite{}.
+A seemingly related problem is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} when \emph{a single system} has to be trained  and tested with data from multiple domains.
+As pointed out eg.\ by \newcite{Dredze08online}, this setting borrows much from DA, but also from multi-task learning \cite{}.
 If the intuitions behind MD may seem simple, the exact specifications of MDMTs system are rarely spelled out, which makes their evaluation difficult.
@@
 
+\subsection{MDMT: a tentative definition}
+Informally, a multi-domain system shoud be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
+\begin{itemize}
+\item the more common situation is when both training and testing documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other documents in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
+It is usally assumed that all the set of test domains is the same as the set of training domains -- if this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT \cite{Firat16multiway,Johnson17google,Platanios18contextual}  This is the more general setting, already proposed in \cite{Dredze09multidomain}.\todo{do we need a name: such as supervised MDMT ?
+and mostly supervised MDMT for the second case ?}
+
+\item It can also be the case that domains are only known in training, while test inputs may or may not be associated with explicit domains.
+This is for instance the setting adopted \cite{Farajian17multidomain}, where a data-based similarity measure is used instead of a categorical domain assignment.
+Note that automatic domain identification can also be effectively performed in the first setting, especially when domains are related.
+
+\item A last setting, that has not received much attention for MT (see however \cite{Sennrich13multidomain}), is where train and test domains are automatically infered from data, using clustering techniques.
+Even though domain labels are typically available in most training scenario, it might be preferable, owing to the multiple sources of intra-corpus and inter-corpus variabilities, to resort to automatic domain assignments.
+\end{itemize}
+
+In the rest of the paper, we will mostly adhere to the first definition, but will also experiment with the other two alternative scenarios, as they also provide useful context to evaluate the actual robustness of MDMT systems.
+
+\subsection{Why adopt MDMT? \label{ssec:whymdmt}}
+The motivations for considering MDMT are many. Looking at the literature, the main arguments in support of MDMT seem to be the following:
+\begin{enumerate}
+\item[M1] computational / practical arguments.
+A first motivation for MDMT is practical: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop only one system instead of having to optimize (and maintain) multiple MT engines \cite{Sennrich13multidomain}.
+Such arguments are all the more convincing that the number of domain considered is large.
+
+\item[M2] statistical arguments.
+In addition, the training data available for each domain is very unevenly distributed, and domain trained or adapted on small datasets will yield large estimator variance.
+Building mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}.
+Under this view, MDMT would be especially useful for domains with little training data: this is for instance what we observe for multilingual MT from English: an increase for under-resource languages, at the cost of a decrease for well-resourced languages \cite{Massively paper}.
+A slightly distinct line of reasoning is that mixing domain has a positive regularization effect for all domains.
+By introducing variability in training, it prevents domain adaptation from overfitting the available adapation data, and could help improve generalization even for well-resourced domains.
+
+\item[M3] linguistic arguments.
+A related line of reasoning contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words on the other hand are domain agnostic and remain semantically stable accross domain.
+A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain to improve the representation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
+\end{enumerate}
+
+These multiple reasons for adopting MDMT will also yield a variance of expectations with respect to the performance of this approach: if the motivation is primarily computational, then a drop in MT quality with respect to fine-tuning each domain might be acceptable; if it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some domains, if not all, over individually trained systems.
+
 \section{Evaluation issues \label{sec:evaluation}}
 
-A general consideration is that evaluation should report our best estimates of the test loss, computed on a sample that is representative of the anticipated test conditions.
+A general consideration is that evaluation should report realistic estimates of the test loss, computed on a sample deemed representative of the anticipated test conditions.
 Three scenarios need be distinguished:
@@
 \item the test distribution is neither equal to the training set, nor uniform; again if this distribution is known in advance an optimal domain mix should be used in training to compute the baseline scores, rather than using the empirical distribution of all the training data;
-\end{itemize}
+\end{enumerate}
+
+\section{Challenging multi-domain translation systems}
+In this section, we discuss the operational requirements that can be expected from a true multi-domain system, as well as ways to evaluate how much these requirements are met in practice.
+
+\subsection{Multi-domain systems should be effective \label{ssec:effective}}
+% \subsubsection{Performance}
+The first requirement of a MDMT is that it should perform well in the face of mix-domain data. These yields the following desired properties.
+
+\paragraph{[P1]} The first, and most obvious requirement of a MDMT system, is that it should perform well, at least better that the obvious baseline which would simply disregard domain labels.
+Testing this requirement is simple, notwithstanding the previous remarks above regarding evaluation procedures.
+
+\paragraph{[P2]} Additionally, following motivation~(M2) above, one would expect that MDMT improves over fine tuning, at least in the the situation where data is scarce.
+To evaluate this ability, we will perform two measurements, using a real as well as an artificial scenario.
+In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains [P2.1]; in the artificial scenario, we unevenly split a training domain in two parts: the prediction here is that a MDMT should yield a gain for the two subdomains, which both should benefit from the additional training data.
+
+\subsection{Robustness with respect to domain definition \label{ssec:robusness}}
+
+\subsection{Effectiveness in handling a large number of domains \label{ssec:scaling}}
 
@@
 \section{Related work \label{sec:related}}
+
 \cite{Farajian17multidomain} performs on-line adaptation at the sentence-level - processing a test sentence triggers the selection of a dedicated (small) set of fine-tuning instances assumed to be similar to the input text; using these, the generic NMT system is then tuned for some epochs, before delivering its translation.
@@
 
-The authors of \cite{Kobus17domaincontrol} propose to add extra information, either in the form of an extra (initial) domain-token, or in the form of additional domain-feature associated to each word.
+The authors of \cite{Kobus17domaincontrol} propose to inject additional domain feature in the model, either in the form of an extra (initial) domain-token, or in the form of additional domain-feature associated to each word.
 This approach is extended in \cite{Pham19generic} where source word embeddings are likely to vary across domains: this is achieved by spliting the embedding vector into a generic and a domain-specific part (one for each possible domain).
+Such approach strongly relies on earlier proposals from the era of statistical MT \cite{Daume07frustratingly,Clark12onesystem}.
 
-The motivation of \cite{Zeng18multidomain} are to leverage as much as possible the available information in the training data.
+The motivation of \cite{Zeng18multidomain,Su19exploring} are to leverage as much as possible the available information in the training data.
 The main assumption is that translation examples can be useful not only for test sentences of the same domains, but also for other domains.

### 2020-03-17 2020/coling2020/multidomain.tex
End of methodology, start of experiments

@@
 \usepackage{latexsym}
+\usepackage{graphicx}
+\usepackage{xcolor}
 \usepackage[draft]{todo}
@@
 % in the camera-ready version and ask you to change it back.
+\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{orange}{#1}}}
 
@@
 In this setting, \emph{domain adaptation} (DA) methods are required.
-DA has a long history in Machine Learning in general \cite{} and in NLP in particular \cite{Daume,Blitzer}, and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available, or where only samples of source-side (or target-side) sentences are available (see \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
+DA has a long history in Machine Learning in general \cite{Daume06domain} and in NLP in particular \cite{Blitzer07domain}, and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
 % In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
@@
 \item the more common situation is when both training and testing documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other documents in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
-It is usally assumed that all the set of test domains is the same as the set of training domains -- if this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT \cite{Firat16multiway,Johnson17google,Platanios18contextual}  This is the more general setting, already proposed in \cite{Dredze09multidomain}.\todo{do we need a name: such as supervised MDMT ?
+It is usally assumed that all the set of test domains is the same as the set of training domains -- if this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}.
+This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name: such as supervised MDMT ?
 and mostly supervised MDMT for the second case ?}
@@
 
-\subsection{Why adopt MDMT? \label{ssec:whymdmt}}
+\subsection{Why implement MDMT? \label{ssec:whymdmt}}
 The motivations for considering MDMT are many. Looking at the literature, the main arguments in support of MDMT seem to be the following:
@@
 Building mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}.
-Under this view, MDMT would be especially useful for domains with little training data: this is for instance what we observe for multilingual MT from English: an increase for under-resource languages, at the cost of a decrease for well-resourced languages \cite{Massively paper}.
+Under this view, MDMT would be especially useful for domains with little training data: this is for instance what we observe for multilingual MT from English: an increase for under-resource languages due to positive transfer, at the cost of a decrease for well-resourced languages \cite{Arivazhagan19massively}.
 A slightly distinct line of reasoning is that mixing domain has a positive regularization effect for all domains.
@@
 A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain to improve the representation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
+It is expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related, and which can share more information.
 \end{enumerate}
 
-These multiple reasons for adopting MDMT will also yield a variance of expectations with respect to the performance of this approach: if the motivation is primarily computational, then a drop in MT quality with respect to fine-tuning each domain might be acceptable; if it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some domains, if not all, over individually trained systems.
+These various reasons for adopting MDMT also yield a variance of expectations with respect to the performance of this approach: if the motivation is primarily computational, then a drop in MT quality with respect to fine-tuning each domain might be acceptable; if it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some domains, if not all, over individually trained systems.
 
@@
 
-A general consideration is that evaluation should report realistic estimates of the test loss, computed on a sample deemed representative of the anticipated test conditions.
+In theory, evaluation procedures should report realistic estimates of the test loss, computed on a sample deemed representative of the anticipated test conditions.
 Three scenarios need be distinguished:
@@
 \item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
-\item all domains are equally likely and important in testing (the prior over domains is uniform), which warrants the computation of per domain scores, and of unweighted averages; note that in this situation, the optimal training regime is to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data;
+\item all domains are equally likely and important in testing (the prior over test domains is uniform), which warrants the computation of per domain scores and of unweighted averages; note that in this situation, the optimal training regime is to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data;
 \item the test distribution is neither equal to the training set, nor uniform; again if this distribution is known in advance an optimal domain mix should be used in training to compute the baseline scores, rather than using the empirical distribution of all the training data;
@@
 
-\paragraph{[P1]} The first, and most obvious requirement of a MDMT system, is that it should perform well, at least better that the obvious baseline which would simply disregard domain labels.
-Testing this requirement is simple, notwithstanding the previous remarks above regarding evaluation procedures.
+\paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} The first, and most obvious requirement of a MDMT system, is that it should perform well, at least better that the obvious baseline which would simply disregard domain labels.
+Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
 
-\paragraph{[P2]} Additionally, following motivation~(M2) above, one would expect that MDMT improves over fine tuning, at least in the the situation where data is scarce.
-To evaluate this ability, we will perform two measurements, using a real as well as an artificial scenario.
-In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains [P2.1]; in the artificial scenario, we unevenly split a training domain in two parts: the prediction here is that a MDMT should yield a gain for the two subdomains, which both should benefit from the additional training data.
+\paragraph{[P2-TUN]} Additionally, following motivation~(M2) above, one would expect that MDMT improves over fine tuning, at least in domains where data is scarce.
+To evaluate this ability, we perform two measurements, using a real as well as an artificial scenario.
+In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains [P2.1]; in the artificial scenario, we unevenly split a single training domain in two parts which are considered as distinct in training.
+The prediction here is that a MDMT should yield a clear gain for the two ``sub-domains'' as compared to fine-tuning, as both should benefit from the supplementary of relevant training.
 
 \subsection{Robustness with respect to domain definition \label{ssec:robusness}}
+A second set of requirements is related to the definition of domains.
+As repeatedly pointed out in the literature, training corpora for MT are often collected opportunistically, and the view that each each corpora constitues a single domain is a gross approximation.
+MDMT should aim to make the best of the available data, and be robust to domain assignments.
+To challenge these requirements we propose to perform the following tests.
+
+\paragraph{[P3-HET]}
+The notion of a domain is often a fragile one and an effective MDMT system should be able to discover not only when cross-domain sharing is most useful [P2.2], but also when intra-domain heterogeneity is hurting.
+Requirement [P3] is tested by artificially conjoining two separate domains in one during training.
+The loss in performance with respect to the baseline (correct domain tags) should then be as small as possible.
+
+\paragraph{[P4-ERR]}
+A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of domain tag errors.
+This is in sharp contrast with what is commonly observed with fine-tuning, where giving the wrong domain tag can yield ``catastrophic'' drop in performance.
+This requirement is measured by decoding the test set for domain A with all other domain tags, and report average drop in performance.
+
+\paragraph{[P5-UNK]}
+A related situation occurs when the domain of a test document is unknown.
+Several alternatives are in order: (a) automatically assigning one of the domain tag for decoding should not be much worse than using the true domain tag, assuming this domain is seen in training.
+For test documents from other domains, the same procedure can apply, and should outperform the use of a single mix-domain system.
+These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.
+
+\paragraph{[P6-DYN]}
+Another requirement, more of an operational nature, is that a MDMT system should smoothly evolve to accomodate a larger number of domains, without having to retrain the full system each time new data is available.
+This is requirement [P7] that we challenge by dynamically changing the number of training and test domains.
 
@@
 
+\paragraph{[P7-NUM]} As mentioned above, the computational argument (M1) is all the more sensible as the number of domains increases, making the optimization of many individual systems both ineffective and undesirable.
+For lack of having access to corpora containing a large (eg.\ in the order of 100-1000 domains), we experiment with automatically learned domains, considering a varying number of clusters.
+
 \section{Experimental settings \label{sec:experiments}}
-\subsection{Data}
-\subsection{MD systems}
+
+\subsection{Data and metrics}
+\fyTodo{This is a big mess, yet}
+We experiment with the language pair (English-French) \fyTodo{Both ways ?} and data originating from 8 domains, corresponding to texts from the following data sources: the European Parlement (EPPS) \cite{Koehn05europarl},
+the European Medicines Agency (EMEA), the European Central Bank (ECB) \cite{Tiedemann09news}; documentations for KDE, Ubuntu, GNOME and PHP from the OPUS web site\footnote{\url{http://opus.nlpl.eu}}, collectively merged in a IT-Domain, Koran texts from \cite{}.
+%
+We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).
+Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.
+
+% Using internal test sets corresponds to somewhat ideal conditions.
+
+%for contrast experiments, we also use supplementary test sets from three other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
+This enables us to evaluate the loss in performance when the test set is from a domain not seen in training.
+% The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
+
+\fyDone{Check which corpus are useful}
+\begin{table}[h]
+  \centering
+  \begin{tabular}{ |llll|} %*{4}{|r|}}
+    \hline
+    Corpus & Train & Valid & Test \\ 
+    \hline
+    \multicolumn{4}{l}{English $\rightarrow$ French }\\
+    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
+    \hline
+    EMEA  & 1.09M & 1,000 & 1,000$_{(300)}$\\
+    ECB    & 0.19M & 1,000 & 1,000     \\
+    EPPS   & 2.01M  & 1,000 & 1,000  \\
+    IT         & 0.54M  & 1,000 & 1,000 \\  
+    \hline
+    \multicolumn{4}{l}{English $\rightarrow$ German}\\
+    %\multicolumn{4}{|l|}{Vocab size -  En:30,159, De: 30,698}\\ 
+    \hline
+    EMEA  & 1.11M & 1,000 & 1,000$_{(300)}$ \\
+    ECB     &  0.11M & 1,000 & 1,000  \\
+    EPPS   & 1.92M & 1,000 & 1,000 \\ 
+    \hline
+\end{tabular}
+\caption{Corpora statistics.}
+\label{tab:Corpora}
+\end{table}
+
+Note that the EMEA dataset distributed on the OPUS site contains multiple sentence duplicates. 
+We therefore report below two numbers as $S_{(T)}$: the first ($S$) is comparable to what has been published on earlier studies (eg.\ \cite{Zeng18multidomain}), the second one ($T$) is obtained by making the test entirely disjoint from the training (700~duplicated sentences are discarded).
+
+To reduce the number of lexical units and make our systems open-vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations. \fyTodo{I need explanations here}
+
+\subsection{Baselines \label{ssec:baselines}}
+
+Our baselines correspond to standard practices for mix-domain systems.
+Using the ransformer architecture of \shortcite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
+
+\begin{itemize}
+\item a generic model trained on a concatenation of all corpora (\texttt{Mixed}).
+We develop two versions of this system, one with domains are unbalanced, and one where all domains are equally represented in training (cf.\ section~\ref{sec:evaluation});
+\item models tuned separately on each domain for respectively (10000, 15000, 5000)\fyTodo{fix this} iterations using in-domain data (\texttt{ft$_{EMEA}$}, \texttt{ft$_{EPPS}$}, \texttt{ft$_{ECB}$});
+\end{itemize}
+
+For all models, we set the embeddings size and the size of hidden layers to~512.
+Our Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
+
+% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
+ %
+Training is performed using Adam, with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$).
+In all cases, we use a batch size of~128 and a dropout rate of 0.1 for all layers.
+All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
+
+\subsection{Multi-domain systems \label{ssec:systems}}
+Our comparison of multidomain systems includes the following reimplementations of multiple proposals from the recent literature:
+\begin{itemize}
+\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced as an additional feature for each word \fyTodo{Check this} (\texttt{DC});
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part;
+\item the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT}), using the transformer-based version of \cite{Su19exploring}: in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.
+\item The layer adaptive layer of \cite{Jiang19multidomain} ...
+\end{itemize}
+
+Note that we did not consider including the system of \cite{Farajian17multidomain} for the following reasons. This system has no problem with scaling to a large number of domains.\fyTodo{TBC}
 

### 2020-03-17 2020/coling2020/multidomain.tex
Last update for tonight

@@
 \end{enumerate}
+The papers in the literature are seldom very explicit regarding their assumption regarding the test distribution.
+In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 is the most common situation.
+
+Another aspect of the evaluation that is hardly, if ever, discussed, is the computational cost of training one single vs multiple systems.
 
@@
 
-Our baselines correspond to standard practices for mix-domain systems.
-Using the ransformer architecture of \shortcite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
+Our baselines correspond to standard practices for mix-domain systems.\footnote{We however omit domain-specific systems trained with a subset of the data, which are often inferior to the mix-domain strategy eg.\ \cite{Britz2017mixing}.} Using the Transformer architecture of \shortcite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
 
@@
 We develop two versions of this system, one with domains are unbalanced, and one where all domains are equally represented in training (cf.\ section~\ref{sec:evaluation});
-\item models tuned separately on each domain for respectively (10000, 15000, 5000)\fyTodo{fix this} iterations using in-domain data (\texttt{ft$_{EMEA}$}, \texttt{ft$_{EPPS}$}, \texttt{ft$_{ECB}$});
+\item models fine-tuned \cite{Freitag16fast} separately on each domain for respectively (10000, 15000, 5000)\fyTodo{fix this} iterations using in-domain data (\texttt{ft$_{EMEA}$}, \texttt{ft$_{EPPS}$}, \texttt{ft$_{ECB}$}).
+In our setting, fine-tuning may update all the parameters of the initial generic model.\fyTodo{Check This}.
 \end{itemize}
@@
 For all models, we set the embeddings size and the size of hidden layers to~512.
-Our Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
+Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
 
@@
 \subsection{Multi-domain systems \label{ssec:systems}}
-Our comparison of multidomain systems includes the following reimplementations of multiple proposals from the recent literature:
+Our comparison of multidomain systems includes the following reimplementations of recent proposals from the literature:
 \begin{itemize}
@@
 \item the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT}), using the transformer-based version of \cite{Su19exploring}: in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.
-\item The layer adaptive layer of \cite{Jiang19multidomain} ...
+\item The layer adaptive approach of \cite{Jiang19multidomain} ...
 \end{itemize}
 
-Note that we did not consider including the system of \cite{Farajian17multidomain} for the following reasons. This system has no problem with scaling to a large number of domains.\fyTodo{TBC}
+Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence, and slightly diverges from our notion of MDMT, which should be able to handle domain information in training and testing.
+This system has shown to meet many of our requirements, and would also have no problem scaling to an large number of domains.\fyTodo{TBC}
 
@@
 
-\cite{Farajian17multidomain} performs on-line adaptation at the sentence-level - processing a test sentence triggers the selection of a dedicated (small) set of fine-tuning instances assumed to be similar to the input text; using these, the generic NMT system is then tuned for some epochs, before delivering its translation.
-No clear notion of a domain, so it would pass all our tests, presumably.
-Also improves upon systems fine tuned individually, sometimes by a wide margin.
-Test distribution is different from the training distribution (for domains).
-
 The authors of \cite{Kobus17domaincontrol} propose to inject additional domain feature in the model, either in the form of an extra (initial) domain-token, or in the form of additional domain-feature associated to each word.
 This approach is extended in \cite{Pham19generic} where source word embeddings are likely to vary across domains: this is achieved by spliting the embedding vector into a generic and a domain-specific part (one for each possible domain).
-Such approach strongly relies on earlier proposals from the era of statistical MT \cite{Daume07frustratingly,Clark12onesystem}.
+Such feature-based approaches strongly relies on earlier proposals from the era of statistical MT \cite{Daume07frustratingly,Clark12onesystem}.
 
-The motivation of \cite{Zeng18multidomain,Su19exploring} are to leverage as much as possible the available information in the training data.
-The main assumption is that translation examples can be useful not only for test sentences of the same domains, but also for other domains.
-In this approach, the upper layers the MT aimed to sort out domain specific information on the one hand, and domain-agnostic information on the other hand, using auxiliary tasks, hoping that the domain-agnostic part will be useful across the board.
-These two sets of representations are then input to a conventional attentional sequence to sequence models; scores are reported separately per domain, suggesting that all domains are equally important at test time.
+In \cite{Britz17mixing}, alternative methods are considered which aim to ensure that domain information is actually used in a mix-domain system.
+Three alternatives are considered, using either domain classification (or domain normalization, via adversarial training) on source or target representations.
+There is no clear winner in either of their three language pairs considered, which only distinguish a limited number of domains.
+The same kinds of techniques are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the upper layers of the MT aimed to sort out domain specific information on the one hand, and domain-agnostic information on the other hand, using auxiliary tasks, hoping that the domain-agnostic part will be useful across the board.
+These two sets of representations are then input to a conventional attentional sequence-to -sequence models in \cite{Zeng18multidomain}, and with Transformers in \cite{Su19exploring}; scores are reported separately per domain, suggesting that all domains are equally important at test time.
 
@@
 In their experiments, they consider that all domains are equally important and report individual scores for each known domain; further experiments are performed with automatically induced domain tags, which show that working with automatically induced tags (at the sentence level) is a viable alternative to using corpus labels as domain tags.
+
+\cite{Farajian17multidomain} performs on-line adaptation at the sentence-level - processing a test sentence triggers the selection of a dedicated (small) set of fine-tuning instances assumed to be similar to the input text; using these, the generic NMT system is then tuned for some epochs, before delivering its translation.
+No clear notion of a domain, so it would pass all our tests, presumably.
+Also improves upon systems fine-tuned individually, sometimes by a wide margin.
+Test distribution is different from the training distribution (for domains).
 

### 2020-03-18 2020/coling2020/multidomain.tex
Updates following the discussion with Minh

@@
 \item the more common situation is when both training and testing documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other documents in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
-It is usally assumed that all the set of test domains is the same as the set of training domains -- if this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}.
+It is usally assumed that all the set of test domains is the same as the set of training domains -- when this is not the case, we can have zero-shot domain adaptation, similar to what is studied in multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}.
 This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name: such as supervised MDMT ?
@@
 \item It can also be the case that domains are only known in training, while test inputs may or may not be associated with explicit domains.
-This is for instance the setting adopted \cite{Farajian17multidomain}, where a data-based similarity measure is used instead of a categorical domain assignment.
-Note that automatic domain identification can also be effectively performed in the first setting, especially when domains are related.
+This is for instance the setting adopted \cite{Zeng18multidomain}, where automatic domain identification is performed for test sentences.
+Note that automatic domain identification can also be effectively performed in the first setting, especially in the presence of related domains in training.
 
-\item A last setting, that has not received much attention for MT (see however \cite{Sennrich13multidomain}), is where train and test domains are automatically infered from data, using clustering techniques.
-Even though domain labels are typically available in most training scenario, it might be preferable, owing to the multiple sources of intra-corpus and inter-corpus variabilities, to resort to automatic domain assignments.
+\item A last setting, that has not received much attention for MT (see however \cite{Sennrich13multidomain}), is where both train and test domains are automatically infered from data, using clustering techniques.
+Even though domain labels are typically available in most training scenario, these approaches assumes that might be preferable, owing to the multiple sources of intra-corpus and inter-corpus variabilities, to resort to automatic domain assignments.
+The online approach of \cite{Farajian17multidomain}, where surface similarities cues, rather that categorical tags, are used for fine-tuning, is a variant of this strategy.
 \end{itemize}
 
-In the rest of the paper, we will mostly adhere to the first definition, but will also experiment with the other two alternative scenarios, as they also provide useful context to evaluate the actual robustness of MDMT systems.
+In the rest of the paper, we mostly adhere to the first definition, and assume that multidomain MT is mostly about effectively using domain tags in training and testing.
+We also experiment in Section~\ref{sec:results} with the other two alternative scenarios, as they also provide useful contexts to evaluate the actual robustness of MDMT systems.
 
-\subsection{Why implement MDMT? \label{ssec:whymdmt}}
+\subsection{Reasons for using MDMT \label{ssec:whymdmt}}
 The motivations for considering MDMT are many. Looking at the literature, the main arguments in support of MDMT seem to be the following:
@@
 A first motivation for MDMT is practical: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop only one system instead of having to optimize (and maintain) multiple MT engines \cite{Sennrich13multidomain}.
-Such arguments are all the more convincing that the number of domain considered is large.
+Such arguments are all the more convincing that the number of domains considered is large.
 
 \item[M2] statistical arguments.
-In addition, the training data available for each domain is very unevenly distributed, and domain trained or adapted on small datasets will yield large estimator variance.
+The training data available for each domain is very unevenly distributed, and domain trained or adapted on small datasets will yield large estimator variance.
 Building mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}.
@@
 It is expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related, and which can share more information.
+\fyTodo{polysemy}
 \end{enumerate}
 
-These various reasons for adopting MDMT also yield a variance of expectations with respect to the performance of this approach: if the motivation is primarily computational, then a drop in MT quality with respect to fine-tuning each domain might be acceptable; if it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some domains, if not all, over individually trained systems.
+These various reasons for adopting MDMT are not mutually exclusive.
+Each however yields variable expectations with respect to the performance of this approach: if the motivation is primarily computational, then a drop in MT quality with respect to fine-tuning each domain might be acceptable; if it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some domains, if not all, over individually trained systems.
 
@@
 \item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
-\item all domains are equally likely and important in testing (the prior over test domains is uniform), which warrants the computation of per domain scores and of unweighted averages; note that in this situation, the optimal training regime is to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data;
+\item all domains are equally likely and important in testing (the prior over test domains is uniform), which would warrant the computation of per domain scores and of unweighted averages, as reported in \cite{Farajian17multidomain,Su19exploring}.
+Note that in this situation, the optimal training regime would be to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data.
 \item the test distribution is neither equal to the training set, nor uniform; again if this distribution is known in advance an optimal domain mix should be used in training to compute the baseline scores, rather than using the empirical distribution of all the training data;
 \end{enumerate}
+
 The papers in the literature are seldom very explicit regarding their assumption regarding the test distribution.
-In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 is the most common situation.
+In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 most faithfully reflects the expected testing condition.
 
-Another aspect of the evaluation that is hardly, if ever, discussed, is the computational cost of training one single vs multiple systems.
+Another aspect of the evaluation that is hardly, if ever, discussed, is the computational cost of training one single versus multiple systems.\fyTodo{Keep this ?}
 
@@
 In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains [P2.1]; in the artificial scenario, we unevenly split a single training domain in two parts which are considered as distinct in training.
-The prediction here is that a MDMT should yield a clear gain for the two ``sub-domains'' as compared to fine-tuning, as both should benefit from the supplementary of relevant training.
+The prediction here is that a MDMT should yield a clear gain for the both pseudo sub-domains as compared to fine-tuning, as both systems should benefit from the supplementary amount of relevant training.
 
@@
 A second set of requirements is related to the definition of domains.
-As repeatedly pointed out in the literature, training corpora for MT are often collected opportunistically, and the view that each each corpora constitues a single domain is a gross approximation.
-MDMT should aim to make the best of the available data, and be robust to domain assignments.
+As repeatedly pointed out in the literature, training corpora for MT are often collected opportunistically, and the view that each each corpora constitues a single domain is a gross approximation.\footnote{Two of our own domains are composed of several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data, and be robust to domain assignments.
 To challenge these requirements we propose to perform the following tests.
@@
 For test documents from other domains, the same procedure can apply, and should outperform the use of a single mix-domain system.
-These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.
+\fyTodo{Systems react to unknown domains}These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.
 
@@
 Another requirement, more of an operational nature, is that a MDMT system should smoothly evolve to accomodate a larger number of domains, without having to retrain the full system each time new data is available.
-This is requirement [P7] that we challenge by dynamically changing the number of training and test domains.
+This is a requirement [P6] that we challenge by dynamically changing the number of training and test domains.
 
@@
 
-\subsection{Data and metrics}
-\fyTodo{This is a big mess, yet}
-We experiment with the language pair (English-French) \fyTodo{Both ways ?} and data originating from 8 domains, corresponding to texts from the following data sources: the European Parlement (EPPS) \cite{Koehn05europarl},
-the European Medicines Agency (EMEA), the European Central Bank (ECB) \cite{Tiedemann09news}; documentations for KDE, Ubuntu, GNOME and PHP from the OPUS web site\footnote{\url{http://opus.nlpl.eu}}, collectively merged in a IT-Domain, Koran texts from \cite{}.
+\subsection{Data and metrics \label{ssec:corpora}}
+
+We experiment with translation from English into French and use data originating from \fyTodo{8} domains, corresponding to texts from the following data sources: the European Parlement (EPPS)\fyTodo{which version, link} \cite{Koehn05europarl}, the UFAL Medical corpus V1.0 (MED)\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (ECB) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (LAW) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a IT-Domain, Ted Talks [Talks] \cite{Cettolo12}, and the Koran \cite{}.
+Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenised with in-house tools.
+\fyTodo{Newsco, Sciello}
+
 %
-We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}).
-Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.
-
-% Using internal test sets corresponds to somewhat ideal conditions.
+We randomly split each corpus into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}). Validation sets are used to chose the best model according to the average BLEU score.
 
@@
 
-\fyDone{Check which corpus are useful}
 \begin{table}[h]
@@
     \hline
-    \multicolumn{4}{l}{English $\rightarrow$ German}\\
-    %\multicolumn{4}{|l|}{Vocab size -  En:30,159, De: 30,698}\\ 
-    \hline
-    EMEA  & 1.11M & 1,000 & 1,000$_{(300)}$ \\
-    ECB     &  0.11M & 1,000 & 1,000  \\
-    EPPS   & 1.92M & 1,000 & 1,000 \\ 
-    \hline
 \end{tabular}
@@
 
-Note that the EMEA dataset distributed on the OPUS site contains multiple sentence duplicates. 
-We therefore report below two numbers as $S_{(T)}$: the first ($S$) is comparable to what has been published on earlier studies (eg.\ \cite{Zeng18multidomain}), the second one ($T$) is obtained by making the test entirely disjoint from the training (700~duplicated sentences are discarded).
-
-To reduce the number of lexical units and make our systems open-vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations. \fyTodo{I need explanations here}
+To reduce the number of lexical units and make our systems open-vocabulary, we finally compute a Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a joint corpus containing all sentences (in both languages).
 
@@
 
-Our baselines correspond to standard practices for mix-domain systems.\footnote{We however omit domain-specific systems trained with a subset of the data, which are often inferior to the mix-domain strategy eg.\ \cite{Britz2017mixing}.} Using the Transformer architecture of \shortcite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
+Our baselines correspond to standard practices for multi-domain systems.\footnote{We however omit domain-specific systems trained with a subset of the data, which are often inferior to the mix-domain strategy eg.\ \cite{Britz17mixing}.} Using the Transformer architecture of \shortcite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
 
@@
 \item a generic model trained on a concatenation of all corpora (\texttt{Mixed}).
-We develop two versions of this system, one with domains are unbalanced, and one where all domains are equally represented in training (cf.\ section~\ref{sec:evaluation});
-\item models fine-tuned \cite{Freitag16fast} separately on each domain for respectively (10000, 15000, 5000)\fyTodo{fix this} iterations using in-domain data (\texttt{ft$_{EMEA}$}, \texttt{ft$_{EPPS}$}, \texttt{ft$_{ECB}$}).
-In our setting, fine-tuning may update all the parameters of the initial generic model.\fyTodo{Check This}.
+We develop two versions of this system, one where domains are unbalanced (\texttt{Mixed-Nat}) and one where all domains are equally represented in training (\texttt{Mixed-Bal}) (cf.\ section~\ref{sec:evaluation});
+\item models fine-tuned \cite{Freitag16fast} separately on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
+In our setting, we again contrast two versions: full fine-tuning (\texttt{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates an adaptation block (\texttt{FT-Block}).
 \end{itemize}
@@
 \begin{itemize}
-\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced as an additional feature for each word \fyTodo{Check this} (\texttt{DC});
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part;
-\item the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT}), using the transformer-based version of \cite{Su19exploring}: in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.
-\item The layer adaptive approach of \cite{Jiang19multidomain} ...
+\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\texttt{DC-Tag}), or introduced in the form of a supplementary feature for each word (\texttt{DC-Feat}).
+\item the proposal of \cite{Britz17mixing}, where the domain tag is introduced in the target side, and is learned from the reference in training, and infered as a regular owrd in testing (\texttt{TTM});
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\texttt{DC-Feat});
+\item the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT}), using the improved transformer-based version of \cite{Su19exploring}: in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.
+\item the layer adaptive approach of \cite{Jiang19multidomain} ...
+\item Several original variant of a multidomain version of \cite{Bapna19simple} \fyTodo{explain}
 \end{itemize}
@@
 \section{Results and discussion \label{sec:results}}
+\subsection{Evaluation of raw performance}
 
@@
 
-In \cite{Britz17mixing}, alternative methods are considered which aim to ensure that domain information is actually used in a mix-domain system.
+In \cite{Britz17mixing}, several methods are considered which aim to ensure that domain information is actually used in a mix-domain system.
 Three alternatives are considered, using either domain classification (or domain normalization, via adversarial training) on source or target representations.

### 2020-03-24 2020/coling2020/multidomain.tex
more discussions and related work - to be continued with more results

@@
 \newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
+\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
+\newcommand{\system}[1]{\texttt{\textbf{#1}}}
 
@@
 \item[M1] computational / practical arguments.
-A first motivation for MDMT is practical: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop only one system instead of having to optimize (and maintain) multiple MT engines \cite{Sennrich13multidomain}.
+A first motivation for MDMT is practical: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop only one system instead of having to optimize (and maintain) multiple MT engines \cite{Sennrich13multidomain;Farajian17neural}.
 Such arguments are all the more convincing that the number of domains considered is large.
@@
 
-To reduce the number of lexical units and make our systems open-vocabulary, we finally compute a Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a joint corpus containing all sentences (in both languages).
+To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encodings \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.
 
@@
 
-Our baselines correspond to standard practices for multi-domain systems.\footnote{We however omit domain-specific systems trained with a subset of the data, which are often inferior to the mix-domain strategy eg.\ \cite{Britz17mixing}.} Using the Transformer architecture of \shortcite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
+Baselines correspond to standard practices for multi-domain systems.\footnote{We however omit domain-specific systems trained with a subset of the data, which are always inferior to the mix-domain strategy eg.\ \cite{Britz17mixing}.} Using the Transformer architecture of \shortcite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
 
@@
 \item a generic model trained on a concatenation of all corpora (\texttt{Mixed}).
-We develop two versions of this system, one where domains are unbalanced (\texttt{Mixed-Nat}) and one where all domains are equally represented in training (\texttt{Mixed-Bal}) (cf.\ section~\ref{sec:evaluation});
+We develop two versions of this system, one where domains are unbalanced (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}) (cf.\ section~\ref{sec:evaluation});
 \item models fine-tuned \cite{Freitag16fast} separately on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
-In our setting, we again contrast two versions: full fine-tuning (\texttt{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates an adaptation block (\texttt{FT-Block}).
+In our setting, we again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a simple adaptation module that is added (with residual connexions) on top of every Transformer layer (\system{FT-Block}).
 \end{itemize}
@@
 Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
-
 % The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
  %
-Training is performed using Adam, with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$).
+Training uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$).
 In all cases, we use a batch size of~128 and a dropout rate of 0.1 for all layers.
-All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
-
+% All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
+\fyTodo{Describe the block adaptation layer}
 \subsection{Multi-domain systems \label{ssec:systems}}
@@
 \begin{itemize}
-\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\texttt{DC-Tag}), or introduced in the form of a supplementary feature for each word (\texttt{DC-Feat}).
-\item the proposal of \cite{Britz17mixing}, where the domain tag is introduced in the target side, and is learned from the reference in training, and infered as a regular owrd in testing (\texttt{TTM});
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\texttt{DC-Feat});
-\item the multi-domain model of \cite{Zeng18multidomain} (\texttt{WDCMT}), using the improved transformer-based version of \cite{Su19exploring}: in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.
-\item the layer adaptive approach of \cite{Jiang19multidomain} ...
-\item Several original variant of a multidomain version of \cite{Bapna19simple} \fyTodo{explain}
+\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
+\item the proposal of \cite{Britz17mixing}, where the domain tag is introduced as an extra word \textsl{on the target side}.
+Training is performed with reference tags and inference is performed with predicted tags, just like for any other target words (\system{TTM});
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});
+\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}), using the improved transformer-based version of \cite{Su19exploring}: in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.
+\item Several original multidomain variants of the approach by \newcite{Bapna19simple}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; residual connexions make it possible to by-pass these modules.
+In a variant, we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
+Note that contrarily to \cite{Bapna19simple} we do not start with a generic system, but train the multi-domain from scratch.\fyTodo{Check this.}
 \end{itemize}
@@
 \section{Results and discussion \label{sec:results}}
-\subsection{Evaluation of raw performance}
+\subsection{Performance of MDMT systems}
+In this section, we discuss the raw performance of MDMT systems trained and tested on 6~domains.
+Results are in Table~\ref{tab:performance}.
+As expected, balancing data in the generic setting makes a great difference (the average is 2 BP better, notably owing to the much better score for \domain{religion}).
+Fine-tuning each domain separately provides us with a better baseline, significantly outperforming \system{generic} for all domains.
+A complete fine-tuning, which updates all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to lack of tuning the capacity of the adapter module.
+\fyTodo{Add 2 averages ?}
+A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on th source is better than on the target, and very comparable to repeating the tag for every source word.
+All improve over the generic baseline (for most domains), but yield weaker results than fine tuning.
+The use of an decidated adaptation component in each of the transformer layer proves an even better strategy and closes the gap with fine-tuning for half of the domains: the use of a gating meachanism yielding a clear improvement over the board.
+The effect of the adaptation layer is especially significant for the smaller domains (\domain{bank}, \domain{IT} and \domain{rel}) for which unplugging this layers causes dramatic performance drops.
+
+\begin{table}
+  \centering
+  \begin{tabular}{|p{4cm}|*{6}{r|}} \hline
+%     &&&&&& \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{med}} & \multicolumn{1}{c|}{\domain{law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talks}} & \multicolumn{1}{c|}{\domain{IT}} & \multicolumn{1}{c|}{\domain{rel}} \\ \hline
+    \system{Mixed-Nat}  & 37.33 & 54.56 & 50.05 & 33.47 & 43.23 & 77.51 \\
+    \system{Mixed-Bal}   &  35.26 & 54.09 & 52.49 & 31.86 & 44.94 & 89.54 \\
+    \system{FT-Full}       & 37.74 & 59.21 & 54.49 & 33.98 & 46.81 & 90.77\\
+   \system{FT-Block}     & 37.3 & 57.93 & 53.91 & 33.79 & 46.69 & 90.17 \\
+
+    Full-finetuned on extended in-domain corpora (news) & && 33.96&&& \\
+
+    \system{DC-Tag}   & 38.06 & 55.27 & 49.9   & 33.24 & 43.54 & 80.45 \\
+    \system{DC-Feat}  & 37.3   & 54.87 & 49.5   & 32.94 & 43.56 & 79.91 \\
+    \system{TTM}        & 37.71 & 54.76 & 49.72 & 33.43 & 43.81 & 79.98 \\
+    \system{LDC-Tag} &&&&& \\
+
+    MDL residual        & 37.86 & 56.01 & 51.2   & 33.5   &  44.4  & 88.34 \\
+    MDL no adapt       & 37.72 & 51.02 & 33.98 & 30.35 & 34.17 & 15.17 \\
+    MDL gated-adapt & 37.68 & 56.49 & 53.17 & 34.05 & 44.57 & 90.73 \\
+
+    ADM (Denny Britz)  &&&&& \\
+    ML (Denny Britz) &&&&& \\
+     \hline
+  \end{tabular}
+  \caption{Translation performance of various MDMT systems }
+  \label{tab:performance}
+\end{table}
+
+\subsection{Redefining domains \label{ssec:redomains}}
+Table~\ref{fig:redomains} summaries the results of XX experiments where we artificially redefine the limits of domains.
+In the three \textsl{split} experiments, we randomly split a domain corpus in two part: a good MDMT system should detect that these two ``domains'' should be mutually beneficial, and should be hardly affected by this change wrt.\ the baseline scenario.\fyTodo{In fact we want more: we want improvements wrt.
+to the finetuned situation with distinct domains} In the merge experiment, we do the reverse and merge two corpora in training, in order to assess the robustness wrt heterogenous domains.
+In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.
+
+\begin{table}
+  \centering
+  \begin{tabular}{|p{2.5cm}|*{9}{r|}} \hline
+    % &&&&&& \\
+    Model / Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & Wrong \\ % \hline
+     Model / Set-up & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.75 / 0.25)}} & \multicolumn{2}{c|}{\domain{bank} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} & domain \\ \hline
+    & \domain{med-1} & \domain{med-2} & \domain{med-1} & \domain{med-2} &  \domain{bank-1} & \domain{bank-2} & \domain{bank} & \domain{law} & \\
+
+    \system{DC-Tag}   & & & & & & & & & \\
+    \system{DC-Feat}  & & & & & & & & & \\
+    \system{TTM}        & & & & & & & & &  \\
+    \system{LDC-Tag} & & & & & & & & & \\
+
+    MDL residual        & & & & & & & & & \\
+    MDL gated-adapt & & & & & & & & & \\
+
+    ADM (Denny Britz) & & & & & & & & & \\
+    ML (Denny Britz)    & & & & & & & & & \\
+     \hline
+  \end{tabular}
+  \caption{Translation performance with variable domain definitions.
+For the Split experiments, we only report the loss in performance for the \domain{med} test set; for the merge experiments for domains \domain{bank} and \domain{Law}.\fyTodo{How about wrong ?}}
+  \label{tab:performance}
+\end{table}
+
+Findings:
+\begin{itemize}
+\item Split UFAL half / half: domain-tag and Gated-res are OK, all systems loose wrt baseline; 
+\item Split UFAL 0,75 / 0.25: same thing, and we even see improvements for domain tags - they are small though
+\item Split Law : only domain feature works ? all the rest have losses, from large to small
+\item Merge domain: bad for many, OK for TTM (as the domain is learnt this is not so bad ?); bad for domain tag, domain feature is again OK ?
+\end{itemize}
+
+\fyTodo{We need significance testing}
+
+\subsection{Automatic domains \label{ssec:autodomains}}
 
@@
 
+A slightly more complex parameter-scheme is presented in \cite{Jiang19multidomain}, which augments a transformer model with domain specific heads, whose individual contributions are regulated at the word (position-level): some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads.
+The results for three language pair outperform the standard baselines for a 2-domain systems (in fr:en and ge:en) and a 4-domain system (zh:en)\footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
+
 \section{Conclusion and outlook \label{sec:conclusion}}

### 2020-03-25 2020/coling2020/multidomain.tex
More results, comments, tables, etc

@@
 A seemingly related problem is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} when \emph{a single system} has to be trained  and tested with data from multiple domains.
-As pointed out eg.\ by \newcite{Dredze08online}, this setting borrows much from DA, but also from multi-task learning \cite{}.
+As pointed out eg.\ by \newcite{Dredze08online}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask}.
 If the intuitions behind MD may seem simple, the exact specifications of MDMTs system are rarely spelled out, which makes their evaluation difficult.
@@
 In this paper, we try to answer these questions and to spell out in a more precise fashion the requirements that a MDMT system should meet.
-Our main contribution are (a) ;  (b) adapted metrics aimed to evaluate how well these expectations are met; (c) a thorough reanalysis of recent approaches from the SOTA based on these new metrics.
+Our main contribution is of methodological nature and is list of expected properties of MDMT system, and associated measurements to evaluate them.
+In addition, we provide a thorough reanalysis of most recent multi-domain approaches from the literature, and also propose two variants of an original model.
 
@@
 \section{Requirements of multi-domain MT \label{sec:requirements}}
+In this section, we try to clearly spell out the reasons for considering a multi-domain scenario, and discuss the implications in terms of performance evaluation.
 
 \subsection{MDMT: a tentative definition}
-Informally, a multi-domain system shoud be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
+Informally, a multi-domain system should be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
 \begin{itemize}
@@
 A first motivation for MDMT is practical: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop only one system instead of having to optimize (and maintain) multiple MT engines \cite{Sennrich13multidomain;Farajian17neural}.
-Such arguments are all the more convincing that the number of domains considered is large.
+Such arguments are all the more convincing that the number of domains considered is large, which is not always the case: for instance \cite{Britz17mixing,Zeng18multidomain,Jiang19multidomain} only consider 2-4 domains in their experiments.
 
@@
 Under this view, MDMT would be especially useful for domains with little training data: this is for instance what we observe for multilingual MT from English: an increase for under-resource languages due to positive transfer, at the cost of a decrease for well-resourced languages \cite{Arivazhagan19massively}.
-A slightly distinct line of reasoning is that mixing domain has a positive regularization effect for all domains.
+A slightly distinct line of reasoning is that mixing domains has a positive regularization effect for all domains.
 By introducing variability in training, it prevents domain adaptation from overfitting the available adapation data, and could help improve generalization even for well-resourced domains.
@@
 
-\section{Evaluation issues \label{sec:evaluation}}
+\subsection{Evaluation issues \label{ssec:evaluation}}
 
@@
 Note that in this situation, the optimal training regime would be to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data.
-\item the test distribution is neither equal to the training set, nor uniform; again if this distribution is known in advance an optimal domain mix should be used in training to compute the baseline scores, rather than using the empirical distribution of all the training data;
+\item the test distribution is neither equal to the training set, nor uniform; again if this distribution is known in advance an optimal domain mix should be used in training to compute the baseline scores, rather than using the empirical distribution of all the training data.
+A more challenging case is when the test includes data from domains unseen in training, that is test instance that have a null probability in training.
+The likeliness of such events should be clearly formulated when describing the evaluation protocol.
 \end{enumerate}
@@
 
-Another aspect of the evaluation that is hardly, if ever, discussed, is the computational cost of training one single versus multiple systems.\fyTodo{Keep this ?}
+Note that when computational cost is the main motivation for considering MDMT, evaluation should report the cost of training one single versus multiple systems, which is hardly even done in practice.
+\fyTodo{Keep this ?}
 
@@
 
-We experiment with translation from English into French and use data originating from \fyTodo{8} domains, corresponding to texts from the following data sources: the European Parlement (EPPS)\fyTodo{which version, link} \cite{Koehn05europarl}, the UFAL Medical corpus V1.0 (MED)\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (ECB) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (LAW) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a IT-Domain, Ted Talks [Talks] \cite{Cettolo12}, and the Koran \cite{}.
+We experiment with translation from English into French and use data originating from 7~domains, corresponding to texts from the following data sources: the UFAL Medical corpus V1.0 (\domain{MED})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{Bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (LAW) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{IT}-Domain, Ted Talks (\domain{Talks}) \cite{Cettolo12wit}, and the Koran (\domain{Rel}).
 Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenised with in-house tools.
-\fyTodo{Newsco, Sciello}
-
+Complementary experiments also use v12 of the News Commentary corpus (\domain{News}).
+Corpus statistics in Table~\ref{tab:Corpora}).
 %
-We randomly split each corpus into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}). Validation sets are used to chose the best model according to the average BLEU score.
+We randomly select in each corpus a development and a test set of 1000 lines, and keep the rest for training.
+Validation sets are used to chose the best model according to the average BLEU score.
+Significance tests are performed as follows...
+\fyTodo{Fix this}.
 
@@
   \centering
-  \begin{tabular}{ |llll|} %*{4}{|r|}}
+  \begin{tabular}{ |lllllll|} %*{4}{|r|}}
     \hline
-    Corpus & Train & Valid & Test \\ 
+    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
+    \domain{Med} & \domain{Law} & \domain{Bank} & \domain{IT} & \domain{Talk} & \domain{Rel} & \domain{News} \\
     \hline
-    \multicolumn{4}{l}{English $\rightarrow$ French }\\
-    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
-    \hline
-    EMEA  & 1.09M & 1,000 & 1,000$_{(300)}$\\
-    ECB    & 0.19M & 1,000 & 1,000     \\
-    EPPS   & 2.01M  & 1,000 & 1,000  \\
-    IT         & 0.54M  & 1,000 & 1,000 \\  
+    2609 (0.68) & 190 (0.05)  & 501 (0.13) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260  \\
     \hline
   \end{tabular}
-\caption{Corpora statistics.}
+\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mix.}
 \label{tab:Corpora}
@@
 % All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
-\fyTodo{Describe the block adaptation layer}
+\fyTodo{Describe the block adaptation layer - voir slides}
+
 \subsection{Multi-domain systems \label{ssec:systems}}
@@
 \item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
-\item the proposal of \cite{Britz17mixing}, where the domain tag is introduced as an extra word \textsl{on the target side}.
-Training is performed with reference tags and inference is performed with predicted tags, just like for any other target words (\system{TTM});
+\item the three proposals of \cite{Britz17mixing}: \system{TTM}, another tag-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
+Training is performed with reference tags and inference is performed with predicted tags, just like for any other target words; \system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations;
 \item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});
-\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}), using the improved transformer-based version of \cite{Su19exploring}: in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.
-\item Several original multidomain variants of the approach by \newcite{Bapna19simple}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; residual connexions make it possible to by-pass these modules.
-In a variant, we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
-Note that contrarily to \cite{Bapna19simple} we do not start with a generic system, but train the multi-domain from scratch.\fyTodo{Check this.}
+\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDC}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\fyTodo{URLs}); the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
+\item A multidomain version of the approach by \newcite{Bapna19simple} denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; residual connexions make it possible to by-pass these modules.
+In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
+Note that contrarily to \cite{Bapna19simple} we do not start with a trained generic system, but train the multi-domain from scratch.
+%\fyTodo{Check this.}
 \end{itemize}
@@
 \section{Results and discussion \label{sec:results}}
-\subsection{Performance of MDMT systems}
+\subsection{Performance of MDMT systems \label{ssec:rawperformance}}
 In this section, we discuss the raw performance of MDMT systems trained and tested on 6~domains.
 Results are in Table~\ref{tab:performance}.
-As expected, balancing data in the generic setting makes a great difference (the average is 2 BP better, notably owing to the much better score for \domain{religion}).
-Fine-tuning each domain separately provides us with a better baseline, significantly outperforming \system{generic} for all domains.
+As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2 BP better, notably owing to the much better score for \domain{religion}).
+As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains.
+Fine-tuning each domain separately provides us with better baselines, significantly outperforming \system{generic} for all domains.
 A complete fine-tuning, which updates all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to lack of tuning the capacity of the adapter module.
-\fyTodo{Add 2 averages ?}
-A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on th source is better than on the target, and very comparable to repeating the tag for every source word.
+\fyDone{Add 2 averages ?}
+A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on the source is better than on the target, and very comparable to repeating the tag for every source word.
 All improve over the generic baseline (for most domains), but yield weaker results than fine tuning.
-The use of an decidated adaptation component in each of the transformer layer proves an even better strategy and closes the gap with fine-tuning for half of the domains: the use of a gating meachanism yielding a clear improvement over the board.
-The effect of the adaptation layer is especially significant for the smaller domains (\domain{bank}, \domain{IT} and \domain{rel}) for which unplugging this layers causes dramatic performance drops.
+\fyTodo{More comments when we have all the results}
+The use of an decidated adaptation component in each of the transformer layer proves an even better strategy and closes the gap with fine-tuning for half of the domains: the use of a gating mechanism yielding a clear improvement over the board.
+The effect of the adaptation layer is especially significant for the smaller domains (\domain{bank}, \domain{IT} and \domain{rel}) for which unplugging this layer causes dramatic performance drops.
 
@@
   \centering
-  \begin{tabular}{|p{4cm}|*{6}{r|}} \hline
+  \begin{tabular}{|p{4cm}|*{8}{r|}} \hline
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{med}} & \multicolumn{1}{c|}{\domain{law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talks}} & \multicolumn{1}{c|}{\domain{IT}} & \multicolumn{1}{c|}{\domain{rel}} \\ \hline
-    \system{Mixed-Nat}  & 37.33 & 54.56 & 50.05 & 33.47 & 43.23 & 77.51 \\
-    \system{Mixed-Bal}   &  35.26 & 54.09 & 52.49 & 31.86 & 44.94 & 89.54 \\
-    \system{FT-Full}       & 37.74 & 59.21 & 54.49 & 33.98 & 46.81 & 90.77\\
-   \system{FT-Block}     & 37.3 & 57.93 & 53.91 & 33.79 & 46.69 & 90.17 \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{med}} & \multicolumn{1}{c|}{\domain{law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talks}} & \multicolumn{1}{c|}{\domain{IT}} & \multicolumn{1}{c|}{\domain{rel}} & \domain{wAVG} & \domain{AVG}\\ \hline
+    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1&49.4\\
+    \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3&51.4\\
+    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8   & 42.8&53.8\\
+   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2    & 42.3&53.3\\
 
-    Full-finetuned on extended in-domain corpora (news) & && 33.96&&& \\
+ %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\
 
-    \system{DC-Tag}   & 38.06 & 55.27 & 49.9   & 33.24 & 43.54 & 80.45 \\
-    \system{DC-Feat}  & 37.3   & 54.87 & 49.5   & 32.94 & 43.56 & 79.91 \\
-    \system{TTM}        & 37.71 & 54.76 & 49.72 & 33.43 & 43.81 & 79.98 \\
+    \system{DC-Tag}   & 38.06 & 55.27 & 49.9   & 33.24 & 43.54 & 80.45 &41.7 & 50.1\\
+    \system{DC-Feat}  & 37.3   & 54.87 & 49.5   & 32.94 & 43.56 & 79.91 &41.4 & 49.9\\
+    \system{TTM}        & 37.71 & 54.76 & 49.72 & 33.43 & 43.81 & 79.98 &41.1 & 49.7\\
     \system{LDC-Tag} &&&&& \\
+    \system{DM}  & 36.5& 52.4& 47.7& 32.1 & 41.0 & 70.4 & 39.7 & 46.7 \\
+    \system{ADM}    &&&&& \\
 
-    MDL residual        & 37.86 & 56.01 & 51.2   & 33.5   &  44.4  & 88.34 \\
-    MDL no adapt       & 37.72 & 51.02 & 33.98 & 30.35 & 34.17 & 15.17 \\
-    MDL gated-adapt & 37.68 & 56.49 & 53.17 & 34.05 & 44.57 & 90.73 \\
-
-    ADM (Denny Britz)  &&&&& \\
-    ML (Denny Britz) &&&&& \\
+    \system{MDL Res}        & 37.86 & 56.01 & 51.2   & 33.5   &  44.4  & 88.34 & 42.1 & 51.9\\
+    MDL no adapt       & 37.72 & 51.02 & 33.98 & 30.35 & 34.17 & 15.17 & 36.4 & 33.7\\
+    \system{MDL Gated} & 37.68 & 56.49 & 53.17 & 34.05 & 44.57 & 90.73 & 42.3 & 53.3\\
      \hline
   \end{tabular}
-  \caption{Translation performance of various MDMT systems }
+  \caption{Translation performance of various MDMT systems. We report the BLEU score for each domain, as well as weighted and unweighted averages.}
   \label{tab:performance}
@@
 \subsection{Redefining domains \label{ssec:redomains}}
-Table~\ref{fig:redomains} summaries the results of XX experiments where we artificially redefine the limits of domains.
-In the three \textsl{split} experiments, we randomly split a domain corpus in two part: a good MDMT system should detect that these two ``domains'' should be mutually beneficial, and should be hardly affected by this change wrt.\ the baseline scenario.\fyTodo{In fact we want more: we want improvements wrt.
-to the finetuned situation with distinct domains} In the merge experiment, we do the reverse and merge two corpora in training, in order to assess the robustness wrt heterogenous domains.
+Table~\ref{fig:redomains} summaries the results of \fyTodo{XX} experiments where we artificially redefine the limits of domains.
+In the three \textsl{split} experiments, we randomly split a domain corpus in two part: a good MDMT system should detect that these two ``domains'' should be mutually beneficial, and should be hardly affected by this change wrt.\ the baseline scenario.
+This is the situation where we should see clear improvements even over finetuning, which only exploits part of the available relevant data.
+In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains.
 In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.
@@
     & \domain{med-1} & \domain{med-2} & \domain{med-1} & \domain{med-2} &  \domain{bank-1} & \domain{bank-2} & \domain{bank} & \domain{law} & \\
-
+    \system{FT-Full}       & & & & & & & & & \\
     \system{DC-Tag}      & & & & & & & & & \\
@@
 
-    MDL residual        & & & & & & & & & \\
-    MDL gated-adapt & & & & & & & & & \\
+    \system{MDL Res}    & & & & & & & & & \\
+    \system{MDL Gated} & & & & & & & & & \\
 
-    ADM (Denny Britz) & & & & & & & & & \\
-    ML (Denny Britz)    & & & & & & & & & \\
+    \system{ADM} & & & & & & & & & \\
+    \system{DM}    & & & & & & & & & \\
+    \system{WDC}  & & & & & & & & & \\
     \hline
@@
   \caption{Translation performance with variable domain definitions.
-For the Split experiments, we only report the loss in performance for the \domain{med} test set; for the merge experiments for domains \domain{bank} and \domain{Law}.\fyTodo{How about wrong ?}}
+For the Split experiments, we only report the loss in performance for the \domain{med} test set, (*) denote significant differences, (+) denotes a significant improvement over \system{Full FT}; for the merge experiments for domains \domain{bank} and \domain{Law}.\fyTodo{How about wrong ?}}
   \label{tab:performance}
@@
 
-Findings:
+Findings so far:
 \begin{itemize}
@@
 
+Another series of experiments aims to evaluate requirement [P6-DYN] as follows: starting with the existing MDMT systems of Section~\ref{ssec: ssec:rawperformance}, we introduce an additional domain (\domain{News}) and resume training with a mixture of data for an additional XXX epochs.
+We contrast this approach with retraining all the systems from scratch, which would be the ideal situation, and report the difference in performance in Table~\ref{tab:XXX}.
+We expect that a good MDMT system should not be significantly affected by the addition of new data.
+
 \subsection{Automatic domains \label{ssec:autodomains}}
+\begin{todo}
+  Distance entre domaines: KL -->
+\end{todo}
 

### 2020-03-26 2020/coling2020/multidomain.tex
jc questions

@@
 Data-based Machine Translation (MT), whether statistical or neural, rests on well understood machine learning principles.
-Given a training sample of paired source-target sentences $(\src,trg)$ drawn from an underlying distribution $\mathcal{D}$, a model parameterized by $\theta$ (here a translation function $f_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(f_\theta(\src, \trg))$.
+Given a training sample of paired source-target sentences $(\src,trg)$ .\jcTodo{shouldnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}$, a model parameterized by $\theta$ (here a translation function $f_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(f_\theta(\src, \trg))$.
 This approach ensures that the expectation of the translation loss will remain low when translating new (test) sentences \emph{drawn from the same distribution}.
@@
 We also experiment in Section~\ref{sec:results} with the other two alternative scenarios, as they also provide useful contexts to evaluate the actual robustness of MDMT systems.
+
+\jcTodo{...a very common scenario in MT industry where clients dispose of mulitiple data corresponding to different domains (differing in genre, thematic, register, style, etc.). }
 
@@
 To evaluate this ability, we perform two measurements, using a real as well as an artificial scenario.
-In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains [P2.1]; in the artificial scenario, we unevenly split a single training domain in two parts which are considered as distinct in training.
+In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains [P2.1] \jcTodo{what is P2.1} ; in the artificial scenario, we unevenly split a single training domain in two parts which are considered as distinct in training.
 The prediction here is that a MDMT should yield a clear gain for the both pseudo sub-domains as compared to fine-tuning, as both systems should benefit from the supplementary amount of relevant training.
+\jcTodo{I dont understand the prediction...
+why running MDMT on 2 artificial subdomains should show gains over fine-tunning?}
 
@@
 A second set of requirements is related to the definition of domains.
-As repeatedly pointed out in the literature, training corpora for MT are often collected opportunistically, and the view that each each corpora constitues a single domain is a gross approximation.\footnote{Two of our own domains are composed of several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data, and be robust to domain assignments.
+As repeatedly pointed out in the literature, training corpora for MT are often collected opportunistically, and the view that each corpora constitues a single domain is a gross approximation.\footnote{Two of our own domains are composed of several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data, and be robust to domain assignments.
 To challenge these requirements we propose to perform the following tests.
@@
 \paragraph{[P3-HET]}
-The notion of a domain is often a fragile one and an effective MDMT system should be able to discover not only when cross-domain sharing is most useful [P2.2], but also when intra-domain heterogeneity is hurting.
+The notion of a domain is often a fragile one and an effective MDMT system should be able to discover not only when cross-domain sharing is most useful [P2.2] \jcTodo{what is P2.2?}, but also when intra-domain heterogeneity is hurting.
 Requirement [P3] is tested by artificially conjoining two separate domains in one during training.
@@
 \item a generic model trained on a concatenation of all corpora (\texttt{Mixed}).
-We develop two versions of this system, one where domains are unbalanced (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}) (cf.\ section~\ref{sec:evaluation});
-\item models fine-tuned \cite{Freitag16fast} separately on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
+We develop two versions of this system, one where domains are unbalanced (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}) (cf.\ section~\ref{sec:evaluation}); \jcTodo{All systems cand be learned following both Bal and Nat.
+does Nat corresponds to wAVG in test?
+is wAVG defined somewhere?}
+\item models fine-tuned \cite{Freitag16fast} \jcTodo{i would use Luong and Maning 2015} separately on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
 In our setting, we again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a simple adaptation module that is added (with residual connexions) on top of every Transformer layer (\system{FT-Block}).
@@
 Results are in Table~\ref{tab:performance}.
-As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2 BP better, notably owing to the much better score for \domain{religion}).
+As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2 BP \jcTodo{BP?} better, notably owing to the much better score for \domain{religion}).
 As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains.

### 2020-03-30 2020/coling2020/multidomain.tex
more experimental results
more comments and ansers to Josep - waiting for numbers

--- 2020-03-25 2020/coling2020/multidomain.tex
@@
 \newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{orange}{#1}}}
+\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{blue}{#1}}}
+\newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{blue}{#1}}}
 
@@
 Data-based Machine Translation (MT), whether statistical or neural, rests on well understood machine learning principles.
-Given a training sample of paired source-target sentences $(\src,trg)$ drawn from an underlying distribution $\mathcal{D}$, a model parameterized by $\theta$ (here a translation function $f_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(f_\theta(\src, \trg))$.
+Given a training sample of paired source-target sentences $(\src,\trg)$\jcDone{shouldnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}$, a model parameterized by $\theta$ (here a translation function $f_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(f_\theta(\src, \trg))$.
 This approach ensures that the expectation of the translation loss will remain low when translating new (test) sentences \emph{drawn from the same distribution}.
@@
 In this setting, \emph{domain adaptation} (DA) methods are required.
-DA has a long history in Machine Learning in general \cite{Daume06domain} and in NLP in particular \cite{Blitzer07domain}, and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
+DA has a long history in Machine Learning in general \cite{Shimodaira00improving,Ben-David09atheory} and in NLP in particular \cite{Daume06domain,Blitzer07domain}, and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
 % In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
 
-A seemingly related problem is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} when \emph{a single system} has to be trained  and tested with data from multiple domains.
+A seemingly related problem is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where \emph{a single system} has to be trained  and tested with data from multiple domains.
 As pointed out eg.\ by \newcite{Dredze08online}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask}.
-If the intuitions behind MD may seem simple, the exact specifications of MDMTs system are rarely spelled out, which makes their evaluation difficult.
-For instance, should a simple score (eg.\ BLEU) averaged across domain be reported, as in \cite{Farajian17multidomain}, or a domain-weighted average as in \cite{}?
-How about domains not seen in training or domain labelling errors?
-How should they be taken into account in the evaluation?
+If the intuitions behind MD may seem simple, the exact specifications of MDMT systems are rarely spelled out, which makes their evaluation difficult: should they perform well when the test data is distributed like the train data or when it is equally distributed across training domains?
+should MDMT also be robust to new domains?
+How should they handle domain labeling errors?
 
-A related question: Can a MD learning improve over DA ?
+A related question concerns the relationship between domain-adaptation and multi-domain translation.
+The latter setting seems more challenging and one would expect that adapting a system toward a known domain $A$ will be better than training a MDMT system where only part of the data from $A$.
+Are there situations where the latter approach can surpass the former as is sometimes expected?
 
 In this paper, we try to answer these questions and to spell out in a more precise fashion the requirements that a MDMT system should meet.
-Our main contribution is of methodological nature and is list of expected properties of MDMT system, and associated measurements to evaluate them.
-In addition, we provide a thorough reanalysis of most recent multi-domain approaches from the literature, and also propose two variants of an original model.
+Our main contribution is of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them.
+Our second contribution is experimental, and we provide a thorough reanalysis of most recent multi-domain approaches from the literature, including two variants of an original model.
+We show that\fyTodo{Spell out conclusions}.
 
@@
 \item the more common situation is when both training and testing documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other documents in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
-It is usally assumed that all the set of test domains is the same as the set of training domains -- when this is not the case, we can have zero-shot domain adaptation, similar to what is studied in multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}.
+It is usally assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}.
 This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name: such as supervised MDMT ?
@@
 
-\item It can also be the case that domains are only known in training, while test inputs may or may not be associated with explicit domains.
-This is for instance the setting adopted \cite{Zeng18multidomain}, where automatic domain identification is performed for test sentences.
+\item It can also be the case that domain labels are only known in training, while test inputs may or may not be associated with explicit domains.
+This is, for instance, the setting adopted \cite{Zeng18multidomain}, where automatic domain identification is performed for test sentences.
 Note that automatic domain identification can also be effectively performed in the first setting, especially in the presence of related domains in training.
@@
 Even though domain labels are typically available in most training scenario, these approaches assumes that might be preferable, owing to the multiple sources of intra-corpus and inter-corpus variabilities, to resort to automatic domain assignments.
-The online approach of \cite{Farajian17multidomain}, where surface similarities cues, rather that categorical tags, are used for fine-tuning, is a variant of this strategy.
+The online approach of \newcite{Farajian17multidomain}, where distributional cues, rather that categorical tags, are used for fine-tuning, is a variant of this strategy.
 \end{itemize}
 
-In the rest of the paper, we mostly adhere to the first definition, and assume that multidomain MT is mostly about effectively using domain tags in training and testing.
-We also experiment in Section~\ref{sec:results} with the other two alternative scenarios, as they also provide useful contexts to evaluate the actual robustness of MDMT systems.
+In the rest of the paper, we mostly adhere to the first definition which correspond to a situation of practical interest for the industry, and assume that multi-domain MT is mostly about effectively using domain tags in training and testing.
+We also experiment in Section~\ref{sec:results} with the other two alternative scenarios, as they also define useful contexts to evaluate the actual robustness of MDMT systems.
+
+\jcDone{...a very common scenario in MT industry where clients dispose of mulitiple data corresponding to different domains (differing in genre, thematic, register, style, etc.). }
 
@@
 \begin{enumerate}
-\item[M1] computational / practical arguments.
-A first motivation for MDMT is practical: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop only one system instead of having to optimize (and maintain) multiple MT engines \cite{Sennrich13multidomain;Farajian17neural}.
+\item[[M1\hspace{-0.5em}]] computational / practical arguments.
+A first motivation for MDMT is practical: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop only one system instead of having to optimize (and maintain) multiple MT engines \cite{Sennrich13multidomain,Farajian17neural}.
 Such arguments are all the more convincing that the number of domains considered is large, which is not always the case: for instance \cite{Britz17mixing,Zeng18multidomain,Jiang19multidomain} only consider 2-4 domains in their experiments.
 
-\item[M2] statistical arguments.
+\item[[M2\hspace{-0.5em}]] statistical arguments.
 The training data available for each domain is very unevenly distributed, and domain trained or adapted on small datasets will yield large estimator variance.
@@
 A slightly distinct line of reasoning is that mixing domains has a positive regularization effect for all domains.
-By introducing variability in training, it prevents domain adaptation from overfitting the available adapation data, and could help improve generalization even for well-resourced domains.
+By introducing variability in training, it prevents domain adaptation from overfitting the available adaptation data and could help improve generalization even for well-resourced domains.
 
-\item[M3] linguistic arguments.
-A related line of reasoning contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words on the other hand are domain agnostic and remain semantically stable accross domain.
-A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain to improve the representation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
-It is expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related, and which can share more information.
+\item[[M3\hspace{-0.5em}]] linguistic arguments.
+A related line of reasoning contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and remain semantically stable accross domain.
+A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the representation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
+It is expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related and which can share more information.
 \fyTodo{polysemy}
@@
 Three scenarios need be distinguished:
-\begin{enumerate}
+\begin{itemize}
 \item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
@@
 Note that in this situation, the optimal training regime would be to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data.
-\item the test distribution is neither equal to the training set, nor uniform; again if this distribution is known in advance an optimal domain mix should be used in training to compute the baseline scores, rather than using the empirical distribution of all the training data.
-A more challenging case is when the test includes data from domains unseen in training, that is test instance that have a null probability in training.
-The likeliness of such events should be clearly formulated when describing the evaluation protocol.
-\end{enumerate}
+\item the test distribution is neither equal to the training set, nor uniform.
+Here again, if this distribution is known in advance, an optimal domain mix should be used in training to compute baseline scores, rather than using the empirical distribution of all the training data.
+A more challenging case is when the test set includes data from domains unseen in training: this situation, where test instances have a null probability in training, goes agains the basic statistical assumptions of DA \cite{Shimodaira00improving}.
+It is thus important that the likeliness of such events should be clearly formulated when describing the evaluation protocol.
+\end{itemize}
 
-The papers in the literature are seldom very explicit regarding their assumption regarding the test distribution.
-In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 most faithfully reflects the expected testing condition.
+The papers in the literature are seldom explicit regarding their assumptions regarding the test distribution.
+In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 most faithfully reflects the expected testing conditions.
+Another important issue is the choice of the evaluation metrics: if BLEU is still the standard metric used in most studies, it is not necessary the best choice to measure subtle improvements of lexical choices \cite{Irvine13measuring}.
 
-Note that when computational cost is the main motivation for considering MDMT, evaluation should report the cost of training one single versus multiple systems, which is hardly even done in practice.
+A last issue regarding evaluation is related to the measurement of computational costs: when this is the main a motivation for using MDMT, we expect that computational costs would be reported and compared, which however is hardly ever done in practice.
 \fyTodo{Keep this ?}
@@
 % \subsubsection{Performance}
-The first requirement of a MDMT is that it should perform well in the face of mix-domain data. These yields the following desired properties.
+The first set of requirements for MDMT systems is that they should perform well in the face of mix-domain data. These yields the following desired properties.
 
-\paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} The first, and most obvious requirement of a MDMT system, is that it should perform well, at least better that the obvious baseline which would simply disregard domain labels.
-Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
+\paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} A first expectation for MDMT is that they perform better that the obvious baseline which disregards domain labels.
+% Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
 
-\paragraph{[P2-TUN]} Additionally, following motivation~(M2) above, one would expect that MDMT improves over fine tuning, at least in domains where data is scarce.
-To evaluate this ability, we perform two measurements, using a real as well as an artificial scenario.
-In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains [P2.1]; in the artificial scenario, we unevenly split a single training domain in two parts which are considered as distinct in training.
-The prediction here is that a MDMT should yield a clear gain for the both pseudo sub-domains as compared to fine-tuning, as both systems should benefit from the supplementary amount of relevant training.
+\paragraph{[P2-TUN]} Additionally, following motivation~(M2) above, one would expect that MDMT improves over fine-tuning, at least in domains where data is scarce, or in situations where several domains are close.
+To evaluate this, we perform two measurements, using a real as well as an artificial scenario.
+In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains (property [P2.1-TUN]).
+In the artificial scenario, we split a single training domain in two parts which are considered as distinct in training.
+The expectation here is that a MDMT should yield a clear gain in the both pseudo sub-domains, which should benefit from the supplementary amount of relevant training (this is property [P2.2-TUN]).
+In this situation, MDMT should even outperform fine-tuning on either of the pseudo sub-domain.
+\jcDone{I dont understand the prediction... why running MDMT on 2 artificial subdomains should show  gains over fine-tunning?}\fyTodo{Clearer now ?}
 
 \subsection{Robustness with respect to domain definition \label{ssec:robusness}}
-A second set of requirements is related to the definition of domains.
-As repeatedly pointed out in the literature, training corpora for MT are often collected opportunistically, and the view that each each corpora constitues a single domain is a gross approximation.\footnote{Two of our own domains are composed of several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data, and be robust to domain assignments.
-To challenge these requirements we propose to perform the following tests.
+A second set of requirements is related to the definition of a domain.
+As repeatedly pointed out in the literature, training corpora for MT are often collected opportunistically and the view that each corpora constitutes a single domain is often a gross approximation.\footnote{Two of our own domains are composed of several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments.
+To challenge these requirements we propose the evaluate the following properties.
 
 \paragraph{[P3-HET]}
-The notion of a domain is often a fragile one and an effective MDMT system should be able to discover not only when cross-domain sharing is most useful [P2.2], but also when intra-domain heterogeneity is hurting.
-Requirement [P3] is tested by artificially conjoining two separate domains in one during training.
+The notion of a domain is often a fragile one and an effective MDMT system should be able to discover not only when cross-domain sharing is most useful (cf.\ property [P2.2-TUN]) \jcDone{what is P2.2?}, but also when intra-domain heterogeneity is hurting.
+Requirement [P3-HET] is tested by artificially conjoining two separate domains in one during training.
 The loss in performance with respect to the baseline (correct domain tags) should then be as small as possible.
@@
 We experiment with translation from English into French and use data originating from 7~domains, corresponding to texts from the following data sources: the UFAL Medical corpus V1.0 (\domain{MED})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{Bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (LAW) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{IT}-Domain, Ted Talks (\domain{Talks}) \cite{Cettolo12wit}, and the Koran (\domain{Rel}).
-Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenised with in-house tools.
 Complementary experiments also use v12 of the News Commentary corpus (\domain{News}).
-Corpus statistics in Table~\ref{tab:Corpora}).
-%
+Corpus statistics in Table~\ref{tab:Corpora}.
+Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools.
+To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encodings \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.%
+
 We randomly select in each corpus a development and a test set of 1000 lines, and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score.
+Validation sets are used to chose the best model according to the average BLEU score computed with\fyTodo{??}.
 Significance tests are performed as follows...
@@
 
-To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encodings \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.
+We measure the distance between domains using the $\mathcal{H}$-Divergence introduced in \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains.
+We trained a domain discriminator is a multilayer perceptron trained as follows.
+Based on this, we see that ...
+\fyTodo{TBC}.
 
@@
 \item a generic model trained on a concatenation of all corpora (\texttt{Mixed}).
-We develop two versions of this system, one where domains are unbalanced (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}) (cf.\ section~\ref{sec:evaluation});
-\item models fine-tuned \cite{Freitag16fast} separately on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
+We develop two versions of this system, one where domains unbalance reflects the distribution of our parallel data (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}) (cf.\ section~\ref{sec:evaluation}); \jcTodo{All systems cand be learned following both Bal and Nat.
+does Nat corresponds to wAVG in test?
+is wAVG defined somewhere?}
+\item models fine-tuned \cite{Luong15stanford,Freitag16fast} \jcDone{i would use Luong and Manning 2015} separately on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
 In our setting, we again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a simple adaptation module that is added (with residual connexions) on top of every Transformer layer (\system{FT-Block}).
@@
 Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
+Some systems (see text below) additionally use an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
+The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
 % The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
@@
 % All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
-\fyTodo{Describe the block adaptation layer - voir slides}
+\fyDone{Describe the block adaptation layer - voir slides}
 
@@
 \item the multi-domain model of \cite{Zeng18multidomain} (\system{WDC}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\fyTodo{URLs}); the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
-\item A multidomain version of the approach by \newcite{Bapna19simple} denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; residual connexions make it possible to by-pass these modules.
+\item A multidomain version of the approach of \newcite{Bapna19simple} denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; residual connexions make it possible to by-pass these modules.
 In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
-Note that contrarily to \cite{Bapna19simple} we do not start with a trained generic system, but train the multi-domain from scratch.
+Note that contrarily to \cite{Bapna19simple} we do not start with a trained generic system, but learn the multi-domain from scratch.
 %\fyTodo{Check this.}
@@
 Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence, and slightly diverges from our notion of MDMT, which should be able to handle domain information in training and testing.
-This system has shown to meet many of our requirements, and would also have no problem scaling to an large number of domains.\fyTodo{TBC}
+This system has shown to meet many of our requirements, and would also have no problem scaling to a large number of domains.\fyTodo{TBC}
 
@@
 Results are in Table~\ref{tab:performance}.
-As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2 BP better, notably owing to the much better score for \domain{religion}).
+As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2 BLEU points \jcDone{BP?} better, notably owing to the much better results for \domain{REL}).
 As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains.
+As all other systems are trained with an unbalanced data distribution, use the weighted average to perform global comparisons.
+
 Fine-tuning each domain separately provides us with better baselines, significantly outperforming \system{generic} for all domains.
@@
 A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on the source is better than on the target, and very comparable to repeating the tag for every source word.
-All improve over the generic baseline (for most domains), but yield weaker results than fine tuning.
+All (except \system{DM} and \system{ADM}) improve over the generic baseline (for most domains), matching requirement [P1-LAB], but yield weaker results than fine tuning.
+As expected
 \fyTodo{More comments when we have all the results}
@@
   \centering
-  \begin{tabular}{|p{4cm}|*{8}{r|}} \hline
+  \fyTodo{Fix column size}
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{med}} & \multicolumn{1}{c|}{\domain{law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talks}} & \multicolumn{1}{c|}{\domain{IT}} & \multicolumn{1}{c|}{\domain{rel}} & \domain{wAVG} & \domain{AVG}\\ \hline
+    Model / Domain & \multicolumn{1}{c|}{\domain{med}} & \multicolumn{1}{c|}{\domain{law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talks}} & \multicolumn{1}{c|}{\domain{it}} & \multicolumn{1}{c|}{\domain{rel}} & \domain{wAVG} & \domain{AVG}\\ \hline
     \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1&49.4\\
@@
 
- %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\
+ %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
 
-    \system{DC-Tag}   & 38.06 & 55.27 & 49.9   & 33.24 & 43.54 & 80.45 &41.7 & 50.1\\
-    \system{DC-Feat}  & 37.3   & 54.87 & 49.5   & 32.94 & 43.56 & 79.91 &41.4 & 49.9\\
-    \system{TTM}        & 37.71 & 54.76 & 49.72 & 33.43 & 43.81 & 79.98 &41.1 & 49.7\\
+    \system{DC-Tag}   & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5 &41.7 & 50.1\\
+    \system{DC-Feat}  & 37.3   & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 &41.4 & 49.9\\
+    \system{TTM}        & 37.7 & 54.8 & 49.7 & 33.4 & 43.8 & 80.0 &41.1 & 49.7\\
     \system{LDC-Tag} &&&&& \\
-    \system{DM}  & 36.5& 52.4& 47.7& 32.1 & 41.0 & 70.4 & 39.7 & 46.7 \\
-    \system{ADM}    &&&&& \\
-
-    \system{MDL Res}        & 37.86 & 56.01 & 51.2   & 33.5   &  44.4  & 88.34 & 42.1 & 51.9\\
-    MDL no adapt       & 37.72 & 51.02 & 33.98 & 30.35 & 34.17 & 15.17 & 36.4 & 33.7\\
-    \system{MDL Gated} & 37.68 & 56.49 & 53.17 & 34.05 & 44.57 & 90.73 & 42.3 & 53.3\\
+    \system{DM}             & 36.5 & 52.4  & 47.7& 32.1 & 41.0 & 70.4 & 39.7 & 46.7 \\
+    \system{ADM}           & 37.1 & 53.8  & 48.9 &32.8 & 42.3 & 78.5 & 40.7 & 48.9 \\
+    \system{MDL Res}     & 37.9 & 56.0  & 51.2   & 33.5   &  44.4  & 88.3 & 42.1 & 51.9\\
+    MDL no adapt           & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
+    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3\\
+    \system{WDCMT}      & & & & & & & & \\
      \hline
   \end{tabular}
-  \caption{Translation performance of various MDMT systems. We report the BLEU score for each domain, as well as weighted and unweighted averages.}
+  \caption{Translation performance of various MDMT systems. We report the BLEU score for each domain, as well as domain-weighted (wAVG) and unweighted (AVG) averages.}
   \label{tab:performance}

### 2020-04-01 2020/coling2020/multidomain.tex
more text and numbers

@@
 Our main contribution is of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them.
-Our second contribution is experimental, and we provide a thorough reanalysis of most recent multi-domain approaches from the literature, including two variants of an original model.
+Our second contribution is experimental, and we perform a thorough reanalysis of most recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
 We show that\fyTodo{Spell out conclusions}.
 
-% From Kobus et al
-% Our goal is toallow a model built from a diverse set of trainingdata  to  produce  in-domain  translations.
-This is, to extend the coverage of generic NMT models to specific domains, with their specialized terminology and style, without lowering translation quality on more generic data.
-
 \section{Requirements of multi-domain MT \label{sec:requirements}}
-In this section, we try to clearly spell out the reasons for considering a multi-domain scenario, and discuss the implications in terms of performance evaluation.
+In this section, we try to recap the main reasons for considering a multi-domain scenario and discuss their implications in terms of performance evaluation.
 
@@
 We measure the distance between domains using the $\mathcal{H}$-Divergence introduced in \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains.
-We trained a domain discriminator is a multilayer perceptron trained as follows.
-Based on this, we see that ...
-\fyTodo{TBC}.
+Formally, given test sets of size $m$ for domains $A$ and $B$, and $h(x)$ a trained domain predictor, $\mathcal{H}(A,B)$ is computed as:
+$$
+\mathcal{H}(A,B) = 2(1 - [\frac{1}{m} \sum_{x:h(x) = B} \mathbb{I}( x \in A) + \frac{1}{m} \sum_{x: h(x) = A} \mathbb{I}(x \in B)]),
+$$
+where $\mathbb{I}$ is the indicator function.
+The  domain discriminator is a multilayer perceptron independently trained for each pair of domains with sentence representations derived via mean-pooling from the source side representation of our generic Transformer model.
+The hidden layer has dimension XXX, and the network is trained on balanced sets of YYY instances for ZZZ epochs.\fyTodo{TBC}.
 
@@
 \subsection{Multi-domain systems \label{ssec:systems}}
+\fyDone{Remove Gated residual}
 Our comparison of multidomain systems includes the following reimplementations of recent proposals from the literature:
@@
 \item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});
-\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDC}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\fyTodo{URLs}); the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
-\item A multidomain version of the approach of \newcite{Bapna19simple} denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; residual connexions make it possible to by-pass these modules.
-In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
-Note that contrarily to \cite{Bapna19simple} we do not start with a trained generic system, but learn the multi-domain from scratch.
-%\fyTodo{Check this.}
+\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\fyTodo{URLs}); the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
+\item An original, multi-domain version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; residual connexions make it possible to by-pass these modules.
+  % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
+Note that contrarily to \cite{Bapna19simple} we do not start with a trained generic system, but learn the multi-domain from scratch. \fyDone{Check this.}
 \end{itemize}
@@
 All (except \system{DM} and \system{ADM}) improve over the generic baseline (for most domains), matching requirement [P1-LAB], but yield weaker results than fine tuning.
-As expected
+As expected, the gap with fine-tuning remains large for an ``isolated'' domain such as \domain{rel}, the best contender being overall \system{MDL Res}, which approximately halves this gap as compared to \system{Mixed-Nat}.
 \fyTodo{More comments when we have all the results}
@@
 The effect of the adaptation layer is especially significant for the smaller domains (\domain{bank}, \domain{IT} and \domain{rel}) for which unplugging this layer causes dramatic performance drops.
+
+\fyTodo{Importance of sharing and unsharing}
 
@@
     \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8   & 42.8&53.8\\
-   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2    & 42.3&53.3\\
+   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2    & 42.3&53.3\\ \hline
 
@@
     \system{TTM}        & 37.7 & 54.8 & 49.7 & 33.4 & 43.8 & 80.0 &41.1 & 49.7\\
-    \system{LDC-Tag} &&&&& \\
+    \system{LDC-Tag} &&&&& &&&\\
     \system{DM}             & 36.5 & 52.4  & 47.7& 32.1 & 41.0 & 70.4 & 39.7 & 46.7 \\
     \system{ADM}           & 37.1 & 53.8  & 48.9 &32.8 & 42.3 & 78.5 & 40.7 & 48.9 \\
+    \system{WDCMT}      & & & & & & & & \\
     \system{MDL Res}     & 37.9 & 56.0  & 51.2   & 33.5   &  44.4  & 88.3 & 42.1 & 51.9\\
-    MDL no adapt           & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
-    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3\\
-    \system{WDCMT}      & & & & & & & & \\
+    \system{MDL Res} (gen)           & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
+%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3\\
      \hline
@@
 In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains.
-In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.
+\fyTodo{Fix this}In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.
 
@@
     Model / Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & Wrong \\ % \hline
-     Model / Set-up & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.75 / 0.25)}} & \multicolumn{2}{c|}{\domain{bank} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} & domain \\ \hline
+     Model / Set-up & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{bank} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} & domain \\ \hline
     & \domain{med-1} & \domain{med-2} & \domain{med-1} & \domain{med-2} &  \domain{bank-1} & \domain{bank-2} & \domain{bank} & \domain{law} & \\
-    \system{FT-Full}       & & & & & & & & & \\
-    \system{DC-Tag}      & & & & & & & & & \\
-    \system{DC-Feat}     & & & & & & & & & \\
+    \system{FT-Full}       & -0.13 & -0.6 & -1.51& -0.25& & & & & \\
+    \system{DC-Tag}      & -0.16 & -0.27& +0.13& +0.2& & & & & \\
+    \system{DC-Feat}     & -0.47& 0.0 & +0.27 $^+$ & +0.27 $^+$& & & & & \\
     \system{TTM}           & & & & & & & & &  \\
@@
 
-    \system{MDL Res}    & & & & & & & & & \\
-    \system{MDL Gated} & & & & & & & & & \\
+    \system{MDL Res}    & -0.2$^+$& -0.07$^+$&+0.18$^+$  &+0.0$^+$ & & & & & \\
+%    \system{MDL Gated} & & & & & & & & & \\
 
@@
   \caption{Translation performance with variable domain definitions.
-For the Split experiments, we only report the loss in performance for the \domain{med} test set, (*) denote significant differences, (+) denotes a significant improvement over \system{Full FT}; for the merge experiments for domains \domain{bank} and \domain{Law}.\fyTodo{How about wrong ?}}
+For the Split experiments, we only report the loss in performance for the \domain{med} test set, (*) denote significant differences, (+) denotes a significant improvement over \system{Full FT}; for the Merge experiments for domains \domain{bank} and \domain{Law} we report differences wrt.\ to the initial split for each domain.\fyTodo{How about wrong ?}}
   \label{tab:performance}
@@
 
+Our main findings can be summarized as follows: for the split experiments, we see a general loss with respect to the situation were domains are kept a whole, except in rare cases ().
+As expected, \system{FT} incurs a large loss, especially for the small split.
 Findings so far:
@@
 \end{todo}
+
+We compute an automatic clustering of the training sentence into $k=30$ classes using the k-means algorithm, and ??? as sentence representations. The resulting set of ``domains''
 

### 2020-04-07 2020/coling2020/multidomain.tex
Revert "close url of the first reference, add acknowledgement part"
This reverts commit df431916d0a81178da1b198f0e6c2dd33cfee7a6.

@@
 \usepackage{xcolor}
+\usepackage{longtable}
+\usepackage{tikz}
+\usetikzlibrary{calc}
 \usepackage[draft]{todo}
@@
 \begin{abstract}
-Abstract goes here
+Abstract goes here \fyTodo{Write abstract}
 \end{abstract}
 
-\section{Introduction}
-\label{intro}
+\section{Introduction} \label{sec:intro}
 
@@
 A second set of requirements is related to the definition of a domain.
-As repeatedly pointed out in the literature, training corpora for MT are often collected opportunistically and the view that each corpora constitutes a single domain is often a gross approximation.\footnote{Two of our own domains are composed of several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments.
+As repeatedly pointed out in the literature, training corpora for MT are often collected opportunistically and the view that each corpora constitutes a single domain is often a gross approximation.\footnote{Two of our own domains are actually composed of several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments.
 To challenge these requirements we propose the evaluate the following properties.
@@
 \paragraph{[P7-NUM]} As mentioned above, the computational argument (M1) is all the more sensible as the number of domains increases, making the optimization of many individual systems both ineffective and undesirable.
-For lack of having access to corpora containing a large (eg.\ in the order of 100-1000 domains), we experiment with automatically learned domains, considering a varying number of clusters.
+For lack of having access to corpora containing a large (eg.\ in the order of 100-1000 domains), we experiment with automatically learned domains.\fyTodo{considering a varying number of clusters.}
 
@@
 where $\mathbb{I}$ is the indicator function.
-The  domain discriminator is a multilayer perceptron independently trained for each pair of domains with sentence representations derived via mean-pooling from the source side representation of our generic Transformer model.
+The  domain discriminator is a multilayer perceptron independently trained for each pair of domains with sentence representations derived via mean pooling from the source side representation of our generic Transformer model.
 The hidden layer has dimension XXX, and the network is trained on balanced sets of YYY instances for ZZZ epochs.\fyTodo{TBC}.
@@
 
-Baselines correspond to standard practices for multi-domain systems.\footnote{We however omit domain-specific systems trained with a subset of the data, which are always inferior to the mix-domain strategy eg.\ \cite{Britz17mixing}.} Using the Transformer architecture of \shortcite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
+Baselines correspond to standard practices for multi-domain systems.\footnote{We however omit domain-specific systems trained with the corresponding subset of the data, which are always inferior to the mix-domain strategy eg.\ \cite{Britz17mixing}.} Using the Transformer architecture of \shortcite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
 
@@
 
-Our main findings can be summarized as follows: for the split experiments, we see a general loss with respect to the situation were domains are kept a whole, except in rare cases ().
+Our main findings can be summarized as follows: for the split experiments, we see a general loss with respect to the situation were domains are kept a whole, except in rare cases\fyTodo{To be completed}.
 As expected, \system{FT} incurs a large loss, especially for the small split.
@@
 
-\fyTodo{We need significance testing}
+\fyTodo{We do need significance testing}
 
-Another series of experiments aims to evaluate requirement [P6-DYN] as follows: starting with the existing MDMT systems of Section~\ref{ssec: ssec:rawperformance}, we introduce an additional domain (\domain{News}) and resume training with a mixture of data for an additional XXX epochs.
+Another series of experiments aims to evaluate the ability to handle additional domains (requirement [P6-DYN]) as follows: starting with the existing MDMT systems of Section~\ref{ssec: ssec:rawperformance}, we introduce an additional domain (\domain{News}) and resume training with a mixture of data for an additional XXX epochs.
 We contrast this approach with retraining all the systems from scratch, which would be the ideal situation, and report the difference in performance in Table~\ref{tab:XXX}.
-We expect that a good MDMT system should not be significantly affected by the addition of new data.
+We expect that a good MDMT system should not be significantly affected by the addition of new data, and reach the same performance as if the training included this domain from scratch.
+It should be noted that integrating new domains is straightforward for \system{DC-Tag}, \system{TTC} or \system{DC-Feat} where adding a new domain simply adds a new tag.
+It is less so for \system{LDC} \fyTodo{Complete effect of new domain}, for which the number of possible domains has built-in the architecture, and has to be anticipated from scratch.
+This makes a first difference between domain-bounded MT-systems, for which the number of possible domains is upper bounded; and the open-domain systems, which seamlessly integrate additional domains.
 
 \subsection{Automatic domains \label{ssec:autodomains}}
-\begin{todo}
-  Distance entre domaines: KL -->
-\end{todo}
+In this section, we experiment with automatic domains, obtained by clustering sentences into $k=30$ classes using the k-means algorithm based on generic sentence representations obtained via mean pooling as in section~\ref{ssec:corpora}.
+This allows us to evaluate requirement [P7-scale], training and testing our systems as if these domains were genuine.
+We also note that many of these classes are mere split of the large \domain{med}, while a few number of classes being mix of two existing domains.
+We are thus in a position to reiterate in more realistic conditions the measurements of section~\label{ssec:}, and to test whether multi-domain systems can effectively take advantage from the cross-domain similarities, and to eventually perform better that fine-tuning.
 
-We compute an automatic clustering of the training sentence into $k=30$ classes using the k-means algorithm, and ??? as sentence representations. The resulting set of ``domains''
+Results are in table~\ref{tab:subdomains}.\fyTodo{Fill the table}
 
@@
 These two sets of representations are then input to a conventional attentional sequence-to -sequence models in \cite{Zeng18multidomain}, and with Transformers in \cite{Su19exploring}; scores are reported separately per domain, suggesting that all domains are equally important at test time.
+% Adversarial adpatation is used in many domains
+% for QA @misc{lee2019domainagnostic,
+%    title={Domain-agnostic Question-Answering with Adversarial Training},
+%    author={Seanie Lee and Donggyu Kim and Jangwon Park},
+%    year={2019},
+%    eprint={1910.09342},
+%    archivePrefix={arXiv},
+%    primaryClass={cs.CL}
+%}, NLU and many others
 
@@
 \section{Conclusion and outlook \label{sec:conclusion}}
+\fyTodo{Write conclusions}
 
@@
 
+\appendix{}
+
+\begin{table}
+  \centering
+  \begin{tabular}{|p{3.2cm}|*{30}{r|}} \hline
+    Model / Set-up &d1&d2&d3&d4&d5&d6&d7&d8&d9&d10&d11&d12&d13&d14&d15
+    &d16&d17&d18&d19&d20&d21&d22&d23&d24&d25&d26&d27&d28&d29&d30 \\\hline
+    %
+    & \multicolumn{30}{|p{20cm}|}{%
+      }% \hline
+    \\
+    %
+    domain size ($\times 10^3$) &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\ \hline\hline
+    %
+   \system{FT-Full}      &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\  
+    \system{DC-Tag}    &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\     
+    \system{DC-Feat}   &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\ 
+    \system{TTM}         &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\ 
+    \system{LDC-Tag}  &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\
+
+    \system{MDL Res} 
+%    &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \hline
+  \end{tabular}
+  \caption{Detailed results of the experiments with automatic domains}
+  \label{tab:automatic_domains}
+\end{table}
+
+\todos{}
+
 \end{document}

### 2020-04-08 2020/coling2020/multidomain.tex
close url of the first reference, add acknowledgement part
more refs and text addendums

@@
 This is, for instance, the setting adopted \cite{Zeng18multidomain}, where automatic domain identification is performed for test sentences.
-Note that automatic domain identification can also be effectively performed in the first setting, especially in the presence of related domains in training.
+Note that automatic domain identification can also be effectively performed in the first setting, especially in the presence of close or related domains in training.
 
@@
 
+\cite[eq. (3)]{Wang20general} defines the loss of MDMT as the (macro) average of domain loss - all domains are equally important.
+
 \subsection{Evaluation issues \label{ssec:evaluation}}
@@
 Note that in this situation, the optimal training regime would be to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data.
+\cite{Wang20general} uses an intermediate mixing regime, which tends to reduce part of the unbalance, and improve the performance for underrepresented domains.
 \item the test distribution is neither equal to the training set, nor uniform.

### 2020-04-08 2020/coling2020/multidomain.tex
more todos, as a follow up of Talk with MinhQuang

@@
 Our main contribution is of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them.
-Our second contribution is experimental, and we perform a thorough reanalysis of most recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
+\fyTodo{New problems - continuous learning, automatic domains}Our second contribution is experimental, and we perform a thorough reanalysis of most recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
 We show that\fyTodo{Spell out conclusions}.
@@
 These various reasons for adopting MDMT are not mutually exclusive.
-Each however yields variable expectations with respect to the performance of this approach: if the motivation is primarily computational, then a drop in MT quality with respect to fine-tuning each domain might be acceptable; if it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some domains, if not all, over individually trained systems.
+Each however yields specific expectations with respect to the performance of this approach: if the motivation is primarily computational, then a drop in MT quality with respect to fine-tuning each domain might be acceptable; if it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some domains, if not all, over individually trained systems.
 
@@
 Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
-Some systems (see text below) additionally use an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
+The multi-domain residual system (see description below) additionally use an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
 The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
@@
 \fyDone{Remove Gated residual}
-Our comparison of multidomain systems includes the following reimplementations of recent proposals from the literature:
+Our comparison of multi-domain systems includes the following reimplementations of recent proposals from the literature:
 \begin{itemize}
@@
 \item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});
-\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\fyTodo{URLs}); the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
-\item An original, multi-domain version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; residual connexions make it possible to by-pass these modules.
+\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\fyTodo{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
+\item An original, multi-domain version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; within each layer, residual connexions make it possible to by-pass this module.
   % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
@@
 Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence, and slightly diverges from our notion of MDMT, which should be able to handle domain information in training and testing.
-This system has shown to meet many of our requirements, and would also have no problem scaling to a large number of domains.\fyTodo{TBC}
+This system has shown to meet many of our requirements and would also have no problem scaling to a large number of domains.\fyTodo{TBC}
 
@@
 \subsection{Performance of MDMT systems \label{ssec:rawperformance}}
-In this section, we discuss the raw performance of MDMT systems trained and tested on 6~domains.
+In this section, we discuss the raw performance of MDMT systems trained and tested on $6$~domains.
 Results are in Table~\ref{tab:performance}.
-As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2 BLEU points \jcDone{BP?} better, notably owing to the much better results for \domain{REL}).
+As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points \jcDone{BP?} better, notably owing to the much better results for \domain{REL}).
 As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains.
@@
 A complete fine-tuning, which updates all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to lack of tuning the capacity of the adapter module.
-\fyDone{Add 2 averages ?}
+\fyDone{Add 2 averages ?}\fyTodo{Significance testing wrt Mix Generic, Full Fine-tuned, for each domain}
 A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on the source is better than on the target, and very comparable to repeating the tag for every source word.
@@
 \end{table}
+\fyTodo{Fill the table with all results, including a comparison with *RNNs* for wcmd}
+\ftTodo{A new column for the News domain with predicted domains (except for the generic case).}
 
@@
 In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains.
-\fyTodo{Fix this}In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.
+\fyTodo{Prepare table for the wrong domain analysis (see also last col.
+of the first table)}In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.
 
@@
 
-\fyTodo{We do need significance testing}
+\fyTodo{We do need significance testing (a) each system against itself (non-split); (b) each system agains FT in split  condition}
 
@@
 
+\fyTodo{Instert table, significancy testing against full learning condition, for each domain}
+
 \subsection{Automatic domains \label{ssec:autodomains}}
@@
 
-Results are in table~\ref{tab:subdomains}.\fyTodo{Fill the table}
+Results are in table~\ref{tab:subdomains}.\fyTodo{Fill the tables with automatic and existing domains - we will have significancy results for the existing domains, for the other may be for larger classes we will have enough test instances?}
 
@@
 % \section*{Acknowledgements}
-
+% This work was granted access to the HPC resources of [TGCC/CINES/IDRIS] under the allocation 20XX- [numéro de dossier] made by GENCI (Grand Equipement National de Calcul Intensif)
 % The acknowledgements should go immediately before the references.  Do

### 2020-04-09 2020/coling2020/multidomain.tex
many supplementary bib entries, related work is almost fixed

@@
 In this setting, \emph{domain adaptation} (DA) methods are required.
-DA has a long history in Machine Learning in general \cite{Shimodaira00improving,Ben-David09atheory} and in NLP in particular \cite{Daume06domain,Blitzer07domain}, and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
+DA has a long history in Machine Learning in general \cite{Shimodaira00improving,Ben-David09atheory} and in NLP in particular \cite{Daume06domain,Blitzer07domain,Jiang07instance}, and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Chu18asurvey} for a recent survey of DA in the context of Neural MT).
 % In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
@@
 
-A related question concerns the relationship between domain-adaptation and multi-domain translation.
-The latter setting seems more challenging and one would expect that adapting a system toward a known domain $A$ will be better than training a MDMT system where only part of the data from $A$.
-Are there situations where the latter approach can surpass the former as is sometimes expected?
+A related question concerns the relationship between conventional domain adaptation and multi-domain translation.
+The latter setting seems more challenging and one would expect that adapting a system to a known target domain $A$ will be easier than training a MDMT system, where only part of the data from $A$.
+Are there situations where MD systems can surpass DA as is sometimes expected?
 
@@
 \begin{itemize}
-\item the more common situation is when both training and testing documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other documents in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
-It is usally assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}.
+\item the more common situation is when both train and test documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
+It is usually assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}.
 This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name: such as supervised MDMT ?
@@
 
-\subsection{Reasons for using MDMT \label{ssec:whymdmt}}
+\subsection{Reasons for using MDMT? \label{ssec:whymdmt}}
 The motivations for considering MDMT are many. Looking at the literature, the main arguments in support of MDMT seem to be the following:
@@
 Building mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}.
-Under this view, MDMT would be especially useful for domains with little training data: this is for instance what we observe for multilingual MT from English: an increase for under-resource languages due to positive transfer, at the cost of a decrease for well-resourced languages \cite{Arivazhagan19massively}.
+Under this view, MDMT would be especially useful for domains with little training data.
+This is what is observed for multilingual MT from English: an increase for under-resource languages due to positive transfer, at the cost of a decrease for well-resourced languages \cite{Arivazhagan19massively}.
 A slightly distinct line of reasoning is that mixing domains has a positive regularization effect for all domains.
@@
 Here again, if this distribution is known in advance, an optimal domain mix should be used in training to compute baseline scores, rather than using the empirical distribution of all the training data.
-A more challenging case is when the test set includes data from domains unseen in training: this situation, where test instances have a null probability in training, goes agains the basic statistical assumptions of DA \cite{Shimodaira00improving}.
+A more challenging case is when the test set includes data from domains unseen in training, or when it is chosen adversarially \cite{XXX}.
+% : this situation, where test instances have a null probability in training, goes agains the basic statistical assumptions of DA \cite{Shimodaira00improving}.
 It is thus important that the likeliness of such events should be clearly formulated when describing the evaluation protocol.
@@
 In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 most faithfully reflects the expected testing conditions.
-Another important issue is the choice of the evaluation metrics: if BLEU is still the standard metric used in most studies, it is not necessary the best choice to measure subtle improvements of lexical choices \cite{Irvine13measuring}.
+Another important issue is the choice of the evaluation metrics: if BLEU is still the standard metric used in most studies, it is not necessary the best choice to measure subtle improvements of lexical choices \cite{Irvine13measuring} that are often associated with adapted systems.
 
@@
 
-We experiment with translation from English into French and use data originating from 7~domains, corresponding to texts from the following data sources: the UFAL Medical corpus V1.0 (\domain{MED})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{Bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (LAW) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{IT}-Domain, Ted Talks (\domain{Talks}) \cite{Cettolo12wit}, and the Koran (\domain{Rel}).
+We experiment with translation from English into French and use data initially originating from 6~domains, corresponding to texts from the following data sources: the UFAL Medical corpus V1.0 (\domain{MED})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{Bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (LAW) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{IT}-Domain, Ted Talks (\domain{Talks}) \cite{Cettolo12wit}, and the Koran (\domain{Rel}).
 Complementary experiments also use v12 of the News Commentary corpus (\domain{News}).
@@
 We randomly select in each corpus a development and a test set of 1000 lines, and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score computed with\fyTodo{??}.
-Significance tests are performed as follows...
-\fyTodo{Fix this}.
+Validation sets are used to chose the best model according to the average BLEU score computed with\fyTodo{A word about meta-parameter settings}.
+Significance tests are performed using the bootstrap resampling method of \newcite{Koehn04statistical}.
+We report significant differences at the level of $p=$ \fyTodo{Fix correct p value}.
 
@@
     \hline
-    2609 (0.68) & 190 (0.05)  & 501 (0.13) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260  \\
+    2609 (0.68) & 190 (0.05)  & 501 (0.13) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
     \hline
   \end{tabular}
-\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mix.}
+\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mix (which excludes the \domain{News} domain).}
 \label{tab:Corpora}
@@
 The  domain discriminator is a multilayer perceptron independently trained for each pair of domains with sentence representations derived via mean pooling from the source side representation of our generic Transformer model.
-The hidden layer has dimension XXX, and the network is trained on balanced sets of YYY instances for ZZZ epochs.\fyTodo{TBC}.
+The hidden layer has dimension XXX, and the network is trained on balanced sets of YYY instances for ZZZ epochs.\fyTodo{Inform the classifier details}.
 
@@
 \item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\fyTodo{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
-\item An original, multi-domain version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; within each layer, residual connexions make it possible to by-pass this module.
+\item an original, multi-domain version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; within each layer, residual connexions make it possible to by-pass this module.
   % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
-Note that contrarily to \cite{Bapna19simple} we do not start with a trained generic system, but learn the multi-domain from scratch. \fyDone{Check this.}
+Note that contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch. \fyDone{Check this.}
 \end{itemize}
 
-Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence, and slightly diverges from our notion of MDMT, which should be able to handle domain information in training and testing.
+Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence, and slightly diverges from our notion of MDMT, which should be able to handle available domain information in training and testing.
 This system has shown to meet many of our requirements and would also have no problem scaling to a large number of domains.\fyTodo{TBC}
@@
   \fyTodo{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+  \begin{tabular}{|p{3cm}|*{8}{r|}|r|} \hline
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{med}} & \multicolumn{1}{c|}{\domain{law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talks}} & \multicolumn{1}{c|}{\domain{it}} & \multicolumn{1}{c|}{\domain{rel}} & \domain{wAVG} & \domain{AVG}\\ \hline
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1&49.4\\
-    \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3&51.4\\
-    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8   & 42.8&53.8\\
-   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2    & 42.3&53.3\\ \hline
+    Model / Domain & \multicolumn{1}{c|}{\domain{med}} & \multicolumn{1}{c|}{\domain{law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talks}} & \multicolumn{1}{c|}{\domain{it}} & \multicolumn{1}{c|}{\domain{rel}} & \multicolumn{1}{c|}{\domain{wAVG}} & \multicolumn{1}{c|}{\domain{AVG}} & \multicolumn{1}{|c|}{\domain{News}} \\ \hline
+    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 &\\
+    \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 &\\
+    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8   & 42.8 & 53.8 &\\
+   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2    & 42.3 & 53.3 &\\ \hline
 
@@
 
-    \system{DC-Tag}   & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5 &41.7 & 50.1\\
-    \system{DC-Feat}  & 37.3   & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 &41.4 & 49.9\\
-    \system{TTM}        & 37.7 & 54.8 & 49.7 & 33.4 & 43.8 & 80.0 &41.1 & 49.7\\
+    \system{DC-Tag}   & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5 &41.7 & 50.1     &\\
+    \system{DC-Feat}  & 37.3   & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 &41.4 & 49.9   & \\
+    \system{TTM}        & 37.7 & 54.8 & 49.7 & 33.4 & 43.8 & 80.0 &41.1 & 49.7       & \\
     \system{LDC-Tag} &&&&& &&&\\
-    \system{DM}             & 36.5 & 52.4  & 47.7& 32.1 & 41.0 & 70.4 & 39.7 & 46.7 \\
-    \system{ADM}           & 37.1 & 53.8  & 48.9 &32.8 & 42.3 & 78.5 & 40.7 & 48.9 \\
+    \system{DM}             & 36.5 & 52.4  & 47.7& 32.1 & 41.0 & 70.4 & 39.7 & 46.7 & \\
+    \system{ADM}           & 37.1 & 53.8  & 48.9 &32.8 & 42.3 & 78.5 & 40.7 & 48.9 & \\
     \system{WDCMT}      & & & & & & & & \\
-    \system{MDL Res}     & 37.9 & 56.0  & 51.2   & 33.5   &  44.4  & 88.3 & 42.1 & 51.9\\
-    \system{MDL Res} (gen)           & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
-%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3\\
+    \system{MDL Res}     & 37.9 & 56.0  & 51.2   & 33.5   &  44.4  & 88.3 & 42.1 & 51.9       &\\
+    \system{MDL Res} (gen)           & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7&\\
+%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3&\\
      \hline
   \end{tabular}
-  \caption{Translation performance of various MDMT systems. We report the BLEU score for each domain, as well as domain-weighted (wAVG) and unweighted (AVG) averages.}
+  \caption{Translation performance of various MDMT systems.
+We report the BLEU score for each domain, as well as domain-weighted (wAVG) and unweighted (AVG) averages.
+(*) denotes significant improvements with respect to the \system{Mix-Nat} system.
+Scores for the \domain{News} domain correspond to the out-of-domain scenario (see ~\ref{ssec:redomains}).
+}
   \label{tab:performance}
@@
   \caption{Translation performance with variable domain definitions.
-For the Split experiments, we only report the loss in performance for the \domain{med} test set, (*) denote significant differences, (+) denotes a significant improvement over \system{Full FT}; for the Merge experiments for domains \domain{bank} and \domain{Law} we report differences wrt.\ to the initial split for each domain.\fyTodo{How about wrong ?}}
-  \label{tab:performance}
+For the Split experiments, we only report the loss in performance for the \domain{med} test set, (*) denote significant losses when domains are split or merged; (+) denotes a significant improvement over \system{Full FT}; \fyTodo{How about wrong ?}}
+  \label{tab:redomains}
 \end{table}
@@
 
-Another series of experiments aims to evaluate the ability to handle additional domains (requirement [P6-DYN]) as follows: starting with the existing MDMT systems of Section~\ref{ssec: ssec:rawperformance}, we introduce an additional domain (\domain{News}) and resume training with a mixture of data for an additional XXX epochs.
-We contrast this approach with retraining all the systems from scratch, which would be the ideal situation, and report the difference in performance in Table~\ref{tab:XXX}.
+Another series of experiments evaluate the ability to handle additional domains (requirement [P6-DYN]) as follows.
+Starting with the existing MDMT systems of Section~\ref{ssec: ssec:rawperformance}, we introduce an additional domain (\domain{News}) and resume training with a mixture of data\footnote{The design of a proper mixture is again of utmost importance for achieving optimal performance: since our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy where half the batches are from the new domain, and half of the batches reproduce the initial training distribution.
+This is of course detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for an XXX epochs.\fyTodo{Training regime of continuation}.
+We contrast this approach with retraining all the systems from scratch, and report the difference in performance in Table~\ref{tab:XXX}.
 We expect that a good MDMT system should not be significantly affected by the addition of new data, and reach the same performance as if the training included this domain from scratch.
-It should be noted that integrating new domains is straightforward for \system{DC-Tag}, \system{TTC} or \system{DC-Feat} where adding a new domain simply adds a new tag.
-It is less so for \system{LDC} \fyTodo{Complete effect of new domain}, for which the number of possible domains has built-in the architecture, and has to be anticipated from scratch.
+It should be noted that dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{TTC} or \system{DC-Feat} where adding a new domain simply adds a new tag.
+It is less so for \system{LDC} \fyTodo{Complete effects of new domain}, for which the number of possible domains is built in the architecture and has to be anticipated from the begining.
 This makes a first difference between domain-bounded MT-systems, for which the number of possible domains is upper bounded; and the open-domain systems, which seamlessly integrate additional domains.
@@
 
-The authors of \cite{Kobus17domaincontrol} propose to inject additional domain feature in the model, either in the form of an extra (initial) domain-token, or in the form of additional domain-feature associated to each word.
-This approach is extended in \cite{Pham19generic} where source word embeddings are likely to vary across domains: this is achieved by spliting the embedding vector into a generic and a domain-specific part (one for each possible domain).
-Such feature-based approaches strongly relies on earlier proposals from the era of statistical MT \cite{Daume07frustratingly,Clark12onesystem}.
-
-In \cite{Britz17mixing}, several methods are considered which aim to ensure that domain information is actually used in a mix-domain system.
-Three alternatives are considered, using either domain classification (or domain normalization, via adversarial training) on source or target representations.
-There is no clear winner in either of their three language pairs considered, which only distinguish a limited number of domains.
-The same kinds of techniques are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the upper layers of the MT aimed to sort out domain specific information on the one hand, and domain-agnostic information on the other hand, using auxiliary tasks, hoping that the domain-agnostic part will be useful across the board.
-These two sets of representations are then input to a conventional attentional sequence-to -sequence models in \cite{Zeng18multidomain}, and with Transformers in \cite{Su19exploring}; scores are reported separately per domain, suggesting that all domains are equally important at test time.
+The multi-domain training regime is more the norm than the exception for NLP \cite{Dredze08online}.
+We focus here mostly on multi-domain learning in machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection and weighting, adversarial training, etc) have also been proposed for other language processing tasks.
+%; scores are reported separately per domain, suggesting that all domains are equally important at test time.
 % Adversarial adpatation is used in many domains
@@
 
-The motivations of \cite{Tars18multidomain} conjoins both (a) they argue that fine-tuning is brittle, owing to the reduced size of the tuning data, and can impair generalization; (b) they also argue that having one domain-agnostic system is a practical solution in many contexts
-\begin{quote}
-  In-domain fine-tuning has two main shortcomings: it depends on the availability of sufficientamounts of in-domain data in order to avoid over-fitting and it results in degraded performance for all other domains.
-The latter means that for translating multiple domains one has to run an individual NMT system for each domain.
-\end{quote}
-In their experiments, they consider that all domains are equally important and report individual scores for each known domain; further experiments are performed with automatically induced domain tags, which show that working with automatically induced tags (at the sentence level) is a viable alternative to using corpus labels as domain tags.
+The authors of \cite{Kobus17domaincontrol} propose to inject additional domain feature in the model, either in the form of an extra (initial) domain-token, or in the form of additional domain-feature associated to each word.
+This technique has also been used to control the style of MT outputs in \cite{Sennrich16politeness}, and to the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}.
+This approach is extended in \cite{Pham19generic} where source word embeddings are likely to vary across domains: this is achieved by spliting the embedding vector into a generic and a domain-specific part (one for each possible domain).
+Such feature-based approaches strongly rely on earlier proposals from the era of statistical MT \cite{Daume07frustratingly,Clark12onesystem}.
+Domain control can also be included in the target side, as in \cite{Chen16guided}, where a vector of topic describing the current document is included as an extra context in the softmax layer of the decoder.
 
-\cite{Farajian17multidomain} performs on-line adaptation at the sentence-level - processing a test sentence triggers the selection of a dedicated (small) set of fine-tuning instances assumed to be similar to the input text; using these, the generic NMT system is then tuned for some epochs, before delivering its translation.
-No clear notion of a domain, so it would pass all our tests, presumably.
-Also improves upon systems fine-tuned individually, sometimes by a wide margin.
-Test distribution is different from the training distribution (for domains).
+In \cite{Britz17mixing}, several methods are considered which aim to ensure that domain information is actually used in a mix-domain system.
+Three alternatives are considered, using either domain classification (or domain normalization, via adversarial training) on source or target representations.
+There is no clear winner in either of their three language pairs considered, which only distinguish a limited number of domains.
+The same basic techniques are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the upper layers of the MT aimed to sort out domain specific information on the one hand, and domain-agnostic information on the other hand, using auxiliary classification tasks, hoping that the domain-specific part will capture specific peculiarities of a domain, while the domain-agnostic part will be useful across the board.
+These two sets of representations are then input to a conventional attentional sequence-to -sequence models in \cite{Zeng18multidomain}, and with Transformers in \cite{Su19exploring}.
 
-A slightly more complex parameter-scheme is presented in \cite{Jiang19multidomain}, which augments a transformer model with domain specific heads, whose individual contributions are regulated at the word (position-level): some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads.
-The results for three language pair outperform the standard baselines for a 2-domain systems (in fr:en and ge:en) and a 4-domain system (zh:en)\footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
+A slightly more complex parameter-sharing scheme is presented in \cite{Jiang19multidomain}, which augments a transformer model with domain specific heads, whose individual contributions are regulated at the word (position-level): some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads.
+The results for three language pairs outperform several standard baselines for a 2-domain systems (in fr:en and ge:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
+
+The work of \cite{Tars18multidomain} makes a good case against fine-tuning, which is often found to be too brittle, a defect that MDMT can fix, delivering superior generalization accuracy.
+They perform experiment with reference domain tags, using mostly the domain-control approach.
+They also report results obtained with automatically induced domain tags, which show that working with automatically induced tags (at the sentence level) is a viable alternative to using corpus labels as domain tags, a result that we have confirmed in section~\ref{}.
+
+\newcite{Farajian17multidomain} also used automatic domains, performing on-line adaptation at the sentence-level.
+In their approach, the processing of a test sentence triggers the selection of a dedicated (small) set of fine-tuning instances assumed to be similar to the input text; using these, the generic NMT system is then tuned for some epochs, before delivering its translation.
+Different of most approaches studied here, this approach entirely dispenses with the notion of a domain, relying primarily on data selection/weighting techniques to handle data heterogeneity.
 

### 2020-04-09 2020/coling2020/multidomain.tex
more refs, notes on a theoretical part

@@
 It is usually assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}.
-This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name: such as supervised MDMT ?
+This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name here: such as supervised MDMT ?
 and mostly supervised MDMT for the second case ?}
@@
 
-We experiment with translation from English into French and use data initially originating from 6~domains, corresponding to texts from the following data sources: the UFAL Medical corpus V1.0 (\domain{MED})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{Bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (LAW) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{IT}-Domain, Ted Talks (\domain{Talks}) \cite{Cettolo12wit}, and the Koran (\domain{Rel}).
-Complementary experiments also use v12 of the News Commentary corpus (\domain{News}).
+We experiment with translation from English into French and use data initially originating from 6~domains, corresponding to texts from the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
+Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
 Corpus statistics in Table~\ref{tab:Corpora}.
@@
     %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
-    \domain{Med} & \domain{Law} & \domain{Bank} & \domain{IT} & \domain{Talk} & \domain{Rel} & \domain{News} \\
+    \domain{med} & \domain{law} & \domain{bank} & \domain{it} & \domain{talk} & \domain{rel} & \domain{news} \\
     \hline
@@
   \end{tabular}
-\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mix (which excludes the \domain{News} domain).}
+\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mix (which excludes the \domain{news} domain).}
 \label{tab:Corpora}
@@
 where $\mathbb{I}$ is the indicator function.
-The  domain discriminator is a multilayer perceptron independently trained for each pair of domains with sentence representations derived via mean pooling from the source side representation of our generic Transformer model.
+Our domain discriminator is a multilayer perceptron independently trained for each pair of domains with sentence representations derived via mean pooling from the source side representation of our generic Transformer model.
 The hidden layer has dimension XXX, and the network is trained on balanced sets of YYY instances for ZZZ epochs.\fyTodo{Inform the classifier details}.
@@
 Results are in Table~\ref{tab:performance}.
-As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points \jcDone{BP?} better, notably owing to the much better results for \domain{REL}).
+As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points \jcDone{BP?} better, notably owing to the much better results for \domain{rel}).
 As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains.
@@
 The use of an decidated adaptation component in each of the transformer layer proves an even better strategy and closes the gap with fine-tuning for half of the domains: the use of a gating mechanism yielding a clear improvement over the board.
-The effect of the adaptation layer is especially significant for the smaller domains (\domain{bank}, \domain{IT} and \domain{rel}) for which unplugging this layer causes dramatic performance drops.
+The effect of the adaptation layer is especially significant for the smaller domains (\domain{bank}, \domain{it} and \domain{rel}) for which unplugging this layer causes dramatic performance drops.
 
@@
 
-\begin{table}
+\begin{table*}
   \centering
-  \fyTodo{Fix column size}
+  \fyDone{Fix column size}
   \begin{tabular}{|p{3cm}|*{8}{r|}|r|} \hline
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{med}} & \multicolumn{1}{c|}{\domain{law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talks}} & \multicolumn{1}{c|}{\domain{it}} & \multicolumn{1}{c|}{\domain{rel}} & \multicolumn{1}{c|}{\domain{wAVG}} & \multicolumn{1}{c|}{\domain{AVG}} & \multicolumn{1}{|c|}{\domain{News}} \\ \hline
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} & \multicolumn{1}{c|}{\domain{news}} \\ \hline
     \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 &\\
@@
    \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2    & 42.3 & 53.3 &\\ \hline
-
  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-
     \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5 &41.7 & 50.1     &\\
@@
     \system{TTM}            & 37.7 & 54.8 & 49.7 & 33.4 & 43.8 & 80.0 &41.1 & 49.7       & \\
-    \system{LDC-Tag} &&&&& &&&\\
+    \system{LDC-Tag}     & 37   & 54.7 & 49.9 & 33.9 & 43.6 & 79.9 & & & \\ 
     \system{DM}             & 36.5 & 52.4  & 47.7& 32.1 & 41.0 & 70.4 & 39.7 & 46.7 & \\
     \system{ADM}           & 37.1 & 53.8  & 48.9 &32.8 & 42.3 & 78.5 & 40.7 & 48.9 & \\
-    \system{WDCMT}      & & & & & & & & \\
+    \system{WDCMT}       & & & & & & & & &\\
     \system{MDL Res}     & 37.9 & 56.0  & 51.2   & 33.5   &  44.4  & 88.3 & 42.1 & 51.9       &\\
@@
   \caption{Translation performance of various MDMT systems.
-We report the BLEU score for each domain, as well as domain-weighted (wAVG) and unweighted (AVG) averages.
+We report the BLEU score for each domain, as well as domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages.
 (*) denotes significant improvements with respect to the \system{Mix-Nat} system.
-Scores for the \domain{News} domain correspond to the out-of-domain scenario (see ~\ref{ssec:redomains}).
-}
+Scores for the \domain{news} domain correspond to the out-of-domain scenario (see section~\ref{ssec:performance}).}
   \label{tab:performance}
-\end{table}
+\end{table*}
 \fyTodo{Fill the table with all results, including a comparison with *RNNs* for wcmd}
-\ftTodo{A new column for the News domain with predicted domains (except for the generic case).}
+\fyTodo{A new column for the News domain with predicted domains (except for the generic case).}
 
@@
 
-\fyTodo{Instert table, significancy testing against full learning condition, for each domain}
+\begin{table*}
+  \begin{tabular}{|p{3cm}|*{9}{r|}} \hline
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} \\ \hline
+
+    \system{Mixed-Nat}  & 37.3  & 54.6   & 50.1   & 33.5  & 43.2  & 77.5  & & 41.1 & 49.4 \\
+    \system{FT-Full}       & 37.7  & 59.2   & 54.5   & 34.0  & 46.8  & 90.8  & & 42.8 & 53.8 \\ \hline\hline
+     \system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9 \\
+                                    & 36.9 & 53.7   & 49.1    & 33.7  &  42.8  & 74.1 & 28.9 &\\
+    \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5 \\
+                                   & 38.0  & 54.0   & 48.8    & 33.5  &  42.7  &73.2  & 28.3& \\
+    \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.1  & 29.1 \\
+                                   & 36.8  & 53.4   & 47.9    & 32.4  &  42.5  &72.8  & 28.4& \\
+    \system{LDC-Tag}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 \\
+    continual                 &&&&&& \\
+    \system{DM}            &36.0 &51.3&46.8&31.8&39.8&65.7&27.0 \\
+                                  &36.5&51.6&47.3&31.7&40.3&65.7&25.8& \\
+    \system{ADM}          &36.6&54.2&49.1&32.9&42.1&75.7&28.7 \\
+                                  &  36.9 &53.5&48.3&32.7&41.7&70.7&26.8& \\
+    \system{WDCMT}     &34.8  &51.8&47.8&30.7&37.3&49.3&25.9 \\
+    continual                 &&&&&&&& \\
+    \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & & \\
+     continual                &&&&&&&& \\
+     \hline
+  \end{tabular}
+  \caption{Ability to handle new a domain: BLEU performance in ``resume training'' mode compared to a complete training session.}
+  \label{tab:warmrestart}
+\end{table*}
+\fyTodo{Insert table, significancy testing against full learning condition, for each domain}
+
+Considerations, \fyTodo{To be completed}:
+
+\begin{itemize}
+\item even when trained from entirely scratch, very few systems are able to take advantage of the additional data for the 6 initial domains - this is observed on the averaged scores in Table~\ref{tab:warmrestart} with hardly surpass those in Table~\ref{tab:performance}
+\item restart is computationally cheaper (\fyTodo{how much cheaper here ?}) but underperforms cold start for all systems, sometimes by a wide margin.
+\item the loss is largest for the small domains - how about distance to \domain{news} ?
+\end{itemize}
+\fyTodo{Very close to fine tuning towards News, no ?}:
 
@@
 The same basic techniques are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the upper layers of the MT aimed to sort out domain specific information on the one hand, and domain-agnostic information on the other hand, using auxiliary classification tasks, hoping that the domain-specific part will capture specific peculiarities of a domain, while the domain-agnostic part will be useful across the board.
-These two sets of representations are then input to a conventional attentional sequence-to -sequence models in \cite{Zeng18multidomain}, and with Transformers in \cite{Su19exploring}.
+These techniques are implemented in an attentional sequence-to -sequence architecture in \cite{Zeng18multidomain}, and in Transformers in \cite{Su19exploring}.
 
@@
 Different of most approaches studied here, this approach entirely dispenses with the notion of a domain, relying primarily on data selection/weighting techniques to handle data heterogeneity.
+
+\subsection{Formalizing multi-domain adaptation (MDA)}
+
+In MDA, the typical training situation is to work with $k$ domains with $\mathcal{D}_k(x)$ the distribution of features $x \in \mathcal{X}$ for domain $k$.
+Given a labelling function $f \mathcal{X} \rightarrow{} \mathcal{Y}$, \cite{Mansour09domainadaptation} assumes $k$ hypotheses $h_k(x): \mathcal{X} \rightarrow{} \mathcal{Y}$ with an error smaller than $\epsilon$ when $x\sim \mathcal{D}_k$: $\sum_{x} L(h_k(x), f(y)) \mathcal{D}_k(x) < \epsilon_k$.
+The target distribution is a convex combination of $\{\mathcal{D}_k\}$: $\mathcal{D}_T(x) = \sum_k \lambda_k \mathcal{D}_k(x)$.
+\cite{Mansour09domainadaptation} derives generalization bounds for \textsl{combined hypotheses}, ie.\ hypotheses for which each prediction $h(x)$ is a linear combination of domain specific hypotheses $h(x_k)$.
+In this setting, one can study:
+\begin{itemize}
+\item generalization bounds when $\mathcal{D}_T(x)$ is fixed in advance, or worst case bounds valid for any mixture target distribution; the optimal solution has the form $h_{z,\eta}(x) = \sum_k \frac{z_kD_k(x) + U(x)/\gamma}{\sum_{k'} z_kD_k(x) + \eta U(x)/l}$ and is termed the \emph{distribution weighted combined rule}.
+\item diverging from this simple case, subsequent work in \cite{Mansour09multiple} generalizes this bound to the case where $\mathcal{D}_T$ is not a mixture of existing domains (in that case we will find a mixture that is close in the Renyì-divergence sense ); where $\mathcal{D}_k$ are unknown and estimated from the data; or when the labelling function is not the same across domains.
+More recent work study the case where the labelling function is stochastic, meaning that domains are defined by joint ($\mathcal{D}_k(x,y)$) rather than marginal distributions \cite{Hoffman18algorithms}, and provides an algorithm to compute the optimal weights for the combined rule based on estimates of the joint distributions.
+
+\item finite-sample bounds for the case where the target domain is known (\emph{multi-source single target adaptation} where we adapt from multiple sources studied in \cite{Cortes19AdaptationBO,Wen19domain}) - the issue is to optimize the selection / weighting of training data
+\item finite-sample bounds and algorithm for unsupervised MDA
+  % (http://papers.nips.cc/paper/8075-adversarial-multiple-source-domain-adaptation.pdf), test data are unlabeled;
+  This setting is also the focus of (Domain Adaptation from Multiple Sources via Auxiliary Classifiers)
+  %https://dl.acm.org/doi/pdf/10.1145/1553374.1553411 
+\end{itemize}
 

### 2020-04-10 2020/coling2020/multidomain.tex
more theoretical thoughts, will integrate in the first / second section

@@
 Data-based Machine Translation (MT), whether statistical or neural, rests on well understood machine learning principles.
-Given a training sample of paired source-target sentences $(\src,\trg)$\jcDone{shouldnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}$, a model parameterized by $\theta$ (here a translation function $f_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(f_\theta(\src, \trg))$.
-This approach ensures that the expectation of the translation loss will remain low when translating new (test) sentences \emph{drawn from the same distribution}.
+Given a training sample of matched source-target sentence pairs $(\src,\trg)$\jcDone{shoulnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}$, a model parameterized by $\theta$ (here a translation function $f_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(f_\theta(\src), \trg)$.
+This approach ensures that the expectation of the translation loss will remain low when translating new sentences \emph{drawn from the same distribution}.
 
@@
 In this setting, \emph{domain adaptation} (DA) methods are required.
-DA has a long history in Machine Learning in general \cite{Shimodaira00improving,Ben-David09atheory} and in NLP in particular \cite{Daume06domain,Blitzer07domain,Jiang07instance}, and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Chu18asurvey} for a recent survey of DA in the context of Neural MT).
+DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory}) and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}), and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for early proposals for statistical MT, or \cite{Chu18asurvey} for a recent survey of DA in the context of Neural MT).
 % In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
 
-A seemingly related problem is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where \emph{a single system} has to be trained  and tested with data from multiple domains.
+A seemingly related problem is \emph{multi-domain (MD) machine translation} \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where a single system has to be trained  and tested with data from multiple domains.
 As pointed out eg.\ by \newcite{Dredze08online}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask}.
-If the intuitions behind MD may seem simple, the exact specifications of MDMT systems are rarely spelled out, which makes their evaluation difficult: should they perform well when the test data is distributed like the train data or when it is equally distributed across training domains?
+If the intuitions behind MD may seem simple, the exact specifications of MDMT systems are rarely spelled out, which makes their evaluation difficult: for instance, should they perform well when the test data is distributed like the train data, when it is equally distributed across training domains or when the test distribution is unknown?
 should MDMT also be robust to new domains?
@@
 
-A related question concerns the relationship between conventional domain adaptation and multi-domain translation.
-The latter setting seems more challenging and one would expect that adapting a system to a known target domain $A$ will be easier than training a MDMT system, where only part of the data from $A$.
-Are there situations where MD systems can surpass DA as is sometimes expected?
+A related question concerns the relationship between domain adaptation and multi-domain translation.
+The latter setting seems more challenging and one would expect that adapting a system to a known target domain $A$ will be easier than training a MDMT system, where only part of the data belongs to domain $A$.
+Are there situations where MD systems can surpass DA, as is sometimes expected?
 
@@
 In this section, we try to recap the main reasons for considering a multi-domain scenario and discuss their implications in terms of performance evaluation.
+\fyTodo{Relation to theory, discussion of worst case scenario}
 
@@
 
-In MDA, the typical training situation is to work with $k$ domains with $\mathcal{D}_k(x)$ the distribution of features $x \in \mathcal{X}$ for domain $k$.
-Given a labelling function $f \mathcal{X} \rightarrow{} \mathcal{Y}$, \cite{Mansour09domainadaptation} assumes $k$ hypotheses $h_k(x): \mathcal{X} \rightarrow{} \mathcal{Y}$ with an error smaller than $\epsilon$ when $x\sim \mathcal{D}_k$: $\sum_{x} L(h_k(x), f(y)) \mathcal{D}_k(x) < \epsilon_k$.
-The target distribution is a convex combination of $\{\mathcal{D}_k\}$: $\mathcal{D}_T(x) = \sum_k \lambda_k \mathcal{D}_k(x)$.
-\cite{Mansour09domainadaptation} derives generalization bounds for \textsl{combined hypotheses}, ie.\ hypotheses for which each prediction $h(x)$ is a linear combination of domain specific hypotheses $h(x_k)$.
-In this setting, one can study:
-\begin{itemize}
-\item generalization bounds when $\mathcal{D}_T(x)$ is fixed in advance, or worst case bounds valid for any mixture target distribution; the optimal solution has the form $h_{z,\eta}(x) = \sum_k \frac{z_kD_k(x) + U(x)/\gamma}{\sum_{k'} z_kD_k(x) + \eta U(x)/l}$ and is termed the \emph{distribution weighted combined rule}.
-\item diverging from this simple case, subsequent work in \cite{Mansour09multiple} generalizes this bound to the case where $\mathcal{D}_T$ is not a mixture of existing domains (in that case we will find a mixture that is close in the Renyì-divergence sense ); where $\mathcal{D}_k$ are unknown and estimated from the data; or when the labelling function is not the same across domains.
+In MDA, the typical training situation (for a classifier) considers $k$ domains with $\mathcal{D}_k(x,y)$ the joint distribution of features $x \in \mathcal{X} and$ labels $y \in \mathcal{Y}$ for domain $k$.
+In \cite{Mansour09domainadaptation}, the labeling function $f$ is deterministic and the joint distribution is entirely defined by the marginal $\mathcal{D}_k(x)$ and $f()$.This work further assumes $k$ hypotheses $h_k(x): \mathcal{X} \rightarrow{} \mathcal{Y}$ with an error smaller than $\epsilon$ when $x\sim \mathcal{D}_k$: $\sum_{x} L(h_k(x), f(y)) \mathcal{D}_k(x) < \epsilon_k$ and that the target distribution is a convex combination of $\{\mathcal{D}_k\}$: $\mathcal{D}_T(x) = \sum_k \lambda_k \mathcal{D}_k(x)$.
+\cite{Mansour09domainadaptation} focuses on \textsl{combined hypotheses}, ie.\ hypotheses for which each prediction $h(x)$ is a linear combination of domain specific hypotheses $h(x_k)$, and derives generalization bounds for the case where $\mathcal{D}_T(x)$ is fixed in advance, or worst case bounds valid for any mixture target distribution.
+The optimal solution has the form $h_{z,\eta}(x) = \sum_k \frac{z_kD_k(x) + U(x)/\gamma}{\sum_{k'} z_kD_k(x) + \eta U(x)/l}$ and is termed the \emph{distribution weighted combined rule}.
+Further work in \cite{Mansour09multiple} generalizes these bounds to the case where $\mathcal{D}_T$ is not a mixture of existing domains (in that case we will find a mixture that is close to $\mathcal{}$in the Renyì-divergence sense ); where $\mathcal{D}_k$ are unknown and estimated from the data; or when the labelling function is not the same across domains.
 More recent work study the case where the labelling function is stochastic, meaning that domains are defined by joint ($\mathcal{D}_k(x,y)$) rather than marginal distributions \cite{Hoffman18algorithms}, and provides an algorithm to compute the optimal weights for the combined rule based on estimates of the joint distributions.
+A guiding principle is that this rule should tend towards uniform performance across domains, to ensure robustness in the worst case scenario \cite{Oren19distributionally}.
 
-\item finite-sample bounds for the case where the target domain is known (\emph{multi-source single target adaptation} where we adapt from multiple sources studied in \cite{Cortes19AdaptationBO,Wen19domain}) - the issue is to optimize the selection / weighting of training data
-\item finite-sample bounds and algorithm for unsupervised MDA
-  % (http://papers.nips.cc/paper/8075-adversarial-multiple-source-domain-adaptation.pdf), test data are unlabeled;
-  This setting is also the focus of (Domain Adaptation from Multiple Sources via Auxiliary Classifiers)
-  %https://dl.acm.org/doi/pdf/10.1145/1553374.1553411 
-\end{itemize}
+A related problem is \emph{Multiple source single domain adaptation} which is a variation of conventional DA from a mixture of source domains.
+In this setting it is advantageous to take advantage of the similarities between each source and the fixed target domain (see eg.\ \cite{Duan12domain,Cortes19AdaptationBO,Wen19domain} for studies of the supervised case, \cite{Zhao18adversarial} for the unsupervised case).
 

### 2020-04-10 2020/coling2020/multidomain.tex
clarify concepts, add references, start reshaping the first part

@@
 In this setting, \emph{domain adaptation} (DA) methods are required.
-DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory}) and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}), and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for early proposals for statistical MT, or \cite{Chu18asurvey} for a recent survey of DA in the context of Neural MT).
+DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory}) and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}), and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for early proposals of the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA in the context of Neural MT).
 % In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
 
-A seemingly related problem is \emph{multi-domain (MD) machine translation} \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where a single system has to be trained  and tested with data from multiple domains.
-As pointed out eg.\ by \newcite{Dredze08online}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask}.
-If the intuitions behind MD may seem simple, the exact specifications of MDMT systems are rarely spelled out, which makes their evaluation difficult: for instance, should they perform well when the test data is distributed like the train data, when it is equally distributed across training domains or when the test distribution is unknown?
-should MDMT also be robust to new domains?
+A seemingly related problem is \emph{multi-domain (MD) machine translation} \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system has to be trained and tested with data from multiple domains.
+As pointed out by \newcite{Dredze08online}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}; note that it also share commonalties with ensemble learning \cite{XXX}.
+If the intuitions behind MDMT are simple, the exact specifications of MDMT systems are rarely spelled out, which makes their evaluation fragile: for instance, should they perform well when the test data is distributed like the train data, when it is equally distributed across training domains or when the test distribution is unknown?
+Should MDMT also be robust to new domains?
 How should they handle domain labeling errors?
+etc.
 
@@
 \fyTodo{Relation to theory, discussion of worst case scenario}
+\fyTodo{Insist on the work of Dredze}
 
-\subsection{MDMT: a tentative definition}
+\subsection{Formalizing multi-domain learning \label{ssec:formalization}}
+
+We conventionally define a domain as a joint distribution $\mathcal{D}_d(x,y)$ over input-output pairs in $(x,y) \in \mathcal{X}\times{}\mathcal{Y}$.
+Assuming that the training data is sampled from a mixture of $D$ source domains $\mathcal{D}_s(x,y) = \sum_d \lambda_{s,d} \mathcal{D}_d(x,y)$, several cases are studied in the ML literature:
+
+\begin{itemize}
+\item the test distribution $\mathcal{D}_t$ is known 
+  \begin{itemize}
+  \item $\mathcal{D}_t = \mathcal{D}_{s}$ - this is standard ML, with mixture models
+  \item $\mathcal{D}_t(x,y) = \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) $ with $\lambda_s \neq \lambda_t$, a scenario deemed \emph{multiple source adaptation} (MSA) \cite{Mansour09domainadaptation}, with a different mixture of domains in training and testing
+  \item $\mathcal{D}_t(x,y) \neq \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) \forall \lambda$, a more complex case of MSA initially studied in \cite{Mansour09multiple}
+  \item variants of cases 3) and 4) only assume knowledge of the marginal distribution $\mathcal{D}_t(x) = \sum_{y} = \mathcal{D}_t(x) $
+  \item a slightly less favorable situation is when the test and train distributions are estimated from a finite sample, also studied in \cite{Mansour09multiple}.
+  \item we recover standard supervised DA is when there is only one active domain in the training and test mixtures, and multiple-source single domain adaptation when the source is a mixture but the test domain is homogeneous.
+Unsupervised DA \cite{Daume06domain} corresponds to the case where only the marginal distribution of the test domain is available, while semi-supervised DA is the reverse situation.
+  \end{itemize}
+\item the test distribution is not known and training needs to be distributionally robust, ie.\ minimize the error against all possible distributions, with variants depending on whether it is or not assumed to be a mixture of existing domains, and whether it is perfectly or imperfectly known.
+This setting is also studied in eg.\ \cite{Mansour09multiple,Hoffman18algorithms}.
+This adversarial setting is also considered for training language models in \cite{Oren19distributionally}.
+\item with respect to the labelling function $f$:
+  \begin{itemize}
+  \item $y=f(x)$ is deterministic, meaning that the joint distribution is entirely specified by the marginal and $f$; a more complex case is when $f$ also varies across domains;  
+  \item the labelling function $P(y|x)$ is stochastic, but shared across all domains, which is the typical covariate-shift scenario; it can also vary across domains \cite{???}
+  \end{itemize}
+\end{itemize}
+
+Multi-domain learning, as defined in \cite{Dredze08online} additionally makes the requirement that categorial domain tags are available in training and testing and that the learner should use them effectively \cite{Joshi12multidomain}.
+This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyTodo{explain zero shot}.
+
+\subsection{MDMT: a tentative definition \label{ssec:definition}}
 Informally, a multi-domain system should be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
@@
 By introducing variability in training, it prevents domain adaptation from overfitting the available adaptation data and could help improve generalization even for well-resourced domains.
+Following \cite{Joshi12multidomain}, one could even expect a benefit from ensembling several MT systems in decoding, irrespective of domain tags.
 
@@
 
-\paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} A first expectation for MDMT is that they perform better that the obvious baseline which disregards domain labels.
+\paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} A first expectation for MDMT is that they perform better that the obvious baseline which would disregards domain labels, or re-assign them in a random fashion \cite{Joshi12multidomain}.
 % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.

### 2020-04-15 2020/coling2020/multidomain.tex
review first sections

--- 2020-04-10 2020/coling2020/multidomain.tex
@@
 \usepackage[draft]{todo}
+\usepackage{xspace}
+\usepackage{amsmath,amsfonts,amssymb}
 %\setlength\titlebox{5cm}
@@
 \newcommand{\system}[1]{\texttt{\textbf{#1}}}
+\newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
 
@@
 Given a training sample of matched source-target sentence pairs $(\src,\trg)$\jcDone{shoulnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}$, a model parameterized by $\theta$ (here a translation function $f_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(f_\theta(\src), \trg)$.
-This approach ensures that the expectation of the translation loss will remain low when translating new sentences \emph{drawn from the same distribution}.
+This approach ensures that the expectation of the translation loss will remain low when translating new sentences drawn from the same distribution.
 
@@
 In this setting, \emph{domain adaptation} (DA) methods are required.
-DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory}) and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}), and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for early proposals for statistical MT, or \cite{Chu18asurvey} for a recent survey of DA in the context of Neural MT).
+DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory})\fyTodo{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}), and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA in the context of Neural MT).
 % In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
 
-A seemingly related problem is \emph{multi-domain (MD) machine translation} \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where a single system has to be trained  and tested with data from multiple domains.
-As pointed out eg.\ by \newcite{Dredze08online}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask}.
-If the intuitions behind MD may seem simple, the exact specifications of MDMT systems are rarely spelled out, which makes their evaluation difficult: for instance, should they perform well when the test data is distributed like the train data, when it is equally distributed across training domains or when the test distribution is unknown?
-should MDMT also be robust to new domains?
+A seemingly related problem, that has important practical implications, is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system has to be trained and tested with data from multiple domains.
+As pointed out both by \newcite{Dredze08online} and \cite{Finkel09hierarchical}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}; note that it also share commonalties with ensemble learning \cite{XXX}.\fyTodo{Missing ref} If the intuitions behind MDMT are simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should they perform well when the test data is distributed like the train data, when it is equally distributed across training domains or when the test distribution is unknown?
+Should MDMT also be robust to new domains?
 How should they handle domain labeling errors?
+etc.
+% , which makes their evaluation fragile
 
@@
 \fyTodo{Relation to theory, discussion of worst case scenario}
+\fyTodo{Insist on the work of Dredze}
 
-\subsection{MDMT: a tentative definition}
+\subsection{Formalizing multi-domain learning \label{ssec:formalization}}
+
+We define a domain $d$ as a joint distribution $\mathcal{D}_d(x,y)$ over input-output pairs $(x,y) \in \mathcal{X}\times{}\mathcal{Y}$.
+Assuming that the training data is sampled from a known mixture of $D$ source domains $\mathcal{D}_s(x,y) = \sum_d \lambda_{s,d} \mathcal{D}_d(x,y)$, several cases are studied in the ML literature:
+
+\begin{itemize}
+\item the test distribution $\mathcal{D}_t$ is known 
+  \begin{itemize}
+  \item $\mathcal{D}_t = \mathcal{D}_{s}$: this is standard ML, with mixture densities
+  \item $\mathcal{D}_t(x,y) = \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) $ with a different mixture of domains in training and testing ($\vlambda_s \neq \vlambda_t$), corresponding to multiple source adaptation (MSA) \cite{Mansour09domainadaptation},
+  \item $\mathcal{D}_t(x,y) \neq \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) \forall \lambda$, where the target domain for adapation is not a mixture of source domains, studied in \cite{Mansour09multiple}
+  \item variants of MSA which only assume knowledge of the marginal distribution $\mathcal{D}_t(x) = \sum_{y} \lambda_{t,d} \mathcal{D}_d(x) $
+  \item more realistic situations where  the train and test distributions are estimated from finite samples, also studied in \cite{Mansour09multiple}.
+  \end{itemize}
+   Note that we recover standard supervised DA is when there is only one component in the training and test mixtures, and multiple-source single domain adaptation when the source is a mixture but the test domain is homogeneous.
+Unsupervised DA \cite{Daume06domain} corresponds to the case where only the marginal distribution of the test domain inputs is available, while semi-supervised DA is the reverse situation.
+
+\item the test distribution is not known and training needs to be distributionally robust, ie.\ minimize the error against all possible distributions, with variants depending on whether it is or not assumed to be a mixture of existing domains, and whether it is perfectly or imperfectly known.
+This setting is also studied in eg.\ \cite{Mansour09multiple,Hoffman18algorithms}, and  in \cite{Oren19distributionally} for adversarial language model learning.
+\item with respect to the conditional label distribution $P(y|x)$:
+  \begin{itemize}
+  \item $P(y|x)$ is deterministic, meaning that the joint distribution is entirely specified by the marginal and a labelling function $y=h(x)$; a more complex case is when $h()$ also varies across domains;
+  \item $P(y|x)$ is stochastic, but shared across all domains, which is the typical covariate-shift scenario \cite{Shimodaira00improving}; it can also vary across domains \cite{???}\fyTodo{Missing citation vary accross domains}
+  \end{itemize}
+\end{itemize}
+
+Multi-domain learning, as defined in \cite{Dredze08online} additionally makes the requirement that categorial domain tags are available in training and testing; a multi-domain learner should use them effectively \cite{Joshi12multidomain} to improve simultaneously on all domains \cite{Finkel09hierarchical}.
+This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyTodo{explain zero shot}.
+
+\subsection{MDMT: a tentative definition \label{ssec:definition}}
 Informally, a multi-domain system should be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
@@
 \item[[M1\hspace{-0.5em}]] computational / practical arguments.
-A first motivation for MDMT is practical: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop only one system instead of having to optimize (and maintain) multiple MT engines \cite{Sennrich13multidomain,Farajian17neural}.
+A first motivation for MDMT is practical  \cite{Sennrich13multidomain,Farajian17neural}: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop only one system instead of having to optimize (and maintain) multiple MT engines.
 Such arguments are all the more convincing that the number of domains considered is large, which is not always the case: for instance \cite{Britz17mixing,Zeng18multidomain,Jiang19multidomain} only consider 2-4 domains in their experiments.
@@
 By introducing variability in training, it prevents domain adaptation from overfitting the available adaptation data and could help improve generalization even for well-resourced domains.
+A related case is made in \cite{Joshi12multidomain},
+  where the benefits of MD are attributed to an ensembling effect, where several MT systems are simultaneously used in decoding; this effect may subsist even in the absence of clear domain separations.
 
 \item[[M3\hspace{-0.5em}]] linguistic arguments.
-A related line of reasoning contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and remain semantically stable accross domain.
+A third line of reasoning contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and remain semantically stable accross domain.
 A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the representation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
-It is expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related and which can share more information.
+It is here expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related and which can share more information.
 \fyTodo{polysemy}
@@
 
-These various reasons for adopting MDMT are not mutually exclusive.
+\fyTodo{recap}
+These various motivations for adopting MDMT are not mutually exclusive.
 Each however yields specific expectations with respect to the performance of this approach: if the motivation is primarily computational, then a drop in MT quality with respect to fine-tuning each domain might be acceptable; if it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some domains, if not all, over individually trained systems.
 
-\cite[eq. (3)]{Wang20general} defines the loss of MDMT as the (macro) average of domain loss - all domains are equally important.
+% \cite[eq. (3)]{Wang20general} defines the loss of MDMT as the (macro) average of domain loss - all domains are equally important.
 
@@
 \section{Challenging multi-domain translation systems}
-In this section, we discuss the operational requirements that can be expected from a true multi-domain system, as well as ways to evaluate how much these requirements are met in practice.
+In this section, we discuss seven\fyTodo{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.
+% in practice.
 
@@
 % \subsubsection{Performance}
-The first set of requirements for MDMT systems is that they should perform well in the face of mix-domain data. These yields the following desired properties.
+The first set of requirements for MDMT systems is that they should perform well in the face of mix-domain data. These yields the following requirements.
 
-\paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} A first expectation for MDMT is that they perform better that the obvious baseline which disregards domain labels.
-% Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
+\paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} A first expectation for MDMT is that they perform better that the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}.\fyTodo{Try random label assignments ?} % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
 
@@
 To evaluate this, we perform two measurements, using a real as well as an artificial scenario.
-In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains (property [P2.1-TUN]).
-In the artificial scenario, we split a single training domain in two parts which are considered as distinct in training.
-The expectation here is that a MDMT should yield a clear gain in the both pseudo sub-domains, which should benefit from the supplementary amount of relevant training (this is property [P2.2-TUN]).
+In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains (requirement [P2.1-TUN]).
+In the artificial scenario, we split a single domain in two parts which are considered as distinct in training.
+The expectation here is that a MDMT should yield a clear gain for both pseudo sub-domains, which should benefit from the supplementary amount of relevant training (requirement [P2.2-TUN]).
 In this situation, MDMT should even outperform fine-tuning on either of the pseudo sub-domain.
-\jcDone{I dont understand the prediction... why running MDMT on 2 artificial subdomains should show  gains over fine-tunning?}\fyTodo{Clearer now ?}
+\jcDone{I dont understand the prediction... why running MDMT on 2 artificial subdomains should show  gains over fine-tunning?}\fyTodo{small loss with respect to non split ?}
 
@@
 As repeatedly pointed out in the literature, training corpora for MT are often collected opportunistically and the view that each corpora constitutes a single domain is often a gross approximation.\footnote{Two of our own domains are actually composed of several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments.
-To challenge these requirements we propose the evaluate the following properties.
+To challenge these requirements we propose the evaluate the following requirements.
 
 \paragraph{[P3-HET]}
-The notion of a domain is often a fragile one and an effective MDMT system should be able to discover not only when cross-domain sharing is most useful (cf.\ property [P2.2-TUN]) \jcDone{what is P2.2?}, but also when intra-domain heterogeneity is hurting.
-Requirement [P3-HET] is tested by artificially conjoining two separate domains in one during training.
+The notion of a domain is often a fragile one and an effective MDMT system should be able to discover not only when cross-domain sharing is useful (cf.\ property [P2.2-TUN])\jcDone{what is P2.2?}, but also when intra-domain heterogeneity is hurting.
+This requirement is tested by artificially conjoining two separate domains into one during training.
 The loss in performance with respect to the baseline (correct domain tags) should then be as small as possible.
@@
 This is in sharp contrast with what is commonly observed with fine-tuning, where giving the wrong domain tag can yield ``catastrophic'' drop in performance.
-This requirement is measured by decoding the test set for domain A with all other domain tags, and report average drop in performance.
+This requirement is measured by decoding the test set for domain $A$ with all other domain tags, and report average drop in performance.
 
@@
 Several alternatives are in order: (a) automatically assigning one of the domain tag for decoding should not be much worse than using the true domain tag, assuming this domain is seen in training.
-For test documents from other domains, the same procedure can apply, and should outperform the use of a single mix-domain system.
-\fyTodo{Systems react to unknown domains}These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.
+For test documents from other domains, the same procedure can apply, and should outperform the use of a single mix-domain system.\fyTodo{Systems react to unknown domains}These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.
 

### 2020-04-15 2020/coling2020/multidomain.tex
full rewrite of sec 2, additional results and comments in section 4

@@
 \newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
+\newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}
 
@@
 Data-based Machine Translation (MT), whether statistical or neural, rests on well understood machine learning principles.
-Given a training sample of matched source-target sentence pairs $(\src,\trg)$\jcDone{shoulnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}$, a model parameterized by $\theta$ (here a translation function $f_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(f_\theta(\src), \trg)$.
-This approach ensures that the expectation of the translation loss will remain low when translating new sentences drawn from the same distribution.
+Given a training sample of matched source-target sentence pairs $(\src,\trg)$\jcDone{shoulnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}_s$, a model parameterized by $\theta$ (here, a translation function $h_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(h_\theta(\src), \trg)$.
+This approach ensures that the translation loss remains low when translating more sentences drawn from the same distribution.
 
-Owing to the great variability of language data, this ideal situation is rarely met in practice, warranting the study of an alternative scenario, where the test distribution $\mathcal{D'}$ differs from $\mathcal{D}$.
-In this setting, \emph{domain adaptation} (DA) methods are required.
-DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory})\fyTodo{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}), and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D'}$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA in the context of Neural MT).
+Owing to the great variability of language data, this ideal situation is rarely met in practice, warranting the study of an alternative scenario, where the test distribution $\mathcal{D}_t$ differs from $\mathcal{D}_s$.
+In this setting, \emph{domain adaptation} (DA) methods are in order.
+DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory,Pan10asurvey})\fyDone{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}) and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D}_t$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
 % In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
 
-A seemingly related problem, that has important practical implications, is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system has to be trained and tested with data from multiple domains.
-As pointed out both by \newcite{Dredze08online} and \cite{Finkel09hierarchical}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}; note that it also share commonalties with ensemble learning \cite{XXX}.\fyTodo{Missing ref} If the intuitions behind MDMT are simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should they perform well when the test data is distributed like the train data, when it is equally distributed across training domains or when the test distribution is unknown?
+A seemingly related problem is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system has to be trained as well as tested with data from multiple domains.
+MD machine translation corresponds to a very common situation, where all available data, no matter its origin, is used to train a system that will be robust to domain variability in testing.
+% As pointed out both by \newcite{Dredze08online} and \cite{Finkel09hierarchical}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}.
+If the intuitions behind MDMT are quite simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should MDMT perform well when the test data is distributed like the training data, when it is equally distributed across domains or when the test distribution is unknown?
 Should MDMT also be robust to new domains?
-How should they handle domain labeling errors?
-etc.
+How should it handle domain labeling errors?
 % , which makes their evaluation fragile
@@
 
-In this paper, we try to answer these questions and to spell out in a more precise fashion the requirements that a MDMT system should meet.
-Our main contribution is of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them.
-\fyTodo{New problems - continuous learning, automatic domains}Our second contribution is experimental, and we perform a thorough reanalysis of most recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
+In this paper, we try to study these questions and to spell out in a more precise fashion the requirements that a MDMT system should meet.
+Our first contribution is of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them.
+In doing so, we also shed lights on new problems that arise in this context, regarding for instance the accomodation of new domains in the course of training, or the computation of automating domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental, and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
 We show that\fyTodo{Spell out conclusions}.
@@
 
-\subsection{Formalizing multi-domain learning \label{ssec:formalization}}
+\subsection{Complements - to be removed \label{ssec:oldformalization}}
 
@@
 
-\subsection{MDMT: a tentative definition \label{ssec:definition}}
-Informally, a multi-domain system should be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
-\begin{itemize}
-\item the more common situation is when both train and test documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
-It is usually assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}.
+% ----
+\subsection{Formalizing multi-domain translation \label{ssec:formalization}}
+
+We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for source sentences; each domain corresponds to a specific source of data, and differs from the other data sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
+Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
+$h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$, but can differ from one domain to the other.
+
+A typical learning scenario in MT is to have access to a samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$.
+Multi-domain learning, as defined in \cite{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ of several domains, making the problem distinct from mere domain adaption.
+A multi-domain learner is then expected to use these tags effectively \cite{Joshi12multidomain} when computing the combined translation function $h(x,d)$, and to perform well in all domains \cite{Finkel09hierarchical}.
+Note that this setting is closely related to the multi-source adaptation problem studied in \cite{Mansour09domainadaptation,Mansour09multiple}.
+
+This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is in \cite{Zeng18multidomain}, where test translations rely on a similarity score between sentences, rather than domain labels.}, and one that we also adopt here.
+Note that in the absence of further specification, the naive answer to the MD setting should be to train one distinct translation function $\hat{h}_d(x)$ for each domain, then translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$.
+We now discuss the arguments that are usually put forward to proceed differently.
+% --
+
+% This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyTodo{explain zero shot}.
+
+\subsection{Reasons for building MDMT systems \label{ssec:whymdmt}}
+
+A first motivation for moving away from the one-domain / one-system solution are practical  \cite{Sennrich13multidomain,Farajian17neural}: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop one single system instead of having to optimize and maintain multiple engines.
+The underlying assumption here is that the number of domains of interests can be large, a limiting scenario being fully personalized machine translation \cite{Michel2018extreme}.
+% which is not always the case: for instance \cite{Britz17mixing,Zeng18multidomain,Jiang19multidomain} only consider 2-4 domains in their experiments.\fyTodo{Keep this last part ?}
+
+A second line of reasoning rests on linguistic properties of the translation function and contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and tend remain semantically stable accross domains, motivating some cross-domain parameter sharing.
+A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the translation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
+It is here expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related and which can share more information.
+
+A third series of motivations are of statistical nature.
+The training data available for each domain is usually unevenly distributed, and domain-specific systems trained or adapted on small datasets are likely to have a high variance and perform poorly.
+Training mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}.
+Under this view, MDMT would be especially beneficial for domains with little training data.
+This is observed for multilingual MT from English: an increase for under-resource languages due to positive transfer, at the cost of a decrease in performance for well-resourced languages \cite{Arivazhagan19massively}.
+
+Combining multiple domain-specific MTs can also been justified in the sake of distributional robustness \cite{Mansour09domainadaptation,Mansour09multiple}, for instance when the test mixture differs from the train mixture, or when it includes new domains unseen in training.
+Such scenario is already well documented for zero-shot multi-lingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}, where mixing languages has more than demonstrated its usefulness.
+An even more challenging case is when the MT would need to perform well for any test distribution (as considered eg.\ in \cite{Hoffman18algorithms,Oren19distributionally}  for other tasks).
+In all these cases, mixing domains in training and/or testing is likely to improve robustness against unexpected inputs.
+
+A distinct line of reasoning is that mixing domains can a positive regularization effect for all domains.
+By introducing variability in training, it prevents domain adaptation from overfitting the available adaptation data and could help improve generalization even for well-resourced domains.
+A related case is made in \cite{Joshi12multidomain}, where some of the benefits of MD training are attributed to an ensembling effect, where several systems from multiple domains are simultaneously used in the prediction phase\cite{Saunders19ucam}; this effect may subsist even in the absence of clear domain separations.
+
+\fyDone{recap}
+To recap, there are multiple arguments for adopting MDMT, some already used in DA settings, and some are new.
+These arguments are not mutually exclusive; however each yields specific expectations with respect to the performance of this approach, and should also yield appropriate evaluation procedure.
+If the motivation is primarily computational, then a drop in MT quality with respect to multiple individual domain might be acceptable if compensated by the computational savings.
+If it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some under-resourced domains, over individually trained systems.
+If finally, it is to make the system more robust to unexpected or adversarial test distributions, then this is the setting that should be used to evaluate MDMT.
+The next section discusses ways in which these various requireents of MDMT systems can ben challenged.
+
+% \cite[eq. (3)]{Wang20general} defines the loss of MDMT as the (macro) average of domain loss - all domains are equally important.
+
+% \subsection{MDMT: a tentative definition \label{ssec:definition}}
+
+% Informally, a multi-domain system should be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
+% \begin{itemize}
+% \item the more common situation is when both train and test documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
+It is usually assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT.
 This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name here: such as supervised MDMT ?
@@
 
-\item It can also be the case that domain labels are only known in training, while test inputs may or may not be associated with explicit domains.
+% \item It can also be the case that domain labels are only known in training, while test inputs may or may not be associated with explicit domains.
 This is, for instance, the setting adopted \cite{Zeng18multidomain}, where automatic domain identification is performed for test sentences.
@@
 
-\item A last setting, that has not received much attention for MT (see however \cite{Sennrich13multidomain}), is where both train and test domains are automatically infered from data, using clustering techniques.
+% \item A last setting, that has not received much attention for MT (see however \cite{Sennrich13multidomain}), is where both train and test domains are automatically infered from data, using clustering techniques.
 Even though domain labels are typically available in most training scenario, these approaches assumes that might be preferable, owing to the multiple sources of intra-corpus and inter-corpus variabilities, to resort to automatic domain assignments.
 The online approach of \newcite{Farajian17multidomain}, where distributional cues, rather that categorical tags, are used for fine-tuning, is a variant of this strategy.
-\end{itemize}
+% \end{itemize}
 
-In the rest of the paper, we mostly adhere to the first definition which correspond to a situation of practical interest for the industry, and assume that multi-domain MT is mostly about effectively using domain tags in training and testing.
+% In the rest of the paper, we mostly adhere to the first definition which correspond to a situation of practical interest for the industry, and assume that multi-domain MT is mostly about effectively using domain tags in training and testing.
 We also experiment in Section~\ref{sec:results} with the other two alternative scenarios, as they also define useful contexts to evaluate the actual robustness of MDMT systems.
 
-\jcDone{...a very common scenario in MT industry where clients dispose of mulitiple data corresponding to different domains (differing in genre, thematic, register, style, etc.). }
+% \jcDone{...a very common scenario in MT industry where clients dispose of mulitiple data corresponding to different domains (differing in genre, thematic, register, style, etc.). }
 
-\subsection{Reasons for using MDMT? \label{ssec:whymdmt}}
-The motivations for considering MDMT are many. Looking at the literature, the main arguments in support of MDMT seem to be the following:
-\begin{enumerate}
-\item[[M1\hspace{-0.5em}]] computational / practical arguments.
-A first motivation for MDMT is practical  \cite{Sennrich13multidomain,Farajian17neural}: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop only one system instead of having to optimize (and maintain) multiple MT engines.
-Such arguments are all the more convincing that the number of domains considered is large, which is not always the case: for instance \cite{Britz17mixing,Zeng18multidomain,Jiang19multidomain} only consider 2-4 domains in their experiments.
+% \subsection{Evaluation issues \label{ssec:evaluation}}
 
-\item[[M2\hspace{-0.5em}]] statistical arguments.
-The training data available for each domain is very unevenly distributed, and domain trained or adapted on small datasets will yield large estimator variance.
-Building mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}.
-Under this view, MDMT would be especially useful for domains with little training data.
-This is what is observed for multilingual MT from English: an increase for under-resource languages due to positive transfer, at the cost of a decrease for well-resourced languages \cite{Arivazhagan19massively}.
-A slightly distinct line of reasoning is that mixing domains has a positive regularization effect for all domains.
-By introducing variability in training, it prevents domain adaptation from overfitting the available adaptation data and could help improve generalization even for well-resourced domains.
-A related case is made in \cite{Joshi12multidomain},
-  where the benefits of MD are attributed to an ensembling effect, where several MT systems are simultaneously used in decoding; this effect may subsist even in the absence of clear domain separations.
-
-\item[[M3\hspace{-0.5em}]] linguistic arguments.
-A third line of reasoning contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and remain semantically stable accross domain.
-A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the representation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
-It is here expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related and which can share more information.
-\fyTodo{polysemy}
-\end{enumerate}
-
-\fyTodo{recap}
-These various motivations for adopting MDMT are not mutually exclusive.
-Each however yields specific expectations with respect to the performance of this approach: if the motivation is primarily computational, then a drop in MT quality with respect to fine-tuning each domain might be acceptable; if it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some domains, if not all, over individually trained systems.
-
-% \cite[eq. (3)]{Wang20general} defines the loss of MDMT as the (macro) average of domain loss - all domains are equally important.
-
-\subsection{Evaluation issues \label{ssec:evaluation}}
-
-In theory, evaluation procedures should report realistic estimates of the test loss, computed on a sample deemed representative of the anticipated test conditions.
+% In theory, evaluation procedures should report realistic estimates of the test loss, computed on a sample deemed representative of the anticipated test conditions.
 Three scenarios need be distinguished:
-\begin{itemize}
-\item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
-\item all domains are equally likely and important in testing (the prior over test domains is uniform), which would warrant the computation of per domain scores and of unweighted averages, as reported in \cite{Farajian17multidomain,Su19exploring}.
+% \begin{itemize}
+% \item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
+% \item all domains are equally likely and important in testing (the prior over test domains is uniform), which would warrant the computation of per domain scores and of unweighted averages, as reported in \cite{Farajian17multidomain,Su19exploring,Wang20general}.
 Note that in this situation, the optimal training regime would be to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data.
 \cite{Wang20general} uses an intermediate mixing regime, which tends to reduce part of the unbalance, and improve the performance for underrepresented domains.
-\item the test distribution is neither equal to the training set, nor uniform.
+% \item the test distribution is neither equal to the training set, nor uniform.
 Here again, if this distribution is known in advance, an optimal domain mix should be used in training to compute baseline scores, rather than using the empirical distribution of all the training data.
 A more challenging case is when the test set includes data from domains unseen in training, or when it is chosen adversarially \cite{XXX}.
-% : this situation, where test instances have a null probability in training, goes agains the basic statistical assumptions of DA \cite{Shimodaira00improving}.
+% % : this situation, where test instances have a null probability in training, goes agains the basic statistical assumptions of DA \cite{Shimodaira00improving}.
 It is thus important that the likeliness of such events should be clearly formulated when describing the evaluation protocol.
-\end{itemize}
+% \end{itemize}
 
-The papers in the literature are seldom explicit regarding their assumptions regarding the test distribution.
+% The papers in the literature are seldom explicit regarding their assumptions regarding the test distribution.
 In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 most faithfully reflects the expected testing conditions.
@@
 
-A last issue regarding evaluation is related to the measurement of computational costs: when this is the main a motivation for using MDMT, we expect that computational costs would be reported and compared, which however is hardly ever done in practice.
+% A last issue regarding evaluation is related to the measurement of computational costs: when this is the main a motivation for using MDMT, we expect that computational costs would be reported and compared, which however is hardly ever done in practice.
 \fyTodo{Keep this ?}
 
-\section{Challenging multi-domain translation systems}
-In this section, we discuss seven\fyTodo{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.
-% in practice.
+\section{Challenging multi-domain translation systems \label{sec:challenging}}
+In this section, we discuss seven\fyTodo{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.\fyTodo{Insert discussion about scores}
 
@@
 % \subsubsection{Performance}
-The first set of requirements for MDMT systems is that they should perform well in the face of mix-domain data. These yields the following requirements.
+A first expectation is that MDMT systems should perform well in the face of mixed-domain test data. This yields the following three requirements.
 
-\paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} A first expectation for MDMT is that they perform better that the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}.\fyTodo{Try random label assignments ?} % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
+\paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} A MDMT should perform better than the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}.
+Evaluating this requirement is a matter of a mere comparison, assuming the test distribution of domains is known: if all domains are equally important, performance averages can be reported; if they are not, weighted averages should be used instead.
 
-\paragraph{[P2-TUN]} Additionally, following motivation~(M2) above, one would expect that MDMT improves over fine-tuning, at least in domains where data is scarce, or in situations where several domains are close.
+\fyTodo{Try random label assignments ?} % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
+
+\paragraph{[P2-TUN]} Additionally, one can expect that MDMT will improve over fine-tuning \cite{Luong15stanford,Freitag16fast}, at least in domains where data is scarce, or in situations where several domains are close.
 To evaluate this, we perform two measurements, using a real as well as an artificial scenario.
-In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains (requirement [P2.1-TUN]).
+In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains.
 In the artificial scenario, we split a single domain in two parts which are considered as distinct in training.
-The expectation here is that a MDMT should yield a clear gain for both pseudo sub-domains, which should benefit from the supplementary amount of relevant training (requirement [P2.2-TUN]).
+The expectation here is that a MDMT should yield a clear gain for both pseudo sub-domains, which should benefit from the supplementary amount of relevant training.
 In this situation, MDMT should even outperform fine-tuning on either of the pseudo sub-domain.
@@
 A second set of requirements is related to the definition of a domain.
-As repeatedly pointed out in the literature, training corpora for MT are often collected opportunistically and the view that each corpora constitutes a single domain is often a gross approximation.\footnote{Two of our own domains are actually composed of several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments.
+As repeatedly pointed out in the literature, parallel corpora used MT are often collected opportunistically and the view that each corpus constitutes a single domain is often a gross approximation.\footnote{Two of our own ``domains'' actually comprise several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments.
 To challenge these requirements we propose the evaluate the following requirements.
@@
 \paragraph{[P3-HET]}
-The notion of a domain is often a fragile one and an effective MDMT system should be able to discover not only when cross-domain sharing is useful (cf.\ property [P2.2-TUN])\jcDone{what is P2.2?}, but also when intra-domain heterogeneity is hurting.
+The notion of a domain being a fragile one, an effective MDMT system should be able to discover not only when cross-domain sharing is useful (cf.\ requirement [P2-TUN])\jcDone{what is P2.2?}, but also when intra-domain heterogeneity is hurting.
 This requirement is tested by artificially conjoining two separate domains into one during training.
-The loss in performance with respect to the baseline (correct domain tags) should then be as small as possible.
+The loss in performance with respect to the baseline (using correct domain tags) should then be as small as possible.
 
@@
 A related situation occurs when the domain of a test document is unknown.
-Several alternatives are in order: (a) automatically assigning one of the domain tag for decoding should not be much worse than using the true domain tag, assuming this domain is seen in training.
-For test documents from other domains, the same procedure can apply, and should outperform the use of a single mix-domain system.\fyTodo{Systems react to unknown domains}These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.
+Several situations need be considered: when the domain is seen in training, then using automatically predicted domain labels should not be much worse than using the correct one.
+For test documents from other domains, the same procedure can be used, and should outperform the use of a single mixed-domain system.\fyTodo{Systems react to unknown domains}
+% These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.
 
 \paragraph{[P6-DYN]}
-Another requirement, more of an operational nature, is that a MDMT system should smoothly evolve to accomodate a larger number of domains, without having to retrain the full system each time new data is available.
-This is a requirement [P6] that we challenge by dynamically changing the number of training and test domains.
+Another requirement, more of an operational nature, is that a MDMT system should smoothly evolve to accomodate a growing number of domains, without having to retrain the full system each time new data is available.
+This is a requirement [P6-DYN] that we challenge by dynamically changing the number of training and test domains.
 
@@
 
-\paragraph{[P7-NUM]} As mentioned above, the computational argument (M1) is all the more sensible as the number of domains increases, making the optimization of many individual systems both ineffective and undesirable.
-For lack of having access to corpora containing a large (eg.\ in the order of 100-1000 domains), we experiment with automatically learned domains.\fyTodo{considering a varying number of clusters.}
+\paragraph{[P7-NUM]} As mentioned above, MDMT systems have often been motivated by computational arguments.
+This argument is all the more sensible as the number of domains increases, making the optimization of many individual systems both ineffective and undesirable.
+For lack of having access to corpora containing very large sets (eg.\ in the order of 100-1000) domains, we experiment with automatically learned domains.\fyTodo{considering a varying number of clusters.}
 
@@
 
-We randomly select in each corpus a development and a test set of 1000 lines, and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score computed with\fyTodo{A word about meta-parameter settings}.
-Significance tests are performed using the bootstrap resampling method of \newcite{Koehn04statistical}.
-We report significant differences at the level of $p=$ \fyTodo{Fix correct p value}.
+We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
+Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\fyDone{A word about meta-parameter settings}, computed with\fyTodo{}.
+Significance tests are performed using the bootstrap resampling method of \newcite{Koehn04statistical} and implementation of compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+We report significant differences at the level of $p=0.05$\fyDone{Fix correct p value}.
 
@@
 where $\mathbb{I}$ is the indicator function.
-Our domain discriminator is a multilayer perceptron independently trained for each pair of domains with sentence representations derived via mean pooling from the source side representation of our generic Transformer model.
-The hidden layer has dimension XXX, and the network is trained on balanced sets of YYY instances for ZZZ epochs.\fyTodo{Inform the classifier details}.
+Our domain discriminator is SVM independently trained for each pair of domains with sentence representations derived via mean pooling from the source side representation of our generic Transformer model.
+We used the sci-kit learn implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well defined and separated from the all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
+
+\begin{table}\centering
+  \begin{tabular}{|l*{5}{|r}|} \hline
+    & law & bank & talk & IT & rel \\ \hline
+    med &1.93 &1.97 &1.9 &1.93 &1.97 \\
+    law   && 1.94 & 1.97 &1.93 & 1.99 \\
+    bank &&&1.98 &1.94 &1.99 \\
+    talk   &&&&1.92 &1.93 \\
+     IT     &&&&& 1.99 \\ \hline
+  \end{tabular}
+  \caption{The $\mathcal{H}$-divergence between domains}
+  \label{tab:domaindist}
+\end{table}
 
@@
 \item a generic model trained on a concatenation of all corpora (\texttt{Mixed}).
-We develop two versions of this system, one where domains unbalance reflects the distribution of our parallel data (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}) (cf.\ section~\ref{sec:evaluation}); \jcTodo{All systems cand be learned following both Bal and Nat.
+We develop two versions of this system, one where domain unbalance reflects the distribution of our parallel data (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
+As explained above, depending whether MDMT are expected to perform specially well on large domains, or to treat all domains equally, one or the other training distribution should be prefered.\jcTodo{All systems cand be learned following both Bal and Nat.
 does Nat corresponds to wAVG in test?
 is wAVG defined somewhere?}
-\item models fine-tuned \cite{Luong15stanford,Freitag16fast} \jcDone{i would use Luong and Manning 2015} separately on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
+\item models fine-tuned \cite{Luong15stanford,Freitag16fast}\jcDone{i would use Luong and Manning 2015} separately on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
 In our setting, we again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a simple adaptation module that is added (with residual connexions) on top of every Transformer layer (\system{FT-Block}).
@@
 \item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});
-\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\fyTodo{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
+\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
 \item an original, multi-domain version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; within each layer, residual connexions make it possible to by-pass this module.
@@
 Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence, and slightly diverges from our notion of MDMT, which should be able to handle available domain information in training and testing.
-This system has shown to meet many of our requirements and would also have no problem scaling to a large number of domains.\fyTodo{TBC}
+This system has shown to meet many of our requirements and would also have no problem scaling to a large number of domains.\fyDone{TBC}
 
 \section{Results and discussion \label{sec:results}}
+
 \subsection{Performance of MDMT systems \label{ssec:rawperformance}}
@@
 Results are in Table~\ref{tab:performance}.
-As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points \jcDone{BP?} better, notably owing to the much better results for \domain{rel}).
+As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points\jcDone{BP?} better, notably owing to the much better results for \domain{rel}).
 As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains.
-As all other systems are trained with an unbalanced data distribution, use the weighted average to perform global comparisons.
+As all other systems are trained with an unbalanced data distribution, we use the weighted average to perform global comparisons.
 
-Fine-tuning each domain separately provides us with better baselines, significantly outperforming \system{generic} for all domains.
-A complete fine-tuning, which updates all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to lack of tuning the capacity of the adapter module.
-\fyDone{Add 2 averages ?}\fyTodo{Significance testing wrt Mix Generic, Full Fine-tuned, for each domain}
+Fine-tuning each domain separately provides us with better baselines, outperforming \system{generic} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
+A complete fine-tuning, updating all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to keeping fixed the capacity of the adapter module.\fyDone{Add 2 averages ?}\fyDone{Significance testing wrt Mix Generic, Full Fine-tuned, for each domain}
+
 A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on the source is better than on the target, and very comparable to repeating the tag for every source word.
-All (except \system{DM} and \system{ADM}) improve over the generic baseline (for most domains), matching requirement [P1-LAB], but yield weaker results than fine tuning.
-As expected, the gap with fine-tuning remains large for an ``isolated'' domain such as \domain{rel}, the best contender being overall \system{MDL Res}, which approximately halves this gap as compared to \system{Mixed-Nat}.
+If all MDMTs (except \system{DM} and \system{ADM}) slightly improve over the generic baseline (for most domains), these improvements are rarely significant, the best contender overall being \system{MDL Res}, which significantly improves over \system{Mixed Nat} for three domains, and approximately halves the gap between the baseline and \system{FT-Full}, and can be considered to match [P1-LAB].
+
+All participating systems fail to outperform fine-tuning, failing to match [P2-TUN].
+This might be expected, even with all significantly outperform the baseline for \domain{rel}, the gap with fine-tuning remains large for an ``isolated'' domain such as \domain{rel},
 \fyTodo{More comments when we have all the results}
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} & \multicolumn{1}{c|}{\domain{news}} \\ \hline
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 &\\
+    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 & 23.5\\
     \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 &\\
-    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8   & 42.8 & 53.8 &\\
-   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2    & 42.3 & 53.3 &\\ \hline
+    \system{FT-Full}       & 37.7 & \bf 59.2 & \bf 54.5 & 34.0 & \bf 46.8 & \bf 90.8   & 42.8 & 53.8 &\\
+   \system{FT-Block}     & 37.3 & \bf 57.9 & 53.9 & 33.8 & \bf 46.7 & \bf 90.2    & 42.3 & 53.3 &\\ \hline
  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5 &41.7 & 50.1     &\\
-    \system{DC-Feat}      & 37.3  & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 &41.4 & 49.9   & \\
-    \system{TTM}            & 37.7 & 54.8 & 49.7 & 33.4 & 43.8 & 80.0 &41.1 & 49.7       & \\
-    \system{LDC-Tag}     & 37   & 54.7 & 49.9 & 33.9 & 43.6 & 79.9 & & & \\ 
-    \system{DM}             & 36.5 & 52.4  & 47.7& 32.1 & 41.0 & 70.4 & 39.7 & 46.7 & \\
-    \system{ADM}           & 37.1 & 53.8  & 48.9 &32.8 & 42.3 & 78.5 & 40.7 & 48.9 & \\
-    \system{WDCMT}       & & & & & & & & &\\
-    \system{MDL Res}     & 37.9 & 56.0  & 51.2   & 33.5   &  44.4  & 88.3 & 42.1 & 51.9       &\\
-    \system{MDL Res} (gen)           & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7&\\
+    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \bf 80.5 &41.7 & 50.1     & 21.8 \\
+    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \bf 79.9 &41.4 & 49.9   & \sl 21.7 \\
+    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \bf 79.9 &41.1 & 49.7       & \\
+    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \bf 79.9 & & 49.8 & 22.1\\ 
+    \system{DM}             & \sl 35.6 & \sl 49.5  & \sl 45.6& \sl 29.9 & \sl 37.1 & \sl 62.4 & 39.7 & 43.8 & \\
+    \system{ADM}           & 36.4 & \sl 53.5  & \sl 48.3 & \sl 32 & \sl 41.5 & \sl 73.4 & 40.7 & 47.5 & \\
+    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 & \\
+    \system{MDL Res}     & 37.9 & \bf 56.0  & \bf 51.2   & 33.5   &  44.4  & \bf 88.3 & 42.1 & 51.9  & \sl 21.2 \\
+    \hfill MDL Res (gen)           & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7&\\
 %    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3&\\
@@
 We report the BLEU score for each domain, as well as domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages.
-(*) denotes significant improvements with respect to the \system{Mix-Nat} system.
-Scores for the \domain{news} domain correspond to the out-of-domain scenario (see section~\ref{ssec:performance}).}
+Boldface denotes significant differences with respect to the \system{Mix-Nat} system.
+Scores for the \domain{news} domain correspond to the out-of-domain scenario (see Section~\ref{ssec:redomains}).}
   \label{tab:performance}
@@
 \fyTodo{Fill the table with all results, including a comparison with *RNNs* for wcmd}
-\fyTodo{A new column for the News domain with predicted domains (except for the generic case).}
+\fyDone{A new column for the News domain with predicted domains (except for the generic case).}
 
 \subsection{Redefining domains \label{ssec:redomains}}
-Table~\ref{fig:redomains} summaries the results of \fyTodo{XX} experiments where we artificially redefine the limits of domains.
+Table~\ref{tab:redomains} summaries the results of four experiments where we artificially redefine the limits of domains.
 In the three \textsl{split} experiments, we randomly split a domain corpus in two part: a good MDMT system should detect that these two ``domains'' should be mutually beneficial, and should be hardly affected by this change wrt.\ the baseline scenario.
@@
     % &&&&&& \\
-    Model / Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & Wrong \\ % \hline
-     Model / Set-up & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{bank} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} & domain \\ \hline
+    \hfill Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & Wrong \\ % \hline
+     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{bank} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} & domain \\ \hline
     & \domain{med-1} & \domain{med-2} & \domain{med-1} & \domain{med-2} &  \domain{bank-1} & \domain{bank-2} & \domain{bank} & \domain{law} & \\
@@
 Another series of experiments evaluate the ability to handle additional domains (requirement [P6-DYN]) as follows.
-Starting with the existing MDMT systems of Section~\ref{ssec: ssec:rawperformance}, we introduce an additional domain (\domain{News}) and resume training with a mixture of data\footnote{The design of a proper mixture is again of utmost importance for achieving optimal performance: since our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy where half the batches are from the new domain, and half of the batches reproduce the initial training distribution.
-This is of course detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for an XXX epochs.\fyTodo{Training regime of continuation}.
+Starting with the existing MDMT systems of Section~\ref{ssec:rawperformance}, we introduce an additional domain (\domain{News}) and resume training with a mixture of data\footnote{The design of a proper mixture is again of utmost importance for achieving optimal performance: since our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy where.
+This is of course detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 2 additional epochs.\fyTodo{Training regime of continuation}.
 We contrast this approach with retraining all the systems from scratch, and report the difference in performance in Table~\ref{tab:XXX}.
@@
 \end{itemize}
-\fyTodo{Very close to fine tuning towards News, no ?}:
+\fyDone{Very close to fine tuning towards News, no ?}
 
@@
 
-In MDA, the typical training situation (for a classifier) considers $k$ domains with $\mathcal{D}_k(x,y)$ the joint distribution of features $x \in \mathcal{X} and$ labels $y \in \mathcal{Y}$ for domain $k$.
+In MDA, the typical training situation (for a classifier) considers $k$ domains with $\mathcal{D}_k(x,y)$ the joint distribution of features $x \in \mathcal{X}$ and labels $y \in \mathcal{Y}$ for domain $k$.
 In \cite{Mansour09domainadaptation}, the labeling function $f$ is deterministic and the joint distribution is entirely defined by the marginal $\mathcal{D}_k(x)$ and $f()$.This work further assumes $k$ hypotheses $h_k(x): \mathcal{X} \rightarrow{} \mathcal{Y}$ with an error smaller than $\epsilon$ when $x\sim \mathcal{D}_k$: $\sum_{x} L(h_k(x), f(y)) \mathcal{D}_k(x) < \epsilon_k$ and that the target distribution is a convex combination of $\{\mathcal{D}_k\}$: $\mathcal{D}_T(x) = \sum_k \lambda_k \mathcal{D}_k(x)$.

### 2020-04-17 2020/coling2020/multidomain.tex
wrote more comments in the result section

@@
 \newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}
+\newcommand{\SB}[1]{\textcolor{green}{#1}}
+\newcommand{\SW}[1]{\textcolor{red}{#1}}
 
@@
 
-A seemingly related problem is multi-domain (MD) machine translation \cite{Hassan17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system has to be trained as well as tested with data from multiple domains.
+A seemingly related problem is multi-domain (MD) machine translation \cite{Sajjad17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system has to be trained as well as tested with data from multiple domains.
 MD machine translation corresponds to a very common situation, where all available data, no matter its origin, is used to train a system that will be robust to domain variability in testing.
@@
 
-This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is in \cite{Zeng18multidomain}, where test translations rely on a similarity score between sentences, rather than domain labels.}, and one that we also adopt here.
-Note that in the absence of further specification, the naive answer to the MD setting should be to train one distinct translation function $\hat{h}_d(x)$ for each domain, then translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$.
+This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is \cite{Farajian17multidomain}, where test translations rely on a similarity score between sentences, rather than domain labels.}, and one that we also adopt here.
+Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function.
 We now discuss the arguments that are usually put forward to proceed differently.
@@
 A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the translation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
-It is here expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related and which can share more information.
+It is here expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related and can share more information.
 
 A third series of motivations are of statistical nature.
-The training data available for each domain is usually unevenly distributed, and domain-specific systems trained or adapted on small datasets are likely to have a high variance and perform poorly.
+The training data available for each domain is usually unevenly distributed, and domain-specific systems trained or adapted on small datasets are likely to have a high variance and generalize poorly.
+For some test domains, there may even be no data at all \cite{Farajian17neural}.
 Training mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}.
@@
 
-Combining multiple domain-specific MTs can also been justified in the sake of distributional robustness \cite{Mansour09domainadaptation,Mansour09multiple}, for instance when the test mixture differs from the train mixture, or when it includes new domains unseen in training.
+Combining multiple domain-specific MTs can also be justified in the sake of distributional robustness \cite{Mansour09domainadaptation,Mansour09multiple}, for instance when the test mixture differs from the train mixture, or when it includes new domains unseen in training.
 Such scenario is already well documented for zero-shot multi-lingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}, where mixing languages has more than demonstrated its usefulness.
@@
 
-A distinct line of reasoning is that mixing domains can a positive regularization effect for all domains.
+A distinct line of reasoning is that mixing domains can have a positive regularization effect for all domains.
 By introducing variability in training, it prevents domain adaptation from overfitting the available adaptation data and could help improve generalization even for well-resourced domains.
@@
 Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools.
-To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encodings \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.%
+To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.%
 
@@
 Significance tests are performed using the bootstrap resampling method of \newcite{Koehn04statistical} and implementation of compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
-We report significant differences at the level of $p=0.05$\fyDone{Fix correct p value}.
+We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
 
@@
   \end{tabular}
-\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mix (which excludes the \domain{news} domain).}
+\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain).
+\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.}
 \label{tab:Corpora}
@@
 $$
-where $\mathbb{I}$ is the indicator function.
-Our domain discriminator is SVM independently trained for each pair of domains with sentence representations derived via mean pooling from the source side representation of our generic Transformer model.
-We used the sci-kit learn implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well defined and separated from the all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
+% where $\mathbb{I}$ is the indicator function.
+The domain discriminator is a SVM independently trained for each pair of domains, with sentence representations derived via mean pooling from the source side representation of the generic Transformer model.
+We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well defined and separated from the all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
 
@@
   \begin{tabular}{|l*{5}{|r}|} \hline
-    & law & bank & talk & IT & rel \\ \hline
-    med &1.93 &1.97 &1.9 &1.93 &1.97 \\
-    law   && 1.94 & 1.97 &1.93 & 1.99 \\
-    bank &&&1.98 &1.94 &1.99 \\
-    talk   &&&&1.92 &1.93 \\
-     IT     &&&&& 1.99 \\ \hline
+    & \domain{law} & \domain{bank} & \domain{talk} & \domain{IT} & \domain{rel} \\ \hline
+    \domain{med} &1.93 &1.97 &1.9 &1.93 &1.97 \\
+    \domain{law}   && 1.94 & 1.97 &1.93 & 1.99 \\
+    \domain{bank} &&&1.98 &1.94 &1.99 \\
+    \domain{talk}   &&&&1.92 &1.93 \\
+     \domain{IT}     &&&&& 1.99 \\ \hline
   \end{tabular}
@@
 
-Baselines correspond to standard practices for multi-domain systems.\footnote{We however omit domain-specific systems trained with the corresponding subset of the data, which are always inferior to the mix-domain strategy eg.\ \cite{Britz17mixing}.} Using the Transformer architecture of \shortcite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
+Baselines correspond to standard practices for multi-domain systems.\footnote{We however omit domain-specific systems trained with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using the Transformer architecture of \shortcite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
 
@@
 We develop two versions of this system, one where domain unbalance reflects the distribution of our parallel data (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
-As explained above, depending whether MDMT are expected to perform specially well on large domains, or to treat all domains equally, one or the other training distribution should be prefered.\jcTodo{All systems cand be learned following both Bal and Nat.
+The former is the best option when the train mixture $\mathcal{D}^s$ is expected to be also observed in testing; the latter should be used when the test distribution is uniform across domains.
+Accordingly, we report two aggrete scores: a weighted average reflecting the default distributio, and an unweighted average (all test domains are equally important).
+%   As explained above, depending whether MDMT are expected to perform specially well on large domains, or to treat all domains equally, one or the other training distribution should be prefered.\jcDone{All systems cand be learned following both Bal and Nat.
 does Nat corresponds to wAVG in test?
@@
 
-For all models, we set the embeddings size and the size of hidden layers to~512.
+For all models, we set the embeddings size and the hidden layers size to~512.
 Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
@@
  %
-Training uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$).
-In all cases, we use a batch size of~128 and a dropout rate of 0.1 for all layers.
+Training use a batch size of~128; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
  % All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
-\fyDone{Describe the block adaptation layer - voir slides}
-
 \subsection{Multi-domain systems \label{ssec:systems}}
@@
 \item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
-\item an original, multi-domain version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; within each layer, residual connexions make it possible to by-pass this module.
+\item an original, multi-domain version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connexions make it possible to by-pass this module.
   % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
-Note that contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch. \fyDone{Check this.}
+Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
 \end{itemize}
 
-Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence, and slightly diverges from our notion of MDMT, which should be able to handle available domain information in training and testing.
-This system has shown to meet many of our requirements and would also have no problem scaling to a large number of domains.\fyDone{TBC}
+Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges from our notion of MDMT, which should be able to handle available domain labels in training and testing.
+% This system has shown to meet several of requirements and would also have no problem scaling to a large number of domains.\fyDone{TBC}
 
@@
 \subsection{Performance of MDMT systems \label{ssec:rawperformance}}
-In this section, we discuss the raw performance of MDMT systems trained and tested on $6$~domains.
+In this section, we discuss the basic performance of MDMT systems trained and tested on $6$~domains.
 Results are in Table~\ref{tab:performance}.
@@
 
-Fine-tuning each domain separately provides us with better baselines, outperforming \system{generic} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
+Fine-tuning each domain separately provides us with better baselines, outperforming \system{Mixed-Nat} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
 A complete fine-tuning, updating all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to keeping fixed the capacity of the adapter module.\fyDone{Add 2 averages ?}\fyDone{Significance testing wrt Mix Generic, Full Fine-tuned, for each domain}
@@
 A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on the source is better than on the target, and very comparable to repeating the tag for every source word.
-If all MDMTs (except \system{DM} and \system{ADM}) slightly improve over the generic baseline (for most domains), these improvements are rarely significant, the best contender overall being \system{MDL Res}, which significantly improves over \system{Mixed Nat} for three domains, and approximately halves the gap between the baseline and \system{FT-Full}, and can be considered to match [P1-LAB].
+If all MDMTs (except \system{DM} and \system{ADM}) slightly improve over the generic baseline (for most domains), these improvements are rarely significant.
+The best contender overall is \system{MDL Res}, which significantly improves over \system{Mixed-Nat} for three domains, and is the only one to clearly fulfill [P1-LAB].
+The use of an decidated adaptation component proves here better than feature-based strategies.
+The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}), for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.
 
-All participating systems fail to outperform fine-tuning, failing to match [P2-TUN].
-This might be expected, even with all significantly outperform the baseline for \domain{rel}, the gap with fine-tuning remains large for an ``isolated'' domain such as \domain{rel},
-\fyTodo{More comments when we have all the results}
-The use of an decidated adaptation component in each of the transformer layer proves an even better strategy and closes the gap with fine-tuning for half of the domains: the use of a gating mechanism yielding a clear improvement over the board.
-The effect of the adaptation layer is especially significant for the smaller domains (\domain{bank}, \domain{it} and \domain{rel}) for which unplugging this layer causes dramatic performance drops.
-
-\fyTodo{Importance of sharing and unsharing}
+In this study, all systems fail to outperform fine-tuning, sometimes by a wide margin: this is observed for an ``isolated'' domain such as \domain{rel}.
+This might be expected, since all domains are clearly separated (cf.\ Section~\ref{ssec:corpora}) and are hardly helping each other.
+In this situation, MDMT systems should make sure that each domain is allocated a sufficient number of specific parameters, so as to close the gap with fine-tuning.
+\fyDone{More comments when we have all the results}.\fyDone{Importance of sharing and unsharing}
 
@@
     \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \bf 79.9 & & 49.8 & 22.1\\ 
-    \system{DM}             & \sl 35.6 & \sl 49.5  & \sl 45.6& \sl 29.9 & \sl 37.1 & \sl 62.4 & 39.7 & 43.8 & \\
-    \system{ADM}           & 36.4 & \sl 53.5  & \sl 48.3 & \sl 32 & \sl 41.5 & \sl 73.4 & 40.7 & 47.5 & \\
+    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 & \\
+    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 40.7 & 47.5 & \\
     \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 & \\
-    \system{MDL Res}     & 37.9 & \bf 56.0  & \bf 51.2   & 33.5   &  44.4  & \bf 88.3 & 42.1 & 51.9  & \sl 21.2 \\
+    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.1 & 51.9  & \SW{21.2} \\
     \hfill MDL Res (gen)           & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7&\\
@@
 \end{table*}
-\fyTodo{Fill the table with all results, including a comparison with *RNNs* for wcmd}
+\fyTodo{Fill the table with all results, including a comparison with *RNNs* for wcmd}\fyTodo{Do we have significancy tests for averages?}
 \fyDone{A new column for the News domain with predicted domains (except for the generic case).}
@@
 \subsection{Redefining domains \label{ssec:redomains}}
+
 Table~\ref{tab:redomains} summaries the results of four experiments where we artificially redefine the limits of domains.
@@
      Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{bank} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} & domain \\ \hline
-    & \domain{med-1} & \domain{med-2} & \domain{med-1} & \domain{med-2} &  \domain{bank-1} & \domain{bank-2} & \domain{bank} & \domain{law} & \\
+    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law} & \\
     \system{FT-Full}      & -0.13 & -0.6 & -1.51& -0.25& & & & & \\
@@
 \end{itemize}
+\fyDone{We do need significance testing (a) each system against itself (non-split); (b) each system agains FT in split  condition}
 
-\fyTodo{We do need significance testing (a) each system against itself (non-split); (b) each system agains FT in split  condition}
+\subsubsection{Handling unknown domains \label{sssec:unknowns}}
+To evaluate requirement [P5-UNK], we translate sentences from a domain unseen in training (\domain{news}).
+For each test sentence, we automatically predict the most likely domain tag, and use it for decoding.
+Results are in the rightmost column of Table~\ref{tab:performance}.
 
+\subsubsection{Coping with a growing number of domains}
 Another series of experiments evaluate the ability to handle additional domains (requirement [P6-DYN]) as follows.
-Starting with the existing MDMT systems of Section~\ref{ssec:rawperformance}, we introduce an additional domain (\domain{News}) and resume training with a mixture of data\footnote{The design of a proper mixture is again of utmost importance for achieving optimal performance: since our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy where.
-This is of course detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 2 additional epochs.\fyTodo{Training regime of continuation}.
-We contrast this approach with retraining all the systems from scratch, and report the difference in performance in Table~\ref{tab:XXX}.
-We expect that a good MDMT system should not be significantly affected by the addition of new data, and reach the same performance as if the training included this domain from scratch.
-It should be noted that dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{TTC} or \system{DC-Feat} where adding a new domain simply adds a new tag.
-It is less so for \system{LDC} \fyTodo{Complete effects of new domain}, for which the number of possible domains is built in the architecture and has to be anticipated from the begining.
-This makes a first difference between domain-bounded MT-systems, for which the number of possible domains is upper bounded; and the open-domain systems, which seamlessly integrate additional domains.
+Starting with the existing MDMT systems of Section~\ref{ssec:rawperformance}, we introduce an additional domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is again of utmost importance for achieving optimal performance: since our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy where.
+This is of course detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation}.
+We contrast this approach with retraining all the systems from scratch, and report the difference in performance in Table~\ref{tab:warmrestart}.
+We expect that a good MDMT system should not be significantly affected by the addition of a new domain, and reach about the same performance as if the training included this domain from scratch.
+Note that from a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{TTC} or \system{DC-Feat} where new domains merely add new labels.
+It is less simple for \system{LDC} and \system{MDL~Res}\fyTodo{Complete effects of new domain}, for which the number of possible domains is built in the architecture and has to be anticipated from the begining.
+This makes another difference between domain-bounded MT-systems, for which the number of possible domains is upper bounded and truely open-domain systems.% , which seamlessly integrate additional domains.
 
@@
   \begin{tabular}{|p{3cm}|*{9}{r|}} \hline
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} \\ \hline
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
 
-    \system{Mixed-Nat}  & 37.3  & 54.6   & 50.1   & 33.5  & 43.2  & 77.5  & & 41.1 & 49.4 \\
-    \system{FT-Full}       & 37.7  & 59.2   & 54.5   & 34.0  & 46.8  & 90.8  & & 42.8 & 53.8 \\ \hline\hline
+    \system{Mixed-Nat}  & 37.3  & 54.6   & 50.1   & 33.5  & 43.2  & 77.5  & 23.5 & 41.1 & 49.4 \\
+    \system{FT-Full}       & 37.7  & 59.2   & 54.5   & 34.0  & 46.8  & 90.8  & **** & 42.8 & 53.8 \\ \hline\hline
      \system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9 \\
@@
 \end{table*}
-\fyTodo{Insert table, significancy testing against full learning condition, for each domain}
+\fyTodo{Fill the table, significancy testing against full learning condition, for each domain}
 
-Considerations, \fyTodo{To be completed}:
+We can first compare the results of coldstart training with six (Table~\ref{tab:performance}) or seven domains (Table~\label{tab:warmrestart}): the most striking observation is that the extra training data is detrimental for all systems and all domains,\fyTodo{is this significant ?}.
+Increasing the heterogeneity of the mix hurts performance in all domains, again the sign that the domain are quite well separated.
+Comparing now the coldstart with the warmstart scenario, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training can negatively impact the performance on other domains - this happens notably for \system{DC-Tag} and \system{TTM}.
+In this setting, and, XXX\fyTodo{To be completed} and especially \system{MDL Res} show the smaller average loss, with the latter achieving the better balance of total training cost and average translation performance over all systems.
 
-\begin{itemize}
-\item even when trained from entirely scratch, very few systems are able to take advantage of the additional data for the 6 initial domains - this is observed on the averaged scores in Table~\ref{tab:warmrestart} with hardly surpass those in Table~\ref{tab:performance}
-\item restart is computationally cheaper (\fyTodo{how much cheaper here ?}) but underperforms cold start for all systems, sometimes by a wide margin.
-\item the loss is largest for the small domains - how about distance to \domain{news} ?
-\end{itemize}
-\fyDone{Very close to fine tuning towards News, no ?}
+% \begin{itemize}
+% \item even when trained from entirely scratch, very few systems are able to take advantage of the additional data for the 6 initial domains - this is observed on the averaged scores in Table~\ref{tab:warmrestart} with hardly surpass those in Table~\ref{tab:performance}
+% \item restart is computationally cheaper (\fyTodo{how much cheaper here ?}) but underperforms cold start for all systems, sometimes by a wide margin.
+% \item the loss is largest for the small domains - how about distance to \domain{news} ?
+% \end{itemize}
+% \fyDone{Very close to fine tuning towards News, no ?}
 

### 2020-04-17 2020/coling2020/multidomain.tex
start working on related work

@@
 
-We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for source sentences; each domain corresponds to a specific source of data, and differs from the other data sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
+We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for source sentences; each domain corresponds to a specific source of data, and differs from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
 Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
@@
 
-This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is \cite{Farajian17multidomain}, where test translations rely on a similarity score between sentences, rather than domain labels.}, and one that we also adopt here.
+This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is \cite{Farajian17multidomain}, where test translations rely on a similarity scores between test and train sentences, rather than on domain labels.}, and one that we also adopt here.
 Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function.
@@
 \section{Challenging multi-domain translation systems \label{sec:challenging}}
-In this section, we discuss seven\fyTodo{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.\fyTodo{Insert discussion about scores}
+In this section, we discuss seven\fyTodo{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.
+All these evaluations will rest on comparison of translation performance, and do not depend on the choice a particular metrics.
+To make our results comparable with the literature, we will only use the BLEU score \cite{Papineni02bleu} in Section~\ref{sec:experiments}, noting it may not be the best choice to measure subtle improvements of lexical choices that are often associated with domain adapted systems \cite{Irvine13measuring}.
+Other important figure of merits for MDMT systems are the computational cost and the total number of parameters.\fyDone{Insert discussion about scores}
 
@@
 In this situation, MDMT systems should make sure that each domain is allocated a sufficient number of specific parameters, so as to close the gap with fine-tuning.
-\fyDone{More comments when we have all the results}.\fyDone{Importance of sharing and unsharing}
+\fyDone{More comments when we have all the results}\fyDone{Importance of sharing and unsharing}
 
@@
     \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 &\\
-    \system{FT-Full}       & 37.7 & \bf 59.2 & \bf 54.5 & 34.0 & \bf 46.8 & \bf 90.8   & 42.8 & 53.8 &\\
-   \system{FT-Block}     & 37.3 & \bf 57.9 & 53.9 & 33.8 & \bf 46.7 & \bf 90.2    & 42.3 & 53.3 &\\ \hline
+    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8 &\\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & 42.3 & 53.3 &\\ \hline
  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \bf 80.5 &41.7 & 50.1     & 21.8 \\
-    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \bf 79.9 &41.4 & 49.9   & \sl 21.7 \\
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \bf 79.9 &41.1 & 49.7       & \\
-    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \bf 79.9 & & 49.8 & 22.1\\ 
+    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.7 & 50.1     & 21.8 \\
+    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   & \SW{21.7} \\
+    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7       & \\
+    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8 & 22.1\\ 
     \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 & \\
@@
 
-Table~\ref{tab:redomains} summaries the results of four experiments where we artificially redefine the limits of domains.
-In the three \textsl{split} experiments, we randomly split a domain corpus in two part: a good MDMT system should detect that these two ``domains'' should be mutually beneficial, and should be hardly affected by this change wrt.\ the baseline scenario.
-This is the situation where we should see clear improvements even over finetuning, which only exploits part of the available relevant data.
+Table~\ref{tab:redomains} summaries the results of four experiments where we artificially redefine the limits of domains, with the aim to challenge requirements [P2-TUN], [P3-HET], and [P4-ERR].
+In the three \emph{split} experiments, we randomly split a domain corpus in two part: a good MDMT system should detect that these two ``domains'' should be mutually beneficial, and should be hardly affected by this change wrt.\ the baseline scenario.
+This is the situation where we should expect to MDMT even surpass fine-tuning separately on each of these dummy domain, as MDMT exploits all data, where fine-tuning only sees a subpart.
+
 In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains.
@@
     & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law} & \\
-    \system{FT-Full}      & -0.13 & -0.6 & -1.51& -0.25& & & & & \\
-    \system{DC-Tag}     & -0.16 & -0.27& +0.13& +0.2& & & & & \\
-    \system{DC-Feat}    & -0.47& 0.0 & +0.27 $^+$ & +0.27 $^+$& & & & & \\
-    \system{TTM}          & & & & & & & & &  \\
-    \system{LDC-Tag}   & & & & & & & & & \\
-
-    \system{MDL Res}   & -0.2$^+$& -0.07$^+$&+0.18$^+$  &+0.0$^+$ & & & & & \\
+    \system{FT-Full}      & -0.13 & -0.6 & -1.51& -0.25& & &\SW{-1.6} & \SW{-1.4}& \\
+    \system{DC-Tag}     & -0.16 & -0.27& +0.13& +0.2& & & -0.5 & -0.4 & \\
+    \system{DC-Feat}    & -0.47& 0.0 & \SB{0.27} $^+$ & +0.27 $^+$& & & 0.3 & 0.1 & \\
+    \system{TTM}          & & & & & & &  0.0 & -0.3 &  \\
+    \system{LDC-Tag}   & & & & & & &  0.0 & 0.1& \\ 
+    \system{MDL Res}   & -0.2$^+$& -0.07$^+$&+0.18$^+$  &+0.0$^+$ & & & \SW{-2.4} & \SW{-3.5} & \\
 %    \system{MDL Gated} & & & & & & & & &  \\
-
-    \system{ADM} & & & & & & & & & \\
-    \system{DM}    & & & & & & & & & \\
+    \system{ADM} & & & & & & &  0.1 & -0.4 & \\
+    \system{DM}    & & & & & & & 0.9 & 0.1 & \\
     \system{WDC}  & & & & & & & & & \\
@@
 
+It is finally interesting to note that all systems, but \system{MDL Res}, are quite robust to domain merge: we see even small gains in performance (eg.\ for \system{DM}).
+While robustness is generally a good thing, it is also a sign of a poor overall usage of domain labels.
+\system{MDL Res}\fyTodo{To be completed}
+
 \subsubsection{Handling unknown domains \label{sssec:unknowns}}
@@
 For each test sentence, we automatically predict the most likely domain tag, and use it for decoding.
-Results are in the rightmost column of Table~\ref{tab:performance}.
+Results are in the rightmost column of Table~\ref{tab:performance}.\fyTodo{To be completed}
 
@@
 
-The multi-domain training regime is more the norm than the exception for NLP \cite{Dredze08online}.
-We focus here mostly on multi-domain learning in machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection and weighting, adversarial training, etc) have also been proposed for other language processing tasks.
+The multi-domain training regime is more the norm than the exception for NLP \cite{Dredze08online,Finkel09hierarchical}.
+We focus here mostly on multi-domain learning in machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection and weighting, adversarial training, etc) have also been proposed for most language processing tasks.
 %; scores are reported separately per domain, suggesting that all domains are equally important at test time.
@@
 
-The authors of \cite{Kobus17domaincontrol} propose to inject additional domain feature in the model, either in the form of an extra (initial) domain-token, or in the form of additional domain-feature associated to each word.
-This technique has also been used to control the style of MT outputs in \cite{Sennrich16politeness}, and to the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}.
-This approach is extended in \cite{Pham19generic} where source word embeddings are likely to vary across domains: this is achieved by spliting the embedding vector into a generic and a domain-specific part (one for each possible domain).
+Multi-domain translation was introduced for statistical MT (eg.\ \cite{Banerjee10combining,Sennrich13multidomain}): as for domain adaptation, two main strategies where considered consisting at weighting train and test instances based on their domain-relatedness on the one hand; or to inject domain labels as additional features in learning.
+
+This latter strategy has also been widely used in MNT: the authors of \cite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token, or in the form of additional domain-feature associated to each word.
+Note his technique has also been used to control the style of MT outputs in \cite{Sennrich16politeness,Niu18multitask}, and to encode the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}.
+This approach is extended in \cite{Pham19generic} where source word embeddings are likely to vary across domains: this is achieved by splitting the embedding vector into a generic and a domain-specific part (one for each possible domain); a similar approach, applied to hidden states, is documented in \cite{Chu18multilingual}.
 Such feature-based approaches strongly rely on earlier proposals from the era of statistical MT \cite{Daume07frustratingly,Clark12onesystem}.
-Domain control can also be included in the target side, as in \cite{Chen16guided}, where a vector of topic describing the current document is included as an extra context in the softmax layer of the decoder.
+Domain control can also be performed on the target side, as in \cite{Chen16guided}, where a vector of topic describing the current document is included as an extra context in the softmax layer of the decoder.
 

### 2020-04-17 2020/coling2020/multidomain.tex
review related work - we still miss a couple of results, the conclusions and abstract

@@
 In this section, we try to recap the main reasons for considering a multi-domain scenario and discuss their implications in terms of performance evaluation.
-\fyTodo{Relation to theory, discussion of worst case scenario}
-\fyTodo{Insist on the work of Dredze}
+\fyDone{Relation to theory, discussion of worst case scenario}
+\fyDone{Insist on the work of Dredze}
 
-\subsection{Complements - to be removed \label{ssec:oldformalization}}
-
-We define a domain $d$ as a joint distribution $\mathcal{D}_d(x,y)$ over input-output pairs $(x,y) \in \mathcal{X}\times{}\mathcal{Y}$.
-Assuming that the training data is sampled from a known mixture of $D$ source domains $\mathcal{D}_s(x,y) = \sum_d \lambda_{s,d} \mathcal{D}_d(x,y)$, several cases are studied in the ML literature:
-
-\begin{itemize}
-\item the test distribution $\mathcal{D}_t$ is known 
-  \begin{itemize}
-  \item $\mathcal{D}_t = \mathcal{D}_{s}$: this is standard ML, with mixture densities
-  \item $\mathcal{D}_t(x,y) = \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) $ with a different mixture of domains in training and testing ($\vlambda_s \neq \vlambda_t$), corresponding to multiple source adaptation (MSA) \cite{Mansour09domainadaptation},
-  \item $\mathcal{D}_t(x,y) \neq \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) \forall \lambda$, where the target domain for adapation is not a mixture of source domains, studied in \cite{Mansour09multiple}
-  \item variants of MSA which only assume knowledge of the marginal distribution $\mathcal{D}_t(x) = \sum_{y} \lambda_{t,d} \mathcal{D}_d(x) $
-  \item more realistic situations where  the train and test distributions are estimated from finite samples, also studied in \cite{Mansour09multiple}.
-  \end{itemize}
-   Note that we recover standard supervised DA is when there is only one component in the training and test mixtures, and multiple-source single domain adaptation when the source is a mixture but the test domain is homogeneous.
-Unsupervised DA \cite{Daume06domain} corresponds to the case where only the marginal distribution of the test domain inputs is available, while semi-supervised DA is the reverse situation.
-
-\item the test distribution is not known and training needs to be distributionally robust, ie.\ minimize the error against all possible distributions, with variants depending on whether it is or not assumed to be a mixture of existing domains, and whether it is perfectly or imperfectly known.
-This setting is also studied in eg.\ \cite{Mansour09multiple,Hoffman18algorithms}, and  in \cite{Oren19distributionally} for adversarial language model learning.
-\item with respect to the conditional label distribution $P(y|x)$:
-  \begin{itemize}
-  \item $P(y|x)$ is deterministic, meaning that the joint distribution is entirely specified by the marginal and a labelling function $y=h(x)$; a more complex case is when $h()$ also varies across domains;
-  \item $P(y|x)$ is stochastic, but shared across all domains, which is the typical covariate-shift scenario \cite{Shimodaira00improving}; it can also vary across domains \cite{???}\fyTodo{Missing citation vary accross domains}
-  \end{itemize}
-\end{itemize}
-
-Multi-domain learning, as defined in \cite{Dredze08online} additionally makes the requirement that categorial domain tags are available in training and testing; a multi-domain learner should use them effectively \cite{Joshi12multidomain} to improve simultaneously on all domains \cite{Finkel09hierarchical}.
-This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyTodo{explain zero shot}.
-
-% ----
 \subsection{Formalizing multi-domain translation \label{ssec:formalization}}
@@
 A multi-domain learner is then expected to use these tags effectively \cite{Joshi12multidomain} when computing the combined translation function $h(x,d)$, and to perform well in all domains \cite{Finkel09hierarchical}.
-Note that this setting is closely related to the multi-source adaptation problem studied in \cite{Mansour09domainadaptation,Mansour09multiple}.
+Note that this setting is closely related to the multi-source adaptation problem formalized in \cite{Mansour09domainadaptation,Mansour09multiple,Hoffman18algorithms}.
 
-This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is \cite{Farajian17multidomain}, where test translations rely on a similarity scores between test and train sentences, rather than on domain labels.}, and one that we also adopt here.
+This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is \cite{Farajian17multidomain}, where test translations rely on a similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here.
 Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function.
@@
 Such scenario is already well documented for zero-shot multi-lingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}, where mixing languages has more than demonstrated its usefulness.
-An even more challenging case is when the MT would need to perform well for any test distribution (as considered eg.\ in \cite{Hoffman18algorithms,Oren19distributionally}  for other tasks).
-In all these cases, mixing domains in training and/or testing is likely to improve robustness against unexpected inputs.
+An even more challenging case is when the MT would need to perform well for any test distribution (as studied for statistical MT in \cite{Huck15mixeddomain}).
+In all these cases, mixing domains in training and/or testing is likely to improve robustness against unexpected or adversarial inputs \cite{,Oren19distributionally}.
 
@@
 % \cite[eq. (3)]{Wang20general} defines the loss of MDMT as the (macro) average of domain loss - all domains are equally important.
-
-% \subsection{MDMT: a tentative definition \label{ssec:definition}}
-
-% Informally, a multi-domain system should be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
-% \begin{itemize}
-% \item the more common situation is when both train and test documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
-It is usually assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT.
-This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name here: such as supervised MDMT ?
-and mostly supervised MDMT for the second case ?}
-
-% \item It can also be the case that domain labels are only known in training, while test inputs may or may not be associated with explicit domains.
-This is, for instance, the setting adopted \cite{Zeng18multidomain}, where automatic domain identification is performed for test sentences.
-Note that automatic domain identification can also be effectively performed in the first setting, especially in the presence of close or related domains in training.
-
-% \item A last setting, that has not received much attention for MT (see however \cite{Sennrich13multidomain}), is where both train and test domains are automatically infered from data, using clustering techniques.
-Even though domain labels are typically available in most training scenario, these approaches assumes that might be preferable, owing to the multiple sources of intra-corpus and inter-corpus variabilities, to resort to automatic domain assignments.
-The online approach of \newcite{Farajian17multidomain}, where distributional cues, rather that categorical tags, are used for fine-tuning, is a variant of this strategy.
-% \end{itemize}
-
-% In the rest of the paper, we mostly adhere to the first definition which correspond to a situation of practical interest for the industry, and assume that multi-domain MT is mostly about effectively using domain tags in training and testing.
-We also experiment in Section~\ref{sec:results} with the other two alternative scenarios, as they also define useful contexts to evaluate the actual robustness of MDMT systems.
-
-% \jcDone{...a very common scenario in MT industry where clients dispose of mulitiple data corresponding to different domains (differing in genre, thematic, register, style, etc.). }
-
-% \subsection{Evaluation issues \label{ssec:evaluation}}
-
-% In theory, evaluation procedures should report realistic estimates of the test loss, computed on a sample deemed representative of the anticipated test conditions.
-Three scenarios need be distinguished:
-% \begin{itemize}
-% \item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
-% \item all domains are equally likely and important in testing (the prior over test domains is uniform), which would warrant the computation of per domain scores and of unweighted averages, as reported in \cite{Farajian17multidomain,Su19exploring,Wang20general}.
-Note that in this situation, the optimal training regime would be to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data.
-\cite{Wang20general} uses an intermediate mixing regime, which tends to reduce part of the unbalance, and improve the performance for underrepresented domains.
-% \item the test distribution is neither equal to the training set, nor uniform.
-Here again, if this distribution is known in advance, an optimal domain mix should be used in training to compute baseline scores, rather than using the empirical distribution of all the training data.
-A more challenging case is when the test set includes data from domains unseen in training, or when it is chosen adversarially \cite{XXX}.
-% % : this situation, where test instances have a null probability in training, goes agains the basic statistical assumptions of DA \cite{Shimodaira00improving}.
-It is thus important that the likeliness of such events should be clearly formulated when describing the evaluation protocol.
-% \end{itemize}
-
-% The papers in the literature are seldom explicit regarding their assumptions regarding the test distribution.
-In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 most faithfully reflects the expected testing conditions.
-Another important issue is the choice of the evaluation metrics: if BLEU is still the standard metric used in most studies, it is not necessary the best choice to measure subtle improvements of lexical choices \cite{Irvine13measuring} that are often associated with adapted systems.
-
-% A last issue regarding evaluation is related to the measurement of computational costs: when this is the main a motivation for using MDMT, we expect that computational costs would be reported and compared, which however is hardly ever done in practice.
-\fyTodo{Keep this ?}
-
 \section{Challenging multi-domain translation systems \label{sec:challenging}}
@@
 We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\fyDone{A word about meta-parameter settings}, computed with\fyTodo{}.
+Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\\footnote{We use multibleu}fyDone{A word about meta-parameter settings}.
 Significance tests are performed using the bootstrap resampling method of \newcite{Koehn04statistical} and implementation of compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
@@
     \system{LDC-Tag}   & & & & & & &  0.0 & 0.1& \\ 
-    \system{MDL Res}   & -0.2$^+$& -0.07$^+$&+0.18$^+$  &+0.0$^+$ & & & \SW{-2.4} & \SW{-3.5} & \\
+    \system{MDL Res}   & -0.2$^+$& -0.07$^+$&+0.18$^+$  &+0.0$^+$ & & & 0.7 & -0.3 & \\
 %    \system{MDL Gated} & & & & & & & & &  \\
@@
 
-It is finally interesting to note that all systems, but \system{MDL Res}, are quite robust to domain merge: we see even small gains in performance (eg.\ for \system{DM}).
+It is finally interesting to note that all systems are quite robust to domain merge: we even see some small gains in performance (eg.\ for \system{DM}).
 While robustness is generally a good thing, it is also a sign of a poor overall usage of domain labels.
-\system{MDL Res}\fyTodo{To be completed}
+\fyTodo{To be completed}
 
@@
 This is of course detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation}.
-We contrast this approach with retraining all the systems from scratch, and report the difference in performance in Table~\ref{tab:warmrestart}.
+We contrast this approach with retraining all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.
 We expect that a good MDMT system should not be significantly affected by the addition of a new domain, and reach about the same performance as if the training included this domain from scratch.
@@
 The multi-domain training regime is more the norm than the exception for NLP \cite{Dredze08online,Finkel09hierarchical}.
-We focus here mostly on multi-domain learning in machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection and weighting, adversarial training, etc) have also been proposed for most language processing tasks.
+Our systematic study of domain use in MDMT owes much to the study of MD opinion mining systems of \newcite{Joshi12multidomain}.
+We focus here mostly on MD machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection and weighting, adversarial training, etc) have also been proposed for most language processing tasks.
 %; scores are reported separately per domain, suggesting that all domains are equally important at test time.
@@
 
-Multi-domain translation was introduced for statistical MT (eg.\ \cite{Banerjee10combining,Sennrich13multidomain}): as for domain adaptation, two main strategies where considered consisting at weighting train and test instances based on their domain-relatedness on the one hand; or to inject domain labels as additional features in learning.
+Multi-domain translation was already proposed for statistical MT (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}): as for domain adaptation, two main strategies where considered consisting at weighting train and test instances based on their domain-relatedness on the one hand; or to inject domain labels as additional features in learning.
 
 This latter strategy has also been widely used in MNT: the authors of \cite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token, or in the form of additional domain-feature associated to each word.
-Note his technique has also been used to control the style of MT outputs in \cite{Sennrich16politeness,Niu18multitask}, and to encode the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}.
-This approach is extended in \cite{Pham19generic} where source word embeddings are likely to vary across domains: this is achieved by splitting the embedding vector into a generic and a domain-specific part (one for each possible domain); a similar approach, applied to hidden states, is documented in \cite{Chu18multilingual}.
+These results are reproduced in \cite{Tars18multidomain}, which also considers, as we do, automatically induced domain tags.
+Such technique also helps control the style of MT outputs in \cite{Sennrich16politeness,Niu18multitask}, and to encode the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}.
+Domain control can also be performed on the target side, as in \cite{Chen16guided}, where a vector of topic describing the current document is included as an extra context in the softmax layer of the decoder.
 Such feature-based approaches strongly rely on earlier proposals from the era of statistical MT \cite{Daume07frustratingly,Clark12onesystem}.
-Domain control can also be performed on the target side, as in \cite{Chen16guided}, where a vector of topic describing the current document is included as an extra context in the softmax layer of the decoder.
+Such ideas are further developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and commonalties are encoded in the network architecture: some parameters are shared across domains (the recurrent layers and the attention in \cite{Pham19generic}), while others are domain-specific (part of the embeddings in \cite{Pham19generic}).
 
-In \cite{Britz17mixing}, several methods are considered which aim to ensure that domain information is actually used in a mix-domain system.
+The techniques introduced in \cite{Britz17mixing} aim to ensure that domain information is actually used in a mix-domain system.
 Three alternatives are considered, using either domain classification (or domain normalization, via adversarial training) on source or target representations.
 There is no clear winner in either of their three language pairs considered, which only distinguish a limited number of domains.
-The same basic techniques are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the upper layers of the MT aimed to sort out domain specific information on the one hand, and domain-agnostic information on the other hand, using auxiliary classification tasks, hoping that the domain-specific part will capture specific peculiarities of a domain, while the domain-agnostic part will be useful across the board.
-These techniques are implemented in an attentional sequence-to -sequence architecture in \cite{Zeng18multidomain}, and in Transformers in \cite{Su19exploring}.
+One contribution of this work was to introduce the idea of normalizing representations through adversarial training, so has to make the mixture of heterogeneous data more effective; representation normalization has since proven a key ingredient in multilingual transfer learning.
+The same basic techniques (partial parameter sharing, automatic domain identification / normalization) are simultaneously at play in \cite{Zeng18multidomain}: in this approach, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific representations on the one hand from domain-agnostic representation on the other hand.
+These representations are then processed as two separate inputs, then recombined to compute the translation.
+These techniques are extended and ported to the Transformer model by \newcite{Su19exploring}.
 
-A slightly more complex parameter-sharing scheme is presented in \cite{Jiang19multidomain}, which augments a transformer model with domain specific heads, whose individual contributions are regulated at the word (position-level): some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads.
-The results for three language pairs outperform several standard baselines for a 2-domain systems (in fr:en and ge:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
+A slightly more complex parameter-sharing scheme is presented in \cite{Jiang19multidomain}, which augments a Transformer model with domain specific heads, whose individual contributions are regulated at the word (position-level): some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling that is used in \cite{Huck15mixeddomain,Saunders19ucam}.
+The results for three language pairs outperform several standard baselines for a two-domain systems (in fr:en and ge:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
 
-The work of \cite{Tars18multidomain} makes a good case against fine-tuning, which is often found to be too brittle, a defect that MDMT can fix, delivering superior generalization accuracy.
-They perform experiment with reference domain tags, using mostly the domain-control approach.
-They also report results obtained with automatically induced domain tags, which show that working with automatically induced tags (at the sentence level) is a viable alternative to using corpus labels as domain tags, a result that we have confirmed in section~\ref{}.
-
-\newcite{Farajian17multidomain} also used automatic domains, performing on-line adaptation at the sentence-level.
+As already explained, \newcite{Farajian17multidomain} differs from most studies, as it does not use domain labels.
 In their approach, the processing of a test sentence triggers the selection of a dedicated (small) set of fine-tuning instances assumed to be similar to the input text; using these, the generic NMT system is then tuned for some epochs, before delivering its translation.
-Different of most approaches studied here, this approach entirely dispenses with the notion of a domain, relying primarily on data selection/weighting techniques to handle data heterogeneity.
-
-\subsection{Formalizing multi-domain adaptation (MDA)}
-
-In MDA, the typical training situation (for a classifier) considers $k$ domains with $\mathcal{D}_k(x,y)$ the joint distribution of features $x \in \mathcal{X}$ and labels $y \in \mathcal{Y}$ for domain $k$.
-In \cite{Mansour09domainadaptation}, the labeling function $f$ is deterministic and the joint distribution is entirely defined by the marginal $\mathcal{D}_k(x)$ and $f()$.This work further assumes $k$ hypotheses $h_k(x): \mathcal{X} \rightarrow{} \mathcal{Y}$ with an error smaller than $\epsilon$ when $x\sim \mathcal{D}_k$: $\sum_{x} L(h_k(x), f(y)) \mathcal{D}_k(x) < \epsilon_k$ and that the target distribution is a convex combination of $\{\mathcal{D}_k\}$: $\mathcal{D}_T(x) = \sum_k \lambda_k \mathcal{D}_k(x)$.
-\cite{Mansour09domainadaptation} focuses on \textsl{combined hypotheses}, ie.\ hypotheses for which each prediction $h(x)$ is a linear combination of domain specific hypotheses $h(x_k)$, and derives generalization bounds for the case where $\mathcal{D}_T(x)$ is fixed in advance, or worst case bounds valid for any mixture target distribution.
-The optimal solution has the form $h_{z,\eta}(x) = \sum_k \frac{z_kD_k(x) + U(x)/\gamma}{\sum_{k'} z_kD_k(x) + \eta U(x)/l}$ and is termed the \emph{distribution weighted combined rule}.
-Further work in \cite{Mansour09multiple} generalizes these bounds to the case where $\mathcal{D}_T$ is not a mixture of existing domains (in that case we will find a mixture that is close to $\mathcal{}$in the Renyì-divergence sense ); where $\mathcal{D}_k$ are unknown and estimated from the data; or when the labelling function is not the same across domains.
-More recent work study the case where the labelling function is stochastic, meaning that domains are defined by joint ($\mathcal{D}_k(x,y)$) rather than marginal distributions \cite{Hoffman18algorithms}, and provides an algorithm to compute the optimal weights for the combined rule based on estimates of the joint distributions.
-A guiding principle is that this rule should tend towards uniform performance across domains, to ensure robustness in the worst case scenario \cite{Oren19distributionally}.
-
-A related problem is \emph{Multiple source single domain adaptation} which is a variation of conventional DA from a mixture of source domains.
-In this setting it is advantageous to take advantage of the similarities between each source and the fixed target domain (see eg.\ \cite{Duan12domain,Cortes19AdaptationBO,Wen19domain} for studies of the supervised case, \cite{Zhao18adversarial} for the unsupervised case).
+This approach entirely dispenses with the notion of a domain, and primarily relies on data selection/weighting techniques to handle data heterogeneity.
 
@@
 \end{document}
+
+\subsection{Formalizing multi-domain adaptation (MDA)}
+
+In MDA, the typical training situation (for a classifier) considers $k$ domains with $\mathcal{D}_k(x,y)$ the joint distribution of features $x \in \mathcal{X}$ and labels $y \in \mathcal{Y}$ for domain $k$.
+In \cite{Mansour09domainadaptation}, the labeling function $f$ is deterministic and the joint distribution is entirely defined by the marginal $\mathcal{D}_k(x)$ and $f()$.This work further assumes $k$ hypotheses $h_k(x): \mathcal{X} \rightarrow{} \mathcal{Y}$ with an error smaller than $\epsilon$ when $x\sim \mathcal{D}_k$: $\sum_{x} L(h_k(x), f(y)) \mathcal{D}_k(x) < \epsilon_k$ and that the target distribution is a convex combination of $\{\mathcal{D}_k\}$: $\mathcal{D}_T(x) = \sum_k \lambda_k \mathcal{D}_k(x)$.
+\cite{Mansour09domainadaptation} focuses on \textsl{combined hypotheses}, ie.\ hypotheses for which each prediction $h(x)$ is a linear combination of domain specific hypotheses $h(x_k)$, and derives generalization bounds for the case where $\mathcal{D}_T(x)$ is fixed in advance, or worst case bounds valid for any mixture target distribution.
+The optimal solution has the form $h_{z,\eta}(x) = \sum_k \frac{z_kD_k(x) + U(x)/\gamma}{\sum_{k'} z_kD_k(x) + \eta U(x)/l}$ and is termed the \emph{distribution weighted combined rule}.
+Further work in \cite{Mansour09multiple} generalizes these bounds to the case where $\mathcal{D}_T$ is not a mixture of existing domains (in that case we will find a mixture that is close to $\mathcal{}$in the Renyì-divergence sense ); where $\mathcal{D}_k$ are unknown and estimated from the data; or when the labelling function is not the same across domains.
+More recent work study the case where the labelling function is stochastic, meaning that domains are defined by joint ($\mathcal{D}_k(x,y)$) rather than marginal distributions \cite{Hoffman18algorithms}, and provides an algorithm to compute the optimal weights for the combined rule based on estimates of the joint distributions.
+A guiding principle is that this rule should tend towards uniform performance across domains, to ensure robustness in the worst case scenario \cite{Oren19distributionally}.
+
+A related problem is \emph{Multiple source single domain adaptation} which is a variation of conventional DA from a mixture of source domains.
+In this setting it is advantageous to take advantage of the similarities between each source and the fixed target domain (see eg.\ \cite{Duan12domain,Cortes19AdaptationBO,Wen19domain} for studies of the supervised case, \cite{Zhao18adversarial} for the unsupervised case).
+
+\subsection{Complements - to be removed \label{ssec:oldformalization}}
+
+We define a domain $d$ as a joint distribution $\mathcal{D}_d(x,y)$ over input-output pairs $(x,y) \in \mathcal{X}\times{}\mathcal{Y}$.
+Assuming that the training data is sampled from a known mixture of $D$ source domains $\mathcal{D}_s(x,y) = \sum_d \lambda_{s,d} \mathcal{D}_d(x,y)$, several cases are studied in the ML literature:
+
+\begin{itemize}
+\item the test distribution $\mathcal{D}_t$ is known 
+  \begin{itemize}
+  \item $\mathcal{D}_t = \mathcal{D}_{s}$: this is standard ML, with mixture densities
+  \item $\mathcal{D}_t(x,y) = \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) $ with a different mixture of domains in training and testing ($\vlambda_s \neq \vlambda_t$), corresponding to multiple source adaptation (MSA) \cite{Mansour09domainadaptation},
+  \item $\mathcal{D}_t(x,y) \neq \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) \forall \lambda$, where the target domain for adapation is not a mixture of source domains, studied in \cite{Mansour09multiple}
+  \item variants of MSA which only assume knowledge of the marginal distribution $\mathcal{D}_t(x) = \sum_{y} \lambda_{t,d} \mathcal{D}_d(x) $
+  \item more realistic situations where  the train and test distributions are estimated from finite samples, also studied in \cite{Mansour09multiple}.
+  \end{itemize}
+   Note that we recover standard supervised DA is when there is only one component in the training and test mixtures, and multiple-source single domain adaptation when the source is a mixture but the test domain is homogeneous.
+Unsupervised DA \cite{Daume06domain} corresponds to the case where only the marginal distribution of the test domain inputs is available, while semi-supervised DA is the reverse situation.
+
+\item the test distribution is not known and training needs to be distributionally robust, ie.\ minimize the error against all possible distributions, with variants depending on whether it is or not assumed to be a mixture of existing domains, and whether it is perfectly or imperfectly known.
+This setting is also studied in eg.\ \cite{Mansour09multiple,Hoffman18algorithms}, and  in \cite{Oren19distributionally} for adversarial language model learning.
+\item with respect to the conditional label distribution $P(y|x)$:
+  \begin{itemize}
+  \item $P(y|x)$ is deterministic, meaning that the joint distribution is entirely specified by the marginal and a labelling function $y=h(x)$; a more complex case is when $h()$ also varies across domains;
+  \item $P(y|x)$ is stochastic, but shared across all domains, which is the typical covariate-shift scenario \cite{Shimodaira00improving}; it can also vary across domains \cite{???}\fyTodo{Missing citation vary accross domains}
+  \end{itemize}
+\end{itemize}
+
+Multi-domain learning, as defined in \cite{Dredze08online} additionally makes the requirement that categorial domain tags are available in training and testing; a multi-domain learner should use them effectively \cite{Joshi12multidomain} to improve simultaneously on all domains \cite{Finkel09hierarchical}.
+This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyDone{explain zero shot}.
+% ----
+
+% \subsection{MDMT: a tentative definition \label{ssec:definition}}
+
+% Informally, a multi-domain system should be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
+% \begin{itemize}
+% \item the more common situation is when both train and test documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
+It is usually assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT.
+This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name here: such as supervised MDMT ?
+and mostly supervised MDMT for the second case ?}
+
+% \item It can also be the case that domain labels are only known in training, while test inputs may or may not be associated with explicit domains.
+This is, for instance, the setting adopted \cite{Zeng18multidomain}, where automatic domain identification is performed for test sentences.
+Note that automatic domain identification can also be effectively performed in the first setting, especially in the presence of close or related domains in training.
+
+% \item A last setting, that has not received much attention for MT (see however \cite{Sennrich13multidomain}), is where both train and test domains are automatically infered from data, using clustering techniques.
+Even though domain labels are typically available in most training scenario, these approaches assumes that might be preferable, owing to the multiple sources of intra-corpus and inter-corpus variabilities, to resort to automatic domain assignments.
+The online approach of \newcite{Farajian17multidomain}, where distributional cues, rather that categorical tags, are used for fine-tuning, is a variant of this strategy.
+% \end{itemize}
+
+% In the rest of the paper, we mostly adhere to the first definition which correspond to a situation of practical interest for the industry, and assume that multi-domain MT is mostly about effectively using domain tags in training and testing.
+We also experiment in Section~\ref{sec:results} with the other two alternative scenarios, as they also define useful contexts to evaluate the actual robustness of MDMT systems.
+
+% \jcDone{...a very common scenario in MT industry where clients dispose of mulitiple data corresponding to different domains (differing in genre, thematic, register, style, etc.). }
+
+% \subsection{Evaluation issues \label{ssec:evaluation}}
+
+% In theory, evaluation procedures should report realistic estimates of the test loss, computed on a sample deemed representative of the anticipated test conditions.
+Three scenarios need be distinguished:
+% \begin{itemize}
+% \item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
+% \item all domains are equally likely and important in testing (the prior over test domains is uniform), which would warrant the computation of per domain scores and of unweighted averages, as reported in \cite{Farajian17multidomain,Su19exploring,Wang20general}.
+Note that in this situation, the optimal training regime would be to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data.
+\cite{Wang20general} uses an intermediate mixing regime, which tends to reduce part of the unbalance, and improve the performance for underrepresented domains.
+% \item the test distribution is neither equal to the training set, nor uniform.
+Here again, if this distribution is known in advance, an optimal domain mix should be used in training to compute baseline scores, rather than using the empirical distribution of all the training data.
+A more challenging case is when the test set includes data from domains unseen in training, or when it is chosen adversarially \cite{XXX}.
+% % : this situation, where test instances have a null probability in training, goes agains the basic statistical assumptions of DA \cite{Shimodaira00improving}.
+It is thus important that the likeliness of such events should be clearly formulated when describing the evaluation protocol.
+% \end{itemize}
+
+% The papers in the literature are seldom explicit regarding their assumptions regarding the test distribution.
+In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 most faithfully reflects the expected testing conditions.
+Another important issue is the choice of the evaluation metrics: if BLEU is still the standard metric used in most studies, it is not necessary the best choice to measure subtle improvements of lexical choices \cite{Irvine13measuring} that are often associated with adapted systems.
+
+% A last issue regarding evaluation is related to the measurement of computational costs: when this is the main a motivation for using MDMT, we expect that computational costs would be reported and compared, which however is hardly ever done in practice.
+\fyTodo{Keep this ?}

### 2020-04-20 2020/coling2020/multidomain.tex
revised related work, draft conclusion and abstract

@@
 \begin{abstract}
-Abstract goes here \fyTodo{Write abstract}
+When building machine translation systems, it is often the case that one need to make the best out of the heterogeneous set of parallel data in training, and to robustly handle inputs from unexpected domains in testing.
+This multi-domain scenario has attracted a lot of recent work, that fall under the general umbrella of transfert learning.
+In this study, we revisit multi-domain machine translation, trying to precisely formulate the motivations for developing such systems, and the associated expectations in terms of performance.
+Our experiments with a representative sample of recent multi-domain systems show that most of these requirements are hardly met in our settings, and suggest that further work is needed to better analyze the current behoviour of multi-domain system, and to make them fully holding their operational promises.
+  \fyDone{Write abstract}
 \end{abstract}
@@
 
-A seemingly related problem is multi-domain (MD) machine translation \cite{Sajjad17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system has to be trained as well as tested with data from multiple domains.
-MD machine translation corresponds to a very common situation, where all available data, no matter its origin, is used to train a system that will be robust to domain variability in testing.
+A seemingly related problem is multi-domain (MD) machine translation \cite{Sajjad17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system is trained and tested with data from multiple domains.
+MD machine translation corresponds to a very common situation, where all available data, no matter its origin, is used to train a robust system that perform well for any kind of new input.
 % As pointed out both by \newcite{Dredze08online} and \cite{Finkel09hierarchical}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}.
@@
 
-In this paper, we try to study these questions and to spell out in a more precise fashion the requirements that a MDMT system should meet.
-Our first contribution is of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them.
-In doing so, we also shed lights on new problems that arise in this context, regarding for instance the accomodation of new domains in the course of training, or the computation of automating domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental, and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
-We show that\fyTodo{Spell out conclusions}.
+In this paper, we try to spell out in a more precise fashion the requirements that an effective MDMT system should meet (section~\ref{sec:requirements}).
+Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them (section~\ref{sec:challenging}).
+In doing so, we also shed lights on new problems that arise in this context, regarding for instance the accomodation of new domains in the course of training, or the computation of automating domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
+We show in section~\ref{sec:experiments}  existing approaches still fall short to match many of these requirements, notably with respect to handling a large amount of heterogeneous domains and to dynamically integrating new domains in training.\fyDone{Spell out conclusions}
 
@@
 An even more challenging case is when the MT would need to perform well for any test distribution (as studied for statistical MT in \cite{Huck15mixeddomain}).
-In all these cases, mixing domains in training and/or testing is likely to improve robustness against unexpected or adversarial inputs \cite{,Oren19distributionally}.
+In all these cases, mixing domains in training and/or testing is likely to improve robustness against unexpected or adversarial inputs \cite{Oren19distributionally}.
 
@@
 By introducing variability in training, it prevents domain adaptation from overfitting the available adaptation data and could help improve generalization even for well-resourced domains.
-A related case is made in \cite{Joshi12multidomain}, where some of the benefits of MD training are attributed to an ensembling effect, where several systems from multiple domains are simultaneously used in the prediction phase\cite{Saunders19ucam}; this effect may subsist even in the absence of clear domain separations.
+A related case is made in \cite{Joshi12multidomain}, where some of the benefits of MD training are attributed to an ensembling effect, where several systems from multiple domains are simultaneously used in the prediction phase; this effect may subsist even in the absence of clear domain separations.
 
 \fyDone{recap}
-To recap, there are multiple arguments for adopting MDMT, some already used in DA settings, and some are new.
+To recap, there are multiple arguments for adopting MDMT, some already used in DA settings, and some original.
 These arguments are not mutually exclusive; however each yields specific expectations with respect to the performance of this approach, and should also yield appropriate evaluation procedure.
@@
 If finally, it is to make the system more robust to unexpected or adversarial test distributions, then this is the setting that should be used to evaluate MDMT.
-The next section discusses ways in which these various requireents of MDMT systems can ben challenged.
+The next section discusses ways in which these requirements of MDMT systems could ben challenged.
 
@@
 To make our results comparable with the literature, we will only use the BLEU score \cite{Papineni02bleu} in Section~\ref{sec:experiments}, noting it may not be the best choice to measure subtle improvements of lexical choices that are often associated with domain adapted systems \cite{Irvine13measuring}.
-Other important figure of merits for MDMT systems are the computational cost and the total number of parameters.\fyDone{Insert discussion about scores}
+Other important figure of merits for MDMT systems are the computational training cost and the total number of parameters.\fyDone{Insert discussion about scores}
 
@@
 
-The multi-domain training regime is more the norm than the exception for NLP \cite{Dredze08online,Finkel09hierarchical}.
-Our systematic study of domain use in MDMT owes much to the study of MD opinion mining systems of \newcite{Joshi12multidomain}.
-We focus here mostly on MD machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection and weighting, adversarial training, etc) have also been proposed for most language processing tasks.
+The multi-domain training regime is more the norm than the exception for natural language processing \cite{Dredze08online,Finkel09hierarchical}, and the design of multi-domain systems has been proposed for many language processing tasks.
+We focus here exclusively on MD machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection / weighting, adversarial training, etc) may have been studied in other contexts.
+
 %; scores are reported separately per domain, suggesting that all domains are equally important at test time.
@@
 
-Multi-domain translation was already proposed for statistical MT (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}): as for domain adaptation, two main strategies where considered consisting at weighting train and test instances based on their domain-relatedness on the one hand; or to inject domain labels as additional features in learning.
+Multi-domain translation was already proposed for statistical MT (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}).
+Two main strategies were considered: instance-based, involving a measure of similarities between train and test domains; feature-based, where domain labels give rise to additional features.
 
-This latter strategy has also been widely used in MNT: the authors of \cite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token, or in the form of additional domain-feature associated to each word.
+This latter strategy has also been widely used in MNT: the authors of \cite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token or in the form of additional domain-feature associated to each word.
 These results are reproduced in \cite{Tars18multidomain}, which also considers, as we do, automatically induced domain tags.
-Such technique also helps control the style of MT outputs in \cite{Sennrich16politeness,Niu18multitask}, and to encode the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}.
+This technique also helps control the style of MT outputs in \cite{Sennrich16politeness,Niu18multitask}, and to encode the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}.
 Domain control can also be performed on the target side, as in \cite{Chen16guided}, where a vector of topic describing the current document is included as an extra context in the softmax layer of the decoder.
-Such feature-based approaches strongly rely on earlier proposals from the era of statistical MT \cite{Daume07frustratingly,Clark12onesystem}.
-Such ideas are further developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and commonalties are encoded in the network architecture: some parameters are shared across domains (the recurrent layers and the attention in \cite{Pham19generic}), while others are domain-specific (part of the embeddings in \cite{Pham19generic}).
+Such ideas are further developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and commonalties are encoded in the network architecture: some parameters
+are shared across all domains, while others are domain-specific. % (part of the embeddings in \cite{Pham19generic}).
 
@@
 There is no clear winner in either of their three language pairs considered, which only distinguish a limited number of domains.
-One contribution of this work was to introduce the idea of normalizing representations through adversarial training, so has to make the mixture of heterogeneous data more effective; representation normalization has since proven a key ingredient in multilingual transfer learning.
-The same basic techniques (partial parameter sharing, automatic domain identification / normalization) are simultaneously at play in \cite{Zeng18multidomain}: in this approach, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific representations on the one hand from domain-agnostic representation on the other hand.
+One contribution of this work was to introduce the idea of normalizing representations through adversarial training, in order to make the mixture of heterogeneous data more effective; representation normalization has since proven a key ingredient in multilingual transfer learning.
+The same basic techniques (partial parameter sharing, automatic domain identification / normalization) are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific representations on the one hand from domain-agnostic representation on the other hand.
 These representations are then processed as two separate inputs, then recombined to compute the translation.
-These techniques are extended and ported to the Transformer model by \newcite{Su19exploring}.
 
-A slightly more complex parameter-sharing scheme is presented in \cite{Jiang19multidomain}, which augments a Transformer model with domain specific heads, whose individual contributions are regulated at the word (position-level): some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling that is used in \cite{Huck15mixeddomain,Saunders19ucam}.
+Another parameter-sharing scheme is presented in \cite{Jiang19multidomain}, which augments a Transformer model with domain specific heads, whose individual contributions are regulated at the word (position-level): some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling that the core of \cite{Huck15mixeddomain,Saunders19ucam}.
 The results for three language pairs outperform several standard baselines for a two-domain systems (in fr:en and ge:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
@@
 As already explained, \newcite{Farajian17multidomain} differs from most studies, as it does not use domain labels.
-In their approach, the processing of a test sentence triggers the selection of a dedicated (small) set of fine-tuning instances assumed to be similar to the input text; using these, the generic NMT system is then tuned for some epochs, before delivering its translation.
-This approach entirely dispenses with the notion of a domain, and primarily relies on data selection/weighting techniques to handle data heterogeneity.
+In their approach, the processing of a test sentence triggers the selection of a small set of similar fine-tuning instances; using these, the generic NMT system is tuned for some epochs, before delivering its translation.
+This approach thus entirely dispenses with the notion of a domain, and primarily relies on data selection/weighting techniques to handle data heterogeneity.
 
 \section{Conclusion and outlook \label{sec:conclusion}}
-\fyTodo{Write conclusions}
+\fyDone{Write conclusions}
+In this study, we have carefully reconsidered the idea of multi-domain machine translation, which seems to be taken for granted in many recent studies.
+We have spelled out the various motivations for building such systems, and associated expectations in terms of system performance.
+We have then designed a series of requirements that MDMT systems should meet, and designed associated testing procedures.
+In our experiments with multiple MDMTs, we have found that most requirements were not met for our experimental conditions.
+If MDMT systems are able to outperform the mixed-domain baseline, at least for some domains, they all fall short to match the performance of fine-tuning on each individual domain, which remains the best choice in multi-source single domain adaptation.
+In this comparison, architecture-based strategies seem to show the best promises.
+As expected however, MDMT are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend,
+dynamically accommodate additional domains:this is especially easy for feature-based methodologies.
+Our experiments finally suggest that all methods show decreasing performance when the number of domain increases (for a fixed data size).
+
+Two other main conclusions can be drawed from this study: first, it seems that more work is needed to make MDMT systems make the best out of the variety of the available data, both to effectively share what needs to be shared while at the same time separating what needs to be separated.
+We notably see two areas worthy of further explorations: the development of parameter sharing strategies when the number of domains is large; the design of training strategies that can effectively handle a change of the training mixture, including an increase in the number of domains.
+Both problems are of practical relevance in industrial settings.
+Second, and may be more importantly, there is a general need to adopt better evaluation methodologies for evaluating MDMT systems, require systems developers to clearly spell out the testing conditions and the associated expected distribution of testing instances, and to report more than mere comparisons with simple baselines on a fixed and known pair of domains.
 

### 2020-04-22 2020/coling2020/multidomain.tex
missing parenthesis
complete some blanks in table 3,4,5

@@
     \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 & 23.5\\
-    \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 &\\
+    \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 & \\
     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8 &\\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & 42.3 & 53.3 &\\ \hline
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & 42.3 & 53.3 & 20.51\\ \hline
  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
@@
     \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   & \SW{21.7} \\
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7       & \\
-    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8 & 22.1\\ 
-    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 & \\
-    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 40.7 & 47.5 & \\
-    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 & \\
+    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7       &  23.43 \\
+    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8 & 22.1 \\ 
+    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 & 22.64\\
+    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 40.7 & 47.5 & 23.34 \\
+    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 & 20.43 \\
     \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.1 & 51.9  & \SW{21.2} \\
@@
     \hfill Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & Wrong \\ % \hline
-     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{bank} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} & domain \\ \hline
+     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{law} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} & domain \\ \hline
     & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law} & \\
-    \system{FT-Full}      & -0.13 & -0.6 & -1.51& -0.25& & &\SW{-1.6} & \SW{-1.4}& \\
-    \system{DC-Tag}     & -0.16 & -0.27& +0.13& +0.2& & & -0.5 & -0.4 & \\
-    \system{DC-Feat}    & -0.47& 0.0 & \SB{0.27} $^+$ & +0.27 $^+$& & & 0.3 & 0.1 & \\
-    \system{TTM}          & & & & & & &  0.0 & -0.3 &  \\
-    \system{LDC-Tag}   & & & & & & &  0.0 & 0.1& \\ 
-    \system{MDL Res}   & -0.2$^+$& -0.07$^+$&+0.18$^+$  &+0.0$^+$ & & & 0.7 & -0.3 & \\
+    \system{FT-Full}      & -0.13 & -0.6 & -1.51& -0.25& -2.28 & -5.08 &\SW{-1.6} & \SW{-1.4}& \\
+    \system{DC-Tag}     & -0.16 & -0.27& +\SB{0.13}$^+$& +0.2& -2.25 & -2.92 & -0.5 & -0.4 & \\
+    \system{DC-Feat}    & -0.47 & 0.0 & +\SB{0.28} $^+$ & +0.28 & +0.32 & +0.33 & 0.3 & 0.1 & \\
+    \system{TTM}          & -0.23 &  & -0.2 & & -0.3 & &  0.0 & -0.3 &  \\
+    \system{LDC-Tag} & +0.08 & +0.15 & +0.36 & +0.28 & -0.04 & -0.01 &  0.0 & 0.1& \\ 
+    \system{MDL Res}   & -0.2 & -0.07 & +\SB{0.18}$^+$ &+0.0 & -0.89 & -0.89 & +\SB{0.7} & -0.35 & \\
 %    \system{MDL Gated} & & & & & & & & &  \\
-    \system{ADM} & & & & & & &  0.1 & -0.4 & \\
-    \system{DM}    & & & & & & & 0.9 & 0.1 & \\
-    \system{WDC}  & & & & & & & & & \\
+    \system{ADM} & +0.6 & & +0.4 & & +0.44 & &  0.1 & -0.4 & \\
+    \system{DM}    & -0.35 & & +0.4 & & +0.33 & & 0.9 & 0.1 & \\
+    \system{WDC}  & -0.95 & & -1.1 & & -2.25 & & & & \\
     \hline
@@
     \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & & \\
-     continual                &&&&&&&& \\
+     continual                & 37.45 & 56.09 & 51.11 & 34.04  &  44.29 & 87.32 & 28.35 & \\
      \hline

### 2020-04-22 2020/coling2020/multidomain.tex
TACL style

--- 2020-04-20 2020/coling2020/multidomain.tex
@@
 
-\documentclass[11pt]{article}
-\usepackage{coling2020}
+\documentclass[11pt,a4paper]{article}
+% \usepackage{coling2020}
+\usepackage{tacl2018v2}
 \usepackage{times}
@@
 % 
-\blfootnote{
+\footnote{
     %
@@
 
-In this paper, we try to spell out in a more precise fashion the requirements that an effective MDMT system should meet (section~\ref{sec:requirements}).
+In this paper, we spell out in a more precise fashion the requirements that an effective MDMT system should meet (section~\ref{sec:requirements}).
 Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them (section~\ref{sec:challenging}).
@@
 \section{Requirements of multi-domain MT \label{sec:requirements}}
-In this section, we try to recap the main reasons for considering a multi-domain scenario and discuss their implications in terms of performance evaluation.
+In this section, we recap the main reasons for considering a multi-domain scenario and discuss their implications in terms of performance evaluation.
 \fyDone{Relation to theory, discussion of worst case scenario}
@@
 Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
-$h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$, but can differ from one domain to the other.
+$h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$), but can differ from one domain to the other.
 
@@
 Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function.
-We now discuss the arguments that are usually put forward to proceed differently.
+We now discuss the arguments that are put forward to proceed differently.
 % --
@@
 % include your own bib file like this:
-\bibliographystyle{coling}
+% \bibliographystyle{coling}
+\bibliographystyle{acl_natbib}
 \bibliography{multidomain}

### 2020-04-23 2020/coling2020/multidomain.tex
more typos, space reductions etc

@@
 \newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}
+}\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyFuture}[1]{\done[FY]\Todo[FY:]{\textcolor{red}{#1}}}
 \newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
@@
 This multi-domain scenario has attracted a lot of recent work, that fall under the general umbrella of transfert learning.
-In this study, we revisit multi-domain machine translation, trying to precisely formulate the motivations for developing such systems, and the associated expectations in terms of performance.
-Our experiments with a representative sample of recent multi-domain systems show that most of these requirements are hardly met in our settings, and suggest that further work is needed to better analyze the current behoviour of multi-domain system, and to make them fully holding their operational promises.
+In this study, we revisit multi-domain machine translation, with the aim to formulate the motivations for developing such systems and the associated expectations with respect to performance.
+Our experiments with a large sample of multi-domain systems show that most of these expectations are hardly met and suggest that further work is needed to better analyze the current behoviour of multi-domain systems and to make them fully hold their promises.
   \fyDone{Write abstract}
@@
 % 
-\footnote{
-    %
-    % for review submission
-    %
-    \hspace{-0.65cm}  % space normally used by the marker
-    Place licence statement here for the camera-ready version. See
-    Section~\ref{licence} of the instructions for preparing a
-    manuscript.
-    %
-    % % final paper: en-uk version 
-    %
+% \footnote{
+%     %
+%     % for review submission
+%     %
 %     \hspace{-0.65cm}  % space normally used by the marker
-    % This work is licensed under a Creative Commons 
-    % Attribution 4.0 International Licence.
-    % Licence details:
-    % \url{http://creativecommons.org/licenses/by/4.0/}.
-    % 
-    % % final paper: en-us version 
-    %
-    % \hspace{-0.65cm}  % space normally used by the marker
-    % This work is licensed under a Creative Commons 
-    % Attribution 4.0 International License.
-    % License details:
-    % \url{http://creativecommons.org/licenses/by/4.0/}.
-}
+%     Place licence statement here for the camera-ready version. See
+%     Section~\ref{licence} of the instructions for preparing a
+%     manuscript.
+%     %
+%     % % final paper: en-uk version 
+%     %
+%     % \hspace{-0.65cm}  % space normally used by the marker
+%     % This work is licensed under a Creative Commons 
+%     % Attribution 4.0 International Licence.
+%     % Licence details:
+%     % \url{http://creativecommons.org/licenses/by/4.0/}.
+%     % 
+%     % % final paper: en-us version 
+%     %
+%     % \hspace{-0.65cm}  % space normally used by the marker
+%     % This work is licensed under a Creative Commons 
+%     % Attribution 4.0 International License.
+%     % License details:
+%     % \url{http://creativecommons.org/licenses/by/4.0/}.
+% }
 
@@
 In this setting, \emph{domain adaptation} (DA) methods are in order.
-DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory,Pan10asurvey})\fyDone{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}) and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D}_t$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
+DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory, Quinonero08dataset,Pan10asurvey})\fyDone{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}) and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D}_t$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
 % In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
 
-A seemingly related problem is multi-domain (MD) machine translation \cite{Sajjad17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system is trained and tested with data from multiple domains.
-MD machine translation corresponds to a very common situation, where all available data, no matter its origin, is used to train a robust system that perform well for any kind of new input.
+A seemingly related problem is \emph{multi-domain} (MD) machine translation \cite{Sajjad17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system is trained and tested with data from multiple domains.
+MD machine translation (MDMT) corresponds to a very common situation, where all available data, no matter its origin, is used to train a robust system that perform well for any kind of new input.
 % As pointed out both by \newcite{Dredze08online} and \cite{Finkel09hierarchical}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}.
-If the intuitions behind MDMT are quite simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should MDMT perform well when the test data is distributed like the training data, when it is equally distributed across domains or when the test distribution is unknown?
+If the intuitions behind MDMT are quite simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should MDMT performs well when the test data is distributed like the training data, when it is equally distributed across domains or when the test distribution is unknown?
 Should MDMT also be robust to new domains?
@@
 
-A related question concerns the relationship between domain adaptation and multi-domain translation.
-The latter setting seems more challenging and one would expect that adapting a system to a known target domain $A$ will be easier than training a MDMT system, where only part of the data belongs to domain $A$.
+A related question concerns the relationship between supervised domain adaptation and multi-domain translation.
+The latter task seems more challenging as adapting a system to a known target domain $A$ looks easier than training a MDMT system, which should do also well on domains other than $A$.
 Are there situations where MD systems can surpass DA, as is sometimes expected?
 
-In this paper, we spell out in a more precise fashion the requirements that an effective MDMT system should meet (section~\ref{sec:requirements}).
+In this paper, we formulate in a more precise fashion the requirements that an effective MDMT system should meet (section~\ref{sec:requirements}).
 Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them (section~\ref{sec:challenging}).
@@
 A distinct line of reasoning is that mixing domains can have a positive regularization effect for all domains.
-By introducing variability in training, it prevents domain adaptation from overfitting the available adaptation data and could help improve generalization even for well-resourced domains.
-A related case is made in \cite{Joshi12multidomain}, where some of the benefits of MD training are attributed to an ensembling effect, where several systems from multiple domains are simultaneously used in the prediction phase; this effect may subsist even in the absence of clear domain separations.
+By introducing variability in training, it prevents DA from overfitting the available adaptation data and could help improve generalization even for well-resourced domains.
+A related case is made in \cite{Joshi12multidomain}, which shows that part of the benefits of MD training is due to an ensembling effect, where systems from multiple domains are simultaneously used in the prediction phase; this effect may subsist even in the absence of clear domain separations.
 
@@
 % \cite[eq. (3)]{Wang20general} defines the loss of MDMT as the (macro) average of domain loss - all domains are equally important.
-\section{Challenging multi-domain translation systems \label{sec:challenging}}
-In this section, we discuss seven\fyTodo{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.
+\section{Challenging multi-domain systems \label{sec:challenging}}
+In this section, we propose seven\fyDone{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.
 All these evaluations will rest on comparison of translation performance, and do not depend on the choice a particular metrics.
-To make our results comparable with the literature, we will only use the BLEU score \cite{Papineni02bleu} in Section~\ref{sec:experiments}, noting it may not be the best choice to measure subtle improvements of lexical choices that are often associated with domain adapted systems \cite{Irvine13measuring}.
+To make our results comparable with the literature, we will only use the BLEU score \cite{Papineni02bleu} in Section~\ref{sec:experiments}, noting it may not be the best yardstick to assess subtle improvements of lexical choices that are often associated with domain adapted systems \cite{Irvine13measuring}.
 Other important figure of merits for MDMT systems are the computational training cost and the total number of parameters.\fyDone{Insert discussion about scores}
@@
 \paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} A MDMT should perform better than the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}.
-Evaluating this requirement is a matter of a mere comparison, assuming the test distribution of domains is known: if all domains are equally important, performance averages can be reported; if they are not, weighted averages should be used instead.
-
-\fyTodo{Try random label assignments ?} % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
+Evaluating this requirement is a matter of a mere comparison, assuming the test distribution of domains is known: if all domains are equally important, performance averages can be reported; if they are not, weighted averages should be used instead.\fyFuture{Try random label assignments ?} % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
 
@@
 In this situation, MDMT should even outperform fine-tuning on either of the pseudo sub-domain.
-\jcDone{I dont understand the prediction... why running MDMT on 2 artificial subdomains should show  gains over fine-tunning?}\fyTodo{small loss with respect to non split ?}
+\jcDone{I dont understand the prediction... why running MDMT on 2 artificial subdomains should show  gains over fine-tunning?}\fyDone{small loss with respect to non split ?}
 
-\subsection{Robustness with respect to domain definition \label{ssec:robusness}}
+\subsection{Robustness to fuzzy domain separation \label{ssec:robusness}}
 A second set of requirements is related to the definition of a domain.
@@
 A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of domain tag errors.
-This is in sharp contrast with what is commonly observed with fine-tuning, where giving the wrong domain tag can yield ``catastrophic'' drop in performance.
-This requirement is measured by decoding the test set for domain $A$ with all other domain tags, and report average drop in performance.
+This is in sharp contrast with what is observed with fine-tuning, where using the wrong domain tag can yield ``catastrophic'' drops in performance.
+This requirement can be measured by translating test texts for domain $A$ with all erroneous or random domain tags and measure the correlated decrease in performance.
 
@@
 Several situations need be considered: when the domain is seen in training, then using automatically predicted domain labels should not be much worse than using the correct one.
-For test documents from other domains, the same procedure can be used, and should outperform the use of a single mixed-domain system.\fyTodo{Systems react to unknown domains}
+This procedure can also be used test documents from unknown domains (zero-shot transfer).
+A good MD system should there outperform the default baseline that merges all available data.\fyDone{Systems react to unknown domains}
 % These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.
@@
 
-\subsection{Effectiveness in handling a large number of domains \label{ssec:scaling}}
+\subsection{Scaling to a large number of domains \label{ssec:scaling}}
 
@@
 This argument is all the more sensible as the number of domains increases, making the optimization of many individual systems both ineffective and undesirable.
-For lack of having access to corpora containing very large sets (eg.\ in the order of 100-1000) domains, we experiment with automatically learned domains.\fyTodo{considering a varying number of clusters.}
+For lack of having access to corpora containing very large sets (eg.\ in the order of 100-1000) domains, we experiment with automatically learned domains.\fyFuture{considering a varying number of clusters.}
 
@@
 
-We experiment with translation from English into French and use data initially originating from 6~domains, corresponding to texts from the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
+We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
@@
 We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\\footnote{We use multibleu}fyDone{A word about meta-parameter settings}.
-Significance tests are performed using the bootstrap resampling method of \newcite{Koehn04statistical} and implementation of compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
 We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
@@
 
-\begin{table}[h]
+\begin{table*}[htbp]
   \centering
@@
 \label{tab:Corpora}
-\end{table}
+\end{table*}
 
-We measure the distance between domains using the $\mathcal{H}$-Divergence introduced in \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains.
-Formally, given test sets of size $m$ for domains $A$ and $B$, and $h(x)$ a trained domain predictor, $\mathcal{H}(A,B)$ is computed as:
-$$
-\mathcal{H}(A,B) = 2(1 - [\frac{1}{m} \sum_{x:h(x) = B} \mathbb{I}( x \in A) + \frac{1}{m} \sum_{x: h(x) = A} \mathbb{I}(x \in B)]),
-$$
+We measure the distance between domains using the $\mathcal{H}$-Divergence \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains.
+%Formally, given test sets of size $m$ for domains $A$ and $B$, and $h(x)$ a trained domain predictor, $\mathcal{H}(A,B)$ is computed %as:
+%$$
+%\mathcal{H}(A,B) = 2(1 - [\frac{1}{m} \sum_{x:h(x) = B} \mathbb{I}( x \in A) + \frac{1}{m} \sum_{x: h(x) = A} \mathbb{I}(x \in B)]),
+% $$
 % where $\mathbb{I}$ is the indicator function.
-The domain discriminator is a SVM independently trained for each pair of domains, with sentence representations derived via mean pooling from the source side representation of the generic Transformer model.
-We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well defined and separated from the all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
+Our discriminator is a SVM independently trained for each pair of domains, with sentence representations derived via mean pooling from the source side representation of the generic Transformer model.
+We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well separated from the all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
 
@@
 
-Baselines correspond to standard practices for multi-domain systems.\footnote{We however omit domain-specific systems trained with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using the Transformer architecture of \shortcite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
+Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
 
@@
 We develop two versions of this system, one where domain unbalance reflects the distribution of our parallel data (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
-The former is the best option when the train mixture $\mathcal{D}^s$ is expected to be also observed in testing; the latter should be used when the test distribution is uniform across domains.
-Accordingly, we report two aggrete scores: a weighted average reflecting the default distributio, and an unweighted average (all test domains are equally important).
+The former is the best option when the train mixture $\mathcal{D}^s$ is also observed in testing; the latter should be used when the test distribution is uniform across domains.
+Accordingly, we report two aggrate scores: a weighted average reflecting the default distribution, and an unweighted average (test domains are equally important).
 %   As explained above, depending whether MDMT are expected to perform specially well on large domains, or to treat all domains equally, one or the other training distribution should be prefered.\jcDone{All systems cand be learned following both Bal and Nat.
@@
 is wAVG defined somewhere?}
-\item models fine-tuned \cite{Luong15stanford,Freitag16fast}\jcDone{i would use Luong and Manning 2015} separately on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
-In our setting, we again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a simple adaptation module that is added (with residual connexions) on top of every Transformer layer (\system{FT-Block}).
+\item fine-tuned models \cite{Luong15stanford,Freitag16fast}\jcDone{i would use Luong and Manning 2015}, based on the \system{Mixed-Nat} system, further trained on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
+We again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a simple adaptation module that is added (with residual connexions) on top of every Transformer layer (\system{FT-Block}).
 \end{itemize}
@@
 \item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
-\item the three proposals of \cite{Britz17mixing}: \system{TTM}, another tag-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
-Training is performed with reference tags and inference is performed with predicted tags, just like for any other target words; \system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations;
+\item the three proposals of \cite{Britz17mixing}: \system{TTM}, another feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
+Training is performed with reference tags and inference is performed with predicted tags, just like for any other target words; \system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
+Note that these three methods only use domain information during training;
 \item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});
-\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
+\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach, two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly, improved results.}\fyTodo{Check this}
 \item an original, multi-domain version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connexions make it possible to by-pass this module.
@@
     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8 &\\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & 42.3 & 53.3 & 20.51\\ \hline
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & 42.3 & 53.3 & 20.5\\ \hline
  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
@@
     \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   & \SW{21.7} \\
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7       &  23.43 \\
+    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7       &  23.4 \\
     \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8 & 22.1 \\ 
-    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 & 22.64\\
-    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 40.7 & 47.5 & 23.34 \\
-    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 & 20.43 \\
+    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 & 22.6\\
+    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 40.7 & 47.5 & 23.3 \\
+    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 & 20.4 \\
     \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.1 & 51.9  & \SW{21.2} \\
@@
 
-\begin{table}
+\begin{table*}
   \centering
@@
   \label{tab:redomains}
-\end{table}
+\end{table*}
 

### 2020-04-23 2020/coling2020/multidomain.tex
fix table, more style and typos

@@
 \item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
-\item the three proposals of \cite{Britz17mixing}: \system{TTM}, another feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
-Training is performed with reference tags and inference is performed with predicted tags, just like for any other target words; \system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
-Note that these three methods only use domain information during training;
 \item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});
+\item the three proposals of \cite{Britz17mixing}.
+\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
+Training is performed with reference tags and inference is performed with predicted tags, just like for regular target words.
+\system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
+Note that these methods only use domain tags during training.
 \item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach, two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly, improved results.}\fyTodo{Check this}
-\item an original, multi-domain version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connexions make it possible to by-pass this module.
+\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connexions make it possible to by-pass this module.
   % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
@@
 The best contender overall is \system{MDL Res}, which significantly improves over \system{Mixed-Nat} for three domains, and is the only one to clearly fulfill [P1-LAB].
-The use of an decidated adaptation component proves here better than feature-based strategies.
+The use of a dedicated adaptation component proves here better than feature-based strategies.
 The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}), for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.
 
-In this study, all systems fail to outperform fine-tuning, sometimes by a wide margin: this is observed for an ``isolated'' domain such as \domain{rel}.
-This might be expected, since all domains are clearly separated (cf.\ Section~\ref{ssec:corpora}) and are hardly helping each other.
+In this study, all systems fail to outperform fine-tuning, sometimes by a wide margin, as is the case for an ``isolated'' domain such as \domain{rel}.
+This might be expected, since all domains are well separated (cf.\ Section~\ref{ssec:corpora}) and are hardly helping each other.
 In this situation, MDMT systems should make sure that each domain is allocated a sufficient number of specific parameters, so as to close the gap with fine-tuning.
@@
     \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   & \SW{21.7} \\
+    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8 & 22.1 \\ 
     \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7       &  23.4 \\
-    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8 & 22.1 \\ 
     \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 & 22.6\\
@@
 
-Table~\ref{tab:redomains} summaries the results of four experiments where we artificially redefine the limits of domains, with the aim to challenge requirements [P2-TUN], [P3-HET], and [P4-ERR].
-In the three \emph{split} experiments, we randomly split a domain corpus in two part: a good MDMT system should detect that these two ``domains'' should be mutually beneficial, and should be hardly affected by this change wrt.\ the baseline scenario.
-This is the situation where we should expect to MDMT even surpass fine-tuning separately on each of these dummy domain, as MDMT exploits all data, where fine-tuning only sees a subpart.
+Table~\ref{tab:redomains} summaries the results of four experiments where we artificially redefine the boundaries of domains, with the aim to challenge requirements [P2-TUN], [P3-HET], and [P4-ERR].
+In first three, we randomly \emph{split} one corpus in two parts and proceed as if this corresponded to two actual domains.
+A MD system should detect that these two pseudo-domains should be mutually beneficial, and should be hardly affected by this change with respect to the baseline scenario (no split).
+In this situation, we should expect MDMT to even surpass fine-tuning separately on each of these dummy domain, as MDMT exploits all data, where fine-tuning focuses only on a subpart.
+In testing, we decode the test set twice, once with each pseudo-domain tag.
+Note that this makes no difference for \system{TMM}, \system{ADM}, \system{ADM} and \system{WDCMT}, which do not use domain tags in testing.
+In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains.
 
-In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains.
 \fyTodo{Prepare table for the wrong domain analysis (see also last col.
@@
 \begin{table*}
-  \centering
+  \centering% \small
   \begin{tabular}{|p{2.5cm}|*{9}{r|}} \hline
@@
     & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \\
-    \system{FT-Full}      & -0.13 & -0.6 & -1.51& -0.25& -2.28 & -5.08 &\SW{-1.6} & \SW{-1.4}& \\
-    \system{DC-Tag}     & -0.16 & -0.27& +\SB{0.13}$^+$& +0.2& -2.25 & -2.92 & -0.5 & -0.4 & \\
-    \system{DC-Feat}    & -0.47 & 0.0 & +\SB{0.28} $^+$ & +0.28 & +0.32 & +0.33 & 0.3 & 0.1 & \\
-    \system{TTM}          & -0.23 &  & -0.2 & & -0.3 & &  0.0 & -0.3 &  \\
-    \system{LDC-Tag} & +0.08 & +0.15 & +0.36 & +0.28 & -0.04 & -0.01 &  0.0 & 0.1& \\ 
+    \system{FT-Full}      & -0.1 & -0.6 & -1.5& -0.2& -2.3 & -5.1 &\SW{-1.6} & \SW{-1.4}& \\
+    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}$^+$& +0.2& -2.2 & -2.9 & -0.5 & -0.4 & \\
+    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3} $^+$ & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & \\
+    \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \\ 
+    \system{TTM} (!)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 &  \\
+%    \system{MDL Gated} & & & & & & & & &  \\
+    \system{ADM} (!)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & \\
+    \system{DM} (!)           & -0.3   & -0.3  & +0.4 & +0.4 & +0.3 & +0.3 & +0.9 & +0.1 & \\
+    \system{WDCMT} (!)     &-0.9    & -0.9  & -1.1 & -1.1  & -2.2 & -2.2  & & & \\
     \system{MDL Res}   & -0.2   & -0.07 & +\SB{0.18}$^+$ &+0.0 & -0.89 & -0.89 & +\SB{0.7} & -0.35 & \\
-%    \system{MDL Gated} & & & & & & & & &  \\
-    \system{ADM} & +0.6 & & +0.4 & & +0.44 & &  0.1 & -0.4 & \\
-    \system{DM}    & -0.35 & & +0.4 & & +0.33 & & 0.9 & 0.1 & \\
-    \system{WDC}  & -0.95 & & -1.1 & & -2.25 & & & & \\
     \hline
@@
   \caption{Translation performance with variable domain definitions.
-For the Split experiments, we only report the loss in performance for the \domain{med} test set, (*) denote significant losses when domains are split or merged; (+) denotes a significant improvement over \system{Full FT}; \fyTodo{How about wrong ?}}
+For the Split experiments, we only report the loss in performance for the concerned test set (\domain{med} or \domain{law}), (*) denote significant losses when domains are split or merged; (+) denotes a significant improvement over \system{Full FT}; \fyTodo{Check significance}\fyTodo{How about wrong ?}}
   \label{tab:redomains}
@@
 Starting with the existing MDMT systems of Section~\ref{ssec:rawperformance}, we introduce an additional domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is again of utmost importance for achieving optimal performance: since our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy where.
-This is of course detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation}.
-We contrast this approach with retraining all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.
-We expect that a good MDMT system should not be significantly affected by the addition of a new domain, and reach about the same performance as if the training included this domain from scratch.
+This is of course detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with retraining all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.
+We expect that a proper MDMT system should not be significantly affected by the addition of a new domain, and reach about the same performance as if the training included this domain from scratch.
 Note that from a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{TTC} or \system{DC-Feat} where new domains merely add new labels.

### 2020-04-23 2020/coling2020/multidomain.tex
reviewd till section 5.3

@@
   \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}|r|} \hline
+  \begin{tabular}{|p{3cm}|*{8}{r|}|} \hline
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} & \multicolumn{1}{c|}{\domain{news}} \\ \hline
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 & 23.5\\
-    \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 & \\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8 &\\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & 42.3 & 53.3 & 20.5\\ \hline
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
+    \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 \\ %& \\
+    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8\\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & 42.3 & 53.3 \\ \hline % & 20.5\\ \hline
  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.7 & 50.1     & 21.8 \\
-    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   & \SW{21.7} \\
-    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8 & 22.1 \\ 
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7       &  23.4 \\
-    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 & 22.6\\
-    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 40.7 & 47.5 & 23.3 \\
-    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 & 20.4 \\
-    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.1 & 51.9  & \SW{21.2} \\
-    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7&\\
+    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.7 & 50.1    \\%    & 21.8 \\
+    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   \\% & \SW{21.7} \\
+    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8          \\% & 22.1 \\ 
+    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7     \\% &  23.4 \\
+    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 \\ % & 22.6\\
+    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 40.7 & 47.5 \\% & 23.3 \\
+    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 \\ % & 20.4 \\
+    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.1 & 51.9 \\%  & \SW{21.2} \\
+    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
 %    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3&\\
@@
 In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains.
-
-\fyTodo{Prepare table for the wrong domain analysis (see also last col.
-of the first table)}In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.
+We then translate the two corresponding test sets with the same (merged) system.
 
@@
   \centering% \small
-  \begin{tabular}{|p{2.5cm}|*{9}{r|}} \hline
+  \begin{tabular}{|p{1.8cm}|*{10}{r|}} \hline
     % &&&&&& \\
-    \hfill Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & Wrong \\ % \hline
-     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{law} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} & domain \\ \hline
-    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \\
-    \system{FT-Full}      & -0.1 & -0.6 & -1.5& -0.2& -2.3 & -5.1 &\SW{-1.6} & \SW{-1.4}& \\
-    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}$^+$& +0.2& -2.2 & -2.9 & -0.5 & -0.4 & \\
-    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3} $^+$ & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & \\
-    \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \\ 
-    \system{TTM} (!)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 &  \\
+    \hfill Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & \multicolumn{2}{c|}{Wrong} \\ % \hline
+     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{law} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} &  \multicolumn{1}{c|}{old} &  \multicolumn{1}{c|}{new}\\ \hline
+    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
+    \system{FT-Full}      & -0.1 & -0.6 & -1.5& -0.2& -2.3 & -5.1 &\SW{-1.6} & \SW{-1.4}& &\\
+    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}$^+$& +0.2& -2.2 & -2.9 & -0.5 & -0.4 & &-1.7\\
+    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3} $^+$ & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & &\SW{-1.8}\\
+    \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& &-1.4\\ 
+    \system{TTM} (!)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & & -0.1\\
 %    \system{MDL Gated} & & & & & & & & &  \\
-    \system{ADM} (!)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & \\
-    \system{DM} (!)           & -0.3   & -0.3  & +0.4 & +0.4 & +0.3 & +0.3 & +0.9 & +0.1 & \\
-    \system{WDCMT} (!)     &-0.9    & -0.9  & -1.1 & -1.1  & -2.2 & -2.2  & & & \\
-    \system{MDL Res}   & -0.2   & -0.07 & +\SB{0.18}$^+$ &+0.0 & -0.89 & -0.89 & +\SB{0.7} & -0.35 & \\
+    \system{DM} (!)           & -0.3   & -0.3  & +0.4 & +0.4 & +0.3 & +0.3 & +0.9 & +0.1 & &-0.9\\
+    \system{ADM} (!)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & &-0.2\\
+    \system{WDCMT} (!)     &-0.9    & -0.9  & -1.1 & -1.1  & -2.2 & -2.2  & & & ***\\
+    \system{MDL Res}   & -0.2   & -0.07 & +\SB{0.18}$^+$ &+0.0 & -0.89 & -0.89 & +\SB{0.7} & -0.35 & & \SW{-1.3}\\
     \hline
@@
 
-Our main findings can be summarized as follows: for the split experiments, we see a general loss with respect to the situation were domains are kept a whole, except in rare cases\fyTodo{To be completed}.
-As expected, \system{FT} incurs a large loss, especially for the small split.
-Findings so far:
-\begin{itemize}
-\item Split UFAL half / half: domain-tag and Gated-res are OK, all systems loose wrt baseline; 
-\item Split UFAL 0,75 / 0.25: same thing, and we even see improvements for domain tags - they are small though
-\item Split Law : only domain feature works ? all the rest have losses, from large to small
-\item Merge domain: bad for many, OK for TTM (as the domain is learnt this is not so bad ?); bad for domain tag, domain feature is again OK ?
-\end{itemize}
-\fyDone{We do need significance testing (a) each system against itself (non-split); (b) each system agains FT in split  condition}
+Our findings can be summarized as follows.
+For the split experiments, we see small variations that can be positive or negative compared to the situation were domains are kept a whole, bu these are hardly significative.
+All systems show here some robustness with respect to fuzzy domain boundaries; this is mostly notable for \system{ADM} suggesting that when domain are close, the neutralization of domain diferences is effective.
+In contrary, \system{FT-full} incurs large losses across the board, especially for the small data condition.
+Even in this very favourable case, very few MDMT systems are able to significantly outperform FT, and this is only observed for the small portion of the \domain{med} domain.
+The merge condition is hardly different, with large losses for \system{FT-full}, and small variations for all systems.
+We even observe improvements with respect to the situation where we use actual domains.\fyDone{More comments ?}
 
-It is finally interesting to note that all systems are quite robust to domain merge: we even see some small gains in performance (eg.\ for \system{DM}).
-While robustness is generally a good thing, it is also a sign of a poor overall usage of domain labels.
-\fyTodo{To be completed}
+\subsubsection{Handling wrong or unknown domains \label{sssec:unknowns}}
+\fyTodo{Prepare table for the wrong domain analysis (see also last col.
+of the first table)}In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.
 
-\subsubsection{Handling unknown domains \label{sssec:unknowns}}
-To evaluate requirement [P5-UNK], we translate sentences from a domain unseen in training (\domain{news}).
+We consider two situations: test data from the domains seen in training, but with an erroneous tag (random domain);\fyTodo{Comments?}
+
+We now turn to requirement [P5-UNK], and translate test sentences from a domain unseen in training (\domain{news}).
 For each test sentence, we automatically predict the most likely domain tag, and use it for decoding.
-Results are in the rightmost column of Table~\ref{tab:performance}.\fyTodo{To be completed}
+Results are in the rightmost column of Table~\ref{tab:performance}.
+In this configuration, all systems do worse than the baseline, sometimes by a large margin.\fyTodo{Check significance, To be completed}
 
 \subsubsection{Coping with a growing number of domains}
-Another series of experiments evaluate the ability to handle additional domains (requirement [P6-DYN]) as follows.
-Starting with the existing MDMT systems of Section~\ref{ssec:rawperformance}, we introduce an additional domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is again of utmost importance for achieving optimal performance: since our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy where.
+Another series of experiments evaluate the ability to dynamically handle suplementary domains (requirement [P6-DYN]) as follows.
+Starting with the existing MD systems of Section~\ref{ssec:rawperformance}, we introduce an extra domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is again of utmost importance for achieving optimal performance: since our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy based on the new training distribution.
 This is of course detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with retraining all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.
 We expect that a proper MDMT system should not be significantly affected by the addition of a new domain, and reach about the same performance as if the training included this domain from scratch.
-Note that from a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{TTC} or \system{DC-Feat} where new domains merely add new labels.
-It is less simple for \system{LDC} and \system{MDL~Res}\fyTodo{Complete effects of new domain}, for which the number of possible domains is built in the architecture and has to be anticipated from the begining.
-This makes another difference between domain-bounded MT-systems, for which the number of possible domains is upper bounded and truely open-domain systems.% , which seamlessly integrate additional domains.
+Note that from a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{DC-Feat} or \system{TTM} where new domains merely add new labels.
+It is less simple for \system{LDC-Tag} and \system{MDL~Res}\fyDone{Complete effects of new domain}, for which the number of possible domains is built in the architecture and has to be anticipated from the begining.
+This makes another difference between domain-bounded MT-systems, for which the number of possible domains is limited and truely open-domain systems.% , which seamlessly integrate additional domains.
 

### 2020-04-23 2020/coling2020/multidomain.tex
one more comment

@@
 The multi-domain training regime is more the norm than the exception for natural language processing \cite{Dredze08online,Finkel09hierarchical}, and the design of multi-domain systems has been proposed for many language processing tasks.
-We focus here exclusively on MD machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection / weighting, adversarial training, etc) may have been studied in other contexts.
+We focus here exclusively on MD machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection / weighting, adversarial training, etc) have been studied in other contexts.
 
@@
 
-Multi-domain translation was already proposed for statistical MT (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}).
-Two main strategies were considered: instance-based, involving a measure of similarities between train and test domains; feature-based, where domain labels give rise to additional features.
+Multi-domain translation was already proposed for statistical MT (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}).\fyTodo{Add also topic models refs} Two main strategies were considered: instance-based, involving a measure of similarities between train and test domains; feature-based, where domain labels give rise to additional features.
 
-This latter strategy has also been widely used in MNT: the authors of \cite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token or in the form of additional domain-feature associated to each word.
+This latter strategy has also been widely used in NMT: the authors of \cite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token or in the form of additional domain-feature associated to each word.
 These results are reproduced in \cite{Tars18multidomain}, which also considers, as we do, automatically induced domain tags.

### 2020-04-24 2020/coling2020/multidomain.tex
more work on continuous learning and formatting, we still lack results

@@
 \item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});
-\item the three proposals of \cite{Britz17mixing}.
+\item the three proposals of \newcite{Britz17mixing}.
 \system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
-Training is performed with reference tags and inference is performed with predicted tags, just like for regular target words.
+Training uses reference tags and inference is performed with predicted tags, just like for regular target words.
 \system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
-Note that these methods only use domain tags during training.
+These methods thus only use domain tags in training.
 \item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach, two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly, improved results.}\fyTodo{Check this}
@@
 We report the BLEU score for each domain, as well as domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages.
-Boldface denotes significant differences with respect to the \system{Mix-Nat} system.
-Scores for the \domain{news} domain correspond to the out-of-domain scenario (see Section~\ref{ssec:redomains}).}
+Boldface denotes significant differences with respect to the \system{Mix-Nat} system.}
   \label{tab:performance}
+  \fyTodo{Fill the table with all results, including a comparison with *RNNs* for wcmd}
+  \fyTodo{Do we have significancy tests for averages?}
 \end{table*}
-\fyTodo{Fill the table with all results, including a comparison with *RNNs* for wcmd}\fyTodo{Do we have significancy tests for averages?}
-\fyDone{A new column for the News domain with predicted domains (except for the generic case).}
 
@@
 In testing, we decode the test set twice, once with each pseudo-domain tag.
-Note that this makes no difference for \system{TMM}, \system{ADM}, \system{ADM} and \system{WDCMT}, which do not use domain tags in testing.
+This makes no difference for \system{TMM}, \system{ADM}, \system{ADM} and \system{WDCMT}, which do not use domain tags in testing.
 In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains.
@@
 For each test sentence, we automatically predict the most likely domain tag, and use it for decoding.
-Results are in the rightmost column of Table~\ref{tab:performance}.
-In this configuration, all systems do worse than the baseline, sometimes by a large margin.\fyTodo{Check significance, To be completed}
+Results are in the rightmost column of Table~\ref{tab:redomains}.
+In this configuration, all systems do worse than the \system{Mixed-Nat} baseline (Bleu=23.5), sometimes by a large margin.\fyTodo{Check significance, To be completed}
 
-\subsubsection{Coping with a growing number of domains}
+\subsubsection{Handling growing numbers of domains}
 Another series of experiments evaluate the ability to dynamically handle suplementary domains (requirement [P6-DYN]) as follows.
 Starting with the existing MD systems of Section~\ref{ssec:rawperformance}, we introduce an extra domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is again of utmost importance for achieving optimal performance: since our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy based on the new training distribution.
-This is of course detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with retraining all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.
-We expect that a proper MDMT system should not be significantly affected by the addition of a new domain, and reach about the same performance as if the training included this domain from scratch.
-Note that from a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{DC-Feat} or \system{TTM} where new domains merely add new labels.
-It is less simple for \system{LDC-Tag} and \system{MDL~Res}\fyDone{Complete effects of new domain}, for which the number of possible domains is built in the architecture and has to be anticipated from the begining.
-This makes another difference between domain-bounded MT-systems, for which the number of possible domains is limited and truely open-domain systems.% , which seamlessly integrate additional domains.
+This is of course detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with training all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.
+We expect that a proper MDMT system should not be significantly affected by the addition of a new domain, and reach about the same performance as if training with this domain from scratch.
+Note that from a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{DC-Feat} or \system{TTM} and the like, for which new domains merely add new labels.
+It is less simple for \system{LDC-Tag} and \system{MDL~Res}\fyDone{Complete effects of new domain}, for which the number of possible domains is built in the architecture and has to be anticipated from the beginning.
+This makes another difference between domain-bounded MT-systems, for which the number of domains is limited and truly open-domain systems.% , which seamlessly integrate additional domains.
 
 \begin{table*}
-  \begin{tabular}{|p{3cm}|*{9}{r|}} \hline
+  \begin{tabular}{|p{2.5cm}|*{9}{r|}} \hline \footnotesize
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
@@
     \system{FT-Full}       & 37.7  & 59.2   & 54.5   & 34.0  & 46.8  & 90.8  & **** & 42.8 & 53.8 \\ \hline\hline
-     \system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9 \\
-                                    & 36.9 & 53.7   & 49.1    & 33.7  &  42.8  & 74.1 & 28.9 &\\
-    \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5 \\
-                                   & 38.0  & 54.0   & 48.8    & 33.5  &  42.7  &73.2  & 28.3& \\
-    \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.1  & 29.1 \\
-                                   & 36.8  & 53.4   & 47.9    & 32.4  &  42.5  &72.8  & 28.4& \\
-    \system{LDC-Tag}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 \\
-    continual                 &&&&&& \\
-    \system{DM}            &36.0 &51.3&46.8&31.8&39.8&65.7&27.0 \\
-                                  &36.5&51.6&47.3&31.7&40.3&65.7&25.8& \\
-    \system{ADM}          &36.6&54.2&49.1&32.9&42.1&75.7&28.7 \\
-                                  &  36.9 &53.5&48.3&32.7&41.7&70.7&26.8& \\
-    \system{WDCMT}     &34.8  &51.8&47.8&30.7&37.3&49.3&25.9 \\
-    continual                 &&&&&&&& \\
-    \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & & \\
-     continual                & 37.45 & 56.09 & 51.11 & 34.04  &  44.29 & 87.32 & 28.35 & \\
+
+     \system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9  & & 50.1\\[-2pt]
+%                   & 37.2 & 54.7   & 49.9    & 34.2  &  43.0  & 79.9 & \SW{28.1}  & &\\
+                   & -0.2 & -0.1  & -0.1 & -0.6 & -0.9 & +0.3 & \SW{-0.8} & & -0.3\\
+
+    \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5  & & 49.9\\[-2pt]
+%                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
+                   & +0.3  & -0.1   & -0.6    & \SW{-1.1}  &  \SW{-1.3}  &\SW{-3.5}  & \SW{-1.4}  & & -1.2\\
+
+    \system{LDC-Tag}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & & 46.3 \\[-2pt]
+                                   &&&&&&&& \\
+
+    \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.2  & 29.1 & & 49.4\\[-2pt]
+%                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
+                    & -0.3  & -0.3  & -0.5    & \SW{-1.1}  &  -1.0  & \SW{-4.0}  & \SW{-1.4}& &-1.2\\
+
+    \system{DM}            &36.0 &51.3&46.8&31.8&39.8&65.7&27.0 & & 45.2\\[-2pt]
+%                   &36.5&51.7&47.4&31.7&40.3&65.7&\SW{25.8}& & \\
+                   &+0.6&+0.4&+0.6&-0.1&0.5&0.0&\SW{-1.2}& & +0.3\\
+
+    \system{ADM}          &36.6&54.2&49.1&32.9&42.1&75.7&28.7 & &4.4 \\[-2pt]
+%                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
+                   & +0.3 & -0.8 & -0.8 & -0.2 & -0.4 & \SW{-5.0} & \SW{-1.9} & & -1.1\\
+
+    \system{WDCMT}     &34.8  &51.8&47.8&30.7&37.3&49.3&25.9 & &\\[-2pt]
+                    &&&&&&&& \\
+
+    \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & & 51.8 \\[-2pt]
+%                      & 37.5 & 56.1 & 51.1 & 34.0  &  44.3 & 87.3 & \SW{28.4} & & 51.7 & \\
+                        & -0.2 & +0.5 & 0.0 & -0.4  &  -0.2 & -0.2 & \SW{-0.8} & & -0.1 \\
      \hline
   \end{tabular}
-  \caption{Ability to handle new a domain: BLEU performance in ``resume training'' mode compared to a complete training session.}
+  \caption{Ability to handle new a domain: BLEU performance in ``resume training'' mode compared to a complete training session.
+Averages only take into account the six initial domains (\domain{News} excluded).}
   \label{tab:warmrestart}
+  \fyTodo{Fill the table, significancy testing against initial learning condition, for each domain}
+  \fyTodo{Need correct number / differences for Mixed-Nat}
+  \fyDone{averages include one more domain ?}
 \end{table*}
-\fyTodo{Fill the table, significancy testing against full learning condition, for each domain}
 
-We can first compare the results of coldstart training with six (Table~\ref{tab:performance}) or seven domains (Table~\label{tab:warmrestart}): the most striking observation is that the extra training data is detrimental for all systems and all domains,\fyTodo{is this significant ?}.
-Increasing the heterogeneity of the mix hurts performance in all domains, again the sign that the domain are quite well separated.
-Comparing now the coldstart with the warmstart scenario, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training can negatively impact the performance on other domains - this happens notably for \system{DC-Tag} and \system{TTM}.
-In this setting, and, XXX\fyTodo{To be completed} and especially \system{MDL Res} show the smaller average loss, with the latter achieving the better balance of total training cost and average translation performance over all systems.
+We can first compare the results of coldstart training with six (Table~\ref{tab:performance}) or seven domains (Table~\ref{tab:warmrestart}): the most striking observation is that the extra training data is detrimental for all systems and all domains,\fyTodo{is this significant ?}.
+Increasing the heterogeneity of the mix hurts performance across the board.
+This confirms that adding more out-of-domain data is not always helping, especially when domains are well separated.
+Comparing now the coldstart with the warmstart scenario, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training can negatively impact the performance on other domains - this happens notably for \system{DC-Tag}, \system{TTM} and \system{ADM}.
+In this setting \system{MDL Res} and \system{DM} show the smaller average loss, with the former achieving the better balance of total training cost and average translation performance of all systems.
 
@@
 We are thus in a position to reiterate in more realistic conditions the measurements of section~\label{ssec:}, and to test whether multi-domain systems can effectively take advantage from the cross-domain similarities, and to eventually perform better that fine-tuning.
+Results are in table~\ref{tab:subdomains}.
 
-Results are in table~\ref{tab:subdomains}.\fyTodo{Fill the tables with automatic and existing domains - we will have significancy results for the existing domains, for the other may be for larger classes we will have enough test instances?}
+\begin{table*}
+  \centering
+  \fyDone{Fix column size}
+  \begin{tabular}{|p{3cm}|*{8}{r|}|} \hline
+%     &&&&&& \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
+    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8\\
+ %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
+    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.7 & 50.1    \\%    & 21.8 \\
+    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   \\% & \SW{21.7} \\
+    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8          \\% & 22.1 \\ 
+    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7     \\% &  23.4 \\
+    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 \\ % & 22.6\\
+    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 40.7 & 47.5 \\% & 23.3 \\
+    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 \\ % & 20.4 \\
+    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.1 & 51.9 \\%  & \SW{21.2} \\
+     \hline
+  \end{tabular}
+  \caption{Translation performance with automatic domains, computed with the original test sets.}
+  \label{tab:subdomains}
+  \fyTodo{Fill the table with correct results,}
+\end{table*}
 
@@
 
-Multi-domain translation was already proposed for statistical MT (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}).\fyTodo{Add also topic models refs} Two main strategies were considered: instance-based, involving a measure of similarities between train and test domains; feature-based, where domain labels give rise to additional features.
+Multi-domain translation was already proposed for statistical MT, either considering as we do multiple sources of training data (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}), or topically heterogeneous domains \cite{Eidelman12topic,Hasler14dynamic-topic}.\fyDone{Add also topic models refs} Two main strategies were considered: instance-based, involving a measure of similarities between train and test domains; feature-based, where domain/topic labels give rise to additional features.
 

### 2020-04-28 2020/coling2020/multidomain.tex
new todo for tests
add more results and comments

@@
     \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 \\ %& \\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8\\
+    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.7 & 53.8\\
    \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & 42.3 & 53.3 \\ \hline % & 20.5\\ \hline
  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.7 & 50.1    \\%    & 21.8 \\
+    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.6 & 50.1    \\%    & 21.8 \\
     \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   \\% & \SW{21.7} \\
-    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8          \\% & 22.1 \\ 
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7     \\% &  23.4 \\
-    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 \\ % & 22.6\\
-    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 40.7 & 47.5 \\% & 23.3 \\
-    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 \\ % & 20.4 \\
-    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.1 & 51.9 \\%  & \SW{21.2} \\
+    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} &40.9 & 49.8          \\% & 22.1 \\ 
+    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.0 & 49.7     \\% &  23.4 \\
+    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 38.1 & 43.4 \\ % & 22.6\\
+    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 38.9 & 47.5 \\% & 23.3 \\
+    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & 39.0 & 41.5 \\ % & 20.4 \\
+    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & 51.9 \\%  & \SW{21.2} \\
     \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
@@
   \fyTodo{Do we have significancy tests for averages?}
+  \fyTodo{Add tests wrt to RNN}
 \end{table*}
@@
 %                   & 37.2 & 54.7   & 49.9    & 34.2  &  43.0  & 79.9 & \SW{28.1}  & &\\
-                   & -0.2 & -0.1  & -0.1 & -0.6 & -0.9 & +0.3 & \SW{-0.8} & & -0.3\\
+                   & -0.2 & -0.1  & -0.1 & -0.6 & -0.9 & +0.3 & \SW{-0.8} & -0.2 & -0.3\\
 
@@
 %                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
-                   & +0.3  & -0.1   & -0.6    & \SW{-1.1}  &  \SW{-1.3}  &\SW{-3.5}  & \SW{-1.4}  & & -1.2\\
+                   & +0.3  & -0.1   & -0.6    & \SW{-1.1}  &  \SW{-1.3}  &\SW{-3.5}  & \SW{-1.4}  & -0.1 & -1.2\\
 
@@
 %                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
-                    & -0.3  & -0.3  & -0.5    & \SW{-1.1}  &  -1.0  & \SW{-4.0}  & \SW{-1.4}& &-1.2\\
+                    & -0.3  & -0.3  & -0.5    & \SW{-1.1}  &  -1.0  & \SW{-4.0}  & \SW{-1.4}& -0.5 &-1.2\\
 
@@
 %                   &36.5&51.7&47.4&31.7&40.3&65.7&\SW{25.8}& & \\
-                   &+0.6&+0.4&+0.6&-0.1&0.5&0.0&\SW{-1.2}& & +0.3\\
+                   &+0.6&+0.4&+0.6&-0.1&0.5&0.0&\SW{-1.2}& +0.5 & +0.3\\
 
@@
 %                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
-                   & +0.3 & -0.8 & -0.8 & -0.2 & -0.4 & \SW{-5.0} & \SW{-1.9} & & -1.1\\
+                   & +0.3 & -0.8 & -0.8 & -0.2 & -0.4 & \SW{-5.0} & \SW{-1.9} & -0.2 & -1.1\\
 
@@
 %                      & 37.5 & 56.1 & 51.1 & 34.0  &  44.3 & 87.3 & \SW{28.4} & & 51.7 & \\
-                        & -0.2 & +0.5 & 0.0 & -0.4  &  -0.2 & -0.2 & \SW{-0.8} & & -0.1 \\
+                        & -0.2 & +0.5 & 0.0 & -0.4  &  -0.2 & -0.2 & \SW{-0.8} & -0.2 & -0.1 \\
      \hline
@@
   \fyTodo{Need correct number / differences for Mixed-Nat}
-  \fyDone{averages include one more domain ?}
+  \fyDone{averages include one more domain ? No for comparison}
 \end{table*}
@@
 \subsection{Automatic domains \label{ssec:autodomains}}
-In this section, we experiment with automatic domains, obtained by clustering sentences into $k=30$ classes using the k-means algorithm based on generic sentence representations obtained via mean pooling as in section~\ref{ssec:corpora}.
+In this section, we experiment with automatic domains, obtained by clustering sentences into $k=30$ classes using the k-means algorithm based on generic sentence representations obtained via mean pooling as in Section~\ref{ssec:corpora}.
 This allows us to evaluate requirement [P7-scale], training and testing our systems as if these domains were genuine.
-We also note that many of these classes are mere split of the large \domain{med}, while a few number of classes being mix of two existing domains.
-We are thus in a position to reiterate in more realistic conditions the measurements of section~\label{ssec:}, and to test whether multi-domain systems can effectively take advantage from the cross-domain similarities, and to eventually perform better that fine-tuning.
-Results are in table~\ref{tab:subdomains}.
+We also note that many of these classes are mere split of the large \domain{med}, while a fewer number of classes are mix of two existing domains.
+We are thus in a position to reiterate in more realistic conditions the measurements of Section~\label{ssec:} and to test whether multi-domain systems can effectively take advantage from the cross-domain similarities, and to eventually perform better that fine-tuning.
 
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8\\
+  %    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
+  %     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8\\
   %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.7 & 50.1    \\%    & 21.8 \\
-    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   \\% & \SW{21.7} \\
-    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8          \\% & 22.1 \\ 
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7     \\% &  23.4 \\
-    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 \\ % & 22.6\\
-    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 40.7 & 47.5 \\% & 23.3 \\
-    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 \\ % & 20.4 \\
-    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.1 & 51.9 \\%  & \SW{21.2} \\
+    \system{DC-Tag}       & 38.5 & \SW{54.0} & 49.0   & 33.6 & \SW{42.2} & \SW{76.7} & 49.0 & 41.6 \\%    & 21.8 \\
+    \system{DC-Feat}      & 37.3  & 54.2 & 49.3   & 33.6 & \SW{41.9} & \SW{75.8} & 48.3 & 40.7  \\% & \SW{21.7} \\
+    \system{LDC-Tag}     & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 48.3 & 39.1         \\% & 22.1 \\ 
+    \system{TTM}            & 37.4 & \SW{53.7} & 48.9 & 32.8 & 41.3 & \SW{75.8} & 48.3 & 40.7   \\% &  23.4 \\
+    \system{DM}             & 35.4 & 49.3  & 45.2 & 29.7 & 37.1 & \SW{60.0} & 42.8 & 37.8 \\ % & 22.6\\
+    \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 47.1 & 39.5\\% & 23.3 \\
+    \system{WDCMT}       & 36.0 & 53.8 & 48.5 & 31.0 & 39.0 & 63.5 & 45.3 & 39.1\\ % & 20.4 \\
+    \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 50.7& 41.2\\%  & \SW{21.2} \\
      \hline
@@
   \label{tab:subdomains}
-  \fyTodo{Fill the table with correct results,}
+  \fyDone{Fill the table with correct results,}
 \end{table*}
+
+Table~\ref{tab:subdomains} reports the effect of using automatic domain tags on the 6 domain-specific test sets: each sentence was first assigned to an automatic class, translated with the corresponding multi-domain system; aggregate numbers were then computed, as well as comparisons with the 6-domain scenarios.
+The results are very clear and confirm previous observations: even though some clusters are very close, the net effect is a decrease in performance for almost all systems and conditions.
+In this setting, the best MDMT of our pool (MDL-Res) is no longer able to surpass the \system{Mix-Nat} baseline.\fyTodo{Note about scaling?}\fyTodo{More comparison?}
 

### 2020-04-28 2020/coling2020/multidomain.tex
orthographic correction, and new results

--- 2020-04-24 2020/coling2020/multidomain.tex
@@
   When building machine translation systems, it is often the case that one need to make the best out of the heterogeneous set of parallel data in training, and to robustly handle inputs from unexpected domains in testing.
-This multi-domain scenario has attracted a lot of recent work, that fall under the general umbrella of transfert learning.
+This multi-domain scenario has attracted a lot of recent work, that fall under the general umbrella of transfer learning.
 In this study, we revisit multi-domain machine translation, with the aim to formulate the motivations for developing such systems and the associated expectations with respect to performance.
-Our experiments with a large sample of multi-domain systems show that most of these expectations are hardly met and suggest that further work is needed to better analyze the current behoviour of multi-domain systems and to make them fully hold their promises.
+Our experiments with a large sample of multi-domain systems show that most of these expectations are hardly met and suggest that further work is needed to better analyze the current behaviour of multi-domain systems and to make them fully hold their promises.
   \fyDone{Write abstract}
@@
     \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   \\% & \SW{21.7} \\
-    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8          \\% & 22.1 \\ 
+    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 40.9 & 49.8          \\% & 22.1 \\ 
     \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7     \\% &  23.4 \\
@@
     & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
-    \system{FT-Full}      & -0.1 & -0.6 & -1.5& -0.2& -2.3 & -5.1 &\SW{-1.6} & \SW{-1.4}& &\\
+    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5}& -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& &\\
     \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}$^+$& +0.2& -2.2 & -2.9 & -0.5 & -0.4 & &-1.7\\

### 2020-04-29 2020/coling2020/multidomain.tex
scores of wrong domain, scores of WDCNMT, scores of full finetune on news
fix a glitch in section 2

--- 2020-04-28 2020/coling2020/multidomain.tex
@@
 \newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}
-}\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
 \newcommand{\fyFuture}[1]{\done[FY]\Todo[FY:]{\textcolor{red}{#1}}}
@@
   \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}|} \hline
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
 %     &&&&&& \\
@@
 The results are very clear and confirm previous observations: even though some clusters are very close, the net effect is a decrease in performance for almost all systems and conditions.
-In this setting, the best MDMT of our pool (MDL-Res) is no longer able to surpass the \system{Mix-Nat} baseline.\fyTodo{Note about scaling?}\fyTodo{More comparison?}
+In this setting, the best MDMT in our pool (MDL-Res) is no longer able to surpass the \system{Mix-Nat} baseline.\fyTodo{Note about scaling?}\fyTodo{More comparison?}
 

### 2020-04-29 2020/coling2020/multidomain.tex

@@
 \fyDone{More comments when we have all the results}\fyDone{Importance of sharing and unsharing}
-
+36.81	53.85	47.21	30	35.74	60.24	39.2	44.0	21.06
 \begin{table*}
@@
     \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
+    \system{Rnn}  & 36.81 & 53.85 & 47.21 & 30 & 35.74 & 60.24  & 39.2  & 44.0 \\
     \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 \\ %& \\
@@
     \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 38.9 & 47.5 \\% & 23.3 \\
-    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & 39.0 & 41.5 \\ % & 20.4 \\
+    \system{WDCMT}       & 35.96 & 53.31 & 48.83 & 31.15 & 38.81 & 58.52 & 39.0 & 44.4 \\ % & 20.4 \\
     \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & 51.9 \\%  & \SW{21.2} \\
@@
     & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
-    \system{FT-Full}      & -0.1 & -0.6 & -1.5& -0.2& -2.3 & -5.1 &\SW{-1.6} & \SW{-1.4}& &\\
-    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}$^+$& +0.2& -2.2 & -2.9 & -0.5 & -0.4 & &-1.7\\
-    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3} $^+$ & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & &\SW{-1.8}\\
-    \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& &-1.4\\ 
+    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5}& -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& 32.55 & \SW{-3.28}\\
+    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}$^+$& +0.2& -2.2 & -2.9 & -0.5 & -0.4 & 35.92 & \SW{-1.7}\\
+    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3} $^+$ & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & 34.91 &\SW{-1.8}\\
+    \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& 37.01 & \SW{-1.4}\\ 
     \system{TTM} (!)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & & -0.1\\
@@
     \system{ADM} (!)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & &-0.2\\
-    \system{WDCMT} (!)     &-0.9    & -0.9  & -1.1 & -1.1  & -2.2 & -2.2  & & & ***\\
-    \system{MDL Res}   & -0.2   & -0.07 & +\SB{0.18}$^+$ &+0.0 & -0.89 & -0.89 & +\SB{0.7} & -0.35 & & \SW{-1.3}\\
+    \system{WDCMT} (!)     &-0.9    & -0.9  & -1.1 & -1.1  & -2.2 & -2.2  & -0.45 & -0.85 & *** & 0.19 \\
+    \system{MDL Res}   & -0.2   & -0.07 & +\SB{0.18}$^+$ &+0.0 & -0.89 & -0.89 & +\SB{0.7} & -0.35 & 31.78 & \SW{-1.3}\\
     \hline
@@
     \system{LDC-Tag}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & & 46.3 \\[-2pt]
-                                   &&&&&&&& \\
+                    & -0.65 &	0.46 &	-0.44 &	-0.58 &	0.49 &	3.8 &	-0.88 &	0.3 &	-0.4 \\
 

### 2020-04-29 2020/coling2020/multidomain.tex
complete table 3,4,6; missing WDCNMT scores in table 5

@@
 \newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}
-\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+}\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
 \newcommand{\fyFuture}[1]{\done[FY]\Todo[FY:]{\textcolor{red}{#1}}}
@@
 Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
-$h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$), but can differ from one domain to the other.
+$h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$, but can differ from one domain to the other.
 
@@
     \system{ADM} (!)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & &-0.2\\
-    \system{WDCMT} (!)     &-0.9    & -0.9  & -1.1 & -1.1  & -2.2 & -2.2  & -0.45 & -0.85 & *** & 0.19 \\
+    \system{WDCMT} (!)     & -0.03    & -0.03  & +0.2 & +0.2  & +0.83 & +0.83  & -0.45 & -0.85 & *** & 0.19 \\
     \system{MDL Res}   & -0.2   & -0.07 & +\SB{0.18}$^+$ &+0.0 & -0.89 & -0.89 & +\SB{0.7} & -0.35 & 31.78 & \SW{-1.3}\\
@@
 
-     \system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9  & & 50.1\\[-2pt]
+     \system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9  & 41.2 & 50.1\\[-2pt]
 %                   & 37.2 & 54.7   & 49.9    & 34.2  &  43.0  & 79.9 & \SW{28.1}  & &\\
@@
 
-    \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5  & & 49.9\\[-2pt]
+    \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5  & 41.4 & 49.9\\[-2pt]
 %                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
@@
 
-    \system{LDC-Tag}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & & 46.3 \\[-2pt]
-                    & -0.65 &	0.46 &	-0.44 &	-0.58 &	0.49 &	3.8 &	-0.88 &	0.3 &	-0.4 \\
+    \system{LDC-Tag}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
+                    & -0.65 &	+0.46 &	-0.44 &	-0.58 &	+0.49 &	+\SB{3.8} &	-0.88 &	+0.5 &	-0.4 \\
 
-    \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.2  & 29.1 & & 49.4\\[-2pt]
+    \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.2  & 29.1 & 41 & 49.4\\[-2pt]
 %                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
-                    & -0.3  & -0.3  & -0.5    & \SW{-1.1}  &  -1.0  & \SW{-4.0}  & \SW{-1.4}& -0.5 &-1.2\\
+                    & -0.3  & -0.3  & -0.5    & \SW{-1.1}  &  \SW{-1.0}  & \SW{-4.0}  & \SW{-1.4}& -0.5 & -1.2\\
 
-    \system{DM}            &36.0 &51.3&46.8&31.8&39.8&65.7&27.0 & & 45.2\\[-2pt]
+    \system{DM}            &36.0 &51.3&46.8&31.8&39.8&65.7&27.0 & 38.9 & 45.2\\[-2pt]
 %                   &36.5&51.7&47.4&31.7&40.3&65.7&\SW{25.8}& & \\
@@
 
-    \system{ADM}          &36.6&54.2&49.1&32.9&42.1&75.7&28.7 & &4.4 \\[-2pt]
+    \system{ADM}          &36.6&54.2&49.1&32.9&42.1&75.7&28.7 & 40.2 & 48.4 \\[-2pt]
 %                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
@@
 
-    \system{WDCMT}     &34.8  &51.8&47.8&30.7&37.3&49.3&25.9 & &\\[-2pt]
-                    &&&&&&&& \\
+    \system{WDCMT}     &&&&&&& & &\\[-2pt]
+                    &&&&&&&&& \\
 
-    \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & & 51.8 \\[-2pt]
+    \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & 41.9 & 51.8 \\[-2pt]
 %                      & 37.5 & 56.1 & 51.1 & 34.0  &  44.3 & 87.3 & \SW{28.4} & & 51.7 & \\
@@
   \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+  \begin{tabular}{|p{3cm}|*{8}{r|}|} \hline
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c||}{w\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
   %    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
@@
 The results are very clear and confirm previous observations: even though some clusters are very close, the net effect is a decrease in performance for almost all systems and conditions.
-In this setting, the best MDMT in our pool (MDL-Res) is no longer able to surpass the \system{Mix-Nat} baseline.\fyTodo{Note about scaling?}\fyTodo{More comparison?}
+In this setting, the best MDMT of our pool (MDL-Res) is no longer able to surpass the \system{Mix-Nat} baseline.\fyTodo{Note about scaling?}\fyTodo{More comparison?}
 

### 2020-05-04 2020/coling2020/multidomain.tex
10 pages limits, last results are almost there

@@
 \newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}
+\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
 \newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
@@
 \newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}
-\newcommand{\SB}[1]{\textcolor{green}{#1}}
-\newcommand{\SW}[1]{\textcolor{red}{#1}}
+% \newcommand{\SB}[1]{\textcolor{green}{#1}}
+% \newcommand{\SW}[1]{\textcolor{red}{#1}}
+\newcommand{\SB}[1]{\textbf{#1}}
+\newcommand{\SW}[1]{\underline{#1}}
+\renewcommand\textfraction{.1}
+\renewcommand\floatpagefraction{.95}
 
@@
 In doing so, we also shed lights on new problems that arise in this context, regarding for instance the accomodation of new domains in the course of training, or the computation of automating domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
-We show in section~\ref{sec:experiments}  existing approaches still fall short to match many of these requirements, notably with respect to handling a large amount of heterogeneous domains and to dynamically integrating new domains in training.\fyDone{Spell out conclusions}
+We show in in Section~\ref{sec:experiments} that existing approaches still fall short to match many of these requirements, notably with respect to handling a large amount of heterogeneous domains and to dynamically integrating new domains in training.\fyDone{Spell out conclusions}
 
@@
 Combining multiple domain-specific MTs can also be justified in the sake of distributional robustness \cite{Mansour09domainadaptation,Mansour09multiple}, for instance when the test mixture differs from the train mixture, or when it includes new domains unseen in training.
-Such scenario is already well documented for zero-shot multi-lingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}, where mixing languages has more than demonstrated its usefulness.
+Such scenario is already well documented for zero-shot multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}, where mixing languages has more than demonstrated its usefulness.
 An even more challenging case is when the MT would need to perform well for any test distribution (as studied for statistical MT in \cite{Huck15mixeddomain}).
@@
 The notion of a domain being a fragile one, an effective MDMT system should be able to discover not only when cross-domain sharing is useful (cf.\ requirement [P2-TUN])\jcDone{what is P2.2?}, but also when intra-domain heterogeneity is hurting.
-This requirement is tested by artificially conjoining two separate domains into one during training.
-The loss in performance with respect to the baseline (using correct domain tags) should then be as small as possible.
+This requirement is tested by artificially conjoining separate domains into one during training, hoping that the loss in performance with respect to the baseline (using correct domain tags) will remain small.
 
 \paragraph{[P4-ERR]}
-A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of domain tag errors.
-This is in sharp contrast with what is observed with fine-tuning, where using the wrong domain tag can yield ``catastrophic'' drops in performance.
-This requirement can be measured by translating test texts for domain $A$ with all erroneous or random domain tags and measure the correlated decrease in performance.
+A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, it has been observed that could yield ``catastrophic'' drops in performance.
+This requirement can be measured by translating test texts for domain $A$ with erroneous or random domain tags and measure the correlated loss.
 
@@
 A related situation occurs when the domain of a test document is unknown.
-Several situations need be considered: when the domain is seen in training, then using automatically predicted domain labels should not be much worse than using the correct one.
-This procedure can also be used test documents from unknown domains (zero-shot transfer).
-A good MD system should there outperform the default baseline that merges all available data.\fyDone{Systems react to unknown domains}
+Several situations need be considered: for domain is seen in training, then using automatically predicted domain labels should not be much worse than using the correct one.
+For test documents from unknown domains (zero-shot transfer), a good MD system should ideally outperform the default baseline that merges all available data.\fyDone{Systems react to unknown domains}
 % These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.
@@
 \paragraph{[P6-DYN]}
-Another requirement, more of an operational nature, is that a MDMT system should smoothly evolve to accomodate a growing number of domains, without having to retrain the full system each time new data is available.
+Another requirement, more of an operational nature, is that a MDMT system should smoothly evolve to handle a growing number of domains, without having to retrain the full system each time new data is available.
 This is a requirement [P6-DYN] that we challenge by dynamically changing the number of training and test domains.
@@
 \item a generic model trained on a concatenation of all corpora (\texttt{Mixed}).
-We develop two versions of this system, one where domain unbalance reflects the distribution of our parallel data (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
+We develop two versions\footnote{In fact three: to enable a fair comparison with WDCMT, a RNN-based variant is also trained and evaluated.} of this system, one where domain unbalance reflects the distribution of our parallel data (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
 The former is the best option when the train mixture $\mathcal{D}^s$ is also observed in testing; the latter should be used when the test distribution is uniform across domains.
-Accordingly, we report two aggrate scores: a weighted average reflecting the default distribution, and an unweighted average (test domains are equally important).
+Accordingly, we report two aggregate scores: a weighted average reflecting the default distribution, and an unweighted average, meaning that test domains are equally important.
 %   As explained above, depending whether MDMT are expected to perform specially well on large domains, or to treat all domains equally, one or the other training distribution should be prefered.\jcDone{All systems cand be learned following both Bal and Nat.
@@
 These methods thus only use domain tags in training.
-\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach, two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly, improved results.}\fyTodo{Check this}
+\item the multi-domain model of \newcite{Zeng18multidomain} (\system{WDCMT}), where a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly, improved results.}\fyDone{Check this}
 \item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connexions make it possible to by-pass this module.
@@
 
-In this study, all systems fail to outperform fine-tuning, sometimes by a wide margin, as is the case for an ``isolated'' domain such as \domain{rel}.
-This might be expected, since all domains are well separated (cf.\ Section~\ref{ssec:corpora}) and are hardly helping each other.
-In this situation, MDMT systems should make sure that each domain is allocated a sufficient number of specific parameters, so as to close the gap with fine-tuning.
+All systems fail to outperform fine-tuning, sometimes by a wide margin, especially for an ``isolated'' domain like \domain{rel}.
+This might be due to the fact that domains are well separated (cf.\ Section~\ref{ssec:corpora}) and are hardly helping each other.
+In this situation, MDMT systems should instead make sure that each domain is allocated a sufficient number of parameters, so as to close the gap with fine-tuning.
 \fyDone{More comments when we have all the results}\fyDone{Importance of sharing and unsharing}
-36.81	53.85	47.21	30	35.74	60.24	39.2	44.0	21.06
+% 36.81	53.85	47.21	30	35.74	60.24	39.2	44.0	21.06
 \begin{table*}
@@
   \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}|} \hline
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
     \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
-    \system{Rnn}  & 36.81 & 53.85 & 47.21 & 30 & 35.74 & 60.24  & 39.2  & 44.0 \\
+    \system{Mixed-Nat-RNN}  & 36.8 & 53.8 & 47.2 & 30.0 & 35.7 & 60.2  & 39.2  & 44.0 \\
     \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 \\ %& \\
@@
     \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 38.9 & 47.5 \\% & 23.3 \\
-    \system{WDCMT}       & 35.96 & 53.31 & 48.83 & 31.15 & 38.81 & 58.52 & 39.0 & 44.4 \\ % & 20.4 \\
+    \system{WDCMT}       & 36.0 & 53.3 & 48.8 & 31.1 & 38.8 & 58.5 & 39.0 & 44.4 \\ % & 20.4 \\
     \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & 51.9 \\%  & \SW{21.2} \\
@@
   \caption{Translation performance of various MDMT systems.
-We report the BLEU score for each domain, as well as domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages.
-Boldface denotes significant differences with respect to the \system{Mix-Nat} system.}
+We report BLEU scores for each domain, as well as domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages.
+Boldface denotes significant gains with respect to \system{Mix-Nat}, underline denotes significant losses.}
   \label{tab:performance}
-  \fyTodo{Fill the table with all results, including a comparison with *RNNs* for wcmd}
-  \fyTodo{Do we have significancy tests for averages?}
-  \fyTodo{Add tests wrt to RNN}
+%  \fyDone{Fill the table with all results, including a comparison with *RNNs* for wcmd}\fyTodo{Do we have significancy tests for averages?} \fyTodo{Add tests wrt to RNN}
 \end{table*}
@@
 A MD system should detect that these two pseudo-domains should be mutually beneficial, and should be hardly affected by this change with respect to the baseline scenario (no split).
-In this situation, we should expect MDMT to even surpass fine-tuning separately on each of these dummy domain, as MDMT exploits all data, where fine-tuning focuses only on a subpart.
+In this situation, we should expect MDMT to even surpass fine-tuning separately on each of these dummy domains, as MDMT exploits all data, while fine-tuning focuses only on a subpart.
 In testing, we decode the test set twice, once with each pseudo-domain tag.
 This makes no difference for \system{TMM}, \system{ADM}, \system{ADM} and \system{WDCMT}, which do not use domain tags in testing.
-In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains.
-We then translate the two corresponding test sets with the same (merged) system.
+In the \textsl{merge} experiment, we merge two corpora in training, in order to assess the robustness with respect to heterogenous domains [P3-HET].
+We then translate the two corresponding tests with the same system.
 
@@
     & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
-    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5}& -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& 32.55 & \SW{-3.28}\\
-    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}$^+$& +0.2& -2.2 & -2.9 & -0.5 & -0.4 & 35.92 & \SW{-1.7}\\
-    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3} $^+$ & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & 34.91 &\SW{-1.8}\\
+    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5} & -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& 32.55 & \SW{-3.3}\\
+    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -2.2 & -2.9 & -0.5 & -0.4 & 35.92 & \SW{-1.7}\\
+    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & 34.91 &\SW{-1.8}\\
     \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& 37.01 & \SW{-1.4}\\ 
-    \system{TTM} (!)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & & -0.1\\
+    \system{TTM} (*)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & 0.0 & -0.1\\
 %    \system{MDL Gated} & & & & & & & & &  \\
-    \system{DM} (!)           & -0.3   & -0.3  & +0.4 & +0.4 & +0.3 & +0.3 & +0.9 & +0.1 & &-0.9\\
-    \system{ADM} (!)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & &-0.2\\
-    \system{WDCMT} (!)     & -0.03    & -0.03  & +0.2 & +0.2  & +0.83 & +0.83  & -0.45 & -0.85 & *** & 0.19 \\
-    \system{MDL Res}   & -0.2   & -0.07 & +\SB{0.18}$^+$ &+0.0 & -0.89 & -0.89 & +\SB{0.7} & -0.35 & 31.78 & \SW{-1.3}\\
+    \system{DM} (*)           & -0.3   & -0.3  & +0.4 & +0.4 & +0.3 & +0.3 & +0.9 & +0.1 & 0.0 &-0.9\\
+    \system{ADM} (*)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & 0.0&-0.2\\
+    \system{WDCMT} (*)     & -0.0    & -0.0  & +0.2 & +0.2  & +0.8 & +0.8  & -0.4 & -0.8 & 0.0 & +0.2 \\
+    \system{MDL Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +\SB{0.7} & -0.3 & 31.8 & \SW{-1.3}\\
     \hline
@@
   \caption{Translation performance with variable domain definitions.
-For the Split experiments, we only report the loss in performance for the concerned test set (\domain{med} or \domain{law}), (*) denote significant losses when domains are split or merged; (+) denotes a significant improvement over \system{Full FT}; \fyTodo{Check significance}\fyTodo{How about wrong ?}}
+In the Split/Merge experiments, we report the loss in performance for the concerned test set(s).
+Underline denotes significant loss when domains are changed, bold a significant improvement over \system{Full FT}; (*) flags systems that ignore domain tags in testing.
+    % \fyTodo{Check wrong and significance of diff}\fyDone{How about wrong ?}
+  }
   \label{tab:redomains}
@@
 Our findings can be summarized as follows.
-For the split experiments, we see small variations that can be positive or negative compared to the situation were domains are kept a whole, bu these are hardly significative.
-All systems show here some robustness with respect to fuzzy domain boundaries; this is mostly notable for \system{ADM} suggesting that when domain are close, the neutralization of domain diferences is effective.
-In contrary, \system{FT-full} incurs large losses across the board, especially for the small data condition.
-Even in this very favourable case, very few MDMT systems are able to significantly outperform FT, and this is only observed for the small portion of the \domain{med} domain.
+For the split experiments, we see small variations that can be positive or negative compared to the baseline situation, but these are hardly significative.
+All systems show some robustness with respect to fuzzy domain boundaries; this is mostly notable for \system{ADM}, suggesting that when domain are close, ignoring domain differences is effective.
+In contrary, \system{FT-Full} incurs large losses across the board, especially for the small data condition.
+Even in this very favourable case, very few MDMT systems are able to significantly outperform \system{FT-Full} and this is only observed for the smaller part of the \domain{med} domain.
 The merge condition is hardly different, with large losses for \system{FT-full}, and small variations for all systems.
-We even observe improvements with respect to the situation where we use actual domains.\fyDone{More comments ?}
+We even observe some rare improvements with respect to the situation where we use actual domains.\fyDone{More comments ?}
 
 \subsubsection{Handling wrong or unknown domains \label{sssec:unknowns}}
-\fyTodo{Prepare table for the wrong domain analysis (see also last col.
-of the first table)}In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.
+\fyDone{Prepare table for the wrong domain analysis (see also last col.
+of the first table)}In the last two columns of Table~\label{tab:redomains}, we report the averaged (across domains) drop in performance when the domain information is erroneous.
+In the first (old), we use test data from the domains seen in training, presented with a random domain tag.
+In this situation, the loss is generally large (more than 10 BLEU points), except for systems trained to disregard domain tags in testing.\fyDone{Comments?}
 
-We consider two situations: test data from the domains seen in training, but with an erroneous tag (random domain);\fyTodo{Comments?}
-
-We now turn to requirement [P5-UNK], and translate test sentences from a domain unseen in training (\domain{news}).
-For each test sentence, we automatically predict the most likely domain tag, and use it for decoding.
-Results are in the rightmost column of Table~\ref{tab:redomains}.
-In this configuration, all systems do worse than the \system{Mixed-Nat} baseline (Bleu=23.5), sometimes by a large margin.\fyTodo{Check significance, To be completed}
+In the second (new), we assess requirement [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).
+For each sentence, we automatically predict the most likely domain tag and use it for decoding.
+% Results are in the rightmost column of Table~\ref{tab:redomains}.
+In this configuration, again, systems using domain tags in testing do significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5), sometimes by a large margin.\fyDone{Check significance, To be completed}
 
@@
 Another series of experiments evaluate the ability to dynamically handle suplementary domains (requirement [P6-DYN]) as follows.
-Starting with the existing MD systems of Section~\ref{ssec:rawperformance}, we introduce an extra domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is again of utmost importance for achieving optimal performance: since our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy based on the new training distribution.
-This is of course detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with training all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.
+Starting with the existing MD systems of Section~\ref{ssec:rawperformance}, we introduce an extra domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is critical for achieving optimal performance: as our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy based on the new training distribution.
+This is detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with training all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.
 We expect that a proper MDMT system should not be significantly affected by the addition of a new domain, and reach about the same performance as if training with this domain from scratch.
@@
     \system{LDC-Tag}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
-                    & -0.65 &	+0.46 &	-0.44 &	-0.58 &	+0.49 &	+\SB{3.8} &	-0.88 &	+0.5 &	-0.4 \\
+                    & -0.6 &	+0.5 &	-0.4 &	-0.6 &	+0.5 &	+\SB{3.8} &	-0.9 &	+0.5 &	-0.4 \\
 
@@
   \end{tabular}
-  \caption{Ability to handle new a domain: BLEU performance in ``resume training'' mode compared to a complete training session.
-Averages only take into account the six initial domains (\domain{News} excluded).}
+  \caption{Ability to handle new a domain: BLEU scores for the ``resume training'' mode compared to a complete training session.
+Averages only take into account the six initial domains (\domain{News} excluded).
+Underline denotes a significant loss between warmstart and coldstart, bold for a significant gain.}
   \label{tab:warmrestart}
-  \fyTodo{Fill the table, significancy testing against initial learning condition, for each domain}
+  \fyTodo{Fill the table, significancy testing against initial learning condition, for each domain, and averages ? }
   \fyTodo{Need correct number / differences for Mixed-Nat}
@@
 
-We can first compare the results of coldstart training with six (Table~\ref{tab:performance}) or seven domains (Table~\ref{tab:warmrestart}): the most striking observation is that the extra training data is detrimental for all systems and all domains,\fyTodo{is this significant ?}.
-Increasing the heterogeneity of the mix hurts performance across the board.
-This confirms that adding more out-of-domain data is not always helping, especially when domains are well separated.
-Comparing now the coldstart with the warmstart scenario, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training can negatively impact the performance on other domains - this happens notably for \system{DC-Tag}, \system{TTM} and \system{ADM}.
-In this setting \system{MDL Res} and \system{DM} show the smaller average loss, with the former achieving the better balance of total training cost and average translation performance of all systems.
+We can first compare the results of coldstart training with six (Table~\ref{tab:performance}) or seven domains (Table~\ref{tab:warmrestart}): the most striking observation is that the extra training data is detrimental for all systems and domains,\fyTodo{is this significant ?}.
+Increasing the heterogeneity of the mix with supplementary out-of-domain data hurts performance across the board.
+% Adding is not helping here, especially when domains are well separated.
+Comparing now the coldstart with the warmstart scenarios, we see that the former is always significantly better for \domain{news}, as expected and that resuming training can negatively impact the performance for other domains.
+This happens notably for \system{DC-Tag}, \system{TTM} and \system{ADM}.
+In this setting \system{MDL Res} and \system{DM} show the smaller average loss, with the former achieving the best balance of training cost and average translation performance.% of all systems.
 
@@
 
-\begin{table*}
+Table~\ref{tab:subdomains} reports the effect of using automatic domain tags for the 6 domain-specific test sets: each sentence was first assigned to an automatic class, translated with the corresponding multi-domain system with 30 classes; aggregate numbers were then computed, and contrasted with the 6-domain scenario.
+Results are clear and confirm previous observations: even though some clusters are very close, the net effect is a loss in performance for almost all systems and conditions.
+In this setting, the best MDMT in our pool (MDL-Res) is no longer able to surpass the \system{Mix-Nat} baseline.%\fyTodo{Note about scaling?}\fyTodo{More comparison?}
+
+\begin{table*}[t]
   \centering
@@
 
-Table~\ref{tab:subdomains} reports the effect of using automatic domain tags on the 6 domain-specific test sets: each sentence was first assigned to an automatic class, translated with the corresponding multi-domain system; aggregate numbers were then computed, as well as comparisons with the 6-domain scenarios.
-The results are very clear and confirm previous observations: even though some clusters are very close, the net effect is a decrease in performance for almost all systems and conditions.
-In this setting, the best MDMT in our pool (MDL-Res) is no longer able to surpass the \system{Mix-Nat} baseline.\fyTodo{Note about scaling?}\fyTodo{More comparison?}
-
 \section{Related work \label{sec:related}}
@@
 
-This latter strategy has also been widely used in NMT: the authors of \cite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token or in the form of additional domain-feature associated to each word.
-These results are reproduced in \cite{Tars18multidomain}, which also considers, as we do, automatically induced domain tags.
+The latter strategy has been widely used in NMT: \newcite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token or in the form of additional domain-feature associated to each word.
+These results are reproduced by \newcite{Tars18multidomain}, who also consider automatically induced domain tags.
 This technique also helps control the style of MT outputs in \cite{Sennrich16politeness,Niu18multitask}, and to encode the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}.
-Domain control can also be performed on the target side, as in \cite{Chen16guided}, where a vector of topic describing the current document is included as an extra context in the softmax layer of the decoder.
-Such ideas are further developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and commonalties are encoded in the network architecture: some parameters
-are shared across all domains, while others are domain-specific. % (part of the embeddings in \cite{Pham19generic}).
+Domain control can also be performed on the target side, as in \cite{Chen16guided}, where a topic vector describing the whole document serves as an extra context in the softmax layer of the decoder.
+Such ideas are further developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and commonalties are encoded in the network architecture: some parameters are shared across domains, while others are domain-specific.
+% (part of the embeddings in \cite{Pham19generic}).
 
-The techniques introduced in \cite{Britz17mixing} aim to ensure that domain information is actually used in a mix-domain system.
-Three alternatives are considered, using either domain classification (or domain normalization, via adversarial training) on source or target representations.
-There is no clear winner in either of their three language pairs considered, which only distinguish a limited number of domains.
-One contribution of this work was to introduce the idea of normalizing representations through adversarial training, in order to make the mixture of heterogeneous data more effective; representation normalization has since proven a key ingredient in multilingual transfer learning.
-The same basic techniques (partial parameter sharing, automatic domain identification / normalization) are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific representations on the one hand from domain-agnostic representation on the other hand.
+Techniques proposed by \cite{Britz17mixing} aim to ensure that domain information is actually used in a mix-domain system.
+Three methods are considered, using either domain classification (or domain normalization, via adversarial training) on the source or target side.
+There is no clear winner in either of the three language pairs considered.
+One contribution of this work is the idea of normalizing representations through adversarial training, so as to make the mixture of heterogeneous data more effective; representation normalization has since proven a key ingredient in multilingual transfer learning.
+The same basic techniques (parameter sharing, automatic domain identification / normalization) are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific representations on the one hand from domain-agnostic representation on the other hand.
 These representations are then processed as two separate inputs, then recombined to compute the translation.
 
-Another parameter-sharing scheme is presented in \cite{Jiang19multidomain}, which augments a Transformer model with domain specific heads, whose individual contributions are regulated at the word (position-level): some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling that the core of \cite{Huck15mixeddomain,Saunders19ucam}.
+Another parameter-sharing scheme is \cite{Jiang19multidomain}, which augments a Transformer model with domain-specific heads, whose contributions are regulated at the word/position level: some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling that the core of \cite{Huck15mixeddomain,Saunders19ucam}.
 The results for three language pairs outperform several standard baselines for a two-domain systems (in fr:en and ge:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
 
-As already explained, \newcite{Farajian17multidomain} differs from most studies, as it does not use domain labels.
-In their approach, the processing of a test sentence triggers the selection of a small set of similar fine-tuning instances; using these, the generic NMT system is tuned for some epochs, before delivering its translation.
-This approach thus entirely dispenses with the notion of a domain, and primarily relies on data selection/weighting techniques to handle data heterogeneity.
+\newcite{Farajian17multidomain} does not use domain tags in training.
+The processing of each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some epochs, before delivering its output.
+This approach entirely dispenses with the notion of domain and relies on data selection techniques to handle data heterogeneity.
 
@@
 In this comparison, architecture-based strategies seem to show the best promises.
+\fyTodo{Mitigate this}
 As expected however, MDMT are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend,
@@
 \bibliography{multidomain}
+
+\todos{}
+
+\end{document}
 
@@
 \end{table}
-
-\todos{}
-
-\end{document}
 

### 2020-05-06 2020/coling2020/multidomain.tex
debug
fix a couple of remaining issues in tables, 10 pages sharp

@@
 
-\paragraph{[P1-LAB]}\fyTodo{Decide naming scheme} A MDMT should perform better than the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}.
+\paragraph{[P1-LAB]}\fyDone{Decide naming scheme}A MDMT should perform better than the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}.
 Evaluating this requirement is a matter of a mere comparison, assuming the test distribution of domains is known: if all domains are equally important, performance averages can be reported; if they are not, weighted averages should be used instead.\fyFuture{Try random label assignments ?} % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
@@
 The multi-domain residual system (see description below) additionally use an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
-The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
+% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
 % The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
- %
 Training use a batch size of~128; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
@@
 If all MDMTs (except \system{DM} and \system{ADM}) slightly improve over the generic baseline (for most domains), these improvements are rarely significant.
-The best contender overall is \system{MDL Res}, which significantly improves over \system{Mixed-Nat} for three domains, and is the only one to clearly fulfill [P1-LAB].
-The use of a dedicated adaptation component proves here better than feature-based strategies.
+The best contender overall is \system{MDL Res}, which significantly improves over \system{Mixed-Nat} for three domains, and is the only one to clearly fulfill [P1-LAB]; \system{WDCMT} also improves on three domains, but regresses on one.
+The use of a dedicated adaptation component seems here to prove better than feature-based strategies.
 The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}), for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.
@@
     \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 38.9 & 47.5 \\% & 23.3 \\
-    \system{WDCMT}       & 36.0 & 53.3 & 48.8 & 31.1 & 38.8 & 58.5 & 39.0 & 44.4 \\ % & 20.4 \\
+    \system{WDCMT}       & 36.0 & 53.3 & \SB{48.8} & 31.1 & \SB{38.8} & \SW{58.5} & 39.0 & 44.4 \\ % & 20.4 \\
     \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & 51.9 \\%  & \SW{21.2} \\
@@
 We report BLEU scores for each domain, as well as domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages.
-Boldface denotes significant gains with respect to \system{Mix-Nat}, underline denotes significant losses.}
+Boldface denotes significant gains with respect to \system{Mix-Nat} (or \system{Mix-Nat-RNN}, for WDCMT), underline denotes significant losses.}
   \label{tab:performance}
-%  \fyDone{Fill the table with all results, including a comparison with *RNNs* for wcmd}\fyTodo{Do we have significancy tests for averages?} \fyTodo{Add tests wrt to RNN}
+  % \fyDone{Fill the table with all results, including a comparison with *RNNs* for wcmd}
+  \fyTodo{Do we have significancy tests for averages?} %\fyTodo{Add tests wrt to RNN}
 \end{table*}
@@
     & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
-    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5} & -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& 32.55 & \SW{-3.3}\\
-    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -2.2 & -2.9 & -0.5 & -0.4 & 35.92 & \SW{-1.7}\\
-    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & 34.91 &\SW{-1.8}\\
-    \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& 37.01 & \SW{-1.4}\\ 
+    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5} & -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& \SW{-21.8} & \SW{-3.3}\\% [32.5]
+    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -2.2 & -2.9 & -0.5 & -0.4 & \SW{-14.2} & \SW{-1.7}\\% [35.9]
+    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & \SW{-15.0} &\SW{-1.8}\\ % [34.9] 
+    \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-12.8} & \SW{-1.4}\\ % [37.0]
     \system{TTM} (*)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & 0.0 & -0.1\\
@@
     \system{WDCMT} (*)     & -0.0    & -0.0  & +0.2 & +0.2  & +0.8 & +0.8  & -0.4 & -0.8 & 0.0 & +0.2 \\
-    \system{MDL Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +\SB{0.7} & -0.3 & 31.8 & \SW{-1.3}\\
+    \system{MDL Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +\SB{0.7} & -0.3 & \SW{-20.1} & \SW{-1.3}\\ % [31.8]
     \hline
@@
   \caption{Translation performance with variable domain definitions.
-In the Split/Merge experiments, we report the loss in performance for the concerned test set(s).
-Underline denotes significant loss when domains are changed, bold a significant improvement over \system{Full FT}; (*) flags systems that ignore domain tags in testing.
-    % \fyTodo{Check wrong and significance of diff}\fyDone{How about wrong ?}
+In the Split/Merge experiments, we report the change in performance for the concerned test set(s).
+Underline denotes significant loss when domains are changed wrt.\ the baseline situation; bold denotes a significant improvement over \system{FT-Full}; (*) flags systems that ignore domain tags in testing.
+    \fyDone{Check wrong and significance of diff for ``old''}\fyDone{How about wrong ?}
   }
@@
 This is detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with training all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.
-We expect that a proper MDMT system should not be significantly affected by the addition of a new domain, and reach about the same performance as if training with this domain from scratch.
+We expect that a MDMT system should not be significantly impacted by the addition of a new domain and should reach about the same performance as when training with this domain from scratch.
 Note that from a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{DC-Feat} or \system{TTM} and the like, for which new domains merely add new labels.
-It is less simple for \system{LDC-Tag} and \system{MDL~Res}\fyDone{Complete effects of new domain}, for which the number of possible domains is built in the architecture and has to be anticipated from the beginning.
+It is less easy for \system{LDC-Tag} and \system{MDL~Res}\fyDone{Complete effects of new domain}, for which the number of possible domains is built in the architecture and has to be anticipated from the beginning.
 This makes another difference between domain-bounded MT-systems, for which the number of domains is limited and truly open-domain systems.% , which seamlessly integrate additional domains.
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-
-    \system{Mixed-Nat}  & 37.3  & 54.6   & 50.1   & 33.5  & 43.2  & 77.5  & 23.5 & 41.1 & 49.4 \\
-    \system{FT-Full}       & 37.7  & 59.2   & 54.5   & 34.0  & 46.8  & 90.8  & **** & 42.8 & 53.8 \\ \hline\hline
-
+     \system{Mixed-Nat}  & 37.1  & 54.1  & 49.6	& 34.1  & 42.1	& 77.0 & 28.9 & 40.8	& 49.0 \\ 
+%    \system{Mixed-Nat}  & 37.3  & 54.6   & 50.1   & 33.5  & 43.2  & 77.5  & 23.5 & 41.1 & 49.4 \\
+%    \system{FT-Full}       & 37.7  & 59.2   & 54.5   & 34.0  & 46.8  & 90.8  & **** & 42.8 & 53.8 \\
+    \hline%\hline
      \system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9  & 41.2 & 50.1\\[-2pt]
@@
   \fyTodo{Fill the table, significancy testing against initial learning condition, for each domain, and averages ? }
-  \fyTodo{Need correct number / differences for Mixed-Nat}
+  \fyDone{Need correct number / differences for Mixed-Nat}
   \fyDone{averages include one more domain ? No for comparison}
@@
 
-We can first compare the results of coldstart training with six (Table~\ref{tab:performance}) or seven domains (Table~\ref{tab:warmrestart}): the most striking observation is that the extra training data is detrimental for all systems and domains,\fyTodo{is this significant ?}.
+We can first compare the results of coldstart training with six (Table~\ref{tab:performance}) or seven domains (Table~\ref{tab:warmrestart}): the most striking observation is that the extra training data is detrimental for almost all systems and domains\fyDone{is this significant ?}.
 Increasing the heterogeneity of the mix with supplementary out-of-domain data hurts performance across the board.
 % Adding is not helping here, especially when domains are well separated.
-Comparing now the coldstart with the warmstart scenarios, we see that the former is always significantly better for \domain{news}, as expected and that resuming training can negatively impact the performance for other domains.
+Comparing now the coldstart with the warmstart scenarios, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training can negatively impact the performance for other domains.
 This happens notably for \system{DC-Tag}, \system{TTM} and \system{ADM}.
@@
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c||}{w\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
   %    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
@@
   %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag}       & 38.5 & \SW{54.0} & 49.0   & 33.6 & \SW{42.2} & \SW{76.7} & 49.0 & 41.6 \\%    & 21.8 \\
-    \system{DC-Feat}      & 37.3  & 54.2 & 49.3   & 33.6 & \SW{41.9} & \SW{75.8} & 48.3 & 40.7  \\% & \SW{21.7} \\
-    \system{LDC-Tag}     & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 48.3 & 39.1         \\% & 22.1 \\ 
-    \system{TTM}            & 37.4 & \SW{53.7} & 48.9 & 32.8 & 41.3 & \SW{75.8} & 48.3 & 40.7   \\% &  23.4 \\
-    \system{DM}             & 35.4 & 49.3  & 45.2 & 29.7 & 37.1 & \SW{60.0} & 42.8 & 37.8 \\ % & 22.6\\
-    \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 47.1 & 39.5\\% & 23.3 \\
-    \system{WDCMT}       & 36.0 & 53.8 & 48.5 & 31.0 & 39.0 & 63.5 & 45.3 & 39.1\\ % & 20.4 \\
-    \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 50.7& 41.2\\%  & \SW{21.2} \\
+    \system{DC-Tag}       & 38.5 & \SW{54.0} & 49.0   & 33.6 & \SW{42.2} & \SW{76.7} & 41.6 & 49.0 \\%    & 21.8 \\
+    \system{DC-Feat}      & 37.3  & 54.2 & 49.3   & 33.6 & \SW{41.9} & \SW{75.8} & 40.7 & 48.3  \\% & \SW{21.7} \\
+    \system{LDC-Tag}     & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 39.1 & 48.3         \\% & 22.1 \\ 
+    \system{TTM}            & 37.4 & \SW{53.7} & 48.9 & 32.8 & 41.3 & \SW{75.8} & 40.7 & 48.3   \\% &  23.4 \\
+    \system{DM}             & 35.4 & 49.3  & 45.2 & 29.7 & 37.1 & \SW{60.0} & 37.8 & 42.8 \\ % & 22.6\\
+    \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 39.5 & 47.1\\% & 23.3 \\
+    \system{WDCMT}       & 36.0 & 53.8 & 48.5 & 31.0 & 39.0 & 63.5 & 39.1 & 45.3\\ % & 20.4 \\
+    \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & 50.7\\%  & \SW{21.2} \\
      \hline
@@
   \label{tab:subdomains}
-  \fyDone{Fill the table with correct results,}
+  \fyDone{Fill the table with correct results,}\fyDone{Change wavg and avg}
 \end{table*}
@@
 \section{Conclusion and outlook \label{sec:conclusion}}
-\fyDone{Write conclusions}
+\fyDone{Write conclusions}\fyTodo{no mixture of experts}
 In this study, we have carefully reconsidered the idea of multi-domain machine translation, which seems to be taken for granted in many recent studies.

### 2020-05-11 2020/coling2020/multidomain.tex
take in JC's comments

@@
-%
 % File coling2020.tex
@@
 \newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{blue}{#1}}}
-\newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{blue}{#1}}}
+\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{red}{#1}}}
+\newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{red}{#1}}}
 
@@
 In this setting, \emph{domain adaptation} (DA) methods are in order.
-DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory, Quinonero08dataset,Pan10asurvey})\fyDone{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}) and various techniques exist to handle both the situations where a (small) training sample drawn from $\mathcal{D}_t$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
+DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory,Quinonero08dataset,Pan10asurvey})\fyDone{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}).
+Various techniques thus exist to handle both the situations where a (small) training sample drawn from $\mathcal{D}_t$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
 % In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
@@
 A seemingly related problem is \emph{multi-domain} (MD) machine translation \cite{Sajjad17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system is trained and tested with data from multiple domains.
-MD machine translation (MDMT) corresponds to a very common situation, where all available data, no matter its origin, is used to train a robust system that perform well for any kind of new input.
+MD machine translation (MDMT) corresponds to a very common situation, where all available data, no matter its origin, is used to train a robust system that performs well for any kind of new input.
 % As pointed out both by \newcite{Dredze08online} and \cite{Finkel09hierarchical}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}.
-If the intuitions behind MDMT are quite simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should MDMT performs well when the test data is distributed like the training data, when it is equally distributed across domains or when the test distribution is unknown?
+If the intuitions behind MDMT are quite simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should MDMT perform well when the test data is distributed like the training data, when it is equally distributed across domains or when the test distribution is unknown?
 Should MDMT also be robust to new domains?
@@
 In this paper, we formulate in a more precise fashion the requirements that an effective MDMT system should meet (section~\ref{sec:requirements}).
-Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them (section~\ref{sec:challenging}).
+Our first contribution is thus of methodological nature and consists of lists of\jcTodo{enumerate?} expected properties of MDMT systems and associated measurements to evaluate them (section~\ref{sec:challenging}).
 In doing so, we also shed lights on new problems that arise in this context, regarding for instance the accomodation of new domains in the course of training, or the computation of automating domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
-We show in in Section~\ref{sec:experiments} that existing approaches still fall short to match many of these requirements, notably with respect to handling a large amount of heterogeneous domains and to dynamically integrating new domains in training.\fyDone{Spell out conclusions}
+We show in Section~\ref{sec:experiments} that existing approaches still fall short to match many of these requirements, notably with respect to handling a large amount of heterogeneous domains and to dynamically integrating new domains in training.\fyDone{Spell out conclusions}
 
@@
 
-A typical learning scenario in MT is to have access to a samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$.
+A typical learning scenario in MT is to have access to samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$.
 Multi-domain learning, as defined in \cite{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ of several domains, making the problem distinct from mere domain adaption.
@@
 
-This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is \cite{Farajian17multidomain}, where test translations rely on a similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here.
+This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is \cite{Farajian17multidomain}, where test translations rely on similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here.
 Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function.
@@
 
-A second line of reasoning rests on linguistic properties of the translation function and contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and tend remain semantically stable accross domains, motivating some cross-domain parameter sharing.
+A second line of reasoning rests on linguistic properties of the translation function and contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and tend to remain semantically stable accross domains, motivating some cross-domain parameter sharing.
 A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the translation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
@@
 In this section, we propose seven\fyDone{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.
-All these evaluations will rest on comparison of translation performance, and do not depend on the choice a particular metrics.
+All these evaluations will rest on comparison of translation performance, and do not depend on the choice of a particular metric.
 To make our results comparable with the literature, we will only use the BLEU score \cite{Papineni02bleu} in Section~\ref{sec:experiments}, noting it may not be the best yardstick to assess subtle improvements of lexical choices that are often associated with domain adapted systems \cite{Irvine13measuring}.
@@
 \paragraph{[P4-ERR]}
-A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, it has been observed that could yield ``catastrophic'' drops in performance.
+A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, it has been observed that could yield ``catastrophic'' drops in performance\jcTodo{cite needed?}.
 This requirement can be measured by translating test texts for domain $A$ with erroneous or random domain tags and measure the correlated loss.
@@
 A related situation occurs when the domain of a test document is unknown.
-Several situations need be considered: for domain is seen in training, then using automatically predicted domain labels should not be much worse than using the correct one.
+Several situations need be considered: for domains seen in training, then using automatically predicted domain labels should not be much worse than using the correct one.
 For test documents from unknown domains (zero-shot transfer), a good MD system should ideally outperform the default baseline that merges all available data.\fyDone{Systems react to unknown domains}
@@
 Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
-The multi-domain residual system (see description below) additionally use an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
+The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
 % The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
@@
 \item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});\jcTodo{why not LDR as in Minh's paper?}
 \item the three proposals of \newcite{Britz17mixing}.
@@
 The use of a dedicated adaptation component seems here to prove better than feature-based strategies.
-The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}), for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.
+The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}), for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.\jcTodo{why LDC does not outperform DC as in Minh's paper (for en2fr MED) ?}
 
@@
 \fyDone{More comments when we have all the results}\fyDone{Importance of sharing and unsharing}
-% 36.81	53.85	47.21	30	35.74	60.24	39.2	44.0	21.06
+
 \begin{table*}
@@
 
-Table~\ref{tab:redomains} summaries the results of four experiments where we artificially redefine the boundaries of domains, with the aim to challenge requirements [P2-TUN], [P3-HET], and [P4-ERR].
+Table~\ref{tab:redomains} summarizes the results of four experiments where we artificially redefine the boundaries of domains, with the aim to challenge requirements [P2-TUN], [P3-HET], and [P4-ERR].
 In first three, we randomly \emph{split} one corpus in two parts and proceed as if this corresponded to two actual domains.
-A MD system should detect that these two pseudo-domains should be mutually beneficial, and should be hardly affected by this change with respect to the baseline scenario (no split).
-In this situation, we should expect MDMT to even surpass fine-tuning separately on each of these dummy domains, as MDMT exploits all data, while fine-tuning focuses only on a subpart.
+A MD system should detect that these two pseudo-domains are mutually beneficial, and should be hardly affected by this change with respect to the baseline scenario (no split).
+In this situation, we expect MDMT to even surpass fine-tuning separately on each of these dummy domains, as MDMT exploits all data, while fine-tuning focuses only on a subpart.
 In testing, we decode the test set twice, once with each pseudo-domain tag.
@@
   \caption{Translation performance with variable domain definitions.
-In the Split/Merge experiments, we report the change in performance for the concerned test set(s).
-Underline denotes significant loss when domains are changed wrt.\ the baseline situation; bold denotes a significant improvement over \system{FT-Full}; (*) flags systems that ignore domain tags in testing.
+In the Split/Merge experiments, we report BLEU differences for the concerned test set(s).
+Underline denotes significant loss when domains are changed wrt.\ the baseline situation; bold for a significant improvement over \system{FT-Full}; (*) flags systems ignoring test domains.
     \fyDone{Check wrong and significance of diff for ``old''}\fyDone{How about wrong ?}
@@
 \subsubsection{Handling wrong or unknown domains \label{sssec:unknowns}}
+
 \fyDone{Prepare table for the wrong domain analysis (see also last col.
-of the first table)}In the last two columns of Table~\label{tab:redomains}, we report the averaged (across domains) drop in performance when the domain information is erroneous.
-In the first (old), we use test data from the domains seen in training, presented with a random domain tag.
+of the first table)}In the last two columns of Table~\ref{tab:redomains}, we report the averaged (across domains) drop in performance when the domain information is erroneous.
+In the first (old), we use test data from the domains seen in training, presented with a random domain tag\jcTodo{all configs use the same tag or is randomization is run for each config}.
 In this situation, the loss is generally large (more than 10 BLEU points), except for systems trained to disregard domain tags in testing.\fyDone{Comments?}
 
-In the second (new), we assess requirement [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).
+In the second (new), we assess requirement [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\jcTodo{i would use RND / UNK rather than OLD / NEW}
 For each sentence, we automatically predict the most likely domain tag and use it for decoding.
 % Results are in the rightmost column of Table~\ref{tab:redomains}.
-In this configuration, again, systems using domain tags in testing do significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5), sometimes by a large margin.\fyDone{Check significance, To be completed}
+In this configuration, again, systems using domain tags in testing do significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\jcTodo{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}
 
@@
 %                   & 37.2 & 54.7   & 49.9    & 34.2  &  43.0  & 79.9 & \SW{28.1}  & &\\
-                   & -0.2 & -0.1  & -0.1 & -0.6 & -0.9 & +0.3 & \SW{-0.8} & -0.2 & -0.3\\
+                   & \scriptsize -0.2 & \scriptsize -0.1  & \scriptsize -0.1 & \scriptsize -0.6 & \scriptsize -0.9 & \scriptsize +0.3 & \scriptsize \SW{-0.8} & \scriptsize -0.2 & \scriptsize -0.3\\
 
@@
 %                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
-                   & +0.3  & -0.1   & -0.6    & \SW{-1.1}  &  \SW{-1.3}  &\SW{-3.5}  & \SW{-1.4}  & -0.1 & -1.2\\
+                   & \scriptsize +0.3  & \scriptsize -0.1   & \scriptsize -0.6    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.3}  & \scriptsize \SW{-3.5}  & \scriptsize \SW{-1.4}  & \scriptsize -0.1 & \scriptsize -1.2\\
 
     \system{LDC-Tag}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
-                    & -0.6 &	+0.5 &	-0.4 &	-0.6 &	+0.5 &	+\SB{3.8} &	-0.9 &	+0.5 &	-0.4 \\
+                    & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize -0.4 & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize +\SB{3.8} & \scriptsize -0.9 & \scriptsize +0.5 & \scriptsize -0.4 \\
 
@@
 %                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
-                    & -0.3  & -0.3  & -0.5    & \SW{-1.1}  &  \SW{-1.0}  & \SW{-4.0}  & \SW{-1.4}& -0.5 & -1.2\\
+                    & \scriptsize -0.3  & \scriptsize -0.3  & \scriptsize -0.5    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.0}  & \scriptsize \SW{-4.0}  & \scriptsize \SW{-1.4}& \scriptsize -0.5 & \scriptsize -1.2\\
 
@@
 %                   &36.5&51.7&47.4&31.7&40.3&65.7&\SW{25.8}& & \\
-                   &+0.6&+0.4&+0.6&-0.1&0.5&0.0&\SW{-1.2}& +0.5 & +0.3\\
+                   & \scriptsize +0.6& \scriptsize +0.4& \scriptsize +0.6& \scriptsize -0.1& \scriptsize +0.5 & \scriptsize 0.0& \scriptsize \SW{-1.2}& \scriptsize +0.5 & \scriptsize +0.3\\
 
@@
 %                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
-                   & +0.3 & -0.8 & -0.8 & -0.2 & -0.4 & \SW{-5.0} & \SW{-1.9} & -0.2 & -1.1\\
+                   & \scriptsize +0.3 & \scriptsize -0.8 & \scriptsize -0.8 & \scriptsize -0.2 & \scriptsize -0.4 & \scriptsize \SW{-5.0} & \scriptsize \SW{-1.9} & \scriptsize -0.2 & \scriptsize -1.1\\
 
@@
 %                      & 37.5 & 56.1 & 51.1 & 34.0  &  44.3 & 87.3 & \SW{28.4} & & 51.7 & \\
-                        & -0.2 & +0.5 & 0.0 & -0.4  &  -0.2 & -0.2 & \SW{-0.8} & -0.2 & -0.1 \\
+                        & \scriptsize -0.2 & \scriptsize +0.5 & \scriptsize 0.0 & \scriptsize -0.4  &  \scriptsize -0.2 & \scriptsize -0.2 & \scriptsize \SW{-0.8} & \scriptsize -0.2 & \scriptsize -0.1 \\
      \hline
   \end{tabular}
-  \caption{Ability to handle new a domain: BLEU scores for the ``resume training'' mode compared to a complete training session.
+  \caption{Ability to handle a new domain: BLEU scores for the ``resume training'' mode compared to a complete training session.
 Averages only take into account the six initial domains (\domain{News} excluded).
@@
   \fyDone{averages include one more domain ? No for comparison}
+  \fyDone{i used smaller number for differences (easier to read...)}
 \end{table*}
@@
 We also note that many of these classes are mere split of the large \domain{med}, while a fewer number of classes are mix of two existing domains.
-We are thus in a position to reiterate in more realistic conditions the measurements of Section~\label{ssec:} and to test whether multi-domain systems can effectively take advantage from the cross-domain similarities, and to eventually perform better that fine-tuning.
+We are thus in a position to reiterate at a larger scale the measurements of Section~\ref{ssec:redomains} and test whether multi-domain systems can effectively take advantage from the cross-domain similarities, and to eventually perform better that fine-tuning.
 
@@
 
-Multi-domain translation was already proposed for statistical MT, either considering as we do multiple sources of training data (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}), or topically heterogeneous domains \cite{Eidelman12topic,Hasler14dynamic-topic}.\fyDone{Add also topic models refs} Two main strategies were considered: instance-based, involving a measure of similarities between train and test domains; feature-based, where domain/topic labels give rise to additional features.
+Multi-domain translation was already proposed for statistical MT, either considering as we do multiple sources of training data (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}), or domains made of multiple topics\fyDone{you mean multiple topics?} \cite{Eidelman12topic,Hasler14dynamic-topic}.\fyDone{Add also topic models refs} Two main strategies were considered: instance-based, involving a measure of similarities between train and test domains; feature-based, where domain/topic labels give rise to additional features.
 
@@
 This approach entirely dispenses with the notion of domain and relies on data selection techniques to handle data heterogeneity.
+\jcTodo{cite Xu et al 2019, Lexical Micro-adaptation for Neural Machine Translation}
 

### 2020-05-13 2020/coling2020/multidomain.tex
significative test
delete enumerate

@@
 In this paper, we formulate in a more precise fashion the requirements that an effective MDMT system should meet (section~\ref{sec:requirements}).
-Our first contribution is thus of methodological nature and consists of lists of\jcTodo{enumerate?} expected properties of MDMT systems and associated measurements to evaluate them (section~\ref{sec:challenging}).
+Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT systems and associated measurements to evaluate them (section~\ref{sec:challenging}).
 In doing so, we also shed lights on new problems that arise in this context, regarding for instance the accomodation of new domains in the course of training, or the computation of automating domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
@@
     \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 \\ %& \\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.7 & 53.8\\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & 42.3 & 53.3 \\ \hline % & 20.5\\ \hline
+    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & \SB{42.7} & \SB{53.8} \\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & \SB{42.3} & \SB{53.3} \\ \hline % & 20.5\\ \hline
  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
@@
     \system{WDCMT}       & 36.0 & 53.3 & \SB{48.8} & 31.1 & \SB{38.8} & \SW{58.5} & 39.0 & 44.4 \\ % & 20.4 \\
-    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & 51.9 \\%  & \SW{21.2} \\
+    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & \SB{51.9} \\%  & \SW{21.2} \\
     \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
@@
     \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5} & -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& \SW{-21.8} & \SW{-3.3}\\% [32.5]
-    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -2.2 & -2.9 & -0.5 & -0.4 & \SW{-14.2} & \SW{-1.7}\\% [35.9]
+    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -0.4 & -0.4 & -0.5 & -0.4 & \SW{-14.2} & \SW{-1.7}\\% [35.9]
     \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & \SW{-15.0} &\SW{-1.8}\\ % [34.9] 
@@
 %                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
-                   & \scriptsize +0.3  & \scriptsize -0.1   & \scriptsize -0.6    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.3}  & \scriptsize \SW{-3.5}  & \scriptsize \SW{-1.4}  & \scriptsize -0.1 & \scriptsize -1.2\\
+                   & \scriptsize +0.3  & \scriptsize -0.1   & \scriptsize -0.6    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.3}  & \scriptsize \SW{-3.5}  & \scriptsize \SW{-1.4}  & \scriptsize -0.1 & \scriptsize \SW{-1.2}\\
 
     \system{LDC-Tag}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
-                    & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize -0.4 & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize +\SB{3.8} & \scriptsize -0.9 & \scriptsize +0.5 & \scriptsize -0.4 \\
+                    & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize -0.4 & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize +\SB{3.8} & \scriptsize \SW{-0.9} & \scriptsize +0.5 & \scriptsize -0.4 \\
 
@@
 %                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
-                    & \scriptsize -0.3  & \scriptsize -0.3  & \scriptsize -0.5    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.0}  & \scriptsize \SW{-4.0}  & \scriptsize \SW{-1.4}& \scriptsize -0.5 & \scriptsize -1.2\\
+                    & \scriptsize -0.3  & \scriptsize -0.3  & \scriptsize -0.5    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.0}  & \scriptsize \SW{-4.0}  & \scriptsize \SW{-1.4}& \scriptsize -0.5 & \scriptsize \SW{-1.2}\\
 
@@
 %                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
-                   & \scriptsize +0.3 & \scriptsize -0.8 & \scriptsize -0.8 & \scriptsize -0.2 & \scriptsize -0.4 & \scriptsize \SW{-5.0} & \scriptsize \SW{-1.9} & \scriptsize -0.2 & \scriptsize -1.1\\
+                   & \scriptsize +0.3 & \scriptsize -0.8 & \scriptsize -0.8 & \scriptsize -0.2 & \scriptsize -0.4 & \scriptsize \SW{-5.0} & \scriptsize \SW{-1.9} & \scriptsize -0.2 & \scriptsize \SW{-1.1}\\
 
@@
     \system{DC-Tag}       & 38.5 & \SW{54.0} & 49.0   & 33.6 & \SW{42.2} & \SW{76.7} & 41.6 & 49.0 \\%    & 21.8 \\
-    \system{DC-Feat}      & 37.3  & 54.2 & 49.3   & 33.6 & \SW{41.9} & \SW{75.8} & 40.7 & 48.3  \\% & \SW{21.7} \\
-    \system{LDC-Tag}     & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 39.1 & 48.3         \\% & 22.1 \\ 
-    \system{TTM}            & 37.4 & \SW{53.7} & 48.9 & 32.8 & 41.3 & \SW{75.8} & 40.7 & 48.3   \\% &  23.4 \\
+    \system{DC-Feat}      & 37.3  & 54.2 & 49.3   & 33.6 & \SW{41.9} & \SW{75.8} & 40.8 & \SW{48.7}  \\% & \SW{21.7} \\
+    \system{LDC-Tag}     & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 39.1 & \SW{48.3}         \\% & 22.1 \\ 
+    \system{TTM}            & 37.4 & \SW{53.7} & 48.9 & 32.8 & 41.3 & \SW{75.8} & 40.7 & \SW{48.3}   \\% &  23.4 \\
     \system{DM}             & 35.4 & 49.3  & 45.2 & 29.7 & 37.1 & \SW{60.0} & 37.8 & 42.8 \\ % & 22.6\\
     \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 39.5 & 47.1\\% & 23.3 \\
-    \system{WDCMT}       & 36.0 & 53.8 & 48.5 & 31.0 & 39.0 & 63.5 & 39.1 & 45.3\\ % & 20.4 \\
-    \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & 50.7\\%  & \SW{21.2} \\
+    \system{WDCMT}       & 35.57 & 53.08 & 48.43 & 30.55 & \SW{37.73} & \SW{56.02} & 38.5 & 43.6 \\ % & 20.4 \\ 
+    \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SB{50.7}\\%  & \SW{21.2} \\
      \hline

### 2020-05-13 2020/coling2020/multidomain.tex
cite catastrophic

@@
 \paragraph{[P4-ERR]}
-A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, it has been observed that could yield ``catastrophic'' drops in performance\jcTodo{cite needed?}.
+A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, it has been observed that could yield ``catastrophic'' drops in performance\jcTodo{cite \cite{McCloskey89catastrophic}}.
 This requirement can be measured by translating test texts for domain $A$ with erroneous or random domain tags and measure the correlated loss.

### 2020-05-13 2020/coling2020/multidomain.tex
Fixes after our discussion with JC

--- 2020-05-13 2020/coling2020/multidomain.tex
@@
 \paragraph{[P4-ERR]}
-A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, it has been observed that could yield ``catastrophic'' drops in performance\jcTodo{cite needed?}.
+A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, catastrophic' drops in performance can be observerd\fyDone{cite \cite{McCloskey89catastrophic}}.
 This requirement can be measured by translating test texts for domain $A$ with erroneous or random domain tags and measure the correlated loss.
@@
 \item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});\jcTodo{why not LDR as in Minh's paper?}
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});\fyDone{why not LDR as in Minh's paper?}
 \item the three proposals of \newcite{Britz17mixing}.
@@
 Results are in Table~\ref{tab:performance}.
-As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points\jcDone{BP?} better, notably owing to the much better results for \domain{rel}).
+As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points\fyDone{BP?} better, notably owing to the much better results for \domain{rel}).
 As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains.
@@
 The use of a dedicated adaptation component seems here to prove better than feature-based strategies.
-The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}), for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.\jcTodo{why LDC does not outperform DC as in Minh's paper (for en2fr MED) ?}
+The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}), for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.\fyDone{why LDC does not outperform DC as in Minh's paper (for en2fr MED) ?}
 
@@
     \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   \\% & \SW{21.7} \\
-    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} &40.9 & 49.8          \\% & 22.1 \\ 
+    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} &40.9 & 49.8          \\% & 22.1 \\ 
     \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.0 & 49.7     \\% &  23.4 \\
@@
   % \fyDone{Fill the table with all results, including a comparison with *RNNs* for wcmd}
-  \fyTodo{Do we have significancy tests for averages?} %\fyTodo{Add tests wrt to RNN}
+  \fyDone{Do we have significancy tests for averages?} %\fyTodo{Add tests wrt to RNN}
 \end{table*}
@@
 In testing, we decode the test set twice, once with each pseudo-domain tag.
-This makes no difference for \system{TMM}, \system{ADM}, \system{ADM} and \system{WDCMT}, which do not use domain tags in testing.
+This makes no difference for \system{TTM}, \system{DM}, \system{ADM} and \system{WDCMT}, which do not use domain tags in testing.
 In the \textsl{merge} experiment, we merge two corpora in training, in order to assess the robustness with respect to heterogenous domains [P3-HET].
@@
     \hfill Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & \multicolumn{2}{c|}{Wrong} \\ % \hline
-     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{law} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} &  \multicolumn{1}{c|}{old} &  \multicolumn{1}{c|}{new}\\ \hline
+     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{law} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} &  \multicolumn{1}{c|}{\domain{rnd}} &  \multicolumn{1}{c|}{\domain{new}}\\ \hline
     & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
@@
     \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & \SW{-15.0} &\SW{-1.8}\\ % [34.9] 
-    \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-12.8} & \SW{-1.4}\\ % [37.0]
+    \system{LDR}           & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-12.8} & \SW{-1.4}\\ % [37.0]
     \system{TTM} (*)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & 0.0 & -0.1\\
@@
 
-\fyDone{Prepare table for the wrong domain analysis (see also last col.
-of the first table)}In the last two columns of Table~\ref{tab:redomains}, we report the averaged (across domains) drop in performance when the domain information is erroneous.
-In the first (old), we use test data from the domains seen in training, presented with a random domain tag\jcTodo{all configs use the same tag or is randomization is run for each config}.
-In this situation, the loss is generally large (more than 10 BLEU points), except for systems trained to disregard domain tags in testing.\fyDone{Comments?}
+\fyTodo{Numbers for the rand domain analysis}In the last two columns of Table~\ref{tab:redomains}, we report the drop in performance when the domain information is not correct.
+In the first (\domain{rnd}), we use test data from the domains seen in training, presented with a random domain tag\fyDone{all configs use the same tag or is randomization is run for each config}.
+In this situation, the loss with respect to using the correct domain is generally large (more than 10 BLEU points), showing an overall failure to meet requirement [P4-ERR], except for systems which disregard domain tags in testing.
+\fyDone{Comments?}
 
-In the second (new), we assess requirement [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\jcTodo{i would use RND / UNK rather than OLD / NEW}
-For each sentence, we automatically predict the most likely domain tag and use it for decoding.
+In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the most likely domain tag and use it for decoding.
 % Results are in the rightmost column of Table~\ref{tab:redomains}.
-In this configuration, again, systems using domain tags in testing do significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\jcTodo{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}
+In this configuration, again, systems using domain tags in testing perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\fyDone{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}
 
@@
 Starting with the existing MD systems of Section~\ref{ssec:rawperformance}, we introduce an extra domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is critical for achieving optimal performance: as our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy based on the new training distribution.
-This is detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with training all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.
-We expect that a MDMT system should not be significantly impacted by the addition of a new domain and should reach about the same performance as when training with this domain from scratch.
-Note that from a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{DC-Feat} or \system{TTM} and the like, for which new domains merely add new labels.
-It is less easy for \system{LDC-Tag} and \system{MDL~Res}\fyDone{Complete effects of new domain}, for which the number of possible domains is built in the architecture and has to be anticipated from the beginning.
-This makes another difference between domain-bounded MT-systems, for which the number of domains is limited and truly open-domain systems.% , which seamlessly integrate additional domains.
+This is detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with training all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.\footnote{WDCMT results are excluded from this Table, as resuming training proved difficult to implement.}\fyDone{Ecrire un commentaire}
+We expect that MDMT systems should not to be significantly impacted by the addition of a new domain and reach about the same performance as when training with this domain from scratch.
+From a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{DC-Feat} or \system{TTM}, for which new domains merely add new labels.
+It is less easy for \system{DM}, \system{ADM} and \system{WDCMT}, which include a built-in domain classifier whose outputs have to be given, or for \system{LDR} and \system{MDL~Res} for which the number of possible domains is built in the architecture and has to be anticipated from the start.
+This makes a difference between domain-bounded systems, for which the number of domains is limited and truly open-domain systems.% , which seamlessly integrate additional domains.
 
@@
 
-    \system{WDCMT}     &&&&&&& & &\\[-2pt]
-                    &&&&&&&&& \\
+%    \system{WDCMT}     &&&&&&& & &\\[-2pt]
+%                    &&&&&&&&& \\
 
@@
   \label{tab:warmrestart}
-  \fyTodo{Fill the table, significancy testing against initial learning condition, for each domain, and averages ? }
+  \fyDone{Fill the table, significancy testing against initial learning condition, for each domain, and averages ? }
   \fyDone{Need correct number / differences for Mixed-Nat}
@@
 Results are clear and confirm previous observations: even though some clusters are very close, the net effect is a loss in performance for almost all systems and conditions.
-In this setting, the best MDMT in our pool (MDL-Res) is no longer able to surpass the \system{Mix-Nat} baseline.%\fyTodo{Note about scaling?}\fyTodo{More comparison?}
+In this setting, the best MDMT in our pool (\system{MDL-Res}) is no longer able to surpass the \system{Mix-Nat} baseline.%\fyTodo{Note about scaling?}\fyTodo{More comparison?}
 
@@
     \system{DC-Feat}      & 37.3  & 54.2 & 49.3   & 33.6 & \SW{41.9} & \SW{75.8} & 40.8 & \SW{48.7}  \\% & \SW{21.7} \\
-    \system{LDC-Tag}     & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 39.1 & \SW{48.3}         \\% & 22.1 \\ 
+    \system{LDR}     & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 39.1 & \SW{48.3}         \\% & 22.1 \\ 
     \system{TTM}            & 37.4 & \SW{53.7} & 48.9 & 32.8 & 41.3 & \SW{75.8} & 40.7 & \SW{48.3}   \\% &  23.4 \\
@@
     \system{TTM}         &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\ 
-    \system{LDC-Tag}  &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\
+    \system{LDR}  &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\
 

### 2020-05-13 2020/coling2020/multidomain.tex
related work

@@
 
-\newcite{Farajian17multidomain} does not use domain tags in training.
-The processing of each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some epochs, before delivering its output.
+Finally, \newcite{Farajian17multidomain,li-etal-2018-one,xu2019} follow a different strategy.
+The processing of each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some iterations, before delivering its output.
 This approach entirely dispenses with the notion of domain and relies on data selection techniques to handle data heterogeneity.
-\jcTodo{cite Xu et al 2019, Lexical Micro-adaptation for Neural Machine Translation}
 

### 2020-05-13 2020/coling2020/multidomain.tex
related work
wrong domain difference

--- 2020-05-13 2020/coling2020/multidomain.tex
@@
 In this paper, we formulate in a more precise fashion the requirements that an effective MDMT system should meet (section~\ref{sec:requirements}).
-Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT systems and associated measurements to evaluate them (section~\ref{sec:challenging}).
+Our first contribution is thus of methodological nature and consists of lists of\jcTodo{enumerate?} expected properties of MDMT systems and associated measurements to evaluate them (section~\ref{sec:challenging}).
 In doing so, we also shed lights on new problems that arise in this context, regarding for instance the accomodation of new domains in the course of training, or the computation of automating domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
@@
 \paragraph{[P4-ERR]}
-A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, it has been observed that could yield ``catastrophic'' drops in performance\jcTodo{cite \cite{McCloskey89catastrophic}}.
+A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, it has been observed that could yield ``catastrophic'' drops in performance\jcTodo{cite needed?}.
 This requirement can be measured by translating test texts for domain $A$ with erroneous or random domain tags and measure the correlated loss.
@@
     & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
-    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5} & -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& \SW{-21.8} & \SW{-3.3}\\% [32.5]
-    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -0.4 & -0.4 & -0.5 & -0.4 & \SW{-14.2} & \SW{-1.7}\\% [35.9]
-    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & \SW{-15.0} &\SW{-1.8}\\ % [34.9] 
-    \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-12.8} & \SW{-1.4}\\ % [37.0]
+    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5} & -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& \SW{} & \SW{-3.3}\\% [32.5]
+    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -0.4 & -0.4 & -0.5 & -0.4 & \SW{-13.4} & \SW{-1.7}\\% [35.9]
+    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & \SW{-14.21} &\SW{-1.8}\\ % [34.9] 
+    \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-11.99} & \SW{-1.4}\\ % [37.0]
     \system{TTM} (*)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & 0.0 & -0.1\\
@@
     \system{WDCMT} (*)     & -0.0    & -0.0  & +0.2 & +0.2  & +0.8 & +0.8  & -0.4 & -0.8 & 0.0 & +0.2 \\
-    \system{MDL Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +\SB{0.7} & -0.3 & \SW{-20.1} & \SW{-1.3}\\ % [31.8]
+    \system{MDL Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +\SB{0.7} & -0.3 & \SW{-18.6} & \SW{-1.3}\\ % [31.8]
     \hline

### 2020-05-13 2020/coling2020/multidomain.tex

@@
 
-\newcite{Farajian17multidomain} does not use domain tags in training.
-The processing of each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some epochs, before delivering its output.
+Finally, \newcite{Farajian17multidomain}, \newcite{li-etal-2018-one} and \newcite{xu2019} follow a different strategy.
+The processing of each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some iterations, before delivering its output.
 This approach entirely dispenses with the notion of domain and relies on data selection techniques to handle data heterogeneity.
-\jcTodo{cite Xu et al 2019, Lexical Micro-adaptation for Neural Machine Translation}
 

### 2020-05-13 2020/coling2020/multidomain.tex
wrong domain scores
fix wrong commit

--- 2020-05-13 2020/coling2020/multidomain.tex
@@
 \paragraph{[P4-ERR]}
-A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, catastrophic' drops in performance can be observerd\fyDone{cite \cite{McCloskey89catastrophic}}.
+A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, catastrophic drops in performance can be observerd\fyDone{cite \cite{McCloskey89catastrophic}}.
 This requirement can be measured by translating test texts for domain $A$ with erroneous or random domain tags and measure the correlated loss.
@@
     & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
-    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5} & -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& \SW{-21.8} & \SW{-3.3}\\% [32.5]
-    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -0.4 & -0.4 & -0.5 & -0.4 & \SW{-14.2} & \SW{-1.7}\\% [35.9]
-    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & \SW{-15.0} &\SW{-1.8}\\ % [34.9] 
-    \system{LDR}           & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-12.8} & \SW{-1.4}\\ % [37.0]
+    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5} & -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& \SW{-19.6} & \SW{-3.3}\\% [32.5]
+    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -0.4 & -0.4 & -0.5 & -0.4 & \SW{-13.4} & \SW{-1.7}\\% [35.9]
+    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & \SW{-14.2} &\SW{-1.8}\\ % [34.9] 
+    \system{LDR}           & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-12} & \SW{-1.4}\\ % [37.0]
     \system{TTM} (*)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & 0.0 & -0.1\\
@@
 
-    \system{LDC-Tag}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
+    \system{LDR}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
                     & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize -0.4 & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize +\SB{3.8} & \scriptsize \SW{-0.9} & \scriptsize +0.5 & \scriptsize -0.4 \\
@@
 
-Finally, \newcite{Farajian17multidomain,li-etal-2018-one,xu2019} follow a different strategy.
+Finally, \newcite{Farajian17multidomain}, \newcite{li-etal-2018-one} and \newcite{xu2019} follow a different strategy.
 The processing of each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some iterations, before delivering its output.

### 2020-05-15 2020/coling2020/multidomain.tex
Final version - for now

@@
 \title{Revisiting Multi-Domain Machine Translation}
-
+\fyFuture{A plea for more thoroughness in evaluation}
 \author{First Author \\
@@
 \begin{abstract}
-  When building machine translation systems, it is often the case that one need to make the best out of the heterogeneous set of parallel data in training, and to robustly handle inputs from unexpected domains in testing.
+  When building machine translation systems, one often needs to make the best out of heterogeneous sets of parallel data in training, and to robustly handle inputs from unexpected domains in testing.
 This multi-domain scenario has attracted a lot of recent work, that fall under the general umbrella of transfer learning.
@@
 These methods thus only use domain tags in training.
-\item the multi-domain model of \newcite{Zeng18multidomain} (\system{WDCMT}), where a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly, improved results.}\fyDone{Check this}
+\item the multi-domain model of \newcite{Zeng18multidomain} (\system{WDCMT}), where a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly improved, results.}\fyDone{Check this}
 \item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connexions make it possible to by-pass this module.
@@
 
-\fyTodo{Numbers for the rand domain analysis}In the last two columns of Table~\ref{tab:redomains}, we report the drop in performance when the domain information is not correct.
+\fyDone{Numbers for the rand domain analysis}In the last two columns of Table~\ref{tab:redomains}, we report the drop in performance when the domain information is not correct.
 In the first (\domain{rnd}), we use test data from the domains seen in training, presented with a random domain tag\fyDone{all configs use the same tag or is randomization is run for each config}.
@@
 Averages only take into account the six initial domains (\domain{News} excluded).
-Underline denotes a significant loss between warmstart and coldstart, bold for a significant gain.}
+Underline denotes a significant loss between warmstart and coldstart, bold a significant gain.}
   \label{tab:warmrestart}
@@
     \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 39.5 & 47.1\\% & 23.3 \\
-    \system{WDCMT}       & 35.57 & 53.08 & 48.43 & 30.55 & \SW{37.73} & \SW{56.02} & 38.5 & 43.6 \\ % & 20.4 \\ 
+    \system{WDCMT}       & 35.6 & 53.1 & 48.4 & 30.5 & \SW{37.7} & \SW{56.0} & 38.5 & 43.6 \\ % & 20.4 \\ 
     \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SB{50.7}\\%  & \SW{21.2} \\
@@
   \end{tabular}
-  \caption{Translation performance with automatic domains, computed with the original test sets.}
+  \caption{Translation performance with automatic domains, computed with the original test sets. Significancy tests are for comparisons with the 6-domain scenario.}
   \label{tab:subdomains}
@@
 Finally, \newcite{Farajian17multidomain}, \newcite{li-etal-2018-one} and \newcite{xu2019} follow a different strategy.
-The processing of each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some iterations, before delivering its output.
+Each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some iterations, before delivering its output.
 This approach entirely dispenses with the notion of domain and relies on data selection techniques to handle data heterogeneity.
@@
 \section{Conclusion and outlook \label{sec:conclusion}}
-\fyDone{Write conclusions}\fyTodo{no mixture of experts}
+\fyDone{Write conclusions}\fyFuture{no mixture of experts}
 In this study, we have carefully reconsidered the idea of multi-domain machine translation, which seems to be taken for granted in many recent studies.
-We have spelled out the various motivations for building such systems, and associated expectations in terms of system performance.
-We have then designed a series of requirements that MDMT systems should meet, and designed associated testing procedures.
-In our experiments with multiple MDMTs, we have found that most requirements were not met for our experimental conditions.
+We have spelled out the various motivations for building such systems and the associated expectations in terms of system performance.
+We have then designed a series of requirements that MDMT systems should meet, and proposed a series of test procedures.
+In our experiments with a representative sample of MDMTs, we have found that most requirements were hardly met for our experimental conditions.
 If MDMT systems are able to outperform the mixed-domain baseline, at least for some domains, they all fall short to match the performance of fine-tuning on each individual domain, which remains the best choice in multi-source single domain adaptation.
-In this comparison, architecture-based strategies seem to show the best promises.
-\fyTodo{Mitigate this}
-As expected however, MDMT are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend,
-dynamically accommodate additional domains:this is especially easy for feature-based methodologies.
+In this scenario, architecture-based strategies aimed at parameter sharing and/or representation normalization obtain the best results.
+\fyDone{Mitigate this}
+As expected however, MDMTs are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend, dynamically accommodate additional domains:this is especially easy for feature-based methodologies.
 Our experiments finally suggest that all methods show decreasing performance when the number of domain increases (for a fixed data size).
 
-Two other main conclusions can be drawed from this study: first, it seems that more work is needed to make MDMT systems make the best out of the variety of the available data, both to effectively share what needs to be shared while at the same time separating what needs to be separated.
+Two other main conclusions can be drawed from this study: first, it seems that more work is needed to make MDMT systems make the best out of the variety of the available data, both to effectively share what needs to be shared while at the same time separating what needs to be kept separated.
 We notably see two areas worthy of further explorations: the development of parameter sharing strategies when the number of domains is large; the design of training strategies that can effectively handle a change of the training mixture, including an increase in the number of domains.
 Both problems are of practical relevance in industrial settings.
-Second, and may be more importantly, there is a general need to adopt better evaluation methodologies for evaluating MDMT systems, require systems developers to clearly spell out the testing conditions and the associated expected distribution of testing instances, and to report more than mere comparisons with simple baselines on a fixed and known pair of domains.
+Second, and may be more importantly, there is a general need to adopt better evaluation methodologies for evaluating MDMT systems, which require systems developers to clearly spell out the testing conditions and the associated expected distribution of testing instances, and to report more than mere comparisons with simple baselines on a fixed and known handful of domains.
 
@@
 
+\end{document}
+
 \todos{}
-
-\end{document}
 

### 2020-05-15 2020/coling2020/multidomain.tex
significative test not correct compared to excel table
batch size of 128 to 12228 tokens

@@
 \documentclass[11pt,a4paper]{article}
-% \usepackage{coling2020}
 \usepackage{tacl2018v2}
@@
 % The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training use a batch size of~128; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
+Training use a batch size of~12288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
  % All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
@@
     \system{WDCMT} (*)     & -0.0    & -0.0  & +0.2 & +0.2  & +0.8 & +0.8  & -0.4 & -0.8 & 0.0 & +0.2 \\
-    \system{MDL Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +\SB{0.7} & -0.3 & \SW{-18.6} & \SW{-1.3}\\ % [31.8]
+    \system{MDL Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +0.7 & -0.3 & \SW{-18.6} & \SW{-1.3}\\ % [31.8]
     \hline

### 2020-06-03 2020/phd_midterm_report/report.tex
phd midterm report

\documentclass[12pt,a4paper,twoside]{report}
\usepackage[left=2cm,right=2cm,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{titlesec}
\titleformat{\chapter}[display]   
{\normalfont\huge\bfseries}{\chaptertitlename\ \thechapter}{20pt}{\Huge}   
\titlespacing*{\chapter}{0pt}{-55pt}{40pt}
\titlespacing*{\section}{0pt}{\baselineskip}{\baselineskip}
\vspace{-\baselineskip}
\usepackage{pagenote}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}	% Para caracteres en español
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{multirow,booktabs}
\usepackage[table]{xcolor}
\usepackage{fullpage}
\usepackage{lastpage}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{mathrsfs}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{calc}
\usepackage{color,soul}
\usepackage{multicol}
\usepackage{cancel}
\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{xcolor}
\colorlet{shadecolor}{orange!15}
\parindent 0in
\parskip 12pt
\geometry{margin=1in, headsep=0.25in}
\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{reg}{Rule}
\newtheorem{exer}{Exercise}
\newtheorem{note}{Note}
\usepackage{listings}
\usepackage{spverbatim}
\usepackage{fancyvrb}
\usepackage{hyperref}
\usepackage{float} 
\usepackage{natbib}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{soulutf8}
\usepackage{color,soul}
\usepackage{xcolor}
\usepackage{tabularx}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{definition}{Definition}[section]
\DeclareMathOperator*{\argmax}{argmax}
\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
\newcommand\minus{\mathop{\mbox{$\mathit{.e}$-}}}
\usepackage[draft]{todo}
\usepackage{mathtools,xparse}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\NewDocumentCommand{\normL}{ s O{} m }{%
  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{L_1(\Omega)}%
}
\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\pgfkeys{
    /tr/rowfilter/.style 2 args={
        /pgfplots/x filter/.append code={
            \edef\arga{\thisrow{#1}}
            \edef\argb{#2}
            \ifx\arga\argb
            \else
                \def\pgfmathresult{}
            \fi
        }
    }
}

\begin{document}
\setlength{\belowdisplayskip}{8pt} \setlength{\belowdisplayshortskip}{8pt}
\setlength{\abovedisplayskip}{8pt} \setlength{\abovedisplayshortskip}{8pt}
\setlist{nosep}
\setlength{\parskip}{0.1cm}
\setlength{\parindent}{1em}
\section*{Introduction}
Neural Machine Translation (will be referred to as NMT in the rest of the article) has achieved the state of the art performance in Automatic Translation since 2015 and became the main architecture in industrial development and natural language processing research. Neural Machine Translation’s quality is assured both theoretically and empirically under the assumption of the matching between training data distribution and testing data distribution. However, this situation is hardly true in real contexts where the NMT model is often trained with big training data which is heterogeneous in terms of topic and size of the size of topics data, and one needs to make the best to robustly handle inputs from unexpected domains in the testing phase.
\subsection*{Overview}
Data-based Machine Translation whether statistical or neural is based on well-studied machine learning principles. Given a training set, $\mathbf{S_{trn}}$ drawn from a distribution $\mathbf{D_s}$ consisting of sentence pairs $(f,e)$, the parameter of the probabilistic model $P_{\theta}$ is optimized by minimizing empirical expectation of a loss function
$$l(h_{\theta}(f,e)) = \frac{1}{N}\displaystyle{\mathop{\sum_{(f,e)\in \mathbf{S_{trn}}}l(h_{\theta}(f,e))}}$$
Then in the testing phase, given an example drawn from the same underlying distribution $\mathbf{D_s}$, the accuracy of the trained model is assured by a statistical upper bound composed of 2 terms as empirical training loss and the capacity of hypothesis space $h_{\theta}$ divided by a factor of training set size. However, if testing samples are drawn from a different source, the upper bound can no longer be applied.

In this thesis, we are interested in several problems arising in multi-domain contexts
\begin{itemize}
	\item Multi-domain Adaptation.
		\begin{itemize}
			\item with labeled data.
			\item with unlabeled data (unsupervised adaptation).
		\end{itemize}
	\item Robustness to out of domain data.
\end{itemize}

The first problem was first studied under the concept of Domain Adaptation where one aims to improve the poor performance of the model trained in a low resource domain by leveraging different high resource domains. In general, the methods for Domain Adaptation problem can be categorized into 2 classes: data-based methods and model-based methods.

\subsection*{State of the art}

\subsubsection*{Data-based adaptation methods}
Methods in this category improve the low resource model by collecting domain-related examples from out-of-domain resources. For the statistical model, \cite{mansour12simple} investigated the use of the Moore-Lewis score to re-estimate the phrase table.\cite{Duh13selection,Axelrod11domain,silva2018extracting} proposed several methods to select examples for training NMT model by scoring based on the cosine similarity of distributed sentence embedding or based on the probability calculated by in-domain Neural Language Model and the cosine similarity of $tf\_idf$ embedding adjusted by Feature decay algorithm. Another way to get more training is to create pseudo parallel data, \cite{Junjie19domain} proposed an unsupervised method to create pseudo in-domain training data using word by word translation.

\subsubsection*{Model-based adaptation methods}
These approaches take interest in the model architecture and model training process. We can further categorize them into 3 classes: feature-augmentation, module-augmentation, surrogate losses, and finetuning. In the class of feature-augmentation, \cite{Kobus17domain}, \cite{kenji17multi}, \cite{Pham19generic} proposed to add domain-tags and learn-able domain-embeddings to existing word feature of NMT model. These methods are cheap and efficient in terms of gain over model-size.

Module-augmentation is more expensive then feature-augmentation because one adds one or several complementary layers into standard NMT architecture; \cite{bapna19simple},\cite{Michel18extreme},\cite{Zheng18multi},\cite{Vilar18learning}, \cite{Britz2017mixing} are included in this category. Residual module in \cite{bapna19simple} and Hidden units in \cite{Vilar18learning} (\ref{fig:overview}) are used as a plug-in which is used to fine-tune the model without
changing the value of the main model. \cite{Zheng18multi} (\ref{fig:overview}) added 2 gates to extract domain agnostic representations and domain-specific representations from the output of encoder which are then plugged to the decoder. \cite{Britz2017mixing} (\ref{fig:overview}) also plugged a domain-classifier on top of the encoder to force the domain-specific representation. \cite{domhan2017using} added a Language Model-like module between target embedding and the decoder so that the decoder can benefit from the monolingual data of the target domain in the target language. The architecture is shown in figure (\ref{fig:overview}). \cite{dakwle17fine} (\ref{fig:overview}) proposed a multi-output layer for domain-model which are trained by minimizing the cross-entropy for in-domain examples and Kullback-Leibler distance to
the conditional distribution of a teacher generic model. The proposed model has 2 output layers, each of them optimizing one of 2 losses respectively. \cite{Dou19Unsupervised} attempted to use only monolingual target data for adapting to a new domain by using additional learnable domain embedding and task embedding and introducing language modeling tasks in the target language.

In machine learning theory and practice, surrogate losses are used to stabilize the training and to improve model generalization. In Domain Adaptation NMT, for example, \cite{Chen17cost} used domain-classification loss in addition to cross-entropy loss to leverage out of domain data while adapting to the target domain.

The last is fine-tuning category according to which models are pretrained with out-of-domain data, then continue being trained with in-domain data during a few training epochs. \cite{Thomas11limsi}  proposed this practice to adapt the neural language model in the statistical Machine Translation model to a new domain. \cite{Luong15stanford} used fine-tuning to adapt a neural machine translation model to a new domain. while the model is easily improved for the in-domain test, it loses dramatically quality in the previous training domain. This phenomenon has been reported as Catastrophic Forgetting by \cite{Michael89catastrophic}. In the MT context, \cite{brian19overcoming}, \cite{khayrallah2018regularized} added
a normalization term to cross-entropy loss in order to retain the generalization of the model while being adapted to target domain. \cite{Farajian17multidomain}, \cite{li2018one} proposed to use the on-the-fly fast fine-tuning on a few related examples in the decoding phase. According to on-the-fly strategy, given a sentence to be translated, first, a retrieval system searches most related training examples. Then the new NMT model is built by fine-tuning a pretrained model on the retrieved examples. Finally, the instantly fine-tuned model is used to translate the given sentence.

\begin{figure}[h]
\includegraphics[scale=0.3]{DA_overview.png}
\caption{Domain Adaptation MT's overview}
\label{fig:overview}
\end{figure}

\subsection*{Realized work}
The work of \cite{Kobus17domain} and \cite{Daume07frustratingly} showed promising results for learning in a multi-domain context. Inspired by the simplicity of the methods, we extend their works to a new version of domain embedding. In this method, each word has its domain embedding which amounts to a small share of the word embedding.

Multi-domain learning can be considered as a multi-task problem in which prediction on each domain is a task. \cite{Caruana97multitask} suggested that multi-task training can help to obtain a better generalization of a learner. A multi-task learner consists of one or stack of multiple shared lower layers and task-related modules above these later layers. By optimizing k objective functions at once, the shared layers are pulled to a compromising region between task-optimal regions. However, multi-task training has to overcome several problems such as the heterogeneity of data size, convergence rates, etc. We investigated a multi-task system based on a Transformer architecture \cite{Vaswani17attention} in a large scale multi-domain context. We reported promising results and proposed another version more adaptive.

Many methods have been proposed to deal with multi-domain contexts but they were mostly evaluated in a very simple context which is not realistic. We took interest in designing new evaluation setups that access several desirable properties of a multi-domain system. We revisited the work of \cite{Britz17effective} \cite{Zheng18multi} \cite{Kobus17domain} and \cite{bapna19simple} in a large scale context in which there are 6-7 domains.

\section*{Lexicalized Domain Representation}
\cite{Kobus17domain} proposed to stack a domain-feature embedding to each word embedding whenever the domain appears. This way, the model pays attention to the domain from which each sample originates. However, words do not equally require domain adaptation because some of them are domain-invariant such as the articles. This property can
be handled by equipping each word its domain feature, and because these additional features are small (4 or 8 units), the method is still cheap.
\begin{figure}[h!]
  \center
  \includegraphics[width=0.48\textwidth]{embeddings}
  \caption{Lexicalized domain embeddings. When processing a sample from domain $2$, we only activate the corresponding parameter region ($\theta_2$) in the input embeddings; the remaining domain-specific parts are zeroed out and do not receive any update. The generic part is always active and is updated irrespective of the input domain.} 
  \label{fig:network}
\end{figure}

To evaluate the performance of the proposed lexicalized domain representation, experiments were performed on 2 language pairs: English $\rightarrow$ French and English $\rightarrow$ German with 3 corpora: EMEA, Europarl and ECB. The statistics of the corpora are shown in the table \ref{tab:Corpora3}
\begin{table}[h!]
  \centering
  \begin{tabular}{ |llll|} %*{4}{|r|}}
    \hline
    Corpus & Train & Valid & Test \\ 
    \hline
    \multicolumn{4}{l}{English $\rightarrow$ French }\\
    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
    \hline
    EMEA  & 1.09M & 1,000 & 1,000$_{(300)}$\\
    ECB    & 0.19M & 1,000 & 1,000     \\
    EPPS   & 2.01M  & 1,000 & 1,000  \\
    IT         & 0.54M  & 1,000 & 1,000 \\  
    \hline
    \multicolumn{4}{l}{English $\rightarrow$ German}\\
    %\multicolumn{4}{|l|}{Vocab size -  En:30,159, De: 30,698}\\ 
    \hline
    EMEA  & 1.11M & 1,000 & 1,000$_{(300)}$ \\
    ECB     &  0.11M & 1,000 & 1,000  \\
    EPPS   & 1.92M & 1,000 & 1,000 \\ 
    \hline
\end{tabular}
\caption{Corpora statistics.}
\label{tab:Corpora3}
\end{table}
\begin{table}[h!]
\parbox{.42\linewidth}{
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|lcc|c|}
\hline
Model & EMEA & EPPS & ECB & Avg. \\
\hline
%\multicolumn{8}{l}{} \\[-9pt]  
\multicolumn{5}{l}{English$\rightarrow$French} \\
\hline
$\mathtt{Mixed}$          & 67.69$_{47.60}$ & 37.50 & 53.49 & 52.89\\
\hline
$\mathtt{FT}_{EMEA}$ & 76.77$_{49.43}$ & 17.16 & 11.99 & 35.30\\
$\mathtt{FT}_{EPPS}$  & 20.86 & 37.04 & 24.53 & 27.47\\
$\mathtt{FT}_{ECB}$    & 26.93 & 27.09 & \bf 56.52 & 36.84\\
\hline
$\mathtt{DC}$              & 67.87$_{45.42}$ & 37.31 & 54.14 & 53.10\\
\hline
$\mathtt{LDR}_{oracle}$            & 74.26$_{\bf 49.90}$ & 37.67 & 54.07 & 55.33\\
$\mathtt{LDR}_{oracle}^{0.5}$   & \bf 74.95$_{49.38}$ & 37.35 & 55.91 & \bf 56.07\\
%$\mathtt{LDR}_{generic}$          & 73.97$_{49.54}$ & 37.81 & 53.67 & \\
$\mathtt{LDR}_{pred}$               & 74.29$_{49.84}$ & \bf 37.73 & 54.01 & 55.34\\
$\mathtt{LDR}_{wrong}$            & 72.95$_{49.78}$ & 37.62 & 53.35 & 54.64\\
  \hline
%\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{5}{l}{English$\rightarrow$German} \\
\hline
$\mathtt{Mixed}$         & 64.57$_{42.99}$ & 26.47 & 68.67 & 53.23\\
\hline
$\mathtt{FT}_{EMEA}$ & 68.35$_{42.97}$ & 17.02 & 32.87 & 39.41\\
$\mathtt{FT}_{EPPS}$  & 36.19 & 26.29 & 40.71 & 34.39\\
$\mathtt{FT}_{ECB}$   & 24.72 & 18.36 & \bf 74.05 & 39.04\\
\hline
$\mathtt{DC}$ & 63.48$_{42.98}$ & 26.27 & 66.95 & 52.23\\
\hline
$\mathtt{LDR}_{oracle}$ & 70.90$_{\bf 46.12}$ & 26.30 & 68.90 & 55.36\\
$\mathtt{LDR}_{oracle}^{0.5}$ & \bf 71.31$_{45.23}$ & 25.98 & 73.74 & \bf 57.01\\
%$\mathtt{LDR}_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & \\
$\mathtt{LDR}_{pred}$ & 70.89$_{\bf 46.12}$ & \bf 26.53 & 68.63 & 55.35\\
$\mathtt{LDR}_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & 54.22\\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BLEU scores for  Transformer systems \label{tab:results-trsf}}
}
\hfill
\parbox{.42\linewidth}{
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|lcc|c|}
\hline
Model & EMEA & EPPS & ECB & Avg. \\
\hline
%\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{5}{l}{English$\rightarrow$French} \\
\hline
$Mixed$                         & 65.42$_{45.11}$ & 34.70 & 51.38 & 50.50\\
\hline
$\mathtt{FT}_{EMEA}$   & 72.06$_{\bf 47.33}$ & 18.62 & 16.78 & 35.82\\
$\mathtt{FT}_{EPPS}$    & 35.47$ $ & 34.61 & 39.56 & 36.55\\
$\mathtt{FT}_{ECB}$      & 21.93$ $ & 22.60 & 51.53 & 32.02\\
\hline
$\mathtt{DC}$                            & 68.26$_{43.76}$ & 35.13 & 50.09 & 51.16\\
% \hline
%$\mathtt{WDCMT}$           & 68.76$_{45.29}$ & 35.71 & 52.75 & \\
\hline
$\mathtt{LDR}_{oracle}$            & 71.73$_{46.30}$ & \bf 35.21 & 50.91 & 52.62\\
%$\mathtt{LDR_{condgru}}_{pred}$   & 71.7$_{46.21}$ & 35.09 & 51.22 & \\
$\mathtt{LDR}_{oracle}^{0.5}$   & 71.70$_{46.41}$ & 34.24& \bf 52.37 & \bf 52.77\\
%$\mathtt{LDR}_{generic}$          & 70.59$_{46.37}$ & 35.22 & 49.68 & \\
$\mathtt{LDR}_{pred}$               &\bf 72.76$_{46.35}$ & 35.10 & 50.38 & 52.75\\
$\mathtt{LDR}_{wrong}$            & 62.10$_{43.29}$ & 34.17 & 48.79 & 48.35\\
\hline
%\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{5}{l}{English$\rightarrow$German} \\
\hline
$Mixed$                        & 57.37$_{37.94}$ & 23.10 & 63.54 & 48.00\\
\hline
$\mathtt{FT}_{EMEA}$  & 65.64$_{\bf 44.71}$ & 12.36 & 15.93 & 31.31\\
$\mathtt{FT}_{EPPS}$   & 24.90$ $ & 22.98 & 26.26 & 24.71\\
$\mathtt{FT}_{ECB}$     & 41.80$ $ & 15.97 & 71.07 & 42.95\\
\hline
$\mathtt{DC}$                & 62.53$_{39.25}$ & \bf 23.74 & 65.71 & 50.66\\
%\hline
%$\mathtt{WDCMT}$ & 64.05$_{38.54}$ & 22.82 & 67.00 & \\
\hline
$\mathtt{LDR}_{oracle}$     & \bf 63.43$_{40.04}$ & 22.66 & 64.40 & 50.16\\
$\mathtt{LDR}_{oracle}^{0.5}$   & 63.27$_{38.16}$ & 21.83 & \bf 69.55 & \bf 51.55\\
%$\mathtt{LDR}_{generic}$   & 63.27 $_{39.75}$ & 22.62 & 63.70 & \\
$\mathtt{LDR}_{pred}$        & 63.17$_{39.92}$ & 22.51 & 64.00 & 49.89\\
$\mathtt{LDR}_{wrong}$     & 56.84$_{37.05}$ & 22.06 & 61.66 & 46.85\\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BLEU scores for RNN systems\label{tab:results-rnn}}
}
\end{table}
\begin{table}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|c|}
\hline
Model & EMEA & EPPS & ECB & Avg. \\
\hline
%\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{5}{l}{English$\rightarrow$French} \\
\hline
$\mathtt{LDR}_{pred}$                   & \bf 72.76$_{46.35}$ & 35.10 & 50.38 & \bf 52.75\\
$\mathtt{LDR}_{pred}^{condgru}$  & 71.70$_{46.21}$ & 35.09 & 51.22 & 52.67\\
$\mathtt{WDCMT}$                        & 68.76$_{45.29}$ & \bf 35.71 & \bf 52.75 & 52.40\\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BLEU scores for RNN systems. Comparison between \texttt{WDCMT} and $\texttt{LDR}_{pred}$ built using conditional GRUs.\label{tab:results-rnn-wdcmt}}
\end{table}

The baseline models consist of standard NMT model (RNN or Transformer) trained with mixing of 3 corpora; and in-domain fine-tuned models; and the last is the method proposed by \cite{Zheng18multi}. In the Transformer setting, LDR improves by a large margin the performance on domain medical and achieves on-par performance compared to the WDCNMT of \cite{Zheng18multi} on RNN architecture. More detailed results have been published in \cite{Pham19generic}.

\section*{Revisiting multi-domain NMT}
Even though many methods are proposed to solve the multi-domain problem, they were not fully evaluated. We introduced several new objectives that the multi-domain NMT system is supposed to achieve. The following are basic properties we want an NMT model to have.
\begin{enumerate}
	\item \label{ax:1} Robustness to unbalancing data.
	\item \label{ax:2} Ability to learning from other domains to improve one domain.
	\item \label{ax:3} Robustness to new domain.
	\item \label{ax:4} Ability to handle unlabeled training data (eg via clustering).
	\item \label{ax:5} Ability to perform continuous learning.
\end{enumerate}

The first property is basic because, in the real context, in-domain corpora often have a very variable size. Indeed, among OPUS corpora, the religion domain with only 130K parallel sentences (Tanzil) is very small compared to the medical domain with more than 2.5M parallel sentences (UFAL). In general, the practice of mixing corpora to train NMT down-samples the small corpora and results in low performance for these domains.

The second is the essence of multi-domain NMT: that each domain in the mixture can benefit to the data from the other domains.

The third property is needed to deal with unknown domains in the test phase. This situation is very common in real contexts where test data can originate from any domains.

The fourth property is needed to deal with training data without domain information. For example, a corpus such as Paracrawl does not contain any specific topic or domain information.

The last property is also an important and active trend in machine learning. The model needs to preserve the learned knowledge while learning as being exposed to new data.

We use the following 5 setups in our study:
\begin{enumerate}
	\item \label{setup:1} Variable-size in-domain corpora in training sets.
	\item \label{setup:2} In-domain and Out-domain test sets.
	\item \label{setup:3} Fuzzy domain separation.
	\item \label{setup:4} Training in considering unsupervised clusters as domains.
	\item \label{setup:5} Continuation of training with new domain data.
\end{enumerate}

According to our study on the state of the art of Multi-domain NMT, this is the first large scale experiment. We experimented with translation from English into French and use texts initially originating from 6 domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in an \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}). Complementary experiments \ref{setup:2} and \ref{ax:5}  also use v12 of the News Commentary corpus (\domain{news}). Corpus statistics in Table~\ref{tab:Corpora7}. Most corpora are available from the Opus web site. \footnote{\url{http://opus.nlpl.eu}}. The heterogeneity of the domain’s population is assured by the choice of in-domain corpora, therefore the property \ref{ax:1} is evaluated. (see table \ref{tab:Corpora7})

\begin{table*}[htbp]
  \centering
  \begin{tabular}{ |lllllll|} %*{4}{|r|}}
    \hline
    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
    \domain{med} & \domain{law} & \domain{bank} & \domain{it} & \domain{talk} & \domain{rel} & \domain{news} \\
    \hline
    2609 (0.68) & 190 (0.05)  & 501 (0.13) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
    \hline
  \end{tabular}
\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain). \domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.}
\label{tab:Corpora7}
\end{table*}

We used a test set split from the original corpus of \domain{news} to evaluate the robustness before an unknown domain. For methods requiring domain-information to predict, given a source sentence, we predict the closest domain to the sentence among already known domains.

In setup \ref{setup:3}, we randomly \emph{split} one corpus into two parts and proceed as if this corresponded to two actual domains. An MD system should detect that these two pseudo-domains are mutually beneficial, and should be hardly affected by this change with respect to the baseline scenario (no split). This setup evaluates the property \ref{ax:2}.

We also merge two corpora in training, in order to assess the robustness with respect to heterogeneous domains. We then translate the two corresponding tests with the same system. The following are 5 setups we used in the study.
\begin{itemize}
	\item Randomly splitting med data into 2 equal-sized sets.
 	\item Randomly splitting med data into 2 sets with proportions: $\frac{3}{4}$, $\frac{1}{4}$ respectively.
	\item Randomly splitting bank data into 2 equal-sized sets.
	\item Randomly splitting law data into 2 equal-sized sets.
	\item Merging bank, law into one pseudo-domain.
\end{itemize}

Setup \ref{setup:4} is used to evaluate property \ref{ax:4}. In this setup, we used k-means clustering to cluster sentence embeddings for the whole training into 30 clusters and considered them as pseudo domains. We then performed Multi-domain NMT methods with these pseudo domains. To translate a sentence, we predict the cluster which is closest to the sentence by comparing the distance between sentence embedding and the cluster means.

The last setup \ref{setup:5} is designed to access the ability of incremental learning \ref{ax:5}. We choose \domain{news} as the 7-th domain which arrives after training with 6 domains \domain{med}, \domain{bank}, \domain{law}, \domain{rel}, \domain{talk}. We then compared the performance of the model trained by continuing training and of the model trained from scratch with 7 domains.

\section*{Multi-task network}
We can approach the multi-domain problem by multi-task learning by considering adaptation to one domain as a task. \cite{Caruana97multitask} showed multi-task training improving the performance of the model for each independent task. \cite{rebuffi18efficient} proposed the first use of a residual adapter which is a small layer plugged on top each layer to adapt classifier to an arbitrary domain. \cite{bapna19simple} reapplied the latter idea to domain adaptation in NMT. Inspired by the mentioned work in multi-task learning and domain adapter, we applied residual adapters to encoder and decoder and train the whole set of parameters from scratch. There is a small difference, instead of plugging the adapter in a sequential scheme, we propose a high-way scheme in which each adapter bridges directly from each intermediate layer to the output of the encoder. This scheme preserves the integrity of the stack of self-attention layers.

Multi-task training for K domains involves K losses on K empirical distributions. 
\begin{equation}
\begin{split}
L(\theta_{s},\theta_{i \in [1..K]}) = \frac{1}{K} \displaystyle{\mathop{\sum}_{i=[1..K]} \mathop{\sum}_{(x_i^{k},y_i^{k}) \in D_i}L(x_i^{k},y_i^{k},\theta_s,\theta_i)}
\end{split}
\label{loss:multi}
\end{equation}

We trained end-to-end the loss \ref{loss:multi}, and overcame the heterogeneity of population of the domains by using "natural" batch sampling frequency. More precisely, for each batch of stochastic training, we sample B examples from an in-domain corpus, the frequency of being picked of each corpus is proportional to the size of the corpus.
$$freq_i = \frac{\#D_i}{\displaystyle{\mathop{\sum}_{j=[1..K]}\#D_j}}$$
To avoid over-fitting for the low resource domain, we used layer activation regularization, i.e, the actual training loss is 
\begin{equation}
\begin{split}
\bar{L}(\theta_{s},\theta_{i \in [1..K]}) = L(\theta_{s},\theta_{i \in [1..K]}) + \displaystyle{\mathop{\sum}_{i \in [1..K]} \mathop{\sum}_{j \in [1..6]} \lambda_i * \normL{Adap^{j}_i(h_j)}}
\end{split}
\label{loss:multi-reg}
\end{equation}
where $\lambda_i$ is regularization scale for the adapter of domain i, $Adapt^j_i$ is the adapter for layer j in domain i. We could also adjust the capacity of each adapter to match the size of the domain training data. The adapter is composed of 5 layers:
$$Adapt(h_i) = W^{d_{model}}_{d_{Bottleneck}} * (RELU( Layernorm ( W^{d_{Bottleneck}}_{d_{model}} * Layernorm(h_i) + b_{d_{Bottleneck}} ) )) + b_{d_{model}}$$
We chosen $d_{Bottleneck}$ by grid search on set of values $\{128, 256,512,1024\}$ and $\lambda_i$ in by grid search on set of values $\times^{K}[1\minus{3}, 1\minus{4},1\minus{5},1\minus{6}]$. Optimal values of $\lambda$ could be heuristically localized by using the rule of thumb: bigger domains should yield larger $\lambda$.
\begin{figure}[h!]
\parbox{.45\linewidth}{
\begin{center}
\includegraphics[scale=0.45]{highway_residual}
\caption{Highway residual adapter}
\label{fig:highwayres}
\end{center}
}
\hfill
\parbox{.45\linewidth}{
\begin{center}
\includegraphics[scale=0.45]{adapter}
\caption{Adapter}
\label{fig:adapter}
\end{center}}
\end{figure}

\section*{Future work}
\subsection*{Adaptive Domain Adaptation}
Generalization is a basic requirement of any learner, even when in-domain data is available. NMT system often loses performance dramatically in the previous training domain after being fine-tuned to the target domain. The phenomenon is well known under the name of catastrophic forgetting which was first reported in \cite{Michael89catastrophic}. This phenomenon appears mostly in neural network systems. The reason does not until now have a well-structured answer because of the low interpretability of this architecture. In the case of machine translation, we would like to improve the generalization of the expert
system by enhancing the stability of the intermediate representation (inside the network) of word whose meaning does not change across domain while maintaining a high degree of freedom for in-domain specific words. The practice of multi-task training might have
this effect because the lower stack of encoding modules shared across tasks is pushed to stay in a region thereby realizing a trade-off between in-domain optimal regions. Words frequently shared between domains, i.e popular in most of the domains will probably fall in
this ”average” region, therefore they will have low variance with respect to domain. We would like to reinforce this behavior of the multi-domain system

\subsection*{Multi-source adaptation}
The second problem of our interest is robustness with respect to out-of-domain data. Given several in-domain datasets in training, \cite{hoffman18algorithms}, and a test set from an unknown domain, we would like to build a good predictor from several out-of-domain expert predictors. In the machine-learning literature, the problem can be referred to as Multi-source Domain Adaptation. \cite{yishay09multiple}, \cite{mansour09domain} studied effective methods to combine learned predictors to predict test examples that come from an arbitrary unknown domain. They proposed a convex combination of learned in-domain predictors whose weights are computed by the in-domain probabilistic model. The authors also provided the upper bound of the risk of the new predictors on an arbitrary test domain. If the test distribution is not too far from the sets of a convex combination of k source distributions (in terms of Rényi divergence), then the risk of the proposed predictor on the test domain is close to the risk of the in-domain predictor on in-domain distribution. We would like to study these ideas and adapt them to our practical settings for robust multi-domain systems.

\subsection*{BPE-vocabulary's transfer effect}
One way to increase the sharing of knowledge between domains is to increase the shared vocabulary between them. By the fact that words share prefixes, suffixes between them, using smaller sequence units than words might help transfer translation information from more confident translation in high resource domain to less confident translation in low resource domains. Until now, there is not any investigation on the effect of BPE in transfer learning between domain. Therefore, we find that it is pertinent to evaluate the effect.

\section*{Other research}
\subsection*{Semantic similarity}
\cite{pham18fixing} proposed a method to predict the semantic similarity between 2 sentences from 2 different languages. The method is used in the context of filtering noisy parallel corpora such as Paracrawl. The method was implemented for corpus filtering shared task  \cite{koehn18findings}. \cite{pham18fixing} used a siamese network to encode a bilingual pair of sentences whose embedding is used to compute the semantic similarity by cosine similarity. The training of the network is done by optimizing the aggregation of soft word alignment scores between 2 sentences. To avoid being over-fitted by positive examples in which every sentence pair is correctly matched, we proposed to use the same amount of artificial negative examples in which words are randomly removed or added and changed in several positions.
\section*{Working plan}
\begin{itemize}
	\item Completing work on residual adapters and submission to WMT 2020 in the end of July 2020.
	\item Completing system submission to Machine Translation Robustness task of WMT 2020 in the end of July 2020.
	\item Completing work on contextualized Machine Translation with Systran's team in the end of July 2020.
	\item Completing work on Multi-source Adaptation between September - December 2020
	\item Completing work on BPE-vocabulary's transfer learning effect between January and March 2021.
	\item 
\end{itemize}
\bibliography{bibliography}
\bibliographystyle{apa-good}
\end{document}

### 2020-06-03 2020/phd_midterm_report/resume_phd.tex
phd midterm report

\documentclass[11pt, letterpaper]{article}
\usepackage[left=2cm,right=2cm,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{enumitem}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\begin{document}
\setlength{\belowdisplayskip}{3pt} \setlength{\belowdisplayshortskip}{3pt}
\setlength{\abovedisplayskip}{3pt} \setlength{\abovedisplayshortskip}{3pt}
\setlist{nosep}
\setlength{\parskip}{0.1cm}
\setlength{\parindent}{1em}
\section*{Minh Quang PHAM}
\begin{itemize}
	\item Formation: 
		\begin{itemize}
			\item Diplôme d'ingénieur polytechnicien, Ecole Polytechnique (2013-2016)
			\item Diplôme d'ingénieur, Telecom ParisTech (2016-2018).
			\item Master 2, Data Science, Paris-Saclay (2016-2017).
		\end{itemize}  
	\item Doctorat: Computer Science.
	\item Contract CIFRE  with Systran Group.
	\item Ecole doctorale: Sciences et Technologies de l'Information et de la Communication.
	\item Date of the first inscription: 1 November 2018.
	\item Date expected for the defense: 
	\item Supervisor: François YVON.
	\item Topic: Neural Machine Translation in Multi-Domain Context.
\end{itemize}
\section*{Publications}
\begin{itemize}
	\item Elise Michon, Minh Quang Pham, Josep M Crego, Jean Senellart. Neural Network Architectures for Arabic Dialect Identification. Proceedings of the Fifth Workshop on NLP for Similar Languages, Varieties and Dialects (VarDial 2018). \url{https://www.aclweb.org/anthology/W18-3914}
	\item Minh Quang Pham, Josep M Crego, Jean Senellart, François Yvon. Fixing Translation Divergences in Parallel Corpora for Neural Machine Translation. In Proceedings of the 2018 Conference on Empirical Methods in Natural Language Processing, Brussels, Belgium, October 31 - November 4, 2018 , (pp. 2967$-$2973). \url{https://www.aclweb.org/anthology/D18-1328}
	\item Minh Quang Pham, Josep M Crego, Jean Senellart. SYSTRAN Participation to the WMT2018 Shared Task on Parallel Corpus Filtering. Proceedings of the Third Conference on Machine Translation: Shared Task Papers (pp. 934$-$938). \url{https://www.aclweb.org/anthology/W18-6485}
	\item Minh Quang Pham, Josep M Crego, Jean Senellart, François Yvon. Generic and Specialized Word Embeddings for Multi-Domain Machine Translation. In Proceedings of the 16th International Workshop on Spoken Language Translation, IWSLT, (p. 9p). Hong-Kong, CN. \url{https://hal.archives-ouvertes.fr/hal-02343215/file/IWSLT2019_paper_10.pdf}
	\item Minh Quang Pham, Josep M Crego, Jean Senellart. SYSTRAN @ WAT2019: Russian-Japanese news commentary task. In Proceedings of the 6th Workshop on Asian Translation, (pp. 189–194). Hong Kong, China: Association for Computational Linguistics. \url{https://www.aclweb.org/anthology/D19-5225}
\end{itemize}
\subsection*{Submissions}
\begin{itemize}
	\item Minh Quang PHAM, Josep CREGO, François YVON. Revisiting Multi-Domain Machine Translation. Submitted to Transactions of the Association for Computational Linguistics.
\end{itemize}
\section*{Phd courses (38 hours)}
\subsection*{Completed courses}
\begin{itemize}
	\item Mouvement Brownien et calculs stochastique (30 hours) (M2 Mathématique Aléatoire, Paris Sud/ Paris Saclay)
	\item Séquenceur Slurm pour accéder à des clusters de calcul (LRI) (8 hours)
	\item "Formation traverse" is exempted.
\end{itemize}
\subsection*{On-going courses}
\begin{itemize}
	\item Ethics \& STICs (session 2 - 2020).
\end{itemize}
\end{document}

### 2020-06-03 2020/wmt_20/emnlp2020.tex
wmt20 submission

--- 2020-05-13 2020/coling2020/multidomain.tex
@@
-% File coling2020.tex
 %
-% Contact: feiliu@cs.ucf.edu & liang.huang.sh@gmail.com
-%% Based on the style files for COLING-2018, which were, in turn,
-%% Based on the style files for COLING-2016, which were, in turn,
-%% Based on the style files for COLING-2014, which were, in turn,
+% File emnlp2020.tex
+%
+%% Based on the style files for ACL 2020, which were
+%% Based on the style files for ACL 2018, NAACL 2018/19, which were
+%% Based on the style files for ACL-2015, with some improvements
+%%  taken from the NAACL-2016 style
 %% Based on the style files for ACL-2014, which were, in turn,
-%% Based on the style files for ACL-2013, which were, in turn,
-%% Based on the style files for ACL-2012, which were, in turn,
-%% based on the style files for ACL-2011, which were, in turn, 
-%% based on the style files for ACL-2010, which were, in turn, 
-%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
-%% based on the style files for EACL-2009 and IJCNLP-2008...
-
+%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
+%% EACL-2009, IJCNLP-2008...
 %% Based on the style files for EACL 2006 by 
@@
 \documentclass[11pt,a4paper]{article}
-% \usepackage{coling2020}
-\usepackage{tacl2018v2}
+\usepackage[hyperref]{emnlp2020}
 \usepackage{times}
-\usepackage{url}
 \usepackage{latexsym}
+\renewcommand{\UrlFont}{\ttfamily\small}
+
 \usepackage{graphicx}
@@
 \usepackage{amsmath,amsfonts,amssymb}
+
+% This is not strictly necessary, and may be commented out,
+% but it will improve the layout of the manuscript,
+% and will typically save some space.
+\usepackage{microtype}
+
+%\aclfinalcopy % Uncomment this line for the final submission
+%\def\aclpaperid{***} %  Enter the acl Paper ID here
+
 %\setlength\titlebox{5cm}
-%\colingfinalcopy % Uncomment this line for the final submission
-
 % You can expand the titlebox if you need extra space
@@
 % in the camera-ready version and ask you to change it back.
+
 \newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
@@
 \newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{red}{#1}}}
+
+\newcommand{\mpTodo}[1]{\Todo[MP:]{\textcolor{green}{#1}}}
+\newcommand{\mpDone}[1]{\done[MP]\Todo[MP:]{\textcolor{green}{#1}}}
 
@@
 
-\title{Revisiting Multi-Domain Machine Translation}
+\newcommand\BibTeX{B\textsc{ib}\TeX}
+
+\title{Less catastrophic forgetting domain adaptation}
 
@@
   Affiliation / Address line 3 \\
-  {\tt email@domain} \\\And
+  \texttt{email@domain} \\\And
   Second Author \\
@@
   Affiliation / Address line 3 \\
-  {\tt email@domain} \\}
+  \texttt{email@domain} \\}
 
@@
 \begin{abstract}
-  When building machine translation systems, it is often the case that one need to make the best out of the heterogeneous set of parallel data in training, and to robustly handle inputs from unexpected domains in testing.
-This multi-domain scenario has attracted a lot of recent work, that fall under the general umbrella of transfer learning.
-In this study, we revisit multi-domain machine translation, with the aim to formulate the motivations for developing such systems and the associated expectations with respect to performance.
-Our experiments with a large sample of multi-domain systems show that most of these expectations are hardly met and suggest that further work is needed to better analyze the current behaviour of multi-domain systems and to make them fully hold their promises.
-  \fyDone{Write abstract}
+\mpTodo{write abstract}
 \end{abstract}
+\section{Introduction } \label{sec:intro}
+\mpTodo{write introduction}
+\citet{Bapna19simple}
+\section{Word-Level Adaptive Domain Adaptation \label{sec:wada}}
+\mpTodo{Formalizing problem, network design, training algorithm}
 
-\section{Introduction} \label{sec:intro}
-
-%
-% The following footnote without marker is needed for the camera-ready
-% version of the paper.
-% Comment out the instructions (first text) and uncomment the 8 lines
-% under "final paper" for your variant of English.
-% 
-% \footnote{
-%     %
-%     % for review submission
-%     %
-%     \hspace{-0.65cm}  % space normally used by the marker
-%     Place licence statement here for the camera-ready version. See
-%     Section~\ref{licence} of the instructions for preparing a
-%     manuscript.
-%     %
-%     % % final paper: en-uk version 
-%     %
-%     % \hspace{-0.65cm}  % space normally used by the marker
-%     % This work is licensed under a Creative Commons 
-%     % Attribution 4.0 International Licence.
-%     % Licence details:
-%     % \url{http://creativecommons.org/licenses/by/4.0/}.
-%     % 
-%     % % final paper: en-us version 
-%     %
-%     % \hspace{-0.65cm}  % space normally used by the marker
-%     % This work is licensed under a Creative Commons 
-%     % Attribution 4.0 International License.
-%     % License details:
-%     % \url{http://creativecommons.org/licenses/by/4.0/}.
-% }
-
-Data-based Machine Translation (MT), whether statistical or neural, rests on well understood machine learning principles.
-Given a training sample of matched source-target sentence pairs $(\src,\trg)$\jcDone{shoulnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}_s$, a model parameterized by $\theta$ (here, a translation function $h_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(h_\theta(\src), \trg)$.
-This approach ensures that the translation loss remains low when translating more sentences drawn from the same distribution.
-
-Owing to the great variability of language data, this ideal situation is rarely met in practice, warranting the study of an alternative scenario, where the test distribution $\mathcal{D}_t$ differs from $\mathcal{D}_s$.
-In this setting, \emph{domain adaptation} (DA) methods are in order.
-DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory,Quinonero08dataset,Pan10asurvey})\fyDone{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}).
-Various techniques thus exist to handle both the situations where a (small) training sample drawn from $\mathcal{D}_t$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
-% In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
-
-A seemingly related problem is \emph{multi-domain} (MD) machine translation \cite{Sajjad17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system is trained and tested with data from multiple domains.
-MD machine translation (MDMT) corresponds to a very common situation, where all available data, no matter its origin, is used to train a robust system that performs well for any kind of new input.
-% As pointed out both by \newcite{Dredze08online} and \cite{Finkel09hierarchical}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}.
-If the intuitions behind MDMT are quite simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should MDMT perform well when the test data is distributed like the training data, when it is equally distributed across domains or when the test distribution is unknown?
-Should MDMT also be robust to new domains?
-How should it handle domain labeling errors?
-% , which makes their evaluation fragile
-
-A related question concerns the relationship between supervised domain adaptation and multi-domain translation.
-The latter task seems more challenging as adapting a system to a known target domain $A$ looks easier than training a MDMT system, which should do also well on domains other than $A$.
-Are there situations where MD systems can surpass DA, as is sometimes expected?
-
-In this paper, we formulate in a more precise fashion the requirements that an effective MDMT system should meet (section~\ref{sec:requirements}).
-Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT systems and associated measurements to evaluate them (section~\ref{sec:challenging}).
-In doing so, we also shed lights on new problems that arise in this context, regarding for instance the accomodation of new domains in the course of training, or the computation of automating domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
-We show in Section~\ref{sec:experiments} that existing approaches still fall short to match many of these requirements, notably with respect to handling a large amount of heterogeneous domains and to dynamically integrating new domains in training.\fyDone{Spell out conclusions}
-
-\section{Requirements of multi-domain MT \label{sec:requirements}}
-In this section, we recap the main reasons for considering a multi-domain scenario and discuss their implications in terms of performance evaluation.
-\fyDone{Relation to theory, discussion of worst case scenario}
-\fyDone{Insist on the work of Dredze}
-
-\subsection{Formalizing multi-domain translation \label{ssec:formalization}}
-
-We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for source sentences; each domain corresponds to a specific source of data, and differs from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
-Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
-$h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$), but can differ from one domain to the other.
-
-A typical learning scenario in MT is to have access to samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$.
-Multi-domain learning, as defined in \cite{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ of several domains, making the problem distinct from mere domain adaption.
-A multi-domain learner is then expected to use these tags effectively \cite{Joshi12multidomain} when computing the combined translation function $h(x,d)$, and to perform well in all domains \cite{Finkel09hierarchical}.
-Note that this setting is closely related to the multi-source adaptation problem formalized in \cite{Mansour09domainadaptation,Mansour09multiple,Hoffman18algorithms}.
-
-This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is \cite{Farajian17multidomain}, where test translations rely on similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here.
-Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function.
-We now discuss the arguments that are put forward to proceed differently.
-% --
-
-% This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyTodo{explain zero shot}.
-
-\subsection{Reasons for building MDMT systems \label{ssec:whymdmt}}
-
-A first motivation for moving away from the one-domain / one-system solution are practical  \cite{Sennrich13multidomain,Farajian17neural}: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop one single system instead of having to optimize and maintain multiple engines.
-The underlying assumption here is that the number of domains of interests can be large, a limiting scenario being fully personalized machine translation \cite{Michel2018extreme}.
-% which is not always the case: for instance \cite{Britz17mixing,Zeng18multidomain,Jiang19multidomain} only consider 2-4 domains in their experiments.\fyTodo{Keep this last part ?}
-
-A second line of reasoning rests on linguistic properties of the translation function and contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and tend to remain semantically stable accross domains, motivating some cross-domain parameter sharing.
-A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the translation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
-It is here expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related and can share more information.
-
-A third series of motivations are of statistical nature.
-The training data available for each domain is usually unevenly distributed, and domain-specific systems trained or adapted on small datasets are likely to have a high variance and generalize poorly.
-For some test domains, there may even be no data at all \cite{Farajian17neural}.
-Training mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}.
-Under this view, MDMT would be especially beneficial for domains with little training data.
-This is observed for multilingual MT from English: an increase for under-resource languages due to positive transfer, at the cost of a decrease in performance for well-resourced languages \cite{Arivazhagan19massively}.
-
-Combining multiple domain-specific MTs can also be justified in the sake of distributional robustness \cite{Mansour09domainadaptation,Mansour09multiple}, for instance when the test mixture differs from the train mixture, or when it includes new domains unseen in training.
-Such scenario is already well documented for zero-shot multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}, where mixing languages has more than demonstrated its usefulness.
-An even more challenging case is when the MT would need to perform well for any test distribution (as studied for statistical MT in \cite{Huck15mixeddomain}).
-In all these cases, mixing domains in training and/or testing is likely to improve robustness against unexpected or adversarial inputs \cite{Oren19distributionally}.
-
-A distinct line of reasoning is that mixing domains can have a positive regularization effect for all domains.
-By introducing variability in training, it prevents DA from overfitting the available adaptation data and could help improve generalization even for well-resourced domains.
-A related case is made in \cite{Joshi12multidomain}, which shows that part of the benefits of MD training is due to an ensembling effect, where systems from multiple domains are simultaneously used in the prediction phase; this effect may subsist even in the absence of clear domain separations.
-
-\fyDone{recap}
-To recap, there are multiple arguments for adopting MDMT, some already used in DA settings, and some original.
-These arguments are not mutually exclusive; however each yields specific expectations with respect to the performance of this approach, and should also yield appropriate evaluation procedure.
-If the motivation is primarily computational, then a drop in MT quality with respect to multiple individual domain might be acceptable if compensated by the computational savings.
-If it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some under-resourced domains, over individually trained systems.
-If finally, it is to make the system more robust to unexpected or adversarial test distributions, then this is the setting that should be used to evaluate MDMT.
-The next section discusses ways in which these requirements of MDMT systems could ben challenged.
-
-% \cite[eq. (3)]{Wang20general} defines the loss of MDMT as the (macro) average of domain loss - all domains are equally important.
-\section{Challenging multi-domain systems \label{sec:challenging}}
-In this section, we propose seven\fyDone{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.
-All these evaluations will rest on comparison of translation performance, and do not depend on the choice of a particular metric.
-To make our results comparable with the literature, we will only use the BLEU score \cite{Papineni02bleu} in Section~\ref{sec:experiments}, noting it may not be the best yardstick to assess subtle improvements of lexical choices that are often associated with domain adapted systems \cite{Irvine13measuring}.
-Other important figure of merits for MDMT systems are the computational training cost and the total number of parameters.\fyDone{Insert discussion about scores}
-
-\subsection{Multi-domain systems should be effective \label{ssec:effective}}
-% \subsubsection{Performance}
-A first expectation is that MDMT systems should perform well in the face of mixed-domain test data. This yields the following three requirements.
-
-\paragraph{[P1-LAB]}\fyDone{Decide naming scheme}A MDMT should perform better than the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}.
-Evaluating this requirement is a matter of a mere comparison, assuming the test distribution of domains is known: if all domains are equally important, performance averages can be reported; if they are not, weighted averages should be used instead.\fyFuture{Try random label assignments ?} % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
-
-\paragraph{[P2-TUN]} Additionally, one can expect that MDMT will improve over fine-tuning \cite{Luong15stanford,Freitag16fast}, at least in domains where data is scarce, or in situations where several domains are close.
-To evaluate this, we perform two measurements, using a real as well as an artificial scenario.
-In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains.
-In the artificial scenario, we split a single domain in two parts which are considered as distinct in training.
-The expectation here is that a MDMT should yield a clear gain for both pseudo sub-domains, which should benefit from the supplementary amount of relevant training.
-In this situation, MDMT should even outperform fine-tuning on either of the pseudo sub-domain.
-\jcDone{I dont understand the prediction... why running MDMT on 2 artificial subdomains should show  gains over fine-tunning?}\fyDone{small loss with respect to non split ?}
-
-\subsection{Robustness to fuzzy domain separation \label{ssec:robusness}}
-A second set of requirements is related to the definition of a domain.
-As repeatedly pointed out in the literature, parallel corpora used MT are often collected opportunistically and the view that each corpus constitutes a single domain is often a gross approximation.\footnote{Two of our own ``domains'' actually comprise several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments.
-To challenge these requirements we propose the evaluate the following requirements.
-
-\paragraph{[P3-HET]}
-The notion of a domain being a fragile one, an effective MDMT system should be able to discover not only when cross-domain sharing is useful (cf.\ requirement [P2-TUN])\jcDone{what is P2.2?}, but also when intra-domain heterogeneity is hurting.
-This requirement is tested by artificially conjoining separate domains into one during training, hoping that the loss in performance with respect to the baseline (using correct domain tags) will remain small.
-
-\paragraph{[P4-ERR]}
-A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, catastrophic drops in performance can be observerd\fyDone{cite \cite{McCloskey89catastrophic}}.
-This requirement can be measured by translating test texts for domain $A$ with erroneous or random domain tags and measure the correlated loss.
-
-\paragraph{[P5-UNK]}
-A related situation occurs when the domain of a test document is unknown.
-Several situations need be considered: for domains seen in training, then using automatically predicted domain labels should not be much worse than using the correct one.
-For test documents from unknown domains (zero-shot transfer), a good MD system should ideally outperform the default baseline that merges all available data.\fyDone{Systems react to unknown domains}
-% These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.
-
-\paragraph{[P6-DYN]}
-Another requirement, more of an operational nature, is that a MDMT system should smoothly evolve to handle a growing number of domains, without having to retrain the full system each time new data is available.
-This is a requirement [P6-DYN] that we challenge by dynamically changing the number of training and test domains.
-
-\subsection{Scaling to a large number of domains \label{ssec:scaling}}
-
-\paragraph{[P7-NUM]} As mentioned above, MDMT systems have often been motivated by computational arguments.
-This argument is all the more sensible as the number of domains increases, making the optimization of many individual systems both ineffective and undesirable.
-For lack of having access to corpora containing very large sets (eg.\ in the order of 100-1000) domains, we experiment with automatically learned domains.\fyFuture{considering a varying number of clusters.}
-
-\section{Experimental settings \label{sec:experiments}}
+\section{Experimental settings \label{sec:exp}}
 
 \subsection{Data and metrics \label{ssec:corpora}}
+\mpTodo{English - French, English - German}
+\subsection{Baselines \label{ssec:baseline}}
+\mpTodo{\system{FT-block}; \system{FT-full}, \system{FT-full Brian Thompson}, \system{DC}, \system{Domain Mixing}}
 
-We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
-Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
-Corpus statistics in Table~\ref{tab:Corpora}.
-Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools.
-To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.%
+\subsection{Proposed systems \label{ssec:wada}}
 
-We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
-We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
+\section{Related Work \label{sec:related}}
 
-%for contrast experiments, we also use supplementary test sets from three other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
-This enables us to evaluate the loss in performance when the test set is from a domain not seen in training.
-% The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
+\section{Discussion \label{sec:discussion}}
 
-\begin{table*}[htbp]
-  \centering
-  \begin{tabular}{ |lllllll|} %*{4}{|r|}}
-    \hline
-    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
-    \domain{med} & \domain{law} & \domain{bank} & \domain{it} & \domain{talk} & \domain{rel} & \domain{news} \\
-    \hline
-    2609 (0.68) & 190 (0.05)  & 501 (0.13) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
-    \hline
-  \end{tabular}
-\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain).
-\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.}
-\label{tab:Corpora}
-\end{table*}
+\section*{Acknowledgments}
 
-We measure the distance between domains using the $\mathcal{H}$-Divergence \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains.
-%Formally, given test sets of size $m$ for domains $A$ and $B$, and $h(x)$ a trained domain predictor, $\mathcal{H}(A,B)$ is computed %as:
-%$$
-%\mathcal{H}(A,B) = 2(1 - [\frac{1}{m} \sum_{x:h(x) = B} \mathbb{I}( x \in A) + \frac{1}{m} \sum_{x: h(x) = A} \mathbb{I}(x \in B)]),
-% $$
-% where $\mathbb{I}$ is the indicator function.
-Our discriminator is a SVM independently trained for each pair of domains, with sentence representations derived via mean pooling from the source side representation of the generic Transformer model.
-We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well separated from the all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
-
-\begin{table}\centering
-  \begin{tabular}{|l*{5}{|r}|} \hline
-    & \domain{law} & \domain{bank} & \domain{talk} & \domain{IT} & \domain{rel} \\ \hline
-    \domain{med} &1.93 &1.97 &1.9 &1.93 &1.97 \\
-    \domain{law}   && 1.94 & 1.97 &1.93 & 1.99 \\
-    \domain{bank} &&&1.98 &1.94 &1.99 \\
-    \domain{talk}   &&&&1.92 &1.93 \\
-     \domain{IT}     &&&&& 1.99 \\ \hline
-  \end{tabular}
-  \caption{The $\mathcal{H}$-divergence between domains}
-  \label{tab:domaindist}
-\end{table}
-
-\subsection{Baselines \label{ssec:baselines}}
-
-Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
-
-\begin{itemize}
-\item a generic model trained on a concatenation of all corpora (\texttt{Mixed}).
-We develop two versions\footnote{In fact three: to enable a fair comparison with WDCMT, a RNN-based variant is also trained and evaluated.} of this system, one where domain unbalance reflects the distribution of our parallel data (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
-The former is the best option when the train mixture $\mathcal{D}^s$ is also observed in testing; the latter should be used when the test distribution is uniform across domains.
-Accordingly, we report two aggregate scores: a weighted average reflecting the default distribution, and an unweighted average, meaning that test domains are equally important.
-%   As explained above, depending whether MDMT are expected to perform specially well on large domains, or to treat all domains equally, one or the other training distribution should be prefered.\jcDone{All systems cand be learned following both Bal and Nat.
-does Nat corresponds to wAVG in test?
-is wAVG defined somewhere?}
-\item fine-tuned models \cite{Luong15stanford,Freitag16fast}\jcDone{i would use Luong and Manning 2015}, based on the \system{Mixed-Nat} system, further trained on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
-We again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a simple adaptation module that is added (with residual connexions) on top of every Transformer layer (\system{FT-Block}).
-\end{itemize}
-
-For all models, we set the embeddings size and the hidden layers size to~512.
-Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
-The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
-% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
-% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training use a batch size of~128; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
- % All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
-\subsection{Multi-domain systems \label{ssec:systems}}
-\fyDone{Remove Gated residual}
-Our comparison of multi-domain systems includes the following reimplementations of recent proposals from the literature:
-\begin{itemize}
-\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});\fyDone{why not LDR as in Minh's paper?}
-\item the three proposals of \newcite{Britz17mixing}.
-\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
-Training uses reference tags and inference is performed with predicted tags, just like for regular target words.
-\system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
-These methods thus only use domain tags in training.
-\item the multi-domain model of \newcite{Zeng18multidomain} (\system{WDCMT}), where a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly, improved results.}\fyDone{Check this}
-\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connexions make it possible to by-pass this module.
-  % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
-Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
-\end{itemize}
-
-Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges from our notion of MDMT, which should be able to handle available domain labels in training and testing.
-% This system has shown to meet several of requirements and would also have no problem scaling to a large number of domains.\fyDone{TBC}
-
-\section{Results and discussion \label{sec:results}}
-
-\subsection{Performance of MDMT systems \label{ssec:rawperformance}}
-In this section, we discuss the basic performance of MDMT systems trained and tested on $6$~domains.
-Results are in Table~\ref{tab:performance}.
-As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points\fyDone{BP?} better, notably owing to the much better results for \domain{rel}).
-As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains.
-As all other systems are trained with an unbalanced data distribution, we use the weighted average to perform global comparisons.
-
-Fine-tuning each domain separately provides us with better baselines, outperforming \system{Mixed-Nat} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
-A complete fine-tuning, updating all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to keeping fixed the capacity of the adapter module.\fyDone{Add 2 averages ?}\fyDone{Significance testing wrt Mix Generic, Full Fine-tuned, for each domain}
-
-A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on the source is better than on the target, and very comparable to repeating the tag for every source word.
-If all MDMTs (except \system{DM} and \system{ADM}) slightly improve over the generic baseline (for most domains), these improvements are rarely significant.
-The best contender overall is \system{MDL Res}, which significantly improves over \system{Mixed-Nat} for three domains, and is the only one to clearly fulfill [P1-LAB]; \system{WDCMT} also improves on three domains, but regresses on one.
-The use of a dedicated adaptation component seems here to prove better than feature-based strategies.
-The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}), for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.\fyDone{why LDC does not outperform DC as in Minh's paper (for en2fr MED) ?}
-
-All systems fail to outperform fine-tuning, sometimes by a wide margin, especially for an ``isolated'' domain like \domain{rel}.
-This might be due to the fact that domains are well separated (cf.\ Section~\ref{ssec:corpora}) and are hardly helping each other.
-In this situation, MDMT systems should instead make sure that each domain is allocated a sufficient number of parameters, so as to close the gap with fine-tuning.
-\fyDone{More comments when we have all the results}\fyDone{Importance of sharing and unsharing}
-
-\begin{table*}
-  \centering
-  \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
-%     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
-    \system{Mixed-Nat-RNN}  & 36.8 & 53.8 & 47.2 & 30.0 & 35.7 & 60.2  & 39.2  & 44.0 \\
-    \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 \\ %& \\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & \SB{42.7} & \SB{53.8} \\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & \SB{42.3} & \SB{53.3} \\ \hline % & 20.5\\ \hline
- %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.6 & 50.1    \\%    & 21.8 \\
-    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   \\% & \SW{21.7} \\
-    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} &40.9 & 49.8          \\% & 22.1 \\ 
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.0 & 49.7     \\% &  23.4 \\
-    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 38.1 & 43.4 \\ % & 22.6\\
-    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 38.9 & 47.5 \\% & 23.3 \\
-    \system{WDCMT}       & 36.0 & 53.3 & \SB{48.8} & 31.1 & \SB{38.8} & \SW{58.5} & 39.0 & 44.4 \\ % & 20.4 \\
-    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & \SB{51.9} \\%  & \SW{21.2} \\
-    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
-%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3&\\
-     \hline
-  \end{tabular}
-  \caption{Translation performance of various MDMT systems.
-We report BLEU scores for each domain, as well as domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages.
-Boldface denotes significant gains with respect to \system{Mix-Nat} (or \system{Mix-Nat-RNN}, for WDCMT), underline denotes significant losses.}
-  \label{tab:performance}
-  % \fyDone{Fill the table with all results, including a comparison with *RNNs* for wcmd}
-  \fyDone{Do we have significancy tests for averages?} %\fyTodo{Add tests wrt to RNN}
-\end{table*}
-
-\subsection{Redefining domains \label{ssec:redomains}}
-
-Table~\ref{tab:redomains} summarizes the results of four experiments where we artificially redefine the boundaries of domains, with the aim to challenge requirements [P2-TUN], [P3-HET], and [P4-ERR].
-In first three, we randomly \emph{split} one corpus in two parts and proceed as if this corresponded to two actual domains.
-A MD system should detect that these two pseudo-domains are mutually beneficial, and should be hardly affected by this change with respect to the baseline scenario (no split).
-In this situation, we expect MDMT to even surpass fine-tuning separately on each of these dummy domains, as MDMT exploits all data, while fine-tuning focuses only on a subpart.
-In testing, we decode the test set twice, once with each pseudo-domain tag.
-This makes no difference for \system{TTM}, \system{DM}, \system{ADM} and \system{WDCMT}, which do not use domain tags in testing.
-In the \textsl{merge} experiment, we merge two corpora in training, in order to assess the robustness with respect to heterogenous domains [P3-HET].
-We then translate the two corresponding tests with the same system.
-
-\begin{table*}
-  \centering% \small
-  \begin{tabular}{|p{1.8cm}|*{10}{r|}} \hline
-    % &&&&&& \\
-    \hfill Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & \multicolumn{2}{c|}{Wrong} \\ % \hline
-     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{law} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} &  \multicolumn{1}{c|}{\domain{rnd}} &  \multicolumn{1}{c|}{\domain{new}}\\ \hline
-    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
-    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5} & -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& \SW{-19.6} & \SW{-3.3}\\% [32.5]
-    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -0.4 & -0.4 & -0.5 & -0.4 & \SW{-13.4} & \SW{-1.7}\\% [35.9]
-    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & \SW{-14.2} &\SW{-1.8}\\ % [34.9] 
-    \system{LDR}           & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-12} & \SW{-1.4}\\ % [37.0]
-    \system{TTM} (*)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & 0.0 & -0.1\\
-%    \system{MDL Gated} & & & & & & & & &  \\
-    \system{DM} (*)           & -0.3   & -0.3  & +0.4 & +0.4 & +0.3 & +0.3 & +0.9 & +0.1 & 0.0 &-0.9\\
-    \system{ADM} (*)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & 0.0&-0.2\\
-    \system{WDCMT} (*)     & -0.0    & -0.0  & +0.2 & +0.2  & +0.8 & +0.8  & -0.4 & -0.8 & 0.0 & +0.2 \\
-    \system{MDL Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +\SB{0.7} & -0.3 & \SW{-18.6} & \SW{-1.3}\\ % [31.8]
-    \hline
-  \end{tabular}
-  \caption{Translation performance with variable domain definitions.
-In the Split/Merge experiments, we report BLEU differences for the concerned test set(s).
-Underline denotes significant loss when domains are changed wrt.\ the baseline situation; bold for a significant improvement over \system{FT-Full}; (*) flags systems ignoring test domains.
-    \fyDone{Check wrong and significance of diff for ``old''}\fyDone{How about wrong ?}
-  }
-  \label{tab:redomains}
-\end{table*}
-
-Our findings can be summarized as follows.
-For the split experiments, we see small variations that can be positive or negative compared to the baseline situation, but these are hardly significative.
-All systems show some robustness with respect to fuzzy domain boundaries; this is mostly notable for \system{ADM}, suggesting that when domain are close, ignoring domain differences is effective.
-In contrary, \system{FT-Full} incurs large losses across the board, especially for the small data condition.
-Even in this very favourable case, very few MDMT systems are able to significantly outperform \system{FT-Full} and this is only observed for the smaller part of the \domain{med} domain.
-The merge condition is hardly different, with large losses for \system{FT-full}, and small variations for all systems.
-We even observe some rare improvements with respect to the situation where we use actual domains.\fyDone{More comments ?}
-
-\subsubsection{Handling wrong or unknown domains \label{sssec:unknowns}}
-
-\fyTodo{Numbers for the rand domain analysis}In the last two columns of Table~\ref{tab:redomains}, we report the drop in performance when the domain information is not correct.
-In the first (\domain{rnd}), we use test data from the domains seen in training, presented with a random domain tag\fyDone{all configs use the same tag or is randomization is run for each config}.
-In this situation, the loss with respect to using the correct domain is generally large (more than 10 BLEU points), showing an overall failure to meet requirement [P4-ERR], except for systems which disregard domain tags in testing.
-\fyDone{Comments?}
-
-In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the most likely domain tag and use it for decoding.
-% Results are in the rightmost column of Table~\ref{tab:redomains}.
-In this configuration, again, systems using domain tags in testing perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\fyDone{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}
-
-\subsubsection{Handling growing numbers of domains}
-Another series of experiments evaluate the ability to dynamically handle suplementary domains (requirement [P6-DYN]) as follows.
-Starting with the existing MD systems of Section~\ref{ssec:rawperformance}, we introduce an extra domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is critical for achieving optimal performance: as our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy based on the new training distribution.
-This is detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with training all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.\footnote{WDCMT results are excluded from this Table, as resuming training proved difficult to implement.}\fyDone{Ecrire un commentaire}
-We expect that MDMT systems should not to be significantly impacted by the addition of a new domain and reach about the same performance as when training with this domain from scratch.
-From a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{DC-Feat} or \system{TTM}, for which new domains merely add new labels.
-It is less easy for \system{DM}, \system{ADM} and \system{WDCMT}, which include a built-in domain classifier whose outputs have to be given, or for \system{LDR} and \system{MDL~Res} for which the number of possible domains is built in the architecture and has to be anticipated from the start.
-This makes a difference between domain-bounded systems, for which the number of domains is limited and truly open-domain systems.% , which seamlessly integrate additional domains.
-
-\begin{table*}
-  \begin{tabular}{|p{2.5cm}|*{9}{r|}} \hline \footnotesize
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-     \system{Mixed-Nat}  & 37.1  & 54.1  & 49.6	& 34.1  & 42.1	& 77.0 & 28.9 & 40.8	& 49.0 \\ 
-%    \system{Mixed-Nat}  & 37.3  & 54.6   & 50.1   & 33.5  & 43.2  & 77.5  & 23.5 & 41.1 & 49.4 \\
-%    \system{FT-Full}       & 37.7  & 59.2   & 54.5   & 34.0  & 46.8  & 90.8  & **** & 42.8 & 53.8 \\
-    \hline%\hline
-     \system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9  & 41.2 & 50.1\\[-2pt]
-%                   & 37.2 & 54.7   & 49.9    & 34.2  &  43.0  & 79.9 & \SW{28.1}  & &\\
-                   & \scriptsize -0.2 & \scriptsize -0.1  & \scriptsize -0.1 & \scriptsize -0.6 & \scriptsize -0.9 & \scriptsize +0.3 & \scriptsize \SW{-0.8} & \scriptsize -0.2 & \scriptsize -0.3\\
-
-    \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5  & 41.4 & 49.9\\[-2pt]
-%                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
-                   & \scriptsize +0.3  & \scriptsize -0.1   & \scriptsize -0.6    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.3}  & \scriptsize \SW{-3.5}  & \scriptsize \SW{-1.4}  & \scriptsize -0.1 & \scriptsize \SW{-1.2}\\
-
-    \system{LDR}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
-                    & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize -0.4 & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize +\SB{3.8} & \scriptsize \SW{-0.9} & \scriptsize +0.5 & \scriptsize -0.4 \\
-
-    \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.2  & 29.1 & 41 & 49.4\\[-2pt]
-%                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
-                    & \scriptsize -0.3  & \scriptsize -0.3  & \scriptsize -0.5    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.0}  & \scriptsize \SW{-4.0}  & \scriptsize \SW{-1.4}& \scriptsize -0.5 & \scriptsize \SW{-1.2}\\
-
-    \system{DM}            &36.0 &51.3&46.8&31.8&39.8&65.7&27.0 & 38.9 & 45.2\\[-2pt]
-%                   &36.5&51.7&47.4&31.7&40.3&65.7&\SW{25.8}& & \\
-                   & \scriptsize +0.6& \scriptsize +0.4& \scriptsize +0.6& \scriptsize -0.1& \scriptsize +0.5 & \scriptsize 0.0& \scriptsize \SW{-1.2}& \scriptsize +0.5 & \scriptsize +0.3\\
-
-    \system{ADM}          &36.6&54.2&49.1&32.9&42.1&75.7&28.7 & 40.2 & 48.4 \\[-2pt]
-%                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
-                   & \scriptsize +0.3 & \scriptsize -0.8 & \scriptsize -0.8 & \scriptsize -0.2 & \scriptsize -0.4 & \scriptsize \SW{-5.0} & \scriptsize \SW{-1.9} & \scriptsize -0.2 & \scriptsize \SW{-1.1}\\
-
-%    \system{WDCMT}     &&&&&&& & &\\[-2pt]
-%                    &&&&&&&&& \\
-
-    \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & 41.9 & 51.8 \\[-2pt]
-%                      & 37.5 & 56.1 & 51.1 & 34.0  &  44.3 & 87.3 & \SW{28.4} & & 51.7 & \\
-                        & \scriptsize -0.2 & \scriptsize +0.5 & \scriptsize 0.0 & \scriptsize -0.4  &  \scriptsize -0.2 & \scriptsize -0.2 & \scriptsize \SW{-0.8} & \scriptsize -0.2 & \scriptsize -0.1 \\
-     \hline
-  \end{tabular}
-  \caption{Ability to handle a new domain: BLEU scores for the ``resume training'' mode compared to a complete training session.
-Averages only take into account the six initial domains (\domain{News} excluded).
-Underline denotes a significant loss between warmstart and coldstart, bold for a significant gain.}
-  \label{tab:warmrestart}
-  \fyDone{Fill the table, significancy testing against initial learning condition, for each domain, and averages ? }
-  \fyDone{Need correct number / differences for Mixed-Nat}
-  \fyDone{averages include one more domain ? No for comparison}
-  \fyDone{i used smaller number for differences (easier to read...)}
-\end{table*}
-
-We can first compare the results of coldstart training with six (Table~\ref{tab:performance}) or seven domains (Table~\ref{tab:warmrestart}): the most striking observation is that the extra training data is detrimental for almost all systems and domains\fyDone{is this significant ?}.
-Increasing the heterogeneity of the mix with supplementary out-of-domain data hurts performance across the board.
-% Adding is not helping here, especially when domains are well separated.
-Comparing now the coldstart with the warmstart scenarios, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training can negatively impact the performance for other domains.
-This happens notably for \system{DC-Tag}, \system{TTM} and \system{ADM}.
-In this setting \system{MDL Res} and \system{DM} show the smaller average loss, with the former achieving the best balance of training cost and average translation performance.% of all systems.
-
-% \begin{itemize}
-% \item even when trained from entirely scratch, very few systems are able to take advantage of the additional data for the 6 initial domains - this is observed on the averaged scores in Table~\ref{tab:warmrestart} with hardly surpass those in Table~\ref{tab:performance}
-% \item restart is computationally cheaper (\fyTodo{how much cheaper here ?}) but underperforms cold start for all systems, sometimes by a wide margin.
-% \item the loss is largest for the small domains - how about distance to \domain{news} ?
-% \end{itemize}
-% \fyDone{Very close to fine tuning towards News, no ?}
-
-\subsection{Automatic domains \label{ssec:autodomains}}
-In this section, we experiment with automatic domains, obtained by clustering sentences into $k=30$ classes using the k-means algorithm based on generic sentence representations obtained via mean pooling as in Section~\ref{ssec:corpora}.
-This allows us to evaluate requirement [P7-scale], training and testing our systems as if these domains were genuine.
-We also note that many of these classes are mere split of the large \domain{med}, while a fewer number of classes are mix of two existing domains.
-We are thus in a position to reiterate at a larger scale the measurements of Section~\ref{ssec:redomains} and test whether multi-domain systems can effectively take advantage from the cross-domain similarities, and to eventually perform better that fine-tuning.
-
-Table~\ref{tab:subdomains} reports the effect of using automatic domain tags for the 6 domain-specific test sets: each sentence was first assigned to an automatic class, translated with the corresponding multi-domain system with 30 classes; aggregate numbers were then computed, and contrasted with the 6-domain scenario.
-Results are clear and confirm previous observations: even though some clusters are very close, the net effect is a loss in performance for almost all systems and conditions.
-In this setting, the best MDMT in our pool (\system{MDL-Res}) is no longer able to surpass the \system{Mix-Nat} baseline.%\fyTodo{Note about scaling?}\fyTodo{More comparison?}
-
-\begin{table*}[t]
-  \centering
-  \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
-%     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-  %    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
-  %     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8\\
-  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag}       & 38.5 & \SW{54.0} & 49.0   & 33.6 & \SW{42.2} & \SW{76.7} & 41.6 & 49.0 \\%    & 21.8 \\
-    \system{DC-Feat}      & 37.3  & 54.2 & 49.3   & 33.6 & \SW{41.9} & \SW{75.8} & 40.8 & \SW{48.7}  \\% & \SW{21.7} \\
-    \system{LDR}     & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 39.1 & \SW{48.3}         \\% & 22.1 \\ 
-    \system{TTM}            & 37.4 & \SW{53.7} & 48.9 & 32.8 & 41.3 & \SW{75.8} & 40.7 & \SW{48.3}   \\% &  23.4 \\
-    \system{DM}             & 35.4 & 49.3  & 45.2 & 29.7 & 37.1 & \SW{60.0} & 37.8 & 42.8 \\ % & 22.6\\
-    \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 39.5 & 47.1\\% & 23.3 \\
-    \system{WDCMT}       & 35.57 & 53.08 & 48.43 & 30.55 & \SW{37.73} & \SW{56.02} & 38.5 & 43.6 \\ % & 20.4 \\ 
-    \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SB{50.7}\\%  & \SW{21.2} \\
-     \hline
-  \end{tabular}
-  \caption{Translation performance with automatic domains, computed with the original test sets.}
-  \label{tab:subdomains}
-  \fyDone{Fill the table with correct results,}\fyDone{Change wavg and avg}
-\end{table*}
-
-\section{Related work \label{sec:related}}
-
-The multi-domain training regime is more the norm than the exception for natural language processing \cite{Dredze08online,Finkel09hierarchical}, and the design of multi-domain systems has been proposed for many language processing tasks.
-We focus here exclusively on MD machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection / weighting, adversarial training, etc) have been studied in other contexts.
-
-%; scores are reported separately per domain, suggesting that all domains are equally important at test time.
-% Adversarial adpatation is used in many domains
-% for QA @misc{lee2019domainagnostic,
-%    title={Domain-agnostic Question-Answering with Adversarial Training},
-%    author={Seanie Lee and Donggyu Kim and Jangwon Park},
-%    year={2019},
-%    eprint={1910.09342},
-%    archivePrefix={arXiv},
-%    primaryClass={cs.CL}
-%}, NLU and many others
-
-Multi-domain translation was already proposed for statistical MT, either considering as we do multiple sources of training data (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}), or domains made of multiple topics\fyDone{you mean multiple topics?} \cite{Eidelman12topic,Hasler14dynamic-topic}.\fyDone{Add also topic models refs} Two main strategies were considered: instance-based, involving a measure of similarities between train and test domains; feature-based, where domain/topic labels give rise to additional features.
-
-The latter strategy has been widely used in NMT: \newcite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token or in the form of additional domain-feature associated to each word.
-These results are reproduced by \newcite{Tars18multidomain}, who also consider automatically induced domain tags.
-This technique also helps control the style of MT outputs in \cite{Sennrich16politeness,Niu18multitask}, and to encode the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}.
-Domain control can also be performed on the target side, as in \cite{Chen16guided}, where a topic vector describing the whole document serves as an extra context in the softmax layer of the decoder.
-Such ideas are further developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and commonalties are encoded in the network architecture: some parameters are shared across domains, while others are domain-specific.
-% (part of the embeddings in \cite{Pham19generic}).
-
-Techniques proposed by \cite{Britz17mixing} aim to ensure that domain information is actually used in a mix-domain system.
-Three methods are considered, using either domain classification (or domain normalization, via adversarial training) on the source or target side.
-There is no clear winner in either of the three language pairs considered.
-One contribution of this work is the idea of normalizing representations through adversarial training, so as to make the mixture of heterogeneous data more effective; representation normalization has since proven a key ingredient in multilingual transfer learning.
-The same basic techniques (parameter sharing, automatic domain identification / normalization) are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific representations on the one hand from domain-agnostic representation on the other hand.
-These representations are then processed as two separate inputs, then recombined to compute the translation.
-
-Another parameter-sharing scheme is \cite{Jiang19multidomain}, which augments a Transformer model with domain-specific heads, whose contributions are regulated at the word/position level: some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling that the core of \cite{Huck15mixeddomain,Saunders19ucam}.
-The results for three language pairs outperform several standard baselines for a two-domain systems (in fr:en and ge:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
-
-Finally, \newcite{Farajian17multidomain}, \newcite{li-etal-2018-one} and \newcite{xu2019} follow a different strategy.
-The processing of each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some iterations, before delivering its output.
-This approach entirely dispenses with the notion of domain and relies on data selection techniques to handle data heterogeneity.
-
-\section{Conclusion and outlook \label{sec:conclusion}}
-\fyDone{Write conclusions}\fyTodo{no mixture of experts}
-In this study, we have carefully reconsidered the idea of multi-domain machine translation, which seems to be taken for granted in many recent studies.
-We have spelled out the various motivations for building such systems, and associated expectations in terms of system performance.
-We have then designed a series of requirements that MDMT systems should meet, and designed associated testing procedures.
-In our experiments with multiple MDMTs, we have found that most requirements were not met for our experimental conditions.
-If MDMT systems are able to outperform the mixed-domain baseline, at least for some domains, they all fall short to match the performance of fine-tuning on each individual domain, which remains the best choice in multi-source single domain adaptation.
-In this comparison, architecture-based strategies seem to show the best promises.
-\fyTodo{Mitigate this}
-As expected however, MDMT are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend,
-dynamically accommodate additional domains:this is especially easy for feature-based methodologies.
-Our experiments finally suggest that all methods show decreasing performance when the number of domain increases (for a fixed data size).
-
-Two other main conclusions can be drawed from this study: first, it seems that more work is needed to make MDMT systems make the best out of the variety of the available data, both to effectively share what needs to be shared while at the same time separating what needs to be separated.
-We notably see two areas worthy of further explorations: the development of parameter sharing strategies when the number of domains is large; the design of training strategies that can effectively handle a change of the training mixture, including an increase in the number of domains.
-Both problems are of practical relevance in industrial settings.
-Second, and may be more importantly, there is a general need to adopt better evaluation methodologies for evaluating MDMT systems, require systems developers to clearly spell out the testing conditions and the associated expected distribution of testing instances, and to report more than mere comparisons with simple baselines on a fixed and known pair of domains.
-
-% --------------
-% \section*{Acknowledgements}
-% This work was granted access to the HPC resources of [TGCC/CINES/IDRIS] under the allocation 20XX- [numéro de dossier] made by GENCI (Grand Equipement National de Calcul Intensif)
-% The acknowledgements should go immediately before the references.  Do
-% not number the acknowledgements section. Do not include this section
-% when submitting your paper for review.
-
-% include your own bib file like this:
-% \bibliographystyle{coling}
 \bibliographystyle{acl_natbib}
-\bibliography{multidomain}
-
-\todos{}
+\bibliography{emnlp2020}
+\appendix
+\section{Appendices}
+\label{sec:appendix}
+jkk
+\section{Supplemental Material}
+\label{sec:supplemental}
 
 \end{document}
-
-\appendix{}
-
-\begin{table}
-  \centering
-  \begin{tabular}{|p{3.2cm}|*{30}{r|}} \hline
-    Model / Set-up &d1&d2&d3&d4&d5&d6&d7&d8&d9&d10&d11&d12&d13&d14&d15
-    &d16&d17&d18&d19&d20&d21&d22&d23&d24&d25&d26&d27&d28&d29&d30 \\\hline
-    %
-    & \multicolumn{30}{|p{20cm}|}{%
-      }% \hline
-    \\
-    %
-    domain size ($\times 10^3$) &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\ \hline\hline
-    %
-   \system{FT-Full}      &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\  
-    \system{DC-Tag}    &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\     
-    \system{DC-Feat}   &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\ 
-    \system{TTM}         &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\ 
-    \system{LDR}  &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\
-
-    \system{MDL Res} 
-%    &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \hline
-  \end{tabular}
-  \caption{Detailed results of the experiments with automatic domains}
-  \label{tab:automatic_domains}
-\end{table}
-
-\subsection{Formalizing multi-domain adaptation (MDA)}
-
-In MDA, the typical training situation (for a classifier) considers $k$ domains with $\mathcal{D}_k(x,y)$ the joint distribution of features $x \in \mathcal{X}$ and labels $y \in \mathcal{Y}$ for domain $k$.
-In \cite{Mansour09domainadaptation}, the labeling function $f$ is deterministic and the joint distribution is entirely defined by the marginal $\mathcal{D}_k(x)$ and $f()$.This work further assumes $k$ hypotheses $h_k(x): \mathcal{X} \rightarrow{} \mathcal{Y}$ with an error smaller than $\epsilon$ when $x\sim \mathcal{D}_k$: $\sum_{x} L(h_k(x), f(y)) \mathcal{D}_k(x) < \epsilon_k$ and that the target distribution is a convex combination of $\{\mathcal{D}_k\}$: $\mathcal{D}_T(x) = \sum_k \lambda_k \mathcal{D}_k(x)$.
-\cite{Mansour09domainadaptation} focuses on \textsl{combined hypotheses}, ie.\ hypotheses for which each prediction $h(x)$ is a linear combination of domain specific hypotheses $h(x_k)$, and derives generalization bounds for the case where $\mathcal{D}_T(x)$ is fixed in advance, or worst case bounds valid for any mixture target distribution.
-The optimal solution has the form $h_{z,\eta}(x) = \sum_k \frac{z_kD_k(x) + U(x)/\gamma}{\sum_{k'} z_kD_k(x) + \eta U(x)/l}$ and is termed the \emph{distribution weighted combined rule}.
-Further work in \cite{Mansour09multiple} generalizes these bounds to the case where $\mathcal{D}_T$ is not a mixture of existing domains (in that case we will find a mixture that is close to $\mathcal{}$in the Renyì-divergence sense ); where $\mathcal{D}_k$ are unknown and estimated from the data; or when the labelling function is not the same across domains.
-More recent work study the case where the labelling function is stochastic, meaning that domains are defined by joint ($\mathcal{D}_k(x,y)$) rather than marginal distributions \cite{Hoffman18algorithms}, and provides an algorithm to compute the optimal weights for the combined rule based on estimates of the joint distributions.
-A guiding principle is that this rule should tend towards uniform performance across domains, to ensure robustness in the worst case scenario \cite{Oren19distributionally}.
-
-A related problem is \emph{Multiple source single domain adaptation} which is a variation of conventional DA from a mixture of source domains.
-In this setting it is advantageous to take advantage of the similarities between each source and the fixed target domain (see eg.\ \cite{Duan12domain,Cortes19AdaptationBO,Wen19domain} for studies of the supervised case, \cite{Zhao18adversarial} for the unsupervised case).
-
-\subsection{Complements - to be removed \label{ssec:oldformalization}}
-
-We define a domain $d$ as a joint distribution $\mathcal{D}_d(x,y)$ over input-output pairs $(x,y) \in \mathcal{X}\times{}\mathcal{Y}$.
-Assuming that the training data is sampled from a known mixture of $D$ source domains $\mathcal{D}_s(x,y) = \sum_d \lambda_{s,d} \mathcal{D}_d(x,y)$, several cases are studied in the ML literature:
-
-\begin{itemize}
-\item the test distribution $\mathcal{D}_t$ is known 
-  \begin{itemize}
-  \item $\mathcal{D}_t = \mathcal{D}_{s}$: this is standard ML, with mixture densities
-  \item $\mathcal{D}_t(x,y) = \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) $ with a different mixture of domains in training and testing ($\vlambda_s \neq \vlambda_t$), corresponding to multiple source adaptation (MSA) \cite{Mansour09domainadaptation},
-  \item $\mathcal{D}_t(x,y) \neq \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) \forall \lambda$, where the target domain for adapation is not a mixture of source domains, studied in \cite{Mansour09multiple}
-  \item variants of MSA which only assume knowledge of the marginal distribution $\mathcal{D}_t(x) = \sum_{y} \lambda_{t,d} \mathcal{D}_d(x) $
-  \item more realistic situations where  the train and test distributions are estimated from finite samples, also studied in \cite{Mansour09multiple}.
-  \end{itemize}
-   Note that we recover standard supervised DA is when there is only one component in the training and test mixtures, and multiple-source single domain adaptation when the source is a mixture but the test domain is homogeneous.
-Unsupervised DA \cite{Daume06domain} corresponds to the case where only the marginal distribution of the test domain inputs is available, while semi-supervised DA is the reverse situation.
-
-\item the test distribution is not known and training needs to be distributionally robust, ie.\ minimize the error against all possible distributions, with variants depending on whether it is or not assumed to be a mixture of existing domains, and whether it is perfectly or imperfectly known.
-This setting is also studied in eg.\ \cite{Mansour09multiple,Hoffman18algorithms}, and  in \cite{Oren19distributionally} for adversarial language model learning.
-\item with respect to the conditional label distribution $P(y|x)$:
-  \begin{itemize}
-  \item $P(y|x)$ is deterministic, meaning that the joint distribution is entirely specified by the marginal and a labelling function $y=h(x)$; a more complex case is when $h()$ also varies across domains;
-  \item $P(y|x)$ is stochastic, but shared across all domains, which is the typical covariate-shift scenario \cite{Shimodaira00improving}; it can also vary across domains \cite{???}\fyTodo{Missing citation vary accross domains}
-  \end{itemize}
-\end{itemize}
-
-Multi-domain learning, as defined in \cite{Dredze08online} additionally makes the requirement that categorial domain tags are available in training and testing; a multi-domain learner should use them effectively \cite{Joshi12multidomain} to improve simultaneously on all domains \cite{Finkel09hierarchical}.
-This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyDone{explain zero shot}.
-% ----
-
-% \subsection{MDMT: a tentative definition \label{ssec:definition}}
-
-% Informally, a multi-domain system should be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
-% \begin{itemize}
-% \item the more common situation is when both train and test documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
-It is usually assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT.
-This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name here: such as supervised MDMT ?
-and mostly supervised MDMT for the second case ?}
-
-% \item It can also be the case that domain labels are only known in training, while test inputs may or may not be associated with explicit domains.
-This is, for instance, the setting adopted \cite{Zeng18multidomain}, where automatic domain identification is performed for test sentences.
-Note that automatic domain identification can also be effectively performed in the first setting, especially in the presence of close or related domains in training.
-
-% \item A last setting, that has not received much attention for MT (see however \cite{Sennrich13multidomain}), is where both train and test domains are automatically infered from data, using clustering techniques.
-Even though domain labels are typically available in most training scenario, these approaches assumes that might be preferable, owing to the multiple sources of intra-corpus and inter-corpus variabilities, to resort to automatic domain assignments.
-The online approach of \newcite{Farajian17multidomain}, where distributional cues, rather that categorical tags, are used for fine-tuning, is a variant of this strategy.
-% \end{itemize}
-
-% In the rest of the paper, we mostly adhere to the first definition which correspond to a situation of practical interest for the industry, and assume that multi-domain MT is mostly about effectively using domain tags in training and testing.
-We also experiment in Section~\ref{sec:results} with the other two alternative scenarios, as they also define useful contexts to evaluate the actual robustness of MDMT systems.
-
-% \jcDone{...a very common scenario in MT industry where clients dispose of mulitiple data corresponding to different domains (differing in genre, thematic, register, style, etc.). }
-
-% \subsection{Evaluation issues \label{ssec:evaluation}}
-
-% In theory, evaluation procedures should report realistic estimates of the test loss, computed on a sample deemed representative of the anticipated test conditions.
-Three scenarios need be distinguished:
-% \begin{itemize}
-% \item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
-% \item all domains are equally likely and important in testing (the prior over test domains is uniform), which would warrant the computation of per domain scores and of unweighted averages, as reported in \cite{Farajian17multidomain,Su19exploring,Wang20general}.
-Note that in this situation, the optimal training regime would be to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data.
-\cite{Wang20general} uses an intermediate mixing regime, which tends to reduce part of the unbalance, and improve the performance for underrepresented domains.
-% \item the test distribution is neither equal to the training set, nor uniform.
-Here again, if this distribution is known in advance, an optimal domain mix should be used in training to compute baseline scores, rather than using the empirical distribution of all the training data.
-A more challenging case is when the test set includes data from domains unseen in training, or when it is chosen adversarially \cite{XXX}.
-% % : this situation, where test instances have a null probability in training, goes agains the basic statistical assumptions of DA \cite{Shimodaira00improving}.
-It is thus important that the likeliness of such events should be clearly formulated when describing the evaluation protocol.
-% \end{itemize}
-
-% The papers in the literature are seldom explicit regarding their assumptions regarding the test distribution.
-In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 most faithfully reflects the expected testing conditions.
-Another important issue is the choice of the evaluation metrics: if BLEU is still the standard metric used in most studies, it is not necessary the best choice to measure subtle improvements of lexical choices \cite{Irvine13measuring} that are often associated with adapted systems.
-
-% A last issue regarding evaluation is related to the measurement of computational costs: when this is the main a motivation for using MDMT, we expect that computational costs would be reported and compared, which however is hardly ever done in practice.
-\fyTodo{Keep this ?}

### 2020-06-04 2020/phd_midterm_report/report.tex
spelling error

@@
 To avoid being over-fitted by positive examples in which every sentence pair is correctly matched, we proposed to use the same amount of artificial negative examples in which words are randomly removed or added and changed in several positions.
-\section*{Working plan}
+\section*{Provisional planning}
 \begin{itemize}
-	\item Completing work on residual adapters and submission to WMT 2020 in the end of July 2020.
-	\item Completing system submission to Machine Translation Robustness task of WMT 2020 in the end of July 2020.
-	\item Completing work on contextualized Machine Translation with Systran's team in the end of July 2020.
-	\item Completing work on Multi-source Adaptation between September - December 2020
-	\item Completing work on BPE-vocabulary's transfer learning effect between January and March 2021.
-	\item 
+	\item Completing work on residual adapters and submission to WMT 2020 at the end of July 2020.
+	\item Completing system submission to Machine Translation Robustness task of WMT 2020 at the end of July 2020.
+	\item Completing work on contextualized Machine Translation with Systran's team at the end of July 2020.
+	\item Completing work on Multi-source Adaptation between August - December 2020
+	\item Completing work on BPE-vocabulary's transfer learning effect between January and April 2021.
 \end{itemize}

### 2020-06-04 2020/phd_midterm_report/report.tex
spelling error
correct pointed errors

@@
 In the class of feature-augmentation, \cite{Kobus17domain}, \cite{kenji17multi}, \cite{Pham19generic} proposed to add domain-tags and learn-able domain-embeddings to existing word feature of NMT model.
-These methods are cheap and efficient in terms of gain over model-size.
+These methods achieve a good trade-off between performance gain and the number of additional parameters.
 
@@
 
-\subsection*{Realized work}
+\subsection*{Past contribution}
 The work of \cite{Kobus17domain} and \cite{Daume07frustratingly} showed promising results for learning in a multi-domain context.
@@
 	\item Completing work on contextualized Machine Translation with Systran's team at the end of July 2020.
-	\item Completing work on Multi-source Adaptation between August - December 2020
+	\item Completing work on Multi-source Adaptation between August - December 2020.
 	\item Completing work on BPE-vocabulary's transfer learning effect between January and April 2021.

### 2020-06-04 2020/phd_midterm_report/report.tex
last comments

@@
 Neural Machine Translation’s quality is assured both theoretically and empirically under the assumption of the matching between training data distribution and testing data distribution.
-However, this situation is hardly true in real contexts where the NMT model is often trained with big training data which is heterogeneous in terms of topic and size of the size of topics data, and one needs to make the best to robustly handle inputs from unexpected domains in the testing phase.
+However, this situation is hardly true in real contexts where the NMT model is often trained with large amounts of training data which are heterogeneous in terms of topic and size, and one needs to make the best to robustly handle inputs from unexpected domains in the testing phase.
 \subsection*{Overview}
@@
 Methods in this category improve the low resource model by collecting domain-related examples from out-of-domain resources.
-For the statistical model, \cite{mansour12simple} investigated the use of the Moore-Lewis score to re-estimate the phrase table.\cite{Duh13selection,Axelrod11domain,silva2018extracting} proposed several methods to select examples for training NMT model by scoring based on the cosine similarity of distributed sentence embedding or based on the probability calculated by in-domain Neural Language Model and the cosine similarity of $tf\_idf$ embedding adjusted by Feature decay algorithm.
+For the statistical model, \citep{mansour12simple} investigated the use of the Moore-Lewis score to re-estimate the phrase table.\citep{Duh13selection,Axelrod11domain,silva2018extracting} proposed several methods to select examples for training NMT model by scoring based on the cosine similarity of distributed sentence embedding or based on the probability calculated by in-domain Neural Language Model and the cosine similarity of $tf\_idf$ embedding adjusted by Feature decay algorithm.
 Another way to get more training is to create pseudo parallel data, \cite{Junjie19domain} proposed an unsupervised method to create pseudo in-domain training data using word by word translation.
@@
 \subsubsection*{Model-based adaptation methods}
-These approaches take interest in the model architecture and model training process.
+These approaches take interest in the model architecture and in the training process.
 We can further categorize them into 3 classes: feature-augmentation, module-augmentation, surrogate losses, and finetuning.
-In the class of feature-augmentation, \cite{Kobus17domain}, \cite{kenji17multi}, \cite{Pham19generic} proposed to add domain-tags and learn-able domain-embeddings to existing word feature of NMT model.
+Regarding feature-augmentation, \cite{Kobus17domain,kenji17multi,Pham19generic} proposed to add domain-tags and learn-able domain-embeddings to existing word feature of NMT model.
 These methods achieve a good trade-off between performance gain and the number of additional parameters.
 
-Module-augmentation is more expensive then feature-augmentation because one adds one or several complementary layers into standard NMT architecture; \cite{bapna19simple},\cite{Michel18extreme},\cite{Zheng18multi},\cite{Vilar18learning}, \cite{Britz2017mixing} are included in this category.
+Model augmentation is more expensive than feature-augmentation because one adds one or several complementary layers into standard NMT architecture; \cite{bapna19simple,Michel18extreme,Zheng18multi,Vilar18learning,Britz2017mixing} are included in this category.
 Residual module in \cite{bapna19simple} and Hidden units in \cite{Vilar18learning} (\ref{fig:overview}) are used as a plug-in which is used to fine-tune the model without
-changing the value of the main model.
-\cite{Zheng18multi} (\ref{fig:overview}) added 2 gates to extract domain agnostic representations and domain-specific representations from the output of encoder which are then plugged to the decoder.
+changing the parameters of the main model.
+\citet{Zheng18multi} (\ref{fig:overview}) added 2 gates to extract domain agnostic representations and domain-specific representations from the output of encoder which are then plugged to the decoder.
 \cite{Britz2017mixing} (\ref{fig:overview}) also plugged a domain-classifier on top of the encoder to force the domain-specific representation.
@@
 The architecture is shown in figure (\ref{fig:overview}).
-\cite{dakwle17fine} (\ref{fig:overview}) proposed a multi-output layer for domain-model which are trained by minimizing the cross-entropy for in-domain examples and Kullback-Leibler distance to
+\cite{dakwle17fine} (\ref{fig:overview}) proposed a multi-output layer for domain-model which are trained by minimizing the cross-entropy for in-domain examples and the Kullback-Leibler distance to
 the conditional distribution of a teacher generic model.
-The proposed model has 2 output layers, each of them optimizing one of 2 losses respectively.
+The proposed model has 2 output layers, each of them optimizing one of two losses respectively.
 \cite{Dou19Unsupervised} attempted to use only monolingual target data for adapting to a new domain by using additional learnable domain embedding and task embedding and introducing language modeling tasks in the target language.
@@
 
-\subsection*{Past contribution}
+\section*{Contributions to date}
 The work of \cite{Kobus17domain} and \cite{Daume07frustratingly} showed promising results for learning in a multi-domain context.
@@
 A multi-task learner consists of one or stack of multiple shared lower layers and task-related modules above these later layers.
-By optimizing k objective functions at once, the shared layers are pulled to a compromising region between task-optimal regions.
+By optimizing $k$ objective functions at once, the shared layers are pulled to a region which represents a trade-off between task-optimal regions.
 However, multi-task training has to overcome several problems such as the heterogeneity of data size, convergence rates, etc.
@@
 We reported promising results and proposed another version more adaptive.
+This contribution has been published in \cite{Pham19generic}.
 
@@
 We took interest in designing new evaluation setups that access several desirable properties of a multi-domain system.
-We revisited the work of \cite{Britz17effective} \cite{Zheng18multi} \cite{Kobus17domain} and \cite{bapna19simple} in a large scale context in which there are 6-7 domains.
+We revisited the work of \cite{Britz17effective,Zheng18multi,Kobus17domain} and \cite{bapna19simple} in a large scale context in which there are 6-7 domains.
+This contribution has been published in a submitted long paper \citep{Pham20revisiting}.
 
-\section*{Lexicalized Domain Representation}
+\subsection*{Lexicalized Domain Representation}
 \cite{Kobus17domain} proposed to stack a domain-feature embedding to each word embedding whenever the domain appears.
@@
 \begin{figure}[h!]
-  \center
+  \centering
   \includegraphics[width=0.48\textwidth]{embeddings}
@@
 In the Transformer setting, LDR improves by a large margin the performance on domain medical and achieves on-par performance compared to the WDCNMT of \cite{Zheng18multi} on RNN architecture.
-More detailed results have been published in \cite{Pham19generic}.
+Detailed results have been published in \cite{Pham19generic}.
 
-\section*{Revisiting multi-domain NMT}
-Even though many methods are proposed to solve the multi-domain problem, they were not fully evaluated.
+\subsection*{Revisiting multi-domain NMT}
+Even though many methods are proposed to solve the multi-domain problem, in our view they were not fully evaluated.
 We introduced several new objectives that the multi-domain NMT system is supposed to achieve.
-The following are basic properties we want an NMT model to have.
+The following are basic properties we would like an NMT model to have.
 \begin{enumerate}
@@
 
-The first property is basic because, in the real context, in-domain corpora often have a very variable size.
+The first property is basic because, in real contexts, in-domain corpora often have a very variable size.
 Indeed, among OPUS corpora, the religion domain with only 130K parallel sentences (Tanzil) is very small compared to the medical domain with more than 2.5M parallel sentences (UFAL).
@@
 
-The last property is also an important and active trend in machine learning. The model needs to preserve the learned knowledge while learning as being exposed to new data.
+The last property is also an important and an active trend in machine learning. The model needs to preserve the learned knowledge while learning when exposed to new data.
 
@@
 
-According to our study on the state of the art of Multi-domain NMT, this is the first large scale experiment.
+According to our study on the state of the art of Multi-domain NMT, ours is the first large scale experiment.
 We experimented with translation from English into French and use texts initially originating from 6 domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in an \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
@@
 
-We used a test set split from the original corpus of \domain{news} to evaluate the robustness before an unknown domain.
+We also used a test set from the \domain{news} domain to evaluate the robustness with respect to unknown domain.
 For methods requiring domain-information to predict, given a source sentence, we predict the closest domain to the sentence among already known domains.
@@
 
-\section*{Multi-task network}
-We can approach the multi-domain problem by multi-task learning by considering adaptation to one domain as a task.
-\cite{Caruana97multitask} showed multi-task training improving the performance of the model for each independent task.
+\subsection*{Multi-task network}
+We can approach the multi-domain problem by multi-task learning by considering adaptation to each domain as a separate task.
+\cite{Caruana97multitask} showed that multi-task training can improve the performance of the model for each independent task.
 \cite{rebuffi18efficient} proposed the first use of a residual adapter which is a small layer plugged on top each layer to adapt classifier to an arbitrary domain.
 \cite{bapna19simple} reapplied the latter idea to domain adaptation in NMT.
-Inspired by the mentioned work in multi-task learning and domain adapter, we applied residual adapters to encoder and decoder and train the whole set of parameters from scratch.
+Inspired by the mentioned work in multi-task learning and domain adaptation, we applied residual adapters to encoder and decoder and train the whole set of parameters from scratch.
 There is a small difference, instead of plugging the adapter in a sequential scheme, we propose a high-way scheme in which each adapter bridges directly from each intermediate layer to the output of the encoder.
@@
 
-Multi-task training for K domains involves K losses on K empirical distributions. 
+Multi-task training for $K$ domains involves $K$ losses on $K$ empirical distributions. 
 \begin{equation}
@@
 \end{equation}
-where $\lambda_i$ is regularization scale for the adapter of domain i, $Adapt^j_i$ is the adapter for layer j in domain i.
+where $\lambda_i$ is regularization scale for the adapter of domain i, $Adapt^j_i$ is the adapter for layer $j$ in domain $i$.
 We could also adjust the capacity of each adapter to match the size of the domain training data.
@@
 Generalization is a basic requirement of any learner, even when in-domain data is available.
-NMT system often loses performance dramatically in the previous training domain after being fine-tuned to the target domain.
+NMT system often dramatically loses performance in the previous training domain after being fine-tuned to the target domain.
 The phenomenon is well known under the name of catastrophic forgetting which was first reported in \cite{Michael89catastrophic}.
@@
 The second problem of our interest is robustness with respect to out-of-domain data.
-Given several in-domain datasets in training, \cite{hoffman18algorithms}, and a test set from an unknown domain, we would like to build a good predictor from several out-of-domain expert predictors.
+Given several in-domain datasets in training and a test set from an unknown domain, we would like to build a good predictor from several out-of-domain expert predictors; this setting is studied notably in \citep{hoffman18algorithms}.
 In the machine-learning literature, the problem can be referred to as Multi-source Domain Adaptation.
-\cite{yishay09multiple}, \cite{mansour09domain} studied effective methods to combine learned predictors to predict test examples that come from an arbitrary unknown domain.
+\cite{yishay09multiple,mansour09domain} studied effective methods to combine learned predictors to predict test examples that come from an arbitrary unknown domain.
 They proposed a convex combination of learned in-domain predictors whose weights are computed by the in-domain probabilistic model.
@@
 
-\section*{Other research}
+\section*{Other completed works}
+
 \subsection*{Semantic similarity}
@@
 The training of the network is done by optimizing the aggregation of soft word alignment scores between 2 sentences.
-To avoid being over-fitted by positive examples in which every sentence pair is correctly matched, we proposed to use the same amount of artificial negative examples in which words are randomly removed or added and changed in several positions.
-\section*{Provisional planning}
+To avoid being over-fitted by positive examples in which every sentence pair is correctly matched, we proposed to use the same amount of artificial negative examples in which words are randomly removed or added and changed in several positions
+.
+\section*{Planning for the next 18 months}
+
 \begin{itemize}
-	\item Completing work on residual adapters and submission to WMT 2020 at the end of July 2020.
-	\item Completing system submission to Machine Translation Robustness task of WMT 2020 at the end of July 2020.
-	\item Completing work on contextualized Machine Translation with Systran's team at the end of July 2020.
-	\item Completing work on Multi-source Adaptation between August - December 2020.
-	\item Completing work on BPE-vocabulary's transfer learning effect between January and April 2021.
+\item Complete work on residual adapters and submission to WMT 2020 at the end of July 2020.
+\item Complete system submission to Machine Translation Robustness task of WMT 2020 at the end of July 2020.
+\item Complete work on contextualized Machine Translation with Systran's team at the end of July 2020.
+\item Complete  work on Multi-source Adaptation between August - December 2020.
+\item Complete work on BPE-vocabulary's transfer learning effect between January and March 2021.
+\item Finalize publications, write PhD memoire between March and August 2021.
+\item Defend PhD September 2021.
 \end{itemize}
+
 \bibliography{bibliography}

### 2020-06-19 2020/wada/adaptive_domain_adaptation.tex
wada summary

\documentclass[12pt,a4paper,twoside]{report}
\usepackage[left=3cm,right=3cm,top=2cm,bottom=3cm]{geometry}
\usepackage{pagenote}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}	% Para caracteres en español
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{multirow,booktabs}
\usepackage[table]{xcolor}
\usepackage{fullpage}
\usepackage{lastpage}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{mathrsfs}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{calc}
\usepackage{color,soul}
\usepackage{multicol}
\usepackage{cancel}
\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{xcolor}
\colorlet{shadecolor}{orange!15}
\parindent 0in
\parskip 12pt
\geometry{margin=1in, headsep=0.25in}
\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{reg}{Rule}
\newtheorem{exer}{Exercise}
\newtheorem{note}{Note}
\usepackage{listings}
\usepackage{spverbatim}
\usepackage{fancyvrb}
\usepackage{hyperref}
\usepackage{float} 
\usepackage{natbib}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{soulutf8}
\usepackage{color,soul}
\usepackage{xcolor}
\usepackage{tabularx}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{definition}{Definition}[section]
\DeclareMathOperator*{\argmax}{argmax}
\usepackage[draft]{todo}
\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\pgfkeys{
    /tr/rowfilter/.style 2 args={
        /pgfplots/x filter/.append code={
            \edef\arga{\thisrow{#1}}
            \edef\argb{#2}
            \ifx\arga\argb
            \else
                \def\pgfmathresult{}
            \fi
        }
    }
}
\usepackage{mathtools,xparse}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\NewDocumentCommand{\normL}{ s O{} m }{%
  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{L_1(\Omega)}%
}

\begin{document}
\begin{center}
{\LARGE \bf Combination of generic and specific}\\
\end{center}
\setlength{\belowdisplayskip}{8pt} \setlength{\belowdisplayshortskip}{8pt}
\setlength{\abovedisplayskip}{8pt} \setlength{\abovedisplayshortskip}{8pt}
\setlist{nosep}
\setlength{\parskip}{0.1cm}
\setlength{\parindent}{1em}
\section*{Word Level Adaptive Domain Adaptation}
Let $h$ is the output of the encoder which is pretrained, the adapted representation will be $h' = h + a_k(h)$. In gating version, $h' = h * (1-z_k) + a_k(h) * z_k$ where $z_k$ is also a function of $h$ taking value in $[0,1]$. More precisely, $h$ and $h'$ is a sequence of context vectors , and the combination is element wise, i.e, $h'(w) = h(w) * z_k(h(w)) + a_k(h(w)) * z_k(h(w))$. \\
In Word Level Adaptive Domain Adaptation version, $z_k(h(w))$ is designed so that the less probably the word $w$ is in domain $k$ the less $z_k(h(w))$ is. The Word Level Adaptive Domain Adaptation aims to reduce the variation caused by $a_k$ of adapted context vector $h'$ compared to $h$ on words that do not surely belong to domain k. For a domain, there are 2 type of "mistrusted" words: out-of-domain words that rarely or never appear in the in-domain data; or popular words that appear frequently in many domains. By reusing learned representation of the "mistrusted" words, we can at least guarantee the risk of bad prediction in case of out-of-domain words by the risk of the learned generic model.
\section*{Training process}
There 3 steps:
\begin{itemize}
	\item Pretraining a generic model with mixed corpora.
	\item Training a domain classifier on top of the encoder and decoder; freezing the parameters of the generic model.
	\item Training parameters of adapters with in-domain data separately while freezing the parameters of the generic model and of the domain classifiers.
\end{itemize}

\section*{Weakness of the method}
The new method depends essentially to the quality of the classifier. Moreover, the classifier can fail to detect in-domain words in the case of polysemous words. For example, the word "joint" in joint stereo, joint sprain and joint project has 2 different translations as "joint" and "articulaire" in french. In the unbalanced dataset as the mixed corpora (6 in-domain corpora),  "joint stereo"

\section*{Analysis on the risk of new method}

The combination of generic representation and the residual adapter in the adapted model is as following
\begin{equation}
h' = h + a_k(h)
\label{eq:1}
\end{equation}

where h is output of sharing layer, i.e $h=M(x,y)$, $M: \mathbb{X} \times \mathbb{Y} \rightarrow \mathbf{\Omega}\subset \mathbb{R}^{d_{model}}$, $a_k$ is the residual adapter which is a function $\mathbb{R}^{d_{model}} \rightarrow \mathbb{R}^{d_{model}}$. \\

The problem of catastrophic forgetting happens when the value of $a_k(h)$ has large magnitude for value of h which is output of an out-of-domain example $(x^{adv}, y^{adv})$ (adv means adversarial). Indeed, finetuning such adapter with only in-domain data, $a_k$ never encounters adversarial examples, thus does not guarantee the risk of prediction on out-of-domain examples. However, if we control the weight of $a_k(h)$ is the combination, we might reduce the wrong belief of $a_k$ on adversarial examples. We then investigate a form more flexible of the latter combination.

\begin{equation}
h' = h * (1-f(z_k)) + a_k(h) * f(z_k)
\label{eq:2}
\end{equation}

where 
\begin{itemize}
	\item $z_k = P(d=k | h=M(x,y)) = p(h)$ where $p: \mathbf{\Omega} \rightarrow \mathbf{\Delta}^{K}$
	\item f is an activation function $f: [0,1] \rightarrow [0,1]$ 
\end{itemize}
We define some important constants:
\begin{itemize}
	\item Domain inseparability $\mathbb{\varepsilon}_{insep} = \displaystyle{\mathop{\sum}_{k' \neq k} \mathbb{E}_{x,y}[\mathbf{1}_{d=k'}*p_k(h)]}$. where $h=M(x,y)$
	\item Risk of generic model $\mathbb{R} = \mathbb{E}_{x,y}[l(h,y)]$
	\item $\mathbf{\Omega_{k,p_{0}}} = \lbrace h | z_k = p_k(h)<0.8\rbrace$.
	\item Risk of adapted model in in-domain test set $R_k = \displaystyle{\mathbb{E}_{x,y \sim p(.|d=k)}[l(h * (1-f(z_k)) + a_k(h) * f(z_k), y)]}$
\end{itemize}
We also have 2 assumptions on the regularization of activation function:
\begin{itemize}
	\item $a_k$ is bounded in sense of $\normL{a_k(h)} < M$
	\item loss function is bounded in sense of $l(h * (1-f(z_k)) + a_k(h) * f(z_k),y) < A$ $\forall x,y$
\end{itemize}
We then access the risk of of the model adapted to domain k with the residual $a_k$ to arbitrary examples of K domains.
\begin{equation}
\begin{split}
R(M,a_k) &= \mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y)] \\
		&= \mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] + \\
		&+ \mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h)]
\end{split}
\end{equation}

By designing a suitable activation function, we can map $f: [0,p_{0}] \rightarrow [0,\delta]$. Then with $\delta$ is small enough we can have below approximation of the first term:
\begin{equation}
\mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] \sim \mathbb{E}_{x,y}[l(h,y)\mathbf{1}_\mathbf{\Omega_{k,p_{0}}}(h)] \leq \mathbb{E}_{x,y}[l(h,y)] = R
\label{eq:4}
\end{equation}
The second term:
\begin{equation}
\begin{split}
\mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h)] &= \displaystyle{\mathop{\sum}_{k'}\mathbb{E}_{x,y \sim p(.|d=k')}[l(h * (1-f(z_k))} \\
	& + a_k(h) * f(z_k),y) *\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h)]P(d=k') \\
	& \leq \displaystyle{\mathop{\sum}_{k' \neq k}} A * \mathbb{E}_{x,y \sim p(.|d=k')} [\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h)]P(d=k') \\
	&+ \mathbb{E}_{x,y \sim p(.|d=k)}[l(h * (1-f(z_k)) + \\
	&+ a_k(h) * f(z_k),y) * \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h)]P(d=k) \\
	&\leq A * \frac{\mathbb{\varepsilon}_{insep}}{p_{0}} + R_k * P(d=k)
\end{split}
\label{eq:5}
\end{equation}

The last inequality is verified by the below inequality:
\begin{equation}
\begin{split}
\displaystyle{\mathop{\sum}_{k' \neq k}} A * \mathbb{E}_{x,y \sim p(.|d=k')} [\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h)]P(d=k') &\leq A * \displaystyle{\mathop{\sum}_{k' \neq k}} \mathbb{E}_{x,y \sim p(.|d=k') } [\frac{p_k(h)}{p_0}]P(d=k') \\
		& = \frac{A}{p_0} \mathbb{E}[\mathbf{1}_{d\neq k}p_k(h)] \\
		& \leq \frac{A}{p_0} \mathbb{\varepsilon}_{insep}
\end{split}
\end{equation}
because
$$ \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h) \leq \frac{p_k(h)}{p_0} $$
Combining the inequality \ref{eq:4} and the inequality \ref{eq:5} we obtain the risk in general case of the adapted model $(M,a_k)$ $$R(M,a_k) \leq R + \frac{A * \mathbb{\varepsilon}_{insep}}{p_0} + P(d=k)R_k$$

\end{document}

### 2020-06-19 2020/phd_midterm_report/report.tex
rewriting

@@
 The training of the network is done by optimizing the aggregation of soft word alignment scores between 2 sentences.
-To avoid being over-fitted by positive examples in which every sentence pair is correctly matched, we proposed to use the same amount of artificial negative examples in which words are randomly removed or added and changed in several positions
+To avoid being over-fitted to positive examples in which every sentence pair is correctly matched, we proposed to use the same amount of artificial negative examples in which words are randomly removed or added and changed in several positions.
 .
@@
 \item Complete work on BPE-vocabulary's transfer learning effect between January and March 2021.
-\item Finalize publications, write PhD memoire between March and August 2021.
-\item Defend PhD September 2021.
+\item Finalize publications, write PhD memoire between April and September 2021.
+\item Defend PhD October 2021.
 \end{itemize}

### 2020-06-19 2020/wada/adaptive_domain_adaptation.tex
rewriting

@@
 
-The problem of catastrophic forgetting happens when the value of $a_k(h)$ has large magnitude for value of h which is output of an out-of-domain example $(x^{adv}, y^{adv})$ (adv means adversarial).
-Indeed, finetuning such adapter with only in-domain data, $a_k$ never encounters adversarial examples, thus does not guarantee the risk of prediction on out-of-domain examples.
+The bad performance of fine-tuned model happens as the value of $a_k(h)$ of the out-of-domain example $(x^{adv}, y^{adv})$ is adversarial to the model.
+Indeed, by fine-tuning such adapter with only in-domain data, $a_k$ never encounters adversarial examples, thus does not guarantee the risk of prediction on out-of-domain examples.
 However, if we control the weight of $a_k(h)$ is the combination, we might reduce the wrong belief of $a_k$ on adversarial examples.
-We then investigate a form more flexible of the latter combination.
+To execute the idea, we  investigate the version of convex combination.
 

### 2020-06-25 2020/wada/adaptive_domain_adaptation.tex
correct errors
fix math and English, some comments

@@
 \section*{Word Level Adaptive Domain Adaptation}
-Let $h$ is the output of the encoder which is pretrained, the adapted representation will be $h' = h + a_k(h)$.
-In gating version, $h' = h * (1-z_k) + a_k(h) * z_k$ where $z_k$ is also a function of $h$ taking value in $[0,1]$.
-More precisely, $h$ and $h'$ is a sequence of context vectors , and the combination is element wise, i.e, $h'(w) = h(w) * z_k(h(w)) + a_k(h(w)) * z_k(h(w))$.
-\\
-In Word Level Adaptive Domain Adaptation version, $z_k(h(w))$ is designed so that the less probably the word $w$ is in domain $k$ the less $z_k(h(w))$ is.
-The Word Level Adaptive Domain Adaptation aims to reduce the variation caused by $a_k$ of adapted context vector $h'$ compared to $h$ on words that do not surely belong to domain k.
-For a domain, there are 2 type of "mistrusted" words: out-of-domain words that rarely or never appear in the in-domain data; or popular words that appear frequently in many domains.
-By reusing learned representation of the "mistrusted" words, we can at least guarantee the risk of bad prediction in case of out-of-domain words by the risk of the learned generic model.
+Let $h$ the output of the pretrained encoder, the adapted representation for domain $k$ will be $h' = h + a_k(h)$, meaning that all words in all sentences for domain $k$ will use the adapter module $a_k$.
+A ``gated'' version uses $h' = h * (1-z_k) + a_k(h) * z_k$\footnote{More precisely, $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure XXX...}\fyTodo{$a_k(h)=$ somme tous les composants résiduels} where $z_k$ is also a function of $h$ taking values in $[0,1]$.
+More precisely, $h$ and $h'$ are sequences of context vectors and the combination is performed element-wise, yielding:
+\begin{equation}
+  h'(w) = h(w) * (1 - z_k(h(w))) + a_k(h(w)) * z_k(h(w)). \label{eq:gated-residual}
+\end{equation}
+
+In Word Level Adaptive Domain Adaptation, $z_k(h(w))$ is designed to reflect the ``topicality'' of  word $w$ is in domain $k$: the more likely $w$ is in domain $k$, the larger $z_k(h(w))$ is.
+The word-level adaptive domain adaptation aims to reduce the variation caused by $a_k$ to the adapted context vector $h'$ (compared to $h$) for words that are not typical of domain $k$.
+For a domain, there are 2 type of non-topical words: out-of-domain words that rarely or never appear in the in-domain texts; or frequent words that appear equally frequently in many domains.
+\fyTodo{This is like a poor Tf-idf - can be because tf is small or because idf is small} By reusing a learned generic representation for the non-topical words, we can at least bound the risk of poor predictions in case of out-of-domain words by the risk of the learned generic model (see below).
+
 \section*{Training process}
-There 3 steps:
+The training process comprises thre main steps:
 \begin{itemize}
-	\item Pretraining a generic model with mixed corpora.
-	\item Training a domain classifier on top of the encoder and decoder; freezing the parameters of the generic model.
+\item Pretraining a generic model with a mixed corpora.
+\item Training a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen.
+This model computes the posterior domain probability $p(k|h(w))$ for each word $w$ based on the representation computed by the last layer.
+\fyTodo{Which layer, which classifier ?}
 \item Training parameters of adapters with in-domain data separately while freezing the parameters of the generic model and of the domain classifiers.
 \end{itemize}
+During the inference, $z_k$ is used to regulate the strength of the adapter module as suggested in equation~\ref{eq:gated-residual}.
 
 \section*{Weakness of the method}
-The new method depends essentially to the quality of the classifier.
+This method depends crucially to the quality of the classifier $p(k|h(w))$.
 Moreover, the classifier can fail to detect in-domain words in the case of polysemous words.
-For example, the word "joint" in joint stereo, joint sprain and joint project has 2 different translations as "joint" and "articulaire" in french.
-In the unbalanced dataset as the mixed corpora (6 in-domain corpora),  "joint stereo"
+For example, the word "joint" in ``joint stereo'', ``joint sprain'' and ``joint project'' correspond to multiple translations such as "joint" and "articulaire" in French.
+In the unbalanced dataset as the mixed corpora (6 in-domain corpora),  "joint stereo" \fyTodo{unfinished sentence}
 
@@
 
-The combination of generic representation and the residual adapter in the adapted model is as following
+The combination of a generic representation and a residual adapter is defined as follows:
 \begin{equation}
-h' = h + a_k(h)
-\label{eq:1}
+  h' = h + a_k(h) \label{eq:residual-adapter}
 \end{equation}
 
-where h is output of sharing layer, i.e $h=M(x,y)$, $M: \mathbb{X} \times \mathbb{Y} \rightarrow \mathbf{\Omega}\subset \mathbb{R}^{d_{model}}$, $a_k$ is the residual adapter which is a function $\mathbb{R}^{d_{model}} \rightarrow \mathbb{R}^{d_{model}}$.
-\\
+where $h$ is the output of shared ``generic'' layers, i.e $h=M(x,y)$, $M: \mathbb{X} \times \mathbb{Y} \rightarrow \mathbf{\Omega}\subset \mathbb{R}^{d_{model}}$, $a_k$ is the residual adapter defined as a function $a_k: \mathbb{R}^{d_{model}} \rightarrow \mathbb{R}^{d_{model}}$.
 
-The bad performance of fine-tuned model happens as the value of $a_k(h)$ of the out-of-domain example $(x^{adv}, y^{adv})$ is adversarial to the model.
-Indeed, by fine-tuning such adapter with only in-domain data, $a_k$ never encounters adversarial examples, thus does not guarantee the risk of prediction on out-of-domain examples.
-However, if we control the weight of $a_k(h)$ is the combination, we might reduce the wrong belief of $a_k$ on adversarial examples.
-To execute the idea, we  investigate the version of convex combination.
+The poor performance of fine-tuned models with a residual layer happens as the value of $a_k(h)$ of an out-of-domain example $(x^{adv}, y^{adv})$ is adversarial to the model.
+Indeed, by fine-tuning the adapter with only in-domain data, $a_k$ never encounters adversarial examples, which makes it error-prone in the face of out-of-domain examples.
+The intuition of wada is that we control the weight of $a_k(h)$ in the combination, we can hope to reduce the wrong belief of $a_k$ on adversarial examples.
+To implement this idea, we  investigate a convex combination of the generic and the residual layers, and define the adapted representation to be:
 
 \begin{equation}
-h' = h * (1-f(z_k)) + a_k(h) * f(z_k)
+h' = h * (1-f(z_k)) + a_k(h) * f(z_k), 
 \label{eq:2}
@@
 
-where 
+\noindent{}where
 \begin{itemize}
-	\item $z_k = P(d=k | h=M(x,y)) = p(h)$ where $p: \mathbf{\Omega} \rightarrow \mathbf{\Delta}^{K}$
+\item $z_k = p(d=k | h=M(x,y)) = p_k(h)$ with $p: \mathbf{\Omega} \rightarrow \mathbf{\Delta}^{K}$\fyTodo{Attention à $x$ et $y$}
 \item f is an activation function $f: [0,1] \rightarrow [0,1]$ 
 \end{itemize}
+
 We define some important constants:
 \begin{itemize}
-	\item Domain inseparability $\mathbb{\varepsilon}_{insep} = \displaystyle{\mathop{\sum}_{k' \neq k} \mathbb{E}_{x,y}[\mathbf{1}_{d=k'}*p_k(h)]}$. where $h=M(x,y)$
-	\item Risk of generic model $\mathbb{R} = \mathbb{E}_{x,y}[l(h,y)]$
-	\item $\mathbf{\Omega_{k,p_{0}}} = \lbrace h | z_k = p_k(h)<0.8\rbrace$.
-	\item Risk of adapted model in in-domain test set $R_k = \displaystyle{\mathbb{E}_{x,y \sim p(.|d=k)}[l(h * (1-f(z_k)) + a_k(h) * f(z_k), y)]}$
+\item Domain inseparability:  \fyTodo{Esperance de classification error - is it }
+  $$
+  \mathbb{\varepsilon}_{insep}(\theta, p) = \sum_{k' \neq k} \mathbb{E}_{x,y \sim \mathcal{D}_{k'}}(p(k|h),
+  $$
+\item Risk of the generic model: $\mathbb{R} = \mathbb{E}_{(x,y) \sim \mathcal{D}}(l(M(x,y_{<t}), y_t))$ \fyDone{wrt to which distribution ?}
+\item $\mathbf{\Omega_{k,p_{0}}} = \lbrace h | z_k = p_\theta(k|h) < p_0\rbrace$.
+\item Risk of the adapted model for the in-domain test set:
+  $$R_k = \mathbb{E}_{x,y \sim \mathcal{D}_k} [l(h * (1 - f(z_k)) + a_k(h) * f(z_k), y)]$$
 \end{itemize}
-We also have 2 assumptions on the regularization of activation function:
+
+We also have two assumptions on the regularity of the activation and the loss functions:
 \begin{itemize}
 \item $a_k$ is bounded in sense of $\normL{a_k(h)} < M$
-	\item loss function is bounded in sense of $l(h * (1-f(z_k)) + a_k(h) * f(z_k),y) < A$ $\forall x,y$
+\item the loss function is bounded in sense of $l(h * (1-f(z_k)) + a_k(h) * f(z_k), y) < A$ \fyTodo{why not $l(h,y) < A$ ?}
+  $\forall x,y$
 \end{itemize}
-We then access the risk of of the model adapted to domain k with the residual $a_k$ to arbitrary examples of K domains.
+We then evaluate the risk of the model adapted to domain $k$ with residuals $a_k$ for arbitrary examples from $K$ domains:
 \begin{equation}
@@
   &= \mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] + \\
-		&+ \mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h)]
+  &+ \mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]
 \end{split}
@@
 
-By designing a suitable activation function, we can map $f: [0,p_{0}] \rightarrow [0,\delta]$. Then with $\delta$ is small enough we can have below approximation of the first term:
+By designing a suitable activation function, we can map $f: [0,p_{0}] \rightarrow [0,\delta]$. Then with $\delta$ is small enough we can have an approximation of the first term:
 \begin{equation}
-\mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] \sim \mathbb{E}_{x,y}[l(h,y)\mathbf{1}_\mathbf{\Omega_{k,p_{0}}}(h)] \leq \mathbb{E}_{x,y}[l(h,y)] = R
+\mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] \approx{} \mathbb{E}_{x,y}[l(h,y)\mathbf{1}_\mathbf{\Omega_{k,p_{0}}}(h)] \leq \mathbb{E}_{x,y}[l(h,y)] = R
 \label{eq:4}
 \end{equation}
-The second term:
-\begin{equation}
+The second term: $ L = \mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]$ can be bounded as follows:
+\begin{equation*}
 \begin{split}
-\mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h)] &= \displaystyle{\mathop{\sum}_{k'}\mathbb{E}_{x,y \sim p(.|d=k')}[l(h * (1-f(z_k))} \\
-	& + a_k(h) * f(z_k),y) *\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h)]P(d=k') \\
-	& \leq \displaystyle{\mathop{\sum}_{k' \neq k}} A * \mathbb{E}_{x,y \sim p(.|d=k')} [\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h)]P(d=k') \\
-	&+ \mathbb{E}_{x,y \sim p(.|d=k)}[l(h * (1-f(z_k)) + \\
-	&+ a_k(h) * f(z_k),y) * \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h)]P(d=k) \\
+L &= \displaystyle{\mathop{\sum}_{k'}\mathbb{E}_{x,y \sim p(.|d=k')}[l(h * (1-f(z_k))} + a_k(h) * f(z_k),y) * \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]p(d=k') \\
+	& \leq \displaystyle{\mathop{\sum}_{k' \neq k}} A * \mathbb{E}_{x,y \sim p(.|d=k')} [\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]P(d=k') \\
+	& \quad + \mathbb{E}_{x,y \sim p(.|d=k)}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y) * \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]P(d=k) \\
 	&\leq A * \frac{\mathbb{\varepsilon}_{insep}}{p_{0}} + R_k * P(d=k)
@@
 \label{eq:5}
-\end{equation}
+\end{equation*}
 
-The last inequality is verified by the below inequality:
+The last inequality is verified since:
 \begin{equation}
 \begin{split}
-\displaystyle{\mathop{\sum}_{k' \neq k}} A * \mathbb{E}_{x,y \sim p(.|d=k')} [\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h)]P(d=k') &\leq A * \displaystyle{\mathop{\sum}_{k' \neq k}} \mathbb{E}_{x,y \sim p(.|d=k') } [\frac{p_k(h)}{p_0}]P(d=k') \\
+\displaystyle{\mathop{\sum}_{k' \neq k}} A * \mathbb{E}_{x,y \sim p(.|d=k')} [\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}} (h)]P(d=k') &\leq A * \displaystyle{\mathop{\sum}_{k' \neq k}} \mathbb{E}_{x,y \sim p(.|d=k') } [\frac{p_k(h)}{p_0}]P(d=k') \\
 		& = \frac{A}{p_0} \mathbb{E}[\mathbf{1}_{d\neq k}p_k(h)] \\
@@
 because
-$$ \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}^{c}}(h) \leq \frac{p_k(h)}{p_0} $$
+$$ \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h) \leq \frac{p_k(h)}{p_0} $$
 Combining the inequality \ref{eq:4} and the inequality \ref{eq:5} we obtain the risk in general case of the adapted model $(M,a_k)$ $$R(M,a_k) \leq R + \frac{A * \mathbb{\varepsilon}_{insep}}{p_0} + P(d=k)R_k$$
 
+\fyTodo{I sort of buy the equations but then so what ? We should may be compare with non-adaptive ?}
+
 \end{document}

### 2020-06-26 2020/wada/adaptive_domain_adaptation.tex
corrections

--- 2020-06-19 2020/wada/adaptive_domain_adaptation.tex
@@
 \section*{Word Level Adaptive Domain Adaptation}
-Let $h$ is the output of the encoder which is pretrained, the adapted representation will be $h' = h + a_k(h)$.
-In gating version, $h' = h * (1-z_k) + a_k(h) * z_k$ where $z_k$ is also a function of $h$ taking value in $[0,1]$.
-More precisely, $h$ and $h'$ is a sequence of context vectors , and the combination is element wise, i.e, $h'(w) = h(w) * z_k(h(w)) + a_k(h(w)) * z_k(h(w))$.
+Let $h$ is the output of the encoder which is pretrained, the representation adapted to the domain k using the residual adapter will be $$h' = h + a_k(h)$$ In gating version, $$h' = h * (1-z_k) + a_k(h) * z_k$$ where $z_k$ is also a function of $h$ taking value in $[0,1]$.
+More precisely, in MT task, $h$ and $h'$ is a sequence of context vectors , and the combination is element wise, i.e, $h'(w) = h(w) * z_k(h(w)) + a_k(h(w)) * z_k(h(w))$ where $w$ is a token in the sequence.
 \\
-In Word Level Adaptive Domain Adaptation version, $z_k(h(w))$ is designed so that the less probably the word $w$ is in domain $k$ the less $z_k(h(w))$ is.
-The Word Level Adaptive Domain Adaptation aims to reduce the variation caused by $a_k$ of adapted context vector $h'$ compared to $h$ on words that do not surely belong to domain k.
-For a domain, there are 2 type of "mistrusted" words: out-of-domain words that rarely or never appear in the in-domain data; or popular words that appear frequently in many domains.
-By reusing learned representation of the "mistrusted" words, we can at least guarantee the risk of bad prediction in case of out-of-domain words by the risk of the learned generic model.
+
+In Word Level Adaptive Domain Adaptation version, $z_k(h(w))$ is designed so that the less probably the word $w$ is in the domain $k$ the smaller $z_k(h(w))$ is.
+The Word Level Adaptive Domain Adaptation aims to reduce the variation caused by $a_k$ of the adapted context vector $h'$ compared to $h$ of a word that does not surely belong to domain k.
+\\
+
+For a domain, there are 2 type of "less specific" words: out-of-domain words that rarely or never appear in the domain but is popular in the other domains; or popular words that appear frequently in many domains.
+By reusing learned representation of the "less specific" words, we can at least guarantee the risk of the bad prediction of out-of-domain words by the risk of the learned generic model.
 \section*{Training process}
-There 3 steps:
+There are 3 steps:
 \begin{itemize}
 	\item Pretraining a generic model with mixed corpora.
-	\item Training a domain classifier on top of the encoder and decoder; freezing the parameters of the generic model.
-	\item Training parameters of adapters with in-domain data separately while freezing the parameters of the generic model and of the domain classifiers.
+	\item Training 2 domain classifiers on top of the encoder and on top of the decoder while freezing the parameters of the generic model.
+	\item Training parameters of the adapters with in-domain data separately while freezing the parameters of the generic model and of the domain classifiers.
 \end{itemize}
@@
 For example, the word "joint" in joint stereo, joint sprain and joint project has 2 different translations as "joint" and "articulaire" in french.
-In the unbalanced dataset as the mixed corpora (6 in-domain corpora),  "joint stereo"
+In the unbalanced dataset as the mixed corpora (6 in-domain corpora), the classifier somehow fails to predict IT as the most probable domain of "joint stereo".
 
@@
 	\item $z_k = P(d=k | h=M(x,y)) = p(h)$ where $p: \mathbf{\Omega} \rightarrow \mathbf{\Delta}^{K}$
-	\item f is an activation function $f: [0,1] \rightarrow [0,1]$ 
+	\item f is an activation function $f: [0,1] \rightarrow [0,1]$ which is monotonically increasing.
 \end{itemize}
@@
 	\item Risk of generic model $\mathbb{R} = \mathbb{E}_{x,y}[l(h,y)]$
-	\item $\mathbf{\Omega_{k,p_{0}}} = \lbrace h | z_k = p_k(h)<0.8\rbrace$.
+	\item $\mathbf{\Omega_{k,p_{0}}} = \lbrace h | z_k = p_k(h)<p_0\rbrace$.
 	\item Risk of adapted model in in-domain test set $R_k = \displaystyle{\mathbb{E}_{x,y \sim p(.|d=k)}[l(h * (1-f(z_k)) + a_k(h) * f(z_k), y)]}$
@@
 \end{itemize}
-We then access the risk of of the model adapted to domain k with the residual $a_k$ to arbitrary examples of K domains.
+We then access the risk of of the model adapted to domain k with the residual $a_k$ in the prediction of arbitrary examples of K domains.
 \begin{equation}

### 2020-06-26 2020/wada/adaptive_domain_adaptation.tex
corrections

--- 2020-06-25 2020/wada/adaptive_domain_adaptation.tex
@@
 Let $h$ the output of the pretrained encoder, the adapted representation for domain $k$ will be $h' = h + a_k(h)$, meaning that all words in all sentences for domain $k$ will use the adapter module $a_k$.
-A ``gated'' version uses $h' = h * (1-z_k) + a_k(h) * z_k$\footnote{More precisely, $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure XXX...}\fyTodo{$a_k(h)=$ somme tous les composants résiduels} where $z_k$ is also a function of $h$ taking values in $[0,1]$.
+A ``gated'' version uses $h' = h * (1-z_k) + a_k(h) * z_k$\footnote{More precisely, $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure \ref{fig:1}} \fyTodo{$a_k(h)=$ somme tous les composants résiduels} where $z_k$ is also a function of $h$ taking values in $[0,1]$.
 More precisely, $h$ and $h'$ are sequences of context vectors and the combination is performed element-wise, yielding:
@@
 \fyTodo{This is like a poor Tf-idf - can be because tf is small or because idf is small} By reusing a learned generic representation for the non-topical words, we can at least bound the risk of poor predictions in case of out-of-domain words by the risk of the learned generic model (see below).
+\begin{figure}[h!]
+\includegraphics[scale=0.5]{highway_residual}
+\caption{Highway residual layer}
+\label{fig:1}
+\end{figure}
 
@@
 \section*{Weakness of the method}
+
 This method depends crucially to the quality of the classifier $p(k|h(w))$.

### 2020-06-30 2020/wada/adaptive_domain_adaptation.tex
answer fyTodo questions

@@
 \NewDocumentCommand{\normL}{ s O{} m }{%
-  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{L_1(\Omega)}%
+  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{L_1}%
 }
@@
 Let $h$ the output of the pretrained encoder, the adapted representation for domain $k$ will be $h' = h + a_k(h)$, meaning that all words in all sentences for domain $k$ will use the adapter module $a_k$.
-A ``gated'' version uses $h' = h * (1-z_k) + a_k(h) * z_k$\footnote{More precisely, $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure \ref{fig:1}} \fyTodo{$a_k(h)=$ somme tous les composants résiduels} where $z_k$ is also a function of $h$ taking values in $[0,1]$.
+A ``gated'' version uses $h' = h + a_k(h) * z_k$\footnote{More precisely, $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure \ref{fig:1}} \fyTodo{$a_k(h)=$ somme tous les composants résiduels} where $z_k$ is also a function of $h$ taking values in $[0,1]$.
 More precisely, $h$ and $h'$ are sequences of context vectors and the combination is performed element-wise, yielding:
 \begin{equation}
-  h'(w) = h(w) * (1 - z_k(h(w))) + a_k(h(w)) * z_k(h(w)). \label{eq:gated-residual}
+  h'(w) = h(w) + a_k(h(w)) * z_k(h(w)). \label{eq:gated-residual}
 \end{equation}
@@
 For a domain, there are 2 type of non-topical words: out-of-domain words that rarely or never appear in the in-domain texts; or frequent words that appear equally frequently in many domains.
-\fyTodo{This is like a poor Tf-idf - can be because tf is small or because idf is small} By reusing a learned generic representation for the non-topical words, we can at least bound the risk of poor predictions in case of out-of-domain words by the risk of the learned generic model (see below).
+\fyTodo{This is like a poor Tf-idf - can be because tf is small or because idf is small}
+
+\fyTodo{Experiment: compute posterior probability of words that have poor $tf-idf$ to verify the gating operation}.
+
+By reusing the learned generic representation for the non-topical words, we can at least bound the risk of poor predictions in case of out-of-domain words by the risk of the learned generic model (see below).
 \begin{figure}[h!]
@@
 This model computes the posterior domain probability $p(k|h(w))$ for each word $w$ based on the representation computed by the last layer.
+We applied domain classifiers to the output of the encoder and to the output of the decoder, i.e, at the last self-attention layer of the encoder and the attention layer of the decoder.There are 2 classifiers which are placed in the top of the encoder and in the top of the decoder.
+$$z^{enc}_k(w^{src}_t) = p(d=k| h_{enc}(w^{src}_t))$$
+where $h_{enc}(w^{src}_t)=Enc(x)(w^{src}_t)$ and 
+$$z^{dec}_k(w^{tgt}_t) = p(d=k| h_{dec}(w^{tgt}_t))$$  
+where $h_{dec}(w^{tgt}_t)=Dec(x,y_{i<t})(w^{tgt}_t)$
+
 \fyTodo{Which layer, which classifier ?}
@@
 For example, the word "joint" in ``joint stereo'', ``joint sprain'' and ``joint project'' correspond to multiple translations such as "joint" and "articulaire" in French.
-In the unbalanced dataset as the mixed corpora (6 in-domain corpora),  "joint stereo" \fyTodo{unfinished sentence}
+In the unbalanced dataset as the mixed corpora (6 in-domain corpora),  "joint stereo" is hardly to be detected as in IT domain, therefore "joint stereo" does not profit the finetuning's effect.
 
@@
 \begin{equation}
-h' = h * (1-f(z_k)) + a_k(h) * f(z_k), 
+h' = h + a_k(h) * f(z_k), 
 \label{eq:2}
@@
 \begin{itemize}
-\item $z_k = p(d=k | h=M(x,y)) = p_k(h)$ with $p: \mathbf{\Omega} \rightarrow \mathbf{\Delta}^{K}$\fyTodo{Attention à $x$ et $y$}
+\item $z_k = p(d=k | h=M(x,y)) = p_k(h)$ with $p: \mathbf{\Omega} \rightarrow \mathbf{\Delta}^{K}$.
+In the sequence to sequence version, there are 2 classifiers which are placed in the top of the encoder and in the top of the decoder.
+$z^{enc}_k(w_t) = p(d=k| h_{enc}(w_t)=Enc(x)(w_t))$ and $z^{dec}_k(w_t) = p(d=k| h_{dec}(w_t)=Dec(y_t,x,y_{i<t})(w_t))$
+\fyTodo{Attention à $x$ et $y$}
 \item f is an activation function $f: [0,1] \rightarrow [0,1]$ 
@@
 
+We assume that the hidden distribution of a multi-domain problem limited to an unknown mixture of $K$ in-domain distribution, i.e, $$P(x,y) = \sum_{k\in[1..K]}P(x,y|d=k)P(d=k)$$.
+We denote also $D_k$ the distribution of domain k $P(x,y|d=k)$
+
 We define some important constants:
 \begin{itemize}
-\item Domain inseparability:  \fyTodo{Esperance de classification error - is it }
-  $$
-  \mathbb{\varepsilon}_{insep}(\theta, p) = \sum_{k' \neq k} \mathbb{E}_{x,y \sim \mathcal{D}_{k'}}(p(k|h),
-  $$
-\item Risk of the generic model: $\mathbb{R} = \mathbb{E}_{(x,y) \sim \mathcal{D}}(l(M(x,y_{<t}), y_t))$ \fyDone{wrt to which distribution ?}
+\item Inseparability of domains over a given fixed representation space $\Omega$
+\begin{equation}
+  \begin{split}
+  \mathbb{\varepsilon}_{insep} = \min_{p} \sum_{k\in[1..K]} \mathbb{E}_{x,y \sim \sum_{i\in [1..K]} \mathcal{D}_{i}*p(d=i)}(\mathbf{1}_{d\neq k} p(k|h))
+  \end{split}
+  \end{equation}
+\item Inseparability of domain k from other domains over a given fixed representation space $\Omega$
+  \begin{equation}
+  \begin{split}
+  \mathbb{\varepsilon}_{insep}^k = \min_{p} \mathbb{E}_{x,y \sim \sum_{i\in [1..K]} \mathcal{D}_{i}*p(d=i)}(\mathbf{1}_{d\neq k} p(k|h))
+  \end{split}
+  \end{equation}
+\item Risk of the generic model: $\mathbb{R} = \mathbb{E}_{(x,y \sim \sum_{k}p(x,y|d=k)p(d=k)) \sim \mathcal{D}}(l(M(x,y_{<t}), y_t))$ \fyTodo{wrt to which distribution ?}
 \item $\mathbf{\Omega_{k,p_{0}}} = \lbrace h | z_k = p_\theta(k|h) < p_0\rbrace$.
 \item Risk of the adapted model for the in-domain test set:
-  $$R_k = \mathbb{E}_{x,y \sim \mathcal{D}_k} [l(h * (1 - f(z_k)) + a_k(h) * f(z_k), y)]$$
+  $$R_k = \mathbb{E}_{x,y \sim \mathcal{D}_k} [l(h + a_k(h) * f(z_k), y)]$$
 \end{itemize}
@@
 \item $a_k$ is bounded in sense of $\normL{a_k(h)} < M$
-\item the loss function is bounded in sense of $l(h * (1-f(z_k)) + a_k(h) * f(z_k), y) < A$ \fyTodo{why not $l(h,y) < A$ ?}
+\item the loss function is bounded in sense of $l(h,y) < A$ $\forall h \in \Omega + \mathbb{B}(M)$ where $\mathbb{B}(M)=\lbrace x\in \mathbb{R}^{d_{model}} | \normL{x}<M \rbrace$ and $A+B = \lbrace x+ \alpha y  | \alpha \in [0..1], x \in A, y \in B \rbrace$  \fyTodo{why not $l(h,y) < A$ ?}
   $\forall x,y$
 \end{itemize}
+
+Finally, we assume that the classifier is good enough, i.e, $$err_k(p) = \mathbb{E}_{x,y \sim \sum_{i\in [1..K]} \mathcal{D}_{i}*p(d=i)}(\mathbf{1}_{d\neq k} p(k|h)) < \varepsilon_{insep}^k + \gamma$$ $\forall k$ with $\gamma$ sufficiently small.
 We then evaluate the risk of the model adapted to domain $k$ with residuals $a_k$ for arbitrary examples from $K$ domains:
@@
 \begin{split}
-  R(M,a_k) &= \mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y)] \\
-  &= \mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] + \\
-  &+ \mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]
+  R(M,a_k) &= \mathbb{E}_{x,y}[l(h + a_k(h) * f(z_k),y)] \\
+  &= \mathbb{E}_{x,y}[l(h + a_k(h) * f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] + \\
+  &+ \mathbb{E}_{x,y}[l(h + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]
 \end{split}
@@
 \begin{equation}
-\mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] \approx{} \mathbb{E}_{x,y}[l(h,y)\mathbf{1}_\mathbf{\Omega_{k,p_{0}}}(h)] \leq \mathbb{E}_{x,y}[l(h,y)] = R
+\mathbb{E}_{x,y}[l(h + a_k(h) * f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] \approx{} \mathbb{E}_{x,y}[l(h,y)\mathbf{1}_\mathbf{\Omega_{k,p_{0}}}(h)] \leq \mathbb{E}_{x,y}[l(h,y)] = R
 \label{eq:4}
 \end{equation}
-The second term: $ L = \mathbb{E}_{x,y}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]$ can be bounded as follows:
+The second term: $ L = \mathbb{E}_{x,y}[l(h + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]$ can be bounded as follows:
 \begin{equation*}
 \begin{split}
-L &= \displaystyle{\mathop{\sum}_{k'}\mathbb{E}_{x,y \sim p(.|d=k')}[l(h * (1-f(z_k))} + a_k(h) * f(z_k),y) * \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]p(d=k') \\
+L &= \displaystyle{\mathop{\sum}_{k'}\mathbb{E}_{x,y \sim p(.|d=k')}[l(h } + a_k(h) * f(z_k),y) * \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]p(d=k') \\
 	& \leq \displaystyle{\mathop{\sum}_{k' \neq k}} A * \mathbb{E}_{x,y \sim p(.|d=k')} [\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]P(d=k') \\
-	& \quad + \mathbb{E}_{x,y \sim p(.|d=k)}[l(h * (1-f(z_k)) + a_k(h) * f(z_k),y) * \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]P(d=k) \\
-	&\leq A * \frac{\mathbb{\varepsilon}_{insep}}{p_{0}} + R_k * P(d=k)
+	& \quad + \mathbb{E}_{x,y \sim p(.|d=k)}[l(h + a_k(h) * f(z_k),y) * \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]P(d=k) \\
+	&\leq A * \frac{\mathbb{\varepsilon}_{insep}^k + \gamma}{p_{0}} + R_k * P(d=k)
 \end{split}
@@
 		& = \frac{A}{p_0} \mathbb{E}[\mathbf{1}_{d\neq k}p_k(h)] \\
-		& \leq \frac{A}{p_0} \mathbb{\varepsilon}_{insep}
+		& \leq \frac{A}{p_0} (\mathbb{\varepsilon}_{insep}^k + \gamma)
 \end{split}
@@
 $$ \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h) \leq \frac{p_k(h)}{p_0} $$
-Combining the inequality \ref{eq:4} and the inequality \ref{eq:5} we obtain the risk in general case of the adapted model $(M,a_k)$ $$R(M,a_k) \leq R + \frac{A * \mathbb{\varepsilon}_{insep}}{p_0} + P(d=k)R_k$$
+Combining the inequality \ref{eq:4} and the inequality \ref{eq:5} we obtain the risk in general case of the adapted model $(M,a_k)$ $$R(M,a_k) \leq R + A * \frac{\mathbb{\varepsilon}_{insep}^k + \gamma}{p_0} + P(d=k)R_k$$
 
+If we the original version of the residual adapter, with the same assumption in the regularization of the adapter, the bound of the risk of translating out-of-domain is only A (the bound of loss function).
+By using adaptive version, we reduce A by factor $\frac{\epsilon^k_{insep}+\gamma}{p_0}$ which is small if we choose $p_0$ is high enough, the domains are distant and the classifier is good enough.
+Furthermore $ \varepsilon^k_{insep} < \varepsilon_{insep} $ which is the minimum of the error expectation over the hypothesis space of the classifier.
+With a trivial prediction $[\frac{1}{K},..,\frac{1}{K}]$ we have error expectation of 0.5.
+Therefore, $\varepsilon_{insep} \leq 0.5$, i.e, $\varepsilon^k_{insep} < 0.5$.
 \fyTodo{I sort of buy the equations but then so what ? We should may be compare with non-adaptive ?}

### 2020-06-30 2020/wada/adaptive_domain_adaptation.tex
correct some errors

@@
 \item $a_k$ is bounded in sense of $\normL{a_k(h)} < M$
-\item the loss function is bounded in sense of $l(h,y) < A$ $\forall h \in \Omega + \mathbb{B}(M)$ where $\mathbb{B}(M)=\lbrace x\in \mathbb{R}^{d_{model}} | \normL{x}<M \rbrace$ and $A+B = \lbrace x+ \alpha y  | \alpha \in [0..1], x \in A, y \in B \rbrace$  \fyTodo{why not $l(h,y) < A$ ?}
+\item the loss function is bounded in sense of $l(h,y) < A$ $\forall h \in \Omega + \mathbb{B}(M)$ where $\mathbb{B}(M)=\lbrace x\in \mathbb{R}^{d_{model}} | \normL{x}<M \rbrace$ and $A+_{\alpha}B = \lbrace x+ \alpha y  | \alpha \in [0..1], x \in A, y \in B \rbrace$  \fyTodo{why not $l(h,y) < A$ ?}
   $\forall x,y$
@@
 If we the original version of the residual adapter, with the same assumption in the regularization of the adapter, the bound of the risk of translating out-of-domain is only A (the bound of loss function).
-By using adaptive version, we reduce A by factor $\frac{\epsilon^k_{insep}+\gamma}{p_0}$ which is small if we choose $p_0$ is high enough, the domains are distant and the classifier is good enough.
+By using adaptive version, we reduce A by factor $\frac{p_0}{\epsilon^k_{insep}+\gamma}$ which is high if we choose $p_0$ is high enough, the domains are distant and the classifier is good enough.
+
 Furthermore $ \varepsilon^k_{insep} < \varepsilon_{insep} $ which is the minimum of the error expectation over the hypothesis space of the classifier.
-With a trivial prediction $[\frac{1}{K},..,\frac{1}{K}]$ we have error expectation of 0.5.
+With a trivial prediction $[\frac{1}{K},..,\frac{1}{K}]$ we have an error expectation equal to 0.5.
 Therefore, $\varepsilon_{insep} \leq 0.5$, i.e, $\varepsilon^k_{insep} < 0.5$.
+In the worst case where domains in the mixture are very close, we are still able to reduce the risk of predicting out-of-domain examples by a factor of 2.
 \fyTodo{I sort of buy the equations but then so what ? We should may be compare with non-adaptive ?}

### 2020-06-30 2020/wada/adaptive_domain_adaptation.tex
correct some errors

@@
 For example, the word "joint" in ``joint stereo'', ``joint sprain'' and ``joint project'' correspond to multiple translations such as "joint" and "articulaire" in French.
-In the unbalanced dataset as the mixed corpora (6 in-domain corpora),  "joint stereo" is hardly to be detected as in IT domain, therefore "joint stereo" does not profit the finetuning's effect.
+In the unbalanced dataset as the mixed corpora (6 in-domain corpora),  "joint stereo" is hardly to be detected as in IT domain, therefore "joint stereo" does not profit the finetuning's effect even being translated with the adapters of IT domain.
 

### 2020-06-30 2020/wada/adaptive_domain_adaptation.tex
correct some errors

@@
 \begin{itemize}
-\item Inseparability of domains over a given fixed representation space $\Omega$
+\item Inseparability of domains over a fixed representation space $\Omega$
 \begin{equation}
@@
   \end{equation}
-\item Inseparability of domain k from other domains over a given fixed representation space $\Omega$
+\item Inseparability of domain k from other domains over a fixed representation space $\Omega$
   \begin{equation}

### 2020-06-30 2020/wada/adaptive_domain_adaptation.tex
correct some errors

@@
 
-Furthermore $ \varepsilon^k_{insep} < \varepsilon_{insep} $ which is the minimum of the error expectation over the hypothesis space of the classifier.
-With a trivial prediction $[\frac{1}{K},..,\frac{1}{K}]$ we have an error expectation equal to 0.5.
-Therefore, $\varepsilon_{insep} \leq 0.5$, i.e, $\varepsilon^k_{insep} < 0.5$.
-In the worst case where domains in the mixture are very close, we are still able to reduce the risk of predicting out-of-domain examples by a factor of 2.
 \fyTodo{I sort of buy the equations but then so what ? We should may be compare with non-adaptive ?}

### 2020-06-30 2020/wada/adaptive_domain_adaptation.tex
correct some errors

--- 2020-06-30 2020/wada/adaptive_domain_adaptation.tex
@@
 If we the original version of the residual adapter, with the same assumption in the regularization of the adapter, the bound of the risk of translating out-of-domain is only A (the bound of loss function).
-By using adaptive version, we reduce A by factor $\frac{p_0}{\epsilon^k_{insep}+\gamma}$ which is high if we choose $p_0$ is high enough, the domains are distant and the classifier is good enough.
+By using adaptive version, we reduce A by factor $\frac{p_0}{\epsilon^k_{insep}+\gamma}$ which is high if we choose $p_0$ is high enough, the domains are distant in the fixed representation space and the classifier is good enough.
 

### 2020-07-01 2020/wada/adaptive_domain_adaptation.tex
correct some errors
mise en forme

--- 2020-06-30 2020/wada/adaptive_domain_adaptation.tex
@@
-\documentclass[12pt,a4paper,twoside]{report}
+\documentclass[12pt,times,a4paper,twoside]{article}
+
+\usepackage[utf8]{inputenc}	% Para caracteres en español
 \usepackage[left=3cm,right=3cm,top=2cm,bottom=3cm]{geometry}
 \usepackage{pagenote}
+
 \usepackage{algorithm}
 \usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage[utf8]{inputenc}	% Para caracteres en español
+
 \usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
+\usepackage{mathtools,xparse}
+\usepackage{mathrsfs}
+
 \usepackage{multirow,booktabs}
@@
 \usepackage{fancyhdr}
-\usepackage{mathrsfs}
+
+\usepackage{graphicx}
 \usepackage{wrapfig}
@@
 \usepackage{calc}
-\usepackage{color,soul}
+% \usepackage{color,soul}
 \usepackage{multicol}
@@
 \usepackage[retainorgcmds]{IEEEtrantools}
-\usepackage{xcolor}
-\colorlet{shadecolor}{orange!15}
-\parindent 0in
-\parskip 12pt
-\geometry{margin=1in, headsep=0.25in}
-\theoremstyle{definition}
-\newtheorem{defn}{Definition}
-\newtheorem{reg}{Rule}
-\newtheorem{exer}{Exercise}
-\newtheorem{note}{Note}
+% \usepackage{xcolor} -- implied by xcolor
+
 \usepackage{listings}
@@
 \usepackage{natbib}
+
 \usepackage{tikz}
@@
 \usepackage{pgfplotstable}
+
 \usepackage{soulutf8}
-\usepackage{color,soul}
-\usepackage{xcolor}
 \usepackage{tabularx}
+
+\usepackage[draft]{todo}
+\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+
+\colorlet{shadecolor}{orange!15}
+\parindent 0in
+\parskip 12pt
+\geometry{margin=1in, headsep=0.25in}
+\setlength{\belowdisplayskip}{8pt} \setlength{\belowdisplayshortskip}{8pt}
+\setlength{\abovedisplayskip}{8pt} \setlength{\abovedisplayshortskip}{8pt}
+\setlist{nosep}
+\setlength{\parskip}{0.1cm}
+\setlength{\parindent}{1em}
+
+\theoremstyle{definition}
+\newtheorem{defn}{Definition}
 \newtheorem{lemma}{Lemma}[section]
@@
 \DeclareMathOperator*{\argmax}{argmax}
-\usepackage[draft]{todo}
-\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\R}{\ensuremath{\mathbb{R}}}
+
 \pgfkeys{
@@
 }
-\usepackage{mathtools,xparse}
+
 \DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
@@
 
+\title{Combination of generic and specific layers in multidomain learning}
+\author{}
+\date{}
+
 \begin{document}
+
+\maketitle
 \begin{center}
-{\LARGE \bf Combination of generic and specific}\\
+{\LARGE \bf}\fyDone{Make a proper title}
 \end{center}
-\setlength{\belowdisplayskip}{8pt} \setlength{\belowdisplayshortskip}{8pt}
-\setlength{\abovedisplayskip}{8pt} \setlength{\abovedisplayshortskip}{8pt}
-\setlist{nosep}
-\setlength{\parskip}{0.1cm}
-\setlength{\parindent}{1em}
+
 \section*{Word Level Adaptive Domain Adaptation}
-Let $h$ the output of the pretrained encoder, the adapted representation for domain $k$ will be $h' = h + a_k(h)$, meaning that all words in all sentences for domain $k$ will use the adapter module $a_k$.
-A ``gated'' version uses $h' = h + a_k(h) * z_k$\footnote{More precisely, $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure \ref{fig:1}} \fyTodo{$a_k(h)=$ somme tous les composants résiduels} where $z_k$ is also a function of $h$ taking values in $[0,1]$.
-More precisely, $h$ and $h'$ are sequences of context vectors and the combination is performed element-wise, yielding:
+We consider a generic encoder-decoder architecture for a multi-domain sequence to sequence learning.
+Let $h \in \R^d$ the output of the encoder, \cite{} proposes to use an adapted representation for domain $k$ defined as $h' = h + a_k(h)$.
+This means that all words in all sentences in domain $k$ will use the adapter module represented by $a_k$.
+In a multilayer version with highway connections (see Figure~\ref{fig:hrl-architecture}), $a_k(h) = \sum_{l} a_{kl}(h^l)$ where $h^l$ is the distributed representation of the sequence at the $l^{th}$ layer.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[scale=0.5]{fig/highway_residual}
+  \caption{Highway multi-domain network with residual layers}
+\label{fig:hrl-architecture}
+\end{figure}
+
+ A ``gated'' version uses $h' = h + a_k(h) \odot{} z_k(h)$\footnote{Or $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure~\ref{fig:architecture}} \fyDone{$a_k(h)=$ somme tous les composants résiduels} where $z_k$ is a function of $h$ taking values in $[0,1]$.
+More precisely, $h$ and $h'$ are sequences of context vectors and the combination is performed element-wise, yielding:\fyDone{pourquoi pas gating: $(1-a_k(h(w))*hh(w)$ ?
+Notation mult.
+component wise ?}
+
 \begin{equation}
-  h'(w) = h(w) + a_k(h(w)) * z_k(h(w)). \label{eq:gated-residual}
+   \forall t \in [1 \dots{} T], h'(w_t) = h(w_t) + a_k(h(w_t)) \odot{} z_k(h(w_t)). \label{eq:gated-residual}
 \end{equation}
 
-In Word Level Adaptive Domain Adaptation, $z_k(h(w))$ is designed to reflect the ``topicality'' of  word $w$ is in domain $k$: the more likely $w$ is in domain $k$, the larger $z_k(h(w))$ is.
-The word-level adaptive domain adaptation aims to reduce the variation caused by $a_k$ to the adapted context vector $h'$ (compared to $h$) for words that are not typical of domain $k$.
-For a domain, there are 2 type of non-topical words: out-of-domain words that rarely or never appear in the in-domain texts; or frequent words that appear equally frequently in many domains.
+In Word Level Adaptive Domain Adaptation (WADA), $z_k(h(w))$ is designed to reflect the ``topicality'' of  word $w$ is in domain $k$: the more likely $w$ is in domain $k$, the larger $z_k(h(w))$ should be.
+Word-level adaptive domain adaptation aims to reduce the variation caused by $a_k$ to the adapted context vector $h'$ (compared to $h$) for words that are not typical of domain $k$.
+For a domain, there are 2 types of non-topical words: out-of-domain words that rarely or never appear in the in-domain texts and frequent words that appear equally frequently in many domains.
 \fyTodo{This is like a poor Tf-idf - can be because tf is small or because idf is small}
@@
 
-By reusing the learned generic representation for the non-topical words, we can at least bound the risk of poor predictions in case of out-of-domain words by the risk of the learned generic model (see below).
-\begin{figure}[h!]
-\includegraphics[scale=0.5]{highway_residual}
-\caption{Highway residual layer}
-\label{fig:1}
-\end{figure}
+By reusing the learned generic representation for non-topical words, we can bound the risk of poor predictions in case of out-of-domain words by the risk of the learned generic model (see below).
 
-\section*{Training process}
-The training process comprises thre main steps:
-\begin{itemize}
-\item Pretraining a generic model with a mixed corpora.
+\section*{The training process}
+The training process comprises three main steps:
+\begin{enumerate}
+\item Pretraining a generic model with a mixed corpora comprising data from multiple domains;
 \item Training a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen.
-This model computes the posterior domain probability $p(k|h(w))$ for each word $w$ based on the representation computed by the last layer.
-We applied domain classifiers to the output of the encoder and to the output of the decoder, i.e, at the last self-attention layer of the encoder and the attention layer of the decoder.There are 2 classifiers which are placed in the top of the encoder and in the top of the decoder.
-$$z^{enc}_k(w^{src}_t) = p(d=k| h_{enc}(w^{src}_t))$$
-where $h_{enc}(w^{src}_t)=Enc(x)(w^{src}_t)$ and 
-$$z^{dec}_k(w^{tgt}_t) = p(d=k| h_{dec}(w^{tgt}_t))$$  
-where $h_{dec}(w^{tgt}_t)=Dec(x,y_{i<t})(w^{tgt}_t)$
+This model computes the posterior domain probability $p(k|h(w_t))$ for each word $w_t$ based on the representation computed by the last layer.
 
-\fyTodo{Which layer, which classifier ?}
-\item Training parameters of adapters with in-domain data separately while freezing the parameters of the generic model and of the domain classifiers.
-\end{itemize}
-During the inference, $z_k$ is used to regulate the strength of the adapter module as suggested in equation~\ref{eq:gated-residual}.
+  Domain classifiers take as input the output of the encoder and the output of the decoder, i.e at the last self-attention layer of the encoder and the attention layer of the decoder.
+\fyDone{Next sentence is a repetition}
+  % Two classifiers are placed respectively on top of the last encoder and decoder layers:
+  \begin{align}
+    \forall t \in [1 \dots T], h_{enc}(w^{src}_t) = Enc(x)(w^{src}_t), z^{enc}_k(w^{src}_t) = p(k| h^{enc}(w^{src}_t))
+  \end{align}
+  \fyTodo{what are x and y ?}
+  and
+  \begin{align}
+    h_{dec}(w^{tgt}_t)=Dec(x, y_{i<t})(w^{tgt}_t)  z^{dec}_k(w^{tgt}_t) = p(d=k| h_{dec}(w^{tgt}_t))
+  \end{align}
+
+\item Training the parameters of adapters with in-domain data separately for each domain while freezing the parameters of the generic model and of the domain classifiers.
+\end{enumerate}
+
+During inference, $z_k$ is used to regulate the strength of the adapter module as suggested in equation~\ref{eq:gated-residual}.
 
@@
 
-This method depends crucially to the quality of the classifier $p(k|h(w))$.
+This method\fyTodo{Do not start with the weaknesses} depends crucially to the quality of the classifier $p(k|h(w))$.
 Moreover, the classifier can fail to detect in-domain words in the case of polysemous words.
@@
 
-The combination of a generic representation and a residual adapter is defined as follows:
+Recall that the combination of a generic representation and a residual adapter introduced in \citep{BapnaXX} is defined as:
 \begin{equation}
-  h' = h + a_k(h) \label{eq:residual-adapter}
+  h' = h + a_k(h) \label{eq:residual-adapter},
 \end{equation}
 
-where $h$ is the output of shared ``generic'' layers, i.e $h=M(x,y)$, $M: \mathbb{X} \times \mathbb{Y} \rightarrow \mathbf{\Omega}\subset \mathbb{R}^{d_{model}}$, $a_k$ is the residual adapter defined as a function $a_k: \mathbb{R}^{d_{model}} \rightarrow \mathbb{R}^{d_{model}}$.
+where $h \in \R^{d_{model}}$ is the output of shared ``generic'' layers, i.e $h=M(x,y)$, $M: \mathbb{X} \times \mathbb{Y} \rightarrow \mathbf{\Omega}\subset \R^{d_{model}}$, $a_k$ is the residual adapter defined as a function $a_k: \R^{d_{model}} \rightarrow \R^{d_{model}}$.\fyTodo{$\R^d$ might be enough.}
 
-The poor performance of fine-tuned models with a residual layer happens as the value of $a_k(h)$ of an out-of-domain example $(x^{adv}, y^{adv})$ is adversarial to the model.
+The poor performance of fine-tuned models with a residual layer happens when the value of $a_k(h)$ of an out-of-domain example $(x^{adv}, y^{adv})$ is adversarial to the model.
 Indeed, by fine-tuning the adapter with only in-domain data, $a_k$ never encounters adversarial examples, which makes it error-prone in the face of out-of-domain examples.
-The intuition of wada is that we control the weight of $a_k(h)$ in the combination, we can hope to reduce the wrong belief of $a_k$ on adversarial examples.
-To implement this idea, we  investigate a convex combination of the generic and the residual layers, and define the adapted representation to be:
+The intuition of WADA is that by controlling the weight of $a_k(h)$ in the combination, we can hope to reduce the wrong impact of $a_k$ on adversarial examples.
+To implement this idea, we  investigate a convex combination\fyDone{Entre $h$ et $h + a h$} of the generic and the residual layers and define the adapted representation of a training sample from domain $k$ as: \fyTodo{Use real labels in equation}
 
 \begin{equation}
-h' = h + a_k(h) * f(z_k), 
+  h' = h + a_k(h) \odot f(z_k), 
 \label{eq:2}
@@
 \begin{itemize}
-\item $z_k = p(d=k | h=M(x,y)) = p_k(h)$ with $p: \mathbf{\Omega} \rightarrow \mathbf{\Delta}^{K}$.
-In the sequence to sequence version, there are 2 classifiers which are placed in the top of the encoder and in the top of the decoder.
+\item $z_k = p(d=k | h=M(x,y)) = p_k(h)$ with $p: \mathbf{\Omega} \rightarrow \mathbf{\Delta}^{K}$\fyTodo{This is an overkill}.
+In the sequence to sequence version, there are 2 classifiers which are placed respectively on top of the encoder and of the decoder.
 $z^{enc}_k(w_t) = p(d=k| h_{enc}(w_t)=Enc(x)(w_t))$ and $z^{dec}_k(w_t) = p(d=k| h_{dec}(w_t)=Dec(y_t,x,y_{i<t})(w_t))$
 \fyTodo{Attention à $x$ et $y$}
-\item f is an activation function $f: [0,1] \rightarrow [0,1]$ 
+\item f is an activation function $f: [0,1] \rightarrow [0,1]$.
 \end{itemize}
 
-We assume that the hidden distribution of a multi-domain problem limited to an unknown mixture of $K$ in-domain distribution, i.e, $$P(x,y) = \sum_{k\in[1..K]}P(x,y|d=k)P(d=k)$$.
-We denote also $D_k$ the distribution of domain k $P(x,y|d=k)$
+We assume that the hidden distribution of a multi-domain problem limited to an unknown mixture of $K$ in-domain distribution, i.e, $$\mathcal{D}(x,y) = \sum_{k=1}^{K} P(d=k) P(x,y|d=k)$$.
+We denote also $\mathcal{D}_k$ the data distribution for domain $k$: $\mathcal{D}_k = P(x,y|d=k)$.
 
@@
 \begin{itemize}
-\item Inseparability of domains over a fixed representation space $\Omega$
-\begin{equation}
-  \begin{split}
-  \mathbb{\varepsilon}_{insep} = \min_{p} \sum_{k\in[1..K]} \mathbb{E}_{x,y \sim \sum_{i\in [1..K]} \mathcal{D}_{i}*p(d=i)}(\mathbf{1}_{d\neq k} p(k|h))
-  \end{split}
-  \end{equation}
-\item Inseparability of domain k from other domains over a fixed representation space $\Omega$
-  \begin{equation}
-  \begin{split}
-  \mathbb{\varepsilon}_{insep}^k = \min_{p} \mathbb{E}_{x,y \sim \sum_{i\in [1..K]} \mathcal{D}_{i}*p(d=i)}(\mathbf{1}_{d\neq k} p(k|h))
-  \end{split}
-  \end{equation}
-\item Risk of the generic model: $\mathbb{R} = \mathbb{E}_{(x,y \sim \sum_{k}p(x,y|d=k)p(d=k)) \sim \mathcal{D}}(l(M(x,y_{<t}), y_t))$ \fyTodo{wrt to which distribution ?}
+\item Inseparability of domains over a fixed representation space $\Omega$\fyTodo{ugly equation, is my version better ?}
+\begin{align}
+  \varepsilon_{insep} =& \min_{p} \sum_{k\in[1..K]} \mathbb{E}_{x,y \sim \sum_{i\in [1..K]} \mathcal{D}_{i}*p(d=i)}(\mathbf{1}_{d\neq k} p(k|h)) \\
+  \varepsilon_{insep} =& \sum_{k=1}^K P(k)  \mathbb{E}_{x,y \sim \mathcal{D}_k} (\sum_{k' \neq k} P(k'|x,y))
+\end{align}
+
+\fyTodo{This does not depend on anything, this is how are data is but we can not know it} 
+\item Inseparability of domain $k$ from other domains over a fixed representation space $\Omega$
+  \begin{align}
+    \varepsilon_{insep}(k) &= \min_{p} \mathbb{E}_{x,y \sim \sum_{i\in [1..K]} \mathcal{D}_{i}*p(d=i)}(\mathbf{1}_{d\neq k} p(k|h)) \\
+    \varepsilon_{insep}(k) &= \mathbb{E}_{x,y \sim \mathcal{D}_k} (\sum_{k' \neq k} P(k'|x,y))    
+  \end{align}
+\item Risk of the generic model: $$R = \mathbb{E}_{x,y \sim\mathcal{D}}(l(h = M(x, y_{<t}), y_t))$$.\fyDone{wrt to which distribution ?}\fyTodo{is $x$ a sequence ?
+then we need to sum over $t$ to get the loss}
 \item $\mathbf{\Omega_{k,p_{0}}} = \lbrace h | z_k = p_\theta(k|h) < p_0\rbrace$.
-\item Risk of the adapted model for the in-domain test set:
-  $$R_k = \mathbb{E}_{x,y \sim \mathcal{D}_k} [l(h + a_k(h) * f(z_k), y)]$$
+\item Risk of the adapted model for the in-domain test set:\fyTodo{Use same notations as before}
+  $$R_k = \mathbb{E}_{x,y \sim \mathcal{D}_k} [l(h + a_k(h) \odot f(z_k), y)]$$.
 \end{itemize}
 
-We also have two assumptions on the regularity of the activation and the loss functions:
+We also need two assumptions on the regularity of the activation and the loss functions:\fyTodo{Are these standard in such models - if yes, cite}
 \begin{itemize}
-\item $a_k$ is bounded in sense of $\normL{a_k(h)} < M$
-\item the loss function is bounded in sense of $l(h,y) < A$ $\forall h \in \Omega + \mathbb{B}(M)$ where $\mathbb{B}(M)=\lbrace x\in \mathbb{R}^{d_{model}} | \normL{x}<M \rbrace$ and $A+_{\alpha}B = \lbrace x+ \alpha y  | \alpha \in [0..1], x \in A, y \in B \rbrace$  \fyTodo{why not $l(h,y) < A$ ?}
+\item $a_k$ is bounded in sense of $\normL{a_k(h)} < M$\fyTodo{$M$ is already used}
+\item the loss function is bounded:  $l(h,y) < A, \forall h \in \Omega + \mathbb{B}(M)$, where $\mathbb{B}(M)=\lbrace x\in \mathbb{R}^{d_{model}} | \normL{x}<M \rbrace$.
+We note $A +_{\alpha}B = \lbrace x+ \alpha y  | \alpha \in [0 \in 1], x \in A, y \in B \rbrace$ \fyTodo{Notation $A$ is a scalar, not a set}
   $\forall x,y$
@@
 
-Finally, we assume that the classifier is good enough, i.e, $$err_k(p) = \mathbb{E}_{x,y \sim \sum_{i\in [1..K]} \mathcal{D}_{i}*p(d=i)}(\mathbf{1}_{d\neq k} p(k|h)) < \varepsilon_{insep}^k + \gamma$$ $\forall k$ with $\gamma$ sufficiently small.
-We then evaluate the risk of the model adapted to domain $k$ with residuals $a_k$ for arbitrary examples from $K$ domains:
+Finally, we assume that our domain classifier $P_\theta$ is good enough, i.e that\fyDone{Attention à $p$}
+$$\exists \gamma, \forall k, err(k, \theta) = \mathbb{E}_{x,y \sim \mathcal{D}_{k}} \sum_{k' \neq k} p_\theta(k'|h=M(x,y)) < \varepsilon_{insep}(k) + \gamma$$ with $\gamma$ sufficiently small.
+We then evaluate the risk of the model adapted to domain $smallk$ with residuals $a_k$ for arbitrary examples from $K$ domains:\fyTodo{(x,y) is sampled from some distribution}
 \begin{equation}
 \begin{split}
-  R(M,a_k) &= \mathbb{E}_{x,y}[l(h + a_k(h) * f(z_k),y)] \\
-  &= \mathbb{E}_{x,y}[l(h + a_k(h) * f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] + \\
-  &+ \mathbb{E}_{x,y}[l(h + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]
+  R(M,a_k) &= \mathbb{E}_{(x,y) \sim \mathcal{D}}[l(h + a_k(h) \odot f(z_k),y)] \\
+  &= \mathbb{E}_{(x,y) \sim \mathcal{D}}[l(h + a_k(h) \odot f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] + \\
+  &\quad + \mathbb{E}_{(x,y) \sim \mathcal{D}}[l(h + a_k(h) \odot f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]
 \end{split}
@@
 
-By designing a suitable activation function, we can map $f: [0,p_{0}] \rightarrow [0,\delta]$. Then with $\delta$ is small enough we can have an approximation of the first term:
-\begin{equation}
-\mathbb{E}_{x,y}[l(h + a_k(h) * f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] \approx{} \mathbb{E}_{x,y}[l(h,y)\mathbf{1}_\mathbf{\Omega_{k,p_{0}}}(h)] \leq \mathbb{E}_{x,y}[l(h,y)] = R
+By designing a suitable activation function, we can map $f: [0,p_{0}] \rightarrow [0,\delta]$.
+Then with $\delta$ is small enough we can have an approximation of the first term:\fyTodo{We need $l$ to be smooth here, no ?}
+\begin{align*}
+  \mathbb{E}_{x,y \sim \mathcal{D}}[l(h + a_k(h) \odot f(z_k),y)\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}}}(h)] \approx{}&
+                                                                                                                      \mathbb{E}_{x,y \sim \mathcal{D}}[l(h,y)\mathbf{1}_\mathbf{\Omega_{k,p_{0}}}(h)] \\
+  \leq& \mathbb{E}_{x,y \sim \mathcal{D}}[l(h,y)] = R
 \label{eq:4}
-\end{equation}
-The second term: $ L = \mathbb{E}_{x,y}[l(h + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]$ can be bounded as follows:
+\end{align*}
+The second term: $ L(M,a_k) = \mathbb{E}_{x,y \sim \mathcal{D}}[l(h + a_k(h) * f(z_k),y) \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]$ can be bounded as follows:
 \begin{equation*}
 \begin{split}
-L &= \displaystyle{\mathop{\sum}_{k'}\mathbb{E}_{x,y \sim p(.|d=k')}[l(h } + a_k(h) * f(z_k),y) * \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]p(d=k') \\
-	& \leq \displaystyle{\mathop{\sum}_{k' \neq k}} A * \mathbb{E}_{x,y \sim p(.|d=k')} [\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]P(d=k') \\
-	& \quad + \mathbb{E}_{x,y \sim p(.|d=k)}[l(h + a_k(h) * f(z_k),y) * \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]P(d=k) \\
+L(M,a_k) &= \displaystyle{\mathop{\sum}_{k'}\mathbb{E}_{x,y \sim \mathcal{D}_{k'}}[l(h } + a_k(h) * f(z_k),y) * \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]p(d=k') \\
+	& \leq \displaystyle{\mathop{\sum}_{k' \neq k}} A * \mathbb{E}_{x,y \sim \mathcal{D}_{k'}} [\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]P(d=k') \\
+	& \quad + \mathbb{E}_{x,y \sim \mathcal{D}_k}[l(h + a_k(h) * f(z_k),y) * \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h)]P(d=k) \\
 	&\leq A * \frac{\mathbb{\varepsilon}_{insep}^k + \gamma}{p_{0}} + R_k * P(d=k)
@@
 \begin{split}
-\displaystyle{\mathop{\sum}_{k' \neq k}} A * \mathbb{E}_{x,y \sim p(.|d=k')} [\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}} (h)]P(d=k') &\leq A * \displaystyle{\mathop{\sum}_{k' \neq k}} \mathbb{E}_{x,y \sim p(.|d=k') } [\frac{p_k(h)}{p_0}]P(d=k') \\
+\displaystyle{\mathop{\sum}_{k' \neq k}} A * \mathbb{E}_{x,y \sim \mathcal{D}_{k'}} [\mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}} (h)]P(d=k') &\leq A * \displaystyle{\mathop{\sum}_{k' \neq k}} \mathbb{E}_{x,y \sim \mathcal{D}_{k'} } [\frac{p_k(h)}{p_0}]P(d=k') \\
 		& = \frac{A}{p_0} \mathbb{E}[\mathbf{1}_{d\neq k}p_k(h)] \\
@@
 $$ \mathbf{1}_{\mathbf{\Omega_{k,p_{0}}^{c}}}(h) \leq \frac{p_k(h)}{p_0} $$
-Combining the inequality \ref{eq:4} and the inequality \ref{eq:5} we obtain the risk in general case of the adapted model $(M,a_k)$ $$R(M,a_k) \leq R + A * \frac{\mathbb{\varepsilon}_{insep}^k + \gamma}{p_0} + P(d=k)R_k$$
+Combining inequalities \ref{eq:4} and \ref{eq:5} we obtain the risk in general case of the adapted model $(M,a_k)$ $$R(M,a_k) \leq R + A * \frac{\mathbb{\varepsilon}_{insep}^k + \gamma}{p_0} + P(d=k)R_k$$
 
-If we the original version of the residual adapter, with the same assumption in the regularization of the adapter, the bound of the risk of translating out-of-domain is only A (the bound of loss function).
-By using adaptive version, we reduce A by factor $\frac{p_0}{\epsilon^k_{insep}+\gamma}$ which is high if we choose $p_0$ is high enough, the domains are distant and the classifier is good enough.
+If we consider the original version of the residual adapter, with the same assumption in the regularization of the adapter, the bound of the risk of translating out-of-domain (the second term in the inequality above) is only $A$ (the bound of loss function).
+By using an adaptive version, we reduce A by factor $\frac{p_0}{\epsilon^k_{insep}+\gamma}$, which is high if we choose $p_0$ is high enough, the domains are distant in the fixed representation space and the classifier is good enough.
 
+Furthermore $ \varepsilon^k_{insep} < \varepsilon_{insep} $ which is the minimum of the error expectation over the hypothesis space of the classifier.
+With a trivial prediction $[\frac{1}{K},..,\frac{1}{K}]$ we have an error expectation equal to $0.5$.
+Therefore, $\varepsilon_{insep} \leq 0.5$, i.e, $\varepsilon_{insep}(k) < 0.5$.\fyTodo{Je ne comprends pas.} In the worst case where domains in the mixture are very close, we are still able to reduce the risk of predicting out-of-domain examples by a factor of 2.
 \fyTodo{I sort of buy the equations but then so what ? We should may be compare with non-adaptive ?}

### 2020-07-06 2020/coling2020/rebuttal.tex
rebuttal session

\documentclass[12pt,times,a4paper,twoside]{article}

\usepackage[utf8]{inputenc}	% Para caracteres en español
\usepackage[left=3cm,right=3cm,top=2cm,bottom=3cm]{geometry}
\usepackage{pagenote}

\usepackage{algorithm}
\usepackage{algorithmic}

\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{mathtools,xparse}
\usepackage{mathrsfs}

\usepackage{multirow,booktabs}
\usepackage[table]{xcolor}
\usepackage{fullpage}
\usepackage{lastpage}
\usepackage{enumitem}
\usepackage{fancyhdr}

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{calc}
% \usepackage{color,soul}
\usepackage{multicol}
\usepackage{cancel}
\usepackage[retainorgcmds]{IEEEtrantools}
% \usepackage{xcolor} -- implied by xcolor

\usepackage{listings}
\usepackage{spverbatim}
\usepackage{fancyvrb}
\usepackage{hyperref}
\usepackage{float} 
\usepackage{natbib}

\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}

\usepackage{soulutf8}
\usepackage{tabularx}

\usepackage[draft]{todo}
\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}

\colorlet{shadecolor}{orange!15}
\parindent 0in
\parskip 12pt
\geometry{margin=1in, headsep=0.25in}
\setlength{\belowdisplayskip}{8pt} \setlength{\belowdisplayshortskip}{8pt}
\setlength{\abovedisplayskip}{8pt} \setlength{\abovedisplayshortskip}{8pt}
\setlist{nosep}
\setlength{\parskip}{0.1cm}
\setlength{\parindent}{1em}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{definition}{Definition}[section]
\DeclareMathOperator*{\argmax}{argmax}
\newcommand{\R}{\ensuremath{\mathbb{R}}}

\pgfkeys{
    /tr/rowfilter/.style 2 args={
        /pgfplots/x filter/.append code={
            \edef\arga{\thisrow{#1}}
            \edef\argb{#2}
            \ifx\arga\argb
            \else
                \def\pgfmathresult{}
            \fi
        }
    }
}

\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\NewDocumentCommand{\normL}{ s O{} m }{%
  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{L_1}%
}

\title{Combination of generic and specific layers in multidomain learning}
\author{}
\date{}

\begin{document}

\maketitle
\begin{center}
{\LARGE \bf}Rebuttal session
\end{center}

\section*{General comments}
As TACL action editor for submission 2327, "Revisiting Multi-Domain Machine Translation", I am happy to tell you that I am accepting your paper subject (conditional) to your making specific revisions within two months.

Please address all comments by the reviewers, paying special attention to the following:
\begin{itemize}
	\item Address the concern by reviewer A about the difference between MDL Res and FT-Block, and include FT-block (and ideally FT-Full also) in the later experiments (5.2.2 and 5.3).
	\item Carefully specify the number of parameters for all models, including the proportions that are domain-specific and domain-independent.
	\item Provide all the experimental details requested, and in particular give enough information to allow the corpora to be reproduced, or (ideally) make them publicly available.
\end{itemize}

\section*{Reviewer A}
Reviewer A:
\\
Overall, I'm a bit ambivalent about the paper. It is a well-written paper with a carefully-executed set of experiments. An important contribution of the paper is to formulate a set of expected properties of multi-domain MT (MDMT) (against which existing approaches are evaluated), and I generally like the choice of properties and that the paper draws attention to them, for example systematically testing how multi-domain approaches behave when gold domain labels are suboptimal. However, it is ambitious to systematically compare domain adaptation or multi-domain approaches which were developed with various conditions in mind. In several cases, I get the feeling that in an attempt to be broad and systematic, the authors have drawn boundaries and distinctions where none need exist, or drew conclusions that are too general. Details follow below.
\\
The authors have chosen a setting where domain labels are generally
available at training and test time, and in the the empirical comparison, focus on approaches that use domain labels at training and test time (533-537), but this is only partially true: their comparison does include systems which only use domain labels at training time (TTM/DM/ADM/WDCMT), and one set of experiments operates with automatically induced domain labels based on k-means clustering of sentence representations. Systems which were designed to work without domain labels, such as that by Farajian et al.(2017) would be a very relevant comparison at least in the latter setting without gold domain labels.
\\
The authors consider fine-tuning separate from MDMT, and declare failure in that no MDMT system meets the expectation that it should outperform fine-tuning. I'm confused why fine-tuning is specifically excluded, since it meets many of the reasons for building MDMT systems. Authors do mention that it would be cheaper to develop a single system instead of having to maintain multiple systems, but they include "MDL Res", which has domain-specific adaptation modules and is very similar to "FT-Block" (a baseline that also combines a general model with domain-specific adaptation modules). The line
between a single system and multiple systems gets fuzzy here, and I see no clear motivation to draw this clear distinction and only consider "MDL Res" a multi-domain system.
\\
The large number of approaches considered is admirable. However, the depth of the discussion suffers as a consequence, and the paper risks making over-general statements about the weaknesses of different approaches, despite related work that has already addressed them. The authors conclude that fine-tuning "incurs large losses [when applied on] the small data condition", and fails to deteriorate gracefully in the presence of tag errors, issues that have been addressed by Miceli Barone et al. (2017) and Thompson et al. (2019), respectively. Similarly, the negative impact of adding data from a "new" domain seems slightly overblown. Firstly, this is probably highly dependent on data mixtures, length of training and regularization strategies, so I'd be wary of drawing general conclusions from a single set of experiments. Secondly, when looking at the models where such a performance deterioration is said to happen (DC-Tag, TTM, ADM; lines
770-771), this deterioration is actually small or non-existent when
comparing table 3 and table 5 averages (-0.2 BLEU for DC-Tag; -0.3 BLEU for TTM; +0.9 BLEU for ADM).
\\
Question to authors:
\\
- can you describe in more detail the differences between FT-Block and MDL Res, apart from the fact that the former is fine-tuned from a generic model, the latter trained from scratch?
\\
\fyTodo{system's description}
\\
minor issues:
\\
- I disagree that generalisation to unseen domains is similar to zero-shot multilingual MT. In zero-shot multilingual MT, both the source and target distribution have been observed at training time, just paired with other languages. In the generalisation to unseen domains, there may be characteristics in the data that were never seen in the training data, including new vocabulary (in the source text, target text, or both).
\\
- 567-568: "having the domain tag on the source is better than on the
target". I assume this refers to the comparison of "DC-Tag" and "TTM". However, TTM actually uses the predicted domain label on the target side, and could perform better with forced decoding of the gold domain label. More precisely, the statement should be that "having the gold tag is better than a predicted tag", but this is hardly surprising.
\section*{Reviewer B}
The paper presents a thorough review of Multi-Domain Machine Translation (MDMT) approaches, and the authors provide comprehensive experiments to study different aspects of MDMT. The innovation of this work can be seen in two main categories:
\begin{itemize}
	\item suggesting a methodology to evaluate MDMT models based on seven properties that they defined in the paper.
	\item empirical studies on eight recent MTMD approaches.
\end{itemize}
In general, the paper has written well, and it is a good read for those who are researching in the domain adaptation field. 
\\
I have a few comments:
\\
* You provide some useful experimental results on various MDMT approaches. To benefit from your results, we need access to the same segmentation of train/dev/test sets that I hope you will provide to the community.
\\
* On your discussion on the number of parameters (L. 591), could you share the number of parameters per each MDMT approach? It will be helpful to see  the impact of the number of parameters on the MT results. 
\\
* Thinking about your seven operational requirements, I wonder if you could also define a domain or the minimum number of samples to specify a domain. I like your selections of domains that you include multiple domains with varying sizes, but what we expect in extreme cases. Having your Split experiments on REL (Table 4), or providing further details on your last analysis in Section 5.3, if you have tiny clusters could help.
\\
* Please define Lambda coefficients in lines 142 and 146.
\\
* On dynamically changing the number of training and test domains, there is a new related work published recently that might worth citing it: Sharaf A, Hassan H, Daumé III H. Meta-Learning for Few-Shot NMT Adaptation.
\\
* I’d prefer to see the number of running words in Table 1, especially when the batch size is given as the number of tokens (L. 485).
\\
* L. 503: It seems you refer to this method as LDR in the rest of the paper. Please make them consistent.
\\
* L. 576: Although you describe in footnote 11, it is better to mention that you compare WDCMT results with Mixed-NAT-RNN explicitly.
\\
* L. 681-683: I don’t see it as a failure, and I think it is a property (requirement) as you describe in the definition of [P4-ERR], we expect deterioration in the presence of tag error. Am I right?
\\
* L. 687: Please describe how you predict the domain tag, and please share the tag prediction error rate on each domain test set. It is helpful to see the amount and the impact of the error on MDMT results.
\\
I spotted a few typos in the text:
\\
* L 100: section 2 $\rightarrow$ Section 2
\\
* L 104: section 3 $\rightarrow$ Section 3
\\
* L 256: could ben $\rightarrow$ could be (?)
\\
* L 702, Table 4, Col. 5: MED1 $\rightarrow$ MED2
\\
* L 790: that $\rightarrow$ than
\\
* L 802-805: It’s better to bring DC-Feat results after DC-Tag to keep the consistency with other tables 
\\
* L 928: ge:en, do you mean de:en (German-English)? 
\\
* L 973: domains:this $\rightarrow$ domains this
\\
* L 979: drawed $\rightarrow$ drawn 
\section*{Reviewer D}
The authors deliver on their promise given in the title and abstract of the paper by doing exactly what I was hoping for: They provide a dense but comprehensive review of approaches to multi-domain MT. In a commendably straight and structured manner, they identify requirements and challenges of multi-domain MT. And they then proceed with empirically evaluating lots of multi-domain MT approaches.
\\
The empirical evaluation replicates more than ten multi-domain translation approaches from the literature and directly compares them against each other on the English-French language pair. The experiments are conducted using a coherent scenario with respect to both the amount of training data and the amount and type of distinct domains that the authors test on.
\\
The authors come up with a series of experiments that carefully considers the initially identified requirements and challenges. The (seven) identified challenges fall under the three categories "effectiveness", "robustness", and "scalability". There can be competing objectives when building multi-domain translation engines, and the results and discussion in the paper reveal the strengths and weaknesses of the various approaches in the considered scenario.
\\
Even though I cannot easily name the one striking contribution of this work, it's quite obvious to me that many fellow researchers will benefit from the clear presentation of multi-domain MT as well as from the extensive empirical results provided in this paper. I'd like to see it published.
\\
Let me nevertheless formulate a few questions or points of criticism:
\\
- I'd like to see some analysis of vocabulary coverages (such as
cross-domain OOV rates), and of "seen errors" and "new sense errors" as defined by Irvine et al. in "Measuring machine translation errors in new domains" (TACL 2013).
\\
- Maybe the (possible) effect of early stopping (and the choice of the validation set) should be discussed? E.g., the validation set for
finetuning can originate from one single domain for each finetuned model. The way you're doing early stopping may even have had an impact on the results in Table 5. Is your validation set always a concatenation of 1000 lines from the six domains, without NEWS? Are you actually always training until the early stopping criterion is reached, and are you then using the model from the best checkpoint? (There are contradictory statements in the paper about "50,000  additional iterations" (?) for the experiments in Table 5.)
\\
- Report the learning rate. Is it decreased for finetuning?
\\
- To some extent, I dislike Table 5. When not reading carefully, the
difference values could be interpreted as differences wrt. scores in Table 3. (Except that the numbers would be wrong...) Differences wrt. to Table 3 are important and it took me a while to obtain an overview for myself. Furthermore, it should be immediately obvious which number is "resumed training" and which is training from scratch. (It isn't from the table caption alone.)
\\
- How is "Mixed-Bal" achieved? By means of upsampling of the corpora from the less highly represented domains?
\\
- For "DC-Feat", is the feature added for each word or for each subword token? Is it attached to the word (increasing the vocabulary size a lot) or does it remain a separate token (increasing the input length a lot)?
\\
- How is the domain-weighted average (WAVG) calculated? Is the weight determined by proportion of the respective domain in the training data?
\\
- Wasn't it possible to use standard test sets? The "REL" Bleu scores are suspiciously high and there might be considerable redundancy in the data, or rather: training data contaminated with test data. (Since duplicates are possible, have you verified that the test sentences don't still appear in the remaining training data?)
\\
Other (minor) remarks:
- From the UFAL medical corpus, please say that you're using the indomain sections only (which is mostly PatTR Medical).
\\
- The "DC-Feat" and "DC-Tag" lines in Table 5 should be swapped in order to be in line with Tables 3 and 4.
\\
- l. 503: "LDC-Feat" $\rightarrow$ "LDR"
\\
- l. 108: "automating" $\rightarrow$ "automatic"
\\
- l. 183: "accross" $\rightarrow$ "across"
\\
- l. 208: "under-resource" $\rightarrow$ "under-resourced"
\\
- l. 256: "ben" $\rightarrow$ "be"
\\
- l. 310: "used MT" $\rightarrow$ "used in MT"
\\
- l. 379: "Corpus statistics in Table 1." $\rightarrow$ "Corpus statistics are given in
Table 1."
\\
- l. 380: "variable" $\rightarrow$ "available"
\\
- l. 426: "the all others" $\rightarrow$ "all others"
\\
- l. 696: "suplementary" $\rightarrow$ "supplementary"
\\
- l. 979: "drawed" $\rightarrow$ "drawn"
\\
- l. 991: "may be" $\rightarrow$ "maybe"
\end{document}

### 2020-07-07 2020/coling2020/rebuttal.tex
correcting spelling errors, adding answers

@@
 
-\title{Combination of generic and specific layers in multidomain learning}
+\title{Rebuttal: Revisiting Multi-Domain Machine Translation}
 \author{}
@@
 \begin{document}
-
 \maketitle
-\begin{center}
-{\LARGE \bf}Rebuttal session
-\end{center}
-
 \section*{General comments}
@@
 \\
-\fyTodo{system's description}
+\textcolor{red}{The two methods use the same architecture composing of one standard Transformer model and several domain adapters that is the same as the model presented in the paper of \cite{Bapna19simple}.
+The only difference between FT-block and MDL Res is their training process.
+\\
+To train FT-block, we plug residual adapters to the output of each self-attention layer of the model Mixed-NAT, then resume the training with the learning rate captured in the end of the training of the model Mixed-NAT while freezing all of the parameters of the model Mixed-NAT.
+We use only the in-domain data as the training data of the finetuning procedure.
+\\
+To train MDL res, we optimize the parameters of the Transformer model and the parameters of the residual adapters from scratch.
+Precisely, to create of training batch, we randomly pick a domain with a frequency proportional to the size of its data, then we sample a batch from its corpora.
+Given a batch from domain i, we optimize the parameters of Transformer model and the parameters of the residual adapters corresponding to the domain i by plugging the residual adapters of domain i to the Transformer model and calculating gradients of the activated parameters.}
 \\
@@
 To benefit from your results, we need access to the same segmentation of train/dev/test sets that I hope you will provide to the community.
+\fyTodo{to share data}
 \\
@@
 It will be helpful to see  the impact of the number of parameters on the MT results.
+\fyTodo{system's description}
 \\
@@
 \\
+\textcolor{red}{We formalize the data distribution as a mixture of several in-domain distribution.
+$\lambda$ is the prior distribution on the set of domains.
+We hypothesize that training distribution is different from testing distribution, so we have 2 different mixtures.}
+\\
 * On dynamically changing the number of training and test domains, there is a new related work published recently that might worth citing it: Sharaf A, Hassan H, Daumé III H.
@@
 \\
+\textcolor{red}{This is a very interesting work that we would like to cite in our final version of the paper.}
+\\
 * I’d prefer to see the number of running words in Table 1, especially when the batch size is given as the number of tokens (L. 485).
+\\
+\fyTodo{to report number of tokens of the datasets}
 \\
@@
 \\
+\textcolor{red}{We have noted this remark. We agree to change the notation correctly.}
+\\
 * L. 576: Although you describe in footnote 11, it is better to mention that you compare WDCMT results with Mixed-NAT-RNN explicitly.
+\\
+\fyTodo{to explicitly compare WDCNMT with Mixed-NAT-RNN}
 \\
@@
 It is helpful to see the amount and the impact of the error on MDMT results.
+\\
+\fyTodo{to report in details the tag prediction.}
 \\
@@
 \\
+\fyTodo{extra experiments to do.}
+\\
 - Maybe the (possible) effect of early stopping (and the choice of the validation set) should be discussed? E.g., the validation set for
@@
 \\
+\textcolor{red}{For each domain, we create train/dev/test sets by randomly splitting the respective corpora.
+We maintain the size of validation sets and of test sets equal to 1000 lines for every domain.
+In the table 3 and 4, to train MDMT models except FT-full and FT-block we use a validation set which is a concatenation of 6 validation sets of 6 domains.
+In the table 5, to train MDMT models except FT-full and FT-block we concatenate the validation set of NEWS to the mixture of 6 validation sets.
+To train FT-full model and FT-block model we only use the in-domain validation set.
+We stop training if either the training iteration reach the maximal number of iterations or the score on the validation set does not increase in 3 consecutive evaluations.
+We average 5 checkpoints up to the best checkpoint to get the final model.}
+\\
 - Report the learning rate. Is it decreased for finetuning?
+\textcolor{red}{The learning rate is determined by this formula $lr = d_{model}^{-0.5}*\min(step\_num^{-0.5},step\_num * warmup\_steps^{-1.5})$ where $d_{model}=512$,$warmup\_steps=4000$.
+In the finetuning procedure, we continue the training using the same learning rate schedule and the same $step\_num$}
 \\
@@
 \\
+\textcolor{red}{In the table 5, we report the difference of the model 's performance in 2 situations including resuming training and training from beginning with 7 domains.
+We would like to report the difference wrt to table 3 in the later version of the paper.
+We agree that the comparison to table 3 is also important, because it could assess the impact of new domain to the previously learned domains.}
+\\
 - How is "Mixed-Bal" achieved? By means of upsampling of the corpora from the less highly represented domains?
+\\
+\textcolor{red}{In all of our experiments, to get a training batch, we first sample a domain from a set of available domains in training procedure with a fixed probability, the we sample a from the corpora corresponding to the previously picked domain.
+Therefore, between our notations, "Mixed-NAT" uses a domain probability equal to proportion of domain in training data, and "Mixed-Bal" uses a domain probability equal to uniform probability.}
 \\
@@
 \\
+\textcolor{red}{DC-Feat is a small embedding representing the domain.
+It is concatenated to every word embedding whenever its corresponding domain is detected.
+The use of DC-Feat increase the size of actual embedding, but does not increase the vocabulary size nor the input length.}
+\\
 - How is the domain-weighted average (WAVG) calculated? Is the weight determined by proportion of the respective domain in the training data?
+\\
+\textcolor{red}{The domain-weighted average is calculated by weighting the performance of the model in each domain test with the proportion of the respective domain in the training data}
 \\
@@
 (Since duplicates are possible, have you verified that the test sentences don't still appear in the remaining training data?)
+\\
+\textcolor{red}{There are standard test sets for the domains including medical, IT, news and talk, but we could not find any standard test sets for the other domains including religion, law and bank.
+We carefully eliminated every duplicate in all of the experimental corpora.
+In our opinion, the reason the model achieved extremely high score in the domain religion is because the text in this domain is much less variable}
 \\
@@
 - l. 991: "may be" $\rightarrow$ "maybe"
+
+\bibliographystyle{acl_natbib}
+\bibliography{multidomain}
 \end{document}

### 2020-07-07 2020/coling2020/multidomain.tex
correcting spelling errors, adding answers

@@
 
-In this paper, we formulate in a more precise fashion the requirements that an effective MDMT system should meet (section~\ref{sec:requirements}).
-Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT systems and associated measurements to evaluate them (section~\ref{sec:challenging}).
-In doing so, we also shed lights on new problems that arise in this context, regarding for instance the accomodation of new domains in the course of training, or the computation of automating domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
+In this paper, we formulate in a more precise fashion the requirements that an effective MDMT system should meet (Section~\ref{sec:requirements}).
+Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT systems and associated measurements to evaluate them (Section~\ref{sec:challenging}).
+In doing so, we also shed lights on new problems that arise in this context, regarding for instance the accommodation of new domains in the course of training, or the computation of automatic domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
 We show in Section~\ref{sec:experiments} that existing approaches still fall short to match many of these requirements, notably with respect to handling a large amount of heterogeneous domains and to dynamically integrating new domains in training.\fyDone{Spell out conclusions}
@@
 
-A second line of reasoning rests on linguistic properties of the translation function and contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and tend to remain semantically stable accross domains, motivating some cross-domain parameter sharing.
+A second line of reasoning rests on linguistic properties of the translation function and contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and tend to remain semantically stable across domains, motivating some cross-domain parameter sharing.
 A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the translation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
@@
 Under this view, MDMT would be especially beneficial for domains with little training data.
-This is observed for multilingual MT from English: an increase for under-resource languages due to positive transfer, at the cost of a decrease in performance for well-resourced languages \cite{Arivazhagan19massively}.
+This is observed for multilingual MT from English: an increase for under-resourced languages due to positive transfer, at the cost of a decrease in performance for well-resourced languages \cite{Arivazhagan19massively}.
 
@@
 If finally, it is to make the system more robust to unexpected or adversarial test distributions, then this is the setting that should be used to evaluate MDMT.
-The next section discusses ways in which these requirements of MDMT systems could ben challenged.
+The next section discusses ways in which these requirements of MDMT systems could be challenged.
 
@@
 A second set of requirements is related to the definition of a domain.
-As repeatedly pointed out in the literature, parallel corpora used MT are often collected opportunistically and the view that each corpus constitutes a single domain is often a gross approximation.\footnote{Two of our own ``domains'' actually comprise several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments.
-To challenge these requirements we propose the evaluate the following requirements.
+As repeatedly pointed out in the literature, parallel corpora in MT are often collected opportunistically and the view that each corpus constitutes a single domain is often a gross approximation.\footnote{Two of our own ``domains'' actually comprise several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments.
+To challenge these requirements we propose evaluating the following requirements.
 
@@
 Our discriminator is a SVM independently trained for each pair of domains, with sentence representations derived via mean pooling from the source side representation of the generic Transformer model.
-We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well separated from the all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
+We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well separated from all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
 
@@
 \item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});\fyDone{why not LDR as in Minh's paper?}
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDR});\fyDone{why not LDR as in Minh's paper?}
 \item the three proposals of \newcite{Britz17mixing}.
@@
      Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{law} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} &  \multicolumn{1}{c|}{\domain{rnd}} &  \multicolumn{1}{c|}{\domain{new}}\\ \hline
-    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
+    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_2$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
     \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5} & -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& \SW{-19.6} & \SW{-3.3}\\% [32.5]
@@
 \subsubsection{Handling growing numbers of domains}
-Another series of experiments evaluate the ability to dynamically handle suplementary domains (requirement [P6-DYN]) as follows.
+Another series of experiments evaluate the ability to dynamically handle supplementary domains (requirement [P6-DYN]) as follows.
 Starting with the existing MD systems of Section~\ref{ssec:rawperformance}, we introduce an extra domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is critical for achieving optimal performance: as our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy based on the new training distribution.
@@
     \hline%\hline
+    \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5  & 41.4 & 49.9\\[-2pt]
+%                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
+                   & \scriptsize +0.3  & \scriptsize -0.1   & \scriptsize -0.6    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.3}  & \scriptsize \SW{-3.5}  & \scriptsize \SW{-1.4}  & \scriptsize -0.1 & \scriptsize \SW{-1.2}\\
+
 	\system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9  & 41.2 & 50.1\\[-2pt]
@@
                    & \scriptsize -0.2 & \scriptsize -0.1  & \scriptsize -0.1 & \scriptsize -0.6 & \scriptsize -0.9 & \scriptsize +0.3 & \scriptsize \SW{-0.8} & \scriptsize -0.2 & \scriptsize -0.3\\
-
-    \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5  & 41.4 & 49.9\\[-2pt]
-%                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
-                   & \scriptsize +0.3  & \scriptsize -0.1   & \scriptsize -0.6    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.3}  & \scriptsize \SW{-3.5}  & \scriptsize \SW{-1.4}  & \scriptsize -0.1 & \scriptsize \SW{-1.2}\\
 
@@
 We also note that many of these classes are mere split of the large \domain{med}, while a fewer number of classes are mix of two existing domains.
-We are thus in a position to reiterate at a larger scale the measurements of Section~\ref{ssec:redomains} and test whether multi-domain systems can effectively take advantage from the cross-domain similarities, and to eventually perform better that fine-tuning.
+We are thus in a position to reiterate at a larger scale the measurements of Section~\ref{ssec:redomains} and test whether multi-domain systems can effectively take advantage from the cross-domain similarities, and to eventually perform better than fine-tuning.
 
@@
 Another parameter-sharing scheme is \cite{Jiang19multidomain}, which augments a Transformer model with domain-specific heads, whose contributions are regulated at the word/position level: some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling that the core of \cite{Huck15mixeddomain,Saunders19ucam}.
-The results for three language pairs outperform several standard baselines for a two-domain systems (in fr:en and ge:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
+The results for three language pairs outperform several standard baselines for a two-domain systems (in fr:en and de:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
 
@@
 \fyDone{Mitigate this}
-As expected however, MDMTs are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend, dynamically accommodate additional domains:this is especially easy for feature-based methodologies.
+As expected however, MDMTs are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend, dynamically accommodate additional domains, this is especially easy for feature-based methodologies.
 Our experiments finally suggest that all methods show decreasing performance when the number of domain increases (for a fixed data size).
 
-Two other main conclusions can be drawed from this study: first, it seems that more work is needed to make MDMT systems make the best out of the variety of the available data, both to effectively share what needs to be shared while at the same time separating what needs to be kept separated.
+Two other main conclusions can be drawn from this study: first, it seems that more work is needed to make MDMT systems make the best out of the variety of the available data, both to effectively share what needs to be shared while at the same time separating what needs to be kept separated.
 We notably see two areas worthy of further explorations: the development of parameter sharing strategies when the number of domains is large; the design of training strategies that can effectively handle a change of the training mixture, including an increase in the number of domains.
 Both problems are of practical relevance in industrial settings.
-Second, and may be more importantly, there is a general need to adopt better evaluation methodologies for evaluating MDMT systems, which require systems developers to clearly spell out the testing conditions and the associated expected distribution of testing instances, and to report more than mere comparisons with simple baselines on a fixed and known handful of domains.
+Second, and maybe more importantly, there is a general need to adopt better evaluation methodologies for evaluating MDMT systems, which require systems developers to clearly spell out the testing conditions and the associated expected distribution of testing instances, and to report more than mere comparisons with simple baselines on a fixed and known handful of domains.
 

### 2020-07-07 2020/coling2020/rebuttal.tex
adding answers

@@
 \\
-\textcolor{red}{The two methods use the same architecture composing of one standard Transformer model and several domain adapters that is the same as the model presented in the paper of \cite{Bapna19simple}.
+\textcolor{red}{The two methods use the same architecture composing of one standard Transformer model and domain adapters presented in the paper of \cite{Bapna19simple}.
 The only difference between FT-block and MDL Res is their training process.
@@
 To train MDL res, we optimize the parameters of the Transformer model and the parameters of the residual adapters from scratch.
-Precisely, to create of training batch, we randomly pick a domain with a frequency proportional to the size of its data, then we sample a batch from its corpora.
+Precisely, to create a training batch, we randomly pick a domain with a probability equal to the proportion of the domain in training data, then we sample a batch from its corpora.
 Given a batch from domain i, we optimize the parameters of Transformer model and the parameters of the residual adapters corresponding to the domain i by plugging the residual adapters of domain i to the Transformer model and calculating gradients of the activated parameters.}
@@
 \\
-- I disagree that generalisation to unseen domains is similar to zero-shot multilingual MT.
+- I disagree that generalization to unseen domains is similar to zero-shot multilingual MT.
 In zero-shot multilingual MT, both the source and target distribution have been observed at training time, just paired with other languages.
-In the generalisation to unseen domains, there may be characteristics in the data that were never seen in the training data, including new vocabulary (in the source text, target text, or both).
+In the generalization to unseen domains, there may be characteristics in the data that were never seen in the training data, including new vocabulary (in the source text, target text, or both).
 \\
@@
 More precisely, the statement should be that "having the gold tag is better than a predicted tag", but this is hardly surprising.
+\\
+\textcolor{red}{We agree that the conclusion is not really fair for "TTM" because "TTM" intend to predict domain label rather than to use gold domain label and TTM could definitively translate with gold domain label as prefix of each hypothesis.
+We will report the performance of "TTM" when translating using gold domain label as prefix in target side.}
 \section*{Reviewer B}
@@
 \\
-* On your discussion on the number of parameters (L.
-591), could you share the number of parameters per each MDMT approach?
-It will be helpful to see  the impact of the number of parameters on the MT results.
+* On your discussion on the number of parameters (L. 591), could you share the number of parameters per each MDMT approach? It will be helpful to see  the impact of the number of parameters on the MT 
+results.
+\\ 
+\textcolor{red}{The description of model in each MDMT approach is as follows:
+\begin{enumerate}
+	\item FT-full, Mixed-NAT, TTM, DC-Tag use medium Transformer model.
+	\item FT-block and MDL res use medium Transformer model, and residual layers with the bottleneck dimension of size 1024.
+\cite{Bapna19simple} showed that if the bottleneck dimension is large enough, the gap between FT-block and FT-full will be small.
+	\item ADM, DM use medium Tranformer model and a domain classifier composing of 3 dense layers of size $512 \times 2048$, $2048 \times 2048$ and $2048 \times domain\_num$.
+The two first layers of the classifier use relu as activation function, the last layer uses tanh as activation function.
+	\item DC-Feat uses medium Transformer model, and domain embeddings of size 4.
+Given a sentence of domain i in training batch, the embedding of domain i is concatenated to the embedding of each token in the sentence.
+	\item LDR uses medium Transformer model and for each token we introduce a LDR feature of size $4 \times domain\_num$.
+Given a sentence of domain i $i\in[1,..,K]$ in training batch, for each token of the sentence, the LDR units of the indexes outside of the range $[4(i-1),..,4i-1]$ are masked to 0, the masked LDR feature will be concatenated to the embedding of the token.
+	\item WDCNMT uses one layer GRU bidirectional as encoder and one layer GRU-conditional as decoder. The size of each hidden layer is 1024, the size of word embedding is 512.
+	\item Mixed-NAT-RNN uses one layer LSTM bidirectional as encoder and one layer LSTM as decoder. The size of each hidden layer is 1024, the size of word embedding is 512. 
+\end{enumerate}
+}
 \fyTodo{system's description}
@@
 \\
-\textcolor{red}{In the table 5, we report the difference of the model 's performance in 2 situations including resuming training and training from beginning with 7 domains.
+\textcolor{red}{In the table 5, we report the difference of the model's performance in 2 situations including resuming training and training from beginning with 7 domains.
 We would like to report the difference wrt to table 3 in the later version of the paper.
@@
 It is concatenated to every word embedding whenever its corresponding domain is detected.
-The use of DC-Feat increase the size of actual embedding, but does not increase the vocabulary size nor the input length.}
+The use of DC-Feat increase the size of the input embedding, but does not increase the vocabulary size nor the input length.}
 \\
@@
 \\
-\textcolor{red}{The domain-weighted average is calculated by weighting the performance of the model in each domain test with the proportion of the respective domain in the training data}
+\textcolor{red}{The domain-weighted average is calculated by weighting the performance of the model in each domain test with the proportion of the respective domain in the training data.}
 \\
@@
 We carefully eliminated every duplicate in all of the experimental corpora.
-In our opinion, the reason the model achieved extremely high score in the domain religion is because the text in this domain is much less variable}
+In our opinion, the reason why the models achieve extremely high scores in the domain religion is because the text in this domain is much less variable.}
 \\
@@
 \\
+\textcolor{red}{From the UFAL medical corpus of the language pair En-Fr, we select only sentences in categories including EMEA, ParTR, CESTA, ECDC.}
+\\
 - The "DC-Feat" and "DC-Tag" lines in Table 5 should be swapped in order to be in line with Tables 3 and 4.
+\\
+\textcolor{red}{We agree to change the order of "DC-Feat" and "DC-Tag" to be consistent with tables 3 and 4.}
 \\

### 2020-07-07 2020/coling2020/rebuttal.tex
small change

@@
 \\
-\fyTodo{to report in details the tag prediction.}
+\textcolor{red}{We train a language model in source language for each domain using respective source training corpus.
+To predict domain tag of a sentence, we compute probability of the sentence according to each pretrained language model and choose the domain corresponding to the highest probability as domain tag of sentence.} \fyTodo{to compute error rate of domain prediction over each test set}
 \\
@@
 - Report the learning rate. Is it decreased for finetuning?
+\\
 \textcolor{red}{The learning rate is determined by this formula $lr = d_{model}^{-0.5}*\min(step\_num^{-0.5},step\_num * warmup\_steps^{-1.5})$ where $d_{model}=512$,$warmup\_steps=4000$.

### 2020-08-11 2020/wmt_20/emnlp2020.tex
changed title
first draft

@@
 \usepackage{amsmath,amsfonts,amssymb}
+\usepackage{mathrsfs}
+\newcommand{\R}{\ensuremath{\mathbb{R}}}
 
@@
 % but it will improve the layout of the manuscript,
-% and will typically save some space.
+% and will typically save some space.Boldface denotes significant gains with respect to \system{Mix-Nat} (or \system{Mix-Nat-RNN}, for WDCMT), underline denotes significant losses.
 \usepackage{microtype}
@@
 \newcommand{\mpDone}[1]{\done[MP]\Todo[MP:]{\textcolor{green}{#1}}}
-
+\usepackage{mathtools,xparse}
+\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
+\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
+\NewDocumentCommand{\normL}{ s O{} m }{%
+  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{L_2}%
+}
 \newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
@@
 
-\title{Less catastrophic forgetting domain adaptation}
+\title{Ablation study on the residual adapter in Neural Machine Translation}
 
@@
 \begin{abstract}
-\mpTodo{write abstract}
+Among the approaches for adapting a pretrained NMT model to a specific domain, finetuning is the dominant method.
+Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{bapna19simple, Vilar18learning}.
+The second approach has several advantages including preserving the pretrained model and the legerity of the adapters.
+However, the behavior of these adapters has not been further studied.
+The objective of this paper is to study the behavior of the adapters, then to extend the application of the adapters to improve the robustness of the model before domain shift and noisy data.
+\mpTodo{correcting abstract}
 \end{abstract}
+
 \section{Introduction } \label{sec:intro}
 \mpTodo{write introduction}
-\citet{Bapna19simple}
-\section{Word-Level Adaptive Domain Adaptation \label{sec:wada}}
+Neural Machine Translation(NMT) has achieved significant performance in many language pairs \cite{Bahdanau15learning,Vaswani17attention,Kalchbrenner13recurrent,Sutskever14sequence}.
+However, as many deep learning model, NMT model need to be trained with a sufficiently large amount of data to reach a high performance; therefore the quality of translation of NMT model is still limited in low-resourced language, and low-resourced domain \cite{duh13adaptation,koehn17six,zoph16transfer}.
+While many approaches have been proposed To improve the quality of NMT model in low-resourced domain (\cite{Chu18asurvey}), full finetuning of model parameters remains the dominant approach for adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
+One first trains an NMT model with a large parallel corpora which is always a mix of many available parallel corpora.
+Given a domain of interest and a parallel corpora of the domain, one could adapt the pretrained model by continue training the model with the in-domain corpora.
+It is a conjecture that pretrained model is a better initialization of training NMT model with small corpora than the random initialization.
+Indeed, studies in transfer learning in NMT such as \cite{artetxe20cross,aji20neural} confirms this claim by extensive experiments.
+Finetuning by training all the parameters (\system{FT-full}) of the model improves significantly the quality of the NMT model on a specific domain on the one hand, loses the translation quality on other domains on the other hand.
+This phenomenon is referred to Catastrophic Forgetting in training neural networks \cite{McCloskey89catastrophic}.
+Therefore, full finetuned model is usefull only to the target domain.
+As the number of domains grow, training and maintaining finetuned models consumes a lot of resource.
+\cite{Vilar18learning},\citep{bapna19simple} propose a simple method to preserve the value of pretrained model by plugging small adapters to hidden layers which will be trained only with the in-domain data while the pretrained model is frozen.
+Because these additional adapters are very small compared to the size of the model, we could reduce significantly the cost of training and maintaining finetuned models.
+
+Because of the extensively demonstrated efficacy of the adapters, we would like to extend the use of this architecture to improve NMT model in several particular situations.
+In translation task, there are several factors that challenge NMT model, including: noise, topical text.
+Residual adapters allow to adapt NMT model to any specific domain in a computationally economical way.
+Could we adapt NMT model to noisy text and topical text at once?
+While residual adapters are so good to adapt to a specific domain, it still loses dramatically performance in the previously training domains as we show in our experiments.
+Therefore, we have to decide whether to use residual adapters manually, i.e, we have to know to which domain the text belongs a priori.
+Therefore, we would like to fuse a domain classifier to the architecture in order to weight the contribution of the adapters with respect to the topic-relatedness of the text.
+
+Our contribution is an ablation study on the use of residual adapter in NMT domain adaptation by answering 3 following question:
+\begin{itemize}
+	\item How the number and position of adapters affect the performance.
+	\item How to train the residual adapter.
+	\item How to be robust to out-of-domain examples.
+\end{itemize}
+
+\section{Residual adapters \label{sec:res}}
+
+\subsection{Architecture \label{ssec:architecture}}
+Residual adapters modifies context vectors of each layer by applying several transformations as follow:
+$$ h^{i}_1 = \mathbf{W}_{d_{model}}^{d_{bottleneck}}h^{i} + b_{1}$$
+$$ h^{i}_2 = \mathbf{RELU}(h_1)$$
+$$ h^{i}_3 = \mathbf{W}_{d_{bottleneck}}^{d_{model}}h_2 + b_{2}$$
+$$ \bar{h}^{i} = h^i_3 + h $$
+For the sake of brevity, we denote $ADAP^{(i)}(h_i) = h^i_3$.
+
+The "adapted" context vector $\bar{h}^i$ goes to the $i-th$ self-attention layer.
+There are 12 residual adapters corresponding to total 12 attention layers of the encoder and the decoder (in the decoder, the stack of self-attention layer and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter).
+\subsection{Ablation study on position and number of residual adapters \label{ssec:ablatation}}
+In this section, we would like to analyze the impact of position and number of residual adapter involved in the adapted model.
+The importance of hidden layers varies with respect to their position in NMT model.
+It is conjectured that the higher layer represents more global pattern such as syntax while the lower layer represents more local pattern such as semantic.
+The domain shift in local patterns and global patterns has not yet studied.
+In this paper, we do not intend to study this aspect of domain adaptation problem.
+We would like to give an extensive comparison between domain adaptation in different levels in NMT model.
+Because the set of possible configurations is large, we only perform experiments on layers at position: 2, 4, 6 which covers a wide range in the hierarchy of attention layers.
+
+Secondly, we would like to observe how much the number of residual adapters can affect the performance of the adapted model.
+By the limit of computation resource, we limit our study to the case of 3 adapters at positions: 2,4,6.
+By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting).
+
+\subsection{Regularization of residual adapters \label{ssec:reg}}
+Finetuning on small corpora can still lead to overfitted model even when the finetuning is applied to a few residual adapters.
+To avoid this situation, we have several options including: reduce size of the adapter, apply weight decay to the adapter during the finetuning, and apply layer regularization to the adapter during the finetuning.
+The first option is not compatible if we consider the probable growth of in-domain data.
+The second and third choices are preferable to regularize a large adapter which will scale if the in-domain corpora extends.
+
+Weight decay applies a penalization on weights of the adapters.
+\begin{equation}
+\begin{split}
+\bar{L} & = \mathop{\sum}_{x,y} (log(P(y|x)) \\
+		  & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{ADAP^{(i)}}})
+\end{split}
+\end{equation}
+Layer regularization applies a penalization on the output of the adapters.
+\begin{equation}
+\begin{split}
+\bar{L} & = \mathop{\sum}_{x,y} (log(P(y|x)) \\
+		  & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{ADAP^{(i)}(h_i(x,y))})
+\end{split}
+\end{equation}
+
+\subsection{Multi-task training\label{ssec:multitask}}
+In the experiments of \cite{Caruana97multitask}, the authors proposed multi-task network in which a ground layer is shared between tasks and is followed by task-related layer in the next level.
+The network is then trained from scratch by fueling task-related example in a batch and updating all parameters including shared network's parameters and task-related network's parameters at the same time.
+The multi-domain context is similar to multi-task context if we consider a domain as a task.
+Therefore, we can also use multi-task training to optimize the parameters of NMT model and the parameters of the residual adapters at once from beginning.
+To assess the efficacy of multi-task training in multi-domain context, we would like to compare the finetuning approach with multi-task training approach.
+
+\subsection{Word-Level Adaptive Domain Adaptation \label{ssec:wada}}
 \mpTodo{Formalizing problem, network design, training algorithm}
+Residual adapters allow us to adapt NMT model to a domain without changing pretrained parameters.
+However, we still have to decide in which domain a sentence is translated to chose a suitable adapter.
+Therefore, we face to another problem which is the error of domain prediction.
+Choosing wrong adapter to translate a sentence can be dramatically bad because a specialized model usually has bad performance in out-of-domain examples according to well-known Catastrophic forgetting \cite{McCloskey89catastrophic}.
+However, we still have another option that is to use pretrained generic model.
+Therefore, we could aim to achieve a performance at least as good as the generic model.
+We would like to present an adaptive version of residual adapter below.
+\\
+We consider a generic encoder-decoder architecture for a multi-domain sequence to sequence learning.
+Let $h \in \R^d$ the output of the encoder, \cite{bapna19simple} proposes to use an adapted representation for domain $k$ defined as $h' = h + a_k(h)$.
+This means that all words in all sentences in domain $k$ will use the adapter module represented by $a_k$.
+In a multilayer version with highway connections (see Figure~\ref{fig:hrl-architecture}), $a_k(h) = \sum_{l} a_{kl}(h^l)$ where $h^l$ is the distributed representation of the sequence at the $l^{th}$ layer.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[scale=0.3]{fig/highway_residual}
+  \caption{Highway multi-domain network with residual layers}
+\label{fig:hrl-architecture}
+\end{figure}
+ A "gated" version uses $h' = h + a_k(h) \odot{} z_k(h)$\footnote{Or $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure~\ref{fig:hrl-architecture}} where $z_k$ is a function of $h$ taking values in $[0,1]$.
+More precisely, $h$ and $h'$ are sequences of context vectors and the combination is performed element-wise, yielding:
+\begin{equation}
+   \forall t \in [1 \dots{} T], h'(w_t) = h(w_t) + a_k(h(w_t)) \odot{} z_k(h(w_t)). \label{eq:gated-residual}
+\end{equation}
+In Word Level Adaptive Domain Adaptation (WADA), $z_k(h(w))$ is designed to reflect the "topicality" of  word $w$ is in domain $k$: the more likely $w$ is in domain $k$, the larger $z_k(h(w))$ should be.
+Word-level adaptive domain adaptation aims to reduce the variation caused by $a_k$ to the adapted context vector $h'$ (compared to $h$) for words that are not typical of domain $k$.
+By reusing the learned generic representation for non-topical words, we can bound the risk of poor predictions in case of out-of-domain words by the risk of the learned generic model.
+\subsubsection*{The training process \label{sssec:train}}
+The training process comprises three main steps:
+\begin{itemize}
+	\item Pretraining a generic model with a mixed corpora comprising data from multiple domains.
+	\item Training a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen.
+This model computes the posterior domain probability $p(k|h(w_t))$ for each word $w_t$ based on the representation computed by the last layer.
+	\item Training the parameters of adapters with in-domain data separately for each domain while freezing the parameters of the generic model and of the domain classifiers.
+\end{itemize}
+During inference, $z_k$ is used to regulate the strength of the adapter module as suggested in equation~\ref{eq:gated-residual}.
+\section{Experimental settings \label{sec:exp}}
+\subsection{Data and metrics \label{ssec:corpora}}
+We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
+Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
+Corpus statistics in Table~\ref{tab:Corpora}.
+Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools.
+To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.%
 
-\section{Experimental settings \label{sec:exp}}
+\begin{table*}[htbp]
+  \centering
+  \begin{tabular}{ |lllllll|} %*{4}{|r|}}
+    \hline
+    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
+    \domain{med} & \domain{law} & \domain{bank} & \domain{it} & \domain{talk} & \domain{rel} & \domain{news} \\
+    \hline
+    2609 (0.68) & 190 (0.05)  & 501 (0.13) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
+    \hline
+  \end{tabular}
+\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain).
+\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.}
+\label{tab:Corpora-en-fr}
+\end{table*}
 
-\subsection{Data and metrics \label{ssec:corpora}}
-\mpTodo{English - French, English - German}
-\subsection{Baselines \label{ssec:baseline}}
-\mpTodo{\system{FT-block}; \system{FT-full}, \system{FT-full Brian Thompson}, \system{DC}, \system{Domain Mixing}}
+We also report a part of experiments in language pair English, German.
+We use available corpora in the robustness task of WMT20 \footnote{\url{http://www.statmt.org/wmt20/robustness.html}} including: European Central Bank corpus (\domain{bank}),  European Economic and Social Committee corpus (\domain{economic}), European Medicines Agency corpus (\domain{med}) \footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Press Release Database of European Commission corpus, News Commentary v15 corpus, Common Crawl corpus (\domain{news}), Europarl v10 (\domain{gouv}) Tilde MODEL - czechtourism (\domain{tourism})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Paracrawl and Wikipedia Matrix (\domain{others}) \footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}.
+We also build a joint BPE vocabulary of size 33K for both languages.
 
-\subsection{Proposed systems \label{ssec:wada}}
+\begin{table*}[htbp]
+  \centering
+  \begin{tabular}{ |lllllll|} %*{4}{|r|}}
+    \hline
+    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
+    \domain{bank} & \domain{economic} & \domain{med} & \domain{gouv} & \domain{news} & \domain{tourism} & \domain{others} \\
+    \hline
+    4(0.00022) & 2857(0.15) & 347(0.018) & 1828(0.095) & 3696(0.19) & 7(0.00039) & 10472771(0.54) \\
+    \hline
+  \end{tabular}
+\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture.
+\domain{news} is the largest domain, containing almost 19\% of the sentences, while \domain{bank} is the smallest, with only 0.02\% of the data.}
+\label{tab:Corpora-en-de}
+\end{table*}
+
+We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
+Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
+
+\subsection{Ablatation study on positions and number of residual adapters}
+
+In table \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} we report the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} in language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} in language pair En-De.
+In most cases, the performance increase with respect to the number of residual adapters used in architecture.
+Setting \system{FT-block} using residual adapter for all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels that outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only on level.
+However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.
+\begin{table*}
+  \centering
+  \fyDone{Fix column size}
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+%     &&&&&& \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
+    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} & \SB{53.8} \\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2} & \SB{53.3} \\ 
+   \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\ 
+   \system{FT-Block$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 \\
+   \system{FT-Block$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.39 & 88.73 & 52 \\
+   \system{FT-Block$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 \\
+   \system{FT-Block$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 \\
+     \hline
+  \end{tabular}
+  \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
+  \label{tab:performance-en-fr}
+\end{table*}
+
+\begin{table*}
+  \centering
+  \fyDone{Fix column size}
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+%     &&&&&& \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{gouv}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tourism}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    \system{Mixed-Nat}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
+    \system{FT-Full}       & 29.8 & 30.97 & \SB{19.81} & \SB{53.43} & \SB{49.98} & 20.84 & 34.14 \\
+   \system{FT-Block}     & 29.65 &  & \SB{19.21} & \SB{48.99} & 47.22 & 20.63 & 33.14 \\ 
+   \system{FT-HW-Block}   & 29.54 & 30.42 & 18.59 & 50.78 & 47.13 & 20.51 & 32.83 \\ 
+   \system{FT-Block$_{(6)}$}     & 29.47 & 30.39 & 18.13 & 49.14 & 46.95 & 20.45 & 32.42 \\
+   \system{FT-Block$_{(4)}$}     & 29.69 & 30.4 & 18.07 & 49.61 & 47.05 & 20.64 & 32.58 \\
+   \system{FT-Block$_{(2)}$}   & 29.64 & 30.4 & 18.29 & 49.41 & 46.71 & 20.59 & 32.51  \\
+   \system{FT-Block$_{(2,4,6)}$}  & 29.68  & 30.55 & 18.85 & 49.57 & 47.09 & 20.63 &  32.73  \\
+     \hline
+  \end{tabular}
+  \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
+  \label{tab:performance-en-de}
+\end{table*}
+
+\subsection{Regularization of finetuning}
+
+\begin{table*}
+  \centering
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+%     &&&&&& \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2} & \SB{53.3} \\ 
+   \system{FT-Block-WD}     & 37.18 & \SB{55.99} & 52.93 & 33.36 & \SB{46.03} & \SB{90.65} & \SB{52.7} \\
+   \system{FT-Block-LR}     & 37.45 & \SB{56.09} & 51.84 & 33.29 & \SB{45.02} & \SB{89.7} & \SB{52.2} \\
+     \hline
+  \end{tabular}
+  \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
+  \label{tab:performance-en-fr}
+\end{table*}
+
+\begin{table*}
+  \centering
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+%     &&&&&& \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{gouv}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tourism}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    \system{Mixed-Nat}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
+   \system{FT-Block}     & 29.65 &  & \SB{19.21} & \SB{48.99} & 47.22 & 20.63 & 33.14 \\
+   \system{FT-Block-WD}     & 29.68 & 30.77 & 20.42 & 50.19 & 47.68 & 20.64 & 33.23 \\
+   \system{FT-Block-LR}     & 29.65 &  & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
+     \hline
+  \end{tabular}
+  \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
+  \label{tab:performance-en-de}
+\end{table*}
+
+\subsection{Multi-task learning}
+
+\subsection{Word-Level Adaptive Domain Adaptation \label{sec:wada}}
 

### 2020-08-12 2020/wmt_20/emnlp2020.tex
Revised abstract, begining part of the intro

@@
 \mpTodo{correcting abstract}
+
+****
+
+\fyTodo{Citation-free abstract}
+Domain adaptation is an old and vexing problem for machine translation systems.
+The most common approach and successful to supervised adaptation is to fine-tune a baseline system with in-domain parallel data.
+Standard fine-tuning however modifies all the network parameters, which makes this approach computationally costly and prone to overfitting.
+A recent, lightweigth approach, instead augments a baseline model with supplementary (small) adapter layers, keeping the rest of the mode unchanged.
+This has the additional merit to leave the baseline model intact, and adaptable to multiple domains.
+In this paper, we conduct a thorough analysis of the adapter model in the context of a multidomain machine translation task.
+We contrast multiple implementations of this idea on two language pairs.
+Our main conclusions is ..\fyTodo{abstract to be continued}
+
 \end{abstract}
@@
 \section{Introduction } \label{sec:intro}
-\mpTodo{write introduction}
-Neural Machine Translation(NMT) has achieved significant performance in many language pairs \cite{Bahdanau15learning,Vaswani17attention,Kalchbrenner13recurrent,Sutskever14sequence}.
-However, as many deep learning model, NMT model need to be trained with a sufficiently large amount of data to reach a high performance; therefore the quality of translation of NMT model is still limited in low-resourced language, and low-resourced domain \cite{duh13adaptation,koehn17six,zoph16transfer}.
-While many approaches have been proposed To improve the quality of NMT model in low-resourced domain (\cite{Chu18asurvey}), full finetuning of model parameters remains the dominant approach for adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
-One first trains an NMT model with a large parallel corpora which is always a mix of many available parallel corpora.
-Given a domain of interest and a parallel corpora of the domain, one could adapt the pretrained model by continue training the model with the in-domain corpora.
-It is a conjecture that pretrained model is a better initialization of training NMT model with small corpora than the random initialization.
-Indeed, studies in transfer learning in NMT such as \cite{artetxe20cross,aji20neural} confirms this claim by extensive experiments.
-Finetuning by training all the parameters (\system{FT-full}) of the model improves significantly the quality of the NMT model on a specific domain on the one hand, loses the translation quality on other domains on the other hand.
-This phenomenon is referred to Catastrophic Forgetting in training neural networks \cite{McCloskey89catastrophic}.
-Therefore, full finetuned model is usefull only to the target domain.
-As the number of domains grow, training and maintaining finetuned models consumes a lot of resource.
-\cite{Vilar18learning},\citep{bapna19simple} propose a simple method to preserve the value of pretrained model by plugging small adapters to hidden layers which will be trained only with the in-domain data while the pretrained model is frozen.
-Because these additional adapters are very small compared to the size of the model, we could reduce significantly the cost of training and maintaining finetuned models.
+\mpTodo{write introduction} \fyTodo{Citations in chronological order}\fyTodo{Split long sentences}
+Owing to multiple architectural improvements, Neural Machine Translation(NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
+However, as many deep learning model, NMT models need to be trained with sufficiently large amounts of data to reach their best performance.
+Therefore the quality of translation of NMT model is still limited in low-resource language and low-resourced domain conditions \cite{duh13adaptation, zoph16transfer,koehn17six}.
+While many approaches have been proposed to improve the quality of NMT models in low-resource domains (see the recent survey of \citet{Chu18asurvey}), full fine-tuning of a generic baseline model remains the dominant supervised approach when adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
 
-Because of the extensively demonstrated efficacy of the adapters, we would like to extend the use of this architecture to improve NMT model in several particular situations.
+Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pretrained model by resuming training with the sole in-domain corpus.
+It is a conjecture that pretrained model constitute a better initialization a the random initialization, especially when the adaptation data is scarce.
+Indeed, studies in transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim by extensive experiments.
+Fine-tuning by training all the parameters (\system{FT-full}) of the baseline model usally improves significantly the quality of the NMT for the chosen domain.
+However, it also implies losses in translation quality for other domains, a phenomenon referred to as Catastrophic Forgetting in the neural networks literature \cite{McCloskey89catastrophic}.
+Therefore, a full fine-tuned model is only useful to one target domain.
+As the number of domains to handle grows, training and maintaining such models can consume a lot of resource.\fyTodo{Fix this sentence}
+
+\cite{Vilar18learning,bapna19simple} have recently proposed a simple and lightweight method to preserve the value of pretrained models by plugging small adapters to hidden layers.
+These adapters are trained only with the in-domain data, keeping the pretrained model frozen.
+Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering performance that are on par with full fine-tuning.
+
+In this paper, we would like to extend this architecture to improve NMT in several specific situations.
 In translation task, there are several factors that challenge NMT model, including: noise, topical text.

### 2020-08-12 2020/wmt_20/emnlp2020.tex
add experiments

--- 2020-08-11 2020/wmt_20/emnlp2020.tex
@@
 Among the approaches for adapting a pretrained NMT model to a specific domain, finetuning is the dominant method.
-Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{bapna19simple, Vilar18learning}.
+Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{Bapna19simple, Vilar18learning}.
 The second approach has several advantages including preserving the pretrained model and the legerity of the adapters.
 However, the behavior of these adapters has not been further studied.
-The objective of this paper is to study the behavior of the adapters, then to extend the application of the adapters to improve the robustness of the model before domain shift and noisy data.
-\mpTodo{correcting abstract}
+The objective of this paper is to give an ablation study on the use of residual adapter in the domain adaptation problem.
 \end{abstract}
-
 \section{Introduction } \label{sec:intro}
-\mpTodo{write introduction}
 Neural Machine Translation(NMT) has achieved significant performance in many language pairs \cite{Bahdanau15learning,Vaswani17attention,Kalchbrenner13recurrent,Sutskever14sequence}.
-However, as many deep learning model, NMT model need to be trained with a sufficiently large amount of data to reach a high performance; therefore the quality of translation of NMT model is still limited in low-resourced language, and low-resourced domain \cite{duh13adaptation,koehn17six,zoph16transfer}.
-While many approaches have been proposed To improve the quality of NMT model in low-resourced domain (\cite{Chu18asurvey}), full finetuning of model parameters remains the dominant approach for adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
-One first trains an NMT model with a large parallel corpora which is always a mix of many available parallel corpora.
-Given a domain of interest and a parallel corpora of the domain, one could adapt the pretrained model by continue training the model with the in-domain corpora.
+However, as many deep learning models, NMT model need to be trained with a sufficiently large amount of data to reach a high performance; therefore the quality of the translation of NMT model is still limited in low-resourced language, and low-resourced domain \cite{duh13adaptation,koehn17six,zoph16transfer}.
+While many approaches have been proposed To improve the quality of NMT model in low-resourced domains (\cite{Chu18asurvey}), full finetuning of model parameters remains the dominant approach for adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
+One first trains an NMT model with a large parallel corpus which is always a mix of many available parallel corpora.
+Given a domain of interest and a parallel corpus of the domain, one could adapt the pretrained model by continue training the model with the in-domain corpora.
 It is a conjecture that pretrained model is a better initialization of training NMT model with small corpora than the random initialization.
-Indeed, studies in transfer learning in NMT such as \cite{artetxe20cross,aji20neural} confirms this claim by extensive experiments.
-Finetuning by training all the parameters (\system{FT-full}) of the model improves significantly the quality of the NMT model on a specific domain on the one hand, loses the translation quality on other domains on the other hand.
-This phenomenon is referred to Catastrophic Forgetting in training neural networks \cite{McCloskey89catastrophic}.
-Therefore, full finetuned model is usefull only to the target domain.
-As the number of domains grow, training and maintaining finetuned models consumes a lot of resource.
-\cite{Vilar18learning},\citep{bapna19simple} propose a simple method to preserve the value of pretrained model by plugging small adapters to hidden layers which will be trained only with the in-domain data while the pretrained model is frozen.
+Indeed, studies in transfer learning in NMT such as \cite{artetxe20cross,aji20neural} confirm this claim by extensive experiments.
+Finetuning by training all the parameters (\system{FT-full}) of the model improves significantly the quality of the NMT model on a specific domain, on the one hand, loses the translation quality on other domains on the other hand.
+This phenomenon is referred to as Catastrophic Forgetting in training neural networks \cite{McCloskey89catastrophic}.
+Therefore, the full finetuned model is useful only to the target domain.
+As the number of domains grows, training and maintaining finetuned models consume a lot of resources.
+\cite{Vilar18learning},\citep{Bapna19simple} propose a simple method to preserve the value of pretrained model by plugging small adapters to hidden layers that will be trained only with the in-domain data while the pretrained model is frozen.
 Because these additional adapters are very small compared to the size of the model, we could reduce significantly the cost of training and maintaining finetuned models.
-
-Because of the extensively demonstrated efficacy of the adapters, we would like to extend the use of this architecture to improve NMT model in several particular situations.
-In translation task, there are several factors that challenge NMT model, including: noise, topical text.
-Residual adapters allow to adapt NMT model to any specific domain in a computationally economical way.
-Could we adapt NMT model to noisy text and topical text at once?
-While residual adapters are so good to adapt to a specific domain, it still loses dramatically performance in the previously training domains as we show in our experiments.
+\\
+Because of the extensively demonstrated efficacy of the adapters, we would like to study further the performance of residual adapters in large range of variants.
+Besides, while residual adapters are so good to adapt to a specific domain, it still loses dramatically performance in the previously training domains as we show in our experiments.
 Therefore, we have to decide whether to use residual adapters manually, i.e, we have to know to which domain the text belongs a priori.
-Therefore, we would like to fuse a domain classifier to the architecture in order to weight the contribution of the adapters with respect to the topic-relatedness of the text.
-
+Therefore, we would like the model to automatically scale the contribution of the adapters with respect to the topic-relatedness of the text.
+\\
 Our contribution is an ablation study on the use of residual adapter in NMT domain adaptation by answering 3 following question:
 \begin{itemize}
-	\item How the number and position of adapters affect the performance.
-	\item How to train the residual adapter.
-	\item How to be robust to out-of-domain examples.
+	\item How the number and position of residual adapters affect the performance.
+	\item How to train residual adapters.
+	\item How to render the combination of NMT model and residual adapters more robust to out-of-domain examples.
 \end{itemize}
-
 \section{Residual adapters \label{sec:res}}
@@
 \subsection{Architecture \label{ssec:architecture}}
-Residual adapters modifies context vectors of each layer by applying several transformations as follow:
+Residual adapters modify context vectors of each layer by applying several transformations as follow:
 $$ h^{i}_1 = \mathbf{W}_{d_{model}}^{d_{bottleneck}}h^{i} + b_{1}$$
@@
 
-The "adapted" context vector $\bar{h}^i$ goes to the $i-th$ self-attention layer.
-There are 12 residual adapters corresponding to total 12 attention layers of the encoder and the decoder (in the decoder, the stack of self-attention layer and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter).
+The "adapted" context vector $\bar{h}^i$ goes to the $(i+1)-th$ self-attention layer.
+12 residual adapters are corresponding to total 12 attention layers of the encoder and the decoder (in the decoder, the stack of self-attention layer and cross encoder-decoder attention only count as one attention layer and serve only one residual adapter).
 \subsection{Ablation study on position and number of residual adapters \label{ssec:ablatation}}
-In this section, we would like to analyze the impact of position and number of residual adapter involved in the adapted model.
+In this section, we would like to analyze the impact of position and number of residual adapters involved in the adapted model.
 The importance of hidden layers varies with respect to their position in NMT model.
@@
 We consider a generic encoder-decoder architecture for a multi-domain sequence to sequence learning.
-Let $h \in \R^d$ the output of the encoder, \cite{bapna19simple} proposes to use an adapted representation for domain $k$ defined as $h' = h + a_k(h)$.
+Let $h \in \R^d$ the output of the encoder, \cite{Bapna19simple} proposes to use an adapted representation for domain $k$ defined as $h' = h + a_k(h)$.
 This means that all words in all sentences in domain $k$ will use the adapter module represented by $a_k$.
 In a multilayer version with highway connections (see Figure~\ref{fig:hrl-architecture}), $a_k(h) = \sum_{l} a_{kl}(h^l)$ where $h^l$ is the distributed representation of the sequence at the $l^{th}$ layer.
-\begin{figure}[h!]
+\begin{figure}[htbp]
   \centering
@@
 By reusing the learned generic representation for non-topical words, we can bound the risk of poor predictions in case of out-of-domain words by the risk of the learned generic model.
-\subsubsection*{The training process \label{sssec:train}}
+\subsubsection{The training process \label{sssec:train}}
 The training process comprises three main steps:
@@
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
-Corpus statistics in Table~\ref{tab:Corpora}.
+Corpus statistics in Table~\ref{tab:Corpora-en-fr}.
 Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools.
@@
 We also build a joint BPE vocabulary of size 33K for both languages.
+The statistics of experimental En-De corpora are reported in table \ref{tab:Corpora-en-de}
 
@@
 We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
+\subsection*{Baseline model}
+Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
 
+\begin{itemize}
+\item a generic model trained on a concatenation of all corpora, denoted \system{Mixed}
+\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing significantly in 3 consecutive epochs.
+We again contrast two versions: full fine-tuning (\system{FT-Full}), which update all the parameters of the initial generic model \system{Mixed}; and the variant of \cite{Bapna19simple} (\system{FT-Block}).
+\end{itemize}
+
+For all models, we set the embeddings size and the hidden layers size to~512.
+Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
+The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
+% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
+% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
+Training use a batch size of~12288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
 \subsection{Ablatation study on positions and number of residual adapters}
 
-In table \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} we report the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} in language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} in language pair En-De.
+In table \ref{tab:performance-en-fr} and \ref{tab:performance-en-de} we report the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} in language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} in language pair En-De.
 In most cases, the performance increase with respect to the number of residual adapters used in architecture.
@@
     \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} & \SB{53.8} \\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2} & \SB{53.3} \\ 
+    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
+   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
    \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\ 
@@
     \system{Mixed-Nat}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
-    \system{FT-Full}       & 29.8 & 30.97 & \SB{19.81} & \SB{53.43} & \SB{49.98} & 20.84 & 34.14 \\
-   \system{FT-Block}     & 29.65 &  & \SB{19.21} & \SB{48.99} & 47.22 & 20.63 & 33.14 \\ 
+    \system{FT-Full}       & 29.8 & 30.97 & 19.81 & 53.43 & 49.98 & 20.84 & 34.14 \\
+   \system{FT-Block}     & 29.65 &  & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
    \system{FT-HW-Block}   & 29.54 & 30.42 & 18.59 & 50.78 & 47.13 & 20.51 & 32.83 \\ 
@@
 \subsection{Regularization of finetuning}
-
+In the translation task En-De, we find that domain \domain{tourism}, \domain{ecb} are extremely small with population percentage only 0.039\% and 0.022\% respectively.
+Finetuning on these domain can lead to overfitting problem.
+We assess two regularization techniques, that help avoiding overfitting, including: weight-decay and layer regularization.
+In table \ref{tab:performance-en-de-reg}, weight-decay results in system \system{FT-Block-WD} outperforms \system{FT-Block}.
 \begin{table*}
@@
     \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2} & \SB{53.3} \\ 
-   \system{FT-Block-WD}     & 37.18 & \SB{55.99} & 52.93 & 33.36 & \SB{46.03} & \SB{90.65} & \SB{52.7} \\
-   \system{FT-Block-LR}     & 37.45 & \SB{56.09} & 51.84 & 33.29 & \SB{45.02} & \SB{89.7} & \SB{52.2} \\
+   \system{FT-Block}     & 37.3	& 57.93 &	53.91 &	33.79 &	46.69 &	90.17 & 42.5 \\ 
+   \system{FT-Block-WD}     & 37.18 & 55.99 & 52.93 & 33.36 & 46.03 & 90.65 & 52.7 \\
+   \system{FT-Block-LR}     & 37.45 & 56.09 & 51.84 & 33.29 & 45.02 & 89.7 & 52.2 \\
      \hline
@@
   \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
-  \label{tab:performance-en-fr}
+  \label{tab:performance-en-fr-reg}
 \end{table*}
@@
     \system{Mixed-Nat}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
-   \system{FT-Block}     & 29.65 &  & \SB{19.21} & \SB{48.99} & 47.22 & 20.63 & 33.14 \\
+   \system{FT-Block}      \\
    \system{FT-Block-WD}     & 29.68 & 30.77 & 20.42 & 50.19 & 47.68 & 20.64 & 33.23 \\
-   \system{FT-Block-LR}     & 29.65 &  & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
+   \system{FT-Block-LR}     & 29.65 & 30.45 & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
      \hline
@@
   \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
-  \label{tab:performance-en-de}
+  \label{tab:performance-en-de-reg}
+\end{table*}
+\subsection{Multi-task learning}
+To assess the efficacy of multi-task learning in training both NMT model and residual adapters at once from beginning, we use several contrast methods in multi-domain learning including:
+\begin{itemize}
+\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDR});
+\item the three proposals of \newcite{Britz17mixing}.
+\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
+Training uses reference tags and inference is performed with predicted tags, just like for regular target words.
+\system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
+These methods thus only use domain tags in training.
+\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connections make it possible to by-pass this module.
+  % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
+Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
+\end{itemize}
+
+\begin{table*}
+  \centering
+  \fyDone{Fix column size}
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+%     &&&&&& \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
+    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & \SB{42.7} & \SB{53.8} \\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & \SB{42.3} & \SB{53.3} \\ \hline % & 20.5\\ \hline
+ %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
+    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.6 & 50.1    \\%    & 21.8 \\
+    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   \\% & \SW{21.7} \\
+    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} &40.9 & 49.8          \\% & 22.1 \\ 
+    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.0 & 49.7     \\% &  23.4 \\
+    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 38.1 & 43.4 \\ % & 22.6\\
+    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 38.9 & 47.5 \\% & 23.3 \\
+    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & \SB{51.9} \\%  & \SW{21.2} \\
+    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
+%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3&\\
+     \hline
+  \end{tabular}
+  \caption{Translation performance of various MDMT systems. We report BLEU scores for each domain, as well as domain-weighted (w\domain{avg})}
+  \label{tab:performance-multi}
 \end{table*}
 
-\subsection{Multi-task learning}
-
 \subsection{Word-Level Adaptive Domain Adaptation \label{sec:wada}}
-
+\mpTodo{wada}
 \section{Related Work \label{sec:related}}
-
+\mpTodo{related work}
 \section{Discussion \label{sec:discussion}}
-
+\mpTodo{discussion}
 \section*{Acknowledgments}
@@
 \label{sec:appendix}
-jkk
+\mpTodo{appendix}
 \section{Supplemental Material}

### 2020-08-12 2020/wmt_20/emnlp2020.tex

@@
 The objective of this paper is to give an ablation study on the use of residual adapter in the domain adaptation problem.
+\mpTodo{correcting abstract}
+
+****
+
+\fyTodo{Citation-free abstract}
+Domain adaptation is an old and vexing problem for machine translation systems.
+The most common approach and successful to supervised adaptation is to fine-tune a baseline system with in-domain parallel data.
+Standard fine-tuning however modifies all the network parameters, which makes this approach computationally costly and prone to overfitting.
+A recent, lightweigth approach, instead augments a baseline model with supplementary (small) adapter layers, keeping the rest of the mode unchanged.
+This has the additional merit to leave the baseline model intact, and adaptable to multiple domains.
+In this paper, we conduct a thorough analysis of the adapter model in the context of a multidomain machine translation task.
+We contrast multiple implementations of this idea on two language pairs.
+Our main conclusions is ..\fyTodo{abstract to be continued}
+
 \end{abstract}
 \section{Introduction } \label{sec:intro}
-Neural Machine Translation(NMT) has achieved significant performance in many language pairs \cite{Bahdanau15learning,Vaswani17attention,Kalchbrenner13recurrent,Sutskever14sequence}.
-However, as many deep learning models, NMT model need to be trained with a sufficiently large amount of data to reach a high performance; therefore the quality of the translation of NMT model is still limited in low-resourced language, and low-resourced domain \cite{duh13adaptation,koehn17six,zoph16transfer}.
-While many approaches have been proposed To improve the quality of NMT model in low-resourced domains (\cite{Chu18asurvey}), full finetuning of model parameters remains the dominant approach for adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
-One first trains an NMT model with a large parallel corpus which is always a mix of many available parallel corpora.
-Given a domain of interest and a parallel corpus of the domain, one could adapt the pretrained model by continue training the model with the in-domain corpora.
-It is a conjecture that pretrained model is a better initialization of training NMT model with small corpora than the random initialization.
-Indeed, studies in transfer learning in NMT such as \cite{artetxe20cross,aji20neural} confirm this claim by extensive experiments.
-Finetuning by training all the parameters (\system{FT-full}) of the model improves significantly the quality of the NMT model on a specific domain, on the one hand, loses the translation quality on other domains on the other hand.
-This phenomenon is referred to as Catastrophic Forgetting in training neural networks \cite{McCloskey89catastrophic}.
-Therefore, the full finetuned model is useful only to the target domain.
-As the number of domains grows, training and maintaining finetuned models consume a lot of resources.
-\cite{Vilar18learning},\citep{Bapna19simple} propose a simple method to preserve the value of pretrained model by plugging small adapters to hidden layers that will be trained only with the in-domain data while the pretrained model is frozen.
-Because these additional adapters are very small compared to the size of the model, we could reduce significantly the cost of training and maintaining finetuned models.
-\\
-Because of the extensively demonstrated efficacy of the adapters, we would like to study further the performance of residual adapters in large range of variants.
-Besides, while residual adapters are so good to adapt to a specific domain, it still loses dramatically performance in the previously training domains as we show in our experiments.
+\mpTodo{write introduction} \fyTodo{Citations in chronological order}\fyTodo{Split long sentences}
+Owing to multiple architectural improvements, Neural Machine Translation(NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
+However, as many deep learning models, NMT models need to be trained with sufficiently large amounts of data to reach their best performance.
+Therefore the quality of the translation of NMT model is still limited in low-resource language and low-resourced domain conditions \cite{duh13adaptation, zoph16transfer,koehn17six}.
+While many approaches have been proposed to improve the quality of NMT models in low-resource domains (see the recent survey of \citet{Chu18asurvey}), full fine-tuning of a generic baseline model remains the dominant supervised approach when adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
+
+Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pretrained model by resuming training with the sole in-domain corpus.
+It is a conjecture that pretrained model constitutes a better initialization a the random initialization, especially when the adaptation data is scarce.
+Indeed, studies in transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim by extensive experiments.
+Fine-tuning by training all the parameters (\system{FT-full}) of the baseline model usually improves significantly the quality of the NMT for the chosen domain.
+However, it also implies losses in translation quality for other domains, a phenomenon referred to as Catastrophic Forgetting in the neural networks literature \cite{McCloskey89catastrophic}.
+Therefore, a full fine-tuned model is only useful to one target domain.
+As the number of domains to handle grows, training and maintaining such models can consume a lot of resources .\fyTodo{Fix this sentence}
+
+\cite{Vilar18learning,bapna19simple} have recently proposed a simple and lightweight method to preserve the value of pretrained models by plugging small adapters to hidden layers.
+These adapters are trained only with the in-domain data, keeping the pretrained model frozen.
+Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering a performance that is on par with full fine-tuning.
+
+In this paper, we would like to extend this architecture to improve NMT in several specific situations.
+In the translation task, several factors challenge NMT model, including noise, topical text.
+Residual adapters allow to adapt NMT model to any specific domain in a computationally economical way.
+Could we adapt NMT model to noisy text and topical text at once?
+While residual adapters are so good to adapt to a specific domain, it's performance drops dramatically in the previous training domains as we show in our experiments.
 Therefore, we have to decide whether to use residual adapters manually, i.e, we have to know to which domain the text belongs a priori.
-Therefore, we would like the model to automatically scale the contribution of the adapters with respect to the topic-relatedness of the text.
-\\
+Therefore, we would like to fuse a domain classifier to the architecture to weight the contribution of the adapters with respect to the topic-relatedness of the text.
+
 Our contribution is an ablation study on the use of residual adapter in NMT domain adaptation by answering 3 following question:
 \begin{itemize}
-	\item How the number and position of residual adapters affect the performance.
-	\item How to train residual adapters.
+\item How the number and position of adapters affect the performance.
+\item How to train the residual adapters.
 \item How to render the combination of NMT model and residual adapters more robust to out-of-domain examples.
 \end{itemize}
+
 \section{Residual adapters \label{sec:res}}
@@
 However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.
-\begin{table*}
+\begin{table*}[htbp]
   \centering
@@
 
-\begin{table*}
+\begin{table*}[htbp]
   \centering
@@
 In table \ref{tab:performance-en-de-reg}, weight-decay results in system \system{FT-Block-WD} outperforms \system{FT-Block}.
-\begin{table*}
+\begin{table*}[htbp]
   \centering
@@
 \item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connections make it possible to by-pass this module.
-  % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
 Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
@@
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & \SB{42.7} & \SB{53.8} \\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & \SB{42.3} & \SB{53.3} \\ \hline % & 20.5\\ \hline
- %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.6 & 50.1    \\%    & 21.8 \\
-    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   \\% & \SW{21.7} \\
-    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} &40.9 & 49.8          \\% & 22.1 \\ 
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.0 & 49.7     \\% &  23.4 \\
-    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 38.1 & 43.4 \\ % & 22.6\\
-    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 38.9 & 47.5 \\% & 23.3 \\
-    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & \SB{51.9} \\%  & \SW{21.2} \\
-    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
-%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3&\\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\% & 23.5\\
+    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} \\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} \\ \hline 
+    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1    \\
+    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} & 49.9   \\
+    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 49.8          \\
+    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} & 49.7   \\
+    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 43.4 \\ 
+    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 47.5 \\
+    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} \\
+    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 33.7\\
      \hline
@@
 \mpTodo{related work}
+dsdsd
 \section{Discussion \label{sec:discussion}}

### 2020-08-12 2020/wmt_20/emnlp2020.tex
add wada

@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
+    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
     \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{gouv}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tourism}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
+    \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
     \system{FT-Full}       & 29.8 & 30.97 & 19.81 & 53.43 & 49.98 & 20.84 & 34.14 \\
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
-   \system{FT-Block}     & 37.3	& 57.93 &	53.91 &	33.79 &	46.69 &	90.17 & 42.5 \\ 
+    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
+   \system{FT-Block}     & 37.3	& 57.93 &	53.91 &	33.79 &	46.69 &	90.17 & 53.3 \\ 
    \system{FT-Block-WD}     & 37.18 & 55.99 & 52.93 & 33.36 & 46.03 & 90.65 & 52.7 \\
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{gouv}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tourism}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
+    \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
    \system{FT-Block}      \\
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\% & 23.5\\
+    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\% & 23.5\\
     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} \\
@@
     \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} \\
+    \system{MDL-HW-res}   & 37.38&	56.43&	52.13&	33.7&	44.81&	89.77&	52.4 \\ 
     \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 33.7\\
@@
   \end{tabular}
-  \caption{Translation performance of various MDMT systems. We report BLEU scores for each domain, as well as domain-weighted (w\domain{avg})}
+  \caption{Translation performance of various MDMT systems.}
   \label{tab:performance-multi}
@@
 \mpTodo{wada}
+To measure the robustness against out-of-domain examples, we perform translation with wrong domain information to trick the systems that need to know domain a priori.
+In the table \ref{tab:performance-random}, the column \domain{RND} stores the score of multi-domain model, that use domain information to translate, in the situation in which we assign random domain tag to each test sentence.
+The phenomenon Catastrophic Forgetting manifest clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
+Our proposed model \system{wada} maintains equivalent performance to generic model in \system{RND} thanks to application of adaptive weight on the output of residual adapter.
+\system{wada} avoid relying in the residual adapters when predicting examples that come from domain far from the domain of the adapters.
+
+\begin{table*}[htbp]
+  \centering
+  \fyDone{Fix column size}
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+%     &&&&&& \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{RND}} \\ \hline
+    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 & 49.4 \\
+    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} & 32.55 \\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} & 38.38 \\ \hline 
+   \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 & 27.13\\ 
+    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1 & 35.92   \\
+    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} & 49.9 & 34.91 \\
+    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 49.8          & 37.01 \\
+    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} & 31.78 \\
+    \system{wada}    & 37.98 &	57.46&	53.05&	33.55&	46.01&	90.07&	53.0&	48.96 \\
+     \hline
+  \end{tabular}
+  \caption{Translation performance of various MDMT systems}
+  \label{tab:performance-random}
+\end{table*}
+
 \section{Related Work \label{sec:related}}
 \mpTodo{related work}
-dsdsd
 \section{Discussion \label{sec:discussion}}

### 2020-08-12 2020/wmt_20/emnlp2020.tex
add missing score
Preliminary rewrite of the intro

--- 2020-08-12 2020/wmt_20/emnlp2020.tex
@@
 Among the approaches for adapting a pretrained NMT model to a specific domain, finetuning is the dominant method.
-Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{Bapna19simple, Vilar18learning}.
+Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{bapna19simple, Vilar18learning}.
 The second approach has several advantages including preserving the pretrained model and the legerity of the adapters.
 However, the behavior of these adapters has not been further studied.
-The objective of this paper is to give an ablation study on the use of residual adapter in the domain adaptation problem.
+The objective of this paper is to study the behavior of the adapters, then to extend the application of the adapters to improve the robustness of the model before domain shift and noisy data.
 \mpTodo{correcting abstract}
@@
 \end{abstract}
+
 \section{Introduction } \label{sec:intro}
@@
 Owing to multiple architectural improvements, Neural Machine Translation(NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
-However, as many deep learning models, NMT models need to be trained with sufficiently large amounts of data to reach their best performance.
-Therefore the quality of the translation of NMT model is still limited in low-resource language and low-resourced domain conditions \cite{duh13adaptation, zoph16transfer,koehn17six}.
+However, as many deep learning model, NMT models need to be trained with sufficiently large amounts of data to reach their best performance.
+Therefore the quality of translation of NMT model is still limited in low-resource language and low-resourced domain conditions \cite{duh13adaptation, zoph16transfer,koehn17six}.
 While many approaches have been proposed to improve the quality of NMT models in low-resource domains (see the recent survey of \citet{Chu18asurvey}), full fine-tuning of a generic baseline model remains the dominant supervised approach when adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
@@
 Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pretrained model by resuming training with the sole in-domain corpus.
-It is a conjecture that pretrained model constitutes a better initialization a the random initialization, especially when the adaptation data is scarce.
+It is a conjecture that pretrained model constitute a better initialization a the random initialization, especially when the adaptation data is scarce.
 Indeed, studies in transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim by extensive experiments.
-Fine-tuning by training all the parameters (\system{FT-full}) of the baseline model usually improves significantly the quality of the NMT for the chosen domain.
+Fine-tuning by training all the parameters (\system{FT-full}) of the baseline model usally improves significantly the quality of the NMT for the chosen domain.
 However, it also implies losses in translation quality for other domains, a phenomenon referred to as Catastrophic Forgetting in the neural networks literature \cite{McCloskey89catastrophic}.
 Therefore, a full fine-tuned model is only useful to one target domain.
-As the number of domains to handle grows, training and maintaining such models can consume a lot of resources .\fyTodo{Fix this sentence}
+As the number of domains to handle grows, training and maintaining such models can consume a lot of resource.\fyTodo{Fix this sentence}
 
@@
 These adapters are trained only with the in-domain data, keeping the pretrained model frozen.
-Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering a performance that is on par with full fine-tuning.
+Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering performance that are on par with full fine-tuning.
 
-In this paper, we would like to extend this architecture to improve NMT in several specific situations.
-In the translation task, several factors challenge NMT model, including noise, topical text.
-Residual adapters allow to adapt NMT model to any specific domain in a computationally economical way.
+In this paper, we would like to extend this architecture to improve NMT in several specific situations that still challenge automatic translation, such as translating noisy inputs or translating topical texts\fyTodo{??}.
+Residual adapters allow us to adapt NMT model to any specific domain in a computationally economical way.
 Could we adapt NMT model to noisy text and topical text at once?
-While residual adapters are so good to adapt to a specific domain, it's performance drops dramatically in the previous training domains as we show in our experiments.
+While residual adapters are good at adapting to one specific domain, their performance still dramatically decrease for the other domains seen in training, as we show in our experiments.
 Therefore, we have to decide whether to use residual adapters manually, i.e, we have to know to which domain the text belongs a priori.
-Therefore, we would like to fuse a domain classifier to the architecture to weight the contribution of the adapters with respect to the topic-relatedness of the text.
+Therefore, we would like to fuse a domain classifier to the architecture in order to weight the contribution of the adapters with respect to the topic-relatedness of the text.\fyTodo{Say differently: various implementations}
 
-Our contribution is an ablation study on the use of residual adapter in NMT domain adaptation by answering 3 following question:
+Our contribution is an thorough experimental study on the use of residual adapters in multidomain adapation. We notably try to answer the following three questions:
 \begin{itemize}
 \item How the number and position of adapters affect the performance.
-\item How to train the residual adapters.
-\item How to render the combination of NMT model and residual adapters more robust to out-of-domain examples.
+\item How to train the residual adapter.
+\item How to be robust to out-of-domain examples.
 \end{itemize}
+\fyTodo{One bit of a conclusion here}
 
@@
 \subsection{Architecture \label{ssec:architecture}}
-Residual adapters modify context vectors of each layer by applying several transformations as follow:
+Residual adapters modifies context vectors of each layer by applying several transformations as follow:
 $$ h^{i}_1 = \mathbf{W}_{d_{model}}^{d_{bottleneck}}h^{i} + b_{1}$$
@@
 
-The "adapted" context vector $\bar{h}^i$ goes to the $(i+1)-th$ self-attention layer.
-12 residual adapters are corresponding to total 12 attention layers of the encoder and the decoder (in the decoder, the stack of self-attention layer and cross encoder-decoder attention only count as one attention layer and serve only one residual adapter).
+The "adapted" context vector $\bar{h}^i$ goes to the $i-th$ self-attention layer.
+There are 12 residual adapters corresponding to total 12 attention layers of the encoder and the decoder (in the decoder, the stack of self-attention layer and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter).
 \subsection{Ablation study on position and number of residual adapters \label{ssec:ablatation}}
-In this section, we would like to analyze the impact of position and number of residual adapters involved in the adapted model.
+In this section, we would like to analyze the impact of position and number of residual adapter involved in the adapted model.
 The importance of hidden layers varies with respect to their position in NMT model.
@@
 We consider a generic encoder-decoder architecture for a multi-domain sequence to sequence learning.
-Let $h \in \R^d$ the output of the encoder, \cite{Bapna19simple} proposes to use an adapted representation for domain $k$ defined as $h' = h + a_k(h)$.
+Let $h \in \R^d$ the output of the encoder, \cite{bapna19simple} proposes to use an adapted representation for domain $k$ defined as $h' = h + a_k(h)$.
 This means that all words in all sentences in domain $k$ will use the adapter module represented by $a_k$.
 In a multilayer version with highway connections (see Figure~\ref{fig:hrl-architecture}), $a_k(h) = \sum_{l} a_{kl}(h^l)$ where $h^l$ is the distributed representation of the sequence at the $l^{th}$ layer.
-\begin{figure}[htbp]
+\begin{figure}[h!]
   \centering
@@
 By reusing the learned generic representation for non-topical words, we can bound the risk of poor predictions in case of out-of-domain words by the risk of the learned generic model.
-\subsubsection{The training process \label{sssec:train}}
+\subsubsection*{The training process \label{sssec:train}}
 The training process comprises three main steps:
@@
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
-Corpus statistics in Table~\ref{tab:Corpora-en-fr}.
+Corpus statistics in Table~\ref{tab:Corpora}.
 Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools.
@@
 We also build a joint BPE vocabulary of size 33K for both languages.
-The statistics of experimental En-De corpora are reported in table \ref{tab:Corpora-en-de}
 
@@
 We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
-\subsection*{Baseline model}
-Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
 
-\begin{itemize}
-\item a generic model trained on a concatenation of all corpora, denoted \system{Mixed}
-\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing significantly in 3 consecutive epochs.
-We again contrast two versions: full fine-tuning (\system{FT-Full}), which update all the parameters of the initial generic model \system{Mixed}; and the variant of \cite{Bapna19simple} (\system{FT-Block}).
-\end{itemize}
-
-For all models, we set the embeddings size and the hidden layers size to~512.
-Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
-The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
-% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
-% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training use a batch size of~12288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
 \subsection{Ablatation study on positions and number of residual adapters}
 
-In table \ref{tab:performance-en-fr} and \ref{tab:performance-en-de} we report the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} in language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} in language pair En-De.
+In table \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} we report the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} in language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} in language pair En-De.
 In most cases, the performance increase with respect to the number of residual adapters used in architecture.
@@
 However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.
-\begin{table*}[htbp]
+\begin{table*}
   \centering
@@
     \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
-    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
-   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
+    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} & \SB{53.8} \\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2} & \SB{53.3} \\ 
    \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\ 
@@
 
-\begin{table*}[htbp]
+\begin{table*}
   \centering
@@
     \system{Mixed-Nat}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
-    \system{FT-Full}       & 29.8 & 30.97 & 19.81 & 53.43 & 49.98 & 20.84 & 34.14 \\
-   \system{FT-Block}     & 29.65 &  & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
+    \system{FT-Full}       & 29.8 & 30.97 & \SB{19.81} & \SB{53.43} & \SB{49.98} & 20.84 & 34.14 \\
+   \system{FT-Block}     & 29.65 &  & \SB{19.21} & \SB{48.99} & 47.22 & 20.63 & 33.14 \\ 
    \system{FT-HW-Block}   & 29.54 & 30.42 & 18.59 & 50.78 & 47.13 & 20.51 & 32.83 \\ 
@@
 \subsection{Regularization of finetuning}
-In the translation task En-De, we find that domain \domain{tourism}, \domain{ecb} are extremely small with population percentage only 0.039\% and 0.022\% respectively.
-Finetuning on these domain can lead to overfitting problem.
-We assess two regularization techniques, that help avoiding overfitting, including: weight-decay and layer regularization.
-In table \ref{tab:performance-en-de-reg}, weight-decay results in system \system{FT-Block-WD} outperforms \system{FT-Block}.
-\begin{table*}[htbp]
+
+\begin{table*}
   \centering
@@
     \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
-   \system{FT-Block}     & 37.3	& 57.93 &	53.91 &	33.79 &	46.69 &	90.17 & 42.5 \\ 
-   \system{FT-Block-WD}     & 37.18 & 55.99 & 52.93 & 33.36 & 46.03 & 90.65 & 52.7 \\
-   \system{FT-Block-LR}     & 37.45 & 56.09 & 51.84 & 33.29 & 45.02 & 89.7 & 52.2 \\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2} & \SB{53.3} \\ 
+   \system{FT-Block-WD}     & 37.18 & \SB{55.99} & 52.93 & 33.36 & \SB{46.03} & \SB{90.65} & \SB{52.7} \\
+   \system{FT-Block-LR}     & 37.45 & \SB{56.09} & 51.84 & 33.29 & \SB{45.02} & \SB{89.7} & \SB{52.2} \\
      \hline
@@
   \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
-  \label{tab:performance-en-fr-reg}
+  \label{tab:performance-en-fr}
 \end{table*}
 
-\begin{table*}[htbp]
+\begin{table*}
   \centering
@@
     \system{Mixed-Nat}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
-   \system{FT-Block}      \\
+   \system{FT-Block}     & 29.65 &  & \SB{19.21} & \SB{48.99} & 47.22 & 20.63 & 33.14 \\
    \system{FT-Block-WD}     & 29.68 & 30.77 & 20.42 & 50.19 & 47.68 & 20.64 & 33.23 \\
-   \system{FT-Block-LR}     & 29.65 & 30.45 & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
+   \system{FT-Block-LR}     & 29.65 &  & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
      \hline
@@
   \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
-  \label{tab:performance-en-de-reg}
-\end{table*}
-\subsection{Multi-task learning}
-To assess the efficacy of multi-task learning in training both NMT model and residual adapters at once from beginning, we use several contrast methods in multi-domain learning including:
-\begin{itemize}
-\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDR});
-\item the three proposals of \newcite{Britz17mixing}.
-\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
-Training uses reference tags and inference is performed with predicted tags, just like for regular target words.
-\system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
-These methods thus only use domain tags in training.
-\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connections make it possible to by-pass this module.
-Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
-\end{itemize}
-
-\begin{table*}[htbp]
-  \centering
-  \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
-%     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\% & 23.5\\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} \\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} \\ \hline 
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1    \\
-    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} & 49.9   \\
-    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 49.8          \\
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} & 49.7   \\
-    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 43.4 \\ 
-    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 47.5 \\
-    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} \\
-    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 33.7\\
-     \hline
-  \end{tabular}
-  \caption{Translation performance of various MDMT systems. We report BLEU scores for each domain, as well as domain-weighted (w\domain{avg})}
-  \label{tab:performance-multi}
+  \label{tab:performance-en-de}
 \end{table*}
 
+\subsection{Multi-task learning}
+
 \subsection{Word-Level Adaptive Domain Adaptation \label{sec:wada}}
-\mpTodo{wada}
+
 \section{Related Work \label{sec:related}}
-\mpTodo{related work}
-dsdsd
+
 \section{Discussion \label{sec:discussion}}
-\mpTodo{discussion}
+
 \section*{Acknowledgments}
@@
 \label{sec:appendix}
-\mpTodo{appendix}
+jkk
 \section{Supplemental Material}

### 2020-08-12 2020/wmt_20/emnlp2020.tex
More fixes and todos

@@
 \mpTodo{write introduction} \fyTodo{Citations in chronological order}\fyTodo{Split long sentences}
-Owing to multiple architectural improvements, Neural Machine Translation(NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
+Owing to multiple architectural improvements, Neural Machine Translation(NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence,Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
 However, as many deep learning model, NMT models need to be trained with sufficiently large amounts of data to reach their best performance.
@@
 Fine-tuning by training all the parameters (\system{FT-full}) of the baseline model usally improves significantly the quality of the NMT for the chosen domain.
-However, it also implies losses in translation quality for other domains, a phenomenon referred to as Catastrophic Forgetting in the neural networks literature \cite{McCloskey89catastrophic}.
+However, it also implies losses in translation quality for other domains, a phenomenon referred to as ``catastrophic forgetting'' in the neural network literature \cite{McCloskey89catastrophic}.
 Therefore, a full fine-tuned model is only useful to one target domain.
@@
 \subsection{Architecture \label{ssec:architecture}}
-Residual adapters modifies context vectors of each layer by applying several transformations as follow:
-$$ h^{i}_1 = \mathbf{W}_{d_{model}}^{d_{bottleneck}}h^{i} + b_{1}$$
-$$ h^{i}_2 = \mathbf{RELU}(h_1)$$
-$$ h^{i}_3 = \mathbf{W}_{d_{bottleneck}}^{d_{model}}h_2 + b_{2}$$
-$$ \bar{h}^{i} = h^i_3 + h $$
-For the sake of brevity, we denote $ADAP^{(i)}(h_i) = h^i_3$.
+\fyTodo{More contexts and notations from the transformer}
 
-The "adapted" context vector $\bar{h}^i$ goes to the $i-th$ self-attention layer.
-There are 12 residual adapters corresponding to total 12 attention layers of the encoder and the decoder (in the decoder, the stack of self-attention layer and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter).
+Residual adapters modify context vectors in each layer by applying several transformations as follows:\fyTodo{Use align env}
+\begin{align*}
+  h^{i}_1 &= \mathbf{W}_{d_{model}}^{d_{bottleneck}}h^{i} + b_{1} \\
+  h^{i}_2 &= \mathbf{RELU}(h_1) \\
+  h^{i}_3 &= \mathbf{W}_{d_{bottleneck}}^{d_{model}}h_2 + b_{2} \\
+  \bar{h}^{i} &= h^i_3 + h
+\end{align*}
+For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$.\fyTodo{or is it $h_i$ ?}
+
+The "adapted" context vector $\bar{h}^i$ goes to\fyTodo{Is the output of ?} the $i^{\text{th}}$ self-attention\fyTodo{Self attention ?} layer.
+In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}.
+
 \subsection{Ablation study on position and number of residual adapters \label{ssec:ablatation}}
-In this section, we would like to analyze the impact of position and number of residual adapter involved in the adapted model.
-The importance of hidden layers varies with respect to their position in NMT model.
-It is conjectured that the higher layer represents more global pattern such as syntax while the lower layer represents more local pattern such as semantic.
+In this section, we would like to analyze the impact of position and number of residual adapters involved in the adapted model.
+In fact, the importance of hidden layers varies with respect to their position in the model.
+It is conjectured that the higher layer represents more global pattern such as syntax while the lower layer represents more local pattern such as semantic\fyTodo{Missing reference here}.
 The domain shift in local patterns and global patterns has not yet studied.
 In this paper, we do not intend to study this aspect of domain adaptation problem.
-We would like to give an extensive comparison between domain adaptation in different levels in NMT model.
-Because the set of possible configurations is large, we only perform experiments on layers at position: 2, 4, 6 which covers a wide range in the hierarchy of attention layers.
+We would like to give an extensive comparison between domain adaptation in different levels in NMT model.\fyTodo{Fix style here} Because the set of possible configurations is large, we only perform experiments on layers at position: 2, 4, 6 which covers a wide range in the hierarchy of attention layers.
 
@@
 \subsection{Regularization of residual adapters \label{ssec:reg}}
-Finetuning on small corpora can still lead to overfitted model even when the finetuning is applied to a few residual adapters.
-To avoid this situation, we have several options including: reduce size of the adapter, apply weight decay to the adapter during the finetuning, and apply layer regularization to the adapter during the finetuning.
-The first option is not compatible if we consider the probable growth of in-domain data.
-The second and third choices are preferable to regularize a large adapter which will scale if the in-domain corpora extends.
+Fine-tuning on small corpora can still lead to overfitted model even when the fine-tuning is applied to a few residual adapters.
+To avoid this situation, we have several options including: reduce the size of the adapter module, apply weight decay to the adapter during adaptation, or apply layer regularization to the adapter modules.
+The first option is not compatible\fyTodo{compatible ???} if we consider the probable growth of in-domain data.
+The second and third choices are preferable to regularize a large adapter which will scale if the in-domain corpora extends.\fyTodo{Style}
 
@@
     \bar{L} & = \mathop{\sum}_{x,y} (log(P(y|x)) \\
-		  & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{ADAP^{(i)}}})
+    & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{\operatorname{ADAP}^{(i)}}})
   \end{split}
 \end{equation}
+
 Layer regularization applies a penalization on the output of the adapters.
@@
     \bar{L} & = \mathop{\sum}_{x,y} (log(P(y|x)) \\
-		  & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{ADAP^{(i)}(h_i(x,y))})
+    & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\operatorname{ADAP}^{(i)}(h_i(x,y))})
   \end{split}
@@
 \subsection{Multi-task training\label{ssec:multitask}}
-In the experiments of \cite{Caruana97multitask}, the authors proposed multi-task network in which a ground layer is shared between tasks and is followed by task-related layer in the next level.
-The network is then trained from scratch by fueling task-related example in a batch and updating all parameters including shared network's parameters and task-related network's parameters at the same time.
-The multi-domain context is similar to multi-task context if we consider a domain as a task.
+In the experiments of \cite{Caruana97multitask}, the authors proposed an architecture suitable for multi-task learning.
+The corresponding network comprises a ground layer, which is shared between tasks and is followed by task-related layers in the next level.
+The network is then trained from scratch by fueling task-related examples in a batch and updating all parameters (shared and task-specific) at the same time.
+The multi-domain context is similar to multi-task context if we consider a domain as a task\cite{Dredze08}.
 Therefore, we can also use multi-task training to optimize the parameters of NMT model and the parameters of the residual adapters at once from beginning.
 To assess the efficacy of multi-task training in multi-domain context, we would like to compare the finetuning approach with multi-task training approach.
+\fyTodo{Change the training regime}
 
@@
 \mpTodo{Formalizing problem, network design, training algorithm}
-Residual adapters allow us to adapt NMT model to a domain without changing pretrained parameters.
-However, we still have to decide in which domain a sentence is translated to chose a suitable adapter.
-Therefore, we face to another problem which is the error of domain prediction.
-Choosing wrong adapter to translate a sentence can be dramatically bad because a specialized model usually has bad performance in out-of-domain examples according to well-known Catastrophic forgetting \cite{McCloskey89catastrophic}.
-However, we still have another option that is to use pretrained generic model.
+Residual adapters allow us to adapt NMT model to a new domain without altering the pretrained parameters.
+However, we still have to decide in which domain a sentence is translated to chose a suitable adapter.\fyTodo{Lack of context - domain errors} Therefore, we face to another problem which is the error of domain prediction.
+Choosing wrong adapter to translate a sentence can be dramatically bad because a specialized model usually has bad performance for out-of-domain instances \cite{McCloskey89catastrophic}.
+However, we still have another option that is to use the pretrained generic model without adaptation.
 Therefore, we could aim to achieve a performance at least as good as the generic model.
 We would like to present an adaptive version of residual adapter below.
-\\
+
 We consider a generic encoder-decoder architecture for a multi-domain sequence to sequence learning.
@@
   \centering
-  \fyDone{Fix column size}
   \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
@@
    \system{FT-Block$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 \\
-   \system{FT-Block$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.39 & 88.73 & 52 \\
+   \system{FT-Block$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.4 & 88.7 & 52 \\
    \system{FT-Block$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 \\

### 2020-08-12 2020/wmt_20/emnlp2020.tex
correct spelling errors

--- 2020-08-12 2020/wmt_20/emnlp2020.tex
@@
     \system{FT-Full}       & 29.8 & 30.97 & 19.81 & 53.43 & 49.98 & 20.84 & 34.14 \\
-   \system{FT-Block}     & 29.65 &  & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
+   \system{FT-Block}     & 29.65 & 30.45 & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
    \system{FT-HW-Block}   & 29.54 & 30.42 & 18.59 & 50.78 & 47.13 & 20.51 & 32.83 \\ 
@@
 \mpTodo{wada}
-To measure the robustness against out-of-domain examples, we perform translation with wrong domain information to trick the systems that need to know domain a priori.
-In the table \ref{tab:performance-random}, the column \domain{RND} stores the score of multi-domain model, that use domain information to translate, in the situation in which we assign random domain tag to each test sentence.
-The phenomenon Catastrophic Forgetting manifest clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
-Our proposed model \system{wada} maintains equivalent performance to generic model in \system{RND} thanks to application of adaptive weight on the output of residual adapter.
-\system{wada} avoid relying in the residual adapters when predicting examples that come from domain far from the domain of the adapters.
+To measure the robustness against out-of-domain examples, we perform translations with wrong domain information to trick the systems that need to know domain a priori.
+In the table \ref{tab:performance-random}, the column \domain{RND} stores the score of multi-domain models, that use domain information to translate, in the situation in which we assign random domain tag to each test sentence.
+The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
+Our proposed model \system{wada} maintains equivalent performance to the generic model in \system{RND} thanks to the application of adaptive weight on the output of the residual adapter.
+\system{wada} avoids relying on the residual adapters when predicting examples that come from domain far from the domain of the adapters.
 

### 2020-08-12 2020/wmt_20/emnlp2020.tex
more todos

--- 2020-08-12 2020/wmt_20/emnlp2020.tex
@@
 
-Secondly, we would like to observe how much the number of residual adapters can affect the performance of the adapted model.
+Second, we would like to observe how much the number of residual adapters affects the performance of the adapted model.
 By the limit of computation resource, we limit our study to the case of 3 adapters at positions: 2,4,6.
-By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting).
+By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting).\fyTodo{We have 12 layers, have we not ?}
 
@@
 By reusing the learned generic representation for non-topical words, we can bound the risk of poor predictions in case of out-of-domain words by the risk of the learned generic model.
+
 \subsubsection*{The training process \label{sssec:train}}
@@
 During inference, $z_k$ is used to regulate the strength of the adapter module as suggested in equation~\ref{eq:gated-residual}.
+
 \section{Experimental settings \label{sec:exp}}
@@
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
-Corpus statistics in Table~\ref{tab:Corpora}.
-Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools.
+Corpus statistics are in Table~\ref{tab:Corpora-en-fr}.
+Most corpora are available from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools.
 To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.%
@@
 
-We also report a part of experiments in language pair English, German.
-We use available corpora in the robustness task of WMT20 \footnote{\url{http://www.statmt.org/wmt20/robustness.html}} including: European Central Bank corpus (\domain{bank}),  European Economic and Social Committee corpus (\domain{economic}), European Medicines Agency corpus (\domain{med}) \footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Press Release Database of European Commission corpus, News Commentary v15 corpus, Common Crawl corpus (\domain{news}), Europarl v10 (\domain{gouv}) Tilde MODEL - czechtourism (\domain{tourism})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Paracrawl and Wikipedia Matrix (\domain{others}) \footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}.
+We also report a part of experiments in language pair English into German.
+We use available corpora in the News task of WMT20\footnote{\url{http://www.statmt.org/wmt20/news.html}} including: European Central Bank corpus (\domain{bank}),  European Economic and Social Committee corpus (\domain{eco}), European Medicines Agency corpus (\domain{med})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Press Release Database of European Commission corpus, News Commentary v15 corpus, Common Crawl corpus (\domain{news}), Europarl v10 (\domain{gouv}) Tilde MODEL - czechtourism (\domain{tourism})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Paracrawl and Wikipedia Matrix (\domain{others}) \footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}.
 We also build a joint BPE vocabulary of size 33K for both languages.
@@
     %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
-    \domain{bank} & \domain{economic} & \domain{med} & \domain{gouv} & \domain{news} & \domain{tourism} & \domain{others} \\
+    \domain{bank} & \domain{eco} & \domain{med} & \domain{gouv} & \domain{news} & \domain{tourism} & \domain{other} \\
     \hline
@@
 We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+Development sets are used to choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
 We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
 
-\subsection{Ablatation study on positions and number of residual adapters}
+\subsection{Baseline architecture \label{ssec:baseline}}
+\fyTodo{Write this - settings and parameters for Mix-Nat and Full-FT}.
 
-In table \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} we report the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} in language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} in language pair En-De.
-In most cases, the performance increase with respect to the number of residual adapters used in architecture.
-Setting \system{FT-block} using residual adapter for all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels that outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only on level.
-However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.
+\subsection{Varying the positions and number of residual adapters}
+
+Tables \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} for En-De.
+In most cases, the performance increases with respect to the number of residual adapters used in architecture.
+The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only one level.
+However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyTodo{What does bold mean ?}\fyTodo{Number of parameters ?}
 \begin{table*}
@@
     \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\ 
+    \system{FT-Block$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 \\
     \system{FT-Block$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 \\
@@
    \system{FT-Block$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 \\
-   \system{FT-Block$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 \\
      \hline
   \end{tabular}
-  \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
+  \caption{Translation performance of various finetuned systems (En:Fr). We report BLEU scores for each domain, as well as averages.}
   \label{tab:performance-en-fr}
@@
   \end{tabular}
-  \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
+  \caption{Translation performance of various fine-tuned systems (En:De). We report BLEU scores for each domain, as well as averages accross domains.}
   \label{tab:performance-en-de}
@@
 
-\subsection{Regularization of finetuning}
+\fyTodo{Why HW worse than standard version ?} \fyTodo{Why is regularization not helping ?}
+\subsection{Regularing of fine-tuning}
 

### 2020-08-13 2020/wmt_20/emnlp2020.tex
add scores

--- 2020-08-12 2020/wmt_20/emnlp2020.tex
@@
 By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting).
+
+Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of attention layers.
+We graphically demonstrate this version in the figure \ref{fig:hrl-architecture}.
+The benefit of this variant is that we can manipulate all values of the adapters at once.
+An disadvantage of th
 
@@
     \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
-   \system{FT-Block}      \\
+   \system{FT-Block} & 29.69 &	30.49 &	19.2 &	49.61 &	49.33 &	20.53 &	33.14     \\
    \system{FT-Block-WD}     & 29.68 & 30.77 & 20.42 & 50.19 & 47.68 & 20.64 & 33.23 \\

### 2020-08-13 2020/wmt_20/emnlp2020.tex

--- 2020-08-12 2020/wmt_20/emnlp2020.tex
@@
 Among the approaches for adapting a pretrained NMT model to a specific domain, finetuning is the dominant method.
-Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{bapna19simple, Vilar18learning}.
+Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{Bapna19simple, Vilar18learning}.
 The second approach has several advantages including preserving the pretrained model and the legerity of the adapters.
 However, the behavior of these adapters has not been further studied.
-The objective of this paper is to study the behavior of the adapters, then to extend the application of the adapters to improve the robustness of the model before domain shift and noisy data.
+The objective of this paper is to give an ablation study on the use of residual adapter in the domain adaptation problem.
 \mpTodo{correcting abstract}
@@
 \end{abstract}
-
 \section{Introduction } \label{sec:intro}
@@
 However, as many deep learning model, NMT models need to be trained with sufficiently large amounts of data to reach their best performance.
-Therefore the quality of translation of NMT model is still limited in low-resource language and low-resourced domain conditions \cite{duh13adaptation, zoph16transfer,koehn17six}.
+Therefore, the quality of translation of NMT model is still limited in low-resource language and low-resourced domain conditions \cite{duh13adaptation, zoph16transfer,koehn17six}.
 While many approaches have been proposed to improve the quality of NMT models in low-resource domains (see the recent survey of \citet{Chu18asurvey}), full fine-tuning of a generic baseline model remains the dominant supervised approach when adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
@@
 Indeed, studies in transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim by extensive experiments.
-Fine-tuning by training all the parameters (\system{FT-full}) of the baseline model usally improves significantly the quality of the NMT for the chosen domain.
-However, it also implies losses in translation quality for other domains, a phenomenon referred to as ``catastrophic forgetting'' in the neural network literature \cite{McCloskey89catastrophic}.
-Therefore, a full fine-tuned model is only useful to one target domain.
+Full fine-tuning, that is adapting all the parameters of a baseline model usally significantly improves the quality of the NMT for the chosen domain.
+However, it also implies large losses in translation quality for other domains, a phenomenon referred to as ``catastrophic forgetting'' in the neural network literature \cite{McCloskey89catastrophic}.
+Therefore, a fully fine-tuned model is only useful to one target domain.
 As the number of domains to handle grows, training and maintaining such models can consume a lot of resource.\fyTodo{Fix this sentence}
@@
 These adapters are trained only with the in-domain data, keeping the pretrained model frozen.
-Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering performance that are on par with full fine-tuning.
+Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering a performance that is on par with full fine-tuning.
 
@@
 \item How the number and position of adapters affect the performance.
-\item How to train the residual adapter.
-\item How to be robust to out-of-domain examples.
+\item How to train the residual adapters.
+\item How to render the combination of NMT model and residual adapters more robust to out-of-domain examples.
 \end{itemize}
@@
   h^{i}_3 &= \mathbf{W}_{d_{bottleneck}}^{d_{model}}h_2 + b_{2} \\
-  \bar{h}^{i} &= h^i_3 + h
+  \bar{h}^{i} &= h^i_3 + h^i
 \end{align*}
-For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$.\fyTodo{or is it $h_i$ ?}
+For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$.\fyDone{or is it $h_i$ ?}
 
-The "adapted" context vector $\bar{h}^i$ goes to\fyTodo{Is the output of ?} the $i^{\text{th}}$ self-attention\fyTodo{Self attention ?} layer.
+The "adapted" context vector $\bar{h}^i$ goes to\fyTodo{Is the output of ?} the $(i+1)^{\text{th}}$ attention\fyDone{Self attention ?} layer.
 In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}.
@@
 By the limit of computation resource, we limit our study to the case of 3 adapters at positions: 2,4,6.
-By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting).\fyTodo{We have 12 layers, have we not ?}
+By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting) in the encoder or the decoder.\fyTodo{We have 12 layers, have we not ?}
+
+Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of attention layers.
+We graphically demonstrate this version in the figure \ref{fig:hrl-architecture}.
+The benefit of this variant is that we can manipulate all values of the adapters at once.
+An disadvantage of this variant is a lower complexity than original version because the change of context vector only occurs in the end of the stack of layers.
 
@@
 To avoid this situation, we have several options including: reduce the size of the adapter module, apply weight decay to the adapter during adaptation, or apply layer regularization to the adapter modules.
-The first option is not compatible\fyTodo{compatible ???} if we consider the probable growth of in-domain data.
+The first option is not flexible \fyTodo{compatible ???} if we consider the probable growth of in-domain data.
 The second and third choices are preferable to regularize a large adapter which will scale if the in-domain corpora extends.\fyTodo{Style}
@@
 \mpTodo{Formalizing problem, network design, training algorithm}
-Residual adapters allow us to adapt NMT model to a new domain without altering the pretrained parameters.
+Residual adapters allow us to adapt NMT models to a new domain without altering the pretrained parameters.
 However, we still have to decide in which domain a sentence is translated to chose a suitable adapter.\fyTodo{Lack of context - domain errors} Therefore, we face to another problem which is the error of domain prediction.
@@
 In a multilayer version with highway connections (see Figure~\ref{fig:hrl-architecture}), $a_k(h) = \sum_{l} a_{kl}(h^l)$ where $h^l$ is the distributed representation of the sequence at the $l^{th}$ layer.
-\begin{figure}[h!]
+\begin{figure}[htbp]
   \centering
@@
 \end{figure}
+
 A "gated" version uses $h' = h + a_k(h) \odot{} z_k(h)$\footnote{Or $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure~\ref{fig:hrl-architecture}} where $z_k$ is a function of $h$ taking values in $[0,1]$.
@@
 \end{equation}
+
 In Word Level Adaptive Domain Adaptation (WADA), $z_k(h(w))$ is designed to reflect the "topicality" of  word $w$ is in domain $k$: the more likely $w$ is in domain $k$, the larger $z_k(h(w))$ should be.
@@
 
-\subsubsection*{The training process \label{sssec:train}}
+\subsubsection{The training process \label{sssec:train}}
 The training process comprises three main steps:
@@
 We also build a joint BPE vocabulary of size 33K for both languages.
+The statistics for the En-De corpora are in Table~\ref{tab:Corpora-en-de}.
 
@@
 \fyTodo{Write this - settings and parameters for Mix-Nat and Full-FT}.
+Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
+\begin{itemize}
+\item a generic model trained on a concatenation of all corpora, denoted \system{Mixed}\fyTodo{Or mixed nat ?}
+\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing significantly in 3 consecutive epochs.
+We again contrast two versions: full fine-tuning (\system{FT-Full}), which update all the parameters of the initial generic model \system{Mixed}; and the variant of \cite{Bapna19simple} (\system{FT-Block}).
+\end{itemize}
+
+For all models, we set the embeddings size and the hidden layers size to~512.
+Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
+The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
+% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
+% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
+Training use a batch size of~12288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
 
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{gouv}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tourism}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
-    \system{FT-Full}       & 29.8 & 30.97 & \SB{19.81} & \SB{53.43} & \SB{49.98} & 20.84 & 34.14 \\
-   \system{FT-Block}     & 29.65 &  & \SB{19.21} & \SB{48.99} & 47.22 & 20.63 & 33.14 \\ 
+    \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
+    \system{FT-Full}       & 29.8 & 30.97 & 19.81 & 53.43 & 49.98 & 20.84 & 34.14 \\
+   \system{FT-Block}     & 29.65 & 30.45 & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
    \system{FT-HW-Block}   & 29.54 & 30.42 & 18.59 & 50.78 & 47.13 & 20.51 & 32.83 \\ 
@@
 \fyTodo{Why HW worse than standard version ?} \fyTodo{Why is regularization not helping ?}
-\subsection{Regularing of fine-tuning}
+\subsection{Regularization of fine-tuning}
+In the translation task En-De, we find that domain \domain{tourism}, \domain{ecb} are extremely small with population percentage only 0.039\% and 0.022\% respectively.
+Finetuning on these domain can lead to overfitting problem.
+We assess two regularization techniques, that help avoiding overfitting, including: weight-decay and layer regularization.
+In table \ref{tab:performance-en-de-reg}, weight-decay results in system \system{FT-Block-WD} outperforms \system{FT-Block}.
 
-\begin{table*}
+\begin{table*}[htbp]
   \centering
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2} & \SB{53.3} \\ 
-   \system{FT-Block-WD}     & 37.18 & \SB{55.99} & 52.93 & 33.36 & \SB{46.03} & \SB{90.65} & \SB{52.7} \\
-   \system{FT-Block-LR}     & 37.45 & \SB{56.09} & 51.84 & 33.29 & \SB{45.02} & \SB{89.7} & \SB{52.2} \\
+    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
+   \system{FT-Block}     & 37.3	& 57.93 &	53.91 &	33.79 &	46.69 &	90.17 & 53.3 \\ 
+   \system{FT-Block-WD}     & 37.18 & 55.99 & 52.93 & 33.36 & 46.03 & 90.65 & 52.7 \\
+   \system{FT-Block-LR}     & 37.45 & 56.09 & 51.84 & 33.29 & 45.02 & 89.7 & 52.2 \\
      \hline
@@
   \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
-  \label{tab:performance-en-fr}
+  \label{tab:performance-en-fr-reg}
 \end{table*}
 
-\begin{table*}
+\begin{table*}[htbp]
   \centering
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{gouv}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tourism}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
-   \system{FT-Block}     & 29.65 &  & \SB{19.21} & \SB{48.99} & 47.22 & 20.63 & 33.14 \\
+    \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
+   \system{FT-Block} & 29.69 &	30.49 &	19.2 &	49.61 &	49.33 &	20.53 &	33.14     \\
    \system{FT-Block-WD}     & 29.68 & 30.77 & 20.42 & 50.19 & 47.68 & 20.64 & 33.23 \\
-   \system{FT-Block-LR}     & 29.65 &  & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
+   \system{FT-Block-LR}     & 29.65 & 30.45 & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
      \hline
@@
   \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
-  \label{tab:performance-en-de}
+  \label{tab:performance-en-de-reg}
+\end{table*}
+\subsection{Multi-task learning}
+To assess the efficacy of multi-task learning in training both NMT model and residual adapters at once from beginning, we use several contrast methods in multi-domain learning including:
+\begin{itemize}
+\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDR});
+\item the three proposals of \newcite{Britz17mixing}.
+\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
+Training uses reference tags and inference is performed with predicted tags, just like for regular target words.
+\system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
+These methods thus only use domain tags in training.
+\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connections make it possible to by-pass this module.
+Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
+\end{itemize}
+
+\begin{table*}[htbp]
+  \centering
+  \fyDone{Fix column size}
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+%     &&&&&& \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\% & 23.5\\
+    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} \\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} \\ \hline 
+    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1    \\
+    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} & 49.9   \\
+    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 49.8          \\
+    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} & 49.7   \\
+    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 43.4 \\ 
+    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 47.5 \\
+    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} \\
+    \system{MDL-HW-res}   & 37.38&	56.43&	52.13&	33.7&	44.81&	89.77&	52.4 \\ 
+    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 33.7\\
+     \hline
+  \end{tabular}
+  \caption{Translation performance of various MDMT systems.}
+  \label{tab:performance-multi}
 \end{table*}
 
-\subsection{Multi-task learning}
+\subsection{Word-Level Adaptive Domain Adaptation \label{sec:wada}}
+\mpTodo{wada}
+To measure the robustness against out-of-domain examples, we perform translations with wrong domain information to trick the systems that need to know domain a priori.
+In the table \ref{tab:performance-random}, the column \domain{RND} stores the score of multi-domain models, that use domain information to translate, in the situation in which we assign random domain tag to each test sentence.
+The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
+Our proposed model \system{wada} maintains equivalent performance to the generic model in \system{RND} thanks to the application of adaptive weight on the output of the residual adapter.
+\system{wada} avoids relying on the residual adapters when predicting examples that come from domain far from the domain of the adapters.
 
-\subsection{Word-Level Adaptive Domain Adaptation \label{sec:wada}}
+\begin{table*}[htbp]
+  \centering
+  \fyDone{Fix column size}
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+%     &&&&&& \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{RND}} \\ \hline
+    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 & 49.4 \\
+    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} & 32.55 \\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} & 38.38 \\ \hline 
+   \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 & 27.13\\ 
+    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1 & 35.92   \\
+    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} & 49.9 & 34.91 \\
+    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 49.8          & 37.01 \\
+    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} & 31.78 \\
+    \system{wada}    & 37.98 &	57.46&	53.05&	33.55&	46.01&	90.07&	53.0&	48.96 \\
+     \hline
+  \end{tabular}
+  \caption{Translation performance of various MDMT systems}
+  \label{tab:performance-random}
+\end{table*}
 
 \section{Related Work \label{sec:related}}
-
+\mpTodo{related work}
 \section{Discussion \label{sec:discussion}}
-
+\mpTodo{discussion}
 \section*{Acknowledgments}

### 2020-08-13 2020/wmt_20/emnlp2020.tex
discussion with minh-quang

@@
 
-\cite{Vilar18learning,bapna19simple} have recently proposed a simple and lightweight method to preserve the value of pretrained models by plugging small adapters to hidden layers.
+\cite{Vilar18learning,Bapna19simple} have recently proposed a simple and lightweight method to preserve the value of pretrained models by plugging small adapters to hidden layers.
 These adapters are trained only with the in-domain data, keeping the pretrained model frozen.
@@
 
-In this paper, we would like to extend this architecture to improve NMT in several specific situations that still challenge automatic translation, such as translating noisy inputs or translating topical texts\fyTodo{??}.
+In this paper, we would like to extend this architecture to improve NMT in several specific situations that still challenge automatic translation, such as translating noisy inputs or translating topical texts\fyTodo{multidomain, multitopic, unbalance data}.
 Residual adapters allow us to adapt NMT model to any specific domain in a computationally economical way.
@@
   h^{i}_1 &= \mathbf{W}_{d_{model}}^{d_{bottleneck}}h^{i} + b_{1} \\
-  h^{i}_2 &= \mathbf{RELU}(h_1) \\
-  h^{i}_3 &= \mathbf{W}_{d_{bottleneck}}^{d_{model}}h_2 + b_{2} \\
-  \bar{h}^{i} &= h^i_3 + h^i
+  h^{i}_2 &= \mathbf{RELU}(h_1^{i}) \\
+  h^{i}_3 &= \mathbf{W}_{d_{bottleneck}}^{d_{model}}h_2^{i} + b_{2} \\
+  \bar{h}^{i} &= h^{i}_3 + h^i
 \end{align*}
-For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$.\fyDone{or is it $h_i$ ?}
+For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$.\fyDone{or is it $h_i$ ?}\fyTodo{attention aux matrices $W_i$}
 
-The "adapted" context vector $\bar{h}^i$ goes to\fyTodo{Is the output of ?} the $(i+1)^{\text{th}}$ attention\fyDone{Self attention ?} layer.
+The "adapted" context vector $\bar{h}^i$ is the input of the $(i+1)^{\text{th}}$ attention\fyDone{Self attention ?} layer.
 In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}.
@@
 \subsection{Ablation study on position and number of residual adapters \label{ssec:ablatation}}
+\fyTodo{Move this}
 In this section, we would like to analyze the impact of position and number of residual adapters involved in the adapted model.
 In fact, the importance of hidden layers varies with respect to their position in the model.
-It is conjectured that the higher layer represents more global pattern such as syntax while the lower layer represents more local pattern such as semantic\fyTodo{Missing reference here}.
+It is conjectured that the higher layers represent more global patterns such as semantic while the lower layers represent more local patterns such as syntactic\fyTodo{Missing reference here}.
 The domain shift in local patterns and global patterns has not yet studied.
@@
 
-Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of attention layers.
-We graphically demonstrate this version in the figure \ref{fig:hrl-architecture}.
-The benefit of this variant is that we can manipulate all values of the adapters at once.
-An disadvantage of this variant is a lower complexity than original version because the change of context vector only occurs in the end of the stack of layers.
+\subsection{Multi-task training\label{ssec:multitask}}
+In the experiments of \cite{Caruana97multitask}, the authors proposed an architecture suitable for multi-task learning.
+The corresponding network comprises a ground layer, which is shared between tasks and is followed by task-related layers in the next level.
+The network is then trained from scratch by fueling task-related examples in a batch and updating all parameters (shared and task-specific) at the same time.
+The multi-domain context is similar to multi-task context if we consider a domain as a task\cite{Dredze08}.
+Therefore, we can also use multi-task training to optimize the parameters of NMT model and the parameters of the residual adapters at once from beginning.
+To assess the efficacy of multi-task training in multi-domain context, we would like to compare the finetuning approach with multi-task training approach.
+\fyTodo{Change the training regime}
 
@@
 
-\subsection{Multi-task training\label{ssec:multitask}}
-In the experiments of \cite{Caruana97multitask}, the authors proposed an architecture suitable for multi-task learning.
-The corresponding network comprises a ground layer, which is shared between tasks and is followed by task-related layers in the next level.
-The network is then trained from scratch by fueling task-related examples in a batch and updating all parameters (shared and task-specific) at the same time.
-The multi-domain context is similar to multi-task context if we consider a domain as a task\cite{Dredze08}.
-Therefore, we can also use multi-task training to optimize the parameters of NMT model and the parameters of the residual adapters at once from beginning.
-To assess the efficacy of multi-task training in multi-domain context, we would like to compare the finetuning approach with multi-task training approach.
-\fyTodo{Change the training regime}
+\subsection{Highway ...}
+Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of attention layers.
+We graphically demonstrate this version in Figure~\ref{fig:hrl-architecture}.
+The benefit of this variant is that we can manipulate all values of the adapters at once.
+An disadvantage of this variant is a lower complexity\fyTodo{Capicity ??
+justify this ?} than original version because the change of context vector only occurs in the end of the stack of layers.
 
@@
 We would like to present an adaptive version of residual adapter below.
-
+\fyTodo{Consistency of notations wrt section 2.1}
 We consider a generic encoder-decoder architecture for a multi-domain sequence to sequence learning.
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
+    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} & \SB{53.8} \\
@@
   \end{tabular}
-  \caption{Translation performance of various finetuned systems (En:Fr). We report BLEU scores for each domain, as well as averages.}
+  \caption{Translation performance of various finetuned systems (En:Fr). We report BLEU scores for each domain, as well as averages.}\fyTodo{Boldface ?}
   \label{tab:performance-en-fr}
@@
 
-\fyTodo{Why HW worse than standard version ?} \fyTodo{Why is regularization not helping ?}
+\fyTodo{Why HW worse than standard version ?} \fyTodo{Why is regularization not helping ? It helps for small domain - domain specific regularization ??}
+
 \subsection{Regularization of fine-tuning}
@@
 Finetuning on these domain can lead to overfitting problem.
-We assess two regularization techniques, that help avoiding overfitting, including: weight-decay and layer regularization.
-In table \ref{tab:performance-en-de-reg}, weight-decay results in system \system{FT-Block-WD} outperforms \system{FT-Block}.
+We assess two regularization techniques, that help avoiding overfitting, including: weight decay and layer regularization.
+In table \ref{tab:performance-en-de-reg}, weight decay results in system \system{FT-Block-WD} outperforms \system{FT-Block}.
+\fyTodo{How is weight decay ?}
 

### 2020-08-13 2020/wmt_20/emnlp2020.tex
small changes
related work plus some fixes

@@
 \subsection{Regularization of fine-tuning}
-In the translation task En-De, we find that domain \domain{tourism}, \domain{ecb} are extremely small with population percentage only 0.039\% and 0.022\% respectively.
-Finetuning on these domain can lead to overfitting problem.
-We assess two regularization techniques, that help avoiding overfitting, including: weight decay and layer regularization.
-In table \ref{tab:performance-en-de-reg}, weight decay results in system \system{FT-Block-WD} outperforms \system{FT-Block}.
-\fyTodo{How is weight decay ?}
+In the translation from English into German, we find that domains \domain{tourism}, \domain{ecb} are extremely small and account for a very small percentage of the training data, with respective proportions of 0.039\% and 0.022\%.
+Fine-tuning on these domains can lead to serious overfitting.
+We assess two well-known regularization techniques for adapter modules, that could help avoid the overfitting issues: weight decay and layer regularization.
+Results in Table~\ref{tab:performance-en-de-reg} show that weight decay (\system{FT-Block-WD}) outperforms the more basic \system{FT-Block}.
+\fyTodo{How is the weight decay parameter set ?}
 
@@
 \end{table*}
+
 \subsection{Multi-task learning}
-To assess the efficacy of multi-task learning in training both NMT model and residual adapters at once from beginning, we use several contrast methods in multi-domain learning including:
+To assess the effectiveness of multi-task learning in training both the NMT model and the residual adapters at once from scratch, we contrast our approach with several proposal from the literature on multi-domain MT including:
+
 \begin{itemize}
@@
 \mpTodo{related work}
+
+Training with data from multiple, heterogeneous sources is a common scenario in natural language processing \cite{Dredze08online,Finkel09hierarchical}.
+It is thus no wonder that the design of multi-domain systems has been proposed for many tasks.
+In this short survey, we exclusively focus on machine translation; it is likely that similar methods (parameter sharing, instance selection / weighting, adversarial training, etc) have also been proposed for other tasks.
+
+%; scores are reported separately per domain, suggesting that all domains are equally important at test time.
+% Adversarial adpatation is used in many domains
+% for QA @misc{lee2019domainagnostic,
+%    title={Domain-agnostic Question-Answering with Adversarial Training},
+%    author={Seanie Lee and Donggyu Kim and Jangwon Park},
+%    year={2019},
+%    eprint={1910.09342},
+%    archivePrefix={arXiv},
+%    primaryClass={cs.CL}
+%}, NLU and many others
+
+Early approaches to the multi-domain scenario were proposed for statistical MT, either considering multiple datansources (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}), or domains containing of several topics \cite{Eidelman12topic,Hasler14dynamic-topic}.
+Two main strategies are considered: feature-based methods, where domain/topic labels are integratetd through supplementary feature and instance-based methods, involving a measure of similarities between train and test domains.
+
+The former approach has also been adapted to NMT: \newcite{Kobus17domaincontrol,Tars18multidomain} use an additional domain feature in an RNN model, either in the form of an extra (initial) domain-token or in the form of additional domain-feature associated to each word.
+\citet{Chen16guided} apply domain control on the \emph{target} side, using a topic vector to describe the whole document context.
+Similar ideas are developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and commonalties are enforced through parameter sharing schemes.
+Parameter-sharing is also used by \citet{Jiang19multidomain}, who consider a Transformer model containing both domain-specific heads and domain-agnostic heads.
+
+\citet{Britz17mixing} studies three general techniques to take domain information into account in training: the rely on either domain classification (or domain normalization) on the source or target side.
+One contribution of this study is the use of adversarial training to normalize representations and make the combination of multiple data sources more effective.
+Similar techniques (parameter sharing, automatic domain classification / normalization) are at play in \cite{Zeng18multidomain}: in this work, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific from domain-agnostic representations.
+These two representations are then processed separately, then merged to compute the final translation.
+
+\citet{Farajian17multidomain,li-etal-2018-one} are two recent representatives of the instance-based approach: for each test sentence, a small adaptation corpus is collected based on similarity measures, and used to fine-tune a mix domain model.
+
 \section{Discussion \label{sec:discussion}}
 \mpTodo{discussion}
+
 \section*{Acknowledgments}

### 2020-08-14 2020/wmt_20/emnlp2020.tex
developping

--- 2020-08-13 2020/wmt_20/emnlp2020.tex
@@
 \fyTodo{More contexts and notations from the transformer}
-
-Residual adapters modify context vectors in each layer by applying several transformations as follows:\fyTodo{Use align env}
+For $i^{\text{th}}$layer of the encoder or the decoder, let $h^i$ be the output of the layer, of dimension d.
+The $i^{\text{th}}$ residual adapter modifies $h^i$ by applying several transformations as follows:\fyTodo{Use align env}
 \begin{align*}
-  h^{i}_1 &= \mathbf{W}_{d_{model}}^{d_{bottleneck}}h^{i} + b_{1} \\
+  h^{i}_1 &= \mathbf{W}_{db}^{i}h^{i} + b^i_{1} \\
   h^{i}_2 &= \mathbf{RELU}(h_1^{i}) \\
-  h^{i}_3 &= \mathbf{W}_{d_{bottleneck}}^{d_{model}}h_2^{i} + b_{2} \\
+  h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + b^i_{2} \\
   \bar{h}^{i} &= h^{i}_3 + h^i
 \end{align*}
+Where matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, b is a bottleneck dimension.
 For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$.\fyDone{or is it $h_i$ ?}\fyTodo{attention aux matrices $W_i$}
 
-The "adapted" context vector $\bar{h}^i$ is the input of the $(i+1)^{\text{th}}$ attention\fyDone{Self attention ?} layer.
-In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}.
+The "adapted" context vector $\bar{h}^i_{ 1\leq i \leq n-1}$, where n is number of layer  is the input of the $(i+1)^{\text{th}}$ \fyDone{Self attention ?} layer, $\bar{h}^n$ is passed to the decoder if it belongs to the encoder side, or the input of output layer if it belongs to the decoder side.
+In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder \footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}.
 
@@
 The benefit of this variant is that we can manipulate all values of the adapters at once.
-An disadvantage of this variant is a lower complexity\fyTodo{Capicity ??
-justify this ?} than original version because the change of context vector only occurs in the end of the stack of layers.
 
@@
 However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyTodo{What does bold mean ?}\fyTodo{Number of parameters ?}
+
+\begin{table*}[htbp]
+  \centering
+  \begin{tabular}{|l|l|} \hline
+    Model & params \\ \hline 
+    \system{Transformer}  & 65 \\
+   \system{Residual Adapter} & 1 \\
+     \hline
+  \end{tabular}
+  \caption{Number of parameters($\times 10^3$)}
+  \label{tab:params}
+\end{table*}
+
 \begin{table*}
@@
     \system{MDL-HW-res}   & 37.38&	56.43&	52.13&	33.7&	44.81&	89.77&	52.4 \\ 
-    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 33.7\\
+    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 33.7 \\
      \hline
@@
     \system{wada}    & 37.98 &	57.46&	53.05&	33.55&	46.01&	90.07&	53.0&	48.96 \\
+    \system{FT-full-UEW} &	37.94 &	55.97 &	52.1 &	33.71 &	44.91 &	89.08 &	52.3 &	46.48 \\
      \hline
@@
 \section{Related Work \label{sec:related}}
-\mpTodo{related work}
+
 \section{Discussion \label{sec:discussion}}
 \mpTodo{discussion}
-\section*{Acknowledgments}
+%\section*{Acknowledgments}
 
@@
 \bibliography{emnlp2020}
-\appendix
-\section{Appendices}
-\label{sec:appendix}
-\mpTodo{appendix}
-\section{Supplemental Material}
-\label{sec:supplemental}
+%\appendix
+%\section{Appendices}
+%\label{sec:appendix}
+%\mpTodo{appendix}
+%\section{Supplemental Material}
+%\label{sec:supplemental}
 

### 2020-08-14 2020/wmt_20/emnlp2020.tex
reviewed related work

--- 2020-08-13 2020/wmt_20/emnlp2020.tex
@@
 
-%; scores are reported separately per domain, suggesting that all domains are equally important at test time.
-% Adversarial adpatation is used in many domains
-% for QA @misc{lee2019domainagnostic,
-%    title={Domain-agnostic Question-Answering with Adversarial Training},
-%    author={Seanie Lee and Donggyu Kim and Jangwon Park},
-%    year={2019},
-%    eprint={1910.09342},
-%    archivePrefix={arXiv},
-%    primaryClass={cs.CL}
-%}, NLU and many others
+Early approaches to the multi-domain setting were proposed for statistical MT, either considering multiple data sources (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}) or domains containing several topics \cite{Eidelman12topic,Hasler14dynamic-topic}.
+Two main strategies are considered: feature-based methods, where domain/topic labels are integratetd through supplementary feature and instance-based methods, involving a measure of similarity between train and test domains.
 
-Early approaches to the multi-domain scenario were proposed for statistical MT, either considering multiple datansources (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}), or domains containing of several topics \cite{Eidelman12topic,Hasler14dynamic-topic}.
-Two main strategies are considered: feature-based methods, where domain/topic labels are integratetd through supplementary feature and instance-based methods, involving a measure of similarities between train and test domains.
+The former approach has also been adapted to NMT: \newcite{Kobus17domaincontrol,Tars18multidomain} use an additional domain feature in an RNN model, in the form of an extra domain-token or of additional domain-features associated with each word.
+\citet{Chen16guided} apply domain control on the \emph{target} side, using a topic vector to describe the whole document context.
+Similar ideas are developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and similarities are enforced through parameter sharing schemes.
+Parameter-sharing also lies at the core of the work by \citet{Jiang19multidomain}, who consider a Transformer model containing both domain-specific heads and domain-agnostic heads.
 
-The former approach has also been adapted to NMT: \newcite{Kobus17domaincontrol,Tars18multidomain} use an additional domain feature in an RNN model, either in the form of an extra (initial) domain-token or in the form of additional domain-feature associated to each word.
-\citet{Chen16guided} apply domain control on the \emph{target} side, using a topic vector to describe the whole document context.
-Similar ideas are developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and commonalties are enforced through parameter sharing schemes.
-Parameter-sharing is also used by \citet{Jiang19multidomain}, who consider a Transformer model containing both domain-specific heads and domain-agnostic heads.
-
-\citet{Britz17mixing} studies three general techniques to take domain information into account in training: the rely on either domain classification (or domain normalization) on the source or target side.
-One contribution of this study is the use of adversarial training to normalize representations and make the combination of multiple data sources more effective.
+\citet{Britz17mixing} studies three general techniques to take domain information into account in training: they rely on either domain classification or domain normalization on the source or target side.
+One contribution of this study is the proposal of an adversarial training scheme to normalize representations across domains and make the combination of multiple data sources more effective.
 Similar techniques (parameter sharing, automatic domain classification / normalization) are at play in \cite{Zeng18multidomain}: in this work, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific from domain-agnostic representations.
@@
 
-\citet{Farajian17multidomain,li-etal-2018-one} are two recent representatives of the instance-based approach: for each test sentence, a small adaptation corpus is collected based on similarity measures, and used to fine-tune a mix domain model.
+\citet{Farajian17multidomain,li-etal-2018-one} are two recent representatives of the instance-based approach: for each test sentence, a small adaptation corpus is collected based on similarity measures, and used to fine-tune a mix-domain model.
 

### 2020-08-14 2020/wmt_20/emnlp2020.tex

--- 2020-08-13 2020/wmt_20/emnlp2020.tex
@@
 \fyTodo{More contexts and notations from the transformer}
-
-Residual adapters modify context vectors in each layer by applying several transformations as follows:\fyTodo{Use align env}
+For $i^{\text{th}}$layer of the encoder or the decoder, let $h^i$ be the output of the layer, of dimension d.
+The $i^{\text{th}}$ residual adapter modifies $h^i$ by applying several transformations as follows:\fyTodo{Use align env}
 \begin{align*}
-  h^{i}_1 &= \mathbf{W}_{d_{model}}^{d_{bottleneck}}h^{i} + b_{1} \\
+  h^{i}_1 &= \mathbf{W}_{db}^{i}h^{i} + b^i_{1} \\
   h^{i}_2 &= \mathbf{RELU}(h_1^{i}) \\
-  h^{i}_3 &= \mathbf{W}_{d_{bottleneck}}^{d_{model}}h_2^{i} + b_{2} \\
+  h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + b^i_{2} \\
   \bar{h}^{i} &= h^{i}_3 + h^i
 \end{align*}
+Where matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, b is a bottleneck dimension.
 For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$.\fyDone{or is it $h_i$ ?}\fyTodo{attention aux matrices $W_i$}
 
-The "adapted" context vector $\bar{h}^i$ is the input of the $(i+1)^{\text{th}}$ attention\fyDone{Self attention ?} layer.
-In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}.
+The "adapted" context vector $\bar{h}^i_{ 1\leq i \leq n-1}$, where n is the number of layers, is the input of the $(i+1)^{\text{th}}$ \fyDone{Self attention ?} layer, $\bar{h}^n$ is passed to the decoder if it belongs to the encoder side, or the input of output layer if it belongs to the decoder side.
+In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder \footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}.
 

### 2020-08-14 2020/wmt_20/emnlp2020.tex
re-write introduction, mostly

--- 2020-08-14 2020/wmt_20/emnlp2020.tex
@@
 Among the approaches for adapting a pretrained NMT model to a specific domain, finetuning is the dominant method.
-Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{Bapna19simple, Vilar18learning}.
+Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{Bapna19simple,Vilar18learning}.
 The second approach has several advantages including preserving the pretrained model and the legerity of the adapters.
@@
 \section{Introduction } \label{sec:intro}
-\mpTodo{write introduction} \fyTodo{Citations in chronological order}\fyTodo{Split long sentences}
-Owing to multiple architectural improvements, Neural Machine Translation(NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence,Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
+\mpTodo{write introduction} \fyDone{Citations in chronological order}\fyDone{Split long sentences}
+Owing to multiple improvements, Neural Machine Translation(NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence,Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
 However, as many deep learning model, NMT models need to be trained with sufficiently large amounts of data to reach their best performance.
-Therefore, the quality of translation of NMT model is still limited in low-resource language and low-resourced domain conditions \cite{duh13adaptation, zoph16transfer,koehn17six}.
+Therefore, the quality of translation of NMT model is still limited in low-resource language and low-resourced domain conditions \cite{duh13adaptation,zoph16transfer,koehn17six}.
 While many approaches have been proposed to improve the quality of NMT models in low-resource domains (see the recent survey of \citet{Chu18asurvey}), full fine-tuning of a generic baseline model remains the dominant supervised approach when adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
 
-Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pretrained model by resuming training with the sole in-domain corpus.
-It is a conjecture that pretrained model constitute a better initialization a the random initialization, especially when the adaptation data is scarce.
-Indeed, studies in transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim by extensive experiments.
-Full fine-tuning, that is adapting all the parameters of a baseline model usally significantly improves the quality of the NMT for the chosen domain.
-However, it also implies large losses in translation quality for other domains, a phenomenon referred to as ``catastrophic forgetting'' in the neural network literature \cite{McCloskey89catastrophic}.
+Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pre-trained model by resuming training with the sole in-domain corpus.
+It is a conjecture that pre-trained model constitute a better initialization than a random one, especially when the adaptation data is scarce.
+Indeed, studies of transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim by extensive experiments.
+Full fine-tuning, that adapts all the parameters of a baseline model usually significantly improves the quality of the NMT for the chosen domain.
+However, it also yields large losses in translation quality for other domains, a phenomenon referred to as ``catastrophic forgetting'' in the neural network literature \cite{McCloskey89catastrophic}.
 Therefore, a fully fine-tuned model is only useful to one target domain.
-As the number of domains to handle grows, training and maintaining such models can consume a lot of resource.\fyTodo{Fix this sentence}
+As the number of domains to handle grows, training and maintaining a separate model for each task can quickly becomes tedious and resource-expensive .\fyDone{Fix this sentence.}
 
-\cite{Vilar18learning,Bapna19simple} have recently proposed a simple and lightweight method to preserve the value of pretrained models by plugging small adapters to hidden layers.
-These adapters are trained only with the in-domain data, keeping the pretrained model frozen.
-Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering a performance that is on par with full fine-tuning.
+\cite{Vilar18learning,Bapna19simple} have recently proposed a simple and lightweight domain adaptation method that also preserves the value of pre-trained models, based on small adapters component that are plugged in each hidden layer.
+These adapters are trained only with the in-domain data, keeping the pre-trained model frozen.
+Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering a performance that is close to that of full fine-tuning.
 
-In this paper, we would like to extend this architecture to improve NMT in several specific situations that still challenge automatic translation, such as translating noisy inputs or translating topical texts\fyTodo{multidomain, multitopic, unbalance data}.
-Residual adapters allow us to adapt NMT model to any specific domain in a computationally economical way.
+In this paper, we would like to extend this architecture to improve NMT in several settings that still challenge automatic translation, such as translating texts from multiple topics, genre or domains, in the face of unbalanced data distribution.
+Furthermore, as the notion of ``domains''  is not always well established, another practical setting is the translation of texts mixing several topics / domains.
+Finally, another requirement is the need to also translate texts from domain unseen in training, based only on the unadapted system, which should be made as strong as possible.
+% Residual adapters allow us to adapt NMT model to any specific domain in a computationally cheap way.
 Could we adapt NMT model to noisy text and topical text at once?
@@
 Therefore, we have to decide whether to use residual adapters manually, i.e, we have to know to which domain the text belongs a priori.
-Therefore, we would like to fuse a domain classifier to the architecture in order to weight the contribution of the adapters with respect to the topic-relatedness of the text.\fyTodo{Say differently: various implementations}
+Therefore, we would like to fuse a domain classifier to the architecture in order to weight the contribution of the adapters with respect to the topic-relatedness of the text.
+\fyDone{Say differently: various implementations}
 
-Our contribution is an thorough experimental study on the use of residual adapters in multidomain adapation. We notably try to answer the following three questions:
-\begin{itemize}
-\item How the number and position of adapters affect the performance.
-\item How to train the residual adapters.
-\item How to render the combination of NMT model and residual adapters more robust to out-of-domain examples.
-\end{itemize}
-\fyTodo{One bit of a conclusion here}
+In this context, our main contribution is a thorough experimental study of the use of residual adapters in a multi-domain translation adaptation.
+We notably explore ways to adjust and/or regularize adapter modules to handle situations where the adaptation data is very small.
+We also propose and contrast two new variants of the residual architecture: in the first one (highway residual adapters), adaptation still affects each layer of the architecture, but its effect is delayed till the last layer, thus making the architecture more modular and adaptive; in the second one (gated residual adapters), we explore ways to improve the performance in the face of train-test data mismatch.
+We experiment with two langage pair and report results that demonstrate the flexibility and effectiveness of these architectures.
+\fyDone{One bit of a conclusion here}\fyFuture{Build a proper training scenario for these two conditions}.
 
@@
 We consider a generic encoder-decoder architecture for a multi-domain sequence to sequence learning.
-Let $h \in \R^d$ the output of the encoder, \cite{bapna19simple} proposes to use an adapted representation for domain $k$ defined as $h' = h + a_k(h)$.
+Let $h \in \R^d$ the output of the encoder, \cite{Bapna19simple} proposes to use an adapted representation for domain $k$ defined as $h' = h + a_k(h)$.
 This means that all words in all sentences in domain $k$ will use the adapter module represented by $a_k$.

### 2020-08-14 2020/wmt_20/emnlp2020.tex
changed title

@@
 \NewDocumentCommand{\normL}{ s O{} m }{%
-  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{L_2}%
+  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{2}%
 }
@@
 
-\title{Ablation study on the residual adapter in Neural Machine Translation}
+% -- \title{Ablation study on the residual adapter in Neural Machine Translation}
+\title{A Study of Residual Adapters for Neural Machine Translation}
 
@@
 
-Weight decay applies a penalization on weights of the adapters.
+Weight decay applies a penalization on the weights of the adapters.\fyTodo{The second summation also runs over $x,y$ ? I think not}
 \begin{equation}

### 2020-08-14 2020/wmt_20/emnlp2020.tex
après discussion avec Minh

@@
 % -- \title{Ablation study on the residual adapter in Neural Machine Translation}
-\title{A Study of Residual Adapters for Neural Machine Translation}
+\title{A Study of Residual Adapters for Multi-Domain Neural Machine Translation}
 
@@
 \begin{abstract}
-Among the approaches for adapting a pretrained NMT model to a specific domain, finetuning is the dominant method.
+% Among the approaches for adapting a pretrained NMT model to a specific domain, finetuning is the dominant method.
 Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{Bapna19simple,Vilar18learning}.
@@
 The objective of this paper is to give an ablation study on the use of residual adapter in the domain adaptation problem.
-\mpTodo{correcting abstract}
-
-****
-
-\fyTodo{Citation-free abstract}
+% \mpTodo{correcting abstract}
+% **** 
+\fyDone{Citation-free abstract}
 Domain adaptation is an old and vexing problem for machine translation systems.
@@
 We experiment with two langage pair and report results that demonstrate the flexibility and effectiveness of these architectures.
-\fyDone{One bit of a conclusion here}\fyFuture{Build a proper training scenario for these two conditions}.
+\fyDone{One bit of a conclusion here}\fyFuture{Build a proper training scenario for these two conditions}
 
@@
 \subsection{Architecture \label{ssec:architecture}}
-\fyTodo{More contexts and notations from the transformer}
-For $i^{\text{th}}$layer of the encoder or the decoder, let $h^i$ be the output of the layer, of dimension d.
-The $i^{\text{th}}$ residual adapter modifies $h^i$ by applying several transformations as follows:\fyTodo{Use align env}
+\fyTodo{More contexts and notations from the transformer}\fyTodo{Encoder / decoder layers}
+For $i^{\text{th}}$layer of the encoder or the decoder, let $h^i$ be the output of the layer, of dimension $d$.
+The $i^{\text{th}}$ residual adapter modifies $h^i$ by applying several transformations as follows:\fyDone{Use align env}
 \begin{align*}
@@
   h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + b^i_{2} \\
-  \bar{h}^{i} &= h^{i}_3 + h^i
+  \bar{h}^{i} &= h^{i}_3 + h^i.
 \end{align*}
-Where matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, b is a bottleneck dimension.
-For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$.\fyDone{or is it $h_i$ ?}\fyTodo{attention aux matrices $W_i$}
+Matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, b is a bottleneck dimension.
+For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the set of parameters for the $i^{th}$ adapters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}
 
 The "adapted" context vector $\bar{h}^i_{ 1\leq i \leq n-1}$, where n is the number of layers, is the input of the $(i+1)^{\text{th}}$ \fyDone{Self attention ?} layer, $\bar{h}^n$ is passed to the decoder if it belongs to the encoder side, or the input of output layer if it belongs to the decoder side.
-In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder \footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}.
+In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder.\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}
+\fyTodo{Many question opens: how to train ? how many adapters ? which size ? etc}
 
@@
 
-Weight decay applies a penalization on the weights of the adapters.\fyTodo{The second summation also runs over $x,y$ ? I think not}
+Weight decay applies a penalization on the weights of the adapters.\fyDone{The second summation also runs over $x,y$ ? I think not}
 \begin{equation}
   \begin{split}
-    \bar{L} & = \mathop{\sum}_{x,y} (log(P(y|x)) \\
+    \bar{L} & = \mathop{\sum}_{x,y} - \log P(y|x) \\
     & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{\operatorname{ADAP}^{(i)}}})
@@
   \begin{split}
-    \bar{L} & = \mathop{\sum}_{x,y} (log(P(y|x)) \\
+    \bar{L} & = \mathop{\sum}_{x,y} (-\log(P(y|x)) \\
     & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\operatorname{ADAP}^{(i)}(h_i(x,y))})
@@
 
-\subsection{Highway ...}
+\subsection{Highway Residual Adapters \label{ssec:highway}}
+\fyTodo{Mettre mieux en valeur} \fyTodo{Meilleur entrainement du fait des connections highway ?}
 Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of attention layers.
@@
 
-\subsection{Word-Level Adaptive Domain Adaptation \label{ssec:wada}}
+\subsection{Gated Residual Adaptaters \label{ssec:wada}}
 \mpTodo{Formalizing problem, network design, training algorithm}
@@
   \end{tabular}
-  \caption{Number of parameters($\times 10^3$)}
+  \caption{Number of parameters ($\times 10^3$)}
   \label{tab:params}
@@
   \end{tabular}
-  \caption{Translation performance of various fine-tuned systems (En:De). We report BLEU scores for each domain, as well as averages accross domains.}
+  \caption{Translation performance of various fine-tuned systems (En:De). We report BLEU scores for each domain, as well as averages across domains.}
   \label{tab:performance-en-de}
@@
 \section{Related Work \label{sec:related}}
-\mpTodo{related work}
+\mpDone{related work}
 

### 2020-08-14 2020/wmt_20/emnlp2020.tex
local changes

--- 2020-08-14 2020/wmt_20/emnlp2020.tex
@@
 Among the approaches for adapting a pretrained NMT model to a specific domain, finetuning is the dominant method.
-Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{Bapna19simple,Vilar18learning}.
+Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{Bapna19simple, Vilar18learning}.
 The second approach has several advantages including preserving the pretrained model and the legerity of the adapters.
@@
 \section{Introduction } \label{sec:intro}
-\mpTodo{write introduction} \fyDone{Citations in chronological order}\fyDone{Split long sentences}
-Owing to multiple improvements, Neural Machine Translation(NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence,Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
+\mpTodo{write introduction} \fyTodo{Citations in chronological order}\fyTodo{Split long sentences}
+Owing to multiple architectural improvements, Neural Machine Translation(NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence,Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
 However, as many deep learning model, NMT models need to be trained with sufficiently large amounts of data to reach their best performance.
-Therefore, the quality of translation of NMT model is still limited in low-resource language and low-resourced domain conditions \cite{duh13adaptation,zoph16transfer,koehn17six}.
+Therefore, the quality of translation of NMT model is still limited in low-resource language and low-resourced domain conditions \cite{duh13adaptation, zoph16transfer,koehn17six}.
 While many approaches have been proposed to improve the quality of NMT models in low-resource domains (see the recent survey of \citet{Chu18asurvey}), full fine-tuning of a generic baseline model remains the dominant supervised approach when adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
 
-Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pre-trained model by resuming training with the sole in-domain corpus.
-It is a conjecture that pre-trained model constitute a better initialization than a random one, especially when the adaptation data is scarce.
-Indeed, studies of transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim by extensive experiments.
-Full fine-tuning, that adapts all the parameters of a baseline model usually significantly improves the quality of the NMT for the chosen domain.
-However, it also yields large losses in translation quality for other domains, a phenomenon referred to as ``catastrophic forgetting'' in the neural network literature \cite{McCloskey89catastrophic}.
+Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pretrained model by resuming training with the sole in-domain corpus.
+It is a conjecture that pretrained model constitute a better initialization a the random initialization, especially when the adaptation data is scarce.
+Indeed, studies in transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim by extensive experiments.
+Full fine-tuning, that is adapting all the parameters of a baseline model usally significantly improves the quality of the NMT for the chosen domain.
+However, it also implies large losses in translation quality for other domains, a phenomenon referred to as ``catastrophic forgetting'' in the neural network literature \cite{McCloskey89catastrophic}.
 Therefore, a fully fine-tuned model is only useful to one target domain.
-As the number of domains to handle grows, training and maintaining a separate model for each task can quickly becomes tedious and resource-expensive .\fyDone{Fix this sentence.}
+As the number of domains to handle grows, training and maintaining such models can consume a lot of resource.\fyTodo{Fix this sentence}
 
-\cite{Vilar18learning,Bapna19simple} have recently proposed a simple and lightweight domain adaptation method that also preserves the value of pre-trained models, based on small adapters component that are plugged in each hidden layer.
-These adapters are trained only with the in-domain data, keeping the pre-trained model frozen.
-Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering a performance that is close to that of full fine-tuning.
+\cite{Vilar18learning,Bapna19simple} have recently proposed a simple and lightweight method to preserve the value of pretrained models by plugging small adapters to hidden layers.
+These adapters are trained only with the in-domain data, keeping the pretrained model frozen.
+Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering a performance that is on par with full fine-tuning.
 
-In this paper, we would like to extend this architecture to improve NMT in several settings that still challenge automatic translation, such as translating texts from multiple topics, genre or domains, in the face of unbalanced data distribution.
-Furthermore, as the notion of ``domains''  is not always well established, another practical setting is the translation of texts mixing several topics / domains.
-Finally, another requirement is the need to also translate texts from domain unseen in training, based only on the unadapted system, which should be made as strong as possible.
-% Residual adapters allow us to adapt NMT model to any specific domain in a computationally cheap way.
+In this paper, we would like to extend this architecture to improve NMT in several specific situations that still challenge automatic translation, such as translating noisy inputs or translating topical texts\fyTodo{multidomain, multitopic, unbalance data}.
+Residual adapters allow us to adapt NMT model to any specific domain in a computationally economical way.
 Could we adapt NMT model to noisy text and topical text at once?
@@
 Therefore, we have to decide whether to use residual adapters manually, i.e, we have to know to which domain the text belongs a priori.
-Therefore, we would like to fuse a domain classifier to the architecture in order to weight the contribution of the adapters with respect to the topic-relatedness of the text.
-\fyDone{Say differently: various implementations}
+Therefore, we would like to fuse a domain classifier to the architecture in order to weight the contribution of the adapters with respect to the topic-relatedness of the text.\fyTodo{Say differently: various implementations}
 
-In this context, our main contribution is a thorough experimental study of the use of residual adapters in a multi-domain translation adaptation.
-We notably explore ways to adjust and/or regularize adapter modules to handle situations where the adaptation data is very small.
-We also propose and contrast two new variants of the residual architecture: in the first one (highway residual adapters), adaptation still affects each layer of the architecture, but its effect is delayed till the last layer, thus making the architecture more modular and adaptive; in the second one (gated residual adapters), we explore ways to improve the performance in the face of train-test data mismatch.
-We experiment with two langage pair and report results that demonstrate the flexibility and effectiveness of these architectures.
-\fyDone{One bit of a conclusion here}\fyFuture{Build a proper training scenario for these two conditions}.
+Our contribution is an thorough experimental study on the use of residual adapters in multidomain adapation. We notably try to answer the following three questions:
+\begin{itemize}
+\item How the number and position of adapters affect the performance.
+\item How to train the residual adapters.
+\item How to render the combination of NMT model and residual adapters more robust to out-of-domain examples.
+\end{itemize}
+\fyTodo{One bit of a conclusion here}
 
@@
 In fact, the importance of hidden layers varies with respect to their position in the model.
-It is conjectured that the higher layers represent more global patterns such as semantic while the lower layers represent more local patterns such as syntactic\fyTodo{Missing reference here}.
+It is conjectured that the higher layers represent more global patterns such as semantic while the lower layers represent more local patterns such as syntactic \cite{raganato18analysis}\fyTodo{Missing reference here}.
 The domain shift in local patterns and global patterns has not yet studied.
 In this paper, we do not intend to study this aspect of domain adaptation problem.
-We would like to give an extensive comparison between domain adaptation in different levels in NMT model.\fyTodo{Fix style here} Because the set of possible configurations is large, we only perform experiments on layers at position: 2, 4, 6 which covers a wide range in the hierarchy of attention layers.
+We would like to give an extensive comparison between domain adaptation in different levels in NMT model.\fyTodo{Fix style here} Because the set of possible configurations is large, we only perform experiments on layers at position: 2, 4, 6 in each side of NMT model (encoder and decoder) which covers a wide range in the hierarchy of attention layers.
 
 Second, we would like to observe how much the number of residual adapters affects the performance of the adapted model.
-By the limit of computation resource, we limit our study to the case of 3 adapters at positions: 2,4,6.
-By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting) in the encoder or the decoder.\fyTodo{We have 12 layers, have we not ?}
+By the limit of computation resource, we limit our study to the case of 3 adapters at positions: 2,4,6 in each side of the NMT model.
+By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting) in each side (encoder and decoder).\fyTodo{We have 12 layers, have we not ?}
 
@@
 
-\subsection{Highway ...}
-Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of attention layers.
+\subsection{Highway connection}
+Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of layers.
 We graphically demonstrate this version in Figure~\ref{fig:hrl-architecture}.
+In this version, only the last layers of the encoder and the decoder are modified.
+
+\begin{equation}
+\bar{h}^n = h^n + \displaystyle{\mathop{\sum}_{1\leq i\leq n}} ADAP^i(h^i)
+\end{equation}
+
 The benefit of this variant is that we can manipulate all values of the adapters at once.
@@
 We consider a generic encoder-decoder architecture for a multi-domain sequence to sequence learning.
-Let $h \in \R^d$ the output of the encoder, \cite{Bapna19simple} proposes to use an adapted representation for domain $k$ defined as $h' = h + a_k(h)$.
+Let $h \in \R^d$ the output of the encoder, \cite{bapna19simple} proposes to use an adapted representation for domain $k$ defined as $h' = h + a_k(h)$.
 This means that all words in all sentences in domain $k$ will use the adapter module represented by $a_k$.
@@
 
-Early approaches to the multi-domain setting were proposed for statistical MT, either considering multiple data sources (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}) or domains containing several topics \cite{Eidelman12topic,Hasler14dynamic-topic}.
-Two main strategies are considered: feature-based methods, where domain/topic labels are integratetd through supplementary feature and instance-based methods, involving a measure of similarity between train and test domains.
+%; scores are reported separately per domain, suggesting that all domains are equally important at test time.
+% Adversarial adpatation is used in many domains
+% for QA @misc{lee2019domainagnostic,
+%    title={Domain-agnostic Question-Answering with Adversarial Training},
+%    author={Seanie Lee and Donggyu Kim and Jangwon Park},
+%    year={2019},
+%    eprint={1910.09342},
+%    archivePrefix={arXiv},
+%    primaryClass={cs.CL}
+%}, NLU and many others
 
-The former approach has also been adapted to NMT: \newcite{Kobus17domaincontrol,Tars18multidomain} use an additional domain feature in an RNN model, in the form of an extra domain-token or of additional domain-features associated with each word.
+Early approaches to the multi-domain scenario were proposed for statistical MT, either considering multiple datansources (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}), or domains containing of several topics \cite{Eidelman12topic,Hasler14dynamic-topic}.
+Two main strategies are considered: feature-based methods, where domain/topic labels are integratetd through supplementary feature and instance-based methods, involving a measure of similarities between train and test domains.
+
+The former approach has also been adapted to NMT: \newcite{Kobus17domaincontrol,Tars18multidomain} use an additional domain feature in an RNN model, either in the form of an extra (initial) domain-token or in the form of additional domain-feature associated to each word.
 \citet{Chen16guided} apply domain control on the \emph{target} side, using a topic vector to describe the whole document context.
-Similar ideas are developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and similarities are enforced through parameter sharing schemes.
-Parameter-sharing also lies at the core of the work by \citet{Jiang19multidomain}, who consider a Transformer model containing both domain-specific heads and domain-agnostic heads.
+Similar ideas are developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and commonalties are enforced through parameter sharing schemes.
+Parameter-sharing is also used by \citet{Jiang19multidomain}, who consider a Transformer model containing both domain-specific heads and domain-agnostic heads.
 
-\citet{Britz17mixing} studies three general techniques to take domain information into account in training: they rely on either domain classification or domain normalization on the source or target side.
-One contribution of this study is the proposal of an adversarial training scheme to normalize representations across domains and make the combination of multiple data sources more effective.
+\citet{Britz17mixing} studies three general techniques to take domain information into account in training: the rely on either domain classification (or domain normalization) on the source or target side.
+One contribution of this study is the use of adversarial training to normalize representations and make the combination of multiple data sources more effective.
 Similar techniques (parameter sharing, automatic domain classification / normalization) are at play in \cite{Zeng18multidomain}: in this work, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific from domain-agnostic representations.
@@
 
-\citet{Farajian17multidomain,li-etal-2018-one} are two recent representatives of the instance-based approach: for each test sentence, a small adaptation corpus is collected based on similarity measures, and used to fine-tune a mix-domain model.
+\citet{Farajian17multidomain,li-etal-2018-one} are two recent representatives of the instance-based approach: for each test sentence, a small adaptation corpus is collected based on similarity measures, and used to fine-tune a mix domain model.
 

### 2020-08-14 2020/wmt_20/emnlp2020.tex

--- 2020-08-14 2020/wmt_20/emnlp2020.tex
@@
 
+\iffalse   % removes draft highlighting for final version
+\renewcommand{\draftAdd}[2]{#1}
+\fi
+
 % -- \title{Ablation study on the residual adapter in Neural Machine Translation}
@@
 We also propose and contrast two new variants of the residual architecture: in the first one (highway residual adapters), adaptation still affects each layer of the architecture, but its effect is delayed till the last layer, thus making the architecture more modular and adaptive; in the second one (gated residual adapters), we explore ways to improve the performance in the face of train-test data mismatch.
-We experiment with two langage pair and report results that demonstrate the flexibility and effectiveness of these architectures.
+We experiment with two language pairs and report results that demonstrate the flexibility and effectiveness of these architectures.
 \fyDone{One bit of a conclusion here}\fyFuture{Build a proper training scenario for these two conditions}
@@
 Matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, b is a bottleneck dimension.
-For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the set of parameters for the $i^{th}$ adapters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}
+For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the set of parameters for the $i^{\text{th}}$ adapters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}
 
-The "adapted" context vector $\bar{h}^i_{ 1\leq i \leq n-1}$, where n is the number of layers, is the input of the $(i+1)^{\text{th}}$ \fyDone{Self attention ?} layer, $\bar{h}^n$ is passed to the decoder if it belongs to the encoder side, or the input of output layer if it belongs to the decoder side.
+The "adapted" hidden vectors $\bar{h}^i_{ 1\leq i \leq n-1}$, where n is the number of layers, is the input of the $(i+1)^{\text{th}}$ \fyDone{Self attention ?} layer, $\bar{h}^n$ is passed to the decoder if it belongs to the encoder side, or the input of output layer if it belongs to the decoder side.
 In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder.\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}
@@
 
-\subsection{Ablation study on position and number of residual adapters \label{ssec:ablatation}}
-\fyTodo{Move this}
-In this section, we would like to analyze the impact of position and number of residual adapters involved in the adapted model.
-In fact, the importance of hidden layers varies with respect to their position in the model.
-It is conjectured that the higher layers represent more global patterns such as semantic while the lower layers represent more local patterns such as syntactic\fyTodo{Missing reference here}.
+The size of available corpora for each domain is extremely varying.
+For example, the corpus of domain \domain{tourism} in En-De contains only around $7\times10^3$ sentence pairs while the corpus of domain \domain{news} in En-De contains approximately $3\times10^6$ sentence pairs.
+For extremely small domains, we could economize the number of parameters by reducing the number of residual adapters in the architecture.
+While only a limited number of adapters are incorporated to the NMT model, which positions in the model are more important in the domain adaptation task?
+We would like to analyze the impact of position and number of residual adapters involved in the adapted model.
+It is conjectured that the higher layers represent more global patterns such as semantic while the lower layers represent more local patterns such as syntactic \cite{raganato18analysis}\fyTodo{Missing reference here}.
 The domain shift in local patterns and global patterns has not yet studied.
-In this paper, we do not intend to study this aspect of domain adaptation problem.
-We would like to give an extensive comparison between domain adaptation in different levels in NMT model.\fyTodo{Fix style here} Because the set of possible configurations is large, we only perform experiments on layers at position: 2, 4, 6 which covers a wide range in the hierarchy of attention layers.
+In this paper, we do not intend to study this aspect of the domain adaptation problem.
+We would like to give an extensive comparison between domain adaptation in different levels in NMT model.
 
-Second, we would like to observe how much the number of residual adapters affects the performance of the adapted model.
-By the limit of computation resource, we limit our study to the case of 3 adapters at positions: 2,4,6.
-By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting) in the encoder or the decoder.\fyTodo{We have 12 layers, have we not ?}
-
-\subsection{Multi-task training\label{ssec:multitask}}
-In the experiments of \cite{Caruana97multitask}, the authors proposed an architecture suitable for multi-task learning.
-The corresponding network comprises a ground layer, which is shared between tasks and is followed by task-related layers in the next level.
-The network is then trained from scratch by fueling task-related examples in a batch and updating all parameters (shared and task-specific) at the same time.
-The multi-domain context is similar to multi-task context if we consider a domain as a task\cite{Dredze08}.
-Therefore, we can also use multi-task training to optimize the parameters of NMT model and the parameters of the residual adapters at once from beginning.
-To assess the efficacy of multi-task training in multi-domain context, we would like to compare the finetuning approach with multi-task training approach.
-\fyTodo{Change the training regime}
-
-\subsection{Regularization of residual adapters \label{ssec:reg}}
-Fine-tuning on small corpora can still lead to overfitted model even when the fine-tuning is applied to a few residual adapters.
-To avoid this situation, we have several options including: reduce the size of the adapter module, apply weight decay to the adapter during adaptation, or apply layer regularization to the adapter modules.
+Fine-tuning on small corpora can still lead to an overfitted model even when the fine-tuning is applied to a few residual adapters.
+To avoid this situation, we have several options including: reducing the size of the adapter module, applying weight decay or layer regularization to the adapter.
 The first option is not flexible \fyTodo{compatible ???} if we consider the probable growth of in-domain data.
-The second and third choices are preferable to regularize a large adapter which will scale if the in-domain corpora extends.\fyTodo{Style}
+The second and third choices are preferable to regularize a large adapter which will scale if the in-domain corpora extend .\fyTodo{Style}
 
@@
   \begin{split}
-    \bar{L} & = \mathop{\sum}_{x,y} - \log P(y|x) \\
-    & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{\operatorname{ADAP}^{(i)}}})
+    \bar{L} & = \frac{1}{\#(x,y)}\mathop{\sum}_{x,y}( - \log(P(y|x))) \\
+    & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{\operatorname{ADAP}^{(i)}}}
   \end{split}
 \end{equation}
+Where $\theta_{\operatorname{ADAP}^{(i)}}$ are parameters of $i^{\text{th}}$ adapter
 
@@
   \begin{split}
-    \bar{L} & = \mathop{\sum}_{x,y} (-\log(P(y|x)) \\
+    \bar{L} & = \frac{1}{\#(x,y)}\mathop{\sum}_{x,y} (-\log(P(y|x)) \\
     & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\operatorname{ADAP}^{(i)}(h_i(x,y))})
@@
 
+In a multi-domain context, there are 2 ways to train domain-specific NMT model using adapters including pretraining generic NMT then finetuning adapters; or training both NMT model and adapters at once from the beginning using multi-task learning.
+In the literature of Machine Learning, \cite{Caruana97multitask} proposed an architecture suitable for multi-task learning.
+The corresponding network comprises a ground layer, which is shared between tasks and is followed by task-related layers.
+The network is then trained from scratch by fueling task-related examples in a batch and updating all parameters (shared and task-specific) at the same time.
+The multi-domain context is similar to the multi-task context if we consider a domain as a task\cite{Dredze08online}.
+Therefore, we can also use multi-task training to optimize the parameters of NMT model and the parameters of the residual adapters at once from the beginning.
+To assess the efficacy of multi-task training in the multi-domain context, we would like to compare the finetuning approach with the multi-task training approach.
+\fyTodo{Change the training regime}
+
 \subsection{Highway Residual Adapters \label{ssec:highway}}
-\fyTodo{Mettre mieux en valeur} \fyTodo{Meilleur entrainement du fait des connections highway ?}
 Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of attention layers.
 We graphically demonstrate this version in Figure~\ref{fig:hrl-architecture}.
-The benefit of this variant is that we can manipulate all values of the adapters at once.
-
-\subsection{Gated Residual Adaptaters \label{ssec:wada}}
-\mpTodo{Formalizing problem, network design, training algorithm}
-Residual adapters allow us to adapt NMT models to a new domain without altering the pretrained parameters.
-However, we still have to decide in which domain a sentence is translated to chose a suitable adapter.\fyTodo{Lack of context - domain errors} Therefore, we face to another problem which is the error of domain prediction.
-Choosing wrong adapter to translate a sentence can be dramatically bad because a specialized model usually has bad performance for out-of-domain instances \cite{McCloskey89catastrophic}.
-However, we still have another option that is to use the pretrained generic model without adaptation.
-Therefore, we could aim to achieve a performance at least as good as the generic model.
-We would like to present an adaptive version of residual adapter below.
-\fyTodo{Consistency of notations wrt section 2.1}
-We consider a generic encoder-decoder architecture for a multi-domain sequence to sequence learning.
-Let $h \in \R^d$ the output of the encoder, \cite{Bapna19simple} proposes to use an adapted representation for domain $k$ defined as $h' = h + a_k(h)$.
-This means that all words in all sentences in domain $k$ will use the adapter module represented by $a_k$.
-In a multilayer version with highway connections (see Figure~\ref{fig:hrl-architecture}), $a_k(h) = \sum_{l} a_{kl}(h^l)$ where $h^l$ is the distributed representation of the sequence at the $l^{th}$ layer.
+In this version, only the last hidden vector of the encoder (decoder) is modified as follow
+\begin{equation}
+\bar{h}^n = h^n + \displaystyle{\mathop{\sum}_{1 \leq i \leq n} ADAP^i(h^i)}
+\end{equation}
+The benefit of this variant is that it allows us to reuse hidden vectors $h^i$ of all hidden layers while adapting to multiple domains.
+In the forward step, the hidden vectors $h^i$ are calculated once, then adapters take these vectors to modify the last layer.
+Therefore, we can finetune the model to multiple domains at once without recomputing $h^i$.
+The second minor benefit of this variant is that it allows implementing the Gated Residual Adaptaters version that will be described in the following section \ref{ssec:gate}.
 \begin{figure}[htbp]
@@
 \end{figure}
+\subsection{Gated Residual Adaptaters \label{ssec:gate}}
+\mpTodo{Formalizing problem, network design, training algorithm}
+To translate a sentence, we have 2 options: using a domain-adapted model or using the generic model.
+In the case of choosing the wrong domain, the performance is dramatically worse than using the generic model.
+Therefore, we would like to design a version of applying residual adapter that allows us to automatically accept or reject given domain information, i.e, even if we choose a wrong adapter, the NMT model is still able to reject the change output from the adapter.
+In this version, we reuse the highway version of residual adapters.
+\fyTodo{Consistency of notations wrt section 2.1}
+Given domain $k$ with set of adapters $ADAP^i_k$, in the model adapted to domain k, instead of using the formulation $\bar{h}^n = h^n + \displaystyle{\mathop{\sum}_{1 \leq i \leq n} ADAP_k^i(h^i)}$ (encoder or decoder's side), we propose "gated" version uses $\bar{h}^n = h^n + \displaystyle{\mathop{\sum}_{1 \leq i \leq n} ADAP_k^i(h^i) \odot{} z_k(h^n)}$.
+Where $z_k(h^n(w))$ is designed to reflect the relatedness of word $w$ to domain $k$: the more likely $w$ is in domain $k$, the larger $z_k(h^n(w))$ should be.
+The gated version aims to reduce the change $\displaystyle{\mathop{\sum}_{1 \leq i \leq n} ADAP_k^i(h^i)}$ added to $h^n$ for words that are not typical of domain $k$.
+In our implementation, we incorporate 2 domain classifiers on top of the encoder and the decoder that take the last hidden layer of encoder (decoder) as input, we take the posterior probability $P(d=k|w)$ as value of $z_k(h^n(w))$.
 
-A "gated" version uses $h' = h + a_k(h) \odot{} z_k(h)$\footnote{Or $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure~\ref{fig:hrl-architecture}} where $z_k$ is a function of $h$ taking values in $[0,1]$.
-More precisely, $h$ and $h'$ are sequences of context vectors and the combination is performed element-wise, yielding:
-\begin{equation}
-   \forall t \in [1 \dots{} T], h'(w_t) = h(w_t) + a_k(h(w_t)) \odot{} z_k(h(w_t)). \label{eq:gated-residual}
-\end{equation}
-
-In Word Level Adaptive Domain Adaptation (WADA), $z_k(h(w))$ is designed to reflect the "topicality" of  word $w$ is in domain $k$: the more likely $w$ is in domain $k$, the larger $z_k(h(w))$ should be.
-Word-level adaptive domain adaptation aims to reduce the variation caused by $a_k$ to the adapted context vector $h'$ (compared to $h$) for words that are not typical of domain $k$.
-By reusing the learned generic representation for non-topical words, we can bound the risk of poor predictions in case of out-of-domain words by the risk of the learned generic model.
-
-\subsubsection{The training process \label{sssec:train}}
-The training process comprises three main steps:
+The training of Gated Residual Adaptaters comprises three main steps:
 \begin{itemize}
-	\item Pretraining a generic model with a mixed corpora comprising data from multiple domains.
+	\item Pretraining a generic model with mixed corpora comprising data from multiple domains.
 	\item Training a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen.
@@
 \end{itemize}
-During inference, $z_k$ is used to regulate the strength of the adapter module as suggested in equation~\ref{eq:gated-residual}.
 
@@
     \hline
-    4(0.00022) & 2857(0.15) & 347(0.018) & 1828(0.095) & 3696(0.19) & 7(0.00039) & 10472771(0.54) \\
+    4(0.00022) & 2857(0.15) & 347(0.018) & 1828(0.095) & 3696(0.19) & 7(0.00039) & 10473(0.54) \\
     \hline
@@
 
-For all models, we set the embeddings size and the hidden layers size to~512.
+For all models in En-Fr, we set the embeddings size and the hidden layers size to~512.
 Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
-The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
+The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of a 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
 % The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
@@
 
-\subsection{Varying the positions and number of residual adapters}
+For all models in En-De, we set the embeddings size and the hidden layers size to~1024.
+Transformers use multi-head attention with 16 heads in each of the 6 layers; the inner feedforward layer contains 4096 cells.
+The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of a 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 2048.
+Training uses a batch size of~32768 tokens.
 
-Tables \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} for En-De.
-In most cases, the performance increases with respect to the number of residual adapters used in architecture.
-The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only one level.
-However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyTodo{What does bold mean ?}\fyTodo{Number of parameters ?}
+The size of models are reported in the table \ref{tab:params}
 
@@
     Model & params \\ \hline 
-    \system{Transformer}  & 65 \\
-   \system{Residual Adapter} & 1 \\
+    \system{Transformer-En-Fr}  & 65 \\
+    \system{Residual Adapter-En-Fr} & 1 \\
+    \system{Transformer-En-De}  & 213 \\
+    \system{Residual Adapter-En-De} & 8 \\
      \hline
   \end{tabular}
-  \caption{Number of parameters ($\times 10^3$)}
+  \caption{Number of parameters ($\times 10^6$)}
   \label{tab:params}
 \end{table*}
+\subsection{Varying the positions and number of residual adapters}
+Tables \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} for En-De.
+The original version \system{FT-Block} outperforms the baseline \system{mixed} in almost all domains while stays equivalent to the baseline in \domain{med} in En-Fr and \domain{news}, \domain{gouv}, \domain{eco}, \domain{med} in En-De.
+However, \system{FT-full} still stay as best model in all domain tests.
+
+By varying the number and positions of residual adapters as aforementioned in section \ref{ssec:architecture}, we assess several variants of applying residual adapters.
+\fyTodo{Fix style here} Because the set of possible configurations is large, we only perform experiments on layers at position: 2, 4, 6 on each side of NMT model (encoder and decoder) which covers a wide range in the hierarchy of attention layers.
+To observe how much the number of residual adapters affects the performance of the adapted model.
+By the limit of computation resources, we limit our study to the case of 3 adapters at positions: 2,4,6 in each side of the NMT model.
+By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting) in each side (encoder and decoder).\fyTodo{We have 12 layers, have we not ?}
+
+ In most cases, the performance increases with respect to the number of residual adapters used in architecture.
+The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only one level.
+However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyTodo{What does bold mean ?}\fyTodo{Number of parameters ?}
 
@@
     \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} & \SB{53.8} \\
-    \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2} & \SB{53.3} \\ 
+    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
+    \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
     \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\ 
@@
 
-\fyTodo{Why HW worse than standard version ?} \fyTodo{Why is regularization not helping ? It helps for small domain - domain specific regularization ??}
-
+\fyTodo{Why HW worse than standard version ?} \fyTodo{Why is regularization not helping ? It helps for small domain - domain-specific regularization ??}
 \subsection{Regularization of fine-tuning}
@@
 We assess two well-known regularization techniques for adapter modules, that could help avoid the overfitting issues: weight decay and layer regularization.
-Results in Table~\ref{tab:performance-en-de-reg} show that weight decay (\system{FT-Block-WD}) outperforms the more basic \system{FT-Block}.
-\fyTodo{How is the weight decay parameter set ?}
 
+We choose optimal hyper-parameter $\lambda$ by grid search with a small set of values $\{ 10^{-3}, 10^{-4}, 10^{-5}$.
+
+Results in Table~\ref{tab:performance-en-de-reg} show that weight decay (\system{FT-Block-WD}) outperforms the more basic \system{FT-Block}. \fyTodo{How is the weight decay parameter set ?}
+
+\iffalse{
 \begin{table*}[htbp]
@@
 \end{table*}
+}
+\fi
 
@@
 \subsection{Multi-task learning}
-To assess the effectiveness of multi-task learning in training both the NMT model and the residual adapters at once from scratch, we contrast our approach with several proposal from the literature on multi-domain MT including:
+To assess the effectiveness of multi-task learning in training both the NMT model and the residual adapters at once from scratch, we contrast our approach with several proposals from the literature on multi-domain MT including:
 
 \begin{itemize}
-\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDR});
+\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}) or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain-specific part (\system{LDR});
 \item the three proposals of \newcite{Britz17mixing}.
 \system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
-Training uses reference tags and inference is performed with predicted tags, just like for regular target words.
-\system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
+The training uses reference tags and inference is performed with predicted tags, just like for regular target words.
+\system{DM} is a multi-task learner where a domain classifier is trained on top of the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
 These methods thus only use domain tags in training.
@@
 Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
+\item highway variant trained with multi-task learning denoted \system{MDL-HW-Res}.
 \end{itemize}
+
+The table \ref{tab:performance-multi} shows that \system{FT-Block} outperforms \system{MDL-Res}, \system{MDL-HW-Res}.
+It means that individually finetuning generic model is better than multi-task training.
+However, \system{MDL-Res} and \system{MDL-HW-Res} are better than other methods in multi-domain learning.
 
@@
     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} \\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} \\ \hline 
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} \\ 
+   \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\
+   \hline 
     \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1    \\
@@
     \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 47.5 \\
-    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} \\
-    \system{MDL-HW-res}   & 37.38&	56.43&	52.13&	33.7&	44.81&	89.77&	52.4 \\ 
-    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 33.7 \\
+    \system{MDL-Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} \\
+    \system{MDL-HW-Res}   & 37.38&	56.43&	52.13&	33.7&	44.81&	89.77&	52.4 \\ 
+    \hfill MDL-Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 33.7 \\
      \hline
@@
 
-\subsection{Word-Level Adaptive Domain Adaptation \label{sec:wada}}
-\mpTodo{wada}
+We also report the number of parameters of each model in table \ref{tab:params-multi}
+
+\begin{table*}[htbp]
+  \centering
+  \begin{tabular}{|l|l|} \hline
+    Model & params \\ \hline 
+    \system{Mixed}  & 65000000 \\
+    \system{DC-Feat} & 65240000 \\
+    \system{DC-Tag}  & 65003072 \\
+    \system{LDR} & 66440000 \\
+    \system{ADM}  & 65000000 \\
+    \system{DM} & 65240000 \\
+    \system{MDL-Res}  & 77601344 \\
+    \system{FT-Block} & 77601344 \\
+     \hline
+  \end{tabular}
+  \caption{Number of parameters ($\times 10^6$)}
+  \label{tab:params-multi}
+\end{table*}
+
+\subsection{Gated Residual Adaptaters \label{ssec:gate-exp}}
 To measure the robustness against out-of-domain examples, we perform translations with wrong domain information to trick the systems that need to know domain a priori.
@@
 The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
-Our proposed model \system{wada} maintains equivalent performance to the generic model in \system{RND} thanks to the application of adaptive weight on the output of the residual adapter.
-\system{wada} avoids relying on the residual adapters when predicting examples that come from domain far from the domain of the adapters.
+Our proposed model \system{FT-gate-block} maintains equivalent performance to the generic model in \system{RND} thanks to the application of adaptive weight on the output of the residual adapter.
+\system{FT-gate-block} avoids relying on the residual adapters when predicting examples that come from domain far from the domain of the adapters.
 
@@
     \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} & 31.78 \\
-    \system{wada}    & 37.98 &	57.46&	53.05&	33.55&	46.01&	90.07&	53.0&	48.96 \\
+    \system{FT-gate-block}    & 37.98 &	57.46&	53.05&	33.55&	46.01&	90.07&	53.0&	48.96 \\
     \system{FT-full-UEW} &	37.94 &	55.97 &	52.1 &	33.71 &	44.91 &	89.08 &	52.3 &	46.48 \\

### 2020-08-14 2020/wmt_20/emnlp2020.tex
Reviewed new section 2

@@
 
-In this context, our main contribution is a thorough experimental study of the use of residual adapters in a multi-domain translation adaptation.
+In this context, our main contribution is a thorough experimental study of the use of residual adapters in multi-domain translation.
 We notably explore ways to adjust and/or regularize adapter modules to handle situations where the adaptation data is very small.
@@
 
-\subsection{Architecture \label{ssec:architecture}}
-\fyTodo{More contexts and notations from the transformer}\fyTodo{Encoder / decoder layers}
-For $i^{\text{th}}$layer of the encoder or the decoder, let $h^i$ be the output of the layer, of dimension $d$.
-The $i^{\text{th}}$ residual adapter modifies $h^i$ by applying several transformations as follows:\fyDone{Use align env}
+\subsection{Basic architecture \label{ssec:architecture}}
+\fyDone{More contexts and notations from the transformer}\fyDone{Encoder / decoder layers}
+
+\subsubsection{The computation of adapter layers}
+Our reference architecture is the Transformer model of \cite{Vaswani17attention}, which we assume contains a stack of layers both in the encoder and the decoder side.
+Each layer contains two subparts, an attention layer and a dense layer.
+Details may vary from one implementation to another, we simply content here that each layer $i \in \{1 \dots L\}$ (in the encoder or the decoder) computes a transform of a fixed-length sequence of $d$-dimensional input vectors $h^{i-1}$ into a sequence of output vectors $h^{i}$.
+
+The residual adapter architecture extends this model by adding one new computational module for each layer.
+The $i^{\text{th}}$ such residual adapter thus modifies $h^i$ by computing the following transformations:\fyDone{Use align env}
 \begin{align*}
   h^{i}_1 &= \mathbf{W}_{db}^{i}h^{i} + b^i_{1} \\
-  h^{i}_2 &= \mathbf{RELU}(h_1^{i}) \\
+  h^{i}_2 &= \mathbf{ReLU}(h_1^{i}) \\
   h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + b^i_{2} \\
@@
 \end{align*}
-Matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, b is a bottleneck dimension.
-For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the set of parameters for the $i^{\text{th}}$ adapters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}
+Overall, the  $i^{\text{th}}$ adapter is thus parameterized by matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, with $b$ the dimension of the adapter (with $d \gg b$)\fyTodo{Check this}.
+For the sake of brevity, we will simply denote $h^{i}_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the corresponding set of parameters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}
 
-The "adapted" hidden vectors $\bar{h}^i_{ 1\leq i \leq n-1}$, where n is the number of layers, is the input of the $(i+1)^{\text{th}}$ \fyDone{Self attention ?} layer, $\bar{h}^n$ is passed to the decoder if it belongs to the encoder side, or the input of output layer if it belongs to the decoder side.
-In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder.\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}
-\fyTodo{Many question opens: how to train ? how many adapters ? which size ? etc}
+The "adapted" hidden vectors $\bar{h}^i_{ 1\leq i \leq L-1}$, where $L$ is the number of layers, will then be the input of the $(i+1)^{\text{th}}$\fyDone{Self attention ?} layer; $\bar{h}^L$ is passed to the decoder if it belongs to the encoder side, or is the input of output layer if it belongs to the decoder side.
+Note that zeroing out all adapters enables us to recover the basic Transformer, with $\bar{h}^{i} = h^i$ for all $i$.
 
-The size of available corpora for each domain is extremely varying.
+In the experiments reported in Section~\ref{sec:exp}, we will use $2\times{}L=12$ residual adapters, one for each of the $L=6$ attention layers of the encoder and similarly for the decoder.\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and only serves one residual adapter.}
+
+\subsubsection{Design space and variants}
+This general architecture leaves open many design choices pertaining to the details of the network organization, the training procedure and the corresponding objective function.
+
+A first question concerns the number of adapter layers.
+While in principle, all Transformer layers can be subject to adaptation, it is nonetheless worthwhile to consider simpler adaptation schemes, which would only alter a limited number of layers.
+Such strategy might be especially relevant when the training corpus contains very small domains, such as the ones considered in our experiments, and for which a complete adaptation may not be necessary or/and or prone to overfitting.
+Likewise, it might be meaningful to explore ways to share subsets of adapters across domains.
+This in turn raises the issues of which layer(s) to adapt, a question that can be approached in the light of recent analyses of the Transformer architecture which conjectured that the higher layers encode more global patterns with a more ``semantic'' interpretation, while the lower layers encode more local patterns akin to morpho-syntactic information \cite{raganato18analysis}.
+
+% The size of available corpora for each domain is extremely varying.
 For example, the corpus of domain \domain{tourism} in En-De contains only around $7\times10^3$ sentence pairs while the corpus of domain \domain{news} in En-De contains approximately $3\times10^6$ sentence pairs.
@@
 
-Fine-tuning on small corpora can still lead to an overfitted model even when the fine-tuning is applied to a few residual adapters.
-To avoid this situation, we have several options including: reducing the size of the adapter module, applying weight decay or layer regularization to the adapter.
-The first option is not flexible \fyTodo{compatible ???} if we consider the probable growth of in-domain data.
-The second and third choices are preferable to regularize a large adapter which will scale if the in-domain corpora extend .\fyTodo{Style}
-
-Weight decay applies a penalization on the weights of the adapters.\fyDone{The second summation also runs over $x,y$ ? I think not}
-\begin{equation}
+A related question concerns the regularization of adapter layers to mitigate overfitting.
+Reducing the number of adapters, or their dimensions, is simple, but these choices are difficult to optimize numerically -- an issue that becomes important as the number of domain grows.
+Less naïve alternatives can also be entertained, such as applying weight decay or layer regularization to the adapter.
+Implementing these requires to modify the objective function in a way that still allows for a smooth optimization problem.
+For instance, weight decay applies a penalization on the weights of the adapters, complementing the cross-entropy term with a function of the norm of the parameters: \fyDone{The second summation also runs over $x,y$ ?
+I think not}
+\begin{equation*}
   \begin{split}
     \bar{L} & = \frac{1}{\#(x,y)}\mathop{\sum}_{x,y}( - \log(P(y|x))) \\
-    & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{\operatorname{ADAP}^{(i)}}}
+    & + \lambda  \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{\operatorname{ADAP}^{(i)}}}
   \end{split}
-\end{equation}
-Where $\theta_{\operatorname{ADAP}^{(i)}}$ are parameters of $i^{\text{th}}$ adapter
-
-Layer regularization applies a penalization on the output of the adapters.
-\begin{equation}
+\end{equation*}
+An alternative scheme is \emph{layer regularization}, which penalizes the output of the adapters, corresponding to the following new training objective:
+\begin{equation*}
   \begin{split}
     \bar{L} & = \frac{1}{\#(x,y)}\mathop{\sum}_{x,y} (-\log(P(y|x)) \\
-    & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\operatorname{ADAP}^{(i)}(h_i(x,y))})
+    & + \lambda \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\operatorname{ADAP}^{(i)}(h_i(x,y))})
   \end{split}
+\end{equation*}
+
+Finally, another independant design choice relates to the training strategy for adapters.
+A first option is to generalize supervised domain adaptation to multi-domain adaptation and to proceed in two steps: (a) train a generic model with all the available data; (b) train eacj adapter layer with domain-specific data, keeping the generic model parameters unchanged.
+Another strategy is to adopt the view of \citet{Dredze08online}, where the multi-domain setting is viewed as an instance of multi-task learning \cite{Caruana97multitask}, where each domain corresponds to a specific task.
+This suggest to train the models as it were in a multi-task mode, training all the parameters from scratch: the generic parameters values will still depend on all the available data, while each adapter will only be trained with the corresponding in-domain data.
+
+\subsection{Highway Residual Adapters \label{ssec:highway}}
+
+In the basic architecture described in Section~\ref{ssec:architecture}, the computation performed by lower level layers will impact all the subsequent layers.
+In this section, we introduce an alternative implementation of the same idea, which however delays the adaptation of each layer to the last layer (of the encoder or the decoder) as represented on Figure~\ref{fig:hrl-architecture}.
+While the basic architecture performs adaptation in sequence, we propose here to perform it in parallel.
+This means that in this version, only the last hidden vector of the encoder (decoder) is modified according to:
+\begin{equation}
+  \bar{h}^L = h^L + \displaystyle{\mathop{\sum}_{1 \leq i \leq L} ADAP^i(h^i)} \label{eq:highway-output}
 \end{equation}
 
-In a multi-domain context, there are 2 ways to train domain-specific NMT model using adapters including pretraining generic NMT then finetuning adapters; or training both NMT model and adapters at once from the beginning using multi-task learning.
-In the literature of Machine Learning, \cite{Caruana97multitask} proposed an architecture suitable for multi-task learning.
-The corresponding network comprises a ground layer, which is shared between tasks and is followed by task-related layers.
-The network is then trained from scratch by fueling task-related examples in a batch and updating all parameters (shared and task-specific) at the same time.
-The multi-domain context is similar to the multi-task context if we consider a domain as a task\cite{Dredze08online}.
-Therefore, we can also use multi-task training to optimize the parameters of NMT model and the parameters of the residual adapters at once from the beginning.
-To assess the efficacy of multi-task training in the multi-domain context, we would like to compare the finetuning approach with the multi-task training approach.
-\fyTodo{Change the training regime}
+One obvious benefit of this variant is that it allows us to reuse the hidden vectors $h^i$ of all hidden layers when computing an adapted output for several domains during the inference.
+In this situation, the forward step needs only to compute the hidden vectors $h^i$ once for the inner encoder layers, before an adapted sequence of vectors is computed at the topmost layer.
+Therefore, we can finetune the model to multiple domains at once without recomputing $h^i$.
+This variant also opens the way to more parameter sharing across adapters, a perspective that we will not explore further in this work.
+Instead, we use it to develop a second variation of the adapter model, that is presented in the next section.
 
-\subsection{Highway Residual Adapters \label{ssec:highway}}
-Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of attention layers.
-We graphically demonstrate this version in Figure~\ref{fig:hrl-architecture}.
-In this version, only the last hidden vector of the encoder (decoder) is modified as follow
-\begin{equation}
-\bar{h}^n = h^n + \displaystyle{\mathop{\sum}_{1 \leq i \leq n} ADAP^i(h^i)}
-\end{equation}
-The benefit of this variant is that it allows us to reuse hidden vectors $h^i$ of all hidden layers while adapting to multiple domains.
-In the forward step, the hidden vectors $h^i$ are calculated once, then adapters take these vectors to modify the last layer.
-Therefore, we can finetune the model to multiple domains at once without recomputing $h^i$.
-The second minor benefit of this variant is that it allows implementing the Gated Residual Adaptaters version that will be described in the following section \ref{ssec:gate}.
 \begin{figure}[htbp]
@@
 \end{figure}
-\subsection{Gated Residual Adaptaters \label{ssec:gate}}
+
+\subsection{Gated Residual Adapters \label{ssec:gate}}
 \mpTodo{Formalizing problem, network design, training algorithm}
-To translate a sentence, we have 2 options: using a domain-adapted model or using the generic model.
-In the case of choosing the wrong domain, the performance is dramatically worse than using the generic model.
-Therefore, we would like to design a version of applying residual adapter that allows us to automatically accept or reject given domain information, i.e, even if we choose a wrong adapter, the NMT model is still able to reject the change output from the adapter.
-In this version, we reuse the highway version of residual adapters.
-\fyTodo{Consistency of notations wrt section 2.1}
-Given domain $k$ with set of adapters $ADAP^i_k$, in the model adapted to domain k, instead of using the formulation $\bar{h}^n = h^n + \displaystyle{\mathop{\sum}_{1 \leq i \leq n} ADAP_k^i(h^i)}$ (encoder or decoder's side), we propose "gated" version uses $\bar{h}^n = h^n + \displaystyle{\mathop{\sum}_{1 \leq i \leq n} ADAP_k^i(h^i) \odot{} z_k(h^n)}$.
-Where $z_k(h^n(w))$ is designed to reflect the relatedness of word $w$ to domain $k$: the more likely $w$ is in domain $k$, the larger $z_k(h^n(w))$ should be.
-The gated version aims to reduce the change $\displaystyle{\mathop{\sum}_{1 \leq i \leq n} ADAP_k^i(h^i)}$ added to $h^n$ for words that are not typical of domain $k$.
-In our implementation, we incorporate 2 domain classifiers on top of the encoder and the decoder that take the last hidden layer of encoder (decoder) as input, we take the posterior probability $P(d=k|w)$ as value of $z_k(h^n(w))$.
+The basic architecture presented above rests on a rather simplistic view of ``domains'' as made of well-separated and unrelated pieces of texts that are processed independently during the adaptation step.
+Likewise, when translating test documents, we need to make a choice between either using a specific domain-adapted model or resorting to the generic, unadapted model.
+In this context, errors on the domain label can have a strong (negative) impact on the trained model, or on the translation performance.
 
-The training of Gated Residual Adaptaters comprises three main steps:
+Therefore, we would like to design a version of applying residual adapter that is more robust to such domains error.
+This variant, called the \emph{gated residual adapter models}, relies on the training of a supplementary trained component that will help decide whether to activate, on a word per word basis, a given residual layer, and to regulate the strength of this activation.
+To this end , we extend the highway version of residual adapters as follows.
+\fyDone{Consistency of notations wrt section 2.1}
+
+Formally, we replace the adapter computation of equation~\eqref{eq:highway-output} and take the adapted hidden (topmost) layer to be computed as (this is for domain $k$):
+\begin{equation}
+  \bar{h}^L = h^L + \displaystyle{\mathop{\sum}_{1 \leq i \leq L} \operatorname{ADAP}_k^i(h^i) \odot{} z_k(h^L)}, \label{eq:gated-output}
+\end{equation}
+where $z_k(h^L(w))$ measures the relatedness of word $w$ to the domain $k$.
+The more likely $w$ is in domain $k$, the larger $z_k(h^n(w))$ should be; conversely, for words that are not typical of any domain $k$ (eg.\ function words),  adaptation is minimum and the adapted corresponding encoder output will remain close to the output of the generic model ($h^L$).
+In our implementation, we incorporate two domain classifiers on top of the encoder and the decoder that take the last hidden layer of encoder (resp.\ decoder) as input, and use the posterior probability $P(d=k|w)$ as the value of $z_k(h^n(w))$.
+
+The training of gated residual adapters thus comprises three main steps:
 \begin{itemize}
-	\item Pretraining a generic model with mixed corpora comprising data from multiple domains.
-	\item Training a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen.
+\item Learn a generic model with mixed corpora comprising data from multiple domains.
+\item Train a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen.
 This model computes the posterior domain probability $p(k|h(w_t))$ for each word $w_t$ based on the representation computed by the last layer.
-	\item Training the parameters of adapters with in-domain data separately for each domain while freezing the parameters of the generic model and of the domain classifiers.
+\item Train the parameters of adapters with in-domain data separately for each domain, while freezing the generic model and the domain classifiers.
 \end{itemize}

### 2020-08-15 2020/wmt_20/emnlp2020.tex
start working on the experiments section

@@
 \end{equation}
-where $z_k(h^L(w))$ measures the relatedness of word $w$ to the domain $k$.
-The more likely $w$ is in domain $k$, the larger $z_k(h^n(w))$ should be; conversely, for words that are not typical of any domain $k$ (eg.\ function words),  adaptation is minimum and the adapted corresponding encoder output will remain close to the output of the generic model ($h^L$).
-In our implementation, we incorporate two domain classifiers on top of the encoder and the decoder that take the last hidden layer of encoder (resp.\ decoder) as input, and use the posterior probability $P(d=k|w)$ as the value of $z_k(h^n(w))$.
+where $z_k(h^L[t]) \in [0,1]$ measures the relatedness of the $t^{\text{th}}$ word $w_t$ to domain $k$.
+The more likely $w_t$ is in domain $k$, the larger $z_k(h^L[t])$ should be; conversely, for words\footnote{The term ``words'' here is employed by mere convenience as the system only manipulates sub-lexical BPE units; furthermore, the values of the hidden representations $h^{i}$ at position $t$ depends from all the other positions in the sentence.} that are not typical of any domain $k$ (eg.\ function words),  adaptation is minimum and the adapted corresponding encoder output ($\bar{h}^L[t]$) will remain close to the output of the generic model ($h^L[t]$).
+In our implementation, we incorporate two domain classifiers on top of the encoder and the decoder, that take the last hidden layer of the encoder (resp.\ decoder) as input and use the posterior probability $P(d=k|h^L[t]); \theta)$, with $\theta$ the set of all parameters, as the value of $z_k(h^L[t])$.
 
-The training of gated residual adapters thus comprises three main steps:
-\begin{itemize}
-\item Learn a generic model with mixed corpora comprising data from multiple domains.
-\item Train a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen.
-This model computes the posterior domain probability $p(k|h(w_t))$ for each word $w_t$ based on the representation computed by the last layer.
-\item Train the parameters of adapters with in-domain data separately for each domain, while freezing the generic model and the domain classifiers.
-\end{itemize}
+The training of gated residual adapters thus comprises three main steps, instead of two for the baseline version:
+\begin{enumerate}
+\item learn a generic model with mixed corpora comprising data from multiple domains.
+\item train a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen.
+This model computes the posterior domain probability $P(k|h^L[t]; \theta)$ for each word $w_t$, based on the representation computed by the last layer.
+\item train the parameters of adapters with in-domain data separately for each domain, while freezing the generic model and the domain classifiers.
+\end{enumerate}
+\fyFuture{is this classifier important, can we train with the rest of the system ?}
 
 \section{Experimental settings \label{sec:exp}}
+
 \subsection{Data and metrics \label{ssec:corpora}}
-We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
+We perform our experiments with two translation pairs involving multiple domains: English-French (En$\rightarrow$Fr and English-German (En$\rightarrow$De).
+For the former pair, we use texts initially from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
 Corpus statistics are in Table~\ref{tab:Corpora-en-fr}.
-Most corpora are available from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools.
-To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.%
+Most corpora are available from the Opus web site.\footnote{\url{http://opus.nlpl.eu}}
 
@@
   \end{tabular}
-\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain).
+\caption{Corpora statistics for En$\rightarrow$Fr : number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain).
 \domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.}
@@
 
-We also report a part of experiments in language pair English into German.
-We use available corpora in the News task of WMT20\footnote{\url{http://www.statmt.org/wmt20/news.html}} including: European Central Bank corpus (\domain{bank}),  European Economic and Social Committee corpus (\domain{eco}), European Medicines Agency corpus (\domain{med})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Press Release Database of European Commission corpus, News Commentary v15 corpus, Common Crawl corpus (\domain{news}), Europarl v10 (\domain{gouv}) Tilde MODEL - czechtourism (\domain{tourism})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Paracrawl and Wikipedia Matrix (\domain{others}) \footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}.
-We also build a joint BPE vocabulary of size 33K for both languages.
-The statistics for the En-De corpora are in Table~\ref{tab:Corpora-en-de}.
-
+For En$\rightarrow$De, we use corpora distributed for the News task of WMT20\footnote{\url{http://www.statmt.org/wmt20/news.html}} including: European Central Bank corpus (\domain{bank}),  European Economic and Social Committee corpus (\domain{eco}), European Medicines Agency corpus (\domain{med})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Press Release Database of European Commission corpus, News Commentary v15 corpus, Common Crawl corpus (\domain{news}), Europarl v10 (\domain{gouv}) Tilde MODEL - czechtourism (\domain{tourism})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Paracrawl and Wikipedia Matrix (\domain{others}) \footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}.
+Statistics are in Table~\ref{tab:Corpora-en-de}.
 \begin{table*}[htbp]
@@
     %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
-    \domain{bank} & \domain{eco} & \domain{med} & \domain{gouv} & \domain{news} & \domain{tourism} & \domain{other} \\
+    \domain{bank} & \domain{eco} & \domain{med} & \domain{gov} & \domain{news} & \domain{tourism} & \domain{other} \\
     \hline
-    4(0.00022) & 2857(0.15) & 347(0.018) & 1828(0.095) & 3696(0.19) & 7(0.00039) & 10473(0.54) \\
+    4(0.00022) & 2857 (0.15) & 347 (0.018) & 1828 (0.095) & 3696 (0.19) & 7 (0.00039) & 10473 (0.54) \\
     \hline
@@
 
-We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Development sets are used to choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
-We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
+We randomly select in each corpus a development and a test set of 1,000 lines each and keep the rest for training.
+Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+We report significant differences at the level of $p=0.05$.\fyTodo{Is this part still accurate ?}
 
-\subsection{Baseline architecture \label{ssec:baseline}}
+\subsection{Baseline architectures \label{ssec:baseline}}
 \fyTodo{Write this - settings and parameters for Mix-Nat and Full-FT}.
-Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
+% Our baselines are standard for multi-domain systems.
+% \footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.}
+Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we train the following baselines:
 \begin{itemize}
 \item a generic model trained on a concatenation of all corpora, denoted \system{Mixed}\fyTodo{Or mixed nat ?}
-\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing significantly in 3 consecutive epochs.
+\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing during 3 consecutive epochs.
 We again contrast two versions: full fine-tuning (\system{FT-Full}), which update all the parameters of the initial generic model \system{Mixed}; and the variant of \cite{Bapna19simple} (\system{FT-Block}).
@@
 
-For all models in En-Fr, we set the embeddings size and the hidden layers size to~512.
-Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
-The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of a 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
+For all models in En$\rightarrow{}$Fr, we set the embeddings size and the hidden layers size to~512.
+Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2,048 cells.
+Residual adapters additionally uses an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$.
 % The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
 % The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training use a batch size of~12288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
+Training use a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4,000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
 
-For all models in En-De, we set the embeddings size and the hidden layers size to~1024.
-Transformers use multi-head attention with 16 heads in each of the 6 layers; the inner feedforward layer contains 4096 cells.
-The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of a 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 2048.
-Training uses a batch size of~32768 tokens.
-
-The size of models are reported in the table \ref{tab:params}
+Models for En$\rightarrow{}$Fr are larger and rely on embeddings size as well as hidden layers size of~1024; each Transformers layer contains 16~attention heads; the inner feedforward layer contains 4,096 cells.
+Adapter modules have the same architecture as for the other language pair, except for their size, that is doubled ($b=2,048$).
+The number of parameters in each model is reported in Table~\ref{tab:params}
 
@@
 \end{table*}
+
 \subsection{Varying the positions and number of residual adapters}
-Tables \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} for En-De.
-The original version \system{FT-Block} outperforms the baseline \system{mixed} in almost all domains while stays equivalent to the baseline in \domain{med} in En-Fr and \domain{news}, \domain{gouv}, \domain{eco}, \domain{med} in En-De.
+Tables \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gov}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} for En-De.
+The original version \system{FT-Block} outperforms the baseline \system{mixed} in almost all domains while stays equivalent to the baseline in \domain{med} in En-Fr and \domain{news}, \domain{gov}, \domain{eco}, \domain{med} in En-De.
 However, \system{FT-full} still stay as best model in all domain tests.
@@
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{gouv}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tourism}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tourism}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
     \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
@@
   \end{tabular}
-  \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
+  \caption{Translation performance of various fine-tuned systems. We report BLEU scores for each domain, as well as averages.}
   \label{tab:performance-en-fr-reg}

### 2020-08-15 2020/wmt_20/emnlp2020.tex
corrections
more experimental results

@@
 The residual adapter architecture extends this model by adding one new computational module for each layer.
-The $i^{\text{th}}$ such residual adapter thus modifies $h^i$ by computing the following transformations:\fyDone{Use align env}
+The $i^{\text{th}}$ residual adapter thus modifies $h^i$ by computing the following transformations:\fyDone{Use align env}
 \begin{align*}
   h^{i}_1 &= \mathbf{W}_{db}^{i}h^{i} + b^i_{1} \\
-  h^{i}_2 &= \mathbf{ReLU}(h_1^{i}) \\
+  h^{i}_2 &= \mathbf{ReLU}(h_1^{i})
+  \\
+%\end{align*}
+%\begin{align*}
   h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + b^i_{2} \\
@@
 While in principle, all Transformer layers can be subject to adaptation, it is nonetheless worthwhile to consider simpler adaptation schemes, which would only alter a limited number of layers.
-Such strategy might be especially relevant when the training corpus contains very small domains, such as the ones considered in our experiments, and for which a complete adaptation may not be necessary or/and or prone to overfitting.
+Such strategy might be especially relevant when the training data contains very small domains, as in the experiments of Section~\ref{sec:exp}, and for which a complete adaptation may not be necessary or/and or prone to overfitting.
 Likewise, it might be meaningful to explore ways to share subsets of adapters across domains.
-This in turn raises the issues of which layer(s) to adapt, a question that can be approached in the light of recent analyses of the Transformer architecture which conjectured that the higher layers encode more global patterns with a more ``semantic'' interpretation, while the lower layers encode more local patterns akin to morpho-syntactic information \cite{raganato18analysis}.
+This in turn raises the issue of which layer(s) to adapt, a question that can be approached in the light of recent analyses of Transformers models, which conjecture that the higher layers encode global patterns with a more ``semantic'' interpretation, while the lower layers encode local patterns akin to morpho-syntactic information \cite{raganato18analysis}.
 
@@
 While the basic architecture performs adaptation in sequence, we propose here to perform it in parallel.
-This means that in this version, only the last hidden vector of the encoder (decoder) is modified according to:
+In this version, only the last hidden vector of the encoder (decoder) is thus modified according to:
 \begin{equation}
@@
   \includegraphics[scale=0.3]{fig/highway_residual}
-  \caption{Highway multi-domain network with residual layers}
+  \caption{Highway residual adapter network}
   \label{fig:hrl-architecture}
@@
 \mpTodo{Formalizing problem, network design, training algorithm}
-The basic architecture presented above rests on a rather simplistic view of ``domains'' as made of well-separated and unrelated pieces of texts that are processed independently during the adaptation step.
-Likewise, when translating test documents, we need to make a choice between either using a specific domain-adapted model or resorting to the generic, unadapted model.
-In this context, errors on the domain label can have a strong (negative) impact on the trained model, or on the translation performance.
+The basic architecture presented above rests on a rather simplistic view of ``domains'' as made of well-separated and unrelated pieces of texts that are processed independently during adaptation.
+Likewise, when translating test documents, one needs to make a choice between either using one specific domain-adapted model or resorting to the generic model.
+In this context, using wrong domain labels can have a strong (negative) on translation performance.
 
@@
 where $z_k(h^L[t]) \in [0,1]$ measures the relatedness of the $t^{\text{th}}$ word $w_t$ to domain $k$.
-The more likely $w_t$ is in domain $k$, the larger $z_k(h^L[t])$ should be; conversely, for words\footnote{The term ``words'' here is employed by mere convenience as the system only manipulates sub-lexical BPE units; furthermore, the values of the hidden representations $h^{i}$ at position $t$ depends from all the other positions in the sentence.} that are not typical of any domain $k$ (eg.\ function words),  adaptation is minimum and the adapted corresponding encoder output ($\bar{h}^L[t]$) will remain close to the output of the generic model ($h^L[t]$).
-In our implementation, we incorporate two domain classifiers on top of the encoder and the decoder, that take the last hidden layer of the encoder (resp.\ decoder) as input and use the posterior probability $P(d=k|h^L[t]); \theta)$, with $\theta$ the set of all parameters, as the value of $z_k(h^L[t])$.
+The more likely $w_t$ is in domain $k$, the larger $z_k(h^L[t])$ should be; conversely, for words\footnote{The term ``word'' is employed here by mere convenience, as the system only manipulates sub-lexical BPE units; furthermore, the values of the hidden representations $h^{i}$ at position $t$ depend upon all the other positions in the sentence.} that are not typical of any domain $k$ (eg.\ function words),  adaptation is minimum and the adapted corresponding encoder output ($\bar{h}^L[t]$) will remain close to the output of the generic model ($h^L[t]$).
+In our implementation, we incorporate two domain classifiers on top of the encoder and the decoder, that take the last hidden layer of the encoder (resp.\ decoder) as input and use the posterior probability $P(k|h^L[t]);)$ of domain $k$ as the value for $z_k(h^L[t])$.
 
-The training of gated residual adapters thus comprises three main steps, instead of two for the baseline version:
+Training gated residual adapters thus comprises three steps, instead of two for the baseline version:
 \begin{enumerate}
@@
 \item train a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen.
-This model computes the posterior domain probability $P(k|h^L[t]; \theta)$ for each word $w_t$, based on the representation computed by the last layer.
-\item train the parameters of adapters with in-domain data separately for each domain, while freezing the generic model and the domain classifiers.
+This model computes the posterior domain probability $P(k|h^L[t])$ for each word $w_t$, based on the representation computed by the last layer.
+\item train the parameters of adapters with in-domain data separately for each domain, while freezing all the other parameters.
 \end{enumerate}
@@
 We perform our experiments with two translation pairs involving multiple domains: English-French (En$\rightarrow$Fr and English-German (En$\rightarrow$De).
-For the former pair, we use texts initially from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
+For the former pair, we use texts\footnote{Most corpora are available from the Opus web site: \url{http://opus.nlpl.eu}} initially from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
 Corpus statistics are in Table~\ref{tab:Corpora-en-fr}.
-Most corpora are available from the Opus web site.\footnote{\url{http://opus.nlpl.eu}}
 
@@
 
-For En$\rightarrow$De, we use corpora distributed for the News task of WMT20\footnote{\url{http://www.statmt.org/wmt20/news.html}} including: European Central Bank corpus (\domain{bank}),  European Economic and Social Committee corpus (\domain{eco}), European Medicines Agency corpus (\domain{med})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Press Release Database of European Commission corpus, News Commentary v15 corpus, Common Crawl corpus (\domain{news}), Europarl v10 (\domain{gouv}) Tilde MODEL - czechtourism (\domain{tourism})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Paracrawl and Wikipedia Matrix (\domain{others}) \footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}.
+En$\rightarrow$De is a much larger task, for which we use corpora distributed for the News task of WMT20\footnote{\url{http://www.statmt.org/wmt20/news.html}} including: European Central Bank corpus (\domain{bank}),  European Economic and Social Committee corpus (\domain{eco}), European Medicines Agency corpus (\domain{med})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Press Release Database of European Commission corpus, News Commentary v15 corpus, Common Crawl corpus (\domain{news}), Europarl v10 (\domain{gov}), Tilde MODEL - czechtourism (\domain{tourism})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Paracrawl and Wikipedia Matrix (\domain{web}).
 Statistics are in Table~\ref{tab:Corpora-en-de}.
@@
     %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
-    \domain{bank} & \domain{eco} & \domain{med} & \domain{gov} & \domain{news} & \domain{tourism} & \domain{other} \\
+    \domain{bank} & \domain{eco} & \domain{med} & \domain{gov} & \domain{news} & \domain{tourism} & \domain{web} \\
     \hline
-    4(0.00022) & 2857 (0.15) & 347 (0.018) & 1828 (0.095) & 3696 (0.19) & 7 (0.00039) & 10473 (0.54) \\
+    4 (0.00022) & 2857 (0.15) & 347 (0.018) & 1828 (0.095) & 3696 (0.19) & 7 (0.00039) & 10473 (0.54) \\
     \hline
   \end{tabular}
-\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture.
-\domain{news} is the largest domain, containing almost 19\% of the sentences, while \domain{bank} is the smallest, with only 0.02\% of the data.}
+\caption{Corpora statistics for En$\rightarrow$De: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture.
+\domain{web} is the largest domain, containing about 54\% of the sentences, while \domain{bank} and \domain{tourism} are very small.}
 \label{tab:Corpora-en-de}
@@
 Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
-We report significant differences at the level of $p=0.05$.\fyTodo{Is this part still accurate ?}
+We report significant differences at the level of $p=0.05$.\fyTodo{Is this part on significance testing still accurate ?}
 
 \subsection{Baseline architectures \label{ssec:baseline}}
-\fyTodo{Write this - settings and parameters for Mix-Nat and Full-FT}.
+\fyTodo{Write this - settings and parameters for Mixed and Full-FT}.
 % Our baselines are standard for multi-domain systems.
@@
 \begin{itemize}
-\item a generic model trained on a concatenation of all corpora, denoted \system{Mixed}\fyTodo{Or mixed nat ?}
+\item a generic model trained on a concatenation of all corpora, denoted \system{Mixed}\fyDone{Or mixed nat ?}
 \item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing during 3 consecutive epochs.
@@
 
-For all models in En$\rightarrow{}$Fr, we set the embeddings size and the hidden layers size to~512.
+For all En$\rightarrow$Fr models, we set the embeddings size and the hidden layers size to~512.
 Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2,048 cells.
@@
 
-Models for En$\rightarrow{}$Fr are larger and rely on embeddings size as well as hidden layers size of~1024; each Transformers layer contains 16~attention heads; the inner feedforward layer contains 4,096 cells.
+Models for En$\rightarrow$De are larger and rely on embeddings size as well as hidden layers size of~1024; each Transformers layer contains 16~attention heads; the inner feedforward layer contains 4,096 cells.
 Adapter modules have the same architecture as for the other language pair, except for their size, that is doubled ($b=2,048$).
-The number of parameters in each model is reported in Table~\ref{tab:params}
+The number of parameters in each model is reported in Table~\ref{tab:params}.
 
@@
   \end{tabular}
-  \caption{Translation performance of various finetuned systems (En:Fr). We report BLEU scores for each domain, as well as averages.}\fyTodo{Boldface ?}
+  \caption{Translation performance of various fine-tuned systems (En:Fr). We report BLEU scores for each domain, as well as averages across languages.}\fyTodo{Boldface ?}
   \label{tab:performance-en-fr}
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tourism}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
-   \system{FT-Block} & 29.69 &	30.49 &	19.2 &	49.61 &	49.33 &	20.53 &	33.14     \\
-   \system{FT-Block-WD}     & 29.68 & 30.77 & 20.42 & 50.19 & 47.68 & 20.64 & 33.23 \\
-   \system{FT-Block-LR}     & 29.65 & 30.45 & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
+    \system{Mixed}  & 29.3 & 30.5 & 17.6 & 38.1 & 47.9 & 20.9  & 30.6 \\
+   \system{FT-Block} & 29.7 &	30.5 &	19.2 &	49.6 &	49.3 &	20.5 &	33.1     \\
+   \system{FT-Block-WD}     & 29.68 & 30.8 & 20.4 & 50.2 & 47.68 & 20.6 & 33.2 \\
+   \system{FT-Block-LR}     & 29.65 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 \\ 
      \hline
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{RND}} \\ \hline
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 & 49.4 \\
+    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 & 49.4 \\
     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} & 32.55 \\

### 2020-08-15 2020/wmt_20/emnlp2020.tex
more changes in the expe section

@@
 
-En$\rightarrow$De is a much larger task, for which we use corpora distributed for the News task of WMT20\footnote{\url{http://www.statmt.org/wmt20/news.html}} including: European Central Bank corpus (\domain{bank}),  European Economic and Social Committee corpus (\domain{eco}), European Medicines Agency corpus (\domain{med})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Press Release Database of European Commission corpus, News Commentary v15 corpus, Common Crawl corpus (\domain{news}), Europarl v10 (\domain{gov}), Tilde MODEL - czechtourism (\domain{tourism})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Paracrawl and Wikipedia Matrix (\domain{web}).
+En$\rightarrow$De is a much larger task, for which we use corpora distributed for the News task of WMT20\footnote{\url{http://www.statmt.org/wmt20/news.html}} including: European Central Bank corpus (\domain{bank}),  European Economic and Social Committee corpus (\domain{eco}), European Medicines Agency corpus (\domain{med})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Press Release Database of European Commission corpus, News Commentary v15 corpus, Common Crawl corpus (\domain{news}), Europarl v10 (\domain{gov}), Tilde MODEL - czechtourism (\domain{tour})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Paracrawl and Wikipedia Matrix (\domain{web}).
 Statistics are in Table~\ref{tab:Corpora-en-de}.
@@
     %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
-    \domain{bank} & \domain{eco} & \domain{med} & \domain{gov} & \domain{news} & \domain{tourism} & \domain{web} \\
+    \domain{bank} & \domain{eco} & \domain{med} & \domain{gov} & \domain{news} & \domain{tour} & \domain{web} \\
     \hline
@@
 \caption{Corpora statistics for En$\rightarrow$De: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture.
-\domain{web} is the largest domain, containing about 54\% of the sentences, while \domain{bank} and \domain{tourism} are very small.}
+\domain{web} is the largest domain, containing about 54\% of the sentences, while \domain{bank} and \domain{tour} are very small.}
 \label{tab:Corpora-en-de}
@@
 \subsection{Varying the positions and number of residual adapters}
-Tables \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gov}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} for En-De.
+Tables \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gov}, \domain{eco}, \domain{tour}, \domain{bank}, \domain{med} and \domain{news} for En-De.
 The original version \system{FT-Block} outperforms the baseline \system{mixed} in almost all domains while stays equivalent to the baseline in \domain{med} in En-Fr and \domain{news}, \domain{gov}, \domain{eco}, \domain{med} in En-De.
@@
   \end{tabular}
-  \caption{Translation performance of various fine-tuned systems (En:Fr). We report BLEU scores for each domain, as well as averages across languages.}\fyTodo{Boldface ?}
+  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$Fr). We report BLEU scores for each domain, as well as averages across languages.}\fyTodo{Boldface ?}
   \label{tab:performance-en-fr}
@@
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tourism}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
     \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
@@
   \end{tabular}
-  \caption{Translation performance of various fine-tuned systems (En:De). We report BLEU scores for each domain, as well as averages across domains.}
+  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$De). We report BLEU scores for each domain, as well as averages across domains.}
   \label{tab:performance-en-de}
@@
 \subsection{Regularization of fine-tuning}
-In the translation from English into German, we find that domains \domain{tourism}, \domain{ecb} are extremely small and account for a very small percentage of the training data, with respective proportions of 0.039\% and 0.022\%.
+In the translation from English into German, we find that domains \domain{tour}, \domain{ecb} are extremely small and account for a very small percentage of the training data, with respective proportions of 0.039\% and 0.022\%.
 Fine-tuning on these domains can lead to serious overfitting.
@@
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tourism}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{med}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
     \system{Mixed}  & 29.3 & 30.5 & 17.6 & 38.1 & 47.9 & 20.9  & 30.6 \\
    \system{FT-Block} & 29.7 &	30.5 &	19.2 &	49.6 &	49.3 &	20.5 &	33.1     \\
-   \system{FT-Block-WD}     & 29.68 & 30.8 & 20.4 & 50.2 & 47.68 & 20.6 & 33.2 \\
-   \system{FT-Block-LR}     & 29.65 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 \\ 
+   \system{FT-Block-WD}     & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 \\
+   \system{FT-Block-LR}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 \\ 
      \hline
   \end{tabular}
-  \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
+  \caption{Translation performance of various fine-tuned systems. We report BLEU scores for each domain, as well as averages.}
   \label{tab:performance-en-de-reg}

### 2020-08-15 2020/wmt_20/emnlp2020.tex
corrections

--- 2020-08-15 2020/wmt_20/emnlp2020.tex
@@
 
-We choose optimal hyper-parameter $\lambda$ by grid search with a small set of values $\{ 10^{-3}, 10^{-4}, 10^{-5}$.
+We choose optimal hyper-parameter $\lambda$ by grid search with a small set of values $\{ 10^{-3}, 10^{-4}, 10^{-5} \}$.
 
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\% & 23.5\\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} \\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} \\ 
+    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\
+    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
+   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
    \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\
    \hline 
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1    \\
-    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} & 49.9   \\
-    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 49.8          \\
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} & 49.7   \\
-    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 43.4 \\ 
-    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 47.5 \\
-    \system{MDL-Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} \\
+    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5  & 50.1    \\
+    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 & 49.9   \\
+    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & 79.9 & 49.8          \\
+    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & 79.9 & 49.7   \\
+    \system{DM}             & 35.6 & 49.5  & 45.6 & 29.9 & 37.1 & 62.4 & 43.4 \\ 
+    \system{ADM}           & 36.4 & 53.5  & 48.3 & 32.0 & 41.5 & 73.4 & 47.5 \\
+    \system{MDL-Res}     & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
     \system{MDL-HW-Res}   & 37.38&	56.43&	52.13&	33.7&	44.81&	89.77&	52.4 \\ 
@@
 In the table \ref{tab:performance-random}, the column \domain{RND} stores the score of multi-domain models, that use domain information to translate, in the situation in which we assign random domain tag to each test sentence.
+
+Beside multi-domain approaches reported in previous section, we report model \system{FT-full-UEW} which is obtained by full-finetuning generic model using uniform elastic weight to avoid catatrophic forgetting.
+
 The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
@@
     \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 & 49.4 \\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} & 32.55 \\
+    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 &  53.8 & 32.55 \\
    \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} & 38.38 \\ \hline 

### 2020-08-15 2020/wmt_20/emnlp2020.tex

--- 2020-08-15 2020/wmt_20/emnlp2020.tex
@@
 \item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing during 3 consecutive epochs.
-We again contrast two versions: full fine-tuning (\system{FT-Full}), which update all the parameters of the initial generic model \system{Mixed}; and the variant of \cite{Bapna19simple} (\system{FT-Block}).
+We again contrast two versions: full fine-tuning (\system{FT-Full}), which update all the parameters of the initial generic model \system{Mixed}; and the variants of \cite{Bapna19simple} (\system{FT-Block}).
 \end{itemize}
@@
 Adapter modules have the same architecture as for the other language pair, except for their size, that is doubled ($b=2,048$).
-The number of parameters in each model is reported in Table~\ref{tab:params}.
+%The number of parameters in each model is reported in Table~\ref{tab:params}.
 
+\iffalse{
 \begin{table*}[htbp]
@@
 \end{table*}
+}
+\fi
+
+\subsection{Multi-domain systems}
+we compare several proposals from the literature on multi-domain MT to finetuning method. The methods included in our experiments are described as follow
+\begin{itemize}
+\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}) or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain-specific part (\system{LDR});
+\item the three proposals of \newcite{Britz17mixing}.
+\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
+The training uses reference tags and inference is performed with predicted tags, just like for regular target words.
+\system{DM} is a multi-task learner where a domain classifier is trained on top of the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
+These methods thus only use domain tags in training.
+\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connections make it possible to by-pass this module.
+Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
+\end{itemize}
+
+Because the limit of our computational resource, we limit our experiments in this section to one language pair En-Fr.
+
+The table \ref{tab:performance-multi} shows that the full-finetuning outperforms all other systems in the in-domain tests.
+\system{FT-Block} is able to reduce the margin in several domains, showing the efficacy of residual adapters.
+Multi-task training system \system{MDL-Res} gains important improvement compared to baseline \system{Mixed}.
+However, \system{MDL-Res} is outperformed by \system{FT-Block}.
+It means that individually finetuning generic model is better than multi-task training.
+However, \system{MDL-Res} is better than other methods in multi-domain learning including:  \system{DC-Tag},  \system{DC-Feat}, \system{LDR}, \system{TTM}, \system{DM}, \system{ADM}.
+
+\begin{table*}[htbp]
+  \centering
+  \fyDone{Fix column size}
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+%     &&&&&& \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
+    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\
+    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
+   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
+   \hline 
+    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5  & 50.1 \\
+    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 & 49.9 \\
+    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & 79.9 & 49.8 \\
+    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & 79.9 & 49.7  \\
+    \system{DM}             & 35.6 & 49.5  & 45.6 & 29.9 & 37.1 & 62.4 & 43.4 \\ 
+    \system{ADM}           & 36.4 & 53.5  & 48.3 & 32.0 & 41.5 & 73.4 & 47.5\\
+    \system{MDL-Res}     & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
+     \hline
+  \end{tabular}
+  \caption{Translation performance of various MDMT systems in En-Fr. }
+  \label{tab:performance-multi}
+\end{table*}
 
 \subsection{Varying the positions and number of residual adapters}
-Tables \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gov}, \domain{eco}, \domain{tour}, \domain{bank}, \domain{med} and \domain{news} for En-De.
+Tables \ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gov}, \domain{eco}, \domain{tour}, \domain{bank}, \domain{med} and \domain{news} for En-De.
 The original version \system{FT-Block} outperforms the baseline \system{mixed} in almost all domains while stays equivalent to the baseline in \domain{med} in En-Fr and \domain{news}, \domain{gov}, \domain{eco}, \domain{med} in En-De.
@@
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
-    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
-    \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
-    \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\ 
-    \system{FT-Block$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 \\
-    \system{FT-Block$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 \\
-    \system{FT-Block$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.4 & 88.7 & 52 \\
-   \system{FT-Block$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline
+    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 & 65M/0 \\
+    \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 & 65M/12M\\
+    \system{FT-Block-WD}     & 37.18 & 55.99 & 52.93 & 33.36 & 46.03 & 90.65 & 52.7 & 65M/12M \\
+   \system{FT-Block-LR}     & 37.45 & 56.09 & 51.84 & 33.29 & 45.02 & 89.7 & 52.2 & 65M/12M \\  
+    \system{FT-Block$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 & 65M/6M\\
+    \system{FT-Block$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 & 65M/2M \\
+    \system{FT-Block$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.4 & 88.7 & 52 & 65M/2M\\
+   \system{FT-Block$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 & 65M/2M\\
      \hline
   \end{tabular}
-  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$Fr). We report BLEU scores for each domain, as well as averages across languages.}\fyTodo{Boldface ?}
-  \label{tab:performance-en-fr}
+  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$Fr).
+We report BLEU scores for each domain, as well as averages across languages.
+In column \domain{params}, we report number of domain-agnostic/domain-specific parameters} \fyDone{Boldface ?}
+  \label{tab:performance-en-fr-pos-reg}
 \end{table*}
@@
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 \\
-    \system{FT-Full}       & 29.8 & 30.97 & 19.81 & 53.43 & 49.98 & 20.84 & 34.14 \\
-   \system{FT-Block}     & 29.65 & 30.45 & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 \\ 
-   \system{FT-HW-Block}   & 29.54 & 30.42 & 18.59 & 50.78 & 47.13 & 20.51 & 32.83 \\ 
-   \system{FT-Block$_{(6)}$}     & 29.47 & 30.39 & 18.13 & 49.14 & 46.95 & 20.45 & 32.42 \\
-   \system{FT-Block$_{(4)}$}     & 29.69 & 30.4 & 18.07 & 49.61 & 47.05 & 20.64 & 32.58 \\
-   \system{FT-Block$_{(2)}$}   & 29.64 & 30.4 & 18.29 & 49.41 & 46.71 & 20.59 & 32.51  \\
-   \system{FT-Block$_{(2,4,6)}$}  & 29.68  & 30.55 & 18.85 & 49.57 & 47.09 & 20.63 &  32.73  \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline
+    \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 & 213M/0M\\
+   \system{FT-Block}     & 29.65 & 30.45 & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 & 213M/96M \\
+   \system{FT-Block-WD}     & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 & 213M/96M \\
+   \system{FT-Block-LR}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 & 213M/96M\\
+   \system{FT-Block$_{(6)}$}     & 29.47 & 30.39 & 18.13 & 49.14 & 46.95 & 20.45 & 32.42 & 213M/16M \\
+   \system{FT-Block$_{(4)}$}     & 29.69 & 30.4 & 18.07 & 49.61 & 47.05 & 20.64 & 32.58 & 213/16M\\
+   \system{FT-Block$_{(2)}$}   & 29.64 & 30.4 & 18.29 & 49.41 & 46.71 & 20.59 & 32.51  & 213M/16M\\
+   \system{FT-Block$_{(2,4,6)}$}  & 29.68  & 30.55 & 18.85 & 49.57 & 47.09 & 20.63 &  32.73 & 213M/48M \\
      \hline
   \end{tabular}
-  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$De). We report BLEU scores for each domain, as well as averages across domains.}
-  \label{tab:performance-en-de}
+  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$De).
+We report BLEU scores for each domain, as well as averages across domains.
+In column \domain{params}, we report number of domain-agnostic/domain-specific parameters}
+  \label{tab:performance-en-de-pos-reg}
 \end{table*}
@@
 \fyTodo{Why HW worse than standard version ?} \fyTodo{Why is regularization not helping ? It helps for small domain - domain-specific regularization ??}
+
 \subsection{Regularization of fine-tuning}
@@
 
-We choose optimal hyper-parameter $\lambda$ by grid search with a small set of values $\{ 10^{-3}, 10^{-4}, 10^{-5}$.
+We choose optimal hyper-parameters: weight decay's coefficient, layer regularizer's coefficient $\lambda$ by grid search with a small set of values $\{ 10^{-3}, 10^{-4}, 10^{-5} \}$.
 
-Results in Table~\ref{tab:performance-en-de-reg} show that weight decay (\system{FT-Block-WD}) outperforms the more basic \system{FT-Block}. \fyTodo{How is the weight decay parameter set ?}
+Results in Table~\ref{tab:performance-en-de-pos-reg} show that weight decay (\system{FT-Block-WD}) outperforms the more basic \system{FT-Block} while Layer regularization does not show any improvement compared to \system{FT-Block}.
+\fyTodo{How is the weight decay parameter set ?}
 
-\iffalse{
-\begin{table*}[htbp]
-  \centering
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
-%     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
-   \system{FT-Block}     & 37.3	& 57.93 &	53.91 &	33.79 &	46.69 &	90.17 & 53.3 \\ 
-   \system{FT-Block-WD}     & 37.18 & 55.99 & 52.93 & 33.36 & 46.03 & 90.65 & 52.7 \\
-   \system{FT-Block-LR}     & 37.45 & 56.09 & 51.84 & 33.29 & 45.02 & 89.7 & 52.2 \\
-     \hline
-  \end{tabular}
-  \caption{Translation performance of various fine-tuned systems. We report BLEU scores for each domain, as well as averages.}
-  \label{tab:performance-en-fr-reg}
-\end{table*}
-}
-\fi
+\subsection{Highway and Gated Residual Adaptaters \label{ssec:gate-exp}}
 
-\begin{table*}[htbp]
-  \centering
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
-%     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{med}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed}  & 29.3 & 30.5 & 17.6 & 38.1 & 47.9 & 20.9  & 30.6 \\
-   \system{FT-Block} & 29.7 &	30.5 &	19.2 &	49.6 &	49.3 &	20.5 &	33.1     \\
-   \system{FT-Block-WD}     & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 \\
-   \system{FT-Block-LR}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 \\ 
-     \hline
-  \end{tabular}
-  \caption{Translation performance of various fine-tuned systems. We report BLEU scores for each domain, as well as averages.}
-  \label{tab:performance-en-de-reg}
-\end{table*}
+We assess the efficacy of proposed variants: Highway residual adapters \system{FT-HW-Block} and Gated residual adapter \system{FT-Gate-Block} by the same experiments in previous section except that we only perform them in language pair En-Fr.
 
-\subsection{Multi-task learning}
-To assess the effectiveness of multi-task learning in training both the NMT model and the residual adapters at once from scratch, we contrast our approach with several proposals from the literature on multi-domain MT including:
+To measure the robustness against out-of-domain examples, we perform translations with wrong domain information to trick the systems that need to know domain a priori.
+In the table \ref{tab:performance-random}, the column \domain{RND} stores the score of multi-domain models, that use domain information to translate, in the situation in which we assign random domain tag to each test sentence.
+Beside the \system{RND} test, we use a new domain \domain{news} to assess the performance of systems in out-of-domain examples.
+Given a sentence of \domain{news}, we predict the closest domain to it by choosing the lowest perplexity of the sentence among those computed by in-domain language models and use the corresponding adapters to translate the sentence.
 
-\begin{itemize}
-\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}) or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain-specific part (\system{LDR});
-\item the three proposals of \newcite{Britz17mixing}.
-\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
-The training uses reference tags and inference is performed with predicted tags, just like for regular target words.
-\system{DM} is a multi-task learner where a domain classifier is trained on top of the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
-These methods thus only use domain tags in training.
-\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connections make it possible to by-pass this module.
-Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
-\item highway variant trained with multi-task learning denoted \system{MDL-HW-Res}.
-\end{itemize}
+We also report model \system{FT-Full-UEW} which is obtained by full-finetuning generic model using uniform elastic weight to avoid catatrophic forgetting.
 
-The table \ref{tab:performance-multi} shows that \system{FT-Block} outperforms \system{MDL-Res}, \system{MDL-HW-Res}.
-It means that individually finetuning generic model is better than multi-task training.
-However, \system{MDL-Res} and \system{MDL-HW-Res} are better than other methods in multi-domain learning.
+Table \ref{tab:performance-random} shows that variants \system{FT-HW-Block} and \system{FT-Gate-Block} achieve equivalent performances to the original version \system{FT-Block} of \cite{Bapna19simple}.
+
+The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
+Our proposed model \system{FT-gate-block} maintains equivalent performance to the generic model in \system{RND} thanks to the application of adaptive weight on the output of the residual adapter.
+\system{FT-gate-block} avoids relying on the residual adapters when predicting examples that come from domain far from the domain of the adapters.
+
+In the new domain test \domain{news}, generic model outperforms other systems.
+Finetuning systems including \system{FT-Full}, \system{FT-Block} and \system{FT-HW-Block} underperforms generic model by large margin.
+More regularized variants including \system{FT-Full-UEW}, and \system{FT-Gate-Block} are able to reduce the gap of performance compared to generic.
+
+The column \domain{PROC} shows that highway residual adapter \system{FT-HW-Block} process examples faster than original version \system{FT-Block}.
 
@@
   \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+  \begin{tabular}{|p{3cm}|*{10}{r|}} \hline
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\% & 23.5\\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} \\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} \\ 
-   \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{RND}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{\domain{PROC}} \\ \hline
+    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 & 49.4 & 23.47 & 34s \\
+   \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 & 32.55 & 20.19 & 34s \\
+   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} & 38.38 & 20.51 & 22s\\ 
+   \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 & 36.62 & 20.22 & 19s \\
    \hline 
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1    \\
-    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} & 49.9   \\
-    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 49.8          \\
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} & 49.7   \\
-    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 43.4 \\ 
-    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 47.5 \\
-    \system{MDL-Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} \\
-    \system{MDL-HW-Res}   & 37.38&	56.43&	52.13&	33.7&	44.81&	89.77&	52.4 \\ 
-    \hfill MDL-Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 33.7 \\
+    \system{FT-Gate-Block}    & 37.98 &	57.46&	53.05&	33.55&	46.01&	90.07&	53.0&	48.96 & 22.51 & 21s \\
+    \system{FT-Full-UEW} &	37.94 &	55.97 &	52.1 &	33.71 &	44.91 &	89.08 &	52.3 &	46.48 & 22.11& 34s \\
      \hline
   \end{tabular}
-  \caption{Translation performance of various MDMT systems.}
-  \label{tab:performance-multi}
-\end{table*}
-
-We also report the number of parameters of each model in table \ref{tab:params-multi}
-
-\begin{table*}[htbp]
-  \centering
-  \begin{tabular}{|l|l|} \hline
-    Model & params \\ \hline 
-    \system{Mixed}  & 65000000 \\
-    \system{DC-Feat} & 65240000 \\
-    \system{DC-Tag}  & 65003072 \\
-    \system{LDR} & 66440000 \\
-    \system{ADM}  & 65000000 \\
-    \system{DM} & 65240000 \\
-    \system{MDL-Res}  & 77601344 \\
-    \system{FT-Block} & 77601344 \\
-     \hline
-  \end{tabular}
-  \caption{Number of parameters ($\times 10^6$)}
-  \label{tab:params-multi}
-\end{table*}
-
-\subsection{Gated Residual Adaptaters \label{ssec:gate-exp}}
-To measure the robustness against out-of-domain examples, we perform translations with wrong domain information to trick the systems that need to know domain a priori.
-In the table \ref{tab:performance-random}, the column \domain{RND} stores the score of multi-domain models, that use domain information to translate, in the situation in which we assign random domain tag to each test sentence.
-The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
-Our proposed model \system{FT-gate-block} maintains equivalent performance to the generic model in \system{RND} thanks to the application of adaptive weight on the output of the residual adapter.
-\system{FT-gate-block} avoids relying on the residual adapters when predicting examples that come from domain far from the domain of the adapters.
-
-\begin{table*}[htbp]
-  \centering
-  \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
-%     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{RND}} \\ \hline
-    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 & 49.4 \\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} & 32.55 \\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} & 38.38 \\ \hline 
-   \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 & 27.13\\ 
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1 & 35.92   \\
-    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} & 49.9 & 34.91 \\
-    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 49.8          & 37.01 \\
-    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} & 31.78 \\
-    \system{FT-gate-block}    & 37.98 &	57.46&	53.05&	33.55&	46.01&	90.07&	53.0&	48.96 \\
-    \system{FT-full-UEW} &	37.94 &	55.97 &	52.1 &	33.71 &	44.91 &	89.08 &	52.3 &	46.48 \\
-     \hline
-  \end{tabular}
-  \caption{Translation performance of various MDMT systems}
+  \caption{Translation performance of highway and gated variants. The column \domain{PROC} is time (in seconds) of 100 iterations with batch size 12288 tokens}
   \label{tab:performance-random}
@@
 Early approaches to the multi-domain setting were proposed for statistical MT, either considering multiple data sources (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}) or domains containing several topics \cite{Eidelman12topic,Hasler14dynamic-topic}.
-Two main strategies are considered: feature-based methods, where domain/topic labels are integratetd through supplementary feature and instance-based methods, involving a measure of similarity between train and test domains.
+Two main strategies are considered: feature-based methods, where domain/topic labels are integrated through supplementary feature and instance-based methods, involving a measure of similarity between train and test domains.
 
@@
 \mpTodo{discussion}
+Residual Adapters allows us to quickly adapt a model to a domain without creating lots of domain-specific parameters even though full finetuning still outperforms this method by an important gap in in-domain test sets.
+For extremely small domains, we are able to reduce the number of adapter incorporated to NMT model without losing performance, also using Weight-Decay in finetuning over small domains demonstrates significant improvement.
+The proposed highway variant shows an equivalent performance to the original version of \cite{Bapna19simple} which allows us to process hidden layer of NMT model only once.
+The extended version of highway residual adapter to gated residual adapter shows the benefit in the robustness against wrong domain prediction and out-of-domain examples.
 

### 2020-08-15 2020/wmt_20/emnlp2020.tex
completion in abstract

@@
 We contrast multiple implementations of this idea on two language pairs.
-Our main conclusions is ..\fyTodo{abstract to be continued}
+Our main conclusions is that residual adapter is fast and cheap method to adapt a model to a domain; reducing number of adapters incorporated to NMT model can achieve equivalent to using all adapters; residual adapters can be bridged directly to the end of the stack of layers without loosing performance.
+Our work also proposed a gated version that render NMT model more robust against out-of-domain examples.
+\fyTodo{abstract to be continued}
 

### 2020-08-15 2020/wmt_20/emnlp2020.tex
remove some Todos

--- 2020-08-15 2020/wmt_20/emnlp2020.tex
@@
 Standard fine-tuning however modifies all the network parameters, which makes this approach computationally costly and prone to overfitting.
-A recent, lightweigth approach, instead augments a baseline model with supplementary (small) adapter layers, keeping the rest of the mode unchanged.
+A recent, lightweight approach, instead augments a baseline model with supplementary (small) adapter layers, keeping the rest of the mode unchanged.
 This has the additional merit to leave the baseline model intact, and adaptable to multiple domains.
@@
 We contrast multiple implementations of this idea on two language pairs.
-Our main conclusions is ..\fyTodo{abstract to be continued}
+Our main conclusion is that residual adapter is a fast and cheap method to adapt a model to a domain; reducing the number of adapters incorporated to NMT model can achieve equivalent to using all adapters; residual adapters can be bridged directly to the end of the stack of layers without losing performance.
+Our work also proposed a gated version that renders the NMT model more robust against out-of-domain examples.
+\fyTodo{abstract to be continued}
 
@@
 \section{Introduction } \label{sec:intro}
-\mpTodo{write introduction} \fyDone{Citations in chronological order}\fyDone{Split long sentences}
+\mpDone{write introduction} \fyDone{Citations in chronological order}\fyDone{Split long sentences}
 Owing to multiple improvements, Neural Machine Translation(NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence,Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
@@
 \subsection{Gated Residual Adapters \label{ssec:gate}}
-\mpTodo{Formalizing problem, network design, training algorithm}
+\mpDone{Formalizing problem, network design, training algorithm}
 The basic architecture presented above rests on a rather simplistic view of ``domains'' as made of well-separated and unrelated pieces of texts that are processed independently during adaptation.
@@
 The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
-Our proposed model \system{FT-gate-block} maintains equivalent performance to the generic model in \system{RND} thanks to the application of adaptive weight on the output of the residual adapter.
-\system{FT-gate-block} avoids relying on the residual adapters when predicting examples that come from domain far from the domain of the adapters.
+Our proposed model \system{FT-Gate-Block} maintains equivalent performance to the generic model in \system{RND} thanks to the application of adaptive weight on the output of the residual adapter.
+\system{FT-Gate-Block} avoids relying on the residual adapters when predicting examples that come from domain far from the domain of the adapters.
 
@@
 \section{Discussion \label{sec:discussion}}
-\mpTodo{discussion}
+\mpDone{discussion}
 Residual Adapters allows us to quickly adapt a model to a domain without creating lots of domain-specific parameters even though full finetuning still outperforms this method by an important gap in in-domain test sets.
@@
 
-\section*{Acknowledgments}
+%\section{Acknowledgments}
 

### 2020-08-15 2020/wmt_20/emnlp2020.tex
correct spelling errors

@@
 Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2,048 cells.
-Residual adapters additionally uses an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$.
+Residual adapters additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$.
 % The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
@@
 Models for En$\rightarrow$De are larger and rely on embeddings size as well as hidden layers size of~1024; each Transformers layer contains 16~attention heads; the inner feedforward layer contains 4,096 cells.
-Adapter modules have the same architecture as for the other language pair, except for their size, that is doubled ($b=2,048$).
+Adapter modules have the same architecture as for the other language pair, except for their size, which is doubled ($b=2,048$).
 %The number of parameters in each model is reported in Table~\ref{tab:params}.
@@
 
-Because the limit of our computational resource, we limit our experiments in this section to one language pair En-Fr.
+Because of the limit of our computational resources, we limit our experiments in this section to one language pair En-Fr.
 
@@
 
-We assess the efficacy of proposed variants: Highway residual adapters \system{FT-HW-Block} and Gated residual adapter \system{FT-Gate-Block} by the same experiments in previous section except that we only perform them in language pair En-Fr.
+We assess the efficacy of proposed variants: Highway residual adapters \system{FT-HW-Block} and Gated residual adapter \system{FT-Gate-Block} by the same experiments in the previous section except that we only perform them in language pair En-Fr.
 
@@
 In the table \ref{tab:performance-random}, the column \domain{RND} stores the score of multi-domain models, that use domain information to translate, in the situation in which we assign random domain tag to each test sentence.
-Beside the \system{RND} test, we use a new domain \domain{news} to assess the performance of systems in out-of-domain examples.
+Besides the \system{RND} test, we use a new domain \domain{news} to assess the performance of systems in out-of-domain examples.
 Given a sentence of \domain{news}, we predict the closest domain to it by choosing the lowest perplexity of the sentence among those computed by in-domain language models and use the corresponding adapters to translate the sentence.
 
-We also report model \system{FT-Full-UEW} which is obtained by full-finetuning generic model using uniform elastic weight to avoid catatrophic forgetting.
+We also report model \system{FT-Full-UEW} which is obtained by full-finetuning the generic model using uniform elastic weight to avoid catastrophic forgetting.
 
@@
 
-The column \domain{PROC} shows that highway residual adapter \system{FT-HW-Block} process examples faster than original version \system{FT-Block}.
+The column \domain{PROC} shows that highway residual adapter \system{FT-HW-Block} process examples faster than the original version \system{FT-Block}.
 
@@
    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 & 32.55 & 20.19 & 34s \\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} & 38.38 & 20.51 & 22s\\ 
+   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2  & 53.3 & 38.38 & 20.51 & 22s\\ 
    \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 & 36.62 & 20.22 & 19s \\
@@
 Residual Adapters allows us to quickly adapt a model to a domain without creating lots of domain-specific parameters even though full finetuning still outperforms this method by an important gap in in-domain test sets.
-For extremely small domains, we are able to reduce the number of adapter incorporated to NMT model without losing performance, also using Weight-Decay in finetuning over small domains demonstrates significant improvement.
-The proposed highway variant shows an equivalent performance to the original version of \cite{Bapna19simple} which allows us to process hidden layer of NMT model only once.
-The extended version of highway residual adapter to gated residual adapter shows the benefit in the robustness against wrong domain prediction and out-of-domain examples.
+For extremely small domains, we are able to reduce the number of adapters incorporated to the NMT model without losing performance, also using Weight-Decay in finetuning over small domains demonstrates significant improvement.
+The proposed highway variant shows an equivalent performance to the original version of \cite{Bapna19simple} which allows us to process the hidden layer of the NMT model only once.
+The extended version of the highway residual adapter to gated residual adapter shows the benefit in the robustness against wrong domain prediction and out-of-domain examples.
 

### 2020-08-15 2020/wmt_20/emnlp2020.tex
End of abstract

@@
 We contrast multiple implementations of this idea on two language pairs.
-Our main conclusion is that residual adapter is a fast and cheap method to adapt a model to a domain; reducing the number of adapters incorporated to NMT model can achieve equivalent to using all adapters; residual adapters can be bridged directly to the end of the stack of layers without losing performance.
-Our work also proposed a gated version that renders the NMT model more robust against out-of-domain examples.
-\fyTodo{abstract to be continued}
+Our main conclusions are that residual adapters provide a fast and cheap method for supervised multi-domain adaptation; our two variants prove as effective as the original adapter model, and open perspective to also make adapted models more robust to label domain errors.
+\fyDone{abstract to be continued}
 

### 2020-08-15 2020/wmt_20/emnlp2020.tex
rewrote section 3

@@
 \renewcommand\textfraction{.1}
-\renewcommand\floatpagefraction{.95}
+\renewcommand\floatpagefraction{.8}
 
@@
 \section{Residual adapters \label{sec:res}}
+In this section, we describe the basic version of the residual adapter architectures, as well as two original variants of this model.
 
@@
 
-\subsubsection{Design space and variants}
+\subsubsection{Design space and variants \label{sssec:design-space}}
 This general architecture leaves open many design choices pertaining to the details of the network organization, the training procedure and the corresponding objective function.
@@
 \item a generic model trained on a concatenation of all corpora, denoted \system{Mixed}\fyDone{Or mixed nat ?}
-\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing during 3 consecutive epochs.
-We again contrast two versions: full fine-tuning (\system{FT-Full}), which update all the parameters of the initial generic model \system{Mixed}; and the variants of \cite{Bapna19simple} (\system{FT-Block}).
+\item a fine-tuned model \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing during 3 consecutive epochs.
+%  We again contrast two versions: full fine-tuning (\system{FT-Full}), which update all the parameters of the initial generic model \system{Mixed}; and the variants of \cite{Bapna19simple} (\system{FT-Block}).
 \end{itemize}
@@
 \subsection{Multi-domain systems}
-we compare several proposals from the literature on multi-domain MT to finetuning method. The methods included in our experiments are described as follow
-\begin{itemize}
-\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}) or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain-specific part (\system{LDR});
-\item the three proposals of \newcite{Britz17mixing}.
-\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
-The training uses reference tags and inference is performed with predicted tags, just like for regular target words.
-\system{DM} is a multi-task learner where a domain classifier is trained on top of the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
-These methods thus only use domain tags in training.
-\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connections make it possible to by-pass this module.
-Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
-\end{itemize}
+In this section, we evaluate several proposals from the literature on multi-domain adaptation and compare them to full fine-tuning on the one hand, and to two variants of the residual adapter architecture on the other hand.
 
-Because of the limit of our computational resources, we limit our experiments in this section to one language pair En-Fr.
-
-The table \ref{tab:performance-multi} shows that the full-finetuning outperforms all other systems in the in-domain tests.
-\system{FT-Block} is able to reduce the margin in several domains, showing the efficacy of residual adapters.
-Multi-task training system \system{MDL-Res} gains important improvement compared to baseline \system{Mixed}.
-However, \system{MDL-Res} is outperformed by \system{FT-Block}.
-It means that individually finetuning generic model is better than multi-task training.
-However, \system{MDL-Res} is better than other methods in multi-domain learning including:  \system{DC-Tag},  \system{DC-Feat}, \system{LDR}, \system{TTM}, \system{DM}, \system{ADM}.
-
-\begin{table*}[htbp]
+\begin{table*}[t!]
   \centering
-  \fyDone{Fix column size}
   \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
@@
     \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
-   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
     \hline 
@@
     \system{DM}            & 35.6  & 49.5  & 45.6 & 29.9 & 37.1 & 62.4 & 43.4   \\ 
-    \system{ADM}           & 36.4 & 53.5  & 48.3 & 32.0 & 41.5 & 73.4 & 47.5\\
-    \system{MDL-Res}     & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
+    \system{ADM}          & 36.4  & 53.5  & 48.3 & 32.0 & 41.5 & 73.4 & 47.5   \\
+    \hline
+    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
+    \system{Res-Adap-MT}  & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
     \hline
   \end{tabular}
-  \caption{Translation performance of various MDMT systems in En-Fr. }
+  \caption{Baseline translation performance of various multi-domain MT systems (En$\rightarrow$Fr).}
   \label{tab:performance-multi}
@@
 
+The reference methods included in our experiments are the following:
+\begin{itemize}
+\item a system using ``domain control'' \cite{Kobus17domaincontrol}.
+In this approach, domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}) or in the form of a supplementary feature for each word (\system{DC-Feat});
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain-specific part (\system{LDR});
+\item the three proposals of \newcite{Britz17mixing}.
+\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
+The training uses reference tags and inference is performed with predicted tags, just like for regular target words.
+\system{DM} is a multi-task learner where a domain classifier is trained on top of the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
+These methods thus only use domain labels in training.
+\end{itemize}
+
+The two variants of the residual adapter model included in this first round of experiment have presented in Section~\ref{ssec:architecture}: \system{Res-Adap} is the multi-domain version of the approach of \citet{Bapna19simple} based on a two step training procedure; while \system{Res-Adap-MT} is the ``multi-task'' version, where the parameters of the generic model and of the adapter  are learned simultaneously from scratch.
+
+Because of the limit of our computational resources, we restrict our experiments in this section to the En$\rightarrow$Fr task. Results are in Table~\ref{tab:performance-multi}.
+
+These results first show that full fine-tuning outperforms all other methods for the in-domain test sets.
+However, \system{Res-Adap} is able to reduce the gap with this approach for several domains, showing the effectiveness of residual adapters.
+The ``multi-task'' variant is slightly less effective in our experiments than the basic version, where optimization is performed in two steps.
+As it turns out, using residual adapters prove here to be much more effective on average than the other reference multi-domain systems; it is also much better than the generic system for translating data from known domains, outperforming the \system{Mixed} system by more than 4 BLEU points in average.
+Gains are especially large for small domains such as \domain{law} and \domain{rel}.
+
+% Multi-task training system \system{MDL-Res} gains important improvement compared to baseline \system{Mixed}.
+However, \system{MDL-Res} is outperformed by \system{Res-Adap}.
+It means that independently fine tuning the generic models is better than multi-task training.
+However, \system{MDL-Res} is better than other methods in multi-domain learning including:  \system{DC-Tag},  \system{DC-Feat}, \system{LDR}, \system{TTM}, \system{DM}, \system{ADM}.
+
 \subsection{Varying the positions and number of residual adapters}
-Tables \ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gov}, \domain{eco}, \domain{tour}, \domain{bank}, \domain{med} and \domain{news} for En-De.
-The original version \system{FT-Block} outperforms the baseline \system{mixed} in almost all domains while stays equivalent to the baseline in \domain{med} in En-Fr and \domain{news}, \domain{gov}, \domain{eco}, \domain{med} in En-De.
-However, \system{FT-full} still stay as best model in all domain tests.
 
-By varying the number and positions of residual adapters as aforementioned in section \ref{ssec:architecture}, we assess several variants of applying residual adapters.
-\fyTodo{Fix style here} Because the set of possible configurations is large, we only perform experiments on layers at position: 2, 4, 6 on each side of NMT model (encoder and decoder) which covers a wide range in the hierarchy of attention layers.
+Tables~\ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} report the BLEU scores for 6 domains in each language pair: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for En$\rightarrow$Fr; \domain{gov}, \domain{eco}, \domain{tour}, \domain{bank}, \domain{med} and \domain{news} for En$\rightarrow$De.
+We first see that for the latter direction, the basic version \system{Res-Adap} also outperforms the \system{mixed} baseline on average, with large gains for the small domains \domain{tour}, \domain{bank} and comparable results for the other domains.
+% in almost all domains while stays equivalent to the baseline in \domain{med} in En$\rightarrow$Fr and \domain{news}, \domain{gov}, \domain{eco}, \domain{med} in En$\rightarrow$De.
+   %    However, \system{FT-full} still remains the best model for all test domains.
+\fyFuture{Full FT for en->DE ? }
+
+By varying the number and position of residual adapters (see Section~\ref{ssec:architecture}), we then contrast several implementation of residual adapters.
+\fyDone{Fix style here} Because the set of possible configurations is large, we only perform experiments for layers $i= 2, 4, 6$ (both for the encoder and decoder).
+Two settings are considered: keeping just one adapter, or keeping the three.
+The trend is the same for the two language directions: suppressing adapters always decrease the overall performance, albeit by a small margin: having six adapters is better than three, which is better than keeping only one.
+With only one adapter active, we observe small, unsignificant changes in performance.
+Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.
+
+% which covers a wide range in the hierarchy of attention layers.
 To observe how much the number of residual adapters affects the performance of the adapted model.
@@
 
- In most cases, the performance increases with respect to the number of residual adapters used in architecture.
-The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only one level.
-However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyTodo{What does bold mean ?}\fyTodo{Number of parameters ?}
+%  In most cases, the performance increases with respect to the number of residual adapters used in architecture.
+The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{Res-Adap$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{Res-Adap$_{(2)}$}, \system{Res-Adap$_{(4)}$} , \system{Res-Adap$_{(6)}$} using residual adapter at only one level.
+However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyDone{What does bold mean ?}\fyDone{Number of parameters ?}
 
-\begin{table*}
+\begin{table*}[htbp]
   \centering
@@
     \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 & 65M/0 \\
-    \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 & 65M/12M\\
-    \system{FT-Block-WD}     & 37.18 & 55.99 & 52.93 & 33.36 & 46.03 & 90.65 & 52.7 & 65M/12M \\
-   \system{FT-Block-LR}     & 37.45 & 56.09 & 51.84 & 33.29 & 45.02 & 89.7 & 52.2 & 65M/12M \\  
-    \system{FT-Block$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 & 65M/6M\\
-    \system{FT-Block$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 & 65M/2M \\
-    \system{FT-Block$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.4 & 88.7 & 52 & 65M/2M\\
-   \system{FT-Block$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 & 65M/2M\\
+    \system{Res-Adap}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 & 65M/12M\\ \hline
+    \system{Res-Adap$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 & 65M/6M\\
+    \system{Res-Adap$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 & 65M/2M \\
+    \system{Res-Adap$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.4 & 88.7 & 52 & 65M/2M\\
+    \system{Res-Adap$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 & 65M/2M\\ \hline
+    \system{Res-Adap-WD}     & 37.2 & 56.0 & 52.9 & 33.4 & 46.0 & 90.6 & 52.7 & 65M/12M \\
+    \system{Res-Adap-LR}      & 37.4 & 56.1 & 51.8 & 33.3 & 45.0 & 89.7 & 52.2 & 65M/12M \\  
      \hline
@@
 We report BLEU scores for each domain, as well as averages across languages.
-In column \domain{params}, we report number of domain-agnostic/domain-specific parameters} \fyDone{Boldface ?}
-  \label{tab:performance-en-fr-pos-reg}
+Column \domain{params} reports the number of domain-agnostic/domain-specific parameters \label{tab:performance-en-fr-pos-reg}} \fyDone{Boldface ?}
 \end{table*}
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline
-    \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 & 213M/0M\\
-   \system{FT-Block}     & 29.65 & 30.45 & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 & 213M/96M \\
-   \system{FT-Block-WD}     & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 & 213M/96M \\
-   \system{FT-Block-LR}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 & 213M/96M\\
-   \system{FT-Block$_{(6)}$}     & 29.47 & 30.39 & 18.13 & 49.14 & 46.95 & 20.45 & 32.42 & 213M/16M \\
-   \system{FT-Block$_{(4)}$}     & 29.69 & 30.4 & 18.07 & 49.61 & 47.05 & 20.64 & 32.58 & 213/16M\\
-   \system{FT-Block$_{(2)}$}   & 29.64 & 30.4 & 18.29 & 49.41 & 46.71 & 20.59 & 32.51  & 213M/16M\\
-   \system{FT-Block$_{(2,4,6)}$}  & 29.68  & 30.55 & 18.85 & 49.57 & 47.09 & 20.63 &  32.73 & 213M/48M \\
+    \system{Mixed}          & 29.3 & 30.5 & 17.6 & 38.1 & 47.9 & 20.9  & 30.6 & 213M/0M\\
+   \system{Res-Adap}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 & 213M/96M \\ \hline
+    \system{Res-Adap$_{(2,4,6)}$} & 29.7  & 30.5 & 18.8 & 49.6 & 47.1 & 20.6 &  32.7 & 213M/48M \\ \hline
+    \system{Res-Adap$_{(6)}$}      & 29.5 & 30.4 & 18.1 & 49.1 & 46.9 & 20.4 & 32.4 & 213M/16M \\
+   \system{Res-Adap$_{(4)}$}       & 29.7 & 30.4 & 18.1 & 49.6 & 47.0 & 20.6 & 32.6 & 213/16M\\
+   \system{Res-Adap$_{(2)}$}       & 29.6 & 30.4 & 18.3 & 49.4 & 46.7 & 20.6 & 32.5  & 213M/16M\\
+    \system{Res-Adap-WD}         & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 & 213M/96M \\
+    \system{Res-Adap-LR}           & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1  & 213M/96M\\
     \hline
@@
 We report BLEU scores for each domain, as well as averages across domains.
-In column \domain{params}, we report number of domain-agnostic/domain-specific parameters}
+In column \domain{params}, we report the number of domain-agnostic/domain-specific parameters}
   \label{tab:performance-en-de-pos-reg}
@@
 
-\fyTodo{Why HW worse than standard version ?} \fyTodo{Why is regularization not helping ? It helps for small domain - domain-specific regularization ??}
+\subsection{Regularizing fine-tuning \label{ssec:regularization-exp}}
 
-\subsection{Regularization of fine-tuning}
-In the translation from English into German, we find that domains \domain{tour}, \domain{ecb} are extremely small and account for a very small percentage of the training data, with respective proportions of 0.039\% and 0.022\%.
+The translation from English into German includes two domains (\domain{tour} and \domain{bank}) that are extremely small and account only for a very small fraction of the training data (respectively for 0.039\% and 0.022\% of the total number of sentences).
 Fine-tuning on these domains can lead to serious overfitting.
-We assess two well-known regularization techniques for adapter modules, that could help avoid the overfitting issues: weight decay and layer regularization.
+We assess two well-known regularization techniques for adapter modules, that could help mitigate this problem: weight decay and layer regularization.
 
-We choose optimal hyper-parameters: weight decay's coefficient, layer regularizer's coefficient $\lambda$ by grid search with a small set of values $\{ 10^{-3}, 10^{-4}, 10^{-5} \}$.
+For each method, the optimal hyper-parameters (weight decay or layer regularization coefficient, see Section~\ref{sssec:design-space}) $\lambda$ by chosen by grid search among a small set of values ($\{ 10^{-3}, 10^{-4}, 10^{-5} \}$).
 
-Results in Table~\ref{tab:performance-en-de-pos-reg} show that weight decay (\system{FT-Block-WD}) outperforms the more basic \system{FT-Block} while Layer regularization does not show any improvement compared to \system{FT-Block}.
-\fyTodo{How is the weight decay parameter set ?}
+Results in Tables~\ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} show that regularizing the adapter model can positively affect the test performance for the smallest domains (this is especially clear for weight-decay (\system{Res-Adap-WD}) in En$\rightarrow$De), at the cost however of a small drop in performance for the other domains.
+Using layer regularization proves here to be comparatively less effective.
+Finding better ways to set the regularization parameters, for instance by varying $\lambda$ for each domain based on the available supervision data, is left for future work.
+\fyDone{How is the weight decay parameter set ?}
+\fyDone{Why is regularization not helping ? It helps for small domain - domain-specific regularization ??}
 
@@
 
-We assess the efficacy of proposed variants: Highway residual adapters \system{FT-HW-Block} and Gated residual adapter \system{FT-Gate-Block} by the same experiments in the previous section except that we only perform them in language pair En-Fr.
+We now turn to the evaluation of our new architectural variants: Highway residual adapters \system{Res-Adap-HW} on the one hand, and Gated residual adapters \system{Res-Adap-Gated} on the other hand.
+We use the same domains and settings as before, focusing here exclusively on the language direction En$\rightarrow$Fr.
 
-To measure the robustness against out-of-domain examples, we perform translations with wrong domain information to trick the systems that need to know domain a priori.
-In the table \ref{tab:performance-random}, the column \domain{RND} stores the score of multi-domain models, that use domain information to translate, in the situation in which we assign random domain tag to each test sentence.
-Besides the \system{RND} test, we use a new domain \domain{news} to assess the performance of systems in out-of-domain examples.
-Given a sentence of \domain{news}, we predict the closest domain to it by choosing the lowest perplexity of the sentence among those computed by in-domain language models and use the corresponding adapters to translate the sentence.
+To also evaluate the robustness with respect to out-of-domain examples, we perform two additional experiments.
+We first generate translations with erroneous (more precisely: randomly assigned) domain information: the corresponding results appear in Table~\ref{tab:performance-random} under the column \domain{rnd}.
+We also compute translation for a domain unseen in training (\domain{news}) as follows.
+For each sentence of this test set, we automatically evaluate the closest domain\footnote{As measured by the perplexity of a language model trained with only in-domain data.\fyFuture{More on this for reproducibility}.}, then use the predicted domain label to compute the translation.
+This is an error prone process, which also challenges the robustness of our multi-domain systems.
+Results are in Table~\ref{tab:performance-random}.
+\fyFuture{This is a baseline, and we have not discussed it - keep for next time: We also report model \system{FT-Full-UEW} which is obtained by full-finetuning the generic model using uniform elastic weight to avoid catastrophic forgetting.}
 
-We also report model \system{FT-Full-UEW} which is obtained by full-finetuning the generic model using uniform elastic weight to avoid catastrophic forgetting.
+A first observation is that for domains seen in training, our variants \system{Res-Adap-HW} and \system{Res-Adap-Gated} achieve BLEU scores that are on a par to those of the original version (\system{Res-Adap}), with insignificant variations across test sets.
+\fyFuture{Comment on processing time}
 
-Table \ref{tab:performance-random} shows that variants \system{FT-HW-Block} and \system{FT-Gate-Block} achieve equivalent performances to the original version \system{FT-Block} of \cite{Bapna19simple}.
+The two other settings are instructive in several ways: they first clearly illustrate that the brittleness of domain-adapted systems, for which large drops in performance (more than 15 BLEU points on average) are observed when the domain label is randomly chosen.
+Our gated variant however proves much more robust than the other adaptation strategy, and performs almost on par to the generic system for that test condition.
+The same trend holds for the unseen \domain{news} domain, with \system{Res-Adap-Gated} being the best domain adapted system in our set, outperforming the other variants by about 2 BLEU points.
 
-The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
+% The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
 Our proposed model \system{FT-Gate-Block} maintains equivalent performance to the generic model in \system{RND} thanks to the application of adaptive weight on the output of the residual adapter.
@@
 
-In the new domain test \domain{news}, generic model outperforms other systems.
-Finetuning systems including \system{FT-Full}, \system{FT-Block} and \system{FT-HW-Block} underperforms generic model by large margin.
+% In the new domain test \domain{news}, generic model outperforms other systems.
+Finetuning systems including \system{FT-Full}, \system{Res-Adap} and \system{FT-HW-Block} underperforms generic model by large margin.
 More regularized variants including \system{FT-Full-UEW}, and \system{FT-Gate-Block} are able to reduce the gap of performance compared to generic.
 
-The column \domain{PROC} shows that highway residual adapter \system{FT-HW-Block} process examples faster than the original version \system{FT-Block}.
+% The column \domain{PROC} shows that highway residual adapter \system{FT-HW-Block} process examples faster than the original version \system{Res-Adap}.
+
+% \fyTodo{Why HW worse than standard version ?}
 
@@
   \centering
-  \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{10}{r|}} \hline
+  \begin{tabular}{|p{3cm}|*{7}{r|}|r|r|} \hline
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{RND}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{\domain{PROC}} \\ \hline
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 & 49.4 & 23.47 & 34s \\
-   \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 & 32.55 & 20.19 & 34s \\
-   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2  & 53.3 & 38.38 & 20.51 & 22s\\ 
-   \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 & 36.62 & 20.22 & 19s \\
-   \hline 
-    \system{FT-Gate-Block}    & 37.98 &	57.46&	53.05&	33.55&	46.01&	90.07&	53.0&	48.96 & 22.51 & 21s \\
-    \system{FT-Full-UEW} &	37.94 &	55.97 &	52.1 &	33.71 &	44.91 &	89.08 &	52.3 &	46.48 & 22.11& 34s \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c||}{\domain{avg}} & \multicolumn{1}{c|}{\domain{rnd}} & \multicolumn{1}{c|}{\domain{news}} \\ \hline % & \multicolumn{1}{c|}{\domain{PROC}} \\ \hline
+    \system{Mixed}             & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5     &  49.4 & 49.4 & 23.5 \\ %& 34s \\ %
+    \system{FT-Full}             & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8   & 53.8 & 32.5 & 20.2  \\ %& 34s \\
+    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2   & 53.3 & 38.4 & 20.5 \\ %& 22s\\ 
+    \system{Res-Adap-HW}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91.0  & 53.1 & 36.6 & 20.2 \\ %& 19s \\
+    \system{Res-Adap-Gate}  & 38.0 & 57.5& 53.0 & 33.5 & 46.0 & 90.1  & 53.0 & 49.0 & 22.5 \\ %& 21s \\
+%    \system{FT-Full-UEW}      & 37.9 & 56.0 & 52.1 & 33.7 & 44.9 & 89.1 &	52.3 & 46.5 & 22.11 \\ %& 34s \\
     \hline
   \end{tabular}
-  \caption{Translation performance of highway and gated variants. The column \domain{PROC} is time (in seconds) of 100 iterations with batch size 12288 tokens}
+  \caption{Translation performance of highway and gated variants for En$\rightarrow$Fr.
+    % Column \domain{PROC} is time (in seconds) of 100 iterations with batch size 12,288 tokens
+  }
   \label{tab:performance-random}
 \end{table*}
+\fyFuture{Computation time should also report the time it takes for the complete tuning process, not one iteration which will be more or less the same}.
 
@@
 Similar ideas are developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and similarities are enforced through parameter sharing schemes.
-Parameter-sharing also lies at the core of the work by \citet{Jiang19multidomain}, who consider a Transformer model containing both domain-specific heads and domain-agnostic heads.
+Parameter-sharing also lies at the core of the work by \citet{Jiang19multidomain}, who consider a Transformer model containing both domain-specific and domain-agnostic heads.
 
@@
 
-\section{Discussion \label{sec:discussion}}
+\section{Conclusion and outlook \label{sec:discussion}}
 \mpDone{discussion}
-Residual Adapters allows us to quickly adapt a model to a domain without creating lots of domain-specific parameters even though full finetuning still outperforms this method by an important gap in in-domain test sets.
+In this paper, we have performed an experimental study of the residual adapter architecture in the context of multi-domain adaptation, where the goal is to build one single systems that (a) performs well for domain seen in training, ideally as well as full fine-tuning; (b) is also able to robustly handle translations for new, unseen domains.
+We have shown that this architecture allowed us to quickly adapt a model to a specific domain, delivering BLEU performance than are much better than the generic, mixed domain baseline, and close the gap with the full-finetuning approach, at a modest computational cost.
+Several new variants have been introduced and evaluated for two language directions: if none that able to clearly surpass the baseline residual adapter models, they provide directions for improving this model in practical settings: unbalanced data condition, noise in label domains, etc.
+In our future work, we would like to continue the development of the Gated variant, which, it seems to us, provides a flexible and robust tool to address the various challenges of multi-domain machine translation.
+
+% without creating lots of domain-specific parameters even though full finetuning still outperforms this method by an important gap in in-domain test sets.
 For extremely small domains, we are able to reduce the number of adapters incorporated to the NMT model without losing performance, also using Weight-Decay in finetuning over small domains demonstrates significant improvement.
@@
 
-%\section{Acknowledgments}
+%\section*Acknowledgments}
+\fyFuture{Thank you Jean Zay}
 

### 2020-08-15 2020/wmt_20/emnlp2020.tex
remove last comments

@@
 \end{align*}
-Overall, the  $i^{\text{th}}$ adapter is thus parameterized by matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, with $b$ the dimension of the adapter (with $d \gg b$)\fyTodo{Check this}.
+Overall, the  $i^{\text{th}}$ adapter is thus parameterized by matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, with $b$ the dimension of the adapter (with $d \gg b$)\fyDone{Check this}.
 For the sake of brevity, we will simply denote $h^{i}_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the corresponding set of parameters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}
@@
 We randomly select in each corpus a development and a test set of 1,000 lines each and keep the rest for training.
-Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
-We report significant differences at the level of $p=0.05$.\fyTodo{Is this part on significance testing still accurate ?}
+Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings}
+   %    Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+We report significant differences at the level of $p=0.05$.
+\fyFuture{Is this part on significance testing still accurate ?}
 
 \subsection{Baseline architectures \label{ssec:baseline}}
-\fyTodo{Write this - settings and parameters for Mixed and Full-FT}.
+\fyDone{Write this - settings and parameters for Mixed and Full-FT}.
 % Our baselines are standard for multi-domain systems.

### 2020-08-16 2020/wmt_20/emnlp2020.tex
local

--- 2020-08-15 2020/wmt_20/emnlp2020.tex
@@
 We contrast multiple implementations of this idea on two language pairs.
-Our main conclusions are that residual adapters provide a fast and cheap method for supervised multi-domain adaptation; our two variants prove as effective as the original adapter model, and open perspective to also make adapted models more robust to label domain errors.
-\fyDone{abstract to be continued}
+Our main conclusion is that residual adapter is a fast and cheap method to adapt a model to a domain; reducing the number of adapters incorporated to NMT model can achieve equivalent to using all adapters; residual adapters can be bridged directly to the end of the stack of layers without losing performance.
+Our work also proposed a gated version that renders the NMT model more robust against out-of-domain examples.
+\fyTodo{abstract to be continued}
 
@@
 \subsection{Varying the positions and number of residual adapters}
-Tables \ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gov}, \domain{eco}, \domain{tour}, \domain{bank}, \domain{med} and \domain{news} for En-De.
-The original version \system{FT-Block} outperforms the baseline \system{mixed} in almost all domains while stays equivalent to the baseline in \domain{med} in En-Fr and \domain{news}, \domain{gov}, \domain{eco}, \domain{med} in En-De.
-However, \system{FT-full} still stay as best model in all domain tests.
-
 By varying the number and positions of residual adapters as aforementioned in section \ref{ssec:architecture}, we assess several variants of applying residual adapters.
@@
 
- In most cases, the performance increases with respect to the number of residual adapters used in architecture.
-The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only one level.
+ Tables ~\ref{tab:performance-en-fr} and ~\ref{tab:performance-en-de-pos-reg} show that in most cases, the performance increases with respect to the number of residual adapters used in architecture.
+The setting \system{FT-Block} using residual adapters at all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only one level.
 However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyTodo{What does bold mean ?}\fyTodo{Number of parameters ?}
 
-\begin{table*}
+\begin{table*}[htbp]
   \centering
@@
 We report BLEU scores for each domain, as well as averages across languages.
-In column \domain{params}, we report number of domain-agnostic/domain-specific parameters} \fyDone{Boldface ?}
-  \label{tab:performance-en-fr-pos-reg}
+In column \domain{params}, we report number of domain-agnostic/domain-specific parameters}
+  \label{tab:performance-en-fr}
 \end{table*}

### 2020-08-16 2020/wmt_20/emnlp2020.tex

--- 2020-08-15 2020/wmt_20/emnlp2020.tex
@@
 \subsection{Varying the positions and number of residual adapters}
-
 Tables~\ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} report the BLEU scores for 6 domains in each language pair: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for En$\rightarrow$Fr; \domain{gov}, \domain{eco}, \domain{tour}, \domain{bank}, \domain{med} and \domain{news} for En$\rightarrow$De.
@@
 Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.
-
-% which covers a wide range in the hierarchy of attention layers.
-To observe how much the number of residual adapters affects the performance of the adapted model.
-By the limit of computation resources, we limit our study to the case of 3 adapters at positions: 2,4,6 in each side of the NMT model.
-By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting) in each side (encoder and decoder).\fyTodo{We have 12 layers, have we not ?}
-
-%  In most cases, the performance increases with respect to the number of residual adapters used in architecture.
-The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{Res-Adap$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{Res-Adap$_{(2)}$}, \system{Res-Adap$_{(4)}$} , \system{Res-Adap$_{(6)}$} using residual adapter at only one level.
-However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyDone{What does bold mean ?}\fyDone{Number of parameters ?}
 

### 2020-08-19 2020/tacl2020/rebuttal.tex
new name + new reference
starting working on reviewer A

--- 2020-07-07 2020/coling2020/rebuttal.tex
@@
 
-\title{Rebuttal: Revisiting Multi-Domain Machine Translation}
-\author{}
+\title{Responses to the reviews of TACL paper \#2327:
+  ``Revisiting Multi-Domain Machine Translation''
+}
+\author{Action Editor: G. Foster}
 \date{}
@@
 \maketitle
-\section*{General comments}
+
+\section*{Cover mail}
 As TACL action editor for submission 2327, "Revisiting Multi-Domain Machine Translation", I am happy to tell you that I am accepting your paper subject (conditional) to your making specific revisions within two months.
@@
 
+You are allowed one to two extra pages of content to accommodate these
+revisions. To submit your revised version, follow the instructions in the
+"Revision and Resubmission Policy for TACL Submissions" section of the
+Author Guidelines at
+\url{https://transacl.org/ojs/index.php/tacl/about/submissions\#authorGuidelines}.
+
+Thank you for submitting to TACL, and I look forward to your revised
+version! \\[2\parskip]
+
+\noindent George Foster\\
+\noindent Google Research\\
+\noindent geord.foster@gmail.com
+
 \section*{Reviewer A}
-Reviewer A:
-\\
+
 Overall, I'm a bit ambivalent about the paper.
@@
 \\
-Question to authors:
+\noindent Question to authors:
+\begin{itemize}
+\item can you describe in more detail the differences between FT-Block and MDL Res, apart from the fact that the former is fine-tuned from a generic model, the latter trained from scratch~?
+\end{itemize}
+\textcolor{red}{The two methods use the same architecture composed of one standard Transformer model extended with domain adapters following the design of \cite{Bapna19simple}.
+As the reviewer correctly points out, the only difference between FT-block and MDL Res is their training process and training data distribution.
   \\
-- can you describe in more detail the differences between FT-Block and MDL Res, apart from the fact that the former is fine-tuned from a generic model, the latter trained from scratch?
-\\
-\textcolor{red}{The two methods use the same architecture composing of one standard Transformer model and domain adapters presented in the paper of \cite{Bapna19simple}.
-The only difference between FT-block and MDL Res is their training process.
-\\
-To train FT-block, we plug residual adapters to the output of each self-attention layer of the model Mixed-NAT, then resume the training with the learning rate captured in the end of the training of the model Mixed-NAT while freezing all of the parameters of the model Mixed-NAT.
-We use only the in-domain data as the training data of the finetuning procedure.
+  To train FT-block, we plug residual adapters to the output of each self-attention layer of the generic model Mixed-NAT, then resume the training with the learning rate attained at the end of the training the generic model while freezing all of the other parameters of the model Mixed-NAT.
+For this step, we only use only the in-domain data, as we do during the full-tuning procedure.
   \\
   To train MDL res, we optimize the parameters of the Transformer model and the parameters of the residual adapters from scratch.
-Precisely, to create a training batch, we randomly pick a domain with a probability equal to the proportion of the domain in training data, then we sample a batch from its corpora.
-Given a batch from domain i, we optimize the parameters of Transformer model and the parameters of the residual adapters corresponding to the domain i by plugging the residual adapters of domain i to the Transformer model and calculating gradients of the activated parameters.}
+More precisely, to create a training batch, we randomly pick a domain with a probability equal to the proportion of the domain in training data, and then sample a batch from the corresponding corpora.
+Given a batch from domain $i$, we simultaneously optimize the parameters of Transformer model and the parameters of the residual adapters for that domain.
 \\
-minor issues:
-\\
-- I disagree that generalization to unseen domains is similar to zero-shot multilingual MT.
+
+  We understand that our naming schemes was insisting on procedures, rather than architectures, thus possibly introducing some confusions.
+We have accordingly changed the name of FT-Block to XXX\fyTodo{fix this}, clarified the description see lines YYY\fyTodo{fix this}, and reported FT-Block as another multidomain system in all tables.\fyTodo{Make sure it is the case}.
+}
+\noindent Minor issues:
+\begin{itemize}
+\item I disagree that generalization to unseen domains is similar to zero-shot multilingual MT.
 In zero-shot multilingual MT, both the source and target distribution have been observed at training time, just paired with other languages.
 In the generalization to unseen domains, there may be characteristics in the data that were never seen in the training data, including new vocabulary (in the source text, target text, or both).
-\\
-- 567-568: "having the domain tag on the source is better than on the
-target".
+\item 567-568: "having the domain tag on the source is better than on the target".
 I assume this refers to the comparison of "DC-Tag" and "TTM".
@@
 More precisely, the statement should be that "having the gold tag is better than a predicted tag", but this is hardly surprising.
-\\
+\end{itemize}
+
 \textcolor{red}{We agree that the conclusion is not really fair for "TTM" because "TTM" intend to predict domain label rather than to use gold domain label and TTM could definitively translate with gold domain label as prefix of each hypothesis.

### 2020-08-19 2020/tacl2020/rebuttal.tex
revision of the rebutal and of the main text

@@
 
+\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
+\newcommand{\system}[1]{\texttt{\textbf{#1}}}
+
 \title{Responses to the reviews of TACL paper \#2327:
@@
 
+The authors wish thank the reviewers and editor for their comments and suggestions, and have tried to address their questions and concerned in the revised version.
+Details are given below, with pointers to the main changes in the text.
+Typos and presentations errors have also been corrected in the new submission.
+
 \section*{Cover mail}
@@
 \begin{itemize}
-	\item Address the concern by reviewer A about the difference between MDL Res and FT-Block, and include FT-block (and ideally FT-Full also) in the later experiments (5.2.2 and 5.3).
-	\item Carefully specify the number of parameters for all models, including the proportions that are domain-specific and domain-independent.
-	\item Provide all the experimental details requested, and in particular give enough information to allow the corpora to be reproduced, or (ideally) make them publicly available.
+\item Address the concern by reviewer~A about the difference between \system{MDL Res} and \system{FT-Block}, and include \system{FT-block} (and ideally \system{FT-Full} also) in the later experiments (5.2.2 and 5.3).
+\\
+  \textcolor{blue}{discusion XXX.}
+\item Carefully specify the number of parameters for all models, including the proportions that are domain-specific and domain-independent. \\
+  \textcolor{blue}{New table ...}
+\item Provide all the experimental details requested, and in particular give enough information to allow the corpora to be reproduced, or (ideally) make them publicly available.\\
+  \textcolor{blue}{Appendix ....}
 \end{itemize}
@@
 \item can you describe in more detail the differences between FT-Block and MDL Res, apart from the fact that the former is fine-tuned from a generic model, the latter trained from scratch~?
-\end{itemize}
-\textcolor{red}{The two methods use the same architecture composed of one standard Transformer model extended with domain adapters following the design of \cite{Bapna19simple}.
-As the reviewer correctly points out, the only difference between FT-block and MDL Res is their training process and training data distribution.
   \\
-  To train FT-block, we plug residual adapters to the output of each self-attention layer of the generic model Mixed-NAT, then resume the training with the learning rate attained at the end of the training the generic model while freezing all of the other parameters of the model Mixed-NAT.
+  \textcolor{red}{%
+    The two methods use the same architecture composed of one standard Transformer model extended with domain adapters following the design of \cite{Bapna19simple}.
+As the reviewer correctly points out, the only difference between \system{FT-block} and \system{MDL Res} is their training process and training data distribution.
+    \\
+    To train \system{FT-block}, we plug residual adapters to the output of each self-attention layer of the generic model Mixed-NAT, then resume the training with the learning rate attained at the end of the training the generic model while freezing all of the other parameters of the model \system{Mixed-NAT}.
 For this step, we only use only the in-domain data, as we do during the full-tuning procedure.
@@
     To train MDL res, we optimize the parameters of the Transformer model and the parameters of the residual adapters from scratch.
-More precisely, to create a training batch, we randomly pick a domain with a probability equal to the proportion of the domain in training data, and then sample a batch from the corresponding corpora.
+More precisely, to create a training batch, we randomly pick a domain with a probability equal to the proportion of the domain in training data, then sample a batch from the corresponding corpus.
 Given a batch from domain $i$, we simultaneously optimize the parameters of Transformer model and the parameters of the residual adapters for that domain.
     \\
+    We understand that our naming schemes was insisting on training procedures, rather than on architectures, thus possibly introducing some confusions.
+We have accordingly changed the name of \system{FT-Block} to XXX\fyTodo{fix this}, clarified the description of the two systems (see lines YYY\fyTodo{fix this}), and reported \system{FT-Block}\fyTodo{or new name} as another multi-domain system in all tables.\fyTodo{Make sure it is the case}.}
+  %}
+\end{itemize}
 
-  We understand that our naming schemes was insisting on procedures, rather than architectures, thus possibly introducing some confusions.
-We have accordingly changed the name of FT-Block to XXX\fyTodo{fix this}, clarified the description see lines YYY\fyTodo{fix this}, and reported FT-Block as another multidomain system in all tables.\fyTodo{Make sure it is the case}.
-}
-\noindent Minor issues:
+\noindent\textbf{Minor issues:}
 \begin{itemize}
@@
 In the generalization to unseen domains, there may be characteristics in the data that were never seen in the training data, including new vocabulary (in the source text, target text, or both).
+  \\
+  \textcolor{red}{We agree lines XXX}
 \item 567-568: "having the domain tag on the source is better than on the target".
@@
 More precisely, the statement should be that "having the gold tag is better than a predicted tag", but this is hardly surprising.
+  \\
+  \textcolor{red}{We agree that the conclusion is not really fair for "TTM" because "TTM" intend to predict domain label rather than to use gold domain label and TTM could definitively translate with gold domain label as prefix of each hypothesis.}
+    \fyTodo{Check prediction quality, moderate conclusion}
 \end{itemize}
 
-\textcolor{red}{We agree that the conclusion is not really fair for "TTM" because "TTM" intend to predict domain label rather than to use gold domain label and TTM could definitively translate with gold domain label as prefix of each hypothesis.
-We will report the performance of "TTM" when translating using gold domain label as prefix in target side.}
 \section*{Reviewer B}
@@
 \\
-* You provide some useful experimental results on various MDMT approaches.
+\begin{itemize}
+\item[*] You provide some useful experimental results on various MDMT approaches.
 To benefit from your results, we need access to the same segmentation of train/dev/test sets that I hope you will provide to the community.
-\fyTodo{to share data}
+\textcolor{red}{Supplementary details regarding the settings used in our experiments have been added in a new Appendix YYY.\fyTodo{Fix this.}.
+We will also distribute the detailed composition of our train / dev / test split upon acceptance of the paper.}
 \\
-* On your discussion on the number of parameters (L. 591), could you share the number of parameters per each MDMT approach? It will be helpful to see  the impact of the number of parameters on the MT 
-results.
+\item[*] On your discussion on the number of parameters (L.
+591), could you share the number of parameters per each MDMT approach~?
+It will be helpful to see  the impact of the number of parameters on the MT results.
+  \textcolor{red}{This information has been added in a new Table~ZZZ \fyTodo{Table number, page and line number and content}.}
 \\ 
-\textcolor{red}{The description of model in each MDMT approach is as follows:
-\begin{enumerate}
-	\item FT-full, Mixed-NAT, TTM, DC-Tag use medium Transformer model.
-	\item FT-block and MDL res use medium Transformer model, and residual layers with the bottleneck dimension of size 1024.
-\cite{Bapna19simple} showed that if the bottleneck dimension is large enough, the gap between FT-block and FT-full will be small.
-	\item ADM, DM use medium Tranformer model and a domain classifier composing of 3 dense layers of size $512 \times 2048$, $2048 \times 2048$ and $2048 \times domain\_num$.
-The two first layers of the classifier use relu as activation function, the last layer uses tanh as activation function.
-	\item DC-Feat uses medium Transformer model, and domain embeddings of size 4.
-Given a sentence of domain i in training batch, the embedding of domain i is concatenated to the embedding of each token in the sentence.
-	\item LDR uses medium Transformer model and for each token we introduce a LDR feature of size $4 \times domain\_num$.
-Given a sentence of domain i $i\in[1,..,K]$ in training batch, for each token of the sentence, the LDR units of the indexes outside of the range $[4(i-1),..,4i-1]$ are masked to 0, the masked LDR feature will be concatenated to the embedding of the token.
-	\item WDCNMT uses one layer GRU bidirectional as encoder and one layer GRU-conditional as decoder. The size of each hidden layer is 1024, the size of word embedding is 512.
-	\item Mixed-NAT-RNN uses one layer LSTM bidirectional as encoder and one layer LSTM as decoder. The size of each hidden layer is 1024, the size of word embedding is 512. 
-\end{enumerate}
-}
-\fyTodo{system's description}
-\\
-* Thinking about your seven operational requirements, I wonder if you could also define a domain or the minimum number of samples to specify a domain.
+\item[*] Thinking about your seven operational requirements, I wonder if you could also define a domain or the minimum number of samples to specify a domain.
 I like your selections of domains that you include multiple domains with varying sizes, but what we expect in extreme cases.
@@
 \\
-* Please define Lambda coefficients in lines 142 and 146.
+\item[*] Please define Lambda coefficients in lines 142 and 146.
 \\
 \textcolor{red}{We formalize the data distribution as a mixture of several in-domain distribution.
-$\lambda$ is the prior distribution on the set of domains.
-We hypothesize that training distribution is different from testing distribution, so we have 2 different mixtures.}
+$\lambda$ correspond to the mixing weights, that also define the prior distribution on domains.
+We hypothesize that training distribution is different from testing distribution, so we have need notation for two different mixtures.
+This has been clarified in the text (see line~XXX) \fyTodo{add lines}}
 \\
-* On dynamically changing the number of training and test domains, there is a new related work published recently that might worth citing it: Sharaf A, Hassan H, Daumé III H.
+\item[*] On dynamically changing the number of training and test domains, there is a new related work published recently that might worth citing it: Sharaf A, Hassan H, Daumé III H.
 Meta-Learning for Few-Shot NMT Adaptation.
 \\
-\textcolor{red}{This is a very interesting work that we would like to cite in our final version of the paper.}
+\textcolor{red}{This reference has been added to the related work section.}\fyTodo{Add citation}
 \\
-* I’d prefer to see the number of running words in Table 1, especially when the batch size is given as the number of tokens (L. 485).
+\item[*] I’d prefer to see the number of running words in Table~1, especially when the batch size is given as the number of tokens (L. 485).
   \\
-\fyTodo{to report number of tokens of the datasets}
+  \textcolor{blue}{The table has been updated accordingly.}\fyTodo{to report number of tokens of the datasets}
 \\
-* L. 503: It seems you refer to this method as LDR in the rest of the paper. Please make them consistent.
+\item[*] L. 503: It seems you refer to this method as LDR in the rest of the paper. Please make them consistent.
 \\
-\textcolor{red}{We have noted this remark. We agree to change the notation correctly.}
+\textcolor{red}{Thanks for pointing this out. We have made this notation more coherent.}
 \\
-* L. 576: Although you describe in footnote 11, it is better to mention that you compare WDCMT results with Mixed-NAT-RNN explicitly.
+\item[*] L. 576: Although you describe in footnote 11, it is better to mention that you compare WDCMT results with Mixed-NAT-RNN explicitly.
 \\
@@
 \\
-* L. 681-683: I don’t see it as a failure, and I think it is a property (requirement) as you describe in the definition of [P4-ERR], we expect deterioration in the presence of tag error. Am I right?
+\item[*] L.
+681-683: I don’t see it as a failure, and I think it is a property (requirement) as you describe in the definition of [P4-ERR], we expect deterioration in the presence of tag error.
+Am I right?
 \\
-* L.
+\item[* L].
 687: Please describe how you predict the domain tag, and please share the tag prediction error rate on each domain test set.
@@
 To predict domain tag of a sentence, we compute probability of the sentence according to each pretrained language model and choose the domain corresponding to the highest probability as domain tag of sentence.} \fyTodo{to compute error rate of domain prediction over each test set}
-\\
-I spotted a few typos in the text:
-\\
-* L 100: section 2 $\rightarrow$ Section 2
-\\
-* L 104: section 3 $\rightarrow$ Section 3
-\\
-* L 256: could ben $\rightarrow$ could be (?)
-\\
-* L 702, Table 4, Col. 5: MED1 $\rightarrow$ MED2
-\\
-* L 790: that $\rightarrow$ than
-\\
-* L 802-805: It’s better to bring DC-Feat results after DC-Tag to keep the consistency with other tables 
-\\
-* L 928: ge:en, do you mean de:en (German-English)? 
-\\
-* L 973: domains:this $\rightarrow$ domains this
-\\
-* L 979: drawed $\rightarrow$ drawn 
+\end{itemize}
+
 \section*{Reviewer D}

### 2020-08-19 2020/tacl2020/decision.tex
revision of the rebutal and of the main text

\documentclass[12pt,times,a4paper,twoside]{article}

\usepackage[utf8]{inputenc}	% Para caracteres en español
\usepackage[left=3cm,right=3cm,top=2cm,bottom=3cm]{geometry}
\usepackage{pagenote}

\usepackage{algorithm}
\usepackage{algorithmic}

\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{mathtools,xparse}
\usepackage{mathrsfs}

\usepackage{multirow,booktabs}
\usepackage[table]{xcolor}
\usepackage{fullpage}
\usepackage{lastpage}
\usepackage{enumitem}
\usepackage{fancyhdr}

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{calc}
% \usepackage{color,soul}
\usepackage{multicol}
\usepackage{cancel}
\usepackage[retainorgcmds]{IEEEtrantools}
% \usepackage{xcolor} -- implied by xcolor

\usepackage{hyperref}

\colorlet{shadecolor}{orange!15}
\parindent 0in
\parskip 12pt
\geometry{margin=1in, headsep=0.25in}
\setlength{\belowdisplayskip}{8pt} \setlength{\belowdisplayshortskip}{8pt}
\setlength{\abovedisplayskip}{8pt} \setlength{\abovedisplayshortskip}{8pt}
\setlist{nosep}
\setlength{\parskip}{0.1cm}
\setlength{\parindent}{1em}

\title{Original decision letter}
\author{}
\date{}

\begin{document}
\maketitle
% \section*{General comments}
As TACL action editor for submission 2327, "Revisiting Multi-Domain Machine Translation", I am happy to tell you that I am accepting your paper subject (conditional) to your making specific revisions within two months.

Please address all comments by the reviewers, paying special attention to the following:
\begin{itemize}
	\item Address the concern by reviewer A about the difference between MDL Res and FT-Block, and include FT-block (and ideally FT-Full also) in the later experiments (5.2.2 and 5.3).
	\item Carefully specify the number of parameters for all models, including the proportions that are domain-specific and domain-independent.
	\item Provide all the experimental details requested, and in particular give enough information to allow the corpora to be reproduced, or (ideally) make them publicly available.
\end{itemize}

You are allowed one to two extra pages of content to accommodate these
revisions. To submit your revised version, follow the instructions in the
"Revision and Resubmission Policy for TACL Submissions" section of the
Author Guidelines at
\url{https://transacl.org/ojs/index.php/tacl/about/submissions\#authorGuidelines}.

Thank you for submitting to TACL, and I look forward to your revised
version! \\[2\parskip]

\noindent George Foster\\
\noindent Google Research\\
\noindent geord.foster@gmail.com

\section*{Reviewer A}

Overall, I'm a bit ambivalent about the paper. It is a well-written paper with a carefully-executed set of experiments. An important contribution of the paper is to formulate a set of expected properties of multi-domain MT (MDMT) (against which existing approaches are evaluated), and I generally like the choice of properties and that the paper draws attention to them, for example systematically testing how multi-domain approaches behave when gold domain labels are suboptimal. However, it is ambitious to systematically compare domain adaptation or multi-domain approaches which were developed with various conditions in mind. In several cases, I get the feeling that in an attempt to be broad and systematic, the authors have drawn boundaries and distinctions where none need exist, or drew conclusions that are too general. Details follow below.
\\
The authors have chosen a setting where domain labels are generally
available at training and test time, and in the the empirical comparison, focus on approaches that use domain labels at training and test time (533-537), but this is only partially true: their comparison does include systems which only use domain labels at training time (TTM/DM/ADM/WDCMT), and one set of experiments operates with automatically induced domain labels based on k-means clustering of sentence representations. Systems which were designed to work without domain labels, such as that by Farajian et al.(2017) would be a very relevant comparison at least in the latter setting without gold domain labels.
\\
The authors consider fine-tuning separate from MDMT, and declare failure in that no MDMT system meets the expectation that it should outperform fine-tuning. I'm confused why fine-tuning is specifically excluded, since it meets many of the reasons for building MDMT systems. Authors do mention that it would be cheaper to develop a single system instead of having to maintain multiple systems, but they include "MDL Res", which has domain-specific adaptation modules and is very similar to "FT-Block" (a baseline that also combines a general model with domain-specific adaptation modules). The line
between a single system and multiple systems gets fuzzy here, and I see no clear motivation to draw this clear distinction and only consider "MDL Res" a multi-domain system.
\\
The large number of approaches considered is admirable. However, the depth of the discussion suffers as a consequence, and the paper risks making over-general statements about the weaknesses of different approaches, despite related work that has already addressed them. The authors conclude that fine-tuning "incurs large losses [when applied on] the small data condition", and fails to deteriorate gracefully in the presence of tag errors, issues that have been addressed by Miceli Barone et al. (2017) and Thompson et al. (2019), respectively. Similarly, the negative impact of adding data from a "new" domain seems slightly overblown. Firstly, this is probably highly dependent on data mixtures, length of training and regularization strategies, so I'd be wary of drawing general conclusions from a single set of experiments. Secondly, when looking at the models where such a performance deterioration is said to happen (DC-Tag, TTM, ADM; lines
770-771), this deterioration is actually small or non-existent when
comparing table 3 and table 5 averages (-0.2 BLEU for DC-Tag; -0.3 BLEU for TTM; +0.9 BLEU for ADM). \\

\noindent Question to authors:
\begin{itemize}
\item can you describe in more detail the differences between FT-Block and MDL Res, apart from the fact that the former is fine-tuned from a generic model, the latter trained from scratch?  
\end{itemize}

\noindent Minor issues:
\begin{itemize}
\item I disagree that generalization to unseen domains is similar to zero-shot multilingual MT. In zero-shot multilingual MT, both the source and target distribution have been observed at training time, just paired with other languages. In the generalization to unseen domains, there may be characteristics in the data that were never seen in the training data, including new vocabulary (in the source text, target text, or both).
\item 567-568: "having the domain tag on the source is better than on the target". I assume this refers to the comparison of "DC-Tag" and "TTM". However, TTM actually uses the predicted domain label on the target side, and could perform better with forced decoding of the gold domain label. More precisely, the statement should be that "having the gold tag is better than a predicted tag", but this is hardly surprising.
\end{itemize}

\section*{Reviewer B}
The paper presents a thorough review of Multi-Domain Machine Translation (MDMT) approaches, and the authors provide comprehensive experiments to study different aspects of MDMT. The innovation of this work can be seen in two main categories:
\begin{itemize}
\item suggesting a methodology to evaluate MDMT models based on seven properties that they defined in the paper.
\item empirical studies on eight recent MTMD approaches.
\end{itemize}
In general, the paper has written well, and it is a good read for those who are researching in the domain adaptation field. 
\\
I have a few comments:
\begin{itemize}
\item[*] You provide some useful experimental results on various MDMT approaches. To benefit from your results, we need access to the same segmentation of train/dev/test sets that I hope you will provide to the community.
\item[*] On your discussion on the number of parameters (L. 591), could you share the number of parameters per each MDMT approach? It will be helpful to see the impact of the number of parameters on the MT results.  * Thinking about your seven operational requirements, I wonder if you could also define a domain or the minimum number of samples to specify a domain. I like your selections of domains that you include multiple domains with varying sizes, but what we expect in extreme cases. Having your Split experiments on REL (Table 4), or providing further details on your last analysis in Section 5.3, if you have tiny clusters could help.
\item[*] Please define Lambda coefficients in lines 142 and 146.
\item[*] On dynamically changing the number of training and test domains, there is a new related work published recently that might worth citing it: Sharaf A, Hassan H, Daumé III H. Meta-Learning for Few-Shot NMT Adaptation.
\item[*] I’d prefer to see the number of running words in Table 1, especially when the batch size is given as the number of tokens (L. 485).
\item[*] L. 503: It seems you refer to this method as LDR in the rest of the paper. Please make them consistent.
\item[*] L. 576: Although you describe in footnote 11, it is better to mention that you compare WDCMT results with Mixed-NAT-RNN explicitly.
\item[*] L. 681-683: I don’t see it as a failure, and I think it is a property (requirement) as you describe in the definition of [P4-ERR], we expect deterioration in the presence of tag error. Am I right?
\item[*]  L. 687: Please describe how you predict the domain tag, and please share the tag prediction error rate on each domain test set. It is helpful to see the amount and the impact of the error on MDMT results.
\end{itemize}

I spotted a few typos in the text:
\begin{itemize}
\item[*] L 100: section 2 $\rightarrow$ Section 2
\item[*] L 104: section 3 $\rightarrow$ Section 3
\item[*] L 256: could ben $\rightarrow$ could be (?)
\item[*] L 702, Table 4, Col. 5: MED1 $\rightarrow$ MED2
\item[*] L 790: that $\rightarrow$ than
\item[*] L 802-805: It’s better to bring DC-Feat results after DC-Tag to keep the consistency with other tables
\item[*] L 928: ge:en, do you mean de:en (German-English)?
\item[*] L 973: domains:this $\rightarrow$ domains this
\item[*] L 979: drawed $\rightarrow$ drawn
\end{itemize}

\section*{Reviewer D}
The authors deliver on their promise given in the title and abstract of the paper by doing exactly what I was hoping for: They provide a dense but comprehensive review of approaches to multi-domain MT. In a commendably straight and structured manner, they identify requirements and challenges of multi-domain MT. And they then proceed with empirically evaluating lots of multi-domain MT approaches.
\\
The empirical evaluation replicates more than ten multi-domain translation approaches from the literature and directly compares them against each other on the English-French language pair. The experiments are conducted using a coherent scenario with respect to both the amount of training data and the amount and type of distinct domains that the authors test on.
\\
The authors come up with a series of experiments that carefully considers the initially identified requirements and challenges. The (seven) identified challenges fall under the three categories "effectiveness", "robustness", and "scalability". There can be competing objectives when building multi-domain translation engines, and the results and discussion in the paper reveal the strengths and weaknesses of the various approaches in the considered scenario.
\\
Even though I cannot easily name the one striking contribution of this work, it's quite obvious to me that many fellow researchers will benefit from the clear presentation of multi-domain MT as well as from the extensive empirical results provided in this paper. I'd like to see it published.
\\
Let me nevertheless formulate a few questions or points of criticism:
\begin{itemize}
\item I'd like to see some analysis of vocabulary coverages (such as cross-domain OOV rates), and of "seen errors" and "new sense errors" as defined by Irvine et al. in "Measuring machine translation errors in new domains" (TACL 2013).
\item Maybe the (possible) effect of early stopping (and the choice of the validation set) should be discussed? E.g., the validation set for finetuning can originate from one single domain for each finetuned model. The way you're doing early stopping may even have had an impact on the results in Table 5. Is your validation set always a concatenation of 1000 lines from the six domains, without NEWS? Are you actually always training until the early stopping criterion is reached, and are you then using the model from the best checkpoint? (There are contradictory statements in the paper about "50,000 additional iterations" (?) for the experiments in Table 5.)
\item Report the learning rate. Is it decreased for finetuning?
\item To some extent, I dislike Table 5. When not reading carefully, the difference values could be interpreted as differences wrt. scores in Table 3. (Except that the numbers would be wrong...) Differences wrt. to Table 3 are important and it took me a while to obtain an overview for myself. Furthermore, it should be immediately obvious which number is "resumed training" and which is training from scratch. (It isn't from the table caption alone.)
\item How is "Mixed-Bal" achieved? By means of upsampling of the corpora from the less highly represented domains?
\item For "DC-Feat", is the feature added for each word or for each subword token? Is it attached to the word (increasing the vocabulary size a lot) or does it remain a separate token (increasing the input length a lot)?
\item How is the domain-weighted average (WAVG) calculated? Is the weight determined by proportion of the respective domain in the training data?
\item Wasn't it possible to use standard test sets? The "REL" Bleu scores are suspiciously high and there might be considerable redundancy in the data, or rather: training data contaminated with test data. (Since duplicates are possible, have you verified that the test sentences don't still appear in the remaining training data?)
\end{itemize}

Other (minor) remarks:
\begin{itemize}
\item From the UFAL medical corpus, please say that you're using the indomain sections only (which is mostly PatTR Medical).
\item The "DC-Feat" and "DC-Tag" lines in Table 5 should be swapped in order to be in line with Tables 3 and 4.
\item l. 503: "LDC-Feat" $\rightarrow$ "LDR"
\item l. 108: "automating" $\rightarrow$ "automatic"
\item l. 183: "accross" $\rightarrow$ "across"
\item l. 208: "under-resource" $\rightarrow$ "under-resourced"
\item l. 256: "ben" $\rightarrow$ "be"
\item l. 310: "used MT" $\rightarrow$ "used in MT"
\item l. 379: "Corpus statistics in Table 1." $\rightarrow$ "Corpus statistics are given in Table 1."
\item l. 380: "variable" $\rightarrow$ "available"
\item l. 426: "the all others" $\rightarrow$ "all others"
\item l. 696: "suplementary" $\rightarrow$ "supplementary"
\item l. 979: "drawed" $\rightarrow$ "drawn"
\item l. 991: "may be" $\rightarrow$ "maybe"
\end{itemize}
\end{document}

### 2020-08-19 2020/tacl2020/multidomain.tex
revision of the rebutal and of the main text

--- 2020-07-07 2020/coling2020/multidomain.tex
@@
 
-A typical learning scenario in MT is to have access to samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$.
+A typical learning scenario in MT is to have access to samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$\textcolor{red}{, with $\{\lambda^{s}_d, d=1 \dots n_d\}$ the corresponding mixture weights ($\sum_d \lambda^{s}_d=1$)}.
 Multi-domain learning, as defined in \cite{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ of several domains, making the problem distinct from mere domain adaption.
@@
 To make our results comparable with the literature, we will only use the BLEU score \cite{Papineni02bleu} in Section~\ref{sec:experiments}, noting it may not be the best yardstick to assess subtle improvements of lexical choices that are often associated with domain adapted systems \cite{Irvine13measuring}.
-Other important figure of merits for MDMT systems are the computational training cost and the total number of parameters.\fyDone{Insert discussion about scores}
+Other important figures of merit for MDMT systems are the computational training cost and the total number of parameters.\fyDone{Insert discussion about scores}
 
@@
 % \subsubsection{Performance}
-A first expectation is that MDMT systems should perform well in the face of mixed-domain test data. This yields the following three requirements.
+A first expectation is that MDMT systems should perform well in the face of mixed-domain test data. From this, we derive the following three requirements.
 
@@
 A second set of requirements is related to the definition of a domain.
-As repeatedly pointed out in the literature, parallel corpora in MT are often collected opportunistically and the view that each corpus constitutes a single domain is often a gross approximation.\footnote{Two of our own ``domains'' actually comprise several subcorpora (IT and MED), see details in section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments.
+As repeatedly pointed out in the literature, parallel corpora in MT are often collected opportunistically and the view that each corpus constitutes a single domain is often a gross approximation.\footnote{Two of our own ``domains'' actually comprise several subcorpora (IT and MED), see details in Section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments.
 To challenge these requirements we propose evaluating the following requirements.
@@
 Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools.
-To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.%
+To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyTodo{Add \# number of tokens, also specificity ?}%
 
@@
   \centering
-  \begin{tabular}{ |lllllll|} %*{4}{|r|}}
+  \begin{tabular}{|l|rrrrrrr|} %*{4}{|r|}}
+    \cline{2-8} 
+    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
+    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} & \multicolumn{1}{c|}{\domain{news}} \\
     \hline 
-    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
-    \domain{med} & \domain{law} & \domain{bank} & \domain{it} & \domain{talk} & \domain{rel} & \domain{news} \\
-    \hline
-    2609 (0.68) & 190 (0.05)  & 501 (0.13) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
+    \# lines & 2609 (0.68) & 190 (0.05)  & 501 (0.13) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
+    \# \textcolor{red}{tokens} &  XX  & XX &  XX &  XX &  XX &  XX &  XX  \\
     \hline
   \end{tabular}
-\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain).
+\caption{Corpora statistics: number of tokens ($\times 10^6$), number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain).
 \domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.}
@@
 \item fine-tuned models \cite{Luong15stanford,Freitag16fast}\jcDone{i would use Luong and Manning 2015}, based on the \system{Mixed-Nat} system, further trained on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
-We again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a simple adaptation module that is added (with residual connexions) on top of every Transformer layer (\system{FT-Block}).
+We again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a simple adaptation module that is added (with residual connections) on top of every Transformer layer (\system{FT-Block}).
 \end{itemize}
@@
 \item the multi-domain model of \newcite{Zeng18multidomain} (\system{WDCMT}), where a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly improved, results.}\fyDone{Check this}
-\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connexions make it possible to by-pass this module.
+\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connections make it possible to by-pass this module.
   % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
@@
 
+\newpage\newpage
+\section*{Appendices}
+
+\subsection*{A - Description of multi-domain systems}
+
+The description of model in each MDMT approach is as follows:
+\begin{itemize}
+\item \system{Mixed-NAT}, \system{FT-full}, \system{TTM}, \system{DC-Tag} use the medium Transformer model (\fyTodo{ref})
+\item \system{FT-block} and \system{MDL res} use medium Transformer model, and residual layers with the bottleneck dimension of size 1024.
+\cite{Bapna19simple} showed that if the bottleneck dimension is large enough, the gap between \system{FT-block} and \system{FT-full} will be small.\fyTodo{This comment is confusing}.
+\item \system{ADM}, \system{DM} use medium Tranformer model and a domain classifier composing of 3 dense layers of size $512 \times 2048$, $2048 \times 2048$ and $2048 \times domain\_num$.
+The two first layers of the classifier use the $\operatorname{ReLU}()$ as activation function, the last layer uses $\tanh()$ as activation function.
+\item \system{DC-Feat} uses medium Transformer model and domain embeddings of size 4.
+Given a sentence of domain~$i$ in a training batch, the embedding of domain $i$ is concatenated to the embedding of each token in the sentence.
+\item \system{LDR} uses medium Transformer model and for each token we introduce a \system{LDR} feature of size $4 \times domain\_num$.
+Given a sentence of domain $i\in[1,..,K]$ in the training batch, for each token of the sentence, the \system{LDR} units of the indexes outside of the range $[4(i-1),..,4i-1]$ are masked to 0, the masked LDR feature will be concatenated to the embedding of the token.
+Details are in \cite{Pham19generic}.
+\item \system{WDCNMT} uses one layer GRU bidirectional as encoder and one layer GRU-conditional as decoder. The size of each hidden layer is 1024, the size of word embedding is 512.
+\item \system{Mixed-NAT-RNN} uses one layer LSTM bidirectional as encoder and one layer LSTM as decoder. The size of each hidden layer is 1024, the size of word embeddings is 512. 
+\end{itemize}
+
 \end{document}
@@
 \todos{}
-
-\appendix{}
 

### 2020-08-20 2020/tacl2020/rebuttal.tex
More fixes and responses in the rebuttal file

@@
 I like your selections of domains that you include multiple domains with varying sizes, but what we expect in extreme cases.
-Having your Split experiments on REL (Table 4), or providing further details on your last analysis in Section 5.3, if you have tiny clusters could help.
+Having your Split experiments on REL (Table 4), or providing further details on your last analysis in Section 5.3, if you have tiny clusters could help.\fyTodo{Answer this with more experiments ?}
 \\
@@
 \\
-\textcolor{red}{Thanks for pointing this out. We have made this notation more coherent.}
+\textcolor{red}{Thanks for pointing this out. We have made this notation more consistent.}
 \\
@@
 \\
-\fyTodo{to explicitly compare WDCNMT with Mixed-NAT-RNN}
+\textcolor{blue}{This has been added in the text (see line YYY).}\fyTodo{to explicitly compare WDCNMT with Mixed-NAT-RNN. This is in the legend - change presentations ?}
 \\
@@
 \\
+\textcolor{blue}{This is correct - a drop is performance is expected when using wrong domain tags; [P4-ERR] is however about robustness against domain errors and a good MDMT system should minimize this loss.
+This is not what we see for the systems that use domain tags in testing.}
+
 \item[* L].
@@
 \\
-\textcolor{red}{We train a language model in source language for each domain using respective source training corpus.
-To predict domain tag of a sentence, we compute probability of the sentence according to each pretrained language model and choose the domain corresponding to the highest probability as domain tag of sentence.} \fyTodo{to compute error rate of domain prediction over each test set}
+\textcolor{blue}{We train a language model in source language for each domain using respective source training corpus.
+To predict domain tag of a sentence, we compute probability of the sentence according to each pretrained language model and choose the domain corresponding to the highest probability as domain tag of sentence.
+This is now explained in footnote ZZZ.}\fyTodo{to compute error rate of domain prediction over each test set, check comment, add footnote number.}
 \end{itemize}
@@
 \\
-- I'd like to see some analysis of vocabulary coverages (such as
+\begin{itemize}
+\item I'd like to see some analysis of vocabulary coverages (such as
 cross-domain OOV rates), and of "seen errors" and "new sense errors" as defined by Irvine et al. in "Measuring machine translation errors in new domains" (TACL 2013).
 \\
-\fyTodo{extra experiments to do.}
+\textcolor{blue}{. See table ZZZ.}\fyTodo{extra experiments to do.}
 \\
-- Maybe the (possible) effect of early stopping (and the choice of the validation set) should be discussed? E.g., the validation set for
-finetuning can originate from one single domain for each finetuned model.
+\item Maybe the (possible) effect of early stopping (and the choice of the validation set) should be discussed~?
+E.g., the validation set for finetuning can originate from one single domain for each finetuned model.
 The way you're doing early stopping may even have had an impact on the results in Table 5.
-Is your validation set always a concatenation of 1000 lines from the six domains, without NEWS?
+Is your validation set always a concatenation of 1000 lines from the six domains, without NEWS~?
 Are you actually always training until the early stopping criterion is reached, and are you then using the model from the best checkpoint?
@@
 \textcolor{red}{For each domain, we create train/dev/test sets by randomly splitting the respective corpora.
-We maintain the size of validation sets and of test sets equal to 1000 lines for every domain.
-In the table 3 and 4, to train MDMT models except FT-full and FT-block we use a validation set which is a concatenation of 6 validation sets of 6 domains.
-In the table 5, to train MDMT models except FT-full and FT-block we concatenate the validation set of NEWS to the mixture of 6 validation sets.
-To train FT-full model and FT-block model we only use the in-domain validation set.
+We maintain the size of validation sets and of test sets equal to 1,000 lines for every domain.
+In Tables 3 and 4, to train MDMT models except FT-full and FT-block, we use a validation set which is a concatenation of the 6 validation sets of 6 domains.
+In Table 5, to train MDMT models except FT-full and FT-block we concatenate the validation set of NEWS to the mixture of 6 validation sets.
+To train FT-full model and FT-block model we only use the in-domain validation sets.
 We stop training if either the training iteration reach the maximal number of iterations or the score on the validation set does not increase in 3 consecutive evaluations.
@@
 \\
-- Report the learning rate. Is it decreased for finetuning?
+\item Report the learning rate. Is it decreased for finetuning?
 \\
-\textcolor{red}{The learning rate is determined by this formula $lr = d_{model}^{-0.5}*\min(step\_num^{-0.5},step\_num * warmup\_steps^{-1.5})$ where $d_{model}=512$,$warmup\_steps=4000$.
-In the finetuning procedure, we continue the training using the same learning rate schedule and the same $step\_num$}
+\textcolor{red}{The learning rate is determined according to $lr = d_{model}^{-0.5}*\min(step\_num^{-0.5},step\_num * warmup\_steps^{-1.5})$ where $d_{model}=512$,$warmup\_steps=4000$, as in the original Transformer paper.
+For the finetuning procedure, we continue training using the same learning rate schedule and the same $step\_num$.}\fyTodo{Fix this, add comments in appendix}
 \\
-- To some extent, I dislike Table 5. When not reading carefully, the
-difference values could be interpreted as differences wrt.
+\item To some extent, I dislike Table~5.
+When not reading carefully, the difference values could be interpreted as differences wrt.
 scores in Table 3.
@@
 \\
-\textcolor{red}{In the table 5, we report the difference of the model's performance in 2 situations including resuming training and training from beginning with 7 domains.
-We would like to report the difference wrt to table 3 in the later version of the paper.
-We agree that the comparison to table 3 is also important, because it could assess the impact of new domain to the previously learned domains.}
+\textcolor{red}{In Table 5, we report the difference in performance between 2 situations: resuming training and training from beginning with 7 domains.
+We would like to report the difference wrt to Table 3 in the later version of the paper.
+We agree that the comparison to Table 3 is also important, because helps assess the impact of a new domain to the previously learned domains.}
 \\
-- How is "Mixed-Bal" achieved? By means of upsampling of the corpora from the less highly represented domains?
+\item How is \system{Mixed-Bal} achieved? By means of upsampling of the corpora from the less highly represented domains?
 \\
-\textcolor{red}{In all of our experiments, to get a training batch, we first sample a domain from a set of available domains in training procedure with a fixed probability, the we sample a from the corpora corresponding to the previously picked domain.
-Therefore, between our notations, "Mixed-NAT" uses a domain probability equal to proportion of domain in training data, and "Mixed-Bal" uses a domain probability equal to uniform probability.}
+\textcolor{blue}{In all of our experiments, to get a training batch, we first sample a domain from a set of available domains in training procedure with some fixed probability, then we sample a batch from the corpora corresponding to the previously picked domain.
+\system{Mixed-NAT} samples a domain based one the observed proportions of the training data (in Table~1), whereas \system{Mixed-Bal} samples domains with uniform probability.}
 \\
-- For "DC-Feat", is the feature added for each word or for each subword token?
+\item For \system{DC-Feat}, is the feature added for each word or for each subword token?
 Is it attached to the word (increasing the vocabulary size a lot) or does it remain a separate token (increasing the input length a lot)?
 \\
-\textcolor{red}{DC-Feat is a small embedding representing the domain.
-It is concatenated to every word embedding whenever its corresponding domain is detected.
-The use of DC-Feat increase the size of the input embedding, but does not increase the vocabulary size nor the input length.}
+\textcolor{red}{\system{DC-Feat} uses a small fixed embedding representing the domain.
+It is concatenated to every word embedding.
+The use of \system{DC-Feat} increases the size of the input embedding, but does not increase the vocabulary size nor the input length.
+This is clarified in the Appendix containing supplementary implementation details.}
 \\
-- How is the domain-weighted average (WAVG) calculated? Is the weight determined by proportion of the respective domain in the training data?
+\item How is the domain-weighted average (WAVG) calculated? Is the weight determined by proportion of the respective domain in the training data?
 \\
@@
 \\
-- Wasn't it possible to use standard test sets?
+\item Wasn't it possible to use standard test sets?
 The "REL" Bleu scores are suspiciously high and there might be considerable redundancy in the data, or rather: training data contaminated with test data.
@@
 \\
-\textcolor{red}{There are standard test sets for the domains including medical, IT, news and talk, but we could not find any standard test sets for the other domains including religion, law and bank.
-We carefully eliminated every duplicate in all of the experimental corpora.
-In our opinion, the reason why the models achieve extremely high scores in the domain religion is because the text in this domain is much less variable.}
+\textcolor{red}{There are standard test sets for some domains such as medical, IT, news and talk, but we could not find any standard test sets for the other domains including religion, law and bank.
+We carefully eliminated every duplicate sentence in all the corpora.
+In our opinion, the reason why the models achieve extremely high scores in the \domain{rel} domain is because texts in this domain is very repetitive and often predictable.}\fyTodo{sentence length ?, diversity index of vocab ?}
 \\
 Other (minor) remarks:
-- From the UFAL medical corpus, please say that you're using the indomain sections only (which is mostly PatTR Medical).
+\item From the UFAL medical corpus, please say that you're using the indomain sections only (which is mostly PatTR Medical).
 \\
-\textcolor{red}{From the UFAL medical corpus of the language pair En-Fr, we select only sentences in categories including EMEA, ParTR, CESTA, ECDC.}
+\textcolor{red}{From the UFAL medical corpus of the language pair En-Fr, we select only sentences in categories including EMEA, ParTR, CESTA, ECDC. This is now reflected in the text (lines XXX)}
 \\
-- The "DC-Feat" and "DC-Tag" lines in Table 5 should be swapped in order to be in line with Tables 3 and 4.
+\item The \system{DC-Feat} and "\item{DC-Tag} lines in Table 5 should be swapped in order to be in line with Tables 3 and 4.
   \\
   \textcolor{red}{We agree to change the order of "DC-Feat" and "DC-Tag" to be consistent with tables 3 and 4.}
-\\
-- l. 503: "LDC-Feat" $\rightarrow$ "LDR"
-\\
-- l. 108: "automating" $\rightarrow$ "automatic"
-\\
-- l. 183: "accross" $\rightarrow$ "across"
-\\
-- l. 208: "under-resource" $\rightarrow$ "under-resourced"
-\\
-- l. 256: "ben" $\rightarrow$ "be"
-\\
-- l. 310: "used MT" $\rightarrow$ "used in MT"
-\\
-- l. 379: "Corpus statistics in Table 1." $\rightarrow$ "Corpus statistics are given in
-Table 1."
-\\
-- l. 380: "variable" $\rightarrow$ "available"
-\\
-- l. 426: "the all others" $\rightarrow$ "all others"
-\\
-- l. 696: "suplementary" $\rightarrow$ "supplementary"
-\\
-- l. 979: "drawed" $\rightarrow$ "drawn"
-\\
-- l. 991: "may be" $\rightarrow$ "maybe"
+\end{itemize}
+The remaining style issues have been corrected.
 

### 2020-08-20 2020/tacl2020/multidomain.tex
More fixes and responses in the rebuttal file

@@
 \newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{red}{#1}}}
+\newcommand{\revision}[1]{\textcolor{red}{#1}}
 
@@
 
-A typical learning scenario in MT is to have access to samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$\textcolor{red}{, with $\{\lambda^{s}_d, d=1 \dots n_d\}$ the corresponding mixture weights ($\sum_d \lambda^{s}_d=1$)}.
+A typical learning scenario in MT is to have access to samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$\revision{, with $\{\lambda^{s}_d, d=1 \dots n_d\}$ the corresponding mixture weights ($\sum_d \lambda^{s}_d=1$)}.
 Multi-domain learning, as defined in \cite{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ of several domains, making the problem distinct from mere domain adaption.
@@
 \paragraph{[P4-ERR]}
-A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, catastrophic drops in performance can be observerd\fyDone{cite \cite{McCloskey89catastrophic}}.
-This requirement can be measured by translating test texts for domain $A$ with erroneous or random domain tags and measure the correlated loss.
+A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, catastrophic drops in performance can be observed\fyDone{cite \cite{McCloskey89catastrophic}}.
+This requirement can be assessed by translating test texts with erroneous domain tags and reporting the correlated decrease of performance.
 
@@
 
-We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
+We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.
+\revision{We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
-Corpus statistics in Table~\ref{tab:Corpora}.
+Corpus statistics are in Table~\ref{tab:Corpora}.
 Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools.
@@
     \# lines & 2609 (0.68) & 190 (0.05)  & 501 (0.13) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
-    \# \textcolor{red}{tokens} &  XX  & XX &  XX &  XX &  XX &  XX &  XX  \\
+    \# \revision{tokens} &  XX  & XX &  XX &  XX &  XX &  XX &  XX  \\
     \hline
@@
 \item a generic model trained on a concatenation of all corpora (\texttt{Mixed}).
-We develop two versions\footnote{In fact three: to enable a fair comparison with WDCMT, a RNN-based variant is also trained and evaluated.} of this system, one where domain unbalance reflects the distribution of our parallel data (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
-The former is the best option when the train mixture $\mathcal{D}^s$ is also observed in testing; the latter should be used when the test distribution is uniform across domains.
-Accordingly, we report two aggregate scores: a weighted average reflecting the default distribution, and an unweighted average, meaning that test domains are equally important.
+We develop two versions\footnote{In fact three: to enable a fair comparison with WDCMT, a RNN-based variant is also trained and evaluated.
+\revision{This variant appears as \system{Mixed-Nat-RNN} in Table~\ref{tab:performance}}.} of this system, one where the domain unbalance reflects the distribution of our training data \revision{given in Table~\ref{tab:Corpora}} (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
+The former is the best option when the train mixture $\mathcal{D}^s$ is also expected in testing; the latter should be used when the test distribution is uniform across domains.
+Accordingly, we report two aggregate scores: a weighted average reflecting the training distribution, and an unweighted average, meaning that test domains are equally important.
 %   As explained above, depending whether MDMT are expected to perform specially well on large domains, or to treat all domains equally, one or the other training distribution should be prefered.\jcDone{All systems cand be learned following both Bal and Nat.
@@
 Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
-The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024.
+The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024.
 % The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
@@
 % All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
+
 \subsection{Multi-domain systems \label{ssec:systems}}
 \fyDone{Remove Gated residual}
-Our comparison of multi-domain systems includes the following reimplementations of recent proposals from the literature:
+Our comparison of multi-domain systems includes the following reimplementations of recent proposals from the literature:\footnote{\revision{Further implementation details are in Appendix~A.}}
 \begin{itemize}
@@
 These methods thus only use domain tags in training.
-\item the multi-domain model of \newcite{Zeng18multidomain} (\system{WDCMT}), where a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly improved, results.}\fyDone{Check this}
+\item the multi-domain model of \newcite{Zeng18multidomain} (\system{WDCMT}), where a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification and adversarial training are used to compute these representations.
+\revision{As for the three previous proposals, inference does not use domain tags.}\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly improved, results.}\fyDone{Check this}
 \item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connections make it possible to by-pass this module.
@@
 
-Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges from our notion of MDMT, which should be able to handle available domain labels in training and testing.
+Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges from our notion of MDMT, which should be able to handle available domain labels in training and\revision{/or} testing.
 % This system has shown to meet several of requirements and would also have no problem scaling to a large number of domains.\fyDone{TBC}
@@
 
-In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the most likely domain tag and use it for decoding.
+In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the most likely domain tag and use it for decoding.\footnote{\revision{To automatically predict domain tags, we train a language model in each domain and assign tags on a per-sentence basis, using the language model log-probability (assuming uniform domain priors).
+Our domain classifier has a prediction error of XXX.}}
 % Results are in the rightmost column of Table~\ref{tab:redomains}.
-In this configuration, again, systems using domain tags in testing perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\fyDone{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}
+In this configuration, again, systems using domain tags in the inference step perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\fyDone{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}
 
@@
 
-\subsection*{A - Description of multi-domain systems}
-
+\subsection*{A - Description of multi-domain systems \label{ssec:implementation-details}}
+\revision{%
 The description of model in each MDMT approach is as follows:
@@
 \end{itemize}
-
+}
 \end{document}
@@
   \item $P(y|x)$ is deterministic, meaning that the joint distribution is entirely specified by the marginal and a labelling function $y=h(x)$; a more complex case is when $h()$ also varies across domains;
-  \item $P(y|x)$ is stochastic, but shared across all domains, which is the typical covariate-shift scenario \cite{Shimodaira00improving}; it can also vary across domains \cite{???}\fyTodo{Missing citation vary accross domains}
+  \item $P(y|x)$ is stochastic, but shared across all domains, which is the typical covariate-shift scenario \cite{Shimodaira00improving}; it can also vary across domains \cite{???}\fyTodo{Missing citation vary accoss domains}
   \end{itemize}

### 2020-08-20 2020/tacl2020/resubmission-paper2327.tex
More fixes and responses in the rebuttal file

%% TACL resubmission-pkg.tex v1.1, Michael Collins and Lillian Lee
% fixed page-size handling (thanks to Ion Madrazo), added anonymous authors, simplified.
%

\documentclass[11pt,a4paper]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% INPUTS FOR THE NEW PAPER -- AUTHORS, CHANGE APPROPRIATELY
\newcommand{\subtitle}{Revisiting Multi-Domain Machine Translation",}
\title{Resubmission of TACL \#2327, \\ \subtitle.   \\}

\newcommand{\coverfile}{rebuttal.pdf}  %% name of anonymized cover-letter file
\newcommand{\resubmission}{multidomain.pdf}  %% revised version
\newcommand{\origdecision}{decision.pdf}  %% name of anonymized original-decision-letter file

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{hyperref}
\hypersetup{
    colorlinks=true, %set true if you want colored links
    linktoc=all,     %set to all if you want both sections and subsections linked
    linkcolor=blue,  %choose some color if you want links to stand out
}

\usepackage{times}
\usepackage{pdfpages}
\pagestyle{plain}

\author{(Anonymous TACL resubmission)}
\setlength{\footskip}{55pt}

\begin{document}
\maketitle

\tableofcontents

\section{Author(s) cover letter responding to the original reviews} Starts on next page.
\includepdf[pages=-,fitpaper=true]{\coverfile}
\section{Revised submission} Starts on next page.
\includepdf[pages=-,fitpaper=true]{\resubmission}
\section{Original decision letter and reviews} Starts on next page.
\includepdf[pages=-,fitpaper=true]{\origdecision}

\end{document}

### 2020-08-20 2020/tacl2020/rebuttal.tex
more fixes and rebuttal - new reference

@@
 To benefit from your results, we need access to the same segmentation of train/dev/test sets that I hope you will provide to the community.
-\textcolor{red}{Supplementary details regarding the settings used in our experiments have been added in a new Appendix YYY.\fyTodo{Fix this.}.
+\textcolor{blue}{Supplementary details regarding the settings used in our experiments have been added in a new Appendix YYY.\fyTodo{Fix the appendix.}.
 We will also distribute the detailed composition of our train / dev / test split upon acceptance of the paper.}
@@
 \\
-\textcolor{red}{We formalize the data distribution as a mixture of several in-domain distribution.
+\textcolor{blue}{We formalize the data distribution as a mixture of several in-domain distribution.
 $\lambda$ correspond to the mixing weights, that also define the prior distribution on domains.
@@
 \\
-\textcolor{red}{Thanks for pointing this out. We have made this notation more consistent.}
+\textcolor{blue}{Thanks for pointing this out. We have made this notation more consistent.}
 \\
@@
 \\
-\textcolor{blue}{This is correct - a drop is performance is expected when using wrong domain tags; [P4-ERR] is however about robustness against domain errors and a good MDMT system should minimize this loss.
+\textcolor{blue}{This is correct - a drop is performance is expected when using wrong domain tags; [P4-ERR] is however about robustness against domain errors and a good MDMT system should keep this loss small.
 This is not what we see for the systems that use domain tags in testing.}
@@
 We would like to report the difference wrt to Table 3 in the later version of the paper.
-We agree that the comparison to Table 3 is also important, because helps assess the impact of a new domain to the previously learned domains.}
+We agree that the comparison to Table 3 is also important, because helps assess the impact of a new domain to the previously learned domains.}fytodo{What should we do about it ?}
 \\
@@
 \\
-\textcolor{red}{\system{DC-Feat} uses a small fixed embedding representing the domain.
+\textcolor{blue}{\system{DC-Feat} uses a small fixed embedding representing the domain.
 It is concatenated to every word embedding.
@@
 \\
-\textcolor{red}{The domain-weighted average is calculated by weighting the performance of the model in each domain test with the proportion of the respective domain in the training data.}
+\textcolor{blue}{This is correct: the domain-weighted average is computed by weighting the performance of the model in each domain test with the proportion of the respective domain in the training data.}
 \\
@@
 \\
-\textcolor{red}{There are standard test sets for some domains such as medical, IT, news and talk, but we could not find any standard test sets for the other domains including religion, law and bank.
-We carefully eliminated every duplicate sentence in all the corpora.
+\textcolor{blue}{There are standard test sets for some domains such as medical, IT, news and talk, but we could not find any standard test sets for the other domains including religion, law and bank.
+We carefully removed every duplicate sentence in all the corpora.
 In our opinion, the reason why the models achieve extremely high scores in the \domain{rel} domain is because texts in this domain is very repetitive and often predictable.}\fyTodo{sentence length ?, diversity index of vocab ?}
-\\
+\end{itemize}
+
 Other (minor) remarks:
+
+\begin{itemize}
 \item From the UFAL medical corpus, please say that you're using the indomain sections only (which is mostly PatTR Medical).
 \\
-\textcolor{red}{From the UFAL medical corpus of the language pair En-Fr, we select only sentences in categories including EMEA, ParTR, CESTA, ECDC. This is now reflected in the text (lines XXX)}
+\textcolor{blue}{From the UFAL medical corpus of the language pair En-Fr, we select only sentences in categories including EMEA, ParTR, CESTA, ECDC.
+This is now mentioned in the text (lines XXX)}\fyTodo{Add lines num}
 \\
-\item The \system{DC-Feat} and "\item{DC-Tag} lines in Table 5 should be swapped in order to be in line with Tables 3 and 4.
+\item The \system{DC-Feat} and \system{DC-Tag} lines in Table 5 should be swapped in order to be in line with Tables 3 and 4.
   \\
-  \textcolor{red}{We agree to change the order of "DC-Feat" and "DC-Tag" to be consistent with tables 3 and 4.}
+  \textcolor{blue}{We have changed the order of "DC-Feat" and "DC-Tag" to be consistent with tables 3 and 4.}
 \end{itemize}
-The remaining style issues have been corrected.
+
+The remaining style issues have also been corrected.
 

### 2020-08-20 2020/tacl2020/multidomain.tex
more fixes and rebuttal - new reference

@@
 \newcommand{\domain}[1]{\texttt{\textsc{#1}}}
-\newcommand{\system}[1]{\texttt{\textbf{#1}}}
+\newcommand{\system}[1]{\texttt{{#1}}}
 \newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
@@
 Training mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}.
-Under this view, MDMT would be especially beneficial for domains with little training data.
+Under this view, MDMT would be especially beneficial for domains with little training data \revision{(see~\cite{Sharaf20metalearning} for a discussion of DA in low resource conditions and a meta-learning based approach to this problem)}.
 This is observed for multilingual MT from English: an increase for under-resourced languages due to positive transfer, at the cost of a decrease in performance for well-resourced languages \cite{Arivazhagan19massively}.
@@
 \begin{table*}
+  \centering
   \begin{tabular}{|p{2.5cm}|*{9}{r|}} \hline \footnotesize
@@
 This allows us to evaluate requirement [P7-scale], training and testing our systems as if these domains were genuine.
-We also note that many of these classes are mere split of the large \domain{med}, while a fewer number of classes are mix of two existing domains.
+We also note that many of these classes are mere split of the large \domain{med}, while a fewer number of classes are mixtures of two existing domains.
 We are thus in a position to reiterate at a larger scale the measurements of Section~\ref{ssec:redomains} and test whether multi-domain systems can effectively take advantage from the cross-domain similarities, and to eventually perform better than fine-tuning.
@@
     \system{WDCMT}       & 35.6 & 53.1 & 48.4 & 30.5 & \SW{37.7} & \SW{56.0} & 38.5 & 43.6 \\ % & 20.4 \\ 
+    \revision{\system{FT-Block}}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SB{50.7}\\%  & \SW{21.2} \\
     \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SB{50.7}\\%  & \SW{21.2} \\
@@
 %}, NLU and many others
-
 Multi-domain translation was already proposed for statistical MT, either considering as we do multiple sources of training data (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}), or domains made of multiple topics\fyDone{you mean multiple topics?} \cite{Eidelman12topic,Hasler14dynamic-topic}.\fyDone{Add also topic models refs} Two main strategies were considered: instance-based, involving a measure of similarities between train and test domains; feature-based, where domain/topic labels give rise to additional features.
@@
 One contribution of this work is the idea of normalizing representations through adversarial training, so as to make the mixture of heterogeneous data more effective; representation normalization has since proven a key ingredient in multilingual transfer learning.
-The same basic techniques (parameter sharing, automatic domain identification / normalization) are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific representations on the one hand from domain-agnostic representation on the other hand.
+The same basic techniques (parameter sharing, automatic domain identification / normalization) are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific representations on the one hand from domain-agnostic representations on the other hand.
 These representations are then processed as two separate inputs, then recombined to compute the translation.

### 2020-08-24 2020/tacl2020/multidomain.tex
More fixes
new tables and numbers for reviewer B

@@
     \# lines & 2609 (0.68) & 190 (0.05)  & 501 (0.13) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
-    \# \revision{tokens} &  XX  & XX &  XX &  XX &  XX &  XX &  XX  \\
+    \# \revision{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 & 7.8 / 9.2   \\
+    \# \revision{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 & - \\
+    \# \revision{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 & - \\
     \hline
   \end{tabular}
-\caption{Corpora statistics: number of tokens ($\times 10^6$), number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain).
-\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.}
+  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$).
+\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
+  }
 \label{tab:Corpora}
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
+    \system{Mixed-Nat}  \hfill{\footnotesize[65m]} & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
+    \system{Mixed-Bal}   \hfill{\footnotesize[65m]} &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 \\ %& \\
+    \system{FT-Full}       \hfill{\footnotesize[65m]} & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & \SB{42.7} & \SB{53.8} \\ \hline
+ %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
+    \system{DC-Tag} \hfill{\footnotesize[+4k]}        & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.6 & 50.1    \\%    & 21.8 \\
+    \system{DC-Feat} \hfill{\footnotesize[+140k]}    & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   \\% & \SW{21.7} \\
+    \system{LDR}       \hfill{\footnotesize[+1.4m]}    & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} &40.9 & 49.8          \\% & 22.1 \\ 
+    \system{TTM}      \hfill{\footnotesize[+4k]}        & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.0 & 49.7     \\% &  23.4 \\
+    \system{DM}        \hfill{\footnotesize[+0]}         & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 38.1 & 43.4 \\ % & 22.6\\
+    \system{ADM}      \hfill{\footnotesize[+0]}         & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 38.9 & 47.5 \\% & 23.3 \\
+    \system{FT Res}   \hfill{\footnotesize[+12.4m]}  & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & \SB{42.3} & \SB{53.3} \\ % & 20.5\\ \hline
+    \system{MDL Res} \hfill{\footnotesize[+12.4m]}    & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & \SB{51.9} \\%  & \SW{21.2} \\
+%     \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
+%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3&\\
+     \hline \hline
     \system{Mixed-Nat-RNN}  & 36.8 & 53.8 & 47.2 & 30.0 & 35.7 & 60.2  & 39.2  & 44.0 \\
-    \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 \\ %& \\
-    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & \SB{42.7} & \SB{53.8} \\
-   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & \SB{42.3} & \SB{53.3} \\ \hline % & 20.5\\ \hline
- %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.6 & 50.1    \\%    & 21.8 \\
-    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   \\% & \SW{21.7} \\
-    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} &40.9 & 49.8          \\% & 22.1 \\ 
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.0 & 49.7     \\% &  23.4 \\
-    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 38.1 & 43.4 \\ % & 22.6\\
-    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 38.9 & 47.5 \\% & 23.3 \\
+    \hline
     \system{WDCMT}       & 36.0 & 53.3 & \SB{48.8} & 31.1 & \SB{38.8} & \SW{58.5} & 39.0 & 44.4 \\ % & 20.4 \\
-    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & \SB{51.9} \\%  & \SW{21.2} \\
-    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
-%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3&\\
     \hline
   \end{tabular}
-  \caption{Translation performance of various MDMT systems.
-We report BLEU scores for each domain, as well as domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages.
+  \caption{Translation performance of various MDMT systems \revision{based on the same Transformer (top) or RNN (bottom) architecture.
+The former contains 65m parameters, the latter has ???.
+For each system, we report the number of domain specific parameters,}  BLEU scores for each domain, as well as domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages.
+\revision{In the former case, we take the domain proportions from Table~\ref{tab:Corpora}}.
 Boldface denotes significant gains with respect to \system{Mix-Nat} (or \system{Mix-Nat-RNN}, for WDCMT), underline denotes significant losses.}
@@
     \system{WDCMT} (*)     & -0.0    & -0.0  & +0.2 & +0.2  & +0.8 & +0.8  & -0.4 & -0.8 & 0.0 & +0.2 \\
+    \revision{\system{FT Res}}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +0.7 & -0.3 & \SW{-18.6} & \SW{-1.3}\\ % [31.8]
     \system{MDL Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +0.7 & -0.3 & \SW{-18.6} & \SW{-1.3}\\ % [31.8]
@@
 
-In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the most likely domain tag and use it for decoding.\footnote{\revision{To automatically predict domain tags, we train a language model in each domain and assign tags on a per-sentence basis, using the language model log-probability (assuming uniform domain priors).
-Our domain classifier has a prediction error of XXX.}}
+In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the most likely domain tag and use it for decoding.\footnote{\revision{To automatically predict domain tags, we train a language model in each domain and assign tag on a per-sentence basis based the language model log-probability (assuming uniform domain priors).
+Our domain classifier has a average prediction error of 16.4\% for in-domain data.}}
 % Results are in the rightmost column of Table~\ref{tab:redomains}.
@@
 Starting with the existing MD systems of Section~\ref{ssec:rawperformance}, we introduce an extra domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is critical for achieving optimal performance: as our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy based on the new training distribution.
-This is detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with training all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.\footnote{WDCMT results are excluded from this Table, as resuming training proved difficult to implement.}\fyDone{Ecrire un commentaire}
+This is detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with training all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.\footnote{WDCMT results are excluded from this table, as resuming training proved difficult to implement.}\fyDone{Ecrire un commentaire}
 We expect that MDMT systems should not to be significantly impacted by the addition of a new domain and reach about the same performance as when training with this domain from scratch.
@@
 %                    &&&&&&&&& \\
+    \revision{\system{FT Res}}  &  \\
 

### 2020-08-24 2020/tacl2020/rebuttal.tex
More numbers in the tables

@@
     We understand that our naming schemes was insisting on training procedures, rather than on architectures, thus possibly introducing some confusions.
-We have accordingly changed the name of \system{FT-Block} to XXX\fyTodo{fix this}, clarified the description of the two systems (see lines YYY\fyTodo{fix this}), and reported \system{FT-Block}\fyTodo{or new name} as another multi-domain system in all tables.\fyTodo{Make sure it is the case}.}
+We have accordingly changed the name of \system{FT-Block} to \system{FT Res}, clarified the description of the two systems (see lines YYY\fyTodo{fix this}), and reported \system{FT Res}\fyTodo{or new name} as another multi-domain system in all tables.\fyTodo{Make sure it is the case}.}
   %}
@@
 It will be helpful to see  the impact of the number of parameters on the MT results.
-  \textcolor{red}{This information has been added in a new Table~ZZZ \fyTodo{Table number, page and line number and content}.}
+  \textcolor{blue}{This information has been added in Table~3\fyTodo{Check number, page and line number and content}.}
 \\ 
@@
   \\
+  \textcolor{red}{Answer this}
 \item[*] Please define Lambda coefficients in lines 142 and 146.
@@
 \\
-\textcolor{red}{This reference has been added to the related work section.}\fyTodo{Add citation}
+\textcolor{blue}{This reference has been added to the related work section.}\fyTodo{Add citation}
 \\
@@
 To predict domain tag of a sentence, we compute probability of the sentence according to each pretrained language model and choose the domain corresponding to the highest probability as domain tag of sentence.
+Note that this classifier is only used for the \domain{news} test data.
 This is now explained in footnote ZZZ.}\fyTodo{to compute error rate of domain prediction over each test set, check comment, add footnote number.}
@@
 \\
-\textcolor{blue}{. See table ZZZ.}\fyTodo{extra experiments to do.}
+\textcolor{blue}{Table 1 now contains the number of types as well as the number of types that are unique in a domain.
+We fill that performing an error analysis for each of the system / domain compared in this study would need to refocus more on a detailled analysis of each system, while our mail goal is to propose new methods to evaluate the actual abilities of MDMT systems.}\fyTodo{Will this be enough ?}
 \\
@@
 \\
-\textcolor{red}{For each domain, we create train/dev/test sets by randomly splitting the respective corpora.
+\textcolor{red}{For each domain, we create train/dev/test sets by randomly each corpus.
 We maintain the size of validation sets and of test sets equal to 1,000 lines for every domain.
-In Tables 3 and 4, to train MDMT models except FT-full and FT-block, we use a validation set which is a concatenation of the 6 validation sets of 6 domains.
-In Table 5, to train MDMT models except FT-full and FT-block we concatenate the validation set of NEWS to the mixture of 6 validation sets.
-To train FT-full model and FT-block model we only use the in-domain validation sets.
+In Tables 3 and 4, to train MDMT systems (except FT-full and FT Res), we use a validation set which is a concatenation of the 6 validation sets of 6 domains.
+In Table 5, to train MDMT systems (except \system{FT-Full} and \system{FT Res}) we append the validation set of \domain{news} to the concatenation of 6 validation sets.
+To train \system{FT-full} and \system{FT Res} model we only use the in-domain validation sets.
 We stop training if either the training iteration reach the maximal number of iterations or the score on the validation set does not increase in 3 consecutive evaluations.
-We average 5 checkpoints up to the best checkpoint to get the final model.}
+We average 5 checkpoints up to the best checkpoint to get the final model.}\fyTodo{up to the best ?}
 \\
@@
 \\
-\textcolor{red}{The learning rate is determined according to $lr = d_{model}^{-0.5}*\min(step\_num^{-0.5},step\_num * warmup\_steps^{-1.5})$ where $d_{model}=512$,$warmup\_steps=4000$, as in the original Transformer paper.
-For the finetuning procedure, we continue training using the same learning rate schedule and the same $step\_num$.}\fyTodo{Fix this, add comments in appendix}
+\textcolor{blue}{The learning rate is determined according to $lr = d_{model}^{-0.5}*\min(step\_num^{-0.5},step\_num * warmup\_steps^{-1.5})$ where $d_{model}=512$,$warmup\_steps=4000$, as in the original Transformer paper.
+For the finetuning procedure, we continue training using the same learning rate schedule and the same $step\_num$.
+This is now clarified in the appendix}\fyDone{Fix this, add comments in appendix}
 \\

### 2020-08-24 2020/tacl2020/multidomain.tex
More numbers in the tables

@@
 Fine-tuning each domain separately provides us with better baselines, outperforming \system{Mixed-Nat} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
-A complete fine-tuning, updating all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to keeping fixed the capacity of the adapter module.\fyDone{Add 2 averages ?}\fyDone{Significance testing wrt Mix Generic, Full Fine-tuned, for each domain}
+% A complete fine-tuning, updating all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to keeping fixed the capacity of the adapter module.
+\fyDone{Add 2 averages ?}\fyDone{Significance testing wrt Mix Generic, Full Fine-tuned, for each domain}
 
@@
 If all MDMTs (except \system{DM} and \system{ADM}) slightly improve over the generic baseline (for most domains), these improvements are rarely significant.
-The best contender overall is \system{MDL Res}, which significantly improves over \system{Mixed-Nat} for three domains, and is the only one to clearly fulfill [P1-LAB]; \system{WDCMT} also improves on three domains, but regresses on one.
-The use of a dedicated adaptation component seems here to prove better than feature-based strategies.
-The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}), for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.\fyDone{why LDC does not outperform DC as in Minh's paper (for en2fr MED) ?}
+\revision{The best contender overall are  \system{FT Res} and \system{MDL Res}, which significantly improve over \system{Mixed-Nat} for a majority of domains, and are the only ones to clearly fulfill [P1-LAB];} \system{WDCMT} also improves on three domains, but regresses on one.
+The use of a dedicated adaptation component, seems here to prove better than feature-based strategies; \revision{this comes at the cost of a large increase of the number of parameters.} The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}).
+% for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.\fyDone{why LDC does not outperform DC as in Minh's paper (for en2fr MED) ?}
 
@@
     \system{Mixed-Bal}   \hfill{\footnotesize[65m]} &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 \\ %& \\
-    \system{FT-Full}       \hfill{\footnotesize[65m]} & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & \SB{42.7} & \SB{53.8} \\ \hline
+    \system{FT-Full}       \hfill{\footnotesize[6$\times$65m]} & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & \SB{42.7} & \SB{53.8} \\ \hline
  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
@@
     \system{ADM}      \hfill{\footnotesize[+0]}         & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 38.9 & 47.5 \\% & 23.3 \\
-    \system{FT Res}   \hfill{\footnotesize[+12.4m]}  & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & \SB{42.3} & \SB{53.3} \\ % & 20.5\\ \hline
+    \revision{\system{FT Res}}   \hfill{\footnotesize[+12.4m]}  & 37.3 & \SB{57.9} & \SB{53.9} & 33.8 & \SB{46.7} & \SB{90.2}  & \SB{42.3} & \SB{53.3} \\ % & 20.5\\ \hline
     \system{MDL Res} \hfill{\footnotesize[+12.4m]}    & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & \SB{51.9} \\%  & \SW{21.2} \\
@@
 In first three, we randomly \emph{split} one corpus in two parts and proceed as if this corresponded to two actual domains.
-A MD system should detect that these two pseudo-domains are mutually beneficial, and should be hardly affected by this change with respect to the baseline scenario (no split).
+A MD system should detect that these two pseudo-domains are mutually beneficial and should be hardly affected by this change with respect to the baseline scenario (no split).
 In this situation, we expect MDMT to even surpass fine-tuning separately on each of these dummy domains, as MDMT exploits all data, while fine-tuning focuses only on a subpart.
@@
 In the \textsl{merge} experiment, we merge two corpora in training, in order to assess the robustness with respect to heterogenous domains [P3-HET].
-We then translate the two corresponding tests with the same system.
+We then translate the two corresponding tests with the same (merged) system.
 
@@
     \system{ADM} (*)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & 0.0&-0.2\\
+    \revision{\system{FT Res}}   & -0.1   & -0.4 & -0.3 &-0.3 & \SW{-2.2} & \SW{-2.9} & \SW{-2.4} & -\SW{3.2} & \SW{-13.3} & \SW{-3.0}\\ % [31.8]
+    \system{MDL Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +0.7 & -0.3 & \SW{-18.6} & \SW{-1.3}\\ % [31.8]
     \system{WDCMT} (*)     & -0.0    & -0.0  & +0.2 & +0.2  & +0.8 & +0.8  & -0.4 & -0.8 & 0.0 & +0.2 \\
-    \revision{\system{FT Res}}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +0.7 & -0.3 & \SW{-18.6} & \SW{-1.3}\\ % [31.8]
-    \system{MDL Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +0.7 & -0.3 & \SW{-18.6} & \SW{-1.3}\\ % [31.8]
     \hline
@@
   \caption{Translation performance with variable domain definitions.
-In the Split/Merge experiments, we report BLEU differences for the concerned test set(s).
-Underline denotes significant loss when domains are changed wrt.\ the baseline situation; bold for a significant improvement over \system{FT-Full}; (*) flags systems ignoring test domains.
+In the Split/Merge experiments, we report BLEU differences for the related test set(s).
+Underline denotes significant loss when domains are changed wrt.\ the baseline situation; bold for a significant improvement over \system{FT-Full}; (*) tags systems ignoring test domains.
     \fyDone{Check wrong and significance of diff for ``old''}\fyDone{How about wrong ?}
@@
 In contrary, \system{FT-Full} incurs large losses across the board, especially for the small data condition.
-Even in this very favourable case, very few MDMT systems are able to significantly outperform \system{FT-Full} and this is only observed for the smaller part of the \domain{med} domain.
+Even in this very favourable case however, very few MDMT systems are able to significantly outperform \system{FT-Full} and this is only observed for the smaller part of the \domain{med} domain.
 The merge condition is hardly different, with large losses for \system{FT-full}, and small variations for all systems.
@@
 \subsubsection{Handling growing numbers of domains}
-Another series of experiments evaluate the ability to dynamically handle supplementary domains (requirement [P6-DYN]) as follows.
+Another set of experiments evaluate the ability to dynamically handle supplementary domains (requirement [P6-DYN]) as follows.
 Starting with the existing MD systems of Section~\ref{ssec:rawperformance}, we introduce an extra domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is critical for achieving optimal performance: as our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy based on the new training distribution.
@@
     \system{WDCMT}       & 35.6 & 53.1 & 48.4 & 30.5 & \SW{37.7} & \SW{56.0} & 38.5 & 43.6 \\ % & 20.4 \\ 
-    \revision{\system{FT-Block}}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SB{50.7}\\%  & \SW{21.2} \\
+    \revision{\system{FT Res}}     & 37.5 & 55.7  & 51.1   & 33.1   &  44.1  & \SW{86.7} & 41.6 & 51.4\\%  & \SW{21.2} \\
     \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SB{50.7}\\%  & \SW{21.2} \\
@@
 Details are in \cite{Pham19generic}.
+\item \system{Mixed-NAT-RNN} uses one layer LSTM bidirectional as encoder and one layer LSTM as decoder. The size of each hidden layer is 1024, the size of word embeddings is 512. 
 \item \system{WDCNMT} uses one layer GRU bidirectional as encoder and one layer GRU-conditional as decoder. The size of each hidden layer is 1024, the size of word embedding is 512.
-\item \system{Mixed-NAT-RNN} uses one layer LSTM bidirectional as encoder and one layer LSTM as decoder. The size of each hidden layer is 1024, the size of word embeddings is 512. 
 \end{itemize}
+\paragraph{Training} For each domain, we create train/dev/test sets by randomly each corpus. We maintain the size of validation sets and of test sets equal to 1,000 lines for every domain.
+The learning rate is set to $lr = d_{model}^{-0.5}*\min(step\_num^{-0.5},step\_num * warmup\_steps^{-1.5})$ where $d_{model}=512$,$warmup\_steps=4000$, as in \cite{Vaswani17attention}.
+For the fine-tuning procedure, we continue training using the same learning rate schedule, continuing the incrementation of $step\_num$.
+In Tables~\ref{tab:performance} and \ref{tab:redomains}, to train MDMT systems (except FT-full and FT Res), we use a validation set which is a concatenation of the 6 validation sets.
+In Table~\ref{tab:warmrestart}, to train MDMT systems (except \system{FT-Full} and \system{FT Res}) we append the validation set of \domain{news} to the concatenation of 6 validation sets.
+To train \system{FT-full} and \system{FT Res} model we only use the in-domain validation sets.
+We stop training if either the training iteration reach the maximal number of iterations or the score on the validation set does not increase in 3 consecutive evaluations.
+We average 5 checkpoints up to the best checkpoint to get the final model.
 }
+
 \end{document}

### 2020-08-24 2020/tacl2020/rebuttal.tex
More answers to reviewers A and C

@@
   \\
-  \textcolor{red}{We agree lines XXX}
+  \textcolor{blue}{We agree with the reviewer that the analogy is misleading.
+As this is not central to our discussion, we have repharsed and dropped the comparison.
+See changes in line XXX}\fytodo{Changes line in.}
 \item 567-568: "having the domain tag on the source is better than on the target".
 I assume this refers to the comparison of "DC-Tag" and "TTM".
-However, TTM actually uses the predicted domain label on the target side, and could perform better with forced decoding of the gold domain label.
+However, TTM actually uses the predicted domain label on the target side and could perform better with forced decoding of the gold domain label.
 More precisely, the statement should be that "having the gold tag is better than a predicted tag", but this is hardly surprising.
   \\
-  \textcolor{red}{We agree that the conclusion is not really fair for "TTM" because "TTM" intend to predict domain label rather than to use gold domain label and TTM could definitively translate with gold domain label as prefix of each hypothesis.}
+  \textcolor{red}{We agree that the conclusion is not really fair for "TTM" because "TTM" intend to predict domain label rather than to use gold domain label and TTM could definitively translate with gold domain label as prefix of each hypothesis.
+In our experiments, we found that TTM is pretty accurate in its prediction, making about 2\% errors in the predicted tags, which make explain the small difference with observe.
+We have rephrased acccordingly}
     \fyTodo{Check prediction quality, moderate conclusion}

### 2020-08-24 2020/tacl2020/multidomain.tex
More answers to reviewers A and C

@@
 \usepackage[draft]{todo}
+\usepackage[normalem]{ulem}
 \usepackage{xspace}
@@
 \newcommand{\revision}[1]{\textcolor{red}{#1}}
-
+\newcommand{\revisiondel}[1]{}
 \newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
@@
 Combining multiple domain-specific MTs can also be justified in the sake of distributional robustness \cite{Mansour09domainadaptation,Mansour09multiple}, for instance when the test mixture differs from the train mixture, or when it includes new domains unseen in training.
-Such scenario is already well documented for zero-shot multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}, where mixing languages has more than demonstrated its usefulness.
-An even more challenging case is when the MT would need to perform well for any test distribution (as studied for statistical MT in \cite{Huck15mixeddomain}).
-In all these cases, mixing domains in training and/or testing is likely to improve robustness against unexpected or adversarial inputs \cite{Oren19distributionally}.
+% \sout{Such scenario is already well documented for zero-shot multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual, where mixing languages has more than demonstrated its usefulness.}
+An even more challenging case is when the MT would need to perform well for any test distribution, as studied for statistical MT in \cite{Huck15mixeddomain}.
+In all these cases, mixing domains in training and/or testing is likely to improve robustness against unexpected or adversarial test distribution \cite{Oren19distributionally}.
 
@@
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
-Corpus statistics are in Table~\ref{tab:Corpora}.
-Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools.
-To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyTodo{Add \# number of tokens, also specificity ?}%
+Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
+To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
 
@@
 We develop two versions\footnote{In fact three: to enable a fair comparison with WDCMT, a RNN-based variant is also trained and evaluated.
-\revision{This variant appears as \system{Mixed-Nat-RNN} in Table~\ref{tab:performance}}.} of this system, one where the domain unbalance reflects the distribution of our training data \revision{given in Table~\ref{tab:Corpora}} (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
+\revision{This system appears as \system{Mixed-Nat-RNN} in Table~\ref{tab:performance}}.} of this system, one where the domain unbalance reflects the distribution of our training data \revision{given in Table~\ref{tab:Corpora}} (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
 The former is the best option when the train mixture $\mathcal{D}^s$ is also expected in testing; the latter should be used when the test distribution is uniform across domains.
@@
 \item fine-tuned models \cite{Luong15stanford,Freitag16fast}\jcDone{i would use Luong and Manning 2015}, based on the \system{Mixed-Nat} system, further trained on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
-We again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a simple adaptation module that is added (with residual connections) on top of every Transformer layer (\system{FT-Block}).
+The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems adapted for one domain, with no parameter sharing across domains.
+\revisiondel{We again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a small adaptation module that is added (with residual connections) on top of every Transformer layer (\system{FT Res}).}
 \end{itemize}
@@
 % The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training use a batch size of~12288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
+Training use a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
 % All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
@@
 These methods thus only use domain tags in training.
+\item \revision{two multi-domain versions of the approach of \newcite{Bapna19simple}, denoted \system{FT Res} and \system{MDL Res}, where a domain-specific adaptation module is added to all the Transformer layers; within each layer, residual connections enable to by-pass this module.
+The former variant corresponds to the original proposal of \citet{Bapna19simple} and fine-tunes the adapter modules of a \system{Mixed-Nat} system independently for each domain, keeping all the other parameters frozen.
+The latter is novel; it uses the same architecture, but a different training procedure, as all parameters are jointly learned from scratch from a mix-domain corpus.
+}
+  % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
+% the entire multi-domain system from scratch.\fyDone{Check this.}
 \item the multi-domain model of \newcite{Zeng18multidomain} (\system{WDCMT}), where a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification and adversarial training are used to compute these representations.
 \revision{As for the three previous proposals, inference does not use domain tags.}\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly improved, results.}\fyDone{Check this}
-\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connections make it possible to by-pass this module.
-  % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
-Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
 \end{itemize}
@@
 \subsection{Performance of MDMT systems \label{ssec:rawperformance}}
-In this section, we discuss the basic performance of MDMT systems trained and tested on $6$~domains.
-Results are in Table~\ref{tab:performance}.
-As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points\fyDone{BP?} better, notably owing to the much better results for \domain{rel}).
-As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains.
-As all other systems are trained with an unbalanced data distribution, we use the weighted average to perform global comparisons.
-
-Fine-tuning each domain separately provides us with better baselines, outperforming \system{Mixed-Nat} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
-% A complete fine-tuning, updating all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to keeping fixed the capacity of the adapter module.
-\fyDone{Add 2 averages ?}\fyDone{Significance testing wrt Mix Generic, Full Fine-tuned, for each domain}
-
-A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on the source is better than on the target, and very comparable to repeating the tag for every source word.
-If all MDMTs (except \system{DM} and \system{ADM}) slightly improve over the generic baseline (for most domains), these improvements are rarely significant.
-\revision{The best contender overall are  \system{FT Res} and \system{MDL Res}, which significantly improve over \system{Mixed-Nat} for a majority of domains, and are the only ones to clearly fulfill [P1-LAB];} \system{WDCMT} also improves on three domains, but regresses on one.
-The use of a dedicated adaptation component, seems here to prove better than feature-based strategies; \revision{this comes at the cost of a large increase of the number of parameters.} The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}).
-% for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.\fyDone{why LDC does not outperform DC as in Minh's paper (for en2fr MED) ?}
-
-All systems fail to outperform fine-tuning, sometimes by a wide margin, especially for an ``isolated'' domain like \domain{rel}.
-This might be due to the fact that domains are well separated (cf.\ Section~\ref{ssec:corpora}) and are hardly helping each other.
-In this situation, MDMT systems should instead make sure that each domain is allocated a sufficient number of parameters, so as to close the gap with fine-tuning.
-\fyDone{More comments when we have all the results}\fyDone{Importance of sharing and unsharing}
 
@@
 \end{table*}
+
+In this section, we discuss the basic performance of MDMT systems trained and tested on $6$~domains.
+Results are in Table~\ref{tab:performance}.
+As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points\fyDone{BP?} better, notably owing to the much better results for \domain{rel}).
+As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains.
+As all other systems are trained with an unbalanced data distribution, we use the weighted average to perform global comparisons.
+
+Fine-tuning each domain separately provides us with better baselines, outperforming \system{Mixed-Nat} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
+% A complete fine-tuning, updating all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to keeping fixed the capacity of the adapter module.
+\fyDone{Add 2 averages ?}\fyDone{Significance testing wrt Mix Generic, Full Fine-tuned, for each domain}
+
+A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on the source is better than on the target, and very comparable to repeating the tag for every source word.
+If all MDMTs (except \system{DM} and \system{ADM}) slightly improve over the generic baseline (for most domains), these improvements are rarely significant.
+\revision{The best contender overall are  \system{FT Res} and \system{MDL Res}, which significantly improve over \system{Mixed-Nat} for a majority of domains, and are the only ones to clearly fulfill [P1-LAB];} \system{WDCMT} also improves on three domains, but regresses on one.
+The use of a dedicated adaptation component, seems here to prove better than feature-based strategies; \revision{this comes at the cost of a large increase of the number of parameters.} The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}).
+% for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.\fyDone{why LDC does not outperform DC as in Minh's paper (for en2fr MED) ?}
+
+All systems fail to outperform fine-tuning, sometimes by a wide margin, especially for an ``isolated'' domain like \domain{rel}.
+This might be due to the fact that domains are well separated (cf.\ Section~\ref{ssec:corpora}) and are hardly helping each other.
+In this situation, MDMT systems should instead make sure that each domain is allocated a sufficient number of parameters, so as to close the gap with fine-tuning.
+\fyDone{More comments when we have all the results}\fyDone{Importance of sharing and unsharing}
 
@@
 \begin{itemize}
-\item \system{Mixed-NAT}, \system{FT-full}, \system{TTM}, \system{DC-Tag} use the medium Transformer model (\fyTodo{ref})
-\item \system{FT-block} and \system{MDL res} use medium Transformer model, and residual layers with the bottleneck dimension of size 1024.
-\cite{Bapna19simple} showed that if the bottleneck dimension is large enough, the gap between \system{FT-block} and \system{FT-full} will be small.\fyTodo{This comment is confusing}.
+\item \system{Mixed-NAT}, \system{FT-full}, \system{TTM}, \system{DC-Tag} use the a medium Transformer model of \cite{Vaswani17attention} with the following settings: embeddings size and hidden layers size are set to~512.
+Multi-head attention comprises 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
+% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
+% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
+Training use a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.
+\item \system{FT Res} and \system{MDL res} use the same medium Transformer model, and add residual layers with a bottleneck dimension of size 1024.
+  % \cite{Bapna19simple} showed that if the bottleneck dimension is large enough, the gap between \system{FT-block} and \system{FT-full} will be small.\fyTodo{This comment is confusing}.
 \item \system{ADM}, \system{DM} use medium Tranformer model and a domain classifier composing of 3 dense layers of size $512 \times 2048$, $2048 \times 2048$ and $2048 \times domain\_num$.
@@
 The learning rate is set to $lr = d_{model}^{-0.5}*\min(step\_num^{-0.5},step\_num * warmup\_steps^{-1.5})$ where $d_{model}=512$,$warmup\_steps=4000$, as in \cite{Vaswani17attention}.
-For the fine-tuning procedure, we continue training using the same learning rate schedule, continuing the incrementation of $step\_num$.
-In Tables~\ref{tab:performance} and \ref{tab:redomains}, to train MDMT systems (except FT-full and FT Res), we use a validation set which is a concatenation of the 6 validation sets.
-In Table~\ref{tab:warmrestart}, to train MDMT systems (except \system{FT-Full} and \system{FT Res}) we append the validation set of \domain{news} to the concatenation of 6 validation sets.
-To train \system{FT-full} and \system{FT Res} model we only use the in-domain validation sets.
-We stop training if either the training iteration reach the maximal number of iterations or the score on the validation set does not increase in 3 consecutive evaluations.
-We average 5 checkpoints up to the best checkpoint to get the final model.
+For the fine-tuning procedures used for \system{FT-full} and \system{FT Res}, we continue training using the same learning rate schedule, continuing the incrementation of $step\_num$.
+All other MDMT systems reported in Tables~\ref{tab:performance} and \ref{tab:redomains} use a combined validation set comprising 6,000 lines, obtained by merging the six development sets.
+For the results in Table~\ref{tab:warmrestart} we also append the validation set of \domain{news} to the multi-domain validation set.
+In any case, training stops if either the training iteration reach a maximum number of XXX\fyTodo{How many ?} or the score on the validation set does not increase in three consecutive evaluations.
+We average five checkpoints up to the best checkpoint to get the final model.\fyTodo{up to the best ?}
 }
-
 \end{document}

### 2020-08-26 2020/tacl2020/rebuttal.tex
Answer most reviewer questions, fix latex

@@
 I'm confused why fine-tuning is specifically excluded, since it meets many of the reasons for building MDMT systems.
-Authors do mention that it would be cheaper to develop a single system instead of having to maintain multiple systems, but they include "MDL Res", which has domain-specific adaptation modules and is very similar to "FT-Block" (a baseline that also combines a general model with domain-specific adaptation modules).
+Authors do mention that it would be cheaper to develop a single system instead of having to maintain multiple systems, but they include "\system{MDL Res}", which has domain-specific adaptation modules and is very similar to "FT-Block" (a baseline that also combines a general model with domain-specific adaptation modules).
 The line
-between a single system and multiple systems gets fuzzy here, and I see no clear motivation to draw this clear distinction and only consider "MDL Res" a multi-domain system.
+between a single system and multiple systems gets fuzzy here, and I see no clear motivation to draw this clear distinction and only consider "\system{MDL Res}" a multi-domain system.
 \\
@@
 Firstly, this is probably highly dependent on data mixtures, length of training and regularization strategies, so I'd be wary of drawing general conclusions from a single set of experiments.
-Secondly, when looking at the models where such a performance deterioration is said to happen (DC-Tag, TTM, ADM; lines
-770-771), this deterioration is actually small or non-existent when
-comparing table 3 and table 5 averages (-0.2 BLEU for DC-Tag; -0.3 BLEU for TTM; +0.9 BLEU for ADM).
+Secondly, when looking at the models where such a performance deterioration is said to happen (\system{DC-Tag}, \system{TTM}, \system{ADM}; lines 770-771), this deterioration is actually small or non-existent when comparing table 3 and table 5 averages (-0.2 BLEU for \system{DC-Tag}; -0.3 BLEU for \system{TTM}; +0.9 BLEU for \system{ADM}).
 \\
+\textcolor{red}{%
+  We have tried to address these commentaries in the revised version of the paper.
+Notably, we have clarified the distinction between \system{FT-block} and \system{MDL Res} (see details below).
+We have also tried to clarify the distinction between multi-domain MT and fine-tuning; while the former is designed with an a priori knowledge that the test data is sampled from a single domain, the latter assumes that the test data will be sampled from multiple domains, using domain probabilities that are not known a priori.
+This raises the issue of robustness wrt to adversarial test distribution and even new domains, an issue that is irrelevant for DA.
+
+In our experiments, fine-tuned systems share no parameter and show large drops in performance when translating data from the wrong domain (even though this could be somewhat mitigated, as the reviewer points out).
+By contrast, the majority of the parameters of our MDMT systems are shared for all domains.
+Arguably, one could see here a continuum between no parameter sharing and full parameter sharing, with \system{MDL Res} and \system{FT Block} corresponding to some intermediay situation.
+
+We also agree that the results for the ``new domain'' experiment should not be over-blown and we have rephrased our conclusions somehow.
+Yet, this experiments highlights the fact that increasing the heterogeneity of the data pool could be harmful
+}
 \noindent Question to authors:
 \begin{itemize}
-\item can you describe in more detail the differences between FT-Block and MDL Res, apart from the fact that the former is fine-tuned from a generic model, the latter trained from scratch~?
+\item can you describe in more detail the differences between \system{FT-Block} and \system{MDL Res}, apart from the fact that the former is fine-tuned from a generic model, the latter trained from scratch~?
   \\
@@
 As this is not central to our discussion, we have repharsed and dropped the comparison.
-See changes in line XXX}\fytodo{Changes line in.}
+See changes in line XXX}\fyTodo{Changes line in.}
 \item 567-568: "having the domain tag on the source is better than on the target".
@@
   \\
-  \textcolor{red}{We agree that the conclusion is not really fair for "TTM" because "TTM" intend to predict domain label rather than to use gold domain label and TTM could definitively translate with gold domain label as prefix of each hypothesis.
-In our experiments, we found that TTM is pretty accurate in its prediction, making about 2\% errors in the predicted tags, which make explain the small difference with observe.
-We have rephrased acccordingly}
-    \fyTodo{Check prediction quality, moderate conclusion}
+  \textcolor{blue}{We agree that the conclusion is not really fair for "TTM" because "TTM" intend to predict domain label rather than to use gold domain label and TTM could definitively translate with gold domain label as prefix of each hypothesis.
+In our experiments, we found that TTM is pretty accurate in its prediction, making about 2\% errors in the predicted tags, which may explain the small difference with observe in Table~3.
+We have rephrased acccordingly in lines WWW.}\fyTodo{Add lines}
 \end{itemize}
@@
   \\
-  \textcolor{red}{Answer this}
+  \textcolor{blue}{We agree that carefully analyzing the effects of using corpora of varying size is important, as this corresponds to practical training scenario.
+From our results in Table 6 (and Table XXX in appendix), we see that some techniques can work well with small in-domain corpora.
+However, this analysis must also take in account the similarity between domains, as these improvements will be stronger if similar or related data are also present in the training mixture -- a situation that we have artificially created in our split conditions.
+Precisely sorting out these two effects is beyond the scope of this study and left for future work.}
+
 \item[*] Please define Lambda coefficients in lines 142 and 146.
@@
 \\
-\textcolor{blue}{Table 1 now contains the number of types as well as the number of types that are unique in a domain.
-We fill that performing an error analysis for each of the system / domain compared in this study would need to refocus more on a detailled analysis of each system, while our mail goal is to propose new methods to evaluate the actual abilities of MDMT systems.}\fyTodo{Will this be enough ?}
+\textcolor{blue}{Table~1 now contains the number of types as well as the number of types that are unique in a domain.
+We fill that performing an error analysis for each of the system / domain compared in this study would require us to focus more on a detailled analysis of each system, while our main goal is to propose new ways to assess the actual abilities of existing MDMT systems.}\fyTodo{Will this be enough ?}
 \\
@@
 For the finetuning procedure, we continue training using the same learning rate schedule and the same $step\_num$.
-This is now clarified in the appendix}\fyDone{Fix this, add comments in appendix}
+This is now clarified in the appendix.}\fyDone{Fix this, add comments in appendix}
 \\
@@
 \\
-\textcolor{red}{In Table 5, we report the difference in performance between 2 situations: resuming training and training from beginning with 7 domains.
-We would like to report the difference wrt to Table 3 in the later version of the paper.
-We agree that the comparison to Table 3 is also important, because helps assess the impact of a new domain to the previously learned domains.}fytodo{What should we do about it ?}
+\textcolor{blue}{In Table 5, we report the difference in performance between 2 situations: resuming training and training from beginning with 7 domains.
+We agree that the comparison to Table 3 is also important, because helps assess the impact of a new domain to the previously learned domains, and have now reported this comparison in table 5.
+We have also modified the caption to clarify our reference situation (coldstart, with 7 domains).
+See changes in Table 5.}\fyDone{What should we do about it ?}
 \\

### 2020-08-26 2020/tacl2020/multidomain.tex
Answer most reviewer questions, fix latex

@@
 \renewcommand\floatpagefraction{.95}
+\newcommand{\sbcl}[2]{{\scriptsize #1 \hfill $|$ \hfill  #2}}
 
@@
 Training mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}.
-Under this view, MDMT would be especially beneficial for domains with little training data \revision{(see~\cite{Sharaf20metalearning} for a discussion of DA in low resource conditions and a meta-learning based approach to this problem)}.
-This is observed for multilingual MT from English: an increase for under-resourced languages due to positive transfer, at the cost of a decrease in performance for well-resourced languages \cite{Arivazhagan19massively}.
+Under this view, MDMT would be especially beneficial for domains with little training data \revision{(see also \cite{Sharaf20metalearning} for a discussion of DA in low-resource conditions and a meta-learning based approach to this problem)}\fyTodo{Best place ?}.
+This is observed for multilingual MT from English: an improvement for under-resourced languages due to positive transfer, at the cost of a decrease in performance for well-resourced languages \cite{Arivazhagan19massively}.
 
 Combining multiple domain-specific MTs can also be justified in the sake of distributional robustness \cite{Mansour09domainadaptation,Mansour09multiple}, for instance when the test mixture differs from the train mixture, or when it includes new domains unseen in training.
-% \sout{Such scenario is already well documented for zero-shot multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual, where mixing languages has more than demonstrated its usefulness.}
+\sout{Such scenario is already well documented for zero-shot multilingual MT} \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}\sout{where mixing languages has more than demonstrated its usefulness.}
 An even more challenging case is when the MT would need to perform well for any test distribution, as studied for statistical MT in \cite{Huck15mixeddomain}.
@@
     \hline 
-    \# lines & 2609 (0.68) & 190 (0.05)  & 501 (0.13) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
+    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
     \# \revision{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 & 7.8 / 9.2   \\
@@
 
-Fine-tuning each domain separately provides us with better baselines, outperforming \system{Mixed-Nat} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
+Fine-tuning each domain separately provides us with a better baseline, outperforming \system{Mixed-Nat} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
 % A complete fine-tuning, updating all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to keeping fixed the capacity of the adapter module.
@@
 
-A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on the source is better than on the target, and very comparable to repeating the tag for every source word.
-If all MDMTs (except \system{DM} and \system{ADM}) slightly improve over the generic baseline (for most domains), these improvements are rarely significant.
-\revision{The best contender overall are  \system{FT Res} and \system{MDL Res}, which significantly improve over \system{Mixed-Nat} for a majority of domains, and are the only ones to clearly fulfill [P1-LAB];} \system{WDCMT} also improves on three domains, but regresses on one.
+All MDMTs (except \system{DM} and \system{ADM}) slightly improve over the generic baseline (for most domains), these improvements are rarely significant.
+\revision{Among systems using an extra domain feature, \system{DC-Tags} has a small edge over \system{DC-Feats} and also requires less paramaters; it also outperforms \system{TTM}, which however uses predicted rather than gold domain tags.
+\system{TTM} is also the best choice among the systems that to not use domain tags in inference.} \revision{The best contender overall are  \system{FT Res} and \system{MDL Res}, which significantly improve over \system{Mixed-Nat} for a majority of domains, and are the only ones to clearly fulfill [P1-LAB];} \system{WDCMT} also improves on three domains, but regresses on one.
 The use of a dedicated adaptation component, seems here to prove better than feature-based strategies; \revision{this comes at the cost of a large increase of the number of parameters.} The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}).
+
 % for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.\fyDone{why LDC does not outperform DC as in Minh's paper (for en2fr MED) ?}
@@
 All systems show some robustness with respect to fuzzy domain boundaries; this is mostly notable for \system{ADM}, suggesting that when domain are close, ignoring domain differences is effective.
-In contrary, \system{FT-Full} incurs large losses across the board, especially for the small data condition.
+In contrary, \system{FT-Full} incurs clear losses across the board, especially for the small data condition \cite{Miceli-barone17regularization}.
 Even in this very favourable case however, very few MDMT systems are able to significantly outperform \system{FT-Full} and this is only observed for the smaller part of the \domain{med} domain.
@@
 In the first (\domain{rnd}), we use test data from the domains seen in training, presented with a random domain tag\fyDone{all configs use the same tag or is randomization is run for each config}.
-In this situation, the loss with respect to using the correct domain is generally large (more than 10 BLEU points), showing an overall failure to meet requirement [P4-ERR], except for systems which disregard domain tags in testing.
+In this situation, the loss with respect to using the correct tag is generally large (more than 10 BLEU points), showing an overall failure to meet requirement [P4-ERR], except for systems that ignore domain tags in testing.
 \fyDone{Comments?}
@@
 In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the most likely domain tag and use it for decoding.\footnote{\revision{To automatically predict domain tags, we train a language model in each domain and assign tag on a per-sentence basis based the language model log-probability (assuming uniform domain priors).
-Our domain classifier has a average prediction error of 16.4\% for in-domain data.}}
+Our domain classifier has an average prediction error of 16.4\% for in-domain data.}}
 % Results are in the rightmost column of Table~\ref{tab:redomains}.
-In this configuration, again, systems using domain tags in the inference step perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\fyDone{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}
+In this configuration, again, systems using domain tags in the inference step perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\fyDone{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}t
 
@@
   \centering
-  \begin{tabular}{|p{2.5cm}|*{9}{r|}} \hline \footnotesize
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-     \system{Mixed-Nat}  & 37.1  & 54.1  & 49.6	& 34.1  & 42.1	& 77.0 & 28.9 & 40.8	& 49.0 \\ 
+  \begin{tabular}{|p{1.7cm}|*{9}{r|}} \hline
+    \footnotesize
+   \hfill Domain  Model \hfill & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
+    \footnotesize\system{Mixed-Nat}  & 37.1  & 54.1  & 49.6	& 34.1  & 42.1	& 77.0 & 28.9 & 40.8	& 49.0 \\[-2pt]
+                   & \sbcl{-0.2}{--}  &\sbcl{-0.5}{---} &\sbcl{-0.5}{--} &\sbcl{+0.6}{--} &\sbcl{-1.1}{--} &\sbcl{-6.5}{--} &\sbcl{+5.4}{--} &\sbcl{-0.3}{--} & \sbcl{-0.4}{--} \\
 %    \system{Mixed-Nat}  & 37.3  & 54.6   & 50.1   & 33.5  & 43.2  & 77.5  & 23.5 & 41.1 & 49.4 \\
@@
     \hline%\hline
-    \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5  & 41.4 & 49.9\\[-2pt]
+    \footnotesize \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5  & 41.4 & 49.9\\[-2pt]
 %                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
-                   & \scriptsize +0.3  & \scriptsize -0.1   & \scriptsize -0.6    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.3}  & \scriptsize \SW{-3.5}  & \scriptsize \SW{-1.4}  & \scriptsize -0.1 & \scriptsize \SW{-1.2}\\
+                   & \sbcl{-0.4}{+0.3}  & \sbcl{-0.8}{-0.1}   & \sbcl{-0.1}{-0.6}  & \sbcl{+1.6}{\SW{-1.1}}  &  \sbcl{+0.4}{\SW{-1.3}}  & \sbcl{-1.7}{\SW{-3.5}}  & \sbcl{+7.7}{\SW{-1.4}}  & \sbcl{-0.2}{-0.1} & \sbcl{-0.2}{\SW{-1.2}}\\
 
-	\system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9  & 41.2 & 50.1\\[-2pt]
+    \footnotesize \system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9  & 41.2 & 50.1\\[-2pt]
 %                   & 37.2 & 54.7   & 49.9    & 34.2  &  43.0  & 79.9 & \SW{28.1}  & &\\
@@
 
-    \system{LDR}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
+    \footnotesize \system{LDR}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
                     & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize -0.4 & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize +\SB{3.8} & \scriptsize \SW{-0.9} & \scriptsize +0.5 & \scriptsize -0.4 \\
 
-    \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.2  & 29.1 & 41 & 49.4\\[-2pt]
+    \footnotesize \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.2  & 29.1 & 41 & 49.4\\[-2pt]
 %                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
@@
 
-    \system{DM}            &36.0 &51.3&46.8&31.8&39.8&65.7&27.0 & 38.9 & 45.2\\[-2pt]
+    \footnotesize \system{DM}            &36.0 &51.3&46.8&31.8&39.8&65.7&27.0 & 38.9 & 45.2\\[-2pt]
 %                   &36.5&51.7&47.4&31.7&40.3&65.7&\SW{25.8}& & \\
@@
 
-    \system{ADM}          &36.6&54.2&49.1&32.9&42.1&75.7&28.7 & 40.2 & 48.4 \\[-2pt]
+    \footnotesize \system{ADM}          &36.6&54.2&49.1&32.9&42.1&75.7&28.7 & 40.2 & 48.4 \\[-2pt]
 %                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
@@
 %                    &&&&&&&&& \\
-    \revision{\system{FT Res}}  &  \\
-
-    \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & 41.9 & 51.8 \\[-2pt]
+    \footnotesize \revision{\system{FT Res}}  & 37.3 & 57.9 & 53.9 & 33.8 &	46.7 & 90.1 & 20.5 &	53.3 & 42.2 \\[-2pt]
+                               & \sbcl{}{+0.2} & \sbcl{}{-2.2} & \sbcl{}{-2.8} & \sbcl{}{-0.7} & \sbcl{}{-2.6} & \sbcl{}{-3.4} & \sbcl{}{-3.0} & \sbcl{}{-1.9} & \sbcl{}{-0.6} \\
+    % 37.53	55.74	51.15	33.06	44.15	86.67	23.05	51.38	41.64
+    \footnotesize \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & 41.9 & 51.8 \\[-2pt]
 %                      & 37.5 & 56.1 & 51.1 & 34.0  &  44.3 & 87.3 & \SW{28.4} & & 51.7 & \\
@@
   \end{tabular}
-  \caption{Ability to handle a new domain: BLEU scores for the ``resume training'' mode compared to a complete training session.
-Averages only take into account the six initial domains (\domain{News} excluded).
+  \caption{Ability to handle a new domain.
+\revision{We report BLEU scores for a complete training session with 7 domains, as well as differences with (left) training with 6 domains (from Table~\ref{tab:performance}); (right) continuous training mode.} Averages only take into account the six initial domains (\domain{News} excluded).
 Underline denotes a significant loss between warmstart and coldstart, bold a significant gain.}
@@
 
-We can first compare the results of coldstart training with six (Table~\ref{tab:performance}) or seven domains (Table~\ref{tab:warmrestart}): the most striking observation is that the extra training data is detrimental for almost all systems and domains\fyDone{is this significant ?}.
-Increasing the heterogeneity of the mix with supplementary out-of-domain data hurts performance across the board.
+We can first compare the results of coldstart training with six or seven domains (Table~\ref{tab:warmrestart}): the most striking observation is that the extra training data is detrimental for most systems and domains, except for \domain{News} where we see a large gain.\fyDone{is this significant ?}.
+Increasing the heterogeneity of the mix with supplementary out-of-domain data thus seems to hurt performance across the board.
 % Adding is not helping here, especially when domains are well separated.
-Comparing now the coldstart with the warmstart scenarios, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training can negatively impact the performance for other domains.
+Comparing now the coldstart with the warmstart scenario, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training also negatively impact the performance for other domains.
 This happens notably for \system{DC-Tag}, \system{TTM} and \system{ADM}.
 In this setting \system{MDL Res} and \system{DM} show the smaller average loss, with the former achieving the best balance of training cost and average translation performance.% of all systems.
-
+\fyTodo{Numbers for MDL Res need significancy testing; and they are weird - why do we see this ? Because the generic is bad ?}
 % \begin{itemize}
@@
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+    Domain / Model  & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
   %    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
@@
     \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 39.5 & 47.1\\% & 23.3 \\
+    \revision{\system{FT Res}}     & 37.5 & \SW{55.7}  & \SW{51.1}   & 33.1   &  \SW{44.1}  & \SW{86.7} & 41.6 & \SW{51.4}\\%  & \SW{21.2} \\
+    \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SW{50.7}\\%  & \SW{21.2} \\
     \system{WDCMT}       & 35.6 & 53.1 & 48.4 & 30.5 & \SW{37.7} & \SW{56.0} & 38.5 & 43.6 \\ % & 20.4 \\ 
-    \revision{\system{FT Res}}     & 37.5 & 55.7  & 51.1   & 33.1   &  44.1  & \SW{86.7} & 41.6 & 51.4\\%  & \SW{21.2} \\
-    \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SB{50.7}\\%  & \SW{21.2} \\
     \hline
   \end{tabular}
-  \caption{Translation performance with automatic domains, computed with the original test sets. Significancy tests are for comparisons with the 6-domain scenario.}
+  \caption{Translation performance with automatic domains, computed with the original test sets.
+Significancy tests are for comparisons with the 6-domain scenario \revision{(Table~\ref{tab:performance}).}}
   \label{tab:subdomains}
@@
 If MDMT systems are able to outperform the mixed-domain baseline, at least for some domains, they all fall short to match the performance of fine-tuning on each individual domain, which remains the best choice in multi-source single domain adaptation.
-In this scenario, architecture-based strategies aimed at parameter sharing and/or representation normalization obtain the best results.
-\fyDone{Mitigate this}
+In this scenario, architecture-based strategies aimed at parameter sharing and/or representation normalization obtain the best results.\fyDone{Mitigate this}
 As expected however, MDMTs are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend, dynamically accommodate additional domains, this is especially easy for feature-based methodologies.
@@
 \revision{%
-The description of model in each MDMT approach is as follows:
+The description of the systems used in this study :
 \begin{itemize}

### 2020-08-27 2020/tacl2020/rebuttal.tex
fixes in the bib, plus some small changes in the text

@@
 We have also tried to clarify the distinction between multi-domain MT and fine-tuning; while the former is designed with an a priori knowledge that the test data is sampled from a single domain, the latter assumes that the test data will be sampled from multiple domains, using domain probabilities that are not known a priori.
-This raises the issue of robustness wrt to adversarial test distribution and even new domains, an issue that is irrelevant for DA.
+This raises the issues of robustness wrt to adversarial test distribution and even new domains, two issues that are somewhat irrelevant for DA.
 
-In our experiments, fine-tuned systems share no parameter and show large drops in performance when translating data from the wrong domain (even though this could be somewhat mitigated, as the reviewer points out).
+In our experiments, fine-tuned systems share no parameter and show large drops in performance when translating data from the wrong domain (even though this could be somewhat mitigated through regularization, as the reviewer points out).
 By contrast, the majority of the parameters of our MDMT systems are shared for all domains.

### 2020-08-27 2020/tacl2020/multidomain.tex
fixes in the bib, plus some small changes in the text

@@
 A related question concerns the relationship between supervised domain adaptation and multi-domain translation.
-The latter task seems more challenging as adapting a system to a known target domain $A$ looks easier than training a MDMT system, which should do also well on domains other than $A$.
-Are there situations where MD systems can surpass DA, as is sometimes expected?
+The latter task seems more challenging \revision{as it tries to optimize MT performance for a more diverse set of potential inputs, with the additional uncertainty regarding the distribution of test data.} Are there still situations where where MD systems can surpass single domain adaptation, as is sometimes expected?
 
@@
 A multi-domain learner is then expected to use these tags effectively \cite{Joshi12multidomain} when computing the combined translation function $h(x,d)$, and to perform well in all domains \cite{Finkel09hierarchical}.
-Note that this setting is closely related to the multi-source adaptation problem formalized in \cite{Mansour09domainadaptation,Mansour09multiple,Hoffman18algorithms}.
+This setting is closely related to the multi-source adaptation problem formalized in \cite{Mansour09domainadaptation,Mansour09multiple,Hoffman18algorithms}.
 
-This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is \cite{Farajian17multidomain}, where test translations rely on similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here.
+This definition seems to be the most accepted view of a multi-domain MT\footnote{An exception is \citep{Farajian17multidomain}, where test translations rely on similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here.
 Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function.
@@
 \fyDone{Remove Gated residual}
-Our comparison of multi-domain systems includes the following reimplementations of recent proposals from the literature:\footnote{\revision{Further implementation details are in Appendix~A.}}
+Our comparison of multi-domain systems includes our own reimplementations of recent proposals from the literature:\footnote{\revision{Further implementation details are in Appendix~A.}}
 \begin{itemize}
@@
 \system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
-Training uses reference tags and inference is performed with predicted tags, just like for regular target words.
+Training uses reference tags and inference is usually performed with predicted tags, just like for regular target words.
 \system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
@@
 \end{itemize}
-
-Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges from our notion of MDMT, which should be able to handle available domain labels in training and\revision{/or} testing.
+\revision{Note that this list includes systems that slightly depart from our definition of MDMT: standard implementations of \system{TTM} and \system{WDCMT} rely on infered, rather than on gold, domain tags - which must somewhat affect their predictions; \system{DM} and {ADM} make no use of domain tags at all.} We however did not consider the proposal of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges more strongly from our notion of MDMT.
+% and should be able to handle available domain labels in training and testing.
 % This system has shown to meet several of requirements and would also have no problem scaling to a large number of domains.\fyDone{TBC}
@@
 This happens notably for \system{DC-Tag}, \system{TTM} and \system{ADM}.
-In this setting \system{MDL Res} and \system{DM} show the smaller average loss, with the former achieving the best balance of training cost and average translation performance.% of all systems.
+In this setting \system{MDL Res} and \system{DM} show the smaller average loss, with the former achieving the best balance of training cost and average BLEU score.% of all systems.
 \fyTodo{Numbers for MDL Res need significancy testing; and they are weird - why do we see this ? Because the generic is bad ?}
@@
 
-Finally, \newcite{Farajian17multidomain}, \newcite{li-etal-2018-one} and \newcite{xu2019} follow a different strategy.
+Finally, \citet{Farajian17multidomain}, \citet{Li18onesentence} and \citet{xu2019} follow a different strategy.
 Each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some iterations, before delivering its output.
@@
 If MDMT systems are able to outperform the mixed-domain baseline, at least for some domains, they all fall short to match the performance of fine-tuning on each individual domain, which remains the best choice in multi-source single domain adaptation.
-In this scenario, architecture-based strategies aimed at parameter sharing and/or representation normalization obtain the best results.\fyDone{Mitigate this}
-As expected however, MDMTs are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend, dynamically accommodate additional domains, this is especially easy for feature-based methodologies.
-Our experiments finally suggest that all methods show decreasing performance when the number of domain increases (for a fixed data size).
+% In this scenario, architecture-based strategies aimed at parameter sharing and/or representation normalization obtain the best results.\fyDone{Mitigate this}
+As expected however, MDMTs are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend, dynamically accommodate additional domains, this being especially easy for feature-based methodologies.
+Our experiments finally suggest that all methods show decreasing performance when the number of domains or the diversity of the domain mixture increases.
+% (for a fixed data size).
 
@@
 Both problems are of practical relevance in industrial settings.
-Second, and maybe more importantly, there is a general need to adopt better evaluation methodologies for evaluating MDMT systems, which require systems developers to clearly spell out the testing conditions and the associated expected distribution of testing instances, and to report more than mere comparisons with simple baselines on a fixed and known handful of domains.
+Second, and maybe more importantly, there is a general need to adopt better evaluation methodologies for evaluating MDMT systems, which require systems developers to clearly spell out the testing conditions and the associated expected distribution of testing instances, and to report more than comparisons with simple baselines on a fixed and known handful of domains.
 

### 2020-08-27 2020/tacl2020/multidomain.tex
revised conclusion

@@
 If MDMT systems are able to outperform the mixed-domain baseline, at least for some domains, they all fall short to match the performance of fine-tuning on each individual domain, which remains the best choice in multi-source single domain adaptation.
-% In this scenario, architecture-based strategies aimed at parameter sharing and/or representation normalization obtain the best results.\fyDone{Mitigate this}
+\revisiondel{In this scenario, architecture-based strategies aimed at parameter sharing and/or representation normalization obtain the best results.}\fyDone{Mitigate this}
 As expected however, MDMTs are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend, dynamically accommodate additional domains, this being especially easy for feature-based methodologies.

### 2020-08-27 2020/tacl2020/multidomain.tex
table 5
section new domain

@@
 This is detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with training all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.\footnote{WDCMT results are excluded from this table, as resuming training proved difficult to implement.}\fyDone{Ecrire un commentaire}
-We expect that MDMT systems should not to be significantly impacted by the addition of a new domain and reach about the same performance as when training with this domain from scratch.
+We expect that MDMT systems should not be too significantly impacted by the addition of a new domain and reach about the same performance as when training with this domain from scratch.
 From a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{DC-Feat} or \system{TTM}, for which new domains merely add new labels.
-It is less easy for \system{DM}, \system{ADM} and \system{WDCMT}, which include a built-in domain classifier whose outputs have to be given, or for \system{LDR} and \system{MDL~Res} for which the number of possible domains is built in the architecture and has to be anticipated from the start.
+It is less easy for \system{DM}, \system{ADM} and \system{WDCMT}, which include a built-in domain classifier whose outputs have to be pre-specified, or for \system{LDR}, \system{FT~Res} and \system{MDL~Res} for which the number of possible domains is built in the architecture and has to be anticipated from the start.
 This makes a difference between domain-bounded systems, for which the number of domains is limited and truly open-domain systems.% , which seamlessly integrate additional domains.

### 2020-08-27 2020/tacl2020/multidomain.tex
removed a typo

@@
 % Results are in the rightmost column of Table~\ref{tab:redomains}.
-In this configuration, again, systems using domain tags in the inference step perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\fyDone{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}t
+In this configuration, again, systems using domain tags in the inference step perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\fyDone{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}
 

### 2020-08-27 2020/tacl2020/multidomain.tex
30 cluster table

@@
 \newcommand{\sbcl}[2]{{\scriptsize #1 \hfill $|$ \hfill  #2}}
-
+\usepackage{multirow}
 \title{Revisiting Multi-Domain Machine Translation}
@@
 %                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
-                   & \sbcl{-0.4}{+0.3}  & \sbcl{-0.8}{-0.1}   & \sbcl{-0.1}{-0.6}  & \sbcl{+1.6}{\SW{-1.1}}  &  \sbcl{+0.4}{\SW{-1.3}}  & \sbcl{-1.7}{\SW{-3.5}}  & \sbcl{+7.7}{\SW{-1.4}}  & \sbcl{-0.2}{-0.1} & \sbcl{-0.2}{\SW{-1.2}}\\
+                   & \sbcl{+0.3}{+0.3}  & \sbcl{\SB{+0.8}}{-0.1}   & \sbcl{-0.04}{-0.6}  & \sbcl{\SW{-1.6}}{\SW{-1.1}}  &  \sbcl{-0.4}{\SW{-1.3}}  & \sbcl{\SB{+1.7}}{\SW{-3.5}}  & \sbcl{\SW{-7.7}}{\SW{-1.4}}  & \sbcl{0.2}{-0.1} & \sbcl{0.1}{\SW{-1.1}}\\
 
@@
 %                   & 37.2 & 54.7   & 49.9    & 34.2  &  43.0  & 79.9 & \SW{28.1}  & &\\
-                   & \scriptsize -0.2 & \scriptsize -0.1  & \scriptsize -0.1 & \scriptsize -0.6 & \scriptsize -0.9 & \scriptsize +0.3 & \scriptsize \SW{-0.8} & \scriptsize -0.2 & \scriptsize -0.3\\
+                   & \scriptsize \sbcl{+0.3}{-0.2} & \scriptsize \sbcl{-0.1}{-0.1}  & \scriptsize \sbcl{-0.3}{-0.1} & \scriptsize \sbcl{\SW{-1.3}}{-0.6} & \scriptsize \sbcl{-0.1}{-0.9} & \scriptsize \sbcl{+0.4}{+0.3} & \scriptsize \sbcl{\SW{-7.3}}{\SW{-0.8}} & \scriptsize \sbcl{+0.1}{-0.2} & \scriptsize \sbcl{-0.2}{-0.3}\\
 
     \footnotesize \system{LDR}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
-                    & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize -0.4 & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize +\SB{3.8} & \scriptsize \SW{-0.9} & \scriptsize +0.5 & \scriptsize -0.4 \\
+                    & \scriptsize \sbcl{0.0}{-0.6} &	\scriptsize \sbcl{+0.1}{+0.5} & \scriptsize \sbcl{+0.2}{-0.4} & \scriptsize \sbcl{-0.4}{-0.6} &	\scriptsize \sbcl{+0.5}{+0.5} & \scriptsize \sbcl{\SB{+2.9}}{+\SB{3.8}} & \scriptsize \sbcl{\SW{-6.6}}{\SW{-0.9}} & \scriptsize \sbcl{+0.6}{+0.5} & \scriptsize \sbcl{+0.1}{-0.4} \\
 
@@
 %                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
-                    & \scriptsize -0.3  & \scriptsize -0.3  & \scriptsize -0.5    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.0}  & \scriptsize \SW{-4.0}  & \scriptsize \SW{-1.4}& \scriptsize -0.5 & \scriptsize \SW{-1.2}\\
+                    & \scriptsize \sbcl{0.0}{-0.3}  & \scriptsize \sbcl{+0.4}{-0.3}  & \scriptsize \sbcl{-0.1}{-0.5}    & \scriptsize \sbcl{-0.9}{\SW{-1.1}}  &  \scriptsize \sbcl{+0.6}{\SW{-1.0}}  & \scriptsize \sbcl{\SB{+1.8}}{\SW{-4.0}} & \scriptsize \sbcl{\SW{-5.7}}{\SW{-1.4}} & \scriptsize \sbcl{0.0}{-0.5} & \scriptsize \sbcl{+0.3}{\SW{-1.2}}\\
 
@@
 %                   &36.5&51.7&47.4&31.7&40.3&65.7&\SW{25.8}& & \\
-                   & \scriptsize +0.6& \scriptsize +0.4& \scriptsize +0.6& \scriptsize -0.1& \scriptsize +0.5 & \scriptsize 0.0& \scriptsize \SW{-1.2}& \scriptsize +0.5 & \scriptsize +0.3\\
+                   & \scriptsize \sbcl{-0.4}{+0.6} & \scriptsize \sbcl{\SW{-1.8}}{+0.4} & \scriptsize \sbcl{\SW{-1.2}}{+0.6} & \scriptsize \sbcl{\SW{-1.8}}{-0.1} & \scriptsize \sbcl{\SW{-2.6}}{+0.5} & \scriptsize \sbcl{\SW{-3.3}}{0.0} & \scriptsize \sbcl{\SW{-4.4}}{\SW{-1.2}}& \scriptsize \sbcl{-0.8}{+0.5} & \scriptsize \sbcl{\SW{-1.8}}{+0.3}\\
 
@@
 %                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
-                   & \scriptsize +0.3 & \scriptsize -0.8 & \scriptsize -0.8 & \scriptsize -0.2 & \scriptsize -0.4 & \scriptsize \SW{-5.0} & \scriptsize \SW{-1.9} & \scriptsize -0.2 & \scriptsize \SW{-1.1}\\
+                   & \scriptsize \sbcl{-0.2}{+0.3} & \scriptsize \sbcl{-0.7}{-0.8} & \scriptsize \sbcl{-0.8}{-0.8} & \scriptsize \sbcl{-0.9}{-0.2} & \scriptsize \sbcl{-0.5}{-0.4} & \scriptsize \sbcl{\SW{-2.3}}{\SW{-5.0}} & \scriptsize \sbcl{\SW{-5.4}}{\SW{-1.9}} & \scriptsize \sbcl{-0.5}{-0.2} & \scriptsize \sbcl{\SW{-0.9}}{\SW{-1.1}}\\
 
@@
     \footnotesize \revision{\system{FT Res}}  & 37.3 & 57.9 & 53.9 & 33.8 &	46.7 & 90.1 & 20.5 &	53.3 & 42.2 \\[-2pt]
-                               & \sbcl{}{+0.2} & \sbcl{}{-2.2} & \sbcl{}{-2.8} & \sbcl{}{-0.7} & \sbcl{}{-2.6} & \sbcl{}{-3.4} & \sbcl{}{-3.0} & \sbcl{}{-1.9} & \sbcl{}{-0.6} \\
+                               & \sbcl{+0.3}{+0.3} & \sbcl{+0.4}{+0.4} & \sbcl{+0.1}{+0.1} & \sbcl{-0.7}{-0.7} & \sbcl{+0.5}{+0.5} & \sbcl{-0.9}{-0.9} & \sbcl{\SW{-9}}{-0.6} & \sbcl{+0.2}{+0.2} & \sbcl{-0.1}{-0.1} \\
     % 37.53	55.74	51.15	33.06	44.15	86.67	23.05	51.38	41.64
@@
 %                      & 37.5 & 56.1 & 51.1 & 34.0  &  44.3 & 87.3 & \SW{28.4} & & 51.7 & \\
-                        & \scriptsize -0.2 & \scriptsize +0.5 & \scriptsize 0.0 & \scriptsize -0.4  &  \scriptsize -0.2 & \scriptsize -0.2 & \scriptsize \SW{-0.8} & \scriptsize -0.2 & \scriptsize -0.1 \\
+                        & \scriptsize \sbcl{+0.2}{-0.2} & \scriptsize \sbcl{+0.4}{+0.5} & \scriptsize \sbcl{+0.1}{0.0} & \scriptsize \sbcl{-0.9}{-0.4}  &  \scriptsize \sbcl{-0.1}{-0.2} & \scriptsize \sbcl{0.9}{-0.2} & \scriptsize \sbcl{\SW{-8.0}}{\SW{-0.8}} & \scriptsize \sbcl{0.1}{-0.2} & \scriptsize \sbcl{0.1}{-0.1} \\
      \hline
@@
 }
+
+\begin{table*}[t]
+  \centering
+  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
+   Cluster/ & \multirow{2}{*}{size} &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
+   Model& & \system{-Nat} & \system{Full} & \system{Res} &\system{Res} & \system{-Feat}& \system{-Tag}& & & & \\ \hline
+24(med)&8129 / 3(law)&90.4&90.4&90.4&90.4&100.0&65.6&100.0&90.4&100.0&100.0 \\
+13(law)&17260 / 52&67.6&75.4&74.3&74.3&75.0&54.7&74.7&75.9&65.9&76.9 \\
+28(med)&25561 / 54&71.6&68.7&68.1&70.2&71.0&42.5&72.0&71.3&65.6&72.6 \\
+19(IT)&27227 / 88&58.5&63.0&60.9&63.9&63.7&57.2&59.4&61.1&60.5&60.3 \\
+0(IT)&27436 / 72&43.9&33.3&45.4&45.4&49.9&15.4&46.8&49.2&46.6&47.8 \\
+22(bank)&27534 / 103&91.5&93.7&93.4&93.9&92.5&72.8&92.3&93.2&91.4&93.4 \\
+25(med)&28237 / 56&57.0&44.8&48.2&49.1&54.6&47.2&49.8&54.2&45.1&52.4 \\
+16(med)&30443 / 18&57.2&70.4&77.4&73.5&61.8&54.2&58.4&58.1&52.5&58.3 \\
+23(med)&47065 / 23&24.5&27.2&26.5&28.5&30.5&27.3&32.0&24.4&29.0&29.8 \\
+17(med)&54378 / 26&39.9&40.3&41.6&38.0&37.1&36.6&35.2&35.4&31.3&33.7 \\
+8(IT)&61353 / 214&46.9&53.1&55.8&53.6&48.9&45.1&48.8&50.9&43.0&46.7 \\
+1(med)&68108 / 122&47.2&47.5&48.7&45.1&46.8&39.1&45.4&44.2&40.7&44.9 \\
+7(med)&91464 / 30&41.3&35.5&41.4&39.9&41.4&36.5&37.3&37.1&40.7&41.8 \\
+11(med)&92982 / 38&31.6&42.6&31.8&35.4&36.0&29.6&36.7&32.7&26.5&36.6 \\
+29(law)&109260 / 242&65.9&69.2&67.6&67.7&66.0&63.8&65.1&64.7&62.4&65.9 \\
+27(med)&109324 / 49&11.0&9.6&8.7&9.2&10.0&19.4&9.4&7.9&10.7&10.6 \\
+5(med)&109949 / 267&46.3&47.4&46.9&45.4&44.0&42.9&43.7&44.3&40.9&45.7 \\
+6(med)&133395 / 73&37.2&38.9&38.7&36.8&37.5&27.5&38.0&37.2&31.3&35.9 \\
+26(med)&134828 / 428&31.8&30.8&31.8&31.2&31.9&32.6&32.2&30.5&29.6&31.2 \\
+15(bank)&136944 / 674&46.5&51.5&47.9&48.0&46.6&46.0&45.8&45.7&42.9&46.0 \\
+4(law)&137451 / 1016&77.1&85.3&83.5&83.3&75.8&46.1&74.2&73.3&63.2&75.9 \\
+2(med)&182594 / 85&70.6&75.8&71.7&69.4&68.2&67.3&67.3&68.6&65.6&68.2 \\
+20(med)&182990 / 71&47.4&47.2&46.8&47.2&48.4&47.5&48.8&47.3&47.1&46.8 \\
+21(IT)&222802 / 868&38.7&38.8&39.0&37.2&37.5&35.9&36.9&37.1&33.4&37.0 \\
+10(med)&225347 / 115&40.0&42.6&40.0&38.2&39.9&35.8&39.5&39.1&36.3&40.7 \\
+18(med)&245000 / 106&57.7&60.3&58.7&58.6&58.4&56.3&57.3&56.1&54.9&55.9 \\
+9(med)&301639 / 145&37.2&37.3&36.5&36.1&36.4&37.7&36.4&35.2&34.2&37.0 \\
+3(law)&323474 / 680&50.1&52.0&50.8&50.1&49.1&48.3&49.0&48.2&44.4&49.1 \\
+14(med)&333957 / 146&31.6&31.4&31.9&33.0&32.5&34.1&31.4&32.1&30.5&31.8 \\
+12(med)&356377 / 148&36.3&36.6&35.9&35.9&35.8&37.0&36.4&35.4&34.2&36.3 \\ \hline
+  \end{tabular}
+  \caption{Detailed results of the experiments with automatic domains}
+  \label{tab:automatic_domains}
+\end{table*}
+
 \end{document}

### 2020-08-27 2020/tacl2020/multidomain.tex
30 cluster table

@@
   \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
-   Cluster/ & \multirow{2}{*}{size} &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
-   Model& & \system{-Nat} & \system{Full} & \system{Res} &\system{Res} & \system{-Feat}& \system{-Tag}& & & & \\ \hline
+   Model/ & \multirow{2}{*}{size} &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
+   Cluster & & \system{-Nat} & \system{Full} & \system{Res} &\system{Res} & \system{-Feat}& \system{-Tag}& & & & \\ \hline
 24(med)&8129 / 3(law)&90.4&90.4&90.4&90.4&100.0&65.6&100.0&90.4&100.0&100.0 \\
@@
   \label{tab:automatic_domains}
+\end{table*}
+
+\begin{table*}[t]
+\centering
+  \begin{tabular}{|p{2.3cm}|*{11}{c|}} \hline
+   Model/ & \multirow{2}{*}{Averaged size} &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
+   Group cluster & & \system{-Nat} & \system{Full} & \system{Res} &\system{Res} & \system{-Feat}& \system{-Tag}& & & & \\ \hline
+10 small&49.5&60.2&60.7&62.6&62.7&63.6&47.3&62.1&61.3&58.8&62.5\\
+10 mid&213.7&40.6&42.6&41.9&41.2&40.9&38.3&40.2&39.5&36.9&40.5\\
+10 large&338&48.7&50.7&49.5&48.9&48.2&44.6&47.7&47.2&44.4&47.9\\
+Avg& &49.8&51.3&51.3&50.9&50.9&43.4&50.0&49.4&46.7&50.29 \\ \hline
+
+  \end{tabular}
+   \caption{Average over group of 10 large, medium, small clusters}
+  \label{tab:avg_automatic_domains}
 \end{table*}

### 2020-08-27 2020/tacl2020/multidomain.tex
more changes in final sections

--- 2020-08-27 2020/tacl2020/multidomain.tex
@@
 Training mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}.
-Under this view, MDMT would be especially beneficial for domains with little training data \revision{(see also \cite{Sharaf20metalearning} for a discussion of DA in low-resource conditions and a meta-learning based approach to this problem)}\fyTodo{Best place ?}.
+Under this view, MDMT would be especially beneficial for domains with little training data.
 This is observed for multilingual MT from English: an improvement for under-resourced languages due to positive transfer, at the cost of a decrease in performance for well-resourced languages \cite{Arivazhagan19massively}.
@@
 Combining multiple domain-specific MTs can also be justified in the sake of distributional robustness \cite{Mansour09domainadaptation,Mansour09multiple}, for instance when the test mixture differs from the train mixture, or when it includes new domains unseen in training.
-\sout{Such scenario is already well documented for zero-shot multilingual MT} \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual}\sout{where mixing languages has more than demonstrated its usefulness.}
+\revisiondel{Such scenario is already well documented for zero-shot multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual} where mixing languages has more than demonstrated its usefulness.}
 An even more challenging case is when the MT would need to perform well for any test distribution, as studied for statistical MT in \cite{Huck15mixeddomain}.
@@
 % \subsubsection{Performance}
-A first expectation is that MDMT systems should perform well in the face of mixed-domain test data. From this, we derive the following three requirements.
+A first expectation is that MDMT systems should perform well in the face of mixed-domain test data. We thus derive the following requirements.
 
@@
 \paragraph{[P4-ERR]}
-A MDMT should perform best when the true domain tag is known, but deteriorate gracefully in the presence of tag errors; in this situation, catastrophic drops in performance can be observed\fyDone{cite \cite{McCloskey89catastrophic}}.
-This requirement can be assessed by translating test texts with erroneous domain tags and reporting the correlated decrease of performance.
+MDMTs should perform best when the true domain tag is known, but deteriorate gracefully in the face of tag errors; in this situation, catastrophic drops in performance are often observed\fyDone{cite \cite{McCloskey89catastrophic}}.
+This requirement can be assessed by translating test texts with erroneous domain tags and reporting the subsequent loss in performance.
 
@@
 The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems adapted for one domain, with no parameter sharing across domains.
-\revisiondel{We again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a small adaptation module that is added (with residual connections) on top of every Transformer layer (\system{FT Res}).}
+\revisiondel{We again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a small adaptation module that is added (with residualconnections) on top of every Transformer layer (\system{FT-Res}).}
 \end{itemize}
 
-For all models, we set the embeddings size and the hidden layers size to~512.
-Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
-The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024.
+All models use embeddings and the hidden layers sizes of dimension~512.
+Transformers contain with 8 attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells.
+The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024.
 % The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
 % The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training use a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
+Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.\fyDone{Describe the block adaptation layer - voir slides}
 % All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
@@
 \begin{itemize}
-\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
+\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or as a supplementary feature for each word (\system{DC-Feat}).
 \item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDR});\fyDone{why not LDR as in Minh's paper?}
@@
 These methods thus only use domain tags in training.
-\item \revision{two multi-domain versions of the approach of \newcite{Bapna19simple}, denoted \system{FT Res} and \system{MDL Res}, where a domain-specific adaptation module is added to all the Transformer layers; within each layer, residual connections enable to by-pass this module.
-The former variant corresponds to the original proposal of \citet{Bapna19simple} and fine-tunes the adapter modules of a \system{Mixed-Nat} system independently for each domain, keeping all the other parameters frozen.
-The latter is novel; it uses the same architecture, but a different training procedure, as all parameters are jointly learned from scratch from a mix-domain corpus.
-}
+\item the multi-domain model of \newcite{Zeng18multidomain} (\system{WDCMT}), where a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification and adversarial training are used to compute these representations.
+\revision{Again, inference does not use domain tags.}\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly improved, results.}\fyDone{Check this}
+\item \revision{two multi-domain versions of the approach of \newcite{Bapna19simple}, denoted \system{FT-Res} and \system{MDL-Res}, where a domain-specific adaptation module is added to all the Transformer layers; within each layer, residual connections enable to by-pass this module.
+The former variant corresponds to the original proposal of \citet{Bapna19simple} \revision{(see also \cite{Sharaf20metalearning})}.
+It fine-tunes the adapter modules of a \system{Mixed-Nat} system independently for each domain, keeping all the other parameters frozen.
+The latter uses the same architecture, but a different training procedure and learns all parameters jointly from scratch with a mix-domain corpus.}
   % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
 % the entire multi-domain system from scratch.\fyDone{Check this.}
-\item the multi-domain model of \newcite{Zeng18multidomain} (\system{WDCMT}), where a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification and adversarial training are used to compute these representations.
-\revision{As for the three previous proposals, inference does not use domain tags.}\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly improved, results.}\fyDone{Check this}
 \end{itemize}
-\revision{Note that this list includes systems that slightly depart from our definition of MDMT: standard implementations of \system{TTM} and \system{WDCMT} rely on infered, rather than on gold, domain tags - which must somewhat affect their predictions; \system{DM} and {ADM} make no use of domain tags at all.} We however did not consider the proposal of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges more strongly from our notion of MDMT.
+\revision{This list includes systems that slightly depart from our definition of MDMT: standard implementations of \system{TTM} and \system{WDCMT} rely on infered, rather than on gold, domain tags - which must somewhat affect their predictions; \system{DM} and {ADM} make no use of domain tags at all.} We however did not consider the proposal of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges more strongly from our notion of MDMT.
 % and should be able to handle available domain labels in training and testing.
@@
 
-Fine-tuning each domain separately provides us with a better baseline, outperforming \system{Mixed-Nat} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
+Fine-tuning each domain separately yields a better baseline, outperforming \system{Mixed-Nat} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
 % A complete fine-tuning, updating all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to keeping fixed the capacity of the adapter module.
@@
 
-All MDMTs (except \system{DM} and \system{ADM}) slightly improve over the generic baseline (for most domains), these improvements are rarely significant.
-\revision{Among systems using an extra domain feature, \system{DC-Tags} has a small edge over \system{DC-Feats} and also requires less paramaters; it also outperforms \system{TTM}, which however uses predicted rather than gold domain tags.
-\system{TTM} is also the best choice among the systems that to not use domain tags in inference.} \revision{The best contender overall are  \system{FT Res} and \system{MDL Res}, which significantly improve over \system{Mixed-Nat} for a majority of domains, and are the only ones to clearly fulfill [P1-LAB];} \system{WDCMT} also improves on three domains, but regresses on one.
-The use of a dedicated adaptation component, seems here to prove better than feature-based strategies; \revision{this comes at the cost of a large increase of the number of parameters.} The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}).
+All MDMTs (except \system{DM} and \system{ADM}) slightly improve over \system{Mixed-Nat}(for most domains), but these gains are rarely significant.
+\revision{Among systems using an extra domain feature, \system{DC-Tag} has a small edge over \system{DC-Feat} and also requires less paramaters; it also outperforms \system{TTM}, which however uses predicted rather than gold domain tags.
+\system{TTM} is also the best choice among the systems that to not use domain tags in inference.} \revision{The best contenders overall are  \system{FT-Res} and \system{MDL-Res}, which significantly improve over \system{Mixed-Nat} for a majority of domains, and are the only ones to clearly fulfill [P1-LAB];} \system{WDCMT} also improves on three domains, but regresses on one.
+The use of a dedicated adaptation component, seems here better than feature-based strategies, \revision{but yields a large increase of the number of parameters.} The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}).
 
@@
 This might be due to the fact that domains are well separated (cf.\ Section~\ref{ssec:corpora}) and are hardly helping each other.
-In this situation, MDMT systems should instead make sure that each domain is allocated a sufficient number of parameters, so as to close the gap with fine-tuning.
+In this situation, MDMT systems should dedicate a sufficient number of parameters to each domain, so as to close the gap with fine-tuning.
 \fyDone{More comments when we have all the results}\fyDone{Importance of sharing and unsharing}
@@
 
-In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the most likely domain tag and use it for decoding.\footnote{\revision{To automatically predict domain tags, we train a language model in each domain and assign tag on a per-sentence basis based the language model log-probability (assuming uniform domain priors).
-Our domain classifier has an average prediction error of 16.4\% for in-domain data.}}
+In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the domain tag and use it for decoding.\footnote{\revision{Domain tags are assigned as follows: we train a language model for each domain and assign tag on a per-sentence basis based the language model log-probability (assuming uniform domain priors).
+The domain classifier has an average prediction error of 16.4\% for in-domain data.}}
 % Results are in the rightmost column of Table~\ref{tab:redomains}.
-In this configuration, again, systems using domain tags in the inference step perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\fyDone{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}
+In this configuration, again, systems using domain tags during inference perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\fyDone{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}
 
@@
 From a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{DC-Feat} or \system{TTM}, for which new domains merely add new labels.
-It is less easy for \system{DM}, \system{ADM} and \system{WDCMT}, which include a built-in domain classifier whose outputs have to be pre-specified, or for \system{LDR}, \system{FT~Res} and \system{MDL~Res} for which the number of possible domains is built in the architecture and has to be anticipated from the start.
+It is less easy for \system{DM}, \system{ADM} and \system{WDCMT}, which include a built-in domain classifier whose outputs have to be pre-specified, or for \system{LDR}, \system{FT-Res} and \system{MDL-Res} for which the number of possible domains is built in the architecture and has to be anticipated from the start.
 This makes a difference between domain-bounded systems, for which the number of domains is limited and truly open-domain systems.% , which seamlessly integrate additional domains.
@@
   \centering
-  \begin{tabular}{|p{1.7cm}|*{9}{r|}} \hline
+  \begin{tabular}{|p{1.6cm}|*{9}{r|}} \hline
     \footnotesize
@@
     \footnotesize\system{Mixed-Nat}  & 37.1  & 54.1  & 49.6	& 34.1  & 42.1	& 77.0 & 28.9 & 40.8	& 49.0 \\[-2pt]
-                   & \sbcl{-0.2}{--}  &\sbcl{-0.5}{---} &\sbcl{-0.5}{--} &\sbcl{+0.6}{--} &\sbcl{-1.1}{--} &\sbcl{-6.5}{--} &\sbcl{+5.4}{--} &\sbcl{-0.3}{--} & \sbcl{-0.4}{--} \\
+                   & \sbcl{+0.2}{--}  &\sbcl{+0.5}{---} &\sbcl{+0.5}{--} &\sbcl{-0.6}{--} &\sbcl{+1.1}{--} &\sbcl{+0.5}{--} &\sbcl{-5.4}{--} &\sbcl{+0.3}{--} & \sbcl{+0.4}{--} \\
 %    \system{Mixed-Nat}  & 37.3  & 54.6   & 50.1   & 33.5  & 43.2  & 77.5  & 23.5 & 41.1 & 49.4 \\
@@
 %                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
-                   & \sbcl{-0.4}{+0.3}  & \sbcl{-0.8}{-0.1}   & \sbcl{-0.1}{-0.6}  & \sbcl{+1.6}{\SW{-1.1}}  &  \sbcl{+0.4}{\SW{-1.3}}  & \sbcl{-1.7}{\SW{-3.5}}  & \sbcl{+7.7}{\SW{-1.4}}  & \sbcl{-0.2}{-0.1} & \sbcl{-0.2}{\SW{-1.2}}\\
+                   & \sbcl{+0.3}{+0.3}  & \sbcl{\SB{+0.8}}{-0.1}   & \sbcl{-0.04}{-0.6}  & \sbcl{\SW{-1.6}}{\SW{-1.1}}  &  \sbcl{-0.4}{\SW{-1.3}}  & \sbcl{\SB{+1.7}}{\SW{-3.5}}  & \sbcl{\SW{-7.7}}{\SW{-1.4}}  & \sbcl{+0.2}{-0.1} & \sbcl{+0.1}{\SW{-1.1}}\\
 
@@
 %                   & 37.2 & 54.7   & 49.9    & 34.2  &  43.0  & 79.9 & \SW{28.1}  & &\\
-                   & \scriptsize -0.2 & \scriptsize -0.1  & \scriptsize -0.1 & \scriptsize -0.6 & \scriptsize -0.9 & \scriptsize +0.3 & \scriptsize \SW{-0.8} & \scriptsize -0.2 & \scriptsize -0.3\\
+                   &  \sbcl{+0.3}{-0.2} & \sbcl{-0.1}{-0.1}  & \sbcl{-0.3}{-0.1} & \sbcl{\SW{-1.3}}{-0.6} & \sbcl{-0.1}{-0.9} & \scriptsize \sbcl{+0.4}{+0.3} & \sbcl{\SW{-7.3}}{\SW{-0.8}} & \sbcl{+0.1}{-0.2} & \sbcl{-0.2}{-0.3}\\
 
     \footnotesize \system{LDR}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
-                    & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize -0.4 & \scriptsize -0.6 &	\scriptsize +0.5 & \scriptsize +\SB{3.8} & \scriptsize \SW{-0.9} & \scriptsize +0.5 & \scriptsize -0.4 \\
+                    & \sbcl{0.0}{-0.6} &  \sbcl{+0.1}{+0.5} & \sbcl{+0.2}{-0.4} &  \sbcl{-0.4}{-0.6} &  \sbcl{+0.5}{+0.5} & \sbcl{\SB{+2.9}}{+\SB{3.8}} & \sbcl{\SW{-6.6}}{\SW{-0.9}} & \sbcl{+0.6}{+0.5} &  \sbcl{+0.1}{-0.4} \\
 
-    \footnotesize \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.2  & 29.1 & 41 & 49.4\\[-2pt]
+    \footnotesize \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.2  & 29.1 & 41.0 & 49.4\\[-2pt]
 %                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
-                    & \scriptsize -0.3  & \scriptsize -0.3  & \scriptsize -0.5    & \scriptsize \SW{-1.1}  &  \scriptsize \SW{-1.0}  & \scriptsize \SW{-4.0}  & \scriptsize \SW{-1.4}& \scriptsize -0.5 & \scriptsize \SW{-1.2}\\
+                    & \sbcl{0.0}{-0.3}  & \sbcl{+0.4}{-0.3}  & \sbcl{-0.1}{-0.5}  & \sbcl{-0.9}{\SW{-1.1}}  & \sbcl{+0.6}{\SW{-1.0}}  & \sbcl{\SB{+1.8}}{\SW{-4.0}} & \sbcl{\SW{-5.7}}{\SW{-1.4}} & \sbcl{0.0}{-0.5} & \sbcl{+0.3}{\SW{-1.2}}\\
 
@@
 %                   &36.5&51.7&47.4&31.7&40.3&65.7&\SW{25.8}& & \\
-                   & \scriptsize +0.6& \scriptsize +0.4& \scriptsize +0.6& \scriptsize -0.1& \scriptsize +0.5 & \scriptsize 0.0& \scriptsize \SW{-1.2}& \scriptsize +0.5 & \scriptsize +0.3\\
+                   & \sbcl{-0.4}{+0.6} & \sbcl{\SW{-1.8}}{+0.4} & \sbcl{\SW{-1.2}}{+0.6} & \sbcl{\SW{-1.8}}{-0.1} & \sbcl{\SW{-2.6}}{+0.5} & \sbcl{\SW{-3.3}}{0.0} & \sbcl{\SW{-4.4}}{\SW{-1.2}} & \sbcl{-0.8}{+0.5} & \sbcl{\SW{-1.8}}{+0.3}\\
 
@@
 %                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
-                   & \scriptsize +0.3 & \scriptsize -0.8 & \scriptsize -0.8 & \scriptsize -0.2 & \scriptsize -0.4 & \scriptsize \SW{-5.0} & \scriptsize \SW{-1.9} & \scriptsize -0.2 & \scriptsize \SW{-1.1}\\
+                   & \sbcl{-0.2}{+0.3} & \sbcl{-0.7}{-0.8} & \sbcl{-0.8}{-0.8} & \sbcl{-0.9}{-0.2} & \sbcl{-0.5}{-0.4} & \sbcl{\SW{-2.3}}{\SW{-5.0}} & \sbcl{\SW{-5.4}}{\SW{-1.9}} & \sbcl{-0.5}{-0.2} & \sbcl{\SW{-0.9}}{\SW{-1.1}}\\
 
@@
 %                    &&&&&&&&& \\
-    \footnotesize \revision{\system{FT Res}}  & 37.3 & 57.9 & 53.9 & 33.8 &	46.7 & 90.1 & 20.5 &	53.3 & 42.2 \\[-2pt]
-                               & \sbcl{}{+0.2} & \sbcl{}{-2.2} & \sbcl{}{-2.8} & \sbcl{}{-0.7} & \sbcl{}{-2.6} & \sbcl{}{-3.4} & \sbcl{}{-3.0} & \sbcl{}{-1.9} & \sbcl{}{-0.6} \\
+    \footnotesize \revision{\system{FT Res}}  & 37.0 & 57.6 & 53.8 & 34.5 &	46.1 & 91.1 & 29.6 &	42.2  & 53.3 \\[-2pt]
+                               & \sbcl{+0.3}{+0.3} & \sbcl{+0.4}{+0.4} & \sbcl{+0.1}{+0.1} & \sbcl{-0.7}{-0.7} & \sbcl{+0.5}{+0.5} & \sbcl{-0.9}{-0.9} & \sbcl{\SW{-9.0}}{-0.6} & \sbcl{-0.1}{-0.1} & \sbcl{+0.2}{+0.2} \\
+
     % 37.53	55.74	51.15	33.06	44.15	86.67	23.05	51.38	41.64
@@
 %                      & 37.5 & 56.1 & 51.1 & 34.0  &  44.3 & 87.3 & \SW{28.4} & & 51.7 & \\
-                        & \scriptsize -0.2 & \scriptsize +0.5 & \scriptsize 0.0 & \scriptsize -0.4  &  \scriptsize -0.2 & \scriptsize -0.2 & \scriptsize \SW{-0.8} & \scriptsize -0.2 & \scriptsize -0.1 \\
+                        &  \sbcl{+0.2}{-0.2} & \sbcl{+0.4}{+0.5} & \sbcl{+0.1}{0.0} & \sbcl{-0.9}{-0.4}  & \sbcl{-0.1}{-0.2} &  \sbcl{+0.9}{-0.2} & \sbcl{\SW{-8.0}}{\SW{-0.8}} & \sbcl{+0.1}{-0.2} & \sbcl{+0.1}{-0.1} \\
      \hline
@@
   \caption{Ability to handle a new domain.
-\revision{We report BLEU scores for a complete training session with 7 domains, as well as differences with (left) training with 6 domains (from Table~\ref{tab:performance}); (right) continuous training mode.} Averages only take into account the six initial domains (\domain{News} excluded).
-Underline denotes a significant loss between warmstart and coldstart, bold a significant gain.}
+\revision{We report BLEU scores for a complete training session with 7 domains, as well as differences with (left) training with 6 domains (from Table~\ref{tab:performance}); (right) continuous training mode.} Averages only take into account six domains (\domain{News} excluded).
+Underline denotes a significant loss, bold a significant gain.}
   \label{tab:warmrestart}
@@
 
-We can first compare the results of coldstart training with six or seven domains (Table~\ref{tab:warmrestart}): the most striking observation is that the extra training data is detrimental for most systems and domains, except for \domain{News} where we see a large gain.\fyDone{is this significant ?}.
-Increasing the heterogeneity of the mix with supplementary out-of-domain data thus seems to hurt performance across the board.
+We can first compare the results of coldstart training with six or seven domains in Table~\ref{tab:warmrestart}: a first observation is that the extra training data \revision{is hardly helping} for most domains, except for \domain{News}, where we see a large gain, and for \domain{talk}.
+The picture is the same when one looks at MDMTs, where only the weakest systems (\system{DM}, \system{ADM}) seem to benefit from more (out-of-domain) data.\fyDone{is this significant ?}
+% Increasing the heterogeneity of the mix with supplementary out-of-domain data does not always improve the performance.% across the board.
 % Adding is not helping here, especially when domains are well separated.
@@
 In this setting \system{MDL Res} and \system{DM} show the smaller average loss, with the former achieving the best balance of training cost and average BLEU score.% of all systems.
-\fyTodo{Numbers for MDL Res need significancy testing; and they are weird - why do we see this ? Because the generic is bad ?}
+\fyDone{Numbers for MDL Res need significancy testing; and they are weird - why do we see this ? Because the generic is bad ?}
 % \begin{itemize}
@@
 In this section, we experiment with automatic domains, obtained by clustering sentences into $k=30$ classes using the k-means algorithm based on generic sentence representations obtained via mean pooling as in Section~\ref{ssec:corpora}.
-This allows us to evaluate requirement [P7-scale], training and testing our systems as if these domains were genuine.
-We also note that many of these classes are mere split of the large \domain{med}, while a fewer number of classes are mixtures of two existing domains.
-We are thus in a position to reiterate at a larger scale the measurements of Section~\ref{ssec:redomains} and test whether multi-domain systems can effectively take advantage from the cross-domain similarities, and to eventually perform better than fine-tuning.
+This allows us to evaluate requirement [P7-scale], training and testing our systems as if these domains were fully separated.
+Many of these clusters are mere splits of the large \domain{med}, while a fewer number of classes are mixtures of two (or more) existing domains \revision{(details are in Appendix B)}.
+We are thus in a position to reiterate, at a larger scale, the measurements of Section~\ref{ssec:redomains} and test whether multi-domain systems can effectively take advantage from the cross-domain similarities and to eventually perform better than fine-tuning.
+\revision{Results in Table~\ref{} show that MDMT actually surpass fine-tuning for the smaller clusters, especially for small subsets of the medical domain; for the large clusters, this is no longer true.}
 
-Table~\ref{tab:subdomains} reports the effect of using automatic domain tags for the 6 domain-specific test sets: each sentence was first assigned to an automatic class, translated with the corresponding multi-domain system with 30 classes; aggregate numbers were then computed, and contrasted with the 6-domain scenario.
+Finally, Table~\ref{tab:subdomains} reports the effect of using automatic domain for each of the 6 test sets: each sentence was first assigned to an automatic class, translated with the corresponding multi-domain system with 30 classes; aggregate numbers were then computed, and contrasted with the 6-domain scenario.
 Results are clear and confirm previous observations: even though some clusters are very close, the net effect is a loss in performance for almost all systems and conditions.
@@
 
-Another parameter-sharing scheme is \cite{Jiang19multidomain}, which augments a Transformer model with domain-specific heads, whose contributions are regulated at the word/position level: some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling that the core of \cite{Huck15mixeddomain,Saunders19ucam}.
+Another parameter-sharing scheme is in \cite{Jiang19multidomain}, which augments a Transformer model with domain-specific heads, whose contributions are regulated at the word/position level: some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling that the core of \cite{Huck15mixeddomain,Saunders19ucam}.
 The results for three language pairs outperform several standard baselines for a two-domain systems (in fr:en and de:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
 
-Finally, \citet{Farajian17multidomain}, \citet{Li18onesentence} and \citet{xu2019} follow a different strategy.
+Finally, \citet{Farajian17multidomain}, \citet{Li18onesentence} and \citet{Xu19lexical} adopt a different strategy.
 Each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some iterations, before delivering its output.
@@
 \revision{%
-The description of the systems used in this study :
+The systems used in this study are set up as follows:
 \begin{itemize}
@@
 \paragraph{Training} For each domain, we create train/dev/test sets by randomly each corpus. We maintain the size of validation sets and of test sets equal to 1,000 lines for every domain.
-The learning rate is set to $lr = d_{model}^{-0.5}*\min(step\_num^{-0.5},step\_num * warmup\_steps^{-1.5})$ where $d_{model}=512$,$warmup\_steps=4000$, as in \cite{Vaswani17attention}.
-For the fine-tuning procedures used for \system{FT-full} and \system{FT Res}, we continue training using the same learning rate schedule, continuing the incrementation of $step\_num$.
+The learning rate is set
+% to $lr = d_{model}^{-0.5}*\min(step\_num^{-0.5},step\_num * warmup\_steps^{-1.5})$ where $d_{model}=512$,$warmup\_steps=4000$,
+as in \cite{Vaswani17attention}.
+For the fine-tuning procedures used for \system{FT-full} and \system{FT Res}, we continue training using the same learning rate schedule, continuing the incrementation of the number of steps.
 All other MDMT systems reported in Tables~\ref{tab:performance} and \ref{tab:redomains} use a combined validation set comprising 6,000 lines, obtained by merging the six development sets.
 For the results in Table~\ref{tab:warmrestart} we also append the validation set of \domain{news} to the multi-domain validation set.
-In any case, training stops if either the training iteration reach a maximum number of XXX\fyTodo{How many ?} or the score on the validation set does not increase in three consecutive evaluations.
-We average five checkpoints up to the best checkpoint to get the final model.\fyTodo{up to the best ?}
+In any case, training stops if either training reaches the maximum number of iterations (50,000)\fyDone{How many ?} or the score on the validation set does not increase for three consecutive evaluations.
+We average five checkpoints to get the final model.\fyDone{up to the best ?}
 }

### 2020-08-27 2020/tacl2020/multidomain.tex

@@
 }
+
+\begin{table*}[t]
+  \centering
+  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
+   Model/ & \multirow{2}{*}{size} &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
+   Cluster & & \system{-Nat} & \system{Full} & \system{Res} &\system{Res} & \system{-Feat}& \system{-Tag}& & & & \\ \hline
+24(med)&8129 / 3(law)&90.4&90.4&90.4&90.4&100.0&65.6&100.0&90.4&100.0&100.0 \\
+13(law)&17260 / 52&67.6&75.4&74.3&74.3&75.0&54.7&74.7&75.9&65.9&76.9 \\
+28(med)&25561 / 54&71.6&68.7&68.1&70.2&71.0&42.5&72.0&71.3&65.6&72.6 \\
+19(IT)&27227 / 88&58.5&63.0&60.9&63.9&63.7&57.2&59.4&61.1&60.5&60.3 \\
+0(IT)&27436 / 72&43.9&33.3&45.4&45.4&49.9&15.4&46.8&49.2&46.6&47.8 \\
+22(bank)&27534 / 103&91.5&93.7&93.4&93.9&92.5&72.8&92.3&93.2&91.4&93.4 \\
+25(med)&28237 / 56&57.0&44.8&48.2&49.1&54.6&47.2&49.8&54.2&45.1&52.4 \\
+16(med)&30443 / 18&57.2&70.4&77.4&73.5&61.8&54.2&58.4&58.1&52.5&58.3 \\
+23(med)&47065 / 23&24.5&27.2&26.5&28.5&30.5&27.3&32.0&24.4&29.0&29.8 \\
+17(med)&54378 / 26&39.9&40.3&41.6&38.0&37.1&36.6&35.2&35.4&31.3&33.7 \\
+8(IT)&61353 / 214&46.9&53.1&55.8&53.6&48.9&45.1&48.8&50.9&43.0&46.7 \\
+1(med)&68108 / 122&47.2&47.5&48.7&45.1&46.8&39.1&45.4&44.2&40.7&44.9 \\
+7(med)&91464 / 30&41.3&35.5&41.4&39.9&41.4&36.5&37.3&37.1&40.7&41.8 \\
+11(med)&92982 / 38&31.6&42.6&31.8&35.4&36.0&29.6&36.7&32.7&26.5&36.6 \\
+29(law)&109260 / 242&65.9&69.2&67.6&67.7&66.0&63.8&65.1&64.7&62.4&65.9 \\
+27(med)&109324 / 49&11.0&9.6&8.7&9.2&10.0&19.4&9.4&7.9&10.7&10.6 \\
+5(med)&109949 / 267&46.3&47.4&46.9&45.4&44.0&42.9&43.7&44.3&40.9&45.7 \\
+6(med)&133395 / 73&37.2&38.9&38.7&36.8&37.5&27.5&38.0&37.2&31.3&35.9 \\
+26(med)&134828 / 428&31.8&30.8&31.8&31.2&31.9&32.6&32.2&30.5&29.6&31.2 \\
+15(bank)&136944 / 674&46.5&51.5&47.9&48.0&46.6&46.0&45.8&45.7&42.9&46.0 \\
+4(law)&137451 / 1016&77.1&85.3&83.5&83.3&75.8&46.1&74.2&73.3&63.2&75.9 \\
+2(med)&182594 / 85&70.6&75.8&71.7&69.4&68.2&67.3&67.3&68.6&65.6&68.2 \\
+20(med)&182990 / 71&47.4&47.2&46.8&47.2&48.4&47.5&48.8&47.3&47.1&46.8 \\
+21(IT)&222802 / 868&38.7&38.8&39.0&37.2&37.5&35.9&36.9&37.1&33.4&37.0 \\
+10(med)&225347 / 115&40.0&42.6&40.0&38.2&39.9&35.8&39.5&39.1&36.3&40.7 \\
+18(med)&245000 / 106&57.7&60.3&58.7&58.6&58.4&56.3&57.3&56.1&54.9&55.9 \\
+9(med)&301639 / 145&37.2&37.3&36.5&36.1&36.4&37.7&36.4&35.2&34.2&37.0 \\
+3(law)&323474 / 680&50.1&52.0&50.8&50.1&49.1&48.3&49.0&48.2&44.4&49.1 \\
+14(med)&333957 / 146&31.6&31.4&31.9&33.0&32.5&34.1&31.4&32.1&30.5&31.8 \\
+12(med)&356377 / 148&36.3&36.6&35.9&35.9&35.8&37.0&36.4&35.4&34.2&36.3 \\ \hline
+  \end{tabular}
+  \caption{Detailed results of the experiments with automatic domains}
+  \label{tab:automatic_domains}
+\end{table*}
+
+\begin{table*}[t]
+\centering
+  \begin{tabular}{|p{2.3cm}|*{11}{c|}} \hline
+   Model/ & \multirow{2}{*}{Averaged size} &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
+   Group cluster & & \system{-Nat} & \system{Full} & \system{Res} &\system{Res} & \system{-Feat}& \system{-Tag}& & & & \\ \hline
+10 small&49.5&60.2&60.7&62.6&62.7&63.6&47.3&62.1&61.3&58.8&62.5\\
+10 mid&213.7&40.6&42.6&41.9&41.2&40.9&38.3&40.2&39.5&36.9&40.5\\
+10 large&338&48.7&50.7&49.5&48.9&48.2&44.6&47.7&47.2&44.4&47.9\\
+Avg& &49.8&51.3&51.3&50.9&50.9&43.4&50.0&49.4&46.7&50.29 \\ \hline
+
+  \end{tabular}
+   \caption{Average over group of 10 large, medium, small clusters}
+  \label{tab:avg_automatic_domains}
+\end{table*}
+
 \end{document}

### 2020-08-28 2020/tacl2020/multidomain.tex
version de travail presque finale

--- 2020-08-27 2020/tacl2020/multidomain.tex
@@
     \system{ADM}      \hfill{\footnotesize[+0]}         & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 38.9 & 47.5 \\% & 23.3 \\
-    \revision{\system{FT Res}}   \hfill{\footnotesize[+12.4m]}  & 37.3 & \SB{57.9} & \SB{53.9} & 33.8 & \SB{46.7} & \SB{90.2}  & \SB{42.3} & \SB{53.3} \\ % & 20.5\\ \hline
-    \system{MDL Res} \hfill{\footnotesize[+12.4m]}    & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & \SB{51.9} \\%  & \SW{21.2} \\
+    \revision{\system{FT-Res}}   \hfill{\footnotesize[+12.4m]}  & 37.3 & \SB{57.9} & \SB{53.9} & 33.8 & \SB{46.7} & \SB{90.2}  & \SB{42.3} & \SB{53.3} \\ % & 20.5\\ \hline
+    \system{MDL-Res} \hfill{\footnotesize[+12.4m]}    & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & \SB{51.9} \\%  & \SW{21.2} \\
 %     \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
@@
     \system{ADM} (*)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & 0.0&-0.2\\
-    \revision{\system{FT Res}}   & -0.1   & -0.4 & -0.3 &-0.3 & \SW{-2.2} & \SW{-2.9} & \SW{-2.4} & -\SW{3.2} & \SW{-13.3} & \SW{-3.0}\\ % [31.8]
-    \system{MDL Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +0.7 & -0.3 & \SW{-18.6} & \SW{-1.3}\\ % [31.8]
+    \revision{\system{FT-Res}}   & -0.1   & -0.4 & -0.3 &-0.3 & \SW{-2.2} & \SW{-2.9} & \SW{-2.4} & -\SW{3.2} & \SW{-13.3} & \SW{-3.0}\\ % [31.8]
+    \system{MDL-Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +0.7 & -0.3 & \SW{-18.6} & \SW{-1.3}\\ % [31.8]
     \system{WDCMT} (*)     & -0.0    & -0.0  & +0.2 & +0.2  & +0.8 & +0.8  & -0.4 & -0.8 & 0.0 & +0.2 \\
@@
 %                    &&&&&&&&& \\
-    \footnotesize \revision{\system{FT Res}}  & 37.0 & 57.6 & 53.8 & 34.5 &	46.1 & 91.1 & 29.6 &	42.2  & 53.3 \\[-2pt]
+    \footnotesize \revision{\system{FT-Res}}  & 37.0 & 57.6 & 53.8 & 34.5 &	46.1 & 91.1 & 29.6 &	42.2  & 53.3 \\[-2pt]
                                & \sbcl{+0.3}{+0.3} & \sbcl{+0.4}{+0.4} & \sbcl{+0.1}{+0.1} & \sbcl{-0.7}{-0.7} & \sbcl{+0.5}{+0.5} & \sbcl{-0.9}{-0.9} & \sbcl{\SW{-9.0}}{-0.6} & \sbcl{-0.1}{-0.1} & \sbcl{+0.2}{+0.2} \\
@@
 We are thus in a position to reiterate, at a larger scale, the measurements of Section~\ref{ssec:redomains} and test whether multi-domain systems can effectively take advantage from the cross-domain similarities and to eventually perform better than fine-tuning.
-\revision{Results in Table~\ref{} show that MDMT actually surpass fine-tuning for the smaller clusters, especially for small subsets of the medical domain; for the large clusters, this is no longer true.}
+\revision{The results in Table~\ref{tab:avg_automatic_domains} suggest that MDMT actually can surpass fine-tuning for the smaller clusters, especially for small subsets of the medical domain; for the large clusters, this is no longer true.}
 
@@
     \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 39.5 & 47.1\\% & 23.3 \\
-    \revision{\system{FT Res}}     & 37.5 & \SW{55.7}  & \SW{51.1}   & 33.1   &  \SW{44.1}  & \SW{86.7} & 41.6 & \SW{51.4}\\%  & \SW{21.2} \\
-    \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SW{50.7}\\%  & \SW{21.2} \\
+    \revision{\system{FT-Res}}     & 37.5 & \SW{55.7}  & \SW{51.1}   & 33.1   &  \SW{44.1}  & \SW{86.7} & 41.6 & \SW{51.4}\\%  & \SW{21.2} \\
+    \system{MDL-Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SW{50.7}\\%  & \SW{21.2} \\
     \system{WDCMT}       & 35.6 & 53.1 & 48.4 & 30.5 & \SW{37.7} & \SW{56.0} & 38.5 & 43.6 \\ % & 20.4 \\ 
@@
 % \bibliographystyle{coling}
-\bibliographystyle{acl_natbib}
-\bibliography{multidomain}
 
-\newpage\newpage
+% \newpage\newpage
 \section*{Appendices}
@@
 Training use a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.
-\item \system{FT Res} and \system{MDL res} use the same medium Transformer model, and add residual layers with a bottleneck dimension of size 1024.
+\item \system{FT-Res} and \system{MDL-res} use the same medium Transformer model, and add residual layers with a bottleneck dimension of size 1024.
   % \cite{Bapna19simple} showed that if the bottleneck dimension is large enough, the gap between \system{FT-block} and \system{FT-full} will be small.\fyTodo{This comment is confusing}.
@@
 as in \cite{Vaswani17attention}.
-For the fine-tuning procedures used for \system{FT-full} and \system{FT Res}, we continue training using the same learning rate schedule, continuing the incrementation of the number of steps.
+For the fine-tuning procedures used for \system{FT-full} and \system{FT-Res}, we continue training using the same learning rate schedule, continuing the incrementation of the number of steps.
 All other MDMT systems reported in Tables~\ref{tab:performance} and \ref{tab:redomains} use a combined validation set comprising 6,000 lines, obtained by merging the six development sets.
@@
 }
+
+\begin{table*}[t]
+  \centering
+  \footnotesize
+  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
+   Model  & size&\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
+   Cluster & train / test & \system{-Nat} & \system{Full} & \system{Res} &\system{Res} & \system{-Feat}& \system{-Tag}& & & & \\ \hline
+24 \hfill [med]&8.1k / 3 &90.4&90.4&90.4&90.4&100.0&65.6&100.0&90.4&100.0&100.0 \\
+13(law)&17.3k / 52&67.6&75.4&74.3&74.3&75.0&54.7&74.7&75.9&65.9&76.9 \\
+28(med)&25.6k / 54&71.6&68.7&68.1&70.2&71.0&42.5&72.0&71.3&65.6&72.6 \\
+19(IT)&27.2k / 88&58.5&63.0&60.9&63.9&63.7&57.2&59.4&61.1&60.5&60.3 \\
+0(IT)&27.4k / 72&43.9&33.3&45.4&45.4&49.9&15.4&46.8&49.2&46.6&47.8 \\
+22(bank)&27534 / 103&91.5&93.7&93.4&93.9&92.5&72.8&92.3&93.2&91.4&93.4 \\
+25(med)&28237 / 56&57.0&44.8&48.2&49.1&54.6&47.2&49.8&54.2&45.1&52.4 \\
+16(med)&30443 / 18&57.2&70.4&77.4&73.5&61.8&54.2&58.4&58.1&52.5&58.3 \\
+23(med)&47065 / 23&24.5&27.2&26.5&28.5&30.5&27.3&32.0&24.4&29.0&29.8 \\
+17(med)&54378 / 26&39.9&40.3&41.6&38.0&37.1&36.6&35.2&35.4&31.3&33.7 \\
+8(IT)&61353 / 214&46.9&53.1&55.8&53.6&48.9&45.1&48.8&50.9&43.0&46.7 \\
+1(med)&68108 / 122&47.2&47.5&48.7&45.1&46.8&39.1&45.4&44.2&40.7&44.9 \\
+7(med)&91464 / 30&41.3&35.5&41.4&39.9&41.4&36.5&37.3&37.1&40.7&41.8 \\
+11(med)&92982 / 38&31.6&42.6&31.8&35.4&36.0&29.6&36.7&32.7&26.5&36.6 \\
+29(law)&109260 / 242&65.9&69.2&67.6&67.7&66.0&63.8&65.1&64.7&62.4&65.9 \\
+27(med)&109324 / 49&11.0&9.6&8.7&9.2&10.0&19.4&9.4&7.9&10.7&10.6 \\
+5(med)&109949 / 267&46.3&47.4&46.9&45.4&44.0&42.9&43.7&44.3&40.9&45.7 \\
+6(med)&133395 / 73&37.2&38.9&38.7&36.8&37.5&27.5&38.0&37.2&31.3&35.9 \\
+26(med)&134828 / 428&31.8&30.8&31.8&31.2&31.9&32.6&32.2&30.5&29.6&31.2 \\
+15(bank)&136944 / 674&46.5&51.5&47.9&48.0&46.6&46.0&45.8&45.7&42.9&46.0 \\
+4(law)&137451 / 1016&77.1&85.3&83.5&83.3&75.8&46.1&74.2&73.3&63.2&75.9 \\
+2(med)&182594 / 85&70.6&75.8&71.7&69.4&68.2&67.3&67.3&68.6&65.6&68.2 \\
+20(med)&182990 / 71&47.4&47.2&46.8&47.2&48.4&47.5&48.8&47.3&47.1&46.8 \\
+21(IT)&222802 / 868&38.7&38.8&39.0&37.2&37.5&35.9&36.9&37.1&33.4&37.0 \\
+10(med)&225347 / 115&40.0&42.6&40.0&38.2&39.9&35.8&39.5&39.1&36.3&40.7 \\
+18(med)&245000 / 106&57.7&60.3&58.7&58.6&58.4&56.3&57.3&56.1&54.9&55.9 \\
+9(med)&301639 / 145&37.2&37.3&36.5&36.1&36.4&37.7&36.4&35.2&34.2&37.0 \\
+3(law)&323474 / 680&50.1&52.0&50.8&50.1&49.1&48.3&49.0&48.2&44.4&49.1 \\
+14(med)&333957 / 146&31.6&31.4&31.9&33.0&32.5&34.1&31.4&32.1&30.5&31.8 \\
+12(med)&356377 / 148&36.3&36.6&35.9&35.9&35.8&37.0&36.4&35.4&34.2&36.3 \\ \hline
+  \end{tabular}
+  \caption{Full results for the experiments with automatic domains.
+For each cluster, we report the majority domain in the case where one domain accounts for more than XXX\% of the data, the training and test size, and the BLEU scores obtained with the various systems used in this study.
+Boldface denotes the best system in each line; test sets are two small to report significativity tests.}
+  \label{tab:automatic_domains}
+\end{table*}
+
+\begin{table*}[t]
+\centering \footnotesize
+  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
+   Model/ & Test &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
+   Clusters & size & \system{Nat} & \system{Full} & \system{Res} &\system{Res} & \system{Feat}& \system{Tag}& & & & \\ \hline
+10 small&49.5&60.2&60.7&62.6&62.7&\bf 63.6&47.3&62.1&61.3&58.8&62.5\\
+10 mid&213.7&40.6&\bf 42.6&41.9&41.2&40.9&38.3&40.2&39.5&36.9&40.5\\
+10 large&338&48.7&\bf 50.7&49.5&48.9&48.2&44.6&47.7&47.2&44.4&47.9\\
+Avg& &49.8&\bf 51.3&\bf 51.3&50.9&50.9&43.4&50.0&49.4&46.7&50.29 \\ \hline
+
+  \end{tabular}
+   \caption{Average BLEU scores for 10 smaller, medium, and larger clusters.
+Best score for each group is in boldface.
+For the small clusters, full-fine tuning is outperformed by several MDMT system - see detail in Table~\ref{tab:automatic_domains}.}
+  \label{tab:avg_automatic_domains}
+\end{table*}
+
+\flushbottom{}
+
+\newpage\newpage
+
+\bibliographystyle{acl_natbib}
+\bibliography{multidomain}
+
 \end{document}
@@
 \todos{}
-
-\begin{table}
-  \centering
-  \begin{tabular}{|p{3.2cm}|*{30}{r|}} \hline
-    Model / Set-up &d1&d2&d3&d4&d5&d6&d7&d8&d9&d10&d11&d12&d13&d14&d15
-    &d16&d17&d18&d19&d20&d21&d22&d23&d24&d25&d26&d27&d28&d29&d30 \\\hline
-    %
-    & \multicolumn{30}{|p{20cm}|}{%
-      }% \hline
-    \\
-    %
-    domain size ($\times 10^3$) &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\ \hline\hline
-    %
-   \system{FT-Full}      &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\  
-    \system{DC-Tag}    &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\     
-    \system{DC-Feat}   &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\ 
-    \system{TTM}         &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\ 
-    \system{LDR}  &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \\
-
-    \system{MDL Res} 
-%    &&&&&&&&&&&&&&&&&&&&&&&&&&&&&& \hline
-  \end{tabular}
-  \caption{Detailed results of the experiments with automatic domains}
-  \label{tab:automatic_domains}
-\end{table}
 

### 2020-08-31 2020/tacl2020/multidomain.tex
local

--- 2020-08-27 2020/tacl2020/multidomain.tex
@@
 }
+
+\begin{table*}[t]
+  \centering
+  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
+   Model/ & \multirow{2}{*}{size} &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
+   Cluster & & \system{-Nat} & \system{Full} & \system{Res} &\system{Res} & \system{-Feat}& \system{-Tag}& & & & \\ \hline
+24(med)&8129 / 3&90.4&90.4&90.4&90.4&100.0&65.6&100.0&90.4&100.0&100.0 \\
+13(law)&17260 / 52&67.6&75.4&74.3&74.3&75.0&54.7&74.7&75.9&65.9&76.9 \\
+28(med)&25561 / 54&71.6&68.7&68.1&70.2&71.0&42.5&72.0&71.3&65.6&72.6 \\
+19(IT)&27227 / 88&58.5&63.0&60.9&63.9&63.7&57.2&59.4&61.1&60.5&60.3 \\
+0(IT)&27436 / 72&43.9&33.3&45.4&45.4&49.9&15.4&46.8&49.2&46.6&47.8 \\
+22(bank)&27534 / 103&91.5&93.7&93.4&93.9&92.5&72.8&92.3&93.2&91.4&93.4 \\
+25(med)&28237 / 56&57.0&44.8&48.2&49.1&54.6&47.2&49.8&54.2&45.1&52.4 \\
+16(med)&30443 / 18&57.2&70.4&77.4&73.5&61.8&54.2&58.4&58.1&52.5&58.3 \\
+23(med)&47065 / 23&24.5&27.2&26.5&28.5&30.5&27.3&32.0&24.4&29.0&29.8 \\
+17(med)&54378 / 26&39.9&40.3&41.6&38.0&37.1&36.6&35.2&35.4&31.3&33.7 \\
+8(IT)&61353 / 214&46.9&53.1&55.8&53.6&48.9&45.1&48.8&50.9&43.0&46.7 \\
+1(med)&68108 / 122&47.2&47.5&48.7&45.1&46.8&39.1&45.4&44.2&40.7&44.9 \\
+7(med)&91464 / 30&41.3&35.5&41.4&39.9&41.4&36.5&37.3&37.1&40.7&41.8 \\
+11(med)&92982 / 38&31.6&42.6&31.8&35.4&36.0&29.6&36.7&32.7&26.5&36.6 \\
+29(law)&109260 / 242&65.9&69.2&67.6&67.7&66.0&63.8&65.1&64.7&62.4&65.9 \\
+27(med)&109324 / 49&11.0&9.6&8.7&9.2&10.0&19.4&9.4&7.9&10.7&10.6 \\
+5(med)&109949 / 267&46.3&47.4&46.9&45.4&44.0&42.9&43.7&44.3&40.9&45.7 \\
+6(med)&133395 / 73&37.2&38.9&38.7&36.8&37.5&27.5&38.0&37.2&31.3&35.9 \\
+26(med)&134828 / 428&31.8&30.8&31.8&31.2&31.9&32.6&32.2&30.5&29.6&31.2 \\
+15(bank)&136944 / 674&46.5&51.5&47.9&48.0&46.6&46.0&45.8&45.7&42.9&46.0 \\
+4(law)&137451 / 1016&77.1&85.3&83.5&83.3&75.8&46.1&74.2&73.3&63.2&75.9 \\
+2(med)&182594 / 85&70.6&75.8&71.7&69.4&68.2&67.3&67.3&68.6&65.6&68.2 \\
+20(med)&182990 / 71&47.4&47.2&46.8&47.2&48.4&47.5&48.8&47.3&47.1&46.8 \\
+21(IT)&222802 / 868&38.7&38.8&39.0&37.2&37.5&35.9&36.9&37.1&33.4&37.0 \\
+10(med)&225347 / 115&40.0&42.6&40.0&38.2&39.9&35.8&39.5&39.1&36.3&40.7 \\
+18(med)&245000 / 106&57.7&60.3&58.7&58.6&58.4&56.3&57.3&56.1&54.9&55.9 \\
+9(med)&301639 / 145&37.2&37.3&36.5&36.1&36.4&37.7&36.4&35.2&34.2&37.0 \\
+3(law)&323474 / 680&50.1&52.0&50.8&50.1&49.1&48.3&49.0&48.2&44.4&49.1 \\
+14(med)&333957 / 146&31.6&31.4&31.9&33.0&32.5&34.1&31.4&32.1&30.5&31.8 \\
+12(med)&356377 / 148&36.3&36.6&35.9&35.9&35.8&37.0&36.4&35.4&34.2&36.3 \\ \hline
+  \end{tabular}
+  \caption{Detailed results of the experiments with automatic domains}
+  \label{tab:automatic_domains}
+\end{table*}
+
+\begin{table*}[t]
+\centering
+  \begin{tabular}{|p{2.3cm}|*{11}{c|}} \hline
+   Model/ & \multirow{2}{*}{Averaged size} &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
+   Group cluster & & \system{-Nat} & \system{Full} & \system{Res} &\system{Res} & \system{-Feat}& \system{-Tag}& & & & \\ \hline
+10 small&49.5&60.2&60.7&62.6&62.7&63.6&47.3&62.1&61.3&58.8&62.5\\
+10 mid&213.7&40.6&42.6&41.9&41.2&40.9&38.3&40.2&39.5&36.9&40.5\\
+10 large&338&48.7&50.7&49.5&48.9&48.2&44.6&47.7&47.2&44.4&47.9\\
+Avg& &49.8&51.3&51.3&50.9&50.9&43.4&50.0&49.4&46.7&50.29 \\ \hline
+
+  \end{tabular}
+   \caption{Average over group of 10 large, medium, small clusters}
+  \label{tab:avg_automatic_domains}
+\end{table*}
+
 \end{document}

### 2020-08-31 2020/tacl2020/multidomain.tex
table 4

--- 2020-08-28 2020/tacl2020/multidomain.tex
@@
   \centering
-  \begin{tabular}{|l|rrrrrrr|} %*{4}{|r|}}
+  \begin{tabular}{|l|ccccccc|} %*{4}{|r|}}
     \cline{2-8} 
@@
 \begin{table}\centering
-  \begin{tabular}{|l*{5}{|r}|} \hline
-    & \domain{law} & \domain{bank} & \domain{talk} & \domain{IT} & \domain{rel} \\ \hline
+  \begin{tabular}{|l*{5}{|r}|} 
+  \cline{2-6}
+  \multicolumn{1}{c|}{} & \domain{law} & \domain{bank} & \domain{talk} & \domain{IT} & \domain{rel} \\ \hline
     \domain{med} &1.93 &1.97 &1.9 &1.93 &1.97 \\
@@
     \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -0.4 & -0.4 & -0.5 & -0.4 & \SW{-13.4} & \SW{-1.7}\\% [35.9]
-    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & \SW{-14.2} &\SW{-1.8}\\ % [34.9] 
-    \system{LDR}           & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-12} & \SW{-1.4}\\ % [37.0]
+    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & \textcolor{blue}{+}0.3 & +0.1 & \SW{-14.2} &\SW{-1.8}\\ % [34.9] 
+    \system{LDR}           & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-12\textcolor{blue}{.0}} & \SW{-1.4}\\ % [37.0]
     \system{TTM} (*)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & 0.0 & -0.1\\
@@
 Underline denotes significant loss when domains are changed wrt.\ the baseline situation; bold for a significant improvement over \system{FT-Full}; (*) tags systems ignoring test domains.
-    \fyDone{Check wrong and significance of diff for ``old''}\fyDone{How about wrong ?}
+    \fyDone{Check wrong and significance of diff for ``old''}\fyDone{How about wrong ?}\jcTodo{FTres results are odd... split LAW 0.5}
   }

### 2020-09-02 2020/tacl2020/rebuttal.tex
more changes in final sections

@@
 This raises the issues of robustness wrt to adversarial test distribution and even new domains, two issues that are somewhat irrelevant for DA.
-
+\\
 In our experiments, fine-tuned systems share no parameter and show large drops in performance when translating data from the wrong domain (even though this could be somewhat mitigated through regularization, as the reviewer points out).
@@
 Arguably, one could see here a continuum between no parameter sharing and full parameter sharing, with \system{MDL Res} and \system{FT Block} corresponding to some intermediay situation.
-
-We also agree that the results for the ``new domain'' experiment should not be over-blown and we have rephrased our conclusions somehow.
-Yet, this experiments highlights the fact that increasing the heterogeneity of the data pool could be harmful
 }

### 2020-09-02 2020/wada/adaptive_domain_adaptation.tex

@@
 
- A ``gated'' version uses $h' = h + a_k(h) \odot{} z_k(h)$\footnote{Or $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure~\ref{fig:architecture}} \fyDone{$a_k(h)=$ somme tous les composants résiduels} where $z_k$ is a function of $h$ taking values in $[0,1]$.
+ A ``gated'' version uses $h' = h + a_k(h) \odot{} z_k(h)$\footnote{Or $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure~\ref{fig:hrl-architecture}} \fyDone{$a_k(h)=$ somme tous les composants résiduels} where $z_k$ is a function of $h$ taking values in $[0,1]$.
 More precisely, $h$ and $h'$ are sequences of context vectors and the combination is performed element-wise, yielding:\fyDone{pourquoi pas gating: $(1-a_k(h(w))*hh(w)$ ?
@@
   \end{align}
+  Where x is source sentence, y is target sentence, $x=\lbrace w^{src}_0,..,w^{src}_{T_0}\rbrace$, $y=\lbrace w^{tgt}_0,..,w^{tgt}_{T_1} \rbrace$
 
@@
 
-Furthermore $ \varepsilon^k_{insep} < \varepsilon_{insep} $ which is the minimum of the error expectation over the hypothesis space of the classifier.
-With a trivial prediction $[\frac{1}{K},..,\frac{1}{K}]$ we have an error expectation equal to $0.5$.
-Therefore, $\varepsilon_{insep} \leq 0.5$, i.e, $\varepsilon_{insep}(k) < 0.5$.\fyTodo{Je ne comprends pas.} In the worst case where domains in the mixture are very close, we are still able to reduce the risk of predicting out-of-domain examples by a factor of 2.
-\fyTodo{I sort of buy the equations but then so what ? We should may be compare with non-adaptive ?}
-
 \end{document}

### 2020-09-02 2020/tacl2020/rebuttal.tex

@@
 
-The authors wish thank the reviewers and editor for their comments and suggestions, and have tried to address their questions and concerned in the revised version.
+\textcolor{blue}{The authors wish thank to the reviewers and editor for their comments and suggestions.
+We have tried to address their questions and concerns in the revised version.
 Details are given below, with pointers to the main changes in the text.
-Typos and presentations errors have also been corrected in the new submission.
+Typos and presentations errors have also been corrected in the new submission.}
 
@@
 \\
-  \textcolor{blue}{discusion XXX.}
+  \textcolor{blue}{The differences between these two systems is explained in our response to reviewer A below, and has lead us to substantially revise section 4.3 where we describe our multi-domain systems.
+As requested,  \system{FT-Block} has now been included in all the comparative evaluations.
+}
 \item Carefully specify the number of parameters for all models, including the proportions that are domain-specific and domain-independent. \\
-  \textcolor{blue}{New table ...}
+  \textcolor{blue}{These numbers now appear in the revised version of Table 3.}
 \item Provide all the experimental details requested, and in particular give enough information to allow the corpora to be reproduced, or (ideally) make them publicly available.\\
-  \textcolor{blue}{Appendix ....}
+  \textcolor{blue}{We have added an appendix devoted to the system details, giving all the necessary information regarding meta-parameter settings and training procedures.
+The exact split of train / dev / test data used in this study will be distributed together with the final version of the paper.}
 \end{itemize}
 
-You are allowed one to two extra pages of content to accommodate these
-revisions. To submit your revised version, follow the instructions in the
-"Revision and Resubmission Policy for TACL Submissions" section of the
-Author Guidelines at
-\url{https://transacl.org/ojs/index.php/tacl/about/submissions\#authorGuidelines}.
+You are allowed one to two extra pages of content to accommodate these revisions.
+To submit your revised version, follow the instructions in the "Revision and Resubmission Policy for TACL Submissions" section of the
+Author Guidelines at \url{https://transacl.org/ojs/index.php/tacl/about/submissions\#authorGuidelines}.
 
-Thank you for submitting to TACL, and I look forward to your revised
-version! \\[2\parskip]
+\textcolor{blue}{We have joint to this mail two versions of the revised paper: one where the main changes appear in red, and the appendix is placed before the references so has to highlight the differences with the first version; one version that looks like a regular submission.
+We hope that this will make it easier for reviewers to view the differences with the first version of the paper.}
+
+Thank you for submitting to TACL, and I look forward to your revised version! \\[2\parskip]
 
@@
 \textcolor{red}{%
-  We have tried to address these commentaries in the revised version of the paper.
+  We have tried to address these comments in the revised version of the paper.
 Notably, we have clarified the distinction between \system{FT-block} and \system{MDL Res} (see details below).
-We have also tried to clarify the distinction between multi-domain MT and fine-tuning; while the former is designed with an a priori knowledge that the test data is sampled from a single domain, the latter assumes that the test data will be sampled from multiple domains, using domain probabilities that are not known a priori.
-This raises the issues of robustness wrt to adversarial test distribution and even new domains, two issues that are somewhat irrelevant for DA.
-\\
+We have also tried to clarify the distinction between multi-domain MT and fine-tuning; while the former is designed given the (mono-domain) test data distribution, the latter assumes that the test data will be sampled from multiple domains, using domain probabilities that are not known in advance.
+This raises the issues of robustness with respect to adversarial test distribution and even new domains, two issues that are somewhat irrelevant for DA.
+\newline
 In our experiments, fine-tuned systems share no parameter and show large drops in performance when translating data from the wrong domain (even though this could be somewhat mitigated through regularization, as the reviewer points out).
@@
     We understand that our naming schemes was insisting on training procedures, rather than on architectures, thus possibly introducing some confusions.
-We have accordingly changed the name of \system{FT-Block} to \system{FT Res}, clarified the description of the two systems (see lines YYY\fyTodo{fix this}), and reported \system{FT Res}\fyTodo{or new name} as another multi-domain system in all tables.\fyTodo{Make sure it is the case}.}
+We have accordingly changed the name of \system{FT-Block} to \system{FT Res}, clarified the description of the two systems (see lines YYY\fyTodo{fix this}), and reported \system{FT-Res}\fyDone{or new name} as another multi-domain system in all tables.\fyTodo{Make sure it is the case}.}
   %}
@@
   \textcolor{blue}{We agree with the reviewer that the analogy is misleading.
-As this is not central to our discussion, we have repharsed and dropped the comparison.
+As this is not central to our discussion, we have rephrased and dropped the comparison.
 See changes in line XXX}\fyTodo{Changes line in.}
@@
   \textcolor{blue}{We agree that carefully analyzing the effects of using corpora of varying size is important, as this corresponds to practical training scenario.
-From our results in Table 6 (and Table XXX in appendix), we see that some techniques can work well with small in-domain corpora.
+From our results in Table 7 (and Table 8 in Appendix B), we see that some techniques can work well with small in-domain corpora.
 However, this analysis must also take in account the similarity between domains, as these improvements will be stronger if similar or related data are also present in the training mixture -- a situation that we have artificially created in our split conditions.
@@
 
-\item[* L].
+\item[*] L.
 687: Please describe how you predict the domain tag, and please share the tag prediction error rate on each domain test set.
@@
 \\
-\textcolor{blue}{We train a language model in source language for each domain using respective source training corpus.
-To predict domain tag of a sentence, we compute probability of the sentence according to each pretrained language model and choose the domain corresponding to the highest probability as domain tag of sentence.
+\textcolor{blue}{We train a language model in source language for each domain using the corresponding training corpus.
+To predict the domain tag of a sentence, we compute the probability of the sentence according to each pretrained language model and choose the domain having the highest probability as the domain tag for that sentence.
 Note that this classifier is only used for the \domain{news} test data.
@@
 \\
-\textcolor{red}{For each domain, we create train/dev/test sets by randomly each corpus.
+\textcolor{red}{For each domain, we create train/dev/test sets by randomly splitting each corpus.
 We maintain the size of validation sets and of test sets equal to 1,000 lines for every domain.
@@
 We stop training if either the training iteration reach the maximal number of iterations or the score on the validation set does not increase in 3 consecutive evaluations.
-We average 5 checkpoints up to the best checkpoint to get the final model.}\fyTodo{up to the best ?}
+We average 5 checkpoints preceding the best checkpoint to get the final model.
+All these details are in Appendix B}\fyDone{up to the best ?}
 \\

### 2020-09-02 2020/tacl2020/multidomain.tex

@@
 \revision{Again, inference does not use domain tags.}\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly improved, results.}\fyDone{Check this}
-\item \revision{two multi-domain versions of the approach of \newcite{Bapna19simple}, denoted \system{FT-Res} and \system{MDL-Res}, where a domain-specific adaptation module is added to all the Transformer layers; within each layer, residual connections enable to by-pass this module.
+\item \revision{two multi-domain versions of the approach of \newcite{Bapna19simple}, denoted \system{FT-Res} and \system{MDL-Res}, where a domain-specific adaptation module is added to all the Transformer layers; within each layer, residual connections enable to short-cut this adapter.
 The former variant corresponds to the original proposal of \citet{Bapna19simple} \revision{(see also \cite{Sharaf20metalearning})}.
@@
   \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+  \begin{tabular}{|p{4cm}|*{8}{r|}} \hline
 %     &&&&&& \\
@@
      \hline \hline
-    \system{Mixed-Nat-RNN}  & 36.8 & 53.8 & 47.2 & 30.0 & 35.7 & 60.2  & 39.2  & 44.0 \\
+    \system{Mixed-Nat-RNN} \hfill{\footnotesize[51m]}  & 36.8 & 53.8 & 47.2 & 30.0 & 35.7 & 60.2  & 39.2  & 44.0 \\
     \hline
-    \system{WDCMT}       & 36.0 & 53.3 & \SB{48.8} & 31.1 & \SB{38.8} & \SW{58.5} & 39.0 & 44.4 \\ % & 20.4 \\
+    \system{WDCMT}  \hfill{\footnotesize[73m]} & 36.0 & 53.3 & \SB{48.8} & 31.1 & \SB{38.8} & \SW{58.5} & 39.0 & 44.4 \\ % & 20.4 \\
     \hline
   \end{tabular}
-  \caption{Translation performance of various MDMT systems \revision{based on the same Transformer (top) or RNN (bottom) architecture.
-The former contains 65m parameters, the latter has ???.
-For each system, we report the number of domain specific parameters,}  BLEU scores for each domain, as well as domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages.
-\revision{In the former case, we take the domain proportions from Table~\ref{tab:Corpora}}.
+  \caption{Translation performance of MDMT systems \revision{based on the same Transformer (top) or RNN (bottom) architecture.
+The former contains 65m parameters, the latter has 51m.
+For each system, we report the number of additional domain specific parameters,}  BLEU scores for each domain, domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages.
+\revision{For weighted-averages, we take the domain proportions from Table~\ref{tab:Corpora}}.
 Boldface denotes significant gains with respect to \system{Mix-Nat} (or \system{Mix-Nat-RNN}, for WDCMT), underline denotes significant losses.}
@@
 All MDMTs (except \system{DM} and \system{ADM}) slightly improve over \system{Mixed-Nat}(for most domains), but these gains are rarely significant.
-\revision{Among systems using an extra domain feature, \system{DC-Tag} has a small edge over \system{DC-Feat} and also requires less paramaters; it also outperforms \system{TTM}, which however uses predicted rather than gold domain tags.
+\revision{Among systems using an extra domain feature, \system{DC-Tag} has a small edge over \system{DC-Feat} and also requires less parameters; it also outperforms \system{TTM}, which however uses predicted rather than gold domain tags.
 \system{TTM} is also the best choice among the systems that to not use domain tags in inference.} \revision{The best contenders overall are  \system{FT-Res} and \system{MDL-Res}, which significantly improve over \system{Mixed-Nat} for a majority of domains, and are the only ones to clearly fulfill [P1-LAB];} \system{WDCMT} also improves on three domains, but regresses on one.
-The use of a dedicated adaptation component, seems here better than feature-based strategies, \revision{but yields a large increase of the number of parameters.} The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}).
+The use of a dedicated adaptation module thus seems better than feature-based strategies, \revision{but yields a large increase of the number of parameters.} The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}).
 
@@
     \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -0.4 & -0.4 & -0.5 & -0.4 & \SW{-13.4} & \SW{-1.7}\\% [35.9]
-    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & \textcolor{blue}{+}0.3 & +0.1 & \SW{-14.2} &\SW{-1.8}\\ % [34.9] 
-    \system{LDR}           & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-12\textcolor{blue}{.0}} & \SW{-1.4}\\ % [37.0]
+    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & +0.3 & +0.1 & \SW{-14.2} &\SW{-1.8}\\ % [34.9] 
+    \system{LDR}           & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-12.0} & \SW{-1.4}\\ % [37.0]
     \system{TTM} (*)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & 0.0 & -0.1\\
@@
 Even in this very favourable case however, very few MDMT systems are able to significantly outperform \system{FT-Full} and this is only observed for the smaller part of the \domain{med} domain.
-The merge condition is hardly different, with large losses for \system{FT-full}, and small variations for all systems.
+The merge condition is hardly different, with again large losses for \system{FT-full} \revision{and \system{FT-Res}}, and small variations for all systems.
 We even observe some rare improvements with respect to the situation where we use actual domains.\fyDone{More comments ?}
@@
 \subsection{Automatic domains \label{ssec:autodomains}}
-In this section, we experiment with automatic domains, obtained by clustering sentences into $k=30$ classes using the k-means algorithm based on generic sentence representations obtained via mean pooling as in Section~\ref{ssec:corpora}.
+In this section, we experiment with automatic domains, obtained by clustering sentences into $k=30$ classes using the k-means algorithm based on generic sentence representations obtained via mean pooling (cf.\ Section~\ref{ssec:corpora}).
 This allows us to evaluate requirement [P7-scale], training and testing our systems as if these domains were fully separated.
-Many of these clusters are mere splits of the large \domain{med}, while a fewer number of classes are mixtures of two (or more) existing domains \revision{(details are in Appendix B)}.
+Many of these clusters are mere splits of the large \domain{med}, while a fewer number of classes are mixtures of two (or more) existing domains \revision{(Full details are in Appendix B)}.
 We are thus in a position to reiterate, at a larger scale, the measurements of Section~\ref{ssec:redomains} and test whether multi-domain systems can effectively take advantage from the cross-domain similarities and to eventually perform better than fine-tuning.
-\revision{The results in Table~\ref{tab:avg_automatic_domains} suggest that MDMT actually can surpass fine-tuning for the smaller clusters, especially for small subsets of the medical domain; for the large clusters, this is no longer true.}
+\revision{The results in Table~\ref{tab:avg_automatic_domains} also suggest that MDMT can surpass fine-tuning for the smaller clusters; for the large clusters, this is no longer true.
+The complete table (in Appendix B) shows that this effect is more visible for small subsets of the medical domain.}
 
+\begin{table*}[t]
+\centering \footnotesize
+\revision{
+  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
+   Model/ & Train &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
+   Clusters & size & \system{Nat} & \system{Full} & \system{Res} &\system{Res} & \system{Feat}& \system{Tag}& & & & \\ \hline
+10 small&29.3k&60.2&60.7&62.6&62.7&\bf 63.6&47.3&62.1&61.3&58.8&62.5\\
+10 mid&104.7k&40.6&\bf 42.6&41.9&41.2&40.9&38.3&40.2&39.5&36.9&40.5\\
+10 large&251.1k&48.7&\bf 50.7&49.5&48.9&48.2&44.6&47.7&47.2&44.4&47.9\\
+Avg& &128.4k&\bf 51.3&\bf 51.3&50.9&50.9&43.4&50.0&49.4&46.7&50.29 \\ \hline
+  \end{tabular}
+   \caption{Average BLEU scores for 10 smaller, medium, and larger clusters.
+Best score for each group is in boldface.
+For the small clusters, full-fine tuning is outperformed by several MDMT systems - see details in Appendix~B.}
+   \label{tab:avg_automatic_domains}}
+\end{table*}
 Finally, Table~\ref{tab:subdomains} reports the effect of using automatic domain for each of the 6 test sets: each sentence was first assigned to an automatic class, translated with the corresponding multi-domain system with 30 classes; aggregate numbers were then computed, and contrasted with the 6-domain scenario.
@@
 \revisiondel{In this scenario, architecture-based strategies aimed at parameter sharing and/or representation normalization obtain the best results.}\fyDone{Mitigate this}
-As expected however, MDMTs are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend, dynamically accommodate additional domains, this being especially easy for feature-based methodologies.
+As expected however, MDMTs are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend, dynamically accommodate additional domains, this being especially easy for feature-based approaches.
 Our experiments finally suggest that all methods show decreasing performance when the number of domains or the diversity of the domain mixture increases.
@@
 \subsection*{A - Description of multi-domain systems \label{ssec:implementation-details}}
+\begin{table*}[t]
+  \centering
+  \footnotesize
+  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
+   Model  & size&\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
+   Cluster & train / test & \system{Nat} & \system{Full} & \system{Res} &\system{Res} & \system{Feat}& \system{Tag}& & & & \\ \hline
+24 \hfill [med]&8.1k / 3 &90.4&90.4&90.4&90.4&100.0&65.6&100.0&90.4&100.0&100.0 \\
+13 \hfill [-]&17.3k / 52&67.6&75.4&74.3&74.3&75.0&54.7&74.7&75.9&65.9&76.9 \\
+28 \hfill [-]&25.6k / 54&71.6&68.7&68.1&70.2&71.0&42.5&72.0&71.3&65.6&72.6 \\
+19 \hfill [IT]&27.2k / 88&58.5&63.0&60.9&63.9&63.7&57.2&59.4&61.1&60.5&60.3 \\
+0   \hfill [-]&27.4k / 72&43.9&33.3&45.4&45.4&49.9&15.4&46.8&49.2&46.6&47.8 \\
+22 \hfill [-]&27.5k / 103&91.5&93.7&93.4&93.9&92.5&72.8&92.3&93.2&91.4&93.4 \\
+25 \hfill [-]&28.2k / 56&57.0&44.8&48.2&49.1&54.6&47.2&49.8&54.2&45.1&52.4 \\
+16 \hfill [med]&30.4k / 18&57.2&70.4&77.4&73.5&61.8&54.2&58.4&58.1&52.5&58.3 \\
+23 \hfill [med]&47.0k / 23&24.5&27.2&26.5&28.5&30.5&27.3&32.0&24.4&29.0&29.8 \\
+17 \hfill [med]&54.4k / 26&39.9&40.3&41.6&38.0&37.1&36.6&35.2&35.4&31.3&33.7 \\
+8  \hfill [IT]&61.4k / 214&46.9&53.1&55.8&53.6&48.9&45.1&48.8&50.9&43.0&46.7 \\
+1 \hfill [-]&68.1k / 122&47.2&47.5&48.7&45.1&46.8&39.1&45.4&44.2&40.7&44.9 \\
+7 \hfill [med]&91.5k / 30&41.3&35.5&41.4&39.9&41.4&36.5&37.3&37.1&40.7&41.8 \\
+11 \hfill [med]&93.0k / 38&31.6&42.6&31.8&35.4&36.0&29.6&36.7&32.7&26.5&36.6 \\
+29 \hfill [law]&109.2k / 242&65.9&69.2&67.6&67.7&66.0&63.8&65.1&64.7&62.4&65.9 \\
+27 \hfill [med]&109.3k / 49&11.0&9.6&8.7&9.2&10.0&19.4&9.4&7.9&10.7&10.6 \\
+5 \hfill [-]&109.9k / 267&46.3&47.4&46.9&45.4&44.0&42.9&43.7&44.3&40.9&45.7 \\
+6 \hfill [med]&133.4k / 73&37.2&38.9&38.7&36.8&37.5&27.5&38.0&37.2&31.3&35.9 \\
+26 \hfill [-]&134.8k / 428&31.8&30.8&31.8&31.2&31.9&32.6&32.2&30.5&29.6&31.2 \\
+15 \hfill [bank]&136.9k / 674&46.5&51.5&47.9&48.0&46.6&46.0&45.8&45.7&42.9&46.0 \\
+4 \hfill [rel]&137.4k / 1016&77.1&85.3&83.5&83.3&75.8&46.1&74.2&73.3&63.2&75.9 \\
+2 \hfill [med]&182.6k / 85&70.6&75.8&71.7&69.4&68.2&67.3&67.3&68.6&65.6&68.2 \\
+20 \hfill [med]&183.0k / 71&47.4&47.2&46.8&47.2&48.4&47.5&48.8&47.3&47.1&46.8 \\
+21 \hfill [-]&222.8k / 868&38.7&38.8&39.0&37.2&37.5&35.9&36.9&37.1&33.4&37.0 \\
+10 \hfill [med]&225.4k / 115&40.0&42.6&40.0&38.2&39.9&35.8&39.5&39.1&36.3&40.7 \\
+18 \hfill [med]&245.0k / 106&57.7&60.3&58.7&58.6&58.4&56.3&57.3&56.1&54.9&55.9 \\
+9 \hfill [med]&301.6k / 145&37.2&37.3&36.5&36.1&36.4&37.7&36.4&35.2&34.2&37.0 \\
+3 \hfill [law]&323.5k / 680&50.1&52.0&50.8&50.1&49.1&48.3&49.0&48.2&44.4&49.1 \\
+14 \hfill [med]&334.0 / 146&31.6&31.4&31.9&33.0&32.5&34.1&31.4&32.1&30.5&31.8 \\
+12 \hfill [med]&356.4k / 148&36.3&36.6&35.9&35.9&35.8&37.0&36.4&35.4&34.2&36.3 \\ \hline
+  \end{tabular}
+  \caption{Complete results for the experiments with automatic domains.
+For each cluster, we report: the majority domain when one domain accounts for more than 75\% of the class; training and test sizes; and BLEU scores obtained with the various systems used in this study.
+Most test sets are too small to report significance tests.}
+  \label{tab:automatic_domains}
+\end{table*}
+
 \revision{%
-The systems used in this study are set up as follows:
+We use the following setups for MDMT systems.
 \begin{itemize}
-\item \system{Mixed-NAT}, \system{FT-full}, \system{TTM}, \system{DC-Tag} use the a medium Transformer model of \cite{Vaswani17attention} with the following settings: embeddings size and hidden layers size are set to~512.
-Multi-head attention comprises 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells.
+\item \system{Mixed-Nat}, \system{FT-full}, \system{TTM}, \system{DC-Tag} use the a medium Transformer model of \cite{Vaswani17attention} with the following settings: embeddings size and hidden layers size are set to~$512$.
+Multi-head attention comprises 8 heads in each of the 6 layers; the inner feedforward layer contains~$2048$ cells.
 % The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
 % The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training use a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.
-\item \system{FT-Res} and \system{MDL-res} use the same medium Transformer model, and add residual layers with a bottleneck dimension of size 1024.
+Training use a batch size of~$12,288$ tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.
+\item \system{FT-Res} and \system{MDL-res} use the same medium Transformer and add residual layers with a bottleneck dimension of size $1024$.
   % \cite{Bapna19simple} showed that if the bottleneck dimension is large enough, the gap between \system{FT-block} and \system{FT-full} will be small.\fyTodo{This comment is confusing}.
@@
 The two first layers of the classifier use the $\operatorname{ReLU}()$ as activation function, the last layer uses $\tanh()$ as activation function.
-\item \system{DC-Feat} uses medium Transformer model and domain embeddings of size 4.
+\item \system{DC-Feat} uses medium Transformer model and domain embeddings of size~$4$.
 Given a sentence of domain~$i$ in a training batch, the embedding of domain $i$ is concatenated to the embedding of each token in the sentence.
 \item \system{LDR} uses medium Transformer model and for each token we introduce a \system{LDR} feature of size $4 \times domain\_num$.
-Given a sentence of domain $i\in[1,..,K]$ in the training batch, for each token of the sentence, the \system{LDR} units of the indexes outside of the range $[4(i-1),..,4i-1]$ are masked to 0, the masked LDR feature will be concatenated to the embedding of the token.
+Given a sentence of domain $i\in[1,..,K]$ in the training batch, for each token of the sentence, the \system{LDR} units of the indexes outside of the range $[4(i-1),..,4i-1]$ are masked to $0$, the masked LDR feature will be concatenated to the embedding of the token.
 Details are in \cite{Pham19generic}.
-\item \system{Mixed-NAT-RNN} uses one layer LSTM bidirectional as encoder and one layer LSTM as decoder. The size of each hidden layer is 1024, the size of word embeddings is 512. 
-\item \system{WDCNMT} uses one layer GRU bidirectional as encoder and one layer GRU-conditional as decoder. The size of each hidden layer is 1024, the size of word embedding is 512.
+\item \system{Mixed-Nat-RNN} uses one bidirectional LSTM layer in the encoder and one LSTM layer in the decoder. The size of hidden layers is $1024$, the size of word embeddings is $512$. 
+\item \system{WDCNMT} uses one bidirectional GRU layer in the encoder and one GRU-conditional layer in the decoder. The size of hidden layers is $1024$, the size of word embedding is $512$.
 \end{itemize}
@@
 }
+\subsection*{B - Experiments with automatic domains \label{ssec:full-automatic}}
+\revision{
+  This experiment aims to simulate with automatic domains a scenario where the number of ``domains'' is large and where some ``domains'' are close and can effectively share information.
+Full results in Table~\ref{tab:automatic_domains}.
+Cluster size vary from approximately 8k sentences (cluster~24) up to more than 350k sentences.
+More than 2/3 of these clusters comprise mostly texts from one single domain, as for cluster 12 which is predominantly \domain{med}, the remaining clusters typically mix 2 domains.
+Fine-tuning with small domains is often outperformed by other MDMT technique, an issue that a better regularization strategy might mitigate.
+Domain-control (\system{dc feat}) is very effective for small domains, but again less so in larger data conditions.
+Among the MD models, approaches using residual adapters have the best average performance.}
 
-\begin{table*}[t]
-  \centering
-  \footnotesize
-  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
-   Model  & size&\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
-   Cluster & train / test & \system{-Nat} & \system{Full} & \system{Res} &\system{Res} & \system{-Feat}& \system{-Tag}& & & & \\ \hline
-24 \hfill [med]&8.1k / 3 &90.4&90.4&90.4&90.4&100.0&65.6&100.0&90.4&100.0&100.0 \\
-13(law)&17.3k / 52&67.6&75.4&74.3&74.3&75.0&54.7&74.7&75.9&65.9&76.9 \\
-28(med)&25.6k / 54&71.6&68.7&68.1&70.2&71.0&42.5&72.0&71.3&65.6&72.6 \\
-19(IT)&27.2k / 88&58.5&63.0&60.9&63.9&63.7&57.2&59.4&61.1&60.5&60.3 \\
-0(IT)&27.4k / 72&43.9&33.3&45.4&45.4&49.9&15.4&46.8&49.2&46.6&47.8 \\
-22(bank)&27534 / 103&91.5&93.7&93.4&93.9&92.5&72.8&92.3&93.2&91.4&93.4 \\
-25(med)&28237 / 56&57.0&44.8&48.2&49.1&54.6&47.2&49.8&54.2&45.1&52.4 \\
-16(med)&30443 / 18&57.2&70.4&77.4&73.5&61.8&54.2&58.4&58.1&52.5&58.3 \\
-23(med)&47065 / 23&24.5&27.2&26.5&28.5&30.5&27.3&32.0&24.4&29.0&29.8 \\
-17(med)&54378 / 26&39.9&40.3&41.6&38.0&37.1&36.6&35.2&35.4&31.3&33.7 \\
-8(IT)&61353 / 214&46.9&53.1&55.8&53.6&48.9&45.1&48.8&50.9&43.0&46.7 \\
-1(med)&68108 / 122&47.2&47.5&48.7&45.1&46.8&39.1&45.4&44.2&40.7&44.9 \\
-7(med)&91464 / 30&41.3&35.5&41.4&39.9&41.4&36.5&37.3&37.1&40.7&41.8 \\
-11(med)&92982 / 38&31.6&42.6&31.8&35.4&36.0&29.6&36.7&32.7&26.5&36.6 \\
-29(law)&109260 / 242&65.9&69.2&67.6&67.7&66.0&63.8&65.1&64.7&62.4&65.9 \\
-27(med)&109324 / 49&11.0&9.6&8.7&9.2&10.0&19.4&9.4&7.9&10.7&10.6 \\
-5(med)&109949 / 267&46.3&47.4&46.9&45.4&44.0&42.9&43.7&44.3&40.9&45.7 \\
-6(med)&133395 / 73&37.2&38.9&38.7&36.8&37.5&27.5&38.0&37.2&31.3&35.9 \\
-26(med)&134828 / 428&31.8&30.8&31.8&31.2&31.9&32.6&32.2&30.5&29.6&31.2 \\
-15(bank)&136944 / 674&46.5&51.5&47.9&48.0&46.6&46.0&45.8&45.7&42.9&46.0 \\
-4(law)&137451 / 1016&77.1&85.3&83.5&83.3&75.8&46.1&74.2&73.3&63.2&75.9 \\
-2(med)&182594 / 85&70.6&75.8&71.7&69.4&68.2&67.3&67.3&68.6&65.6&68.2 \\
-20(med)&182990 / 71&47.4&47.2&46.8&47.2&48.4&47.5&48.8&47.3&47.1&46.8 \\
-21(IT)&222802 / 868&38.7&38.8&39.0&37.2&37.5&35.9&36.9&37.1&33.4&37.0 \\
-10(med)&225347 / 115&40.0&42.6&40.0&38.2&39.9&35.8&39.5&39.1&36.3&40.7 \\
-18(med)&245000 / 106&57.7&60.3&58.7&58.6&58.4&56.3&57.3&56.1&54.9&55.9 \\
-9(med)&301639 / 145&37.2&37.3&36.5&36.1&36.4&37.7&36.4&35.2&34.2&37.0 \\
-3(law)&323474 / 680&50.1&52.0&50.8&50.1&49.1&48.3&49.0&48.2&44.4&49.1 \\
-14(med)&333957 / 146&31.6&31.4&31.9&33.0&32.5&34.1&31.4&32.1&30.5&31.8 \\
-12(med)&356377 / 148&36.3&36.6&35.9&35.9&35.8&37.0&36.4&35.4&34.2&36.3 \\ \hline
-  \end{tabular}
-  \caption{Full results for the experiments with automatic domains.
-For each cluster, we report the majority domain in the case where one domain accounts for more than XXX\% of the data, the training and test size, and the BLEU scores obtained with the various systems used in this study.
-Boldface denotes the best system in each line; test sets are two small to report significativity tests.}
-  \label{tab:automatic_domains}
-\end{table*}
+\clearpage{}
 
-\begin{table*}[t]
-\centering \footnotesize
-  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
-   Model/ & Test &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
-   Clusters & size & \system{Nat} & \system{Full} & \system{Res} &\system{Res} & \system{Feat}& \system{Tag}& & & & \\ \hline
-10 small&49.5&60.2&60.7&62.6&62.7&\bf 63.6&47.3&62.1&61.3&58.8&62.5\\
-10 mid&213.7&40.6&\bf 42.6&41.9&41.2&40.9&38.3&40.2&39.5&36.9&40.5\\
-10 large&338&48.7&\bf 50.7&49.5&48.9&48.2&44.6&47.7&47.2&44.4&47.9\\
-Avg& &49.8&\bf 51.3&\bf 51.3&50.9&50.9&43.4&50.0&49.4&46.7&50.29 \\ \hline
-
-  \end{tabular}
-   \caption{Average BLEU scores for 10 smaller, medium, and larger clusters.
-Best score for each group is in boldface.
-For the small clusters, full-fine tuning is outperformed by several MDMT system - see detail in Table~\ref{tab:automatic_domains}.}
-  \label{tab:avg_automatic_domains}
-\end{table*}
-
-\flushbottom{}
-
-\newpage\newpage
+% \newpage\newpage \newpage
 

### 2020-09-02 2020/tacl2020/rebuttal.tex
small and final fixes

@@
 \\
-  \textcolor{blue}{The differences between these two systems is explained in our response to reviewer A below, and has lead us to substantially revise section 4.3 where we describe our multi-domain systems.
-As requested,  \system{FT-Block} has now been included in all the comparative evaluations.
+  \textcolor{blue}{The difference between these two systems is explained in our response to reviewer~A below and has lead us to substantially revise Section~4.3 where we describe our multi-domain systems.
+As requested, \system{FT-Block} (now \system{FT-Res}) has now been included in all the comparative evaluations.
 }
 \item Carefully specify the number of parameters for all models, including the proportions that are domain-specific and domain-independent. \\
-  \textcolor{blue}{These numbers now appear in the revised version of Table 3.}
+  \textcolor{blue}{These numbers now appear in the revised version of Table~3.}
 \item Provide all the experimental details requested, and in particular give enough information to allow the corpora to be reproduced, or (ideally) make them publicly available.\\
@@
 \newline
-In our experiments, fine-tuned systems share no parameter and show large drops in performance when translating data from the wrong domain (even though this could be somewhat mitigated through regularization, as the reviewer points out).
+In our experiments, fine-tuned systems share no parameter and show drops in performance when translating data from the wrong domain (even though this could be somewhat mitigated through regularization, as the reviewer points out).
 By contrast, the majority of the parameters of our MDMT systems are shared for all domains.
-Arguably, one could see here a continuum between no parameter sharing and full parameter sharing, with \system{MDL Res} and \system{FT Block} corresponding to some intermediay situation.
+Arguably, one could see here a continuum between no parameter sharing and full parameter sharing, with \system{MDL-Res} and \system{FT-Block} corresponding to some intermediary situation.
+Sharing parameters is important when the number of target domains becomes larger.
 }

### 2020-09-02 2020/tacl2020/multidomain.tex
cancel local changes and revert to revision - improve table and layout, check wording
small and final fixes

@@
   \end{tabular}
-   \caption{Average BLEU scores for 10 smaller, medium, and larger clusters.
+   \caption{BLEU scores computed by merging the 10 smaller, medium, and larger cluster test sets.
 Best score for each group is in boldface.
@@
 We have spelled out the various motivations for building such systems and the associated expectations in terms of system performance.
-We have then designed a series of requirements that MDMT systems should meet, and proposed a series of test procedures.
+We have then designed a series of requirements that MDMT systems should meet, and proposed a series of associated test procedures.
 In our experiments with a representative sample of MDMTs, we have found that most requirements were hardly met for our experimental conditions.

### 2020-09-02 2020/wmt_20/emnlp2020.tex
post discussion

@@
 
-Because of the limit of our computational resources, we restrict our experiments in this section to the En$\rightarrow$Fr task. Results are in Table~\ref{tab:performance-multi}.
+Because of the limit of our computational resources, we restrict the experiments in this section to the En$\rightarrow$Fr task.
+Results are in Table~\ref{tab:performance-multi}.\fyFuture{Perform experiments for de:en}\fyFuture{Restore results where residual are removed from other paper.}
 

### 2020-09-02 2020/tacl2020/multidomain.tex
score of cluster groups

@@
    Clusters & size & \system{Nat} & \system{Full} & \system{Res} &\system{Res} & \system{Feat}& \system{Tag}& & & & \\ \hline
-10 small&29.3k&60.2&60.7&62.6&62.7&\bf 63.6&47.3&62.1&61.3&58.8&62.5\\
-10 mid&104.7k&40.6&\bf 42.6&41.9&41.2&40.9&38.3&40.2&39.5&36.9&40.5\\
-10 large&251.1k&48.7&\bf 50.7&49.5&48.9&48.2&44.6&47.7&47.2&44.4&47.9\\
+10 small&29.3k&68.3&70.0&70.7&\bf 71.2&70.6&53.1&67.3&69.8&67.0&70.2\\
+10 mid&104.7k&44.8&\bf 48.0&46.0&45.7&44.8&44.3&44.5&43.7&41.6&44.5\\
+10 large&251.1k&50.4&\bf 52.9&52.0&51.3&49.6&43.2&49.1&48.5&44.3&49.5\\
 Avg& &128.4k&\bf 51.3&\bf 51.3&50.9&50.9&43.4&50.0&49.4&46.7&50.29 \\ \hline

### 2020-09-03 2020/tacl2020/rebuttal.tex
add score table 6, add lines, minor corrections: dc feat to DC-Feat, {ADM} to \system{ADM}

@@
     We understand that our naming schemes was insisting on training procedures, rather than on architectures, thus possibly introducing some confusions.
-We have accordingly changed the name of \system{FT-Block} to \system{FT Res}, clarified the description of the two systems (see lines YYY\fyTodo{fix this}), and reported \system{FT-Res}\fyDone{or new name} as another multi-domain system in all tables.\fyTodo{Make sure it is the case}.}
+We have accordingly changed the name of \system{FT-Block} to \system{FT Res}, clarified the description of the two systems (see lines~527-541\fyTodo{fix this}), and reported \system{FT-Res}\fyDone{or new name} as another multi-domain system in all tables.\fyTodo{Make sure it is the case}.}
   %}
@@
 As this is not central to our discussion, we have rephrased and dropped the comparison.
-See changes in line XXX}\fyTodo{Changes line in.}
+See changes in line~216-220}\fyTodo{Changes line in.}
 \item 567-568: "having the domain tag on the source is better than on the target".
@@
 In our experiments, we found that TTM is pretty accurate in its prediction, making about 2\% errors in the predicted tags, which may explain the small difference with observe in Table~3.
-We have rephrased acccordingly in lines WWW.}\fyTodo{Add lines}
+We have rephrased acccordingly in lines~542-544, 550-551, 580-582.}\fyTodo{Add lines}
 \end{itemize}
@@
 To benefit from your results, we need access to the same segmentation of train/dev/test sets that I hope you will provide to the community.
-\textcolor{blue}{Supplementary details regarding the settings used in our experiments have been added in a new Appendix YYY.\fyTodo{Fix the appendix.}.
+\textcolor{blue}{Supplementary details regarding the settings used in our experiments have been added in a new Appendix A.\fyTodo{Fix the appendix.}.
 We will also distribute the detailed composition of our train / dev / test split upon acceptance of the paper.}
@@
 It will be helpful to see  the impact of the number of parameters on the MT results.
-  \textcolor{blue}{This information has been added in Table~3\fyTodo{Check number, page and line number and content}.}
+  \textcolor{blue}{This information has been added in Table~3 lines~614-617 \fyTodo{Check number, page and line number and content}.}
 \\ 
@@
 We hypothesize that training distribution is different from testing distribution, so we have need notation for two different mixtures.
-This has been clarified in the text (see line~XXX) \fyTodo{add lines}}
+This has been clarified in the text (see line~143-144) \fyTodo{add lines}}
 \\
@@
 \\
-\textcolor{blue}{This has been added in the text (see line YYY).}\fyTodo{to explicitly compare WDCNMT with Mixed-NAT-RNN. This is in the legend - change presentations ?}
+\textcolor{blue}{This has been added in the text (see line~586-587 ).}\fyTodo{to explicitly compare WDCNMT with Mixed-NAT-RNN. This is in the legend - change presentations ?}
 \\
@@
 Note that this classifier is only used for the \domain{news} test data.
-This is now explained in footnote ZZZ.}\fyTodo{to compute error rate of domain prediction over each test set, check comment, add footnote number.}
+This is now explained in footnote lines~739-743.}\fyTodo{to compute error rate of domain prediction over each test set, check comment, add footnote number.}
 \end{itemize}
@@
 \textcolor{blue}{From the UFAL medical corpus of the language pair En-Fr, we select only sentences in categories including EMEA, ParTR, CESTA, ECDC.
-This is now mentioned in the text (lines XXX)}\fyTodo{Add lines num}
+This is now mentioned in the text (lines~396-397)}\fyTodo{Add lines num}
 \\

### 2020-09-03 2020/tacl2020/multidomain.tex
add score table 6, add lines, minor corrections: dc feat to DC-Feat, {ADM} to \system{ADM}

@@
 \end{itemize}
-\revision{This list includes systems that slightly depart from our definition of MDMT: standard implementations of \system{TTM} and \system{WDCMT} rely on infered, rather than on gold, domain tags - which must somewhat affect their predictions; \system{DM} and {ADM} make no use of domain tags at all.} We however did not consider the proposal of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges more strongly from our notion of MDMT.
+\revision{This list includes systems that slightly depart from our definition of MDMT: standard implementations of \system{TTM} and \system{WDCMT} rely on infered, rather than on gold, domain tags - which must somewhat affect their predictions; \system{DM} and \system{ADM} make no use of domain tags at all.} We however did not consider the proposal of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges more strongly from our notion of MDMT.
 % and should be able to handle available domain labels in training and testing.
@@
 Underline denotes significant loss when domains are changed wrt.\ the baseline situation; bold for a significant improvement over \system{FT-Full}; (*) tags systems ignoring test domains.
-    \fyDone{Check wrong and significance of diff for ``old''}\fyDone{How about wrong ?}\jcTodo{FTres results are odd... split LAW 0.5}
+    \fyDone{Check wrong and significance of diff for ``old''}\fyDone{How about wrong ?}\jcDone{FTres results are odd... split LAW 0.5}
   }
@@
 10 large&251.1k&50.4&\bf 52.9&52.0&51.3&49.6&43.2&49.1&48.5&44.3&49.5\\
-Avg& &128.4k&\bf 51.3&\bf 51.3&50.9&50.9&43.4&50.0&49.4&46.7&50.29 \\ \hline
+Avg&128.4k&54.5&\bf 57.0&56.2&56.1&55.0&46.9&53.6&54.0&51.0&54.7\\ \hline
   \end{tabular}
@@
 Fine-tuning with small domains is often outperformed by other MDMT technique, an issue that a better regularization strategy might mitigate.
-Domain-control (\system{dc feat}) is very effective for small domains, but again less so in larger data conditions.
+Domain-control (\system{DC-Feat}) is very effective for small domains, but again less so in larger data conditions.
 Among the MD models, approaches using residual adapters have the best average performance.}

### 2020-09-04 2020/tacl2020/rebuttal.tex
final version for revision

--- 2020-09-02 2020/tacl2020/rebuttal.tex
@@
 
-\textcolor{blue}{We have joint to this mail two versions of the revised paper: one where the main changes appear in red, and the appendix is placed before the references so has to highlight the differences with the first version; one version that looks like a regular submission.
+\textcolor{blue}{We have joint to this mail two versions of the revised paper: one version where the main changes appear in red, and the appendix is placed before the references so as to highlight the differences with the first version; one version that looks like a regular submission, with the appendix following the references.
 We hope that this will make it easier for reviewers to view the differences with the first version of the paper.}
@@
 \\
-\textcolor{red}{%
+\textcolor{blue}{%
   We have tried to address these comments in the revised version of the paper.
@@
   \\
-  \textcolor{red}{%
+  \textcolor{blue}{%
     The two methods use the same architecture composed of one standard Transformer model extended with domain adapters following the design of \cite{Bapna19simple}.
@@
     We understand that our naming schemes was insisting on training procedures, rather than on architectures, thus possibly introducing some confusions.
-We have accordingly changed the name of \system{FT-Block} to \system{FT Res}, clarified the description of the two systems (see lines YYY\fyTodo{fix this}), and reported \system{FT-Res}\fyDone{or new name} as another multi-domain system in all tables.\fyTodo{Make sure it is the case}.}
+We have accordingly changed the name of \system{FT-Block} to \system{FT-Res}, clarified the description of the two systems (see lines~527-541\fyDone{fix this}), and reported \system{FT-Res}\fyDone{or new name} as another multi-domain system in all tables.\fyDone{Make sure it is the case}.}
   %}
@@
 As this is not central to our discussion, we have rephrased and dropped the comparison.
-See changes in line XXX}\fyTodo{Changes line in.}
+See changes in line~216-220}\fyDone{Changes line in.}
 \item 567-568: "having the domain tag on the source is better than on the target".
@@
 In our experiments, we found that TTM is pretty accurate in its prediction, making about 2\% errors in the predicted tags, which may explain the small difference with observe in Table~3.
-We have rephrased acccordingly in lines WWW.}\fyTodo{Add lines}
+We have rephrased acccordingly in lines~542-544, 550-551, 580-582.}\fyDone{Add lines}
 \end{itemize}
@@
 To benefit from your results, we need access to the same segmentation of train/dev/test sets that I hope you will provide to the community.
-\textcolor{blue}{Supplementary details regarding the settings used in our experiments have been added in a new Appendix YYY.\fyTodo{Fix the appendix.}.
+\textcolor{blue}{Supplementary details regarding the settings used in our experiments have been added in a new Appendix A.\fyDone{Fix the appendix.}.
 We will also distribute the detailed composition of our train / dev / test split upon acceptance of the paper.}
@@
 It will be helpful to see  the impact of the number of parameters on the MT results.
-  \textcolor{blue}{This information has been added in Table~3\fyTodo{Check number, page and line number and content}.}
+  \textcolor{blue}{This information has been added in Table~3\fyDone{Check number, page and line number and content}.}
 \\ 
@@
 I like your selections of domains that you include multiple domains with varying sizes, but what we expect in extreme cases.
-Having your Split experiments on REL (Table 4), or providing further details on your last analysis in Section 5.3, if you have tiny clusters could help.\fyTodo{Answer this with more experiments ?}
+Having your Split experiments on \domain{REL} (Table~4), or providing further details on your last analysis in Section 5.3, if you have tiny clusters could help.\fyDone{Answer this with more experiments ?}
   \\
   \textcolor{blue}{We agree that carefully analyzing the effects of using corpora of varying size is important, as this corresponds to practical training scenario.
-From our results in Table 7 (and Table 8 in Appendix B), we see that some techniques can work well with small in-domain corpora.
+From our results in a new Table 6 (and Table 8 in Appendix B), we see that some techniques can work well with small in-domain corpora.
 However, this analysis must also take in account the similarity between domains, as these improvements will be stronger if similar or related data are also present in the training mixture -- a situation that we have artificially created in our split conditions.
@@
 We hypothesize that training distribution is different from testing distribution, so we have need notation for two different mixtures.
-This has been clarified in the text (see line~XXX) \fyTodo{add lines}}
+This has been clarified in the text (see line~143-144) \fyDone{add lines}}
 \\
@@
 \\
-\textcolor{blue}{This reference has been added to the related work section.}\fyTodo{Add citation}
+\textcolor{blue}{This reference has been added to the related work section.}\fyDone{Add citation}
 \\
@@
   \\
-  \textcolor{blue}{The table has been updated accordingly.}\fyTodo{to report number of tokens of the datasets}
+  \textcolor{blue}{The table has been updated accordingly.}\fyDone{to report number of tokens of the datasets}
 \\
@@
 \\
-\textcolor{blue}{This has been added in the text (see line YYY).}\fyTodo{to explicitly compare WDCNMT with Mixed-NAT-RNN. This is in the legend - change presentations ?}
+\textcolor{blue}{This has been added in the text (see line~586-587 ).}\fyDone{to explicitly compare WDCNMT with Mixed-NAT-RNN. This is in the legend - change presentations ?}
 \\
@@
 Note that this classifier is only used for the \domain{news} test data.
-This is now explained in footnote ZZZ.}\fyTodo{to compute error rate of domain prediction over each test set, check comment, add footnote number.}
+This is now explained in footnote lines~739-743.}\fyDone{to compute error rate of domain prediction over each test set, check comment, add footnote number.}
 \end{itemize}
@@
 \textcolor{blue}{Table~1 now contains the number of types as well as the number of types that are unique in a domain.
-We fill that performing an error analysis for each of the system / domain compared in this study would require us to focus more on a detailled analysis of each system, while our main goal is to propose new ways to assess the actual abilities of existing MDMT systems.}\fyTodo{Will this be enough ?}
+We fill that performing an error analysis for each of the system / domain compared in this study would require us to focus more on a detailled analysis of each system, while our main goal is to propose new ways to assess the actual abilities of existing MDMT systems.}\fyDone{Will this be enough ?}
 \\
@@
 \\
-\textcolor{red}{For each domain, we create train/dev/test sets by randomly splitting each corpus.
+\textcolor{blue}{For each domain, we create train/dev/test sets by randomly splitting each corpus.
 We maintain the size of validation sets and of test sets equal to 1,000 lines for every domain.
@@
 We carefully removed every duplicate sentence in all the corpora.
-In our opinion, the reason why the models achieve extremely high scores in the \domain{rel} domain is because texts in this domain is very repetitive and often predictable.}\fyTodo{sentence length ?, diversity index of vocab ?}
+In our opinion, the reason why the models achieve extremely high scores in the \domain{rel} domain is because texts in this domain is very repetitive and often predictable.}\fyDone{sentence length ?, diversity index of vocab ?}
 \end{itemize}
@@
 \textcolor{blue}{From the UFAL medical corpus of the language pair En-Fr, we select only sentences in categories including EMEA, ParTR, CESTA, ECDC.
-This is now mentioned in the text (lines XXX)}\fyTodo{Add lines num}
+This is now mentioned in the text (lines~396-397)}\fyDone{Add lines num}
 \\

### 2020-09-04 2020/tacl2020/multidomain.tex
final version for revision

@@
 \newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{red}{#1}}}
-\newcommand{\revision}[1]{\textcolor{red}{#1}}
+\newcommand{\revision}[1]{\textcolor{black}{#1}}
 \newcommand{\revisiondel}[1]{}
@@
 \revision{The results in Table~\ref{tab:avg_automatic_domains} also suggest that MDMT can surpass fine-tuning for the smaller clusters; for the large clusters, this is no longer true.
-The complete table (in Appendix B) shows that this effect is more visible for small subsets of the medical domain.}
+The complete table (in Appendix~B) shows that this effect is more visible for small subsets of the medical domain.}
 
@@
 10 large&251.1k&50.4&\bf 52.9&52.0&51.3&49.6&43.2&49.1&48.5&44.3&49.5\\
-Avg&128.4k&54.5&\bf 57.0&56.2&56.1&55.0&46.9&53.6&54.0&51.0&54.7\\ \hline
+% Avg&128.4k&54.5&\bf 57.0&56.2&56.1&55.0&46.9&53.6&54.0&51.0&54.7\\ \hline
   \end{tabular}
@@
 
+\bibliographystyle{acl_natbib}
+\bibliography{multidomain}
+
 % \newpage\newpage
@@
   This experiment aims to simulate with automatic domains a scenario where the number of ``domains'' is large and where some ``domains'' are close and can effectively share information.
-Full results in Table~\ref{tab:automatic_domains}.
-Cluster size vary from approximately 8k sentences (cluster~24) up to more than 350k sentences.
+Complete results in Table~\ref{tab:automatic_domains}.
+Cluster sizes vary from approximately 8k sentences (cluster~24) up to more than 350k sentences.
 More than 2/3 of these clusters comprise mostly texts from one single domain, as for cluster 12 which is predominantly \domain{med}, the remaining clusters typically mix 2 domains.
-Fine-tuning with small domains is often outperformed by other MDMT technique, an issue that a better regularization strategy might mitigate.
-Domain-control (\system{DC-Feat}) is very effective for small domains, but again less so in larger data conditions.
-Among the MD models, approaches using residual adapters have the best average performance.}
+Fine-tuning with small domains is often outperformed by other MDMT techniques, an issue that a better regularization strategy might mitigate.
+Domain-control (\system{DC-Feat}) is very effective for small domains, but less so in larger data conditions.
+Among the MDMT models, approaches using residual adapters have the best average performance overall.}
 
-\clearpage{}
+% \clearpage{}
 
 % \newpage\newpage \newpage
-
-\bibliographystyle{acl_natbib}
-\bibliography{multidomain}
 

### 2020-09-10 2020/wmt_20/mdadapt.tex
fixed typos, add some future work

--- 2020-09-02 2020/wmt_20/emnlp2020.tex
@@
 Domain adaptation is an old and vexing problem for machine translation systems.
-The most common approach and successful to supervised adaptation is to fine-tune a baseline system with in-domain parallel data.
+The most common and successful approach to supervised adaptation is to fine-tune a baseline system with in-domain parallel data.
 Standard fine-tuning however modifies all the network parameters, which makes this approach computationally costly and prone to overfitting.
-A recent, lightweight approach, instead augments a baseline model with supplementary (small) adapter layers, keeping the rest of the mode unchanged.
-This has the additional merit to leave the baseline model intact, and adaptable to multiple domains.
+A recent, lightweight approach, instead augments a baseline model with supplementary (small) adapter layers, keeping the rest of the model unchanged.
+This has the additional merit to leave the baseline model intact and adaptable to multiple domains.
 In this paper, we conduct a thorough analysis of the adapter model in the context of a multidomain machine translation task.
-We contrast multiple implementations of this idea on two language pairs.
-Our main conclusions are that residual adapters provide a fast and cheap method for supervised multi-domain adaptation; our two variants prove as effective as the original adapter model, and open perspective to also make adapted models more robust to label domain errors.
+We contrast multiple implementations of this idea using two language pairs.
+Our main conclusions are that residual adapters provide a fast and cheap method for supervised multi-domain adaptation; our two variants prove as effective as the original adapter model and open perspective to also make adapted models more robust to label domain errors.
 \fyDone{abstract to be continued}
@@
 \end{abstract}
-\section{Introduction } \label{sec:intro}
+\section{Introduction \label{sec:intro}}
 \mpDone{write introduction} \fyDone{Citations in chronological order}\fyDone{Split long sentences}
-Owing to multiple improvements, Neural Machine Translation(NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence,Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
-However, as many deep learning model, NMT models need to be trained with sufficiently large amounts of data to reach their best performance.
-Therefore, the quality of translation of NMT model is still limited in low-resource language and low-resourced domain conditions \cite{duh13adaptation,zoph16transfer,koehn17six}.
-While many approaches have been proposed to improve the quality of NMT models in low-resource domains (see the recent survey of \citet{Chu18asurvey}), full fine-tuning of a generic baseline model remains the dominant supervised approach when adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
+Owing to multiple improvements, Neural Machine Translation (NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence,Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
+However, as many deep learning models, NMT systems need to be trained with sufficiently large amounts of data to reach their best performance.
+Therefore, the quality of translation of NMT models is still limited in low-resource language or domain conditions \cite{duh13adaptation,zoph16transfer,koehn17six}.
+While many approaches have been proposed to improve the quality of NMT models in low-resource domains (see the recent survey of \citet{Chu18asurvey}), full fine-tuning \cite{Luong15stanford,neubig18rapid} of a generic baseline model remains the dominant supervised approach when adapting NMT models to specific domains.
 
 Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pre-trained model by resuming training with the sole in-domain corpus.
-It is a conjecture that pre-trained model constitute a better initialization than a random one, especially when the adaptation data is scarce.
-Indeed, studies of transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim by extensive experiments.
+It is a conjecture that pre-trained model constitutes a better initialization than a random one, especially when adaptation data is scarce.
+Indeed, studies of transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim in extensive experiments.
 Full fine-tuning, that adapts all the parameters of a baseline model usually significantly improves the quality of the NMT for the chosen domain.
 However, it also yields large losses in translation quality for other domains, a phenomenon referred to as ``catastrophic forgetting'' in the neural network literature \cite{McCloskey89catastrophic}.
-Therefore, a fully fine-tuned model is only useful to one target domain.
-As the number of domains to handle grows, training and maintaining a separate model for each task can quickly becomes tedious and resource-expensive .\fyDone{Fix this sentence.}
+Therefore, a fully fine-tuned model is \emph{only useful to one target domain}.
+As the number of domains to handle grows, training and maintaining a separate model for each task can quickly become tedious and resource-expensive .\fyDone{Fix this sentence.}
 
@@
 
-In this paper, we would like to extend this architecture to improve NMT in several settings that still challenge automatic translation, such as translating texts from multiple topics, genre or domains, in the face of unbalanced data distribution.
+In this paper, we would like to extend this architecture to improve NMT in several settings that still challenge automatic translation, such as translating texts from multiple topics, genre or domains, in the face of unbalanced data distributions.
 Furthermore, as the notion of ``domains''  is not always well established, another practical setting is the translation of texts mixing several topics / domains.
-Finally, another requirement is the need to also translate texts from domain unseen in training, based only on the unadapted system, which should be made as strong as possible.
+An additional requirement is to translate texts from domains unseen in training, based only on the unadapted system, which should then be made as strong as possible.
 % Residual adapters allow us to adapt NMT model to any specific domain in a computationally cheap way.
@@
 
-In this context, our main contribution is a thorough experimental study of the use of residual adapters in multi-domain translation.
+In this context, our main contribution is a thorough experimental study of the use of residual adapters for multi-domain translation.
 We notably explore ways to adjust and/or regularize adapter modules to handle situations where the adaptation data is very small.
-We also propose and contrast two new variants of the residual architecture: in the first one (highway residual adapters), adaptation still affects each layer of the architecture, but its effect is delayed till the last layer, thus making the architecture more modular and adaptive; in the second one (gated residual adapters), we explore ways to improve the performance in the face of train-test data mismatch.
-We experiment with two language pairs and report results that demonstrate the flexibility and effectiveness of these architectures.
+We also propose and contrast two new variants of the residual architecture: in the first one (\emph{highway residual adapters}), adaptation still affects each layer of the architecture, but its effect is delayed till the last layer, thus making the architecture more modular and adaptive; in the second one (\emph{gated residual adapters}), we explore ways to improve performance in the face of train-test data mismatch.
+We experiment with two language pairs and report results that illustrate the flexibility and effectiveness of these architectures.
 \fyDone{One bit of a conclusion here}\fyFuture{Build a proper training scenario for these two conditions}
@@
 \subsubsection{The computation of adapter layers}
-Our reference architecture is the Transformer model of \cite{Vaswani17attention}, which we assume contains a stack of layers both in the encoder and the decoder side.
+Our reference architecture is the Transformer model of \cite{Vaswani17attention}, which we assume contains a stack of layers both on the encoder and the decoder sides.
 Each layer contains two subparts, an attention layer and a dense layer.
-Details may vary from one implementation to another, we simply content here that each layer $i \in \{1 \dots L\}$ (in the encoder or the decoder) computes a transform of a fixed-length sequence of $d$-dimensional input vectors $h^{i-1}$ into a sequence of output vectors $h^{i}$.
+Details vary from one implementation to another, we simply contend here that each layer $i \in \{1 \dots L\}$ (in the encoder or the decoder) computes a transform of a fixed-length sequence of $d$-dimensional input vectors $h^{i}$ into a sequence of output vectors $h^{i+1}$.
 
-The residual adapter architecture extends this model by adding one new computational module for each layer.
+The residual adapter architecture extends this model by adding one new computation module for each layer.
 The $i^{\text{th}}$ residual adapter thus modifies $h^i$ by computing the following transformations:\fyDone{Use align env}
@@
 \end{align*}
-Overall, the  $i^{\text{th}}$ adapter is thus parameterized by matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, with $b$ the dimension of the adapter (with $d \gg b$)\fyDone{Check this}.
+Overall, the  $i^{\text{th}}$ adapter is thus parameterized by matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, with $b$ the dimension of the adapter ($d \gg b$)\fyDone{Check this}.
 For the sake of brevity, we will simply denote $h^{i}_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the corresponding set of parameters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}
@@
 
-In the experiments reported in Section~\ref{sec:exp}, we will use $2\times{}L=12$ residual adapters, one for each of the $L=6$ attention layers of the encoder and similarly for the decoder.\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and only serves one residual adapter.}
+In the experiments of Section~\ref{sec:exp}, we use $2\times{}L=12$ residual adapters, one for each of the $L=6$ attention layers of the encoder and similarly for the decoder.\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and only serves one residual adapter.}
 
@@
 A related question concerns the regularization of adapter layers to mitigate overfitting.
-Reducing the number of adapters, or their dimensions, is simple, but these choices are difficult to optimize numerically -- an issue that becomes important as the number of domain grows.
+Reducing the number of adapters, or their dimensions, is simple, but such choices are difficult to optimize numerically -- an issue that becomes important as the number of domain grows.
 Less naïve alternatives can also be entertained, such as applying weight decay or layer regularization to the adapter.
@@
 \end{equation*}
-An alternative scheme is \emph{layer regularization}, which penalizes the output of the adapters, corresponding to the following new training objective:
+An alternative scheme is \emph{layer regularization}, which penalizes the output of the adapters, corresponding to the following objective:
 \begin{equation*}
@@
 Finally, another independant design choice relates to the training strategy for adapters.
-A first option is to generalize supervised domain adaptation to multi-domain adaptation and to proceed in two steps: (a) train a generic model with all the available data; (b) train eacj adapter layer with domain-specific data, keeping the generic model parameters unchanged.
-Another strategy is to adopt the view of \citet{Dredze08online}, where the multi-domain setting is viewed as an instance of multi-task learning \cite{Caruana97multitask}, where each domain corresponds to a specific task.
-This suggest to train the models as it were in a multi-task mode, training all the parameters from scratch: the generic parameters values will still depend on all the available data, while each adapter will only be trained with the corresponding in-domain data.
+A first option is to generalize supervised domain adaptation to multi-domain adaptation and to proceed in two steps: (a) train a generic model with all the available data; (b) train each adapter layer with domain-specific data, keeping the generic model parameters unchanged.
+Another strategy is to adopt the view of \citet{Dredze08online}, where the multi-domain setting is viewed as an instance of multi-task learning \cite{Caruana97multitask} with each domain corresponding to a specific task.
+This suggest to train all the parameters from scratch, as we would do in a multi-task mode.
+The generic parameters will still depend on all the available data, while each adapter will only be trained with the corresponding in-domain data.\fyFuture{Does everyone need a domain?}
 
@@
 In the basic architecture described in Section~\ref{ssec:architecture}, the computation performed by lower level layers will impact all the subsequent layers.
-In this section, we introduce an alternative implementation of the same idea, which however delays the adaptation of each layer to the last layer (of the encoder or the decoder) as represented on Figure~\ref{fig:hrl-architecture}.
+In this section, we introduce an alternative implementation of the same idea, which however delays the adaptation of each layer to the last layer (of the encoder or the decoder) as depicted on Figure~\ref{fig:hrl-architecture}.
 While the basic architecture performs adaptation in sequence, we propose here to perform it in parallel.
@@
 The basic architecture presented above rests on a rather simplistic view of ``domains'' as made of well-separated and unrelated pieces of texts that are processed independently during adaptation.
-Likewise, when translating test documents, one needs to make a choice between either using one specific domain-adapted model or resorting to the generic model.
-In this context, using wrong domain labels can have a strong (negative) on translation performance.
+Likewise, when translating test documents, one needs to choose between either using one specific domain-adapted model or resorting to the generic model.
+In this context, using wrong domain labels can have a strong (negative) effect on translation performance.
 
-Therefore, we would like to design a version of applying residual adapter that is more robust to such domains error.
-This variant, called the \emph{gated residual adapter models}, relies on the training of a supplementary trained component that will help decide whether to activate, on a word per word basis, a given residual layer, and to regulate the strength of this activation.
-To this end , we extend the highway version of residual adapters as follows.
+Therefore, we would like to design a version of residual adapters that is more robust to such domain errors.
+This variant, called the \emph{gated residual adapter model}, relies on the training of a supplementary trained component that will help decide whether to activate, on a word per word basis, a given residual layer, and to regulate the strength of this activation.
+To this end, we extend the highway version of residual adapters as follows.
 \fyDone{Consistency of notations wrt section 2.1}
@@
 where $z_k(h^L[t]) \in [0,1]$ measures the relatedness of the $t^{\text{th}}$ word $w_t$ to domain $k$.
-The more likely $w_t$ is in domain $k$, the larger $z_k(h^L[t])$ should be; conversely, for words\footnote{The term ``word'' is employed here by mere convenience, as the system only manipulates sub-lexical BPE units; furthermore, the values of the hidden representations $h^{i}$ at position $t$ depend upon all the other positions in the sentence.} that are not typical of any domain $k$ (eg.\ function words),  adaptation is minimum and the adapted corresponding encoder output ($\bar{h}^L[t]$) will remain close to the output of the generic model ($h^L[t]$).
-In our implementation, we incorporate two domain classifiers on top of the encoder and the decoder, that take the last hidden layer of the encoder (resp.\ decoder) as input and use the posterior probability $P(k|h^L[t]);)$ of domain $k$ as the value for $z_k(h^L[t])$.
+The more likely $w_t$ is in domain $k$, the larger $z_k(h^L[t])$ should be; conversely, for words\footnote{The term ``word'' is employed here by mere convenience, as systems only manipulate sub-lexical BPE units; furthermore, the values of the hidden representations $h^{i}$ at position $t$ depend upon all the other positions in the sentence.} that are not typical of any domain $k$ (eg.\ function words),  adaptation is minimum and the corresponding adapted encoder output ($\bar{h}^L[t]$) will remain close to the output of the generic model ($h^L[t]$).
+In our implementation, we incorporate two domain classifiers on top of the encoder and the decoder, that take the last hidden layer of the encoder (resp.\ decoder) as input and use the posterior probability $P(k|h^L[t])$ of domain $k$ as the value for $z_k(h^L[t])$.
 
@@
 \begin{enumerate}
-\item learn a generic model with mixed corpora comprising data from multiple domains.
+\item learn a generic model with mixed corpora from multiple domains.
 \item train a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen.
@@
 \subsection{Data and metrics \label{ssec:corpora}}
-We perform our experiments with two translation pairs involving multiple domains: English-French (En$\rightarrow$Fr and English-German (En$\rightarrow$De).
+We perform our experiments with two translation pairs involving multiple domains: English-French (En$\rightarrow$Fr) and English-German (En$\rightarrow$De).
 For the former pair, we use texts\footnote{Most corpora are available from the Opus web site: \url{http://opus.nlpl.eu}} initially from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
@@
 \subsection{Baseline architectures \label{ssec:baseline}}
-\fyDone{Write this - settings and parameters for Mixed and Full-FT}.
+\fyDone{Write this - settings and parameters for Mixed and Full-FT}
 % Our baselines are standard for multi-domain systems.
@@
 \begin{itemize}
-\item a generic model trained on a concatenation of all corpora, denoted \system{Mixed}\fyDone{Or mixed nat ?}
+\item a generic model trained on a concatenation of all corpora, denoted \system{Mixed};\fyDone{Or mixed nat ?}
 \item a fine-tuned model \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing during 3 consecutive epochs.
@@
 In this section, we evaluate several proposals from the literature on multi-domain adaptation and compare them to full fine-tuning on the one hand, and to two variants of the residual adapter architecture on the other hand.
+The reference methods included in our experiments are the following:
+\begin{itemize}
+\item a system using ``domain control'' \cite{Kobus17domaincontrol}.
+In this approach, domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}) or in the form of a supplementary feature for each word (\system{DC-Feat});
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain-specific part (\system{LDR});
+\item the three proposals of \newcite{Britz17mixing}.
+\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
+The training uses reference tags and inference is performed with predicted tags, just like for regular target words.
+\system{DM} is a multi-task learner where a domain classifier is trained on top of the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
+These methods only use domain labels in training.
+\end{itemize}
 
@@
 \end{table*}
-
-The reference methods included in our experiments are the following:
-\begin{itemize}
-\item a system using ``domain control'' \cite{Kobus17domaincontrol}.
-In this approach, domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}) or in the form of a supplementary feature for each word (\system{DC-Feat});
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain-specific part (\system{LDR});
-\item the three proposals of \newcite{Britz17mixing}.
-\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
-The training uses reference tags and inference is performed with predicted tags, just like for regular target words.
-\system{DM} is a multi-task learner where a domain classifier is trained on top of the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
-These methods thus only use domain labels in training.
-\end{itemize}
-
-The two variants of the residual adapter model included in this first round of experiment have presented in Section~\ref{ssec:architecture}: \system{Res-Adap} is the multi-domain version of the approach of \citet{Bapna19simple} based on a two step training procedure; while \system{Res-Adap-MT} is the ``multi-task'' version, where the parameters of the generic model and of the adapter  are learned simultaneously from scratch.
+The two variants of the residual adapter model included in this first round of experiment have been presented in Section~\ref{ssec:architecture}: \system{Res-Adap} is the multi-domain version of the approach of \citet{Bapna19simple} based on a two step training procedure; while \system{Res-Adap-MT} is the ``multi-task'' version, where the parameters of the generic model and of the adapters are jointly learned from scratch.
 
@@
 The ``multi-task'' variant is slightly less effective in our experiments than the basic version, where optimization is performed in two steps.
-As it turns out, using residual adapters prove here to be much more effective on average than the other reference multi-domain systems; it is also much better than the generic system for translating data from known domains, outperforming the \system{Mixed} system by more than 4 BLEU points in average.
+As it turns out, using residual adapters proves here more effective on average than the other reference multi-domain systems; it is also much better than the generic system for translating data from known domains, outperforming the \system{Mixed} system by more than 4 BLEU points in average.
 Gains are especially large for small domains such as \domain{law} and \domain{rel}.
@@
 Two settings are considered: keeping just one adapter, or keeping the three.
-The trend is the same for the two language directions: suppressing adapters always decrease the overall performance, albeit by a small margin: having six adapters is better than three, which is better than keeping only one.
-With only one adapter active, we observe small, unsignificant changes in performance.
-Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.
+The trend is the same for the two language directions: suppressing adapters always hurts the overall performance, albeit by a small margin: having six adapters is better than three, which is better than keeping only one.
+With only one adapter active, we observe small, unsignificant changes in performance when varying the adapter's depth.
+Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.\fyFuture{More configuration ?}
 
@@
 
-For each method, the optimal hyper-parameters (weight decay or layer regularization coefficient, see Section~\ref{sssec:design-space}) $\lambda$ by chosen by grid search among a small set of values ($\{ 10^{-3}, 10^{-4}, 10^{-5} \}$).
+For each method, the optimal hyper-parameters (weight decay or layer regularization coefficient, see Section~\ref{sssec:design-space}) $\lambda$ by chosen by grid search among a small set of values ($\{ 10^{-3}, 10^{-4}, 10^{-5} \}$).\fyFuture{Better grid}
 
@@
 We also compute translation for a domain unseen in training (\domain{news}) as follows.
-For each sentence of this test set, we automatically evaluate the closest domain\footnote{As measured by the perplexity of a language model trained with only in-domain data.\fyFuture{More on this for reproducibility}.}, then use the predicted domain label to compute the translation.
+For each sentence of this test set, we automatically evaluate the closest domain,\footnote{As measured by the perplexity of a language model trained with only in-domain data.\fyFuture{More on this for reproducibility}.} then use the predicted domain label to compute the translation.
 This is an error prone process, which also challenges the robustness of our multi-domain systems.
@@
 The same trend holds for the unseen \domain{news} domain, with \system{Res-Adap-Gated} being the best domain adapted system in our set, outperforming the other variants by about 2 BLEU points.
-
 % The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
@@
 \fyFuture{Computation time should also report the time it takes for the complete tuning process, not one iteration which will be more or less the same}.
-
 \section{Related Work \label{sec:related}}
@@
 
-Early approaches to the multi-domain setting were proposed for statistical MT, either considering multiple data sources (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}) or domains containing several topics \cite{Eidelman12topic,Hasler14dynamic-topic}.
-Two main strategies are considered: feature-based methods, where domain/topic labels are integrated through supplementary feature and instance-based methods, involving a measure of similarity between train and test domains.
+Early approaches to multi-domain MT were proposed for statistical MT, either considering multiple data sources (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}) or domains containing several topics \cite{Eidelman12topic,Hasler14dynamic-topic}.
+Two main strategies emerge: feature-based methods, where domain labels are integrated through supplementary features; and instance-based methods, involving a measure of similarity between train and test domains.
 
@@
 
-\citet{Britz17mixing} studies three general techniques to take domain information into account in training: they rely on either domain classification or domain normalization on the source or target side.
-One contribution of this study is the proposal of an adversarial training scheme to normalize representations across domains and make the combination of multiple data sources more effective.
+\citet{Britz17mixing} study three general techniques to take domain information into account in training: they rely on either domain classification or domain normalization on the source or target side.
+A contribution of this study is an adversarial training scheme to normalize representations across domains and make the combination of multiple data sources more effective.
 Similar techniques (parameter sharing, automatic domain classification / normalization) are at play in \cite{Zeng18multidomain}: in this work, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific from domain-agnostic representations.
-These two representations are then processed separately, then merged to compute the final translation.
+These representations are first processed separately, then merged to compute the final translation.
 
 \citet{Farajian17multidomain,li-etal-2018-one} are two recent representatives of the instance-based approach: for each test sentence, a small adaptation corpus is collected based on similarity measures, and used to fine-tune a mix-domain model.
+
+Finally note that a distinct evolution of the residual adapter model of \cite{Bapna19simple} is presented in \cite{Sharaf20metalearning} where meta-learning techniques are used to make fine tuning more effective in a standard domain-adaptation setting.
 
@@
 Several new variants have been introduced and evaluated for two language directions: if none that able to clearly surpass the baseline residual adapter models, they provide directions for improving this model in practical settings: unbalanced data condition, noise in label domains, etc.
-In our future work, we would like to continue the development of the Gated variant, which, it seems to us, provides a flexible and robust tool to address the various challenges of multi-domain machine translation.
+In our future work, we would like to continue the development of the gated variant, which, it seems to us, provides a flexible and robust tool to address the various challenges of multi-domain machine translation.
 
@@
 
+\fyFuture{Example of gating ?}
+\fyFuture{Why MDMT does not work well - full train from generic?}
+\fyFuture{Structured domain? A Bayesian version?}
+\fyFuture{Let the machine decide domains in learning? - domains and their organization - how could it?}
+
 %\section*Acknowledgments}
@@
 \bibliographystyle{acl_natbib}
-\bibliography{emnlp2020}
+\bibliography{mdadapt}
 %\appendix

### 2020-09-11 2020/wmt_20/mdadapt.tex
discussion avec M.Q-P

@@
 This suggest to train all the parameters from scratch, as we would do in a multi-task mode.
-The generic parameters will still depend on all the available data, while each adapter will only be trained with the corresponding in-domain data.\fyFuture{Does everyone need a domain?}
+The generic parameters will still depend on all the available data, while each adapter will only be trained with the corresponding in-domain data.\fyFuture{Does everyone need a domain?
+Do we need this discussion - talk about sentence level adaptation, better classifier pour MDL}
 
@@
 \end{enumerate}
-\fyFuture{is this classifier important, can we train with the rest of the system ?}
+\fyFuture{is this classifier important, can we train with the rest of the system ?
+Yes we could, but do we want the encoder to be good at classification?}\fyFuture{Check writing to indicate label smoothing for this has been used.}\fyFuture{Can we train a more complex training model ?
+should we us emore monolingual data ?
+etc}
 
@@
     \system{Res-Adap-MT}  & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
+    %  \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
     \hline
@@
 With only one adapter active, we observe small, unsignificant changes in performance when varying the adapter's depth.
-Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.\fyFuture{More configuration ?}
+Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.\fyFuture{More configurations ?}
 
@@
 
-For each method, the optimal hyper-parameters (weight decay or layer regularization coefficient, see Section~\ref{sssec:design-space}) $\lambda$ by chosen by grid search among a small set of values ($\{ 10^{-3}, 10^{-4}, 10^{-5} \}$).\fyFuture{Better grid}
+For each method, the optimal hyper-parameters (weight decay or layer regularization coefficient, see Section~\ref{sssec:design-space}) $\lambda$ by chosen by grid search among a small set of values ($\{ 10^{-3}, 10^{-4}, 10^{-5} \}$).\fyFuture{Better grid even better EWC.}
 
@@
 A first observation is that for domains seen in training, our variants \system{Res-Adap-HW} and \system{Res-Adap-Gated} achieve BLEU scores that are on a par to those of the original version (\system{Res-Adap}), with insignificant variations across test sets.
-\fyFuture{Comment on processing time}
+\fyFuture{Comment on processing time.}
 
@@
 \fyFuture{Example of gating ?}
-\fyFuture{Why MDMT does not work well - full train from generic?}
+\fyTodo{Why MDMT does not work well - full train from generic? - add new results}
 \fyFuture{Structured domain? A Bayesian version?}

### 2020-09-21 2020/wmt_20/mdadapt.tex
pretrained init scores, minor corrections

@@
 \newcommand{\R}{\ensuremath{\mathbb{R}}}
+
+\usepackage{footnote}
+\makesavenoteenv{tabular*}
+\makesavenoteenv{table*}
 
@@
 Yes we could, but do we want the encoder to be good at classification?}\fyFuture{Check writing to indicate label smoothing for this has been used.}\fyFuture{Can we train a more complex training model ?
-should we us emore monolingual data ?
+should we use more monolingual data ?
 etc}
@@
   \centering
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+  \begin{tabular}{|p{4.3cm}|*{8}{r|}} \hline
 %     &&&&&& \\
@@
     \system{Res-Adap-MT}  & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
+    \system{Res-Adap-MT-ptr\footnote{We initialized shared parameters by values of parameters of \system{mixed} model and train Res-Adap using multi-task learning}} & 37.47 & 57.06 & 52.39 & 33.72 & 46.17 & 89.47 & 52.7 \\
     %  \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
@@
    \system{Res-Adap}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 & 213M/96M \\ \hline
-    \system{Res-Adap$_{(2,4,6)}$} & 29.7  & 30.5 & 18.8 & 49.6 & 47.1 & 20.6 &  32.7 & 213M/48M \\ \hline
+    \system{Res-Adap$_{(2,4,6)}$} & 29.7  & 30.5 & 18.8 & 49.6 & 47.1 & 20.6 &  32.7 & 213M/48M \\ 
     \system{Res-Adap$_{(6)}$}      & 29.5 & 30.4 & 18.1 & 49.1 & 46.9 & 20.4 & 32.4 & 213M/16M \\
-   \system{Res-Adap$_{(4)}$}       & 29.7 & 30.4 & 18.1 & 49.6 & 47.0 & 20.6 & 32.6 & 213/16M\\
+   \system{Res-Adap$_{(4)}$}       & 29.7 & 30.4 & 18.1 & 49.6 & 47.0 & 20.6 & 32.6 & 213M/16M\\
    \system{Res-Adap$_{(2)}$}       & 29.6 & 30.4 & 18.3 & 49.4 & 46.7 & 20.6 & 32.5  & 213M/16M\\
+   \hline
     \system{Res-Adap-WD}         & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 & 213M/96M \\
@@
   \centering
-  \begin{tabular}{|p{3cm}|*{7}{r|}|r|r|} \hline
+  \begin{tabular}{|p{4.8cm}|*{7}{r|}|r|r|} \hline
 %     &&&&&& \\
@@
     \system{Res-Adap-HW}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91.0  & 53.1 & 36.6 & 20.2 \\ %& 19s \\
+    \system{Res-Adap-HW-MT} & 37.4 & 56.4 & 52.1 & 33.7 & 44.8 & 89.8 & 52.4 & 27.1 & 20.4 \\ %&  \\
+    \system{Res-Adap-HW-MT-ptr \footnote{We initialized shared parameters by values of parameters of \system{mixed} model and train Res-Adap-HW using multi-task learning}} & 37.7 & 57 & 52.5 & 33.5 & 46.1 & 89 & 52.6 & 46.5 & 21.4 \\ %&  \\
     \system{Res-Adap-Gate}  & 38.0 & 57.5& 53.0 & 33.5 & 46.0 & 90.1  & 53.0 & 49.0 & 22.5 \\ %& 21s \\

### 2020-10-06 2020/wmt_20/mdadapt.tex
local corrections following reviews

@@
 \newcommand{\system}[1]{\texttt{\textbf{#1}}}
+\newcommand{\revision}[1]{\textcolor{red}{#1}}
 \newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
@@
 \date{}
-
+%\hypersetup{draft}
 \begin{document}
@@
 However, as many deep learning models, NMT systems need to be trained with sufficiently large amounts of data to reach their best performance.
-Therefore, the quality of translation of NMT models is still limited in low-resource language or domain conditions \cite{duh13adaptation,zoph16transfer,koehn17six}.
+Therefore, the quality of the translation of NMT models is still limited in low-resource language or domain conditions \cite{duh13adaptation,zoph16transfer,koehn17six}.
 While many approaches have been proposed to improve the quality of NMT models in low-resource domains (see the recent survey of \citet{Chu18asurvey}), full fine-tuning \cite{Luong15stanford,neubig18rapid} of a generic baseline model remains the dominant supervised approach when adapting NMT models to specific domains.
@@
 Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pre-trained model by resuming training with the sole in-domain corpus.
-It is a conjecture that pre-trained model constitutes a better initialization than a random one, especially when adaptation data is scarce.
-Indeed, studies of transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim in extensive experiments.
+It is a conjecture that the pretrained model constitutes a better initialization than a random one, especially when adaptation data is scarce.
+Indeed, studies of transfer learning for NMT such as \citet{artetxe20cross,aji20neural} have confirmed this claim in extensive experiments.
 Full fine-tuning, that adapts all the parameters of a baseline model usually significantly improves the quality of the NMT for the chosen domain.
@@
 Therefore, a fully fine-tuned model is \emph{only useful to one target domain}.
-As the number of domains to handle grows, training and maintaining a separate model for each task can quickly become tedious and resource-expensive .\fyDone{Fix this sentence.}
+As the number of domains to handle grows, training, and maintaining a separate model for each task can quickly become tedious and resource-expensive .\fyDone{Fix this sentence.}
 
-\cite{Vilar18learning,Bapna19simple} have recently proposed a simple and lightweight domain adaptation method that also preserves the value of pre-trained models, based on small adapters component that are plugged in each hidden layer.
+\citet{Vilar18learning,Bapna19simple,wuebker18compact,michel18extreme} have recently proposed a simple and lightweight domain adaptation method that also preserves the value of pre-trained models, based on small adapter components that are plugged in each hidden layer.
 These adapters are trained only with the in-domain data, keeping the pre-trained model frozen.
@@
 
-In this paper, we would like to extend this architecture to improve NMT in several settings that still challenge automatic translation, such as translating texts from multiple topics, genre or domains, in the face of unbalanced data distributions.
-Furthermore, as the notion of ``domains''  is not always well established, another practical setting is the translation of texts mixing several topics / domains.
+In this paper, we would like to extend this architecture to improve NMT in several settings that still challenge automatic translation, such as translating texts from multiple topics, genre, or domains, in the face of unbalanced data distributions.
+Furthermore, as the notion of ``domains''  is not always well established, another practical setting is the translation of texts mixing several topics/domains.
 An additional requirement is to translate texts from domains unseen in training, based only on the unadapted system, which should then be made as strong as possible.
-% Residual adapters allow us to adapt NMT model to any specific domain in a computationally cheap way.
-Could we adapt NMT model to noisy text and topical text at once?
-While residual adapters are good at adapting to one specific domain, their performance still dramatically decrease for the other domains seen in training, as we show in our experiments.
+% Residual adapters allow us to adapt the NMT model to any specific domain in a computationally cheap way.
+Could we adapt the NMT model to noisy text and topical text at once?
+While residual adapters are good at adapting to one specific domain, their performance still dramatically decreases for the other domains seen in training, as we show in our experiments.
 Therefore, we have to decide whether to use residual adapters manually, i.e, we have to know to which domain the text belongs a priori.
@@
 \section{Residual adapters \label{sec:res}}
-In this section, we describe the basic version of the residual adapter architectures, as well as two original variants of this model.
+In this section, we describe the basic version of the residual adapter architectures \citep{houlsby19parameter, Bapna19simple}, as well as two original variants of this model.
 
@@
 \subsubsection{The computation of adapter layers}
-Our reference architecture is the Transformer model of \cite{Vaswani17attention}, which we assume contains a stack of layers both on the encoder and the decoder sides.
-Each layer contains two subparts, an attention layer and a dense layer.
+Our reference architecture is the Transformer model of \citet{Vaswani17attention}, which we assume contains a stack of layers both on the encoder and the decoder sides.
+Each layer contains two subparts, an attention layer, and a dense layer.
 Details vary from one implementation to another, we simply contend here that each layer $i \in \{1 \dots L\}$ (in the encoder or the decoder) computes a transform of a fixed-length sequence of $d$-dimensional input vectors $h^{i}$ into a sequence of output vectors $h^{i+1}$.
@@
 The residual adapter architecture extends this model by adding one new computation module for each layer.
-The $i^{\text{th}}$ residual adapter thus modifies $h^i$ by computing the following transformations:\fyDone{Use align env}
+The $i^{\text{th}}$ residual adapter thus modifies $h^i$ by computing the following transformations:\fyDone{Use align env} \fyTodo{Add layer normalization, replace notation b by a}
 \begin{align*}
-  h^{i}_1 &= \mathbf{W}_{db}^{i}h^{i} + b^i_{1} \\
+  h^{i}_0 &= \mathbf{LN}(h^{i}) \\
+  h^{i}_1 &= \mathbf{W}_{db}^{i}h_0^{i} + a^i_{1} \\
   h^{i}_2 &= \mathbf{ReLU}(h_1^{i})
@@
 %\begin{align*}
-  h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + b^i_{2} \\
+  h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + a^i_{2} \\
   \bar{h}^{i} &= h^{i}_3 + h^i.
 \end{align*}
-Overall, the  $i^{\text{th}}$ adapter is thus parameterized by matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, with $b$ the dimension of the adapter ($d \gg b$)\fyDone{Check this}.
+Overall, the  $i^{\text{th}}$ adapter is thus parameterized by matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, with $b$ the dimension of the adapter \fyTodo{($d \gg b$) not true}\fyDone{Check this}.
 For the sake of brevity, we will simply denote $h^{i}_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the corresponding set of parameters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}
@@
 \subsubsection{Design space and variants \label{sssec:design-space}}
-This general architecture leaves open many design choices pertaining to the details of the network organization, the training procedure and the corresponding objective function.
+This general architecture leaves open many design choices pertaining to the details of the network organization, the training procedure, and the corresponding objective function.
 
-A first question concerns the number of adapter layers.
+The first question concerns the number of adapter layers.
 While in principle, all Transformer layers can be subject to adaptation, it is nonetheless worthwhile to consider simpler adaptation schemes, which would only alter a limited number of layers.
@@
 It is conjectured that the higher layers represent more global patterns such as semantic while the lower layers represent more local patterns such as syntactic \cite{raganato18analysis}\fyTodo{Missing reference here}.
-The domain shift in local patterns and global patterns has not yet studied.
+The domain shift in local patterns and global patterns has not yet been studied.
 In this paper, we do not intend to study this aspect of the domain adaptation problem.
-We would like to give an extensive comparison between domain adaptation in different levels in NMT model.
+We would like to give an extensive comparison between domain adaptation in different levels in the NMT model.
 
@@
 Reducing the number of adapters, or their dimensions, is simple, but such choices are difficult to optimize numerically -- an issue that becomes important as the number of domain grows.
-Less naïve alternatives can also be entertained, such as applying weight decay or layer regularization to the adapter.
+Less naive alternatives can also be entertained, such as applying weight decay or layer regularization to the adapter.
 Implementing these requires to modify the objective function in a way that still allows for a smooth optimization problem.
@@
 
-Finally, another independant design choice relates to the training strategy for adapters.
+Finally, another independent design choice relates to the training strategy for adapters.
 A first option is to generalize supervised domain adaptation to multi-domain adaptation and to proceed in two steps: (a) train a generic model with all the available data; (b) train each adapter layer with domain-specific data, keeping the generic model parameters unchanged.
 Another strategy is to adopt the view of \citet{Dredze08online}, where the multi-domain setting is viewed as an instance of multi-task learning \cite{Caruana97multitask} with each domain corresponding to a specific task.
-This suggest to train all the parameters from scratch, as we would do in a multi-task mode.
+This suggests training all the parameters from scratch, as we would do in a multi-task mode.
 The generic parameters will still depend on all the available data, while each adapter will only be trained with the corresponding in-domain data.\fyFuture{Does everyone need a domain?
-Do we need this discussion - talk about sentence level adaptation, better classifier pour MDL}
+Do we need this discussion - talk about sentence-level adaptation, better classifier pour MDL}
 
@@
 Therefore, we would like to design a version of residual adapters that is more robust to such domain errors.
-This variant, called the \emph{gated residual adapter model}, relies on the training of a supplementary trained component that will help decide whether to activate, on a word per word basis, a given residual layer, and to regulate the strength of this activation.
+This variant, called the \emph{gated residual adapter model}, relies on the training of a supplementary trained component that will help decide whether to activate, on a word per word basis, a given residual layer and to regulate the strength of this activation.
 To this end, we extend the highway version of residual adapters as follows.
@@
 \end{equation}
-where $z_k(h^L[t]) \in [0,1]$ measures the relatedness of the $t^{\text{th}}$ word $w_t$ to domain $k$.
+where the scalar $z_k(h^L[t]) \in [0,1]$ measures the relatedness of the $t^{\text{th}}$ word $w_t$ to domain $k$.
 The more likely $w_t$ is in domain $k$, the larger $z_k(h^L[t])$ should be; conversely, for words\footnote{The term ``word'' is employed here by mere convenience, as systems only manipulate sub-lexical BPE units; furthermore, the values of the hidden representations $h^{i}$ at position $t$ depend upon all the other positions in the sentence.} that are not typical of any domain $k$ (eg.\ function words),  adaptation is minimum and the corresponding adapted encoder output ($\bar{h}^L[t]$) will remain close to the output of the generic model ($h^L[t]$).
@@
 \end{enumerate}
-\fyFuture{is this classifier important, can we train with the rest of the system ?
+\fyFuture{is this classifier important, can we train with the rest of the system?
 Yes we could, but do we want the encoder to be good at classification?}\fyFuture{Check writing to indicate label smoothing for this has been used.}\fyFuture{Can we train a more complex training model ?
-should we use more monolingual data ?
+should we use more monolingual data?
 etc}
@@
 Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2,048 cells.
-Residual adapters additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$.
+Residual adapters additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$.\revision{ \citet{Bapna19simple} showed that the performance of adapted model increases with respect to the size of the inner dimension.
+In the experiments of \citet{Bapna19simple}, we note that $b=1024$, which is 2-folds larger than the dimension of Transformer layer, achieved performance closed to fine-tuned model.
+Therefore, we would like to apply this setting to our experiments.}
 % The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
 % The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training use a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4,000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
+Training uses a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4,000$), and a dropout rate of $0.1$ for all layers.\revision{ For \system{mixed} model, we use initial learning 1.0 and the concatenation of the validation sets of 6 domains as the validation set of \system{mixed}.
+In the fine-tuning experiments, we continue training using \system{mixed} as starting point, we use the same learning rate schedule, continuing the incrementation of the number of steps.
+In the multi-task training, we use the same learning rate schedule as \system{mixed}, for each iteration, we sample a domain from the pool of 6 domains with the probability proportional to the size of training data of each domain, then we sample a batch of~12,288 tokens from training data of the picked domain, the shared parameters and the parameters of the corresponding adapter are updated by gradients.}\fyDone{Describe the block adaptation layer - voir slides}
 
@@
     \system{Transformer-En-De}  & 213 \\
-    \system{Residual Adapter-En-De} & 8 \\
+    \system{Residual Adapter-En-De} & 4 \\
      \hline
@@
 \end{table*}
-The two variants of the residual adapter model included in this first round of experiment have been presented in Section~\ref{ssec:architecture}: \system{Res-Adap} is the multi-domain version of the approach of \citet{Bapna19simple} based on a two step training procedure; while \system{Res-Adap-MT} is the ``multi-task'' version, where the parameters of the generic model and of the adapters are jointly learned from scratch.
+The two variants of the residual adapter model included in this first round of experiment have been presented in Section~\ref{ssec:architecture}: \system{Res-Adap} is the multi-domain version of the approach of \citet{Bapna19simple} based on a two-step training procedure; while \system{Res-Adap-MT} is the ``multi-task'' version, where the parameters of the generic model and of the adapters are jointly learned from scratch.
 
@@
 
-By varying the number and position of residual adapters (see Section~\ref{ssec:architecture}), we then contrast several implementation of residual adapters.
+By varying the number and position of residual adapters (see Section~\ref{ssec:architecture}), we then contrast several implementations of residual adapters.
 \fyDone{Fix style here} Because the set of possible configurations is large, we only perform experiments for layers $i= 2, 4, 6$ (both for the encoder and decoder).
-Two settings are considered: keeping just one adapter, or keeping the three.
+Two settings are considered: keeping just one adapter or keeping the three.
 The trend is the same for the two language directions: suppressing adapters always hurts the overall performance, albeit by a small margin: having six adapters is better than three, which is better than keeping only one.
-With only one adapter active, we observe small, unsignificant changes in performance when varying the adapter's depth.
-Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.\fyFuture{More configurations ?}
+With only one adapter active, we observe small, insignificant changes in performance when varying the adapter's depth.
+\revision{Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.}\fyFuture{More configurations ?}
 
@@
     \system{Mixed}          & 29.3 & 30.5 & 17.6 & 38.1 & 47.9 & 20.9  & 30.6 & 213M/0M\\
-   \system{Res-Adap}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 & 213M/96M \\ \hline
-    \system{Res-Adap$_{(2,4,6)}$} & 29.7  & 30.5 & 18.8 & 49.6 & 47.1 & 20.6 &  32.7 & 213M/48M \\ 
-    \system{Res-Adap$_{(6)}$}      & 29.5 & 30.4 & 18.1 & 49.1 & 46.9 & 20.4 & 32.4 & 213M/16M \\
-   \system{Res-Adap$_{(4)}$}       & 29.7 & 30.4 & 18.1 & 49.6 & 47.0 & 20.6 & 32.6 & 213M/16M\\
-   \system{Res-Adap$_{(2)}$}       & 29.6 & 30.4 & 18.3 & 49.4 & 46.7 & 20.6 & 32.5  & 213M/16M\\
+   \system{Res-Adap}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 & 213M/48M \\ \hline
+    \system{Res-Adap$_{(2,4,6)}$} & 29.7  & 30.5 & 18.8 & 49.6 & 47.1 & 20.6 &  32.7 & 213M/24M \\ 
+    \system{Res-Adap$_{(6)}$}      & 29.5 & 30.4 & 18.1 & 49.1 & 46.9 & 20.4 & 32.4 & 213M/8M \\
+   \system{Res-Adap$_{(4)}$}       & 29.7 & 30.4 & 18.1 & 49.6 & 47.0 & 20.6 & 32.6 & 213M/8M\\
+   \system{Res-Adap$_{(2)}$}       & 29.6 & 30.4 & 18.3 & 49.4 & 46.7 & 20.6 & 32.5  & 213M/8M\\
    \hline
-    \system{Res-Adap-WD}         & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 & 213M/96M \\
-    \system{Res-Adap-LR}           & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1  & 213M/96M\\
+    \system{Res-Adap-WD}         & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 & 213M/48M \\
+    \system{Res-Adap-LR}           & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1  & 213M/48M\\
     \hline
@@
 \fyDone{How is the weight decay parameter set ?}
-\fyDone{Why is regularization not helping ? It helps for small domain - domain-specific regularization ??}
+\fyDone{Why is regularization not helping? It helps for the small domain - domain-specific regularization ??}
 
@@
 For each sentence of this test set, we automatically evaluate the closest domain,\footnote{As measured by the perplexity of a language model trained with only in-domain data.\fyFuture{More on this for reproducibility}.} then use the predicted domain label to compute the translation.
-This is an error prone process, which also challenges the robustness of our multi-domain systems.
+This is an error-prone process, which also challenges the robustness of our multi-domain systems.
 Results are in Table~\ref{tab:performance-random}.
@@
 The two other settings are instructive in several ways: they first clearly illustrate that the brittleness of domain-adapted systems, for which large drops in performance (more than 15 BLEU points on average) are observed when the domain label is randomly chosen.
-Our gated variant however proves much more robust than the other adaptation strategy, and performs almost on par to the generic system for that test condition.
+Our gated variant however proves much more robust than the other adaptation strategy and performs almost on par to the generic system for that test condition.
 The same trend holds for the unseen \domain{news} domain, with \system{Res-Adap-Gated} being the best domain adapted system in our set, outperforming the other variants by about 2 BLEU points.
@@
   \centering
-  \begin{tabular}{|p{4.8cm}|*{7}{r|}|r|r|} \hline
+  \begin{tabular}{|p{4cm}|*{8}{r|}|r|r|} \hline
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c||}{\domain{avg}} & \multicolumn{1}{c|}{\domain{rnd}} & \multicolumn{1}{c|}{\domain{news}} \\ \hline % & \multicolumn{1}{c|}{\domain{PROC}} \\ \hline
-    \system{Mixed}             & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5     &  49.4 & 49.4 & 23.5 \\ %& 34s \\ %
-    \system{FT-Full}             & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8   & 53.8 & 32.5 & 20.2  \\ %& 34s \\
-    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2   & 53.3 & 38.4 & 20.5 \\ %& 22s\\ 
-    \system{Res-Adap-HW}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91.0  & 53.1 & 36.6 & 20.2 \\ %& 19s \\
-    \system{Res-Adap-HW-MT} & 37.4 & 56.4 & 52.1 & 33.7 & 44.8 & 89.8 & 52.4 & 27.1 & 20.4 \\ %&  \\
-    \system{Res-Adap-HW-MT-ptr \footnote{We initialized shared parameters by values of parameters of \system{mixed} model and train Res-Adap-HW using multi-task learning}} & 37.7 & 57 & 52.5 & 33.5 & 46.1 & 89 & 52.6 & 46.5 & 21.4 \\ %&  \\
-    \system{Res-Adap-Gate}  & 38.0 & 57.5& 53.0 & 33.5 & 46.0 & 90.1  & 53.0 & 49.0 & 22.5 \\ %& 21s \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c||}{\domain{avg}} & \multicolumn{1}{c|}{\domain{rnd}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{\domain{PARAMS}}  \\ \hline % & \multicolumn{1}{c|}{\domain{PROC}} \\ \hline
+    \system{Mixed}             & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5     &  49.4 & 49.4 & 23.5 & 65M \\ %& 34s \\ %
+    \system{FT-Full}             & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8   & 53.8 & 32.5 & 20.2 & 65M \\ %& 34s \\
+    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2   & 53.3 & 38.4 & 20.5 & 65M/12M \\ %& 22s\\ 
+    \system{Res-Adap-HW}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91.0  & 53.1 & 36.6 & 20.2 & 65M/12M \\ %& 19s \\
+    \system{Res-Adap-HW-MT} & 37.4 & 56.4 & 52.1 & 33.7 & 44.8 & 89.8 & 52.4 & 27.1 & 20.4 & 65M/12M \\ %&  \\
+    \system{Res-Adap-HW-MT-ptr} & 37.7 & 57 & 52.5 & 33.5 & 46.1 & 89 & 52.6 & 46.5 & 21.4 & 65M/12M \\ %&  \\
+    \system{Res-Adap-Gate}  & 38.0 & 57.5& 53.0 & 33.5 & 46.0 & 90.1  & 53.0 & 49.0 & 22.5 & 65M/13M \\ %& 21s \\
 %    \system{FT-Full-UEW}      & 37.9 & 56.0 & 52.1 & 33.7 & 44.9 & 89.1 &	52.3 & 46.5 & 22.11 \\ %& 34s \\
 	\hline
+	%\system{Res-Adap-WD} & 37.18 & 55.99 & 52.93 & 33.36 & 46.03 & 90.65 & 52.7 & 38.73 & 20.69 & 65M/12M \\
+	%\system{Res-Adap-LR}      & 37.4 & 56.1 & 51.8 & 33.3 & 45.0 & 89.7 & 52.2 &  &  & 65M/12M \\  
+    %\hline
   \end{tabular}
-  \caption{Translation performance of highway and gated variants for En$\rightarrow$Fr.
+  \caption{Translation performance of highway and gated variants for En$\rightarrow$Fr.\revision{\domain{news} is excluded from training data and considered as out-of-domain test}.
     % Column \domain{PROC} is time (in seconds) of 100 iterations with batch size 12,288 tokens
@@
 It is thus no wonder that the design of multi-domain systems has been proposed for many tasks.
-In this short survey, we exclusively focus on machine translation; it is likely that similar methods (parameter sharing, instance selection / weighting, adversarial training, etc) have also been proposed for other tasks.
+In this short survey, we exclusively focus on machine translation; it is likely that similar methods (parameter sharing, instance selection/weighting, adversarial training, etc) have also been proposed for other tasks.
 
@@
 \citet{Chen16guided} apply domain control on the \emph{target} side, using a topic vector to describe the whole document context.
-Similar ideas are developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and similarities are enforced through parameter sharing schemes.
+Similar ideas are developed in \citet{Chu18multilingual,Pham19generic}, where domain differences and similarities are enforced through parameter sharing schemes.
 Parameter-sharing also lies at the core of the work by \citet{Jiang19multidomain}, who consider a Transformer model containing both domain-specific and domain-agnostic heads.
@@
 A contribution of this study is an adversarial training scheme to normalize representations across domains and make the combination of multiple data sources more effective.
-Similar techniques (parameter sharing, automatic domain classification / normalization) are at play in \cite{Zeng18multidomain}: in this work, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific from domain-agnostic representations.
+Similar techniques (parameter sharing, automatic domain classification/normalization) are at play in \citet{Zeng18multidomain}: in this work, the lower layers of the MT use auxiliary classification tasks to disentangle domain-specific from domain-agnostic representations.
 These representations are first processed separately, then merged to compute the final translation.
 
-\citet{Farajian17multidomain,li-etal-2018-one} are two recent representatives of the instance-based approach: for each test sentence, a small adaptation corpus is collected based on similarity measures, and used to fine-tune a mix-domain model.
+\citet{Farajian17multidomain,li-etal-2018-one} are two recent representatives of the instance-based approach: for each test sentence, a small adaptation corpus is collected based on similarity measures and used to fine-tune a mix-domain model.
 
-Finally note that a distinct evolution of the residual adapter model of \cite{Bapna19simple} is presented in \cite{Sharaf20metalearning} where meta-learning techniques are used to make fine tuning more effective in a standard domain-adaptation setting.
+Finally, note that a distinct evolution of the residual adapter model of \citet{Bapna19simple} is presented in \citet{Sharaf20metalearning} where meta-learning techniques are used to make fine-tuning more effective in a standard domain-adaptation setting.
 
@@
 \mpDone{discussion}
-In this paper, we have performed an experimental study of the residual adapter architecture in the context of multi-domain adaptation, where the goal is to build one single systems that (a) performs well for domain seen in training, ideally as well as full fine-tuning; (b) is also able to robustly handle translations for new, unseen domains.
+In this paper, we have performed an experimental study of the residual adapter architecture in the context of multi-domain adaptation, where the goal is to build one single system that (a) performs well for domain seen in training, ideally as well as full fine-tuning; (b) is also able to robustly handle translations for new, unseen domains.
 We have shown that this architecture allowed us to quickly adapt a model to a specific domain, delivering BLEU performance than are much better than the generic, mixed domain baseline, and close the gap with the full-finetuning approach, at a modest computational cost.
-Several new variants have been introduced and evaluated for two language directions: if none that able to clearly surpass the baseline residual adapter models, they provide directions for improving this model in practical settings: unbalanced data condition, noise in label domains, etc.
+Several new variants have been introduced and evaluated for two language directions: if none that able to clearly surpass the baseline, residual adapter models, they provide directions for improving this model in practical settings: unbalanced data condition, noise in label domains, etc.
 In our future work, we would like to continue the development of the gated variant, which, it seems to us, provides a flexible and robust tool to address the various challenges of multi-domain machine translation.

### 2020-10-06 2020/wmt_20/mdadapt.tex
reviewed till page 7

@@
 
-%\aclfinalcopy % Uncomment this line for the final submission
-%\def\aclpaperid{***} %  Enter the acl Paper ID here
+\aclfinalcopy % Uncomment this line for the final submission
+\def\aclpaperid{***} %  Enter the acl Paper ID here
 
@@
 
-\author{First Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{email@domain} \\\And
-  Second Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{email@domain} \\}
+\author{MinhQuang Pham$^{\dag\ddag}$, \ \ Josep Crego$^\dag$,\ \ Fran\c cois Yvon$^\ddag$, \ \ Jean Senellart$^\dag$ \\
+  $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France\\
+  {\tt firstname.lastname@systrangroup.com}\\
+  $^\ddag$Universit\'e Paris-Saclay, CNRS, LIMSI,  91405 Orsay, France\\
+  {\tt firstname.lastname@limsi.fr}}
 
@@
 
-Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pre-trained model by resuming training with the sole in-domain corpus.
+Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pre-trained model by resuming training with the sole in-domain corpus.
 It is a conjecture that the pretrained model constitutes a better initialization than a random one, especially when adaptation data is scarce.
-Indeed, studies of transfer learning for NMT such as \citet{artetxe20cross,aji20neural} have confirmed this claim in extensive experiments.
+Indeed, studies of transfer learning for NMT such as \citep{artetxe20cross,aji20neural} have confirmed this claim in extensive experiments.
 Full fine-tuning, that adapts all the parameters of a baseline model usually significantly improves the quality of the NMT for the chosen domain.
@@
 Therefore, a fully fine-tuned model is \emph{only useful to one target domain}.
-As the number of domains to handle grows, training, and maintaining a separate model for each task can quickly become tedious and resource-expensive .\fyDone{Fix this sentence.}
+As the number of domains to handle grows, training, and maintaining a separate model for each task can quickly become tedious and resource-expensive.\fyDone{Fix this sentence.}
 
-\citet{Vilar18learning,Bapna19simple,wuebker18compact,michel18extreme} have recently proposed a simple and lightweight domain adaptation method that also preserves the value of pre-trained models, based on small adapter components that are plugged in each hidden layer.
+Several recent studies (e.g.\ \citep{Vilar18learning,Wuebker18compact,Michel18extreme,Bapna19simple}) have proposed more lightweight schemes to perform domain adaptation, while also preserving  the value of pre-trained models.
+Our main inspiration is the latter work, whose proposal relies on small \emph{adapter components} that are plugged in each hidden layer.
 These adapters are trained only with the in-domain data, keeping the pre-trained model frozen.
-Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering a performance that is close to that of full fine-tuning.
+Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while delivering a performance that remains close to that of full fine-tuning.
 
@@
 \section{Residual adapters \label{sec:res}}
-In this section, we describe the basic version of the residual adapter architectures \citep{houlsby19parameter, Bapna19simple}, as well as two original variants of this model.
+In this section, we describe the basic version of the residual adapter architectures \citep{houlsby19parameter, Bapna19simple}, as well as two novel variants of this model.
 
@@
 Each layer contains two subparts, an attention layer, and a dense layer.
-Details vary from one implementation to another, we simply contend here that each layer $i \in \{1 \dots L\}$ (in the encoder or the decoder) computes a transform of a fixed-length sequence of $d$-dimensional input vectors $h^{i}$ into a sequence of output vectors $h^{i+1}$.
-
-The residual adapter architecture extends this model by adding one new computation module for each layer.
-The $i^{\text{th}}$ residual adapter thus modifies $h^i$ by computing the following transformations:\fyDone{Use align env} \fyTodo{Add layer normalization, replace notation b by a}
+Details vary from one implementation to another, we simply contend here that each layer $i \in \{1 \dots L\}$ (in the encoder or the decoder) computes a transform of a fixed-length sequence of $d$-dimensional input vectors $h^{i}$ into a sequence of output vectors $h^{i+1}$ as follows ($\mathbf{LN}$ denotes the (sub)layer normalization, $\mathbf{ReLU}$ is the ``rectified linear unit'' operator):\fyDone{Use align env}\fyDone{Add layer normalization, replace notation b by a}
 \begin{align*}
@@
 
-In the experiments of Section~\ref{sec:exp}, we use $2\times{}L=12$ residual adapters, one for each of the $L=6$ attention layers of the encoder and similarly for the decoder.\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and only serves one residual adapter.}
+In the experiments of Section~\ref{sec:exp}, we use $2\times{}L=12$ residual adapters, one for each of the $L=6$ attention layers of the encoder and similarly for the decoder.\footnote{In the decoder, the stack of self-attention and cross encoder-decoder attention only counts as one attention layer and only corresponds to one residual adapter.}
 
@@
 
-The first question concerns the number of adapter layers.
+The first question is the number of adapter layers.
 While in principle, all Transformer layers can be subject to adaptation, it is nonetheless worthwhile to consider simpler adaptation schemes, which would only alter a limited number of layers.
@@
 Likewise, it might be meaningful to explore ways to share subsets of adapters across domains.
-This in turn raises the issue of which layer(s) to adapt, a question that can be approached in the light of recent analyses of Transformers models, which conjecture that the higher layers encode global patterns with a more ``semantic'' interpretation, while the lower layers encode local patterns akin to morpho-syntactic information \cite{raganato18analysis}.
+This, in turn, raises the issue of which layer(s) to adapt, a question that can be approached in the light of recent analyses of Transformers models, which conjecture that the higher layers encode global patterns with a more ``semantic'' interpretation, while the lower layers encode local patterns akin to morpho-syntactic information \cite{raganato18analysis}.
 
@@
 In this situation, the forward step needs only to compute the hidden vectors $h^i$ once for the inner encoder layers, before an adapted sequence of vectors is computed at the topmost layer.
-Therefore, we can finetune the model to multiple domains at once without recomputing $h^i$.
+Therefore, we can fine-tune the model to multiple domains at once without recomputing $h^i$.
 This variant also opens the way to more parameter sharing across adapters, a perspective that we will not explore further in this work.
@@
 Therefore, we would like to design a version of residual adapters that is more robust to such domain errors.
-This variant, called the \emph{gated residual adapter model}, relies on the training of a supplementary trained component that will help decide whether to activate, on a word per word basis, a given residual layer and to regulate the strength of this activation.
+This variant, called the \emph{gated residual adapter model}, relies on the training of a supplementary component that will help decide whether to activate, on a word per word basis, a given residual layer and to regulate the strength of this activation.
 To this end, we extend the highway version of residual adapters as follows.
@@
 \item a generic model trained on a concatenation of all corpora, denoted \system{Mixed};\fyDone{Or mixed nat ?}
-\item a fine-tuned model \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing during 3 consecutive epochs.
+\item a fine-tuned model \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the development BLEU score stops increasing during 3 consecutive epochs.
 %  We again contrast two versions: full fine-tuning (\system{FT-Full}), which update all the parameters of the initial generic model \system{Mixed}; and the variants of \cite{Bapna19simple} (\system{FT-Block}).
@@
 Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2,048 cells.
-Residual adapters additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$.\revision{ \citet{Bapna19simple} showed that the performance of adapted model increases with respect to the size of the inner dimension.
-In the experiments of \citet{Bapna19simple}, we note that $b=1024$, which is 2-folds larger than the dimension of Transformer layer, achieved performance closed to fine-tuned model.
-Therefore, we would like to apply this setting to our experiments.}
+Residual adapters additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$.
+\revision{\citet{Bapna19simple} showed that the performance of adapted model increases with respect to the size of the inner dimension and obtained performance close to the full fine-tuned model with $b=1024$, which is twice as large as the dimension of a Transformer layer.
+We used the same setting in our experiments.}\fyDone{Check: is it $b$?}
+
 % The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
 % The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training uses a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4,000$), and a dropout rate of $0.1$ for all layers.\revision{ For \system{mixed} model, we use initial learning 1.0 and the concatenation of the validation sets of 6 domains as the validation set of \system{mixed}.
-In the fine-tuning experiments, we continue training using \system{mixed} as starting point, we use the same learning rate schedule, continuing the incrementation of the number of steps.
-In the multi-task training, we use the same learning rate schedule as \system{mixed}, for each iteration, we sample a domain from the pool of 6 domains with the probability proportional to the size of training data of each domain, then we sample a batch of~12,288 tokens from training data of the picked domain, the shared parameters and the parameters of the corresponding adapter are updated by gradients.}\fyDone{Describe the block adaptation layer - voir slides}
+Training uses a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4,000$), and a dropout rate of $0.1$ for all layers.
+\revision{For the \system{mixed} model, we use an initial learning rate of 1.0 and take the concatenation of the validation sets of 6 domains for development.
+In the fine-tuning experiments, we continue training using \system{mixed} as starting point, using the same learning rate schedule, and continuing the incrementation of the number of steps.
+In the multi-task training, we use the same learning rate schedule as for \system{mixed}: for each iteration, we sample a domain a probability proportional to its size; we then sample a batch of~12,288 tokens that is used to update the shared parameters and the parameters of the corresponding adapter.}\fyDone{Describe the block adaptation layer - voir slides}
 
-Models for En$\rightarrow$De are larger and rely on embeddings size as well as hidden layers size of~1024; each Transformers layer contains 16~attention heads; the inner feedforward layer contains 4,096 cells.
-Adapter modules have the same architecture as for the other language pair, except for their size, which is doubled ($b=2,048$).
+Models for En$\rightarrow$De are larger and rely on embeddings as well as hidden layers of size~1024; each Transformers layer contains 16~attention heads; the inner feedforward layer contains 4,096 cells.
+Adapter modules have the same architecture as for the other language pair, except for their size, which is doubled ($b=2,048$).%\fyTodo{Check: is it $b$?}
 %The number of parameters in each model is reported in Table~\ref{tab:params}.
@@
 The ``multi-task'' variant is slightly less effective in our experiments than the basic version, where optimization is performed in two steps.
-As it turns out, using residual adapters proves here more effective on average than the other reference multi-domain systems; it is also much better than the generic system for translating data from known domains, outperforming the \system{Mixed} system by more than 4 BLEU points in average.
+As it turns out, using residual adapters proves here better on average than the other reference multi-domain systems; it is also much better than the generic system for translating data from known domains, outperforming the \system{Mixed} system by more than 4 BLEU points in average.
 Gains are especially large for small domains such as \domain{law} and \domain{rel}.
@@
 \subsection{Varying the positions and number of residual adapters}
-Tables~\ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} report the BLEU scores for 6 domains in each language pair: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for En$\rightarrow$Fr; \domain{gov}, \domain{eco}, \domain{tour}, \domain{bank}, \domain{med} and \domain{news} for En$\rightarrow$De.
+Tables~\ref{tab:performance-en-fr-pos-reg}-\ref{tab:performance-en-de-pos-reg} report BLEU scores for 6 domains in each language pair: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for En$\rightarrow$Fr; \domain{gov}, \domain{eco}, \domain{tour}, \domain{bank}, \domain{med} and \domain{news} for En$\rightarrow$De.
 We first see that for the latter direction, the basic version \system{Res-Adap} also outperforms the \system{mixed} baseline on average, with large gains for the small domains \domain{tour}, \domain{bank} and comparable results for the other domains.
@@
    %    However, \system{FT-full} still remains the best model for all test domains.
-\fyFuture{Full FT for en->DE ? }
+\fyFuture{Full FT for en->DE ?}
 

### 2020-10-06 2020/wmt_20/mdadapt.tex
up to experiments

@@
 Training uses a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4,000$), and a dropout rate of $0.1$ for all layers.
-\revision{For the \system{mixed} model, we use an initial learning rate of 1.0 and take the concatenation of the validation sets of 6 domains for development.
-In the fine-tuning experiments, we continue training using \system{mixed} as starting point, using the same learning rate schedule, and continuing the incrementation of the number of steps.
-In the multi-task training, we use the same learning rate schedule as for \system{mixed}: for each iteration, we sample a domain a probability proportional to its size; we then sample a batch of~12,288 tokens that is used to update the shared parameters and the parameters of the corresponding adapter.}\fyDone{Describe the block adaptation layer - voir slides}
+\revision{For the \system{Mixed} model, we use an initial learning rate of 1.0 and take the concatenation of the validation sets of 6 domains for development.
+In the fine-tuning experiments, we continue training using \system{Mixed} as starting point, using the same learning rate schedule, and continuing the incrementation of the number of steps.
+In the multi-task training, we use the same learning rate schedule as for \system{Mixed}: for each iteration, we sample a domain a probability proportional to its size; we then sample a batch of~12,288 tokens that is used to update the shared parameters and the parameters of the corresponding adapter.}\fyDone{Describe the block adaptation layer - voir slides}
 
@@
   \centering
-  \begin{tabular}{|p{4.3cm}|*{8}{r|}} \hline
+  \begin{tabular}{|p{3.5cm}|*{8}{r|}} \hline
 %     &&&&&& \\
@@
     \system{Res-Adap-MT}  & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
-    \system{Res-Adap-MT-ptr\footnote{We initialized shared parameters by values of parameters of \system{mixed} model and train Res-Adap using multi-task learning}} & 37.47 & 57.06 & 52.39 & 33.72 & 46.17 & 89.47 & 52.7 \\
+    \system{Res-Adap-MT}$^{+}$ & 37.5 & 57.1 & 52.4 & 33.7 & 46.2 & 89.5 & 52.7 \\
     %  \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
@@
 The two variants of the residual adapter model included in this first round of experiment have been presented in Section~\ref{ssec:architecture}: \system{Res-Adap} is the multi-domain version of the approach of \citet{Bapna19simple} based on a two-step training procedure; while \system{Res-Adap-MT} is the ``multi-task'' version, where the parameters of the generic model and of the adapters are jointly learned from scratch.
+We also report results for the same system, using the the parameters of the \system{Mixed} model as initialization (\system{Res-Adap-MT}${^+}$).
 
@@
 With only one adapter active, we observe small, insignificant changes in performance when varying the adapter's depth.
-\revision{Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.}\fyFuture{More configurations ?}
+%\revision{Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.}\fyFuture{More configurations ?}
 
@@
 We report BLEU scores for each domain, as well as averages across languages.
-Column \domain{params} reports the number of domain-agnostic/domain-specific parameters \label{tab:performance-en-fr-pos-reg}} \fyDone{Boldface ?}
+Column \domain{params} reports the number of domain-agnostic/domain-specific parameters.\label{tab:performance-en-fr-pos-reg}} \fyDone{Boldface ?}
 \end{table*}
@@
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline
+    Model / Domain & \multicolumn{1}{c|}{\domain{ gov}} & \multicolumn{1}{c|}{\domain{ eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline
     \system{Mixed}          & 29.3 & 30.5 & 17.6 & 38.1 & 47.9 & 20.9  & 30.6 & 213M/0M\\
@@
 We report BLEU scores for each domain, as well as averages across domains.
-In column \domain{params}, we report the number of domain-agnostic/domain-specific parameters}
+Column \domain{params} reports the number of domain-agnostic/domain-specific parameters.}
   \label{tab:performance-en-de-pos-reg}
@@
 
-For each method, the optimal hyper-parameters (weight decay or layer regularization coefficient, see Section~\ref{sssec:design-space}) $\lambda$ by chosen by grid search among a small set of values ($\{ 10^{-3}, 10^{-4}, 10^{-5} \}$).\fyFuture{Better grid even better EWC.}
+For each method, the optimal hyper-parameter $\lambda$ (weight decay or layer regularization coefficient, see Section~\ref{sssec:design-space}) are chosen by grid search in a small set of values ($\{ 10^{-3}, 10^{-4}, 10^{-5} \}$).\fyFuture{Better grid even better EWC.}
 
-Results in Tables~\ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} show that regularizing the adapter model can positively affect the test performance for the smallest domains (this is especially clear for weight-decay (\system{Res-Adap-WD}) in En$\rightarrow$De), at the cost however of a small drop in performance for the other domains.
+Results in Tables~\ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} show that regularizing the adapter model can positively impact the test performance for the smallest domains (this is especially clear for weight-decay (\system{Res-Adap-WD}) in En$\rightarrow$De), at the cost however of a small drop in performance for the other domains.
 Using layer regularization proves here to be comparatively less effective.
@@
 To also evaluate the robustness with respect to out-of-domain examples, we perform two additional experiments.
-We first generate translations with erroneous (more precisely: randomly assigned) domain information: the corresponding results appear in Table~\ref{tab:performance-random} under the column \domain{rnd}.
+We first generate translations with erroneous (more precisely: randomly assigned) domain information: the corresponding results appear in Table~\ref{tab:performance-random} under column \domain{rnd}.
 We also compute translation for a domain unseen in training (\domain{news}) as follows.
@@
 
-A first observation is that for domains seen in training, our variants \system{Res-Adap-HW} and \system{Res-Adap-Gated} achieve BLEU scores that are on a par to those of the original version (\system{Res-Adap}), with insignificant variations across test sets.
-\fyFuture{Comment on processing time.}
+A first observation is that for domains seen in training, our variants \system{Res-Adap-HW} and \system{Res-Adap-Gated} achieve BLEU scores that are on a par to those of the original version (\system{Res-Adap}), with insignificant variations across test sets.\fyFuture{Comment on processing time.}
 
-The two other settings are instructive in several ways: they first clearly illustrate that the brittleness of domain-adapted systems, for which large drops in performance (more than 15 BLEU points on average) are observed when the domain label is randomly chosen.
+The two other settings are instructive in several ways: they first clearly illustrate the brittleness of domain-adapted systems, for which large drops in performance (more than 15 BLEU points on average) are observed when the domain label is randomly chosen.
 Our gated variant however proves much more robust than the other adaptation strategy and performs almost on par to the generic system for that test condition.
@@
     \system{Res-Adap-HW-MT} & 37.4 & 56.4 & 52.1 & 33.7 & 44.8 & 89.8 & 52.4 & 27.1 & 20.4 & 65M/12M \\ %&  \\
-    \system{Res-Adap-HW-MT-ptr} & 37.7 & 57 & 52.5 & 33.5 & 46.1 & 89 & 52.6 & 46.5 & 21.4 & 65M/12M \\ %&  \\
+    \system{Res-Adap-HW-MT}$^{+}$ & 37.7 & 57 & 52.5 & 33.5 & 46.1 & 89 & 52.6 & 46.5 & 21.4 & 65M/12M \\ %&  \\
     \system{Res-Adap-Gate}  & 38.0 & 57.5& 53.0 & 33.5 & 46.0 & 90.1  & 53.0 & 49.0 & 22.5 & 65M/13M \\ %& 21s \\
@@
 \end{table*}
-\fyFuture{Computation time should also report the time it takes for the complete tuning process, not one iteration which will be more or less the same}.
+\fyFuture{Computation time should also report the time it takes for the complete tuning process, not one iteration which will be more or less the same.}
 \section{Related Work \label{sec:related}}
@@
 \citet{Chen16guided} apply domain control on the \emph{target} side, using a topic vector to describe the whole document context.
-Similar ideas are developed in \citet{Chu18multilingual,Pham19generic}, where domain differences and similarities are enforced through parameter sharing schemes.
+Similar ideas are developed in \citep{Chu18multilingual,Pham19generic}, where domain differences and similarities are enforced through parameter sharing schemes.
 Parameter-sharing also lies at the core of the work by \citet{Jiang19multidomain}, who consider a Transformer model containing both domain-specific and domain-agnostic heads.
@@
 \citet{Farajian17multidomain,li-etal-2018-one} are two recent representatives of the instance-based approach: for each test sentence, a small adaptation corpus is collected based on similarity measures and used to fine-tune a mix-domain model.
+As shown in the former work, also adapting the training regime is crucial to make these techniques really effective.
 
-Finally, note that a distinct evolution of the residual adapter model of \citet{Bapna19simple} is presented in \citet{Sharaf20metalearning} where meta-learning techniques are used to make fine-tuning more effective in a standard domain-adaptation setting.
+Finally, note that a distinct evolution of the residual adapter model of \citet{Bapna19simple} is presented in \citep{Sharaf20metalearning} where meta-learning techniques are used to make fine-tuning more effective in a standard domain-adaptation setting.
 
@@
 \fyFuture{Example of gating ?}
-\fyTodo{Why MDMT does not work well - full train from generic? - add new results}
+\fyDone{Why MDMT does not work well - full train from generic? - add new results}
 \fyFuture{Structured domain? A Bayesian version?}

### 2020-10-06 2020/wmt_20/mdadapt.tex
Still a couple of decision to make

@@
 We notably explore ways to adjust and/or regularize adapter modules to handle situations where the adaptation data is very small.
-We also propose and contrast two new variants of the residual architecture: in the first one (\emph{highway residual adapters}), adaptation still affects each layer of the architecture, but its effect is delayed till the last layer, thus making the architecture more modular and adaptive; in the second one (\emph{gated residual adapters}), we explore ways to improve performance in the face of train-test data mismatch.
+We also propose and contrast two new variants of the residual architecture: in the first one (\emph{highway residual adapters}), adaptation still affects each layer of the architecture, but its effect is delayed till the last layer, thus making the architecture more modular and adaptive; our second variant (\emph{gated residual adapters}) exploits this modularity and enables us to explore ways to improve performance in the face of train-test data mismatch.
 We experiment with two language pairs and report results that illustrate the flexibility and effectiveness of these architectures.
@@
 We randomly select in each corpus a development and a test set of 1,000 lines each and keep the rest for training.
-Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings}
+Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings}\fyTodo{Answer question of reviewer 3 about splits ?}
    %    Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
@@
 Residual adapters additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$.
-\revision{\citet{Bapna19simple} showed that the performance of adapted model increases with respect to the size of the inner dimension and obtained performance close to the full fine-tuned model with $b=1024$, which is twice as large as the dimension of a Transformer layer.
+\revision{\citet{Bapna19simple} showed that the performance of adapted models increases with respect to the size of the inner dimension and obtained performance close to the full fine-tuned model with $b=1024$, which is twice as large as the dimension of a Transformer layer.
 We used the same setting in our experiments.}\fyDone{Check: is it $b$?}
@@
 \begin{itemize}
-\item a system using ``domain control'' \cite{Kobus17domaincontrol}.
+\item a system using ``domain control'' \citep{Kobus17domaincontrol}.
 In this approach, domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}) or in the form of a supplementary feature for each word (\system{DC-Feat});
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain-specific part (\system{LDR});
+\item a system using lexicalized domain representations \citep{Pham19generic}: word embeddings are composed of a generic and a domain-specific part (\system{LDR});
 \item the three proposals of \newcite{Britz17mixing}.
@@
     \system{Res-Adap-MT}$^{+}$ & 37.5 & 57.1 & 52.4 & 33.7 & 46.2 & 89.5 & 52.7 \\
-    %  \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
+    \hfill {Res-Adap-MT} (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 \\
     \hline
   \end{tabular}
-  \caption{Baseline translation performance of various multi-domain MT systems (En$\rightarrow$Fr).}
+  \caption{Translation performance of various multi-domain MT systems (En$\rightarrow$Fr) compared to variants of the residual adapter models.}
   \label{tab:performance-multi}
@@
 However, \system{MDL-Res} is better than other methods in multi-domain learning including:  \system{DC-Tag},  \system{DC-Feat}, \system{LDR}, \system{TTM}, \system{DM}, \system{ADM}.
+\revision{Our comparison of training regimes (\system{Res-Adap}  vs   \system{Res-Adap-MT} vs \system{Res-Adap-MT}$^+$) suggests that the simultaneous learning of all parameters is detrimental to performance in our settings, where the 2-step procedure implemented in \system{Res-Adap}) yield the best scores, even when initializing with good parameter values .
+This may be because in this setting, the adapters have access to a stable version of the generic system.
+The last line ({Res-Adap-MT} (gen)) gives the results for \system{Res-Adap-MT} trained system in which we cancel the adaptor in inference - comparing this to \system{Mixed} show how differently the generic parts of these two systems behave.}
 
@@
 
-By varying the number and position of residual adapters (see Section~\ref{ssec:architecture}), we then contrast several implementations of residual adapters.
+By varying the number and position of residual adapters (see Section~\ref{ssec:architecture}), we then contrast several implementations.
 \fyDone{Fix style here} Because the set of possible configurations is large, we only perform experiments for layers $i= 2, 4, 6$ (both for the encoder and decoder).
@@
   \caption{Translation performance of various fine-tuned systems (En$\rightarrow$Fr).
-We report BLEU scores for each domain, as well as averages across languages.
+We report BLEU scores for each domain, as well as averages across domains.
 Column \domain{params} reports the number of domain-agnostic/domain-specific parameters.\label{tab:performance-en-fr-pos-reg}} \fyDone{Boldface ?}
@@
   \centering
-  \begin{tabular}{|p{4cm}|*{8}{r|}|r|r|} \hline
+  \begin{tabular}{|p{3.5cm}|*{7}{r|}|r|r|} \hline
 %     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c||}{\domain{avg}} & \multicolumn{1}{c|}{\domain{rnd}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{\domain{PARAMS}}  \\ \hline % & \multicolumn{1}{c|}{\domain{PROC}} \\ \hline
-    \system{Mixed}             & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5     &  49.4 & 49.4 & 23.5 & 65M \\ %& 34s \\ %
-    \system{FT-Full}             & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8   & 53.8 & 32.5 & 20.2 & 65M \\ %& 34s \\
-    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2   & 53.3 & 38.4 & 20.5 & 65M/12M \\ %& 22s\\ 
-    \system{Res-Adap-HW}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91.0  & 53.1 & 36.6 & 20.2 & 65M/12M \\ %& 19s \\
-    \system{Res-Adap-HW-MT} & 37.4 & 56.4 & 52.1 & 33.7 & 44.8 & 89.8 & 52.4 & 27.1 & 20.4 & 65M/12M \\ %&  \\
-    \system{Res-Adap-HW-MT}$^{+}$ & 37.7 & 57 & 52.5 & 33.5 & 46.1 & 89 & 52.6 & 46.5 & 21.4 & 65M/12M \\ %&  \\
-    \system{Res-Adap-Gate}  & 38.0 & 57.5& 53.0 & 33.5 & 46.0 & 90.1  & 53.0 & 49.0 & 22.5 & 65M/13M \\ %& 21s \\
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c||}{\domain{avg}} & \multicolumn{1}{c|}{\domain{rnd}} & \multicolumn{1}{c|}{\domain{news}}
+\\ \hline %    & \multicolumn{1}{c|}{\domain{PARAMS}}  \\ \hline % & \multicolumn{1}{c|}{\domain{PROC}} \\ \hline  
+    \system{Mixed}             & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5     &  49.4 & 49.4 & 23.5 \\ %& 65M \\ %& 34s \\ %
+    \system{FT-Full}             & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8   & 53.8 & 32.5 & 20.2 \\ %& 65M \\ %& 34s \\
+    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2   & 53.3 & 38.4 & 20.5 \\ %& 65M/12M \\ %& 22s\\ 
+    \system{Res-Adap-HW}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91.0  & 53.1 & 36.6 & 20.2 \\ %& 65M/12M \\ %& 19s \\
+    \system{Res-Adap-HW-MT} & 37.4 & 56.4 & 52.1 & 33.7 & 44.8 & 89.8 & 52.4 & 27.1 & 20.4 \\ %& 65M/12M \\ %&  \\
+    \system{Res-Adap-HW-MT}$^{+}$ & 37.7 & 57.0 & 52.5 & 33.5 & 46.1 & 89.0 & 52.6 & 46.5 & 21.4 \\ %& 65M/12M \\ %&  \\
+    \system{Res-Adap-Gate}  & 38.0 & 57.5& 53.0 & 33.5 & 46.0 & 90.1  & 53.0 & 49.0 & 22.5 \\ %& 65M/13M \\ %& 21s \\
 %    \system{FT-Full-UEW}      & 37.9 & 56.0 & 52.1 & 33.7 & 44.9 & 89.1 &	52.3 & 46.5 & 22.11 \\ %& 34s \\
@@
   \end{tabular}
-  \caption{Translation performance of highway and gated variants for En$\rightarrow$Fr.\revision{\domain{news} is excluded from training data and considered as out-of-domain test}.
+  \caption{Translation performance of highway and gated variants for En$\rightarrow$Fr.
+\revision{\domain{news} is excluded from the training data and considered as an out-of-domain test}.\fyTodo{Comparison with a regularized version ?}
     % Column \domain{PROC} is time (in seconds) of 100 iterations with batch size 12,288 tokens
@@
 \fyFuture{Computation time should also report the time it takes for the complete tuning process, not one iteration which will be more or less the same.}
+
 \section{Related Work \label{sec:related}}
@@
 
-The former approach has also been adapted to NMT: \newcite{Kobus17domaincontrol,Tars18multidomain} use an additional domain feature in an RNN model, in the form of an extra domain-token or of additional domain-features associated with each word.
+The former approach has also been adapted to NMT: \citep{Kobus17domaincontrol,Tars18multidomain} use an additional domain feature in an RNN model, in the form of an extra domain-token or of additional domain-features associated with each word.
 \citet{Chen16guided} apply domain control on the \emph{target} side, using a topic vector to describe the whole document context.
@@
 \citet{Farajian17multidomain,li-etal-2018-one} are two recent representatives of the instance-based approach: for each test sentence, a small adaptation corpus is collected based on similarity measures and used to fine-tune a mix-domain model.
-As shown in the former work, also adapting the training regime is crucial to make these techniques really effective.
+As shown in the former work, also adapting the training regime on a per sentence basis is crucial to make these techniques really effective.
 
-Finally, note that a distinct evolution of the residual adapter model of \citet{Bapna19simple} is presented in \citep{Sharaf20metalearning} where meta-learning techniques are used to make fine-tuning more effective in a standard domain-adaptation setting.
+Finally, note that a distinct evolution of the residual adapter model of \citet{Bapna19simple} is presented in \citep{Sharaf20metalearning}, where meta-learning techniques are used to make fine-tuning more effective in a standard domain-adaptation setting.
 

### 2020-10-07 2020/tacl2020/final_multidomain.tex
final version
authors and acknowledgments

--- 2020-09-03 2020/tacl2020/multidomain.tex
@@
 \newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{red}{#1}}}
-\newcommand{\revision}[1]{\textcolor{red}{#1}}
+\newcommand{\revision}[1]{#1}
 \newcommand{\revisiondel}[1]{}
@@
 \fyFuture{A plea for more thoroughness in evaluation}
-\author{First Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain} \\\And
-  Second Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain} \\}
 
+\author{MinhQuang Pham$^{\dag\ddag}$,\ \  Josep Crego$^{\dag}$,\ \  Fran\c cois Yvon$^{\ddag}$, \ \ Jean Senellart$^{\dag}$ \\ \\
+  $^\ddag$Universit\'e Paris-Saclay, CNRS, LIMSI, 91400, Orsay, France \\
+  \texttt{firstname.lastname@limsi.fr} \\
+  $^\dag$SYSTRAN, 5 rue Feydeau, 75002 Paris, France \\
+  \texttt{firstname.lastname@systrangroup.com} \\
+}
 \date{}
@@
 % --------------
-% \section*{Acknowledgements}
+\section*{Acknowledgements}
+
+The work presented in this paper was partially supported by the European Commission under contract H2020-787061 ANITA.
 % This work was granted access to the HPC resources of [TGCC/CINES/IDRIS] under the allocation 20XX- [numéro de dossier] made by GENCI (Grand Equipement National de Calcul Intensif)
@@
 % \newpage\newpage
+
+\bibliographystyle{acl_natbib}
+\bibliography{multidomain}
+
+\clearpage{}
+
 \section*{Appendices}
@@
 Among the MD models, approaches using residual adapters have the best average performance.}
-
-\clearpage{}
-
 % \newpage\newpage \newpage
-
-\bibliographystyle{acl_natbib}
-\bibliography{multidomain}
 

### 2020-10-09 2020/wmt_20/mdadapt.tex
Fin des revisions

@@
 
-We randomly select in each corpus a development and a test set of 1,000 lines each and keep the rest for training.
-Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings}\fyTodo{Answer question of reviewer 3 about splits ?}
+We randomly select in each corpus a development and a test set of 1,000 lines each and keep the rest for training.\footnote{Scripts to replicate these experiments are available at url{}.}\fyTodo{github} Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings}\fyDone{Answer question of reviewer 3 about splits ?}
    %    Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
@@
 Residual adapters additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$.
-\revision{\citet{Bapna19simple} showed that the performance of adapted models increases with respect to the size of the inner dimension and obtained performance close to the full fine-tuned model with $b=1024$, which is twice as large as the dimension of a Transformer layer.
-We used the same setting in our experiments.}\fyDone{Check: is it $b$?}
+\citet{Bapna19simple} showed that the performance of adapted models increases with respect to the size of the inner dimension and obtained performance close to the full fine-tuned model with $b=1024$, which is twice as large as the dimension of a Transformer layer.
+We used the same setting in our experiments.\fyDone{Check: is it $b$?}
 
@@
 Training uses a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4,000$), and a dropout rate of $0.1$ for all layers.
-\revision{For the \system{Mixed} model, we use an initial learning rate of 1.0 and take the concatenation of the validation sets of 6 domains for development.
+For the \system{Mixed} model, we use an initial learning rate of $1.0$ and take the concatenation of the validation sets of 6~domains for development.
 In the fine-tuning experiments, we continue training using \system{Mixed} as starting point, using the same learning rate schedule, and continuing the incrementation of the number of steps.
-In the multi-task training, we use the same learning rate schedule as for \system{Mixed}: for each iteration, we sample a domain a probability proportional to its size; we then sample a batch of~12,288 tokens that is used to update the shared parameters and the parameters of the corresponding adapter.}\fyDone{Describe the block adaptation layer - voir slides}
+In the multi-task training, we use the same learning rate schedule as for \system{Mixed}: for each iteration, we sample a domain a probability proportional to its size; we then sample a batch of~12,288 tokens that is used to update the shared parameters and the parameters of the corresponding adapter.\fyDone{Describe the block adaptation layer - voir slides}
 
@@
 The two variants of the residual adapter model included in this first round of experiment have been presented in Section~\ref{ssec:architecture}: \system{Res-Adap} is the multi-domain version of the approach of \citet{Bapna19simple} based on a two-step training procedure; while \system{Res-Adap-MT} is the ``multi-task'' version, where the parameters of the generic model and of the adapters are jointly learned from scratch.
-We also report results for the same system, using the the parameters of the \system{Mixed} model as initialization (\system{Res-Adap-MT}${^+}$).
+We also report results for the same system, using the the parameters of the \system{Mixed} model as initialization (\system{Res-Adap-MT}${^+}$).\footnote{This system also includes a layer dropout policy that cancels adapter layers with probability $0.5$}
 
@@
 However, \system{MDL-Res} is better than other methods in multi-domain learning including:  \system{DC-Tag},  \system{DC-Feat}, \system{LDR}, \system{TTM}, \system{DM}, \system{ADM}.
-\revision{Our comparison of training regimes (\system{Res-Adap}  vs   \system{Res-Adap-MT} vs \system{Res-Adap-MT}$^+$) suggests that the simultaneous learning of all parameters is detrimental to performance in our settings, where the 2-step procedure implemented in \system{Res-Adap}) yield the best scores, even when initializing with good parameter values .
+Comparing training schemes (\system{Res-Adap} vs \system{Res-Adap-MT} vs \system{Res-Adap-MT}$^+$) suggests that the simultaneous learning of all parameters is detrimental to performance in our settings: we see that the 2-step procedure implemented in \system{Res-Adap} always yields the best scores, even when \system{Res-Adap-MT} is initialized with good parameter values .
 This may be because in this setting, the adapters have access to a stable version of the generic system.
-The last line ({Res-Adap-MT} (gen)) gives the results for \system{Res-Adap-MT} trained system in which we cancel the adaptor in inference - comparing this to \system{Mixed} show how differently the generic parts of these two systems behave.}
+The last line ({Res-Adap-MT} (gen)) gives the results for a \system{Res-Adap-MT} trained system in which we cancel the adapter in inference - comparing this to \system{Mixed} shows how differently the generic parts of these two systems behave.
 

### 2020-10-13 2020/tacl2020/final_multidomain.tex
BANK 1,2 -> LAW 1,2 table 4

@@
      Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{law} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} &  \multicolumn{1}{c|}{\domain{rnd}} &  \multicolumn{1}{c|}{\domain{new}}\\ \hline
-    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_2$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
+    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_2$ &  \domain{law}$_1$ & \domain{law}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
     \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5} & -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& \SW{-19.6} & \SW{-3.3}\\% [32.5]
@@
 The work presented in this paper was partially supported by the European Commission under contract H2020-787061 ANITA.
-% This work was granted access to the HPC resources of [TGCC/CINES/IDRIS] under the allocation 20XX- [numéro de dossier] made by GENCI (Grand Equipement National de Calcul Intensif)
+
+This work was granted access to the HPC resources of [TGCC/CINES/IDRIS] under the allocation 2020- [AD011011270] made by GENCI (Grand Equipement National de Calcul Intensif)
 % The acknowledgements should go immediately before the references.  Do

### 2020-10-13 2020/wmt_20/mdadapt.tex

@@
 It is a conjecture that the pretrained model constitutes a better initialization than a random one, especially when adaptation data is scarce.
-Indeed, studies of transfer learning for NMT such as \citep{artetxe20cross,aji20neural} have confirmed this claim in extensive experiments.
+Indeed, studies of transfer learning for NMT such as \citet{artetxe20cross,aji20neural} have confirmed this claim in extensive experiments.
 Full fine-tuning, that adapts all the parameters of a baseline model usually significantly improves the quality of the NMT for the chosen domain.
@@
 \end{align*}
-Overall, the  $i^{\text{th}}$ adapter is thus parameterized by matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, with $b$ the dimension of the adapter \fyTodo{($d \gg b$) not true}\fyDone{Check this}.
+Overall, the  $i^{\text{th}}$ adapter is thus parameterized by matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, with $b$ the dimension of the adapter \fyDone{($d \gg b$) not true}\fyDone{Check this}.
 For the sake of brevity, we will simply denote $h^{i}_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the corresponding set of parameters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}
@@
 
-We randomly select in each corpus a development and a test set of 1,000 lines each and keep the rest for training.\footnote{Scripts to replicate these experiments are available at url{}.}\fyTodo{github} Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings}\fyDone{Answer question of reviewer 3 about splits ?}
+We randomly select in each corpus a development and a test set of 1,000 lines each and keep the rest for training.\footnote{Scripts to replicate these experiments are available at url{https://github.com/qmpham/experiments.git}.}\fyDone{github} Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings}\fyDone{Answer question of reviewer 3 about splits ?}
    %    Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
@@
   \caption{Translation performance of highway and gated variants for En$\rightarrow$Fr.
-\revision{\domain{news} is excluded from the training data and considered as an out-of-domain test}.\fyTodo{Comparison with a regularized version ?}
+\domain{news} is excluded from the training data and considered as an out-of-domain test.\fyDone{Comparison with a regularized version ?}
     % Column \domain{PROC} is time (in seconds) of 100 iterations with batch size 12,288 tokens
@@
 
-Early approaches to multi-domain MT were proposed for statistical MT, either considering multiple data sources (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}) or domains containing several topics \cite{Eidelman12topic,Hasler14dynamic-topic}.
+Early approaches to multi-domain MT were proposed for statistical MT, either considering multiple data sources (eg.\ \citet{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}) or domains containing several topics \cite{Eidelman12topic,Hasler14dynamic-topic}.
 Two main strategies emerge: feature-based methods, where domain labels are integrated through supplementary features; and instance-based methods, involving a measure of similarity between train and test domains.
 
-The former approach has also been adapted to NMT: \citep{Kobus17domaincontrol,Tars18multidomain} use an additional domain feature in an RNN model, in the form of an extra domain-token or of additional domain-features associated with each word.
+The former approach has also been adapted to NMT: \citet{Kobus17domaincontrol,Tars18multidomain} use an additional domain feature in an RNN model, in the form of an extra domain-token or of additional domain-features associated with each word.
 \citet{Chen16guided} apply domain control on the \emph{target} side, using a topic vector to describe the whole document context.
-Similar ideas are developed in \citep{Chu18multilingual,Pham19generic}, where domain differences and similarities are enforced through parameter sharing schemes.
+Similar ideas are developed in \citet{Chu18multilingual,Pham19generic}, where domain differences and similarities are enforced through parameter sharing schemes.
 Parameter-sharing also lies at the core of the work by \citet{Jiang19multidomain}, who consider a Transformer model containing both domain-specific and domain-agnostic heads.
@@
 
-Finally, note that a distinct evolution of the residual adapter model of \citet{Bapna19simple} is presented in \citep{Sharaf20metalearning}, where meta-learning techniques are used to make fine-tuning more effective in a standard domain-adaptation setting.
+Finally, note that a distinct evolution of the residual adapter model of \citet{Bapna19simple} is presented in \citet{Sharaf20metalearning}, where meta-learning techniques are used to make fine-tuning more effective in a standard domain-adaptation setting.
 
@@
 The extended version of the highway residual adapter to gated residual adapter shows the benefit in the robustness against wrong domain prediction and out-of-domain examples.
+
+\section*{Acknowledgements}
+This work was granted access to the HPC resources of [TGCC/CINES/IDRIS] under the allocation 2020- [AD011011270] made by GENCI (Grand Equipement National de Calcul Intensif)
 

### 2020-10-16 2020/tacl2020/multidomain.tex

@@
 \fyFuture{A plea for more thoroughness in evaluation}
+\author{MinhQuang Pham$^{\dag\ddag}$, \ \ Josep Crego$^\dag$,\ \ Fran\c cois Yvon$^\ddag$, \ \ Jean Senellart$^\dag$ \\
+  $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France\\
+  {\tt firstname.lastname@systrangroup.com}\\
+  $^\ddag$Universit\'e Paris-Saclay, CNRS, LIMSI,  91405 Orsay, France\\
+  {\tt firstname.lastname@limsi.fr}}
+
 \author{First Author \\

### 2020-10-19 2021/mdmt/mdmt.tex
mdmt first commit

\documentclass[12pt,times,a4paper,twoside]{article}

\usepackage[utf8]{inputenc}	% Para caracteres en español
\usepackage[left=3cm,right=3cm,top=2cm,bottom=3cm]{geometry}
\usepackage{pagenote}

\usepackage{algorithm}
\usepackage{algorithmic}

\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{mathtools,xparse}
\usepackage{mathrsfs}

\usepackage{multirow,booktabs}
\usepackage[table]{xcolor}
\usepackage{fullpage}
\usepackage{lastpage}
\usepackage{enumitem}
\usepackage{fancyhdr}

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{calc}
% \usepackage{color,soul}
\usepackage{multicol}
\usepackage{cancel}
\usepackage[retainorgcmds]{IEEEtrantools}
% \usepackage{xcolor} -- implied by xcolor

\usepackage{listings}
\usepackage{spverbatim}
\usepackage{fancyvrb}
\usepackage{hyperref}
\usepackage{float} 
\usepackage{natbib}

\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}

\usepackage{soulutf8}
\usepackage{tabularx}

\usepackage[draft]{todo}
\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}

\colorlet{shadecolor}{orange!15}
\parindent 0in
\parskip 12pt
\geometry{margin=1in, headsep=0.25in}
\setlength{\belowdisplayskip}{8pt} \setlength{\belowdisplayshortskip}{8pt}
\setlength{\abovedisplayskip}{8pt} \setlength{\abovedisplayshortskip}{8pt}
\setlist{nosep}
\setlength{\parskip}{0.1cm}
\setlength{\parindent}{1em}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{definition}{Definition}[section]
\DeclareMathOperator*{\argmax}{argmax}
\newcommand{\R}{\ensuremath{\mathbb{R}}}

\pgfkeys{
    /tr/rowfilter/.style 2 args={
        /pgfplots/x filter/.append code={
            \edef\arga{\thisrow{#1}}
            \edef\argb{#2}
            \ifx\arga\argb
            \else
                \def\pgfmathresult{}
            \fi
        }
    }
}

\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\NewDocumentCommand{\normL}{ s O{} m }{%
  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{L_1}%
}

\title{Multi-task training of multi-domain NMT}
\author{}
\date{}

\begin{document}

\maketitle
\begin{center}
{\LARGE \bf}\fyDone{Make a proper title}
\end{center}
\cite{Caruana97multitask} proposes a multi-task model with large shared stack of layers followed by multiple independent task-specific layers. The training is to simply average the gradients of separate learning objectives. The author admitted that the achieved optimal model is outperformed by far by different "snap shot" taken along the training by applying early stopping independently to each individual task. We might wonder whether this simple strategy can be improved.

Let $\theta_s, \{\theta_i\}_{i=[1..T]}$ be the parameters of our MD-MT model and the data generation process be 
\begin{align*}
Y_k^i | X_k^i : P_{\theta_s, \theta_i}(Y_k^i | X_k^i) \\
X_k^i : D^i(X_k)
\end{align*}

The objective of the training is $$\displaystyle{\mathop{\min \sum_{i=1}^{T}\sum_{X_k^i, Y_k^i} -log(P_{\theta_s,\theta_i}(Y_k^i | X_k^i))}}$$

The loss is a sum of prediction losses over all the domains of interest. However, the gradient is calculated partially over randomly picked batch from one of in-domain corpora. The gradient along the curve of a member loss may be harmful to other member losses of the total multi-domain loss. In order to prevent the bad influence of the partial gradient, we could regularize it by a constraint on the shifting between probabilities $P(\theta_s, \theta_i)$ before and after update.

In the optimization by gradient descent, at each iteration, the parameter is updated by gradient of loss function at current point in the space of hypothesis.

\begin{equation}
\theta^{t+1} = \theta^t - \mu * \frac{\delta L}{\delta \theta}(\theta_t)
\end{equation}

We can interpret this in an other point of view in which at each iteration, we search an displacement, whose magnitude is limited, following which the loss decreases the most.
\begin{align}\label{eq}
	\displaystyle{\mathop{\argmax_{\Delta \theta}} L(\theta^t + \mu \Delta \theta)} \\
	\parallel \Delta \theta \parallel \leq 1  \nonumber
\end{align}
This optimization problem results in $-\frac{\delta L}{\delta \theta}(\theta_t)$ as solution.

Using this approach, we replaces the constraint on magnitude of the displacement by a constraint on the magnitude of the distribution shifting over domains.
\begin{align}
	\displaystyle{\mathop{\argmax_{\Delta \theta}} L_i(\theta_s^t + \mu \Delta \theta_s, \theta_i^t + \mu \Delta \theta_i)} \\
	\mathop{\sum_{j=1}^{T}} E_{x_j}[KL(P_{\theta^t_s, \theta^t_j}(y_j|x_j)|P_{\theta^t_s + \mu \Delta \theta_s, \theta^t_j + \mu \delta_{i}(j) \Delta \theta_i}(y_j|x_j))] \leq c \nonumber
\end{align}

For a small displacement, the KL distance can be approximated by Fisher information matrix.
\begin{align}
KL(P_{\theta^t_s, \theta^t_j}(y_j|x_j)|P_{\theta^t_s - \Delta \theta_s, \theta^t_j}(y_j|x_j)) \approx <\Delta \theta_s , I_{\theta^t_s, \theta^t_j}^{\theta_s} \Delta \theta_s > \\
KL(P_{\theta^t_s, \theta^t_i}(y_i|x_i)|P_{\theta^t_s - \Delta \theta_s, \theta^t_i - \Delta \theta_i}(y_i|x_i)) \approx <\Delta \theta , I_{\theta^t_s, \theta^t_i}^{\theta} \Delta \theta >
\end{align}
Where $I_{\theta^t_s, \theta^t_j}^{\theta_s} = E_{y:P_{\theta^t_s, \theta^t_j}(y|x)}[\frac{\delta P_{\theta_s, \theta_j}(y|x)}{\delta \theta_s} \times \frac{\delta P_{\theta_s, \theta_j}(y|x)}{\delta \theta_s}^T]$; $\theta = (\theta_s, \theta_i)$. As in \citet{Thompson19overcoming} and \citet{Kirkpatrick17overcoming}, matrix Fisher is usually approximated by its diagonal, therefore we could use following approximation $<\Delta \theta , I_{\theta^t_s, \theta^t_i}^{\theta} \Delta \theta > \approx <\Delta \theta_s , I_{\theta^t_s, \theta^t_i}^{\theta_s} \Delta \theta_s > + <\Delta \theta_i , I_{\theta^t_s, \theta^t_i}^{\theta_i} \Delta \theta_i >$

Therefore, we could rewrite the constraint as
\begin{align}
\mathop{\sum_{j=1}^{T}}  E_{x_j}[<\Delta \theta_s , I_{\theta^t_s, \theta^t_j}^{\theta_s} \Delta \theta_s >] + E_{x_i}[<\Delta \theta_i , I_{\theta^t_s, \theta^t_i}^{\theta_i} \Delta \theta_i >] \leq  c \\
\Longleftrightarrow <\Delta \theta_s , \mathop{\sum_{j=1}^{T}}  E_{x_j}[I_{\theta^t_s, \theta^t_j}^{\theta_s}] \Delta \theta_s > + <\Delta \theta_i , E_{x_i}[I_{\theta^t_s, \theta^t_i}^{\theta_i}] \Delta \theta_i > \leq c 
\end{align}

Similar to equation \ref{eq}, we obtain a solution as follow 
\begin{align}
\Delta \theta_s = -(\mathop{\sum_{j=1}^{T}}  E_{x_j}[I_{\theta^t_s, \theta^t_j}^{\theta_s}])^{-1}\frac{\delta L_i}{\delta \theta_s} \\
\Delta \theta_i = -(E_{x_i}[I_{\theta^t_s, \theta^t_i}^{\theta_i}])^{-1}\frac{\delta L_i}{\delta \theta_i}
\end{align}

\bibliographystyle{acl_natbib}
\bibliography{multidomain}
\end{document}

### 2020-10-23 2020/tacl2020/final_multidomain.tex
added magic accepted option
table 5
Elise corrections

@@
 In this setting, \emph{domain adaptation} (DA) methods are in order.
-DA has a long history in Machine Learning in general (eg.\ \cite{Shimodaira00improving,Ben-David09atheory,Quinonero08dataset,Pan10asurvey})\fyDone{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}).
+DA has a long history in Machine Learning in general (e.g.\ \cite{Shimodaira00improving,Ben-David09atheory,Quinonero08dataset,Pan10asurvey})\fyDone{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}).
 Various techniques thus exist to handle both the situations where a (small) training sample drawn from $\mathcal{D}_t$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
@@
 A related question concerns the relationship between supervised domain adaptation and multi-domain translation.
-The latter task seems more challenging \revision{as it tries to optimize MT performance for a more diverse set of potential inputs, with the additional uncertainty regarding the distribution of test data.} Are there still situations where where MD systems can surpass single domain adaptation, as is sometimes expected?
+The latter task seems more challenging \revision{as it tries to optimize MT performance for a more diverse set of potential inputs, with an additional uncertainty regarding the distribution of test data.} Are there still situations where MD systems can surpass single domain adaptation, as is sometimes expected?
 
@@
 Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT systems and associated measurements to evaluate them (Section~\ref{sec:challenging}).
-In doing so, we also shed lights on new problems that arise in this context, regarding for instance the accommodation of new domains in the course of training, or the computation of automatic domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
-We show in Section~\ref{sec:experiments} that existing approaches still fall short to match many of these requirements, notably with respect to handling a large amount of heterogeneous domains and to dynamically integrating new domains in training.\fyDone{Spell out conclusions}
+In doing so, we also shed light on new problems that arise in this context, regarding for instance the accommodation of new domains in the course of training, or the computation of automatic domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
+We show in Section~\ref{sec:experiments} that existing approaches still fall short to match many of these requirements, notably with respect to the handling of a large amount of heterogeneous domains and to dynamically integrating new domains in training.\fyDone{Spell out conclusions}
 
@@
 A typical learning scenario in MT is to have access to samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$\revision{, with $\{\lambda^{s}_d, d=1 \dots n_d\}$ the corresponding mixture weights ($\sum_d \lambda^{s}_d=1$)}.
-Multi-domain learning, as defined in \cite{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ of several domains, making the problem distinct from mere domain adaption.
+Multi-domain learning, as defined in \citet{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ of several domains, making the problem distinct from mere domain adaption.
 A multi-domain learner is then expected to use these tags effectively \cite{Joshi12multidomain} when computing the combined translation function $h(x,d)$, and to perform well in all domains \cite{Finkel09hierarchical}.
@@
 A second line of reasoning rests on linguistic properties of the translation function and contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and tend to remain semantically stable across domains, motivating some cross-domain parameter sharing.
-A MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the translation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
+An MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the translation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
 It is here expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related and can share more information.
@@
 These arguments are not mutually exclusive; however each yields specific expectations with respect to the performance of this approach, and should also yield appropriate evaluation procedure.
-If the motivation is primarily computational, then a drop in MT quality with respect to multiple individual domain might be acceptable if compensated by the computational savings.
+If the motivation is primarily computational, then a drop in MT quality with respect to multiple individual domains might be acceptable if compensated by the computational savings.
 If it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some under-resourced domains, over individually trained systems.
@@
 A related situation occurs when the domain of a test document is unknown.
-Several situations need be considered: for domains seen in training, then using automatically predicted domain labels should not be much worse than using the correct one.
+Several situations need be considered: for domains seen in training, using automatically predicted domain labels should not be much worse than using the correct one.
 For test documents from unknown domains (zero-shot transfer), a good MD system should ideally outperform the default baseline that merges all available data.\fyDone{Systems react to unknown domains}
@@
 \paragraph{[P6-DYN]}
-Another requirement, more of an operational nature, is that a MDMT system should smoothly evolve to handle a growing number of domains, without having to retrain the full system each time new data is available.
+Another requirement, more of an operational nature, is that an MDMT system should smoothly evolve to handle a growing number of domains, without having to retrain the full system each time new data is available.
 This is a requirement [P6-DYN] that we challenge by dynamically changing the number of training and test domains.
@@
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
-Most corpora are variable from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
+Most corpora are available from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
 To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
@@
 \revision{Among systems using an extra domain feature, \system{DC-Tag} has a small edge over \system{DC-Feat} and also requires less parameters; it also outperforms \system{TTM}, which however uses predicted rather than gold domain tags.
-\system{TTM} is also the best choice among the systems that to not use domain tags in inference.} \revision{The best contenders overall are  \system{FT-Res} and \system{MDL-Res}, which significantly improve over \system{Mixed-Nat} for a majority of domains, and are the only ones to clearly fulfill [P1-LAB];} \system{WDCMT} also improves on three domains, but regresses on one.
+\system{TTM} is also the best choice among the systems that do not use domain tags in inference.} \revision{The best contenders overall are  \system{FT-Res} and \system{MDL-Res}, which significantly improve over \system{Mixed-Nat} for a majority of domains, and are the only ones to clearly fulfill [P1-LAB];} \system{WDCMT} also improves on three domains, but regresses on one.
 The use of a dedicated adaptation module thus seems better than feature-based strategies, \revision{but yields a large increase of the number of parameters.} The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}).
@@
 In first three, we randomly \emph{split} one corpus in two parts and proceed as if this corresponded to two actual domains.
-A MD system should detect that these two pseudo-domains are mutually beneficial and should be hardly affected by this change with respect to the baseline scenario (no split).
+A MD system should detect that these two pseudo-domains are mutually beneficial and should hardly be affected by this change with respect to the baseline scenario (no split).
 In this situation, we expect MDMT to even surpass fine-tuning separately on each of these dummy domains, as MDMT exploits all data, while fine-tuning focuses only on a subpart.
@@
 Our findings can be summarized as follows.
-For the split experiments, we see small variations that can be positive or negative compared to the baseline situation, but these are hardly significative.
+For the split experiments, we see small variations that can be positive or negative compared to the baseline situation, but these are hardly significant.
 All systems show some robustness with respect to fuzzy domain boundaries; this is mostly notable for \system{ADM}, suggesting that when domain are close, ignoring domain differences is effective.
@@
 
-In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the domain tag and use it for decoding.\footnote{\revision{Domain tags are assigned as follows: we train a language model for each domain and assign tag on a per-sentence basis based the language model log-probability (assuming uniform domain priors).
+In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the domain tag and use it for decoding.\footnote{\revision{Domain tags are assigned as follows: we train a language model for each domain and assign tag on a per-sentence basis based on the language model log-probability (assuming uniform domain priors).
 The domain classifier has an average prediction error of 16.4\% for in-domain data.}}
@@
 This makes a difference between domain-bounded systems, for which the number of domains is limited and truly open-domain systems.% , which seamlessly integrate additional domains.
+
+\begin{figure*}[h!]
+    \begin{center}
+        \input{table5.pgf}
+    \end{center}
+    \caption{Ability to handle a new domain.
+\revision{We report BLEU scores for a complete training session with 7 domains, as well as differences with (bleu) training with 6 domains (from Table~\ref{tab:performance}); (red) continuous training mode}.}
+\end{figure*}
 
@@
 % Adding is not helping here, especially when domains are well separated.
-Comparing now the coldstart with the warmstart scenario, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training also negatively impact the performance for other domains.
+Comparing now the coldstart with the warmstart scenario, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training also negatively impacts the performance for other domains.
 This happens notably for \system{DC-Tag}, \system{TTM} and \system{ADM}.
-In this setting \system{MDL Res} and \system{DM} show the smaller average loss, with the former achieving the best balance of training cost and average BLEU score.% of all systems.
+In this setting \system{MDL-Res} and \system{DM} show the smaller average loss, with the former achieving the best balance of training cost and average BLEU score.% of all systems.
 \fyDone{Numbers for MDL Res need significancy testing; and they are weird - why do we see this ? Because the generic is bad ?}
@@
   \caption{Translation performance with automatic domains, computed with the original test sets.
-Significancy tests are for comparisons with the 6-domain scenario \revision{(Table~\ref{tab:performance}).}}
+Significance tests are for comparisons with the 6-domain scenario \revision{(Table~\ref{tab:performance}).}}
   \label{tab:subdomains}
@@
 
-The latter strategy has been widely used in NMT: \newcite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token or in the form of additional domain-feature associated to each word.
+The latter strategy has been widely used in NMT: \newcite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token or in the form of an additional domain-feature associated to each word.
 These results are reproduced by \newcite{Tars18multidomain}, who also consider automatically induced domain tags.
@@
 
-Another parameter-sharing scheme is in \cite{Jiang19multidomain}, which augments a Transformer model with domain-specific heads, whose contributions are regulated at the word/position level: some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling that the core of \cite{Huck15mixeddomain,Saunders19ucam}.
+Another parameter-sharing scheme is in \cite{Jiang19multidomain}, which augments a Transformer model with domain-specific heads, whose contributions are regulated at the word/position level: some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling at the core of \cite{Huck15mixeddomain,Saunders19ucam}.
 The results for three language pairs outperform several standard baselines for a two-domain systems (in fr:en and de:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
@@
 \begin{itemize}
-\item \system{Mixed-Nat}, \system{FT-full}, \system{TTM}, \system{DC-Tag} use the a medium Transformer model of \cite{Vaswani17attention} with the following settings: embeddings size and hidden layers size are set to~$512$.
+\item \system{Mixed-Nat}, \system{FT-full}, \system{TTM}, \system{DC-Tag} use a medium Transformer model of \cite{Vaswani17attention} with the following settings: embeddings size and hidden layers size are set to~$512$.
 Multi-head attention comprises 8 heads in each of the 6 layers; the inner feedforward layer contains~$2048$ cells.
@@
 \item \system{LDR} uses medium Transformer model and for each token we introduce a \system{LDR} feature of size $4 \times domain\_num$.
-Given a sentence of domain $i\in[1,..,K]$ in the training batch, for each token of the sentence, the \system{LDR} units of the indexes outside of the range $[4(i-1),..,4i-1]$ are masked to $0$, the masked LDR feature will be concatenated to the embedding of the token.
+Given a sentence of domain $i\in[1,..,K]$ in the training batch, for each token of the sentence, the \system{LDR} units of the indexes outside of the range $[4(i-1),..,4i-1]$ are masked to $0$, and the masked LDR feature will be concatenated to the embedding of the token.
 Details are in \cite{Pham19generic}.
 \item \system{Mixed-Nat-RNN} uses one bidirectional LSTM layer in the encoder and one LSTM layer in the decoder. The size of hidden layers is $1024$, the size of word embeddings is $512$. 
-\item \system{WDCNMT} uses one bidirectional GRU layer in the encoder and one GRU-conditional layer in the decoder. The size of hidden layers is $1024$, the size of word embedding is $512$.
+\item \system{WDCNMT} uses one bidirectional GRU layer in the encoder and one GRU-conditional layer in the decoder. The size of hidden layers is $1024$, the size of word embeddings is $512$.
 \end{itemize}
-\paragraph{Training} For each domain, we create train/dev/test sets by randomly each corpus. We maintain the size of validation sets and of test sets equal to 1,000 lines for every domain.
+\paragraph{Training} For each domain, we create train/dev/test sets by randomly splitting each corpus. We maintain the size of validation sets and of test sets equal to 1,000 lines for every domain.
 The learning rate is set
@@
 Cluster size vary from approximately 8k sentences (cluster~24) up to more than 350k sentences.
-More than 2/3 of these clusters comprise mostly texts from one single domain, as for cluster 12 which is predominantly \domain{med}, the remaining clusters typically mix 2 domains.
-Fine-tuning with small domains is often outperformed by other MDMT technique, an issue that a better regularization strategy might mitigate.
+More than 2/3 of these clusters mostly comprise texts from one single domain, as for cluster 12 which is predominantly \domain{med}, the remaining clusters typically mix 2 domains.
+Fine-tuning with small domains is often outperformed by other MDMT techniques, an issue that a better regularization strategy might mitigate.
 Domain-control (\system{DC-Feat}) is very effective for small domains, but again less so in larger data conditions.
@@
 
-Multi-domain learning, as defined in \cite{Dredze08online} additionally makes the requirement that categorial domain tags are available in training and testing; a multi-domain learner should use them effectively \cite{Joshi12multidomain} to improve simultaneously on all domains \cite{Finkel09hierarchical}.
+Multi-domain learning, as defined in \citet{Dredze08online} additionally makes the requirement that categorial domain tags are available in training and testing; a multi-domain learner should use them effectively \cite{Joshi12multidomain} to improve simultaneously on all domains \cite{Finkel09hierarchical}.
 This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyDone{explain zero shot}.

### 2020-10-23 2020/tacl2020/FinalVersion/main-2327-PhamMinhQuang.tex
Final version

--- 2020-10-23 2020/tacl2020/final_multidomain.tex
@@
 
-\documentclass[11pt,a4paper]{article}
+\documentclass[11pt,a4paper,acceptedWithA,copyedit]{article}
 \usepackage{tacl2018v2}
@@
 % in the camera-ready version and ask you to change it back.
-\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
 \newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
@@
 \newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{red}{#1}}}
-\newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{red}{#1}}}
+
 \newcommand{\revision}[1]{#1}
@@
 \usepackage{multirow}
+
 \title{Revisiting Multi-Domain Machine Translation}
-\fyFuture{A plea for more thoroughness in evaluation}
 
-\author{MinhQuang Pham$^{\dag\ddag}$,\ \  Josep Crego$^{\dag}$,\ \  Fran\c cois Yvon$^{\ddag}$, \ \ Jean Senellart$^{\dag}$ \\ \\
+\author{MinhQuang Pham$^{\dag\ddag}$,\ \  Josep Maria Crego$^{\dag}$,\ \  Fran\c cois Yvon$^{\ddag}$ \\ \\
   $^\ddag$Universit\'e Paris-Saclay, CNRS, LIMSI, 91400, Orsay, France \\
@@
 Our experiments with a large sample of multi-domain systems show that most of these expectations are hardly met and suggest that further work is needed to better analyze the current behaviour of multi-domain systems and to make them fully hold their promises.
-  \fyDone{Write abstract}
 \end{abstract}
@@
 
-%
-% The following footnote without marker is needed for the camera-ready
-% version of the paper.
-% Comment out the instructions (first text) and uncomment the 8 lines
-% under "final paper" for your variant of English.
-% 
-% \footnote{
-%     %
-%     % for review submission
-%     %
-%     \hspace{-0.65cm}  % space normally used by the marker
-%     Place licence statement here for the camera-ready version. See
-%     Section~\ref{licence} of the instructions for preparing a
-%     manuscript.
-%     %
-%     % % final paper: en-uk version 
-%     %
-%     % \hspace{-0.65cm}  % space normally used by the marker
-%     % This work is licensed under a Creative Commons 
-%     % Attribution 4.0 International Licence.
-%     % Licence details:
-%     % \url{http://creativecommons.org/licenses/by/4.0/}.
-%     % 
-%     % % final paper: en-us version 
-%     %
-%     % \hspace{-0.65cm}  % space normally used by the marker
-%     % This work is licensed under a Creative Commons 
-%     % Attribution 4.0 International License.
-%     % License details:
-%     % \url{http://creativecommons.org/licenses/by/4.0/}.
-% }
-
 Data-based Machine Translation (MT), whether statistical or neural, rests on well understood machine learning principles.
-Given a training sample of matched source-target sentence pairs $(\src,\trg)$\jcDone{shoulnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}_s$, a model parameterized by $\theta$ (here, a translation function $h_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(h_\theta(\src), \trg)$.
+Given a training sample of matched source-target sentence pairs $(\src,\trg)$ drawn from an underlying distribution $\mathcal{D}_s$, a model parameterized by $\theta$ (here, a translation function $h_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(h_\theta(\src), \trg)$.
 This approach ensures that the translation loss remains low when translating more sentences drawn from the same distribution.
@@
 In this setting, \emph{domain adaptation} (DA) methods are in order.
-DA has a long history in Machine Learning in general (e.g.\ \cite{Shimodaira00improving,Ben-David09atheory,Quinonero08dataset,Pan10asurvey})\fyDone{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}).
+DA has a long history in Machine Learning in general (e.g.\ \cite{Shimodaira00improving,Ben-David09atheory,Quinonero08dataset,Pan10asurvey}) and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}).
 Various techniques thus exist to handle both the situations where a (small) training sample drawn from $\mathcal{D}_t$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
-% In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
 
@@
 MD machine translation (MDMT) corresponds to a very common situation, where all available data, no matter its origin, is used to train a robust system that performs well for any kind of new input.
-% As pointed out both by \newcite{Dredze08online} and \cite{Finkel09hierarchical}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}.
 If the intuitions behind MDMT are quite simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should MDMT perform well when the test data is distributed like the training data, when it is equally distributed across domains or when the test distribution is unknown?
@@
 How should it handle domain labeling errors?
-% , which makes their evaluation fragile
 
@@
 Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT systems and associated measurements to evaluate them (Section~\ref{sec:challenging}).
-In doing so, we also shed light on new problems that arise in this context, regarding for instance the accommodation of new domains in the course of training, or the computation of automatic domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
-We show in Section~\ref{sec:experiments} that existing approaches still fall short to match many of these requirements, notably with respect to the handling of a large amount of heterogeneous domains and to dynamically integrating new domains in training.\fyDone{Spell out conclusions}
+In doing so, we also shed light on new problems that arise in this context, regarding for instance the accommodation of new domains in the course of training, or the computation of automatic domain tags.
+Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
+We show in Section~\ref{sec:experiments} that existing approaches still fall short to match many of these requirements, notably with respect to the handling of a large amount of heterogeneous domains and to dynamically integrating new domains in training.
 
@@
 In this section, we recap the main reasons for considering a multi-domain scenario and discuss their implications in terms of performance evaluation.
-\fyDone{Relation to theory, discussion of worst case scenario}
-\fyDone{Insist on the work of Dredze}
 
@@
 We now discuss the arguments that are put forward to proceed differently.
-% --
-
-% This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyTodo{explain zero shot}.
 
@@
 The underlying assumption here is that the number of domains of interests can be large, a limiting scenario being fully personalized machine translation \cite{Michel2018extreme}.
-% which is not always the case: for instance \cite{Britz17mixing,Zeng18multidomain,Jiang19multidomain} only consider 2-4 domains in their experiments.\fyTodo{Keep this last part ?}
 
@@
 
-\fyDone{recap}
 To recap, there are multiple arguments for adopting MDMT, some already used in DA settings, and some original.
@@
 
-% \cite[eq. (3)]{Wang20general} defines the loss of MDMT as the (macro) average of domain loss - all domains are equally important.
 \section{Challenging multi-domain systems \label{sec:challenging}}
-In this section, we propose seven\fyDone{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.
+In this section, we propose seven operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.
 All these evaluations will rest on comparison of translation performance, and do not depend on the choice of a particular metric.
 To make our results comparable with the literature, we will only use the BLEU score \cite{Papineni02bleu} in Section~\ref{sec:experiments}, noting it may not be the best yardstick to assess subtle improvements of lexical choices that are often associated with domain adapted systems \cite{Irvine13measuring}.
-Other important figures of merit for MDMT systems are the computational training cost and the total number of parameters.\fyDone{Insert discussion about scores}
+Other important figures of merit for MDMT systems are the computational training cost and the total number of parameters.
 
 \subsection{Multi-domain systems should be effective \label{ssec:effective}}
-% \subsubsection{Performance}
 A first expectation is that MDMT systems should perform well in the face of mixed-domain test data. We thus derive the following requirements.
 
-\paragraph{[P1-LAB]}\fyDone{Decide naming scheme}A MDMT should perform better than the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}.
-Evaluating this requirement is a matter of a mere comparison, assuming the test distribution of domains is known: if all domains are equally important, performance averages can be reported; if they are not, weighted averages should be used instead.\fyFuture{Try random label assignments ?} % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
+\paragraph{[P1-LAB]} A MDMT should perform better than the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}.
+Evaluating this requirement is a matter of a mere comparison, assuming the test distribution of domains is known: if all domains are equally important, performance averages can be reported; if they are not, weighted averages should be used instead.
 
@@
 In this situation, MDMT should even outperform fine-tuning on either of the pseudo sub-domain.
-\jcDone{I dont understand the prediction... why running MDMT on 2 artificial subdomains should show  gains over fine-tunning?}\fyDone{small loss with respect to non split ?}
 
@@
 \paragraph{[P3-HET]}
-The notion of a domain being a fragile one, an effective MDMT system should be able to discover not only when cross-domain sharing is useful (cf.\ requirement [P2-TUN])\jcDone{what is P2.2?}, but also when intra-domain heterogeneity is hurting.
+The notion of a domain being a fragile one, an effective MDMT system should be able to discover not only when cross-domain sharing is useful (cf.\ requirement [P2-TUN]), but also when intra-domain heterogeneity is hurting.
 This requirement is tested by artificially conjoining separate domains into one during training, hoping that the loss in performance with respect to the baseline (using correct domain tags) will remain small.
@@
 \paragraph{[P4-ERR]}
-MDMTs should perform best when the true domain tag is known, but deteriorate gracefully in the face of tag errors; in this situation, catastrophic drops in performance are often observed\fyDone{cite \cite{McCloskey89catastrophic}}.
+MDMTs should perform best when the true domain tag is known, but deteriorate gracefully in the face of tag errors; in this situation, catastrophic drops in performance are often observed.
 This requirement can be assessed by translating test texts with erroneous domain tags and reporting the subsequent loss in performance.
@@
 Several situations need be considered: for domains seen in training, using automatically predicted domain labels should not be much worse than using the correct one.
-For test documents from unknown domains (zero-shot transfer), a good MD system should ideally outperform the default baseline that merges all available data.\fyDone{Systems react to unknown domains}
-% These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.
+For test documents from unknown domains (zero-shot transfer), a good MD system should ideally outperform the default baseline that merges all available data.
 
@@
 Most corpora are available from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
-To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
+To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.
 
-We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
-We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
-
-%for contrast experiments, we also use supplementary test sets from three other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
-This enables us to evaluate the loss in performance when the test set is from a domain not seen in training.
-% The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
+We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.\footnote{The code for reproducing our train, dev and test datasets is available at \url{https://github.com/qmpham/experiments}.} Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+We report significant differences at the level of $p=0.05$.
 
@@
 We measure the distance between domains using the $\mathcal{H}$-Divergence \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains.
-%Formally, given test sets of size $m$ for domains $A$ and $B$, and $h(x)$ a trained domain predictor, $\mathcal{H}(A,B)$ is computed %as:
-%$$
-%\mathcal{H}(A,B) = 2(1 - [\frac{1}{m} \sum_{x:h(x) = B} \mathbb{I}( x \in A) + \frac{1}{m} \sum_{x: h(x) = A} \mathbb{I}(x \in B)]),
-% $$
-% where $\mathbb{I}$ is the indicator function.
 Our discriminator is a SVM independently trained for each pair of domains, with sentence representations derived via mean pooling from the source side representation of the generic Transformer model.
-We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well separated from all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
+We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values.
+Results in Table~\ref{tab:domaindist} show that all domains are well separated from all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
 
@@
 Accordingly, we report two aggregate scores: a weighted average reflecting the training distribution, and an unweighted average, meaning that test domains are equally important.
-%   As explained above, depending whether MDMT are expected to perform specially well on large domains, or to treat all domains equally, one or the other training distribution should be prefered.\jcDone{All systems cand be learned following both Bal and Nat.
-does Nat corresponds to wAVG in test?
-is wAVG defined somewhere?}
-\item fine-tuned models \cite{Luong15stanford,Freitag16fast}\jcDone{i would use Luong and Manning 2015}, based on the \system{Mixed-Nat} system, further trained on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
+\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed-Nat} system, further trained on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
 The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems adapted for one domain, with no parameter sharing across domains.
@@
 The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024.
-% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
-% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.\fyDone{Describe the block adaptation layer - voir slides}
-% All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
+Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.
 
 \subsection{Multi-domain systems \label{ssec:systems}}
-\fyDone{Remove Gated residual}
+
 Our comparison of multi-domain systems includes our own reimplementations of recent proposals from the literature:\footnote{\revision{Further implementation details are in Appendix~A.}}
@@
 \item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or as a supplementary feature for each word (\system{DC-Feat}).
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDR});\fyDone{why not LDR as in Minh's paper?}
+\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDR});
 \item the three proposals of \newcite{Britz17mixing}.
@@
 \item the multi-domain model of \newcite{Zeng18multidomain} (\system{WDCMT}), where a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification and adversarial training are used to compute these representations.
-\revision{Again, inference does not use domain tags.}\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly improved, results.}\fyDone{Check this}
+\revision{Again, inference does not use domain tags.}\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly improved, results.}
 \item \revision{two multi-domain versions of the approach of \newcite{Bapna19simple}, denoted \system{FT-Res} and \system{MDL-Res}, where a domain-specific adaptation module is added to all the Transformer layers; within each layer, residual connections enable to short-cut this adapter.
@@
 The latter uses the same architecture, but a different training procedure and learns all parameters jointly from scratch with a mix-domain corpus.}
-  % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
-% the entire multi-domain system from scratch.\fyDone{Check this.}
+
 \end{itemize}
-\revision{This list includes systems that slightly depart from our definition of MDMT: standard implementations of \system{TTM} and \system{WDCMT} rely on infered, rather than on gold, domain tags - which must somewhat affect their predictions; \system{DM} and \system{ADM} make no use of domain tags at all.} We however did not consider the proposal of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges more strongly from our notion of MDMT.
-% and should be able to handle available domain labels in training and testing.
-% This system has shown to meet several of requirements and would also have no problem scaling to a large number of domains.\fyDone{TBC}
+This list includes systems that slightly depart from our definition of MDMT: standard implementations of \system{TTM} and \system{WDCMT} rely on infered, rather than on gold, domain tags - which must somewhat affect their predictions; \system{DM} and \system{ADM} make no use of domain tags at all.
+We however did not consider the proposal of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges more strongly from our notion of MDMT.
 
@@
   \centering
-  \fyDone{Fix column size}
   \begin{tabular}{|p{4cm}|*{8}{r|}} \hline
@@
   \label{tab:performance}
-  % \fyDone{Fill the table with all results, including a comparison with *RNNs* for wcmd}
-  \fyDone{Do we have significancy tests for averages?} %\fyTodo{Add tests wrt to RNN}
 \end{table*}
@@
 Results are in Table~\ref{tab:performance}.
-As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points\fyDone{BP?} better, notably owing to the much better results for \domain{rel}).
+As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points better, notably owing to the much better results for \domain{rel}).
 As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains.
@@
 Fine-tuning each domain separately yields a better baseline, outperforming \system{Mixed-Nat} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
-% A complete fine-tuning, updating all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to keeping fixed the capacity of the adapter module.
-\fyDone{Add 2 averages ?}\fyDone{Significance testing wrt Mix Generic, Full Fine-tuned, for each domain}
 
@@
 
-% for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.\fyDone{why LDC does not outperform DC as in Minh's paper (for en2fr MED) ?}
-
 All systems fail to outperform fine-tuning, sometimes by a wide margin, especially for an ``isolated'' domain like \domain{rel}.
@@
 In this situation, MDMT systems should dedicate a sufficient number of parameters to each domain, so as to close the gap with fine-tuning.
-\fyDone{More comments when we have all the results}\fyDone{Importance of sharing and unsharing}
 
@@
   \begin{tabular}{|p{1.8cm}|*{10}{r|}} \hline
-    % &&&&&& \\
     \hfill Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & \multicolumn{2}{c|}{Wrong} \\ % \hline
@@
     \system{TTM} (*)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & 0.0 & -0.1\\
-%    \system{MDL Gated} & & & & & & & & &  \\
     \system{DM} (*)           & -0.3   & -0.3  & +0.4 & +0.4 & +0.3 & +0.3 & +0.9 & +0.1 & 0.0 &-0.9\\
@@
 Underline denotes significant loss when domains are changed wrt.\ the baseline situation; bold for a significant improvement over \system{FT-Full}; (*) tags systems ignoring test domains.
-    \fyDone{Check wrong and significance of diff for ``old''}\fyDone{How about wrong ?}\jcDone{FTres results are odd... split LAW 0.5}
   }
@@
 The merge condition is hardly different, with again large losses for \system{FT-full} \revision{and \system{FT-Res}}, and small variations for all systems.
-We even observe some rare improvements with respect to the situation where we use actual domains.\fyDone{More comments ?}
+We even observe some rare improvements with respect to the situation where we use actual domains.
 
@@
 
-\fyDone{Numbers for the rand domain analysis}In the last two columns of Table~\ref{tab:redomains}, we report the drop in performance when the domain information is not correct.
-In the first (\domain{rnd}), we use test data from the domains seen in training, presented with a random domain tag\fyDone{all configs use the same tag or is randomization is run for each config}.
+In the last two columns of Table~\ref{tab:redomains}, we report the drop in performance when the domain information is not correct.
+In the first (\domain{rnd}), we use test data from the domains seen in training, presented with a random domain tag.
 In this situation, the loss with respect to using the correct tag is generally large (more than 10 BLEU points), showing an overall failure to meet requirement [P4-ERR], except for systems that ignore domain tags in testing.
-\fyDone{Comments?}
 
-In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the domain tag and use it for decoding.\footnote{\revision{Domain tags are assigned as follows: we train a language model for each domain and assign tag on a per-sentence basis based on the language model log-probability (assuming uniform domain priors).
+In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).
+For each sentence, we automatically predict the domain tag and use it for decoding.\footnote{\revision{Domain tags are assigned as follows: we train a language model for each domain and assign tag on a per-sentence basis based on the language model log-probability (assuming uniform domain priors).
 The domain classifier has an average prediction error of 16.4\% for in-domain data.}}
-% Results are in the rightmost column of Table~\ref{tab:redomains}.
-In this configuration, again, systems using domain tags during inference perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\fyDone{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}
+In this configuration, again, systems using domain tags during inference perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5).
 
 \subsubsection{Handling growing numbers of domains}
+
 Another set of experiments evaluate the ability to dynamically handle supplementary domains (requirement [P6-DYN]) as follows.
 Starting with the existing MD systems of Section~\ref{ssec:rawperformance}, we introduce an extra domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is critical for achieving optimal performance: as our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy based on the new training distribution.
-This is detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with training all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.\footnote{WDCMT results are excluded from this table, as resuming training proved difficult to implement.}\fyDone{Ecrire un commentaire}
+This is detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.
+We contrast this approach with training all systems from scratch and report differences in performance in Figure~\ref{fig:warmrestart} (see also Table~\ref{tab:warmrestart} in appendix~B).\footnote{WDCMT results are excluded from this table, as resuming training proved difficult to implement.}
 We expect that MDMT systems should not be too significantly impacted by the addition of a new domain and reach about the same performance as when training with this domain from scratch.
@@
 It is less easy for \system{DM}, \system{ADM} and \system{WDCMT}, which include a built-in domain classifier whose outputs have to be pre-specified, or for \system{LDR}, \system{FT-Res} and \system{MDL-Res} for which the number of possible domains is built in the architecture and has to be anticipated from the start.
-This makes a difference between domain-bounded systems, for which the number of domains is limited and truly open-domain systems.% , which seamlessly integrate additional domains.
+This makes a difference between domain-bounded systems, for which the number of domains is limited and truly open-domain systems.
 
@@
     \caption{Ability to handle a new domain.
-\revision{We report BLEU scores for a complete training session with 7 domains, as well as differences with (bleu) training with 6 domains (from Table~\ref{tab:performance}); (red) continuous training mode}.}
+\revision{We report BLEU scores for a complete training session with 7 domains, as well as differences (in blue) with training with 6 domains (from Table~\ref{tab:performance}); and (in red) differences with continual training}.}
+    \label{fig:warmrestart}
 \end{figure*}
+
+We can first compare the results of coldstart training with six or seven domains in Table~\ref{tab:warmrestart}: a first observation is that the extra training data \revision{is hardly helping} for most domains, except for \domain{News}, where we see a large gain, and for \domain{talk}.
+The picture is the same when one looks at MDMTs, where only the weakest systems (\system{DM}, \system{ADM}) seem to benefit from more (out-of-domain) data.
+Comparing now the coldstart with the warmstart scenario, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training also negatively impacts the performance for other domains.
+This happens notably for \system{DC-Tag}, \system{TTM} and \system{ADM}.
+In this setting \system{MDL-Res} and \system{DM} show the smaller average loss, with the former achieving the best balance of training cost and average BLEU score.
+
+\subsection{Automatic domains \label{ssec:autodomains}}
+In this section, we experiment with automatic domains, obtained by clustering sentences into $k=30$ classes using the k-means algorithm based on generic sentence representations obtained via mean pooling (cf.\ Section~\ref{ssec:corpora}).
+This allows us to evaluate requirement [P7-scale], training and testing our systems as if these domains were fully separated.
+Many of these clusters are mere splits of the large \domain{med}, while a fewer number of classes are mixtures of two (or more) existing domains (full details are in Appendix~C).
+We are thus in a position to reiterate, at a larger scale, the measurements of Section~\ref{ssec:redomains} and test whether multi-domain systems can effectively take advantage from the cross-domain similarities and to eventually perform better than fine-tuning.
+\revision{The results in Table~\ref{tab:avg_automatic_domains} also suggest that MDMT can surpass fine-tuning for the smaller clusters; for the large clusters, this is no longer true.
+The complete table (in Appendix~C) shows that this effect is more visible for small subsets of the medical domain.}
+
+\begin{table*}[t]
+\centering \footnotesize
+\revision{
+  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
+   Model/ & Train &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
+   Clusters & size & \system{Nat} & \system{Full} & \system{Res} &\system{Res} & \system{Feat}& \system{Tag}& & & & \\ \hline
+10 small&29.3k&68.3&70.0&70.7&\bf 71.2&70.6&53.1&67.3&69.8&67.0&70.2\\
+10 mid&104.7k&44.8&\bf 48.0&46.0&45.7&44.8&44.3&44.5&43.7&41.6&44.5\\
+10 large&251.1k&50.4&\bf 52.9&52.0&51.3&49.6&43.2&49.1&48.5&44.3&49.5\\
+Avg&128.4k&54.5&\bf 57.0&56.2&56.1&55.0&46.9&53.6&54.0&51.0&54.7\\ \hline
+  \end{tabular}
+   \caption{BLEU scores computed by merging the 10 smaller, medium, and larger cluster test sets.
+Best score for each group is in boldface.
+For the small clusters, full-fine tuning is outperformed by several MDMT systems - see details in Appendix~C.}
+   \label{tab:avg_automatic_domains}}
+\end{table*}
+Finally, Table~\ref{tab:subdomains} reports the effect of using automatic domain for each of the 6 test sets: each sentence was first assigned to an automatic class, translated with the corresponding multi-domain system with 30 classes; aggregate numbers were then computed, and contrasted with the 6-domain scenario.
+Results are clear and confirm previous observations: even though some clusters are very close, the net effect is a loss in performance for almost all systems and conditions.
+In this setting, the best MDMT in our pool (\system{MDL-Res}) is no longer able to surpass the \system{Mix-Nat} baseline.
+
+\begin{table*}[t]
+  \centering
+  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
+%     &&&&&& \\
+    Domain / Model  & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
+  %    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
+  %     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8\\
+  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
+    \system{DC-Tag}       & 38.5 & \SW{54.0} & 49.0   & 33.6 & \SW{42.2} & \SW{76.7} & 41.6 & 49.0 \\%    & 21.8 \\
+    \system{DC-Feat}      & 37.3  & 54.2 & 49.3   & 33.6 & \SW{41.9} & \SW{75.8} & 40.8 & \SW{48.7}  \\% & \SW{21.7} \\
+    \system{LDR}             & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 39.1 & \SW{48.3}         \\% & 22.1 \\ 
+    \system{TTM}            & 37.4 & \SW{53.7} & 48.9 & 32.8 & 41.3 & \SW{75.8} & 40.7 & \SW{48.3}   \\% &  23.4 \\
+    \system{DM}             & 35.4 & 49.3  & 45.2 & 29.7 & 37.1 & \SW{60.0} & 37.8 & 42.8 \\ % & 22.6\\
+    \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 39.5 & 47.1\\% & 23.3 \\
+    \revision{\system{FT-Res}}     & 37.5 & \SW{55.7}  & \SW{51.1}   & 33.1   &  \SW{44.1}  & \SW{86.7} & 41.6 & \SW{51.4}\\%  & \SW{21.2} \\
+    \system{MDL-Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SW{50.7}\\%  & \SW{21.2} \\
+    \system{WDCMT}       & 35.6 & 53.1 & 48.4 & 30.5 & \SW{37.7} & \SW{56.0} & 38.5 & 43.6 \\ % & 20.4 \\ 
+    \hline
+  \end{tabular}
+  \caption{Translation performance with automatic domains, computed with the original test sets.
+Significance tests are for comparisons with the 6-domain scenario \revision{(Table~\ref{tab:performance}).}}
+  \label{tab:subdomains}
+\end{table*}
+
+\section{Related work \label{sec:related}}
+
+The multi-domain training regime is more the norm than the exception for natural language processing \cite{Dredze08online,Finkel09hierarchical}, and the design of multi-domain systems has been proposed for many language processing tasks.
+We focus here exclusively on MD machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection / weighting, adversarial training, etc) have been studied in other contexts.
+
+%; scores are reported separately per domain, suggesting that all domains are equally important at test time.
+% Adversarial adpatation is used in many domains
+% for QA @misc{lee2019domainagnostic,
+%    title={Domain-agnostic Question-Answering with Adversarial Training},
+%    author={Seanie Lee and Donggyu Kim and Jangwon Park},
+%    year={2019},
+%    eprint={1910.09342},
+%    archivePrefix={arXiv},
+%    primaryClass={cs.CL}
+%}, NLU and many others
+Multi-domain translation was already proposed for statistical MT, either considering as we do multiple sources of training data (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}), or domains made of multiple topics \cite{Eidelman12topic,Hasler14dynamic-topic}.
+Two main strategies were considered: instance-based, involving a measure of similarities between train and test domains; feature-based, where domain/topic labels give rise to additional features.
+
+The latter strategy has been widely used in NMT: \newcite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token or in the form of an additional domain-feature associated to each word.
+These results are reproduced by \newcite{Tars18multidomain}, who also consider automatically induced domain tags.
+This technique also helps control the style of MT outputs in \cite{Sennrich16politeness,Niu18multitask}, and to encode the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}.
+Domain control can also be performed on the target side, as in \cite{Chen16guided}, where a topic vector describing the whole document serves as an extra context in the softmax layer of the decoder.
+Such ideas are further developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and commonalties are encoded in the network architecture: some parameters are shared across domains, while others are domain-specific.
+% (part of the embeddings in \cite{Pham19generic}).
+
+Techniques proposed by \cite{Britz17mixing} aim to ensure that domain information is actually used in a mix-domain system.
+Three methods are considered, using either domain classification (or domain normalization, via adversarial training) on the source or target side.
+There is no clear winner in either of the three language pairs considered.
+One contribution of this work is the idea of normalizing representations through adversarial training, so as to make the mixture of heterogeneous data more effective; representation normalization has since proven a key ingredient in multilingual transfer learning.
+The same basic techniques (parameter sharing, automatic domain identification / normalization) are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific representations on the one hand from domain-agnostic representations on the other hand.
+These representations are then processed as two separate inputs, then recombined to compute the translation.
+
+Another parameter-sharing scheme is in \cite{Jiang19multidomain}, which augments a Transformer model with domain-specific heads, whose contributions are regulated at the word/position level: some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling at the core of \cite{Huck15mixeddomain,Saunders19ucam}.
+The results for three language pairs outperform several standard baselines for a two-domain systems (in fr:en and de:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
+
+Finally, \citet{Farajian17multidomain}, \citet{Li18onesentence} and \citet{Xu19lexical} adopt a different strategy.
+Each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some iterations, before delivering its output.
+This approach entirely dispenses with the notion of domain and relies on data selection techniques to handle data heterogeneity.
+
+\section{Conclusion and outlook \label{sec:conclusion}}
+
+In this study, we have carefully reconsidered the idea of multi-domain machine translation, which seems to be taken for granted in many recent studies.
+We have spelled out the various motivations for building such systems and the associated expectations in terms of system performance.
+We have then designed a series of requirements that MDMT systems should meet, and proposed a series of associated test procedures.
+In our experiments with a representative sample of MDMTs, we have found that most requirements were hardly met for our experimental conditions.
+If MDMT systems are able to outperform the mixed-domain baseline, at least for some domains, they all fall short to match the performance of fine-tuning on each individual domain, which remains the best choice in multi-source single domain adaptation.
+\revisiondel{In this scenario, architecture-based strategies aimed at parameter sharing and/or representation normalization obtain the best results.}
+As expected however, MDMTs are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend, dynamically accommodate additional domains, this being especially easy for feature-based approaches.
+Our experiments finally suggest that all methods show decreasing performance when the number of domains or the diversity of the domain mixture increases.
+% (for a fixed data size).
+
+Two other main conclusions can be drawn from this study: first, it seems that more work is needed to make MDMT systems make the best out of the variety of the available data, both to effectively share what needs to be shared while at the same time separating what needs to be kept separated.
+We notably see two areas worthy of further explorations: the development of parameter sharing strategies when the number of domains is large; the design of training strategies that can effectively handle a change of the training mixture, including an increase in the number of domains.
+Both problems are of practical relevance in industrial settings.
+Second, and maybe more importantly, there is a general need to adopt better evaluation methodologies for evaluating MDMT systems, which require systems developers to clearly spell out the testing conditions and the associated expected distribution of testing instances, and to report more than comparisons with simple baselines on a fixed and known handful of domains.
+
+% --------------
+\section*{Acknowledgements}
+
+The work presented in this paper was partially supported by the European Commission under contract H2020-787061 ANITA.
+
+This work was granted access to the HPC resources of [TGCC/CINES/IDRIS] under the allocation 2020- [AD011011270] made by GENCI (Grand Equipement National de Calcul Intensif).
+
+\bibliographystyle{acl_natbib}
+\bibliography{main-2327-PhamMinhQuang}
+
+\clearpage{}
+
+\section*{Appendices}
+
+\subsection*{A - Description of multi-domain systems \label{ssec:implementation-details}}
+
+We use the following setups for MDMT systems.
+\begin{itemize}
+\item \system{Mixed-Nat}, \system{FT-full}, \system{TTM}, \system{DC-Tag} use a medium Transformer model of \cite{Vaswani17attention} with the following settings: embeddings size and hidden layers size are set to~$512$.
+Multi-head attention comprises 8 heads in each of the 6 layers; the inner feedforward layer contains~$2048$ cells.
+Training use a batch size of~$12,288$ tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.
+\item \system{FT-Res} and \system{MDL-res} use the same medium Transformer and add residual layers with a bottleneck dimension of size $1024$.
+\item \system{ADM}, \system{DM} use medium Tranformer model and a domain classifier composing of 3 dense layers of size $512 \times 2048$, $2048 \times 2048$ and $2048 \times domain\_num$.
+The two first layers of the classifier use the $\operatorname{ReLU}()$ as activation function, the last layer uses $\tanh()$ as activation function.
+\item \system{DC-Feat} uses medium Transformer model and domain embeddings of size~$4$.
+Given a sentence of domain~$i$ in a training batch, the embedding of domain $i$ is concatenated to the embedding of each token in the sentence.
+\item \system{LDR} uses medium Transformer model and for each token we introduce a \system{LDR} feature of size $4 \times domain\_num$.
+Given a sentence of domain $i\in[1,..,K]$ in the training batch, for each token of the sentence, the \system{LDR} units of the indexes outside of the range $[4(i-1),..,4i-1]$ are masked to $0$, and the masked LDR feature will be concatenated to the embedding of the token.
+Details are in \cite{Pham19generic}.
+\item \system{Mixed-Nat-RNN} uses one bidirectional LSTM layer in the encoder and one LSTM layer in the decoder. The size of hidden layers is $1024$, the size of word embeddings is $512$. 
+\item \system{WDCNMT} uses one bidirectional GRU layer in the encoder and one GRU-conditional layer in the decoder. The size of hidden layers is $1024$, the size of word embeddings is $512$.
+\end{itemize}
+\paragraph{Training} For each domain, we create train/dev/test sets by randomly splitting each corpus. We maintain the size of validation sets and of test sets equal to 1,000 lines for every domain.
+The learning rate is set
+as in \cite{Vaswani17attention}.
+For the fine-tuning procedures used for \system{FT-full} and \system{FT-Res}, we continue training using the same learning rate schedule, continuing the incrementation of the number of steps.
+All other MDMT systems reported in Tables~\ref{tab:performance} and \ref{tab:redomains} use a combined validation set comprising 6,000 lines, obtained by merging the six development sets.
+For the results in Table~\ref{tab:warmrestart} we also append the validation set of \domain{news} to the multi-domain validation set.
+In any case, training stops if either training reaches the maximum number of iterations (50,000) or the score on the validation set does not increase for three consecutive evaluations.
+We average five checkpoints to get the final model.
+
+\subsection*{B - Experiments with continual learning \label{ssec:full-continual}}
 
@@
     \footnotesize \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.2  & 29.1 & 41.0 & 49.4\\[-2pt]
-%                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
+
                     & \sbcl{0.0}{-0.3}  & \sbcl{+0.4}{-0.3}  & \sbcl{-0.1}{-0.5}  & \sbcl{-0.9}{\SW{-1.1}}  & \sbcl{+0.6}{\SW{-1.0}}  & \sbcl{\SB{+1.8}}{\SW{-4.0}} & \sbcl{\SW{-5.7}}{\SW{-1.4}} & \sbcl{0.0}{-0.5} & \sbcl{+0.3}{\SW{-1.2}}\\
@@
     \footnotesize \system{DM}            &36.0 &51.3&46.8&31.8&39.8&65.7&27.0 & 38.9 & 45.2\\[-2pt]
-%                   &36.5&51.7&47.4&31.7&40.3&65.7&\SW{25.8}& & \\
                    & \sbcl{-0.4}{+0.6} & \sbcl{\SW{-1.8}}{+0.4} & \sbcl{\SW{-1.2}}{+0.6} & \sbcl{\SW{-1.8}}{-0.1} & \sbcl{\SW{-2.6}}{+0.5} & \sbcl{\SW{-3.3}}{0.0} & \sbcl{\SW{-4.4}}{\SW{-1.2}} & \sbcl{-0.8}{+0.5} & \sbcl{\SW{-1.8}}{+0.3}\\
@@
     \footnotesize \system{ADM}          &36.6&54.2&49.1&32.9&42.1&75.7&28.7 & 40.2 & 48.4 \\[-2pt]
-%                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
                    & \sbcl{-0.2}{+0.3} & \sbcl{-0.7}{-0.8} & \sbcl{-0.8}{-0.8} & \sbcl{-0.9}{-0.2} & \sbcl{-0.5}{-0.4} & \sbcl{\SW{-2.3}}{\SW{-5.0}} & \sbcl{\SW{-5.4}}{\SW{-1.9}} & \sbcl{-0.5}{-0.2} & \sbcl{\SW{-0.9}}{\SW{-1.1}}\\
 
-%    \system{WDCMT}     &&&&&&& & &\\[-2pt]
-%                    &&&&&&&&& \\
-    \footnotesize \revision{\system{FT-Res}}  & 37.0 & 57.6 & 53.8 & 34.5 &	46.1 & 91.1 & 29.6 &	42.2  & 53.3 \\[-2pt]
+    \footnotesize \system{FT-Res}  & 37.0 & 57.6 & 53.8 & 34.5 &	46.1 & 91.1 & 29.6 &	42.2  & 53.3 \\[-2pt]
                                & \sbcl{+0.3}{+0.3} & \sbcl{+0.4}{+0.4} & \sbcl{+0.1}{+0.1} & \sbcl{-0.7}{-0.7} & \sbcl{+0.5}{+0.5} & \sbcl{-0.9}{-0.9} & \sbcl{\SW{-9.0}}{-0.6} & \sbcl{-0.1}{-0.1} & \sbcl{+0.2}{+0.2} \\
 
-    % 37.53	55.74	51.15	33.06	44.15	86.67	23.05	51.38	41.64
     \footnotesize \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & 41.9 & 51.8 \\[-2pt]
-%                      & 37.5 & 56.1 & 51.1 & 34.0  &  44.3 & 87.3 & \SW{28.4} & & 51.7 & \\
                         &  \sbcl{+0.2}{-0.2} & \sbcl{+0.4}{+0.5} & \sbcl{+0.1}{0.0} & \sbcl{-0.9}{-0.4}  & \sbcl{-0.1}{-0.2} &  \sbcl{+0.9}{-0.2} & \sbcl{\SW{-8.0}}{\SW{-0.8}} & \sbcl{+0.1}{-0.2} & \sbcl{+0.1}{-0.1} \\
@@
   \caption{Ability to handle a new domain.
-\revision{We report BLEU scores for a complete training session with 7 domains, as well as differences with (left) training with 6 domains (from Table~\ref{tab:performance}); (right) continuous training mode.} Averages only take into account six domains (\domain{News} excluded).
+We report BLEU scores for a complete training session with 7 domains, as well as differences with (left) training with 6 domains (from Table~\ref{tab:performance}); (right) continuous training mode.
+Averages only take into account six domains (\domain{News} excluded).
 Underline denotes a significant loss, bold a significant gain.}
   \label{tab:warmrestart}
-  \fyDone{Fill the table, significancy testing against initial learning condition, for each domain, and averages ? }
-  \fyDone{Need correct number / differences for Mixed-Nat}
-  \fyDone{averages include one more domain ? No for comparison}
-  \fyDone{i used smaller number for differences (easier to read...)}
 \end{table*}
 
-We can first compare the results of coldstart training with six or seven domains in Table~\ref{tab:warmrestart}: a first observation is that the extra training data \revision{is hardly helping} for most domains, except for \domain{News}, where we see a large gain, and for \domain{talk}.
-The picture is the same when one looks at MDMTs, where only the weakest systems (\system{DM}, \system{ADM}) seem to benefit from more (out-of-domain) data.\fyDone{is this significant ?}
-% Increasing the heterogeneity of the mix with supplementary out-of-domain data does not always improve the performance.% across the board.
-% Adding is not helping here, especially when domains are well separated.
-Comparing now the coldstart with the warmstart scenario, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training also negatively impacts the performance for other domains.
-This happens notably for \system{DC-Tag}, \system{TTM} and \system{ADM}.
-In this setting \system{MDL-Res} and \system{DM} show the smaller average loss, with the former achieving the best balance of training cost and average BLEU score.% of all systems.
-\fyDone{Numbers for MDL Res need significancy testing; and they are weird - why do we see this ? Because the generic is bad ?}
-% \begin{itemize}
-% \item even when trained from entirely scratch, very few systems are able to take advantage of the additional data for the 6 initial domains - this is observed on the averaged scores in Table~\ref{tab:warmrestart} with hardly surpass those in Table~\ref{tab:performance}
-% \item restart is computationally cheaper (\fyTodo{how much cheaper here ?}) but underperforms cold start for all systems, sometimes by a wide margin.
-% \item the loss is largest for the small domains - how about distance to \domain{news} ?
-% \end{itemize}
-% \fyDone{Very close to fine tuning towards News, no ?}
+\subsection*{C - Experiments with automatic domains \label{ssec:full-automatic}}
 
-\subsection{Automatic domains \label{ssec:autodomains}}
-In this section, we experiment with automatic domains, obtained by clustering sentences into $k=30$ classes using the k-means algorithm based on generic sentence representations obtained via mean pooling (cf.\ Section~\ref{ssec:corpora}).
-This allows us to evaluate requirement [P7-scale], training and testing our systems as if these domains were fully separated.
-Many of these clusters are mere splits of the large \domain{med}, while a fewer number of classes are mixtures of two (or more) existing domains \revision{(Full details are in Appendix B)}.
-We are thus in a position to reiterate, at a larger scale, the measurements of Section~\ref{ssec:redomains} and test whether multi-domain systems can effectively take advantage from the cross-domain similarities and to eventually perform better than fine-tuning.
-\revision{The results in Table~\ref{tab:avg_automatic_domains} also suggest that MDMT can surpass fine-tuning for the smaller clusters; for the large clusters, this is no longer true.
-The complete table (in Appendix B) shows that this effect is more visible for small subsets of the medical domain.}
-
-\begin{table*}[t]
-\centering \footnotesize
-\revision{
-  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
-   Model/ & Train &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
-   Clusters & size & \system{Nat} & \system{Full} & \system{Res} &\system{Res} & \system{Feat}& \system{Tag}& & & & \\ \hline
-10 small&29.3k&68.3&70.0&70.7&\bf 71.2&70.6&53.1&67.3&69.8&67.0&70.2\\
-10 mid&104.7k&44.8&\bf 48.0&46.0&45.7&44.8&44.3&44.5&43.7&41.6&44.5\\
-10 large&251.1k&50.4&\bf 52.9&52.0&51.3&49.6&43.2&49.1&48.5&44.3&49.5\\
-Avg&128.4k&54.5&\bf 57.0&56.2&56.1&55.0&46.9&53.6&54.0&51.0&54.7\\ \hline
-  \end{tabular}
-   \caption{BLEU scores computed by merging the 10 smaller, medium, and larger cluster test sets.
-Best score for each group is in boldface.
-For the small clusters, full-fine tuning is outperformed by several MDMT systems - see details in Appendix~B.}
-   \label{tab:avg_automatic_domains}}
-\end{table*}
-Finally, Table~\ref{tab:subdomains} reports the effect of using automatic domain for each of the 6 test sets: each sentence was first assigned to an automatic class, translated with the corresponding multi-domain system with 30 classes; aggregate numbers were then computed, and contrasted with the 6-domain scenario.
-Results are clear and confirm previous observations: even though some clusters are very close, the net effect is a loss in performance for almost all systems and conditions.
-In this setting, the best MDMT in our pool (\system{MDL-Res}) is no longer able to surpass the \system{Mix-Nat} baseline.%\fyTodo{Note about scaling?}\fyTodo{More comparison?}
-
-\begin{table*}[t]
-  \centering
-  \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
-%     &&&&&& \\
-    Domain / Model  & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-  %    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
-  %     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8\\
-  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag}       & 38.5 & \SW{54.0} & 49.0   & 33.6 & \SW{42.2} & \SW{76.7} & 41.6 & 49.0 \\%    & 21.8 \\
-    \system{DC-Feat}      & 37.3  & 54.2 & 49.3   & 33.6 & \SW{41.9} & \SW{75.8} & 40.8 & \SW{48.7}  \\% & \SW{21.7} \\
-    \system{LDR}             & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 39.1 & \SW{48.3}         \\% & 22.1 \\ 
-    \system{TTM}            & 37.4 & \SW{53.7} & 48.9 & 32.8 & 41.3 & \SW{75.8} & 40.7 & \SW{48.3}   \\% &  23.4 \\
-    \system{DM}             & 35.4 & 49.3  & 45.2 & 29.7 & 37.1 & \SW{60.0} & 37.8 & 42.8 \\ % & 22.6\\
-    \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 39.5 & 47.1\\% & 23.3 \\
-    \revision{\system{FT-Res}}     & 37.5 & \SW{55.7}  & \SW{51.1}   & 33.1   &  \SW{44.1}  & \SW{86.7} & 41.6 & \SW{51.4}\\%  & \SW{21.2} \\
-    \system{MDL-Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SW{50.7}\\%  & \SW{21.2} \\
-    \system{WDCMT}       & 35.6 & 53.1 & 48.4 & 30.5 & \SW{37.7} & \SW{56.0} & 38.5 & 43.6 \\ % & 20.4 \\ 
-    \hline
-  \end{tabular}
-  \caption{Translation performance with automatic domains, computed with the original test sets.
-Significance tests are for comparisons with the 6-domain scenario \revision{(Table~\ref{tab:performance}).}}
-  \label{tab:subdomains}
-  \fyDone{Fill the table with correct results,}\fyDone{Change wavg and avg}
-\end{table*}
-
-\section{Related work \label{sec:related}}
-
-The multi-domain training regime is more the norm than the exception for natural language processing \cite{Dredze08online,Finkel09hierarchical}, and the design of multi-domain systems has been proposed for many language processing tasks.
-We focus here exclusively on MD machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection / weighting, adversarial training, etc) have been studied in other contexts.
-
-%; scores are reported separately per domain, suggesting that all domains are equally important at test time.
-% Adversarial adpatation is used in many domains
-% for QA @misc{lee2019domainagnostic,
-%    title={Domain-agnostic Question-Answering with Adversarial Training},
-%    author={Seanie Lee and Donggyu Kim and Jangwon Park},
-%    year={2019},
-%    eprint={1910.09342},
-%    archivePrefix={arXiv},
-%    primaryClass={cs.CL}
-%}, NLU and many others
-Multi-domain translation was already proposed for statistical MT, either considering as we do multiple sources of training data (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}), or domains made of multiple topics\fyDone{you mean multiple topics?} \cite{Eidelman12topic,Hasler14dynamic-topic}.\fyDone{Add also topic models refs} Two main strategies were considered: instance-based, involving a measure of similarities between train and test domains; feature-based, where domain/topic labels give rise to additional features.
-
-The latter strategy has been widely used in NMT: \newcite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token or in the form of an additional domain-feature associated to each word.
-These results are reproduced by \newcite{Tars18multidomain}, who also consider automatically induced domain tags.
-This technique also helps control the style of MT outputs in \cite{Sennrich16politeness,Niu18multitask}, and to encode the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}.
-Domain control can also be performed on the target side, as in \cite{Chen16guided}, where a topic vector describing the whole document serves as an extra context in the softmax layer of the decoder.
-Such ideas are further developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and commonalties are encoded in the network architecture: some parameters are shared across domains, while others are domain-specific.
-% (part of the embeddings in \cite{Pham19generic}).
-
-Techniques proposed by \cite{Britz17mixing} aim to ensure that domain information is actually used in a mix-domain system.
-Three methods are considered, using either domain classification (or domain normalization, via adversarial training) on the source or target side.
-There is no clear winner in either of the three language pairs considered.
-One contribution of this work is the idea of normalizing representations through adversarial training, so as to make the mixture of heterogeneous data more effective; representation normalization has since proven a key ingredient in multilingual transfer learning.
-The same basic techniques (parameter sharing, automatic domain identification / normalization) are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific representations on the one hand from domain-agnostic representations on the other hand.
-These representations are then processed as two separate inputs, then recombined to compute the translation.
-
-Another parameter-sharing scheme is in \cite{Jiang19multidomain}, which augments a Transformer model with domain-specific heads, whose contributions are regulated at the word/position level: some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling at the core of \cite{Huck15mixeddomain,Saunders19ucam}.
-The results for three language pairs outperform several standard baselines for a two-domain systems (in fr:en and de:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
-
-Finally, \citet{Farajian17multidomain}, \citet{Li18onesentence} and \citet{Xu19lexical} adopt a different strategy.
-Each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some iterations, before delivering its output.
-This approach entirely dispenses with the notion of domain and relies on data selection techniques to handle data heterogeneity.
-
-\section{Conclusion and outlook \label{sec:conclusion}}
-\fyDone{Write conclusions}\fyFuture{no mixture of experts}
-In this study, we have carefully reconsidered the idea of multi-domain machine translation, which seems to be taken for granted in many recent studies.
-We have spelled out the various motivations for building such systems and the associated expectations in terms of system performance.
-We have then designed a series of requirements that MDMT systems should meet, and proposed a series of associated test procedures.
-In our experiments with a representative sample of MDMTs, we have found that most requirements were hardly met for our experimental conditions.
-If MDMT systems are able to outperform the mixed-domain baseline, at least for some domains, they all fall short to match the performance of fine-tuning on each individual domain, which remains the best choice in multi-source single domain adaptation.
-\revisiondel{In this scenario, architecture-based strategies aimed at parameter sharing and/or representation normalization obtain the best results.}\fyDone{Mitigate this}
-As expected however, MDMTs are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend, dynamically accommodate additional domains, this being especially easy for feature-based approaches.
-Our experiments finally suggest that all methods show decreasing performance when the number of domains or the diversity of the domain mixture increases.
-% (for a fixed data size).
-
-Two other main conclusions can be drawn from this study: first, it seems that more work is needed to make MDMT systems make the best out of the variety of the available data, both to effectively share what needs to be shared while at the same time separating what needs to be kept separated.
-We notably see two areas worthy of further explorations: the development of parameter sharing strategies when the number of domains is large; the design of training strategies that can effectively handle a change of the training mixture, including an increase in the number of domains.
-Both problems are of practical relevance in industrial settings.
-Second, and maybe more importantly, there is a general need to adopt better evaluation methodologies for evaluating MDMT systems, which require systems developers to clearly spell out the testing conditions and the associated expected distribution of testing instances, and to report more than comparisons with simple baselines on a fixed and known handful of domains.
-
-% --------------
-\section*{Acknowledgements}
-
-The work presented in this paper was partially supported by the European Commission under contract H2020-787061 ANITA.
-
-This work was granted access to the HPC resources of [TGCC/CINES/IDRIS] under the allocation 2020- [AD011011270] made by GENCI (Grand Equipement National de Calcul Intensif)
-% The acknowledgements should go immediately before the references.  Do
-% not number the acknowledgements section. Do not include this section
-% when submitting your paper for review.
-
-% include your own bib file like this:
-% \bibliographystyle{coling}
-
-% \newpage\newpage
-
-\bibliographystyle{acl_natbib}
-\bibliography{multidomain}
-
-\clearpage{}
-
-\section*{Appendices}
-
-\subsection*{A - Description of multi-domain systems \label{ssec:implementation-details}}
 \begin{table*}[t]
@@
 
-\revision{%
-We use the following setups for MDMT systems.
-\begin{itemize}
-\item \system{Mixed-Nat}, \system{FT-full}, \system{TTM}, \system{DC-Tag} use a medium Transformer model of \cite{Vaswani17attention} with the following settings: embeddings size and hidden layers size are set to~$512$.
-Multi-head attention comprises 8 heads in each of the 6 layers; the inner feedforward layer contains~$2048$ cells.
-% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
-% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training use a batch size of~$12,288$ tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.
-\item \system{FT-Res} and \system{MDL-res} use the same medium Transformer and add residual layers with a bottleneck dimension of size $1024$.
-  % \cite{Bapna19simple} showed that if the bottleneck dimension is large enough, the gap between \system{FT-block} and \system{FT-full} will be small.\fyTodo{This comment is confusing}.
-\item \system{ADM}, \system{DM} use medium Tranformer model and a domain classifier composing of 3 dense layers of size $512 \times 2048$, $2048 \times 2048$ and $2048 \times domain\_num$.
-The two first layers of the classifier use the $\operatorname{ReLU}()$ as activation function, the last layer uses $\tanh()$ as activation function.
-\item \system{DC-Feat} uses medium Transformer model and domain embeddings of size~$4$.
-Given a sentence of domain~$i$ in a training batch, the embedding of domain $i$ is concatenated to the embedding of each token in the sentence.
-\item \system{LDR} uses medium Transformer model and for each token we introduce a \system{LDR} feature of size $4 \times domain\_num$.
-Given a sentence of domain $i\in[1,..,K]$ in the training batch, for each token of the sentence, the \system{LDR} units of the indexes outside of the range $[4(i-1),..,4i-1]$ are masked to $0$, and the masked LDR feature will be concatenated to the embedding of the token.
-Details are in \cite{Pham19generic}.
-\item \system{Mixed-Nat-RNN} uses one bidirectional LSTM layer in the encoder and one LSTM layer in the decoder. The size of hidden layers is $1024$, the size of word embeddings is $512$. 
-\item \system{WDCNMT} uses one bidirectional GRU layer in the encoder and one GRU-conditional layer in the decoder. The size of hidden layers is $1024$, the size of word embeddings is $512$.
-\end{itemize}
-\paragraph{Training} For each domain, we create train/dev/test sets by randomly splitting each corpus. We maintain the size of validation sets and of test sets equal to 1,000 lines for every domain.
-The learning rate is set
-% to $lr = d_{model}^{-0.5}*\min(step\_num^{-0.5},step\_num * warmup\_steps^{-1.5})$ where $d_{model}=512$,$warmup\_steps=4000$,
-as in \cite{Vaswani17attention}.
-For the fine-tuning procedures used for \system{FT-full} and \system{FT-Res}, we continue training using the same learning rate schedule, continuing the incrementation of the number of steps.
-All other MDMT systems reported in Tables~\ref{tab:performance} and \ref{tab:redomains} use a combined validation set comprising 6,000 lines, obtained by merging the six development sets.
-For the results in Table~\ref{tab:warmrestart} we also append the validation set of \domain{news} to the multi-domain validation set.
-In any case, training stops if either training reaches the maximum number of iterations (50,000)\fyDone{How many ?} or the score on the validation set does not increase for three consecutive evaluations.
-We average five checkpoints to get the final model.\fyDone{up to the best ?}
-}
-\subsection*{B - Experiments with automatic domains \label{ssec:full-automatic}}
-\revision{
 This experiment aims to simulate with automatic domains a scenario where the number of ``domains'' is large and where some ``domains'' are close and can effectively share information.
@@
 Domain-control (\system{DC-Feat}) is very effective for small domains, but again less so in larger data conditions.
-Among the MD models, approaches using residual adapters have the best average performance.}
-% \newpage\newpage \newpage
+Among the MD models, approaches using residual adapters have the best average performance.
 
 \end{document}
-
-\todos{}
-
-\subsection{Formalizing multi-domain adaptation (MDA)}
-
-In MDA, the typical training situation (for a classifier) considers $k$ domains with $\mathcal{D}_k(x,y)$ the joint distribution of features $x \in \mathcal{X}$ and labels $y \in \mathcal{Y}$ for domain $k$.
-In \cite{Mansour09domainadaptation}, the labeling function $f$ is deterministic and the joint distribution is entirely defined by the marginal $\mathcal{D}_k(x)$ and $f()$.This work further assumes $k$ hypotheses $h_k(x): \mathcal{X} \rightarrow{} \mathcal{Y}$ with an error smaller than $\epsilon$ when $x\sim \mathcal{D}_k$: $\sum_{x} L(h_k(x), f(y)) \mathcal{D}_k(x) < \epsilon_k$ and that the target distribution is a convex combination of $\{\mathcal{D}_k\}$: $\mathcal{D}_T(x) = \sum_k \lambda_k \mathcal{D}_k(x)$.
-\cite{Mansour09domainadaptation} focuses on \textsl{combined hypotheses}, ie.\ hypotheses for which each prediction $h(x)$ is a linear combination of domain specific hypotheses $h(x_k)$, and derives generalization bounds for the case where $\mathcal{D}_T(x)$ is fixed in advance, or worst case bounds valid for any mixture target distribution.
-The optimal solution has the form $h_{z,\eta}(x) = \sum_k \frac{z_kD_k(x) + U(x)/\gamma}{\sum_{k'} z_kD_k(x) + \eta U(x)/l}$ and is termed the \emph{distribution weighted combined rule}.
-Further work in \cite{Mansour09multiple} generalizes these bounds to the case where $\mathcal{D}_T$ is not a mixture of existing domains (in that case we will find a mixture that is close to $\mathcal{}$in the Renyì-divergence sense ); where $\mathcal{D}_k$ are unknown and estimated from the data; or when the labelling function is not the same across domains.
-More recent work study the case where the labelling function is stochastic, meaning that domains are defined by joint ($\mathcal{D}_k(x,y)$) rather than marginal distributions \cite{Hoffman18algorithms}, and provides an algorithm to compute the optimal weights for the combined rule based on estimates of the joint distributions.
-A guiding principle is that this rule should tend towards uniform performance across domains, to ensure robustness in the worst case scenario \cite{Oren19distributionally}.
-
-A related problem is \emph{Multiple source single domain adaptation} which is a variation of conventional DA from a mixture of source domains.
-In this setting it is advantageous to take advantage of the similarities between each source and the fixed target domain (see eg.\ \cite{Duan12domain,Cortes19AdaptationBO,Wen19domain} for studies of the supervised case, \cite{Zhao18adversarial} for the unsupervised case).
-
-\subsection{Complements - to be removed \label{ssec:oldformalization}}
-
-We define a domain $d$ as a joint distribution $\mathcal{D}_d(x,y)$ over input-output pairs $(x,y) \in \mathcal{X}\times{}\mathcal{Y}$.
-Assuming that the training data is sampled from a known mixture of $D$ source domains $\mathcal{D}_s(x,y) = \sum_d \lambda_{s,d} \mathcal{D}_d(x,y)$, several cases are studied in the ML literature:
-
-\begin{itemize}
-\item the test distribution $\mathcal{D}_t$ is known 
-  \begin{itemize}
-  \item $\mathcal{D}_t = \mathcal{D}_{s}$: this is standard ML, with mixture densities
-  \item $\mathcal{D}_t(x,y) = \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) $ with a different mixture of domains in training and testing ($\vlambda_s \neq \vlambda_t$), corresponding to multiple source adaptation (MSA) \cite{Mansour09domainadaptation},
-  \item $\mathcal{D}_t(x,y) \neq \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) \forall \lambda$, where the target domain for adapation is not a mixture of source domains, studied in \cite{Mansour09multiple}
-  \item variants of MSA which only assume knowledge of the marginal distribution $\mathcal{D}_t(x) = \sum_{y} \lambda_{t,d} \mathcal{D}_d(x) $
-  \item more realistic situations where  the train and test distributions are estimated from finite samples, also studied in \cite{Mansour09multiple}.
-  \end{itemize}
-   Note that we recover standard supervised DA is when there is only one component in the training and test mixtures, and multiple-source single domain adaptation when the source is a mixture but the test domain is homogeneous.
-Unsupervised DA \cite{Daume06domain} corresponds to the case where only the marginal distribution of the test domain inputs is available, while semi-supervised DA is the reverse situation.
-
-\item the test distribution is not known and training needs to be distributionally robust, ie.\ minimize the error against all possible distributions, with variants depending on whether it is or not assumed to be a mixture of existing domains, and whether it is perfectly or imperfectly known.
-This setting is also studied in eg.\ \cite{Mansour09multiple,Hoffman18algorithms}, and  in \cite{Oren19distributionally} for adversarial language model learning.
-\item with respect to the conditional label distribution $P(y|x)$:
-  \begin{itemize}
-  \item $P(y|x)$ is deterministic, meaning that the joint distribution is entirely specified by the marginal and a labelling function $y=h(x)$; a more complex case is when $h()$ also varies across domains;
-  \item $P(y|x)$ is stochastic, but shared across all domains, which is the typical covariate-shift scenario \cite{Shimodaira00improving}; it can also vary across domains \cite{???}\fyTodo{Missing citation vary accoss domains}
-  \end{itemize}
-\end{itemize}
-
-Multi-domain learning, as defined in \citet{Dredze08online} additionally makes the requirement that categorial domain tags are available in training and testing; a multi-domain learner should use them effectively \cite{Joshi12multidomain} to improve simultaneously on all domains \cite{Finkel09hierarchical}.
-This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyDone{explain zero shot}.
-% ----
-
-% \subsection{MDMT: a tentative definition \label{ssec:definition}}
-
-% Informally, a multi-domain system should be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
-% \begin{itemize}
-% \item the more common situation is when both train and test documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
-It is usually assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT.
-This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name here: such as supervised MDMT ?
-and mostly supervised MDMT for the second case ?}
-
-% \item It can also be the case that domain labels are only known in training, while test inputs may or may not be associated with explicit domains.
-This is, for instance, the setting adopted \cite{Zeng18multidomain}, where automatic domain identification is performed for test sentences.
-Note that automatic domain identification can also be effectively performed in the first setting, especially in the presence of close or related domains in training.
-
-% \item A last setting, that has not received much attention for MT (see however \cite{Sennrich13multidomain}), is where both train and test domains are automatically infered from data, using clustering techniques.
-Even though domain labels are typically available in most training scenario, these approaches assumes that might be preferable, owing to the multiple sources of intra-corpus and inter-corpus variabilities, to resort to automatic domain assignments.
-The online approach of \newcite{Farajian17multidomain}, where distributional cues, rather that categorical tags, are used for fine-tuning, is a variant of this strategy.
-% \end{itemize}
-
-% In the rest of the paper, we mostly adhere to the first definition which correspond to a situation of practical interest for the industry, and assume that multi-domain MT is mostly about effectively using domain tags in training and testing.
-We also experiment in Section~\ref{sec:results} with the other two alternative scenarios, as they also define useful contexts to evaluate the actual robustness of MDMT systems.
-
-% \jcDone{...a very common scenario in MT industry where clients dispose of mulitiple data corresponding to different domains (differing in genre, thematic, register, style, etc.). }
-
-% \subsection{Evaluation issues \label{ssec:evaluation}}
-
-% In theory, evaluation procedures should report realistic estimates of the test loss, computed on a sample deemed representative of the anticipated test conditions.
-Three scenarios need be distinguished:
-% \begin{itemize}
-% \item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
-% \item all domains are equally likely and important in testing (the prior over test domains is uniform), which would warrant the computation of per domain scores and of unweighted averages, as reported in \cite{Farajian17multidomain,Su19exploring,Wang20general}.
-Note that in this situation, the optimal training regime would be to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data.
-\cite{Wang20general} uses an intermediate mixing regime, which tends to reduce part of the unbalance, and improve the performance for underrepresented domains.
-% \item the test distribution is neither equal to the training set, nor uniform.
-Here again, if this distribution is known in advance, an optimal domain mix should be used in training to compute baseline scores, rather than using the empirical distribution of all the training data.
-A more challenging case is when the test set includes data from domains unseen in training, or when it is chosen adversarially \cite{XXX}.
-% % : this situation, where test instances have a null probability in training, goes agains the basic statistical assumptions of DA \cite{Shimodaira00improving}.
-It is thus important that the likeliness of such events should be clearly formulated when describing the evaluation protocol.
-% \end{itemize}
-
-% The papers in the literature are seldom explicit regarding their assumptions regarding the test distribution.
-In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 most faithfully reflects the expected testing conditions.
-Another important issue is the choice of the evaluation metrics: if BLEU is still the standard metric used in most studies, it is not necessary the best choice to measure subtle improvements of lexical choices \cite{Irvine13measuring} that are often associated with adapted systems.
-
-% A last issue regarding evaluation is related to the measurement of computational costs: when this is the main a motivation for using MDMT, we expect that computational costs would be reported and compared, which however is hardly ever done in practice.
-\fyTodo{Keep this ?}

### 2020-11-18 2021/mdmt/mdmt.tex
mdmt

@@
 
+\section{Algorithm}
+\begin{algorithm}[h]
+\caption{Multi-domain Training}
+\label{alg:multidomain}
+\begin{algorithmic}[1]
+\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for $d$ domains, Batch size $B$}%, Optimization algorithm $\operatorname{Opt}$}
+\REPEAT 
+\STATE{Iteration t}
+\STATE{If $t mod 1000 == 0$: Update $H_t$ via \ref{alg:Fisher} else $H_t = H_{t-1}$}
+\STATE{Randomly pick $i \in [1,..,d]$ w.r.t the multinomial distribution $[\frac{|C_i|}{\sum_{i\in [1,..,d]}|C_i|}]$.}
+\STATE{Randomly pick $B$ sentences from $C_i$.}
+\STATE{Activate residual adapters of domain $i$.}
+\STATE{Calculate gradients $\frac{\delta L_i(B_i)}{\delta \theta_i}$, $\frac{\delta L_i(B_i)}{\delta \theta_s}$}
+\STATE{Correct gradient by Fisher diagonal matrix $H_t$: 
+\begin{align*}
+\Delta \theta_i &= -\frac{1}{R} (H_t^i + \mu)^{-1} \frac{\delta L_i}{\delta \theta_i} \\
+\Delta \theta_s &= -\frac{1}{R} (H_t^s + \mu)^{-1} \frac{\delta L_i}{\delta \theta_s} \\
+R &=  (H_t^i + \mu)^{-1} \frac{\delta L_i}{\delta \theta_i} * \frac{\delta L_i}{\delta \theta_i}^T + (H_t^s + \mu)^{-1} \frac{\delta L_i}{\delta \theta_s} * \frac{\delta L_i}{\delta \theta_s}^T \\
+\end{align*}
+}
+\STATE{Update parameters:
+\begin{align*}
+\theta_i^{t+1} = \theta_i^{t} - lr * \Delta \theta_i \\
+\theta_s^{t+1} = \theta_s^{t} - lr * \Delta \theta_s \\
+\end{align*}}
+\UNTIL{convergence}
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[h]
+\caption{Compute Fisher matrix}
+\label{alg:Fisher}
+\begin{algorithmic}[1]
+\REQUIRE {Corpora $C_i, i\in [1,..,d]$ for $d$ domains, Batch size $B$}
+\STATE{Pick batch $B_i$ from $C_i$ $\forall i \in [1,..,d]$}
+\FOR{$i\in [1,..,d]$}
+     \STATE{$H^i = 0$, $H^s = 0$}
+\ENDFOR
+\FOR{$i\in [1,..,d]$}
+   \FOR{$j\in [1,..,B]$}
+        \STATE{Compute gradient wrt example $B_i^j$ of batch $B_i$: $\Delta H_i = \frac{\delta L_i(B_i^j)}{\delta \theta_i}$, $\Delta H_s =\frac{\delta L_i(B_i^j)}{\delta \theta_s}$}
+        \STATE{Update $H_i$, $H_s$:
+\begin{align*}
+H_i &+= \Delta H_i ** 2 \\
+H_s &+= \Delta H_s ** 2 \\
+\end{align*}}
+   \ENDFOR
+\ENDFOR
+\FOR{$i\in [1,..,d]$}
+     \STATE{$H^i \setminus= B $, $H^s \setminus= B*d$}
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
+
 \bibliographystyle{acl_natbib}

### 2021-01-22 2021/acl-ijcnlp2021/mdmt.tex
first push

--- 2020-08-16 2020/wmt_20/emnlp2020.tex
@@
 %
-% File emnlp2020.tex
+% File acl2021.tex
 %
+%% Based on the style files for EMNLP 2020, which were
 %% Based on the style files for ACL 2020, which were
@@
 \documentclass[11pt,a4paper]{article}
-\usepackage[hyperref]{emnlp2020}
+\usepackage[hyperref]{acl2021}
 \usepackage{times}
@@
 
-\usepackage{graphicx}
-\usepackage{xcolor}
-\usepackage{longtable}
-\usepackage{tikz}
-\usetikzlibrary{calc}
-\usepackage[draft]{todo}
-\usepackage{xspace}
-\usepackage{amsmath,amsfonts,amssymb}
-\usepackage{mathrsfs}
-\newcommand{\R}{\ensuremath{\mathbb{R}}}
-
 % This is not strictly necessary, and may be commented out,
 % but it will improve the layout of the manuscript,
-% and will typically save some space.Boldface denotes significant gains with respect to \system{Mix-Nat} (or \system{Mix-Nat-RNN}, for WDCMT), underline denotes significant losses.
+% and will typically save some space.
 \usepackage{microtype}
@@
 
+\newcommand\BibTeX{B\textsc{ib}\TeX}
+
+% Standard package includes
+\usepackage{times}
+\usepackage{url}
+\usepackage{latexsym}
+\usepackage{graphicx}
+\usepackage{xcolor}
+\usepackage{longtable}
+\usepackage{tikz}
+\usetikzlibrary{calc}
+\usepackage[draft]{todo}
+\usepackage[normalem]{ulem}
+\usepackage{xspace}
+\usepackage{amsmath,amsfonts,amssymb}
+\usepackage{algorithm, algorithmic}
 \newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
@@
 \newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{red}{#1}}}
-\newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{red}{#1}}}
-
-\newcommand{\mpTodo}[1]{\Todo[MP:]{\textcolor{green}{#1}}}
-\newcommand{\mpDone}[1]{\done[MP]\Todo[MP:]{\textcolor{green}{#1}}}
-\usepackage{mathtools,xparse}
-\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
-\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
-\NewDocumentCommand{\normL}{ s O{} m }{%
-  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{2}%
-}
+\newcommand{\revision}[1]{#1}
+\newcommand{\revisiondel}[1]{}
 \newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
@@
 \renewcommand\floatpagefraction{.95}
+\newcommand{\sbcl}[2]{{\scriptsize #1 \hfill $|$ \hfill  #2}}
+\usepackage{multirow}
 
-\newcommand\BibTeX{B\textsc{ib}\TeX}
-
-\iffalse   % removes draft highlighting for final version
-\renewcommand{\draftAdd}[2]{#1}
-\fi
-
-% -- \title{Ablation study on the residual adapter in Neural Machine Translation}
-\title{A Study of Residual Adapters for Multi-Domain Neural Machine Translation}
+\title{}
 
@@
 \begin{abstract}
-% Among the approaches for adapting a pretrained NMT model to a specific domain, finetuning is the dominant method.
-Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{Bapna19simple,Vilar18learning}.
-The second approach has several advantages including preserving the pretrained model and the legerity of the adapters.
-However, the behavior of these adapters has not been further studied.
-The objective of this paper is to give an ablation study on the use of residual adapter in the domain adaptation problem.
-% \mpTodo{correcting abstract}
-% **** 
-\fyDone{Citation-free abstract}
-Domain adaptation is an old and vexing problem for machine translation systems.
-The most common approach and successful to supervised adaptation is to fine-tune a baseline system with in-domain parallel data.
-Standard fine-tuning however modifies all the network parameters, which makes this approach computationally costly and prone to overfitting.
-A recent, lightweight approach, instead augments a baseline model with supplementary (small) adapter layers, keeping the rest of the mode unchanged.
-This has the additional merit to leave the baseline model intact, and adaptable to multiple domains.
-In this paper, we conduct a thorough analysis of the adapter model in the context of a multidomain machine translation task.
-We contrast multiple implementations of this idea on two language pairs.
-Our main conclusion is that residual adapter is a fast and cheap method to adapt a model to a domain; reducing the number of adapters incorporated to NMT model can achieve equivalent to using all adapters; residual adapters can be bridged directly to the end of the stack of layers without losing performance.
-Our work also proposed a gated version that renders the NMT model more robust against out-of-domain examples.
-\fyTodo{abstract to be continued}
-
+When building machine translation systems, one often needs to make the best out of heterogeneous sets of parallel data in training to achieve the best performance in certain domains of interest.
+The priority of the in-domain performances in the overall evaluation affects the choice of choosing data.
+The problem falls in the data selection category in the domain adaptation topic.
+Several data selection methods pre-compute the domain-relatedness of training examples and use these scores to build mini-batches in the training while other works propose dynamic strategies to build training mini-batches based on reinforcement learning.
+In this study, we first formulate the most general algorithm for training the Neural Machine Translation (NMT) model given a predefined priority of the domains.
+Then, we apply a recently proposed method Differential Data Selection (DDS) to solve the problem.
+We also report a disadvantage of DDS and fix it by proposing a minor development for DDS.
+Our experiments with a large sample of multi-domain systems show several important benefits of DDS in multi-domain NMT.
 \end{abstract}
-\section{Introduction } \label{sec:intro}
-\mpDone{write introduction} \fyDone{Citations in chronological order}\fyDone{Split long sentences}
-Owing to multiple improvements, Neural Machine Translation(NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence,Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs.
-However, as many deep learning model, NMT models need to be trained with sufficiently large amounts of data to reach their best performance.
-Therefore, the quality of translation of NMT model is still limited in low-resource language and low-resourced domain conditions \cite{duh13adaptation,zoph16transfer,koehn17six}.
-While many approaches have been proposed to improve the quality of NMT models in low-resource domains (see the recent survey of \citet{Chu18asurvey}), full fine-tuning of a generic baseline model remains the dominant supervised approach when adapting NMT models to specific domains \cite{Luong15stanford,neubig18rapid}.
-
-Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pre-trained model by resuming training with the sole in-domain corpus.
-It is a conjecture that pre-trained model constitute a better initialization than a random one, especially when the adaptation data is scarce.
-Indeed, studies of transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim by extensive experiments.
-Full fine-tuning, that adapts all the parameters of a baseline model usually significantly improves the quality of the NMT for the chosen domain.
-However, it also yields large losses in translation quality for other domains, a phenomenon referred to as ``catastrophic forgetting'' in the neural network literature \cite{McCloskey89catastrophic}.
-Therefore, a fully fine-tuned model is only useful to one target domain.
-As the number of domains to handle grows, training and maintaining a separate model for each task can quickly becomes tedious and resource-expensive .\fyDone{Fix this sentence.}
-
-\cite{Vilar18learning,Bapna19simple} have recently proposed a simple and lightweight domain adaptation method that also preserves the value of pre-trained models, based on small adapters component that are plugged in each hidden layer.
-These adapters are trained only with the in-domain data, keeping the pre-trained model frozen.
-Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering a performance that is close to that of full fine-tuning.
-
-In this paper, we would like to extend this architecture to improve NMT in several settings that still challenge automatic translation, such as translating texts from multiple topics, genre or domains, in the face of unbalanced data distribution.
-Furthermore, as the notion of ``domains''  is not always well established, another practical setting is the translation of texts mixing several topics / domains.
-Finally, another requirement is the need to also translate texts from domain unseen in training, based only on the unadapted system, which should be made as strong as possible.
-% Residual adapters allow us to adapt NMT model to any specific domain in a computationally cheap way.
-Could we adapt NMT model to noisy text and topical text at once?
-While residual adapters are good at adapting to one specific domain, their performance still dramatically decrease for the other domains seen in training, as we show in our experiments.
-Therefore, we have to decide whether to use residual adapters manually, i.e, we have to know to which domain the text belongs a priori.
-Therefore, we would like to fuse a domain classifier to the architecture in order to weight the contribution of the adapters with respect to the topic-relatedness of the text.
-\fyDone{Say differently: various implementations}
-
-In this context, our main contribution is a thorough experimental study of the use of residual adapters in multi-domain translation.
-We notably explore ways to adjust and/or regularize adapter modules to handle situations where the adaptation data is very small.
-We also propose and contrast two new variants of the residual architecture: in the first one (highway residual adapters), adaptation still affects each layer of the architecture, but its effect is delayed till the last layer, thus making the architecture more modular and adaptive; in the second one (gated residual adapters), we explore ways to improve the performance in the face of train-test data mismatch.
-We experiment with two language pairs and report results that demonstrate the flexibility and effectiveness of these architectures.
-\fyDone{One bit of a conclusion here}\fyFuture{Build a proper training scenario for these two conditions}
-
-\section{Residual adapters \label{sec:res}}
-
-\subsection{Basic architecture \label{ssec:architecture}}
-\fyDone{More contexts and notations from the transformer}\fyDone{Encoder / decoder layers}
-
-\subsubsection{The computation of adapter layers}
-Our reference architecture is the Transformer model of \cite{Vaswani17attention}, which we assume contains a stack of layers both in the encoder and the decoder side.
-Each layer contains two subparts, an attention layer and a dense layer.
-Details may vary from one implementation to another, we simply content here that each layer $i \in \{1 \dots L\}$ (in the encoder or the decoder) computes a transform of a fixed-length sequence of $d$-dimensional input vectors $h^{i-1}$ into a sequence of output vectors $h^{i}$.
-
-The residual adapter architecture extends this model by adding one new computational module for each layer.
-The $i^{\text{th}}$ residual adapter thus modifies $h^i$ by computing the following transformations:\fyDone{Use align env}
-\begin{align*}
-  h^{i}_1 &= \mathbf{W}_{db}^{i}h^{i} + b^i_{1} \\
-  h^{i}_2 &= \mathbf{ReLU}(h_1^{i})
-  \\
-%\end{align*}
-%\begin{align*}
-  h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + b^i_{2} \\
-  \bar{h}^{i} &= h^{i}_3 + h^i.
-\end{align*}
-Overall, the  $i^{\text{th}}$ adapter is thus parameterized by matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, with $b$ the dimension of the adapter (with $d \gg b$)\fyTodo{Check this}.
-For the sake of brevity, we will simply denote $h^{i}_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the corresponding set of parameters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}
-
-The "adapted" hidden vectors $\bar{h}^i_{ 1\leq i \leq L-1}$, where $L$ is the number of layers, will then be the input of the $(i+1)^{\text{th}}$\fyDone{Self attention ?} layer; $\bar{h}^L$ is passed to the decoder if it belongs to the encoder side, or is the input of output layer if it belongs to the decoder side.
-Note that zeroing out all adapters enables us to recover the basic Transformer, with $\bar{h}^{i} = h^i$ for all $i$.
-
-In the experiments reported in Section~\ref{sec:exp}, we will use $2\times{}L=12$ residual adapters, one for each of the $L=6$ attention layers of the encoder and similarly for the decoder.\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and only serves one residual adapter.}
-
-\subsubsection{Design space and variants}
-This general architecture leaves open many design choices pertaining to the details of the network organization, the training procedure and the corresponding objective function.
-
-A first question concerns the number of adapter layers.
-While in principle, all Transformer layers can be subject to adaptation, it is nonetheless worthwhile to consider simpler adaptation schemes, which would only alter a limited number of layers.
-Such strategy might be especially relevant when the training data contains very small domains, as in the experiments of Section~\ref{sec:exp}, and for which a complete adaptation may not be necessary or/and or prone to overfitting.
-Likewise, it might be meaningful to explore ways to share subsets of adapters across domains.
-This in turn raises the issue of which layer(s) to adapt, a question that can be approached in the light of recent analyses of Transformers models, which conjecture that the higher layers encode global patterns with a more ``semantic'' interpretation, while the lower layers encode local patterns akin to morpho-syntactic information \cite{raganato18analysis}.
-
-% The size of available corpora for each domain is extremely varying.
-For example, the corpus of domain \domain{tourism} in En-De contains only around $7\times10^3$ sentence pairs while the corpus of domain \domain{news} in En-De contains approximately $3\times10^6$ sentence pairs.
-For extremely small domains, we could economize the number of parameters by reducing the number of residual adapters in the architecture.
-While only a limited number of adapters are incorporated to the NMT model, which positions in the model are more important in the domain adaptation task?
-We would like to analyze the impact of position and number of residual adapters involved in the adapted model.
-It is conjectured that the higher layers represent more global patterns such as semantic while the lower layers represent more local patterns such as syntactic \cite{raganato18analysis}\fyTodo{Missing reference here}.
-The domain shift in local patterns and global patterns has not yet studied.
-In this paper, we do not intend to study this aspect of the domain adaptation problem.
-We would like to give an extensive comparison between domain adaptation in different levels in NMT model.
-
-A related question concerns the regularization of adapter layers to mitigate overfitting.
-Reducing the number of adapters, or their dimensions, is simple, but these choices are difficult to optimize numerically -- an issue that becomes important as the number of domain grows.
-Less naïve alternatives can also be entertained, such as applying weight decay or layer regularization to the adapter.
-Implementing these requires to modify the objective function in a way that still allows for a smooth optimization problem.
-For instance, weight decay applies a penalization on the weights of the adapters, complementing the cross-entropy term with a function of the norm of the parameters: \fyDone{The second summation also runs over $x,y$ ?
-I think not}
-\begin{equation*}
-  \begin{split}
-    \bar{L} & = \frac{1}{\#(x,y)}\mathop{\sum}_{x,y}( - \log(P(y|x))) \\
-    & + \lambda  \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{\operatorname{ADAP}^{(i)}}}
-  \end{split}
-\end{equation*}
-An alternative scheme is \emph{layer regularization}, which penalizes the output of the adapters, corresponding to the following new training objective:
-\begin{equation*}
-  \begin{split}
-    \bar{L} & = \frac{1}{\#(x,y)}\mathop{\sum}_{x,y} (-\log(P(y|x)) \\
-    & + \lambda \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\operatorname{ADAP}^{(i)}(h_i(x,y))})
-  \end{split}
-\end{equation*}
-
-Finally, another independant design choice relates to the training strategy for adapters.
-A first option is to generalize supervised domain adaptation to multi-domain adaptation and to proceed in two steps: (a) train a generic model with all the available data; (b) train eacj adapter layer with domain-specific data, keeping the generic model parameters unchanged.
-Another strategy is to adopt the view of \citet{Dredze08online}, where the multi-domain setting is viewed as an instance of multi-task learning \cite{Caruana97multitask}, where each domain corresponds to a specific task.
-This suggest to train the models as it were in a multi-task mode, training all the parameters from scratch: the generic parameters values will still depend on all the available data, while each adapter will only be trained with the corresponding in-domain data.
-
-\subsection{Highway Residual Adapters \label{ssec:highway}}
-
-In the basic architecture described in Section~\ref{ssec:architecture}, the computation performed by lower level layers will impact all the subsequent layers.
-In this section, we introduce an alternative implementation of the same idea, which however delays the adaptation of each layer to the last layer (of the encoder or the decoder) as represented on Figure~\ref{fig:hrl-architecture}.
-While the basic architecture performs adaptation in sequence, we propose here to perform it in parallel.
-In this version, only the last hidden vector of the encoder (decoder) is thus modified according to:
-\begin{equation}
-  \bar{h}^L = h^L + \displaystyle{\mathop{\sum}_{1 \leq i \leq L} ADAP^i(h^i)} \label{eq:highway-output}
-\end{equation}
-
-One obvious benefit of this variant is that it allows us to reuse the hidden vectors $h^i$ of all hidden layers when computing an adapted output for several domains during the inference.
-In this situation, the forward step needs only to compute the hidden vectors $h^i$ once for the inner encoder layers, before an adapted sequence of vectors is computed at the topmost layer.
-Therefore, we can finetune the model to multiple domains at once without recomputing $h^i$.
-This variant also opens the way to more parameter sharing across adapters, a perspective that we will not explore further in this work.
-Instead, we use it to develop a second variation of the adapter model, that is presented in the next section.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[scale=0.3]{fig/highway_residual}
-  \caption{Highway residual adapter network}
-  \label{fig:hrl-architecture}
-\end{figure}
-
-\subsection{Gated Residual Adapters \label{ssec:gate}}
-\mpDone{Formalizing problem, network design, training algorithm}
-The basic architecture presented above rests on a rather simplistic view of ``domains'' as made of well-separated and unrelated pieces of texts that are processed independently during adaptation.
-Likewise, when translating test documents, one needs to make a choice between either using one specific domain-adapted model or resorting to the generic model.
-In this context, using wrong domain labels can have a strong (negative) on translation performance.
-
-Therefore, we would like to design a version of applying residual adapter that is more robust to such domains error.
-This variant, called the \emph{gated residual adapter models}, relies on the training of a supplementary trained component that will help decide whether to activate, on a word per word basis, a given residual layer, and to regulate the strength of this activation.
-To this end , we extend the highway version of residual adapters as follows.
-\fyDone{Consistency of notations wrt section 2.1}
-
-Formally, we replace the adapter computation of equation~\eqref{eq:highway-output} and take the adapted hidden (topmost) layer to be computed as (this is for domain $k$):
-\begin{equation}
-  \bar{h}^L = h^L + \displaystyle{\mathop{\sum}_{1 \leq i \leq L} \operatorname{ADAP}_k^i(h^i) \odot{} z_k(h^L)}, \label{eq:gated-output}
-\end{equation}
-where $z_k(h^L[t]) \in [0,1]$ measures the relatedness of the $t^{\text{th}}$ word $w_t$ to domain $k$.
-The more likely $w_t$ is in domain $k$, the larger $z_k(h^L[t])$ should be; conversely, for words\footnote{The term ``word'' is employed here by mere convenience, as the system only manipulates sub-lexical BPE units; furthermore, the values of the hidden representations $h^{i}$ at position $t$ depend upon all the other positions in the sentence.} that are not typical of any domain $k$ (eg.\ function words),  adaptation is minimum and the adapted corresponding encoder output ($\bar{h}^L[t]$) will remain close to the output of the generic model ($h^L[t]$).
-In our implementation, we incorporate two domain classifiers on top of the encoder and the decoder, that take the last hidden layer of the encoder (resp.\ decoder) as input and use the posterior probability $P(k|h^L[t]);)$ of domain $k$ as the value for $z_k(h^L[t])$.
-
-Training gated residual adapters thus comprises three steps, instead of two for the baseline version:
-\begin{enumerate}
-\item learn a generic model with mixed corpora comprising data from multiple domains.
-\item train a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen.
-This model computes the posterior domain probability $P(k|h^L[t])$ for each word $w_t$, based on the representation computed by the last layer.
-\item train the parameters of adapters with in-domain data separately for each domain, while freezing all the other parameters.
-\end{enumerate}
-\fyFuture{is this classifier important, can we train with the rest of the system ?}
-
-\section{Experimental settings \label{sec:exp}}
-
-\subsection{Data and metrics \label{ssec:corpora}}
-We perform our experiments with two translation pairs involving multiple domains: English-French (En$\rightarrow$Fr and English-German (En$\rightarrow$De).
-For the former pair, we use texts\footnote{Most corpora are available from the Opus web site: \url{http://opus.nlpl.eu}} initially from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
-Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
-Corpus statistics are in Table~\ref{tab:Corpora-en-fr}.
-
-\begin{table*}[htbp]
-  \centering
-  \begin{tabular}{ |lllllll|} %*{4}{|r|}}
-    \hline
-    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
-    \domain{med} & \domain{law} & \domain{bank} & \domain{it} & \domain{talk} & \domain{rel} & \domain{news} \\
-    \hline
-    2609 (0.68) & 190 (0.05)  & 501 (0.13) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
-    \hline
-  \end{tabular}
-\caption{Corpora statistics for En$\rightarrow$Fr : number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain).
-\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.}
-\label{tab:Corpora-en-fr}
-\end{table*}
-
-En$\rightarrow$De is a much larger task, for which we use corpora distributed for the News task of WMT20\footnote{\url{http://www.statmt.org/wmt20/news.html}} including: European Central Bank corpus (\domain{bank}),  European Economic and Social Committee corpus (\domain{eco}), European Medicines Agency corpus (\domain{med})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Press Release Database of European Commission corpus, News Commentary v15 corpus, Common Crawl corpus (\domain{news}), Europarl v10 (\domain{gov}), Tilde MODEL - czechtourism (\domain{tour})\footnote{\url{https://tilde-model.s3-eu-west-1.amazonaws.com/Tilde_MODEL_Corpus.html}}, Paracrawl and Wikipedia Matrix (\domain{web}).
-Statistics are in Table~\ref{tab:Corpora-en-de}.
-\begin{table*}[htbp]
-  \centering
-  \begin{tabular}{ |lllllll|} %*{4}{|r|}}
-    \hline
-    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
-    \domain{bank} & \domain{eco} & \domain{med} & \domain{gov} & \domain{news} & \domain{tour} & \domain{web} \\
-    \hline
-    4 (0.00022) & 2857 (0.15) & 347 (0.018) & 1828 (0.095) & 3696 (0.19) & 7 (0.00039) & 10473 (0.54) \\
-    \hline
-  \end{tabular}
-\caption{Corpora statistics for En$\rightarrow$De: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture.
-\domain{web} is the largest domain, containing about 54\% of the sentences, while \domain{bank} and \domain{tour} are very small.}
-\label{tab:Corpora-en-de}
-\end{table*}
-
-We randomly select in each corpus a development and a test set of 1,000 lines each and keep the rest for training.
-Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
-We report significant differences at the level of $p=0.05$.\fyTodo{Is this part on significance testing still accurate ?}
-
-\subsection{Baseline architectures \label{ssec:baseline}}
-\fyTodo{Write this - settings and parameters for Mixed and Full-FT}.
-% Our baselines are standard for multi-domain systems.
-% \footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.}
-Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we train the following baselines:
-\begin{itemize}
-\item a generic model trained on a concatenation of all corpora, denoted \system{Mixed}\fyDone{Or mixed nat ?}
-\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing during 3 consecutive epochs.
-We again contrast two versions: full fine-tuning (\system{FT-Full}), which update all the parameters of the initial generic model \system{Mixed}; and the variants of \cite{Bapna19simple} (\system{FT-Block}).
-\end{itemize}
-
-For all En$\rightarrow$Fr models, we set the embeddings size and the hidden layers size to~512.
-Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2,048 cells.
-Residual adapters additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$.
-% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
-% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training use a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4,000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}
-
-Models for En$\rightarrow$De are larger and rely on embeddings size as well as hidden layers size of~1024; each Transformers layer contains 16~attention heads; the inner feedforward layer contains 4,096 cells.
-Adapter modules have the same architecture as for the other language pair, except for their size, which is doubled ($b=2,048$).
-%The number of parameters in each model is reported in Table~\ref{tab:params}.
-
-\iffalse{
-\begin{table*}[htbp]
-  \centering
-  \begin{tabular}{|l|l|} \hline
-    Model & params \\ \hline 
-    \system{Transformer-En-Fr}  & 65 \\
-    \system{Residual Adapter-En-Fr} & 1 \\
-    \system{Transformer-En-De}  & 213 \\
-    \system{Residual Adapter-En-De} & 8 \\
-     \hline
-  \end{tabular}
-  \caption{Number of parameters ($\times 10^6$)}
-  \label{tab:params}
-\end{table*}
-}
-\fi
-
-\subsection{Multi-domain systems}
-we compare several proposals from the literature on multi-domain MT to finetuning method. The methods included in our experiments are described as follow
-\begin{itemize}
-\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}) or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain-specific part (\system{LDR});
-\item the three proposals of \newcite{Britz17mixing}.
-\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
-The training uses reference tags and inference is performed with predicted tags, just like for regular target words.
-\system{DM} is a multi-task learner where a domain classifier is trained on top of the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
-These methods thus only use domain tags in training.
-\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connections make it possible to by-pass this module.
-Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
-\end{itemize}
-
-Because of the limit of our computational resources, we limit our experiments in this section to one language pair En-Fr.
-
-The table \ref{tab:performance-multi} shows that the full-finetuning outperforms all other systems in the in-domain tests.
-\system{FT-Block} is able to reduce the margin in several domains, showing the efficacy of residual adapters.
-Multi-task training system \system{MDL-Res} gains important improvement compared to baseline \system{Mixed}.
-However, \system{MDL-Res} is outperformed by \system{FT-Block}.
-It means that individually finetuning generic model is better than multi-task training.
-However, \system{MDL-Res} is better than other methods in multi-domain learning including:  \system{DC-Tag},  \system{DC-Feat}, \system{LDR}, \system{TTM}, \system{DM}, \system{ADM}.
-
-\begin{table*}[htbp]
-  \centering
-  \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
-%     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\
-    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
-   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
-   \hline 
-    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5  & 50.1 \\
-    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 & 49.9 \\
-    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & 79.9 & 49.8 \\
-    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & 79.9 & 49.7  \\
-    \system{DM}             & 35.6 & 49.5  & 45.6 & 29.9 & 37.1 & 62.4 & 43.4 \\ 
-    \system{ADM}           & 36.4 & 53.5  & 48.3 & 32.0 & 41.5 & 73.4 & 47.5\\
-    \system{MDL-Res}     & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
-     \hline
-  \end{tabular}
-  \caption{Translation performance of various MDMT systems in En-Fr. }
-  \label{tab:performance-multi}
-\end{table*}
-
-\subsection{Varying the positions and number of residual adapters}
-By varying the number and positions of residual adapters as aforementioned in section \ref{ssec:architecture}, we assess several variants of applying residual adapters.
-\fyTodo{Fix style here} Because the set of possible configurations is large, we only perform experiments on layers at position: 2, 4, 6 on each side of NMT model (encoder and decoder) which covers a wide range in the hierarchy of attention layers.
-To observe how much the number of residual adapters affects the performance of the adapted model.
-By the limit of computation resources, we limit our study to the case of 3 adapters at positions: 2,4,6 in each side of the NMT model.
-By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting) in each side (encoder and decoder).\fyTodo{We have 12 layers, have we not ?}
-
- Tables ~\ref{tab:performance-en-fr} and ~\ref{tab:performance-en-de-pos-reg} show that in most cases, the performance increases with respect to the number of residual adapters used in architecture.
-The setting \system{FT-Block} using residual adapters at all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only one level.
-However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyTodo{What does bold mean ?}\fyTodo{Number of parameters ?}
-
-\begin{table*}[htbp]
-  \centering
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
-%     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline
-    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 & 65M/0 \\
-    \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 & 65M/12M\\
-    \system{FT-Block-WD}     & 37.18 & 55.99 & 52.93 & 33.36 & 46.03 & 90.65 & 52.7 & 65M/12M \\
-   \system{FT-Block-LR}     & 37.45 & 56.09 & 51.84 & 33.29 & 45.02 & 89.7 & 52.2 & 65M/12M \\  
-    \system{FT-Block$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 & 65M/6M\\
-    \system{FT-Block$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 & 65M/2M \\
-    \system{FT-Block$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.4 & 88.7 & 52 & 65M/2M\\
-   \system{FT-Block$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 & 65M/2M\\
-     \hline
-  \end{tabular}
-  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$Fr).
-We report BLEU scores for each domain, as well as averages across languages.
-In column \domain{params}, we report number of domain-agnostic/domain-specific parameters}
-  \label{tab:performance-en-fr}
-\end{table*}
-
-\begin{table*}[htbp]
-  \centering
-  \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
-%     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline
-    \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 & 213M/0M\\
-   \system{FT-Block}     & 29.65 & 30.45 & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 & 213M/96M \\
-   \system{FT-Block-WD}     & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 & 213M/96M \\
-   \system{FT-Block-LR}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 & 213M/96M\\
-   \system{FT-Block$_{(6)}$}     & 29.47 & 30.39 & 18.13 & 49.14 & 46.95 & 20.45 & 32.42 & 213M/16M \\
-   \system{FT-Block$_{(4)}$}     & 29.69 & 30.4 & 18.07 & 49.61 & 47.05 & 20.64 & 32.58 & 213/16M\\
-   \system{FT-Block$_{(2)}$}   & 29.64 & 30.4 & 18.29 & 49.41 & 46.71 & 20.59 & 32.51  & 213M/16M\\
-   \system{FT-Block$_{(2,4,6)}$}  & 29.68  & 30.55 & 18.85 & 49.57 & 47.09 & 20.63 &  32.73 & 213M/48M \\
-     \hline
-  \end{tabular}
-  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$De).
-We report BLEU scores for each domain, as well as averages across domains.
-In column \domain{params}, we report number of domain-agnostic/domain-specific parameters}
-  \label{tab:performance-en-de-pos-reg}
-\end{table*}
-
-\fyTodo{Why HW worse than standard version ?} \fyTodo{Why is regularization not helping ? It helps for small domain - domain-specific regularization ??}
-
-\subsection{Regularization of fine-tuning}
-In the translation from English into German, we find that domains \domain{tour}, \domain{ecb} are extremely small and account for a very small percentage of the training data, with respective proportions of 0.039\% and 0.022\%.
-Fine-tuning on these domains can lead to serious overfitting.
-We assess two well-known regularization techniques for adapter modules, that could help avoid the overfitting issues: weight decay and layer regularization.
-
-We choose optimal hyper-parameters: weight decay's coefficient, layer regularizer's coefficient $\lambda$ by grid search with a small set of values $\{ 10^{-3}, 10^{-4}, 10^{-5} \}$.
-
-Results in Table~\ref{tab:performance-en-de-pos-reg} show that weight decay (\system{FT-Block-WD}) outperforms the more basic \system{FT-Block} while Layer regularization does not show any improvement compared to \system{FT-Block}.
-\fyTodo{How is the weight decay parameter set ?}
-
-\subsection{Highway and Gated Residual Adaptaters \label{ssec:gate-exp}}
-
-We assess the efficacy of proposed variants: Highway residual adapters \system{FT-HW-Block} and Gated residual adapter \system{FT-Gate-Block} by the same experiments in the previous section except that we only perform them in language pair En-Fr.
-
-To measure the robustness against out-of-domain examples, we perform translations with wrong domain information to trick the systems that need to know domain a priori.
-In the table \ref{tab:performance-random}, the column \domain{RND} stores the score of multi-domain models, that use domain information to translate, in the situation in which we assign random domain tag to each test sentence.
-Besides the \system{RND} test, we use a new domain \domain{news} to assess the performance of systems in out-of-domain examples.
-Given a sentence of \domain{news}, we predict the closest domain to it by choosing the lowest perplexity of the sentence among those computed by in-domain language models and use the corresponding adapters to translate the sentence.
-
-We also report model \system{FT-Full-UEW} which is obtained by full-finetuning the generic model using uniform elastic weight to avoid catastrophic forgetting.
-
-Table \ref{tab:performance-random} shows that variants \system{FT-HW-Block} and \system{FT-Gate-Block} achieve equivalent performances to the original version \system{FT-Block} of \cite{Bapna19simple}.
-
-The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}.
-Our proposed model \system{FT-Gate-Block} maintains equivalent performance to the generic model in \system{RND} thanks to the application of adaptive weight on the output of the residual adapter.
-\system{FT-Gate-Block} avoids relying on the residual adapters when predicting examples that come from domain far from the domain of the adapters.
-
-In the new domain test \domain{news}, generic model outperforms other systems.
-Finetuning systems including \system{FT-Full}, \system{FT-Block} and \system{FT-HW-Block} underperforms generic model by large margin.
-More regularized variants including \system{FT-Full-UEW}, and \system{FT-Gate-Block} are able to reduce the gap of performance compared to generic.
-
-The column \domain{PROC} shows that highway residual adapter \system{FT-HW-Block} process examples faster than the original version \system{FT-Block}.
-
-\begin{table*}[htbp]
-  \centering
-  \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{10}{r|}} \hline
-%     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{RND}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{\domain{PROC}} \\ \hline
-    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 & 49.4 & 23.47 & 34s \\
-   \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 & 32.55 & 20.19 & 34s \\
-   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2  & 53.3 & 38.38 & 20.51 & 22s\\ 
-   \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 & 36.62 & 20.22 & 19s \\
-   \hline 
-    \system{FT-Gate-Block}    & 37.98 &	57.46&	53.05&	33.55&	46.01&	90.07&	53.0&	48.96 & 22.51 & 21s \\
-    \system{FT-Full-UEW} &	37.94 &	55.97 &	52.1 &	33.71 &	44.91 &	89.08 &	52.3 &	46.48 & 22.11& 34s \\
-     \hline
-  \end{tabular}
-  \caption{Translation performance of highway and gated variants. The column \domain{PROC} is time (in seconds) of 100 iterations with batch size 12288 tokens}
-  \label{tab:performance-random}
-\end{table*}
-
-\section{Related Work \label{sec:related}}
-\mpDone{related work}
-
-Training with data from multiple, heterogeneous sources is a common scenario in natural language processing \cite{Dredze08online,Finkel09hierarchical}.
-It is thus no wonder that the design of multi-domain systems has been proposed for many tasks.
-In this short survey, we exclusively focus on machine translation; it is likely that similar methods (parameter sharing, instance selection / weighting, adversarial training, etc) have also been proposed for other tasks.
-
-Early approaches to the multi-domain setting were proposed for statistical MT, either considering multiple data sources (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}) or domains containing several topics \cite{Eidelman12topic,Hasler14dynamic-topic}.
-Two main strategies are considered: feature-based methods, where domain/topic labels are integrated through supplementary feature and instance-based methods, involving a measure of similarity between train and test domains.
-
-The former approach has also been adapted to NMT: \newcite{Kobus17domaincontrol,Tars18multidomain} use an additional domain feature in an RNN model, in the form of an extra domain-token or of additional domain-features associated with each word.
-\citet{Chen16guided} apply domain control on the \emph{target} side, using a topic vector to describe the whole document context.
-Similar ideas are developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and similarities are enforced through parameter sharing schemes.
-Parameter-sharing also lies at the core of the work by \citet{Jiang19multidomain}, who consider a Transformer model containing both domain-specific heads and domain-agnostic heads.
-
-\citet{Britz17mixing} studies three general techniques to take domain information into account in training: they rely on either domain classification or domain normalization on the source or target side.
-One contribution of this study is the proposal of an adversarial training scheme to normalize representations across domains and make the combination of multiple data sources more effective.
-Similar techniques (parameter sharing, automatic domain classification / normalization) are at play in \cite{Zeng18multidomain}: in this work, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific from domain-agnostic representations.
-These two representations are then processed separately, then merged to compute the final translation.
-
-\citet{Farajian17multidomain,li-etal-2018-one} are two recent representatives of the instance-based approach: for each test sentence, a small adaptation corpus is collected based on similarity measures, and used to fine-tune a mix-domain model.
-
-\section{Discussion \label{sec:discussion}}
-\mpDone{discussion}
-Residual Adapters allows us to quickly adapt a model to a domain without creating lots of domain-specific parameters even though full finetuning still outperforms this method by an important gap in in-domain test sets.
-For extremely small domains, we are able to reduce the number of adapters incorporated to the NMT model without losing performance, also using Weight-Decay in finetuning over small domains demonstrates significant improvement.
-The proposed highway variant shows an equivalent performance to the original version of \cite{Bapna19simple} which allows us to process the hidden layer of the NMT model only once.
-The extended version of the highway residual adapter to gated residual adapter shows the benefit in the robustness against wrong domain prediction and out-of-domain examples.
-
-%\section{Acknowledgments}
 
 \bibliographystyle{acl_natbib}
-\bibliography{emnlp2020}
+\bibliography{acl2021}
+
 %\appendix
-%\section{Appendices}
-%\label{sec:appendix}
-%\mpTodo{appendix}
-%\section{Supplemental Material}
-%\label{sec:supplemental}
 

### 2021-01-24 2021/acl-ijcnlp2021/mdmt.tex
rewrite of the abstract

@@
 \begin{abstract}
-When building machine translation systems, one often needs to make the best out of heterogeneous sets of parallel data in training to achieve the best performance in certain domains of interest.
-The priority of the in-domain performances in the overall evaluation affects the choice of choosing data.
+  When building machine translation systems, one often needs to make the best out of heterogeneous sets of parallel data in training to achieve the best test performance in one or several domain(s) of interest.
+This multi-source (multi-)domain) adaptation problem is often approach with instance selection or reweighting strategies, most of which pre-supposes an ad-hoc assessment of relevance for training sentences.
+In this contribution, we study the recently proposed Differential Data Selection (DDS) model  and explore its ability to serve an effective generic framework for these various situations.
+Our experiments for both domain adaptation and multi-domain learning show that DDS often enables to outperform heuristic training strategy; we also introduce a variant that boosts DDS performance in he presence of adapter modules.
+%   This is a well-known issue, which has given raise to a rich litterature in under the umbrellas of domain adaptation or multi-domain
+
+%   The priority of the in-domain performance in the overall evaluation affects the choice of choosing data.
 The problem falls in the data selection category in the domain adaptation topic.
 Several data selection methods pre-compute the domain-relatedness of training examples and use these scores to build mini-batches in the training while other works propose dynamic strategies to build training mini-batches based on reinforcement learning.
-In this study, we first formulate the most general algorithm for training the Neural Machine Translation (NMT) model given a predefined priority of the domains.
-Then, we apply a recently proposed method Differential Data Selection (DDS) to solve the problem.
+% In this study, we first formulate the most general algorithm for training the Neural Machine Translation (NMT) model given a predefined priority of the domains.
+Then, we apply a recently proposed method to solve the problem.
 We also report a disadvantage of DDS and fix it by proposing a minor development for DDS.

### 2021-01-25 2021/acl-ijcnlp2021/mdmt.tex
add algo

--- 2021-01-22 2021/acl-ijcnlp2021/mdmt.tex
@@
 
-\title{}
+\title{Improving Multi-Domain Neural Machine Translation by Differential Data Selection}
 
@@
 
+\section{Introduction}\label{sec:intro}
+
+\section{Formalizing multi-domain translation}
+\begin{algorithm}[h!]
+\caption{Multi-domain Training} \label{alg:mdmt}
+\label{alg:multidomain}
+\begin{algorithmic}[1]
+\REQUIRE {
+\begin{itemize}
+	\item Corpora $C^i, i\in [1,..,K]$ for $K$ domains.
+	\item Dev sets $Dev^i, i\in [1,..,K]$ for $K$ domains.
+	\item Importance weights of domains in the final evaluation $\lambda_{t,d}^i, i\in [1,..,K]$
+	\item Batch size $B$.
+	\item Temporal Data Selection Distribution(T-DSD)$$P_{\system{T-DSD}}(t)(x,y)=\sum_{i}\lambda_{l,d}^i(t)P_i(t)(x,y)$$
+	\item $Eval\_scores = []$
+	\item $Early\_stopping$ criterion.
+\end{itemize}}
+\REPEAT 
+\STATE{Iteration t.}
+\STATE{Randomly pick $i \in [1,..,K]$ w.r.t  $[\lambda_{l,d}^1 \dots \lambda_{l,d}^K](t)$.}
+\STATE{Sample $B$ sentences from $C^i$ with empirical distribution $P_i(t)(x,y)$.}
+\STATE{Update model by applying SGD computed from $B$ sampled sentences.}
+\IF{$t \equiv 0 \mod{eval\_step}$}
+	\STATE{Evaluate current model with K dev sets. Denote $S_t^i$ performance of model at iteration t on domain i.}
+	\STATE{Report weighted score using Domain Importance Weights(DI-Weights) $\lambda_{t,d}^i, i\in [1,..,K]$. $$eval(t) = \displaystyle{\mathop{\sum}_i \lambda_{t,d}^i S_t^i}$$}
+	\STATE{$Eval\_scores.append(eval(t))$.}
+\ENDIF
+\IF{$Early\_stopping(Eval\_scores)$}
+	\STATE{Stop training loop.}
+\ENDIF
+\UNTIL{convergence}
+\end{algorithmic}
+\end{algorithm}
+\section*{Acknowledgments}
 \bibliographystyle{acl_natbib}
-\bibliography{acl2021}
-
+\bibliography{multidomain}
 %\appendix
-
 \end{document}

### 2021-01-25 2021/acl-ijcnlp2021/mdmt.tex
spelling correction

@@
   When building machine translation systems, one often needs to make the best out of heterogeneous sets of parallel data in training to achieve the best test performance in one or several domain(s) of interest.
-This multi-source (multi-)domain) adaptation problem is often approach with instance selection or reweighting strategies, most of which pre-supposes an ad-hoc assessment of relevance for training sentences.
+This multi-source (multi-domain) adaptation problem is often approach with instance selection or reweighting strategies, most of which pre-supposes an ad-hoc assessment of relevance for training sentences.
 In this contribution, we study the recently proposed Differential Data Selection (DDS) model  and explore its ability to serve an effective generic framework for these various situations.
-Our experiments for both domain adaptation and multi-domain learning show that DDS often enables to outperform heuristic training strategy; we also introduce a variant that boosts DDS performance in he presence of adapter modules.
+Our experiments for both domain adaptation and multi-domain learning show that DDS often enables to outperform heuristic training strategy; we also introduce a variant that boosts DDS performance in the presence of adapter modules.
 %   This is a well-known issue, which has given raise to a rich litterature in under the umbrellas of domain adaptation or multi-domain

### 2021-01-25 2021/acl-ijcnlp2021/mdmt.tex
spelling correction

@@
   When building machine translation systems, one often needs to make the best out of heterogeneous sets of parallel data in training to achieve the best test performance in one or several domain(s) of interest.
-This multi-source (multi-domain) adaptation problem is often approach with instance selection or reweighting strategies, most of which pre-supposes an ad-hoc assessment of relevance for training sentences.
+This multi-source (multi-domain) adaptation problem is often approached with instance selection or reweighting strategies, most of which pre-supposes an ad-hoc assessment of relevance for training sentences.
 In this contribution, we study the recently proposed Differential Data Selection (DDS) model  and explore its ability to serve an effective generic framework for these various situations.

### 2021-01-25 2021/acl-ijcnlp2021/mdmt.tex
corrected last sentence

@@
 In this contribution, we study the recently proposed Differential Data Selection (DDS) model  and explore its ability to serve an effective generic framework for these various situations.
-Our experiments for both domain adaptation and multi-domain learning show that DDS often enables to outperform heuristic training strategy; we also introduce a variant that boosts DDS performance in the presence of adapter modules.
+Our experiments for both domain adaptation and multi-domain learning show that DDS often enables to outperform heuristic training strategies; we also introduce a variant that boosts DDS performance of adapter-based multid-omain systems.
 %   This is a well-known issue, which has given raise to a rich litterature in under the umbrellas of domain adaptation or multi-domain

### 2021-01-25 2021/acl-ijcnlp2021/mdmt.tex
spelling correction

--- 2021-01-25 2021/acl-ijcnlp2021/mdmt.tex
@@
 In this contribution, we study the recently proposed Differential Data Selection (DDS) model  and explore its ability to serve an effective generic framework for these various situations.
-Our experiments for both domain adaptation and multi-domain learning show that DDS often enables to outperform heuristic training strategy; we also introduce a variant that boosts DDS performance in the presence of adapter modules.
+Our experiments for both domain adaptation and multi-domain learning show that DDS often enables to outperform heuristic training strategies; we also introduce a variant that boosts DDS performance of adapter-based multi-domain systems.
 %   This is a well-known issue, which has given raise to a rich litterature in under the umbrellas of domain adaptation or multi-domain

### 2021-01-29 2021/acl-ijcnlp2021/mdmt.tex
formalization, multi-dds, baselines

@@
 \usepackage{xspace}
+\usepackage{float}
 \usepackage{amsmath,amsfonts,amssymb}
@@
 \newcommand{\SW}[1]{\underline{#1}}
+% limits underneath
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{arg\,max}
 \renewcommand\textfraction{.1}
@@
 \usepackage{multirow}
+\usepackage{adjustbox}
 
@@
 
-\section{Formalizing multi-domain translation}
-\begin{algorithm}[h!]
+\section{Formalizing multi-domain translation} \label{sec:mdmt}
+We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for source sentences; each domain corresponds to a specific source of data, and differs from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
+Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
+$h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$), but can differ from one domain to the other.
+
+A typical learning scenario in MT is to have access to samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$\revision{, with $\{\lambda^{s}_d, d=1 \dots n_d\}$ the corresponding mixture weights ($\sum_d \lambda^{s}_d=1$)}.
+Multi-domain learning, as defined in \citet{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ of several domains, making the problem distinct from mere domain adaption.
+A multi-domain learner is then expected to use these tags effectively \cite{Joshi12multidomain} when computing the combined translation function $h(x,d)$, and to perform well in all domains \cite{Finkel09hierarchical}.
+This setting is closely related to the multi-source adaptation problem formalized in \cite{Mansour09domainadaptation,Mansour09multiple,Hoffman18algorithms}.
+
+This definition seems to be the most accepted view of a multi-domain MT\footnote{An exception is \citep{Farajian17multidomain}, where test translations rely on similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here.
+Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function.
+
+Given an objective mixture over test distribution $\lambda^{t}_{d}$,$d\in[1,\dots,K]$ and a mixture of training data $\lambda^{s}_{d}$, the Multi-domain training course is determined by the below algorithm \ref{alg:mdmt}.
+The training mixture can be constant or changing through time.
+Optimizing the training mixture is an interesting problem that we would like to address in the paper.
+\begin{algorithm}[H]
 \caption{Multi-domain Training} \label{alg:mdmt}
@@
 \begin{itemize}
-	\item Corpora $C^i, i\in [1,..,K]$ for $K$ domains.
-	\item Dev sets $Dev^i, i\in [1,..,K]$ for $K$ domains.
-	\item Importance weights of domains in the final evaluation $\lambda_{t,d}^i, i\in [1,..,K]$
+	\item Corpora $C^d, d\in [1,..,K]$ for $K$ domains equipped by an empirical distribution $D_d(x,y)$
+	\item Dev sets $Dev^d, d\in [1,..,K]$ for $K$ domains.
+	\item Testing mixture weights in the final evaluation $\lambda^t_d, d\in [1,..,K]$
 	\item Batch size $B$.
-	\item Temporal Data Selection Distribution(T-DSD)$$P_{\system{T-DSD}}(t)(x,y)=\sum_{i}\lambda_{l,d}^i(t)P_i(t)(x,y)$$
+	\item Temporal Data Selection Distribution(T-DSD)$$P_{\system{T-DSD}}(t)(x,y)=\sum_{d}\lambda^s_d(t)D_d(x,y)$$
 	\item $Eval\_scores = []$
@@
 \STATE{Iteration t.}
-\STATE{Randomly pick $i \in [1,..,K]$ w.r.t  $[\lambda_{l,d}^1 \dots \lambda_{l,d}^K](t)$.}
-\STATE{Sample $B$ sentences from $C^i$ with empirical distribution $P_i(t)(x,y)$.}
+\STATE{Randomly pick $d \in [1,..,K]$ w.r.t  $[\lambda^s_1 \dots \lambda^s_K](t)$.}
+\STATE{Sample $B$ sentences from $C^d$ with empirical distribution $D_d(x,y)$.}
 \STATE{Update model by applying SGD computed from $B$ sampled sentences.}
 \IF{$t \equiv 0 \mod{eval\_step}$}
-	\STATE{Evaluate current model with K dev sets. Denote $S_t^i$ performance of model at iteration t on domain i.}
-	\STATE{Report weighted score using Domain Importance Weights(DI-Weights) $\lambda_{t,d}^i, i\in [1,..,K]$. $$eval(t) = \displaystyle{\mathop{\sum}_i \lambda_{t,d}^i S_t^i}$$}
+	\STATE{Evaluate current model with K dev sets. Denote $S^t_d$ performance of model at iteration t on domain d.}
+	\STATE{Report weighted score using testing mixture weights $\lambda^t_d, d\in [1,..,K]$. $$eval(t) = \displaystyle{\mathop{\sum}_d \lambda^t_d S^t_d}$$}
 	\STATE{$Eval\_scores.append(eval(t))$.}
@@
 \end{algorithm}
+\section{Differential Data Selection} \label{sec:dds}
+\cite{Wang20balancing,Wang20learning-multi} formulate the problem optimizing the distribution over the training data in a bi-level optimization problem as follow.
+\begin{equation} \label{eq:bilevel}
+\begin{split}
+D_{trn}^* &= \displaystyle{\mathop{\argmin}_{D_{trn}} J(\theta^*(D_{trn}), D_{dev})} \\
+					& \theta^*(D_{trn}) = \displaystyle{\mathop{\argmin} J(\theta, D_{trn})}
+\end{split}
+\end{equation}
+Where
+\begin{align*}
+J(\theta, D_{dev}) &= \displaystyle{\mathop{\sum}_{x,y \in D_{dev}} l(\theta,x,y)} \\
+J(\theta, D_{trn}) &= \displaystyle{\mathop{E_{x,y \sim D_{trn}}[l(\theta,x,y)]}} \\
+l(\theta,x,y) &= - log P(\theta,x,y)
+\end{align*}
+The problem \ref{eq:bilevel} can be approached by coordinate gradient descent alternating between $D_{trn}$ and $\theta$.
+We limit the search space of $D_{trn}$ to a parameterized differential function $D_{trn}(\psi)$.
+Given $\psi^t$, $\theta^t$ is update via usual gradient descent as follow
+\begin{align*}
+\theta^{t+1} &= \theta^t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta^t, x,y)}{\partial \theta} \\
+x,y &\sim D_{trn}(\psi^t) 
+\\
+\end{align*}
+Given $\theta^t$, $\psi^t$ is updated via Reinforce algorithm as follow
+\begin{align*}
+\psi^{t+1} &= \psi^t + \mathbf{lr}_{data} * R(x,y) * \frac{\partial log(D_{trn}(\psi^t)(x,y))}{\partial \psi} \\
+x,y &\sim D_{trn}(\psi^t) \\
+R(x,y) &= \langle \frac{\partial l(\theta',x,y)}{\partial \theta}, \frac{\partial J(\theta^{t+1},D_{dev})}{\partial \theta} \rangle \\
+\theta' &= \theta^t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta^t, x,y)}{\partial \theta} \\
+\end{align*}
+
+To apply this method to multi-domain training (\system{Multi-DDS}), we limit the search family of training distribution $D_{trn}$ to the family of mixture of K in-domain training distribution such as $\sum_d \lambda^{s}_{d} \mathcal{D}_d(x,y)$.
+We also parameterize $\lambda^{s}_{d}$ using Softmax function so that $\lambda^{s}_{d} = \text{Softmax}(\psi)_d$ where $\psi \in \mathbb{R}^K$.
+
+\section{Experimental settings} \label{sec:exp}
+\subsection{Data and metrics \label{ssec:corpora}}
+We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.
+\revision{We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
+Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
+Most corpora are available from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
+To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
+
+We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
+Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
+
+%for contrast experiments, we also use supplementary test sets from three other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
+This enables us to evaluate the loss in performance when the test set is from a domain not seen in training.
+% The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
+
+\begin{table*}[htbp]
+  \centering
+  \begin{tabular}{|l|ccccccc|} %*{4}{|r|}}
+    \cline{2-8} 
+    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
+    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} & \multicolumn{1}{c|}{\domain{news}} \\
+    \hline 
+    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
+    \# \revision{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 & 7.8 / 9.2   \\
+    \# \revision{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 & - \\
+    \# \revision{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 & - \\
+    \hline
+  \end{tabular}
+  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$).
+\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
+  }
+\label{tab:Corpora}
+\end{table*}
+
+We measure the distance between domains using the $\mathcal{H}$-Divergence \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains.
+%Formally, given test sets of size $m$ for domains $A$ and $B$, and $h(x)$ a trained domain predictor, $\mathcal{H}(A,B)$ is computed %as:
+%$$
+%\mathcal{H}(A,B) = 2(1 - [\frac{1}{m} \sum_{x:h(x) = B} \mathbb{I}( x \in A) + \frac{1}{m} \sum_{x: h(x) = A} \mathbb{I}(x \in B)]),
+% $$
+% where $\mathbb{I}$ is the indicator function.
+Our discriminator is a SVM independently trained for each pair of domains, with sentence representations derived via mean pooling from the source side representation of the generic Transformer model.
+We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well separated from all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
+
+\begin{table}\centering
+  \begin{tabular}{|l*{5}{|r}|} 
+  \cline{2-6}
+  \multicolumn{1}{c|}{} & \domain{law} & \domain{bank} & \domain{talk} & \domain{IT} & \domain{rel} \\ \hline
+    \domain{med} &1.93 &1.97 &1.9 &1.93 &1.97 \\
+    \domain{law}   && 1.94 & 1.97 &1.93 & 1.99 \\
+    \domain{bank} &&&1.98 &1.94 &1.99 \\
+    \domain{talk}   &&&&1.92 &1.93 \\
+     \domain{IT}     &&&&& 1.99 \\ \hline
+  \end{tabular}
+  \caption{The $\mathcal{H}$-divergence between domains}
+  \label{tab:domaindist}
+\end{table}
+\subsection{Baseline architectures \label{ssec:baseline}}
+Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
+
+\begin{itemize}
+\itemsep0em 
+\item Generic models trained with several prefixed mixtures of training data,
+\begin{align} \label{mixture:trn}
+\lambda^s_d = \frac{q_d^{\alpha}}{\displaystyle{\mathop{\sum}_{i=1}^Kq_i^{\alpha}}} &&
+q_i = \frac{\mid D_d \mid}{\displaystyle{\mathop{\sum}_{i=1}^K\mid D_i \mid}}
+\end{align} 
+with $\alpha \in [0,0.25,0.5,0.75,1.0]$. We denote these systems \system{Mixed-$\alpha$}
+%on a concatenation of all corpora (\texttt{Mixed}).
+We develop two versions\footnote{In fact three: to enable a fair comparison with WDCMT, a RNN-based variant is also trained and evaluated.
+\revision{This system appears as \system{Mixed-Nat-RNN} in Table~\ref{tab:performance}}.} of this system, one where the domain unbalance reflects the distribution of our training data \revision{given in Table~\ref{tab:Corpora}} (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
+The former is the best option when the train mixture $\mathcal{D}^s$ is also expected in testing; the latter should be used when the test distribution is uniform across domains.
+Accordingly, we report two aggregate scores: a weighted average reflecting the training distribution, and an unweighted average, meaning that test domains are equally important.
+\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
+The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems adapted for one domain, with no parameter sharing across domains.
+
+\item two multi-domain versions of the approach of \newcite{Bapna19simple}, denoted \system{FT-Res} and \system{MDL-Res}, where a domain-specific adaptation module is added to all the Transformer layers; within each layer, residual connections enable to short-cut this adapter.
+The former variant corresponds to the original proposal of \citet{Bapna19simple} (see also \cite{Sharaf20metalearning}).
+It fine-tunes the adapter modules of a \system{Mixed-$1.0$} system independently for each domain, keeping all the other parameters frozen.
+The latter uses the same architecture, but a different training procedure and learns all parameters jointly from scratch with the prefixed mixtures of training data \ref{mixture:trn}.
+\end{itemize}
+
+All models use embeddings and the hidden layers sizes of dimension~512.
+Transformers contain with 8 attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells.
+The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024.
+Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.
+\subsection{Multi-domain systems} \label{ssec:mdsys}
+Our comparison of multi-domain systems includes our own reimplementations of recent proposals from the literature:\footnote{\revision{Further implementation details are in Appendix~A.}}
+\begin{itemize}
+\itemsep0em 
+\item NMT models trained with Multi-DDS mixture given several objective testing distributions: uniform, uni-domain and natural.
+\item \system{MDL-Res} trained from scratch with Multi-DDS mixture.
+\end{itemize}
+\section{Results and discussion \label{sec:results}}
+\begin{table*}
+  \centering% \small
+  \begin{adjustbox}{width=2\columnwidth,center}
+  \begin{tabular}{|p{3.0cm}|*{13}{r|}} \hline
+    \multirow{2}{*}{Training} & \multicolumn{6}{|c}{$\lambda^t_d$} & \multicolumn{6}{|c|}{BLEU} & \multirow{2}{*}{BLEU} \\ \cline{2-13}	
+  mixture & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & average \\
+    \hline
+  \system{Mixed-$0.0$} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 & 35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
+  \system{Mixed-$0.25$} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 &35.92& 54.87& 52.55& 32.55& 44.98& 90.28& 51.86\\
+  \system{Mixed-$0.5$} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 &36.11& 55.37& 51.76& 33.52& 46.23& 89.99& 52.16\\
+  \system{Mixed-$0.75$} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 &36.53&	55.03& 51.15& 33.98& 44.28& 87.22& 51.365\\
+  \system{Mixed-$1.0$} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+  \system{Multi-DDS} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 & & & & & & & \\
+  \system{Multi-DDS-v1} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 & & & & & & & \\
+  \hline
+  \end{tabular}
+  \end{adjustbox}
+\end{table*}
+\begin{table*}
+  \centering% \small
+  \begin{adjustbox}{width=2\columnwidth,center}
+  \begin{tabular}{|p{2cm}|*{13}{r|}} \hline
+    \multirow{2}{*}{Training} & \multicolumn{6}{|c}{$\lambda^t_d$} & \multicolumn{6}{|c|}{BLEU} & \multirow{2}{*}{BLEU} \\ \cline{2-13}	
+  mixture & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & average \\
+    \hline
+  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
+  \system{Multi-DDS} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
+  \system{FT-Full} & 0.0 &1.0 & 0.0 &0.0 &0.0 & 0.0 &35.92& 54.87& 52.55& 32.55& 44.98& 90.28& 51.86\\
+  \system{Multi-DDS} & 0.0 &1.0 & 0.0 &0.0 &0.0 & 0.0 &35.92& 54.87& 52.55& 32.55& 44.98& 90.28& 51.86\\
+  \system{FT-Full} & 0.0 &0.0 & 1.0 &0.0 &0.0 & 0.0 &36.11& 55.37& 51.76& 33.52& 46.23& 89.99& 52.16\\
+  \system{Multi-DDS} & 0.0 &0.0 & 1.0 &0.0 &0.0 & 0.0 &36.11& 55.37& 51.76& 33.52& 46.23& 89.99& 52.16\\
+  \system{FT-Full} & 0.0 &0.0 & 0.0 &1.0 &0.0 & 0.0 &36.53&	55.03& 51.15& 33.98& 44.28& 87.22& 51.365\\
+  \system{Multi-DDS} & 0.0 &0.0 & 0.0 &1.0 &0.0 & 0.0 &36.53&	55.03& 51.15& 33.98& 44.28& 87.22& 51.365\\
+  \system{FT-Full} & 0.0 &0.0 & 0.0 &0.0 &1.0 & 0.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+  \system{Multi-DDS} & 0.0 &0.0 & 0.0 &0.0 &1.0 & 0.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+  \system{FT-Full} & 0.0 &0.0 & 0.0 &0.0 &0.0 & 1.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+  \system{Multi-DDS} & 0.0 &0.0 & 0.0 &0.0 &0.0 & 1.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+  \hline
+  \end{tabular}
+  \end{adjustbox}
+  \label{tab:redomains}
+\end{table*}
+\section{Related Work \label{sec:related}}
+\section{Conclusion and outlook \label{sec:discussion}}
 \section*{Acknowledgments}

### 2021-01-29 2021/acl-ijcnlp2021/mdmt.tex
Introduction

--- 2021-01-25 2021/acl-ijcnlp2021/mdmt.tex
@@
 
-\title{Improving Multi-Domain Neural Machine Translation by Differential Data Selection}
+\title{Improving Multi-Domain Neural Machine Translation with Differential Data Selection}
 
@@
 \section{Introduction}\label{sec:intro}
+A typical setting in machine translation, is to collect the largest possible collection of parallel data for the chosen language pair, with the intent to achieve optimal performance for a task of interest.
+In such situations, the training data distribution is opportunistic, while the test data distribution is chosen; a key component in training is then to mitigate the detrimental effects of this mismatch between distributions.
+Single-source and multi-source domain adaptation is a well-studied instance of this setting (see \cite{Chu2017comparison} for a review) , and so is multi-domain learning \cite{Chu18multilingual}\cite{Zeng18multidomain}\cite{Jiang19multidomain}.
+A related situation is multi-lingual machine translation \cite{Firat16multiway}\cite{Johnson17google}\cite{Ha16towards}\fytodo{Add more recent work}, where the heterogeneity of training data correspond not only to variation in topic or genre, but also in languages.
+
+\fyTodo{Label or covariate shift ?}
+This problem is often approached with instance selection or reweighting strategies, where the available training data is used in proportion to its relevance for the test conditions.
+[Additional techniques : GANs for feau]
+
+one often needs to make the best out of heterogeneous sets of parallel data in training to achieve the best test performance in one or several domain(s) of interest.
+This multi-source (multi-domain) adaptation problem is often approached with instance selection or reweighting strategies, most of which pre-supposes an ad-hoc assessment of relevance for training sentences.
+In this contribution, we study the recently proposed Differential Data Selection (DDS) model  and explore its ability to serve an effective generic framework for these various situations.
+Our experiments for both domain adaptation and multi-domain learning show that DDS often enables to outperform heuristic training strategies; we also introduce a variant that boosts DDS performance of adapter-based multi-domain systems.
 

### 2021-01-31 2021/acl-ijcnlp2021/mdmt.tex
write section 2, section 3

@@
 \usepackage{microtype}
+\usepackage{graphicx}
 
@@
 In such situations, the training data distribution is opportunistic, while the test data distribution is chosen; a key component in training is then to mitigate the detrimental effects of this mismatch between distributions.
-Single-source and multi-source domain adaptation is a well-studied instance of this setting (see \cite{Chu2017comparison} for a review) , and so is multi-domain learning \cite{Chu18multilingual}\cite{Zeng18multidomain}\cite{Jiang19multidomain}.
-A related situation is multi-lingual machine translation \cite{Firat16multiway}\cite{Johnson17google}\cite{Ha16towards}\fytodo{Add more recent work}, where the heterogeneity of training data correspond not only to variation in topic or genre, but also in languages.
+Single-source and multi-source domain adaptation is a well-studied instance of this setting (see \cite{Chu2017comparison} for a review) , and so is multi-domain learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain}.
+A related situation is multi-lingual machine translation \cite{Firat16multiway,Ha16towards,Johnson17google}\fyTodo{Add more recent work}, where the heterogeneity of training data correspond not only to variation in topic or genre, but also in languages.
 
@@
 
-\section{Formalizing multi-domain translation} \label{sec:mdmt}
-We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for source sentences; each domain corresponds to a specific source of data, and differs from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
+\section{Learning with multiple data sources} \label{sec:mdmt}
+
+We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
 Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
-$h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$), but can differ from one domain to the other.
+$h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$), but might differ from one domain to the other.
 
-A typical learning scenario in MT is to have access to samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$\revision{, with $\{\lambda^{s}_d, d=1 \dots n_d\}$ the corresponding mixture weights ($\sum_d \lambda^{s}_d=1$)}.
-Multi-domain learning, as defined in \citet{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ of several domains, making the problem distinct from mere domain adaption.
-A multi-domain learner is then expected to use these tags effectively \cite{Joshi12multidomain} when computing the combined translation function $h(x,d)$, and to perform well in all domains \cite{Finkel09hierarchical}.
-This setting is closely related to the multi-source adaptation problem formalized in \cite{Mansour09domainadaptation,Mansour09multiple,Hoffman18algorithms}.
+It is usual in MT to opportunistically collect training samples from multiple domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}_{d} \mathcal{D}_d(x)$\revision{, with $\{\lambda^{s}_d, d=1 \dots n_d\}$ the corresponding mixture weights such that $\sum_d \lambda^{s}_d=1$)}.
 
-This definition seems to be the most accepted view of a multi-domain MT\footnote{An exception is \citep{Farajian17multidomain}, where test translations rely on similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here.
-Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function.
+The main challenge is then to make the best of this heterogeneous data, with the aim to achieve the best possible system for the intended test conditions.
+These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unsupervised domain adaptation); in multi-domain adaptation the test distribution is itself a mixture of domains, some of which may also be observed in training.
+Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ - with only one non-null  $\lambda^{t}_{d}$ in the case of domain adaptation.
+These situations are illustrated in Figure~\ref{fig:mdmt-lambdas}.
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.48\textwidth]{mdmt-lambdas}
+  \caption{Training and testing with distribution mismatch.
+We consider just three domains, and represent the mixture weights $\lambda^{s}$ and $\lambda^{t}$ in the 3-dimensional simplex.
+Training with weights in (a) and testing with weights in (c) is standard multi-source domain adaptation to $d_2$, while (b)-(c) is unsupervised adapation; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with configurations (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
+\end{figure}
 
-Given an objective mixture over test distribution $\lambda^{t}_{d}$,$d\in[1,\dots,K]$ and a mixture of training data $\lambda^{s}_{d}$, the Multi-domain training course is determined by the below algorithm \ref{alg:mdmt}.
-The training mixture can be constant or changing through time.
-Optimizing the training mixture is an interesting problem that we would like to address in the paper.
-\begin{algorithm}[H]
-\caption{Multi-domain Training} \label{alg:mdmt}
-\label{alg:multidomain}
-\begin{algorithmic}[1]
-\REQUIRE {
-\begin{itemize}
-	\item Corpora $C^d, d\in [1,..,K]$ for $K$ domains equipped by an empirical distribution $D_d(x,y)$
-	\item Dev sets $Dev^d, d\in [1,..,K]$ for $K$ domains.
-	\item Testing mixture weights in the final evaluation $\lambda^t_d, d\in [1,..,K]$
-	\item Batch size $B$.
-	\item Temporal Data Selection Distribution(T-DSD)$$P_{\system{T-DSD}}(t)(x,y)=\sum_{d}\lambda^s_d(t)D_d(x,y)$$
-	\item $Eval\_scores = []$
-	\item $Early\_stopping$ criterion.
-\end{itemize}}
-\REPEAT 
-\STATE{Iteration t.}
-\STATE{Randomly pick $d \in [1,..,K]$ w.r.t  $[\lambda^s_1 \dots \lambda^s_K](t)$.}
-\STATE{Sample $B$ sentences from $C^d$ with empirical distribution $D_d(x,y)$.}
-\STATE{Update model by applying SGD computed from $B$ sampled sentences.}
-\IF{$t \equiv 0 \mod{eval\_step}$}
-	\STATE{Evaluate current model with K dev sets. Denote $S^t_d$ performance of model at iteration t on domain d.}
-	\STATE{Report weighted score using testing mixture weights $\lambda^t_d, d\in [1,..,K]$. $$eval(t) = \displaystyle{\mathop{\sum}_d \lambda^t_d S^t_d}$$}
-	\STATE{$Eval\_scores.append(eval(t))$.}
-\ENDIF
-\IF{$Early\_stopping(Eval\_scores)$}
-	\STATE{Stop training loop.}
-\ENDIF
-\UNTIL{convergence}
-\end{algorithmic}
-\end{algorithm}
+Theses situations have been amply documented from a theoretical perspective (eg.\ \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}).
+A general recommandation in the DA setting  is to adjust the sampling distribution used to optimize the system so as to compensate for the mismatch between $\mathcal{D}^s(x)$ and $\mathcal{D}^t(x)$.
+This can be approximated by reweighting instances, or more conveniently domains, which are selected during training with a probability $Q^{trnn}(d)$, with $Q^{trn}(d) \neq \lambda^{s}(d)$.
+
+Another practical standard approach to supervised DA is fine-tuning \cite{Luong15stanford,Freitag16fast}, where $q(x)$ is allowed to vary during the course of learning.
+With our notations, this approach amounts to first learning a initial parameter value with all data ($Q^{trn}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain ($Q^{trn}(d) = \lambda^{s}(d) = \lambda^{t}(d)$).
+Note that this strategy may be suboptimal, since some out-of-domain samples may still be useful for the final performance due to eg.\ domain overlap.
+Optimizing the learning distribution in multi-domain learning is even more challenging, as the learner has to best take advantage of domain overlap, and also of the fact that some domains might be easier to learn than others.\fyTodo{How to measure this?}
+
+Differentiable data selection (DDS), introduced in \cite{Wang20optimizing} and later applied to multilingual machine translation in \cite{Wang20balancing}, introduces an additional degree of freedom and allows $Q^{trn}(d)$ to change more often in the course of training, as we now explain.
+
+% === The part on multidomain training has been moved to the end.
+
 \section{Differential Data Selection} \label{sec:dds}
-\cite{Wang20balancing,Wang20learning-multi} formulate the problem optimizing the distribution over the training data in a bi-level optimization problem as follow.
+We assume a training sample consisting of data from $n_d$ domains $d_1 \dots d_{n_d}$; we denote the size of the training subcorpus from domain $d$ as  $N^{trn}_d$, and $N^{trn} = \sum_d N^{trn}_d$ is the total number of training samples.
+We assume a fixed and pre-defined test distribution, approximated with an equivalent number of development corpora as $\widehat{\mathcal{D}^t}(x) = \sum_{d} Q^{dev}(d) \frac{[[x \in d]]}{N^{dev}_d}$, where $[[x \in d]]$ is the indicator function for domain $d$.
+Training is modeled as optimizing the parameters $\theta$ of a neural translation model.
+
+Sampling training instances from a distribution $Q^{trn}(d)$ over domains results in a parameter estimate $\theta(Q^{trn})$, delivering performance that can be estimated using $Q^{dev}$.
+\cite{Wang20optimizing,Wang20balancing} then formulate the problem of finding an optimal sampling $Q^{trn}_{*}$ maximizing performance on the development data as the following bi-level optimization problem.
 \begin{equation} \label{eq:bilevel}
 \begin{split}
-D_{trn}^* &= \displaystyle{\mathop{\argmin}_{D_{trn}} J(\theta^*(D_{trn}), D_{dev})} \\
-					& \theta^*(D_{trn}) = \displaystyle{\mathop{\argmin} J(\theta, D_{trn})}
+Q^{trn}_* &= \displaystyle{\mathop{\argmin}_{Q^{trn}} J(\theta^*(Q^{trn}), Q^{dev})} \\
+					& \theta^*(Q^{trn}) = \displaystyle{\mathop{\argmin} J(\theta, Q^{trn})}, 
 \end{split}
 \end{equation}
-Where
+where we use the following notations:\fyTodo{Overload Q, from domain to instances to pairs}
 \begin{align*}
-J(\theta, D_{dev}) &= \displaystyle{\mathop{\sum}_{x,y \in D_{dev}} l(\theta,x,y)} \\
-J(\theta, D_{trn}) &= \displaystyle{\mathop{E_{x,y \sim D_{trn}}[l(\theta,x,y)]}} \\
+J(\theta, Q^{dev}) &= \displaystyle{\mathop{\sum}_{(x,y) \in Q^{dev}} l(\theta,x,y)} \\
+J(\theta, Q^{trn}) &= \displaystyle{\mathop{E_{(x,y) \sim Q^{trn}}[l(\theta,x,y)]}} \\
 l(\theta,x,y) &= - log P(\theta,x,y)
 \end{align*}
-The problem \ref{eq:bilevel} can be approached by coordinate gradient descent alternating between $D_{trn}$ and $\theta$.
-We limit the search space of $D_{trn}$ to a parameterized differential function $D_{trn}(\psi)$.
-Given $\psi^t$, $\theta^t$ is update via usual gradient descent as follow
+Problem~\eqref{eq:bilevel} can be approached by coordinate gradient descent alternating between $Q^{trn}$ and $\theta$.
+\citet{Wang20balancing} limit the search space of $Q^{trn}$ to a parameterized differentiable function $Q^{trn}(\psi)$.
+Given $\psi_t$ at time step $s$, $\theta_t$ is updated via usual gradient descent as follows:
 \begin{align*}
-\theta^{t+1} &= \theta^t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta^t, x,y)}{\partial \theta} \\
-x,y &\sim D_{trn}(\psi^t) 
-\\
-\end{align*}
-Given $\theta^t$, $\psi^t$ is updated via Reinforce algorithm as follow
-\begin{align*}
-\psi^{t+1} &= \psi^t + \mathbf{lr}_{data} * R(x,y) * \frac{\partial log(D_{trn}(\psi^t)(x,y))}{\partial \psi} \\
-x,y &\sim D_{trn}(\psi^t) \\
-R(x,y) &= \langle \frac{\partial l(\theta',x,y)}{\partial \theta}, \frac{\partial J(\theta^{t+1},D_{dev})}{\partial \theta} \rangle \\
-\theta' &= \theta^t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta^t, x,y)}{\partial \theta} \\
+\theta_{t+1} &= \theta_t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta_t, x,y)}{\partial \theta} \\
+x,y &\sim Q^{trn}(\psi_t) \\
 \end{align*}
 
-To apply this method to multi-domain training (\system{Multi-DDS}), we limit the search family of training distribution $D_{trn}$ to the family of mixture of K in-domain training distribution such as $\sum_d \lambda^{s}_{d} \mathcal{D}_d(x,y)$.
+Given $\theta_t$, $\psi_t$ is updated via the Reinforce algorithm:\fyTodo{Ref Reinforce}
+\begin{align*}
+  \psi_{t+1} &= \psi_t + \mathbf{lr}_{data} * R(x,y) * \\
+  & \quad\quad \frac{\partial log(Q^{trn}(\psi_t)(x,y))}{\partial \psi} \\
+x,y &\sim Q^{trn}(\psi_t) \\
+R(x,y) &= \langle \frac{\partial l(\theta',x,y)}{\partial \theta}, \frac{\partial J(\theta_{t+1},Q^{dev})}{\partial \theta} \rangle \\
+\theta' &= \theta_t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta_t, x,y)}{\partial \theta} \\
+\end{align*}
+
+\fyTodo{Explain lr, may be more}
+% To apply this method to multi-domain training (\system{Multi-DDS}), we limit the search family of training distribution $D_{trn}$ to the family of mixture of K in-domain training distribution such as $\sum_d \lambda^{s}_{d} \mathcal{D}_d(x,y)$.
 We also parameterize $\lambda^{s}_{d}$ using Softmax function so that $\lambda^{s}_{d} = \text{Softmax}(\psi)_d$ where $\psi \in \mathbb{R}^K$.
@@
   \end{tabular}
-  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$).
+  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the training domain mixture (which does not contain \domain{news}), number English and French tokens ($\times 10^6$), number English and French types ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$).
 \domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
@@
 
+\fyTodo{Keep this ?}
 We measure the distance between domains using the $\mathcal{H}$-Divergence \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains.
@@
 \end{table}
+
 \subsection{Baseline architectures \label{ssec:baseline}}
@@
 \itemsep0em 
-\item Generic models trained with several prefixed mixtures of training data,
+\item Generic models trained with various predefined mixtures of the training data taking the form:
 \begin{align} \label{mixture:trn}
-\lambda^s_d = \frac{q_d^{\alpha}}{\displaystyle{\mathop{\sum}_{i=1}^Kq_i^{\alpha}}} &&
-q_i = \frac{\mid D_d \mid}{\displaystyle{\mathop{\sum}_{i=1}^K\mid D_i \mid}}
+Q^{trn}(d) = \frac{q_d^{\alpha}}{\displaystyle{\mathop{\sum}_{d=1}^{n_d}q_d^{\alpha}}} &&
+q_d = \frac{\mid N^{trn}_d \mid}{\displaystyle{N^{trn}}} % \mathop{\sum}_{i=1}^K\mid D_i \mid}}
 \end{align} 
-with $\alpha \in [0,0.25,0.5,0.75,1.0]$. We denote these systems \system{Mixed-$\alpha$}
+with $\alpha \in [0,0.25,0.5,0.75,1.0]$. We denote below these systems \system{Mixed-$\alpha$}
 %on a concatenation of all corpora (\texttt{Mixed}).
@@
 \end{document}
+
+Multi-domain learning, as defined in \citet{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture of several domains, making the problem distinct from mere domain adaption.
+A multi-domain learner is then expected to use these tags effectively \cite{Joshi12multidomain} when computing the combined translation function $h(x,d)$, and to perform well in all domains \cite{Finkel09hierarchical}.
+This setting is closely related to the multi-source adaptation problem formalized in \cite{Mansour09domainadaptation,Mansour09multiple,Hoffman18algorithms}.
+
+This definition seems to be the most accepted view of a multi-domain MT\footnote{An exception is \citep{Farajian17multidomain}, where test translations rely on similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here.
+In the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function.
+
+Given an objective mixture over test distribution $\lambda^{t}_{d}$,$d\in[1,\dots,K]$ and a mixture of training data $\lambda^{s}_{d}$, the Multi-domain training course is determined by the below algorithm \ref{alg:mdmt}.
+The training mixture can be constant or changing through time.
+Optimizing the training mixture is an interesting problem that we would like to address in the paper.
+
+\begin{algorithm}[H]
+\caption{Multi-domain Training} \label{alg:mdmt}
+\label{alg:multidomain}
+\begin{algorithmic}[1]
+\REQUIRE {
+\begin{itemize}
+	\item Corpora $C^d, d\in [1,..,K]$ for $K$ domains equipped by an empirical distribution $D_d(x,y)$
+	\item Dev sets $Dev^d, d\in [1,..,K]$ for $K$ domains.
+	\item Testing mixture weights in the final evaluation $\lambda^t_d, d\in [1,..,K]$
+	\item Batch size $B$.
+	\item Temporal Data Selection Distribution(T-DSD)$$P_{\system{T-DSD}}(t)(x,y)=\sum_{d}\lambda^s_d(t)D_d(x,y)$$
+	\item $Eval\_scores = []$
+	\item $Early\_stopping$ criterion.
+\end{itemize}}
+\REPEAT 
+\STATE{Iteration t.}
+\STATE{Randomly pick $d \in [1,..,K]$ w.r.t  $[\lambda^s_1 \dots \lambda^s_K](t)$.}
+\STATE{Sample $B$ sentences from $C^d$ with empirical distribution $D_d(x,y)$.}
+\STATE{Update model by applying SGD computed from $B$ sampled sentences.}
+\IF{$t \equiv 0 \mod{eval\_step}$}
+	\STATE{Evaluate current model with $K$ dev sets. $S^t_d$ is the performance at iteration $t$ on domain $d$.}
+	\STATE{Report weighted score using testing mixture weights $\lambda^t_d, d\in [1,..,K]$. $$eval(t) = \displaystyle{\mathop{\sum}_d \lambda^t_d S^t_d}$$}
+	\STATE{$Eval\_scores.append(eval(t))$.}
+\ENDIF
+\IF{$Early\_stopping(Eval\_scores)$}
+	\STATE{Stop training loop.}
+\ENDIF
+\UNTIL{convergence}
+\end{algorithmic}
+\end{algorithm}

### 2021-01-31 2021/acl-ijcnlp2021/mdmt.tex
write section 2, section 3

@@
 Single-source and multi-source domain adaptation is a well-studied instance of this setting (see \cite{Chu2017comparison} for a review) , and so is multi-domain learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain}.
-A related situation is multi-lingual machine translation \cite{Firat16multiway,Ha16towards,Johnson17google}\fyTodo{Add more recent work}, where the heterogeneity of training data correspond not only to variation in topic or genre, but also in languages.
+A related situation is multi-lingual machine translation \cite{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work}, where the heterogeneity of training data not only corresponds to variation in topic or genre, but also in language.
 
 \fyTodo{Label or covariate shift ?}
-This problem is often approached with instance selection or reweighting strategies, where the available training data is used in proportion to its relevance for the test conditions.
-[Additional techniques : GANs for feau]
+This problem is often approached with instance selection or reweighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions.
+Finding the optimal balance of training data is however a challenging task due for instance to similarity between domains / languages.
+Several recent proposals \cite{Kumar19reinforcement}
 

### 2021-01-31 2021/acl-ijcnlp2021/mdmt.tex
first introduction

@@
 This problem is often approached with instance selection or reweighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions.
-Finding the optimal balance of training data is however a challenging task due for instance to similarity between domains / languages.
-Several recent proposals \cite{Kumar19reinforcement}
+Finding the optimal balance of training data is however a challenging task due for instance to similarity between domains / languages; it may also be suboptimal, as some domains or languages might be easier to train than others.
+Several recent proposals \cite{Wang17instance,Kumar19reinforcement, Wang20learning-multi} have explored ways to instead consider dynamic data selection strategy.
+We study in this contribution the proposal of \cite{Wang20balancing}, initially introduced for multilingual MT, with the goal to evaluate its potential in various single domain and multi-domain adaptation settings and assess DDS as an effective replacement to standard heuristics approaches.
 
-one often needs to make the best out of heterogeneous sets of parallel data in training to achieve the best test performance in one or several domain(s) of interest.
+Based on experimental results obtained on a diverse set of domains, our main conclusions are that (a) using DDS often yields overall performance that are as good as the standard fine-tuning strategy for domain adaptation; (b) DDS can effectively handle a variety of test target distributions without any meta-parameter search; (c) our extension of DDS is able to boost its performance when used in conjunction to the adapter model of \citet{Bapna19simple}.
+
+% one often needs to make the best out of heterogeneous sets of parallel data in training to achieve the best test performance in one or several domain(s) of interest.
 This multi-source (multi-domain) adaptation problem is often approached with instance selection or reweighting strategies, most of which pre-supposes an ad-hoc assessment of relevance for training sentences.
@@
 \end{itemize}
+
 \section{Results and discussion \label{sec:results}}
@@
 \section{Related Work \label{sec:related}}
+\cite{Wang17instance} uses dynamic weighting (for sentences and domains) for DA.
+
+\cite{Wang20multidomain} for multidomain DA.
+
 \section{Conclusion and outlook \label{sec:discussion}}

### 2021-02-01 2021/acl-ijcnlp2021/mdmt.tex
started related work, more refs

@@
 \fyTodo{Label or covariate shift ?}
-This problem is often approached with instance selection or reweighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions.
-Finding the optimal balance of training data is however a challenging task due for instance to similarity between domains / languages; it may also be suboptimal, as some domains or languages might be easier to train than others.
-Several recent proposals \cite{Wang17instance,Kumar19reinforcement, Wang20learning-multi} have explored ways to instead consider dynamic data selection strategy.
+This problem is often approached with \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions.
+Finding the optimal balance of training data is however a challenging task due for instance to similarity between domains / languages, but also due to regularization effects of out-of-domain data; it may also be suboptimal, as some domains or languages might be easier to train than others.
+Several recent proposals \cite{Wang17instance,Kumar19reinforcement,Wang20learning-multi} have explored ways to instead consider \emph{dynamic} data selection strategy.
 We study in this contribution the proposal of \cite{Wang20balancing}, initially introduced for multilingual MT, with the goal to evaluate its potential in various single domain and multi-domain adaptation settings and assess DDS as an effective replacement to standard heuristics approaches.
@@
 
-\cite{Wang20multidomain} for multidomain DA.
+\cite{Wang20multidomain} adapts curriculum learning \cite{BengioXX} for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
+
+Also \cite{Saunders19adaptive}.
 

### 2021-02-16 2021/acl-ijcnlp2021/mdmt.tex
local changes

--- 2021-01-31 2021/acl-ijcnlp2021/mdmt.tex
@@
   \system{Mixed-$1.0$} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
-  \system{Multi-DDS} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 & & & & & & & \\
-  \system{Multi-DDS-v1} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 & & & & & & & \\
+  \system{Multi-DDS} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 & 37.26 & 55.07&50.96&33.49&43.41&90.76& 51.83 \\
+  \system{Multi-DDS-v1} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 & 36.96 & 55.55 & 52.88 & 33.06 & 44.52 &91.25 & 52.37\\
   \hline
@@
     \hline
-  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
+  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &37.45&51.26&45.47&31.62&37.55&65.22& \\
+  \system{FT-Res} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
   \system{Multi-DDS} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
   \system{FT-Full} & 0.0 &1.0 & 0.0 &0.0 &0.0 & 0.0 &35.92& 54.87& 52.55& 32.55& 44.98& 90.28& 51.86\\
+  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
   \system{Multi-DDS} & 0.0 &1.0 & 0.0 &0.0 &0.0 & 0.0 &35.92& 54.87& 52.55& 32.55& 44.98& 90.28& 51.86\\
   \system{FT-Full} & 0.0 &0.0 & 1.0 &0.0 &0.0 & 0.0 &36.11& 55.37& 51.76& 33.52& 46.23& 89.99& 52.16\\
+  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
   \system{Multi-DDS} & 0.0 &0.0 & 1.0 &0.0 &0.0 & 0.0 &36.11& 55.37& 51.76& 33.52& 46.23& 89.99& 52.16\\
   \system{FT-Full} & 0.0 &0.0 & 0.0 &1.0 &0.0 & 0.0 &36.53&	55.03& 51.15& 33.98& 44.28& 87.22& 51.365\\
+  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
   \system{Multi-DDS} & 0.0 &0.0 & 0.0 &1.0 &0.0 & 0.0 &36.53&	55.03& 51.15& 33.98& 44.28& 87.22& 51.365\\
   \system{FT-Full} & 0.0 &0.0 & 0.0 &0.0 &1.0 & 0.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
   \system{Multi-DDS} & 0.0 &0.0 & 0.0 &0.0 &1.0 & 0.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
   \system{FT-Full} & 0.0 &0.0 & 0.0 &0.0 &0.0 & 1.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
   \system{Multi-DDS} & 0.0 &0.0 & 0.0 &0.0 &0.0 & 1.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\

### 2021-03-04 2021/acl-ijcnlp2021/mdmt.tex
add summary report

--- 2021-02-01 2021/acl-ijcnlp2021/mdmt.tex
@@
 
+\subsection{Sampler's parameterization}
+
+\cite{Wang20balancing} used softmax to parameterize the sampler i.e $Q^{trn}(\psi_t) = \text{Softmax}(\psi_t)$ in which $\psi_t \in \mathbb{R}^K$.
+Parameterizing the sampler by softmax will induce the following gradient.
+\begin{align*}
+\frac{\partial l}{\partial \psi_i} & = \displaystyle{\mathop{\sum}_{j=1}^K} R_j \frac{\partial P_j}{\partial \psi_i} \\
+	& = R_i P_i*(1-P_i) - \displaystyle{\mathop{\sum}_{j\neq i}^K} R_j P_j P_i \\
+	& = P_i(R_i-l)\\
+\end{align*} 
+The fact that the change of $\psi_i$ is proportional to $p_i$ can explain that the optimization often leads to a critical solution in which the sampler ignore almost domains except a few ones.
+In order to avoid the appearance of the
 \section{Experimental settings} \label{sec:exp}
@@
 \end{table*}
+
 \begin{table*}

### 2021-03-04 2021/acl-ijcnlp2021/summary.tex
add summary report

\documentclass[12pt,times,a4paper,twoside]{article}

\usepackage[utf8]{inputenc}	% Para caracteres en español
\usepackage[left=3cm,right=3cm,top=2cm,bottom=3cm]{geometry}
\usepackage{pagenote}

\usepackage{algorithm}
\usepackage{algorithmic}

\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{mathtools,xparse}
\usepackage{mathrsfs}

\usepackage{multirow,booktabs}
\usepackage[table]{xcolor}
\usepackage{fullpage}
\usepackage{lastpage}
\usepackage{enumitem}
\usepackage{fancyhdr}

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{calc}
% \usepackage{color,soul}
\usepackage{multicol}
\usepackage{cancel}
\usepackage[retainorgcmds]{IEEEtrantools}
% \usepackage{xcolor} -- implied by xcolor

\usepackage{listings}
\usepackage{spverbatim}
\usepackage{fancyvrb}
\usepackage{hyperref}
\usepackage{float} 
\usepackage{natbib}

\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}

\usepackage{soulutf8}
\usepackage{tabularx}

\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{caption}
\usepackage{subcaption}

\usepackage[draft]{todo}
\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
\newcommand{\system}[1]{\texttt{{#1}}}
\colorlet{shadecolor}{orange!15}
\parindent 0in
\parskip 12pt
\geometry{margin=1in, headsep=0.25in}
\setlength{\belowdisplayskip}{8pt} \setlength{\belowdisplayshortskip}{8pt}
\setlength{\abovedisplayskip}{8pt} \setlength{\abovedisplayshortskip}{8pt}
\setlist{nosep}
\setlength{\parskip}{0.1cm}
\setlength{\parindent}{1em}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{definition}{Definition}[section]
\DeclareMathOperator*{\argmax}{argmax}
\newcommand{\R}{\ensuremath{\mathbb{R}}}

\pgfkeys{
    /tr/rowfilter/.style 2 args={
        /pgfplots/x filter/.append code={
            \edef\arga{\thisrow{#1}}
            \edef\argb{#2}
            \ifx\arga\argb
            \else
                \def\pgfmathresult{}
            \fi
        }
    }
}

\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\NewDocumentCommand{\normL}{ s O{} m }{%
  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{L_1}%
}

\title{MDMT training}
\author{}
\date{}

\begin{document}
\maketitle

\section{Results}
\subsection{Variants of the original algorithm}
We have found 3 variants of the algorithm which lead to a significant change in the behavior of the algorithm. Those 3 options are the number of updates before computing gradient over validation sets that we denote $sim\_iter$; the parameterization of the sampling distribution; and the regularization on sampling distribution (we chose entropy maximization as regularizing method). We report the performance of the resulting models in table \ref{tab:performance} and the sampling distribution of the domains in figure \ref{fig:sampling}

\begin{table}
  \centering% \small
  \begin{adjustbox}{width=1.0\columnwidth,center}
  \begin{tabular}{|p{3.0cm}|*{13}{r|}} \hline
    \multirow{2}{*}{Name} & \multirow{2}{*}{$sim\_iter$} & \multirow{2}{*}{parameterization} & \multirow{2}{*}{entropy constraint} & \multicolumn{6}{|c|}{BLEU} & \multirow{2}{*}{BLEU average} \\ \cline{5-10}	
   & & & & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} &  \\
    \hline
  \system{$config\_673$} & 10 & linear & yes & 36.02& 56.18& 52.96& 32.06& 45.2 & 90.94 & 52.23 \\
  \system{$config\_675$} & 10 & softmax & yes & 36.24& 55.67& 52.63& 32.74& 44.47& 90.45& 52.03\\
  \system{$config\_616$} & 1 & softmax & yes & 36.96& 55.55& 52.88& 33.06& 44.52& 91.25& 52.37 \\
  \system{$config\_603$} & 1 & softmax & no & 37.26& 55.07& 50.96& 33.49& 43.41& 90.76& 51.83\\
  \hline
  \end{tabular}
  \end{adjustbox}
  \caption{BLEU over multi-domain test sets}
  \label{tab:performance}
\end{table}

\begin{figure}[htb]
\begin{subfigure}{.5\textwidth}
  \centering
  % include first image
  \includegraphics[width=.8\linewidth]{config675.png}  
  \caption{config 675}
  \label{fig:675}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  % include second image
  \includegraphics[width=.8\linewidth]{config673.png}  
  \caption{Config 673}
  \label{fig:673}
\end{subfigure}
\newline
\begin{subfigure}{.5\textwidth}
  \centering
  % include third image
  \includegraphics[width=.8\linewidth]{config616.png}  
  \caption{Config 616}
  \label{fig:616}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  % include fourth image
  \includegraphics[width=.8\linewidth]{config603.png}  
  \caption{Config 603}
  \label{fig:603}
\end{subfigure}
\caption{Sampling distribution}
\label{fig:sampling}
\end{figure}

Changing the number of updates (denoted $sim\_iter$) before computing gradient over the validation sets can be interpreted as a simple approximation of an extension of the original optimization problem. We recall that the sampler's parameter $\psi$ is optimized to reduce the following loss:

\begin{align*}
L(\theta^t,\psi) &=  \displaystyle{\mathop{\sum_{i=1}^k}} L(\theta^{t+1}, Q^{dev}_i) \\
\theta^{t+1} &= \theta^{t+i} - \eta * \frac{\partial l(\theta_{t+i}, x^i,y^i)}{\partial \theta} \\
x^i,y^i &\sim Q^{trn}(\psi)
\end{align*}

We extend the problem to further evaluation point $t+N$ instead of $t+1$ in the training course

\begin{align*}
L(\theta^t,\psi) &=  \displaystyle{\mathop{\sum_{i=1}^k}} L(\theta^{t+N}, Q^{dev}_i) \\
\theta^{t+i+1} &= \theta^t - \eta * \frac{\partial l(\theta_{t+i}, x,y)}{\partial \theta} \forall 0 \leq i \leq N-1\\
x,y &\sim Q^{trn}(\psi)
\end{align*}

By comparing figure \ref{fig:616} and \ref{fig:675} we observe that changing $sim\_iter$ leads to a significant shift in the resulting sampling distribution in which \domain{rel} is reduced while other domains become more frequent.

The second variant is the parameterization of the sampling distribution. We've found that optimizing softmax parameterization without regularization usually leads to a critical point in which one or two domains take all percentages while other domains disappear. Furthermore, the gradient $\psi_i$ in the softmax parameterization has closed-form as follow
\begin{align*}
\frac{\partial l}{\partial \psi_i} & = \displaystyle{\mathop{\sum}_{j=1}^K} R_j \frac{\partial P_j}{\partial \psi_i} \\
	& = R_i P_i*(1-P_i) - \displaystyle{\mathop{\sum}_{j\neq i}^K} R_j P_j P_i \\
	& = P_i(R_i-l)\\
\end{align*} 
which is proportional to the current value $p_i$, which will lead to the "richer get richer" phenomenon. By comparing figure \ref{fig:603} and \ref{fig:673}, we observe that the severe drop in the poor domains is reduced by replacing softmax parameterization with linear parameterization which uses the following updates

\begin{align*}
\psi_i = \frac{\psi_i + \eta * R_i}{ 1 + \eta \displaystyle{\mathop{\sum}_{j=1}^K} R_j}
\end{align*}

Finally, as we forementioned in the previous paragraph, to slow down the phenomenon "richer gets richer", we use entropy maximizing to regularize the sampling distribution, which leads to a better optimal distribution that does not discard poor domains which may be "prematurely" misjudged.

\subsection{Domain adaptation with less catatrophic forgetting}
DDS method could perform domain adaptation by down-weighing the contribution of domains other than the target-domain in the evaluation score to 0. We report fine-tuned systems using the DDS method, $DDS\_FT\_\{ med, bank, law, IT, rel, talk \}$. We report the performances of those systems in the table \ref{tab:ft}. We observe that out-of-interest domains do not lose huge performance as usually reported in the standard fine-tuning method. \fyTodo{to compare with mixed domain adaptation of Dabre}.

\begin{table}[htb]
  \centering% \small
  \begin{adjustbox}{width=1.0\columnwidth,center}
  \begin{tabular}{|p{3.0cm}|*{13}{r|}} \hline
    \multirow{2}{*}{Name} & \multirow{2}{*}{$sim\_iter$} & \multirow{2}{*}{parameterization} & \multirow{2}{*}{entropy constraint} & \multicolumn{6}{|c|}{BLEU} & \multirow{2}{*}{BLEU average} \\ \cline{5-10}	
   & & & & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} &  \\
    \hline
  \system{$DDS\_FT\_med$} & 1 & softmax & no & 36.7&51.14&52&44.32&90.41&33.22&51.3\\
  \system{$DDS\_FT\_law$} & 1 & softmax & no &36.05&56.18&53.57&44.05&91.24&33.09&52.36\\
  \system{$DDS\_FT\_bank$} & 1 & softmax & no &36.15&54.17&54.29&41.33&89.95&31.54&51.24 \\
  \system{$DDS\_FT\_IT$} & 1 & softmax & no &34.39&48.98&52.82&46.8&85.3&31.37&49.94 \\
  \system{$DDS\_FT\_rel$} & 1 & softmax & no & 34.54&52.29&51.46&44.8&91.77&31.84&51.12\\
  \system{$DDS\_FT\_talk$} & 1 & softmax & no & 35.53&50.55&52.64&44.86&85.8&33.47&50.48\\
  \system{$DDS\_FT\_full$} & NA & NA & NA & 37.74&59.21	&54.49&46.81&90.77&33.98&53.83\\
  \hline
  \end{tabular}
  \end{adjustbox}
  \caption{BLEU over multi-domain test sets}
  \label{tab:ft}
\end{table}

\section{Open problems}
\subsection{Convergence rate between domains}
The sampling distribution converges to an optimal point which does not favor hard domains but easy domains. 
\begin{figure}[h!]
\includegraphics[width=\linewidth]{517.png}
\caption{MDMT performance in each domain under uniform sampling distribution}
\label{fig:uniform}
\end{figure}
Under uniform sampling distribution, the mdmt model received equivalent amounts of examples between domains. Figure \ref{fig:uniform} shows that with the same amount of training examples, the performance of the mdmt model approaches FT model's performance in the domains \domain{IT},\domain{bank} and \domain{rel} faster than in the domains \domain{law}, \domain{talk} and \domain{med}. We call domains having fast convergence "easy-domain", otherwise "hard-domain".

The different convergence rate between tasks was also reported in chapter 6 of \cite{Caruana97multitask}

By analyzing the sampling distribution optimized with DDS method, we observe that the resulting distribution is not optimal for harmonizing the heterogeneity of the convergence rate of the domains as an easy-domain such as \domain{rel} have same sampling probability as a hard-domain such as \domain{law}. A better sampling distribution should favor more the harder-domains such as \domain{med} and \domain{law}.

\subsection{Interpreting rewards via the proximity between domains (e.g vocabulary)}
Figure \ref{fig:reward} shows the trending lines of the rewards through the training course. We observe that the rewards of the domains do not have a large gap and they are very close to zero, that means a good domain between those domains simply does not hurt other domains.

Furthermore, the reward of a domain is simply the average of the cosine similarity between the gradient computed from a training batch sampled from that domain and each gradient computed from K dev sets. Reward may reveal the proximity between domains, therefore become a more natural way for instance weighing than such as Moore-Lewis score,
cosine similarity of sentence embedding, etc.

\begin{figure}[h!]
\includegraphics[width=\linewidth]{reward.png}
\caption{rewards of the domains}
\label{fig:reward}
\end{figure}

\section{Conclusion}
DDS method shows an excellent benefit in the domain adaptation problem as the method shows equivalent performance as fine-tuned mt system in the domain of interest while mitigate the significant drop in the other domains.

However, DDS method does not take in account the heterogeneity of the convergence rate of different domains which is also a key problem in mdmt training.
\bibliographystyle{acl_natbib}
\bibliography{multidomain}
\end{document}

### 2021-03-08 2021/acl-ijcnlp2021/mdmt.tex
try to get consistent notations

@@
   When building machine translation systems, one often needs to make the best out of heterogeneous sets of parallel data in training to achieve the best test performance in one or several domain(s) of interest.
-This multi-source (multi-domain) adaptation problem is often approached with instance selection or reweighting strategies, most of which pre-supposes an ad-hoc assessment of relevance for training sentences.
-In this contribution, we study the recently proposed Differential Data Selection (DDS) model  and explore its ability to serve an effective generic framework for these various situations.
-Our experiments for both domain adaptation and multi-domain learning show that DDS often enables to outperform heuristic training strategies; we also introduce a variant that boosts DDS performance of adapter-based multi-domain systems.
+This multi-source (multi-domain) adaptation problem is often approached with instance selection or reweighting strategies, most of which pre-suppose an ad-hoc assessment of the relevance of training instances.
+In this paper, we study the recently proposed Differential Data Selection (DDS) model and explore its ability to serve as a generic framework for these various situations.
+Our experiments study both domain adaptation and multi-domain learning and show that DDS often enables to outperform more heuristic adaptation strategies.
+We also introduce a variant that boosts DDS performance of adapter-based multi-domain systems.\fyTodo{Keep this part ?}
 %   This is a well-known issue, which has given raise to a rich litterature in under the umbrellas of domain adaptation or multi-domain
@@
 \section{Introduction}\label{sec:intro}
-A typical setting in machine translation, is to collect the largest possible collection of parallel data for the chosen language pair, with the intent to achieve optimal performance for a task of interest.
-In such situations, the training data distribution is opportunistic, while the test data distribution is chosen; a key component in training is then to mitigate the detrimental effects of this mismatch between distributions.
-Single-source and multi-source domain adaptation is a well-studied instance of this setting (see \cite{Chu2017comparison} for a review) , and so is multi-domain learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain}.
-A related situation is multi-lingual machine translation \cite{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work}, where the heterogeneity of training data not only corresponds to variation in topic or genre, but also in language.
+A typical setting in machine translation (MT) is to collect the largest possible collection of parallel data for the chosen language pair, with the intent to achieve optimal performance for a task of interest.
+In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key component in training is then to mitigate the detrimental effects of a possible mismatch between these distributions.
+Single-source and multi-source domain adaptation (DA) is a well-studied instance of this setting (see \cite{Chu2017comparison} for a review) , and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+A related situation is multi-lingual machine translation \cite{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
 
@@
 This problem is often approached with \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions.
-Finding the optimal balance of training data is however a challenging task due for instance to similarity between domains / languages, but also due to regularization effects of out-of-domain data; it may also be suboptimal, as some domains or languages might be easier to train than others.
-Several recent proposals \cite{Wang17instance,Kumar19reinforcement,Wang20learning-multi} have explored ways to instead consider \emph{dynamic} data selection strategy.
-We study in this contribution the proposal of \cite{Wang20balancing}, initially introduced for multilingual MT, with the goal to evaluate its potential in various single domain and multi-domain adaptation settings and assess DDS as an effective replacement to standard heuristics approaches.
+Finding the optimal balance of training data is however a challenging task due, for instance, to similarity between domains / languages, but also due to regularization effects of out-of-domain data; it may also be suboptimal, as some target domains or languages might be easier to train than others.
+Several recent proposals \cite{Wang17instance,Kumar19reinforcement,Wang20learning-multi} have explored ways to instead consider \emph{dynamic} data selection strategies.
+We study in this contribution the proposal of \citet{Wang20balancing}, initially introduced for multilingual MT, with the goal to evaluate its potential in various single-domain and multi-domain adaptation settings and assess DDS, both in its basic version and with our improvements, as an effective replacement to standard heuristics approaches.
 
@@
 
-We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
+We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \citep{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
 Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
@@
 
-It is usual in MT to opportunistically collect training samples from multiple domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}_{d} \mathcal{D}_d(x)$\revision{, with $\{\lambda^{s}_d, d=1 \dots n_d\}$ the corresponding mixture weights such that $\sum_d \lambda^{s}_d=1$)}.
+It is usual in MT to opportunistically collect training samples from several domains, which means that training instances are distributed according to the mixture $\mathcal{D}^s$ such that $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}(d) \mathcal{D}_d(x)$, where $\{\lambda^{s}(d), d=1 \dots n_d\}$ are the corresponding mixture weights such that $\sum_d \lambda^{s}(d)=1$).
 
-The main challenge is then to make the best of this heterogeneous data, with the aim to achieve the best possible system for the intended test conditions.
-These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unsupervised domain adaptation); in multi-domain adaptation the test distribution is itself a mixture of domains, some of which may also be observed in training.
-Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ - with only one non-null  $\lambda^{t}_{d}$ in the case of domain adaptation.
+The main challenge is then to make the best of this heterogeneous data, with the aim to achieve the optimal performance for the intended test conditions.
+These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unsupervised domain adaptation); in multi-domain adaptation finally the test distribution is itself a mixture of domains, some of which may also be observed in training.
+Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  $\lambda^{t}(d)$ in the case of domain adaptation.
 These situations are illustrated in Figure~\ref{fig:mdmt-lambdas}.
@@
   \caption{Training and testing with distribution mismatch.
-We consider just three domains, and represent the mixture weights $\lambda^{s}$ and $\lambda^{t}$ in the 3-dimensional simplex.
-Training with weights in (a) and testing with weights in (c) is standard multi-source domain adaptation to $d_2$, while (b)-(c) is unsupervised adapation; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with configurations (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
+We consider just three domains, and represent vectors of mixture weights $\lambda^{s}$ and $\lambda^{t}$ in the 3-dimensional simplex.
+Training with weights in (a) and testing with weights in (c) is standard multi-source domain adaptation to $d_2$, while (b)-(c) is unsupervised adaptation; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with configurations (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
 \end{figure}
 
-Theses situations have been amply documented from a theoretical perspective (eg.\ \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}).
+These situations have been amply documented from a theoretical perspective (eg.\ \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}).
 A general recommandation in the DA setting  is to adjust the sampling distribution used to optimize the system so as to compensate for the mismatch between $\mathcal{D}^s(x)$ and $\mathcal{D}^t(x)$.
-This can be approximated by reweighting instances, or more conveniently domains, which are selected during training with a probability $Q^{trnn}(d)$, with $Q^{trn}(d) \neq \lambda^{s}(d)$.
+This can be approximated by reweighting instances, or more conveniently domains, which are selected during training with a probability $\lambda^{l}(d)$, with $\lambda^{l}(d) \neq \lambda^{s}(d)$.
 
-Another practical standard approach to supervised DA is fine-tuning \cite{Luong15stanford,Freitag16fast}, where $q(x)$ is allowed to vary during the course of learning.
-With our notations, this approach amounts to first learning a initial parameter value with all data ($Q^{trn}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain ($Q^{trn}(d) = \lambda^{s}(d) = \lambda^{t}(d)$).
-Note that this strategy may be suboptimal, since some out-of-domain samples may still be useful for the final performance due to eg.\ domain overlap.
-Optimizing the learning distribution in multi-domain learning is even more challenging, as the learner has to best take advantage of domain overlap, and also of the fact that some domains might be easier to learn than others.\fyTodo{How to measure this?}
+A more practical standard approach to supervised DA is \emph{fine-tuning} \cite{Luong15stanford,Freitag16fast}, where $\lambda^{l}$ is allowed to vary during the course of learning.
+With our notations, this approach amounts to first learning an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$) with $\indic{A}$ the indicator function for predicate $A$.
+Note that this strategy is potentially suboptimal, as some out-of-domain samples may contribute to the final performance due to eg.\ domain overlap.
+Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains are easier to learn than others.\fyTodo{How to measure this?}
 
-Differentiable data selection (DDS), introduced in \cite{Wang20optimizing} and later applied to multilingual machine translation in \cite{Wang20balancing}, introduces an additional degree of freedom and allows $Q^{trn}(d)$ to change more often in the course of training, as we now explain.
+Differentiable data selection (DDS), proposed by \citet{Wang20optimizing} and later applied to multilingual machine translation by \citet{Wang20balancing}, introduces an additional degree of freedom and allows $\lambda^{l}(d)$ to change multiple times in the course of training, as we now explain.
 
@@
 \section{Differential Data Selection} \label{sec:dds}
-We assume a training sample consisting of data from $n_d$ domains $d_1 \dots d_{n_d}$; we denote the size of the training subcorpus from domain $d$ as  $N^{trn}_d$, and $N^{trn} = \sum_d N^{trn}_d$ is the total number of training samples.
-We assume a fixed and pre-defined test distribution, approximated with an equivalent number of development corpora as $\widehat{\mathcal{D}^t}(x) = \sum_{d} Q^{dev}(d) \frac{[[x \in d]]}{N^{dev}_d}$, where $[[x \in d]]$ is the indicator function for domain $d$.
-Training is modeled as optimizing the parameters $\theta$ of a neural translation model.
+Assume training data from $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as  $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
+We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t}_d$ to denote the empirical train and test distributions for domain $d$, and $\widehat{\mathcal{D}}(x;\lambda) = \sum_{d} \lambda(d) \widehat{\mathcal{D}^l_d}(x)$.
+In our setting,  $\lambda^t$, and hence $\widehat{\mathcal{D}^t}(x;\lambda^t)$ are fixed and predefined, approximated with an equivalent number of development corpora.
+Training optimizes the parameters $\theta$ of a neural translation model.
 
-Sampling training instances from a distribution $Q^{trn}(d)$ over domains results in a parameter estimate $\theta(Q^{trn})$, delivering performance that can be estimated using $Q^{dev}$.
-\cite{Wang20optimizing,Wang20balancing} then formulate the problem of finding an optimal sampling $Q^{trn}_{*}$ maximizing performance on the development data as the following bi-level optimization problem.
+Sampling training domains using a distribution $\lambda^{l}(d)$ results in a parameter estimate $\theta(\lambda^{l})$, delivering performance that can be estimated using $\lambda^{t}$.
+\cite{Wang20optimizing,Wang20balancing} then formulate the problem of finding an optimal sampling strategy $\lambda^{l}_{*}$ maximizing performance on the development data as the following bi-level optimization problem:
 \begin{equation} \label{eq:bilevel}
 \begin{split}
-Q^{trn}_* &= \displaystyle{\mathop{\argmin}_{Q^{trn}} J(\theta^*(Q^{trn}), Q^{dev})} \\
-					& \theta^*(Q^{trn}) = \displaystyle{\mathop{\argmin} J(\theta, Q^{trn})}, 
+\lambda^{l}_* &= \displaystyle{\mathop{\argmin}_{\lambda^{l}} J(\theta^*(\lambda^{l}), \lambda^{t})} \\
+					& \theta^*(\lambda^{l}) = \displaystyle{\mathop{\argmin_{\theta}} J(\theta, \lambda^{l})}, 
 \end{split}
 \end{equation}
-where we use the following notations:\fyTodo{Overload Q, from domain to instances to pairs}
+where we use the following notations:% \fyTodo{Overload Q, from domain to instances to pairs}
 \begin{align*}
-J(\theta, Q^{dev}) &= \displaystyle{\mathop{\sum}_{(x,y) \in Q^{dev}} l(\theta,x,y)} \\
-J(\theta, Q^{trn}) &= \displaystyle{\mathop{E_{(x,y) \sim Q^{trn}}[l(\theta,x,y)]}} \\
+J(\theta, \lambda^{t} ) &= \displaystyle{\mathop{\sum}_{(x,y) \sim \widehat{\mathcal{D}^{t}(x)}} \ell(\theta,x,y)} \\
+J(\theta, \lambda^{l}) &= \displaystyle{\mathop{E_{(x,y) \sim \widehat{\mathcal{D}^{l}(x)}}[\ell(\theta,x,y)]}} \\
 l(\theta,x,y) &= - log P(\theta,x,y)
 \end{align*}
-Problem~\eqref{eq:bilevel} can be approached by coordinate gradient descent alternating between $Q^{trn}$ and $\theta$.
-\citet{Wang20balancing} limit the search space of $Q^{trn}$ to a parameterized differentiable function $Q^{trn}(\psi)$.
-Given $\psi_t$ at time step $s$, $\theta_t$ is updated via usual gradient descent as follows:
+Problem~\eqref{eq:bilevel} can be approached by coordinate gradient descent alternating between $\lambda^{l}$ and $\theta$.
+\citet{Wang20balancing} limit the search space of $\lambda^{l}$ to a parameterized differentiable function $\lambda^{l}(d; \psi)$.
+Given $\psi_t$ at time step $t$, $\theta_t$ is updated via usual gradient descent as follows:
 \begin{align*}
 \theta_{t+1} &= \theta_t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta_t, x,y)}{\partial \theta} \\
-x,y &\sim Q^{trn}(\psi_t) \\
+x,y &\sim \widehat{\mathcal{D}^l}(x;\lambda(\psi_t)) \\
 \end{align*}
 
-Given $\theta_t$, $\psi_t$ is updated via the Reinforce algorithm:\fyTodo{Ref Reinforce}
+Given $\theta_t$, $\psi_t$ is updated via the Reinforce algorithm \cite{Williams92reinforce}:\fyTodo{Ref Reinforce}
 \begin{align*}
   \psi_{t+1} &= \psi_t + \mathbf{lr}_{data} * R(x,y) * \\
-  & \quad\quad \frac{\partial log(Q^{trn}(\psi_t)(x,y))}{\partial \psi} \\
-x,y &\sim Q^{trn}(\psi_t) \\
-R(x,y) &= \langle \frac{\partial l(\theta',x,y)}{\partial \theta}, \frac{\partial J(\theta_{t+1},Q^{dev})}{\partial \theta} \rangle \\
+  & \quad\quad \frac{\partial log(\lambda^l(\psi_t)(x,y))}{\partial \psi} \\
+x,y &\sim \lambda^l (\psi_t) \\
+R(x,y) &= \langle \frac{\partial l(\theta',x,y)}{\partial \theta}, \frac{\partial J(\theta_{t+1}, \lambda^l)}{\partial \theta} \rangle \\
 \theta' &= \theta_t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta_t, x,y)}{\partial \theta} \\
@@
 
-\subsection{Sampler's parameterization}
+\subsection{Parameterizing the sampler \label{ssec:sampler}}
 
-\cite{Wang20balancing} used softmax to parameterize the sampler i.e $Q^{trn}(\psi_t) = \text{Softmax}(\psi_t)$ in which $\psi_t \in \mathbb{R}^K$.
-Parameterizing the sampler by softmax will induce the following gradient.
+\cite{Wang20balancing} used a simple softmax function and the following parameterization of the sampler $\lambda^l(d;\psi) = \operatorname{softmax}(\psi)$ where $\psi \in \mathbb{R}^{n_d}$.
+This parameterization of the sampler will induce the following gradient:
 \begin{align*}
-\frac{\partial l}{\partial \psi_i} & = \displaystyle{\mathop{\sum}_{j=1}^K} R_j \frac{\partial P_j}{\partial \psi_i} \\
-	& = R_i P_i*(1-P_i) - \displaystyle{\mathop{\sum}_{j\neq i}^K} R_j P_j P_i \\
-	& = P_i(R_i-l)\\
+\frac{\partial l}{\partial \psi_d} & = \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \frac{\partial \lambda^l(j; \psi)}{\partial \psi_d} \\
+	& = R_d \lambda^l(d; \psi)(1- \lambda^l(d; \psi)) - \displaystyle{\mathop{\sum}_{j\neq d}^{n_d}} R_j \lambda^l(d; \psi) \lambda^l(j; \psi)\\
+	& = P_d(R_d-l)\\
 \end{align*} 
-The fact that the change of $\psi_i$ is proportional to $p_i$ can explain that the optimization often leads to a critical solution in which the sampler ignore almost domains except a few ones.
+The fact that the change of $\psi_d$ is proportional to $\lambda^l(d; \psi)$ can explain that the optimization often leads to a degenerate solution in which the sampler ignore almost domains except a few ones.
 In order to avoid the appearance of the
+
 \section{Experimental settings} \label{sec:exp}
@@
 Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.
+
 \subsection{Multi-domain systems} \label{ssec:mdsys}
-Our comparison of multi-domain systems includes our own reimplementations of recent proposals from the literature:\footnote{\revision{Further implementation details are in Appendix~A.}}
+
+Our comparison of multi-domain systems includes our own reimplementations of recent proposals from the literature:\footnote{\revision{Further implementation details are in supplementary material.}}
 \begin{itemize}
@@
 \end{table*}
+
 \section{Related Work \label{sec:related}}
@@
 
-\cite{Wang20multidomain} adapts curriculum learning \cite{BengioXX} for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
+\cite{Wang20learning-multi} adapts curriculum learning \cite{BengioXX} for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
 

### 2021-03-09 2021/acl-ijcnlp2021/mdmt.tex
continue the integration, one more ref

@@
 \usepackage{algorithm, algorithmic}
+
+\usepackage{tikz}
+\usepackage{pgfplots}
+\usepackage{pgfplotstable}
+
+\usepackage{soulutf8}
+\usepackage{tabularx}
+
+\usepackage{multirow}
+\usepackage{adjustbox}
+\usepackage{caption}
+\usepackage{subcaption}
+
 \newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
@@
 \section{Differential Data Selection} \label{sec:dds}
+\subsection{Basic principles}
 Assume training data from $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as  $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
-We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t}_d$ to denote the empirical train and test distributions for domain $d$, and $\widehat{\mathcal{D}}(x;\lambda) = \sum_{d} \lambda(d) \widehat{\mathcal{D}^l_d}(x)$.
+We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t _d}$ to denote the empirical train and test distributions for domain $d$, and $\widehat{\mathcal{D}}(x;\lambda) = \sum_{d} \lambda(d) \widehat{\mathcal{D}^l_d}(x)$.
 In our setting,  $\lambda^t$, and hence $\widehat{\mathcal{D}^t}(x;\lambda^t)$ are fixed and predefined, approximated with an equivalent number of development corpora.
@@
 
-Sampling training domains using a distribution $\lambda^{l}(d)$ results in a parameter estimate $\theta(\lambda^{l})$, delivering performance that can be estimated using $\lambda^{t}$.
+Sampling training domains using a distribution $\lambda^{l}(d)$ results in a parameter estimate $\theta(\lambda^{l})$, delivering performance that can be estimated on the development set using $\lambda^{t}$.
 \cite{Wang20optimizing,Wang20balancing} then formulate the problem of finding an optimal sampling strategy $\lambda^{l}_{*}$ maximizing performance on the development data as the following bi-level optimization problem:
@@
 
-Given $\theta_t$, $\psi_t$ is updated via the Reinforce algorithm \cite{Williams92reinforce}:\fyTodo{Ref Reinforce}
+Given $\theta_t$, $\psi_t$ is updated via the Reinforce algorithm \cite{Williams92reinforce}:\fyTodo{Rewrite this correctly with domain average reward}
 \begin{align*}
@@
 
-\fyTodo{Explain lr, may be more}
+\fyTodo{Explain lr, may be more}\fyTodo{More refs to the algorithm }
 % To apply this method to multi-domain training (\system{Multi-DDS}), we limit the search family of training distribution $D_{trn}$ to the family of mixture of K in-domain training distribution such as $\sum_d \lambda^{s}_{d} \mathcal{D}_d(x,y)$.
 We also parameterize $\lambda^{s}_{d}$ using Softmax function so that $\lambda^{s}_{d} = \text{Softmax}(\psi)_d$ where $\psi \in \mathbb{R}^K$.
+\subsection{DDS for (multi) domain adaptation}
 
-\subsection{Parameterizing the sampler \label{ssec:sampler}}
+This setting is pretty general and can in principle accomodate the variety of situations mentioned  above, and many more: basic domain adaption, multi-domain adaptation with various target distribution, possibly including domains unseen in training.
+In our experiments we would like to better assess the actual potential of DDS is these settings, and seek to experimentally answer the following questions:
+\begin{itemize}
+\item is DDS a viable alternative to conventional fine-tuning ? In particular does it enable to better take advantage of relevant data from other domains ?
+\item is DDS a viable alternative in multidomain scenarios ?
+\item \fyTodo{TBContinued}
+\end{itemize}
 
-\cite{Wang20balancing} used a simple softmax function and the following parameterization of the sampler $\lambda^l(d;\psi) = \operatorname{softmax}(\psi)$ where $\psi \in \mathbb{R}^{n_d}$.
-This parameterization of the sampler will induce the following gradient:
-\begin{align*}
-\frac{\partial l}{\partial \psi_d} & = \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \frac{\partial \lambda^l(j; \psi)}{\partial \psi_d} \\
-	& = R_d \lambda^l(d; \psi)(1- \lambda^l(d; \psi)) - \displaystyle{\mathop{\sum}_{j\neq d}^{n_d}} R_j \lambda^l(d; \psi) \lambda^l(j; \psi)\\
-	& = P_d(R_d-l)\\
-\end{align*} 
-The fact that the change of $\psi_d$ is proportional to $\lambda^l(d; \psi)$ can explain that the optimization often leads to a degenerate solution in which the sampler ignore almost domains except a few ones.
-In order to avoid the appearance of the
+These questions are further discussed in Section~\ref{}. We now turn to our experimental conditions.
+\section{Experimental settings} \label{sec:exp}
 
-\section{Experimental settings} \label{sec:exp}
 \subsection{Data and metrics \label{ssec:corpora}}
@@
 \begin{align} \label{mixture:trn}
-Q^{trn}(d) = \frac{q_d^{\alpha}}{\displaystyle{\mathop{\sum}_{d=1}^{n_d}q_d^{\alpha}}} &&
-q_d = \frac{\mid N^{trn}_d \mid}{\displaystyle{N^{trn}}} % \mathop{\sum}_{i=1}^K\mid D_i \mid}}
+\lambda^{l}(d;\alpha) = \frac{q_d^{\alpha}}{\displaystyle{\mathop{\sum}_{d=1}^{n_d}q_d^{\alpha}}} &&
+q_d = \frac{\mid N^{s}_d \mid}{\displaystyle{N^{s}}} % \mathop{\sum}_{i=1}^K\mid D_i \mid}}
 \end{align} 
-with $\alpha \in [0,0.25,0.5,0.75,1.0]$. We denote below these systems \system{Mixed-$\alpha$}
+with $\alpha \in [0,0.25,0.5,0.75,1.0]$. These systems are denoted \system{Mixed-$\alpha$} below.
 %on a concatenation of all corpora (\texttt{Mixed}).
@@
 \item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
-The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems adapted for one domain, with no parameter sharing across domains.
+The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems each adapted for one domain, with no parameter sharing across domains.
 
@@
 It fine-tunes the adapter modules of a \system{Mixed-$1.0$} system independently for each domain, keeping all the other parameters frozen.
-The latter uses the same architecture, but a different training procedure and learns all parameters jointly from scratch with the prefixed mixtures of training data \ref{mixture:trn}.
+The latter uses the same architecture, but a different training procedure and learns all parameters jointly from scratch with the prefixed mixtures of training data \ref{mixture:trn}.\fyTodo{Keep this ?}
 \end{itemize}
@@
 All models use embeddings and the hidden layers sizes of dimension~512.
-Transformers contain with 8 attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells.
+Transformer models contain with 8 attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells.
 The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024.
@@
 \section{Results and discussion \label{sec:results}}
+
+\subsection{Fine tuning experiments}
+
+\fyTodo{Fix the table}
+In Table~\ref{tab:finetuning}, we compare the results of Fine tuning a system initially trained with a uniform mixture of domains with the result of learning from scratch, also with a initial uniform distribution ($\phi_0(d) = 0 \forall d$).
+\fyTodo{What do we see ?
+How about domain proximity ?
+- show a graph}.
+Note that the DDS systems are somewhat quicker to build, as we only train for 100K iterations (instead of 100K + 50K for the fine tuning part).
+
+% \begin{table*}
+%   \centering% \small
+%   \begin{adjustbox}{width=2\columnwidth,center}
+%   \begin{tabular}{|p{2cm}|*{13}{r|}} \hline
+%     \multirow{2}{*}{Training} & \multicolumn{6}{|c}{$\lambda^t_d$} & \multicolumn{6}{|c|}{BLEU} & \multirow{2}{*}{BLEU} \\ \cline{2-13}	
+%   mixture & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & average \\
+%     \hline
+%   \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &37.45&51.26&45.47&31.62&37.55&65.22& \\
+%   \system{FT-Res} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
+%   \system{Multi-DDS} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
+%   \system{FT-Full} & 0.0 &1.0 & 0.0 &0.0 &0.0 & 0.0 &35.92& 54.87& 52.55& 32.55& 44.98& 90.28& 51.86\\
+%   \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
+%   \system{Multi-DDS} & 0.0 &1.0 & 0.0 &0.0 &0.0 & 0.0 &35.92& 54.87& 52.55& 32.55& 44.98& 90.28& 51.86\\
+%   \system{FT-Full} & 0.0 &0.0 & 1.0 &0.0 &0.0 & 0.0 &36.11& 55.37& 51.76& 33.52& 46.23& 89.99& 52.16\\
+%   \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
+%   \system{Multi-DDS} & 0.0 &0.0 & 1.0 &0.0 &0.0 & 0.0 &36.11& 55.37& 51.76& 33.52& 46.23& 89.99& 52.16\\
+%   \system{FT-Full} & 0.0 &0.0 & 0.0 &1.0 &0.0 & 0.0 &36.53&	55.03& 51.15& 33.98& 44.28& 87.22& 51.365\\
+%   \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
+%   \system{Multi-DDS} & 0.0 &0.0 & 0.0 &1.0 &0.0 & 0.0 &36.53&	55.03& 51.15& 33.98& 44.28& 87.22& 51.365\\
+%   \system{FT-Full} & 0.0 &0.0 & 0.0 &0.0 &1.0 & 0.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+%   \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
+%   \system{Multi-DDS} & 0.0 &0.0 & 0.0 &0.0 &1.0 & 0.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+%   \system{FT-Full} & 0.0 &0.0 & 0.0 &0.0 &0.0 & 1.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+%   \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
+%   \system{Multi-DDS} & 0.0 &0.0 & 0.0 &0.0 &0.0 & 1.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+%   \hline
+%   \end{tabular}
+%   \end{adjustbox}
+%   \label{tab:redomains}
+% \end{table*}
+
+\begin{table*}[htb]
+  \centering% \small
+  \begin{adjustbox}{width=1.0\textwidth,center}
+  \begin{tabular}{|p{3.0cm}|*{13}{r|}} \hline
+    \multirow{2}{*}{Name} & \multirow{2}{*}{$sim\_iter$} & \multirow{2}{*}{parameterization} & \multirow{2}{*}{entropy constraint} & \multicolumn{6}{|c|}{BLEU} & \multirow{2}{*}{BLEU average} \\ \cline{5-10}
+   & & & & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{it}} & \multicolumn{1}{c|}{\domain{ rel }} & \multicolumn{1}{c|}{\domain{ talk}} &  \\
+    \hline
+  \system{$DDS\_FT (med)$} & 1 & softmax & no & 36.7&51.14&52&44.32&90.41&33.22&51.3\\
+  \system{$DDS\_FT\_law$} & 1 & softmax & no &36.05&56.18&53.57&44.05&91.24&33.09&52.36\\
+  \system{$DDS\_FT\_bank$} & 1 & softmax & no &36.15&54.17&54.29&41.33&89.95&31.54&51.24 \\
+  \system{$DDS\_FT\_IT$} & 1 & softmax & no &34.39&48.98&52.82&46.8&85.3&31.37&49.94 \\
+  \system{$DDS\_FT\_rel$} & 1 & softmax & no & 34.54&52.29&51.46&44.8&91.77&31.84&51.12\\
+  \system{$DDS\_FT\_talk$} & 1 & softmax & no & 35.53&50.55&52.64&44.86&85.8&33.47&50.48\\
+  \system{$DDS\_FT\_full$} & NA & NA & NA & 37.74&59.21	&54.49&46.81&90.77&33.98&53.83\\
+    \hline
+    $\Delta$ & \\
+  \end{tabular}
+  \end{adjustbox}
+  \caption{DDS as a fine-tuning receipe.
+We report scores with DDS and each target domain, and compare with full-fine-tuning.
+Note that the last line reports the results of 6 different fine tuning procedures.}
+  \label{tab:finetuning}
+\end{table*}
+
+\subsection{Multi-domain experiments}
+Results for our multidomain experiments are in Table~\ref{tab:multidomains}.
+
 \begin{table*}
@@
 \end{adjustbox}
+  \label{tab:multidomains}
 \end{table*}
 
+Looking again at the evolution of $\lambda^l$ during training we often observe a rich get richer effect, where a small number of domains manage to attract most of the probability mass, while other are completely ignore after a given number of iterations.
+
+\fyTodo{Explain issues}
+
 \begin{table*}
-  \centering% \small
-  \begin{adjustbox}{width=2\columnwidth,center}
-  \begin{tabular}{|p{2cm}|*{13}{r|}} \hline
-    \multirow{2}{*}{Training} & \multicolumn{6}{|c}{$\lambda^t_d$} & \multicolumn{6}{|c|}{BLEU} & \multirow{2}{*}{BLEU} \\ \cline{2-13}	
-  mixture & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & average \\
+  \centering % \small
+  \begin{adjustbox}{width=1.0\textwidth,center}
+  \begin{tabular}{|p{2.0cm}|*{13}{r|}} \hline
+    \multirow{2}{*}{Name} & \multirow{2}{*}{$sim\_iter$} & \multirow{2}{*}{parameterization} & \multirow{2}{*}{entropy constraint} & \multicolumn{6}{|c|}{BLEU} & \multirow{2}{*}{average} \\ \cline{5-10}
+   & & & & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} &  \\
     \hline
-  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &37.45&51.26&45.47&31.62&37.55&65.22& \\
-  \system{FT-Res} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-  \system{Multi-DDS} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-  \system{FT-Full} & 0.0 &1.0 & 0.0 &0.0 &0.0 & 0.0 &35.92& 54.87& 52.55& 32.55& 44.98& 90.28& 51.86\\
-  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-  \system{Multi-DDS} & 0.0 &1.0 & 0.0 &0.0 &0.0 & 0.0 &35.92& 54.87& 52.55& 32.55& 44.98& 90.28& 51.86\\
-  \system{FT-Full} & 0.0 &0.0 & 1.0 &0.0 &0.0 & 0.0 &36.11& 55.37& 51.76& 33.52& 46.23& 89.99& 52.16\\
-  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-  \system{Multi-DDS} & 0.0 &0.0 & 1.0 &0.0 &0.0 & 0.0 &36.11& 55.37& 51.76& 33.52& 46.23& 89.99& 52.16\\
-  \system{FT-Full} & 0.0 &0.0 & 0.0 &1.0 &0.0 & 0.0 &36.53&	55.03& 51.15& 33.98& 44.28& 87.22& 51.365\\
-  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-  \system{Multi-DDS} & 0.0 &0.0 & 0.0 &1.0 &0.0 & 0.0 &36.53&	55.03& 51.15& 33.98& 44.28& 87.22& 51.365\\
-  \system{FT-Full} & 0.0 &0.0 & 0.0 &0.0 &1.0 & 0.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
-  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-  \system{Multi-DDS} & 0.0 &0.0 & 0.0 &0.0 &1.0 & 0.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
-  \system{FT-Full} & 0.0 &0.0 & 0.0 &0.0 &0.0 & 1.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
-  \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-  \system{Multi-DDS} & 0.0 &0.0 & 0.0 &0.0 &0.0 & 1.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+    \system{$config\_603$} & 1 & softmax & no & 37.26& 55.07& 50.96& 33.49& 43.41& 90.76& 51.83\\
+    \system{$config\_616$} & 1 & softmax & yes & 36.96& 55.55& 52.88& 33.06& 44.52& 91.25& 52.37 \\
+    \system{$config\_675$} & 10 & softmax & yes & 36.24& 55.67& 52.63& 32.74& 44.47& 90.45& 52.03\\
+    \system{$config\_673$} & 10 & linear & yes & 36.02& 56.18& 52.96& 32.06& 45.2 & 90.94 & 52.23 \\
   \hline
@@
   \end{adjustbox}
-  \label{tab:redomains}
+  \caption{Controlling / regularizing the training distribution}
+  \label{tab:performance}
 \end{table*}
+
+\begin{figure*}[htb]
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include first image
+  \includegraphics[width=.8\linewidth]{config603.png}  
+  \caption{config 603}
+  \label{fig:603}
+\end{subfigure}
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=.8\linewidth]{config616.png}  
+  \caption{Config 616}
+  \label{fig:616}
+\end{subfigure}
+\newline
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include third image
+  \includegraphics[width=.8\linewidth]{config675.png}  
+  \caption{Config 675}
+  \label{fig:675}
+\end{subfigure}
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include fourth image
+  \includegraphics[width=.8\linewidth]{config673.png}  
+  \caption{Config 673}
+  \label{fig:673}
+\end{subfigure}
+\caption{Sampling distribution}
+\label{fig:sampling}
+\end{figure*}
+
+\subsection{Stabilizing the sampler \label{ssec:sampler}}
+
+\cite{Wang20balancing} used a simple softmax function and the following parameterization of the sampler $\lambda^l(d;\psi) = \operatorname{softmax}(\psi)$ where $\psi \in \mathbb{R}^{n_d}$.
+This parameterization of the sampler will induce the following gradient:\fyTodo{Change name of obj.
+function}
+\begin{align*}
+\frac{\partial XXX}{\partial \psi_d} & = \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \frac{\partial \lambda^l(j; \psi)}{\partial \psi_d} \\
+%	& = R_d \lambda^l(d; \psi)(1- \lambda^l(d; \psi)) - \displaystyle{\mathop{\sum}_{j\neq d}^{n_d}} R_j \lambda^l(d; \psi) \lambda^l(j; \psi)\\
+	& = \lambda^l(d; \psi) (R_d- XXX)\\
+\end{align*} 
+The fact that the change of $\psi_d$ is proportional to $\lambda^l(d; \psi)$ can explain that the optimization often leads to a degenerate solution in which the sampler ignore almost domains except a few ones.
+In order to avoid the appearance of a ``rich-gets-richer'' effect and degenerate solutions such as Fig.~\ref{}, we consider two alternatives.
+
+The first one is to replace the softmax function by a simpler linear function, yielding the following updates:
+\begin{align*}
+\psi_d = \frac{\psi_d + \eta * R_d}{ 1 + \eta \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j},
+\end{align*}
+where $\eta$ controls the learning rate of $\psi$.
+
+The second one simply adds an entropy constraint which ensures that the entropy of the domain distribution is always greater than a predefined value $\alpha$.\fyTodo{Explain the math}
 

### 2021-03-10 2021/acl-ijcnlp2021/mdmt.tex
more comments on the plan

@@
 
-This setting is pretty general and can in principle accomodate the variety of situations mentioned  above, and many more: basic domain adaption, multi-domain adaptation with various target distribution, possibly including domains unseen in training.
+This setting is pretty general and can in principle accomodate the variety of situations mentioned  above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
 In our experiments we would like to better assess the actual potential of DDS is these settings, and seek to experimentally answer the following questions:
@@
 
-These questions are further discussed in Section~\ref{}. We now turn to our experimental conditions.
+These questions are further discussed in Section~\ref{sec:results}. We now turn to our experimental conditions.
+
 \section{Experimental settings} \label{sec:exp}
@@
 \end{table*}
+\fyTodo{Works best for small domains ?}
+\fyTodo{Curves for bank / medical ou law}
+
+\fyTodo{Fine tune pour 2 domains}
 

### 2021-03-15 2021/acl-ijcnlp2021/mdmt.tex
rewrite reinforce equation

@@
 \newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\revision}[1]{#1}
+\newcommand{\revision}[1]{\textcolor{red}{#1}}
 \newcommand{\revisiondel}[1]{}
@@
 Assume training data from $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as  $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
-We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t _d}$ to denote the empirical train and test distributions for domain $d$, and $\widehat{\mathcal{D}}(x;\lambda) = \sum_{d} \lambda(d) \widehat{\mathcal{D}^l_d}(x)$.
+We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t _d}$ to denote the empirical train and test distributions for domain $d$, and \revision{$\widehat{\mathcal{D}}(x;\lambda) = \sum_{d} \lambda(d) \widehat{\mathcal{D}^l_d}(x)$} \fyTodo{$\widehat{\mathcal{D}}(x;\lambda) = \sum_{d} \lambda(d) \widehat{\mathcal{D}_d}(x)$}.
 In our setting,  $\lambda^t$, and hence $\widehat{\mathcal{D}^t}(x;\lambda^t)$ are fixed and predefined, approximated with an equivalent number of development corpora.
@@
 where we use the following notations:% \fyTodo{Overload Q, from domain to instances to pairs}
+
 \begin{align*}
@@
 \theta_{t+1} &= \theta_t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta_t, x,y)}{\partial \theta} \\
-x,y &\sim \widehat{\mathcal{D}^l}(x;\lambda(\psi_t)) \\
+x,y &\sim \widehat{\mathcal{D}^l}(x;\lambda^l(\psi_t)) = \displaystyle{\mathop{\sum}_{d=1}^K} \lambda^l(d,\psi_t) * \widehat{D_d}(x) \\
 \end{align*}
+Given $\theta_t$, $\psi_t$ is updated via the Reinforce algorithm \cite{Williams92reinforce}:\fyTodo{Rewrite this correctly with domain average reward}
+%\begin{align*}
+%  \psi_{t+1} &= \psi_t + \mathbf{lr}_{data} * R(x,y) * \\
+%  & \quad\quad \frac{\partial log(\lambda^l(\psi_t)(x,y))}{\partial \psi} \\
+%x,y &\sim \lambda^l (\psi_t) \\
+%R(x,y) &= \langle \frac{\partial l(\theta',x,y)}{\partial \theta}, \frac{\partial J(\theta_{t+1}, \lambda^l)}{\partial \theta} \rangle \\
+%\theta' &= \theta_t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta_t, x,y)}{\partial \theta} \\
+%\end{align*}
 
-Given $\theta_t$, $\psi_t$ is updated via the Reinforce algorithm \cite{Williams92reinforce}:\fyTodo{Rewrite this correctly with domain average reward}
 \begin{align*}
-  \psi_{t+1} &= \psi_t + \mathbf{lr}_{data} * R(x,y) * \\
-  & \quad\quad \frac{\partial log(\lambda^l(\psi_t)(x,y))}{\partial \psi} \\
-x,y &\sim \lambda^l (\psi_t) \\
-R(x,y) &= \langle \frac{\partial l(\theta',x,y)}{\partial \theta}, \frac{\partial J(\theta_{t+1}, \lambda^l)}{\partial \theta} \rangle \\
+\psi_{t+1} &= \psi_t + \mathbf{lr}_{data} * \displaystyle{\mathop{\sum}_{d=1}^K} R(d) * \frac{\partial log(\lambda^l(\psi_t)(d))}{\partial \psi} \\
+  R(d) &= E_{x,y \sim \widehat{D_d}}[\langle \frac{\partial l(\theta_t,x,y)}{\partial \theta}, \frac{\partial J(\theta', \lambda^t)}{\partial \theta} \rangle] \\
   \theta' &= \theta_t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta_t, x,y)}{\partial \theta} \\
@@
 
+$\mathbf{lr}_{data}$ is the learning rate of the parameter of the sampler.
+$\mathbf{lr}_{nmt}$ is the learning rate of the parameter of the NMT system.
 \fyTodo{Explain lr, may be more}\fyTodo{More refs to the algorithm }

### 2021-04-05 2021/thesis/layout/firstpage.tex
first commits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Modèle pour la 1ère de couverture des thèses préparées à l'Université Paris-Saclay, basé sur le modèle produit par Guillaume BRIGOT / Template for back cover of thesis made at Université Paris-Saclay, based on the template made by Guillaume BRIGOT
%%% Mis à jour par Aurélien ARNOUX (École polytechnique)/ Updated by Aurélien ARNOUX (École polytechnique)
%%% Les instructions concernant chaque donnée à remplir sont données en bloc de commentaire / Rules to fill this file are given in comment blocks
%%% ATTENTION Ces informations doivent tenir sur une seule page une fois compilées / WARNING These informations must contain in no more than one page once compiled
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Version du 23 mai 2019 (Merci à Thibault CHEVALÉRIAS (CEA) pour ses suggestions et corrections)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\label{form_first}

%%% Formulaire / Form
%%% Remplacer les paramètres des \newcommand par les informations demandées / Replace \newcommand parameters by asked informations
%%%

\newcommand{\NNT}{20XXSACLXXXX} 															%% Numéro National de Thèse (donnée par la bibliothèque à la suite du 1er dépôt)/ National Thesis Number (given by the Library after the first deposit)

\newcommand{\ecodoctitle}{SCIENCES ET TECHNOLOGIES DE L'INFORMATION ET DE LA COMMUNICATION} 				%% Nom de l'ED. Voir site de l'Université Paris-Saclay / Full name of Doctoral School. See Université Paris-Saclay website
\newcommand{\ecodocacro}{STIC}																%% Sigle de l'ED. Voir site de l'Université Paris-Saclay / Acronym of the Doctoral School. See Université Paris-Saclay website
\newcommand{\ecodocnum}{580} 																%% Numéro de l'école doctorale / Doctoral School number
\newcommand{\PhDspeciality}{Big Data, Knowledge, Learning, Interactions} 									%% Spécialité de doctorat / Speciality 
\newcommand{\PhDworkingplace}{l'Université Paris-Sud} 										%% Établissement de préparation / PhD working place : l'Université Paris-Sud, l'Université de Versailles-Saint-Quentin-en-Yvelines, l'Université d'Evry-Val-d'Essonne, l'Institut des sciences et industries du vivant et de l'environnement (AgroParisTech), CentraleSupélec,l'Ecole normale supérieure de Cachan, l'Ecole Polytechnique, l'Ecole nationale supérieure de techniques avancées, l'Ecole nationale de la statistique et de l’administration économique, HEC Paris, l'Institut d'optique théorique et appliquée, Télécom ParisTech, Télécom SudParis   
\newcommand{\defenseplace}{Orsay} 											%% Ville de soutenance / Place of defense
\newcommand{\defensedate}{Date} 															%% Date de soutenance / Date of defense

%%% Établissement / Institution
%%% Si la thèse a été produite dans le cadre d'une co-tutelle, commenter la partie "Pas de co-tutelle" et décommenter la partie "Co-tutelle" / If the thesis has been prepared in guardianship, comment the part "Pas de co-tutelle" and uncomment the part "Co-tutelle"

	%%%%%%%%%%%%%%%%%%%%%%%%%
	%%% Pas de co-tutelle %%%
	%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\logoEtt}{blank}																%% NE PAS MODIFIER / DO NOT MODIFY
\newcommand{\vpostt}{0.1} 																	%% NE PAS MODIFIER / DO NOT MODIFY
\newcommand{\hpostt}{6}																		%% NE PAS MODIFIER / DO NOT MODIFY
\newcommand{\logoEt}{etab} 																	%% Logo de l'établissement de soutenance. Indiquer le sigle / Institution logo. Indicate the acronym : AGRO, CENTSUP, ENS, ENSAE, ENSTA, HEC, IOGS, TPT, TSP, UEVE, UPSUD, UVSQ, X 
\newcommand{\vpos}{0.1}																		%% À modifier au besoin pour aligner le logo verticalement / If needed, modify to align logo vertilcally
\newcommand{\hpos}{11}																		%% À modifier au besoin pour aligner le logo horizontalement / If needed, modify to align logo horizontaly

		%%%%%%%%%%%%%%%%%%
		%%% Co-tutelle %%%
		%%%%%%%%%%%%%%%%%%

%\newcommand{\logoEt}{etab} 																%% Logo de l'université partenaire. Placer le fichier .png dans le répertoire '/media/etab' et indiquer le nom du fichier sans l'extension / Logo of partner university. Place the .png file in the directory '/media/etab' and point the file name without the extension
%\newcommand{\vpos}{0.1}																	%% À modifier au besoin pour aligner les logos verticalement / If needed, modify to align logos vertilcally
%\newcommand{\hpos}{11}																		%% À modifier au besoin pour aligner les logos horizontalement / If needed, modify to align logos horizontaly
%\newcommand{\logoEtt}{etab2}  																%% Logo de l'établissement de soutenance. Le nom du fichier correspond au sigle de l'établissement /  Institution logo. Filename correspond to institution acronym : AGRO, CENTSUP, ENS, ENSAE, ENSTA, HEC, IOGS, TPT, TSP, UEVE, UPSUD, UVSQ, X 
%\newcommand{\vpostt}{0.1} 																	%% À modifier au besoin pour aligner les logos verticalement / If needed, modify to align logos vertilcally
%\newcommand{\hpostt}{6}																	%% À modifier au besoin pour aligner les logos horizontalement / If needed, modify to align logos horizontaly

%%% JURY

% Lors du premier dépôt de la thèse le nom du président nest pas connu, le choix du président se fait par les membres du Jury juste avant la soutenance. La précision est apportée sur la couverture lors du second dépôt / Choice of the jury's president is made during the defense. Thus, it must be specified only for the second file deposition in ADUM.
% Tous les membres du juty listés doivent avoir été présents lors de la soutenance / All the jury members listed here must have been present during the defense.

%%% Membre n°1 (Président) / Member n°1 (President)
\newcommand{\jurynameA}{Prénom Nom}
\newcommand{\juryadressA}{Statut, Établissement (Unité de recherche)}
\newcommand{\juryroleA}{Président}

%%% Membre n°2 (Rapporteur) / Member n°2 (Reviewer)
\newcommand{\jurynameB}{Prénom Nom}
\newcommand{\juryadressB}{Statut, Établissement (Unité de recherche)}
\newcommand{\juryroleB}{Rapporteur}

%%% Membre n°3 (Rapporteur) / Member n°3 (Reviewer)
\newcommand{\jurynameC}{Prénom Nom}
\newcommand{\juryadressC}{Statut, Établissement (Unité de recherche)}
\newcommand{\juryroleC}{Rapporteur}

%%% Membre n°4 (Examinateur) / Member n°4 (Examiner)
\newcommand{\jurynameD}{Prénom Nom}
\newcommand{\juryadressD}{Statut, Établissement (Unité de recherche)}
\newcommand{\juryroleD}{Examinateur}

%%% Membre n°5 (Directeur de thèse) / Member n°5 (Thesis supervisor)
\newcommand{\jurynameE}{Prénom Nom}
\newcommand{\juryadressE}{Statut, Établissement (Unité de recherche)}
\newcommand{\juryroleE}{Directeur de thèse}

%%% Membre n°6 (Co-directeur de thèse) / Member n°6 (Thesis co-supervisor)
\newcommand{\jurynameF}{Prénom Nom}
\newcommand{\juryadressF}{Statut, Établissement (Unité de recherche)}
\newcommand{\juryroleF}{Co-directeur de thèse}

%%% Membre n°7 (Invité) / Member n°7 (Guest)
\newcommand{\jurynameG}{Prénom Nom}
\newcommand{\juryadressG}{Statut, Établissement (Unité de recherche)}
\newcommand{\juryroleG}{Invité}

%%% Membre n°8 (Invité) / Member n°8 (Guest)
\newcommand{\jurynameH}{Prénom Nom}
\newcommand{\juryadressH}{Statut, Établissement (Unité de recherche)}
\newcommand{\juryroleH}{Invité}

%% Il est possible d'ajouter des membres supplémentaires selon le même modèle / More jury members can be added according to the same model

\label{layout_first}
%%% Mise en page / Page layout      
%%% NE RIEN MODIFIER EXCEPTÉ LA PARTIE CONCERNANT LE JURY (voir \label{jury}) SI BESOIN / DO NOT MODIFY EXCEPT SECTION CONCERNING JURY (see \label{jury}) IF NEEDED
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\thispagestyle{empty}
.
\begin{textblock}{5}(0,0)
	\textblockcolour{bordeau}
	\includegraphics [scale=1]{logos/bande.png}
	\vspace{300mm}
\end{textblock}

\begin{textblock}{1}(0.6,9.5)

	\Huge{\rotatebox{90}{\color{white}{\fontsize{38}{54}\selectfont Thèse de doctorat}}}
\end{textblock}

\begin{textblock}{1}(0.6,3)
	\Large{\rotatebox{90}{\color{white}{NNT : \NNT}}}
\end{textblock}

\begin{textblock}{1}(5.5,0)
	\textblockcolour{white}
	\includegraphics[scale=1]{logos/ed/STIC.jpeg}
\end{textblock}

\begin{textblock}{1}(5.5,1.5)
	\textblockcolour{white}
		\includegraphics[width=2.0\textwidth]{logos/etab/UPSUD.png}	
\end{textblock}

\begin{textblock}{1}(8.3,1.5)
	\textblockcolour{white}
		\includegraphics[width=4.5\textwidth]{logos/systran/systran-logo.png}	
\end{textblock}

\begin{textblock}{1}(13.4,1.5)
	\textblockcolour{white}
		\includegraphics[width=2.0\textwidth]{logos/limsi/cnrs-logo.png}	
\end{textblock}

%\begin{textblock}{1}(\pos,\vpos)
%	\textblockcolour{white}
%		\includegraphics[scale=1]{logos/etab/UPSUD.png}	
%\end{textblock}

%% Texte
\begin{singlespace}
\begin{textblock}{10}(5.7,3)
	\textblockcolour{white}

	\color{bordeau}
	\begin{flushright}

		\center \huge{\PhDTitle} \bigskip %% Titre de la thèse 
		\vfill
		\color{black} %% Couleur noire du reste du texte
		\normalsize {Thèse de doctorat de l'Université Paris-Saclay} \\
		préparée à \PhDworkingplace \\ \bigskip
		\vfill
		École doctorale n$^{\circ}$\ecodocnum ~\ecodoctitle ~(\ecodocacro)  \\

		\small{Spécialité de doctorat: \PhDspeciality} \bigskip %% Spécialité 
		\vfill  
		\footnotesize{Thèse présentée et soutenue à \defenseplace, le \defensedate, par} \bigskip
		\vfill
		\Large{\textbf{\textsc{\PhDname}}} %% Nom du docteur
		\vfill
	\end{flushright}

	\color{black}
	%% Jury
	\begin{flushleft}

		\small Composition du Jury :
	\end{flushleft}
	%% Members of the jury

	\small
	%\begin{center}
	\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
	\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}lm{#1}}

	\label{jury} 																				%% Mettre à jour si des membres ont été ajoutés ou retirés / Update if members have been added or removed
	\begin{flushleft}
	\begin{tabular}{@{} L{9.5cm} R{4.5cm}}
		\jurynameA  \\ \juryadressA & \juryroleA \\[5pt]
		\jurynameB  \\ \juryadressB & \juryroleB \\[5pt]
		\jurynameC  \\ \juryadressC & \juryroleC \\[5pt]
		\jurynameD  \\ \juryadressD & \juryroleD \\[5pt]
		\jurynameE  \\ \juryadressE & \juryroleE \\[5pt]
		\jurynameF  \\ \juryadressF & \juryroleF \\[5pt]
		\jurynameG  \\ \juryadressG & \juryroleG \\[5pt]
		\jurynameH  \\ \juryadressH & \juryroleH \\[5pt]
	\end{tabular} 
	\end{flushleft}   
	%\end{center}
\end{textblock}
\end{singlespace}
\afterpage{\blankpage}

### 2021-04-05 2021/thesis/parts/introduction.tex
first commits

\chapter{Introduction}

For a long time, Machine Translation(MT\nomenclature[mt]{MT}{Machine Translation}) has evolved around data-driven methods including Statistical Machine Translation (SMT\nomenclature[smt]{SMT}{Statistical Machine Translation}) and Neural Machine Translation (NMT\nomenclature[nmt]{NMT}{Neural Machine Translation}). Learning from data is the cheapest and the most computationally efficient road to building such intelligent systems for the Automatic Translation task. For nearly one decade, NMT models have surpassed SMT models in the high-resource language pairs such as French(Fr\nomenclature[fr]{Fr}{French}) - English(En\nomenclature[en]{En}{English}), German(De\nomenclature[de]{De}{German}) - English thanks to their extremely complex architecture. Since our first encounter with NMT models, it becomes a unique paradigm both in the research and in the industry that we feed all the available data to the NMT model with the belief that the more training data is fed, the better the model performs. However, recent works in Domain Adaptation (DA\nomenclature[da]{DA}{Domain Adaptation}), e.g \cite{rico13domain}, showed that the belief was no longer true when the NMT model was tested with text that came from a different context than training data. NMT model's quality, as SMT model's, is guaranteed by the relevance between training data and testing data. The goal of this thesis is to point out exactly what can happen if training distribution and testing distribution are mismatched, and to propose new methods to improve the NMT model's performance under such a situation.

\section{What is a domain?}
The notion of a domain is vaguely defined despite the fact that it is the subject of a large group of research works including DA and MDL. 
The richness of data is not only demonstrated by the number of parallel sentences but is also expressed by the high variety of genres, topics, and style. Parallel text is usually grouped into parallel corpora such as the European Central Bank corpus \citep{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus\citep{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \citep{Tiedemann09news}and Ted Talks \citep{Cettolo12wit}. Each corpus is collected from activities of an organization or

\section{Multi-domain problem}

\section{Thesis contribution}

\section{Outline}

### 2021-04-05 2021/thesis/parts/state_of_the_art.tex
first commits

\part{State of the art}
\chapter{Neural Machine Translation}
\section{Recurrent Neural Network models}
\subsection{GRU, LSTM}
\subsection{Attentional mechanism}
\section{Transformer models}
\subsection{self-attention, cross-attention layers}
\subsection{Variants}
\chapter{Data centric domain adaptation}
\chapter{Feature centric domain adaptation}
\chapter{Model centric domain adaptation}

### 2021-04-08 2021/thesis/parts/state_of_the_art.tex
chapter 2

@@
-\part{State of the art}
-\chapter{Neural Machine Translation}
+\chapter{Neural Machine Translation's review \label{chap:2}} 
+This chapter of the thesis is dedicated to reviewing the recent developments in NMT research.
+We will focus particularly on an important part of the field which provides the foundation for the experiments of the thesis.
+Neural Machine Translation was first introduced in 2014 via the work of \cite{bahdanau14neural,cho14properties}.
+Since then, NMT has been largely developed and outperformed old approaches including Rule-based MT, SMT, and Hybrid MT in high-resource languages such as English-French, English-German.
+In general, NMT approaches use a parallel corpus to learn their parameters.
+Each sentence is processed into a sequence of tokens, which can be words, sub-words, or characters, then is indexed by a predetermined vocabulary.
+Given a choice of neural architecture, the parameters of the NMT model are optimized according to a training objective and training examples.
+After training, given any input, the NMT model predicts an output, which approximates the translation of that input, via a decoding algorithm such as beam search \citep{koehn04pharaoh}.
+\section{Text preprocessing for NMT \label{sec:2.1}}
+Typically a sentence can be split into a sequence of tokens and we call this process tokenization.
+There are 3 common types of a token including word, sub-word, and character.
+These tokens will be indexed by a predetermined corresponding vocabulary in the source or target language.
+The sequence of tokens is translated into a sequence of integers $IDs \in V$ where V is the set of the index of the corresponding vocabulary.
+The vocabulary of the NMT model is limited and fixed before and after the training.
+Any out of vocabulary (OOV\nomenclature[oov]{OOV}{Out of vocabulary}) words are mapped to the UNK token.
+The size of the vocabulary is chosen to balance the coverage with the practical constraint on the size of the model.
+An NMT model vocabulary is usually limited to 30 thousand tokens
+\subsection{Word tokenization}
+The first type of token is the most natural because it does not need extra effort to separate the words of a sentence.
+However, this approach has several disadvantages as it treats words as isolated units and therefore can not handle the large vocabulary of the corresponding language and the growing number of unseen words.
+\subsection{Subword tokenization}
+However, the morphological property of languages allows us to go beyond the surface level of words and present a word as a compound of morphemes.
+This presentation increases largely the coverage of the vocabulary, therefore handles beautifully unseen words.
+These tokens are referred to as subwords.
+The vocabulary can be built by applying the morphological rules of the language or can be learned by heuristic algorithms such as Byte pair encoding (BPE\nomenclature[bpe]{BPE}{Byte pair encoding})\citep{sennrich16neural,gage94anew}.
+
+Because BPE tokenization is the most popular tokenization in NMT research, we will review briefly this technique.
+In general, BPE vocabulary is first initialized with a set of characters; words in the collected corpora are represented as sequences of characters.
+BPE algorithm then counts the occurrences of each pair of BPE tokens, and iteratively merges the most frequent pair into a new token, and add it to BPE vocabulary until the vocabulary reaches a predetermined size.
+In the end, most frequent words become tokens in the vocabulary, less frequent words remain sequences of BPE tokens, which can be sequences of characters or characters in the worst case.
+BPE vocabulary can be learned separately from one language either source or target, or jointly from both source and target, or from multiple languages as in multi-lingual NMT.
+
+Previous BPE tokenization is frequency-based; there are alternative paradigms for subword tokenization such as syllabification \cite{assylbekov17syllable}, subword-regularization\cite{taku18subword}, linguistically informed tokenization \cite{ataman17linguistically,huck17target,machcek18morphological}.
+BPE tokenization is not only applied to text representation but also to byte-level representation as \cite{wang19neural} showed comparable performance of byte-level BPE-based NMT compared to BPE-based NMT.
+Furthermore, byte-level BPE is able to represent any Unicode character, there to make an open vocabulary MT possible.
+\subsection{Character tokenization}
+The problem of finding optimal subword segmentation becomes more complex
+when an NMT system must handle multiple source/target languages, as in multilingual translation or zero-shot approaches.
+Translating characters instead of subword tokens avoids these problems and give the MT system
+access to all available information about source and target sentences regardless of their languages.
+Furthermore, character tokenization reduces the size of the vocabulary to a small number of written characters.
+However, because of extremely splitting words into character units, the length of the resulting sequence increases largely.
+Increasing sequence length increases both the computational requirements during training and the decoding time during the prediction.
+
+First attempts of character based NMT including \cite{wang15character,luong16achieving} focused on solving the out-of-vocabulary and softmax bottleneck problems associated with word level models.
+\cite{costa16character,lee17fully,chung16character,costa17byte} proposed different fully character based NMT.
 \section{Recurrent Neural Network models}
+This section reviews the very first NMT architecture which is Recurrent Neural Network.
+In principle, the NMT model consists of 2 parts: encoder and decoder.
+The encoder maps (or encodes) the source sentence to either a sequence of real value vectors or a unique real value vector in a high dimension space called Latent space.
+The decoder iteratively encodes the target sequence from left to right, token by token, to a real value vector at each step and combine it with the output of the encoder into one real vector which then is mapped to a real vector of dimension equal to the size of target vocabulary, then passed to softmax activation function to produce the probability of words in the target vocabulary to appear at that step.
+The variants in the RNN category are different only in functions used to map sentences to Latent spaces.
+\subsection{Word embeddings}
+Each token in either source or target sequences is mapped to a real vector called word embedding.
+Word embeddings corresponding to words in the source$/$ target vocabulary are stored in a lookup table and will be extracted using the index of the corresponding token.
+The lookup table has the size of $V \times d$ where $V$ is the size of the corresponding vocabulary, $d$ is the dimension of word embedding space.
+Typically, d is 512 or 1024.
+
+Word embedding is not only used in the NMT model but also in the Neural language model\citep{bengio03aneural}(NLM\nomenclature{nlm}{NLM}{Neural language model}).
+\cite{le12continuous,schwenk12continuous} used NLM models for phrase-based statistical machine translation.
+
+Moreover, word embedding could be trained alone using Skip-gram model\citep{mikolov13distributed} or Continuous Bag of Word model\citep{mikolov13efficient}.
+After training such models, the resulting word embeddings possess semantic properties so that words having similar meanings or close meanings are mapped to similar vectors in terms of cosine similarity\citep{collobert11natural,mikolov13distributed}.
+This property is one of the ingredients in the success of neural model in text classification, text retrieval, language modeling and machine translation.
+
 \subsection{GRU, LSTM}
+
 \subsection{Attentional mechanism}
+\section{Convolution }
 \section{Transformer models}
-\subsection{self-attention, cross-attention layers}
+\subsection{Self-attention, cross-attention layers}
 \subsection{Variants}
-\chapter{Data centric domain adaptation}
-\chapter{Feature centric domain adaptation}
-\chapter{Model centric domain adaptation}
+\section{NMT training}
+\section{Decoding algorithms}
+\chapter{Data-centric domain adaptation}
+\chapter{Feature-centric domain adaptation}
+\chapter{Model-centric domain adaptation}

### 2021-04-08 2021/thesis/compile.sh

@@ -1,0 +1,6 @@
+pdflatex thesis.tex
+bibtex thesis.aux
+pdflatex thesis.tex
+pdflatex thesis.tex
+makeindex -s nomencl.ist -o thesis.nls thesis.nlo
+pdflatex thesis.tex

### 2021-04-12 2021/emnlp/mdmt.tex
start reshaping text

--- 2021-03-15 2021/acl-ijcnlp2021/mdmt.tex
@@
 \newcommand{\system}[1]{\texttt{{#1}}}
+
 \newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
+\newcommand{\vtheta}{\ensuremath{\boldsymbol\theta}\xspace} % parameters vector for a distribution
+\newcommand{\vpsi}{\ensuremath{\boldsymbol\psi}\xspace} % parameters vector for a distribution
 \newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}
@@
   When building machine translation systems, one often needs to make the best out of heterogeneous sets of parallel data in training to achieve the best test performance in one or several domain(s) of interest.
-This multi-source (multi-domain) adaptation problem is often approached with instance selection or reweighting strategies, most of which pre-suppose an ad-hoc assessment of the relevance of training instances.
+This multi-source / multi-domain adaptation problem is often approached with instance selection or reweighting strategies, most of which pre-suppose an ex-ante assessment of the relevance of training instances.
 In this paper, we study the recently proposed Differential Data Selection (DDS) model and explore its ability to serve as a generic framework for these various situations.
 Our experiments study both domain adaptation and multi-domain learning and show that DDS often enables to outperform more heuristic adaptation strategies.
-We also introduce a variant that boosts DDS performance of adapter-based multi-domain systems.\fyTodo{Keep this part ?}
+We also introduce variants that boosts DDS performance of adapter-based multi-domain systems.\fyTodo{Keep this part ?}
 %   This is a well-known issue, which has given raise to a rich litterature in under the umbrellas of domain adaptation or multi-domain
@@
 Single-source and multi-source domain adaptation (DA) is a well-studied instance of this setting (see \cite{Chu2017comparison} for a review) , and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
-A related situation is multi-lingual machine translation \cite{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
+A related situation is multilingual machine translation \cite{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
 
@@
 
-It is usual in MT to opportunistically collect training samples from several domains, which means that training instances are distributed according to the mixture $\mathcal{D}^s$ such that $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}(d) \mathcal{D}_d(x)$, where $\{\lambda^{s}(d), d=1 \dots n_d\}$ are the corresponding mixture weights such that $\sum_d \lambda^{s}(d)=1$).
+It is usual in MT to opportunistically collect training samples from several domains, which means that training instances are distributed according to the mixture $\mathcal{D}^s$ such that $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}(d) \mathcal{D}_d(x)$, where $\{\lambda^{s}(d), d=1 \dots n_d\}$ are the corresponding mixture weights satisfying $\sum_d \lambda^{s}(d)=1$.
 
@@
 These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unsupervised domain adaptation); in multi-domain adaptation finally the test distribution is itself a mixture of domains, some of which may also be observed in training.
-Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  $\lambda^{t}(d)$ in the case of domain adaptation.
+Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of fine-tuning.
 These situations are illustrated in Figure~\ref{fig:mdmt-lambdas}.
@@
   \caption{Training and testing with distribution mismatch.
-We consider just three domains, and represent vectors of mixture weights $\lambda^{s}$ and $\lambda^{t}$ in the 3-dimensional simplex.
-Training with weights in (a) and testing with weights in (c) is standard multi-source domain adaptation to $d_2$, while (b)-(c) is unsupervised adaptation; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with configurations (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
+We consider just three domains, and represent vectors of mixture weights $\vlambda^{s}$ and $\vlambda^{t}$ in the 3-dimensional simplex.
+Training with weights in (a) and testing with weights in (c) is supervised multi-source domain adaptation to domain~2 ($d_2$), while (b)-(c) is the unsupervised version, with no training data from $d_2$; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with configurations (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
 \end{figure}
@@
 
-A more practical standard approach to supervised DA is \emph{fine-tuning} \cite{Luong15stanford,Freitag16fast}, where $\lambda^{l}$ is allowed to vary during the course of learning.
+A more practical standard approach to supervised DA is \emph{fine-tuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during the course of learning.
 With our notations, this approach amounts to first learning an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$) with $\indic{A}$ the indicator function for predicate $A$.
 Note that this strategy is potentially suboptimal, as some out-of-domain samples may contribute to the final performance due to eg.\ domain overlap.
-Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains are easier to learn than others.\fyTodo{How to measure this?}
+Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains might be easier to learn than others.\fyTodo{How to measure this?}
 
-Differentiable data selection (DDS), proposed by \citet{Wang20optimizing} and later applied to multilingual machine translation by \citet{Wang20balancing}, introduces an additional degree of freedom and allows $\lambda^{l}(d)$ to change multiple times in the course of training, as we now explain.
+Differentiable data selection (DDS), proposed by \citet{Wang20optimizing} and later applied to multilingual machine translation by \citet{Wang20balancing}, introduces an additional degree of freedom and allows to iteratively learn an optimal value for $\lambda^{l}$, as we now explain.
 
@@
 \subsection{Basic principles}
-Assume training data from $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as  $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
-We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t _d}$ to denote the empirical train and test distributions for domain $d$, and \revision{$\widehat{\mathcal{D}}(x;\lambda) = \sum_{d} \lambda(d) \widehat{\mathcal{D}^l_d}(x)$} \fyTodo{$\widehat{\mathcal{D}}(x;\lambda) = \sum_{d} \lambda(d) \widehat{\mathcal{D}_d}(x)$}.
-In our setting,  $\lambda^t$, and hence $\widehat{\mathcal{D}^t}(x;\lambda^t)$ are fixed and predefined, approximated with an equivalent number of development corpora.
-Training optimizes the parameters $\theta$ of a neural translation model.
+Assuming training data from $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as  $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
+We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t _d}$ to denote the empirical train and test distributions for domain $d$, and \revision{$\widehat{\mathcal{D}^{u}}(x;\lambda^{u}) = \sum_{d} \lambda^{u}(d) \widehat{\mathcal{D}^{u}_d}(x)$ for $u\in\{l,t\}$} \fyTodo{$\widehat{\mathcal{D}}(x;\lambda) = \sum_{d} \lambda(d) \widehat{\mathcal{D}_d}(x)$}.
+In our setting,  $\vlambda^t$, and hence $\widehat{\mathcal{D}^t}(x;\vlambda^t)$ are fixed and predefined, approximated with an equivalent number of development corpora.
+Training optimizes the parameters $\vtheta$ of a neural translation model.
 
-Sampling training domains using a distribution $\lambda^{l}(d)$ results in a parameter estimate $\theta(\lambda^{l})$, delivering performance that can be estimated on the development set using $\lambda^{t}$.
-\cite{Wang20optimizing,Wang20balancing} then formulate the problem of finding an optimal sampling strategy $\lambda^{l}_{*}$ maximizing performance on the development data as the following bi-level optimization problem:
+Sampling training domains using a distribution $\lambda^{l}$ results in a parameter estimate $\theta(\lambda^{l})$, delivering performance that can be estimated on the development set using $\lambda^{t}$.
+\citet{Wang20optimizing,Wang20balancing} then formulate the problem of finding an optimal sampling strategy $\lambda^{l}_{*}$ maximizing performance on the development data as the following bi-level optimization problem:
 \begin{equation} \label{eq:bilevel}
 \begin{split}
-\lambda^{l}_* &= \displaystyle{\mathop{\argmin}_{\lambda^{l}} J(\theta^*(\lambda^{l}), \lambda^{t})} \\
-					& \theta^*(\lambda^{l}) = \displaystyle{\mathop{\argmin_{\theta}} J(\theta, \lambda^{l})}, 
+\vlambda^{l}_* &= \displaystyle{\mathop{\argmin}_{\lambda^{l}} J(\theta^*(\vlambda^{l}), \vlambda^{t})} \\
+					& \theta^*(\vlambda^{l}) = \displaystyle{\mathop{\argmin_{\vtheta}} J(\vtheta, \vlambda^{l})}, 
 \end{split}
@@
 \begin{align*}
-J(\theta, \lambda^{t} ) &= \displaystyle{\mathop{\sum}_{(x,y) \sim \widehat{\mathcal{D}^{t}(x)}} \ell(\theta,x,y)} \\
-J(\theta, \lambda^{l}) &= \displaystyle{\mathop{E_{(x,y) \sim \widehat{\mathcal{D}^{l}(x)}}[\ell(\theta,x,y)]}} \\
-l(\theta,x,y) &= - log P(\theta,x,y)
+J(\vtheta, \vlambda^{t} ) &= \displaystyle{\mathop{\sum}_{(x,y) \sim \widehat{\mathcal{D}^{t}(x)}} \ell(\vtheta,x,y)} \\
+J(\vtheta, \vlambda^{l}) &= \displaystyle{\mathop{E_{(x,y) \sim \widehat{\mathcal{D}^{l}(x)}}[\ell(\vtheta,x,y)]}} \\
+l(\vtheta,x,y) &= - log P(x,y;\vtheta)
 \end{align*}
-Problem~\eqref{eq:bilevel} can be approached by coordinate gradient descent alternating between $\lambda^{l}$ and $\theta$.
-\citet{Wang20balancing} limit the search space of $\lambda^{l}$ to a parameterized differentiable function $\lambda^{l}(d; \psi)$.
-Given $\psi_t$ at time step $t$, $\theta_t$ is updated via usual gradient descent as follows:
+Problem~\eqref{eq:bilevel} can be approached by coordinate gradient descent alternating between $\lambda^{l}$ and $\vtheta$.
+\citet{Wang20balancing} limit the search space of $\vlambda^{l}$ to a parameterized differentiable function $\vlambda^{l}(\vpsi)$.
+Given $\vpsi_t$ at time step $t$, $\vtheta_t$ is updated via usual gradient descent as follows:
 \begin{align*}
-\theta_{t+1} &= \theta_t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta_t, x,y)}{\partial \theta} \\
-x,y &\sim \widehat{\mathcal{D}^l}(x;\lambda^l(\psi_t)) = \displaystyle{\mathop{\sum}_{d=1}^K} \lambda^l(d,\psi_t) * \widehat{D_d}(x) \\
+\vtheta_{t+1} &= \vtheta_t - \mathbf{lr}_{nmt} * \frac{\partial l(\vtheta_t, x,y)}{\partial \vtheta} \\
+x,y &\sim \widehat{\mathcal{D}^l}(x;\vlambda^l(\psi_t)) = \displaystyle{\mathop{\sum}_{d=1}^K} \lambda^l(d;\psi_t) * \widehat{D_d}(x) \\
 \end{align*}
-Given $\theta_t$, $\psi_t$ is updated via the Reinforce algorithm \cite{Williams92reinforce}:\fyTodo{Rewrite this correctly with domain average reward}
+Given $\vtheta_t$, $\psi_t$ is updated via the Reinforce algorithm \cite{Williams92reinforce}:\fyTodo{Rewrite this correctly with domain average reward}
 %\begin{align*}
@@
 This setting is pretty general and can in principle accomodate the variety of situations mentioned  above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
-In our experiments we would like to better assess the actual potential of DDS is these settings, and seek to experimentally answer the following questions:
+In our experiments we would like to better assess the actual potential of DDS is these settings and seek to experimentally answer the following questions:
 \begin{itemize}
-\item is DDS a viable alternative to conventional fine-tuning ? In particular does it enable to better take advantage of relevant data from other domains ?
-\item is DDS a viable alternative in multidomain scenarios ?
-\item \fyTodo{TBContinued}
+\item is DDS a viable alternative to conventional fine-tuning? In particular does it enable to better take advantage of relevant data from other domains?
+\item is DDS also a viable option in multidomain adaptation scenarios?
+\item does DDS also enable to perform unsupervised (multi-) domain adaptation?\fyTodo{TBContinued}
 \end{itemize}
@@
 \subsection{Data and metrics \label{ssec:corpora}}
-We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.
-\revision{We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
+We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.\revision{We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}}; the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \cite{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
-Most corpora are available from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
+Most corpora are available from the Opus website.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
 To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
@@
 \end{align} 
-with $\alpha \in [0,0.25,0.5,0.75,1.0]$. These systems are denoted \system{Mixed-$\alpha$} below.
+with $\alpha \in [0,0.25,0.5,0.75,1.0]$.
+These systems are denoted \system{Mixed-$\alpha$} below.
+\system{Mixed-$0$} uses a uniform domain distribution, \system{Mixed-$1.0$} simple uses the observed domain distribution.
 %on a concatenation of all corpora (\texttt{Mixed}).
@@
 
-All models use embeddings and the hidden layers sizes of dimension~512.
-Transformer models contain with 8 attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells.
+All models use embeddings and hidden layers sizes of dimension~512.
+Transformer models contain 8~attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells.
 The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024.
@@
 
-\subsection{Fine tuning experiments}
+\subsubsection{Fine-tuning experiments}
 
 \fyTodo{Fix the table}
-In Table~\ref{tab:finetuning}, we compare the results of Fine tuning a system initially trained with a uniform mixture of domains with the result of learning from scratch, also with a initial uniform distribution ($\phi_0(d) = 0 \forall d$).
+In Table~\ref{tab:finetuning}, we compare the results of fine-tuning a system initially trained with a uniform mixture\fyTodo{uniform or natural?} of domains with the result of learning from scratch, also with a initial uniform distribution ($\phi_0(d) = 0 \forall d$).
 \fyTodo{What do we see ?
@@
 - show a graph}.
-Note that the DDS systems are somewhat quicker to build, as we only train for 100K iterations (instead of 100K + 50K for the fine tuning part).
+Note that the DDS systems are somewhat quicker to build, as we only train for 100K iterations (instead of 100K + 50K for the fine-tuning part).
 
@@
   \begin{tabular}{|p{3.0cm}|*{13}{r|}} \hline
-    \multirow{2}{*}{Name} & \multirow{2}{*}{$sim\_iter$} & \multirow{2}{*}{parameterization} & \multirow{2}{*}{entropy constraint} & \multicolumn{6}{|c|}{BLEU} & \multirow{2}{*}{BLEU average} \\ \cline{5-10}
-   & & & & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{it}} & \multicolumn{1}{c|}{\domain{ rel }} & \multicolumn{1}{c|}{\domain{ talk}} &  \\
+    \multirow{2}{*}{Name} & \multirow{2}{*}{$sim\_iter$} & \multirow{2}{*}{parameterization} & \multirow{2}{*}{entropy constraint} & \multicolumn{7}{|c|}{BLEU} \\ \cline{5-11}	
+   & & & & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{it}} & \multicolumn{1}{c|}{\domain{ rel }} & \multicolumn{1}{c|}{\domain{ talk}} & \multicolumn{1}{c|}{mean} \\
     \hline
@@
   \system{$DDS\_FT\_talk$} & 1 & softmax & no & 35.53&50.55&52.64&44.86&85.8&33.47&50.48\\
-  \system{$DDS\_FT\_full$} & NA & NA & NA & 37.74&59.21	&54.49&46.81&90.77&33.98&53.83\\
+  \system{$FT\_full$} & NA & NA & NA & 37.74&59.21	&54.49&46.81&90.77&33.98&53.83\\
     \hline
@@
 \end{table*}
+
+\fyTodo{This table does not work- One line would be enough, where we compare autoFT with true FT}
+
 \fyTodo{Works best for small domains ?}
@@
 
-\subsection{Multi-domain experiments}
-Results for our multidomain experiments are in Table~\ref{tab:multidomains}.
+\subsubsection{Multi-domain experiments}
+Results for our multidomain experiments are in Table~\ref{tab:multidomains}.\fyTodo{Same table, line autoMD.
+We should be clear about the dev / test distribution - uniform ?
+so I would compare MIX with DDS(UNIF, UNIF) - DDS(NAT,UNIF) and not report / domain scores}
 
@@
 
-Looking again at the evolution of $\lambda^l$ during training we often observe a rich get richer effect, where a small number of domains manage to attract most of the probability mass, while other are completely ignore after a given number of iterations.
+In this scenario, DDS is not as good as using a fixed distribution, with a well-chosen $\alpha$.
 
-\fyTodo{Explain issues}
+Looking again at the evolution of $\vlambda^l$ during training we often observe a rich-get-richer effect, where a small number of domains manage to attract most of the samples, while others are almost completely ignored after a given number of iterations.
+This is illustrated on Figure~\ref{}, where we see..., and also when we look at the entropy of $\vlambda^l$ at the end of training, which also measures the KL divergence between $\vlambda^l$ and a uniform $\vlambda^t$.\fyTodo{Give numbers}
+
+\subsection{Analysis}
+\subsubsection{Winner-takes-all}
+\cite{Wang20balancing} used a simple softmax function and the following parameterization of the sampler $\lambda^l(d;\psi) = \operatorname{softmax}(\psi)$ where $\psi \in \mathbb{R}^{n_d}$.
+This parameterization of the sampler will induce the following gradient:\fyTodo{Change name of obj.
+function}
+\begin{align*}
+\frac{\partial XXX}{\partial \psi_d} & = \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \frac{\partial \lambda^l(j; \psi)}{\partial \psi_d} \\
+%	& = R_d \lambda^l(d; \psi)(1- \lambda^l(d; \psi)) - \displaystyle{\mathop{\sum}_{j\neq d}^{n_d}} R_j \lambda^l(d; \psi) \lambda^l(j; \psi)\\
+	& = \lambda^l(d; \psi) (R_d- XXX)\\
+\end{align*} 
+The fact that the change of $\psi_d$ is proportional to $\lambda^l(d; \psi)$ can explain that the optimization often leads to a degenerate solution in which the sampler ignore almost domains except a few ones.
+In order to avoid the appearance of a ``rich-gets-richer'' effect and degenerate solutions such as Fig.~\ref{},
 
@@
 
-\cite{Wang20balancing} used a simple softmax function and the following parameterization of the sampler $\lambda^l(d;\psi) = \operatorname{softmax}(\psi)$ where $\psi \in \mathbb{R}^{n_d}$.
-This parameterization of the sampler will induce the following gradient:\fyTodo{Change name of obj.
-function}
-\begin{align*}
-\frac{\partial XXX}{\partial \psi_d} & = \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \frac{\partial \lambda^l(j; \psi)}{\partial \psi_d} \\
-%	& = R_d \lambda^l(d; \psi)(1- \lambda^l(d; \psi)) - \displaystyle{\mathop{\sum}_{j\neq d}^{n_d}} R_j \lambda^l(d; \psi) \lambda^l(j; \psi)\\
-	& = \lambda^l(d; \psi) (R_d- XXX)\\
-\end{align*} 
-The fact that the change of $\psi_d$ is proportional to $\lambda^l(d; \psi)$ can explain that the optimization often leads to a degenerate solution in which the sampler ignore almost domains except a few ones.
-In order to avoid the appearance of a ``rich-gets-richer'' effect and degenerate solutions such as Fig.~\ref{}, we consider two alternatives.
-
-The first one is to replace the softmax function by a simpler linear function, yielding the following updates:
+To stabilize the sampler, we consider two alternatives. The first one is to replace the softmax function, yielding the following updates:\fyTodo{List options}
 \begin{align*}

### 2021-04-13 2021/emnlp/mdmt.tex
new organization of experiments

@@
 \begin{align} \label{mixture:trn}
-\lambda^{l}(d;\alpha) = \frac{q_d^{\alpha}}{\displaystyle{\mathop{\sum}_{d=1}^{n_d}q_d^{\alpha}}} &&
+\lambda_{\alpha}(d) = \frac{q_d^{\alpha}}{\displaystyle{\mathop{\sum}_{d=1}^{n_d}q_d^{\alpha}}} &&
 q_d = \frac{\mid N^{s}_d \mid}{\displaystyle{N^{s}}} % \mathop{\sum}_{i=1}^K\mid D_i \mid}}
@@
 The latter uses the same architecture, but a different training procedure and learns all parameters jointly from scratch with the prefixed mixtures of training data \ref{mixture:trn}.\fyTodo{Keep this ?}
+
+\item Our comparison of multi-domain systems includes baselines with fixed data mixtures corresponding to $\vlambda_0$ and $\vlambda_1$, and our own reimplementations of recent proposals from the literature:\footnote{\revision{Further implementation details are in supplementary material.}}\fyTodo{Choose a reasonnable MD baseline }
+
 \end{itemize}
@@
 
-\subsection{Multi-domain systems} \label{ssec:mdsys}
+\subsection{DDS-based systems } \label{ssec:dds-sys}
 
-Our comparison of multi-domain systems includes our own reimplementations of recent proposals from the literature:\footnote{\revision{Further implementation details are in supplementary material.}}
-\begin{itemize}
-\itemsep0em 
-\item NMT models trained with Multi-DDS mixture given several objective testing distributions: uniform, uni-domain and natural.
-\item \system{MDL-Res} trained from scratch with Multi-DDS mixture.
-\end{itemize}
+The architecture of DDS based-systems is identical to that of our baselines, and only differ in the training regime.
+Their behavior only depends on (a) the initial domain distribution at the start of training $\vlambda^{l}_{t=0}$, and (b) the targeted (dev/test) distribution $\vlambda^{t}$.
+We will thus report these systems as \system{DDS($\vlambda^{l}_{t=0}$, $\vlambda^{t}$)}.
+The domain distribution attained at the end of training will be denoted  $\vlambda^{l}_{*}$.
 
 \section{Results and discussion \label{sec:results}}
+
+\subsection{DDS for single-domain and multi-domain adapation}
 
@@
 \fyTodo{Fix the table}
-In Table~\ref{tab:finetuning}, we compare the results of fine-tuning a system initially trained with a uniform mixture\fyTodo{uniform or natural?} of domains with the result of learning from scratch, also with a initial uniform distribution ($\phi_0(d) = 0 \forall d$).
+In Table~\ref{tab:finetuning}, we compare the results of fine-tuning a system initially trained with a uniform mixture\fyTodo{uniform or natural?} of domains with the result of learning from scratch, also with a initial uniform distribution.
 \fyTodo{What do we see ?
@@
 Note that the DDS systems are somewhat quicker to build, as we only train for 100K iterations (instead of 100K + 50K for the fine-tuning part).
+
+\begin{table*}
+  \centering \small
+  \begin{tabular}{|l|*8{r|}} \hline
+    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline
+    Full-FT($d$) & \\
+    \system{DDS($\vlambda_0, \indic{d}$)} & \\
+    $- \log \lambda_*^{l}(d)$ ou \% & \\
+    $ \mathcal{H}(\vlambda_*^{l})$ &  \\ \hline\hline
+    \system{Mixed-0} & \\
+    \system{Mixed-1} & \\
+    \system{DDS($\vlambda_1, \vlambda_0$)} & \\
+    $\operatorname{KL}(\vlambda_*^{l}, \vlambda_0$) & \\
+     $ \mathcal{H}(\vlambda_*^{l})$ &  \\
+    \system{DDS($\vlambda_0, \vlambda_0$)} & \\ \hline
+  \end{tabular}
+  \caption{Baseline results for DDS.
+In the top part, for each domain, we report the scores of DDS targeting solely this domain and compare with full-fine-tuning.
+We also compute the percentage of in-domain data in the resulting mixture and its entropy.}
+  \label{tab:finetuning}
+\end{table*}
+
+A first observation is DDS only better than FT for \domain{rel}, and worse for \domain{med} and \domain{law}.
+There seem to be a connection with the domain size, as \domain{rel} is small and benefits from data from other domains, while the opposite effect is observed for the larger \domain{med} and \domain{law}.
+For the three other domains, we do not see any marked difference, which we view as a positive result.
+
+If we now look at the multi-domain results (bottom part of Table~\ref{}), we see that DDS actually does better than the two baseline strategies, but is not as good as using a fixed distribution with a well-chosen $\alpha$....\fyTodo{To be continued}
 
@@
   \end{adjustbox}
-  \caption{DDS as a fine-tuning receipe.
+  \caption{DDS as a fine-tuning recipe.
 We report scores with DDS and each target domain, and compare with full-fine-tuning.
 Note that the last line reports the results of 6 different fine tuning procedures.}
-  \label{tab:finetuning}
-\end{table*}
-
-\fyTodo{This table does not work- One line would be enough, where we compare autoFT with true FT}
+  \label{tab:finetuning-old}
+\end{table*}\fyTodo{This table does not work- One line would be enough, where we compare autoFT with true FT}
 
@@
 \fyTodo{Curves for bank / medical ou law}
-
-\fyTodo{Fine tune pour 2 domains}
+\fyTodo{Fine tune pour 2 domains ?}
 
@@
 
-In this scenario, DDS is not as good as using a fixed distribution, with a well-chosen $\alpha$.
-
-Looking again at the evolution of $\vlambda^l$ during training we often observe a rich-get-richer effect, where a small number of domains manage to attract most of the samples, while others are almost completely ignored after a given number of iterations.
-This is illustrated on Figure~\ref{}, where we see..., and also when we look at the entropy of $\vlambda^l$ at the end of training, which also measures the KL divergence between $\vlambda^l$ and a uniform $\vlambda^t$.\fyTodo{Give numbers}
+It is however striking to see that the distribution at the end of training is concentrated on few domains, and is quite remote from the targeted test distribution.
+Looking at the evolution of $\vlambda^l$ during training, we often observe a rich-get-richer effect, where a small number of domains manage to attract most of the samples, while others are almost completely ignored after a given number of iterations.
+This is illustrated on Figure~\ref{}, where we see....\fyTodo{Give illustration and numbers}
 
 \subsection{Analysis}
+
 \subsubsection{Winner-takes-all}

### 2021-04-14 2021/emnlp/mdmt.tex
more todos
related work in progress, more refs

@@
 In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key component in training is then to mitigate the detrimental effects of a possible mismatch between these distributions.
-Single-source and multi-source domain adaptation (DA) is a well-studied instance of this setting (see \cite{Chu2017comparison} for a review) , and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+Single-source and multi-source domain adaptation (DA) is a well-studied instance of this setting (see \citep{Chu2017comparison} for a review) , and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
 A related situation is multilingual machine translation \cite{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
@@
 Finding the optimal balance of training data is however a challenging task due, for instance, to similarity between domains / languages, but also due to regularization effects of out-of-domain data; it may also be suboptimal, as some target domains or languages might be easier to train than others.
-Several recent proposals \cite{Wang17instance,Kumar19reinforcement,Wang20learning-multi} have explored ways to instead consider \emph{dynamic} data selection strategies.
+Several recent proposals \cite{Wang17instance,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi} have explored ways to instead consider more \emph{dynamic} data selection and sampling strategies.
 We study in this contribution the proposal of \citet{Wang20balancing}, initially introduced for multilingual MT, with the goal to evaluate its potential in various single-domain and multi-domain adaptation settings and assess DDS, both in its basic version and with our improvements, as an effective replacement to standard heuristics approaches.
@@
     domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline
-    Full-FT($d$) & \\
-    \system{DDS($\vlambda_0, \indic{d}$)} & \\
-    $- \log \lambda_*^{l}(d)$ ou \% & \\
-    $ \mathcal{H}(\vlambda_*^{l})$ &  \\ \hline\hline
-    \system{Mixed-0} & \\
-    \system{Mixed-1} & \\
-    \system{DDS($\vlambda_1, \vlambda_0$)} & \\
+    FT-Full($d$) \hfill [BLEU] & 37.74&59.21	&54.49&46.81&90.77&33.98&53.83\\
+    FT-Res($d$) \\ \hline
+    \hline
+    \system{DDS($\vlambda_0, \vlambda_d$)} \hfill [BLEU] & \\
+    \% domain $d$ & \\ \hline
+    \system{DDS($\vlambda_0, \vlambda_d$)} \hfill [BLEU]&  \\ 
+    \% domain $d$& \\ \hline\hline
+    % Multidomain
+    \system{Mixed-0}      & 35.26 & 54.09 & 52.49 & 31.86 & 44.94 & 89.54& 51.36 \\
+    \system{Mixed-0.25} & 35.92 & 54.87 & 52.55 & 32.55 & 44.98 & 90.28& 51.86 \\
+    \system{Mixed-0.5}   & 36.11 & 55.37 & 51.76 & 33.52 & 46.23 & 89.99& 52.16 \\
+    \system{Mixed-0.75} & 36.53 & 55.03 & 51.15 & 33.98 & 44.28 & 87.22& 51.36 \\
+    \system{Mixed-1}     & 37.33  & 54.56 & 50.05 & 33.47 & 43.23 & 77.51& 49.36 \\
+    \system{FT-Res} & \\\hline
+    \system{DDS($\vlambda_1, \vlambda_0$)} & 55.07&50.96&33.49&43.41&90.76& 51.83 \\
     $\operatorname{KL}(\vlambda_*^{l}, \vlambda_0$) & \\
@@
   \end{tabular}
-  \caption{Baseline results for DDS.
-In the top part, for each domain, we report the scores of DDS targeting solely this domain and compare with full-fine-tuning.
-We also compute the percentage of in-domain data in the resulting mixture and its entropy.}
-  \label{tab:finetuning}
+  \caption{Baseline results for DDS in fine-tuning and multi-domain adaptation.
+In the top part, we report scores with DDS and each target domain, and compare with full-fine-tuning: each column corresponds to a different system.
+We also report the proportion of in-domain data in the final mixture.
+In the bottom part, we run multi-domain experiments: for a given line, all the columns \emph{correspond to the same system}.
+We also report the KL divergence between the optimal and the target mixtures.}
+  \label{tab:basic-dds}
 \end{table*}
 
-A first observation is DDS only better than FT for \domain{rel}, and worse for \domain{med} and \domain{law}.
+A first observation (top part of Table~\ref{tab:basic-dds}) is that DDS is only better than FT for \domain{rel}, and worse for \domain{med} and \domain{law}.
+For the three other domains, we do not see any marked difference, which we view as a positive result.
 There seem to be a connection with the domain size, as \domain{rel} is small and benefits from data from other domains, while the opposite effect is observed for the larger \domain{med} and \domain{law}.
-For the three other domains, we do not see any marked difference, which we view as a positive result.
+Averaged over the 6 domains, we see a small (0.5 BLEU) difference between the two approaches for the domain of interest.
+Note however that DDS also mitigates the catastrophic forgetting effect observed with full fine-tuning (see supplemental material).\fyTodo{Supplementary}
 
-If we now look at the multi-domain results (bottom part of Table~\ref{}), we see that DDS actually does better than the two baseline strategies, but is not as good as using a fixed distribution with a well-chosen $\alpha$....\fyTodo{To be continued}
+\subsubsection{Multi-domain experiments}
 
-% \begin{table*}
+If we now look at the multi-domain experiments (bottom part of Table~\ref{tab:basicdds}), we see that DDS actually does better than the two baseline strategies that use an untuned training distribution, but is not as good as using a fixed distribution with a well-chosen $\alpha$.
+When comparing with an improved (multi-)domain adaptation techniques, based on residual adapters \cite{Bapna19simple,Pham20astudy}, we observe an even larger gap in performance.
+
+In addition we observe ...\fyTodo{To be continued}
+
+\fyTodo{Fixed regime with target mixture weigths ?}
+% \begin{table*}[htb]
 %   \centering% \small
-%   \begin{adjustbox}{width=2\columnwidth,center}
-%   \begin{tabular}{|p{2cm}|*{13}{r|}} \hline
-%     \multirow{2}{*}{Training} & \multicolumn{6}{|c}{$\lambda^t_d$} & \multicolumn{6}{|c|}{BLEU} & \multirow{2}{*}{BLEU} \\ \cline{2-13}	
-%   mixture & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & average \\
+%   \begin{adjustbox}{width=1.0\textwidth,center}
+%   \begin{tabular}{|p{3.0cm}|*{13}{r|}} \hline
+%     \multirow{2}{*}{Name} & \multirow{2}{*}{$sim\_iter$} & \multirow{2}{*}{parameterization} & \multirow{2}{*}{entropy constraint} & \multicolumn{7}{|c|}{BLEU} \\ \cline{5-11}	
+%    & & & & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{it}} & \multicolumn{1}{c|}{\domain{ rel }} & \multicolumn{1}{c|}{\domain{ talk}} & \multicolumn{1}{c|}{mean} \\
 %   \hline
-%   \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &37.45&51.26&45.47&31.62&37.55&65.22& \\
-%   \system{FT-Res} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-%   \system{Multi-DDS} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-%   \system{FT-Full} & 0.0 &1.0 & 0.0 &0.0 &0.0 & 0.0 &35.92& 54.87& 52.55& 32.55& 44.98& 90.28& 51.86\\
-%   \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-%   \system{Multi-DDS} & 0.0 &1.0 & 0.0 &0.0 &0.0 & 0.0 &35.92& 54.87& 52.55& 32.55& 44.98& 90.28& 51.86\\
-%   \system{FT-Full} & 0.0 &0.0 & 1.0 &0.0 &0.0 & 0.0 &36.11& 55.37& 51.76& 33.52& 46.23& 89.99& 52.16\\
-%   \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-%   \system{Multi-DDS} & 0.0 &0.0 & 1.0 &0.0 &0.0 & 0.0 &36.11& 55.37& 51.76& 33.52& 46.23& 89.99& 52.16\\
-%   \system{FT-Full} & 0.0 &0.0 & 0.0 &1.0 &0.0 & 0.0 &36.53&	55.03& 51.15& 33.98& 44.28& 87.22& 51.365\\
-%   \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-%   \system{Multi-DDS} & 0.0 &0.0 & 0.0 &1.0 &0.0 & 0.0 &36.53&	55.03& 51.15& 33.98& 44.28& 87.22& 51.365\\
-%   \system{FT-Full} & 0.0 &0.0 & 0.0 &0.0 &1.0 & 0.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
-%   \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-%   \system{Multi-DDS} & 0.0 &0.0 & 0.0 &0.0 &1.0 & 0.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
-%   \system{FT-Full} & 0.0 &0.0 & 0.0 &0.0 &0.0 & 1.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
-%   \system{FT-Full} & 1.0 &0.0 & 0.0 &0.0 &0.0 & 0.0 &35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-%   \system{Multi-DDS} & 0.0 &0.0 & 0.0 &0.0 &0.0 & 1.0 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
+%   \system{$DDS\_FT (med)$} & 1 & softmax & no & 36.7&51.14&52&44.32&90.41&33.22&51.3\\
+%   \system{$DDS\_FT\_law$} & 1 & softmax & no &36.05&56.18&53.57&44.05&91.24&33.09&52.36\\
+%   \system{$DDS\_FT\_bank$} & 1 & softmax & no &36.15&54.17&54.29&41.33&89.95&31.54&51.24 \\
+%   \system{$DDS\_FT\_IT$} & 1 & softmax & no &34.39&48.98&52.82&46.8&85.3&31.37&49.94 \\
+%   \system{$DDS\_FT\_rel$} & 1 & softmax & no & 34.54&52.29&51.46&44.8&91.77&31.84&51.12\\
+%   \system{$DDS\_FT\_talk$} & 1 & softmax & no & 35.53&50.55&52.64&44.86&85.8&33.47&50.48\\
+%   \system{$FT\_full$} & NA & NA & NA & 37.74&59.21	&54.49&46.81&90.77&33.98&53.83\\
 %   \hline
+%   $\Delta$ & \\
 %   \end{tabular}
 %   \end{adjustbox}
-%   \label{tab:redomains}
+%   \caption{DDS for fine-tuning and multi-domain adaptaion.
+In the top part, we report scores with DDS and each target domain, and compare with full-fine-tuning: each column corresponds to a different system.
+In the bottom part, we run multi-domain experiments: for a given line, all the columns \alert{correspond to the same system}.}
+%   \label{tab:finetuning-old}
 % \end{table*}
-
-\begin{table*}[htb]
-  \centering% \small
-  \begin{adjustbox}{width=1.0\textwidth,center}
-  \begin{tabular}{|p{3.0cm}|*{13}{r|}} \hline
-    \multirow{2}{*}{Name} & \multirow{2}{*}{$sim\_iter$} & \multirow{2}{*}{parameterization} & \multirow{2}{*}{entropy constraint} & \multicolumn{7}{|c|}{BLEU} \\ \cline{5-11}	
-   & & & & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{it}} & \multicolumn{1}{c|}{\domain{ rel }} & \multicolumn{1}{c|}{\domain{ talk}} & \multicolumn{1}{c|}{mean} \\
-    \hline
-  \system{$DDS\_FT (med)$} & 1 & softmax & no & 36.7&51.14&52&44.32&90.41&33.22&51.3\\
-  \system{$DDS\_FT\_law$} & 1 & softmax & no &36.05&56.18&53.57&44.05&91.24&33.09&52.36\\
-  \system{$DDS\_FT\_bank$} & 1 & softmax & no &36.15&54.17&54.29&41.33&89.95&31.54&51.24 \\
-  \system{$DDS\_FT\_IT$} & 1 & softmax & no &34.39&48.98&52.82&46.8&85.3&31.37&49.94 \\
-  \system{$DDS\_FT\_rel$} & 1 & softmax & no & 34.54&52.29&51.46&44.8&91.77&31.84&51.12\\
-  \system{$DDS\_FT\_talk$} & 1 & softmax & no & 35.53&50.55&52.64&44.86&85.8&33.47&50.48\\
-  \system{$FT\_full$} & NA & NA & NA & 37.74&59.21	&54.49&46.81&90.77&33.98&53.83\\
-    \hline
-    $\Delta$ & \\
-  \end{tabular}
-  \end{adjustbox}
-  \caption{DDS as a fine-tuning recipe.
-We report scores with DDS and each target domain, and compare with full-fine-tuning.
-Note that the last line reports the results of 6 different fine tuning procedures.}
-  \label{tab:finetuning-old}
-\end{table*}\fyTodo{This table does not work- One line would be enough, where we compare autoFT with true FT}
-
-\fyTodo{Works best for small domains ?}
+\fyDone{This table does not work- One line would be enough, where we compare autoFT with true FT}
 \fyTodo{Curves for bank / medical ou law}
@@
 
-\subsubsection{Multi-domain experiments}
-Results for our multidomain experiments are in Table~\ref{tab:multidomains}.\fyTodo{Same table, line autoMD.
-We should be clear about the dev / test distribution - uniform ?
-so I would compare MIX with DDS(UNIF, UNIF) - DDS(NAT,UNIF) and not report / domain scores}
+It is notably striking to see that the distribution at the end of training ($\vlambda^l_*$) is concentrated on few domains and is quite remote from the target test distribution.
+Looking at the evolution of $\vlambda^l$ during training, we often observe a rich-get-richer effect, where a small number of domains manage to attract most of the samples, while others are almost completely ignored after a given number of iterations.
+This is illustrated on Figure~\ref{}, where we see....\fyTodo{Provide illustrations and numbers}
 
-\begin{table*}
-  \centering% \small
-  \begin{adjustbox}{width=2\columnwidth,center}
-  \begin{tabular}{|p{3.0cm}|*{13}{r|}} \hline
-    \multirow{2}{*}{Training} & \multicolumn{6}{|c}{$\lambda^t_d$} & \multicolumn{6}{|c|}{BLEU} & \multirow{2}{*}{BLEU} \\ \cline{2-13}	
-  mixture & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & average \\
-    \hline
-  \system{Mixed-$0.0$} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 & 35.26 &54.09 &52.49& 31.86& 44.94& 89.54& 51.36\\
-  \system{Mixed-$0.25$} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 &35.92& 54.87& 52.55& 32.55& 44.98& 90.28& 51.86\\
-  \system{Mixed-$0.5$} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 &36.11& 55.37& 51.76& 33.52& 46.23& 89.99& 52.16\\
-  \system{Mixed-$0.75$} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 &36.53&	55.03& 51.15& 33.98& 44.28& 87.22& 51.365\\
-  \system{Mixed-$1.0$} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 &37.33& 54.56& 50.05& 33.47& 43.23& 77.51& 49.36\\
-  \system{Multi-DDS} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 & 37.26 & 55.07&50.96&33.49&43.41&90.76& 51.83 \\
-  \system{Multi-DDS-v1} & 0.166&0.166 &0.166 &0.166 &0.166 & 0.166 & 36.96 & 55.55 & 52.88 & 33.06 & 44.52 &91.25 & 52.37\\
-  \hline
-  \end{tabular}
-\end{adjustbox}
-  \label{tab:multidomains}
-\end{table*}
+\subsection{Analysis and improvements}
 
-It is however striking to see that the distribution at the end of training is concentrated on few domains, and is quite remote from the targeted test distribution.
-Looking at the evolution of $\vlambda^l$ during training, we often observe a rich-get-richer effect, where a small number of domains manage to attract most of the samples, while others are almost completely ignored after a given number of iterations.
-This is illustrated on Figure~\ref{}, where we see....\fyTodo{Give illustration and numbers}
-
-\subsection{Analysis}
+\subsection{Unsupervised DDS}
+To better analyse the behaviour of DDS, we study a scenario where $\vlambda^s$ and $\vlambda^t$ have different dimensions designed as follows.
+All training data is automatically clustered into $K=30$ of varying sizes using the k-means algorithm based on generic sentence representations obtained via mean pooling.
+The resulting classes contain from a tenths of thousands to several hundreds of thousand training samples.
+Most classes contain a dominant domain, which accounts for more of 90\% of the samples.\footnote{See detail in supplementary material} Using the same DA setting as before, we study (a) whether DDS is able to identify the classes that correspond to the domain of interest (b) whether having more degree of freedoms can yield an improved training performance (c) how DDS scales when more domains are introduced.\fyTodo{TBC - with what experiments?}
 
 \subsubsection{Winner-takes-all}
-\cite{Wang20balancing} used a simple softmax function and the following parameterization of the sampler $\lambda^l(d;\psi) = \operatorname{softmax}(\psi)$ where $\psi \in \mathbb{R}^{n_d}$.
-This parameterization of the sampler will induce the following gradient:\fyTodo{Change name of obj.
+
+Following \citep{Wang20balancing}, our baseline uses the following parameterization of the sampling component $\vlambda^l(\vpsi) = \operatorname{softmax}(\vpsi)$ where $\vpsi \in \mathbb{R}^{n_d}$.
+This parameterization induces the following gradients:\fyTodo{Change name of obj.
 function}
 \begin{align*}
-\frac{\partial XXX}{\partial \psi_d} & = \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \frac{\partial \lambda^l(j; \psi)}{\partial \psi_d} \\
+\frac{\partial XXX}{\partial \vpsi[d]} & = \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \frac{\partial \lambda^l(j; \vpsi)}{\partial \vpsi[d]} \\
 %	& = R_d \lambda^l(d; \psi)(1- \lambda^l(d; \psi)) - \displaystyle{\mathop{\sum}_{j\neq d}^{n_d}} R_j \lambda^l(d; \psi) \lambda^l(j; \psi)\\
@@
 \end{align*}
-The fact that the change of $\psi_d$ is proportional to $\lambda^l(d; \psi)$ can explain that the optimization often leads to a degenerate solution in which the sampler ignore almost domains except a few ones.
-In order to avoid the appearance of a ``rich-gets-richer'' effect and degenerate solutions such as Fig.~\ref{},
+
+The fact that the change of $\psi(d)$ is proportional to $\lambda^l(d; \vpsi)$ explains that the optimization often leads to a degenerate solution in which only a small number of domains are considered.
+This behaviour is even more visible when using 30 automatic domains in training:\fyTodo{What results would be useful ?}.
+
+In order to mitigate this behavior, two variants of the baseline are investigated:
+\begin{itemize}
+\item to ensure that $\vpsi$ remains sufficient diverse or close to the target distribution $\vlambda^t$. This is obtained by adding a constraint when updating $\vpsi$. 
+\item to replace the softmax function for which the rich-get-richer effect are attenuated. We consider functions of the form $f(\vpsi)[d] = \frac{\vpsi[d]^\beta}{\sum_d \vpsi[d]^\beta}$.
+\begin{align*}
+\psi[d] = \frac{\psi[d] + \eta * R_d}{ 1 + \eta \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j},
+\end{align*}
+where $\eta$ controls the learning rate of $\vpsi$.
+\end{itemize}
 
@@
 
-\subsection{Stabilizing the sampler \label{ssec:sampler}}
+\subsubsection{Improving the reward computation}
 
-To stabilize the sampler, we consider two alternatives. The first one is to replace the softmax function, yielding the following updates:\fyTodo{List options}
+Figures such as Figure~\ref{} show that the training signal given by the reward~\ref{eq:reward} can be very noisy and misleading.
+For instance, we observe that even when targeting the \domain{medical} domain, the \domain{religion} data keeps being rewarded, and constitues a significant portion of the final mixture.\fyTodo{Which figure, which equation ?}\fyTodo{Do we have number that show variance and strong divergence wrt initial training point ?}
+
+Again, we have explored several options to improve this aspect of the algorithm.
+The first was to compute better estimates of the gradients in~\eqref{} by using larger batches, or performing multiple updates when computing $\vtheta_t$.
+This did not prove useful, and this option was quickly discontinued.
+The second, more promising alley, was to base the reward computation not on the gradient of the loss, but more directly on the development loss, replacing line~\ref{alg:number}\fyTodo{which line ?} with the following reward computation:
 \begin{align*}
-\psi_d = \frac{\psi_d + \eta * R_d}{ 1 + \eta \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j},
+  \operatorname{TODO}
 \end{align*}
-where $\eta$ controls the learning rate of $\psi$.
+This idea is reminiscent of the work of \citet{Kumar19reinforcement}, who also design their curriculum trainng strategy based on the development loss.
 
-The second one simply adds an entropy constraint which ensures that the entropy of the domain distribution is always greater than a predefined value $\alpha$.\fyTodo{Explain the math}
+\subsection{Improving dynamic data selection}
+
+Experiments with these revised versions are for various settings in Table~\ref{tab:improved-dds}.
+Results for DA directly compare with Table~{tab:basic-dds}; results for unsupervised DA used our 6 domains for training, but test on the \domain{News}, unseen in traning.
+\begin{table*}
+  \centering \small
+  \begin{tabular}{|l|*9{r|}} \hline
+    \multicolumn{9}{|c|}{\sl Supervised domain adaptation} \\ \hline
+    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} \\ \hline \hline
+    \multicolumn{9}{|c|}{\sl Unsupervised domain adaptation} \\ \hline
+    \\  \hline\hline 
+    \multicolumn{9}{|c|}{\sl Multi-domain training} \\ \hline
+    \\
+    \system{DDS($\vlambda_1, \vlambda_0$)} & \\
+    \system{DDS($\vlambda_0, \vlambda_0$)} & \\ \hline
+  \end{tabular}
+  \caption{Improved results for DDS in fine-tuning and various multi-domain adaptation settings.}
+  \label{tab:basic-dds}
+\end{table*}
+
+\fyTodo{Refaire table 3 + (unsupervised domain in source= classification) + new domain}
 
 \section{Related Work \label{sec:related}}
-\cite{Wang17instance} uses dynamic weighting (for sentences and domains) for DA.
 
-\cite{Wang20learning-multi} adapts curriculum learning \cite{BengioXX} for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
+Domain adaptation is an ancient and vexing problem that has been studied from many angles both for SMT and NMT.
+A survey of supervised and unsupervised DA for NMT is in \cite{Chu18asurvey}, where they distinguish between data-centric and model-centric DA, a view also adapted in the more recent review of \cite{Saunders20domain}.
+Our approach to DA in this paper clearly falls under the former category.
+We refer interested readers to these refrences, and do not discuss review DA any further.
 
-Also \cite{Saunders19adaptive}.
+Multi-domain adaptation (MDA) aims to develop systems that simultaneously bode well for several domains.
+Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{KobKobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citep{Jiang19multidomain}.
+Unsupervised MDA is considered in \citep{Farajian17multidomain} as an unsupervised DA problem.
+
+Most approaches to adaptive / dynamic data selection take inspiration from \citep{Bengio09curriculum}, where the notion of \emph{curriculum learning} (CL) is initially introduced.
+CL is relies on a notion of ``easyness'' of a sample to schedule the presentation of the training data so that the easiest examples  presented first and the hardest, last.
+CL has been showed to speed up learning.
+While the initial aim was to improve and speed up training, it has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitons of ``easyness''.
+
+For instance, \citet{Zhang19curriculum} study supervised domain adaptation, and propose a curriculum approach which progressively extends the data used in training: in the early stages only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
+This is somehow opposite to the recommendations of \cite{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyTodo{These have not been compared ?
+and also to what we do ?}
+
+\citet{Kumar19reinforcement} use reinforcement techniques (deep Q-learning) to learn the curriculum strategy: in this study, complexity corresponds to difficulty levels and which are binned based on contrastive data selection.
+The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyTodo{Alert: what do we do during warm up ?} The same technique is recently applied to the training of a multilingual NMT system \cite{Kumar21learning}.
+\citet{Zhou20uncertainty} propose another curriculum-based approach which instead relies on \emph{instance uncertainty} as a measure of their difficulty and presents the data sample starting with the easiest (more predictable) first.
+Another contribution of this paper is an alternative criterium for stopping the training.
+More related to our problems, \citet{Wang20learning-multi} adapt curriculum learning for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization (derivative-free) techniques.
+Each step consists in (a) weighting the instances based on relevance features, (b) fine tuning current model using the (weighted) training set, and is applied iteratively to generate a sequence of models.
+The one that maximizes the development set performance is finally retained.
 

### 2021-04-21 2021/emnlp/mdmt.tex
table 3 DDS score and % domain

@@
 \end{align*}
-Given $\vtheta_t$, $\psi_t$ is updated via the Reinforce algorithm \cite{Williams92reinforce}:\fyTodo{Rewrite this correctly with domain average reward}
+Given $\vtheta_t$, $\psi_t$ is updated via the Reinforce algorithm \cite{Williams92simple}:\fyTodo{Rewrite this correctly with domain average reward}
 %\begin{align*}
@@
     domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline
-    FT-Full($d$) \hfill [BLEU] & 37.74&59.21	&54.49&46.81&90.77&33.98&53.83\\
-    FT-Res($d$) \\ \hline
+    FT-Full($d$) \hfill [BLEU] &37.74&59.21&54.49&33.98&46.81&90.77&53.8\\
+    FT-Res($d$) \hfill &37.3&57.93&53.91&33.79&46.69&90.17&53.3\\ \hline
     \hline
-    \system{DDS($\vlambda_0, \vlambda_d$)} \hfill [BLEU] & \\
-    \% domain $d$ & \\ \hline
+    \system{DDS($\vlambda_0, \vlambda_d$)} \hfill [BLEU] &36.7&56.18&54.29&33.47&46.8&91.77&53.2\\
+    \% domain $d$ &59.7&14.9&86.1&6.4&67.2&6.4\\ \hline
     \system{DDS($\vlambda_0, \vlambda_d$)} \hfill [BLEU]&  \\ 
@@
     \system{Mixed-1}     & 37.33  & 54.56 & 50.05 & 33.47 & 43.23 & 77.51& 49.36 \\
-    \system{FT-Res} & \\\hline
-    \system{DDS($\vlambda_1, \vlambda_0$)} & 55.07&50.96&33.49&43.41&90.76& 51.83 \\
+    \system{MDL-Res} &37.08&57.42&53.57&33.17&46.01&90.65&52.98\\\hline
+    \system{DDS($\vlambda_1, \vlambda_0$)} & \\
     $\operatorname{KL}(\vlambda_*^{l}, \vlambda_0$) & \\
     $ \mathcal{H}(\vlambda_*^{l})$ &  \\
-    \system{DDS($\vlambda_0, \vlambda_0$)} & \\ \hline
+    \system{DDS($\vlambda_0, \vlambda_0$)} &37.26&55.07&50.96&33.49&43.41&90.76&51.83 \\ \hline
   \end{tabular}
@@
 Domain adaptation is an ancient and vexing problem that has been studied from many angles both for SMT and NMT.
-A survey of supervised and unsupervised DA for NMT is in \cite{Chu18asurvey}, where they distinguish between data-centric and model-centric DA, a view also adapted in the more recent review of \cite{Saunders20domain}.
+A survey of supervised and unsupervised DA for NMT is in \cite{Chu18asurvey}, where they distinguish between data-centric and model-centric DA, a view also adapted in the more recent review of \cite{Saunders21domain}.
 Our approach to DA in this paper clearly falls under the former category.
@@
 Multi-domain adaptation (MDA) aims to develop systems that simultaneously bode well for several domains.
-Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{KobKobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citep{Jiang19multidomain}.
+Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citep{Jiang19multidomain}.
 Unsupervised MDA is considered in \citep{Farajian17multidomain} as an unsupervised DA problem.

### 2021-04-21 2021/emnlp/mdmt.tex
After discussion with MQP

@@
     \% domain $d$ &59.7&14.9&86.1&6.4&67.2&6.4\\ \hline
-    \system{DDS($\vlambda_0, \vlambda_d$)} \hfill [BLEU]&  \\ 
+    \system{DDS($\vlambda_1, \vlambda_d$)} \hfill [BLEU]&  \\ 
     \% domain $d$& \\ \hline\hline
@@
     $\operatorname{KL}(\vlambda_*^{l}, \vlambda_0$) & \\
-    $ \mathcal{H}(\vlambda_*^{l})$ &  \\
+%    $ \mathcal{H}(\vlambda_*^{l})$ &  \\
     \system{DDS($\vlambda_0, \vlambda_0$)} &37.26&55.07&50.96&33.49&43.41&90.76&51.83 \\ \hline
@@
 The resulting classes contain from a tenths of thousands to several hundreds of thousand training samples.
-Most classes contain a dominant domain, which accounts for more of 90\% of the samples.\footnote{See detail in supplementary material} Using the same DA setting as before, we study (a) whether DDS is able to identify the classes that correspond to the domain of interest (b) whether having more degree of freedoms can yield an improved training performance (c) how DDS scales when more domains are introduced.\fyTodo{TBC - with what experiments?}
+Most classes contain a dominant domain, which accounts for more of 90\% of the samples.\footnote{See detail in supplementary material} Using the same DA setting as before, we study (a) whether DDS is able to identify the classes that correspond to the domain of interest (b) whether having additional degrees of freedom can yield an improved training performance (c) how DDS scales when a large number of  domains is considered.\fyTodo{TBC - with what experiments?}
 
@@
 \subsection{Improving dynamic data selection}
-
 Experiments with these revised versions are for various settings in Table~\ref{tab:improved-dds}.
@@
 \section{Conclusion and outlook \label{sec:discussion}}
+\fyTodo{Are we learning the dev set ?} Boostrap the dev ?
+
 \section*{Acknowledgments}

### 2021-04-26 2021/emnlp/mdmt.tex
more results, change the formulation of the algo, add the formula of reward derived from dev-loss, add new reference in Automated Curriculumn Learning

@@
 \newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+
 \newcommand{\revision}[1]{\textcolor{red}{#1}}
@@
 \maketitle
+\setlength{\abovedisplayskip}{2pt}
+\setlength{\belowdisplayskip}{2pt}
 \begin{abstract}
@@
 
-\fyTodo{Label or covariate shift ?}
+\fyTodo{Label or covariate shift?} \revision{Answer: Covariate shift happens when D(x) differs between train and test; label shift happens when the conditional distribution $D(y|x)$ differs between train and test}
 This problem is often approached with \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions.
@@
 Note that this strategy is potentially suboptimal, as some out-of-domain samples may contribute to the final performance due to eg.\ domain overlap.
-Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains might be easier to learn than others.\fyTodo{How to measure this?}
+Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains might be easier to learn than others.\fyTodo{How to measure this?} \revision{Answer: The graphic of \system{Mixed-0} shows that with same amount of data(batches) the convergences of the domains vary significantly}
 
@@
 Sampling training domains using a distribution $\lambda^{l}$ results in a parameter estimate $\theta(\lambda^{l})$, delivering performance that can be estimated on the development set using $\lambda^{t}$.
-\citet{Wang20optimizing,Wang20balancing} then formulate the problem of finding an optimal sampling strategy $\lambda^{l}_{*}$ maximizing performance on the development data as the following bi-level optimization problem:
-\begin{equation} \label{eq:bilevel}
-\begin{split}
-\vlambda^{l}_* &= \displaystyle{\mathop{\argmin}_{\lambda^{l}} J(\theta^*(\vlambda^{l}), \vlambda^{t})} \\
-					& \theta^*(\vlambda^{l}) = \displaystyle{\mathop{\argmin_{\vtheta}} J(\vtheta, \vlambda^{l})}, 
-\end{split}
-\end{equation}
-where we use the following notations:% \fyTodo{Overload Q, from domain to instances to pairs}
+\citet{Wang20optimizing,Wang20balancing} then formulate the problem of finding an optimal sampling strategy $\lambda^{l}_{*}$ maximizing performance on the development data as a bi-level optimization problem that can be approached by coordinate gradient descent alternating between $\lambda^{l}$ and $\vtheta$.
+\citet{Wang20balancing} limit the search space of $\vlambda^{l}$ to a parameterized differentiable function $\vlambda^{l}(\vpsi)$.
+Given $\vpsi_t$ at time step $t$, $\vtheta_t$ is updated via usual gradient descent as follows.
+\begin{align*}
+\vtheta_{t+1} &= \vtheta_t - \mathbf{lr}_{nmt} * \frac{1}{N} \displaystyle{\mathop{\sum}_{i=1}^N} \frac{\partial l(\vtheta_t, x_i,y_i)}{\partial \vtheta} \\
+x_i,y_i &\sim \widehat{\mathcal{D}^l}(x;\vlambda^l(\psi_t)) = \displaystyle{\mathop{\sum}_{d=1}^K} \lambda^l(d;\psi_t) * \widehat{D^l_d}(x) \\
+\end{align*}
+Where $l(\theta_t,x,y)$ is the loss on an example.
 
+Given $\vtheta_t$, $\psi_t$ is updated via the Reinforce algorithm \cite{Williams92simple}:\fyTodo{Rewrite this correctly with domain average reward}
 \begin{align*}
-J(\vtheta, \vlambda^{t} ) &= \displaystyle{\mathop{\sum}_{(x,y) \sim \widehat{\mathcal{D}^{t}(x)}} \ell(\vtheta,x,y)} \\
-J(\vtheta, \vlambda^{l}) &= \displaystyle{\mathop{E_{(x,y) \sim \widehat{\mathcal{D}^{l}(x)}}[\ell(\vtheta,x,y)]}} \\
-l(\vtheta,x,y) &= - log P(x,y;\vtheta)
+\psi_{t+1} &= \psi_t + \mathbf{lr}_{data} * \displaystyle{\mathop{\sum}_{d=1}^K} R(d) * \frac{\partial log(\lambda^l(d;\psi_t))}{\partial \psi} \\
+  R(d) &= \langle \frac{\partial J^l_d(\theta_t)}{\partial \theta}, \frac{\partial J^t(\theta', \lambda^t)}{\partial \theta} \rangle \\
 \end{align*}
-Problem~\eqref{eq:bilevel} can be approached by coordinate gradient descent alternating between $\lambda^{l}$ and $\vtheta$.
-\citet{Wang20balancing} limit the search space of $\vlambda^{l}$ to a parameterized differentiable function $\vlambda^{l}(\vpsi)$.
-Given $\vpsi_t$ at time step $t$, $\vtheta_t$ is updated via usual gradient descent as follows:
+Where
+\begingroup
+\allowdisplaybreaks
 \begin{align*}
-\vtheta_{t+1} &= \vtheta_t - \mathbf{lr}_{nmt} * \frac{\partial l(\vtheta_t, x,y)}{\partial \vtheta} \\
-x,y &\sim \widehat{\mathcal{D}^l}(x;\vlambda^l(\psi_t)) = \displaystyle{\mathop{\sum}_{d=1}^K} \lambda^l(d;\psi_t) * \widehat{D_d}(x) \\
+\theta' &= \theta_t - \mathbf{lr}_{nmt} * \displaystyle{\mathop{\sum}_{i=1,x^i_d,y^i_d \in \widehat{D^l_d}}^N} \frac{\partial l(\theta_t,x^i_d,y^i_d)}{\partial \theta} \\
+  J^l_d(\theta_t) &= \displaystyle{\mathop{\sum}_{i=1,x^i_d,y^i_d \in \widehat{D^l_d}}^N} l(\theta_t,x^i_d,y^i_d) \\
+  J^t(\theta',\lambda_t) &= \displaystyle{\mathop{\sum}_{d}}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta',x^t_d,y^t_d)\\
 \end{align*}
-Given $\vtheta_t$, $\psi_t$ is updated via the Reinforce algorithm \cite{Williams92simple}:\fyTodo{Rewrite this correctly with domain average reward}
-%\begin{align*}
-%  \psi_{t+1} &= \psi_t + \mathbf{lr}_{data} * R(x,y) * \\
-%  & \quad\quad \frac{\partial log(\lambda^l(\psi_t)(x,y))}{\partial \psi} \\
-%x,y &\sim \lambda^l (\psi_t) \\
-%R(x,y) &= \langle \frac{\partial l(\theta',x,y)}{\partial \theta}, \frac{\partial J(\theta_{t+1}, \lambda^l)}{\partial \theta} \rangle \\
-%\theta' &= \theta_t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta_t, x,y)}{\partial \theta} \\
-%\end{align*}
-
-\begin{align*}
-\psi_{t+1} &= \psi_t + \mathbf{lr}_{data} * \displaystyle{\mathop{\sum}_{d=1}^K} R(d) * \frac{\partial log(\lambda^l(\psi_t)(d))}{\partial \psi} \\
-  R(d) &= E_{x,y \sim \widehat{D_d}}[\langle \frac{\partial l(\theta_t,x,y)}{\partial \theta}, \frac{\partial J(\theta', \lambda^t)}{\partial \theta} \rangle] \\
-  \theta' &= \theta_t - \mathbf{lr}_{nmt} * \frac{\partial l(\theta_t, x,y)}{\partial \theta} \\
-\end{align*}
-
+\endgroup
 $\mathbf{lr}_{data}$ is the learning rate of the parameter of the sampler.
@@
 Transformer models contain 8~attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells.
-The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024.
+The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$EST activation function operating on normalized entries of dimension~1024.
 Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.
@@
 \fyTodo{Fix the table}
-In Table~\ref{tab:finetuning}, we compare the results of fine-tuning a system initially trained with a uniform mixture\fyTodo{uniform or natural?} of domains with the result of learning from scratch, also with a initial uniform distribution.
+In Table ~\ref{tab:basic-dds}, we compare the results of fine-tuning a system initially trained with a natural mixture, which is proportional to the sizes of domains, of domains with the result of DDS learning from scratch with a initial uniform distribution.
 \fyTodo{What do we see ?
@@
 - show a graph}.
-Note that the DDS systems are somewhat quicker to build, as we only train for 100K iterations (instead of 100K + 50K for the fine-tuning part).
+
+\fyTodo{Only if we compare on an individual domain} \revision{Note that the DDS systems are somewhat quicker to build, as we only train for 200K iterations (instead of 200K + 50K for the fine-tuning part)}.
 
@@
     \system{DDS($\vlambda_0, \vlambda_d$)} \hfill [BLEU] &36.7&56.18&54.29&33.47&46.8&91.77&53.2\\
-    \% domain $d$ &59.7&14.9&86.1&6.4&67.2&6.4\\ \hline
-    \system{DDS($\vlambda_1, \vlambda_d$)} \hfill [BLEU]&  \\ 
+    \% domain $d$ (final / total) & 59.7 / 49.5 & 6.5 / 18.3 & 14.9 / 24.5 & 6.4 / 12.8 & 67.2/47.4 & 67.2 / 45.3&\\ \hline
+    \system{DDS($\vlambda_1, \vlambda_d$)} \hfill [BLEU]&36.4&56.3& 45.4& 32.6&44.5&43.1&43.1\\
+    \% domain $d$& 89 / 86.6 & 50/18.4 & 1.4 / 1.6 & 3.0 / 2.5 & 7.5 / 9.2 & 1.0 / 1.0 \\ \hline
+    \system{CL($d$)} \hfill [BLEU]&  \\ 
     \% domain $d$& \\ \hline\hline
@@
     \system{MDL-Res} &37.08&57.42&53.57&33.17&46.01&90.65&52.98\\\hline
-    \system{DDS($\vlambda_1, \vlambda_0$)} & \\
-    $\operatorname{KL}(\vlambda_*^{l}, \vlambda_0$) & \\
+    \system{DDS($\vlambda_1, \vlambda_0$)} &37.86&50.22&47.22&31.22&40.83&58.56&44.32 \\
+    $\operatorname{KL}(\vlambda_*^{l}, \vlambda_0$) &36.96&55.55&52.88&33.06&44.52&91.25&52.37 \\
 %    $ \mathcal{H}(\vlambda_*^{l})$ &  \\
@@
 \subsection{Unsupervised DDS}
-To better analyse the behaviour of DDS, we study a scenario where $\vlambda^s$ and $\vlambda^t$ have different dimensions designed as follows.
+To better analyze the behavior of DDS, we study a scenario where $\vlambda^s$ and $\vlambda^t$ have different dimensions designed as follows.
 All training data is automatically clustered into $K=30$ of varying sizes using the k-means algorithm based on generic sentence representations obtained via mean pooling.
@@
 \begin{align*}
-\frac{\partial XXX}{\partial \vpsi[d]} & = \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \frac{\partial \lambda^l(j; \vpsi)}{\partial \vpsi[d]} \\
+\Delta \psi[d] & = \eta \times \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \frac{\partial \lambda^l(j; \vpsi)}{\partial \vpsi[d]} \\
 %	& = R_d \lambda^l(d; \psi)(1- \lambda^l(d; \psi)) - \displaystyle{\mathop{\sum}_{j\neq d}^{n_d}} R_j \lambda^l(d; \psi) \lambda^l(j; \psi)\\
-	& = \lambda^l(d; \psi) (R_d- XXX)\\
+	& = \eta \times \lambda^l(d; \psi) \big(R_d-\displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \lambda^l(j;\psi) \big)\\
 \end{align*}
+where $\eta$ controls the learning rate of $\vpsi$.
 
-The fact that the change of $\psi(d)$ is proportional to $\lambda^l(d; \vpsi)$ explains that the optimization often leads to a degenerate solution in which only a small number of domains are considered.
-This behaviour is even more visible when using 30 automatic domains in training:\fyTodo{What results would be useful ?}.
+The fact that the change of $\psi(d)$ is proportional to $\lambda^l(d; \vpsi)$ explains that the optimization often leads to a degenerated solution in which only a small number of domains are considered.
+%This behavior is even more visible when using 30 automatic domains in training:\fyTodo{What results would be useful ?}.
 
@@
 \item to ensure that $\vpsi$ remains sufficient diverse or close to the target distribution $\vlambda^t$. This is obtained by adding a constraint when updating $\vpsi$. 
-\item to replace the softmax function for which the rich-get-richer effect are attenuated. We consider functions of the form $f(\vpsi)[d] = \frac{\vpsi[d]^\beta}{\sum_d \vpsi[d]^\beta}$.
+\item to replace the softmax function for which the rich-get-richer effect are attenuated. We consider functions of the form $\lambda
+^l(d;\vpsi) = \frac{\vpsi[d]^\beta}{\sum_i \vpsi[i]^\beta}$. The gradient of $\psi$ will be as follows.
 \begin{align*}
-\psi[d] = \frac{\psi[d] + \eta * R_d}{ 1 + \eta \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j},
+\Delta\psi[d] = \eta \times \beta \frac{\lambda^l(d;\psi)}{\psi[d]} \big( R_d - \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \lambda^l(j;\psi) \big)
 \end{align*}
@@
 \begin{align*}
-  \operatorname{TODO}
+  R_d = J^t(\theta_{t+k},\lambda^t) - J^t(\theta_t,\lambda^t)
 \end{align*}
-This idea is reminiscent of the work of \citet{Kumar19reinforcement}, who also design their curriculum trainng strategy based on the development loss.
+Where 
+\begin{equation}
+\begin{array}{rcl}
+\theta_{t+i} &=& \theta_{t+i-1} - \mathbf{lr}_{nmt} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta} \\
+x^i_j, y^i_j &\sim& \widehat{D^l_d}(x)
+\end{array}
+\end{equation}
+
+This idea is reminiscent of the work of \citet{alex17automated} and  \citet{Kumar19reinforcement}, who also design their curriculum trainng strategy based on the development loss.
 
@@
     \\
-    \system{DDS($\vlambda_1, \vlambda_0$)} & \\
-    \system{DDS($\vlambda_0, \vlambda_0$)} & \\ \hline
+    \system{DDS-v1($\vlambda_0, \vlambda_0$)} &36.96&55.55&52.88&33.06&44.52&91.25&52.37&- \\
+    \system{DDS-v2($\vlambda_0, \vlambda_0$)}&35.45&55.46&53.7&32.3&45.6&91.11&52.27& - \\
+    \system{DDS-v1($\vlambda_1, \vlambda_0$)} & \\
+    \system{DDS-v2($\vlambda_1, \vlambda_0$)} & \\ \hline
   \end{tabular}
@@
 \end{algorithm}
+
+%\begin{equation} \label{eq:bilevel}
+%\begin{split}
+%\vlambda^{l}_* &= \displaystyle{\mathop{\argmin}_{\lambda^{l}} J(\theta^*(\vlambda^{l}), \vlambda^{t})} \\
+%					& \theta^*(\vlambda^{l}) = \displaystyle{\mathop{\argmin_{\vtheta}} J(\vtheta, \vlambda^{l})}, 
+%\end{split}
+%\end{equation}
+%where we use the following notations:% \fyTodo{Overload Q, from domain to instances to pairs}
+%
+%\begin{align*}
+%J(\vtheta, \vlambda^{t} ) &= \displaystyle{\mathop{\sum}_{(x,y) \sim \widehat{\mathcal{D}^{t}(x)}} \ell(\vtheta,x,y)} \\
+%J(\vtheta, \vlambda^{l}) &= \displaystyle{\mathop{E_{(x,y) \sim \widehat{\mathcal{D}^{l}(x)}}[\ell(\vtheta,x,y)]}} \\
+%l(\vtheta,x,y) &= - log P(x,y;\vtheta)
+%\end{align*}

### 2021-04-29 2021/emnlp/mdmt.tex
more scores

@@
 Accordingly, we report two aggregate scores: a weighted average reflecting the training distribution, and an unweighted average, meaning that test domains are equally important.
-\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
+\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing.
 The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems each adapted for one domain, with no parameter sharing across domains.
@@
 Transformer models contain 8~attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells.
-The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$EST activation function operating on normalized entries of dimension~1024.
+The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024.
 Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.
@@
 \fyTodo{Fix the table}
-In Table ~\ref{tab:basic-dds}, we compare the results of fine-tuning a system initially trained with a natural mixture, which is proportional to the sizes of domains, of domains with the result of DDS learning from scratch with a initial uniform distribution.
+In Table ~\ref{tab:basic-dds}, we compare the results of fine-tuning a system initially trained with a natural mixture, which is proportional to the sizes of the domains, of domains with the result of DDS learning from scratch with a initial uniform distribution.
 \fyTodo{What do we see ?
@@
     domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline
-    FT-Full($d$) \hfill [BLEU] &37.74&59.21&54.49&33.98&46.81&90.77&53.8\\
-    FT-Res($d$) \hfill &37.3&57.93&53.91&33.79&46.69&90.17&53.3\\ \hline
+    FT-Full($d$) \hfill [BLEU] &37.7&59.2&54.5&34&46.8&90.8&53.8\\
+    FT-Res($d$) \hfill &37.3&57.9&53.9&33.8&46.7&90.2&53.3\\ \hline
     \hline
-    \system{DDS($\vlambda_0, \vlambda_d$)} \hfill [BLEU] &36.7&56.18&54.29&33.47&46.8&91.77&53.2\\
+    \system{DDS($\vlambda_0, \vlambda_d$)} \hfill [BLEU] &36.7&56.2&54.3&33.5&46.8&91.8&53.2\\
     \% domain $d$ (final / total) & 59.7 / 49.5 & 6.5 / 18.3 & 14.9 / 24.5 & 6.4 / 12.8 & 67.2/47.4 & 67.2 / 45.3&\\ \hline
@@
     % Multidomain
-    \system{Mixed-0}      & 35.26 & 54.09 & 52.49 & 31.86 & 44.94 & 89.54& 51.36 \\
-    \system{Mixed-0.25} & 35.92 & 54.87 & 52.55 & 32.55 & 44.98 & 90.28& 51.86 \\
-    \system{Mixed-0.5}   & 36.11 & 55.37 & 51.76 & 33.52 & 46.23 & 89.99& 52.16 \\
-    \system{Mixed-0.75} & 36.53 & 55.03 & 51.15 & 33.98 & 44.28 & 87.22& 51.36 \\
-    \system{Mixed-1}     & 37.33  & 54.56 & 50.05 & 33.47 & 43.23 & 77.51& 49.36 \\
-    \system{MDL-Res} &37.08&57.42&53.57&33.17&46.01&90.65&52.98\\\hline
-    \system{DDS($\vlambda_1, \vlambda_0$)} &37.86&50.22&47.22&31.22&40.83&58.56&44.32 \\
-    $\operatorname{KL}(\vlambda_*^{l}, \vlambda_0$) &36.96&55.55&52.88&33.06&44.52&91.25&52.37 \\
+    \system{Mixed-0}      & 35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5& 51.4 \\
+    \system{Mixed-0.25} & 35.9 & 54.9 & 52.6 & 32.6 & 45 & 90.3& 51.9 \\
+    \system{Mixed-0.5}   & 36.1 & 55.4 & 51.8 & 33.5 & 46.2 & 90 & 52.2 \\
+    \system{Mixed-0.75} & 36.5 & 55 & 51.2 & 34 & 44.3 & 87.2& 51.7 \\
+    \system{Mixed-1}     & 37.3  & 54.6 & 50.1 & 33.5 & 43.2 & 77.5& 49.4 \\
+    \system{MDL-Res} &37.1&57.4&53.6&33.2&46&90.7&53\\\hline
+    \system{DDS($\vlambda_1, \vlambda_0$)} &37.9&50.2&47.2&31.2&40.8&58.6&44.3\\
+    $\operatorname{KL}(\vlambda_*^{l}, \vlambda_0$) &37&55.6&52.9&33.1&44.5&91.3&52.4 \\
 %    $ \mathcal{H}(\vlambda_*^{l})$ &  \\
-    \system{DDS($\vlambda_0, \vlambda_0$)} &37.26&55.07&50.96&33.49&43.41&90.76&51.83 \\ \hline
+    \system{DDS($\vlambda_0, \vlambda_0$)} &37.3&55.1&51&33.5&43.4&90.8&51.8\\ \hline
   \end{tabular}
@@
     \multicolumn{9}{|c|}{\sl Unsupervised domain adaptation} \\ \hline
+    \system{DDS-v1($\vlambda_0, \vlambda_{news}$)} & \\  \hline\hline 
+    \multicolumn{9}{|c|}{\sl 30 clusters} \\ \hline
     \\  \hline\hline 
@@
     \\
+    \system{DDS-v1($\vlambda_0, \vlambda_d$)} &36.7&57.6&54.9&33.3&46.1&91.9&53.4&- \\
+    \system{DDS-v1($\vlambda_1, \vlambda_d$)} & \\
     \system{DDS-v1($\vlambda_0, \vlambda_0$)} &36.96&55.55&52.88&33.06&44.52&91.25&52.37&- \\
-    \system{DDS-v2($\vlambda_0, \vlambda_0$)}&35.45&55.46&53.7&32.3&45.6&91.11&52.27& - \\
-    \system{DDS-v1($\vlambda_1, \vlambda_0$)} & \\
-    \system{DDS-v2($\vlambda_1, \vlambda_0$)} & \\ \hline
+    \system{DDS-v1($\vlambda_1, \vlambda_0$)} & \\ \hline
   \end{tabular}

### 2021-05-05 2021/emnlp/mdmt.tex
petits changements

@@
   \caption{Baseline results for DDS in fine-tuning and multi-domain adaptation.
-In the top part, we report scores with DDS and each target domain, and compare with full-fine-tuning: each column corresponds to a different system.
+In the top part, we report scores with DDS and each target domain, and compare with full fine-tuning: each column corresponds to a distinct system.
 We also report the proportion of in-domain data in the final mixture.
@@
 \end{table*}
+\fyTodo{Report CL results / explain in legend}
 
@@
 The resulting classes contain from a tenths of thousands to several hundreds of thousand training samples.
-Most classes contain a dominant domain, which accounts for more of 90\% of the samples.\footnote{See detail in supplementary material} Using the same DA setting as before, we study (a) whether DDS is able to identify the classes that correspond to the domain of interest (b) whether having additional degrees of freedom can yield an improved training performance (c) how DDS scales when a large number of  domains is considered.\fyTodo{TBC - with what experiments?}
+Most classes contain a dominant domain which accounts for more of 90\% of the samples.\footnote{See detail in supplementary material} Using the same DA setting as before, we study (a) whether DDS is able to identify the classes that correspond to the domain of interest (b) whether having additional degrees of freedom\fyTodo{explain better} can yield an improved training performance (c) how DDS scales when a large number of  domains is considered.\fyTodo{TBC - with what experiments?}
 
@@
 
-This idea is reminiscent of the work of \citet{alex17automated} and  \citet{Kumar19reinforcement}, who also design their curriculum trainng strategy based on the development loss.
+This idea is reminiscent of the work of \citet{Graves17automated} and \citet{Kumar19reinforcement}, who also design their curriculum trainng strategy based on the development loss.
 

### 2021-05-05 2021/emnlp/auto_cl_mdmt.tex
scores, second script

--- 2021-04-29 2021/emnlp/mdmt.tex
@@
 
-\title{Improving Multi-Domain Neural Machine Translation with Differential Data Selection}
+\title{Automated curriculum learning for Multi-Domain NMT by using loss-driven reward}
 
@@
 This multi-source / multi-domain adaptation problem is often approached with instance selection or reweighting strategies, most of which pre-suppose an ex-ante assessment of the relevance of training instances.
-In this paper, we study the recently proposed Differential Data Selection (DDS) model and explore its ability to serve as a generic framework for these various situations.
-Our experiments study both domain adaptation and multi-domain learning and show that DDS often enables to outperform more heuristic adaptation strategies.
-We also introduce variants that boosts DDS performance of adapter-based multi-domain systems.\fyTodo{Keep this part ?}
-%   This is a well-known issue, which has given raise to a rich litterature in under the umbrellas of domain adaptation or multi-domain
-
-%   The priority of the in-domain performance in the overall evaluation affects the choice of choosing data.
-The problem falls in the data selection category in the domain adaptation topic.
-Several data selection methods pre-compute the domain-relatedness of training examples and use these scores to build mini-batches in the training while other works propose dynamic strategies to build training mini-batches based on reinforcement learning.
-% In this study, we first formulate the most general algorithm for training the Neural Machine Translation (NMT) model given a predefined priority of the domains.
-Then, we apply a recently proposed method to solve the problem.
-We also report a disadvantage of DDS and fix it by proposing a minor development for DDS.
-Our experiments with a large sample of multi-domain systems show several important benefits of DDS in multi-domain NMT.
+In this paper, we propose a novel dynamic data selection algorithm which re-evaluates the data utility and adapts the data selection policy along the training of the NMT model.
+We explore the ability of our method to serve as a generic framework for the previously mentioned situations.
 \end{abstract}
@@
 
-\fyTodo{Label or covariate shift?} \revision{Answer: Covariate shift happens when D(x) differs between train and test; label shift happens when the conditional distribution $D(y|x)$ differs between train and test}
-This problem is often approached with \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions.
+This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions.
 Finding the optimal balance of training data is however a challenging task due, for instance, to similarity between domains / languages, but also due to regularization effects of out-of-domain data; it may also be suboptimal, as some target domains or languages might be easier to train than others.
-Several recent proposals \cite{Wang17instance,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi} have explored ways to instead consider more \emph{dynamic} data selection and sampling strategies.
-We study in this contribution the proposal of \citet{Wang20balancing}, initially introduced for multilingual MT, with the goal to evaluate its potential in various single-domain and multi-domain adaptation settings and assess DDS, both in its basic version and with our improvements, as an effective replacement to standard heuristics approaches.
+Several recent proposals \cite{Wang17instance,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,alex17automated} have explored ways to instead consider \emph{dynamic} data selection and sampling strategies.
+In this paper, we propose a nouvel method of building an automated curriculum for training Multi-Domain NMT model, that is inspired by the work of \cite{alex17automated}.
+We evaluate our method in various single-domain and multi-domain adaptation settings.
+We compare our method with several contrasts including the work of \cite{Zhang19curriculum} and \cite{Wang20balancing}.
+From now on, we name our method MD-ACL
 
-Based on experimental results obtained on a diverse set of domains, our main conclusions are that (a) using DDS often yields overall performance that are as good as the standard fine-tuning strategy for domain adaptation; (b) DDS can effectively handle a variety of test target distributions without any meta-parameter search; (c) our extension of DDS is able to boost its performance when used in conjunction to the adapter model of \citet{Bapna19simple}.
-
-% one often needs to make the best out of heterogeneous sets of parallel data in training to achieve the best test performance in one or several domain(s) of interest.
-This multi-source (multi-domain) adaptation problem is often approached with instance selection or reweighting strategies, most of which pre-supposes an ad-hoc assessment of relevance for training sentences.
-In this contribution, we study the recently proposed Differential Data Selection (DDS) model  and explore its ability to serve an effective generic framework for these various situations.
-Our experiments for both domain adaptation and multi-domain learning show that DDS often enables to outperform heuristic training strategies; we also introduce a variant that boosts DDS performance of adapter-based multi-domain systems.
-
+Based on experimental results obtained on a diverse set of domains, our main conclusions are that (a) using MD-ACL often yields overall performance that are as good as the standard fine-tuning strategy for domain adaptation; (b) MD-ACL can effectively handle a variety of test target distributions without any meta-parameter search.
+%(c) our extension of DDS is able to boost its performance when used in conjunction to the adapter model of \citet{Bapna19simple}.
 \section{Learning with multiple data sources} \label{sec:mdmt}
@@
 
-Given $\vtheta_t$, $\psi_t$ is updated via the Reinforce algorithm \cite{Williams92simple}:\fyTodo{Rewrite this correctly with domain average reward}
+Given $\vtheta_t$, $\psi_t$ is updated as follows.
+%via the Reinforce algorithm \cite{Williams92simple}:\fyTodo{Rewrite this correctly with domain average reward}
 \begin{align*}
@@
 In Table ~\ref{tab:basic-dds}, we compare the results of fine-tuning a system initially trained with a natural mixture, which is proportional to the sizes of the domains, of domains with the result of DDS learning from scratch with a initial uniform distribution.
-\fyTodo{What do we see ?
-How about domain proximity ?
-- show a graph}.
-
-\fyTodo{Only if we compare on an individual domain} \revision{Note that the DDS systems are somewhat quicker to build, as we only train for 200K iterations (instead of 200K + 50K for the fine-tuning part)}.
 
@@
     \multicolumn{9}{|c|}{\sl 30 clusters} \\ \hline
+    \system{DDS-v1($\vlambda^*, \vlambda_d$)}& &56.9&51.3&33.2&43.4&&&
     \\  \hline\hline 
@@
     \\
-    \system{DDS-v1($\vlambda_0, \vlambda_d$)} &36.7&57.6&54.9&33.3&46.1&91.9&53.4&- \\
+    \system{DDS-v1($\vlambda_0, \vlambda_d$)} &36.7&58.2&55&33.3&46.6&91.9&53.6& 23.1 \\
     \system{DDS-v1($\vlambda_1, \vlambda_d$)} & \\
-    \system{DDS-v1($\vlambda_0, \vlambda_0$)} &36.96&55.55&52.88&33.06&44.52&91.25&52.37&- \\
-    \system{DDS-v1($\vlambda_1, \vlambda_0$)} & \\ \hline
+    \system{DDS-v1($\vlambda_0, \vlambda_0$)}&35.5&55.5&53.7&32.3&45.6&91.1&52.3& - \\
+    \system{DDS-v1($\vlambda_1, \vlambda_0$)}&37.4&54.4&50.3&32.0&42.7&80.2&49.5 \\ \hline
   \end{tabular}

### 2021-05-06 2021/emnlp/auto_cl_mdmt.tex
adding new version of the paper, adding scores, adding figures, adding new references

@@
 \usepackage{latexsym}
-\renewcommand{\UrlFont}{\ttfamily\small}
+%\renewcommand{\UrlFont}{\ttfamily\small}
 
@@
 \usepackage{pgfplotstable}
+
+\let\oldbibitem\bibitem
+\def\bibitem{\vfill\oldbibitem}
 
@@
 This multi-source / multi-domain adaptation problem is often approached with instance selection or reweighting strategies, most of which pre-suppose an ex-ante assessment of the relevance of training instances.
-In this paper, we propose a novel dynamic data selection algorithm which re-evaluates the data utility and adapts the data selection policy along the training of the NMT model.
+In this paper, we propose a novel dynamic data selection algorithm that re-evaluates the data utility and adapts the data selection policy along with the training of the NMT model.
 We explore the ability of our method to serve as a generic framework for the previously mentioned situations.
@@
 In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key component in training is then to mitigate the detrimental effects of a possible mismatch between these distributions.
-Single-source and multi-source domain adaptation (DA) is a well-studied instance of this setting (see \citep{Chu2017comparison} for a review) , and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
-A related situation is multilingual machine translation \cite{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
+Single-source and multi-source domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
 
-This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions.
-Finding the optimal balance of training data is however a challenging task due, for instance, to similarity between domains / languages, but also due to regularization effects of out-of-domain data; it may also be suboptimal, as some target domains or languages might be easier to train than others.
-Several recent proposals \cite{Wang17instance,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,alex17automated} have explored ways to instead consider \emph{dynamic} data selection and sampling strategies.
-In this paper, we propose a nouvel method of building an automated curriculum for training Multi-Domain NMT model, that is inspired by the work of \cite{alex17automated}.
+This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions \citep{Moore10selection,Axelrod11domain}.
+Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data; it may also be suboptimal, as some target domains or languages might be easier to train than others.
+For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not promise optimization of a particular domain.
+
+Several recent proposals \citep{platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Graves17automated,Wang20balancing,Wang20learning-multi,Vanderwees17dynamic} have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies.
+Furthermore \citet{Zhang19curriculum,Vanderwees17dynamic} construct a fixed curriculum while \citet{platanios19competence,Kumar19reinforcement,Graves17automated,Wang20balancing,Wang20learning-multi} build curricula automatically adapted to the training.
+
+In this paper, we propose a novel method of building an automated curriculum for training the Multi-Domain NMT model, that is inspired by the work of \citet{Graves17automated}.
 We evaluate our method in various single-domain and multi-domain adaptation settings.
-We compare our method with several contrasts including the work of \cite{Zhang19curriculum} and \cite{Wang20balancing}.
-From now on, we name our method MD-ACL
+We compare our method with several contrasts including the work of \citet{Zhang19curriculum}, that we name CL, and \citet{Wang20balancing}, that we name DDS.
+From now on, we name our method MDAC which is brief for Multi-Domain Automated Curriculum.
+Based on experimental results obtained on a diverse set of domains, our main conclusions are that (a) using MDAC often yields overall performance that are as good as the standard fine-tuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test target distributions without any meta-parameter search.
 
-Based on experimental results obtained on a diverse set of domains, our main conclusions are that (a) using MD-ACL often yields overall performance that are as good as the standard fine-tuning strategy for domain adaptation; (b) MD-ACL can effectively handle a variety of test target distributions without any meta-parameter search.
-%(c) our extension of DDS is able to boost its performance when used in conjunction to the adapter model of \citet{Bapna19simple}.
 \section{Learning with multiple data sources} \label{sec:mdmt}
@@
 The main challenge is then to make the best of this heterogeneous data, with the aim to achieve the optimal performance for the intended test conditions.
-These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unsupervised domain adaptation); in multi-domain adaptation finally the test distribution is itself a mixture of domains, some of which may also be observed in training.
+These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unsupervised domain adaptation); in multi-domain adaptation finally, the test distribution is itself a mixture of domains, some of which may also be observed in training.
 Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of fine-tuning.
@@
 These situations have been amply documented from a theoretical perspective (eg.\ \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}).
-A general recommandation in the DA setting  is to adjust the sampling distribution used to optimize the system so as to compensate for the mismatch between $\mathcal{D}^s(x)$ and $\mathcal{D}^t(x)$.
+A general recommendation in the DA setting is to adjust the sampling distribution used to optimize the system so as to compensate for the mismatch between $\mathcal{D}^s(x)$ and $\mathcal{D}^t(x)$.
 This can be approximated by reweighting instances, or more conveniently domains, which are selected during training with a probability $\lambda^{l}(d)$, with $\lambda^{l}(d) \neq \lambda^{s}(d)$.
 
-A more practical standard approach to supervised DA is \emph{fine-tuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during the course of learning.
+A more practical standard approach to supervised DA is \emph{fine-tuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
 With our notations, this approach amounts to first learning an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$) with $\indic{A}$ the indicator function for predicate $A$.
 Note that this strategy is potentially suboptimal, as some out-of-domain samples may contribute to the final performance due to eg.\ domain overlap.
-Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains might be easier to learn than others.\fyTodo{How to measure this?} \revision{Answer: The graphic of \system{Mixed-0} shows that with same amount of data(batches) the convergences of the domains vary significantly}
-
-Differentiable data selection (DDS), proposed by \citet{Wang20optimizing} and later applied to multilingual machine translation by \citet{Wang20balancing}, introduces an additional degree of freedom and allows to iteratively learn an optimal value for $\lambda^{l}$, as we now explain.
+Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains might be easier to learn than others.\fyTodo{How to measure this?} \revision{Answer: The graphic of \system{Mixed-0} shows that with the same amount of data(batches) the convergences of the domains vary significantly}
 
@@
 
-\section{Differential Data Selection} \label{sec:dds}
+\section{Multi-Domain Automated Curriculum } \label{sec:mdac}
 \subsection{Basic principles}
 Assuming training data from $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as  $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
-We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t _d}$ to denote the empirical train and test distributions for domain $d$, and \revision{$\widehat{\mathcal{D}^{u}}(x;\lambda^{u}) = \sum_{d} \lambda^{u}(d) \widehat{\mathcal{D}^{u}_d}(x)$ for $u\in\{l,t\}$} \fyTodo{$\widehat{\mathcal{D}}(x;\lambda) = \sum_{d} \lambda(d) \widehat{\mathcal{D}_d}(x)$}.
+We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t _d}$ to denote the empirical train and test distributions for domain $d$, and $\widehat{\mathcal{D}^{u}}(x;\lambda^{u}) = \sum_{d} \lambda^{u}(d) \widehat{\mathcal{D}^{u}_d}(x)$ for $u\in\{l,t\}$.
 In our setting,  $\vlambda^t$, and hence $\widehat{\mathcal{D}^t}(x;\vlambda^t)$ are fixed and predefined, approximated with an equivalent number of development corpora.
-Training optimizes the parameters $\vtheta$ of a neural translation model.
 
-Sampling training domains using a distribution $\lambda^{l}$ results in a parameter estimate $\theta(\lambda^{l})$, delivering performance that can be estimated on the development set using $\lambda^{t}$.
-\citet{Wang20optimizing,Wang20balancing} then formulate the problem of finding an optimal sampling strategy $\lambda^{l}_{*}$ maximizing performance on the development data as a bi-level optimization problem that can be approached by coordinate gradient descent alternating between $\lambda^{l}$ and $\vtheta$.
-\citet{Wang20balancing} limit the search space of $\vlambda^{l}$ to a parameterized differentiable function $\vlambda^{l}(\vpsi)$.
-Given $\vpsi_t$ at time step $t$, $\vtheta_t$ is updated via usual gradient descent as follows.
+Our method MDAC constructs an automatically adapting distribution $\lambda^{l}$ that optimizes the data selection policy along with the training of the NMT model.
+We parameterize $\lambda^{l}$ by a differentiable function $\lambda^l(\psi)$.
+We divide the training into many short sessions; in each session $t$, the model is trained with a fixed data distribution $\lambda^{l}(\psi_t)$.
+After one learning session, we update the data distribution via REINFORCE algorithm \citep{Williams92simple}.
+
 \begin{align*}
-\vtheta_{t+1} &= \vtheta_t - \mathbf{lr}_{nmt} * \frac{1}{N} \displaystyle{\mathop{\sum}_{i=1}^N} \frac{\partial l(\vtheta_t, x_i,y_i)}{\partial \vtheta} \\
-x_i,y_i &\sim \widehat{\mathcal{D}^l}(x;\vlambda^l(\psi_t)) = \displaystyle{\mathop{\sum}_{d=1}^K} \lambda^l(d;\psi_t) * \widehat{D^l_d}(x) \\
+\psi_{t+1} &= \psi_t + \mathbf{lr}_{data} * \displaystyle{\mathop{\sum}_{d=1}^K} R(d) * \frac{\partial \lambda^l(d;\psi_t)}{\partial \psi} \\
 \end{align*}
-Where $l(\theta_t,x,y)$ is the loss on an example.
-
-Given $\vtheta_t$, $\psi_t$ is updated as follows.
-%via the Reinforce algorithm \cite{Williams92simple}:\fyTodo{Rewrite this correctly with domain average reward}
+\begingroup
+\allowdisplaybreaks
+With the following reward computation:
 \begin{align*}
-\psi_{t+1} &= \psi_t + \mathbf{lr}_{data} * \displaystyle{\mathop{\sum}_{d=1}^K} R(d) * \frac{\partial log(\lambda^l(d;\psi_t))}{\partial \psi} \\
-  R(d) &= \langle \frac{\partial J^l_d(\theta_t)}{\partial \theta}, \frac{\partial J^t(\theta', \lambda^t)}{\partial \theta} \rangle \\
+  R(d) = J^t(\theta_{t+k},\lambda^t) - J^t(\theta_t,\lambda^t)
 \end{align*}
 Where 
-\begingroup
-\allowdisplaybreaks
-\begin{align*}
-\theta' &= \theta_t - \mathbf{lr}_{nmt} * \displaystyle{\mathop{\sum}_{i=1,x^i_d,y^i_d \in \widehat{D^l_d}}^N} \frac{\partial l(\theta_t,x^i_d,y^i_d)}{\partial \theta} \\
-  J^l_d(\theta_t) &= \displaystyle{\mathop{\sum}_{i=1,x^i_d,y^i_d \in \widehat{D^l_d}}^N} l(\theta_t,x^i_d,y^i_d) \\
-  J^t(\theta',\lambda_t) &= \displaystyle{\mathop{\sum}_{d}}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta',x^t_d,y^t_d)\\
-\end{align*}
+\begin{equation}
+\begin{array}{rcl}
+\theta_{t+i} &=& \theta_{t+i-1} - \mathbf{lr}_{nmt} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta} \\ \nonumber
+x^i_j, y^i_j &\sim& \widehat{D^l_d}(x) \\
+J^t(\theta,\lambda_t) &=& \displaystyle{\mathop{\sum}_{d=1}^K}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta,x^t_d,y^t_d)
+\end{array}
+\end{equation}
 \endgroup
@@
 $\mathbf{lr}_{nmt}$ is the learning rate of the parameter of the NMT system.
-\fyTodo{Explain lr, may be more}\fyTodo{More refs to the algorithm }
-% To apply this method to multi-domain training (\system{Multi-DDS}), we limit the search family of training distribution $D_{trn}$ to the family of mixture of K in-domain training distribution such as $\sum_d \lambda^{s}_{d} \mathcal{D}_d(x,y)$.
-We also parameterize $\lambda^{s}_{d}$ using Softmax function so that $\lambda^{s}_{d} = \text{Softmax}(\psi)_d$ where $\psi \in \mathbb{R}^K$.
-\subsection{DDS for (multi) domain adaptation}
+$J^t(\theta,\lambda_t)$ is the weighted loss computed over $K$ dev-sets corresponding to $K$ target domains.
 
-This setting is pretty general and can in principle accomodate the variety of situations mentioned  above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
-In our experiments we would like to better assess the actual potential of DDS is these settings and seek to experimentally answer the following questions:
+To compute the reward of using data from domain $d$ to train the model, $R(d)$, we simulate k training steps from the current checkpoint by training the model with k batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
+The reward of our method is inspired by the target prediction gain in \cite{Graves17automated}.
+However, \cite{Graves17automated} used accumulated gains from the past as reward while we use the "future" gain as reward.
+
+The work of \cite{Wang20balancing} based on the Bi-level Optimization framework, which aims to find an optimal fixed distribution $\lambda^{l}$ that will result in the best NMT model regarding a given target dev set at the end of the training, derives also a similar form of update for $\psi$.
+However, their reward is the cosine similarity between the gradient computed with the training data and the gradient computed with the dev set.
+
+\subsection{MDAC for (multi) domain adaptation}
+This setting is pretty general and can in principle accommodate the variety of situations mentioned  above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
+In our experiments we would like to better assess the actual potential of MDAC is these settings and seek to experimentally answer the following questions:
 \begin{itemize}
-\item is DDS a viable alternative to conventional fine-tuning? In particular does it enable to better take advantage of relevant data from other domains?
-\item is DDS also a viable option in multidomain adaptation scenarios?
-\item does DDS also enable to perform unsupervised (multi-) domain adaptation?\fyTodo{TBContinued}
+\item is MDAC a viable alternative to conventional fine-tuning? In particular does it enable to better take advantage of relevant data from other domains?
+\item is MDAC also a viable option in multi-domain adaptation scenarios?
+\item does MDAC also enable to perform unsupervised (multi-) domain adaptation? \fyDone{TBContinued}
 \end{itemize}
@@
 \section{Experimental settings} \label{sec:exp}
-
 \subsection{Data and metrics \label{ssec:corpora}}
-We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.\revision{We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}}; the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \cite{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
+We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.
+We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \cite{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
@@
 We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
-
-%for contrast experiments, we also use supplementary test sets from three other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
-This enables us to evaluate the loss in performance when the test set is from a domain not seen in training.
-% The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
 
@@
 
-\fyTodo{Keep this ?}
-We measure the distance between domains using the $\mathcal{H}$-Divergence \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains.
-%Formally, given test sets of size $m$ for domains $A$ and $B$, and $h(x)$ a trained domain predictor, $\mathcal{H}(A,B)$ is computed %as:
-%$$
-%\mathcal{H}(A,B) = 2(1 - [\frac{1}{m} \sum_{x:h(x) = B} \mathbb{I}( x \in A) + \frac{1}{m} \sum_{x: h(x) = A} \mathbb{I}(x \in B)]),
-% $$
-% where $\mathbb{I}$ is the indicator function.
-Our discriminator is a SVM independently trained for each pair of domains, with sentence representations derived via mean pooling from the source side representation of the generic Transformer model.
-We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well separated from all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
-
-\begin{table}\centering
-  \begin{tabular}{|l*{5}{|r}|} 
-  \cline{2-6}
-  \multicolumn{1}{c|}{} & \domain{law} & \domain{bank} & \domain{talk} & \domain{IT} & \domain{rel} \\ \hline
-    \domain{med} &1.93 &1.97 &1.9 &1.93 &1.97 \\
-    \domain{law}   && 1.94 & 1.97 &1.93 & 1.99 \\
-    \domain{bank} &&&1.98 &1.94 &1.99 \\
-    \domain{talk}   &&&&1.92 &1.93 \\
-     \domain{IT}     &&&&& 1.99 \\ \hline
-  \end{tabular}
-  \caption{The $\mathcal{H}$-divergence between domains}
-  \label{tab:domaindist}
-\end{table}
-
 \subsection{Baseline architectures \label{ssec:baseline}}
@@
 q_d = \frac{\mid N^{s}_d \mid}{\displaystyle{N^{s}}} % \mathop{\sum}_{i=1}^K\mid D_i \mid}}
-\end{align} 
+\end{align} \setlength{\abovedisplayskip}{2pt}
+\setlength{\belowdisplayskip}{2pt} 
 with $\alpha \in [0,0.25,0.5,0.75,1.0]$.
@@
 \system{Mixed-$0$} uses a uniform domain distribution, \system{Mixed-$1.0$} simple uses the observed domain distribution.
-%on a concatenation of all corpora (\texttt{Mixed}).
-We develop two versions\footnote{In fact three: to enable a fair comparison with WDCMT, a RNN-based variant is also trained and evaluated.
-\revision{This system appears as \system{Mixed-Nat-RNN} in Table~\ref{tab:performance}}.} of this system, one where the domain unbalance reflects the distribution of our training data \revision{given in Table~\ref{tab:Corpora}} (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
-The former is the best option when the train mixture $\mathcal{D}^s$ is also expected in testing; the latter should be used when the test distribution is uniform across domains.
-Accordingly, we report two aggregate scores: a weighted average reflecting the training distribution, and an unweighted average, meaning that test domains are equally important.
 \item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing.
 The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems each adapted for one domain, with no parameter sharing across domains.
-
 \item two multi-domain versions of the approach of \newcite{Bapna19simple}, denoted \system{FT-Res} and \system{MDL-Res}, where a domain-specific adaptation module is added to all the Transformer layers; within each layer, residual connections enable to short-cut this adapter.
@@
 It fine-tunes the adapter modules of a \system{Mixed-$1.0$} system independently for each domain, keeping all the other parameters frozen.
-The latter uses the same architecture, but a different training procedure and learns all parameters jointly from scratch with the prefixed mixtures of training data \ref{mixture:trn}.\fyTodo{Keep this ?}
+The latter uses the same architecture, but a different training procedure and learns all parameters jointly from scratch with the prefixed mixtures of training data \ref{mixture:trn}.\fyDone{Keep this ?}
 
-\item Our comparison of multi-domain systems includes baselines with fixed data mixtures corresponding to $\vlambda_0$ and $\vlambda_1$, and our own reimplementations of recent proposals from the literature:\footnote{\revision{Further implementation details are in supplementary material.}}\fyTodo{Choose a reasonnable MD baseline }
-
+\item Our comparison of multi-domain systems includes baselines with fixed data mixtures corresponding to $\vlambda_0$ and $\vlambda_1$, and our reimplementations of recent proposals from the literature: Curriculum Learning of \cite{Zhang19curriculum} and Differential Data Selection of \cite{Wang20balancing}.
 \end{itemize}
@@
 
-\subsection{DDS-based systems } \label{ssec:dds-sys}
+\subsection{MDAC systems} \label{ssec:dds-sys}
+The behavior of MDAC only depends on (a) the initial domain distribution at the start of training $\vlambda^{l}_{t=0}$, and (b) the targeted (dev/test) distribution $\vlambda^{t}$.
+We will thus report these systems as \system{MDAC($\vlambda^{l}_{t=0}$, $\vlambda^{t}$)}.
+We also compare our method with the method DDS under the same conditions.
 
-The architecture of DDS based-systems is identical to that of our baselines, and only differ in the training regime.
-Their behavior only depends on (a) the initial domain distribution at the start of training $\vlambda^{l}_{t=0}$, and (b) the targeted (dev/test) distribution $\vlambda^{t}$.
-We will thus report these systems as \system{DDS($\vlambda^{l}_{t=0}$, $\vlambda^{t}$)}.
-The domain distribution attained at the end of training will be denoted  $\vlambda^{l}_{*}$.
+In our work, we simply parameterize the distribution $\lambda^l$ as follows.
+\begin{equation}
+\lambda
+^l(d;\vpsi) = \frac{\vpsi[d]^\beta}{\sum_i \vpsi[i]^\beta} \nonumber
+\end{equation}
+We choose $\beta=2$ for our experiments.
+
+We provide also in this section our choices of the hyper-parameters of our algorithm including the number of simulation training steps $k=10$, and the update rate $\mathbf{lr}_{data}=0.1$.
+We split the training into many short sessions that last 2000 training steps each.
+\subsection{CL and DDS's re-implementation}
+We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to original code provided by \citet{Wang20balancing} \footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}.
+
+We re-implement the work of \citet{Zhang19curriculum} by following exactly the description of the algorithm.
+For each experiment of adapting the NMT system to one domain, we combine the training data of other domains into one corpus, then compute the cross-entropy difference score of each sentence of this corpus (we only compute the score for the source side).
+We sort and split the corpus into 9 shards and execute curriculum learning with 10 shards in which the first shard is the in-domain corpus.
+\subsection{Experimental tasks}
+We evaluate the efficacy of our methods compared to several aforementioned baselines via 3 tasks including domain adaptation; multi-domain adaptation and unsupervised domain adaptation.
+
+In the domain adaptation task, given the data of 6 domains including \domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}, we aim to build expert NMT models for each domain respectively.
+
+In the Multi-domain adaptation task, given the data of 6 domains, we aim to build an NMT model which gives the best performance regarding a predefined test set compromising equivalently 6 domains.
+
+In the unsupervised domain adaptation task, we design 2 experiments.
+In the first experiment, given the data of 6 domains, we aim to build an NMT model which is robust to an unseen domain, which is News in this case.
+In the second experiment, we cluster the training data of 6 domains into 30 clusters using the KNN algorithm in the same way as in the work of \citet{Tars18multidomain}, then adapt these clusters to one of 6 domains given the corresponding in-domain dev set.
+We compare MDAC to the contrast DDS over 6 domains.
 
@@
 
-\subsection{DDS for single-domain and multi-domain adapation}
+\subsection{Domain Adaptation}
+In this setting, we aim to build an expert NMT model in only one domain. This setting corresponds to Dirac distribution $\lambda^t$, in which the probability of domains is 0 except for one domain.
 
-\subsubsection{Fine-tuning experiments}
+We initialize MDAC and DDS by 2 different distribution $\lambda_0$ and $\lambda_1$.
+According to table \ref{tab:da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen.
+The same conclusion is also true for DDS.
 
-\fyTodo{Fix the table}
-In Table ~\ref{tab:basic-dds}, we compare the results of fine-tuning a system initially trained with a natural mixture, which is proportional to the sizes of the domains, of domains with the result of DDS learning from scratch with a initial uniform distribution.
+According to the table \ref{tab:da}, MDAC is outperformed by finetuning method in large domains including medical and law.
+However, MDAC improves finetuning method by approximately 0.5 BLEU in \domain{bank} and 1.0 BLEU in \domain{rel}.
+This indicates that for small domains, out-of-domain data can improve the generalization of the NMT model and MDAC is capable to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in finetuning.
 
-\begin{table*}
+DDS also demonstrates promising results in this task but is always outperformed by MDAC overall. The same conclusion is also applied to the CL method.
+
+\begin{table*}[htbp]
   \centering \small
@@
     \hline
-    \system{DDS($\vlambda_0, \vlambda_d$)} \hfill [BLEU] &36.7&56.2&54.3&33.5&46.8&91.8&53.2\\
-    \% domain $d$ (final / total) & 59.7 / 49.5 & 6.5 / 18.3 & 14.9 / 24.5 & 6.4 / 12.8 & 67.2/47.4 & 67.2 / 45.3&\\ \hline
-    \system{DDS($\vlambda_1, \vlambda_d$)} \hfill [BLEU]&36.4&56.3& 45.4& 32.6&44.5&43.1&43.1\\
-    \% domain $d$& 89 / 86.6 & 50/18.4 & 1.4 / 1.6 & 3.0 / 2.5 & 7.5 / 9.2 & 1.0 / 1.0 \\ \hline
-    \system{CL($d$)} \hfill [BLEU]&  \\ 
-    \% domain $d$& \\ \hline\hline
+    \system{DDS($\vlambda_0, \vlambda_d$)} &36.7&56.2&54.3&33.5&46.8&91.8&53.2\\
+    \system{DDS($\vlambda_1, \vlambda_d$)} &36.4&56.3& 45.4& 32.6&44.5&43.1&43.1\\ \hline
+    \system{CL($d$)} &37.2&55.4&53.0&31.3&45.3&90.8&52.2\\ \hline
+    \system{MDAC($\vlambda_0, \vlambda_d$)} &36.7&58.2&55&33.3&46.6&91.9&53.6\\
+    \system{MDAC($\vlambda_1, \vlambda_d$)} &37.2&55.1&51.8&32.8&44.9&85.4&51.2\\ \hline
+  \end{tabular}
+  \caption{We report scores of each method with 6 target domains, and compare with full fine-tuning: each column corresponds to a distinct system.}
+  \label{tab:da}
+\end{table*}
+
+\subsection{Multi-domain adaptation}
+
+Multi-domain adaptation task requires best averaged performance of the NMT model in 6 domains.
+This setting corresponds to the uniform test distribution $\lambda^t = \lambda_0$.
+In this situation, the method CL of \citet{Zhang19curriculum} can not be applied.
+We therefore only compare the performance of MDAC, DDS and several fixed training data distribution $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$.
+The definition of those notations can be found in the equation \ref{mixture:trn}.
+
+We initialize MDAC and DDS by 2 different distribution $\lambda_0$ and $\lambda_1$.
+According to table \ref{tab:multi-da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen.
+The same conclusion is also true for DDS.
+
+MDAC outperforms almost fixed training distribution including $\big[ \lambda_0, \lambda_{0.75}, \lambda_{1.0}\big]$ with significant margin while performs slightly better than $\big[ \lambda_{0.25}, \lambda_{0.5} \big]$.
+This observation indicates that MDAC allows us to skip the heuristic choice for the training mixture of different domains or tasks.
+
+The second observation is that DDS is again outperformed by MDAC with a significant margin of 0.5 BLEU on average. However, MDAC is surpassed by far with 1.8 BLEU in the domain \domain{med} by DDS.
+
+We show in the below graphic \ref{fig:sampling}, the evolution of the training mixture through training sessions.
+According to the figure \ref{fig:DDS} of DDS, the proportion of 6 domains converges rapidly to bi-domain mode in which only \domain{med} and \domain{rel} have significant probability.
+In contrast, the proportion of 6 domains computed by MDAC evolves smoothly; small domains including \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} take larger proportion in the beginning then decrease their proportions while larger domains as \domain{med} and \domain{law} increase their proportions in the end where the NMT model might have been already close to the optimal performance in the smaller domains.
+
+\begin{table*}[htbp]
+  \centering \small
+  \begin{tabular}{|l|*8{r|}} \hline
+    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline \hline
     % Multidomain
@@
     \system{Mixed-1}     & 37.3  & 54.6 & 50.1 & 33.5 & 43.2 & 77.5& 49.4 \\
-    \system{MDL-Res} &37.1&57.4&53.6&33.2&46&90.7&53\\\hline
+    \hline \hline
+    \system{DDS($\vlambda_0, \vlambda_0$)} &37.3&55.1&51&33.5&43.4&90.8&51.8\\ 
+    \system{MDAC($\vlambda_0, \vlambda_0$)}&35.5&55.5&53.7&32.3&45.6&91.1&52.3\\ 
+    \hline \hline
     \system{DDS($\vlambda_1, \vlambda_0$)} &37.9&50.2&47.2&31.2&40.8&58.6&44.3\\
-    $\operatorname{KL}(\vlambda_*^{l}, \vlambda_0$) &37&55.6&52.9&33.1&44.5&91.3&52.4 \\
-%    $ \mathcal{H}(\vlambda_*^{l})$ &  \\
-    \system{DDS($\vlambda_0, \vlambda_0$)} &37.3&55.1&51&33.5&43.4&90.8&51.8\\ \hline
+    \system{MDAC($\vlambda_1, \vlambda_0$)}&37.4&54.4&50.3&32.0&42.7&80.2&49.5\\
+    \hline
   \end{tabular}
-  \caption{Baseline results for DDS in fine-tuning and multi-domain adaptation.
-In the top part, we report scores with DDS and each target domain, and compare with full-fine-tuning: each column corresponds to a different system.
-We also report the proportion of in-domain data in the final mixture.
-In the bottom part, we run multi-domain experiments: for a given line, all the columns \emph{correspond to the same system}.
-We also report the KL divergence between the optimal and the target mixtures.}
-  \label{tab:basic-dds}
+  \caption{We report the performance of our method MDAC and the work of \citet{Wang20balancing} DDS in multi-domain adaptation task.
+For a given line, all the columns \emph{correspond to the same system}}
+  \label{tab:multi-da}
 \end{table*}
 
-A first observation (top part of Table~\ref{tab:basic-dds}) is that DDS is only better than FT for \domain{rel}, and worse for \domain{med} and \domain{law}.
-For the three other domains, we do not see any marked difference, which we view as a positive result.
-There seem to be a connection with the domain size, as \domain{rel} is small and benefits from data from other domains, while the opposite effect is observed for the larger \domain{med} and \domain{law}.
-Averaged over the 6 domains, we see a small (0.5 BLEU) difference between the two approaches for the domain of interest.
-Note however that DDS also mitigates the catastrophic forgetting effect observed with full fine-tuning (see supplemental material).\fyTodo{Supplementary}
-
-\subsubsection{Multi-domain experiments}
-
-If we now look at the multi-domain experiments (bottom part of Table~\ref{tab:basicdds}), we see that DDS actually does better than the two baseline strategies that use an untuned training distribution, but is not as good as using a fixed distribution with a well-chosen $\alpha$.
-When comparing with an improved (multi-)domain adaptation techniques, based on residual adapters \cite{Bapna19simple,Pham20astudy}, we observe an even larger gap in performance.
-
-In addition we observe ...\fyTodo{To be continued}
-
-\fyTodo{Fixed regime with target mixture weigths ?}
-% \begin{table*}[htb]
-%   \centering% \small
-%   \begin{adjustbox}{width=1.0\textwidth,center}
-%   \begin{tabular}{|p{3.0cm}|*{13}{r|}} \hline
-%     \multirow{2}{*}{Name} & \multirow{2}{*}{$sim\_iter$} & \multirow{2}{*}{parameterization} & \multirow{2}{*}{entropy constraint} & \multicolumn{7}{|c|}{BLEU} \\ \cline{5-11}	
-%    & & & & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{it}} & \multicolumn{1}{c|}{\domain{ rel }} & \multicolumn{1}{c|}{\domain{ talk}} & \multicolumn{1}{c|}{mean} \\
-%   \hline
-%   \system{$DDS\_FT (med)$} & 1 & softmax & no & 36.7&51.14&52&44.32&90.41&33.22&51.3\\
-%   \system{$DDS\_FT\_law$} & 1 & softmax & no &36.05&56.18&53.57&44.05&91.24&33.09&52.36\\
-%   \system{$DDS\_FT\_bank$} & 1 & softmax & no &36.15&54.17&54.29&41.33&89.95&31.54&51.24 \\
-%   \system{$DDS\_FT\_IT$} & 1 & softmax & no &34.39&48.98&52.82&46.8&85.3&31.37&49.94 \\
-%   \system{$DDS\_FT\_rel$} & 1 & softmax & no & 34.54&52.29&51.46&44.8&91.77&31.84&51.12\\
-%   \system{$DDS\_FT\_talk$} & 1 & softmax & no & 35.53&50.55&52.64&44.86&85.8&33.47&50.48\\
-%   \system{$FT\_full$} & NA & NA & NA & 37.74&59.21	&54.49&46.81&90.77&33.98&53.83\\
-%   \hline
-%   $\Delta$ & \\
-%   \end{tabular}
-%   \end{adjustbox}
-%   \caption{DDS for fine-tuning and multi-domain adaptaion.
-In the top part, we report scores with DDS and each target domain, and compare with full-fine-tuning: each column corresponds to a different system.
-In the bottom part, we run multi-domain experiments: for a given line, all the columns \alert{correspond to the same system}.}
-%   \label{tab:finetuning-old}
-% \end{table*}
-\fyDone{This table does not work- One line would be enough, where we compare autoFT with true FT}
-\fyTodo{Curves for bank / medical ou law}
-\fyTodo{Fine tune pour 2 domains ?}
-
-It is notably striking to see that the distribution at the end of training ($\vlambda^l_*$) is concentrated on few domains and is quite remote from the target test distribution.
-Looking at the evolution of $\vlambda^l$ during training, we often observe a rich-get-richer effect, where a small number of domains manage to attract most of the samples, while others are almost completely ignored after a given number of iterations.
-This is illustrated on Figure~\ref{}, where we see....\fyTodo{Provide illustrations and numbers}
-
-\subsection{Analysis and improvements}
-
-\subsection{Unsupervised DDS}
-To better analyze the behavior of DDS, we study a scenario where $\vlambda^s$ and $\vlambda^t$ have different dimensions designed as follows.
-All training data is automatically clustered into $K=30$ of varying sizes using the k-means algorithm based on generic sentence representations obtained via mean pooling.
-The resulting classes contain from a tenths of thousands to several hundreds of thousand training samples.
-Most classes contain a dominant domain, which accounts for more of 90\% of the samples.\footnote{See detail in supplementary material} Using the same DA setting as before, we study (a) whether DDS is able to identify the classes that correspond to the domain of interest (b) whether having additional degrees of freedom can yield an improved training performance (c) how DDS scales when a large number of  domains is considered.\fyTodo{TBC - with what experiments?}
-
-\subsubsection{Winner-takes-all}
-
-Following \citep{Wang20balancing}, our baseline uses the following parameterization of the sampling component $\vlambda^l(\vpsi) = \operatorname{softmax}(\vpsi)$ where $\vpsi \in \mathbb{R}^{n_d}$.
-This parameterization induces the following gradients:\fyTodo{Change name of obj.
-function}
-\begin{align*}
-\Delta \psi[d] & = \eta \times \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \frac{\partial \lambda^l(j; \vpsi)}{\partial \vpsi[d]} \\
-%	& = R_d \lambda^l(d; \psi)(1- \lambda^l(d; \psi)) - \displaystyle{\mathop{\sum}_{j\neq d}^{n_d}} R_j \lambda^l(d; \psi) \lambda^l(j; \psi)\\
-	& = \eta \times \lambda^l(d; \psi) \big(R_d-\displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \lambda^l(j;\psi) \big)\\
-\end{align*}
-where $\eta$ controls the learning rate of $\vpsi$.
-
-The fact that the change of $\psi(d)$ is proportional to $\lambda^l(d; \vpsi)$ explains that the optimization often leads to a degenerated solution in which only a small number of domains are considered.
-%This behavior is even more visible when using 30 automatic domains in training:\fyTodo{What results would be useful ?}.
-
-In order to mitigate this behavior, two variants of the baseline are investigated:
-\begin{itemize}
-\item to ensure that $\vpsi$ remains sufficient diverse or close to the target distribution $\vlambda^t$. This is obtained by adding a constraint when updating $\vpsi$. 
-\item to replace the softmax function for which the rich-get-richer effect are attenuated. We consider functions of the form $\lambda
-^l(d;\vpsi) = \frac{\vpsi[d]^\beta}{\sum_i \vpsi[i]^\beta}$. The gradient of $\psi$ will be as follows.
-\begin{align*}
-\Delta\psi[d] = \eta \times \beta \frac{\lambda^l(d;\psi)}{\psi[d]} \big( R_d - \displaystyle{\mathop{\sum}_{j=1}^{n_d}} R_j \lambda^l(j;\psi) \big)
-\end{align*}
-where $\eta$ controls the learning rate of $\vpsi$.
-\end{itemize}
-
-\begin{table*}
-  \centering % \small
-  \begin{adjustbox}{width=1.0\textwidth,center}
-  \begin{tabular}{|p{2.0cm}|*{13}{r|}} \hline
-    \multirow{2}{*}{Name} & \multirow{2}{*}{$sim\_iter$} & \multirow{2}{*}{parameterization} & \multirow{2}{*}{entropy constraint} & \multicolumn{6}{|c|}{BLEU} & \multirow{2}{*}{average} \\ \cline{5-10}
-   & & & & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} &  \\
-    \hline
-    \system{$config\_603$} & 1 & softmax & no & 37.26& 55.07& 50.96& 33.49& 43.41& 90.76& 51.83\\
-    \system{$config\_616$} & 1 & softmax & yes & 36.96& 55.55& 52.88& 33.06& 44.52& 91.25& 52.37 \\
-    \system{$config\_675$} & 10 & softmax & yes & 36.24& 55.67& 52.63& 32.74& 44.47& 90.45& 52.03\\
-    \system{$config\_673$} & 10 & linear & yes & 36.02& 56.18& 52.96& 32.06& 45.2 & 90.94 & 52.23 \\
-  \hline
-  \end{tabular}
-  \end{adjustbox}
-  \caption{Controlling / regularizing the training distribution}
-  \label{tab:performance}
-\end{table*}
-
-\begin{figure*}[htb]
+\begin{figure*}[htbp]
 \begin{subfigure}{.5\textwidth}
@@
   % include first image
-  \includegraphics[width=.8\linewidth]{config603.png}  
-  \caption{config 603}
-  \label{fig:603}
+  \includegraphics[width=.8\linewidth]{DDS.png}  
+  \caption{DDS}
+  \label{fig:DDS}
 \end{subfigure}
@@
   % include second image
-  \includegraphics[width=.8\linewidth]{config616.png}  
-  \caption{Config 616}
-  \label{fig:616}
-\end{subfigure}
-\newline
-\begin{subfigure}{.5\textwidth}
-  \centering
-  % include third image
-  \includegraphics[width=.8\linewidth]{config675.png}  
-  \caption{Config 675}
-  \label{fig:675}
-\end{subfigure}
-\begin{subfigure}{.5\textwidth}
-  \centering
-  % include fourth image
-  \includegraphics[width=.8\linewidth]{config673.png}  
-  \caption{Config 673}
-  \label{fig:673}
+  \includegraphics[width=.8\linewidth]{MDAC.png}  
+  \caption{MDAC}
+  \label{fig:MDAC}
 \end{subfigure}
@@
 
-\subsubsection{Improving the reward computation}
-
-Figures such as Figure~\ref{} show that the training signal given by the reward~\ref{eq:reward} can be very noisy and misleading.
-For instance, we observe that even when targeting the \domain{medical} domain, the \domain{religion} data keeps being rewarded, and constitues a significant portion of the final mixture.\fyTodo{Which figure, which equation ?}\fyTodo{Do we have number that show variance and strong divergence wrt initial training point ?}
-
-Again, we have explored several options to improve this aspect of the algorithm.
-The first was to compute better estimates of the gradients in~\eqref{} by using larger batches, or performing multiple updates when computing $\vtheta_t$.
-This did not prove useful, and this option was quickly discontinued.
-The second, more promising alley, was to base the reward computation not on the gradient of the loss, but more directly on the development loss, replacing line~\ref{alg:number}\fyTodo{which line ?} with the following reward computation:
-\begin{align*}
-  R_d = J^t(\theta_{t+k},\lambda^t) - J^t(\theta_t,\lambda^t)
-\end{align*}
-Where 
-\begin{equation}
-\begin{array}{rcl}
-\theta_{t+i} &=& \theta_{t+i-1} - \mathbf{lr}_{nmt} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta} \\
-x^i_j, y^i_j &\sim& \widehat{D^l_d}(x)
-\end{array}
-\end{equation}
-
-This idea is reminiscent of the work of \citet{alex17automated} and  \citet{Kumar19reinforcement}, who also design their curriculum trainng strategy based on the development loss.
-
-\subsection{Improving dynamic data selection}
-Experiments with these revised versions are for various settings in Table~\ref{tab:improved-dds}.
-Results for DA directly compare with Table~{tab:basic-dds}; results for unsupervised DA used our 6 domains for training, but test on the \domain{News}, unseen in traning.
-\begin{table*}
+\subsection{Unsupervised domain adaptation}
+We
+\begin{table*}[htbp]
   \centering \small
@@
     \multicolumn{9}{|c|}{\sl Unsupervised domain adaptation} \\ \hline
-    \system{DDS-v1($\vlambda_0, \vlambda_{news}$)} & \\  \hline\hline 
+    \system{Mixed-0}      & - & - & - & - & - & - & - & 22.4 \\
+    \system{Mixed-0.25} & - & - & - & - & - & - & - & 22.7\\
+    \system{Mixed-0.5}   & - & - & - & - & - & - & - & 23.6\\
+    \system{Mixed-0.75} & - & - & - & - & - & - & - & 23.7\\
+    \system{Mixed-1} & - & - & - & - & - & - & - & 23.5\\
+    \hline \hline
+    \system{MDAC($\vlambda_0, \vlambda_{news}$)} &-&-&-&-&-&-&-&23.1 \\
+     \system{MDAC($\vlambda_0, \vlambda_{news}$)} &-&-&-&-&-&-&-&23.1 \\ \hline\hline 
     \multicolumn{9}{|c|}{\sl 30 clusters} \\ \hline
-    \system{DDS-v1($\vlambda^*, \vlambda_d$)}& &56.9&51.3&33.2&43.4&&&
-    \\  \hline\hline 
-    \multicolumn{9}{|c|}{\sl Multi-domain training} \\ \hline
-    \\
-    \system{DDS-v1($\vlambda_0, \vlambda_d$)} &36.7&58.2&55&33.3&46.6&91.9&53.6& 23.1 \\
-    \system{DDS-v1($\vlambda_1, \vlambda_d$)} & \\
-    \system{DDS-v1($\vlambda_0, \vlambda_0$)}&35.5&55.5&53.7&32.3&45.6&91.1&52.3& - \\
-    \system{DDS-v1($\vlambda_1, \vlambda_0$)}&37.4&54.4&50.3&32.0&42.7&80.2&49.5 \\ \hline
+    \system{MDAC($\vlambda^*, \vlambda_d$)}&36.3&56.9&51.3&33.2&43.4&89.4&51.8&-\\
+     \system{DDS($\vlambda^*, \vlambda_d$)}&35.3&55.2&&&43.3& & &-\\ \hline\hline 
+    \system{MDAC($\vlambda_1, \vlambda_2$)}&36.2&-&-&-&46.1&-&-&- \\
+    \system{DDS($\vlambda_1, \vlambda_2$)}&36.5&-&-&-&44.4&-&-&- \\
+    \system{MDAC($\vlambda_1, \vlambda_2$)}&-&57.2&53.8&-&-&-&-&- \\
+    \system{DDS($\vlambda_1, \vlambda_2$)}&-&56.1&52.6&-&-&-&-&- \\
+    \system{MDAC($\vlambda_1, \vlambda_2$)}&-&-&-&32&-&90&-&- \\
+    \system{DDS($\vlambda_1, \vlambda_2$)}&-&-&-&32.9&-&91.1&-&- \\ \hline
   \end{tabular}
@@
 
-\fyTodo{Refaire table 3 + (unsupervised domain in source= classification) + new domain}
+\subsubsection{Unseen domain}
+\subsubsection{Automatic clustering}
 
@@
 Domain adaptation is an ancient and vexing problem that has been studied from many angles both for SMT and NMT.
-A survey of supervised and unsupervised DA for NMT is in \cite{Chu18asurvey}, where they distinguish between data-centric and model-centric DA, a view also adapted in the more recent review of \cite{Saunders21domain}.
+A survey of supervised and unsupervised DA for NMT is in \citet{Chu18asurvey}, where they distinguish between data-centric and model-centric DA, a view also adapted in the more recent review of \citet{Saunders21domain}.
 Our approach to DA in this paper clearly falls under the former category.
@@
 Multi-domain adaptation (MDA) aims to develop systems that simultaneously bode well for several domains.
-Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citep{Jiang19multidomain}.
-Unsupervised MDA is considered in \citep{Farajian17multidomain} as an unsupervised DA problem.
+Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citet{Jiang19multidomain}.
+Unsupervised MDA is considered in \citet{Farajian17multidomain} as an unsupervised DA problem.
 
-Most approaches to adaptive / dynamic data selection take inspiration from \citep{Bengio09curriculum}, where the notion of \emph{curriculum learning} (CL) is initially introduced.
+Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of \emph{curriculum learning} (CL) is initially introduced.
 CL is relies on a notion of ``easyness'' of a sample to schedule the presentation of the training data so that the easiest examples  presented first and the hardest, last.
@@
 For instance, \citet{Zhang19curriculum} study supervised domain adaptation, and propose a curriculum approach which progressively extends the data used in training: in the early stages only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
-This is somehow opposite to the recommendations of \cite{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyTodo{These have not been compared ?
+This is somehow opposite to the recommendations of \citet{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyTodo{These have not been compared ?
 and also to what we do ?}
@@
 \citet{Kumar19reinforcement} use reinforcement techniques (deep Q-learning) to learn the curriculum strategy: in this study, complexity corresponds to difficulty levels and which are binned based on contrastive data selection.
-The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyTodo{Alert: what do we do during warm up ?} The same technique is recently applied to the training of a multilingual NMT system \cite{Kumar21learning}.
+The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyTodo{Alert: what do we do during warm up ?} The same technique is recently applied to the training of a multilingual NMT system \citep{Kumar21learning}.
 \citet{Zhou20uncertainty} propose another curriculum-based approach which instead relies on \emph{instance uncertainty} as a measure of their difficulty and presents the data sample starting with the easiest (more predictable) first.
@@
 
-\section{Conclusion and outlook \label{sec:discussion}}
-\fyTodo{Are we learning the dev set ?} Boostrap the dev ?
-
 \section*{Acknowledgments}
@@
 \end{document}
-
-Multi-domain learning, as defined in \citet{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture of several domains, making the problem distinct from mere domain adaption.
-A multi-domain learner is then expected to use these tags effectively \cite{Joshi12multidomain} when computing the combined translation function $h(x,d)$, and to perform well in all domains \cite{Finkel09hierarchical}.
-This setting is closely related to the multi-source adaptation problem formalized in \cite{Mansour09domainadaptation,Mansour09multiple,Hoffman18algorithms}.
-
-This definition seems to be the most accepted view of a multi-domain MT\footnote{An exception is \citep{Farajian17multidomain}, where test translations rely on similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here.
-In the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function.
-
-Given an objective mixture over test distribution $\lambda^{t}_{d}$,$d\in[1,\dots,K]$ and a mixture of training data $\lambda^{s}_{d}$, the Multi-domain training course is determined by the below algorithm \ref{alg:mdmt}.
-The training mixture can be constant or changing through time.
-Optimizing the training mixture is an interesting problem that we would like to address in the paper.
-
-\begin{algorithm}[H]
-\caption{Multi-domain Training} \label{alg:mdmt}
-\label{alg:multidomain}
-\begin{algorithmic}[1]
-\REQUIRE {
-\begin{itemize}
-	\item Corpora $C^d, d\in [1,..,K]$ for $K$ domains equipped by an empirical distribution $D_d(x,y)$
-	\item Dev sets $Dev^d, d\in [1,..,K]$ for $K$ domains.
-	\item Testing mixture weights in the final evaluation $\lambda^t_d, d\in [1,..,K]$
-	\item Batch size $B$.
-	\item Temporal Data Selection Distribution(T-DSD)$$P_{\system{T-DSD}}(t)(x,y)=\sum_{d}\lambda^s_d(t)D_d(x,y)$$
-	\item $Eval\_scores = []$
-	\item $Early\_stopping$ criterion.
-\end{itemize}}
-\REPEAT 
-\STATE{Iteration t.}
-\STATE{Randomly pick $d \in [1,..,K]$ w.r.t  $[\lambda^s_1 \dots \lambda^s_K](t)$.}
-\STATE{Sample $B$ sentences from $C^d$ with empirical distribution $D_d(x,y)$.}
-\STATE{Update model by applying SGD computed from $B$ sampled sentences.}
-\IF{$t \equiv 0 \mod{eval\_step}$}
-	\STATE{Evaluate current model with $K$ dev sets. $S^t_d$ is the performance at iteration $t$ on domain $d$.}
-	\STATE{Report weighted score using testing mixture weights $\lambda^t_d, d\in [1,..,K]$. $$eval(t) = \displaystyle{\mathop{\sum}_d \lambda^t_d S^t_d}$$}
-	\STATE{$Eval\_scores.append(eval(t))$.}
-\ENDIF
-\IF{$Early\_stopping(Eval\_scores)$}
-	\STATE{Stop training loop.}
-\ENDIF
-\UNTIL{convergence}
-\end{algorithmic}
-\end{algorithm}
-
-%\begin{equation} \label{eq:bilevel}
-%\begin{split}
-%\vlambda^{l}_* &= \displaystyle{\mathop{\argmin}_{\lambda^{l}} J(\theta^*(\vlambda^{l}), \vlambda^{t})} \\
-%					& \theta^*(\vlambda^{l}) = \displaystyle{\mathop{\argmin_{\vtheta}} J(\vtheta, \vlambda^{l})}, 
-%\end{split}
-%\end{equation}
-%where we use the following notations:% \fyTodo{Overload Q, from domain to instances to pairs}
-%
-%\begin{align*}
-%J(\vtheta, \vlambda^{t} ) &= \displaystyle{\mathop{\sum}_{(x,y) \sim \widehat{\mathcal{D}^{t}(x)}} \ell(\vtheta,x,y)} \\
-%J(\vtheta, \vlambda^{l}) &= \displaystyle{\mathop{E_{(x,y) \sim \widehat{\mathcal{D}^{l}(x)}}[\ell(\vtheta,x,y)]}} \\
-%l(\vtheta,x,y) &= - log P(x,y;\vtheta)
-%\end{align*}

### 2021-05-07 2021/emnlp/auto_cl_mdmt.tex
complete results and discussion, complete tables

@@
     \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
-    \# \revision{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 & 7.8 / 9.2   \\
-    \# \revision{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 & - \\
-    \# \revision{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 & - \\
+    \# tokens &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 & 7.8 / 9.2   \\
+    \# types & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 & - \\
+    \# uniq & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 & - \\
     \hline
@@
 \subsection{Experimental tasks}
-We evaluate the efficacy of our methods compared to several aforementioned baselines via 3 tasks including domain adaptation; multi-domain adaptation and unsupervised domain adaptation.
+We evaluate the efficacy of our methods compared to several aforementioned baselines via 5 tasks including domain adaptation; multi-domain adaptation, bi-domain adaptation, unseen domain adaptation, and unsupervised domain adaptation.
 
@@
 
-In the Multi-domain adaptation task, given the data of 6 domains, we aim to build an NMT model which gives the best performance regarding a predefined test set compromising equivalently 6 domains.
+In the multi-domain adaptation task, given the data of 6 domains, we aim to build an NMT model which gives the best performance regarding a predefined test set compromising equivalently 6 domains.
 
-In the unsupervised domain adaptation task, we design 2 experiments.
-In the first experiment, given the data of 6 domains, we aim to build an NMT model which is robust to an unseen domain, which is News in this case.
-In the second experiment, we cluster the training data of 6 domains into 30 clusters using the KNN algorithm in the same way as in the work of \citet{Tars18multidomain}, then adapt these clusters to one of 6 domains given the corresponding in-domain dev set.
+In the bi-domain adaptation task, given the data of 6 domains, we only focus on 2 domains of them.
+
+In the fourth experiment, given the data of 6 domains and a small dev set of an unseen domain, which is News, in this case, we aim to build an NMT model which is robust to the unseen domain.
+
+In the unsupervised domain adaptation task, we cluster the training data of 6 domains into 30 clusters using the KNN algorithm in the same way as in the work of \citet{Tars18multidomain}, then adapt these clusters to one of 6 domains given the corresponding in-domain dev set.
 We compare MDAC to the contrast DDS over 6 domains.
@@
   \caption{We report the performance of our method MDAC and the work of \citet{Wang20balancing} DDS in multi-domain adaptation task.
-For a given line, all the columns \emph{correspond to the same system}}
+For a given line, all the columns \emph{correspond to the same multi-domain system}}
   \label{tab:multi-da}
@@
 
+\subsection{Bi-domain adaptation}
+
+\begin{table*}[htbp]
+  \centering \small
+  \begin{tabular}{|l|*9{r|}} \hline
+    \multicolumn{9}{|c|}{\sl Supervised domain adaptation} \\ \hline
+    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} \\ \hline \hline
+    \system{MDAC($\vlambda_1, \vlambda_2$)}&36.2&-&-&-&46.1&-&-&- \\
+    \system{DDS($\vlambda_1, \vlambda_2$)}&36.5&-&-&-&44.4&-&-&- \\
+    \system{MDAC($\vlambda_1, \vlambda_2$)}&-&57.2&53.8&-&-&-&-&- \\
+    \system{DDS($\vlambda_1, \vlambda_2$)}&-&56.1&52.6&-&-&-&-&- \\
+    \system{MDAC($\vlambda_1, \vlambda_2$)}&-&-&-&32&-&90&-&- \\
+    \system{DDS($\vlambda_1, \vlambda_2$)}&-&-&-&32.9&-&91.1&-&- \\ \hline
+  \end{tabular}
+  \caption{Non empty columns correspond to the target domains. For a given line, all the non empty columns \emph{correspond to the same bi-domain system}}
+  \label{tab:bi-da}
+\end{table*}
+
 \subsection{Unsupervised domain adaptation}
-We
+
+According the table \ref{tab:bi-da}, MDAC outperforms DDS in 2 over 3 settings including adapting to \domain{med-it} and \domain{law-bank} while being surpassed in \domain{talk-rel}, showing that MDAC is capable to adapt to a variable set of target domains.
+
 \begin{table*}[htbp]
@@
     \system{MDAC($\vlambda_0, \vlambda_{news}$)} &-&-&-&-&-&-&-&23.1 \\
-     \system{MDAC($\vlambda_0, \vlambda_{news}$)} &-&-&-&-&-&-&-&23.1 \\ \hline\hline 
+     \system{DDS($\vlambda_0, \vlambda_{news}$)} &-&-&-&-&-&-&-&23.3 \\ \hline\hline 
     \multicolumn{9}{|c|}{\sl 30 clusters} \\ \hline
     \system{MDAC($\vlambda^*, \vlambda_d$)}&36.3&56.9&51.3&33.2&43.4&89.4&51.8&-\\
-     \system{DDS($\vlambda^*, \vlambda_d$)}&35.3&55.2&&&43.3& & &-\\ \hline\hline 
-    \system{MDAC($\vlambda_1, \vlambda_2$)}&36.2&-&-&-&46.1&-&-&- \\
-    \system{DDS($\vlambda_1, \vlambda_2$)}&36.5&-&-&-&44.4&-&-&- \\
-    \system{MDAC($\vlambda_1, \vlambda_2$)}&-&57.2&53.8&-&-&-&-&- \\
-    \system{DDS($\vlambda_1, \vlambda_2$)}&-&56.1&52.6&-&-&-&-&- \\
-    \system{MDAC($\vlambda_1, \vlambda_2$)}&-&-&-&32&-&90&-&- \\
-    \system{DDS($\vlambda_1, \vlambda_2$)}&-&-&-&32.9&-&91.1&-&- \\ \hline
+     \system{DDS($\vlambda^*, \vlambda_d$)}&35.3&55.2&49.4&30.6&43.3&89.9&50.6&-\\ \hline
   \end{tabular}
-  \caption{Improved results for DDS in fine-tuning and various multi-domain adaptation settings.}
-  \label{tab:basic-dds}
+  \caption{Non empty columns correspond to the target domains. For a given line, each non empty column correspond to one expert system.}
+  \label{tab:unsupervised-da}
 \end{table*}
@@
 \subsubsection{Unseen domain}
+The top of the table \ref{tab:unsupervised-da} demonstrates the performance on the unseen domain \domain{news} of the NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and dynamic data selection MDAC and DDS.
+These systems have insignificant differences in BLEU, showing that dynamic mixture does not improve the robustness of the NMT system against unseen domains.
+However, the fact that the performance of MDAC and DDS is close to the best performance guarantees the harmlessness of these algorithms to the robustness of the NMT model.
+
 \subsubsection{Automatic clustering}
+The bottom of the table \ref{tab:unsupervised-da} shows the performance of NMT systems adapted to each domain in using 30 clusters instead of 6 domains.
+According to the table, MDAC outperforms DDS on every domain, proving the efficacy of the method in this situation where no information of the domain is given a priori.
 

### 2021-05-07 2021/emnlp/auto_cl_mdmt.tex
change title

@@
 
-\title{Automated curriculum learning for Multi-Domain NMT by using loss-driven reward}
+\title{Multi-Domain Adaptation in Neural Machine Translation with dynamic sampling strategy}
 
@@
 Several recent proposals \citep{platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Graves17automated,Wang20balancing,Wang20learning-multi,Vanderwees17dynamic} have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies.
-Furthermore \citet{Zhang19curriculum,Vanderwees17dynamic} construct a fixed curriculum while \citet{platanios19competence,Kumar19reinforcement,Graves17automated,Wang20balancing,Wang20learning-multi} build curricula automatically adapted to the training.
+Furthermore \citet{Zhang19curriculum,Vanderwees17dynamic} construct a static curriculum while \citet{platanios19competence,Kumar19reinforcement,Graves17automated,Wang20balancing,Wang20learning-multi} build curricula automatically adapted to the training.
 
@@
 We parameterize $\lambda^{l}$ by a differentiable function $\lambda^l(\psi)$.
-We divide the training into many short sessions; in each session $t$, the model is trained with a fixed data distribution $\lambda^{l}(\psi_t)$.
+We divide the training into many short sessions; in each session $t$, the model is trained with a static data distribution $\lambda^{l}(\psi_t)$.
 After one learning session, we update the data distribution via REINFORCE algorithm \citep{Williams92simple}.
@@
 \begin{align*}
-  R(d) = J^t(\theta_{t+k},\lambda^t) - J^t(\theta_t,\lambda^t)
+  R(d) = J^t(\theta_{t+k},\vlambda^t) - J^t(\theta_t,\vlambda^t)
 \end{align*}
@@
 x^i_j, y^i_j &\sim& \widehat{D^l_d}(x) \\
-J^t(\theta,\lambda_t) &=& \displaystyle{\mathop{\sum}_{d=1}^K}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta,x^t_d,y^t_d)
+J^t(\theta,\lambda_t) &=& \displaystyle{\mathop{\sum}_{d=1}^{n_d}}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta,x^t_d,y^t_d)
 \end{array}
@@
 
-The work of \cite{Wang20balancing} based on the Bi-level Optimization framework, which aims to find an optimal fixed distribution $\lambda^{l}$ that will result in the best NMT model regarding a given target dev set at the end of the training, derives also a similar form of update for $\psi$.
+The work of \cite{Wang20balancing} based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\lambda^{l}$ that will result in the best NMT model regarding a given target dev set at the end of the training, derives also a similar form of update for $\psi$.
 However, their reward is the cosine similarity between the gradient computed with the training data and the gradient computed with the dev set.
@@
 
-\item Our comparison of multi-domain systems includes baselines with fixed data mixtures corresponding to $\vlambda_0$ and $\vlambda_1$, and our reimplementations of recent proposals from the literature: Curriculum Learning of \cite{Zhang19curriculum} and Differential Data Selection of \cite{Wang20balancing}.
+\item Our comparison of multi-domain systems includes baselines with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning of \cite{Zhang19curriculum} and Differential Data Selection of \cite{Wang20balancing}.
 \end{itemize}
@@
 
-MDAC outperforms almost fixed training distribution including $\big[ \lambda_0, \lambda_{0.75}, \lambda_{1.0}\big]$ with significant margin while performs slightly better than $\big[ \lambda_{0.25}, \lambda_{0.5} \big]$.
+MDAC outperforms static training distributions including $\big[ \lambda_0, \lambda_{0.75}, \lambda_{1.0}\big]$ with significant margin while performs slightly better than $\big[ \lambda_{0.25}, \lambda_{0.5} \big]$.
 This observation indicates that MDAC allows us to skip the heuristic choice for the training mixture of different domains or tasks.
@@
   \begin{tabular}{|l|*9{r|}} \hline
-    \multicolumn{9}{|c|}{\sl Supervised domain adaptation} \\ \hline
     domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} \\ \hline \hline

### 2021-05-07 2021/emnlp/auto_cl_mdmt.tex
abstract + new style

@@
-%
-% File acl2021.tex
-%
-%% Based on the style files for EMNLP 2020, which were
-%% Based on the style files for ACL 2020, which were
-%% Based on the style files for ACL 2018, NAACL 2018/19, which were
-%% Based on the style files for ACL-2015, with some improvements
-%%  taken from the NAACL-2016 style
-%% Based on the style files for ACL-2014, which were, in turn,
-%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
-%% EACL-2009, IJCNLP-2008...
-%% Based on the style files for EACL 2006 by 
-%%e.agirre@ehu.es or Sergi.Balari@uab.es
-%% and that of ACL 08 by Joakim Nivre and Noah Smith
+% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
+\pdfoutput=1
+% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.
 
-\documentclass[11pt,a4paper]{article}
-\usepackage[hyperref]{acl2021}
+\documentclass[11pt]{article}
+
+% Remove the "review" option to generate the final version.
+\usepackage[review]{emnlp2021}
+
+% Standard package includes
 \usepackage{times}
 \usepackage{latexsym}
-%\renewcommand{\UrlFont}{\ttfamily\small}
 
+% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
+\usepackage[T1]{fontenc}
+% For Vietnamese characters
+% \usepackage[T5]{fontenc}
+% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets
+
+% This assumes your files are encoded as UTF8
+\usepackage[utf8]{inputenc}
+\renewcommand{\UrlFont}{\ttfamily\small}
+\usepackage{graphicx}
+\usepackage{grffile}
+\usepackage{multirow}
+\usepackage{xcolor,colortbl}
+\usepackage{amsmath}
+%\usepackage{fdsymbol}
+\usepackage{amssymb}
+\usepackage{makecell}
+\usepackage[super]{nth}
+\usepackage{arydshln}
+
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+\usepackage{regexpatch}
+\usepackage{subcaption}
 % This is not strictly necessary, and may be commented out,
@@
 \usepackage{microtype}
-\usepackage{graphicx}
 
-%\aclfinalcopy % Uncomment this line for the final submission
-%\def\aclpaperid{***} %  Enter the acl Paper ID here
-
-%\setlength\titlebox{5cm}
-% You can expand the titlebox if you need extra space
-% to show all the authors. Please do not make the titlebox
-% smaller than 5cm (the original size); we will check this
-% in the camera-ready version and ask you to change it back.
-
-\newcommand\BibTeX{B\textsc{ib}\TeX}
-
-% Standard package includes
-\usepackage{times}
 \usepackage{url}
-\usepackage{latexsym}
-\usepackage{graphicx}
-\usepackage{xcolor}
 \usepackage{longtable}
@@
 \usepackage{float}
-\usepackage{amsmath,amsfonts,amssymb}
-\usepackage{algorithm, algorithmic}
 
-\usepackage{tikz}
 \usepackage{pgfplots}
@@
 
-\title{Multi-Domain Adaptation in Neural Machine Translation with dynamic sampling strategy}
+\title{Multi-Domain Adaptation in Neural Machine Translation with Dynamic Sampling Strategies}
 
@@
 \begin{abstract}
-  When building machine translation systems, one often needs to make the best out of heterogeneous sets of parallel data in training to achieve the best test performance in one or several domain(s) of interest.
-This multi-source / multi-domain adaptation problem is often approached with instance selection or reweighting strategies, most of which pre-suppose an ex-ante assessment of the relevance of training instances.
-In this paper, we propose a novel dynamic data selection algorithm that re-evaluates the data utility and adapts the data selection policy along with the training of the NMT model.
-We explore the ability of our method to serve as a generic framework for the previously mentioned situations.
+  Building effective Neural Machine Translation models often implies to accomodate diverse sets of heterogeneous data so as to optimize performance for the domain(s) of interest.
+Such multi-source / multi-domain adaptation problems are typically approached through instance selection or reweighting strategies, based on a static assessment of the relevance of training instances with respect to the task at hand.
+In this paper, we study dynamic data selection strategies that are able to automatically re-evaluate the usefulness of data samples and to evolve a data selection policy in the course of training.
+Based on the results of multiple experiments, we show that such methods constitute a generic framework to automatically and effectively cater for a variety of real-world situations, from supervised multi-source domain adaptation to multi-domain learning and unsupervised domain adaptation.
 \end{abstract}

### 2021-05-10 2021/emnlp/auto_cl_mdmt.tex
adding statistical significance, removing residual experiments

@@
 \begin{abstract}
-  Building effective Neural Machine Translation models often implies to accomodate diverse sets of heterogeneous data so as to optimize performance for the domain(s) of interest.
+  Building effective Neural Machine Translation models often imply accommodating diverse sets of heterogeneous data so as to optimize performance for the domain(s) of interest.
 Such multi-source / multi-domain adaptation problems are typically approached through instance selection or reweighting strategies, based on a static assessment of the relevance of training instances with respect to the task at hand.
 In this paper, we study dynamic data selection strategies that are able to automatically re-evaluate the usefulness of data samples and to evolve a data selection policy in the course of training.
-Based on the results of multiple experiments, we show that such methods constitute a generic framework to automatically and effectively cater for a variety of real-world situations, from supervised multi-source domain adaptation to multi-domain learning and unsupervised domain adaptation.
+Based on the results of multiple experiments, we show that such methods constitute a generic framework to automatically and effectively cater to a variety of real-world situations, from supervised multi-source domain adaptation to multi-domain learning and unsupervised domain adaptation.
 \end{abstract}
@@
 We evaluate our method in various single-domain and multi-domain adaptation settings.
-We compare our method with several contrasts including the work of \citet{Zhang19curriculum}, that we name CL, and \citet{Wang20balancing}, that we name DDS.
+We compare our method with several contrasts including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}.
 From now on, we name our method MDAC which is brief for Multi-Domain Automated Curriculum.
-Based on experimental results obtained on a diverse set of domains, our main conclusions are that (a) using MDAC often yields overall performance that are as good as the standard fine-tuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test target distributions without any meta-parameter search.
+Based on experimental results obtained on a diverse set of domains, our main conclusions are that (a) using MDAC often yields overall performance that are as good as the standard fine-tuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test target distributions.
 
@@
 \begin{align*}
-\psi_{t+1} &= \psi_t + \mathbf{lr}_{data} * \displaystyle{\mathop{\sum}_{d=1}^K} R(d) * \frac{\partial \lambda^l(d;\psi_t)}{\partial \psi} \\
+\psi_{t+1} &= \psi_t + \mathbf{lr}_{1} * \displaystyle{\mathop{\sum}_{d=1}^{n_d}} R(d) * \frac{\partial \lambda^l(d;\psi_t)}{\partial \psi} \\
 \end{align*}
@@
 \begin{array}{rcl}
-\theta_{t+i} &=& \theta_{t+i-1} - \mathbf{lr}_{nmt} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta} \\ \nonumber
+\theta_{t+i} &=& Update\big(\theta_{t+i-1},[x^i_j,y^i_j]_{j=1}^N\big) \\ \nonumber
 x^i_j, y^i_j &\sim& \widehat{D^l_d}(x) \\
@@
 \endgroup
-$\mathbf{lr}_{data}$ is the learning rate of the parameter of the sampler.
-$\mathbf{lr}_{nmt}$ is the learning rate of the parameter of the NMT system.
-$J^t(\theta,\lambda_t)$ is the weighted loss computed over $K$ dev-sets corresponding to $K$ target domains.
+%\theta_{t+i-1} - \mathbf{lr_2} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta}
+$N$ is the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $\mathbf{lr_2}$ is the learning rate of the NMT system; $l(\theta,x,y)$ is the loss of the NMT model on the sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss computed over $n_d$ dev-sets corresponding to $n_d$ given domains.
 
-To compute the reward of using data from domain $d$ to train the model, $R(d)$, we simulate k training steps from the current checkpoint by training the model with k batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
+To compute the reward of using data from domain $d$ to train the model, $R(d)$, we simulate $k$ training steps from the current checkpoint by training the model with k batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
 The reward of our method is inspired by the target prediction gain in \cite{Graves17automated}.
 However, \cite{Graves17automated} used accumulated gains from the past as reward while we use the "future" gain as reward.
+Furthermore, the choice of the parameterization of the sampling distribution in our method is different from one in \cite{Graves17automated}, although this choice is empirical.
 
@@
 
-These questions are further discussed in Section~\ref{sec:results}. We now turn to our experimental conditions.
+These questions are further explored in Section~\ref{sec:results}. We now turn to our experimental conditions.
 
@@
 
-\subsection{Baseline architectures \label{ssec:baseline}}
+\subsection{Baseline systems \label{ssec:baseline}}
 Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
@@
 These systems are denoted \system{Mixed-$\alpha$} below.
-\system{Mixed-$0$} uses a uniform domain distribution, \system{Mixed-$1.0$} simple uses the observed domain distribution.
+\system{Mixed-$0$} uses a uniform domain distribution, \system{Mixed-$1.0$} uses the observed domain distribution.
 \item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing.
 The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems each adapted for one domain, with no parameter sharing across domains.
-\item two multi-domain versions of the approach of \newcite{Bapna19simple}, denoted \system{FT-Res} and \system{MDL-Res}, where a domain-specific adaptation module is added to all the Transformer layers; within each layer, residual connections enable to short-cut this adapter.
-The former variant corresponds to the original proposal of \citet{Bapna19simple} (see also \cite{Sharaf20metalearning}).
-It fine-tunes the adapter modules of a \system{Mixed-$1.0$} system independently for each domain, keeping all the other parameters frozen.
-The latter uses the same architecture, but a different training procedure and learns all parameters jointly from scratch with the prefixed mixtures of training data \ref{mixture:trn}.\fyDone{Keep this ?}
 
-\item Our comparison of multi-domain systems includes baselines with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning of \cite{Zhang19curriculum} and Differential Data Selection of \cite{Wang20balancing}.
+\item Our comparison of multi-domain systems includes baselines with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning of \cite{Zhang19curriculum} that we denote CL and Differential Data Selection of \cite{Wang20balancing} that we denote DDS.
 \end{itemize}
@@
 
-We provide also in this section our choices of the hyper-parameters of our algorithm including the number of simulation training steps $k=10$, and the update rate $\mathbf{lr}_{data}=0.1$.
+We also provide in this section our choices of the hyper-parameters of our algorithm including the number of simulation training steps $k=10$, and the update rate $\mathbf{lr}_{data}=0.1$.
 We split the training into many short sessions that last 2000 training steps each.
 \subsection{CL and DDS's re-implementation}
-We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to original code provided by \citet{Wang20balancing} \footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}.
+We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to original code of \citet{Wang20balancing} \footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}.
 
@@
 
-In the multi-domain adaptation task, given the data of 6 domains, we aim to build an NMT model which gives the best performance regarding a predefined test set compromising equivalently 6 domains.
+In the multi-domain adaptation task, given the data of 6 domains, we aim to build an NMT model which gives the best performance regarding a predefined test set containing equivalently 6 domains.
 
@@
 In the unsupervised domain adaptation task, we cluster the training data of 6 domains into 30 clusters using the KNN algorithm in the same way as in the work of \citet{Tars18multidomain}, then adapt these clusters to one of 6 domains given the corresponding in-domain dev set.
-We compare MDAC to the contrast DDS over 6 domains.
+We compare MDAC to DDS over 6 domains.
 
@@
 We initialize MDAC and DDS by 2 different distribution $\lambda_0$ and $\lambda_1$.
-According to table \ref{tab:da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen.
+According to Table \ref{tab:da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen.
 The same conclusion is also true for DDS.
 
-According to the table \ref{tab:da}, MDAC is outperformed by finetuning method in large domains including medical and law.
-However, MDAC improves finetuning method by approximately 0.5 BLEU in \domain{bank} and 1.0 BLEU in \domain{rel}.
-This indicates that for small domains, out-of-domain data can improve the generalization of the NMT model and MDAC is capable to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in finetuning.
+According to Table \ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}.
+However, MDAC improves fine-tuning by approximately 0.5 BLEU in \domain{bank} and 1.0 BLEU in \domain{rel}.
+This indicates that for small domains, out-of-domain data can improve the generalization of the NMT model and MDAC is capable to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
 
@@
     domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline
-    FT-Full($d$) \hfill [BLEU] &37.7&59.2&54.5&34&46.8&90.8&53.8\\
-    FT-Res($d$) \hfill &37.3&57.9&53.9&33.8&46.7&90.2&53.3\\ \hline
+    FT-Full($d$) \hfill [BLEU] &37.7&59.2&54.5&34&46.8&90.8&53.8\\ \hline
     \hline
+    \system{CL($d$)} &37.2&55.4&53.0&31.3&45.3&90.8&52.2\\ \hline
     \system{DDS($\vlambda_0, \vlambda_d$)} &36.7&56.2&54.3&33.5&46.8&91.8&53.2\\
-    \system{DDS($\vlambda_1, \vlambda_d$)} &36.4&56.3& 45.4& 32.6&44.5&43.1&43.1\\ \hline
-    \system{CL($d$)} &37.2&55.4&53.0&31.3&45.3&90.8&52.2\\ \hline
-    \system{MDAC($\vlambda_0, \vlambda_d$)} &36.7&58.2&55&33.3&46.6&91.9&53.6\\
-    \system{MDAC($\vlambda_1, \vlambda_d$)} &37.2&55.1&51.8&32.8&44.9&85.4&51.2\\ \hline
+    \system{MDAC($\vlambda_0, \vlambda_d$)} &36.7&58.2$^{**}$&55$^{*}$&33.3&46.6&91.9$^{***}$&53.6\\
+     \hline
+    \system{DDS($\vlambda_1, \vlambda_d$)} &36.4&56.3& 45.4& 32.6&44.5&43.1&43.1\\
+    \system{MDAC($\vlambda_1, \vlambda_d$)} &37.2&55.1&51.8&32.8&44.9&85.4&51.2 \\ \hline
   \end{tabular}
-  \caption{We report scores of each method with 6 target domains, and compare with full fine-tuning: each column corresponds to a distinct system.}
+  \caption{We report scores of each method with 6 target domains, and compare with full fine-tuning: each column corresponds to a distinct system.
+($^{*}$) MDAC is significantly better than CL, fine-tuning and DDS with $p<0.05$.
+($^{**}$) MDAC is significantly better than CL and DDS with $p<0.05$.
+($^{***}$) MDAC is significantly better than CL, fine-tuning with $p<0.05$}
   \label{tab:da}
@@
 We initialize MDAC and DDS by 2 different distribution $\lambda_0$ and $\lambda_1$.
-According to table \ref{tab:multi-da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen.
+According to Table \ref{tab:multi-da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen.
 The same conclusion is also true for DDS.
@@
 
-We show in the below graphic \ref{fig:sampling}, the evolution of the training mixture through training sessions.
+We show in the below figure \ref{fig:sampling}, the evolution of the training mixture through training sessions.
 According to the figure \ref{fig:DDS} of DDS, the proportion of 6 domains converges rapidly to bi-domain mode in which only \domain{med} and \domain{rel} have significant probability.
@@
     \system{DDS($\vlambda_0, \vlambda_0$)} &37.3&55.1&51&33.5&43.4&90.8&51.8\\ 
-    \system{MDAC($\vlambda_0, \vlambda_0$)}&35.5&55.5&53.7&32.3&45.6&91.1&52.3\\ 
+    \system{MDAC($\vlambda_0, \vlambda_0$)}&35.5&55.5&53.7$^*$&32.3&45.6&91.1$^*$&52.3$^{**}$\\ 
     \hline \hline
@@
   \caption{We report the performance of our method MDAC and the work of \citet{Wang20balancing} DDS in multi-domain adaptation task.
-For a given line, all the columns \emph{correspond to the same multi-domain system}}
+For a given line, all the columns \emph{correspond to the same multi-domain system}.
+($^*$) MDAC is significantly better than \system{Mixed-$\alpha$} with $p<0.05$ and DDS.
+($^{**}$) MDAC is significantly better than DDS with $p<0.05$.}
   \label{tab:multi-da}
@@
     domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} \\ \hline \hline
-    \system{MDAC($\vlambda_1, \vlambda_2$)}&36.2&-&-&-&46.1&-&-&- \\
+    \system{MDAC($\vlambda_1, \vlambda_2$)}&36.2&-&-&-&46.1$^*$&-&-&- \\
     \system{DDS($\vlambda_1, \vlambda_2$)}&36.5&-&-&-&44.4&-&-&- \\
-    \system{MDAC($\vlambda_1, \vlambda_2$)}&-&57.2&53.8&-&-&-&-&- \\
+    \system{MDAC($\vlambda_1, \vlambda_2$)}&-&57.2$^*$&53.8$^*$&-&-&-&-&- \\
     \system{DDS($\vlambda_1, \vlambda_2$)}&-&56.1&52.6&-&-&-&-&- \\
@@
   \end{tabular}
-  \caption{Non empty columns correspond to the target domains. For a given line, all the non empty columns \emph{correspond to the same bi-domain system}}
+  \caption{Non empty columns correspond to the target domains.
+For a given line, all the non empty columns \emph{correspond to the same bi-domain system}.
+($^*$) MDAC is significantly better than DDS with $p<0.05$.}
   \label{tab:bi-da}
 \end{table*}
-
-\subsection{Unsupervised domain adaptation}
-
-According the table \ref{tab:bi-da}, MDAC outperforms DDS in 2 over 3 settings including adapting to \domain{med-it} and \domain{law-bank} while being surpassed in \domain{talk-rel}, showing that MDAC is capable to adapt to a variable set of target domains.
+According to Table \ref{tab:bi-da}, MDAC outperforms DDS in 2 over 3 settings including adapting to \domain{med-it} and \domain{law-bank} while being surpassed in \domain{talk-rel}, showing that MDAC is capable to adapt to a variable set of target domains.
 
@@
     \multicolumn{9}{|c|}{\sl 30 clusters} \\ \hline
-    \system{MDAC($\vlambda^*, \vlambda_d$)}&36.3&56.9&51.3&33.2&43.4&89.4&51.8&-\\
+    \system{MDAC($\vlambda^*, \vlambda_d$)}&36.3$^*$&56.9$^*$&51.3$^*$&33.2$^*$&43.4&89.4&51.8$^*$&-\\
      \system{DDS($\vlambda^*, \vlambda_d$)}&35.3&55.2&49.4&30.6&43.3&89.9&50.6&-\\ \hline
   \end{tabular}
-  \caption{Non empty columns correspond to the target domains. For a given line, each non empty column correspond to one expert system.}
+  \caption{Non empty columns correspond to the target domains. For a given line, each non empty column correspond to one expert system. ($^*$)MDAC is significantly better than DDS.}
   \label{tab:unsupervised-da}
@@
 
-\subsubsection{Unseen domain}
-The top of the table \ref{tab:unsupervised-da} demonstrates the performance on the unseen domain \domain{news} of the NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and dynamic data selection MDAC and DDS.
+\subsection{Unseen domains}
+The top of Table \ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and dynamic data selection MDAC and DDS.
 These systems have insignificant differences in BLEU, showing that dynamic mixture does not improve the robustness of the NMT system against unseen domains.
@@
 
-\subsubsection{Automatic clustering}
-The bottom of the table \ref{tab:unsupervised-da} shows the performance of NMT systems adapted to each domain in using 30 clusters instead of 6 domains.
-According to the table, MDAC outperforms DDS on every domain, proving the efficacy of the method in this situation where no information of the domain is given a priori.
+\subsection{Automatic clustering}
+The bottom of Table \ref{tab:unsupervised-da} shows the performance of NMT systems adapted to each domain in using 30 clusters instead of 6 domains.
+According to Table \ref{tab:unsupervised-da}, MDAC outperforms DDS on every domain, proving the efficacy of the method in this situation where no information of the domain is given a priori.
 

### 2021-05-11 2021/emnlp/auto_cl_mdmt.tex
adding lambda_d, lambda_d DA scores, adding one reference

@@
 Single-source and multi-source domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
-A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
+A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work} \revision{cite also \citep{shen21source}}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
 
@@
 
-In this paper, we propose a novel method of building an automated curriculum for training the Multi-Domain NMT model, that is inspired by the work of \citet{Graves17automated}.
-We evaluate our method in various single-domain and multi-domain adaptation settings.
+The major contribution of our work consists of 2 parts.
+First, we propose a novel framework that covers the domain adaptation problem and the multi-domain adaptation problem.
+Secondly, we propose a novel method of building an automated curriculum for training the Multi-Domain NMT model.
+We evaluate our method in various single-domain adaptation and multi-domain adaptation settings.
 We compare our method with several contrasts including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}.
 From now on, we name our method MDAC which is brief for Multi-Domain Automated Curriculum.
-Based on experimental results obtained on a diverse set of domains, our main conclusions are that (a) using MDAC often yields overall performance that are as good as the standard fine-tuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test target distributions.
+Based on experimental results obtained on a diverse set of domains, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard fine-tuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test target distributions.
 
@@
 The reward of our method is inspired by the target prediction gain in \cite{Graves17automated}.
-However, \cite{Graves17automated} used accumulated gains from the past as reward while we use the "future" gain as reward.
+However, \cite{Graves17automated} used accumulated gains from the past as reward while we predict the utility of each domain for improving the competence of the system at the current state by simulating the training with the data of the domain.
 Furthermore, the choice of the parameterization of the sampling distribution in our method is different from one in \cite{Graves17automated}, although this choice is empirical.
@@
 Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.
+We train all systems with 200~000 iterations.
 
@@
 
-We also provide in this section our choices of the hyper-parameters of our algorithm including the number of simulation training steps $k=10$, and the update rate $\mathbf{lr}_{data}=0.1$.
-We split the training into many short sessions that last 2000 training steps each.
+We also provide in this section our choices of the hyper-parameters of our algorithm including the number of simulation training steps $k=10$, and the learning rate $\mathbf{lr}_{data}=0.001$.
+We update the sampling distribution via 100 gradient descent iterations for almost the experimental settings except that for adaptation with automatic cluster \ref{ssec:clda}, we use 20 gradient descent iterations.
+We split the training into 100 short sessions that last 2000 training steps each.
+The choice of those hyper-parameters is mostly heuristic except for the learning rate $\mathbf{lr}_{data}$ which is optimized via grid search over a range of values $[0.001,0.0025,0.005]$
 \subsection{CL and DDS's re-implementation}
@@
 
-\subsection{Domain Adaptation}
+\subsection{Domain Adaptation}\label{ssec:da}
 In this setting, we aim to build an expert NMT model in only one domain. This setting corresponds to Dirac distribution $\lambda^t$, in which the probability of domains is 0 except for one domain.
 
-We initialize MDAC and DDS by 2 different distribution $\lambda_0$ and $\lambda_1$.
-According to Table \ref{tab:da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen.
+We initialize MDAC and DDS by 3 different distributions $\lambda_0$, $\lambda_1$ and $\lambda_d$.
+According to Table \ref{tab:da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen, though $\lambda_0$  initialization only improve significantly $\lambda_d$ initialization in small domains including \domain{talk},\domain{bank}and\domain{it}.
+$\lambda_d$ initialization is beneficial for adapting the NMT system to large domains such as \domain{med} and \domain{law}.
 The same conclusion is also true for DDS.
 
+We now compare the best MDAC (initialized by $\lambda_0$) to fine-tuning.
 According to Table \ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}.
@@
     \system{MDAC($\vlambda_1, \vlambda_d$)} &37.2&55.1&51.8&32.8&44.9&85.4&51.2 \\ \hline
+    \system{DDS($\vlambda_d, \vlambda_d$)} &37.0&59.6&53.7&30.2&45.7&91.6&53.0 \\
+    \system{MDAC($\vlambda_d, \vlambda_d$)}&37.9&59.6&53.9&30.4&45.8&92.1&53.3 \\
+    \hline
   \end{tabular}
@@
 
-\subsection{Multi-domain adaptation}
+\subsection{Multi-domain adaptation}\label{ssec:mda}
 
@@
 
-\subsection{Bi-domain adaptation}
+\subsection{Bi-domain adaptation}\label{ssec:bida}
 
@@
 
-\subsection{Unseen domains}
+\subsection{Unseen domains}\label{ssec:uda}
 The top of Table \ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and dynamic data selection MDAC and DDS.
@@
 
-\subsection{Automatic clustering}
+\subsection{Automatic clustering}\label{ssec:clda}
 The bottom of Table \ref{tab:unsupervised-da} shows the performance of NMT systems adapted to each domain in using 30 clusters instead of 6 domains.
-According to Table \ref{tab:unsupervised-da}, MDAC outperforms DDS on every domain, proving the efficacy of the method in this situation where no information of the domain is given a priori.
-
+According to Table \ref{tab:unsupervised-da}, MDAC underperforms the same version under the domain adaptation setting \ref{ssec:da}.
+However, automatic clustering could make the situations harder than the conventional settings where
 \section{Related Work \label{sec:related}}

### 2021-05-11 2021/emnlp/auto_cl_mdmt.tex
small changes in intro

@@
 This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions \citep{Moore10selection,Axelrod11domain}.
-Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data; it may also be suboptimal, as some target domains or languages might be easier to train than others.
+Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data.
+It may also be suboptimal, as some target domains or languages might be easier to train than others.
 For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not promise optimization of a particular domain.
 
-Several recent proposals \citep{platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Graves17automated,Wang20balancing,Wang20learning-multi,Vanderwees17dynamic} have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies.
-Furthermore \citet{Zhang19curriculum,Vanderwees17dynamic} construct a static curriculum while \citet{platanios19competence,Kumar19reinforcement,Graves17automated,Wang20balancing,Wang20learning-multi} build curricula automatically adapted to the training.
+Several recent proposals
+% \citep{Vanderwees17dynamic,Graves17automated,platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Wang20balancing,Wang20learning-multi}
+have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies.
+Notably,  \citet{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while\citet{Graves17automated,platanios19competence,Kumar19reinforcement,Wang20balancing,Wang20learning-multi} build curricula that automatically adapt to the training data.
 
-The major contribution of our work consists of 2 parts.
-First, we propose a novel framework that covers the domain adaptation problem and the multi-domain adaptation problem.
-Secondly, we propose a novel method of building an automated curriculum for training the Multi-Domain NMT model.
-We evaluate our method in various single-domain adaptation and multi-domain adaptation settings.
+In this paper, we contribute to this line of research in several ways.
+First, we propose a novel framework (\emph{Multi-Domain Automated Curriculum}, MDAC for short) that simultaneously accounts for the domain adaptation and the multi-domain adaptation problems.
+Second, we propose a novel method of building an automated curriculum for training Multi-Domain NMT models.
+We evaluate our method in various single-domain and multi-domain adaptation settings.
 We compare our method with several contrasts including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}.
-From now on, we name our method MDAC which is brief for Multi-Domain Automated Curriculum.
-Based on experimental results obtained on a diverse set of domains, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard fine-tuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test target distributions.
+% From now on, we name our method MDAC which is brief for .
+Based on these experimental results, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard finetuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test situations, from targeting one single domain to the full multi-domain scenario.
+% distributions
 

### 2021-05-11 2021/emnlp/auto_cl_mdmt.tex
review section 3

@@
 It may also be suboptimal, as some target domains or languages might be easier to train than others.
-For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not promise optimization of a particular domain.
+For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not achieve optimily for any particular domain.
 
@@
 
-We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \citep{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
+We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \citep{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \citep{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
 Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
@@
 
-It is usual in MT to opportunistically collect training samples from several domains, which means that training instances are distributed according to the mixture $\mathcal{D}^s$ such that $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}(d) \mathcal{D}_d(x)$, where $\{\lambda^{s}(d), d=1 \dots n_d\}$ are the corresponding mixture weights satisfying $\sum_d \lambda^{s}(d)=1$.
+It is usual in MT to opportunistically collect training samples from several domains, which means that training instances are distributed according to a mixture $\mathcal{D}^s$ such that $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}(d) \mathcal{D}_d(x)$, with $\{\lambda^{s}(d), d=1 \dots n_d\}$ the mixture weights satisfying $\sum_d \lambda^{s}(d)=1$.
 
@@
 These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unsupervised domain adaptation); in multi-domain adaptation finally, the test distribution is itself a mixture of domains, some of which may also be observed in training.
-Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of fine-tuning.
-These situations are illustrated in Figure~\ref{fig:mdmt-lambdas}.
+Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of pure domain adaptation.
+These various settings are illustrated Figure~\ref{fig:mdmt-lambdas}.
 \begin{figure}[h]
@@
 
-A more practical standard approach to supervised DA is \emph{fine-tuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
+A more practical standard approach to supervised DA is \emph{finetuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
 With our notations, this approach amounts to first learning an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$) with $\indic{A}$ the indicator function for predicate $A$.
@@
 \subsection{Basic principles}
-Assuming training data from $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as  $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
+Assuming training data in each of the $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
 We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t _d}$ to denote the empirical train and test distributions for domain $d$, and $\widehat{\mathcal{D}^{u}}(x;\lambda^{u}) = \sum_{d} \lambda^{u}(d) \widehat{\mathcal{D}^{u}_d}(x)$ for $u\in\{l,t\}$.
@@
 
-Our method MDAC constructs an automatically adapting distribution $\lambda^{l}$ that optimizes the data selection policy along with the training of the NMT model.
+% Our method
+MDAC constructs an adaptative training distribution $\lambda^{l}$ that optimizes the data selection policy along with the training of the NMT model.
 We parameterize $\lambda^{l}$ by a differentiable function $\lambda^l(\psi)$.
 We divide the training into many short sessions; in each session $t$, the model is trained with a static data distribution $\lambda^{l}(\psi_t)$.
-After one learning session, we update the data distribution via REINFORCE algorithm \citep{Williams92simple}.
-
+After one learning session, we update the data distribution using the REINFORCE algorithm of \citet{Williams92simple}.
+The evolution of $\psi$ is thus defined by:
 \begin{align*}
-\psi_{t+1} &= \psi_t + \mathbf{lr}_{1} * \displaystyle{\mathop{\sum}_{d=1}^{n_d}} R(d) * \frac{\partial \lambda^l(d;\psi_t)}{\partial \psi} \\
+\psi_{t+1} &= \psi_t + \mathbf{lr}_{1} * \displaystyle{\mathop{\sum}_{d=1}^{n_d}} R(d) * \frac{\partial \lambda^l(d;\psi_t)}{\partial \psi}, \\
 \end{align*}
@@
 \allowdisplaybreaks
-With the following reward computation:
+where reward $R(d)$ is computed as:
 \begin{align*}
-  R(d) = J^t(\theta_{t+k},\vlambda^t) - J^t(\theta_t,\vlambda^t)
+  R(d) = J^t(\theta_{t+k},\vlambda^t) - J^t(\theta_t,\vlambda^t),
 \end{align*}
-Where 
+and where we also define:
 \begin{equation}
@@
 x^i_j, y^i_j &\sim& \widehat{D^l_d}(x) \\
-J^t(\theta,\lambda_t) &=& \displaystyle{\mathop{\sum}_{d=1}^{n_d}}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta,x^t_d,y^t_d)
+J^t(\theta,\lambda_t) &=& \displaystyle{\mathop{\sum}_{d=1}^{n_d}}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta,x^t_d,y^t_d).
 \end{array}
@@
 \endgroup
-%\theta_{t+i-1} - \mathbf{lr_2} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta}
-$N$ is the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $\mathbf{lr_2}$ is the learning rate of the NMT system; $l(\theta,x,y)$ is the loss of the NMT model on the sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss computed over $n_d$ dev-sets corresponding to $n_d$ given domains.
+% \theta_{t+i-1} - \mathbf{lr_2} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta}
+In these equations, $N$ denotes the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $\mathbf{lr_2}$ is the learning rate of the NMT system; $l(\theta,x,y)$ is the loss of the NMT model on sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss aggregated over $n_d$ dev-sets corresponding to the $n_d$ domains.
 
-To compute the reward of using data from domain $d$ to train the model, $R(d)$, we simulate $k$ training steps from the current checkpoint by training the model with k batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
-The reward of our method is inspired by the target prediction gain in \cite{Graves17automated}.
-However, \cite{Graves17automated} used accumulated gains from the past as reward while we predict the utility of each domain for improving the competence of the system at the current state by simulating the training with the data of the domain.
-Furthermore, the choice of the parameterization of the sampling distribution in our method is different from one in \cite{Graves17automated}, although this choice is empirical.
+To compute the reward $R(d)$ of using data from domain $d$ to train the model, we simulate $k$ training steps from current checkpoint by training the model with $k$ batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
+This computation is inspired by the target prediction gain in \citep{Graves17automated}.
+However, \cite{Graves17automated} used accumulated gains from the past as reward, while we predict the usefulness of each domain for improving the future performance of the system given its current state.
+This is achieved by simulating a round of training with only the data from one domain.
+Furthermore, the choice of the parameterization of the sampling distribution in our method differs from that of \citet{Graves17automated}, although this choice is empirical.
 
-The work of \cite{Wang20balancing} based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\lambda^{l}$ that will result in the best NMT model regarding a given target dev set at the end of the training, derives also a similar form of update for $\psi$.
-However, their reward is the cosine similarity between the gradient computed with the training data and the gradient computed with the dev set.
+The work of \cite{Wang20balancing} is also related: it is based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\lambda^{l}$ that will result in the best NMT model with respect to a given target dev set at the end of training.
+These authors also derive a similar form of update for $\psi$.
+However, their reward is the cosine similarity between the gradient computed with the training data from one domain and the gradient computed with the dev set.
+We compare this approach with ours in the experiment section.
 
 \subsection{MDAC for (multi) domain adaptation}
-This setting is pretty general and can in principle accommodate the variety of situations mentioned  above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
-In our experiments we would like to better assess the actual potential of MDAC is these settings and seek to experimentally answer the following questions:
+The setting developed in previous sections is quite general and can in principle accommodate the variety of situations mentioned above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
+In our experiments, we would like to better assess the true potential of MDAC is these settings and seek to experimentally answer the following questions:
 \begin{itemize}
-\item is MDAC a viable alternative to conventional fine-tuning? In particular does it enable to better take advantage of relevant data from other domains?
+\item is MDAC a viable alternative to conventional finetuning? In particular, does it enable to better take advantage of relevant data from other domains?
 \item is MDAC also a viable option in multi-domain adaptation scenarios?
@@
 \end{itemize}
-
-These questions are further explored in Section~\ref{sec:results}. We now turn to our experimental conditions.
+These questions are further explored in section~\ref{sec:results}. We now turn to our experimental conditions.
 

### 2021-05-11 2021/emnlp/auto_cl_mdmt.tex
review section 3

@@
 \item is MDAC also a viable option in multi-domain adaptation scenarios?
-\item does MDAC also enable to perform unsupervised (multi-) domain adaptation? \fyDone{TBContinued}
+\item does MDAC also enable to perform \emph{unsupervised} (multi-)domain adaptation? \fyDone{TBContinued}
 \end{itemize}
@@
 \system{Mixed-$0$} uses a uniform domain distribution, \system{Mixed-$1.0$} uses the observed domain distribution.
-\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing.
-The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems each adapted for one domain, with no parameter sharing across domains.
+\item finetuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing.\fyTodo{for ??
+iterations} The full finetuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems each adapted for one domain, with no parameter sharing across domains.
 
-\item Our comparison of multi-domain systems includes baselines with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning of \cite{Zhang19curriculum} that we denote CL and Differential Data Selection of \cite{Wang20balancing} that we denote DDS.
+\item Our comparison of multi-domain systems\fyTodo{CL is only dynamic ?} includes baselines with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning (CL) of \citet{Zhang19curriculum} and Differential Data Selection (DDS) of \citet{Wang20balancing}.
 \end{itemize}
@@
 We will thus report these systems as \system{MDAC($\vlambda^{l}_{t=0}$, $\vlambda^{t}$)}.
-We also compare our method with the method DDS under the same conditions.
+We also compare our method with the DDS method under the same conditions.
 

### 2021-05-11 2021/emnlp/auto_cl_mdmt.tex
review section 3

@@
 We will thus report these systems as \system{MDAC($\vlambda^{l}_{t=0}$, $\vlambda^{t}$)}.
-We also compare our method with the DDS method under the same conditions.
+We also compare our method with DDS under the same settings.
 
-In our work, we simply parameterize the distribution $\lambda^l$ as follows.
+In our work, we simply parameterize the distribution $\lambda^l$ as follows (with $\beta=2$ in all our experiments):\fyTodo{Explain why}
 \begin{equation}
 \lambda
-^l(d;\vpsi) = \frac{\vpsi[d]^\beta}{\sum_i \vpsi[i]^\beta} \nonumber
+^l(d;\vpsi) = \frac{\vpsi[d]^\beta}{\sum_i \vpsi[i]^\beta}. \nonumber
 \end{equation}
-We choose $\beta=2$ for our experiments.
 
-We also provide in this section our choices of the hyper-parameters of our algorithm including the number of simulation training steps $k=10$, and the learning rate $\mathbf{lr}_{data}=0.001$.
-We update the sampling distribution via 100 gradient descent iterations for almost the experimental settings except that for adaptation with automatic cluster \ref{ssec:clda}, we use 20 gradient descent iterations.
-We split the training into 100 short sessions that last 2000 training steps each.
-The choice of those hyper-parameters is mostly heuristic except for the learning rate $\mathbf{lr}_{data}$ which is optimized via grid search over a range of values $[0.001,0.0025,0.005]$
-\subsection{CL and DDS's re-implementation}
-We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to original code of \citet{Wang20balancing} \footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}.
+Additional choices for the hyper-parameters of our method include the number of simulation steps $k=10$ and the learning rate $\mathbf{lr}_{data}=0.001$.
+We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section \ref{ssec:clda}), where we use 20 gradient descent iterations.\fyTodo{Why?
+computation?} We split the training into 100 short sessions that last 2000 training steps each.
+The choice of those hyper-parameters is mostly heuristic except for the learning rate $\mathbf{lr}_{data}$ which is optimized via grid search over a range of values $[0.001,0.0025,0.005]$.
 
-We re-implement the work of \citet{Zhang19curriculum} by following exactly the description of the algorithm.
-For each experiment of adapting the NMT system to one domain, we combine the training data of other domains into one corpus, then compute the cross-entropy difference score of each sentence of this corpus (we only compute the score for the source side).
-We sort and split the corpus into 9 shards and execute curriculum learning with 10 shards in which the first shard is the in-domain corpus.
+\subsection{CL and DDS's re-implementation}\fyTodo{Right place for this ?}
+We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to original code of \citet{Wang20balancing}.\footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}
+
+We also re-implement the work of \citet{Zhang19curriculum} according to the authors' description of the algorithm.
+For each domain adaptation experiment, we combine the training data of all other domains into one corpus, then compute the cross-entropy difference score of each source sentence of this combined dataset.
+We then sort and split the corpus into 9 shards and execute curriculum learning with 10 shards, using the in-domain corpus as the first shard.
+
 \subsection{Experimental tasks}

### 2021-05-11 2021/emnlp/auto_cl_mdmt.tex
reviewed till section 5

@@
 \subsection{Experimental tasks}
-We evaluate the efficacy of our methods compared to several aforementioned baselines via 5 tasks including domain adaptation; multi-domain adaptation, bi-domain adaptation, unseen domain adaptation, and unsupervised domain adaptation.
+We evaluate and compare our method with multiple baselines in 5~tasks including: domain adaptation; multi-domain adaptation, bi-domain adaptation, unseen domain adaptation, and unsupervised domain adaptation.
 
-In the domain adaptation task, given the data of 6 domains including \domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}, we aim to build expert NMT models for each domain respectively.
+In the domain adaptation task, given the data from 6 domains (\domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}), we aim to build distinct expert NMT models for each domain.
 
-In the multi-domain adaptation task, given the data of 6 domains, we aim to build an NMT model which gives the best performance regarding a predefined test set containing equivalently 6 domains.
+In the multi-domain adaptation task, given the same 6 domains, we aim to build one single NMT model that would perform optimally assuming a uniform distribution of domains during the test.
 
-In the bi-domain adaptation task, given the data of 6 domains, we only focus on 2 domains of them.
+In the bi-domain adaptation task, given again same 6 domains, we only focus on adapating only to 2 domains.
 
-In the fourth experiment, given the data of 6 domains and a small dev set of an unseen domain, which is News, in this case, we aim to build an NMT model which is robust to the unseen domain.
+In the fourth experiment, given training data in 6 domains and a small development set of an new domain, (\domain{News} in our case) we aim to build an NMT model which will perform well on test data from the unseen domain.
 
-In the unsupervised domain adaptation task, we cluster the training data of 6 domains into 30 clusters using the KNN algorithm in the same way as in the work of \citet{Tars18multidomain}, then adapt these clusters to one of 6 domains given the corresponding in-domain dev set.
+Finally, in the unsupervised domain adaptation task, we cluster all available training data into 30 clusters using the KNN algorithm in the same way as in the work of \citet{Tars18multidomain}, then adapt these clusters to one of 6 domains given the corresponding in-domain dev set.
 We compare MDAC to DDS over 6 domains.

### 2021-05-11 2021/emnlp/auto_cl_mdmt.tex
correcting spelling errors, adding supplemenatary materials including generalized dynamic domain sampling pseudo alg and automatic clusters statistic table

--- 2021-05-11 2021/emnlp/auto_cl_mdmt.tex
@@
 This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions \citep{Moore10selection,Axelrod11domain}.
-Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data.
-It may also be suboptimal, as some target domains or languages might be easier to train than others.
-For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not achieve optimily for any particular domain.
+Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data; it may also be suboptimal, as some target domains or languages might be easier to train than others.
+For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not promise optimization of a particular domain.
 
-Several recent proposals
-% \citep{Vanderwees17dynamic,Graves17automated,platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Wang20balancing,Wang20learning-multi}
-have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies.
-Notably,  \citet{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while\citet{Graves17automated,platanios19competence,Kumar19reinforcement,Wang20balancing,Wang20learning-multi} build curricula that automatically adapt to the training data.
+Several recent proposals \citep{platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Graves17automated,Wang20balancing,Wang20learning-multi,Vanderwees17dynamic} have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies.
+Furthermore \citet{Zhang19curriculum,Vanderwees17dynamic} construct a static curriculum while \citet{platanios19competence,Kumar19reinforcement,Graves17automated,Wang20balancing,Wang20learning-multi} build curricula automatically adapted to the training.
 
-In this paper, we contribute to this line of research in several ways.
-First, we propose a novel framework (\emph{Multi-Domain Automated Curriculum}, MDAC for short) that simultaneously accounts for the domain adaptation and the multi-domain adaptation problems.
-Second, we propose a novel method of building an automated curriculum for training Multi-Domain NMT models.
-We evaluate our method in various single-domain and multi-domain adaptation settings.
+The major contribution of our work consists of 2 parts.
+First, we propose a novel framework that covers the domain adaptation problem and the multi-domain adaptation problem.
+Secondly, we propose a novel method of building an automated curriculum for training the Multi-Domain NMT model.
+We evaluate our method in various single-domain adaptation and multi-domain adaptation settings.
 We compare our method with several contrasts including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}.
-% From now on, we name our method MDAC which is brief for .
-Based on these experimental results, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard finetuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test situations, from targeting one single domain to the full multi-domain scenario.
-% distributions
+From now on, we name our method MDAC which is brief for Multi-Domain Automated Curriculum.
+Based on experimental results obtained on a diverse set of domains, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard fine-tuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test target distributions.
 
@@
 
-We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \citep{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \citep{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
+We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \citep{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
 Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
@@
 
-It is usual in MT to opportunistically collect training samples from several domains, which means that training instances are distributed according to a mixture $\mathcal{D}^s$ such that $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}(d) \mathcal{D}_d(x)$, with $\{\lambda^{s}(d), d=1 \dots n_d\}$ the mixture weights satisfying $\sum_d \lambda^{s}(d)=1$.
+It is usual in MT to opportunistically collect training samples from several domains, which means that training instances are distributed according to the mixture $\mathcal{D}^s$ such that $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}(d) \mathcal{D}_d(x)$, where $\{\lambda^{s}(d), d=1 \dots n_d\}$ are the corresponding mixture weights satisfying $\sum_d \lambda^{s}(d)=1$.
 
@@
 These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unsupervised domain adaptation); in multi-domain adaptation finally, the test distribution is itself a mixture of domains, some of which may also be observed in training.
-Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of pure domain adaptation.
-These various settings are illustrated Figure~\ref{fig:mdmt-lambdas}.
+Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of fine-tuning.
+These situations are illustrated in Figure~\ref{fig:mdmt-lambdas}.
 \begin{figure}[h]
@@
 
-A more practical standard approach to supervised DA is \emph{finetuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
+A more practical standard approach to supervised DA is \emph{fine-tuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
 With our notations, this approach amounts to first learning an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$) with $\indic{A}$ the indicator function for predicate $A$.
@@
 \subsection{Basic principles}
-Assuming training data in each of the $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
+Assuming training data from $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as  $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
 We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t _d}$ to denote the empirical train and test distributions for domain $d$, and $\widehat{\mathcal{D}^{u}}(x;\lambda^{u}) = \sum_{d} \lambda^{u}(d) \widehat{\mathcal{D}^{u}_d}(x)$ for $u\in\{l,t\}$.
@@
 
-% Our method
-MDAC constructs an adaptative training distribution $\lambda^{l}$ that optimizes the data selection policy along with the training of the NMT model.
+Our method MDAC constructs an automatically adapting distribution $\lambda^{l}$ that optimizes the data selection policy along with the training of the NMT model.
 We parameterize $\lambda^{l}$ by a differentiable function $\lambda^l(\psi)$.
 We divide the training into many short sessions; in each session $t$, the model is trained with a static data distribution $\lambda^{l}(\psi_t)$.
-After one learning session, we update the data distribution using the REINFORCE algorithm of \citet{Williams92simple}.
-The evolution of $\psi$ is thus defined by:
+After one learning session, we update the data distribution via REINFORCE algorithm \citep{Williams92simple}.
+
 \begin{align*}
-\psi_{t+1} &= \psi_t + \mathbf{lr}_{1} * \displaystyle{\mathop{\sum}_{d=1}^{n_d}} R(d) * \frac{\partial \lambda^l(d;\psi_t)}{\partial \psi}, \\
+\psi_{t+1} &= \psi_t + \mathbf{lr}_{1} * \displaystyle{\mathop{\sum}_{d=1}^{n_d}} R(d) * \frac{\partial \lambda^l(d;\psi_t)}{\partial \psi} \\
 \end{align*}
@@
 \allowdisplaybreaks
-where reward $R(d)$ is computed as:
+With the following reward computation:
 \begin{align*}
-  R(d) = J^t(\theta_{t+k},\vlambda^t) - J^t(\theta_t,\vlambda^t),
+  R(d) = J^t(\theta_{t+k},\vlambda^t) - J^t(\theta_t,\vlambda^t)
 \end{align*}
-and where we also define:
+Where 
 \begin{equation}
@@
 \theta_{t+i} &=& Update\big(\theta_{t+i-1},[x^i_j,y^i_j]_{j=1}^N\big) \\ \nonumber
-x^i_j, y^i_j &\sim& \widehat{D^l_d}(x) \\
-J^t(\theta,\lambda_t) &=& \displaystyle{\mathop{\sum}_{d=1}^{n_d}}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta,x^t_d,y^t_d).
+x^i_j, y^i_j &\sim& \widehat{D^l_d}(x) \; \; \forall j \in [1,\cdots,N] \\
+J^t(\theta,\lambda_t) &=& \displaystyle{\mathop{\sum}_{d=1}^{n_d}}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta,x^t_d,y^t_d)
 \end{array}
@@
 \endgroup
-% \theta_{t+i-1} - \mathbf{lr_2} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta}
-In these equations, $N$ denotes the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $\mathbf{lr_2}$ is the learning rate of the NMT system; $l(\theta,x,y)$ is the loss of the NMT model on sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss aggregated over $n_d$ dev-sets corresponding to the $n_d$ domains.
+%\theta_{t+i-1} - \mathbf{lr_2} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta}
+$N$ is the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $\mathbf{lr_2}$ is the learning rate of the NMT system; $l(\theta,x,y)$ is the loss of the NMT model on the sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss computed over $n_d$ dev-sets corresponding to $n_d$ given domains.
 
-To compute the reward $R(d)$ of using data from domain $d$ to train the model, we simulate $k$ training steps from current checkpoint by training the model with $k$ batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
-This computation is inspired by the target prediction gain in \citep{Graves17automated}.
-However, \cite{Graves17automated} used accumulated gains from the past as reward, while we predict the usefulness of each domain for improving the future performance of the system given its current state.
-This is achieved by simulating a round of training with only the data from one domain.
-Furthermore, the choice of the parameterization of the sampling distribution in our method differs from that of \citet{Graves17automated}, although this choice is empirical.
+To compute the reward of using data from domain $d$ to train the model, $R(d)$, we simulate $k$ training steps from the current checkpoint by training the model with k batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
+The reward of our method is inspired by the target prediction gain in \cite{Graves17automated}.
+However, \cite{Graves17automated} used accumulated gains from the past as reward while we predict the utility of each domain for improving the competence of the system at the current state by simulating the training with the data of the domain.
+Furthermore, the choice of the parameterization of the sampling distribution in our method is different from one in \cite{Graves17automated}, although this choice is empirical.
 
-The work of \cite{Wang20balancing} is also related: it is based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\lambda^{l}$ that will result in the best NMT model with respect to a given target dev set at the end of training.
-These authors also derive a similar form of update for $\psi$.
-However, their reward is the cosine similarity between the gradient computed with the training data from one domain and the gradient computed with the dev set.
-We compare this approach with ours in the experiment section.
+The work of \cite{Wang20balancing} based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\lambda^{l}$ that will result in the best NMT model regarding a given target dev set at the end of the training, derives also a similar form of update for $\psi$.
+However, their reward is the cosine similarity between the gradient computed with the training data and the gradient computed with the dev set.
 
 \subsection{MDAC for (multi) domain adaptation}
-The setting developed in previous sections is quite general and can in principle accommodate the variety of situations mentioned above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
-In our experiments, we would like to better assess the true potential of MDAC is these settings and seek to experimentally answer the following questions:
+This setting is pretty general and can in principle accommodate the variety of situations mentioned  above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
+In our experiments we would like to better assess the actual potential of MDAC is these settings and seek to experimentally answer the following questions:
 \begin{itemize}
-\item is MDAC a viable alternative to conventional finetuning? In particular, does it enable to better take advantage of relevant data from other domains?
+\item is MDAC a viable alternative to conventional fine-tuning? In particular does it enable to better take advantage of relevant data from other domains?
 \item is MDAC also a viable option in multi-domain adaptation scenarios?
-\item does MDAC also enable to perform \emph{unsupervised} (multi-)domain adaptation? \fyDone{TBContinued}
+\item does MDAC also enable to perform unsupervised (multi-) domain adaptation? \fyDone{TBContinued}
 \end{itemize}
-These questions are further explored in section~\ref{sec:results}. We now turn to our experimental conditions.
+
+These questions are further explored in Section~\ref{sec:results}. We now turn to our experimental conditions.
 
@@
 \system{Mixed-$0$} uses a uniform domain distribution, \system{Mixed-$1.0$} uses the observed domain distribution.
-\item finetuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing.\fyTodo{for ??
-iterations} The full finetuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems each adapted for one domain, with no parameter sharing across domains.
+\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing.
+The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems each adapted for one domain, with no parameter sharing across domains.
 
-\item Our comparison of multi-domain systems\fyTodo{CL is only dynamic ?} includes baselines with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning (CL) of \citet{Zhang19curriculum} and Differential Data Selection (DDS) of \citet{Wang20balancing}.
+\item Our comparison of multi-domain systems includes baselines with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning of \cite{Zhang19curriculum} that we denote CL and Differential Data Selection of \cite{Wang20balancing} that we denote DDS.
 \end{itemize}
@@
 Transformer models contain 8~attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells.
-The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024.
 Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.
@@
 We will thus report these systems as \system{MDAC($\vlambda^{l}_{t=0}$, $\vlambda^{t}$)}.
-We also compare our method with the DDS method under the same conditions.
+We also compare our method with the method DDS under the same conditions.
 
@@
 We also provide in this section our choices of the hyper-parameters of our algorithm including the number of simulation training steps $k=10$, and the learning rate $\mathbf{lr}_{data}=0.001$.
-We update the sampling distribution via 100 gradient descent iterations for almost the experimental settings except that for adaptation with automatic cluster \ref{ssec:clda}, we use 20 gradient descent iterations.
+We update the sampling distribution via 100 gradient descent iterations for almost the experimental settings except that for adaptation with automatic clusters \ref{ssec:clda}, we use 20 gradient descent iterations.
 We split the training into 100 short sessions that last 2000 training steps each.
@@
     domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} \\ \hline \hline
-    \system{MDAC($\vlambda_1, \vlambda_2$)}&36.2&-&-&-&46.1$^*$&-&-&- \\
-    \system{DDS($\vlambda_1, \vlambda_2$)}&36.5&-&-&-&44.4&-&-&- \\
-    \system{MDAC($\vlambda_1, \vlambda_2$)}&-&57.2$^*$&53.8$^*$&-&-&-&-&- \\
-    \system{DDS($\vlambda_1, \vlambda_2$)}&-&56.1&52.6&-&-&-&-&- \\
-    \system{MDAC($\vlambda_1, \vlambda_2$)}&-&-&-&32&-&90&-&- \\
-    \system{DDS($\vlambda_1, \vlambda_2$)}&-&-&-&32.9&-&91.1&-&- \\ \hline
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&36.2&-&-&-&46.1$^*$&-&-&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&36.5&-&-&-&44.4&-&-&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&57.2$^*$&53.8$^*$&-&-&-&-&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&56.1&52.6&-&-&-&-&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&32&-&90&-&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&32.9&-&91.1&-&- \\ \hline
   \end{tabular}
@@
 Domain adaptation is an ancient and vexing problem that has been studied from many angles both for SMT and NMT.
-A survey of supervised and unsupervised DA for NMT is in \citet{Chu18asurvey}, where they distinguish between data-centric and model-centric DA, a view also adapted in the more recent review of \citet{Saunders21domain}.
+A survey of supervised and unsupervised DA for NMT is in \citet{Chu18asurvey}, where they distinguish between data-centric and model-centric DA, a view also adopted in the more recent review of \citet{Saunders21domain}.
 Our approach to DA in this paper clearly falls under the former category.
-We refer interested readers to these refrences, and do not discuss review DA any further.
+We refer interested readers to these references and do not discuss review DA any further.
 
@@
 Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of \emph{curriculum learning} (CL) is initially introduced.
-CL is relies on a notion of ``easyness'' of a sample to schedule the presentation of the training data so that the easiest examples  presented first and the hardest, last.
-CL has been showed to speed up learning.
-While the initial aim was to improve and speed up training, it has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitons of ``easyness''.
+CL relies on the notion of ``easiness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
+CL has been shown to speed up learning.
+While the initial aim was to improve and speed up training, it has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitions of ``easiness''.
 
 For instance, \citet{Zhang19curriculum} study supervised domain adaptation, and propose a curriculum approach which progressively extends the data used in training: in the early stages only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
-This is somehow opposite to the recommendations of \citet{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyTodo{These have not been compared ?
+This is somehow opposite to the recommendations of \citet{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyTodo{These have not been compared?
 and also to what we do ?}
@@
 \citet{Kumar19reinforcement} use reinforcement techniques (deep Q-learning) to learn the curriculum strategy: in this study, complexity corresponds to difficulty levels and which are binned based on contrastive data selection.
-The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyTodo{Alert: what do we do during warm up ?} The same technique is recently applied to the training of a multilingual NMT system \citep{Kumar21learning}.
+The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyTodo{Alert: what do we do during warm-up ?} The same technique is recently applied to the training of a multilingual NMT system \citep{Kumar21learning}.
 \citet{Zhou20uncertainty} propose another curriculum-based approach which instead relies on \emph{instance uncertainty} as a measure of their difficulty and presents the data sample starting with the easiest (more predictable) first.
@@
 More related to our problems, \citet{Wang20learning-multi} adapt curriculum learning for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization (derivative-free) techniques.
-Each step consists in (a) weighting the instances based on relevance features, (b) fine tuning current model using the (weighted) training set, and is applied iteratively to generate a sequence of models.
+Each step consists of (a) weighting the instances based on relevance features, (b) fine-tuning a pretrained model using the (weighted) training set, and is applied iteratively to generate a sequence of models.
 The one that maximizes the development set performance is finally retained.

### 2021-05-12 2021/emnlp/auto_cl_mdmt.tex
fix some lambda, resolve fixmes

--- 2021-05-11 2021/emnlp/auto_cl_mdmt.tex
@@
 In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key component in training is then to mitigate the detrimental effects of a possible mismatch between these distributions.
-Single-source and multi-source domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
-A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work} \revision{cite also \citep{shen21source}}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
+Single-source and multi-source\footnote{In this context, multi-source refers to simultaneously having multiple domains to adapt from and should not be confused with multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Fix ref}\revision{cite also \citep{Shen21source}}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
 
@@
 This computation is inspired by the target prediction gain in \citep{Graves17automated}.
-However, \cite{Graves17automated} used accumulated gains from the past as reward, while we predict the usefulness of each domain for improving the future performance of the system given its current state.
+However, \citep{Graves17automated} used accumulated gains from the past as reward, while we predict the usefulness of each domain for improving the future performance of the system given its current state.
 This is achieved by simulating a round of training with only the data from one domain.
@@
 
-The work of \cite{Wang20balancing} is also related: it is based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\lambda^{l}$ that will result in the best NMT model with respect to a given target dev set at the end of training.
+The work of \citet{Wang20balancing} is also related: it is based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\lambda^{l}$ that will result in the best NMT model with respect to a given target dev set at the end of training.
 These authors also derive a similar form of update for $\psi$.
@@
 \end{itemize}
-These questions are further explored in section~\ref{sec:results}. We now turn to our experimental conditions.
+These questions are further explored in Section~\ref{sec:results}. We now turn to our experimental conditions.
 
@@
 We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
-We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
+Validation sets are used to chose the best model according to the average BLEU score \citep{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \citep{Neubig19compare-mt}.
+We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value.}\fyTodo{Pourquoi pas sacre BLEU?}
 
@@
 \system{Mixed-$0$} uses a uniform domain distribution, \system{Mixed-$1.0$} uses the observed domain distribution.
-\item finetuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing.\fyTodo{for ??
-iterations} The full finetuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems each adapted for one domain, with no parameter sharing across domains.
+\item finetuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ??
+iterations} The full finetuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems, each adapted for one domain, with no parameter sharing across domains.
 
-\item Our comparison of multi-domain systems\fyTodo{CL is only dynamic ?} includes baselines with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning (CL) of \citet{Zhang19curriculum} and Differential Data Selection (DDS) of \citet{Wang20balancing}.
+\item Our comparison of multi-domain systems\fyTodo{CL is only dynamic ?} includes baselines with fixed data mixtures corresponding to $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning (CL) of \citet{Zhang19curriculum} and Differential Data Selection (DDS) of \citet{Wang20balancing}.
 \end{itemize}
@@
 
-In our work, we simply parameterize the distribution $\lambda^l$ as follows (with $\beta=2$ in all our experiments):\fyTodo{Explain why}
+In our work, we simply parameterize the distribution $\vlambda^l$ as follows (with $\beta=2$ in all our experiments):\fyTodo{Explain why}
 \begin{equation}
@@
 Additional choices for the hyper-parameters of our method include the number of simulation steps $k=10$ and the learning rate $\mathbf{lr}_{data}=0.001$.
-We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section \ref{ssec:clda}), where we use 20 gradient descent iterations.\fyTodo{Why?
+We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section~\ref{ssec:clda}), where we use only 20 gradient descent iterations.\footnote{This is intended to regularize the training of $\psi$ and avoid converging towards degenerate solutions where some clusters get no training samples.}\fyDone{Why?
 computation?} We split the training into 100 short sessions that last 2000 training steps each.
@@
 In the domain adaptation task, given the data from 6 domains (\domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}), we aim to build distinct expert NMT models for each domain.
+In the bi-domain adaptation task, given again same 6 domains, we only focus on adapting only to 2 (known) domains.
 
@@
 
-In the bi-domain adaptation task, given again same 6 domains, we only focus on adapating only to 2 domains.
-
-In the fourth experiment, given training data in 6 domains and a small development set of an new domain, (\domain{News} in our case) we aim to build an NMT model which will perform well on test data from the unseen domain.
+In the fourth experiment (few shot adaptation)\fyTodo{or weakly supervised ?}, given training data in 6 domains and a small development set of a new domain, (\domain{News} in our case), our goal is to build an NMT model which will perform well on test data from the unseen domain.
 
@@
 \subsection{Domain Adaptation}\label{ssec:da}
-In this setting, we aim to build an expert NMT model in only one domain. This setting corresponds to Dirac distribution $\lambda^t$, in which the probability of domains is 0 except for one domain.
+In this setting, we aim to build an expert NMT model in only one domain. This setting corresponds to Dirac distribution $\vlambda^t$, in which the probability of domains is 0 except for one domain.
 
-We initialize MDAC and DDS by 3 different distributions $\lambda_0$, $\lambda_1$ and $\lambda_d$.
-According to Table \ref{tab:da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen, though $\lambda_0$  initialization only improve significantly $\lambda_d$ initialization in small domains including \domain{talk},\domain{bank}and\domain{it}.
-$\lambda_d$ initialization is beneficial for adapting the NMT system to large domains such as \domain{med} and \domain{law}.
-The same conclusion is also true for DDS.
+We initialize MDAC and DDS by 3 different distributions $\vlambda_0$, $\vlambda_1$ and $\vlambda_d$.
+According to Table~\ref{tab:da}, MDAC achieves the best performance with $\vlambda_0$ (uniform domain distribution) as the initial value, though $\vlambda_0$  initialization only significantly improves over $\vlambda_d$ initialization in small domains such as \domain{talk},\domain{bank} and\domain{it}.
+Initializing with $\vlambda_d$ seems more appropriate when targeting large domains such as \domain{med} and \domain{law}.
+The same conclusions also hold for DDS.
 
-We now compare the best MDAC (initialized by $\lambda_0$) to fine-tuning.
-According to Table \ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}.
+We now compare the best MDAC system (initialized with $\vlambda_{t=0} = \vlambda_0$) to full finetuning.
+According to Table~\ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}.
 However, MDAC improves fine-tuning by approximately 0.5 BLEU in \domain{bank} and 1.0 BLEU in \domain{rel}.
@@
 
-DDS also demonstrates promising results in this task but is always outperformed by MDAC overall. The same conclusion is also applied to the CL method.
+DDS also demonstrates promising results in this task, but is always outperformed by MDAC overall. The same conclusions also hold for the CL method.\fyTodo{Continue}
 
@@
     domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline
-    FT-Full($d$) \hfill [BLEU] &37.7&59.2&54.5&34&46.8&90.8&53.8\\ \hline
+    FT-Full($d$) &37.7&59.2&54.5&34&46.8&90.8&53.8\\ \hline
     \hline
@@
   \end{tabular}
-  \caption{We report scores of each method with 6 target domains, and compare with full fine-tuning: each column corresponds to a distinct system.
+  \caption{We report BLEU scores of each method with 6 target domains, and compare with full fine-tuning: each column corresponds to a distinct system.
 ($^{*}$) MDAC is significantly better than CL, fine-tuning and DDS with $p<0.05$.
@@
 Multi-domain adaptation task requires best averaged performance of the NMT model in 6 domains.
-This setting corresponds to the uniform test distribution $\lambda^t = \lambda_0$.
-In this situation, the method CL of \citet{Zhang19curriculum} can not be applied.
-We therefore only compare the performance of MDAC, DDS and several fixed training data distribution $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$.
-The definition of those notations can be found in the equation \ref{mixture:trn}.
+This setting corresponds to a uniform test distribution $\vlambda^t = \vlambda_0$.
+In this situation, the CL method of \citet{Zhang19curriculum} does not apply.
+We therefore only compare the performance of MDAC, DDS and several fixed training data distribution $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ (see equation~\eqref{mixture:trn}).
 
@@
   \centering \small
-  \begin{tabular}{|l|*9{r|}} \hline
-    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} \\ \hline \hline
-    \system{MDAC($\vlambda_1, \vlambda_2$)}&36.2&-&-&-&46.1$^*$&-&-&- \\
-    \system{DDS($\vlambda_1, \vlambda_2$)}&36.5&-&-&-&44.4&-&-&- \\
-    \system{MDAC($\vlambda_1, \vlambda_2$)}&-&57.2$^*$&53.8$^*$&-&-&-&-&- \\
-    \system{DDS($\vlambda_1, \vlambda_2$)}&-&56.1&52.6&-&-&-&-&- \\
-    \system{MDAC($\vlambda_1, \vlambda_2$)}&-&-&-&32&-&90&-&- \\
-    \system{DDS($\vlambda_1, \vlambda_2$)}&-&-&-&32.9&-&91.1&-&- \\ \hline
+  \begin{tabular}{|l|*7{r|}} \hline
+    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} \\ \hline \hline%  \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}}
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&36.2&-&-&-&46.1$^*$&-&\\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&36.5&-&-&-&44.4&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&57.2$^*$&53.8$^*$&-&-&-& \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&56.1&52.6&-&-&-&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&32&-&90 \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&32.9&-&91.1 \\ \hline
   \end{tabular}

### 2021-05-12 2021/emnlp/supplementary_material.tex

--- 2020-10-23 2020/tacl2020/final_multidomain.tex
@@
-% File coling2020.tex
-%
-% Contact: feiliu@cs.ucf.edu & liang.huang.sh@gmail.com
-%% Based on the style files for COLING-2018, which were, in turn,
-%% Based on the style files for COLING-2016, which were, in turn,
-%% Based on the style files for COLING-2014, which were, in turn,
-%% Based on the style files for ACL-2014, which were, in turn,
-%% Based on the style files for ACL-2013, which were, in turn,
-%% Based on the style files for ACL-2012, which were, in turn,
-%% based on the style files for ACL-2011, which were, in turn, 
-%% based on the style files for ACL-2010, which were, in turn, 
-%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
-%% based on the style files for EACL-2009 and IJCNLP-2008...
+% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
+\pdfoutput=1
+% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.
 
-%% Based on the style files for EACL 2006 by 
-%%e.agirre@ehu.es or Sergi.Balari@uab.es
-%% and that of ACL 08 by Joakim Nivre and Noah Smith
+\documentclass[11pt]{article}
 
-\documentclass[11pt,a4paper]{article}
-\usepackage{tacl2018v2}
+% Remove the "review" option to generate the final version.
+\usepackage[review]{emnlp2021}
+
+% Standard package includes
 \usepackage{times}
+\usepackage{latexsym}
+
+% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
+\usepackage[T1]{fontenc}
+% For Vietnamese characters
+% \usepackage[T5]{fontenc}
+% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets
+
+% This assumes your files are encoded as UTF8
+\usepackage[utf8]{inputenc}
+\renewcommand{\UrlFont}{\ttfamily\small}
+\usepackage{graphicx}
+\usepackage{grffile}
+\usepackage{multirow}
+\usepackage{xcolor,colortbl}
+\usepackage{amsmath}
+%\usepackage{fdsymbol}
+\usepackage{amssymb}
+\usepackage{makecell}
+\usepackage[super]{nth}
+\usepackage{arydshln}
+\usepackage{algorithm,algorithmic}
+%\usepackage[noend]{algpseudocode}
+\usepackage{regexpatch}
+\usepackage{subcaption}
+% This is not strictly necessary, and may be commented out,
+% but it will improve the layout of the manuscript,
+% and will typically save some space.
+\usepackage{microtype}
+
 \usepackage{url}
-\usepackage{latexsym}
-\usepackage{graphicx}
-\usepackage{xcolor}
 \usepackage{longtable}
@@
 \usepackage{xspace}
-\usepackage{amsmath,amsfonts,amssymb}
-%\setlength\titlebox{5cm}
-%\colingfinalcopy % Uncomment this line for the final submission
+\usepackage{float}
 
-% You can expand the titlebox if you need extra space
-% to show all the authors. Please do not make the titlebox
-% smaller than 5cm (the original size); we will check this
-% in the camera-ready version and ask you to change it back.
+\usepackage{pgfplots}
+\usepackage{pgfplotstable}
+
+\let\oldbibitem\bibitem
+\def\bibitem{\vfill\oldbibitem}
+
+\usepackage{soulutf8}
+\usepackage{tabularx}
+
+\usepackage{multirow}
+\usepackage{adjustbox}
+\usepackage{caption}
+\usepackage{subcaption}
+
 \newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
@@
 \newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{red}{#1}}}
-\newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{red}{#1}}}
-\newcommand{\revision}[1]{#1}
+
+\newcommand{\revision}[1]{\textcolor{red}{#1}}
 \newcommand{\revisiondel}[1]{}
@@
 \usepackage{multirow}
-\title{Revisiting Multi-Domain Machine Translation}
-\fyFuture{A plea for more thoroughness in evaluation}
+\usepackage{adjustbox}
 
-\author{MinhQuang Pham$^{\dag\ddag}$,\ \  Josep Crego$^{\dag}$,\ \  Fran\c cois Yvon$^{\ddag}$, \ \ Jean Senellart$^{\dag}$ \\ \\
-  $^\ddag$Universit\'e Paris-Saclay, CNRS, LIMSI, 91400, Orsay, France \\
-  \texttt{firstname.lastname@limsi.fr} \\
-  $^\dag$SYSTRAN, 5 rue Feydeau, 75002 Paris, France \\
-  \texttt{firstname.lastname@systrangroup.com} \\
-}
+\title{Supplementary materials}
+
+\author{First Author \\
+  Affiliation / Address line 1 \\
+  Affiliation / Address line 2 \\
+  Affiliation / Address line 3 \\
+  \texttt{email@domain} \\\And
+  Second Author \\
+  Affiliation / Address line 1 \\
+  Affiliation / Address line 2 \\
+  Affiliation / Address line 3 \\
+  \texttt{email@domain} \\}
+
 \date{}
@@
 \maketitle
-\begin{abstract}
-  When building machine translation systems, one often needs to make the best out of heterogeneous sets of parallel data in training, and to robustly handle inputs from unexpected domains in testing.
-This multi-domain scenario has attracted a lot of recent work, that fall under the general umbrella of transfer learning.
-In this study, we revisit multi-domain machine translation, with the aim to formulate the motivations for developing such systems and the associated expectations with respect to performance.
-Our experiments with a large sample of multi-domain systems show that most of these expectations are hardly met and suggest that further work is needed to better analyze the current behaviour of multi-domain systems and to make them fully hold their promises.
-  \fyDone{Write abstract}
-\end{abstract}
+\setlength{\abovedisplayskip}{2pt}
+\setlength{\belowdisplayskip}{2pt}
 
-\section{Introduction} \label{sec:intro}
+\section*{A - Generalized Multi-Domain Adaptation Dynamic Sampling Algorithm}
+We provide below in the pseudo algorithm \ref{alg:mdmt}  the generalized Multi-Domain Adaptation Dynamic Sampling Algorithm.
+\begin{algorithm}[htbp]
+\caption{Multi-Domain Adaptation Dynamic Sampling} \label{alg:mdmt}
+\begin{algorithmic}[1]
+\REQUIRE {
+\begin{itemize}
+	\item K corpora $C^d, d\in [1,..,K]$ for $K$ domains equipped by an empirical distribution $D_d(x)$
+	\item K dev sets $Dev^d, d\in [1,..,K]$ for $K$ domains.
+	\item Domain testing distribution $\lambda^t \in \mathbb{R}^K$
+	\item Batch size $B$
+	\item Domain Dynamic Sampling Distribution $\lambda^l(i) \in \mathbb{R}^K$ where $i$ is the enumeration of the iteration.
+	\item $Eval\_scores = []$
+	\item $Early\_stopping$ criterion
+	\item Total training iterations $iter\_num$
+	\item Updating rule of the sampling distribution $\lambda^l(i)$
+\end{itemize}}
+\REPEAT 
+\STATE{Iteration i.}
+\STATE{Randomly pick $d \in [1,..,K]$ from sampling distribution $\lambda^l(i)$}
+\STATE{Sample $B$ sentences from $C^d$ with empirical distribution $D_d(x)$}
+\STATE{Update model by applying SGD computed from $B$ sampled sentences}
+\IF{$i \equiv 0 \mod{eval\_step}$}
+	\STATE{Evaluate current model with $K$ dev sets. $S^i_d$ is the performance at iteration $i-th$ on domain $d$}
+	\STATE{Report weighted score using the domain testing distribution $\lambda^t$. $$eval(i) = \displaystyle{\mathop{\sum}_d^K \lambda^t(d) S^i_d}$$}
+	\STATE{$Eval\_scores.append(eval(i))$.}
+\ENDIF
+\IF{$i \equiv 0 \mod{sampler\_updating\_step}$}
+	\STATE Update $\lambda^l(i)$.
+\ENDIF
+\IF{$Early\_stopping(Eval\_scores)$}
+	\STATE{Break.}
+\ENDIF
+\UNTIL{$i> iter\_num$}
+\end{algorithmic}
+\end{algorithm}
 
-%
-% The following footnote without marker is needed for the camera-ready
-% version of the paper.
-% Comment out the instructions (first text) and uncomment the 8 lines
-% under "final paper" for your variant of English.
-% 
-% \footnote{
-%     %
-%     % for review submission
-%     %
-%     \hspace{-0.65cm}  % space normally used by the marker
-%     Place licence statement here for the camera-ready version. See
-%     Section~\ref{licence} of the instructions for preparing a
-%     manuscript.
-%     %
-%     % % final paper: en-uk version 
-%     %
-%     % \hspace{-0.65cm}  % space normally used by the marker
-%     % This work is licensed under a Creative Commons 
-%     % Attribution 4.0 International Licence.
-%     % Licence details:
-%     % \url{http://creativecommons.org/licenses/by/4.0/}.
-%     % 
-%     % % final paper: en-us version 
-%     %
-%     % \hspace{-0.65cm}  % space normally used by the marker
-%     % This work is licensed under a Creative Commons 
-%     % Attribution 4.0 International License.
-%     % License details:
-%     % \url{http://creativecommons.org/licenses/by/4.0/}.
-% }
-
-Data-based Machine Translation (MT), whether statistical or neural, rests on well understood machine learning principles.
-Given a training sample of matched source-target sentence pairs $(\src,\trg)$\jcDone{shoulnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}_s$, a model parameterized by $\theta$ (here, a translation function $h_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(h_\theta(\src), \trg)$.
-This approach ensures that the translation loss remains low when translating more sentences drawn from the same distribution.
-
-Owing to the great variability of language data, this ideal situation is rarely met in practice, warranting the study of an alternative scenario, where the test distribution $\mathcal{D}_t$ differs from $\mathcal{D}_s$.
-In this setting, \emph{domain adaptation} (DA) methods are in order.
-DA has a long history in Machine Learning in general (e.g.\ \cite{Shimodaira00improving,Ben-David09atheory,Quinonero08dataset,Pan10asurvey})\fyDone{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}).
-Various techniques thus exist to handle both the situations where a (small) training sample drawn from $\mathcal{D}_t$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
-% In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the
-
-A seemingly related problem is \emph{multi-domain} (MD) machine translation \cite{Sajjad17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system is trained and tested with data from multiple domains.
-MD machine translation (MDMT) corresponds to a very common situation, where all available data, no matter its origin, is used to train a robust system that performs well for any kind of new input.
-% As pointed out both by \newcite{Dredze08online} and \cite{Finkel09hierarchical}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}.
-If the intuitions behind MDMT are quite simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should MDMT perform well when the test data is distributed like the training data, when it is equally distributed across domains or when the test distribution is unknown?
-Should MDMT also be robust to new domains?
-How should it handle domain labeling errors?
-% , which makes their evaluation fragile
-
-A related question concerns the relationship between supervised domain adaptation and multi-domain translation.
-The latter task seems more challenging \revision{as it tries to optimize MT performance for a more diverse set of potential inputs, with an additional uncertainty regarding the distribution of test data.} Are there still situations where MD systems can surpass single domain adaptation, as is sometimes expected?
-
-In this paper, we formulate in a more precise fashion the requirements that an effective MDMT system should meet (Section~\ref{sec:requirements}).
-Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT systems and associated measurements to evaluate them (Section~\ref{sec:challenging}).
-In doing so, we also shed light on new problems that arise in this context, regarding for instance the accommodation of new domains in the course of training, or the computation of automatic domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA.
-We show in Section~\ref{sec:experiments} that existing approaches still fall short to match many of these requirements, notably with respect to the handling of a large amount of heterogeneous domains and to dynamically integrating new domains in training.\fyDone{Spell out conclusions}
-
-\section{Requirements of multi-domain MT \label{sec:requirements}}
-In this section, we recap the main reasons for considering a multi-domain scenario and discuss their implications in terms of performance evaluation.
-\fyDone{Relation to theory, discussion of worst case scenario}
-\fyDone{Insist on the work of Dredze}
-
-\subsection{Formalizing multi-domain translation \label{ssec:formalization}}
-
-We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for source sentences; each domain corresponds to a specific source of data, and differs from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
-Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
-$h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$), but can differ from one domain to the other.
-
-A typical learning scenario in MT is to have access to samples from $n_d$ domains, which means that the training distribution $\mathcal{D}^s$ is a mixture $\mathcal{D}^s(x) = \sum_d \lambda^{s}_{d} \mathcal{D}_d(x)$\revision{, with $\{\lambda^{s}_d, d=1 \dots n_d\}$ the corresponding mixture weights ($\sum_d \lambda^{s}_d=1$)}.
-Multi-domain learning, as defined in \citet{Dredze08online} further assumes that domain tags are also available in testing; the implication being that the test distribution is also as a mixture $\mathcal{D}^t(x) = \sum_d \lambda^{t}_{d} \mathcal{D}_d(x)$ of several domains, making the problem distinct from mere domain adaption.
-A multi-domain learner is then expected to use these tags effectively \cite{Joshi12multidomain} when computing the combined translation function $h(x,d)$, and to perform well in all domains \cite{Finkel09hierarchical}.
-This setting is closely related to the multi-source adaptation problem formalized in \cite{Mansour09domainadaptation,Mansour09multiple,Hoffman18algorithms}.
-
-This definition seems to be the most accepted view of a multi-domain MT\footnote{An exception is \citep{Farajian17multidomain}, where test translations rely on similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here.
-Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function.
-We now discuss the arguments that are put forward to proceed differently.
-% --
-
-% This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyTodo{explain zero shot}.
-
-\subsection{Reasons for building MDMT systems \label{ssec:whymdmt}}
-
-A first motivation for moving away from the one-domain / one-system solution are practical  \cite{Sennrich13multidomain,Farajian17neural}: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop one single system instead of having to optimize and maintain multiple engines.
-The underlying assumption here is that the number of domains of interests can be large, a limiting scenario being fully personalized machine translation \cite{Michel2018extreme}.
-% which is not always the case: for instance \cite{Britz17mixing,Zeng18multidomain,Jiang19multidomain} only consider 2-4 domains in their experiments.\fyTodo{Keep this last part ?}
-
-A second line of reasoning rests on linguistic properties of the translation function and contends that domain specificities are mostly expressed lexically and will primarily affect content words or multi-word expressions; function words, on the other hand, are domain agnostic and tend to remain semantically stable across domains, motivating some cross-domain parameter sharing.
-An MDMT system should simultaneously learn lexical domain peculiarities, and leverage cross-domain similarities to improve the translation of generic contexts and words \cite{Zeng18multidomain,Pham19generic}.
-It is here expected that the MDMT scenario should be more profitable when the domain mix includes domains that are closely related and can share more information.
-
-A third series of motivations are of statistical nature.
-The training data available for each domain is usually unevenly distributed, and domain-specific systems trained or adapted on small datasets are likely to have a high variance and generalize poorly.
-For some test domains, there may even be no data at all \cite{Farajian17neural}.
-Training mix-domain systems is likely to reduce this variance, at the expense of a larger statistical bias \cite{Clark12onesystem}.
-Under this view, MDMT would be especially beneficial for domains with little training data.
-This is observed for multilingual MT from English: an improvement for under-resourced languages due to positive transfer, at the cost of a decrease in performance for well-resourced languages \cite{Arivazhagan19massively}.
-
-Combining multiple domain-specific MTs can also be justified in the sake of distributional robustness \cite{Mansour09domainadaptation,Mansour09multiple}, for instance when the test mixture differs from the train mixture, or when it includes new domains unseen in training.
-\revisiondel{Such scenario is already well documented for zero-shot multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Platanios18contextual} where mixing languages has more than demonstrated its usefulness.}
-An even more challenging case is when the MT would need to perform well for any test distribution, as studied for statistical MT in \cite{Huck15mixeddomain}.
-In all these cases, mixing domains in training and/or testing is likely to improve robustness against unexpected or adversarial test distribution \cite{Oren19distributionally}.
-
-A distinct line of reasoning is that mixing domains can have a positive regularization effect for all domains.
-By introducing variability in training, it prevents DA from overfitting the available adaptation data and could help improve generalization even for well-resourced domains.
-A related case is made in \cite{Joshi12multidomain}, which shows that part of the benefits of MD training is due to an ensembling effect, where systems from multiple domains are simultaneously used in the prediction phase; this effect may subsist even in the absence of clear domain separations.
-
-\fyDone{recap}
-To recap, there are multiple arguments for adopting MDMT, some already used in DA settings, and some original.
-These arguments are not mutually exclusive; however each yields specific expectations with respect to the performance of this approach, and should also yield appropriate evaluation procedure.
-If the motivation is primarily computational, then a drop in MT quality with respect to multiple individual domains might be acceptable if compensated by the computational savings.
-If it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some under-resourced domains, over individually trained systems.
-If finally, it is to make the system more robust to unexpected or adversarial test distributions, then this is the setting that should be used to evaluate MDMT.
-The next section discusses ways in which these requirements of MDMT systems could be challenged.
-
-% \cite[eq. (3)]{Wang20general} defines the loss of MDMT as the (macro) average of domain loss - all domains are equally important.
-\section{Challenging multi-domain systems \label{sec:challenging}}
-In this section, we propose seven\fyDone{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.
-All these evaluations will rest on comparison of translation performance, and do not depend on the choice of a particular metric.
-To make our results comparable with the literature, we will only use the BLEU score \cite{Papineni02bleu} in Section~\ref{sec:experiments}, noting it may not be the best yardstick to assess subtle improvements of lexical choices that are often associated with domain adapted systems \cite{Irvine13measuring}.
-Other important figures of merit for MDMT systems are the computational training cost and the total number of parameters.\fyDone{Insert discussion about scores}
-
-\subsection{Multi-domain systems should be effective \label{ssec:effective}}
-% \subsubsection{Performance}
-A first expectation is that MDMT systems should perform well in the face of mixed-domain test data. We thus derive the following requirements.
-
-\paragraph{[P1-LAB]}\fyDone{Decide naming scheme}A MDMT should perform better than the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}.
-Evaluating this requirement is a matter of a mere comparison, assuming the test distribution of domains is known: if all domains are equally important, performance averages can be reported; if they are not, weighted averages should be used instead.\fyFuture{Try random label assignments ?} % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.
-
-\paragraph{[P2-TUN]} Additionally, one can expect that MDMT will improve over fine-tuning \cite{Luong15stanford,Freitag16fast}, at least in domains where data is scarce, or in situations where several domains are close.
-To evaluate this, we perform two measurements, using a real as well as an artificial scenario.
-In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains.
-In the artificial scenario, we split a single domain in two parts which are considered as distinct in training.
-The expectation here is that a MDMT should yield a clear gain for both pseudo sub-domains, which should benefit from the supplementary amount of relevant training.
-In this situation, MDMT should even outperform fine-tuning on either of the pseudo sub-domain.
-\jcDone{I dont understand the prediction... why running MDMT on 2 artificial subdomains should show  gains over fine-tunning?}\fyDone{small loss with respect to non split ?}
-
-\subsection{Robustness to fuzzy domain separation \label{ssec:robusness}}
-A second set of requirements is related to the definition of a domain.
-As repeatedly pointed out in the literature, parallel corpora in MT are often collected opportunistically and the view that each corpus constitutes a single domain is often a gross approximation.\footnote{Two of our own ``domains'' actually comprise several subcorpora (IT and MED), see details in Section~\ref{ssec:corpora}.} MDMT should aim to make the best of the available data and be robust to domain assignments.
-To challenge these requirements we propose evaluating the following requirements.
-
-\paragraph{[P3-HET]}
-The notion of a domain being a fragile one, an effective MDMT system should be able to discover not only when cross-domain sharing is useful (cf.\ requirement [P2-TUN])\jcDone{what is P2.2?}, but also when intra-domain heterogeneity is hurting.
-This requirement is tested by artificially conjoining separate domains into one during training, hoping that the loss in performance with respect to the baseline (using correct domain tags) will remain small.
-
-\paragraph{[P4-ERR]}
-MDMTs should perform best when the true domain tag is known, but deteriorate gracefully in the face of tag errors; in this situation, catastrophic drops in performance are often observed\fyDone{cite \cite{McCloskey89catastrophic}}.
-This requirement can be assessed by translating test texts with erroneous domain tags and reporting the subsequent loss in performance.
-
-\paragraph{[P5-UNK]}
-A related situation occurs when the domain of a test document is unknown.
-Several situations need be considered: for domains seen in training, using automatically predicted domain labels should not be much worse than using the correct one.
-For test documents from unknown domains (zero-shot transfer), a good MD system should ideally outperform the default baseline that merges all available data.\fyDone{Systems react to unknown domains}
-% These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.
-
-\paragraph{[P6-DYN]}
-Another requirement, more of an operational nature, is that an MDMT system should smoothly evolve to handle a growing number of domains, without having to retrain the full system each time new data is available.
-This is a requirement [P6-DYN] that we challenge by dynamically changing the number of training and test domains.
-
-\subsection{Scaling to a large number of domains \label{ssec:scaling}}
-
-\paragraph{[P7-NUM]} As mentioned above, MDMT systems have often been motivated by computational arguments.
-This argument is all the more sensible as the number of domains increases, making the optimization of many individual systems both ineffective and undesirable.
-For lack of having access to corpora containing very large sets (eg.\ in the order of 100-1000) domains, we experiment with automatically learned domains.\fyFuture{considering a varying number of clusters.}
-
-\section{Experimental settings \label{sec:experiments}}
-
-\subsection{Data and metrics \label{ssec:corpora}}
-
-We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.
-\revision{We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
-Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
-Most corpora are available from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
-To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
-
-We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
-We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
-
-%for contrast experiments, we also use supplementary test sets from three other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}.
-This enables us to evaluate the loss in performance when the test set is from a domain not seen in training.
-% The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
-
+\section*{B - Experiments with automatic domains}
 \begin{table*}[htbp]
-  \centering
-  \begin{tabular}{|l|ccccccc|} %*{4}{|r|}}
-    \cline{2-8} 
-    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
-    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} & \multicolumn{1}{c|}{\domain{news}} \\
-    \hline 
-    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
-    \# \revision{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 & 7.8 / 9.2   \\
-    \# \revision{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 & - \\
-    \# \revision{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 & - \\
-    \hline
-  \end{tabular}
-  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$).
-\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
-  }
-\label{tab:Corpora}
-\end{table*}
-
-We measure the distance between domains using the $\mathcal{H}$-Divergence \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains.
-%Formally, given test sets of size $m$ for domains $A$ and $B$, and $h(x)$ a trained domain predictor, $\mathcal{H}(A,B)$ is computed %as:
-%$$
-%\mathcal{H}(A,B) = 2(1 - [\frac{1}{m} \sum_{x:h(x) = B} \mathbb{I}( x \in A) + \frac{1}{m} \sum_{x: h(x) = A} \mathbb{I}(x \in B)]),
-% $$
-% where $\mathbb{I}$ is the indicator function.
-Our discriminator is a SVM independently trained for each pair of domains, with sentence representations derived via mean pooling from the source side representation of the generic Transformer model.
-We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well separated from all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.
-
-\begin{table}\centering
-  \begin{tabular}{|l*{5}{|r}|} 
-  \cline{2-6}
-  \multicolumn{1}{c|}{} & \domain{law} & \domain{bank} & \domain{talk} & \domain{IT} & \domain{rel} \\ \hline
-    \domain{med} &1.93 &1.97 &1.9 &1.93 &1.97 \\
-    \domain{law}   && 1.94 & 1.97 &1.93 & 1.99 \\
-    \domain{bank} &&&1.98 &1.94 &1.99 \\
-    \domain{talk}   &&&&1.92 &1.93 \\
-     \domain{IT}     &&&&& 1.99 \\ \hline
-  \end{tabular}
-  \caption{The $\mathcal{H}$-divergence between domains}
-  \label{tab:domaindist}
-\end{table}
-
-\subsection{Baselines \label{ssec:baselines}}
-
-Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
-
-\begin{itemize}
-\item a generic model trained on a concatenation of all corpora (\texttt{Mixed}).
-We develop two versions\footnote{In fact three: to enable a fair comparison with WDCMT, a RNN-based variant is also trained and evaluated.
-\revision{This system appears as \system{Mixed-Nat-RNN} in Table~\ref{tab:performance}}.} of this system, one where the domain unbalance reflects the distribution of our training data \revision{given in Table~\ref{tab:Corpora}} (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}).
-The former is the best option when the train mixture $\mathcal{D}^s$ is also expected in testing; the latter should be used when the test distribution is uniform across domains.
-Accordingly, we report two aggregate scores: a weighted average reflecting the training distribution, and an unweighted average, meaning that test domains are equally important.
-%   As explained above, depending whether MDMT are expected to perform specially well on large domains, or to treat all domains equally, one or the other training distribution should be prefered.\jcDone{All systems cand be learned following both Bal and Nat.
-does Nat corresponds to wAVG in test?
-is wAVG defined somewhere?}
-\item fine-tuned models \cite{Luong15stanford,Freitag16fast}\jcDone{i would use Luong and Manning 2015}, based on the \system{Mixed-Nat} system, further trained on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing.
-The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems adapted for one domain, with no parameter sharing across domains.
-\revisiondel{We again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a small adaptation module that is added (with residualconnections) on top of every Transformer layer (\system{FT-Res}).}
-\end{itemize}
-
-All models use embeddings and the hidden layers sizes of dimension~512.
-Transformers contain with 8 attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells.
-The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024.
-% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
-% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
-Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.\fyDone{Describe the block adaptation layer - voir slides}
-% All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
-
-\subsection{Multi-domain systems \label{ssec:systems}}
-\fyDone{Remove Gated residual}
-Our comparison of multi-domain systems includes our own reimplementations of recent proposals from the literature:\footnote{\revision{Further implementation details are in Appendix~A.}}
-\begin{itemize}
-\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or as a supplementary feature for each word (\system{DC-Feat}).
-\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDR});\fyDone{why not LDR as in Minh's paper?}
-\item the three proposals of \newcite{Britz17mixing}.
-\system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}.
-Training uses reference tags and inference is usually performed with predicted tags, just like for regular target words.
-\system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations.
-These methods thus only use domain tags in training.
-\item the multi-domain model of \newcite{Zeng18multidomain} (\system{WDCMT}), where a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification and adversarial training are used to compute these representations.
-\revision{Again, inference does not use domain tags.}\footnote{For this system, we use the available RNN-based system from the authors (\url{https://github.com/DeepLearnXMU/WDCNMT}\fyDone{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slightly improved, results.}\fyDone{Check this}
-\item \revision{two multi-domain versions of the approach of \newcite{Bapna19simple}, denoted \system{FT-Res} and \system{MDL-Res}, where a domain-specific adaptation module is added to all the Transformer layers; within each layer, residual connections enable to short-cut this adapter.
-The former variant corresponds to the original proposal of \citet{Bapna19simple} \revision{(see also \cite{Sharaf20metalearning})}.
-It fine-tunes the adapter modules of a \system{Mixed-Nat} system independently for each domain, keeping all the other parameters frozen.
-The latter uses the same architecture, but a different training procedure and learns all parameters jointly from scratch with a mix-domain corpus.}
-  % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
-% the entire multi-domain system from scratch.\fyDone{Check this.}
-\end{itemize}
-\revision{This list includes systems that slightly depart from our definition of MDMT: standard implementations of \system{TTM} and \system{WDCMT} rely on infered, rather than on gold, domain tags - which must somewhat affect their predictions; \system{DM} and \system{ADM} make no use of domain tags at all.} We however did not consider the proposal of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence and diverges more strongly from our notion of MDMT.
-% and should be able to handle available domain labels in training and testing.
-% This system has shown to meet several of requirements and would also have no problem scaling to a large number of domains.\fyDone{TBC}
-
-\section{Results and discussion \label{sec:results}}
-
-\subsection{Performance of MDMT systems \label{ssec:rawperformance}}
-
-\begin{table*}
-  \centering
-  \fyDone{Fix column size}
-  \begin{tabular}{|p{4cm}|*{8}{r|}} \hline
-%     &&&&&& \\
-    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-    \system{Mixed-Nat}  \hfill{\footnotesize[65m]} & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
-    \system{Mixed-Bal}   \hfill{\footnotesize[65m]} &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 \\ %& \\
-    \system{FT-Full}       \hfill{\footnotesize[6$\times$65m]} & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & \SB{42.7} & \SB{53.8} \\ \hline
- %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag} \hfill{\footnotesize[+4k]}        & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.6 & 50.1    \\%    & 21.8 \\
-    \system{DC-Feat} \hfill{\footnotesize[+140k]}    & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   \\% & \SW{21.7} \\
-    \system{LDR}       \hfill{\footnotesize[+1.4m]}    & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} &40.9 & 49.8          \\% & 22.1 \\ 
-    \system{TTM}      \hfill{\footnotesize[+4k]}        & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.0 & 49.7     \\% &  23.4 \\
-    \system{DM}        \hfill{\footnotesize[+0]}         & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 38.1 & 43.4 \\ % & 22.6\\
-    \system{ADM}      \hfill{\footnotesize[+0]}         & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 38.9 & 47.5 \\% & 23.3 \\
-    \revision{\system{FT-Res}}   \hfill{\footnotesize[+12.4m]}  & 37.3 & \SB{57.9} & \SB{53.9} & 33.8 & \SB{46.7} & \SB{90.2}  & \SB{42.3} & \SB{53.3} \\ % & 20.5\\ \hline
-    \system{MDL-Res} \hfill{\footnotesize[+12.4m]}    & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.0 & \SB{51.9} \\%  & \SW{21.2} \\
-%     \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
-%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3&\\
-     \hline \hline
-    \system{Mixed-Nat-RNN} \hfill{\footnotesize[51m]}  & 36.8 & 53.8 & 47.2 & 30.0 & 35.7 & 60.2  & 39.2  & 44.0 \\
-    \hline
-    \system{WDCMT}  \hfill{\footnotesize[73m]} & 36.0 & 53.3 & \SB{48.8} & 31.1 & \SB{38.8} & \SW{58.5} & 39.0 & 44.4 \\ % & 20.4 \\
-    \hline
-  \end{tabular}
-  \caption{Translation performance of MDMT systems \revision{based on the same Transformer (top) or RNN (bottom) architecture.
-The former contains 65m parameters, the latter has 51m.
-For each system, we report the number of additional domain specific parameters,}  BLEU scores for each domain, domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages.
-\revision{For weighted-averages, we take the domain proportions from Table~\ref{tab:Corpora}}.
-Boldface denotes significant gains with respect to \system{Mix-Nat} (or \system{Mix-Nat-RNN}, for WDCMT), underline denotes significant losses.}
-  \label{tab:performance}
-  % \fyDone{Fill the table with all results, including a comparison with *RNNs* for wcmd}
-  \fyDone{Do we have significancy tests for averages?} %\fyTodo{Add tests wrt to RNN}
-\end{table*}
-
-In this section, we discuss the basic performance of MDMT systems trained and tested on $6$~domains.
-Results are in Table~\ref{tab:performance}.
-As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points\fyDone{BP?} better, notably owing to the much better results for \domain{rel}).
-As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains.
-As all other systems are trained with an unbalanced data distribution, we use the weighted average to perform global comparisons.
-
-Fine-tuning each domain separately yields a better baseline, outperforming \system{Mixed-Nat} for all domains, with significant gains for domains that are distant from \domain{med}: \domain{rel}, \domain{it}, \domain{bank}, \domain{law}.
-% A complete fine-tuning, updating all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to keeping fixed the capacity of the adapter module.
-\fyDone{Add 2 averages ?}\fyDone{Significance testing wrt Mix Generic, Full Fine-tuned, for each domain}
-
-All MDMTs (except \system{DM} and \system{ADM}) slightly improve over \system{Mixed-Nat}(for most domains), but these gains are rarely significant.
-\revision{Among systems using an extra domain feature, \system{DC-Tag} has a small edge over \system{DC-Feat} and also requires less parameters; it also outperforms \system{TTM}, which however uses predicted rather than gold domain tags.
-\system{TTM} is also the best choice among the systems that do not use domain tags in inference.} \revision{The best contenders overall are  \system{FT-Res} and \system{MDL-Res}, which significantly improve over \system{Mixed-Nat} for a majority of domains, and are the only ones to clearly fulfill [P1-LAB];} \system{WDCMT} also improves on three domains, but regresses on one.
-The use of a dedicated adaptation module thus seems better than feature-based strategies, \revision{but yields a large increase of the number of parameters.} The effect of the adaptation layer is especially significant for small domains (\domain{bank}, \domain{it} and \domain{rel}).
-
-% for which unplugging this layer (see line [MDL res (gen)] in Table~\ref{tab:performance}) causes dramatic drop in performance.\fyDone{why LDC does not outperform DC as in Minh's paper (for en2fr MED) ?}
-
-All systems fail to outperform fine-tuning, sometimes by a wide margin, especially for an ``isolated'' domain like \domain{rel}.
-This might be due to the fact that domains are well separated (cf.\ Section~\ref{ssec:corpora}) and are hardly helping each other.
-In this situation, MDMT systems should dedicate a sufficient number of parameters to each domain, so as to close the gap with fine-tuning.
-\fyDone{More comments when we have all the results}\fyDone{Importance of sharing and unsharing}
-
-\subsection{Redefining domains \label{ssec:redomains}}
-
-Table~\ref{tab:redomains} summarizes the results of four experiments where we artificially redefine the boundaries of domains, with the aim to challenge requirements [P2-TUN], [P3-HET], and [P4-ERR].
-In first three, we randomly \emph{split} one corpus in two parts and proceed as if this corresponded to two actual domains.
-A MD system should detect that these two pseudo-domains are mutually beneficial and should hardly be affected by this change with respect to the baseline scenario (no split).
-In this situation, we expect MDMT to even surpass fine-tuning separately on each of these dummy domains, as MDMT exploits all data, while fine-tuning focuses only on a subpart.
-In testing, we decode the test set twice, once with each pseudo-domain tag.
-This makes no difference for \system{TTM}, \system{DM}, \system{ADM} and \system{WDCMT}, which do not use domain tags in testing.
-In the \textsl{merge} experiment, we merge two corpora in training, in order to assess the robustness with respect to heterogenous domains [P3-HET].
-We then translate the two corresponding tests with the same (merged) system.
-
-\begin{table*}
-  \centering% \small
-  \begin{tabular}{|p{1.8cm}|*{10}{r|}} \hline
-    % &&&&&& \\
-    \hfill Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & \multicolumn{2}{c|}{Wrong} \\ % \hline
-     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{law} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} &  \multicolumn{1}{c|}{\domain{rnd}} &  \multicolumn{1}{c|}{\domain{new}}\\ \hline
-    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_2$ &  \domain{law}$_1$ & \domain{law}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
-    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5} & -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& \SW{-19.6} & \SW{-3.3}\\% [32.5]
-    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}  & +0.2& -0.4 & -0.4 & -0.5 & -0.4 & \SW{-13.4} & \SW{-1.7}\\% [35.9]
-    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3}   & +0.3 & +0.3 & +0.3 & +0.3 & +0.1 & \SW{-14.2} &\SW{-1.8}\\ % [34.9] 
-    \system{LDR}           & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& \SW{-12.0} & \SW{-1.4}\\ % [37.0]
-    \system{TTM} (*)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & 0.0 & -0.1\\
-%    \system{MDL Gated} & & & & & & & & &  \\
-    \system{DM} (*)           & -0.3   & -0.3  & +0.4 & +0.4 & +0.3 & +0.3 & +0.9 & +0.1 & 0.0 &-0.9\\
-    \system{ADM} (*)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & 0.0&-0.2\\
-    \revision{\system{FT-Res}}   & -0.1   & -0.4 & -0.3 &-0.3 & \SW{-2.2} & \SW{-2.9} & \SW{-2.4} & -\SW{3.2} & \SW{-13.3} & \SW{-3.0}\\ % [31.8]
-    \system{MDL-Res}   & -0.2   & -0.1 & +\SB{0.2} &+0.0 & -0.9 & -0.9 & +0.7 & -0.3 & \SW{-18.6} & \SW{-1.3}\\ % [31.8]
-    \system{WDCMT} (*)     & -0.0    & -0.0  & +0.2 & +0.2  & +0.8 & +0.8  & -0.4 & -0.8 & 0.0 & +0.2 \\
-    \hline
-  \end{tabular}
-  \caption{Translation performance with variable domain definitions.
-In the Split/Merge experiments, we report BLEU differences for the related test set(s).
-Underline denotes significant loss when domains are changed wrt.\ the baseline situation; bold for a significant improvement over \system{FT-Full}; (*) tags systems ignoring test domains.
-    \fyDone{Check wrong and significance of diff for ``old''}\fyDone{How about wrong ?}\jcDone{FTres results are odd... split LAW 0.5}
-  }
-  \label{tab:redomains}
-\end{table*}
-
-Our findings can be summarized as follows.
-For the split experiments, we see small variations that can be positive or negative compared to the baseline situation, but these are hardly significant.
-All systems show some robustness with respect to fuzzy domain boundaries; this is mostly notable for \system{ADM}, suggesting that when domain are close, ignoring domain differences is effective.
-In contrary, \system{FT-Full} incurs clear losses across the board, especially for the small data condition \cite{Miceli-barone17regularization}.
-Even in this very favourable case however, very few MDMT systems are able to significantly outperform \system{FT-Full} and this is only observed for the smaller part of the \domain{med} domain.
-The merge condition is hardly different, with again large losses for \system{FT-full} \revision{and \system{FT-Res}}, and small variations for all systems.
-We even observe some rare improvements with respect to the situation where we use actual domains.\fyDone{More comments ?}
-
-\subsubsection{Handling wrong or unknown domains \label{sssec:unknowns}}
-
-\fyDone{Numbers for the rand domain analysis}In the last two columns of Table~\ref{tab:redomains}, we report the drop in performance when the domain information is not correct.
-In the first (\domain{rnd}), we use test data from the domains seen in training, presented with a random domain tag\fyDone{all configs use the same tag or is randomization is run for each config}.
-In this situation, the loss with respect to using the correct tag is generally large (more than 10 BLEU points), showing an overall failure to meet requirement [P4-ERR], except for systems that ignore domain tags in testing.
-\fyDone{Comments?}
-
-In the second (\domain{new}), we assess [P5-UNK] by translating sentences from a domain unseen in training (\domain{news}).\fyDone{i would use RND / UNK rather than OLD / NEW} For each sentence, we automatically predict the domain tag and use it for decoding.\footnote{\revision{Domain tags are assigned as follows: we train a language model for each domain and assign tag on a per-sentence basis based on the language model log-probability (assuming uniform domain priors).
-The domain classifier has an average prediction error of 16.4\% for in-domain data.}}
-% Results are in the rightmost column of Table~\ref{tab:redomains}.
-In this configuration, again, systems using domain tags during inference perform poorly, significantly worse than the \system{Mixed-Nat} baseline (Bleu=23.5)\fyDone{performance drop is now measured against Mixed-nat (leave it clearer?)}, sometimes by a large margin.\fyDone{Check significance, To be completed}
-
-\subsubsection{Handling growing numbers of domains}
-Another set of experiments evaluate the ability to dynamically handle supplementary domains (requirement [P6-DYN]) as follows.
-Starting with the existing MD systems of Section~\ref{ssec:rawperformance}, we introduce an extra domain (\domain{News}) and resume training with this new mixture of data\footnote{The design of a proper balance between domains in training is critical for achieving optimal performance: as our goal is to evaluate all systems in the same conditions, we consider a basic mixing policy based on the new training distribution.
-This is detrimental to the small domains, for which the ``negative transfer'' effect is stronger than for larger domains.} for 50,000 additional iterations.\fyDone{Training regime of continuation} We contrast this approach with training all systems from scratch and report differences in performance in Table~\ref{tab:warmrestart}.\footnote{WDCMT results are excluded from this table, as resuming training proved difficult to implement.}\fyDone{Ecrire un commentaire}
-We expect that MDMT systems should not be too significantly impacted by the addition of a new domain and reach about the same performance as when training with this domain from scratch.
-From a practical viewpoint, dynamically integrating new domains is straightforward for \system{DC-Tag}, \system{DC-Feat} or \system{TTM}, for which new domains merely add new labels.
-It is less easy for \system{DM}, \system{ADM} and \system{WDCMT}, which include a built-in domain classifier whose outputs have to be pre-specified, or for \system{LDR}, \system{FT-Res} and \system{MDL-Res} for which the number of possible domains is built in the architecture and has to be anticipated from the start.
-This makes a difference between domain-bounded systems, for which the number of domains is limited and truly open-domain systems.% , which seamlessly integrate additional domains.
-
-\begin{figure*}[h!]
-    \begin{center}
-        \input{table5.pgf}
-    \end{center}
-    \caption{Ability to handle a new domain.
-\revision{We report BLEU scores for a complete training session with 7 domains, as well as differences with (bleu) training with 6 domains (from Table~\ref{tab:performance}); (red) continuous training mode}.}
-\end{figure*}
-
-\begin{table*}
-  \centering
-  \begin{tabular}{|p{1.6cm}|*{9}{r|}} \hline
-    \footnotesize
-   \hfill Domain  Model \hfill & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{news}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \footnotesize\system{Mixed-Nat}  & 37.1  & 54.1  & 49.6	& 34.1  & 42.1	& 77.0 & 28.9 & 40.8	& 49.0 \\[-2pt]
-                   & \sbcl{+0.2}{--}  &\sbcl{+0.5}{---} &\sbcl{+0.5}{--} &\sbcl{-0.6}{--} &\sbcl{+1.1}{--} &\sbcl{+0.5}{--} &\sbcl{-5.4}{--} &\sbcl{+0.3}{--} & \sbcl{+0.4}{--} \\
-%    \system{Mixed-Nat}  & 37.3  & 54.6   & 50.1   & 33.5  & 43.2  & 77.5  & 23.5 & 41.1 & 49.4 \\
-%    \system{FT-Full}       & 37.7  & 59.2   & 54.5   & 34.0  & 46.8  & 90.8  & **** & 42.8 & 53.8 \\
-    \hline%\hline
-    \footnotesize \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5  & 41.4 & 49.9\\[-2pt]
-%                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
-                   & \sbcl{+0.3}{+0.3}  & \sbcl{\SB{+0.8}}{-0.1}   & \sbcl{-0.04}{-0.6}  & \sbcl{\SW{-1.6}}{\SW{-1.1}}  &  \sbcl{-0.4}{\SW{-1.3}}  & \sbcl{\SB{+1.7}}{\SW{-3.5}}  & \sbcl{\SW{-7.7}}{\SW{-1.4}}  & \sbcl{+0.2}{-0.1} & \sbcl{+0.1}{\SW{-1.1}}\\
-
-    \footnotesize \system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9  & 41.2 & 50.1\\[-2pt]
-%                   & 37.2 & 54.7   & 49.9    & 34.2  &  43.0  & 79.9 & \SW{28.1}  & &\\
-                   &  \sbcl{+0.3}{-0.2} & \sbcl{-0.1}{-0.1}  & \sbcl{-0.3}{-0.1} & \sbcl{\SW{-1.3}}{-0.6} & \sbcl{-0.1}{-0.9} & \scriptsize \sbcl{+0.4}{+0.3} & \sbcl{\SW{-7.3}}{\SW{-0.8}} & \sbcl{+0.1}{-0.2} & \sbcl{-0.2}{-0.3}\\
-
-    \footnotesize \system{LDR}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
-                    & \sbcl{0.0}{-0.6} &  \sbcl{+0.1}{+0.5} & \sbcl{+0.2}{-0.4} &  \sbcl{-0.4}{-0.6} &  \sbcl{+0.5}{+0.5} & \sbcl{\SB{+2.9}}{+\SB{3.8}} & \sbcl{\SW{-6.6}}{\SW{-0.9}} & \sbcl{+0.6}{+0.5} &  \sbcl{+0.1}{-0.4} \\
-
-    \footnotesize \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.2  & 29.1 & 41.0 & 49.4\\[-2pt]
-%                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
-                    & \sbcl{0.0}{-0.3}  & \sbcl{+0.4}{-0.3}  & \sbcl{-0.1}{-0.5}  & \sbcl{-0.9}{\SW{-1.1}}  & \sbcl{+0.6}{\SW{-1.0}}  & \sbcl{\SB{+1.8}}{\SW{-4.0}} & \sbcl{\SW{-5.7}}{\SW{-1.4}} & \sbcl{0.0}{-0.5} & \sbcl{+0.3}{\SW{-1.2}}\\
-
-    \footnotesize \system{DM}            &36.0 &51.3&46.8&31.8&39.8&65.7&27.0 & 38.9 & 45.2\\[-2pt]
-%                   &36.5&51.7&47.4&31.7&40.3&65.7&\SW{25.8}& & \\
-                   & \sbcl{-0.4}{+0.6} & \sbcl{\SW{-1.8}}{+0.4} & \sbcl{\SW{-1.2}}{+0.6} & \sbcl{\SW{-1.8}}{-0.1} & \sbcl{\SW{-2.6}}{+0.5} & \sbcl{\SW{-3.3}}{0.0} & \sbcl{\SW{-4.4}}{\SW{-1.2}} & \sbcl{-0.8}{+0.5} & \sbcl{\SW{-1.8}}{+0.3}\\
-
-    \footnotesize \system{ADM}          &36.6&54.2&49.1&32.9&42.1&75.7&28.7 & 40.2 & 48.4 \\[-2pt]
-%                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
-                   & \sbcl{-0.2}{+0.3} & \sbcl{-0.7}{-0.8} & \sbcl{-0.8}{-0.8} & \sbcl{-0.9}{-0.2} & \sbcl{-0.5}{-0.4} & \sbcl{\SW{-2.3}}{\SW{-5.0}} & \sbcl{\SW{-5.4}}{\SW{-1.9}} & \sbcl{-0.5}{-0.2} & \sbcl{\SW{-0.9}}{\SW{-1.1}}\\
-
-%    \system{WDCMT}     &&&&&&& & &\\[-2pt]
-%                    &&&&&&&&& \\
-    \footnotesize \revision{\system{FT-Res}}  & 37.0 & 57.6 & 53.8 & 34.5 &	46.1 & 91.1 & 29.6 &	42.2  & 53.3 \\[-2pt]
-                               & \sbcl{+0.3}{+0.3} & \sbcl{+0.4}{+0.4} & \sbcl{+0.1}{+0.1} & \sbcl{-0.7}{-0.7} & \sbcl{+0.5}{+0.5} & \sbcl{-0.9}{-0.9} & \sbcl{\SW{-9.0}}{-0.6} & \sbcl{-0.1}{-0.1} & \sbcl{+0.2}{+0.2} \\
-
-    % 37.53	55.74	51.15	33.06	44.15	86.67	23.05	51.38	41.64
-    \footnotesize \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & 41.9 & 51.8 \\[-2pt]
-%                      & 37.5 & 56.1 & 51.1 & 34.0  &  44.3 & 87.3 & \SW{28.4} & & 51.7 & \\
-                        &  \sbcl{+0.2}{-0.2} & \sbcl{+0.4}{+0.5} & \sbcl{+0.1}{0.0} & \sbcl{-0.9}{-0.4}  & \sbcl{-0.1}{-0.2} &  \sbcl{+0.9}{-0.2} & \sbcl{\SW{-8.0}}{\SW{-0.8}} & \sbcl{+0.1}{-0.2} & \sbcl{+0.1}{-0.1} \\
-     \hline
-  \end{tabular}
-  \caption{Ability to handle a new domain.
-\revision{We report BLEU scores for a complete training session with 7 domains, as well as differences with (left) training with 6 domains (from Table~\ref{tab:performance}); (right) continuous training mode.} Averages only take into account six domains (\domain{News} excluded).
-Underline denotes a significant loss, bold a significant gain.}
-  \label{tab:warmrestart}
-  \fyDone{Fill the table, significancy testing against initial learning condition, for each domain, and averages ? }
-  \fyDone{Need correct number / differences for Mixed-Nat}
-  \fyDone{averages include one more domain ? No for comparison}
-  \fyDone{i used smaller number for differences (easier to read...)}
-\end{table*}
-
-We can first compare the results of coldstart training with six or seven domains in Table~\ref{tab:warmrestart}: a first observation is that the extra training data \revision{is hardly helping} for most domains, except for \domain{News}, where we see a large gain, and for \domain{talk}.
-The picture is the same when one looks at MDMTs, where only the weakest systems (\system{DM}, \system{ADM}) seem to benefit from more (out-of-domain) data.\fyDone{is this significant ?}
-% Increasing the heterogeneity of the mix with supplementary out-of-domain data does not always improve the performance.% across the board.
-% Adding is not helping here, especially when domains are well separated.
-Comparing now the coldstart with the warmstart scenario, we see that the former is always significantly better for \domain{news}, as expected, and that resuming training also negatively impacts the performance for other domains.
-This happens notably for \system{DC-Tag}, \system{TTM} and \system{ADM}.
-In this setting \system{MDL-Res} and \system{DM} show the smaller average loss, with the former achieving the best balance of training cost and average BLEU score.% of all systems.
-\fyDone{Numbers for MDL Res need significancy testing; and they are weird - why do we see this ? Because the generic is bad ?}
-% \begin{itemize}
-% \item even when trained from entirely scratch, very few systems are able to take advantage of the additional data for the 6 initial domains - this is observed on the averaged scores in Table~\ref{tab:warmrestart} with hardly surpass those in Table~\ref{tab:performance}
-% \item restart is computationally cheaper (\fyTodo{how much cheaper here ?}) but underperforms cold start for all systems, sometimes by a wide margin.
-% \item the loss is largest for the small domains - how about distance to \domain{news} ?
-% \end{itemize}
-% \fyDone{Very close to fine tuning towards News, no ?}
-
-\subsection{Automatic domains \label{ssec:autodomains}}
-In this section, we experiment with automatic domains, obtained by clustering sentences into $k=30$ classes using the k-means algorithm based on generic sentence representations obtained via mean pooling (cf.\ Section~\ref{ssec:corpora}).
-This allows us to evaluate requirement [P7-scale], training and testing our systems as if these domains were fully separated.
-Many of these clusters are mere splits of the large \domain{med}, while a fewer number of classes are mixtures of two (or more) existing domains \revision{(Full details are in Appendix B)}.
-We are thus in a position to reiterate, at a larger scale, the measurements of Section~\ref{ssec:redomains} and test whether multi-domain systems can effectively take advantage from the cross-domain similarities and to eventually perform better than fine-tuning.
-\revision{The results in Table~\ref{tab:avg_automatic_domains} also suggest that MDMT can surpass fine-tuning for the smaller clusters; for the large clusters, this is no longer true.
-The complete table (in Appendix B) shows that this effect is more visible for small subsets of the medical domain.}
-
-\begin{table*}[t]
-\centering \footnotesize
-\revision{
-  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
-   Model/ & Train &\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
-   Clusters & size & \system{Nat} & \system{Full} & \system{Res} &\system{Res} & \system{Feat}& \system{Tag}& & & & \\ \hline
-10 small&29.3k&68.3&70.0&70.7&\bf 71.2&70.6&53.1&67.3&69.8&67.0&70.2\\
-10 mid&104.7k&44.8&\bf 48.0&46.0&45.7&44.8&44.3&44.5&43.7&41.6&44.5\\
-10 large&251.1k&50.4&\bf 52.9&52.0&51.3&49.6&43.2&49.1&48.5&44.3&49.5\\
-Avg&128.4k&54.5&\bf 57.0&56.2&56.1&55.0&46.9&53.6&54.0&51.0&54.7\\ \hline
-  \end{tabular}
-   \caption{BLEU scores computed by merging the 10 smaller, medium, and larger cluster test sets.
-Best score for each group is in boldface.
-For the small clusters, full-fine tuning is outperformed by several MDMT systems - see details in Appendix~B.}
-   \label{tab:avg_automatic_domains}}
-\end{table*}
-Finally, Table~\ref{tab:subdomains} reports the effect of using automatic domain for each of the 6 test sets: each sentence was first assigned to an automatic class, translated with the corresponding multi-domain system with 30 classes; aggregate numbers were then computed, and contrasted with the 6-domain scenario.
-Results are clear and confirm previous observations: even though some clusters are very close, the net effect is a loss in performance for almost all systems and conditions.
-In this setting, the best MDMT in our pool (\system{MDL-Res}) is no longer able to surpass the \system{Mix-Nat} baseline.%\fyTodo{Note about scaling?}\fyTodo{More comparison?}
-
-\begin{table*}[t]
-  \centering
-  \fyDone{Fix column size}
-  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
-%     &&&&&& \\
-    Domain / Model  & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}}
-  %    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
-  %     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8\\
-  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
-    \system{DC-Tag}       & 38.5 & \SW{54.0} & 49.0   & 33.6 & \SW{42.2} & \SW{76.7} & 41.6 & 49.0 \\%    & 21.8 \\
-    \system{DC-Feat}      & 37.3  & 54.2 & 49.3   & 33.6 & \SW{41.9} & \SW{75.8} & 40.8 & \SW{48.7}  \\% & \SW{21.7} \\
-    \system{LDR}             & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 39.1 & \SW{48.3}         \\% & 22.1 \\ 
-    \system{TTM}            & 37.4 & \SW{53.7} & 48.9 & 32.8 & 41.3 & \SW{75.8} & 40.7 & \SW{48.3}   \\% &  23.4 \\
-    \system{DM}             & 35.4 & 49.3  & 45.2 & 29.7 & 37.1 & \SW{60.0} & 37.8 & 42.8 \\ % & 22.6\\
-    \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 39.5 & 47.1\\% & 23.3 \\
-    \revision{\system{FT-Res}}     & 37.5 & \SW{55.7}  & \SW{51.1}   & 33.1   &  \SW{44.1}  & \SW{86.7} & 41.6 & \SW{51.4}\\%  & \SW{21.2} \\
-    \system{MDL-Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 41.2 & \SW{50.7}\\%  & \SW{21.2} \\
-    \system{WDCMT}       & 35.6 & 53.1 & 48.4 & 30.5 & \SW{37.7} & \SW{56.0} & 38.5 & 43.6 \\ % & 20.4 \\ 
-    \hline
-  \end{tabular}
-  \caption{Translation performance with automatic domains, computed with the original test sets.
-Significance tests are for comparisons with the 6-domain scenario \revision{(Table~\ref{tab:performance}).}}
-  \label{tab:subdomains}
-  \fyDone{Fill the table with correct results,}\fyDone{Change wavg and avg}
-\end{table*}
-
-\section{Related work \label{sec:related}}
-
-The multi-domain training regime is more the norm than the exception for natural language processing \cite{Dredze08online,Finkel09hierarchical}, and the design of multi-domain systems has been proposed for many language processing tasks.
-We focus here exclusively on MD machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection / weighting, adversarial training, etc) have been studied in other contexts.
-
-%; scores are reported separately per domain, suggesting that all domains are equally important at test time.
-% Adversarial adpatation is used in many domains
-% for QA @misc{lee2019domainagnostic,
-%    title={Domain-agnostic Question-Answering with Adversarial Training},
-%    author={Seanie Lee and Donggyu Kim and Jangwon Park},
-%    year={2019},
-%    eprint={1910.09342},
-%    archivePrefix={arXiv},
-%    primaryClass={cs.CL}
-%}, NLU and many others
-Multi-domain translation was already proposed for statistical MT, either considering as we do multiple sources of training data (eg.\ \cite{Banerjee10combining,Clark12onesystem,Sennrich13multidomain,Huck15mixeddomain}), or domains made of multiple topics\fyDone{you mean multiple topics?} \cite{Eidelman12topic,Hasler14dynamic-topic}.\fyDone{Add also topic models refs} Two main strategies were considered: instance-based, involving a measure of similarities between train and test domains; feature-based, where domain/topic labels give rise to additional features.
-
-The latter strategy has been widely used in NMT: \newcite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token or in the form of an additional domain-feature associated to each word.
-These results are reproduced by \newcite{Tars18multidomain}, who also consider automatically induced domain tags.
-This technique also helps control the style of MT outputs in \cite{Sennrich16politeness,Niu18multitask}, and to encode the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}.
-Domain control can also be performed on the target side, as in \cite{Chen16guided}, where a topic vector describing the whole document serves as an extra context in the softmax layer of the decoder.
-Such ideas are further developed in \cite{Chu18multilingual,Pham19generic}, where domain differences and commonalties are encoded in the network architecture: some parameters are shared across domains, while others are domain-specific.
-% (part of the embeddings in \cite{Pham19generic}).
-
-Techniques proposed by \cite{Britz17mixing} aim to ensure that domain information is actually used in a mix-domain system.
-Three methods are considered, using either domain classification (or domain normalization, via adversarial training) on the source or target side.
-There is no clear winner in either of the three language pairs considered.
-One contribution of this work is the idea of normalizing representations through adversarial training, so as to make the mixture of heterogeneous data more effective; representation normalization has since proven a key ingredient in multilingual transfer learning.
-The same basic techniques (parameter sharing, automatic domain identification / normalization) are simultaneously at play in \cite{Zeng18multidomain,Su19exploring}: in this approach, the lower layers of the MT use auxiliary classification tasks to disentangle domain specific representations on the one hand from domain-agnostic representations on the other hand.
-These representations are then processed as two separate inputs, then recombined to compute the translation.
-
-Another parameter-sharing scheme is in \cite{Jiang19multidomain}, which augments a Transformer model with domain-specific heads, whose contributions are regulated at the word/position level: some words have ``generic'' use and rely on mixed-domain heads, while for some other words it is preferable to use domain specific heads, thereby reintroducing the idea of ensembling at the core of \cite{Huck15mixeddomain,Saunders19ucam}.
-The results for three language pairs outperform several standard baselines for a two-domain systems (in fr:en and de:en) and a 4-domain system (zh:en).% \footnote{The reported results for WDC+WL are very bad, worst than anything \fyTodo{This is a mystery to me}.}
-
-Finally, \citet{Farajian17multidomain}, \citet{Li18onesentence} and \citet{Xu19lexical} adopt a different strategy.
-Each test sentence triggers the selection of a small set of related instances; using these, a generic NMT is tuned for some iterations, before delivering its output.
-This approach entirely dispenses with the notion of domain and relies on data selection techniques to handle data heterogeneity.
-
-\section{Conclusion and outlook \label{sec:conclusion}}
-\fyDone{Write conclusions}\fyFuture{no mixture of experts}
-In this study, we have carefully reconsidered the idea of multi-domain machine translation, which seems to be taken for granted in many recent studies.
-We have spelled out the various motivations for building such systems and the associated expectations in terms of system performance.
-We have then designed a series of requirements that MDMT systems should meet, and proposed a series of associated test procedures.
-In our experiments with a representative sample of MDMTs, we have found that most requirements were hardly met for our experimental conditions.
-If MDMT systems are able to outperform the mixed-domain baseline, at least for some domains, they all fall short to match the performance of fine-tuning on each individual domain, which remains the best choice in multi-source single domain adaptation.
-\revisiondel{In this scenario, architecture-based strategies aimed at parameter sharing and/or representation normalization obtain the best results.}\fyDone{Mitigate this}
-As expected however, MDMTs are less brittle than fine-tuning when domain frontiers are uncertain, and can, to a certain extend, dynamically accommodate additional domains, this being especially easy for feature-based approaches.
-Our experiments finally suggest that all methods show decreasing performance when the number of domains or the diversity of the domain mixture increases.
-% (for a fixed data size).
-
-Two other main conclusions can be drawn from this study: first, it seems that more work is needed to make MDMT systems make the best out of the variety of the available data, both to effectively share what needs to be shared while at the same time separating what needs to be kept separated.
-We notably see two areas worthy of further explorations: the development of parameter sharing strategies when the number of domains is large; the design of training strategies that can effectively handle a change of the training mixture, including an increase in the number of domains.
-Both problems are of practical relevance in industrial settings.
-Second, and maybe more importantly, there is a general need to adopt better evaluation methodologies for evaluating MDMT systems, which require systems developers to clearly spell out the testing conditions and the associated expected distribution of testing instances, and to report more than comparisons with simple baselines on a fixed and known handful of domains.
-
-% --------------
-\section*{Acknowledgements}
-
-The work presented in this paper was partially supported by the European Commission under contract H2020-787061 ANITA.
-
-This work was granted access to the HPC resources of [TGCC/CINES/IDRIS] under the allocation 2020- [AD011011270] made by GENCI (Grand Equipement National de Calcul Intensif)
-% The acknowledgements should go immediately before the references.  Do
-% not number the acknowledgements section. Do not include this section
-% when submitting your paper for review.
-
-% include your own bib file like this:
-% \bibliographystyle{coling}
-
-% \newpage\newpage
-
-\bibliographystyle{acl_natbib}
-\bibliography{multidomain}
-
-\clearpage{}
-
-\section*{Appendices}
-
-\subsection*{A - Description of multi-domain systems \label{ssec:implementation-details}}
-\begin{table*}[t]
   \centering
@@
 More than 2/3 of these clusters mostly comprise texts from one single domain, as for cluster 12 which is predominantly \domain{med}, the remaining clusters typically mix 2 domains.
-Fine-tuning with small domains is often outperformed by other MDMT techniques, an issue that a better regularization strategy might mitigate.
-Domain-control (\system{DC-Feat}) is very effective for small domains, but again less so in larger data conditions.
-Among the MD models, approaches using residual adapters have the best average performance.}
-% \newpage\newpage \newpage
 
 \end{document}
-
-\todos{}
-
-\subsection{Formalizing multi-domain adaptation (MDA)}
-
-In MDA, the typical training situation (for a classifier) considers $k$ domains with $\mathcal{D}_k(x,y)$ the joint distribution of features $x \in \mathcal{X}$ and labels $y \in \mathcal{Y}$ for domain $k$.
-In \cite{Mansour09domainadaptation}, the labeling function $f$ is deterministic and the joint distribution is entirely defined by the marginal $\mathcal{D}_k(x)$ and $f()$.This work further assumes $k$ hypotheses $h_k(x): \mathcal{X} \rightarrow{} \mathcal{Y}$ with an error smaller than $\epsilon$ when $x\sim \mathcal{D}_k$: $\sum_{x} L(h_k(x), f(y)) \mathcal{D}_k(x) < \epsilon_k$ and that the target distribution is a convex combination of $\{\mathcal{D}_k\}$: $\mathcal{D}_T(x) = \sum_k \lambda_k \mathcal{D}_k(x)$.
-\cite{Mansour09domainadaptation} focuses on \textsl{combined hypotheses}, ie.\ hypotheses for which each prediction $h(x)$ is a linear combination of domain specific hypotheses $h(x_k)$, and derives generalization bounds for the case where $\mathcal{D}_T(x)$ is fixed in advance, or worst case bounds valid for any mixture target distribution.
-The optimal solution has the form $h_{z,\eta}(x) = \sum_k \frac{z_kD_k(x) + U(x)/\gamma}{\sum_{k'} z_kD_k(x) + \eta U(x)/l}$ and is termed the \emph{distribution weighted combined rule}.
-Further work in \cite{Mansour09multiple} generalizes these bounds to the case where $\mathcal{D}_T$ is not a mixture of existing domains (in that case we will find a mixture that is close to $\mathcal{}$in the Renyì-divergence sense ); where $\mathcal{D}_k$ are unknown and estimated from the data; or when the labelling function is not the same across domains.
-More recent work study the case where the labelling function is stochastic, meaning that domains are defined by joint ($\mathcal{D}_k(x,y)$) rather than marginal distributions \cite{Hoffman18algorithms}, and provides an algorithm to compute the optimal weights for the combined rule based on estimates of the joint distributions.
-A guiding principle is that this rule should tend towards uniform performance across domains, to ensure robustness in the worst case scenario \cite{Oren19distributionally}.
-
-A related problem is \emph{Multiple source single domain adaptation} which is a variation of conventional DA from a mixture of source domains.
-In this setting it is advantageous to take advantage of the similarities between each source and the fixed target domain (see eg.\ \cite{Duan12domain,Cortes19AdaptationBO,Wen19domain} for studies of the supervised case, \cite{Zhao18adversarial} for the unsupervised case).
-
-\subsection{Complements - to be removed \label{ssec:oldformalization}}
-
-We define a domain $d$ as a joint distribution $\mathcal{D}_d(x,y)$ over input-output pairs $(x,y) \in \mathcal{X}\times{}\mathcal{Y}$.
-Assuming that the training data is sampled from a known mixture of $D$ source domains $\mathcal{D}_s(x,y) = \sum_d \lambda_{s,d} \mathcal{D}_d(x,y)$, several cases are studied in the ML literature:
-
-\begin{itemize}
-\item the test distribution $\mathcal{D}_t$ is known 
-  \begin{itemize}
-  \item $\mathcal{D}_t = \mathcal{D}_{s}$: this is standard ML, with mixture densities
-  \item $\mathcal{D}_t(x,y) = \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) $ with a different mixture of domains in training and testing ($\vlambda_s \neq \vlambda_t$), corresponding to multiple source adaptation (MSA) \cite{Mansour09domainadaptation},
-  \item $\mathcal{D}_t(x,y) \neq \sum_d \lambda_{t,d} \mathcal{D}_d(x,y) \forall \lambda$, where the target domain for adapation is not a mixture of source domains, studied in \cite{Mansour09multiple}
-  \item variants of MSA which only assume knowledge of the marginal distribution $\mathcal{D}_t(x) = \sum_{y} \lambda_{t,d} \mathcal{D}_d(x) $
-  \item more realistic situations where  the train and test distributions are estimated from finite samples, also studied in \cite{Mansour09multiple}.
-  \end{itemize}
-   Note that we recover standard supervised DA is when there is only one component in the training and test mixtures, and multiple-source single domain adaptation when the source is a mixture but the test domain is homogeneous.
-Unsupervised DA \cite{Daume06domain} corresponds to the case where only the marginal distribution of the test domain inputs is available, while semi-supervised DA is the reverse situation.
-
-\item the test distribution is not known and training needs to be distributionally robust, ie.\ minimize the error against all possible distributions, with variants depending on whether it is or not assumed to be a mixture of existing domains, and whether it is perfectly or imperfectly known.
-This setting is also studied in eg.\ \cite{Mansour09multiple,Hoffman18algorithms}, and  in \cite{Oren19distributionally} for adversarial language model learning.
-\item with respect to the conditional label distribution $P(y|x)$:
-  \begin{itemize}
-  \item $P(y|x)$ is deterministic, meaning that the joint distribution is entirely specified by the marginal and a labelling function $y=h(x)$; a more complex case is when $h()$ also varies across domains;
-  \item $P(y|x)$ is stochastic, but shared across all domains, which is the typical covariate-shift scenario \cite{Shimodaira00improving}; it can also vary across domains \cite{???}\fyTodo{Missing citation vary accoss domains}
-  \end{itemize}
-\end{itemize}
-
-Multi-domain learning, as defined in \citet{Dredze08online} additionally makes the requirement that categorial domain tags are available in training and testing; a multi-domain learner should use them effectively \cite{Joshi12multidomain} to improve simultaneously on all domains \cite{Finkel09hierarchical}.
-This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyDone{explain zero shot}.
-% ----
-
-% \subsection{MDMT: a tentative definition \label{ssec:definition}}
-
-% Informally, a multi-domain system should be able to accurately translate heterogeneous sources of data. This loose definition correspond to several practical situations:
-% \begin{itemize}
-% \item the more common situation is when both train and test documents (or sentences) are domain-tagged, where each domain corresponds to a specific source of data, and differ from the other sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
-It is usually assumed that the set of test domains is the same as the set of training domains -- when this is not the case, we would have zero-shot domain adaptation, similar to what is studied in multilingual MT.
-This is the more general setting, already proposed in \cite{Dredze09multidomain}.\fyTodo{do we need a name here: such as supervised MDMT ?
-and mostly supervised MDMT for the second case ?}
-
-% \item It can also be the case that domain labels are only known in training, while test inputs may or may not be associated with explicit domains.
-This is, for instance, the setting adopted \cite{Zeng18multidomain}, where automatic domain identification is performed for test sentences.
-Note that automatic domain identification can also be effectively performed in the first setting, especially in the presence of close or related domains in training.
-
-% \item A last setting, that has not received much attention for MT (see however \cite{Sennrich13multidomain}), is where both train and test domains are automatically infered from data, using clustering techniques.
-Even though domain labels are typically available in most training scenario, these approaches assumes that might be preferable, owing to the multiple sources of intra-corpus and inter-corpus variabilities, to resort to automatic domain assignments.
-The online approach of \newcite{Farajian17multidomain}, where distributional cues, rather that categorical tags, are used for fine-tuning, is a variant of this strategy.
-% \end{itemize}
-
-% In the rest of the paper, we mostly adhere to the first definition which correspond to a situation of practical interest for the industry, and assume that multi-domain MT is mostly about effectively using domain tags in training and testing.
-We also experiment in Section~\ref{sec:results} with the other two alternative scenarios, as they also define useful contexts to evaluate the actual robustness of MDMT systems.
-
-% \jcDone{...a very common scenario in MT industry where clients dispose of mulitiple data corresponding to different domains (differing in genre, thematic, register, style, etc.). }
-
-% \subsection{Evaluation issues \label{ssec:evaluation}}
-
-% In theory, evaluation procedures should report realistic estimates of the test loss, computed on a sample deemed representative of the anticipated test conditions.
-Three scenarios need be distinguished:
-% \begin{itemize}
-% \item test data are mixed in the same proportion as training data; this warrants the design of an unbalanced mix-domain test sets in the same proportions as in training; alternatively, if per domain scores are ever reported, their average should take the mixing proportions into account;
-% \item all domains are equally likely and important in testing (the prior over test domains is uniform), which would warrant the computation of per domain scores and of unweighted averages, as reported in \cite{Farajian17multidomain,Su19exploring,Wang20general}.
-Note that in this situation, the optimal training regime would be to balance the training set in the same way -- which is hardly ever done in practice, as the baseline scores always consider mixing unevenly all available data.
-\cite{Wang20general} uses an intermediate mixing regime, which tends to reduce part of the unbalance, and improve the performance for underrepresented domains.
-% \item the test distribution is neither equal to the training set, nor uniform.
-Here again, if this distribution is known in advance, an optimal domain mix should be used in training to compute baseline scores, rather than using the empirical distribution of all the training data.
-A more challenging case is when the test set includes data from domains unseen in training, or when it is chosen adversarially \cite{XXX}.
-% % : this situation, where test instances have a null probability in training, goes agains the basic statistical assumptions of DA \cite{Shimodaira00improving}.
-It is thus important that the likeliness of such events should be clearly formulated when describing the evaluation protocol.
-% \end{itemize}
-
-% The papers in the literature are seldom explicit regarding their assumptions regarding the test distribution.
-In most cases, per-domain performances are reported, and even sometimes averaged, suggesting that the scenario~2 most faithfully reflects the expected testing conditions.
-Another important issue is the choice of the evaluation metrics: if BLEU is still the standard metric used in most studies, it is not necessary the best choice to measure subtle improvements of lexical choices \cite{Irvine13measuring} that are often associated with adapted systems.
-
-% A last issue regarding evaluation is related to the measurement of computational costs: when this is the main a motivation for using MDMT, we expect that computational costs would be reported and compared, which however is hardly ever done in practice.
-\fyTodo{Keep this ?}

### 2021-05-12 2021/emnlp/auto_cl_mdmt.tex
with last todos

@@
 Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of \emph{curriculum learning} (CL) is initially introduced.
-CL relies on the notion of ``easiness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
+CL relies on the notion of ``easyness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
 CL has been shown to speed up learning.
-While the initial aim was to improve and speed up training, it has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitions of ``easiness''.
+While the initial aim was to improve and speed up training, it has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitions of ``easiness''.\fyTodo{Ajouter Grave}
 
@@
 
+\section{Conclusion and future work}
+\fyTodo{Write conclusion}
+
+Contribution : generic framework for several settings. // comparison of several dynamic strategies.
+
+Future work:
+slow down learning during warm up or normalize reward; handle large number of classes;} 
+too much importance to the dev ? randomize ?
+
 \section*{Acknowledgments}

### 2021-05-12 2021/emnlp/auto_cl_mdmt.tex
local changes

--- 2021-05-12 2021/emnlp/auto_cl_mdmt.tex
@@
 In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key component in training is then to mitigate the detrimental effects of a possible mismatch between these distributions.
-Single-source and multi-source\footnote{In this context, multi-source refers to simultaneously having multiple domains to adapt from and should not be confused with multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
-A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Fix ref}\revision{cite also \citep{Shen21source}}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
+Single-source and multi-source domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work} \revision{cite also \citep{shen21source}}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
 
@@
 This computation is inspired by the target prediction gain in \citep{Graves17automated}.
-However, \citep{Graves17automated} used accumulated gains from the past as reward, while we predict the usefulness of each domain for improving the future performance of the system given its current state.
+However, \cite{Graves17automated} used accumulated gains from the past as reward, while we predict the usefulness of each domain for improving the future performance of the system given its current state.
 This is achieved by simulating a round of training with only the data from one domain.
@@
 
-The work of \citet{Wang20balancing} is also related: it is based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\lambda^{l}$ that will result in the best NMT model with respect to a given target dev set at the end of training.
+The work of \cite{Wang20balancing} is also related: it is based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\lambda^{l}$ that will result in the best NMT model with respect to a given target dev set at the end of training.
 These authors also derive a similar form of update for $\psi$.
@@
 \end{itemize}
-These questions are further explored in Section~\ref{sec:results}. We now turn to our experimental conditions.
+These questions are further explored in section~\ref{sec:results}. We now turn to our experimental conditions.
 
@@
 We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \citep{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \citep{Neubig19compare-mt}.
-We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value.}\fyTodo{Pourquoi pas sacre BLEU?}
+Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
 
@@
 \system{Mixed-$0$} uses a uniform domain distribution, \system{Mixed-$1.0$} uses the observed domain distribution.
-\item finetuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ??
-iterations} The full finetuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems, each adapted for one domain, with no parameter sharing across domains.
+\item finetuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing.\fyTodo{for ??
+iterations} The full finetuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems each adapted for one domain, with no parameter sharing across domains.
 
-\item Our comparison of multi-domain systems\fyTodo{CL is only dynamic ?} includes baselines with fixed data mixtures corresponding to $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning (CL) of \citet{Zhang19curriculum} and Differential Data Selection (DDS) of \citet{Wang20balancing}.
+\item Our comparison of multi-domain systems\fyTodo{CL is only dynamic ?} includes baselines with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning (CL) of \citet{Zhang19curriculum} and Differential Data Selection (DDS) of \citet{Wang20balancing}.
 \end{itemize}
@@
 
-In our work, we simply parameterize the distribution $\vlambda^l$ as follows (with $\beta=2$ in all our experiments):\fyTodo{Explain why}
+In our work, we simply parameterize the distribution $\lambda^l$ as follows (with $\beta=2$ in all our experiments):\fyTodo{Explain why}
 \begin{equation}
@@
 Additional choices for the hyper-parameters of our method include the number of simulation steps $k=10$ and the learning rate $\mathbf{lr}_{data}=0.001$.
-We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section~\ref{ssec:clda}), where we use only 20 gradient descent iterations.\footnote{This is intended to regularize the training of $\psi$ and avoid converging towards degenerate solutions where some clusters get no training samples.}\fyDone{Why?
+We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section \ref{ssec:clda}), where we use 20 gradient descent iterations.\fyTodo{Why?
 computation?} We split the training into 100 short sessions that last 2000 training steps each.
@@
 In the domain adaptation task, given the data from 6 domains (\domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}), we aim to build distinct expert NMT models for each domain.
-In the bi-domain adaptation task, given again same 6 domains, we only focus on adapting only to 2 (known) domains.
 
@@
 
-In the fourth experiment (few shot adaptation)\fyTodo{or weakly supervised ?}, given training data in 6 domains and a small development set of a new domain, (\domain{News} in our case), our goal is to build an NMT model which will perform well on test data from the unseen domain.
+In the bi-domain adaptation task, given again same 6 domains, we only focus on adapating only to 2 domains.
+
+In the fourth experiment, given training data in 6 domains and a small development set of an new domain, (\domain{News} in our case) we aim to build an NMT model which will perform well on test data from the unseen domain.
 
@@
 \subsection{Domain Adaptation}\label{ssec:da}
-In this setting, we aim to build an expert NMT model in only one domain. This setting corresponds to Dirac distribution $\vlambda^t$, in which the probability of domains is 0 except for one domain.
+In this setting, we aim to build an expert NMT model in only one domain. This setting corresponds to Dirac distribution $\lambda^t$, in which the probability of domains is 0 except for one domain.
 
-We initialize MDAC and DDS by 3 different distributions $\vlambda_0$, $\vlambda_1$ and $\vlambda_d$.
-According to Table~\ref{tab:da}, MDAC achieves the best performance with $\vlambda_0$ (uniform domain distribution) as the initial value, though $\vlambda_0$  initialization only significantly improves over $\vlambda_d$ initialization in small domains such as \domain{talk},\domain{bank} and\domain{it}.
-Initializing with $\vlambda_d$ seems more appropriate when targeting large domains such as \domain{med} and \domain{law}.
-The same conclusions also hold for DDS.
+We initialize MDAC and DDS by 3 different distributions $\lambda_0$, $\lambda_1$ and $\lambda_d$.
+According to Table \ref{tab:da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen, though $\lambda_0$  initialization only improve significantly $\lambda_d$ initialization in small domains including \domain{talk},\domain{bank}and\domain{it}.
+$\lambda_d$ initialization is beneficial for adapting the NMT system to large domains such as \domain{med} and \domain{law}.
+The same conclusion is also true for DDS.
 
-We now compare the best MDAC system (initialized with $\vlambda_{t=0} = \vlambda_0$) to full finetuning.
-According to Table~\ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}.
+We now compare the best MDAC (initialized by $\lambda_0$) to fine-tuning.
+According to Table \ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}.
 However, MDAC improves fine-tuning by approximately 0.5 BLEU in \domain{bank} and 1.0 BLEU in \domain{rel}.
@@
 
-DDS also demonstrates promising results in this task, but is always outperformed by MDAC overall. The same conclusions also hold for the CL method.\fyTodo{Continue}
+DDS also demonstrates promising results in this task but is always outperformed by MDAC overall. The same conclusion is also applied to the CL method.
 
@@
     domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline
-    FT-Full($d$) &37.7&59.2&54.5&34&46.8&90.8&53.8\\ \hline
+    FT-Full($d$) \hfill [BLEU] &37.7&59.2&54.5&34&46.8&90.8&53.8\\ \hline
     \hline
@@
   \end{tabular}
-  \caption{We report BLEU scores of each method with 6 target domains, and compare with full fine-tuning: each column corresponds to a distinct system.
+  \caption{We report scores of each method with 6 target domains, and compare with full fine-tuning: each column corresponds to a distinct system.
 ($^{*}$) MDAC is significantly better than CL, fine-tuning and DDS with $p<0.05$.
@@
 Multi-domain adaptation task requires best averaged performance of the NMT model in 6 domains.
-This setting corresponds to a uniform test distribution $\vlambda^t = \vlambda_0$.
-In this situation, the CL method of \citet{Zhang19curriculum} does not apply.
-We therefore only compare the performance of MDAC, DDS and several fixed training data distribution $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ (see equation~\eqref{mixture:trn}).
+This setting corresponds to the uniform test distribution $\lambda^t = \lambda_0$.
+In this situation, the method CL of \citet{Zhang19curriculum} can not be applied.
+We therefore only compare the performance of MDAC, DDS and several fixed training data distribution $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$.
+The definition of those notations can be found in the equation \ref{mixture:trn}.
 
@@
   \centering \small
-  \begin{tabular}{|l|*7{r|}} \hline
-    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} \\ \hline \hline%  \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}}
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&36.2&-&-&-&46.1$^*$&-&\\
-    \system{DDS($\vlambda_0, \vlambda_2$)}&36.5&-&-&-&44.4&- \\
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&57.2$^*$&53.8$^*$&-&-&-& \\
-    \system{DDS($\vlambda_0, \vlambda_2$)}&-&56.1&52.6&-&-&-&- \\
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&32&-&90 \\
-    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&32.9&-&91.1 \\ \hline
+  \begin{tabular}{|l|*9{r|}} \hline
+    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} \\ \hline \hline
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&36.2&-&-&-&46.1$^*$&-&-&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&36.5&-&-&-&44.4&-&-&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&57.2$^*$&53.8$^*$&-&-&-&-&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&56.1&52.6&-&-&-&-&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&32&-&90&-&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&32.9&-&91.1&-&- \\ \hline
   \end{tabular}

### 2021-05-14 2021/emnlp/auto_cl_mdmt.tex
last check at the intro

--- 2021-05-12 2021/emnlp/auto_cl_mdmt.tex
@@
 \section{Introduction}\label{sec:intro}
-A typical setting in machine translation (MT) is to collect the largest possible collection of parallel data for the chosen language pair, with the intent to achieve optimal performance for a task of interest.
+A typical setting in machine translation (MT) is to collect the largest possible collection of parallel data for the chosen language pair, with the intent to achieve optimal performance for the task of interest.
 In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key component in training is then to mitigate the detrimental effects of a possible mismatch between these distributions.
-Single-source and multi-source\footnote{In this context, multi-source refers to simultaneously having multiple domains to adapt from and should not be confused with multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
-A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Fix ref}\revision{cite also \citep{Shen21source}}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
+Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from and should not be confused with multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison,Saunders21domain} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Aharoni19massively},\fyDone{Fix ref}
+% \revision{cite also \citep{Shen21source}} -> I think this is not so appropriate here
+where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
 
 This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions \citep{Moore10selection,Axelrod11domain}.
-Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data.
+Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data \citep{Miceli-barone17regularization}.
 It may also be suboptimal, as some target domains or languages might be easier to train than others.
@@
 have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies.
-Notably,  \citet{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while\citet{Graves17automated,platanios19competence,Kumar19reinforcement,Wang20balancing,Wang20learning-multi} build curricula that automatically adapt to the training data.
+Notably, \citet{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while \citet{Graves17automated,Platanios19competence,Kumar19reinforcement,Wang20balancing,Wang20learning-multi} build curricula that automatically adapt to the training data.
 
@@
 We evaluate our method in various single-domain and multi-domain adaptation settings.
-We compare our method with several contrasts including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}.
+We compare our method with several contrasts including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}, which was previously only applied to multilingual MT.
 % From now on, we name our method MDAC which is brief for .
-Based on these experimental results, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard finetuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test situations, from targeting one single domain to the full multi-domain scenario.
+Based on these experimental results, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard fine-tuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test situations, from targeting one single domain to the full multi-domain scenario.
 % distributions
@@
 \begin{itemize}
-\item is MDAC a viable alternative to conventional finetuning? In particular, does it enable to better take advantage of relevant data from other domains?
+\item is MDAC a viable alternative to conventional fine-tuning? In particular, does it enable to better take advantage of relevant data from other domains?
 \item is MDAC also a viable option in multi-domain adaptation scenarios?
@@
 \item finetuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ??
-iterations} The full finetuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems, each adapted for one domain, with no parameter sharing across domains.
+iterations} The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems, each adapted for one domain, with no parameter sharing across domains.
 
@@
 
-We now compare the best MDAC system (initialized with $\vlambda_{t=0} = \vlambda_0$) to full finetuning.
+We now compare the best MDAC system (initialized with $\vlambda_{t=0} = \vlambda_0$) to full fine-tuning.
 According to Table~\ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}.
@@
 However, automatic clustering could make the situations harder than the conventional settings where
+
 \section{Related Work \label{sec:related}}
@@
 \section{Conclusion and future work}
+
 \fyTodo{Write conclusion}
 
-Contribution : generic framework for several settings. // comparison of several dynamic strategies.
+Contribution : generic framework for several settings ; comparison of several dynamic strategies.
 
 Future work:
-slow down learning during warm up or normalize reward; handle large number of classes;} 
+slow down learning during warm up or normalize reward; handle large number of classes;
 too much importance to the dev ? randomize ?
@@
 \section*{Acknowledgments}
+
 \bibliographystyle{acl_natbib}

### 2021-05-14 2021/emnlp/auto_cl_mdmt.tex
with a conclusion

@@
 Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of pure domain adaptation.
-These various settings are illustrated Figure~\ref{fig:mdmt-lambdas}.
+These settings are illustrated Figure~\ref{fig:mdmt-lambdas}.
 \begin{figure}[h]
@@
 
-A more practical standard approach to supervised DA is \emph{fine-tuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
+A widely used approach to supervised DA is \emph{fine-tuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
 With our notations, this approach amounts to first learning an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$) with $\indic{A}$ the indicator function for predicate $A$.
 Note that this strategy is potentially suboptimal, as some out-of-domain samples may contribute to the final performance due to eg.\ domain overlap.
-Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains might be easier to learn than others.\fyTodo{How to measure this?} \revision{Answer: The graphic of \system{Mixed-0} shows that with the same amount of data(batches) the convergences of the domains vary significantly}
+Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains might be easier to learn than others.\fyDone{How to measure this?}
+% \revision{Answer: The graphic of \system{Mixed-0} shows that with the same amount of data(batches) the convergences of the domains vary significantly}
 
@@
 Our approach to DA in this paper clearly falls under the former category.
-We refer interested readers to these references and do not discuss review DA any further.
+We refer interested readers to these references and do not discuss DA any further.
 
 Multi-domain adaptation (MDA) aims to develop systems that simultaneously bode well for several domains.
-Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citet{Jiang19multidomain}.
-Unsupervised MDA is considered in \citet{Farajian17multidomain} as an unsupervised DA problem.
+Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \citep{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citet{Jiang19multidomain}.
+Unsupervised MDA is considered in \citet{Farajian17multidomain} as an instance of unsupervised DA.
 
-Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of \emph{curriculum learning} (CL) is initially introduced.
+Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of curriculum learning is initially introduced.
 CL relies on the notion of ``easyness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
-CL has been shown to speed up learning.
-While the initial aim was to improve and speed up training, it has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitions of ``easiness''.\fyTodo{Ajouter Grave}
+Various ways to automate CL using the framework of multi-armed bandits are notably explored in \citep{Graves17automated}, which has been an inspiration for our implementation.
+While the initial aim was primarily to improve and speed up training, CL has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
 
@@
 
-\fyTodo{Write conclusion}
+\fyDone{Write conclusion}
+In this study, we have presented a generic framework to perform a variety of standard adaptation tasks for machine translation, ranging from the conventional supervised domain adaptation to multi-domain adaptation and few shot DA.
+By experimenting with all these settings, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during  the course of training, could be used in all these situations.
+This algorithm, we believe, provides us with a more sound approach to (multi-domain) DA that existing heuristics, and dispenses with the costly search of optimal meta-parameters.
+Another contribution has been an experimental comparison of several recent approachs for dynamic data selection.
 
-Contribution : generic framework for several settings ; comparison of several dynamic strategies.
+In our future work, we intend to continue develop this approach and improve its effectiveness.
+One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training, so as to slow down convergence \cite{Kumar19reinforcement}.
+Another area where we need to progress is the unsupervised learning setting of Section~\ref{ssec:clda}, where our results still lag way behind that of supervised DA - this might be due to the inability of our simplistic optimization strategy to handle situations where the number of cluster is too large.
+% slow down learning during warm up or normalize reward; handle large number of classes;
+% too much importance to the dev ? randomize ?
 
-Future work:
-slow down learning during warm up or normalize reward; handle large number of classes;
-too much importance to the dev ? randomize ?
-
-\section*{Acknowledgments}
+% \section*{Acknowledgments}
 

### 2021-05-14 2021/emnlp/auto_cl_mdmt.tex
local changes

--- 2021-05-12 2021/emnlp/auto_cl_mdmt.tex
@@
 According to Table \ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}.
-However, MDAC improves fine-tuning by approximately 0.5 BLEU in \domain{bank} and 1.0 BLEU in \domain{rel}.
+However, MDAC improves fine-tuning by approximately 1.2 BLEU in \domain{bank} and 1.0 BLEU in \domain{rel}.
 This indicates that for small domains, out-of-domain data can improve the generalization of the NMT model and MDAC is capable to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
@@
     domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline
-    FT-Full($d$) \hfill [BLEU] &37.7&59.2&54.5&34&46.8&90.8&53.8\\ \hline
+    FT-Full($d$) \hfill [BLEU] &40.3&63.8&54.4&38.5&52.0&91.0&56.7\\ \hline
     \hline
-    \system{CL($d$)} &37.2&55.4&53.0&31.3&45.3&90.8&52.2\\ \hline
-    \system{DDS($\vlambda_0, \vlambda_d$)} &36.7&56.2&54.3&33.5&46.8&91.8&53.2\\
-    \system{MDAC($\vlambda_0, \vlambda_d$)} &36.7&58.2$^{**}$&55$^{*}$&33.3&46.6&91.9$^{***}$&53.6\\
+    \system{CL($d$)} &40.2&60.2&53.7&36.5&51.1&91.1&55.5\\ \hline
+    \system{DDS($\vlambda_0, \vlambda_d$)} &39.6&60.1&55.0&38.5&52.5&92.0&56.3\\
+    \system{MDAC($\vlambda_0, \vlambda_d$)} &39.6&62.5$^{**}$&55.6$^{*}$&38.5&52.4&92$^{***}$&56.8\\
      \hline
-    \system{DDS($\vlambda_1, \vlambda_d$)} &36.4&56.3& 45.4& 32.6&44.5&43.1&43.1\\
-    \system{MDAC($\vlambda_1, \vlambda_d$)} &37.2&55.1&51.8&32.8&44.9&85.4&51.2 \\ \hline
-    \system{DDS($\vlambda_d, \vlambda_d$)} &37.0&59.6&53.7&30.2&45.7&91.6&53.0 \\
-    \system{MDAC($\vlambda_d, \vlambda_d$)}&37.9&59.6&53.9&30.4&45.8&92.1&53.3 \\
+    \system{DDS($\vlambda_1, \vlambda_d$)} &39.7&53.9&49.6&37.9&43.1&64.3&48.1\\
+    \system{MDAC($\vlambda_1, \vlambda_d$)} &40.2&59.9&52.6&38.5&50.7&79.8&53.6\\ \hline
+    \system{DDS($\vlambda_d, \vlambda_d$)} &39.9&63.9&54.5&35.4&51.2&91.8&56.1\\
+    \system{MDAC($\vlambda_d, \vlambda_d$)}&40.6&63.9&54.5&35.6&51.3&92.3&56.4\\
     \hline
@@
     % Multidomain
-    \system{Mixed-0}      & 35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5& 51.4 \\
-    \system{Mixed-0.25} & 35.9 & 54.9 & 52.6 & 32.6 & 45 & 90.3& 51.9 \\
-    \system{Mixed-0.5}   & 36.1 & 55.4 & 51.8 & 33.5 & 46.2 & 90 & 52.2 \\
-    \system{Mixed-0.75} & 36.5 & 55 & 51.2 & 34 & 44.3 & 87.2& 51.7 \\
-    \system{Mixed-1}     & 37.3  & 54.6 & 50.1 & 33.5 & 43.2 & 77.5& 49.4 \\
+    \system{Mixed-0} &38.6&59.3&53.7&37.3&51.0&90.4&55.1\\
+    \system{Mixed-0.25}&38.9&59.6&53.3&37.6&50.5&90.6&55.1\\
+    \system{Mixed-0.5}&39.0&60.2&52.5&38.5&51.9&90.3&55.4\\
+    \system{Mixed-0.75}&39.4&59.9&51.9&38.8&50.0&87.6&54.6\\
+    \system{Mixed-1}&40.3&59.5&49.8&36.4&49.0&80.0&52.5\\
     \hline \hline
-    \system{DDS($\vlambda_0, \vlambda_0$)} &37.3&55.1&51&33.5&43.4&90.8&51.8\\ 
-    \system{MDAC($\vlambda_0, \vlambda_0$)}&35.5&55.5&53.7$^*$&32.3&45.6&91.1$^*$&52.3$^{**}$\\ 
+    \system{DDS($\vlambda_0, \vlambda_0$)} &40.1&56.9&50.7&37.4&46.8&92.0&54.0\\ 
+    \system{MDAC($\vlambda_0, \vlambda_0$)}&38.5&60.3$^{**}$&54.4$^*$&37.3&51.3$^{**}$&91.4$^*$&55.5$^{**}$\\ 
     \hline \hline
-    \system{DDS($\vlambda_1, \vlambda_0$)} &37.9&50.2&47.2&31.2&40.8&58.6&44.3\\
-    \system{MDAC($\vlambda_1, \vlambda_0$)}&37.4&54.4&50.3&32.0&42.7&80.2&49.5\\
+    \system{DDS($\vlambda_1, \vlambda_0$)} &40.6&55.5&48.0&36.2&46.9&60.1&47.9\\
+    \system{MDAC($\vlambda_1, \vlambda_0$)}&40.2&59.3$^{**}$&51.0$^{**}$&36.9$^{**}$&48.6$^{**}$&80.7$^{**}$&52.8$^{**}$\\
     \hline
@@
 For a given line, all the columns \emph{correspond to the same multi-domain system}.
-($^*$) MDAC is significantly better than \system{Mixed-$\alpha$} with $p<0.05$ and DDS.
+($^*$) MDAC is significantly better than \system{Mixed-$\alpha$} with $p<0.05$.
 ($^{**}$) MDAC is significantly better than DDS with $p<0.05$.}
@@
     domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} \\ \hline \hline
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&36.2&-&-&-&46.1$^*$&-&-&- \\
-    \system{DDS($\vlambda_0, \vlambda_2$)}&36.5&-&-&-&44.4&-&-&- \\
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&57.2$^*$&53.8$^*$&-&-&-&-&- \\
-    \system{DDS($\vlambda_0, \vlambda_2$)}&-&56.1&52.6&-&-&-&-&- \\
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&32&-&90&-&- \\
-    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&32.9&-&91.1&-&- \\ \hline
+    \system{DDS($\vlambda_0, \vlambda_2$)}&39.5&-&-&-&50.1&-&-&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&39.1&-&-&-&51.8$^*$&-&-&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&60.8&53.3&-&-&-&-&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&61.9$^*$&54.5$^*$&-&-&-&-&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&37.9&-&91.3&-&- \\ 
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&36.9&-&90.4&-&- \\
+    \hline
   \end{tabular}
@@
   \centering \small
-  \begin{tabular}{|l|*9{r|}} \hline
-    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} \\ \hline \hline
-    \multicolumn{9}{|c|}{\sl Unsupervised domain adaptation} \\ \hline
-    \system{Mixed-0}      & - & - & - & - & - & - & - & 22.4 \\
-    \system{Mixed-0.25} & - & - & - & - & - & - & - & 22.7\\
-    \system{Mixed-0.5}   & - & - & - & - & - & - & - & 23.6\\
-    \system{Mixed-0.75} & - & - & - & - & - & - & - & 23.7\\
-    \system{Mixed-1} & - & - & - & - & - & - & - & 23.5\\
-    \hline \hline
-    \system{MDAC($\vlambda_0, \vlambda_{news}$)} &-&-&-&-&-&-&-&23.1 \\
-     \system{DDS($\vlambda_0, \vlambda_{news}$)} &-&-&-&-&-&-&-&23.3 \\ \hline\hline 
-    \multicolumn{9}{|c|}{\sl 30 clusters} \\ \hline
-    \system{MDAC($\vlambda^*, \vlambda_d$)}&36.3$^*$&56.9$^*$&51.3$^*$&33.2$^*$&43.4&89.4&51.8$^*$&-\\
-     \system{DDS($\vlambda^*, \vlambda_d$)}&35.3&55.2&49.4&30.6&43.3&89.9&50.6&-\\ \hline
+  \begin{tabular}{|l|*{10}{r|}} \cline{1-2} \cline{4-11}
+    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{news}}& \hfill &domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean}  \\
+\cline{1-2} \cline{4-11}
+    \multicolumn{2}{|c|}{\sl unseen domains} & &\multicolumn{8}{c|}{\sl 30 clusters} \\ 
+\cline{1-2} \cline{4-11}
+    \system{Mixed-0}      & 25.7 & &\system{DDS($\vlambda^*, \vlambda_d$)}&38.3&60.1&50.3&35.8&49.1&90.1&53.9\\
+    \system{Mixed-0.25} & 25.8 & &\system{MDAC($\vlambda^*, \vlambda_d$)}&39.2$^*$&61.6$^*$&52.0$^*$&38.2$^*$&49.1&89.7&55.0$^*$\\ \cline{4-11}
+    \system{Mixed-0.5}   &26.5\\
+    \system{Mixed-0.75} &26.8\\
+    \system{Mixed-1} &26.9 \\
+    \cline{1-2}
+     \system{DDS($\vlambda_0, \vlambda_{news}$)} &26.3 \\
+     \system{MDAC($\vlambda_0, \vlambda_{news}$)} &26.3 \\
+     \cline{1-2}
   \end{tabular}
@@
 Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of \emph{curriculum learning} (CL) is initially introduced.
-CL relies on the notion of ``easyness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
+CL relies on the notion of ``easiness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
 CL has been shown to speed up learning.
-While the initial aim was to improve and speed up training, it has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitions of ``easiness''.\fyTodo{Ajouter Grave}
+While the initial aim was to improve and speed up training, it has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitions of ``easiness''.
 
@@
 
-\section{Conclusion and future work}
-\fyTodo{Write conclusion}
-
-Contribution : generic framework for several settings. // comparison of several dynamic strategies.
-
-Future work:
-slow down learning during warm up or normalize reward; handle large number of classes;} 
-too much importance to the dev ? randomize ?
-
 \section*{Acknowledgments}

### 2021-05-14 2021/emnlp/auto_cl_mdmt.tex

@@
 \section{Introduction}\label{sec:intro}
-A typical setting in machine translation (MT) is to collect the largest possible collection of parallel data for the chosen language pair, with the intent to achieve optimal performance for a task of interest.
+A typical setting in machine translation (MT) is to collect the largest possible collection of parallel data for the chosen language pair, with the intent to achieve optimal performance for the task of interest.
 In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key component in training is then to mitigate the detrimental effects of a possible mismatch between these distributions.
-Single-source and multi-source domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
-A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Add more recent work} \revision{cite also \citep{shen21source}}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
+Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from and should not be confused with multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison,Saunders21domain} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Aharoni19massively},\fyDone{Fix ref}
+% \revision{cite also \citep{Shen21source}} -> I think this is not so appropriate here
+where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
 
 This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions \citep{Moore10selection,Axelrod11domain}.
-Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data.
+Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data \citep{Miceli-barone17regularization}.
 It may also be suboptimal, as some target domains or languages might be easier to train than others.
@@
 
-A more practical standard approach to supervised DA is \emph{finetuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
+A widely used approach to supervised DA is \emph{fine-tuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
 With our notations, this approach amounts to first learning an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$) with $\indic{A}$ the indicator function for predicate $A$.
 Note that this strategy is potentially suboptimal, as some out-of-domain samples may contribute to the final performance due to eg.\ domain overlap.
-Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains might be easier to learn than others.\fyTodo{How to measure this?} \revision{Answer: The graphic of \system{Mixed-0} shows that with the same amount of data(batches) the convergences of the domains vary significantly}
+Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains might be easier to learn than others.\fyDone{How to measure this?}
+% \revision{Answer: The graphic of \system{Mixed-0} shows that with the same amount of data(batches) the convergences of the domains vary significantly}
 
@@
 \system{Mixed-$0$} uses a uniform domain distribution, \system{Mixed-$1.0$} uses the observed domain distribution.
-\item finetuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing.\fyTodo{for ??
-iterations} The full finetuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems each adapted for one domain, with no parameter sharing across domains.
-
+\item finetuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ??
+iterations} The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems, each adapted for one domain, with no parameter sharing across domains.
 \item Our comparison of multi-domain systems\fyTodo{CL is only dynamic ?} includes baselines with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning (CL) of \citet{Zhang19curriculum} and Differential Data Selection (DDS) of \citet{Wang20balancing}.
@@
 
-We now compare the best MDAC (initialized by $\lambda_0$) to fine-tuning.
-According to Table \ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}.
-However, MDAC improves fine-tuning by approximately 1.2 BLEU in \domain{bank} and 1.0 BLEU in \domain{rel}.
+We now compare the best MDAC system (initialized with $\vlambda_{t=0} = \vlambda_0$) to full fine-tuning.
+According to Table~\ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}.
+However, MDAC improves fine-tuning by approximately 0.5 BLEU in \domain{bank} and 1.0 BLEU in \domain{rel}.
 This indicates that for small domains, out-of-domain data can improve the generalization of the NMT model and MDAC is capable to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
@@
 Our approach to DA in this paper clearly falls under the former category.
-We refer interested readers to these references and do not discuss review DA any further.
+We refer interested readers to these references and do not discuss DA any further.
 
 Multi-domain adaptation (MDA) aims to develop systems that simultaneously bode well for several domains.
-Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citet{Jiang19multidomain}.
-Unsupervised MDA is considered in \citet{Farajian17multidomain} as an unsupervised DA problem.
+Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \citep{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citet{Jiang19multidomain}.
+Unsupervised MDA is considered in \citet{Farajian17multidomain} as an instance of unsupervised DA.
 
-Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of \emph{curriculum learning} (CL) is initially introduced.
-CL relies on the notion of ``easiness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
-CL has been shown to speed up learning.
-While the initial aim was to improve and speed up training, it has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitions of ``easiness''.
+Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of curriculum learning is initially introduced.
+CL relies on the notion of ``easyness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
+Various ways to automate CL using the framework of multi-armed bandits are notably explored in \citep{Graves17automated}, which has been an inspiration for our implementation.
+While the initial aim was primarily to improve and speed up training, CL has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
 
@@
 
-\section*{Acknowledgments}
+\section{Conclusion and future work}
+
+\fyDone{Write conclusion}
+In this study, we have presented a generic framework to perform a variety of standard adaptation tasks for machine translation, ranging from the conventional supervised domain adaptation to multi-domain adaptation and few shot DA.
+By experimenting with all these settings, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during  the course of training, could be used in all these situations.
+This algorithm, we believe, provides us with a more sound approach to (multi-domain) DA that existing heuristics, and dispenses with the costly search of optimal meta-parameters.
+Another contribution has been an experimental comparison of several recent approachs for dynamic data selection.
+
+In our future work, we intend to continue develop this approach and improve its effectiveness.
+One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training, so as to slow down convergence \cite{Kumar19reinforcement}.
+Another area where we need to progress is the unsupervised learning setting of Section~\ref{ssec:clda}, where our results still lag way behind that of supervised DA - this might be due to the inability of our simplistic optimization strategy to handle situations where the number of cluster is too large.
+% slow down learning during warm up or normalize reward; handle large number of classes;
+% too much importance to the dev ? randomize ?
+
+% \section*{Acknowledgments}
 \bibliographystyle{acl_natbib}

### 2021-05-14 2021/emnlp/auto_cl_mdmt.tex
remove some spelling errors

@@
 
-The second observation is that DDS is again outperformed by MDAC with a significant margin of 0.5 BLEU on average. However, MDAC is surpassed by far with 1.8 BLEU in the domain \domain{med} by DDS.
+The second observation is that DDS is again outperformed by MDAC with a significant margin of 1.5 BLEU on average. However, MDAC is surpassed by far with 1.6 BLEU in the domain \domain{med} by DDS.
 
@@
   \end{tabular}
-  \caption{Non empty columns correspond to the target domains. For a given line, each non empty column correspond to one expert system. ($^*$)MDAC is significantly better than DDS.}
+  \caption{For a given line, each column correspond to one expert system. ($^*$)MDAC is significantly better than DDS.}
   \label{tab:unsupervised-da}
@@
 According to Table \ref{tab:unsupervised-da}, MDAC underperforms the same version under the domain adaptation setting \ref{ssec:da}.
-However, automatic clustering could make the situations harder than the conventional settings where
+However, automatic clustering could make the situations harder than the conventional settings where \fyTodo{more discussion here}.
 
@@
 Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of curriculum learning is initially introduced.
-CL relies on the notion of ``easyness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
+CL relies on the notion of ``easiness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
 Various ways to automate CL using the framework of multi-armed bandits are notably explored in \citep{Graves17automated}, which has been an inspiration for our implementation.

### 2021-05-14 2021/emnlp/auto_cl_mdmt.tex
change comments in section results

--- 2021-05-14 2021/emnlp/auto_cl_mdmt.tex
@@
 \subsection{Domain Adaptation}\label{ssec:da}
-In this setting, we aim to build an expert NMT model in only one domain. This setting corresponds to Dirac distribution $\lambda^t$, in which the probability of domains is 0 except for one domain.
+In this setting, we aim to build an expert NMT model in only one domain.
+This corresponds to setting $\vlambda^t$ to a deterministic distribution, in which the probability of all domains is $0$ except for one domain.
+We denote this distribution $\vlambda_d$, with $d$ the domain of interest.
 
-We initialize MDAC and DDS by 3 different distributions $\lambda_0$, $\lambda_1$ and $\lambda_d$.
-According to Table \ref{tab:da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen, though $\lambda_0$  initialization only improve significantly $\lambda_d$ initialization in small domains including \domain{talk},\domain{bank}and\domain{it}.
-$\lambda_d$ initialization is beneficial for adapting the NMT system to large domains such as \domain{med} and \domain{law}.
-The same conclusion is also true for DDS.
+We initialize MDAC and DDS in 3 ways, using $\vlambda_0$, $\vlambda_1$ and $\vlambda_d$.
+According to Table~\ref{tab:da}, MDAC achieves the overall best performance when $\vlambda_{t=0} = \vlambda_0$.
+Doing so proves significantly better than initializing with $\lambda_d$ for the small domains: \domain{talk}, \domain{bank} and\domain{it}.
+Conversely, initializing with $\lambda_d$ is beneficial when targeting large domains such as \domain{med} and \domain{law}.
+The same conclusion also holds for DDS.
 

### 2021-05-14 2021/emnlp/auto_cl_mdmt.tex
finish section 5

@@
 This computation is inspired by the target prediction gain in \citep{Graves17automated}.
-However, \cite{Graves17automated} used accumulated gains from the past as reward, while we predict the usefulness of each domain for improving the future performance of the system given its current state.
+However, where \cite{Graves17automated} used accumulated gains from the past as reward, we instead predict the usefulness of each domain for improving the future performance of the system given its current state.
 This is achieved by simulating a round of training with only the data from one domain.
@@
 Additional choices for the hyper-parameters of our method include the number of simulation steps $k=10$ and the learning rate $\mathbf{lr}_{data}=0.001$.
-We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section \ref{ssec:clda}), where we use 20 gradient descent iterations.\fyTodo{Why?
+We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section~\ref{ssec:clda}), where we use 20~gradient descent iterations to slow down convergence.\fyTodo{Why?
 computation?} We split the training into 100 short sessions that last 2000 training steps each.
@@
 \subsection{Experimental tasks}
-We evaluate and compare our method with multiple baselines in 5~tasks including: domain adaptation; multi-domain adaptation, bi-domain adaptation, unseen domain adaptation, and unsupervised domain adaptation.
+We evaluate and compare our method with multiple baselines in the 5 following conditions.
+%including: domain adaptation; multi-domain adaptation, bi-domain adaptation, unseen domain adaptation, and unsupervised domain adaptation.
+In the \emph{supervised domain adaptation task}, given the data from 6 domains (\domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}), we aim to build distinct expert NMT models for each domain.
+To challenge the flexibility of the method, we also consider a \emph{bi-domain adaptation task}, where given the same 6 domains, we only focus on adapating only to 2 domains.
 
-In the domain adaptation task, given the data from 6 domains (\domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}), we aim to build distinct expert NMT models for each domain.
+In the \emph{multi-domain adaptation task}, given the same 6 domains, we aim to build one single NMT model that would perform optimally assuming a uniform distribution of domains during the test.
 
-In the multi-domain adaptation task, given the same 6 domains, we aim to build one single NMT model that would perform optimally assuming a uniform distribution of domains during the test.
+In a fourth experiment (\emph{few shot adaptation}), given training data in 6 domains and a small development set of a new domain, (\domain{News} in our case), we aim to build an NMT model which performs well for the unseen domain.
 
-In the bi-domain adaptation task, given again same 6 domains, we only focus on adapating only to 2 domains.
-
-In the fourth experiment, given training data in 6 domains and a small development set of an new domain, (\domain{News} in our case) we aim to build an NMT model which will perform well on test data from the unseen domain.
-
-Finally, in the unsupervised domain adaptation task, we cluster all available training data into 30 clusters using the KNN algorithm in the same way as in the work of \citet{Tars18multidomain}, then adapt these clusters to one of 6 domains given the corresponding in-domain dev set.
-We compare MDAC to DDS over 6 domains.
+Finally, in the \emph{unsupervised domain adaptation task}, we cluster all available training data into 30 clusters using the KNN algorithm in the same way as in \citep{Tars18multidomain}, then adapt these clusters to one of 6 domains using the corresponding in-domain dev set.
+We compare MDAC to DDS for each our 6 test sets.
 
@@
 
-We initialize MDAC and DDS in 3 ways, using $\vlambda_0$, $\vlambda_1$ and $\vlambda_d$.
+We consider three initializations for MDAC and DDS, using $\vlambda_0$, $\vlambda_1$ and $\vlambda_d$.
 According to Table~\ref{tab:da}, MDAC achieves the overall best performance when $\vlambda_{t=0} = \vlambda_0$.
-Doing so proves significantly better than initializing with $\lambda_d$ for the small domains: \domain{talk}, \domain{bank} and\domain{it}.
-Conversely, initializing with $\lambda_d$ is beneficial when targeting large domains such as \domain{med} and \domain{law}.
-The same conclusion also holds for DDS.
+Doing so proves much better than initializing with $\vlambda_d$ for small domains: \domain{talk}, \domain{bank} and\domain{it}.
+Conversely, initializing with $\vlambda_d$ is beneficial when targeting large domains such as \domain{med} and \domain{law}.
+The same conclusion holds for DDS.
 
-We now compare the best MDAC system (initialized with $\vlambda_{t=0} = \vlambda_0$) to full fine-tuning.
-According to Table~\ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}.
-However, MDAC improves fine-tuning by approximately 0.5 BLEU in \domain{bank} and 1.0 BLEU in \domain{rel}.
-This indicates that for small domains, out-of-domain data can improve the generalization of the NMT model and MDAC is capable to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
-
-DDS also demonstrates promising results in this task but is always outperformed by MDAC overall. The same conclusion is also applied to the CL method.
+We now compare the best MDAC system (using $\vlambda_{t=0} = \vlambda_0$) to full fine-tuning.
+According to Table~\ref{tab:da}, fine-tuning is better for large domains such as \domain{med} and \domain{law}, while MDAC outperforms fine-tuning by approximately 0.5 BLEU for \domain{bank} and 1.0 BLEU for \domain{rel}.
+This indicates that for small domains, out-of-domain data helps improve the generalization and that MDAC is able to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
+The results for DDS display similar trends, but are always outperformed by MDAC.
+The behaviour of CL, which does only well the large domain \domain{med} lag somewhat behind.
 
@@
   \begin{tabular}{|l|*8{r|}} \hline
-    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline
-    FT-Full($d$) \hfill [BLEU] &40.3&63.8&54.4&38.5&52.0&91.0&56.7\\ \hline
+    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{avg.} \\ \hline
+    FT-Full($d$) &40.3&63.8&54.4&38.5&52.0&91.0&56.7\\ \hline
     \hline
@@
   \end{tabular}
-  \caption{We report scores of each method with 6 target domains, and compare with full fine-tuning: each column corresponds to a distinct system.
+  \caption{Domain adaptation experiments.
+We report BLEU scores of each method for 6 target domains and their average: each column corresponds to a distinct system.
 ($^{*}$) MDAC is significantly better than CL, fine-tuning and DDS with $p<0.05$.
 ($^{**}$) MDAC is significantly better than CL and DDS with $p<0.05$.
-($^{***}$) MDAC is significantly better than CL, fine-tuning with $p<0.05$}
+($^{***}$) MDAC is significantly better than CL, fine-tuning with $p<0.05$.}
   \label{tab:da}
@@
 
+\subsection{Bi-domain adaptation}\label{ssec:bida}
+
+\begin{table*}[htbp]
+  \centering \small
+  \begin{tabular}{|l|*7{r|}} \hline
+    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} \\ \hline \hline
+    \system{DDS($\vlambda_0, \vlambda_2$)}&39.5&-&-&-&50.1&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&39.1&-&-&-&51.8$^*$&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&60.8&53.3&-&-&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&61.9$^*$&54.5$^*$&-&-&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&37.9&-&91.3 \\ 
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&36.9&-&90.4 \\
+    \hline
+  \end{tabular}
+  \caption{Adpating to two domains. For a given line, non empty columns correspond to the pair of target domains. ($^*$) MDAC is significantly better than DDS with $p<0.05$.}
+  \label{tab:bi-da}
+\end{table*}
+
+In these control experiments, we showcase the flexibility of dynamic sampling and try to adapt to (arbitrary) pairs of target domains with equal weight and contrast our results with those of DDS.
+Results are in Table~\ref{tab:bi-da}.
+Here, MDAC significantly outperforms DDS in two settings (\domain{med+it} and \domain{law+bank}) while being surpassed in \domain{talk+rel}.
+% Allthough we showing that MDAC is capable to adapt to a variable set of target domains.
+
 \subsection{Multi-domain adaptation}\label{ssec:mda}
 
-Multi-domain adaptation task requires best averaged performance of the NMT model in 6 domains.
-This setting corresponds to the uniform test distribution $\lambda^t = \lambda_0$.
-In this situation, the method CL of \citet{Zhang19curriculum} can not be applied.
-We therefore only compare the performance of MDAC, DDS and several fixed training data distribution $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$.
-The definition of those notations can be found in the equation \ref{mixture:trn}.
+We now turn to a more realistic scenario and consider multi-domain adaptation, which aims to train one single system with optimal performance averaged over 6 domains.
+This setting targets a uniform test distribution $\lambda^t = \lambda_0$.
+In this situation, CL \citep{Zhang19curriculum} does not apply.
+We therefore only contrast the performance of MDAC, DDS and several fixed training data distribution $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$, where $\vlambda_{\alpha}$ is defined according to equation~\eqref{mixture:trn}.
 
-We initialize MDAC and DDS by 2 different distribution $\lambda_0$ and $\lambda_1$.
-According to Table \ref{tab:multi-da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen.
-The same conclusion is also true for DDS.
+We again initialize MDAC and DDS with two distribution $\lambda_0$ and $\lambda_1$.
+According to Table~\ref{tab:multi-da}, MDAC achieves the best performance with initial (uniform) $\lambda_0$.
+The same conclusion holds for DDS.
+For this configuration, MDAC outperforms static training distributions including $\big[ \lambda_0, \lambda_{0.75}, \lambda_{1.0}\big]$ by a significant margin, and performs slightly better than $\big[ \lambda_{0.25}, \lambda_{0.5} \big]$.
+This indicates that MDAC allows us to skip the somewhat heuristic choice of the optimal training mixture.
 
-MDAC outperforms static training distributions including $\big[ \lambda_0, \lambda_{0.75}, \lambda_{1.0}\big]$ with significant margin while performs slightly better than $\big[ \lambda_{0.25}, \lambda_{0.5} \big]$.
-This observation indicates that MDAC allows us to skip the heuristic choice for the training mixture of different domains or tasks.
-
-The second observation is that DDS is again outperformed by MDAC with a significant margin of 0.5 BLEU on average. However, MDAC is surpassed by far with 1.8 BLEU in the domain \domain{med} by DDS.
-
-We show in the below figure \ref{fig:sampling}, the evolution of the training mixture through training sessions.
-According to the figure \ref{fig:DDS} of DDS, the proportion of 6 domains converges rapidly to bi-domain mode in which only \domain{med} and \domain{rel} have significant probability.
-In contrast, the proportion of 6 domains computed by MDAC evolves smoothly; small domains including \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} take larger proportion in the beginning then decrease their proportions while larger domains as \domain{med} and \domain{law} increase their proportions in the end where the NMT model might have been already close to the optimal performance in the smaller domains.
+A second observation is that DDS is again outperformed by MDAC by a significant margin of 1.5 BLEU on average; the only domain where DDS does (much) better is \domain{med}.
+Figure~\ref{fig:sampling}, where we plot the evolution of the training mixture in the course of training sessions, helps understand the difference between the two methods.
+For DDS (Figure~\ref{fig:DDS}), the proportion of 6 domains quicly converges towards a bi-domain mode in which only \domain{med} and \domain{rel} have significant probability -- hence the good performance on the former domain.
+In contrast, the proportion of 6~domains computed by MDAC evolves more smoothly; small domains such as \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} receive a larger proportion of training data in the early stages; their weights then slowly decrease as larger domains such as \domain{med} and \domain{law} increase their share.
+This however only happends at the end of training, when the NMT models might have been already close to their optimal performance for the small domains.
 
@@
   \end{tabular}
-  \caption{We report the performance of our method MDAC and the work of \citet{Wang20balancing} DDS in multi-domain adaptation task.
+  \caption{Multi domain adaptation.
 For a given line, all the columns \emph{correspond to the same multi-domain system}.
@@
 \end{subfigure}
-\caption{Sampling distribution}
+\caption{Evolution of the sampling distribution during training.}
 \label{fig:sampling}
 \end{figure*}
-
-\subsection{Bi-domain adaptation}\label{ssec:bida}
-
-\begin{table*}[htbp]
-  \centering \small
-  \begin{tabular}{|l|*9{r|}} \hline
-    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} \\ \hline \hline
-    \system{DDS($\vlambda_0, \vlambda_2$)}&39.5&-&-&-&50.1&-&-&- \\
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&39.1&-&-&-&51.8$^*$&-&-&- \\
-    \system{DDS($\vlambda_0, \vlambda_2$)}&-&60.8&53.3&-&-&-&-&- \\
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&61.9$^*$&54.5$^*$&-&-&-&-&- \\
-    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&37.9&-&91.3&-&- \\ 
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&36.9&-&90.4&-&- \\
-    \hline
-  \end{tabular}
-  \caption{Non empty columns correspond to the target domains.
-For a given line, all the non empty columns \emph{correspond to the same bi-domain system}.
-($^*$) MDAC is significantly better than DDS with $p<0.05$.}
-  \label{tab:bi-da}
-\end{table*}
-According to Table \ref{tab:bi-da}, MDAC outperforms DDS in 2 over 3 settings including adapting to \domain{med-it} and \domain{law-bank} while being surpassed in \domain{talk-rel}, showing that MDAC is capable to adapt to a variable set of target domains.
 
@@
 \cline{1-2} \cline{4-11}
-    \multicolumn{2}{|c|}{\sl unseen domains} & &\multicolumn{8}{c|}{\sl 30 clusters} \\ 
+    \multicolumn{2}{|c|}{\sl Unseen domain} & &\multicolumn{8}{c|}{\sl Training with 30 clusters} \\ 
 \cline{1-2} \cline{4-11}
@@
   \end{tabular}
-  \caption{Non empty columns correspond to the target domains. For a given line, each non empty column correspond to one expert system. ($^*$)MDAC is significantly better than DDS.}
+  \caption{Few shot adaptation (left) and unsupervised adaptation (right). For a given line, each column correspond to one distinct system. ($^*$)MDAC is significantly better than DDS.}
   \label{tab:unsupervised-da}
@@
 
-\subsection{Unseen domains}\label{ssec:uda}
-The top of Table \ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and dynamic data selection MDAC and DDS.
+\subsection{Unseen domain}\label{ssec:uda}
+The left part of Table~\ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and wiht dynamic data selection (MDAC and DDS).
 These systems have insignificant differences in BLEU, showing that dynamic mixture does not improve the robustness of the NMT system against unseen domains.
-However, the fact that the performance of MDAC and DDS is close to the best performance guarantees the harmlessness of these algorithms to the robustness of the NMT model.
+However, the fact that the performance of MDAC and DDS is close to the best performance is a sign that they can also apply in such challengin situation.
+% the robustness of the NMT model.
 
 \subsection{Automatic clustering}\label{ssec:clda}
-The bottom of Table \ref{tab:unsupervised-da} shows the performance of NMT systems adapted to each domain in using 30 clusters instead of 6 domains.
-According to Table \ref{tab:unsupervised-da}, MDAC underperforms the same version under the domain adaptation setting \ref{ssec:da}.
-However, automatic clustering could make the situations harder than the conventional settings where
+The right part of Table~\ref{tab:unsupervised-da} reports the performance of NMT systems adapted to each domain.
+In comparison to section~\ref{ssec:da}, the training data is distributed in 30~clusters instead of the 6 original domains.
+Using a finer-grain spliting of data provides the learner with extra degrees of freedom when selecting the best distribution.
+However, as these clusters are built automatically, they do not necessarily contain data from just one domain, and are more noisy in nature.
+According to the results in Table~\ref{tab:unsupervised-da}, this scenario is hard both for DDS and MDAC which perform much worse than for the supervised DA setting.
+This again signals the importance of initialization: analyzing the clustering, we find that the data for \domain{rel} mostly correspond to one single cluster.
+Using a uniform initialization, this cluster starts with a very small weight, and never succeeds to yield the (good) performance observed in the DA settting.
 
@@
 Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \citep{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citet{Jiang19multidomain}.
-Unsupervised MDA is considered in \citet{Farajian17multidomain} as an instance of unsupervised DA.
+Unsupervised MDA is studied in \citet{Farajian17multidomain}, as an instance of unsupervised DA.
 
@@
 
-For instance, \citet{Zhang19curriculum} study supervised domain adaptation, and propose a curriculum approach which progressively extends the data used in training: in the early stages only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
-This is somehow opposite to the recommendations of \citet{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyTodo{These have not been compared?
+For instance, \citet{Zhang19curriculum} study supervised domain adaptation and propose a curriculum approach which progressively extends the data used in training: in the early stages only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
+This is somehow opposite to the recommendations of \citet{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyDone{These have not been compared?
 and also to what we do ?}
@@
 \citet{Kumar19reinforcement} use reinforcement techniques (deep Q-learning) to learn the curriculum strategy: in this study, complexity corresponds to difficulty levels and which are binned based on contrastive data selection.
-The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyTodo{Alert: what do we do during warm-up ?} The same technique is recently applied to the training of a multilingual NMT system \citep{Kumar21learning}.
+The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyDone{Alert: what do we do during warm-up ?} The same technique is recently applied to the training of a multilingual NMT system \citep{Kumar21learning}.
 \citet{Zhou20uncertainty} propose another curriculum-based approach which instead relies on \emph{instance uncertainty} as a measure of their difficulty and presents the data sample starting with the easiest (more predictable) first.
@@
 
-\section{Conclusion and future work}
+\section{Conclusion and outlook \label{sec:conclusion}}
 
@@
 This algorithm, we believe, provides us with a more sound approach to (multi-domain) DA that existing heuristics, and dispenses with the costly search of optimal meta-parameters.
-Another contribution has been an experimental comparison of several recent approachs for dynamic data selection.
+Another contribution of our work is an experimental comparison of recent approachs to dynamic data selection.
 
-In our future work, we intend to continue develop this approach and improve its effectiveness.
-One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training, so as to slow down convergence \cite{Kumar19reinforcement}.
-Another area where we need to progress is the unsupervised learning setting of Section~\ref{ssec:clda}, where our results still lag way behind that of supervised DA - this might be due to the inability of our simplistic optimization strategy to handle situations where the number of cluster is too large.
+In the future, we intend to continue develop this approach and improve its effectiveness.
+One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training \cite{Kumar19reinforcement}.
+Another area where we need to progress is the unsupervised learning setting of Section~\ref{ssec:clda}, where our results still lag way behind that of supervised DA - this might be due to the inability of our simplistic optimization strategy to handle situations where the number of clusters is too large.
 % slow down learning during warm up or normalize reward; handle large number of classes;

### 2021-05-14 2021/emnlp/auto_cl_mdmt.tex
still too long

@@
 \usepackage{adjustbox}
+\renewcommand{\floatpagefraction}{.9}
 
@@
 \system{Mixed-$0$} uses a uniform domain distribution, \system{Mixed-$1.0$} uses the observed domain distribution.
-\item finetuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ??
+\item fine-tuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ??
 iterations} The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems, each adapted for one domain, with no parameter sharing across domains.
-\item Our comparison of multi-domain systems\fyTodo{CL is only dynamic ?} includes baselines with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning (CL) of \citet{Zhang19curriculum} and Differential Data Selection (DDS) of \citet{Wang20balancing}.
+\item systems trained with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$; these are used in the multi-domain experiments of Section~\ref{ssec:mda};
+\item  our own implementations of recent dynamic sampling proposals from the literature: Curriculum Learning (CL) of \citet{Zhang19curriculum} and Differential Data Selection (DDS) of \citet{Wang20balancing} (see details below);
 \end{itemize}
 
-All models use embeddings and hidden layers sizes of dimension~512.
+All models use embeddings and hidden layers of dimension~512.
 Transformer models contain 8~attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells.
 Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.
-We train all systems with 200~000 iterations.
+We train all systems with 200k iterations.
+
+\subsection{CL and DDS's re-implementation}\fyDone{Right place for this ?}
+We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to original code of \citet{Wang20balancing}.\footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}
+
+We also re-implement the approach of \citet{Zhang19curriculum} according to the authors' description of the algorithm.
+For each domain adaptation experiment, we combine the training data of all other domains into one corpus, then compute the cross-entropy difference score of each source sentence of this combined dataset.
+We then sort and split the corpus into 9 shards and execute curriculum learning with 10 shards, using the in-domain corpus as the first shard.
 
@@
 The behavior of MDAC only depends on (a) the initial domain distribution at the start of training $\vlambda^{l}_{t=0}$, and (b) the targeted (dev/test) distribution $\vlambda^{t}$.
-We will thus report these systems as \system{MDAC($\vlambda^{l}_{t=0}$, $\vlambda^{t}$)}.
-We also compare our method with DDS under the same settings.
+We thus report these systems as \system{MDAC($\vlambda^{l}_{t=0}$, $\vlambda^{t}$)} and compare with DDS using the same settings.
 
-In our work, we simply parameterize the distribution $\lambda^l$ as follows (with $\beta=2$ in all our experiments):\fyTodo{Explain why}
+In our work, we parameterize the distribution $\lambda^l$ as follows (with $\beta=2$ in all experiments):\fyFuture{Explain why}
 \begin{equation}
@@
 Additional choices for the hyper-parameters of our method include the number of simulation steps $k=10$ and the learning rate $\mathbf{lr}_{data}=0.001$.
-We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section~\ref{ssec:clda}), where we use 20~gradient descent iterations to slow down convergence.\fyTodo{Why?
+We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section~\ref{ssec:clda}), where we use 20~gradient descent iterations to slow down convergence.\fyFuture{Why?
 computation?} We split the training into 100 short sessions that last 2000 training steps each.
 The choice of those hyper-parameters is mostly heuristic except for the learning rate $\mathbf{lr}_{data}$ which is optimized via grid search over a range of values $[0.001,0.0025,0.005]$.
-
-\subsection{CL and DDS's re-implementation}\fyTodo{Right place for this ?}
-We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to original code of \citet{Wang20balancing}.\footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}
-
-We also re-implement the work of \citet{Zhang19curriculum} according to the authors' description of the algorithm.
-For each domain adaptation experiment, we combine the training data of all other domains into one corpus, then compute the cross-entropy difference score of each source sentence of this combined dataset.
-We then sort and split the corpus into 9 shards and execute curriculum learning with 10 shards, using the in-domain corpus as the first shard.
 
@@
 \subsection{Domain Adaptation}\label{ssec:da}
-In this setting, we aim to build an expert NMT model in only one domain.
-This corresponds to setting $\vlambda^t$ to a deterministic distribution, in which the probability of all domains is $0$ except for one domain.
-We denote this distribution $\vlambda_d$, with $d$ the domain of interest.
+In this setting, we aim to build an NMT model for one single domain: we accordingly set $\vlambda^t$ to a deterministic distribution $\vlambda_d$, where the target domain $d$ has probability~1.
 
@@
 
-\begin{table*}[htbp]
-  \centering \small
-  \begin{tabular}{|l|*7{r|}} \hline
-    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} \\ \hline \hline
-    \system{DDS($\vlambda_0, \vlambda_2$)}&39.5&-&-&-&50.1&- \\
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&39.1&-&-&-&51.8$^*$&- \\
-    \system{DDS($\vlambda_0, \vlambda_2$)}&-&60.8&53.3&-&-&- \\
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&61.9$^*$&54.5$^*$&-&-&- \\
-    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&37.9&-&91.3 \\ 
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&36.9&-&90.4 \\
-    \hline
-  \end{tabular}
-  \caption{Adpating to two domains. For a given line, non empty columns correspond to the pair of target domains. ($^*$) MDAC is significantly better than DDS with $p<0.05$.}
-  \label{tab:bi-da}
-\end{table*}
-
 In these control experiments, we showcase the flexibility of dynamic sampling and try to adapt to (arbitrary) pairs of target domains with equal weight and contrast our results with those of DDS.
@@
 This however only happends at the end of training, when the NMT models might have been already close to their optimal performance for the small domains.
+
+\begin{table*}[htbp]
+  \centering \small
+  \begin{tabular}{|l|*7{r|}} \hline
+    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} \\ \hline \hline
+    \system{DDS($\vlambda_0, \vlambda_2$)}&39.5&-&-&-&50.1&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&39.1&-&-&-&51.8$^*$&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&60.8&53.3&-&-&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&61.9$^*$&54.5$^*$&-&-&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&37.9&-&91.3 \\ 
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&36.9&-&90.4 \\
+    \hline
+  \end{tabular}
+  \caption{Adapting to two domains. For a given line, non empty columns correspond to the pair of target domains. ($^*$) MDAC is significantly better than DDS with $p<0.05$.}
+  \label{tab:bi-da}
+\end{table*}
 
@@
 In comparison to section~\ref{ssec:da}, the training data is distributed in 30~clusters instead of the 6 original domains.
-Using a finer-grain spliting of data provides the learner with extra degrees of freedom when selecting the best distribution.
-However, as these clusters are built automatically, they do not necessarily contain data from just one domain, and are more noisy in nature.
+Spliting the train data in small groups provides the learner with extra degrees of freedom when selecting the best distribution.
+However, as these clusters are built automatically, they are more noisy in nature.
 According to the results in Table~\ref{tab:unsupervised-da}, this scenario is hard both for DDS and MDAC which perform much worse than for the supervised DA setting.

### 2021-05-14 2021/emnlp/auto_cl_mdmt.tex
local changes

--- 2021-05-14 2021/emnlp/auto_cl_mdmt.tex
@@
 % \theta_{t+i-1} - \mathbf{lr_2} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta}
-In these equations, $N$ denotes the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $\mathbf{lr_2}$ is the learning rate of the NMT system; $l(\theta,x,y)$ is the loss of the NMT model on sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss aggregated over $n_d$ dev-sets corresponding to the $n_d$ domains.
+In these equations, $N$ denotes the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $l(\theta,x,y)$ is the loss of the NMT model on sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss aggregated over $n_d$ dev-sets corresponding to the $n_d$ domains.
 
@@
 This computation is inspired by the target prediction gain in \citep{Graves17automated}.
-However, where \cite{Graves17automated} used accumulated gains from the past as reward, we instead predict the usefulness of each domain for improving the future performance of the system given its current state.
+However, \cite{Graves17automated} used accumulated gains from the past as reward, while we predict the usefulness of each domain for improving the future performance of the system given its current state.
 This is achieved by simulating a round of training with only the data from one domain.
@@
 Additional choices for the hyper-parameters of our method include the number of simulation steps $k=10$ and the learning rate $\mathbf{lr}_{data}=0.001$.
-We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section~\ref{ssec:clda}), where we use 20~gradient descent iterations to slow down convergence.\fyTodo{Why?
+We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section \ref{ssec:clda}), where we use 20 gradient descent iterations.\fyTodo{Why?
 computation?} We split the training into 100 short sessions that last 2000 training steps each.
@@
 \subsection{Experimental tasks}
-We evaluate and compare our method with multiple baselines in the 5 following conditions.
-%including: domain adaptation; multi-domain adaptation, bi-domain adaptation, unseen domain adaptation, and unsupervised domain adaptation.
-In the \emph{supervised domain adaptation task}, given the data from 6 domains (\domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}), we aim to build distinct expert NMT models for each domain.
-To challenge the flexibility of the method, we also consider a \emph{bi-domain adaptation task}, where given the same 6 domains, we only focus on adapating only to 2 domains.
+We evaluate and compare our method with multiple baselines in 5~tasks including: domain adaptation; multi-domain adaptation, bi-domain adaptation, unseen domain adaptation, and unsupervised domain adaptation.
 
-In the \emph{multi-domain adaptation task}, given the same 6 domains, we aim to build one single NMT model that would perform optimally assuming a uniform distribution of domains during the test.
+In the domain adaptation task, given the data from 6 domains (\domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}), we aim to build distinct expert NMT models for each domain.
 
-In a fourth experiment (\emph{few shot adaptation}), given training data in 6 domains and a small development set of a new domain, (\domain{News} in our case), we aim to build an NMT model which performs well for the unseen domain.
+In the multi-domain adaptation task, given the same 6 domains, we aim to build one single NMT model that would perform optimally assuming a uniform distribution of domains during the test.
 
-Finally, in the \emph{unsupervised domain adaptation task}, we cluster all available training data into 30 clusters using the KNN algorithm in the same way as in \citep{Tars18multidomain}, then adapt these clusters to one of 6 domains using the corresponding in-domain dev set.
-We compare MDAC to DDS for each our 6 test sets.
+In the bi-domain adaptation task, given again same 6 domains, we only focus on adapating only to 2 domains.
+
+In the fourth experiment, given training data in 6 domains and a small development set of an new domain, (\domain{News} in our case) we aim to build an NMT model which will perform well on test data from the unseen domain.
+
+Finally, in the unsupervised domain adaptation task, we cluster all available training data into 30 clusters using the KNN algorithm in the same way as in the work of \citet{Tars18multidomain}, then adapt these clusters to one of 6 domains given the corresponding in-domain dev set.
+We compare MDAC to DDS over 6 domains.
 
@@
 \subsection{Domain Adaptation}\label{ssec:da}
-In this setting, we aim to build an expert NMT model in only one domain.
-This corresponds to setting $\vlambda^t$ to a deterministic distribution, in which the probability of all domains is $0$ except for one domain.
-We denote this distribution $\vlambda_d$, with $d$ the domain of interest.
+In this setting, we aim to build an expert NMT model in only one domain. This setting corresponds to Dirac distribution $\lambda^t$, in which the probability of domains is 0 except for one domain.
 
-We consider three initializations for MDAC and DDS, using $\vlambda_0$, $\vlambda_1$ and $\vlambda_d$.
-According to Table~\ref{tab:da}, MDAC achieves the overall best performance when $\vlambda_{t=0} = \vlambda_0$.
-Doing so proves much better than initializing with $\vlambda_d$ for small domains: \domain{talk}, \domain{bank} and\domain{it}.
-Conversely, initializing with $\vlambda_d$ is beneficial when targeting large domains such as \domain{med} and \domain{law}.
-The same conclusion holds for DDS.
+We initialize MDAC and DDS by 3 different distributions $\lambda_0$, $\lambda_1$ and $\lambda_d$.
+According to Table \ref{tab:da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen, though $\lambda_0$  initialization only improve significantly $\lambda_d$ initialization in small domains including \domain{talk},\domain{bank}and\domain{it}.
+$\lambda_d$ initialization is beneficial for adapting the NMT system to large domains such as \domain{med} and \domain{law}.
+The same conclusion is also true for DDS.
 
-We now compare the best MDAC system (using $\vlambda_{t=0} = \vlambda_0$) to full fine-tuning.
-According to Table~\ref{tab:da}, fine-tuning is better for large domains such as \domain{med} and \domain{law}, while MDAC outperforms fine-tuning by approximately 0.5 BLEU for \domain{bank} and 1.0 BLEU for \domain{rel}.
-This indicates that for small domains, out-of-domain data helps improve the generalization and that MDAC is able to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
-The results for DDS display similar trends, but are always outperformed by MDAC.
-The behaviour of CL, which does only well the large domain \domain{med} lag somewhat behind.
+We now compare the best MDAC system (initialized with $\vlambda_{t=0} = \vlambda_0$) to full fine-tuning.
+According to Table~\ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}.
+However, MDAC improves fine-tuning by approximately 0.5 BLEU in \domain{bank} and 1.0 BLEU in \domain{rel}.
+This indicates that for small domains, out-of-domain data can improve the generalization of the NMT model and MDAC is capable to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
+
+DDS also demonstrates promising results in this task but is always outperformed by MDAC overall. The same conclusion is also applied to the CL method.
 
@@
   \begin{tabular}{|l|*8{r|}} \hline
-    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{avg.} \\ \hline
-    FT-Full($d$) &40.3&63.8&54.4&38.5&52.0&91.0&56.7\\ \hline
+    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline
+    FT-Full($d$) \hfill [BLEU] &40.3&63.8&54.4&38.5&52.0&91.0&56.7\\ \hline
     \hline
@@
   \end{tabular}
-  \caption{Domain adaptation experiments.
-We report BLEU scores of each method for 6 target domains and their average: each column corresponds to a distinct system.
+  \caption{We report scores of each method with 6 target domains, and compare with full fine-tuning: each column corresponds to a distinct system.
 ($^{*}$) MDAC is significantly better than CL, fine-tuning and DDS with $p<0.05$.
 ($^{**}$) MDAC is significantly better than CL and DDS with $p<0.05$.
-($^{***}$) MDAC is significantly better than CL, fine-tuning with $p<0.05$.}
+($^{***}$) MDAC is significantly better than CL, fine-tuning with $p<0.05$}
   \label{tab:da}
@@
 
-\subsection{Bi-domain adaptation}\label{ssec:bida}
-
-\begin{table*}[htbp]
-  \centering \small
-  \begin{tabular}{|l|*7{r|}} \hline
-    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} \\ \hline \hline
-    \system{DDS($\vlambda_0, \vlambda_2$)}&39.5&-&-&-&50.1&- \\
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&39.1&-&-&-&51.8$^*$&- \\
-    \system{DDS($\vlambda_0, \vlambda_2$)}&-&60.8&53.3&-&-&- \\
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&61.9$^*$&54.5$^*$&-&-&- \\
-    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&37.9&-&91.3 \\ 
-    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&36.9&-&90.4 \\
-    \hline
-  \end{tabular}
-  \caption{Adpating to two domains. For a given line, non empty columns correspond to the pair of target domains. ($^*$) MDAC is significantly better than DDS with $p<0.05$.}
-  \label{tab:bi-da}
-\end{table*}
-
-In these control experiments, we showcase the flexibility of dynamic sampling and try to adapt to (arbitrary) pairs of target domains with equal weight and contrast our results with those of DDS.
-Results are in Table~\ref{tab:bi-da}.
-Here, MDAC significantly outperforms DDS in two settings (\domain{med+it} and \domain{law+bank}) while being surpassed in \domain{talk+rel}.
-% Allthough we showing that MDAC is capable to adapt to a variable set of target domains.
-
 \subsection{Multi-domain adaptation}\label{ssec:mda}
 
-We now turn to a more realistic scenario and consider multi-domain adaptation, which aims to train one single system with optimal performance averaged over 6 domains.
-This setting targets a uniform test distribution $\lambda^t = \lambda_0$.
-In this situation, CL \citep{Zhang19curriculum} does not apply.
-We therefore only contrast the performance of MDAC, DDS and several fixed training data distribution $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$, where $\vlambda_{\alpha}$ is defined according to equation~\eqref{mixture:trn}.
+Multi-domain adaptation task requires best averaged performance of the NMT model in 6 domains.
+This setting corresponds to the uniform test distribution $\lambda^t = \lambda_0$.
+In this situation, the method CL of \citet{Zhang19curriculum} can not be applied.
+We therefore only compare the performance of MDAC, DDS and several fixed training data distribution $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$.
+The definition of those notations can be found in the equation \ref{mixture:trn}.
 
-We again initialize MDAC and DDS with two distribution $\lambda_0$ and $\lambda_1$.
-According to Table~\ref{tab:multi-da}, MDAC achieves the best performance with initial (uniform) $\lambda_0$.
-The same conclusion holds for DDS.
-For this configuration, MDAC outperforms static training distributions including $\big[ \lambda_0, \lambda_{0.75}, \lambda_{1.0}\big]$ by a significant margin, and performs slightly better than $\big[ \lambda_{0.25}, \lambda_{0.5} \big]$.
-This indicates that MDAC allows us to skip the somewhat heuristic choice of the optimal training mixture.
+We initialize MDAC and DDS by 2 different distribution $\lambda_0$ and $\lambda_1$.
+According to Table \ref{tab:multi-da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen.
+The same conclusion is also true for DDS.
 
-A second observation is that DDS is again outperformed by MDAC by a significant margin of 1.5 BLEU on average; the only domain where DDS does (much) better is \domain{med}.
-Figure~\ref{fig:sampling}, where we plot the evolution of the training mixture in the course of training sessions, helps understand the difference between the two methods.
-For DDS (Figure~\ref{fig:DDS}), the proportion of 6 domains quicly converges towards a bi-domain mode in which only \domain{med} and \domain{rel} have significant probability -- hence the good performance on the former domain.
-In contrast, the proportion of 6~domains computed by MDAC evolves more smoothly; small domains such as \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} receive a larger proportion of training data in the early stages; their weights then slowly decrease as larger domains such as \domain{med} and \domain{law} increase their share.
-This however only happends at the end of training, when the NMT models might have been already close to their optimal performance for the small domains.
+MDAC outperforms static training distributions including $\big[ \lambda_0, \lambda_{0.75}, \lambda_{1.0}\big]$ with significant margin while performs slightly better than $\big[ \lambda_{0.25}, \lambda_{0.5} \big]$.
+This observation indicates that MDAC allows us to skip the heuristic choice for the training mixture of different domains or tasks.
+
+The second observation is that DDS is again outperformed by MDAC with a significant margin of 1.5 BLEU on average. However, MDAC is surpassed by far with 1.6 BLEU in the domain \domain{med} by DDS.
+
+We show in the below figure \ref{fig:sampling}, the evolution of the training mixture through training sessions.
+According to the figure \ref{fig:DDS} of DDS, the proportion of 6 domains converges rapidly to bi-domain mode in which only \domain{med} and \domain{rel} have significant probability.
+In contrast, the proportion of 6 domains computed by MDAC evolves smoothly; small domains including \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} take larger proportion in the beginning then decrease their proportions while larger domains as \domain{med} and \domain{law} increase their proportions in the end where the NMT model might have been already close to the optimal performance in the smaller domains.
 
@@
   \end{tabular}
-  \caption{Multi domain adaptation.
+  \caption{We report the performance of our method MDAC and the work of \citet{Wang20balancing} DDS in multi-domain adaptation task.
 For a given line, all the columns \emph{correspond to the same multi-domain system}.
@@
 \end{subfigure}
-\caption{Evolution of the sampling distribution during training.}
+\caption{Sampling distribution}
 \label{fig:sampling}
 \end{figure*}
+
+\subsection{Bi-domain adaptation}\label{ssec:bida}
+
+\begin{table*}[htbp]
+  \centering \small
+  \begin{tabular}{|l|*9{r|}} \hline
+    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} \\ \hline \hline
+    \system{DDS($\vlambda_0, \vlambda_2$)}&39.5&-&-&-&50.1&-&-&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&39.1&-&-&-&51.8$^*$&-&-&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&60.8&53.3&-&-&-&-&- \\
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&61.9$^*$&54.5$^*$&-&-&-&-&- \\
+    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&37.9&-&91.3&-&- \\ 
+    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&36.9&-&90.4&-&- \\
+    \hline
+  \end{tabular}
+  \caption{Non empty columns correspond to the target domains.
+For a given line, all the non empty columns \emph{correspond to the same bi-domain system}.
+($^*$) MDAC is significantly better than DDS with $p<0.05$.}
+  \label{tab:bi-da}
+\end{table*}
+According to Table \ref{tab:bi-da}, MDAC outperforms DDS in 2 over 3 settings including adapting to \domain{med-it} and \domain{law-bank} while being surpassed in \domain{talk-rel}, showing that MDAC is capable to adapt to a variable set of target domains.
 
@@
 \cline{1-2} \cline{4-11}
-    \multicolumn{2}{|c|}{\sl Unseen domain} & &\multicolumn{8}{c|}{\sl Training with 30 clusters} \\ 
+    \multicolumn{2}{|c|}{\sl unseen domains} & &\multicolumn{8}{c|}{\sl 30 clusters} \\ 
 \cline{1-2} \cline{4-11}
@@
   \end{tabular}
-  \caption{Few shot adaptation (left) and unsupervised adaptation (right). For a given line, each column correspond to one distinct system. ($^*$)MDAC is significantly better than DDS.}
+  \caption{For a given line, each column correspond to one expert system. ($^*$)MDAC is significantly better than DDS.}
   \label{tab:unsupervised-da}
@@
 
-\subsection{Unseen domain}\label{ssec:uda}
-The left part of Table~\ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and wiht dynamic data selection (MDAC and DDS).
+\subsection{Unseen domains}\label{ssec:uda}
+The top of Table \ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and dynamic data selection MDAC and DDS.
 These systems have insignificant differences in BLEU, showing that dynamic mixture does not improve the robustness of the NMT system against unseen domains.
-However, the fact that the performance of MDAC and DDS is close to the best performance is a sign that they can also apply in such challengin situation.
-% the robustness of the NMT model.
+However, the fact that the performance of MDAC and DDS is close to the best performance guarantees the harmlessness of these algorithms to the robustness of the NMT model.
 
 \subsection{Automatic clustering}\label{ssec:clda}
-The right part of Table~\ref{tab:unsupervised-da} reports the performance of NMT systems adapted to each domain.
-In comparison to section~\ref{ssec:da}, the training data is distributed in 30~clusters instead of the 6 original domains.
-Using a finer-grain spliting of data provides the learner with extra degrees of freedom when selecting the best distribution.
-However, as these clusters are built automatically, they do not necessarily contain data from just one domain, and are more noisy in nature.
-According to the results in Table~\ref{tab:unsupervised-da}, this scenario is hard both for DDS and MDAC which perform much worse than for the supervised DA setting.
-This again signals the importance of initialization: analyzing the clustering, we find that the data for \domain{rel} mostly correspond to one single cluster.
-Using a uniform initialization, this cluster starts with a very small weight, and never succeeds to yield the (good) performance observed in the DA settting.
+The bottom of Table \ref{tab:unsupervised-da} shows the performance of NMT systems adapted to each domain in using 30 clusters instead of 6 domains.
+According to Table \ref{tab:unsupervised-da}, MDAC underperforms the same version under the domain adaptation setting \ref{ssec:da}.
+However, automatic clustering could make the situations harder than the conventional settings where \fyTodo{more discussion here}.
 
@@
 Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \citep{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citet{Jiang19multidomain}.
-Unsupervised MDA is studied in \citet{Farajian17multidomain}, as an instance of unsupervised DA.
+Unsupervised MDA is considered in \citet{Farajian17multidomain} as an instance of unsupervised DA.
 
 Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of curriculum learning is initially introduced.
-CL relies on the notion of ``easyness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
+CL relies on the notion of ``easiness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
 Various ways to automate CL using the framework of multi-armed bandits are notably explored in \citep{Graves17automated}, which has been an inspiration for our implementation.
@@
 
-For instance, \citet{Zhang19curriculum} study supervised domain adaptation and propose a curriculum approach which progressively extends the data used in training: in the early stages only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
-This is somehow opposite to the recommendations of \citet{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyDone{These have not been compared?
+For instance, \citet{Zhang19curriculum} study supervised domain adaptation, and propose a curriculum approach which progressively extends the data used in training: in the early stages only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
+This is somehow opposite to the recommendations of \citet{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyTodo{These have not been compared?
 and also to what we do ?}
@@
 \citet{Kumar19reinforcement} use reinforcement techniques (deep Q-learning) to learn the curriculum strategy: in this study, complexity corresponds to difficulty levels and which are binned based on contrastive data selection.
-The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyDone{Alert: what do we do during warm-up ?} The same technique is recently applied to the training of a multilingual NMT system \citep{Kumar21learning}.
+The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyTodo{Alert: what do we do during warm-up ?} The same technique is recently applied to the training of a multilingual NMT system \citep{Kumar21learning}.
 \citet{Zhou20uncertainty} propose another curriculum-based approach which instead relies on \emph{instance uncertainty} as a measure of their difficulty and presents the data sample starting with the easiest (more predictable) first.
@@
 
-\section{Conclusion and outlook \label{sec:conclusion}}
+\section{Conclusion and future work}
 
@@
 This algorithm, we believe, provides us with a more sound approach to (multi-domain) DA that existing heuristics, and dispenses with the costly search of optimal meta-parameters.
-Another contribution of our work is an experimental comparison of recent approachs to dynamic data selection.
+Another contribution has been an experimental comparison of several recent approachs for dynamic data selection.
 
-In the future, we intend to continue develop this approach and improve its effectiveness.
-One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training \cite{Kumar19reinforcement}.
-Another area where we need to progress is the unsupervised learning setting of Section~\ref{ssec:clda}, where our results still lag way behind that of supervised DA - this might be due to the inability of our simplistic optimization strategy to handle situations where the number of clusters is too large.
+In our future work, we intend to continue develop this approach and improve its effectiveness.
+One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training, so as to slow down convergence \cite{Kumar19reinforcement}.
+Another area where we need to progress is the unsupervised learning setting of Section~\ref{ssec:clda}, where our results still lag way behind that of supervised DA - this might be due to the inability of our simplistic optimization strategy to handle situations where the number of cluster is too large.
 % slow down learning during warm up or normalize reward; handle large number of classes;

### 2021-05-15 2021/emnlp/auto_cl_mdmt.tex
Now the right length

--- 2021-05-14 2021/emnlp/auto_cl_mdmt.tex
@@
 % \revision{Answer: The graphic of \system{Mixed-0} shows that with the same amount of data(batches) the convergences of the domains vary significantly}
-
-% === The part on multidomain training has been moved to the end.
 
@@
   \end{tabular}
-  \caption{Few shot adaptation (left) and unsupervised adaptation (right). For a given line, each column correspond to one distinct system. ($^*$)MDAC is significantly better than DDS.}
+  \caption{Few shot adaptation (left) and unsupervised adaptation (right). For a given line, each column correspond to one distinct system. ($^*$) MDAC is significantly better than DDS.}
   \label{tab:unsupervised-da}
@@
 These systems have insignificant differences in BLEU, showing that dynamic mixture does not improve the robustness of the NMT system against unseen domains.
-However, the fact that the performance of MDAC and DDS is close to the best performance is a sign that they can also apply in such challengin situation.
+However, the fact that the performance of MDAC and DDS is close to the best performance is a sign that they can also apply in such challenging situation.
 % the robustness of the NMT model.
@@
 The right part of Table~\ref{tab:unsupervised-da} reports the performance of NMT systems adapted to each domain.
-In comparison to section~\ref{ssec:da}, the training data is distributed in 30~clusters instead of the 6 original domains.
+In comparison to Section~\ref{ssec:da}, the training data is distributed in 30 automatic clusters instead of the 6 original domains.
 Spliting the train data in small groups provides the learner with extra degrees of freedom when selecting the best distribution.
@@
 
-Domain adaptation is an ancient and vexing problem that has been studied from many angles both for SMT and NMT.
-A survey of supervised and unsupervised DA for NMT is in \citet{Chu18asurvey}, where they distinguish between data-centric and model-centric DA, a view also adopted in the more recent review of \citet{Saunders21domain}.
-Our approach to DA in this paper clearly falls under the former category.
+Domain adaptation is an old problem that has been studied from many angles both for SMT and NMT.
+A survey of supervised and unsupervised DA for NMT is in \citet{Chu18asurvey}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the more recent review of \citet{Saunders21domain}.
+Our approach to DA in this paper falls under the former category.
 We refer interested readers to these references and do not discuss DA any further.
@@
 Multi-domain adaptation (MDA) aims to develop systems that simultaneously bode well for several domains.
-Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \citep{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citet{Jiang19multidomain}.
+Like for DA, techniques for supervised MDA combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \citep{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citet{Jiang19multidomain}.
 Unsupervised MDA is studied in \citet{Farajian17multidomain}, as an instance of unsupervised DA.
 
-Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of curriculum learning is initially introduced.
-CL relies on the notion of ``easiness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
-Various ways to automate CL using the framework of multi-armed bandits are notably explored in \citep{Graves17automated}, which has been an inspiration for our implementation.
+Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of curriculum learning is introduced.
+CL relies on the notion of ``easiness'' of a sample to schedule the presentation of training data so that the easiest examples are presented first and the hardest, last.
+Various ways to automate CL using the framework of multi-armed bandits are explored in \citep{Graves17automated}, which has been an inspiration for our implementation.
 While the initial aim was primarily to improve and speed up training, CL has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
-
-For instance, \citet{Zhang19curriculum} study supervised domain adaptation and propose a curriculum approach which progressively extends the data used in training: in the early stages only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
+For instance, \citet{Zhang19curriculum} study supervised DA and propose a curriculum approach which progressively augments the training data: in the early stages, only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
 This is somehow opposite to the recommendations of \citet{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyDone{These have not been compared?
@@
 
-\citet{Kumar19reinforcement} use reinforcement techniques (deep Q-learning) to learn the curriculum strategy: in this study, complexity corresponds to difficulty levels and which are binned based on contrastive data selection.
-The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyDone{Alert: what do we do during warm-up ?} The same technique is recently applied to the training of a multilingual NMT system \citep{Kumar21learning}.
+\citet{Kumar19reinforcement} use reinforcement techniques (deep Q-learning) to learn the curriculum strategy: in this work, complexity corresponds to difficulty levels which are binned using contrastive data selection.
+The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyDone{Alert: what do we do during warm-up ?} The same technique is recently applied to multilingual NMT in \citep{Kumar21learning}.
 \citet{Zhou20uncertainty} propose another curriculum-based approach which instead relies on \emph{instance uncertainty} as a measure of their difficulty and presents the data sample starting with the easiest (more predictable) first.
 Another contribution of this paper is an alternative criterium for stopping the training.
-More related to our problems, \citet{Wang20learning-multi} adapt curriculum learning for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization (derivative-free) techniques.
-Each step consists of (a) weighting the instances based on relevance features, (b) fine-tuning a pretrained model using the (weighted) training set, and is applied iteratively to generate a sequence of models.
+More related to our problems, \citet{Wang20learning-multi} adapt CL for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
+Each step consists of (a) weighting instances based on relevance features, (b) fine-tuning a pretrained model using the weighted training set, and is applied iteratively to train a sequence of models.
 The one that maximizes the development set performance is finally retained.

### 2021-05-16 2021/emnlp/auto_cl_mdmt_v1.tex
local changes: correcting spelling, adding sacrebleu reference, adding cluster-domain proportion table

--- 2021-05-14 2021/emnlp/auto_cl_mdmt.tex
@@
 \begin{abstract}
-  Building effective Neural Machine Translation models often imply accommodating diverse sets of heterogeneous data so as to optimize performance for the domain(s) of interest.
+  Building effective Neural Machine Translation models often implies accommodating diverse sets of heterogeneous data so as to optimize performance for the domain(s) of interest.
 Such multi-source / multi-domain adaptation problems are typically approached through instance selection or reweighting strategies, based on a static assessment of the relevance of training instances with respect to the task at hand.
@@
 % \revision{cite also \citep{Shen21source}} -> I think this is not so appropriate here
-where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
+where the heterogeneity of training data not only corresponds to variations in the topic, genre, or register but also in language.
 
@@
 It may also be suboptimal, as some target domains or languages might be easier to train than others.
-For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not achieve optimily for any particular domain.
+For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not achieve optimally for any particular domain.
 
@@
 We evaluate our method in various single-domain and multi-domain adaptation settings.
-We compare our method with several contrasts including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}, which was previously only applied to multilingual MT.
-% From now on, we name our method MDAC which is brief for .
+We compare our method with several contrasts, including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}, which was previously only applied to multilingual MT.
+% From now on, we name our method MDAC, which is brief for .
 Based on these experimental results, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard fine-tuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test situations, from targeting one single domain to the full multi-domain scenario.
@@
 
-To compute the reward $R(d)$ of using data from domain $d$ to train the model, we simulate $k$ training steps from current checkpoint by training the model with $k$ batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
+To compute the reward $R(d)$ of using data from domain $d$ to train the model, we simulate $k$ training steps from the current checkpoint by training the model with $k$ batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
 This computation is inspired by the target prediction gain in \citep{Graves17automated}.
-However, where \cite{Graves17automated} used accumulated gains from the past as reward, we instead predict the usefulness of each domain for improving the future performance of the system given its current state.
+However, where \cite{Graves17automated} used accumulated gains from the past as rewards, we instead predict the usefulness of each domain for improving the system's future performance given its current state.
 This is achieved by simulating a round of training with only the data from one domain.
@@
 \subsection{MDAC for (multi) domain adaptation}
-The setting developed in previous sections is quite general and can in principle accommodate the variety of situations mentioned above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
-In our experiments, we would like to better assess the true potential of MDAC is these settings and seek to experimentally answer the following questions:
+The setting developed in previous sections is quite general.
+It can, in principle, accommodate the variety of situations mentioned above and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
+In our experiments, we would like to assess better the true potential of MDAC in these settings and seek to answer the following questions experimentally:
 \begin{itemize}
-\item is MDAC a viable alternative to conventional fine-tuning? In particular, does it enable to better take advantage of relevant data from other domains?
+\item is MDAC a viable alternative to conventional fine-tuning? In particular, does it enable better to take advantage of relevant data from other domains?
 \item is MDAC also a viable option in multi-domain adaptation scenarios?
@@
 We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and \texttt{sacrebleu}\citep{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
 We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
@@
 Additional choices for the hyper-parameters of our method include the number of simulation steps $k=10$ and the learning rate $\mathbf{lr}_{data}=0.001$.
-We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section~\ref{ssec:clda}), where we use 20~gradient descent iterations to slow down convergence.\fyFuture{Why?
+We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section~\ref{ssec:clda}), where we use 20~gradient descent iterations to avoid converging to degenerate distributions.\fyFuture{Why?
 computation?} We split the training into 100 short sessions that last 2000 training steps each.
@@
 
-In a fourth experiment (\emph{few shot adaptation}), given training data in 6 domains and a small development set of a new domain, (\domain{News} in our case), we aim to build an NMT model which performs well for the unseen domain.
+In a fourth experiment (\emph{few-shot adaptation}), given training data in 6 domains and a small development set of a new domain, (\domain{News} in our case), we aim to build an NMT model which performs well for the unseen domain.
 
 Finally, in the \emph{unsupervised domain adaptation task}, we cluster all available training data into 30 clusters using the KNN algorithm in the same way as in \citep{Tars18multidomain}, then adapt these clusters to one of 6 domains using the corresponding in-domain dev set.
-We compare MDAC to DDS for each our 6 test sets.
+We compare MDAC to DDS for each of our 6 test sets.
 
@@
 We now compare the best MDAC system (using $\vlambda_{t=0} = \vlambda_0$) to full fine-tuning.
-According to Table~\ref{tab:da}, fine-tuning is better for large domains such as \domain{med} and \domain{law}, while MDAC outperforms fine-tuning by approximately 0.5 BLEU for \domain{bank} and 1.0 BLEU for \domain{rel}.
+According to Table~\ref{tab:da}, fine-tuning is better for large domains such as \domain{med} and \domain{law}, while MDAC outperforms fine-tuning by approximately 1.2 BLEU for \domain{bank} and 1.0 BLEU for \domain{rel}.
 This indicates that for small domains, out-of-domain data helps improve the generalization and that MDAC is able to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
-The results for DDS display similar trends, but are always outperformed by MDAC.
-The behaviour of CL, which does only well the large domain \domain{med} lag somewhat behind.
+The results for DDS display similar trends but are always outperformed by MDAC.
+The behavior of CL, which does only well on the large domain \domain{med}, lag somewhat behind.
 
@@
 Figure~\ref{fig:sampling}, where we plot the evolution of the training mixture in the course of training sessions, helps understand the difference between the two methods.
-For DDS (Figure~\ref{fig:DDS}), the proportion of 6 domains quicly converges towards a bi-domain mode in which only \domain{med} and \domain{rel} have significant probability -- hence the good performance on the former domain.
+For DDS (Figure~\ref{fig:DDS}), the proportion of 6 domains quickly converges towards a bi-domain mode in which only \domain{med} and \domain{rel} have significant probability -- hence the good performance on the former domain.
 In contrast, the proportion of 6~domains computed by MDAC evolves more smoothly; small domains such as \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} receive a larger proportion of training data in the early stages; their weights then slowly decrease as larger domains such as \domain{med} and \domain{law} increase their share.
-This however only happends at the end of training, when the NMT models might have been already close to their optimal performance for the small domains.
+This, however, only happens at the end of the training, when the NMT models might have already reached their optimal performance for the small domains.
 
@@
 \subsection{Unseen domain}\label{ssec:uda}
-The left part of Table~\ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and wiht dynamic data selection (MDAC and DDS).
+The left part of Table~\ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and with dynamic data selection (MDAC and DDS).
 These systems have insignificant differences in BLEU, showing that dynamic mixture does not improve the robustness of the NMT system against unseen domains.
-However, the fact that the performance of MDAC and DDS is close to the best performance is a sign that they can also apply in such challengin situation.
+However, the fact that the performance of MDAC and DDS is close to the best performance is a sign that they can also apply in such challenging situations.
 % the robustness of the NMT model.
@@
 In comparison to section~\ref{ssec:da}, the training data is distributed in 30~clusters instead of the 6 original domains.
-Spliting the train data in small groups provides the learner with extra degrees of freedom when selecting the best distribution.
-However, as these clusters are built automatically, they are more noisy in nature.
-According to the results in Table~\ref{tab:unsupervised-da}, this scenario is hard both for DDS and MDAC which perform much worse than for the supervised DA setting.
+Splitting the train data into small groups provides the learner with extra degrees of freedom when selecting the best distribution.
+However, as these clusters are built automatically, they are noisier in nature.
+According to the results in Table~\ref{tab:unsupervised-da}, this scenario is hard both for DDS and MDAC, which perform much worse than for the supervised DA setting.
 This again signals the importance of initialization: analyzing the clustering, we find that the data for \domain{rel} mostly correspond to one single cluster.
-Using a uniform initialization, this cluster starts with a very small weight, and never succeeds to yield the (good) performance observed in the DA settting.
+Using a uniform initialization, this cluster starts with a very small weight and never succeeds in yielding the (good) performance observed in the DA setting.
 
@@
 
-Domain adaptation is an ancient and vexing problem that has been studied from many angles both for SMT and NMT.
+Domain adaptation is an ancient and vexing problem that has been studied from many angles, both for SMT and NMT.
 A survey of supervised and unsupervised DA for NMT is in \citet{Chu18asurvey}, where they distinguish between data-centric and model-centric DA, a view also adopted in the more recent review of \citet{Saunders21domain}.
@@
 Multi-domain adaptation (MDA) aims to develop systems that simultaneously bode well for several domains.
-Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \citep{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citet{Jiang19multidomain}.
+Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \citep{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification, e.g., \ \citet{Jiang19multidomain}.
 Unsupervised MDA is studied in \citet{Farajian17multidomain}, as an instance of unsupervised DA.
 
-Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of curriculum learning is initially introduced.
-CL relies on the notion of ``easiness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
+Most approaches to adaptive/dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of curriculum learning is initially introduced.
+CL relies on the notion of the ``easiness'' of a sample to schedule the presentation of the training data so that the easiest examples presented first and the hardest, last.
 Various ways to automate CL using the framework of multi-armed bandits are notably explored in \citep{Graves17automated}, which has been an inspiration for our implementation.
-While the initial aim was primarily to improve and speed up training, CL has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
+While the initial aim was primarily to improve and speed up training, CL has also proven useful for domain adaptive / multi-domain / multilingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
 
-For instance, \citet{Zhang19curriculum} study supervised domain adaptation and propose a curriculum approach which progressively extends the data used in training: in the early stages only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
+For instance, \citet{Zhang19curriculum} study supervised domain adaptation and propose a curriculum approach which progressively extends the data used in training: in the early stages, only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
 This is somehow opposite to the recommendations of \citet{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyDone{These have not been compared?
@@
 \section{Conclusion and outlook \label{sec:conclusion}}
+\fyDone{Write conclusion}
+This study presents a generic framework to perform various standard adaptation tasks for machine translation, ranging from the conventional supervised domain adaptation to multi-domain adaptation and few-shot DA.
+By experimenting with all these settings, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during training, could be used in all these situations.
+This algorithm, we believe, provides us with a more sound approach to (multi-domain) DA than existing heuristics and dispenses with the costly search of optimal meta-parameters.
+Another contribution of our work is an experimental comparison of recent approaches to dynamic data selection.
 
-\fyDone{Write conclusion}
-In this study, we have presented a generic framework to perform a variety of standard adaptation tasks for machine translation, ranging from the conventional supervised domain adaptation to multi-domain adaptation and few shot DA.
-By experimenting with all these settings, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during  the course of training, could be used in all these situations.
-This algorithm, we believe, provides us with a more sound approach to (multi-domain) DA that existing heuristics, and dispenses with the costly search of optimal meta-parameters.
-Another contribution of our work is an experimental comparison of recent approachs to dynamic data selection.
+In the future, we intend to improve this approach by solving several issues.
+One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training \cite{Kumar19reinforcement}.
+We also would like to improve the unsupervised learning setting of Section~\ref{ssec:clda}, where our results still lag behind that of supervised DA - this might be due to the inability of our simplistic optimization strategy to handle situations where the number of clusters is too large.
 
-In the future, we intend to continue develop this approach and improve its effectiveness.
-One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training \cite{Kumar19reinforcement}.
-Another area where we need to progress is the unsupervised learning setting of Section~\ref{ssec:clda}, where our results still lag way behind that of supervised DA - this might be due to the inability of our simplistic optimization strategy to handle situations where the number of clusters is too large.
 % slow down learning during warm up or normalize reward; handle large number of classes;

### 2021-05-16 2021/emnlp/auto_cl_mdmt.tex

@@
 % \revision{cite also \citep{Shen21source}} -> I think this is not so appropriate here
-where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.
+where the heterogeneity of training data not only corresponds to variations in the topic, genre, or register but also in language.
 
 This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions \citep{Moore10selection,Axelrod11domain}.
-Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data \citep{Miceli-barone17regularization}.
+Finding the optimal balance of training data is, however, a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data \citep{Miceli-barone17regularization}.
 It may also be suboptimal, as some target domains or languages might be easier to train than others.
-For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not achieve optimily for any particular domain.
+For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not achieve optimally for any particular domain.
 
@@
 We evaluate our method in various single-domain and multi-domain adaptation settings.
-We compare our method with several contrasts including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}, which was previously only applied to multilingual MT.
+We compare our method with several contrasts, including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}, which was previously only applied to multilingual MT.
 % From now on, we name our method MDAC which is brief for .
@@
 % \theta_{t+i-1} - \mathbf{lr_2} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta}
-In these equations, $N$ denotes the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $\mathbf{lr_2}$ is the learning rate of the NMT system; $l(\theta,x,y)$ is the loss of the NMT model on sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss aggregated over $n_d$ dev-sets corresponding to the $n_d$ domains.
+In these equations, $N$ denotes the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $l(\theta,x,y)$ is the loss of the NMT model on sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss aggregated over $n_d$ dev-sets corresponding to the $n_d$ domains.
 
-To compute the reward $R(d)$ of using data from domain $d$ to train the model, we simulate $k$ training steps from current checkpoint by training the model with $k$ batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
+To compute the reward $R(d)$ of using data from domain $d$ to train the model, we simulate $k$ training steps from the current checkpoint by training the model with $k$ batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
 This computation is inspired by the target prediction gain in \citep{Graves17automated}.
@@
 \subsection{MDAC for (multi) domain adaptation}
-The setting developed in previous sections is quite general and can in principle accommodate the variety of situations mentioned above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
-In our experiments, we would like to better assess the true potential of MDAC is these settings and seek to experimentally answer the following questions:
+The setting developed in previous sections is quite general and can, in principle, accommodate the variety of situations mentioned above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
+In our experiments, we would like to better assess the true potential of MDAC in these settings and seek to experimentally answer the following questions:
 \begin{itemize}
@@
 \subsection{CL and DDS's re-implementation}\fyDone{Right place for this ?}
-We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to original code of \citet{Wang20balancing}.\footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}
+We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to the original code of \citet{Wang20balancing}.\footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}
 
@@
 
-In the \emph{multi-domain adaptation task}, given the same 6 domains, we aim to build one single NMT model that would perform optimally assuming a uniform distribution of domains during the test.
+In the \emph{multi-domain adaptation task}, given the same 6 domains, we aim to build one single NMT model that would perform optimally, assuming a uniform distribution of domains during the test.
 
-In a fourth experiment (\emph{few shot adaptation}), given training data in 6 domains and a small development set of a new domain, (\domain{News} in our case), we aim to build an NMT model which performs well for the unseen domain.
+In a fourth experiment (\emph{few-shot adaptation}), given training data in 6 domains and a small development set of a new domain (\domain{News} in our case), we aim to build an NMT model which performs well for the unseen domain.
 
 Finally, in the \emph{unsupervised domain adaptation task}, we cluster all available training data into 30 clusters using the KNN algorithm in the same way as in \citep{Tars18multidomain}, then adapt these clusters to one of 6 domains using the corresponding in-domain dev set.
-We compare MDAC to DDS for each our 6 test sets.
+We compare MDAC to DDS for each of our 6 test sets.
 
@@
 We now compare the best MDAC system (using $\vlambda_{t=0} = \vlambda_0$) to full fine-tuning.
-According to Table~\ref{tab:da}, fine-tuning is better for large domains such as \domain{med} and \domain{law}, while MDAC outperforms fine-tuning by approximately 0.5 BLEU for \domain{bank} and 1.0 BLEU for \domain{rel}.
+According to Table~\ref{tab:da}, fine-tuning is better for large domains such as \domain{med} and \domain{law}, while MDAC outperforms fine-tuning by approximately 1.2 BLEU for \domain{bank} and 1.0 BLEU for \domain{rel}.
 This indicates that for small domains, out-of-domain data helps improve the generalization and that MDAC is able to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.

### 2021-05-16 2021/emnlp/auto_cl_mdmt.tex
fining writing with grammarly, correcting citep -> citet, few shot --> few-shot, adding supplementary material

@@
 
-We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \citep{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \citep{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
+We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \citep{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \citep{Chen16guided,Zhang16topicinformed}, register \citep{Sennrich16politeness}, style \citep{Niu18multitask}, etc.
 Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
@@
 
-These situations have been amply documented from a theoretical perspective (eg.\ \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}).
+These situations have been amply documented from a theoretical perspective (eg.\ \citet{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}).
 A general recommendation in the DA setting is to adjust the sampling distribution used to optimize the system so as to compensate for the mismatch between $\mathcal{D}^s(x)$ and $\mathcal{D}^t(x)$.
@@
 
-A widely used approach to supervised DA is \emph{fine-tuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
+A widely used approach to supervised DA is \emph{fine-tuning} \citep{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
 With our notations, this approach amounts to first learning an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$) with $\indic{A}$ the indicator function for predicate $A$.
@@
 To compute the reward $R(d)$ of using data from domain $d$ to train the model, we simulate $k$ training steps from the current checkpoint by training the model with $k$ batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
-This computation is inspired by the target prediction gain in \citep{Graves17automated}.
-However, where \cite{Graves17automated} used accumulated gains from the past as reward, we instead predict the usefulness of each domain for improving the future performance of the system given its current state.
+This computation is inspired by the target prediction gain in \citet{Graves17automated}.
+However, where \citet{Graves17automated} used accumulated gains from the past as rewards, we instead predict the usefulness of each domain for improving the future performance of the system given its current state.
 This is achieved by simulating a round of training with only the data from one domain.
@@
 We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.
-We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \cite{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
+We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}) \citep{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \citep{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \citep{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \citep{Cettolo12wit}, and the Koran (\domain{rel}).
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
 Most corpora are available from the Opus website.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
-To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
+To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \citep{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
 
 We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and \texttt{sacrebleu}\citep{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+Validation sets are used to chose the best model according to the average BLEU score \citep{Papineni02bleu}.\footnote{We use truecasing and \texttt{sacrebleu}\citep{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \citep{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \citep{Neubig19compare-mt}.
 We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
@@
 \subsection{Baseline systems \label{ssec:baseline}}
-Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
+Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \citep{Britz17mixing}.} Using Transformers \citep{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \citep{Klein17opennmt}, we build the following systems:
 
@@
 This indicates that for small domains, out-of-domain data helps improve the generalization and that MDAC is able to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
-The results for DDS display similar trends, but are always outperformed by MDAC.
-The behaviour of CL, which does only well the large domain \domain{med} lag somewhat behind.
+The results for DDS display similar trends but are always outperformed by MDAC.
+The behavior of CL, which does only well the large domain \domain{med} lag somewhat behind.
 
@@
 Figure~\ref{fig:sampling}, where we plot the evolution of the training mixture in the course of training sessions, helps understand the difference between the two methods.
-For DDS (Figure~\ref{fig:DDS}), the proportion of 6 domains quicly converges towards a bi-domain mode in which only \domain{med} and \domain{rel} have significant probability -- hence the good performance on the former domain.
+For DDS (Figure~\ref{fig:DDS}), the proportion of 6 domains quickly converges towards a bi-domain mode in which only \domain{med} and \domain{rel} have significant probability -- hence the good performance on the former domain.
 In contrast, the proportion of 6~domains computed by MDAC evolves more smoothly; small domains such as \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} receive a larger proportion of training data in the early stages; their weights then slowly decrease as larger domains such as \domain{med} and \domain{law} increase their share.
-This however only happends at the end of training, when the NMT models might have been already close to their optimal performance for the small domains.
+This, however, only happens at the end of the training, when the NMT models might have already been close to their optimal performance for the small domains.
 
@@
   \end{tabular}
-  \caption{Few shot adaptation (left) and unsupervised adaptation (right). For a given line, each column correspond to one distinct system. ($^*$) MDAC is significantly better than DDS.}
+  \caption{Few-shot adaptation (left) and unsupervised adaptation (right). For a given line, each column correspond to one distinct system. ($^*$) MDAC is significantly better than DDS.}
   \label{tab:unsupervised-da}
@@
 \subsection{Unseen domain}\label{ssec:uda}
-The left part of Table~\ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and wiht dynamic data selection (MDAC and DDS).
+The left part of Table~\ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and with dynamic data selection (MDAC and DDS).
 These systems have insignificant differences in BLEU, showing that dynamic mixture does not improve the robustness of the NMT system against unseen domains.
-However, the fact that the performance of MDAC and DDS is close to the best performance is a sign that they can also apply in such challenging situation.
+However, the fact that the performance of MDAC and DDS is close to the best performance is a sign that they can also apply in such challenging situations.
 % the robustness of the NMT model.
@@
 In comparison to Section~\ref{ssec:da}, the training data is distributed in 30 automatic clusters instead of the 6 original domains.
-Spliting the train data in small groups provides the learner with extra degrees of freedom when selecting the best distribution.
-However, as these clusters are built automatically, they are more noisy in nature.
-According to the results in Table~\ref{tab:unsupervised-da}, this scenario is hard both for DDS and MDAC which perform much worse than for the supervised DA setting.
+Splitting the train data into small groups provides the learner with extra degrees of freedom when selecting the best distribution.
+However, as these clusters are built automatically, they are noisier in nature.
+According to the results in Table~\ref{tab:unsupervised-da}, this scenario is hard both for DDS and MDAC, which perform much worse than for the supervised DA setting.
 This again signals the importance of initialization: analyzing the clustering, we find that the data for \domain{rel} mostly correspond to one single cluster.
-Using a uniform initialization, this cluster starts with a very small weight, and never succeeds to yield the (good) performance observed in the DA settting.
+In using a uniform initialization, this cluster starts with a very small weight and never succeeds in yielding the (good) performance observed in the DA setting.
 
@@
 
-Domain adaptation is an old problem that has been studied from many angles both for SMT and NMT.
+Domain adaptation is an old problem that has been studied from many angles, both for SMT and NMT.
 A survey of supervised and unsupervised DA for NMT is in \citet{Chu18asurvey}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the more recent review of \citet{Saunders21domain}.
@@
 
-Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of curriculum learning is introduced.
-CL relies on the notion of ``easiness'' of a sample to schedule the presentation of training data so that the easiest examples are presented first and the hardest, last.
+Most approaches to adaptive/dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of curriculum learning is introduced.
+CL relies on the notion of the ``easiness'' of a sample to schedule the presentation of training data so that the easiest examples are presented first and the hardest last.
 Various ways to automate CL using the framework of multi-armed bandits are explored in \citep{Graves17automated}, which has been an inspiration for our implementation.
-While the initial aim was primarily to improve and speed up training, CL has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
+While the initial aim was primarily to improve and speed up training, CL has also proven useful for domain adaptive / multi-domain / multilingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
 For instance, \citet{Zhang19curriculum} study supervised DA and propose a curriculum approach which progressively augments the training data: in the early stages, only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
@@
 \section{Conclusion and outlook \label{sec:conclusion}}
+\fyDone{Write conclusion}
+In this study, we have presented a generic framework to perform a variety of standard adaptation tasks for machine translation, ranging from the conventional supervised domain adaptation to multi-domain adaptation and few-shot DA.
+By experimenting with all these settings, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during the course of training, could be used in all these situations.
+This algorithm, we believe, provides us with a more sound approach to (multi-domain) DA than existing heuristics and dispenses with the costly search of optimal meta-parameters.
+Another contribution of our work is an experimental comparison of recent approaches to dynamic data selection.
 
-\fyDone{Write conclusion}
-In this study, we have presented a generic framework to perform a variety of standard adaptation tasks for machine translation, ranging from the conventional supervised domain adaptation to multi-domain adaptation and few shot DA.
-By experimenting with all these settings, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during  the course of training, could be used in all these situations.
-This algorithm, we believe, provides us with a more sound approach to (multi-domain) DA that existing heuristics, and dispenses with the costly search of optimal meta-parameters.
-Another contribution of our work is an experimental comparison of recent approachs to dynamic data selection.
-
-In the future, we intend to continue develop this approach and improve its effectiveness.
-One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training \cite{Kumar19reinforcement}.
+In the future, we intend to continue developing this approach and improve its effectiveness.
+One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training \citep{Kumar19reinforcement}.
 Another area where we need to progress is the unsupervised learning setting of Section~\ref{ssec:clda}, where our results still lag way behind that of supervised DA - this might be due to the inability of our simplistic optimization strategy to handle situations where the number of clusters is too large.

### 2021-05-16 2021/emnlp/supplementary_material.tex

@@
 
-\section*{A - Generalized Multi-Domain Adaptation Dynamic Sampling Algorithm}
+\section*{A - Generalized Multi-Domain Dynamic Adaptation Curriculum Algorithm}
 We provide below in the pseudo algorithm \ref{alg:mdmt}  the generalized Multi-Domain Adaptation Dynamic Sampling Algorithm.
@@
   \footnotesize
-  \begin{tabular}{|p{1.3cm}|*{11}{c|}} \hline
-   Model  & size&\system{Mixed}&\system{FT}&\system{FT}&\system{MDL}&\system{DC}&\system{DC}&\multirow{2}{*}{\system{TTM}}&\multirow{2}{*}{\system{ADM}}&\multirow{2}{*}{\system{DM}}&\multirow{2}{*}{\system{LDR}}  \\
-   Cluster & train / test & \system{Nat} & \system{Full} & \system{Res} &\system{Res} & \system{Feat}& \system{Tag}& & & & \\ \hline
-24 \hfill [med]&8.1k / 3 &90.4&90.4&90.4&90.4&100.0&65.6&100.0&90.4&100.0&100.0 \\
-13 \hfill [-]&17.3k / 52&67.6&75.4&74.3&74.3&75.0&54.7&74.7&75.9&65.9&76.9 \\
-28 \hfill [-]&25.6k / 54&71.6&68.7&68.1&70.2&71.0&42.5&72.0&71.3&65.6&72.6 \\
-19 \hfill [IT]&27.2k / 88&58.5&63.0&60.9&63.9&63.7&57.2&59.4&61.1&60.5&60.3 \\
-0   \hfill [-]&27.4k / 72&43.9&33.3&45.4&45.4&49.9&15.4&46.8&49.2&46.6&47.8 \\
-22 \hfill [-]&27.5k / 103&91.5&93.7&93.4&93.9&92.5&72.8&92.3&93.2&91.4&93.4 \\
-25 \hfill [-]&28.2k / 56&57.0&44.8&48.2&49.1&54.6&47.2&49.8&54.2&45.1&52.4 \\
-16 \hfill [med]&30.4k / 18&57.2&70.4&77.4&73.5&61.8&54.2&58.4&58.1&52.5&58.3 \\
-23 \hfill [med]&47.0k / 23&24.5&27.2&26.5&28.5&30.5&27.3&32.0&24.4&29.0&29.8 \\
-17 \hfill [med]&54.4k / 26&39.9&40.3&41.6&38.0&37.1&36.6&35.2&35.4&31.3&33.7 \\
-8  \hfill [IT]&61.4k / 214&46.9&53.1&55.8&53.6&48.9&45.1&48.8&50.9&43.0&46.7 \\
-1 \hfill [-]&68.1k / 122&47.2&47.5&48.7&45.1&46.8&39.1&45.4&44.2&40.7&44.9 \\
-7 \hfill [med]&91.5k / 30&41.3&35.5&41.4&39.9&41.4&36.5&37.3&37.1&40.7&41.8 \\
-11 \hfill [med]&93.0k / 38&31.6&42.6&31.8&35.4&36.0&29.6&36.7&32.7&26.5&36.6 \\
-29 \hfill [law]&109.2k / 242&65.9&69.2&67.6&67.7&66.0&63.8&65.1&64.7&62.4&65.9 \\
-27 \hfill [med]&109.3k / 49&11.0&9.6&8.7&9.2&10.0&19.4&9.4&7.9&10.7&10.6 \\
-5 \hfill [-]&109.9k / 267&46.3&47.4&46.9&45.4&44.0&42.9&43.7&44.3&40.9&45.7 \\
-6 \hfill [med]&133.4k / 73&37.2&38.9&38.7&36.8&37.5&27.5&38.0&37.2&31.3&35.9 \\
-26 \hfill [-]&134.8k / 428&31.8&30.8&31.8&31.2&31.9&32.6&32.2&30.5&29.6&31.2 \\
-15 \hfill [bank]&136.9k / 674&46.5&51.5&47.9&48.0&46.6&46.0&45.8&45.7&42.9&46.0 \\
-4 \hfill [rel]&137.4k / 1016&77.1&85.3&83.5&83.3&75.8&46.1&74.2&73.3&63.2&75.9 \\
-2 \hfill [med]&182.6k / 85&70.6&75.8&71.7&69.4&68.2&67.3&67.3&68.6&65.6&68.2 \\
-20 \hfill [med]&183.0k / 71&47.4&47.2&46.8&47.2&48.4&47.5&48.8&47.3&47.1&46.8 \\
-21 \hfill [-]&222.8k / 868&38.7&38.8&39.0&37.2&37.5&35.9&36.9&37.1&33.4&37.0 \\
-10 \hfill [med]&225.4k / 115&40.0&42.6&40.0&38.2&39.9&35.8&39.5&39.1&36.3&40.7 \\
-18 \hfill [med]&245.0k / 106&57.7&60.3&58.7&58.6&58.4&56.3&57.3&56.1&54.9&55.9 \\
-9 \hfill [med]&301.6k / 145&37.2&37.3&36.5&36.1&36.4&37.7&36.4&35.2&34.2&37.0 \\
-3 \hfill [law]&323.5k / 680&50.1&52.0&50.8&50.1&49.1&48.3&49.0&48.2&44.4&49.1 \\
-14 \hfill [med]&334.0 / 146&31.6&31.4&31.9&33.0&32.5&34.1&31.4&32.1&30.5&31.8 \\
-12 \hfill [med]&356.4k / 148&36.3&36.6&35.9&35.9&35.8&37.0&36.4&35.4&34.2&36.3 \\ \hline
+  \begin{tabular}{|p{1.3cm}|*{6}{c|}} \hline
+Cluster&MED&ECB&IT&LAW&REL&TALK \\
+\hline
+1&0.24&0.11&0.47&0.17&0.00&0.01 \\
+2&0.48&0.04&0.19&0.28&0.00&0.00 \\
+3&1.00&0.00&0.00&0.00&0.00&0.00 \\
+4&0.05&0.06&0.01&0.87&0.00&0.00 \\
+5&0.03&0.00&0.00&0.00&0.86&0.10 \\
+6&0.44&0.04&0.40&0.07&0.01&0.04 \\
+7&0.92&0.01&0.01&0.05&0.00&0.00 \\
+8&0.98&0.00&0.00&0.02&0.00&0.00 \\
+9&0.02&0.01&0.96&0.01&0.00&0.00 \\
+10&0.98&0.00&0.00&0.02&0.00&0.00 \\
+11&0.93&0.00&0.01&0.04&0.00&0.01 \\
+12&0.98&0.00&0.01&0.01&0.00&0.00 \\
+13&0.99&0.00&0.00&0.01&0.00&0.00 \\
+14&0.03&0.37&0.01&0.59&0.00&0.00 \\
+15&0.98&0.00&0.01&0.01&0.00&0.00 \\
+16&0.02&0.89&0.01&0.08&0.00&0.00 \\
+17&0.96&0.01&0.02&0.01&0.00&0.00 \\
+18&0.93&0.00&0.05&0.02&0.00&0.00 \\
+19&0.99&0.00&0.00&0.00&0.00&0.00 \\
+20&0.15&0.00&0.79&0.05&0.00&0.01 \\
+21&0.99&0.00&0.00&0.01&0.00&0.00 \\
+22&0.21&0.01&0.40&0.04&0.02&0.32 \\
+23&0.11&0.48&0.02&0.39&0.00&0.00 \\
+24&0.99&0.00&0.01&0.00&0.00&0.00 \\
+25&0.95&0.00&0.04&0.01&0.00&0.00 \\
+26&0.59&0.02&0.23&0.11&0.00&0.05 \\
+27&0.53&0.00&0.02&0.00&0.01&0.44 \\
+28&0.99&0.00&0.00&0.01&0.00&0.00 \\
+29&0.56&0.16&0.08&0.16&0.00&0.04 \\
+30&0.01&0.06&0.00&0.92&0.00&0.00 \\
+\hline
 \end{tabular}
-  \caption{Complete results for the experiments with automatic domains.
-For each cluster, we report: the majority domain when one domain accounts for more than 75\% of the class; training and test sizes; and BLEU scores obtained with the various systems used in this study.
-Most test sets are too small to report significance tests.}
+\caption{Domains's proportion in each cluster.}
 \label{tab:automatic_domains}

### 2021-05-16 2021/emnlp/auto_cl_mdmt.tex
added comment on parameterization

@@
 
-This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions \citep{Moore10selection,Axelrod11domain}.
+This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the test conditions \citep{Moore10selection,Axelrod11domain}.
 Finding the optimal balance of training data is, however, a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data \citep{Miceli-barone17regularization}.
 It may also be suboptimal, as some target domains or languages might be easier to train than others.
-For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not achieve optimally for any particular domain.
+Finally, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not achieve optimally for any particular domain.
 
@@
 have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies.
-Notably, \citet{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while \citet{Graves17automated,Platanios19competence,Kumar19reinforcement,Wang20balancing,Wang20learning-multi} build curricula that automatically adapt to the training data.
+Notably, \citet{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while \citet{Graves17automated,Platanios19competence,Kumar19reinforcement,Wang20learning-multi,Wang20balancing} build curricula that automatically adapt to the training data.
 
@@
 Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of pure domain adaptation.
-These settings are illustrated Figure~\ref{fig:mdmt-lambdas}.
+These settings are illustrated on Figure~\ref{fig:mdmt-lambdas}.
 \begin{figure}[h]
@@
 To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \citep{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
-
+%
 We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
@@
 \itemsep0em 
-\item Generic models trained with various predefined mixtures of the training data taking the form:
+\item Generic models trained with predefined mixtures of the training data taking the form:
 \begin{align} \label{mixture:trn}
@@
 q_d = \frac{\mid N^{s}_d \mid}{\displaystyle{N^{s}}} % \mathop{\sum}_{i=1}^K\mid D_i \mid}}
-\end{align} \setlength{\abovedisplayskip}{2pt}
-\setlength{\belowdisplayskip}{2pt} 
-with $\alpha \in [0,0.25,0.5,0.75,1.0]$.
-These systems are denoted \system{Mixed-$\alpha$} below.
-\system{Mixed-$0$} uses a uniform domain distribution, \system{Mixed-$1.0$} uses the observed domain distribution.
+\end{align}
+%\setlength{\abovedisplayskip}{2pt}
+% \setlength{\belowdisplayskip}{2pt} 
+with $\alpha \in \{0,0.25,0.5,0.75,1.0\}$. We denote these as \system{Mixed-$\alpha$} below. \system{Mixed-$0$} uses a uniform distribution, \system{Mixed-$1.0$} the empirical domain distribution.
 \item fine-tuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ??
@@
 Transformer models contain 8~attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells.
-Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.
-We train all systems with 200k iterations.
+Training lasts for 200K iterations, with batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.
 
@@
 We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to the original code of \citet{Wang20balancing}.\footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}
-
-We also re-implement the approach of \citet{Zhang19curriculum} according to the authors' description of the algorithm.
-For each domain adaptation experiment, we combine the training data of all other domains into one corpus, then compute the cross-entropy difference score of each source sentence of this combined dataset.
+% 
+We also re-implement the approach of \citet{Zhang19curriculum} according to the authors' description.
+For each domain adaptation experiment, we combine the training data of all other domains into one corpus then compute the cross-entropy difference score of each source sentence of this combined dataset.
 We then sort and split the corpus into 9 shards and execute curriculum learning with 10 shards, using the in-domain corpus as the first shard.
@@
 
-In our work, we parameterize the distribution $\lambda^l$ as follows (with $\beta=2$ in all experiments):\fyFuture{Explain why}
+In our work, we parameterize the distribution $\vlambda^l$ as follows (with $\beta=2$ in all experiments):\fyDone{Explain why}
 \begin{equation}
-\lambda
-^l(d;\vpsi) = \frac{\vpsi[d]^\beta}{\sum_i \vpsi[i]^\beta}. \nonumber
+\lambda^l(d;\vpsi) = \frac{\vpsi[d]^\beta}{\sum_i \vpsi[i]^\beta}. \nonumber
 \end{equation}
-
-Additional choices for the hyper-parameters of our method include the number of simulation steps $k=10$ and the learning rate $\mathbf{lr}_{data}=0.001$.
+This parameterization avoids the ``rich-get-richer'' effect that we observe when using $\vlambda(\vpsi) = \operatorname{softmax}(\vpsi)$, which yields gradients wrt.
+$\lambda_d$ to be proportional to $\lambda_d$ (see also Figure~\ref{fig:sampling}).
+Additional settings for the hyper-parameters of our method include the number of simulation steps $k=10$ and the learning rate $\mathbf{lr}_{data}=0.001$.
 We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section~\ref{ssec:clda}), where we use 20~gradient descent iterations to avoid converging to degenerate distributions.\fyFuture{Why?
 computation?} We split the training into 100 short sessions that last 2000 training steps each.
-The choice of those hyper-parameters is mostly heuristic except for the learning rate $\mathbf{lr}_{data}$ which is optimized via grid search over a range of values $[0.001,0.0025,0.005]$.
+The choice of those hyper-parameters is mostly heuristic except for the learning rate $\mathbf{lr}_{data}$ which is optimized via grid search over a set of values $\{0.001,0.0025,0.005\}$.
 
@@
 According to Table~\ref{tab:da}, fine-tuning is better for large domains such as \domain{med} and \domain{law}, while MDAC outperforms fine-tuning by approximately 1.2 BLEU for \domain{bank} and 1.0 BLEU for \domain{rel}.
-This indicates that for small domains, out-of-domain data helps improve the generalization and that MDAC is able to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
-The results for DDS display similar trends but are always outperformed by MDAC.
+This indicates that for small domains, out-of-domain data helps improve the generalization and that MDAC is able to exploit both the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
+Results for DDS display similar trends but are always outperformed by MDAC.
 The behavior of CL, which does only well the large domain \domain{med} lag somewhat behind.
@@
 Figure~\ref{fig:sampling}, where we plot the evolution of the training mixture in the course of training sessions, helps understand the difference between the two methods.
-For DDS (Figure~\ref{fig:DDS}), the proportion of 6 domains quickly converges towards a bi-domain mode in which only \domain{med} and \domain{rel} have significant probability -- hence the good performance on the former domain.
-In contrast, the proportion of 6~domains computed by MDAC evolves more smoothly; small domains such as \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} receive a larger proportion of training data in the early stages; their weights then slowly decrease as larger domains such as \domain{med} and \domain{law} increase their share.
+For DDS (Figure~\ref{fig:DDS}), the sampling distribution quickly converges towards a bi-domain mode in which only \domain{med} and \domain{rel} have significant probability -- hence the good performance on the former domain.
+In contrast, the distribution computed by MDAC evolves more smoothly; small domains such as \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} receive a larger proportion of training data in the early stages; their weights then slowly decrease as larger domains such as \domain{med} and \domain{law} increase their share.
 This, however, only happens at the end of the training, when the NMT models might have already been close to their optimal performance for the small domains.
@@
   \end{tabular}
-  \caption{Few-shot adaptation (left) and unsupervised adaptation (right). For a given line, each column correspond to one distinct system. ($^*$) MDAC is significantly better than DDS.}
+  \caption{Few-shot adaptation (left) and unsupervised adaptation (right). For a given line, each column corresponds to one distinct system. ($^*$) MDAC is significantly better than DDS.}
   \label{tab:unsupervised-da}
@@
 This again signals the importance of initialization: analyzing the clustering, we find that the data for \domain{rel} mostly correspond to one single cluster.
-In using a uniform initialization, this cluster starts with a very small weight and never succeeds in yielding the (good) performance observed in the DA setting.
-
+When using a uniform initialization, this cluster starts with a very small weight and never succeeds in yielding the (good) performance observed in the DA setting.
+s
 \section{Related Work \label{sec:related}}

### 2021-05-16 2021/emnlp/supplementary_material.tex
This is ugly - we should do better

@@
 
-\title{Supplementary materials}
-
-\author{First Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{email@domain} \\\And
-  Second Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{email@domain} \\}
+\title{Supplementary material for paper ``Multi-Domain Adaptation in Neural Machine Translation with Dynamic Sampling Strategies''}
 
@@
 \section*{A - Generalized Multi-Domain Dynamic Adaptation Curriculum Algorithm}
-We provide below in the pseudo algorithm \ref{alg:mdmt}  the generalized Multi-Domain Adaptation Dynamic Sampling Algorithm.
-\begin{algorithm}[htbp]
+The pseudo-code for the generalized Multi-Domain Adaptation Dynamic Sampling Algorithm  is in algorithm~\ref{alg:mdmt}.
+
+\begin{algorithm}[h!]
  \caption{Multi-Domain Adaptation Dynamic Sampling} \label{alg:mdmt}
@@
   \begin{itemize}
-	\item K corpora $C^d, d\in [1,..,K]$ for $K$ domains equipped by an empirical distribution $D_d(x)$
-	\item K dev sets $Dev^d, d\in [1,..,K]$ for $K$ domains.
-	\item Domain testing distribution $\lambda^t \in \mathbb{R}^K$
+    \setlength{\itemsep}{-1pt}    \setlength{\parsep}{0pt}
+  \item K corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equipped by an empirical distribution $D_d(x)$
+  \item K dev sets $Dev^d, d\in [1, \dots,n_d]$ for $n_d$ domains.
+  \item Domain testing distribution $\vlambda^t \in \mathbb{R}^K$
   \item Batch size $B$
-	\item Domain Dynamic Sampling Distribution $\lambda^l(i) \in \mathbb{R}^K$ where $i$ is the enumeration of the iteration.
+  \item Domain Dynamic Sampling Distribution $\lambda^l_i \in \mathbb{R}^K$ where $i$ indexes iterations.
   \item $Eval\_scores = []$
@@
   \item Total training iterations $iter\_num$
-	\item Updating rule of the sampling distribution $\lambda^l(i)$
+  \item Update rule for sampling distribution $\vlambda^l_i$
  \end{itemize}}
@@
 \STATE{Iteration i.}
-\STATE{Randomly pick $d \in [1,..,K]$ from sampling distribution $\lambda^l(i)$}
+\STATE{Randomly pick $d \in [1,\dots,n_d]$ from sampling distribution $\lambda^l(i)$}
 \STATE{Sample $B$ sentences from $C^d$ with empirical distribution $D_d(x)$}
@@
 \IF{$i \equiv 0 \mod{eval\_step}$}
-	\STATE{Evaluate current model with $K$ dev sets. $S^i_d$ is the performance at iteration $i-th$ on domain $d$}
-	\STATE{Report weighted score using the domain testing distribution $\lambda^t$. $$eval(i) = \displaystyle{\mathop{\sum}_d^K \lambda^t(d) S^i_d}$$}
+	\STATE{Evaluate current model with $K$ dev sets. $S^d_i$ is the performance at iteration $i^{th}$ on domain $d$}
+	\STATE{Report weighted score using test distribution $\vlambda^t$. $$eval(i) = \displaystyle{\mathop{\sum}_d^K \lambda^t(d) S^d_i}$$}
 	\STATE{$Eval\_scores.append(eval(i))$.}
@@
 \IF{$i \equiv 0 \mod{sampler\_updating\_step}$}
-	\STATE Update $\lambda^l(i)$.
+	\STATE Update $\vlambda^l_i$.
 \ENDIF
@@
 This experiment aims to simulate with automatic domains a scenario where the number of ``domains'' is large and where some ``domains'' are close and can effectively share information.
-Full results in Table~\ref{tab:automatic_domains}.
-Cluster size vary from approximately 8k sentences (cluster~24) up to more than 350k sentences.
+Full results are in Table~\ref{tab:automatic_domains}.
+Cluster sizes vary from approximately 8k sentences (cluster~24) up to more than 350k sentences.
 More than 2/3 of these clusters mostly comprise texts from one single domain, as for cluster 12 which is predominantly \domain{med}, the remaining clusters typically mix 2 domains.

### 2021-05-17 2021/emnlp/supplementary_material.tex
local changes in auto_cl_mdmt, improving Supplementary materials by breaking algorithm, adding cluster size, cluster proportion in each system, underline the most sampled clusters

@@
 
+\makeatletter
+\newenvironment{breakablealgorithm}
+  {% \begin{breakablealgorithm}
+   \begin{center}
+     \refstepcounter{algorithm}% New algorithm
+     \hrule height.8pt depth0pt \kern2pt% \@fs@pre for \@fs@ruled
+     \renewcommand{\caption}[2][\relax]{% Make a new \caption
+       {\raggedright\textbf{\fname@algorithm~\thealgorithm} ##2\par}%
+       \ifx\relax##1\relax % #1 is \relax
+         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##2}%
+       \else % #1 is not \relax
+         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##1}%
+       \fi
+       \kern2pt\hrule\kern2pt
+     }
+  }{% \end{breakablealgorithm}
+     \kern2pt\hrule\relax% \@fs@post for \@fs@ruled
+   \end{center}
+  }
+\makeatother
+
 \title{Supplementary material for paper ``Multi-Domain Adaptation in Neural Machine Translation with Dynamic Sampling Strategies''}
@@
 The pseudo-code for the generalized Multi-Domain Adaptation Dynamic Sampling Algorithm  is in algorithm~\ref{alg:mdmt}.
+\section*{B - Experiments with automatic domains}
+\begin{table*}[htbp]
+  \centering
+  \footnotesize
+  \begin{tabular}{|p{1.3cm}|*{13}{c|}} 
+  \hline
+\multirow{2}*{Cluster}&\multirow{2}*{size}&\multicolumn{6}{c|}{Domain content}&\multicolumn{6}{c|}{MDAC systems}\\
+  \cline{3-14}
+&&MED&ECB&IT&LAW&REL&TALK&MED&ECB&IT&LAW&REL&TALK \\
+\hline
+1&27436&0.24&0.11&0.47&0.17&0.00&0.01&0.01&0.00&0.01&0.00&0.00&0.01 \\
+2&68108&0.48&0.04&0.19&0.28&0.00&0.00&0.02&0.03&0.02&0.01&0.02&0.02 \\
+3&182594&1.00&0.00&0.00&0.00&0.00&0.00&0.03&0.06&0.04&0.06&0.03&0.03 \\
+4&323474&0.05&0.06&0.01&0.87&0.00&0.00&0.03&\underline{0.07}&0.03&\underline{0.20}&0.03&0.04 \\
+5&137451&0.03&0.00&0.00&0.00&0.86&0.10&0.03&0.04&0.04&0.06&\underline{0.17}&0.04 \\
+6&109949&0.44&0.04&0.40&0.07&0.01&0.04&0.04&0.06&0.04&0.05&0.02&0.03 \\
+7&133395&0.92&0.01&0.01&0.05&0.00&0.00&0.03&0.04&0.03&0.05&0.05&0.03 \\
+8&91464&0.98&0.00&0.00&0.02&0.00&0.00&0.04&0.03&0.04&0.04&0.05&0.03 \\
+9&61353&0.02&0.01&0.96&0.01&0.00&0.00&0.03&0.02&0.03&0.01&0.04&0.02 \\
+10&301639&0.98&0.00&0.00&0.02&0.00&0.00&0.02&0.01&0.01&0.01&0.02&0.03 \\
+11&225347&0.93&0.00&0.01&0.04&0.00&0.01&0.03&0.04&0.04&0.02&0.03&0.04 \\
+12&92982&0.98&0.00&0.01&0.01&0.00&0.00&0.04&0.02&0.03&0.01&0.03&0.03 \\
+13&356377&0.99&0.00&0.00&0.01&0.00&0.00&0.03&0.03&0.04&0.03&0.03&0.04 \\
+14&17260&0.03&0.37&0.01&0.59&0.00&0.00&0.03&0.03&0.03&0.02&0.03&0.02 \\
+15&333957&0.98&0.00&0.01&0.01&0.00&0.00&0.04&0.02&0.03&0.01&0.03&0.03 \\
+16&136944&0.02&0.89&0.01&0.08&0.00&0.00&0.04&\underline{0.08}&0.03&0.04&0.03&0.04 \\
+17&30443&0.96&0.01&0.02&0.01&0.00&0.00&0.04&0.03&0.03&0.03&0.04&0.03 \\
+18&54378&0.93&0.00&0.05&0.02&0.00&0.00&0.04&0.04&0.02&0.03&0.07&0.03 \\
+19&245000&0.99&0.00&0.00&0.00&0.00&0.00&0.04&0.04&0.03&0.03&0.04&0.03 \\
+20&27227&0.15&0.00&0.79&0.05&0.00&0.01&0.04&0.03&0.03&0.02&0.04&0.03 \\
+21&182990&0.99&0.00&0.00&0.01&0.00&0.00&0.03&0.02&0.03&0.03&0.03&0.03 \\
+22&222802&0.21&0.01&0.40&0.04&0.02&0.32&0.04&0.04&\underline{0.08}&0.02&0.01&\underline{0.08} \\
+23&27534&0.11&0.48&0.02&0.39&0.00&0.00&0.03&0.03&0.05&0.02&0.01&0.04 \\
+24&47065&0.99&0.00&0.01&0.00&0.00&0.00&0.04&0.02&0.05&0.05&0.03&0.04 \\
+25&8129&0.95&0.00&0.04&0.01&0.00&0.00&0.03&0.03&0.04&0.03&0.02&0.02 \\
+26&28237&0.59&0.02&0.23&0.11&0.00&0.05&0.03&0.02&0.02&0.01&0.03&0.02 \\
+27&134828&0.53&0.00&0.02&0.00&0.01&0.44&0.04&0.03&0.03&0.01&0.01&0.05 \\
+28&109324&0.99&0.00&0.00&0.01&0.00&0.00&0.04&0.03&0.03&0.02&0.01&0.02 \\
+29&25561&0.56&0.16&0.08&0.16&0.00&0.04&0.03&0.03&0.03&0.02&0.02&0.03 \\
+30&109260&0.01&0.06&0.00&0.92&0.00&0.00&0.03&0.03&0.03&0.06&0.04&0.04 \\
+\hline
+\end{tabular}
+\caption{Automatic clustering experiment.
+We report the size of 30 clusters.
+In the 6 left columns, each line provides the proportions of the domains in each cluster.
+In the 6 right columns, each column corresponds to a system; each line provides the proportion of the corresponding cluster in the total sampled training data from the beginning to the end of the training.
+For each system, we underline the most sampled clusters.}
+\label{tab:automatic_domains}
+\end{table*}
 
-\begin{algorithm}[h!]
+This experiment aims to simulate with automatic domains a scenario where the number of ``domains'' is large and where some ``domains'' are close and can effectively share information.
+Full results are in Table~\ref{tab:automatic_domains}.
+Cluster sizes vary from approximately 8k sentences (cluster~24) up to more than 350k sentences.
+More than 2/3 of these clusters mostly comprise texts from one single domain, as for cluster 12, which is predominantly \domain{med}, the remaining clusters typically mix 2 domains.
+According to the table, for each system, the clusters most sampled by MDAC contain most of the data of the corresponding domain.
+This demonstrates that MDAC is able to find related data to the task automatically.
+However, MDAC performance is still far behind compared to that of the standard situation.
+\\
+\\
+\\
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{breakablealgorithm}
 \caption{Multi-Domain Adaptation Dynamic Sampling} \label{alg:mdmt}
@@
 \end{algorithmic}
- \end{algorithm}
-
-\section*{B - Experiments with automatic domains}
-\begin{table*}[htbp]
-  \centering
-  \footnotesize
-  \begin{tabular}{|p{1.3cm}|*{6}{c|}} \hline
-Cluster&MED&ECB&IT&LAW&REL&TALK \\
-\hline
-1&0.24&0.11&0.47&0.17&0.00&0.01 \\
-2&0.48&0.04&0.19&0.28&0.00&0.00 \\
-3&1.00&0.00&0.00&0.00&0.00&0.00 \\
-4&0.05&0.06&0.01&0.87&0.00&0.00 \\
-5&0.03&0.00&0.00&0.00&0.86&0.10 \\
-6&0.44&0.04&0.40&0.07&0.01&0.04 \\
-7&0.92&0.01&0.01&0.05&0.00&0.00 \\
-8&0.98&0.00&0.00&0.02&0.00&0.00 \\
-9&0.02&0.01&0.96&0.01&0.00&0.00 \\
-10&0.98&0.00&0.00&0.02&0.00&0.00 \\
-11&0.93&0.00&0.01&0.04&0.00&0.01 \\
-12&0.98&0.00&0.01&0.01&0.00&0.00 \\
-13&0.99&0.00&0.00&0.01&0.00&0.00 \\
-14&0.03&0.37&0.01&0.59&0.00&0.00 \\
-15&0.98&0.00&0.01&0.01&0.00&0.00 \\
-16&0.02&0.89&0.01&0.08&0.00&0.00 \\
-17&0.96&0.01&0.02&0.01&0.00&0.00 \\
-18&0.93&0.00&0.05&0.02&0.00&0.00 \\
-19&0.99&0.00&0.00&0.00&0.00&0.00 \\
-20&0.15&0.00&0.79&0.05&0.00&0.01 \\
-21&0.99&0.00&0.00&0.01&0.00&0.00 \\
-22&0.21&0.01&0.40&0.04&0.02&0.32 \\
-23&0.11&0.48&0.02&0.39&0.00&0.00 \\
-24&0.99&0.00&0.01&0.00&0.00&0.00 \\
-25&0.95&0.00&0.04&0.01&0.00&0.00 \\
-26&0.59&0.02&0.23&0.11&0.00&0.05 \\
-27&0.53&0.00&0.02&0.00&0.01&0.44 \\
-28&0.99&0.00&0.00&0.01&0.00&0.00 \\
-29&0.56&0.16&0.08&0.16&0.00&0.04 \\
-30&0.01&0.06&0.00&0.92&0.00&0.00 \\
-\hline
-\end{tabular}
-\caption{Domains's proportion in each cluster.}
-\label{tab:automatic_domains}
-\end{table*}
-
-This experiment aims to simulate with automatic domains a scenario where the number of ``domains'' is large and where some ``domains'' are close and can effectively share information.
-Full results are in Table~\ref{tab:automatic_domains}.
-Cluster sizes vary from approximately 8k sentences (cluster~24) up to more than 350k sentences.
-More than 2/3 of these clusters mostly comprise texts from one single domain, as for cluster 12 which is predominantly \domain{med}, the remaining clusters typically mix 2 domains.
+\end{breakablealgorithm}
 

### 2021-05-17 2021/emnlp/auto_cl_mdmt.tex
refining

@@
 In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key component in training is then to mitigate the detrimental effects of a possible mismatch between these distributions.
-Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from and should not be confused with multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison,Saunders21domain} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from and should not be confused with multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison,Saunders21asurvey} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
 A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Aharoni19massively},\fyDone{Fix ref}
@@
 
-In our work, we parameterize the distribution $\vlambda^l$ as follows (with $\beta=2$ in all experiments):\fyDone{Explain why}
+In our work, we parameterize the distribution $\vlambda^l$ as follows (with $\beta=2$\footnote{This setting corresponds to the \emph{spherical softmax} of \cite{Brebisson16anexploration}.} in all experiments):\fyDone{Explain why}
 \begin{equation}
@@
 This parameterization avoids the ``rich-get-richer'' effect that we observe when using $\vlambda(\vpsi) = \operatorname{softmax}(\vpsi)$, which yields gradients wrt.
-$\lambda_d$ to be proportional to $\lambda_d$ (see also Figure~\ref{fig:sampling}).
+$\psi[d]$ that are proportional to $\exp(\psi[d])$ (see also Figure~\ref{fig:sampling}).
 Additional settings for the hyper-parameters of our method include the number of simulation steps $k=10$ and the learning rate $\mathbf{lr}_{data}=0.001$.
@@
 
-Finally, in the \emph{unsupervised domain adaptation task}, we cluster all available training data into 30 clusters using the KNN algorithm in the same way as in \citep{Tars18multidomain}, then adapt these clusters to one of 6 domains using the corresponding in-domain dev set.
+Finally, in the \emph{unsupervised domain adaptation task}, we cluster all available training data into 30~clusters using the KNN algorithm in the same way as in \citep{Tars18multidomain}, then adapt these clusters to one of 6 domains using the corresponding in-domain dev set.
 We compare MDAC to DDS for each of our 6 test sets.
@@
   % include first image
-  \includegraphics[width=.8\linewidth]{DDS.png}  
+  \includegraphics[width=0.97\textwidth]{DDS.png}  
   \caption{DDS}
@@
   % include second image
-  \includegraphics[width=.8\linewidth]{MDAC.png}  
+  \includegraphics[width=0.97\textwidth]{MDAC.png}  
   \caption{MDAC}
@@
 \subsection{Automatic clustering}\label{ssec:clda}
-The right part of Table~\ref{tab:unsupervised-da} reports the performance of NMT systems adapted to each domain.
-In comparison to Section~\ref{ssec:da}, the training data is distributed in 30 automatic clusters instead of the 6 original domains.
+The right part of Table~\ref{tab:unsupervised-da} reports the performance of NMT systems adapted to each domain.\footnote{A more detailed analysis is in the supplementary material.} In comparison to Section~\ref{ssec:da}, the training data is distributed in 30 automatic clusters instead of the 6 original domains.
 Splitting the train data into small groups provides the learner with extra degrees of freedom when selecting the best distribution.
@@
 When using a uniform initialization, this cluster starts with a very small weight and never succeeds in yielding the (good) performance observed in the DA setting.
-s
+
 \section{Related Work \label{sec:related}}
@@
 Domain adaptation is an old problem that has been studied from many angles, both for SMT and NMT.
-A survey of supervised and unsupervised DA for NMT is in \citet{Chu18asurvey}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the more recent review of \citet{Saunders21domain}.
+A survey of supervised and unsupervised DA for NMT is in \citet{Chu18asurvey}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the more recent survey of \citet{Saunders21asurvey}.
 Our approach to DA in this paper falls under the former category.

### 2021-05-17 2021/emnlp/supplementary_material.tex
revised version

@@
   \footnotesize
-  \begin{tabular}{|p{1.3cm}|*{13}{c|}} 
+  \begin{tabular}{|p{0.5cm}|*{7}{c|}|*{6}{c|}} 
   \hline
-\multirow{2}*{Cluster}&\multirow{2}*{size}&\multicolumn{6}{c|}{Domain content}&\multicolumn{6}{c|}{MDAC systems}\\
+\multirow{2}*{Cl. }&\multirow{2}*{size}&\multicolumn{6}{c||}{Domain content}&\multicolumn{6}{c|}{MDAC systems}\\
   \cline{3-14}
@@
 \end{tabular}
-\caption{Automatic clustering experiment.
-We report the size of 30 clusters.
-In the 6 left columns, each line provides the proportions of the domains in each cluster.
-In the 6 right columns, each column corresponds to a system; each line provides the proportion of the corresponding cluster in the total sampled training data from the beginning to the end of the training.
-For each system, we underline the most sampled clusters.}
+\caption{Automatic clustering experiments.
+We report the size of each cluster.
+In the 6 left columns, each line gives the proportions of the domains in each cluster.
+In the 6 right columns, each column corresponds to a MDAC experiment; each line gives the cumulated proportion of the corresponding cluster in the training data.
+For instance, when targeting the domain \domain{ecb}, cluster~4 (mostly \domain{law}) is sampled with a probability of $0.07$, and cluster~16 (mostly \domain{ecb}) is sampled with probability $0.08$.
+For each system, we underline the most often sampled clusters.}
 \label{tab:automatic_domains}
@@
 This demonstrates that MDAC is able to find related data to the task automatically.
-However, MDAC performance is still far behind compared to that of the standard situation.
+However, MDAC performance is still far behind that of the supervised scenario, as we explain in the paper.
 \\
@@
     \setlength{\itemsep}{-1pt}    \setlength{\parsep}{0pt}
-  \item K corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equipped by an empirical distribution $D_d(x)$
-  \item K dev sets $Dev^d, d\in [1, \dots,n_d]$ for $n_d$ domains.
-  \item Domain testing distribution $\vlambda^t \in \mathbb{R}^K$
+  \item[] 
+  \item $n_d$ corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equipped by an empirical distribution $D_d(x)$
+  \item $n_d$  dev sets $Dev^d, d\in [1, \dots,n_d]$ for $n_d$ domains.
+  \item Domain testing distribution $\vlambda^t \in \mathbb{R}^{n_d}$
   \item Batch size $B$
-  \item Domain Dynamic Sampling Distribution $\lambda^l_i \in \mathbb{R}^K$ where $i$ indexes iterations.
+  \item Domain Dynamic Sampling Distribution $\lambda^l_i \in \mathbb{R}^{n_d}$ where $i$ indexes iterations.
   \item $Eval\_scores = []$
@@
 \REPEAT 
-\STATE{Iteration i.}
-\STATE{Randomly pick $d \in [1,\dots,n_d]$ from sampling distribution $\lambda^l(i)$}
+\STATE{// Start of iteration i}
+\STATE{Randomly pick $d \in [1,\dots,n_d]$ from sampling distribution $\lambda^l_i$}
 \STATE{Sample $B$ sentences from $C^d$ with empirical distribution $D_d(x)$}
@@
 	\STATE{Evaluate current model with $K$ dev sets. $S^d_i$ is the performance at iteration $i^{th}$ on domain $d$}
-	\STATE{Report weighted score using test distribution $\vlambda^t$. $$eval(i) = \displaystyle{\mathop{\sum}_d^K \lambda^t(d) S^d_i}$$}
-	\STATE{$Eval\_scores.append(eval(i))$.}
+	\STATE{Report weighted score using test distribution $\vlambda^t$. $$eval(i) = \displaystyle{\mathop{\sum}_d^{n_d} \lambda^t(d) S^d_i}$$}
+	\STATE{$Eval\_scores.append(eval(i))$}
 \ENDIF
 \IF{$i \equiv 0 \mod{sampler\_updating\_step}$}
-	\STATE Update $\vlambda^l_i$.
+	\STATE Update $\vlambda^l_i$
 \ENDIF
 \IF{$Early\_stopping(Eval\_scores)$}
-	\STATE{Break.}
+	\STATE{break}
 \ENDIF

### 2021-08-02 2021/wmt/auto_cl_mdmt.tex
WMT submission

--- 2021-05-17 2021/emnlp/auto_cl_mdmt.tex
@@
 In this paper, we study dynamic data selection strategies that are able to automatically re-evaluate the usefulness of data samples and to evolve a data selection policy in the course of training.
-Based on the results of multiple experiments, we show that such methods constitute a generic framework to automatically and effectively cater to a variety of real-world situations, from supervised multi-source domain adaptation to multi-domain learning and unsupervised domain adaptation.
+Based on the results of multiple experiments, we show that such methods constitute a generic framework to automatically and effectively cater to a variety of real-world situations, from multi-source domain adaptation to multi-domain learning and unsupervised domain adaptation.
 \end{abstract}
@@
 It is usual in MT to opportunistically collect training samples from several domains, which means that training instances are distributed according to a mixture $\mathcal{D}^s$ such that $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}(d) \mathcal{D}_d(x)$, with $\{\lambda^{s}(d), d=1 \dots n_d\}$ the mixture weights satisfying $\sum_d \lambda^{s}(d)=1$.
+To clarify the ambiguity in the notation of $\lambda$ and $\vlambda$, regardless the superscripts and the subscripts, $\lambda$ is used to denote a scalar and $\vlambda$ is used to denote a vector, for example, $\lambda^{t}(d)$ is the $d^{th}$ element of the vector $\vlambda^{t}$, which represents the probability distribution over the set of domains.
 
 The main challenge is then to make the best of this heterogeneous data, with the aim to achieve the optimal performance for the intended test conditions.
-These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unsupervised domain adaptation); in multi-domain adaptation finally, the test distribution is itself a mixture of domains, some of which may also be observed in training.
+These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unseen domain adaptation); in multi-domain adaptation finally, the test distribution is itself a mixture of domains, some of which may also be observed in training.
 Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of pure domain adaptation.
@@
 Assuming training data in each of the $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
-We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t _d}$ to denote the empirical train and test distributions for domain $d$, and $\widehat{\mathcal{D}^{u}}(x;\lambda^{u}) = \sum_{d} \lambda^{u}(d) \widehat{\mathcal{D}^{u}_d}(x)$ for $u\in\{l,t\}$.
+We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t_d}$ to denote the empirical train and test distributions for domain $d$, and $\widehat{\mathcal{D}^{u}}(x;\vlambda^{u}) = \sum_{d} \lambda^{u}(d) \widehat{\mathcal{D}^{u}_d}(x)$ for $u\in\{l,t\}$.
 In our setting,  $\vlambda^t$, and hence $\widehat{\mathcal{D}^t}(x;\vlambda^t)$ are fixed and predefined, approximated with an equivalent number of development corpora.
@@
 % Our method
-MDAC constructs an adaptative training distribution $\lambda^{l}$ that optimizes the data selection policy along with the training of the NMT model.
-We parameterize $\lambda^{l}$ by a differentiable function $\lambda^l(\psi)$.
-We divide the training into many short sessions; in each session $t$, the model is trained with a static data distribution $\lambda^{l}(\psi_t)$.
+MDAC constructs an adaptative training distribution $\vlambda^{l}$ that optimizes the data selection policy along with the training of the NMT model.
+We parameterize $\vlambda^{l}$ by a differentiable function $\vlambda^l(\psi)$, which will be described in detail in Section \ref{ssec:dds-sys}.
+We divide the training into many short sessions; in each session $t$, the model is trained with a static data distribution $\vlambda^{l}(\psi_t)$.
 After one learning session, we update the data distribution using the REINFORCE algorithm of \citet{Williams92simple}.
@@
 x^i_j, y^i_j &\sim& \widehat{D^l_d}(x) \\
-J^t(\theta,\lambda_t) &=& \displaystyle{\mathop{\sum}_{d=1}^{n_d}}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta,x^t_d,y^t_d).
+J^t(\theta,\vlambda^t) &=& \displaystyle{\mathop{\sum}_{d=1}^{n_d}}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta,x^t_d,y^t_d).
 \end{array}
@@
 % \theta_{t+i-1} - \mathbf{lr_2} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta}
-In these equations, $N$ denotes the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $l(\theta,x,y)$ is the loss of the NMT model on sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss aggregated over $n_d$ dev-sets corresponding to the $n_d$ domains.
+In these equations, $N$ denotes the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $l(\theta,x,y)$ is the loss of the NMT model on sample $(x,y)$; $J^t(\theta,\vlambda^t)$ is the weighted loss aggregated over $n_d$ dev-sets corresponding to the $n_d$ domains.
 
@@
 
-The work of \citet{Wang20balancing} is also related: it is based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\lambda^{l}$ that will result in the best NMT model with respect to a given target dev set at the end of training.
+The work of \citet{Wang20balancing} is also related: it is based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\vlambda^{l}$ that will result in the best NMT model with respect to a given target dev set at the end of training.
 These authors also derive a similar form of update for $\psi$.
@@
 iterations} The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems, each adapted for one domain, with no parameter sharing across domains.
-\item systems trained with fixed data mixtures corresponding to $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$; these are used in the multi-domain experiments of Section~\ref{ssec:mda};
+\item systems trained with fixed data mixtures corresponding to $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$; these are used in the multi-domain experiments of Section~\ref{ssec:mda};
 \item  our own implementations of recent dynamic sampling proposals from the literature: Curriculum Learning (CL) of \citet{Zhang19curriculum} and Differential Data Selection (DDS) of \citet{Wang20balancing} (see details below);
@@
 
+The computational cost of our approach is due to the simulation step, which is conducted after every 2000 iterations to compute the reward of each domain.
+During the simulation step, we update the temporary checkpoint with $k$ updates for each domain, which cost as much as $k$ training updates.
+Therefore, we execute $k \times n_d$ updates in total after every 2000 iterations.
+Our algorithm approximately costs $1 + \frac{k \times n_d}{2000}$ times as much as a standard training.
+
 \subsection{Experimental tasks}
@@
 
-In a fourth experiment (\emph{few-shot adaptation}), given training data in 6 domains and a small development set of a new domain (\domain{News} in our case), we aim to build an NMT model which performs well for the unseen domain.
+In a fourth experiment (\emph{unsupervised domain adaptation}), given training data in 6 domains and a small development set of a new domain (\domain{News} in our case), we aim to build an NMT model which performs well for the unseen domain.
 
@@
 We now turn to a more realistic scenario and consider multi-domain adaptation, which aims to train one single system with optimal performance averaged over 6 domains.
-This setting targets a uniform test distribution $\lambda^t = \lambda_0$.
+This setting targets a uniform test distribution $\vlambda^t = \vlambda_0$.
 In this situation, CL \citep{Zhang19curriculum} does not apply.
@@
 
-We again initialize MDAC and DDS with two distribution $\lambda_0$ and $\lambda_1$.
-According to Table~\ref{tab:multi-da}, MDAC achieves the best performance with initial (uniform) $\lambda_0$.
+We again initialize MDAC and DDS with two distribution $\vlambda_0$ and $\vlambda_1$.
+According to Table~\ref{tab:multi-da}, MDAC achieves the best performance with initial (uniform) $\vlambda_0$.
 The same conclusion holds for DDS.
-For this configuration, MDAC outperforms static training distributions including $\big[ \lambda_0, \lambda_{0.75}, \lambda_{1.0}\big]$ by a significant margin, and performs slightly better than $\big[ \lambda_{0.25}, \lambda_{0.5} \big]$.
+For this configuration, MDAC outperforms static training distributions including $\big[ \vlambda_0, \vlambda_{0.75}, \vlambda_{1.0}\big]$ by a significant margin, and performs slightly better than $\big[ \vlambda_{0.25}, \vlambda_{0.5} \big]$.
 This indicates that MDAC allows us to skip the somewhat heuristic choice of the optimal training mixture.
@@
   \end{tabular}
-  \caption{Few-shot adaptation (left) and unsupervised adaptation (right). For a given line, each column corresponds to one distinct system. ($^*$) MDAC is significantly better than DDS.}
+  \caption{Unseen domain adaptation (left) and automatic clustering adaptation (right). For a given line, each column corresponds to one distinct system. ($^*$) MDAC is significantly better than DDS.}
   \label{tab:unsupervised-da}
@@
 \subsection{Unseen domain}\label{ssec:uda}
-The left part of Table~\ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and with dynamic data selection (MDAC and DDS).
+The left part of Table~\ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$ and with dynamic data selection (MDAC and DDS).
 These systems have insignificant differences in BLEU, showing that dynamic mixture does not improve the robustness of the NMT system against unseen domains.
@@
 \fyDone{Write conclusion}
-In this study, we have presented a generic framework to perform a variety of standard adaptation tasks for machine translation, ranging from the conventional supervised domain adaptation to multi-domain adaptation and few-shot DA.
+In this study, we have presented a generic framework to perform a variety of standard adaptation tasks for machine translation, ranging from the conventional supervised domain adaptation to multi-domain adaptation and unseen domain adaptation.
 By experimenting with all these settings, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during the course of training, could be used in all these situations.

### 2021-08-04 2021/wmt/auto_cl_mdmt.tex
quick review: 8 pages

@@
 In this paper, we study dynamic data selection strategies that are able to automatically re-evaluate the usefulness of data samples and to evolve a data selection policy in the course of training.
-Based on the results of multiple experiments, we show that such methods constitute a generic framework to automatically and effectively cater to a variety of real-world situations, from multi-source domain adaptation to multi-domain learning and unsupervised domain adaptation.
+Based on the results of multiple experiments, we show that such methods constitute a generic framework to automatically and effectively handle a variety of real-world situations, from multi-source domain adaptation to multi-domain learning and unsupervised domain adaptation.
 \end{abstract}
@@
 It is usual in MT to opportunistically collect training samples from several domains, which means that training instances are distributed according to a mixture $\mathcal{D}^s$ such that $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}(d) \mathcal{D}_d(x)$, with $\{\lambda^{s}(d), d=1 \dots n_d\}$ the mixture weights satisfying $\sum_d \lambda^{s}(d)=1$.
-To clarify the ambiguity in the notation of $\lambda$ and $\vlambda$, regardless the superscripts and the subscripts, $\lambda$ is used to denote a scalar and $\vlambda$ is used to denote a vector, for example, $\lambda^{t}(d)$ is the $d^{th}$ element of the vector $\vlambda^{t}$, which represents the probability distribution over the set of domains.
+In the sequel, we also use boldface to denote the vector $\vlambda$, and $\lambda(d)$ is used to denote a scalar corresponding to the $d^{th}$ element of the vector $\vlambda$.
 
-The main challenge is then to make the best of this heterogeneous data, with the aim to achieve the optimal performance for the intended test conditions.
+The main challenge in this situation is to make the best of these heterogeneous data, with the aim to achieve the optimal performance for the intended test conditions.
 These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unseen domain adaptation); in multi-domain adaptation finally, the test distribution is itself a mixture of domains, some of which may also be observed in training.
-Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of pure domain adaptation.
-These settings are illustrated on Figure~\ref{fig:mdmt-lambdas}.
+Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of pure domain adaptation (see illustration on Figure~\ref{fig:mdmt-lambdas}).
 \begin{figure}[h]
@@
 We consider just three domains, and represent vectors of mixture weights $\vlambda^{s}$ and $\vlambda^{t}$ in the 3-dimensional simplex.
-Training with weights in (a) and testing with weights in (c) is supervised multi-source domain adaptation to domain~2 ($d_2$), while (b)-(c) is the unsupervised version, with no training data from $d_2$; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with configurations (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
+Training with weights in (a) and testing with weights in (c) is supervised multi-source domain adaptation to domain~2 ($d_2$), while (b)-(c) is the unsupervised\fyTodo{Unseen ?} version, with no training data from $d_2$; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with configurations (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
 \end{figure}
@@
 With our notations, this approach amounts to first learning an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$) with $\indic{A}$ the indicator function for predicate $A$.
-Note that this strategy is potentially suboptimal, as some out-of-domain samples may contribute to the final performance due to eg.\ domain overlap.
+Note that this strategy is potentially suboptimal, as some out-of-domain samples may contribute to the final performance due to eg.\ domain similarity.
 Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains might be easier to learn than others.\fyDone{How to measure this?}
@@
 \begin{align*}
-\psi_{t+1} &= \psi_t + \mathbf{lr}_{1} * \displaystyle{\mathop{\sum}_{d=1}^{n_d}} R(d) * \frac{\partial \lambda^l(d;\psi_t)}{\partial \psi}, \\
+\psi_{t+1} &= \psi_t + \mathbf{lr}_{1} * \displaystyle{\mathop{\sum}_{d=1}^{n_d}} R(d) * \frac{\partial \lambda^l(d;\psi_t)}{\partial \psi},
 \end{align*}
@@
 \allowdisplaybreaks
-where reward $R(d)$ is computed as:
+where the reward $R(d)$ is computed as:
 \begin{align*}
@@
 We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \citep{Papineni02bleu}.\footnote{We use truecasing and \texttt{sacrebleu}\citep{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \citep{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \citep{Neubig19compare-mt}.
+Validation sets are used to chose the best model according to the average BLEU score \citep{Papineni02bleu}.\footnote{We use truecasing and \texttt{sacrebleu} \citep{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \citep{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \citep{Neubig19compare-mt}.
 We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
@@
 % \setlength{\belowdisplayskip}{2pt} 
-with $\alpha \in \{0,0.25,0.5,0.75,1.0\}$. We denote these as \system{Mixed-$\alpha$} below. \system{Mixed-$0$} uses a uniform distribution, \system{Mixed-$1.0$} the empirical domain distribution.
+with $\alpha \in \{0,0.25,0.5,0.75,1.0\}$. We denote these as \system{Mixed-$\alpha$} below. \system{Mixed-$0$} uses a uniform distribution, \system{Mixed-$1.0$} the empirical distribution of domains.
 \item fine-tuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ??
@@
 
-In our work, we parameterize the distribution $\vlambda^l$ as follows (with $\beta=2$\footnote{This setting corresponds to the \emph{spherical softmax} of \cite{Brebisson16anexploration}.} in all experiments):\fyDone{Explain why}
+In our work, we parameterize the distribution $\vlambda^l$ as follows (with $\beta=2$ in all experiments):\footnote{This setting corresponds to the \emph{spherical softmax} of \cite{Brebisson16anexploration}.}\fyDone{Explain why}
 \begin{equation}
@@
 In the \emph{supervised domain adaptation task}, given the data from 6 domains (\domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}), we aim to build distinct expert NMT models for each domain.
-To challenge the flexibility of the method, we also consider a \emph{bi-domain adaptation task}, where given the same 6 domains, we only focus on adapating only to 2 domains.
+To challenge the flexibility of the method, we also consider a \emph{bi-domain adaptation task}, where given the same 6 domains, we only focus on adapting to a mixture of 2~domains.
 
@@
 
-A second observation is that DDS is again outperformed by MDAC by a significant margin of 1.5 BLEU on average; the only domain where DDS does (much) better is \domain{med}.
-Figure~\ref{fig:sampling}, where we plot the evolution of the training mixture in the course of training sessions, helps understand the difference between the two methods.
-For DDS (Figure~\ref{fig:DDS}), the sampling distribution quickly converges towards a bi-domain mode in which only \domain{med} and \domain{rel} have significant probability -- hence the good performance on the former domain.
+A second observation is that MDAC again outperforms DDS by a wide margin (+1.5 BLEU on average); the only domain where DDS does better is \domain{med}.
+Figure~\ref{fig:sampling}, which displays the evolution of the training mixture in the course of training, helps to understand the difference between the two methods.
+For DDS (Figure~\ref{fig:DDS}), the sampling distribution quickly reaches a bi-domain mode in which only \domain{med} and \domain{rel} have significant probability -- hence the good performance on the former domain.
 In contrast, the distribution computed by MDAC evolves more smoothly; small domains such as \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} receive a larger proportion of training data in the early stages; their weights then slowly decrease as larger domains such as \domain{med} and \domain{law} increase their share.
-This, however, only happens at the end of the training, when the NMT models might have already been close to their optimal performance for the small domains.
+This, however, only happens at the end of training, when some NMT models might already be close to their best performance for the small domains.
 
@@
 These systems have insignificant differences in BLEU, showing that dynamic mixture does not improve the robustness of the NMT system against unseen domains.
-However, the fact that the performance of MDAC and DDS is close to the best performance is a sign that they can also apply in such challenging situations.
+However, the fact that the performance of MDAC and DDS is close to the best performance suggests that they also apply in such challenging situations.
 % the robustness of the NMT model.
@@
 However, as these clusters are built automatically, they are noisier in nature.
-According to the results in Table~\ref{tab:unsupervised-da}, this scenario is hard both for DDS and MDAC, which perform much worse than for the supervised DA setting.
+According to the results in Table~\ref{tab:unsupervised-da}, this scenario is hard both for DDS and MDAC, which performs much worse than for the supervised DA setting.
 This again signals the importance of initialization: analyzing the clustering, we find that the data for \domain{rel} mostly correspond to one single cluster.
-When using a uniform initialization, this cluster starts with a very small weight and never succeeds in yielding the (good) performance observed in the DA setting.
+When using a uniform initialization, this cluster starts with a very small weight and never succeeds in matching the (good) performance observed in the DA setting.
 
@@
 \citet{Zhou20uncertainty} propose another curriculum-based approach which instead relies on \emph{instance uncertainty} as a measure of their difficulty and presents the data sample starting with the easiest (more predictable) first.
-Another contribution of this paper is an alternative criterium for stopping the training.
-More related to our problems, \citet{Wang20learning-multi} adapt CL for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
+Another contribution of this paper is an alternative stopping criterium.
+Closest to our problems, \citet{Wang20learning-multi} adapt CL for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
 Each step consists of (a) weighting instances based on relevance features, (b) fine-tuning a pretrained model using the weighted training set, and is applied iteratively to train a sequence of models.
@@
 \fyDone{Write conclusion}
-In this study, we have presented a generic framework to perform a variety of standard adaptation tasks for machine translation, ranging from the conventional supervised domain adaptation to multi-domain adaptation and unseen domain adaptation.
+In this study, we have presented a generic framework to perform a variety of adaptation tasks for machine translation, ranging from conventional supervised domain adaptation to multi-domain adaptation and unseen domain adaptation.
 By experimenting with all these settings, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during the course of training, could be used in all these situations.

### 2021-11-04 2022/acl/acl_latex.tex
latent multi-task group dropout

% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Remove the "review" option to generate the final version.
\usepackage[review]{acl}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

%algo packages
\usepackage{algorithm}
\usepackage{algorithmic}

%color packages
\usepackage{xcolor}
\colorlet{shadecolor}{orange!15}
\usepackage{color,soul}
\usepackage{soulutf8}
\definecolor{bordeau}{rgb}{0.3515625,0,0.234375}

% position of figure
\usepackage[absolute,overlay]{textpos}

%graphic packages
\usepackage{graphicx}
\graphicspath{ {figures/} }
\usepackage{wrapfig}
\usepackage{float} 
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}

% table packages
\usepackage{array}
\usepackage{multirow,booktabs}
\usepackage{multicol}
\usepackage{tabularx}

% font package
\usepackage[utf8]{inputenc}	% Para caracteres en español
\usepackage{textcomp}
%\usepackage{times}
%\usepackage{lmodern}
\usepackage{mathptmx}

% equation packages
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{mathrsfs}

\title{Frustratingly easy multi-domain, multilingual translation with latent multi-task group dropout}

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a seperate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  \texttt{email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  \texttt{email@domain} \\}

\begin{document}
\maketitle
\begin{abstract}
Multi-domain and multilingual translation can be cast under multi-task learning in which each domain, each language pair are considered as a task. Sharing most of the parameters of the model causes a catastrophic interference between unrelated tasks. In this work, we propose equivalently grouping the nodes of each layer and dropping a portion of these groups assigning different sparse sub-networks to different task. We demonstrate a significant improvement in multi-domain and multilingual translation. Furthermore, through an ablation study, we show that the learned dropping patterns are similar in related tasks allowing the knowledge transferring between these task.
\end{abstract}

\section{Introduction}
Multi-domain translation and multilingual translation develop one model for translations in multi-domain and multiple language pairs respectively. These paradigms are motivated by the compactness of the translation model and the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} and between languages in the same family \citep{Tan19multilingual}.

The method we developed here mitigate the catastrophic interference occurring in training a neural machine translation (NMT) model for multiple domains or multiple language pairs.

\section{Multi-task group dropout}
We define a group dropout for a domain $d$ as follow
\begin{equation}
\begin{array}{rcl}
\tilde{h}^l &=& h^l \odot r_l^d ,\\
h^{l+1} &=& \text{LAYER}^{l+1}(h^l) ,\\
l & \in & [0,\cdots,L] ,\\
\end{array}
\end{equation}
where $L$ is number of Transformer layers of our NMT model.

The dropping mask $r_l^d$ of $l^{th}$ layer and domain $d$ is defined as follow
\begin{equation}
\begin{array}{rcl}
r_l^d(i) &=& \begin{cases}
      1, & \text{if}\ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p} \\
      & \text{AND} \  m_l^d(p) == 1 \\
      0, & \text{otherwise}
    \end{cases} \\
p & \in & \{0,\dots,n_p-1 \} \\
d & \in & \{0,\dots,n_d-1 \} \\
i & \in & \{0,\dots,d_k-1 \} \\
m_l^d & \in & \{0,1\}^{n_p} \ \forall d,l \\
\mid m_l^d \mid_{L_1} & = & k  \\
\Delta^{n_p}_k & = & \{ m \in \{0,1\}^{n_p} | \mid m \mid_{L_1} = k \}
\end{array}
\end{equation}
where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$. The group $p$ is retained for domain $d$ if $m_l^d(p)==1$ and is dropped if $m_l^d(p)==0$. We also denote $\Delta^{n_p}_k$ the set of all admissible dropping masks.

We incorporate the latent variables $m_{l\in[0,\cdots,L]}^d$ into the log-likelihood of our model as below. For the sake of brevity, we denote $m_l^d$ the set of latent variables $m_{l\in[0,\cdots,L]}^d$
\begin{equation}
\hspace{-1.5em}
\begin{array}{rcl}
\text{log} P(y|x,\theta,d) & \text{=} & \text{log} P(y, m_l^d |x,\theta,d) \text{-} \text{log} P(m_l^d|y, x,\theta,d) \\
& \text{=} & \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \\
& \text{+} & KL(P(m_l^d | \Phi_l^d) \parallel P(m_l^d|y, x,\theta,d)) \\
& \geq & \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big]  \\
& \text{=} & \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y | x,\theta,m_l^d, d) \big] \\
& \text{-} & KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \\
& \text{=} & \mathcal{L}(\theta, \Phi_l^d, x , y) \label{eq:lower-bound}
\end{array}
\end{equation}
our training objective will be the lowerbound, which is called the evidence lower bound (ELBO) in variational statistic \citep{Diederick14auto}, $\mathcal{L}(\theta, \Phi_l^d, x , y)$. The variational distribution of $m_l^d$, $P(m_l^d|\Phi_l^d)$, is defined as follows
\begin{align*}
\Phi_d & \in \mathbb{R}^{n_p} \\
Ind^d & \text{=} \{ i_1, \cdots, i_k \} \ | \ \Phi_l^d \\
& \sim \ \text{\small Sampling\_without\_replacement} \big(\text{softmax}(\Phi_l^d) \big) \\
m_l^d(i) &\text{=} \mathbb{I}(i \in Ind^d) \\
\end{align*}

The choice of the modeling of the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick \citep{Kool19stochastic}. In addition, we suppose the prior distribution of $m_l^d$ is uniform over the admissible set $\Delta^{n_p}_k$, $P(m_l^d | x, \theta, d) \ \text{=} \ U(\Delta^{n_p}_k)$.

With our assumption over the prior distribution and our choice of the variational distribution, we express the two terms of the lowerbound \eqref{eq:lower-bound} as follows

\begin{equation}
\begin{array}{rcl}
T_1 &=& \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y | x,\theta,m_l^d, d)  \big] \\
&=& \mathbb{E}_{g_l^d \sim Gumbel(0,1) \ \text{i.i.d}} \big[ \text{log} P(y | x,\theta,\tilde{m}_l^d,d) \big] \\
\end{array}
\end{equation}
where
\begin{equation}
\begin{array}{rcl}
g_l^d & \in & \mathbb{R}^{n_p} \\
g_l^d(p) & \sim & Gumbel(0,1) \ \text{i.i.d} \\
&& \forall p \in \{ 0,\cdots,n_p-1 \} \\
i_1, \cdots, i_k &=& Top_{k} \{ g_l^d(0) + \Phi_l^d(0), \cdots,\\ &&g_l^d(n_d-1) + \Phi_l^d(n_d-1) \} \\
\tilde{m}_l^d(p) &=& \mathbb{I}(p \in Ind^d). \\
\end{array}
\end{equation}

\begin{equation}
\begin{array}{rcl}
T_2 &=& - KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \\
	&=& \mathbb{H} \big[ P(m_l^d | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \\ \label{eq:entropy}
\end{array}
\end{equation}
we will prove this inequality in Appendix \ref{appendix:b}.
Our loss function, which is the negative of the lowerbound \eqref{eq:lower-bound}, consists of 2 terms $-T_1$ and $-T_2$. The inequality \eqref{eq:entropy} show that the upperbound of $-T_2$ is the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ with

%\begin{equation}
%\begin{array}{rcl}
%L(\theta, \Phi_0, \cdots, \Phi_{n_d\text{-1}}) &=& \sum_{d=0}^{n_d-1} \mathbb{E}_{x \sim \mathcal{D}_s^d, y \sim g^d(x)} \big[ \mathcal{L}(\theta,\Phi_d,x,y) \big]
%\end{array}
%\end{equation}

%
%\begin{equation}
%\begin{array}{rcl}
% \tilde{m}^d &=&\displaystyle{\mathop{argmin}_{\substack{
%       0 \leq m_i \leq 1 \ \forall 0 \leq i \leq n_d\text{-}1 \\
%        1^{T}.m = k
%      }}} -(g+\Phi_l^d)^{T} . m - \tau H_b(m) \\
%& & \\
%H_b(m) &=& - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i)
%\end{array}
%\end{equation}
%
%\begin{equation}
%\begin{array}{rcl} 
%
%Ind^d &=& Top_k(\Phi_l^d) \\
%m_l^d &=& \mathbb{I}(i\in Ind^d)\\
%
%y &\sim& P(.|x,\theta,m_l^d)
%
%\end{array}
%\end{equation}
\section{Experimental settings}
\section{Results and ablation studies}
\section{Related work}
\section{Conclusions and outlook}
\bibliography{bibliography}
\bibliographystyle{acl_natbib}

\appendix
\section{Appendix A}
\label{appendix:a}
Theorem of implicit differentiation

\section{Appendix B}
\label{appendix:b}
Entropy inequality
\section{Appendix C}
\label{appendix:c}
Data size

%\label{sec:appendix}

\end{document}

### 2021-11-04 2022/acl/acl_latex.tex
finishing theoretical section, appendix, experimental section

@@
 
+%new commands
+\def\-{\raisebox{.75pt}{-}}
+\usepackage{mathtools,xparse}
+\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
+\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
+\NewDocumentCommand{\normL}{ s O{} m }{%
+  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{2}%
+}
+\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
+%\newcommand{\system}[1]{\texttt{{#1}}}
+%\renewcommand{\UrlFont}{\ttfamily\small}
+%\renewcommand{\S}{\scriptsize}
+%\newcommand{\R}{\mathbb{R}}
+%\newcommand{\vect}[1]{\boldsymbol{#1}}
+%\DeclareMathOperator*{\argmax}{argmax}
+%\DeclareMathOperator*{\argmin}{argmin}
+%\usepackage[draft]{todo}
+%\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
+%\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
+%\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+%\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+%\newcommand{\fyFuture}[1]{\done[FY]\Todo[FY:]{\textcolor{red}{#1}}}
+
 \title{Frustratingly easy multi-domain, multilingual translation with latent multi-task group dropout}
@@
 \tilde{h}^l &=& h^l \odot r_l^d ,\\
+r_l^d \in \{ 0,1 \}^{d_k}, \\
 h^{l+1} &=& \text{LAYER}^{l+1}(h^l) ,\\
@@
 \end{equation}
-where $L$ is number of Transformer layers of our NMT model.
-
-The dropping mask $r_l^d$ of $l^{th}$ layer and domain $d$ is defined as follow
+where $L$ is number of Transformer layers of our NMT model, the binary vector $r_l^d$ is a dropping mask. The dropping mask $r_l^d$ of $l^{th}$ layer and domain $d$ is defined as follow
 \begin{equation}
 \begin{array}{rcl}
-r_l^d(i) &=& \begin{cases}
+r_l^d(i) &=& \begin{cases\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
+\newcommand{\system}[1]{\texttt{{#1}}}}
       1, & \text{if}\ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p} \\
-      & \text{AND} \  m_l^d(p) == 1 \\
-      0, & \text{otherwise}
+      & \text{AND} \  m_l^d(p) \text{\small ==} 1 \\
+      0, & \text{otherwise},
     \end{cases} \\
-p & \in & \{0,\dots,n_p-1 \} \\
-d & \in & \{0,\dots,n_d-1 \} \\
-i & \in & \{0,\dots,d_k-1 \} \\
-m_l^d & \in & \{0,1\}^{n_p} \ \forall d,l \\
-\mid m_l^d \mid_{L_1} & = & k  \\
-\Delta^{n_p}_k & = & \{ m \in \{0,1\}^{n_p} | \mid m \mid_{L_1} = k \}
+p & \in & \{0,\dots,n_p\text{-}1 \}, \\
+d & \in & \{0,\dots,n_d\text{-}1 \}, \\
+i & \in & \{0,\dots,d_k\text{-}1 \}, \\
+m_l^d & \in & \{0,1\}^{n_p} \ \forall d,l \, \\
+\mid m_l^d \mid_{L_1} & = & k , \\
+\Delta^{n_p}_k & = & \{ m \in \{0,1\}^{n_p} | \mid m \mid_{L_1} = k \},
 \end{array}
@@
 where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
-The group $p$ is retained for domain $d$ if $m_l^d(p)==1$ and is dropped if $m_l^d(p)==0$.
+The group $p$ is retained for domain $d$ if $m_l^d(p)\text{small ==}1$ and is dropped if $m_l^d(p)\text{small ==}0$.
 We also denote $\Delta^{n_p}_k$ the set of all admissible dropping masks.
@@
 The variational distribution of $m_l^d$, $P(m_l^d|\Phi_l^d)$, is defined as follows
-\begin{align*}
-\Phi_d & \in \mathbb{R}^{n_p} \\
-Ind^d & \text{=} \{ i_1, \cdots, i_k \} \ | \ \Phi_l^d \\
-& \sim \ \text{\small Sampling\_without\_replacement} \big(\text{softmax}(\Phi_l^d) \big) \\
-m_l^d(i) &\text{=} \mathbb{I}(i \in Ind^d) \\
-\end{align*}
+\begin{equation}
+\hspace{-2.em}
+\begin{array}{rcl}
+\Phi_d & \in & \mathbb{R}^{n_p} , \\
+Ind^d & \text{=} & \{ i_1, \cdots, i_k \} \ | \ \Phi_l^d \\
+& \sim & \ \text{\small Sampling\_without\_replacement} \big(\text{softmax}(\Phi_l^d) \big) , \\
+m_l^d(i) &\text{=}& \mathbb{I}(i \in Ind^d) .\\
+\end{array}
+\end{equation}
 
@@
 
-With our assumption over the prior distribution and our choice of the variational distribution, we express the two terms of the lowerbound \eqref{eq:lower-bound} as follows
+With our assumption over the prior distribution and our choice of the variational distribution, we express the two terms of the lower-bound \eqref{eq:lower-bound} as follows
 
@@
 T_1 &=& \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y | x,\theta,m_l^d, d)  \big] \\
-&=& \mathbb{E}_{g_l^d \sim Gumbel(0,1) \ \text{i.i.d}} \big[ \text{log} P(y | x,\theta,\tilde{m}_l^d,d) \big] \\
+&=& \mathbb{E}_{g_l^d(p) \sim Gumbel(0,1) \ \text{i.i.d}} \big[ \text{log} P(y | x,\theta,\tilde{m}_l^d,d) \big] \\
 \end{array}
@@
 g_l^d(p) & \sim & Gumbel(0,1) \ \text{i.i.d} \\
-&& \forall p \in \{ 0,\cdots,n_p-1 \} \\
-i_1, \cdots, i_k &=& Top_{k} \{ g_l^d(0) + \Phi_l^d(0), \cdots,\\ &&g_l^d(n_d-1) + \Phi_l^d(n_d-1) \} \\
+&& \forall p \in \{ 0,\cdots,n_p\text{-}1 \} \\
+i_1, \cdots, i_k &=& Top\_k \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots,\\ &&g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \} \\
 \tilde{m}_l^d(p) &=& \mathbb{I}(p \in Ind^d). \\
-\end{array}
+\end{array} \label{eq:top-k}
 \end{equation}
+
+\begin{align}
+T_2 &= - KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \\ 
+	&= \mathbb{H} \big[ P(m_l^d | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \\ 
+	&= \mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \\ \
+	&\geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \label{eq:entropy}
+\end{align}
+
+we will prove Inequality \eqref{eq:entropy} in Appendix \ref{appendix:b}.
+
+Now we go back to our loss function, which is the negative of the lower-bound \eqref{eq:lower-bound}, consisting of 2 terms $-T_1$ and $-T_2$.
+The inequality \eqref{eq:entropy} shows that the upperbound of $-T_2$ is $\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d))$.
+In the training, we will maximize the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ as a regularization over the parameters $\Phi_l^d$ of the variational distribution.
+
+Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi_l^d$ into log-likelihood function and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}.
+However, the operator Top-K, which defines $\tilde{m}_l^d$ in Equation \eqref{eq:top-k}, is not differentiable.
+Therefore, we approximate this function by the Soft-Top-K function defined as follows
 
@@
 \begin{array}{rcl}
-T_2 &=& - KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \\
-	&=& \mathbb{H} \big[ P(m_l^d | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \\ \label{eq:entropy}
+ \hat{m}_l^d(\tau) &=&\displaystyle{\mathop{argmin}_{\substack{
+       0 \leqslant m_i \leqslant 1 \ \forall 0 \leqslant i \leqslant n_d\text{-}1 \\
+        1^{T}.m = k
+      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \\
+& \text{in which} & \\
+H_b(m) &=& - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i)
 \end{array}
+\label{eq:soft-top-k}
 \end{equation}
-we will prove this inequality in Appendix \ref{appendix:b}.
-Our loss function, which is the negative of the lowerbound \eqref{eq:lower-bound}, consists of 2 terms $-T_1$ and $-T_2$.
-The inequality \eqref{eq:entropy} show that the upperbound of $-T_2$ is the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ with
 
-%\begin{equation}
-%\begin{array}{rcl}
-%L(\theta, \Phi_0, \cdots, \Phi_{n_d\text{-1}}) &=& \sum_{d=0}^{n_d-1} \mathbb{E}_{x \sim \mathcal{D}_s^d, y \sim g^d(x)} \big[ \mathcal{L}(\theta,\Phi_d,x,y) \big]
-%\end{array}
-%\end{equation}
-
-%
-%\begin{equation}
-%\begin{array}{rcl}
-% \tilde{m}^d &=&\displaystyle{\mathop{argmin}_{\substack{
-%       0 \leq m_i \leq 1 \ \forall 0 \leq i \leq n_d\text{-}1 \\
-%        1^{T}.m = k
-%      }}} -(g+\Phi_l^d)^{T} . m - \tau H_b(m) \\
-%& & \\
-%H_b(m) &=& - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i)
-%\end{array}
-%\end{equation}
-%
-%\begin{equation}
-%\begin{array}{rcl} 
-%
-%Ind^d &=& Top_k(\Phi_l^d) \\
-%m_l^d &=& \mathbb{I}(i\in Ind^d)\\
-%
-%y &\sim& P(.|x,\theta,m_l^d)
-%
-%\end{array}
-%\end{equation}
+In Appendix \ref{appendix:a}, we will prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
+Furthermore, we will also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and its differentiation is computed via implicit differentiation theorem.
+During the training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyperparameter $\tau$ to $0$.
+The gradient of the term $T_1$ w.r.t $\Phi_l^d$ is computed via the chain rule as follows
+\begin{equation}
+\frac{\partial T_1}{\Phi_l^d} = \frac{\partial T_1}{\hat{m}_l^d(\tau)} \times \frac{\hat{m}_l^d(\tau)}{\Phi_l^d}
+\end{equation}
+The gradient $\frac{\partial T_1}{\hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\hat{m}_l^d(\tau)}{\Phi_l^d}$ is computed via implicit differentiation, that we will explain carefully in Appendix \ref{appendix:a}.
 \section{Experimental settings}
+\subsection{Baselines}
+\subsubsection{Multi-domain, multilingual translation}
+In both cases, we examinate the performance of three following MT systems.
+\begin{itemize}
+	\item Transformer trained with .
+	\item Adapter based Transformer.
+	\item Transformer with multi-task group dropout.
+\end{itemize}
+For the sake of brevity, we put the detailed descriptions of these systems in Appendix \ref{appendix:c}.
+\subsection{Datasets}
+\subsubsection{Multi-domain translation}
+We reuse the same data of a recent work of \citet{Pham21revisiting} in multi-domain translation. For the sake of brevity, we leave most of the details of data in Appendix \ref{appendix:c}.
+\subsubsection{Multilingual translation}
+We evaluate our model on both one-to-many (O2M) and many-to-one (M2O)
+translation tasks. We evaluate models on widely used multilingual translation datasets.
+\begin{itemize}
+	\item TED8-Related. Following the setting of \citep{Wang20balancing}, we use a subset of translations between English and eight related languages.
+	\item TED8-Diverse. The dataset consists of parallel sentences between English and eight diverse languages \citep{Wang20balancing}.
+\end{itemize}
+The detailed description of the datasets is provided in Appendix \ref{appendix:c}.
 \section{Results and ablation studies}
+\subsection{Multi-domain translation}
+\subsection{Multilingual translation}
+\subsection{Similarity between dropping masks}
 \section{Related work}
@@
 \bibliographystyle{acl_natbib}
-
 \appendix
@@
 \label{appendix:a}
-Theorem of implicit differentiation
-
+Proof of limit. Computation of $\hat{m}_l^d(\tau)$. Theorem of implicit differentiation and Proof of differentiability.
 \section{Appendix B}
 \label{appendix:b}
-Entropy inequality
+In this section, we give a simple proof of Inequality \ref{eq:entropy}.
+In fact, we only need to prove $\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] \geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]$.
+The proof is as follows
+\begin{align*}
+&\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] = - \mathbb{E}_{i_1,\cdots,i_k | \Phi_l^d} \big[ \text{log}P(i_1,\cdots,i_k | \Phi_l^d) \big] \\
+&= -\mathbb{E}_{i_1,\cdots,i_k | \Phi_l^d} \big[ \displaystyle{\sum_{j=2}^k}\text{log}P(i_j | i_1,\cdots,j_{j-1},\Phi_l^d) +  \text{log} P(i_1 | \Phi_l^d) \big] \\
+&\geqslant -\mathbb{E}_{i_1,\cdots,i_k | \Phi_l^d} \big[ \text{log} P(i_1 | \Phi_l^d) \big] \\
+&= -\mathbb{E}_{i_1 | \Phi_l^d} \big[ \text{log} P(i_1 | \Phi_l^d) \big] = \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]
+\end{align*}
 \section{Appendix C}
 \label{appendix:c}
-Data size
+\subsection{Multi-domain translation}
 
-%\label{sec:appendix}
+\begin{table*}[h!]
+  \centering
+  \begin{tabular}{|l|ccccccc|} %*{4}{|r|}}
+    \cline{2-8} 
+    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
+    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} & \multicolumn{1}{c|}{\domain{news}} \\
+    \hline 
+    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
+    \# \revisiondone{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 & 7.8 / 9.2   \\
+    \# \revisiondone{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 & 77.8 / 80.4 \\
+    \# \revisiondone{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 & 31.5 / 23.1 \\
+    \hline
+  \end{tabular}
+  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$).
+\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
+  }
+\label{tab:Corpora-chap4}
+\end{table*}
+
+\subsection{Multilingual translation}
 

### 2021-11-04 2022/acl/acl_latex.tex
add mdmt scores

@@
 
-%new commands
-\def\-{\raisebox{.75pt}{-}}
-\usepackage{mathtools,xparse}
-\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
-\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
-\NewDocumentCommand{\normL}{ s O{} m }{%
-  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{2}%
-}
+\usepackage{latexsym}
+\usepackage{graphicx}
+\usepackage{xcolor}
+\usepackage{longtable}
+\usepackage{tikz}
+\usetikzlibrary{calc}
+\usepackage[draft]{todo}
+\usepackage[normalem]{ulem}
+\usepackage{xspace}
+\usepackage{amsmath,amsfonts,amssymb}
+
+\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\fyFuture}[1]{\done[FY]\Todo[FY:]{\textcolor{red}{#1}}}
+\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
+\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{red}{#1}}}
+\newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{red}{#1}}}
+\newcommand{\revision}[1]{\textcolor{black}{#1}}
+\newcommand{\revisiondone}[1]{\textcolor{black}{#1}}
+\newcommand{\revisiondel}[1]{}
+\newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
+\newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
 \newcommand{\domain}[1]{\texttt{\textsc{#1}}}
-%\newcommand{\system}[1]{\texttt{{#1}}}
-%\renewcommand{\UrlFont}{\ttfamily\small}
-%\renewcommand{\S}{\scriptsize}
-%\newcommand{\R}{\mathbb{R}}
-%\newcommand{\vect}[1]{\boldsymbol{#1}}
-%\DeclareMathOperator*{\argmax}{argmax}
-%\DeclareMathOperator*{\argmin}{argmin}
-%\usepackage[draft]{todo}
-%\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
-%\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
-%\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
-%\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
-%\newcommand{\fyFuture}[1]{\done[FY]\Todo[FY:]{\textcolor{red}{#1}}}
+\newcommand{\system}[1]{\texttt{{#1}}}
 
@@
 \begin{array}{rcl}
-r_l^d(i) &=& \begin{cases\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
-\newcommand{\system}[1]{\texttt{{#1}}}}
+r_l^d(i) &=& \begin{cases}
       1, & \text{if}\ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p} \\
@@
 The variational distribution of $m_l^d$, $P(m_l^d|\Phi_l^d)$, is defined as follows
-\begin{equation}
+
+\begin{align}
 \hspace{-2.em}
-\begin{array}{rcl}
-\Phi_d & \in & \mathbb{R}^{n_p} , \\
-Ind^d & \text{=} & \{ i_1, \cdots, i_k \} \ | \ \Phi_l^d \\
-& \sim & \ \text{\small Sampling\_without\_replacement} \big(\text{softmax}(\Phi_l^d) \big) , \\
-m_l^d(i) &\text{=}& \mathbb{I}(i \in Ind^d) .\\
-\end{array}
-\end{equation}
+&\Phi_d \in \mathbb{R}^{n_p} , \nonumber \\
+&Ind^d \text{=}  \{ i_1, \cdots, i_k \} \nonumber \\
+&\sim \text{\small Sampling\_without\_replacement} \big(\text{softmax}(\Phi_l^d) \big) , \nonumber \\
+&m_l^d(i) \text{=} \mathbb{I}(i \in Ind^d) .\nonumber
+\end{align}
 
@@
 
-\begin{equation}
-\begin{array}{rcl}
-T_1 &=& \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y | x,\theta,m_l^d, d)  \big] \\
-&=& \mathbb{E}_{g_l^d(p) \sim Gumbel(0,1) \ \text{i.i.d}} \big[ \text{log} P(y | x,\theta,\tilde{m}_l^d,d) \big] \\
-\end{array}
-\end{equation}
+\begin{align}
+\hspace{-2.em}
+T_1 &= \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y | x,\theta,m_l^d, d)  \big] \nonumber \\
+&= \mathbb{E}_{g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} Gumbel(0,1) \ } \big[ \text{log} P(y | x,\theta,\tilde{m}_l^d,d) \big] \nonumber
+\end{align}
 where
-\begin{equation}
-\begin{array}{rcl}
-g_l^d & \in & \mathbb{R}^{n_p} \\
-g_l^d(p) & \sim & Gumbel(0,1) \ \text{i.i.d} \\
-&& \forall p \in \{ 0,\cdots,n_p\text{-}1 \} \\
-i_1, \cdots, i_k &=& Top\_k \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots,\\ &&g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \} \\
-\tilde{m}_l^d(p) &=& \mathbb{I}(p \in Ind^d). \\
-\end{array} \label{eq:top-k}
-\end{equation}
-
 \begin{align}
-T_2 &= - KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \\ 
-	&= \mathbb{H} \big[ P(m_l^d | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \\ 
-	&= \mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \\ \
+\hspace{-2.em}
+g_l^d & \in \mathbb{R}^{n_p} \nonumber \\
+g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} Gumbel(0,1) \nonumber \\
+& \forall p \in \{ 0,\cdots,n_p\text{-}1 \} \nonumber \\
+i_1, \cdots, i_k &= Top\_k \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \label{eq:top-k} \\ 
+&g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \} \nonumber \\
+\tilde{m}_l^d(p) &= \mathbb{I}(p \in Ind^d). \nonumber 
+\end{align}
+%
+\begin{align}
+T_2 &= - KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\ 
+	&= \mathbb{H} \big[ P(m_l^d | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\ 
+	&= \mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\ \
 	&\geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \label{eq:entropy}
@@
 
-\begin{equation}
-\begin{array}{rcl}
- \hat{m}_l^d(\tau) &=&\displaystyle{\mathop{argmin}_{\substack{
-       0 \leqslant m_i \leqslant 1 \ \forall 0 \leqslant i \leqslant n_d\text{-}1 \\
+\begin{align}
+\hat{m}_l^d(\tau) &= \displaystyle{\mathop{argmin}_{\substack{
+       0 \leqslant m_i \leqslant 1 \ \forall 0 \leqslant i \leqslant n_d\text{-}1 \label{eq:soft-top-k}\\
         1^{T}.m = k
       }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \\
-& \text{in which} & \\
-H_b(m) &=& - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i)
-\end{array}
-\label{eq:soft-top-k}
-\end{equation}
+& \text{in which} \nonumber \\
+H_b(m) &= - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i) \nonumber 
+\end{align}
 
@@
 \begin{itemize}
-	\item TED8-Related. Following the setting of \citep{Wang20balancing}, we use a subset of translations between English and eight related languages.
-	\item TED8-Diverse. The dataset consists of parallel sentences between English and eight diverse languages \citep{Wang20balancing}.
+	\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations of \citet{qi18when} between English and eight related languages.
+	\item TED8-Diverse. The dataset consists of parallel sentences between English and eight diverse languages as in \citet{Wang20balancing}.
 \end{itemize}
@@
 \subsection{Multi-domain translation}
+According to Table \ref{tab:mdmt}, \system{LaMGD} shows a significant improvement over the generic system \system{Transformer} with zero extra parameters.
+Moreover, \system{LaMGD} achieve equivalent performance in average compared to \system{Adapter}, which is carrefully finetuned with a approximately 12.4M additional parameters.
+\begin{table*}
+  \centering
+  \begin{tabular}{|p{4cm}|*{7}{r|}} \hline
+    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
+    \system{Transformer}  \hfill{\footnotesize[65m]} & 40.3 & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+12.4m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.28 \\ 
+    \system{LaMGD}   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & 39.0 & 52.4 & 87.5 & 55.33 \\ 
+    \hline
+  \end{tabular}
+  \caption{Multi-domain translation}
+  \label{tab:mdmt}
+\end{table*}
 \subsection{Multilingual translation}
+
 \subsection{Similarity between dropping masks}

### 2021-11-04 2022/acl/acl_latex.tex
adding multilingual scores

@@
 \begin{abstract}
-Multi-domain and multilingual translation can be cast under multi-task learning in which each domain, each language pair are considered as a task.
+Multi-domain and multilingual translation can be cast under multi-task learning in which each domain or language pair is considered a task.
 Sharing most of the parameters of the model causes a catastrophic interference between unrelated tasks.
-In this work, we propose equivalently grouping the nodes of each layer and dropping a portion of these groups assigning different sparse sub-networks to different task.
+In this work, we propose equivalently grouping the nodes of each layer and dropping a portion of these groups.
+Via the task-dependent dropout mechanism, we assign different sparse sub-networks to different tasks.
 We demonstrate a significant improvement in multi-domain and multilingual translation.
-Furthermore, through an ablation study, we show that the learned dropping patterns are similar in related tasks allowing the knowledge transferring between these task.
+Furthermore, through an ablation study, we show that the learned dropping patterns are similar in related tasks allowing the knowledge transferring between these tasks.
+
 \end{abstract}
@@
 where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
-The group $p$ is retained for domain $d$ if $m_l^d(p)\text{small ==}1$ and is dropped if $m_l^d(p)\text{small ==}0$.
+The group $p$ is retained for domain $d$ if $m_l^d(p)\text{\small ==}1$ and is dropped if $m_l^d(p)\text{\small ==}0$.
 We also denote $\Delta^{n_p}_k$ the set of all admissible dropping masks.
@@
 \begin{itemize}
-	\item Transformer trained with .
-	\item Adapter based Transformer.
-	\item Transformer with multi-task group dropout.
+	\item Transformer
+	\item Adapter based Transformer
+	\item Transformer with multi-task group dropout
 \end{itemize}
@@
 \subsection{Multi-domain translation}
-According to Table \ref{tab:mdmt}, \system{LaMGD} shows a significant improvement over the generic system \system{Transformer} with zero extra parameters.
+According to Table \ref{tab:mdmt}, \system{LaMGD} shows a significant improvement (+2.78) over the generic system \system{Transformer} with zero extra parameters.
 Moreover, \system{LaMGD} achieve equivalent performance in average compared to \system{Adapter}, which is carrefully finetuned with a approximately 12.4M additional parameters.
-\begin{table*}
+\begin{table*}[h!]
   \centering
@@
 \subsection{Multilingual translation}
+According to Table \ref{tab:multilingual}, \system{LAMGD} shows a boost of 0.42, 0.33, ,0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, O2M-diverse, M2O-diverse respectively.
+Significant gains are observed in \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction).
+\begin{table*}[ht]
+  \centering
+  \begin{tabular}{|p{4cm}|*{9}{r|}} \hline
+    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
+    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.80 &7.30&20.80&21.10&39.70&19.80&22.60&15.20&18.91 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} & & & & & & & & & \\ 
+    \system{LaMGD}  \hfill{\footnotesize[+0m]}  & 5.20&9.40&20.60&22.80&39.60&19.60&22.40&15.00&19.33 \\ 
+	\hline
+    \hline
+    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
+    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.40&16.60&28.50&	27.10&43.70&24.60&30.30&25.60&25.98 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} & & & & & & & & & \\ 
+    \system{LaMGD}   \hfill{\footnotesize[+0m]}  &11.30&17.40&28.60&28.70&	43.70&24.50&30.70&25.60&26.31 \\ 
+    \hline
+    \hline
+    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
+    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.20&4.0&12.70&22.20&31.80&34.00&38.30&8.30&20.19 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  & & & & & & & & & \\ 
+    \system{LaMGD}   \hfill{\footnotesize[+0m]} & & & & & & & & & \\
+    \hline 
+    \hline
+    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
+    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.40&9.70&20.50&31.80&37.50&38.70&39.80&19.00&27.43 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  & & & & & & & & & \\ 
+    \system{LaMGD}   \hfill{\footnotesize[+0m]} &23.50&9.60&21.50&32.20&37.70&38.60&40.00&18.90&27.75 \\
+    \hline
+  \end{tabular}
+  \caption{Multilingual translation}
+  \label{tab:multilingual}
+\end{table*}
+\subsection{Similarity between dropping masks}
 
-\subsection{Similarity between dropping masks}
 \section{Related work}
@@
 \label{appendix:c}
-\subsection{Multi-domain translation}
+\subsection{Multi-domain datasets}
 
@@
 
-\subsection{Multilingual translation}
+\subsection{Multilingual datasets}
+
+\subsection{Multi-domain systems}
+
+\subsection{Multilingual systems}
 

### 2021-11-05 2022/acl/acl_latex.tex
adding reference
local merges
local changes

@@
 These paradigms are motivated by the compactness of the translation model and the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} and between languages in the same family \citep{Tan19multilingual}.
+
+\citet{Christopher21efficiently}
 
@@
 \subsection{Multilingual translation}
-According to Table \ref{tab:multilingual}, \system{LAMGD} shows a boost of 0.42, 0.33, ,0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, O2M-diverse, M2O-diverse respectively.
+According to Table \ref{tab:multilingual}, \system{LaMGD} shows a boost of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, M2O-diverse respectively.
 Significant gains are observed in \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction).

### 2021-11-05 2022/acl/acl_latex.tex
add some insignificant details

@@
 \newcommand{\system}[1]{\texttt{{#1}}}
+\newcommand{\SB}[1]{\textbf{#1}}
+\newcommand{\SW}[1]{\underline{#1}}
 
@@
 For the sake of brevity, we denote $m_l^d$ the set of latent variables $m_{l\in[0,\cdots,L]}^d$
-\begin{equation}
-\hspace{-1.5em}
-\begin{array}{rcl}
-\text{log} P(y|x,\theta,d) & \text{=} & \text{log} P(y, m_l^d |x,\theta,d) \text{-} \text{log} P(m_l^d|y, x,\theta,d) \\
-& \text{=} & \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \\
-& \text{+} & KL(P(m_l^d | \Phi_l^d) \parallel P(m_l^d|y, x,\theta,d)) \\
-& \geq & \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big]  \\
-& \text{=} & \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y | x,\theta,m_l^d, d) \big] \\
-& \text{-} & KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \\
-& \text{=} & \mathcal{L}(\theta, \Phi_l^d, x , y) \label{eq:lower-bound}
-\end{array}
-\end{equation}
-our training objective will be the lowerbound, which is called the evidence lower bound (ELBO) in variational statistic \citep{Diederick14auto}, $\mathcal{L}(\theta, \Phi_l^d, x , y)$.
+\begin{align}
+\hspace{-1.em}
+\text{log} P(y|x,\theta,d) & \text{=} \text{log} P(y, m_l^d |x,\theta,d) \text{-} \text{log} P(m_l^d|y, x,\theta,d) \nonumber \\
+& \text{=} \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
+& \text{+} KL(P(m_l^d | \Phi_l^d) \parallel P(m_l^d|y, x,\theta,d)) \nonumber \\
+& \geqslant \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
+& \text{=} \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y | x,\theta,m_l^d, d) \big] \label{eq:lower-bound} \\
+& \text{-} KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\
+& \text{=} \mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y). \nonumber
+\end{align}
+$\mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y)$ is the lowerbound, which is called the evidence lower bound (ELBO) in variational statistic \citep{Diederick14auto}, of our log-likelihood.
+Our training loss will be $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$.
+
 The variational distribution of $m_l^d$, $P(m_l^d|\Phi_l^d)$, is defined as follows
@@
 \begin{align}
+\hspace{-2.em}
 T_2 &= - KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\ 
@@
 The gradient $\frac{\partial T_1}{\hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\hat{m}_l^d(\tau)}{\Phi_l^d}$ is computed via implicit differentiation, that we will explain carefully in Appendix \ref{appendix:a}.
+
+We jointly train the Transformer parameters $\theta$ and the parameters of the latent dropping masks $\Phi_l^d$ using the following multi-task loss.
+
+\begin{equation}
+\mathcal{L}(\theta,\Phi_l^1,\cdots,\Phi_l^k) = \displaystyle{\mathop{\sum}_{d=1}^{k}} \mathbb{E}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)} \big[-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)\big]
+\end{equation}
+in which $\mathcal{D}_s^d$ is distribution of the task $d$ over the input space $\Omega^d_s$; $g^d: \Omega^d_s \rightarrow \Omega^d_t$ is the mapping function, which our multi-task modern needs to learn, of the task $d$; $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$ is the ELBO loss, which we defined in Equation \ref{eq:lower-bound}.
 \section{Experimental settings}
@@
 \subsubsection{Multi-domain, multilingual translation}
-In both cases, we examinate the performance of three following MT systems.
+In both cases, we examine the performance of three following MT systems.
 \begin{itemize}
@@
 \subsubsection{Multi-domain translation}
-We reuse the same data of a recent work of \citet{Pham21revisiting} in multi-domain translation. For the sake of brevity, we leave most of the details of data in Appendix \ref{appendix:c}.
+We reuse the same data as the recent work of \citet{Pham21revisiting} in multi-domain translation. For the sake of brevity, we leave most of the details of data in Appendix \ref{appendix:c}.
 \subsubsection{Multilingual translation}
@@
 \subsection{Multilingual translation}
-According to Table \ref{tab:multilingual}, \system{LaMGD} shows a boost of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, M2O-diverse respectively.
+According to Table \ref{tab:multilingual}, \system{LaMGD} shows a boost of 0.42, 0.33, ,0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, M2O-diverse respectively.
 Significant gains are observed in \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction).
+However, \system{LaMGD} lags behind multilingual \system{Transformer} in O2M-diverse.
 \begin{table*}[ht]
@@
     \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} & & & & & & & & & \\ 
-    \system{LaMGD}  \hfill{\footnotesize[+0m]}  & 5.20&9.40&20.60&22.80&39.60&19.60&22.40&15.00&19.33 \\ 
+    \system{LaMGD}  \hfill{\footnotesize[+0m]}  & 5.20&\SB{9.40}&20.60&\SB{22.80}&39.60&19.60&22.40&15.00&19.33 \\ 
 	\hline
@@
     \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} & & & & & & & & & \\ 
-    \system{LaMGD}   \hfill{\footnotesize[+0m]}  &11.30&17.40&28.60&28.70&	43.70&24.50&30.70&25.60&26.31 \\ 
+    \system{LaMGD}   \hfill{\footnotesize[+0m]}  &11.30&\SB{17.40}&28.60&\SB{28.70}&43.70&24.50&30.70&25.60&26.31 \\ 
     \hline
@@
     \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  & & & & & & & & & \\ 
-    \system{LaMGD}   \hfill{\footnotesize[+0m]} &23.50&9.60&21.50&32.20&37.70&38.60&40.00&18.90&27.75 \\
+    \system{LaMGD}   \hfill{\footnotesize[+0m]} &\SB{23.50}&9.60&\SB{21.50}&32.20&37.70&38.60&40.00&18.90&27.75 \\
     \hline

### 2021-11-05 2022/acl/acl_latex.tex

@@
 \subsection{Multilingual translation}
-According to Table \ref{tab:multilingual}, \system{LaMGD} shows a boost of 0.42, 0.33, ,0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, M2O-diverse respectively.
+According to Table \ref{tab:multilingual}, \system{LaMGD} shows a boost of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, M2O-diverse respectively.
 Significant gains are observed in \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction).

### 2021-11-08 2022/acl/acl_latex.tex
finishing proofs, adding graphics and references

@@
 
+\usepackage{enumitem}
+\setlist{nosep}
 \usepackage{latexsym}
@@
 
-\title{Frustratingly easy multi-domain, multilingual translation with latent multi-task group dropout}
+\title{Multi-domain, multilingual translation with latent multi-task group dropout}
 
@@
 \begin{document}
+\setlength{\abovedisplayskip}{3pt}
+\setlength{\belowdisplayskip}{3pt}
 \maketitle
@@
 Sharing most of the parameters of the model causes a catastrophic interference between unrelated tasks.
-In this work, we propose equivalently grouping the nodes of each layer and dropping a portion of these groups.
-Via the task-dependent dropout mechanism, we assign different sparse sub-networks to different tasks.
-We demonstrate a significant improvement in multi-domain and multilingual translation.
-Furthermore, through an ablation study, we show that the learned dropping patterns are similar in related tasks allowing the knowledge transferring between these tasks.
-
+Most of the proposed multi-task architectures propose sharing the parameters of an underlying generic model and using different output layers to predict different tasks.
+However, these paradigms do not explicitly exploit the proximity between tasks.
+In this work, we aim to express the task proximity in the multi-task network.
+Instead of using a predefined multi-task network, we jointly learn routing task-dependent sub-networks and the parameters of the whole network.
+More precisely, we propose clustering the nodes of each layer into equal-sized groups.
+For each layer and each domain, a number of groups will be dropped.
+By consequence, we assign different sparse sub-networks to different tasks.
+Our method learns the task-dependent dropping selections and the parameters of the underlying network.
+Our approach allows the proximity between tasks to express in their dropping selections.
+Through an extensive set of experiments, we demonstrate the advantages of our method over strong baselines in multi-domain and multilingual translation.
+Furthermore, through an ablation study, we illustrate a significant correlation between the similarity between learned task-dependent dropping selections and the proximity between tasks.
 \end{abstract}
@@
 \section{Introduction}
-Multi-domain translation and multilingual translation develop one model for translations in multi-domain and multiple language pairs respectively.
-These paradigms are motivated by the compactness of the translation model and the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} and between languages in the same family \citep{Tan19multilingual}.
+Multi-domain translation and multilingual translation develop one model for translations in multi-domain and multiple language pairs, respectively.
+These paradigms are motivated by the compactness of the translation model \cite{dabre20survey,Chu18multilingual} and the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
+However, sharing parameters of the model can cause the catastrophic interference between unrelated tasks \citep{conneau20unsupervised,wang20negative}.
 
-\citet{Christopher21efficiently}
+In multi-domain setting, \citet{Pham21revisiting} pointed out a challenge in how multi-domain systems fully exploit the similarity between domains.
+The authors demonstrated the weakness of these systems in handling the proximity between domains through a simple setting using a pseudo domain separation.
+Domain adapters, which are strong baselines for multi-domain translation \cite{Pham20Study}, failed this test.
+Such predefined allocations of parameters to domains do not take into account the proximity of domains in their choice.
 
-The method we developed here mitigate the catastrophic interference occurring in training a neural machine translation (NMT) model for multiple domains or multiple language pairs.
+The method we developed here mitigates the catastrophic interference in training a neural machine translation (NMT) model for multiple domains or multiple language pairs.
+Our main contribution are summarized below:
+\begin{itemize}
+\setlength{\itemsep}{1pt}
+  \setlength{\parskip}{0pt}
+  \setlength{\parsep}{0pt}
+	\item We provide a rigorous mathematical modeling the problem of jointly learning routing task-dependent sub-network and the parameters of the underlying model using variational probabilistic modeling.
+	\item Our method propose end to end training with very little extra computational cost.
+	\item With an extensive experimental setting, we demonstrate a significant gain of our method against strong baselines.
+	\item We illustrate a significant correlation between the similarity between learned sub-networks and the proximity of their corresponding tasks.
+\end{itemize}
 
@@
 We define a group dropout for a domain $d$ as follow
-\begin{equation}
-\begin{array}{rcl}
-\tilde{h}^l &=& h^l \odot r_l^d ,\\
-r_l^d \in \{ 0,1 \}^{d_k}, \\
-h^{l+1} &=& \text{LAYER}^{l+1}(h^l) ,\\
-l & \in & [0,\cdots,L] ,\\
-\end{array}
-\end{equation}
-where $L$ is number of Transformer layers of our NMT model, the binary vector $r_l^d$ is a dropping mask. The dropping mask $r_l^d$ of $l^{th}$ layer and domain $d$ is defined as follow
-\begin{equation}
-\begin{array}{rcl}
-r_l^d(i) &=& \begin{cases}
+\begin{align*}
+\tilde{h}^l &= h^l \odot r_l^d ,\\
+r_l^d &\in \{ 0,1 \}^{d_k}, \\
+h^{l+1} &= \text{LAYER}^{l+1}(h^l) ,\\
+l & \in [0,\cdots,L] ,\\
+\end{align*}
+where $L$ is number of Transformer layers of our NMT model, the binary vector $r_l^d$ is a dropping mask. We illustrate the group dropout in Figure \ref{fig:group_dropout}
+\begin{figure}
+\includegraphics[width=0.4\textwidth]{group_dropout}
+\caption{Latent group dropout.
+The set of nodes in one layer is divided into equal-sized groups.
+For a domain, we drop a fixed number of groups.
+Our method learns which groups will be dropped for a domain.}
+\label{fig:group_dropout}
+\end{figure}
+ The dropping mask $r_l^d$ of $l^{th}$ layer and domain $d$ is defined as follow
+\begin{align*}
+r_l^d(i) &= \begin{cases}
       1, & \text{if}\ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p} \\
@@
     \end{cases} \\
-p & \in & \{0,\dots,n_p\text{-}1 \}, \\
-d & \in & \{0,\dots,n_d\text{-}1 \}, \\
-i & \in & \{0,\dots,d_k\text{-}1 \}, \\
-m_l^d & \in & \{0,1\}^{n_p} \ \forall d,l \, \\
-\mid m_l^d \mid_{L_1} & = & k , \\
-\Delta^{n_p}_k & = & \{ m \in \{0,1\}^{n_p} | \mid m \mid_{L_1} = k \},
-\end{array}
-\end{equation}
+p & \in  \{0,\dots,n_p\text{-}1 \}, \\
+d & \in  \{0,\dots,n_d\text{-}1 \}, \\
+i & \in  \{0,\dots,d_k\text{-}1 \}, \\
+m_l^d & \in  \{0,1\}^{n_p} \ \forall d,l \, \\
+\mid m_l^d \mid_{L_1} & =  k , \\
+\Delta^{n_p}_k & =  \{ m \in \{0,1\}^{n_p} | \mid m \mid_{L_1} = k \},
+\end{align*}
 where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
@@
 \begin{align}
-\hspace{-1.em}
+\hspace{-3.em}
 \text{log} P(y|x,\theta,d) & \text{=} \text{log} P(y, m_l^d |x,\theta,d) \text{-} \text{log} P(m_l^d|y, x,\theta,d) \nonumber \\
-& \text{=} \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
+& \text{=} \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
 & \text{+} KL(P(m_l^d | \Phi_l^d) \parallel P(m_l^d|y, x,\theta,d)) \nonumber \\
-& \geqslant \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
-& \text{=} \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y | x,\theta,m_l^d, d) \big] \label{eq:lower-bound} \\
+& \geqslant \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
+& \text{=} \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y | x,\theta,m_l^d, d) \big] \label{eq:lower-bound} \\
 & \text{-} KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\
@@
 \hspace{-2.em}
-T_1 &= \mathbb{E}_{P(m_l^d | \Phi_l^d)} \big[ \text{log} P(y | x,\theta,m_l^d, d)  \big] \nonumber \\
-&= \mathbb{E}_{g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} Gumbel(0,1) \ } \big[ \text{log} P(y | x,\theta,\tilde{m}_l^d,d) \big] \nonumber
+T_1 &= \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y | x,\theta,m_l^d, d)  \big] \nonumber \\
+&= \displaystyle{\mathop{\mathbb{E}}_{g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} Gumbel(0,1) \ }} \big[ \text{log} P(y | x,\theta,\tilde{m}_l^d,d) \big] \nonumber
 \end{align}
@@
 \end{align}
-%
+and
 \begin{align}
@@
 	&= \mathbb{H} \big[ P(m_l^d | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\ 
-	&= \mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\ \
+	&= \mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\
 	&\geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \label{eq:entropy}
@@
 
-\begin{equation}
-\mathcal{L}(\theta,\Phi_l^1,\cdots,\Phi_l^k) = \displaystyle{\mathop{\sum}_{d=1}^{k}} \mathbb{E}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)} \big[-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)\big]
-\end{equation}
+\begin{align}
+\hspace{-3.em}
+\mathcal{L}(\theta,\Phi_l^1,\cdots,\Phi_l^k) = \displaystyle{\mathop{\sum}_{d=1}^{k}} \displaystyle{\mathop{\mathbb{E}}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)}} \big[-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)\big]
+\end{align}
+
 in which $\mathcal{D}_s^d$ is distribution of the task $d$ over the input space $\Omega^d_s$; $g^d: \Omega^d_s \rightarrow \Omega^d_t$ is the mapping function, which our multi-task modern needs to learn, of the task $d$; $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$ is the ELBO loss, which we defined in Equation \ref{eq:lower-bound}.
+
+Finally, in the inference, we define the dropping mask for $l^{th}$ layer and domain $d$ as follow
+
+\begin{align*}
+Ind_l^d &= Top_k(\Phi_l^d) \\
+m_l^d &= \mathbb{I}(i\in Ind_l^d)\\
+\end{align*}
+
 \section{Experimental settings}
@@
 \end{table*}
+\subsubsection{Fuzzy domain separation}
+In this experiment, we reuse an experiment proposed by \citet{Pham21revisiting} which measures the efficiency of a multi-domain NMT system exploit the proximity between domains.
+The experiment use the same data as the previous experiment.
+However, the domain \domain{LAW} is now randomly split into two pseudo domains \domain{law$_1$} and \domain{law$_2$}.
+If the multi-domain system made full use of the proximity between \domain{law$_1$} and \domain{law$_2$}, there should be no significant difference between the performance of the system trained with six original domains and with seven domains replacing \domain{law} by \domain{law$_1$} and \domain{law$_2$}.
+\citet{Pham21revisiting} showed a large gap between two training settings when using residual adapters.
+We report below the gap between two settings when using \system{LaMGD}.
+
 \subsection{Multilingual translation}
@@
 However, \system{LaMGD} lags behind multilingual \system{Transformer} in O2M-diverse.
-\begin{table*}[ht]
+\begin{table*}[h!]
   \centering
@@
     \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.80 &7.30&20.80&21.10&39.70&19.80&22.60&15.20&18.91 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} & & & & & & & & & \\ 
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &4.30&6.80&21.10&22.00&39.70&20.00&23.00&15.20&19.01 \\ 
     \system{LaMGD}  \hfill{\footnotesize[+0m]}  & 5.20&\SB{9.40}&20.60&\SB{22.80}&39.60&19.60&22.40&15.00&19.33 \\ 
@@
     \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.40&16.60&28.50&	27.10&43.70&24.60&30.30&25.60&25.98 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} & & & & & & & & & \\ 
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &10.10&15.80&28.40&26.80&43.70&24.50&30.20&25.60&25.64\\ 
     \system{LaMGD}   \hfill{\footnotesize[+0m]}  &11.30&\SB{17.40}&28.60&\SB{28.70}&43.70&24.50&30.70&25.60&26.31 \\ 
@@
     \hline
-    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
+    M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
     \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.40&9.70&20.50&31.80&37.50&38.70&39.80&19.00&27.43 \\
@@
 \subsection{Similarity between dropping masks}
+In this section, we illustrate the similarity between dropping masks.
 
 \section{Related work}
+The most adopted paradigm is to have two groups of parameters including task-agnostic parameters and task-specific parameters.
+Domain adapters and language adapters are strong baselines of this paradigm.
+However, adapters does not take into account the proximity of the task in their designed architectures.
+
+\citep{sen19multilingual,kong21multilingual} used the similar idea of language proximity based parameter sharing.
+The authors proposed a multi-decoder model sharing the same encoder among languages and routed languages in different families to different decoders.
+
+\citet{xian20deep} proposed using different depths for different tasks.
+The authors aimed to match the depth of the sub-network to the complexity of the task.
+\citet{Gong21pay,Gong21adaptive} also take interest in the adaptive sparse Transformers which differ the selection of heads in multi-head attention, of layers and of blocks in feed-forward matrices between different tasks.
+\citet{Christopher21efficiently}.
+Mixture-of-experts (MoE) is another effective approach of sparse modeling.
+Within the transformer model, GShard model replaces a single feedforward (FFN) sub-layer with an MoE module consisting of multiple FFN sub-layers \citep{lepikhin21gshard}.
+
+%\citet{william21switch}
+
+%Adapters
+%
+%With the similar idea of language proximity based parameter sharing, recent works propose multi-decoder model.
+Sharing the same encoder among languages, they route languages in different families to different decoders
+
 \section{Conclusions and outlook}
@@
 \label{appendix:a}
-Proof of limit. Computation of $\hat{m}_l^d(\tau)$. Theorem of implicit differentiation and Proof of differentiability.
+This section explains how we calculate $\hat{m}_l^d(\tau)$ by solving the optimization problem \eqref{eq:soft-top-k} and then how to compute the gradient $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$.
+
+First, to solve \eqref{eq:soft-top-k} we follow the same approach as in \cite{amos19lml,amos20differential} by applying the Karush–Kuhn–Tucker (KKT) conditions to \eqref{eq:soft-top-k}.
+The solution of \eqref{eq:soft-top-k} will have the following form
+
+\begin{align}
+\hat{m}_l^d(\tau) &= \sigma(\frac{g_l^d + \Phi_l^d + \bar{\nu}}{\tau}) \label{eq:soft-m}
+\end{align}
+in which $\sigma(.)$ is sigmoid function and $\bar{\nu}$ is the solution of the following equation
+\begin{align}
+\displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = k \label{eq:nu}
+\end{align}
+
+Because sigmoid is monotonic, the equation \eqref{eq:nu} has unique solution.
+Furthermore,  because of the smoothness of $g(\nu,\Phi_l^d) = \displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau})$ w.r.t $\nu$ and $\Phi_l^d$, we could deduce the implicit differentiation of its solution $\bar{\nu}$ w.r.t $\Phi_l^d$ as below although the solution of \eqref{eq:nu} does not have explicit form.
+
+\begin{align*}
+&\frac{\partial g}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d} + \frac{\partial g}{\partial \Phi_l^d} = 0 \\
+& \Rightarrow \frac{\partial \nu}{\partial \Phi_l^d} = - \big(\frac{\partial g}{\partial \nu}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}
+\end{align*}
+
+Because the differentiation of sigmoid has exact forms, $\frac{\partial g}{\partial \nu}$ and $\frac{\partial g}{\partial \Phi_l^d}$ also have exact form.
+Therefore, we do not need autograd to compute the implicit gradient $\frac{\partial \nu}{\partial \Phi_l^d}$.
+The gradient of $\hat{m}_l^d(\tau)$ w.r.t $\Phi_l^d$ is deduced as follow
+\begin{align}
+\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} = \frac{\partial \hat{m}_l^d(\tau)}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d}
+\end{align}
+
+In our algorithm, we solve \eqref{eq:nu} by binary search.
+The convergence of binary search is extremely fast and assured by the monotonicity of $g(\nu,\Phi_l^d)$.
+In our experiments, we set our search range to $[-100,100]$.
+
+Finally, we need prove the limit $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
+We suppose $g_l^d(i_1) + \Phi_l^d(i_1) > g_l^d(i_2) + \Phi_l^d(i_2) > \cdots > g_l^d(i_{n_p}) + \Phi_l^d(i_{n_p})$.
+
+Because
+
+\begin{align*}
+\lim_{\tau \rightarrow 0} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = \begin{cases}
+      1, & \text{if}\ \tau > - (g_l^d(i) + \Phi_l^d(i)), \\
+      0, & \text{if}\ \tau < - (g_l^d(i) + \Phi_l^d(i)), \\
+      \frac{1}{2} &\text{otherwise}\\
+    \end{cases}
+\end{align*}
+
+and
+
+\begin{align*}
+\displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = k
+\end{align*}
+
+there exist $\epsilon$ such that $\forall \tau < \epsilon$, the solution $\bar{\nu}$ of \eqref{eq:nu} is between $- (g_l^d(i_{k+1}) + \Phi_l^d(i_{k+1})) > \bar{\nu} > - (g_l^d(i_k) + \Phi_l^d(i_k))$.
+Furthermore, because sigmoid is monotonic,
+
+\begin{align*}
+&\sigma(\frac{g_l^d(i) + \Phi_l^d(i) - (g_l^d(i_{k}) + \Phi_l^d(i_{k}))}{\tau}) < \hat{m}_l^d(\tau)(i)  \\
+& < \sigma(\frac{g_l^d(i) + \Phi_l^d(i) - (g_l^d(i_{k+1}) + \Phi_l^d(i_{k+1}))}{\tau}) 
+\end{align*}
+
+By taking the limit of both sides, we have following limits
+
+\begin{align*}
+\lim_{\tau \rightarrow 0} \hat{m}_l^d(\tau)(i_u) = \begin{cases}
+      1, & \text{if}\ u > k \\
+      0, & \text{if}\ u < k \\
+    \end{cases}
+\end{align*}
+
+And, because $ \displaystyle{\mathop{\sum}_{u=1}^{n_p}}\hat{m}_l^d(\tau)(i_u) = k$, by taking the limit at both side, we will have $\lim_{\tau \rightarrow 0} \hat{m}_l^d(\tau)(i_k) = 1$.
+Finally, we have
+
+\begin{align*}
+\lim_{\tau \rightarrow 0} \hat{m}_l^d(\tau)(i_u) = \begin{cases}
+      1, & \text{if} u \geqslant k \\
+      0, & \text{if} u < k \\
+    \end{cases}
+\end{align*}
+
+which is equivalent to $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
+
 \section{Appendix B}
@@
 \begin{align*}
-&\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] = - \mathbb{E}_{i_1,\cdots,i_k | \Phi_l^d} \big[ \text{log}P(i_1,\cdots,i_k | \Phi_l^d) \big] \\
-&= -\mathbb{E}_{i_1,\cdots,i_k | \Phi_l^d} \big[ \displaystyle{\sum_{j=2}^k}\text{log}P(i_j | i_1,\cdots,j_{j-1},\Phi_l^d) +  \text{log} P(i_1 | \Phi_l^d) \big] \\
-&\geqslant -\mathbb{E}_{i_1,\cdots,i_k | \Phi_l^d} \big[ \text{log} P(i_1 | \Phi_l^d) \big] \\
-&= -\mathbb{E}_{i_1 | \Phi_l^d} \big[ \text{log} P(i_1 | \Phi_l^d) \big] = \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]
+&\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] = - \displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \text{log}P(i_1,\cdots,i_k | \Phi_l^d) \big] \\
+&= -\displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \displaystyle{\sum_{j=2}^k}\text{log}P(i_j | i_1,\cdots,j_{j-1},\Phi_l^d) +  \text{log} P(i_1 | \Phi_l^d) \big] \\
+&\geqslant -\displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] \\
+&= -\displaystyle{\mathop{\mathbb{E}}_{i_1 | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] = \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]
 \end{align*}
@@
   \centering
-  \begin{tabular}{|l|ccccccc|} %*{4}{|r|}}
-    \cline{2-8} 
-    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
-    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} & \multicolumn{1}{c|}{\domain{news}} \\
+  \begin{tabular}{|l|cccccc|} %*{4}{|r|}}
+    \cline{2-7} 
+    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} \\
     \hline 
-    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
-    \# \revisiondone{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 & 7.8 / 9.2   \\
-    \# \revisiondone{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 & 77.8 / 80.4 \\
-    \# \revisiondone{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 & 31.5 / 23.1 \\
+    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) \\
+    \# \revisiondone{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 \\
+    \# \revisiondone{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 \\
+    \# \revisiondone{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 \\
     \hline

### 2021-11-09 2022/acl/acl_latex-fy.tex
suggestions by fy

--- 2021-11-08 2022/acl/acl_latex.tex
@@
 \title{Multi-domain, multilingual translation with latent multi-task group dropout}
-
+\title{Latent group dropout for multilingual and multidomain machine translation}
 % Author information can be set in various styles:
@@
 
+\begin{abstract}
+  [FY] Multi-domain and multi-lingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalties of the task at hand, while smaller parts are reserved to model the peculariaties of a language or a domain.
+In adapter-based approaches, these strategies are hard coded in the network architecture.
+In this work, we propose a new method aimed at learning the parameter sharing pattern, using a latent-variable model.
+We show how this model can be trained end to end, and report experimental results which show that the learned pattern are both meaningful, and yield improved translation performance without any increase of the model size.
+\end{abstract}
+
 \section{Introduction}
-Multi-domain translation and multilingual translation develop one model for translations in multi-domain and multiple language pairs, respectively.
-These paradigms are motivated by the compactness of the translation model \cite{dabre20survey,Chu18multilingual} and the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
-However, sharing parameters of the model can cause the catastrophic interference between unrelated tasks \citep{conneau20unsupervised,wang20negative}.
+Multi-domain and multilingual machine translation (MMT) aim to develop one single model to perform translations for multiple domains and multiple language pairs, respectively.
+These paradigms are motivated by the compactness of the resulting translation system \citep{dabre20survey,Chu18multilingual}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
+However, having all the tasks use exactly the same model parameters can cause negative interference between unrelated tasks \citep{conneau20unsupervised,wang20negative}.
+Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using the adapter layer of \citet{Bapna19simple,houlsby19parameter,Pham20Study}.
 

### 2021-11-09 2022/acl/acl_latex-fy.tex
intro rewrite

@@
 \newcommand{\SW}[1]{\underline{#1}}
+\newcommand{\mtsquare}[0]{MT$^2$}
 
@@
 \begin{abstract}
-Multi-domain and multilingual translation can be cast under multi-task learning in which each domain or language pair is considered a task.
+Multidomain and multilingual translation can be cast under multi-task learning in which each domain or language pair is considered a task.
 Sharing most of the parameters of the model causes a catastrophic interference between unrelated tasks.
@@
 \begin{abstract}
-  [FY] Multi-domain and multi-lingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalties of the task at hand, while smaller parts are reserved to model the peculariaties of a language or a domain.
+  [FY] Multidomain and multilingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalties of the task at hand, while smaller parts are reserved to model the peculariaties of a language or a domain.
 In adapter-based approaches, these strategies are hard coded in the network architecture.
@@
 \section{Introduction}
-Multi-domain and multilingual machine translation (MMT) aim to develop one single model to perform translations for multiple domains and multiple language pairs, respectively.
-These paradigms are motivated by the compactness of the resulting translation system \citep{dabre20survey,Chu18multilingual}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
+Multidomain and multilingual machine translation aim to develop one single model to perform translations for multiple domains and multiple language pairs, respectively.\footnote{We will refer to these two situations as 'multi-task MT' (or \mtsquare{} for short), and refer to individual domains and languages as 'tasks'.} These paradigms are motivated by the compactness of the resulting translation system \citep{dabre20survey,Chu18multilingual}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
 However, having all the tasks use exactly the same model parameters can cause negative interference between unrelated tasks \citep{conneau20unsupervised,wang20negative}.
-Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using the adapter layer of \citet{Bapna19simple,houlsby19parameter,Pham20Study}.
+Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using adapter layers as studied in \citet{houlsby19parameter,Bapna19simple,Pham20Study}.
+If these techniques have proven effective to achieve strong baselines, such fail to fully take advantage of the similarities that exist between domains and tasks, as documented eg.\ in \citep{Pham21revisiting}.
+This is because the partition of the parameter space between generic or task-specific sub-parts, and their allocation to each task is hard coded in the network, irrespective of the existing commonalties and differences in the data space.
 
-In multi-domain setting, \citet{Pham21revisiting} pointed out a challenge in how multi-domain systems fully exploit the similarity between domains.
-The authors demonstrated the weakness of these systems in handling the proximity between domains through a simple setting using a pseudo domain separation.
-Domain adapters, which are strong baselines for multi-domain translation \cite{Pham20Study}, failed this test.
-Such predefined allocations of parameters to domains do not take into account the proximity of domains in their choice.
-
-The method we developed here mitigates the catastrophic interference in training a neural machine translation (NMT) model for multiple domains or multiple language pairs.
+In this paper, we try to develop techniques that can take the similarity between tasks into account in a more effective way, by learning the network organization from the data.
+To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the parameter space, and show how training can still be performed end-to-end using a variational surrogate of the standard loss function.
+Our experiments with multilingual and multidomain confirm that this method can automatically detect similarities in the data, and yield related tasks to share more parameters.
+They also show that this method is comparable to using adapter layers in a number of empirical comparisons; contrarily to adapters, these performance are obtained without any increase of the model size.
 Our main contribution are summarized below:
@@
   \setlength{\parsep}{0pt}
-	\item We provide a rigorous mathematical modeling the problem of jointly learning routing task-dependent sub-network and the parameters of the underlying model using variational probabilistic modeling.
-	\item Our method propose end to end training with very little extra computational cost.
-	\item With an extensive experimental setting, we demonstrate a significant gain of our method against strong baselines.
-	\item We illustrate a significant correlation between the similarity between learned sub-networks and the proximity of their corresponding tasks.
+\item We provide a rigorous mathematical formulattion to the problem of jointly learning routing task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling;
+\item We introduce algorithms to train this model end to end with very little extra computational cost.
+\item We report, using an extensive experimental setting, significant gains of this method with respect to strong baselines.
+\item We study how this method can actually exploit the similarities between tasks to learn consistent sub-networks.
 \end{itemize}
+
+\fyTodo{Can you do distillation ? That is only keep the parameter for one domain and still get good results (better than DA ?)}
 

### 2021-11-09 2022/acl/acl_latex.tex
local changes

@@
 \begin{document}
-\setlength{\abovedisplayskip}{3pt}
-\setlength{\belowdisplayskip}{3pt}
+\setlength{\parskip}{0pt}
+\setlength{\abovedisplayskip}{0pt}
+\setlength{\belowdisplayskip}{0pt}
+\setlength{\abovedisplayshortskip}{0pt}
+\setlength{\belowdisplayshortskip}{0pt}
 \maketitle
@@
 Most of the proposed multi-task architectures propose sharing the parameters of an underlying generic model and using different output layers to predict different tasks.
-However, these paradigms do not explicitly exploit the proximity between tasks.
+However, these methods do not explicitly exploit the proximity between tasks.
 In this work, we aim to express the task proximity in the multi-task network.
@@
 The authors demonstrated the weakness of these systems in handling the proximity between domains through a simple setting using a pseudo domain separation.
-Domain adapters, which are strong baselines for multi-domain translation \cite{Pham20Study}, failed this test.
+Domain Adapters, which are strong baselines for multi-domain translation \cite{Pham20Study}, failed this test.
 Such predefined allocations of parameters to domains do not take into account the proximity of domains in their choice.
@@
 \section{Multi-task group dropout}
-We define a group dropout for a domain $d$ as follow
+We define a group dropout for a domain $d$ as follows
 \begin{align*}
@@
 \end{figure}
- The dropping mask $r_l^d$ of $l^{th}$ layer and domain $d$ is defined as follow
+ The dropping mask $r_l^d$ of $l^{th}$ layer and domain $d$ is defined as follows
 \begin{align*}
@@
 \begin{align}
-\hspace{-3.em}
+\hspace{-0.em}
 \mathcal{L}(\theta,\Phi_l^1,\cdots,\Phi_l^k) = \displaystyle{\mathop{\sum}_{d=1}^{k}} \displaystyle{\mathop{\mathbb{E}}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)}} \big[-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)\big]
@@
 
-Finally, in the inference, we define the dropping mask for $l^{th}$ layer and domain $d$ as follow
+Finally, in the inference, we define the dropping mask for $l^{th}$ layer and domain $d$ as follows
 
@@
 	\item Adapter based Transformer
-	\item Transformer with multi-task group dropout
+	\item \system{LaMGD} Tranformer
 \end{itemize}
@@
 \subsection{Multi-domain translation}
-According to Table \ref{tab:mdmt}, \system{LaMGD} shows a significant improvement (+2.78) over the generic system \system{Transformer} with zero extra parameters.
-Moreover, \system{LaMGD} achieve equivalent performance in average compared to \system{Adapter}, which is carrefully finetuned with a approximately 12.4M additional parameters.
+According to Table \ref{tab:mdmt}, \system{LaMGD} Tranformer shows a significant improvement (+2.78) over the generic system \system{Transformer} with zero extra parameters.
+Moreover, \system{LaMGD} Tranformer achieves equivalent performance in average compared to \system{Adapter}, which is carrefully finetuned with a approximately 12.4M additional parameters.
 \begin{table*}[h!]
@@
     \revision{\system{Adapter}}   \hfill{\footnotesize[+12.4m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.28 \\ 
-    \system{LaMGD}   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & 39.0 & 52.4 & 87.5 & 55.33 \\ 
+    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & 39.0 & 52.4 & 87.5 & 55.33 \\ 
     \hline
@@
 \subsubsection{Fuzzy domain separation}
+\label{ssec:fuzzy}
 In this experiment, we reuse an experiment proposed by \citet{Pham21revisiting} which measures the efficiency of a multi-domain NMT system exploit the proximity between domains.
@@
 \citet{Pham21revisiting} showed a large gap between two training settings when using residual adapters.
-We report below the gap between two settings when using \system{LaMGD}.
+We report below the gap between two settings when using \system{LaMGD} Tranformer.
+\begin{table*}[h!]
+  \centering
+  \begin{tabular}{|p{4cm}|*{2}{r|}} \hline
+    Model / Domain & \multicolumn{1}{c|}{\domain{law$_1$}} & \multicolumn{1}{c|}{\domain{law$_2$}} \\ \hline 
+    \system{Adapter}   \hfill{\footnotesize[+25m]}  & -0.6 & -0.8 \\ 
+    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 0.0 & 0.0 \\ 
+    \hline
+  \end{tabular}
+  \caption{Fuzzy domain separation}
+  \label{tab:fuzzy}
+\end{table*}
+
+Table \ref{tab:fuzzy} demonstrates a significant loss when training with two pseudo domain \domain{law$_1$} and \domain{law$_2$} compared to training with the real domain \domain{law}.
+Our system does not have this loss between two situation.
+In section \ref{ssec:abalation}, we show that our algorithm outputs the same sub-network for \domain{law$_1$} and \domain{law$_2$}, that allows a full transfer between two domain.
 
 \subsection{Multilingual translation}
-According to Table \ref{tab:multilingual}, \system{LaMGD} shows a boost of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, M2O-diverse respectively.
-Significant gains are observed in \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction).
-However, \system{LaMGD} lags behind multilingual \system{Transformer} in O2M-diverse.
+According to Table \ref{tab:multilingual}, \system{LaMGD} Tranformer shows a boost of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, M2O-diverse respectively.
+Significant gains are observed in \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resourced languages in the list.
+However, \system{LaMGD} Tranformer lags behind multilingual \system{Transformer} in O2M-diverse.
 \begin{table*}[h!]
@@
     \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &4.30&6.80&21.10&22.00&39.70&20.00&23.00&15.20&19.01 \\ 
-    \system{LaMGD}  \hfill{\footnotesize[+0m]}  & 5.20&\SB{9.40}&20.60&\SB{22.80}&39.60&19.60&22.40&15.00&19.33 \\ 
+    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]}  & 5.20&\SB{9.40}&20.60&\SB{22.80}&39.60&19.60&22.40&15.00&19.33 \\ 
 	\hline
@@
     \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &10.10&15.80&28.40&26.80&43.70&24.50&30.20&25.60&25.64\\ 
-    \system{LaMGD}   \hfill{\footnotesize[+0m]}  &11.30&\SB{17.40}&28.60&\SB{28.70}&43.70&24.50&30.70&25.60&26.31 \\ 
+    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  &11.30&\SB{17.40}&28.60&\SB{28.70}&43.70&24.50&30.70&25.60&26.31 \\ 
     \hline
@@
     O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.20&4.0&12.70&22.20&31.80&34.00&38.30&8.30&20.19 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  & & & & & & & & & \\ 
-    \system{LaMGD}   \hfill{\footnotesize[+0m]} & & & & & & & & & \\
+    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.20&4.0&12.70&22.20&31.80&\SB{34.00}&38.30&8.30&20.19 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &10.20&4.00&\SB{13.30}&21.90&32.20&\SB{34.10}&38.50&8.30&20.31 \\ 
+    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}&10.10&3.80&12.60&\SB{22.80}&31.80&33.40&38.10&8.10&20.09\\
     \hline 
@@
     \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.40&9.70&20.50&31.80&37.50&38.70&39.80&19.00&27.43 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  & & & & & & & & & \\ 
-    \system{LaMGD}   \hfill{\footnotesize[+0m]} &\SB{23.50}&9.60&\SB{21.50}&32.20&37.70&38.60&40.00&18.90&27.75 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.50&9.40&20.00&30.60&37.20&38.20&39.30&19.00&27.03\\ 
+    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]} &\SB{23.50}&9.60&\SB{21.50}&32.20&37.70&38.60&40.00&18.90&27.75 \\
     \hline
@@
 \end{table*}
+
 \subsection{Similarity between dropping masks}
-In this section, we illustrate the similarity between dropping masks.
+\label{ssec:abalation}
+In this section, we compare the sub-networks of the domains or of the language pairs by computing the average of the similarities between the corresponding dropping mask in each layer.
+
+For the multi-domain experiment, we analyze the case of fuzzy domain separation in Section \ref{ssec:fuzzy}.
+Table \ref{tab:fuzzy-sim} shows that the sub-networks of \domain{law$_1$} and \domain{law$_2$} are identical, which is one of the goals of our method.
+\begin{table*}[h!]
+  \centering
+  \begin{tabular}{|p{1cm}|*{7}{r|}} \hline
+    & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{bank}}& \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{law$_1$}} & \multicolumn{1}{c|}{\domain{law$_2$}} & \multicolumn{1}{c|}{\domain{rel}} &\multicolumn{1}{c|}{\domain{talk}} \\ \hline
+    \domain{med} &1.0&0.80&0.85&0.83&0.83&0.78&0.89\\
+    \domain{bank} &0.80&1.0&0.80&0.90&0.90&0.73&0.82\\
+    \domain{it} &0.85&0.80&1.0&0.80&0.80&0.78&0.85\\
+    \domain{law$_1$} &0.83&0.90&0.80&1.0&\SB{1.0}&0.75&0.82\\ 
+    \domain{law$_2$} &0.83&0.90&0.80&\SB{1.0}&1.0&0.75&0.82\\ 
+    \domain{rel} &0.78&0.73&0.78&0.75&0.75&1.0&0.77\\
+    \domain{talk} &0.89&0.82&0.85&0.82&0.82&0.77&1.0\\
+    \hline
+  \end{tabular}
+  \caption{Multi-domain sub-networks' similarities}
+  \label{tab:fuzzy-sim}
+\end{table*}
+
+For multilingual case TED-related, we have four language families including (1) Turkic family with Azerbaijani and Turkish(\domain{aze},\domain{tur}); (2) Slavic family with Belarusian and Russian (\domain{bel},\domain{rus}); (3) Romance family with Glacian and Portuguese (\domain{glg},\domain{por}); and (4) Czech-Slovak family with Slovak and Czech (\domain{ces},\domain{slk}).
+The similarities between the sub-networks of each pair are 0.95, 0.94, 0.99 and 0.96 which are the maximum in each row except the diagonal, which stands for the self-similarity.
+\begin{table*}[h!]
+  \centering
+  \begin{tabular}{|p{1cm}|*{8}{r|}} \hline
+    & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{bel}}& \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} &\multicolumn{1}{c|}{\domain{slk}}&\multicolumn{1}{c|}{\domain{tur}} \\ \hline
+    \domain{aze} &1.0&0.90&0.89&0.86&0.85&0.91&0.89&\SB{0.95}\\
+    \domain{bel} &0.90&1.0&0.94&0.91&0.91&\SB{0.94}&0.94&0.88\\
+    \domain{ces} &0.89&0.94&1.0&0.92&0.94&0.98&\SB{0.99}&0.91\\
+    \domain{glg} &0.86&0.91&0.92&1.0&\SB{0.96}&0.92&0.93&0.86\\ 
+    \domain{por} &0.85&0.91&0.94&\SB{0.96}&1.0&0.93&0.94&0.87\\ 
+    \domain{rus} &0.91&\SB{0.94}&0.98&0.92&0.93&1.0&0.98&0.93\\
+    \domain{slk} &0.89&0.94&\SB{0.99}&0.93&0.94&0.98&1.0&0.91\\
+    \domain{tur} &\SB{0.95}&0.88&0.91&0.86&0.87&0.93&0.91&1.0\\
+    \hline
+  \end{tabular}
+  \caption{Multilingual sub-networks' similarities (M2O Ted-related)}
+  \label{tab:related-sim}
+\end{table*}
 
@@
 \citet{Gong21pay,Gong21adaptive} also take interest in the adaptive sparse Transformers which differ the selection of heads in multi-head attention, of layers and of blocks in feed-forward matrices between different tasks.
-\citet{Christopher21efficiently}.
 Mixture-of-experts (MoE) is another effective approach of sparse modeling.
 Within the transformer model, GShard model replaces a single feedforward (FFN) sub-layer with an MoE module consisting of multiple FFN sub-layers \citep{lepikhin21gshard}.
-
-%\citet{william21switch}
+\citet{william21switch}
 
@@
 \section{Conclusions and outlook}
+In this work, we present a novel method allows jointly searching optimal sub-networks for multiple tasks and the parameters of the underlying network.
+Our method proposed a rigorous mathematical modeling and an end-to-end optimization with very little extra computational cost.
+We demonstrate an improvement of our model with large margin over multi-domain Transformer without zero additional parameters and an equivalent performance compared to a strong baseline such as Domain Adapters consisting of large number of extra parameters.
+In multilingual translation, our model outperforms multilingual Transformer and Language Adapters in 3/4 cases.
+Besides, we provided an thorough ablation study on the similarities between learned sub-networks and demonstrated a strong correlation between the sub-networks' similarity and the proximity of their tasks (domain or language).
+
+Despite several advantages, our method has several weaknesses including the heuristically predefined number of retained groups, $k$ , and the need of domain label in the inference.
+In future work, we would like to provide an extensive variational framework to model both the number of retained groups, $k$ and the dropping selections.
+Secondly, to get rid of the requirement of domain information in the inference, we could condition the dropping selection directly on the input $x$ instead of the domain $d$.
+These two open question are both very interesting and promising.
 \bibliography{bibliography}
@@
 Therefore, we do not need autograd to compute the implicit gradient $\frac{\partial \nu}{\partial \Phi_l^d}$.
-The gradient of $\hat{m}_l^d(\tau)$ w.r.t $\Phi_l^d$ is deduced as follow
+The gradient of $\hat{m}_l^d(\tau)$ w.r.t $\Phi_l^d$ is deduced as follows
 \begin{align}
@@
 \begin{align*}
+\hspace{-3.em}
 \lim_{\tau \rightarrow 0} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = \begin{cases}
@@
 \lim_{\tau \rightarrow 0} \hat{m}_l^d(\tau)(i_u) = \begin{cases}
-      1, & \text{if} u \geqslant k \\
-      0, & \text{if} u < k \\
+      1, & \text{if}\ u \geqslant k \\
+      0, & \text{if}\ u < k \\
     \end{cases}
@@
 \subsection{Multi-domain datasets}
-
-\begin{table*}[h!]
+We describe our datasets for multi-domain translation experiment in Table \ref{tab:Corpora-chap4}
+\begin{table*}
   \centering
@@
 \end{table*}
-
 \subsection{Multilingual datasets}
+We clarify the notation of the languages in multilingual experiments as follows
+\begin{itemize}
+	\item Diverse: bos (bosnia), mar(marathi),
+         hin(hindi), mkd(macedonia), ell(greek),
+         bulgaria(bul), french(fra), korean(kor) 
+	\item Related: azerbajiani (aze), belarusian(bel),
+          glacian(glg), slovak (slk), turkisk (tur),
+		  russian(rus), portugese(por),czech(ces)
+\end{itemize}
+Table \ref{tab:Corpora-multilingual} describes the subset of TED8 that we used in our experiments.
+\begin{table*}
+  \centering
+  \begin{tabular}{l|cccc} 
+  \hline
+    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c}{\domain{test}} \\
+    \hline 
+    \multirow{8}{*}{Related} & Azerbaijani & 5.94k & 671 & 903 \\
+							 & Belarusian & 4.51k & 248 & 664 \\
+							 & Glacian & 10.0k & 682 & 1007 \\
+							 & Slovak & 61.5k & 2271 & 2445 \\
+							 & Turkish & 182k & 4045 & 5029 \\
+                             & Russian & 208k & 4814 & 5483 \\
+                             & Portuguese & 185k & 4035 & 4855 \\
+                             & Czech & 103k & 3462 & 3831 \\
+    \hline
+    \multirow{8}{*}{Related} & Bosnian & 5.64k & 474 & 463 \\
+                             & Marathi & 9.84k & 767 & 1090 \\
+                             & Hindi & 18.79k & 854 & 1243 \\
+                             & Macedonian & 25.33k & 640 & 438 \\
+                             & Greek & 134k & 3344 & 4433 \\
+                             & Bulgarian & 174k & 4082 & 5060 \\
+                             & French & 192k & 4320 & 4866 \\
+                             & Korean & 205k & 4441 & 5637 \\
+  \end{tabular}
+  \caption{Data Statistics of TED8 Datasets}
+\label{tab:Corpora-multilingual}
+\end{table*}
 
 \subsection{Multi-domain systems}
+We describe the choice of our systems in multi-domain experiments as follows
+\begin{itemize}
+	\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048, the multi-head attentions contain with 8 heads.
+	\item Adapter based Transformer: The intermediate feeforward dimension is set as 2048 as in \citet{Pham21Revisiting}.
+	\item \system{LaMGD} Tranformer: There is no change in the architecture.
+We group 512 nodes in each layer into 16 groups of 32 consecutive nodes.
+For each domain, only 12 groups are selected.
+The number of parameters of variational distributions are $k \times L \times n_d$ which is tiny compared to the size of Transformer.
+\end{itemize}
 
+We set the dropout probability as 0.1.
+We train the multi-domain Transformer model in 200~K iterations with a batchsize of 12~K tokens.
+We train \system{LaMGD} Tranformer in 300~K iterations with a batchsize of 12~K tokens since \system{LaMGD} Tranformer has to learn both the sub-networks and the parameters of the underlying Transformer.
+Finally, we plug Domain Adapters to the multi-domain Transformer model and finetune them with 25~K iterations using the same batchsize as the former.
 \subsection{Multilingual systems}
+We describe the choice of our systems in multilingual experiments as follows
+\begin{itemize}
+	\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 1024, the multi-head attentions contain with 8 heads as in \citet{Wang20balancing}.
+	\item Adapter based Transformer: The intermediate feeforward dimension is set as 128 as in \citet{Gong21adaptive}.
+	\item \system{LaMGD} Tranformer: There is no change in the architecture.
+We group 512 nodes in each layer into 16 groups We group 512 nodes in each layer into 16 groups of 32 consecutive nodes.
+For each domain, only 12 groups are selectedof 32 consecutive nodes.
+For each domain, only 12 groups are selected.
+\end{itemize}
+
+We set the dropout probability as 0.3.
+We train the multilingual Transformer model in 40~K iterations with a batchsize of 9600 tokens on 16 GPUs as in \citet{Gong21adaptive}.
+We train \system{LaMGD} Tranformer in 50~K iterations with the same batchsize.
+Finally, we finetune the Language Adapters in 5~K iterations.
 

### 2021-11-09 2022/acl/acl_latex-fy.tex
start working on the model

@@
 \maketitle
-\begin{abstract}
-Multidomain and multilingual translation can be cast under multi-task learning in which each domain or language pair is considered a task.
-Sharing most of the parameters of the model causes a catastrophic interference between unrelated tasks.
-Most of the proposed multi-task architectures propose sharing the parameters of an underlying generic model and using different output layers to predict different tasks.
-However, these paradigms do not explicitly exploit the proximity between tasks.
-In this work, we aim to express the task proximity in the multi-task network.
-Instead of using a predefined multi-task network, we jointly learn routing task-dependent sub-networks and the parameters of the whole network.
-More precisely, we propose clustering the nodes of each layer into equal-sized groups.
-For each layer and each domain, a number of groups will be dropped.
-By consequence, we assign different sparse sub-networks to different tasks.
-Our method learns the task-dependent dropping selections and the parameters of the underlying network.
-Our approach allows the proximity between tasks to express in their dropping selections.
-Through an extensive set of experiments, we demonstrate the advantages of our method over strong baselines in multi-domain and multilingual translation.
-Furthermore, through an ablation study, we illustrate a significant correlation between the similarity between learned task-dependent dropping selections and the proximity between tasks.
-\end{abstract}
 
@@
 
-\section{Multi-task group dropout}
-We define a group dropout for a domain $d$ as follows
-\begin{align*}
-\tilde{h}^l &= h^l \odot r_l^d ,\\
-r_l^d &\in \{ 0,1 \}^{d_k}, \\
-h^{l+1} &= \text{LAYER}^{l+1}(h^l) ,\\
-l & \in [0,\cdots,L] ,\\
-\end{align*}
-where $L$ is number of Transformer layers of our NMT model, the binary vector $r_l^d$ is a dropping mask. We illustrate the group dropout in Figure \ref{fig:group_dropout}
+\section{Multi-task group dropout \label{sec:architecture}}
+\subsection{Network architecture, groups and layers}
+Most architectures for multitask learning are based on the assignment of model parameters with tasks.
+While our model also needs to know the task $d \in [1 \dots n_d$ for each training and test instance, the structure of our Transformer networks \citep{Vaswani17attention} is based on the notion of \emph{groups of parameters}.
+In each Transformer layer $l \in [1 \dots L]$, we identify $n_p$ groups of parameters.\fyTodo{In the attention heads only?
+In the FF?} The assignment of tasks to groups will be learned from the data, under the constraint that each task has only access to exactly $k$ groups of \emph{active parameters}, while the all the other parameter values are deemed to be \emph{dropped} (see Figure~\ref{fig:group_dropout}).
 \begin{figure}
@@
 \caption{Latent group dropout.
-The set of nodes in one layer is divided into equal-sized groups.
-For a domain, we drop a fixed number of groups.
-Our method learns which groups will be dropped for a domain.}
+The set of nodes\fyTodo{parameters ?} in each layer is divided into equal-sized groups.
+For a task, we only keep a fixed number of groups.
+Our method learns which groups will be kept / dropped for each task.}
 \label{fig:group_dropout}
 \end{figure}
- The dropping mask $r_l^d$ of $l^{th}$ layer and domain $d$ is defined as follows
+Formally, we define the group dropout mask $m_d^l$ a $n_p$ dimensional binary vector containing exactly $k$ ones: the group $p$ is retained for task $d$ if $m_l^d(p)\text{\small =}1$ and is dropped if $m_l^d(p)\text{\small =}0$.
+We denote $\Delta^{n_p}_k = \{ m \in \{0,1\}^{n_p} \text{such that} \}$ the set of all admissible dropping masks.
+
+$r_l^d$ for a task $d$ in layer $l$ as follows:
+\begin{align*}
+  r_l^d &\in \{ 0,1 \}^{d_k}, \\
+  \tilde{h}^l &= h^l \odot r_l^d ,\\
+  h^{l+1} &= \text{LAYER}^{l+1}(h^l) ,\\
+  l & \in [0,\cdots, L-1] ,\\
+\end{align*}
+where $\text{LAYER}^l()$ represents all the computations in transformer layer $l$. The dropping mask $r_l^d$ is computed as follows
 \begin{align*}
@@
 where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
-The group $p$ is retained for domain $d$ if $m_l^d(p)\text{\small ==}1$ and is dropped if $m_l^d(p)\text{\small ==}0$.
-We also denote $\Delta^{n_p}_k$ the set of all admissible dropping masks.
 

### 2021-11-09 2022/acl/acl_latex-fy.tex
more rewriting in the Math part

@@
 \subsection{Network architecture, groups and layers}
-Most architectures for multitask learning are based on the assignment of model parameters with tasks.
-While our model also needs to know the task $d \in [1 \dots n_d$ for each training and test instance, the structure of our Transformer networks \citep{Vaswani17attention} is based on the notion of \emph{groups of parameters}.
+Most architectures for multitask learning are based on a matching of subset of model parameters with tasks.
+While our model also needs to know the task $d \in [0 \dots n_d-1$ for each training and test instance, the structure of our Transformer networks \citep{Vaswani17attention} is emphbased on the notion of \emph{groups of parameters}.
 In each Transformer layer $l \in [1 \dots L]$, we identify $n_p$ groups of parameters.\fyTodo{In the attention heads only?
@@
 \begin{figure}
+  \centering
 \includegraphics[width=0.4\textwidth]{group_dropout}
@@
 \end{figure}
-Formally, we define the group dropout mask $m_d^l$ a $n_p$ dimensional binary vector containing exactly $k$ ones: the group $p$ is retained for task $d$ if $m_l^d(p)\text{\small =}1$ and is dropped if $m_l^d(p)\text{\small =}0$.
-We denote $\Delta^{n_p}_k = \{ m \in \{0,1\}^{n_p} \text{such that} \}$ the set of all admissible dropping masks.
+Formally, we define the \emph{group dropout mask} $m_d^l$ as a $n_p$-dimensional binary vector containing exactly $k$ ones: the group $p$ ($\in  [0,\dots,n_p\text{-}1]$) is retained for task $d$ if $m_l^d(p)\text{\small =}1$ and is dropped if $m_l^d(p)\text{\small =} 0$.
+We denote $\Delta^{n_p}_k = \{ m \in \{0,1\}^{n_p} \text{such that} \mid{} m \mid_{L_1} = k \}$, where $\mid{} m \mid_{L_1}$ is the $L_1$ norm, the set of all admissible dropping masks.
 
-$r_l^d$ for a task $d$ in layer $l$ as follows:
+Given $m_l^d$, the parameter mask $r_l^d$ for task $d$ in layer $l$ is then derived as follows:
+\begin{align*}
+  r_l^d(i) &= \begin{cases}
+    1, & \text{if}\ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p} \\
+    & \text{AND} \  m_l^d(p) \text{\small =} 1 \\
+    0, & \text{otherwise},
+  \end{cases} 
+  % p & \in  \{0,\dots,n_p\text{-}1 \}, \\
+  % d & \in  \{0,\dots,n_d\text{-}1 \}, \\
+  % i & \in  \{0,\dots,d_k\text{-}1 \}, \\
+  % m_l^d & \in  \{0,1\}^{n_p} \ \forall d,l \, \\
+  % \mid m_l^d \mid_{L_1} & =  k , \\
+  % \Delta^{n_p}_k & =  \{ m \in \{0,1\}^{n_p} | \mid m \mid_{L_1} = k \},
+\end{align*}
+where $d_k$ is the dimension of the hidden state.\fyTodo{Check this} The propagation of information within the network then depends on the current task value and can be formalized as:
 \begin{align*}
@@
   h^{l+1} &= \text{LAYER}^{l+1}(h^l) ,\\
-  l & \in [0,\cdots, L-1] ,\\
+  l & \in [0,\cdots, L-1] ,
 \end{align*}
-where $\text{LAYER}^l()$ represents all the computations in transformer layer $l$. The dropping mask $r_l^d$ is computed as follows
-\begin{align*}
-r_l^d(i) &= \begin{cases}
-      1, & \text{if}\ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p} \\
-      & \text{AND} \  m_l^d(p) \text{\small ==} 1 \\
-      0, & \text{otherwise},
-    \end{cases} \\
-p & \in  \{0,\dots,n_p\text{-}1 \}, \\
-d & \in  \{0,\dots,n_d\text{-}1 \}, \\
-i & \in  \{0,\dots,d_k\text{-}1 \}, \\
-m_l^d & \in  \{0,1\}^{n_p} \ \forall d,l \, \\
-\mid m_l^d \mid_{L_1} & =  k , \\
-\Delta^{n_p}_k & =  \{ m \in \{0,1\}^{n_p} | \mid m \mid_{L_1} = k \},
-\end{align*}
-where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
-
-We incorporate the latent variables $m_{l\in[0,\cdots,L]}^d$ into the log-likelihood of our model as below.
-For the sake of brevity, we denote $m_l^d$ the set of latent variables $m_{l\in[0,\cdots,L]}^d$
+where $\text{LAYER}^l()$ represents all the computations in Transformer layer $l$.
+% The dropping mask $r_l^d$ is computed as follows
+% where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
+\subsection{Training with latent dropout masks}
+We incorporate the latent variables  $m_{l}^d, l \in[0,\dots,L]$ into the log-likelihood as follows.
+For the sake of brevity, we denote $m_l^d$ the set of latent variables, \fyTodo{what do you mean here - this is all there is?} and use $x$, $y$, and $\theta$ to refer to the input, output, and parameter vectors, respectively; finally $\Phi^d_i \in \mathbb{R}^{n_p}$ denotes the set of variational parameters in the model:\fyTodo{Again this is complicated.}
 \begin{align}
 \hspace{-3.em}
-\text{log} P(y|x,\theta,d) & \text{=} \text{log} P(y, m_l^d |x,\theta,d) \text{-} \text{log} P(m_l^d|y, x,\theta,d) \nonumber \\
-& \text{=} \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
-& \text{+} KL(P(m_l^d | \Phi_l^d) \parallel P(m_l^d|y, x,\theta,d)) \nonumber \\
-& \geqslant \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
-& \text{=} \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y | x,\theta,m_l^d, d) \big] \label{eq:lower-bound} \\
-& \text{-} KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\
-& \text{=} \mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y). \nonumber
+\text{log} P(y|x,\theta,d) \text{=} &\text{log} P(y, m_l^d |x,d;\theta) \text{-} \text{log} P(m_l^d|y, x, d ;\theta) \nonumber \\
+\text{=} & \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,d;\theta) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
+& \quad + KL(P(m_l^d | \Phi_l^d) \parallel P(m_l^d|y, x,\theta,d)) \nonumber \\
+\geqslant & \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
+ = & \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y | x,\theta,m_l^d, d) \big] \label{eq:lower-bound} \\
+& \quad - KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\
+ = & \mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y). \nonumber
 \end{align}
-$\mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y)$ is the lowerbound, which is called the evidence lower bound (ELBO) in variational statistic \citep{Diederick14auto}, of our log-likelihood.
+$\mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y)$ is a lowerbound of the log-likelihood, known as the evidence lower bound (ELBO) in variational statistics \citep{Diederick14auto}, of our log-likelihood.
 Our training loss will be $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$.
 
-The variational distribution of $m_l^d$, $P(m_l^d|\Phi_l^d)$, is defined as follows
-
+The variational distribution of $m_l^d$, $P(m_l^d|\Phi_l^d)$, is defined as follows:
 \begin{align}
 \hspace{-2.em}
-&\Phi_d \in \mathbb{R}^{n_p} , \nonumber \\
-&Ind^d \text{=}  \{ i_1, \cdots, i_k \} \nonumber \\
-&\sim \text{\small Sampling\_without\_replacement} \big(\text{softmax}(\Phi_l^d) \big) , \nonumber \\
-&m_l^d(i) \text{=} \mathbb{I}(i \in Ind^d) .\nonumber
+% &\Phi_d \in \mathbb{R}^{n_p} , \nonumber \\
+  Ind^d =& \{ i_1, \cdots, i_k \} \sim \text{SRS}(\operatorname{softmax}(\Phi_l^d)) , \nonumber \\
+  m_l^d(i) =& \mathbb{I}(i \in Ind^d) ,\nonumber
 \end{align}
+with $\text{SRS}(\pi)$ denoting the process of sampling without replacement from the distribution $\pi$, and $\mathbb{I}$ the indicator function.
+This choice for the modeling of the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick of \citet{Kool19stochastic}.
+In addition, we assume that the prior distribution of $m_l^d$ is uniform over the admissible set $\Delta^{n_p}_k$, $P(m_l^d | x, \theta, d) \ \text{=} \ U(\Delta^{n_p}_k)$.
 
-The choice of the modeling of the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick \citep{Kool19stochastic}.
-In addition, we suppose the prior distribution of $m_l^d$ is uniform over the admissible set $\Delta^{n_p}_k$, $P(m_l^d | x, \theta, d) \ \text{=} \ U(\Delta^{n_p}_k)$.
-
-With our assumption over the prior distribution and our choice of the variational distribution, we express the two terms of the lower-bound \eqref{eq:lower-bound} as follows
+Given these assumptions for the prior distribution and our choice for the variational distribution, the two terms in Eq.~\eqref{eq:lower-bound} can be reformulated as:
 
@@
 Now we go back to our loss function, which is the negative of the lower-bound \eqref{eq:lower-bound}, consisting of 2 terms $-T_1$ and $-T_2$.
-The inequality \eqref{eq:entropy} shows that the upperbound of $-T_2$ is $\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d))$.
-In the training, we will maximize the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ as a regularization over the parameters $\Phi_l^d$ of the variational distribution.
+The inequality \eqref{eq:entropy} shows that an upperbound of $-T_2$ is $\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d))$.
+During training, we maximize the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ as a regularization over the parameters $\Phi_l^d$ of the variational distribution.
 
@@
 
-In Appendix \ref{appendix:a}, we will prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
-Furthermore, we will also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and its differentiation is computed via implicit differentiation theorem.
-During the training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyperparameter $\tau$ to $0$.
-The gradient of the term $T_1$ w.r.t $\Phi_l^d$ is computed via the chain rule as follows
+In Appendix \ref{appendix:a}, we prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
+Furthermore, we also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and that its differentiation can be computed using the implicit differentiation theorem.
+During training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyperparameter $\tau$ to $0$.
+The gradient of the term $T_1$ w.r.t $\Phi_l^d$ is computed via the chain rule as follows:
 \begin{equation}
@@
 \end{equation}
-The gradient $\frac{\partial T_1}{\hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\hat{m}_l^d(\tau)}{\Phi_l^d}$ is computed via implicit differentiation, that we will explain carefully in Appendix \ref{appendix:a}.
+The gradient $\frac{\partial T_1}{\hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\hat{m}_l^d(\tau)}{\Phi_l^d}$ is computed via implicit differentiation, as explained in Appendix~\ref{appendix:a}.
 
-We jointly train the Transformer parameters $\theta$ and the parameters of the latent dropping masks $\Phi_l^d$ using the following multi-task loss.
-
+We jointly train the Transformer parameters $\theta$ and the parameters of the latent dropping masks $\Phi_l^d$ using the following multi-task loss.\fyTodo{This is repeated}
 \begin{align}
@@
 
-in which $\mathcal{D}_s^d$ is distribution of the task $d$ over the input space $\Omega^d_s$; $g^d: \Omega^d_s \rightarrow \Omega^d_t$ is the mapping function, which our multi-task modern needs to learn, of the task $d$; $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$ is the ELBO loss, which we defined in Equation \ref{eq:lower-bound}.
+in which $\mathcal{D}_s^d$ is distribution of the task $d$ over the input space $\Omega^d_s$; $g^d: \Omega^d_s \rightarrow \Omega^d_t$ is the mapping function, which our multi-task modern needs to learn, of the task $d$; $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$ is the ELBO loss, defined in Equation~\ref{eq:lower-bound}.
 
-Finally, in the inference, we define the dropping mask for $l^{th}$ layer and domain $d$ as follows
-
+Finally, during inference, we define the dropping mask for $l^{th}$ layer and task $d$ as follows:
 \begin{align*}
   Ind_l^d &= Top_k(\Phi_l^d) \\
-m_l^d &= \mathbb{I}(i\in Ind_l^d)\\
+  m_l^d &= \mathbb{I}(i\in Ind_l^d),\\
 \end{align*}
+meaning that we simply pick the $k$ most likely parameter groups for the task at hand, and define the dropout mask accordingly.
 

### 2021-11-10 2022/acl/acl_latex-fy.tex
more changes in the math

@@
 Most architectures for multitask learning are based on a matching of subset of model parameters with tasks.
-While our model also needs to know the task $d \in [0 \dots n_d-1$ for each training and test instance, the structure of our Transformer networks \citep{Vaswani17attention} is emphbased on the notion of \emph{groups of parameters}.
+While our model also needs to know the task $d \in [0 \dots n_d-1]$ for each training and test instance, the structure of our Transformer networks \citep{Vaswani17attention} is emphbased on the notion of \emph{groups of parameters}.
 In each Transformer layer $l \in [1 \dots L]$, we identify $n_p$ groups of parameters.\fyTodo{In the attention heads only?
@@
 Formally, we define the \emph{group dropout mask} $m_d^l$ as a $n_p$-dimensional binary vector containing exactly $k$ ones: the group $p$ ($\in  [0,\dots,n_p\text{-}1]$) is retained for task $d$ if $m_l^d(p)\text{\small =}1$ and is dropped if $m_l^d(p)\text{\small =} 0$.
-We denote $\Delta^{n_p}_k = \{ m \in \{0,1\}^{n_p} \text{such that} \mid{} m \mid_{L_1} = k \}$, where $\mid{} m \mid_{L_1}$ is the $L_1$ norm, the set of all admissible dropping masks.
+We denote $\Delta^{n_p}_k = \{ m \in \{0,1\}^{n_p} \text{such that}  \mid{} m \mid_{L_1} = k \}$, where $\mid{} m \mid_{L_1}$ is the $L_1$ norm, the set of all admissible dropout masks.
 
@@
 % where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
+
 \subsection{Training with latent dropout masks}
 We incorporate the latent variables $m_{l}^d, l \in[0,\dots,L]$ into the log-likelihood as follows.
-For the sake of brevity, we denote $m_l^d$ the set of latent variables, \fyTodo{what do you mean here - this is all there is?} and use $x$, $y$, and $\theta$ to refer to the input, output, and parameter vectors, respectively; finally $\Phi^d_i \in \mathbb{R}^{n_p}$ denotes the set of variational parameters in the model:\fyTodo{Again this is complicated.}
+For the sake of brevity, we denote $m_l^d$ the set of latent variables, \fyTodo{what do you mean here - this is all there is?} and use $x$, $y$, and $\theta$ to refer to the input, output, and parameter vectors, respectively; finally $\Phi^d_i \in \mathbb{R}^{n_p}$ denotes the set of variational parameters in the model:\fyTodo{Again this is too complicated.
+How about the notations ?
+Why no $y$ ?}
+Use $\phi$ for all the variational parameters.
 \begin{align}
 \hspace{-3.em}
-\text{log} P(y|x,\theta,d) \text{=} &\text{log} P(y, m_l^d |x,d;\theta) \text{-} \text{log} P(m_l^d|y, x, d ;\theta) \nonumber \\
-\text{=} & \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,d;\theta) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
+\text{log} P(y|x,d;\theta) = &\text{log} P(y, m_l^d |x,d;\theta) \text{-} \text{log} P(m_l^d|y, x, d ;\theta) \nonumber \\
+= & \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,d;\theta) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
 & \quad + KL(P(m_l^d | \Phi_l^d) \parallel P(m_l^d|y, x,\theta,d)) \nonumber \\
@@
 % &\Phi_d \in \mathbb{R}^{n_p} , \nonumber \\
-  Ind^d =& \{ i_1, \cdots, i_k \} \sim \text{SRS}(\operatorname{softmax}(\Phi_l^d)) , \nonumber \\
+  Ind^d =& \{ i_1, \cdots, i_k \} \sim \text{SRS}(\operatorname{softmax}(\Phi_l^d),k) , \nonumber \\
   m_l^d(i) =& \mathbb{I}(i \in Ind^d) ,\nonumber
 \end{align}
-with $\text{SRS}(\pi)$ denoting the process of sampling without replacement from the distribution $\pi$, and $\mathbb{I}$ the indicator function.
+with $\text{SRS}(\pi,k)$ denoting the process of sampling $k$ times without replacement from the distribution $\pi$, and $\mathbb{I}$ the indicator function.
 This choice for the modeling of the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick of \citet{Kool19stochastic}.
-In addition, we assume that the prior distribution of $m_l^d$ is uniform over the admissible set $\Delta^{n_p}_k$, $P(m_l^d | x, \theta, d) \ \text{=} \ U(\Delta^{n_p}_k)$.
+In addition, we assume that the prior distribution of $m_l^d$ is uniform over the admissible set $\Delta^{n_p}_k$, $P(m_l^d | x, \theta, d) \ \text{=} \ \operatorname{Unif}(\Delta^{n_p}_k)$.
 
@@
 
-we will prove Inequality \eqref{eq:entropy} in Appendix \ref{appendix:b}.
+we prove inequality~\eqref{eq:entropy} in Appendix~\ref{appendix:b}.
 

### 2021-11-10 2022/acl/acl_latex.tex
local changes

@@
 \begin{document}
-\setlength{\parskip}{0pt}
-\setlength{\abovedisplayskip}{0pt}
-\setlength{\belowdisplayskip}{0pt}
-\setlength{\abovedisplayshortskip}{0pt}
-\setlength{\belowdisplayshortskip}{0pt}
+%\setlength{\abovedisplayskip}{3pt}
+%\setlength{\belowdisplayskip}{3pt}
+%\setlength{\parskip}{0pt}
+%\setlength{\abovedisplayskip}{0pt}
+%\setlength{\belowdisplayskip}{0pt}
+%\setlength{\abovedisplayshortskip}{0pt}
+%\setlength{\belowdisplayshortskip}{0pt}
 \maketitle
-\begin{abstract}
-Multi-domain and multilingual translation can be cast under multi-task learning in which each domain or language pair is considered a task.
+%\begin{abstract}
+%Multi-domain and multilingual translation can be cast under multi-task learning in which each domain or language pair is considered a task.
 Sharing most of the parameters of the model causes a catastrophic interference between unrelated tasks.
@@
 Furthermore, through an ablation study, we illustrate a significant correlation between the similarity between learned task-dependent dropping selections and the proximity between tasks.
+%\end{abstract}
+
+\begin{abstract}
+  Multidomain and multilingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalties of the task at hand, while smaller parts are reserved to model the peculariaties of a language or a domain.
+In adapter-based approaches, these strategies are hard coded in the network architecture.
+In this work, we propose a new method aimed at learning the parameter sharing pattern, using a latent-variable model.
+We show how this model can be trained end to end, and report experimental results which show that the learned pattern are both meaningful, and yield improved translation performance without any increase of the model size.
 \end{abstract}
@@
 \section{Introduction}
-Multi-domain translation and multilingual translation develop one model for translations in multi-domain and multiple language pairs, respectively.
+
+%Multi-domain translation and multilingual translation develop one model for translations in multi-domain and multiple language pairs, respectively.
 These paradigms are motivated by the compactness of the translation model \cite{dabre20survey,Chu18multilingual} and the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
 However, sharing parameters of the model can cause the catastrophic interference between unrelated tasks \citep{conneau20unsupervised,wang20negative}.
-
-In multi-domain setting, \citet{Pham21revisiting} pointed out a challenge in how multi-domain systems fully exploit the similarity between domains.
+%
+%In multi-domain setting, \citet{Pham21revisiting} pointed out a challenge in how multi-domain systems fully exploit the similarity between domains.
 The authors demonstrated the weakness of these systems in handling the proximity between domains through a simple setting using a pseudo domain separation.
 Domain Adapters, which are strong baselines for multi-domain translation \cite{Pham20Study}, failed this test.
-Such predefined allocations of parameters to domains do not take into account the proximity of domains in their choice.
+Such predefined allocations of parameters to domains do not take into account the proximity of domains in their choice and that motivates us to learn both the .
+%
+%The method we developed here mitigates the catastrophic interference in training a neural machine translation (NMT) model for multiple domains or multiple language pairs.
+Our main contribution are summarized below:
+%\begin{itemize}
+%\setlength{\itemsep}{1pt}
+%  \setlength{\parskip}{0pt}
+%  \setlength{\parsep}{0pt}
+%	\item We provide a rigorous mathematical modeling the problem of jointly learning routing task-dependent sub-network and the parameters of the underlying model using variational probabilistic modeling.
+%	\item Our method propose end to end training with very little extra computational cost.
+%	\item With an extensive experimental setting, we demonstrate a significant gain of our method against strong baselines.
+%	\item We illustrate a significant correlation between the similarity between learned sub-networks and the proximity of their corresponding tasks.
+%\end{itemize}
 
-The method we developed here mitigates the catastrophic interference in training a neural machine translation (NMT) model for multiple domains or multiple language pairs.
+Multidomain and multilingual machine translation aim to develop one single model to perform translations for multiple domains and multiple language pairs, respectively.\footnote{We will refer to these two situations as 'multi-task MT' (or \mtsquare{} for short), and refer to individual domains and languages as 'tasks'.} These paradigms are motivated by the compactness of the resulting translation system \citep{dabre20survey,Chu18multilingual}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
+However, having all the tasks use exactly the same model parameters can cause negative interference between unrelated tasks \citep{conneau20unsupervised,wang20negative}.
+Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using adapter layers as studied in \citet{houlsby19parameter,Bapna19simple,Pham20Study}.
+If these techniques have proven effective to achieve strong baselines, such fail to fully take advantage of the similarities that exist between domains and tasks, as documented eg.\ in \citep{Pham21revisiting}.
+This is because the partition of the parameter space between generic or task-specific sub-parts, and their allocation to each task is hard coded in the network, irrespective of the existing commonalties and differences in the data space.
+
+In this paper, we try to develop techniques that can take the similarity between tasks into account in a more effective way, by learning the network organization from the data.
+To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the parameter space, and show how training can still be performed end-to-end using a variational surrogate of the standard loss function.
+Our experiments with multilingual and multidomain confirm that this method can automatically detect similarities in the data, and yield related tasks to share more parameters.
+They also show that this method is comparable to using adapter layers in a number of empirical comparisons; contrarily to adapters, these performance are obtained without any increase of the model size.
 Our main contribution are summarized below:
@@
   \setlength{\parsep}{0pt}
-	\item We provide a rigorous mathematical modeling the problem of jointly learning routing task-dependent sub-network and the parameters of the underlying model using variational probabilistic modeling.
-	\item Our method propose end to end training with very little extra computational cost.
-	\item With an extensive experimental setting, we demonstrate a significant gain of our method against strong baselines.
-	\item We illustrate a significant correlation between the similarity between learned sub-networks and the proximity of their corresponding tasks.
+\item We provide a rigorous mathematical formulattion to the problem of jointly learning routing task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling;
+\item We introduce algorithms to train this model end to end with very little extra computational cost.
+\item We report, using an extensive experimental setting, significant gains of this method with respect to strong baselines.
+\item We study how this method can actually exploit the similarities between tasks to learn consistent sub-networks.
 \end{itemize}
 
-\section{Multi-task group dropout}
+\section{Latent multi-task group dropout}
 We define a group dropout for a domain $d$ as follows

### 2021-11-10 2022/acl/acl_latex-fy.tex
now revised the full section 2

--- 2021-11-09 2022/acl/acl_latex-fy.tex
@@
 \section{Introduction}
-Multidomain and multilingual machine translation aim to develop one single model to perform translations for multiple domains and multiple language pairs, respectively.\footnote{We will refer to these two situations as 'multi-task MT' (or \mtsquare{} for short), and refer to individual domains and languages as 'tasks'.} These paradigms are motivated by the compactness of the resulting translation system \citep{dabre20survey,Chu18multilingual}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
+Multidomain and multilingual machine translation aim to develop one single model to perform translations for multiple domains and multiple language pairs, respectively.\footnote{We will refer to these two situations as 'multi-task MT' (or \mtsquare{} for short), and refer to individual domains and languages as 'tasks'.} These paradigms are motivated by the compactness of the resulting translation system \citep{Chu18multilingual,dabre20survey}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
 However, having all the tasks use exactly the same model parameters can cause negative interference between unrelated tasks \citep{conneau20unsupervised,wang20negative}.
-Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using adapter layers as studied in \citet{houlsby19parameter,Bapna19simple,Pham20Study}.
-If these techniques have proven effective to achieve strong baselines, such fail to fully take advantage of the similarities that exist between domains and tasks, as documented eg.\ in \citep{Pham21revisiting}.
+Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using adapter layers as studied in \citep{houlsby19parameter,Bapna19simple,Pham20Study}.
+If these techniques have proven effective for building strong baselines, such fail to fully take advantage of the similarities that exist between domains and tasks, as documented eg.\ in \citep{Pham21revisiting}.
 This is because the partition of the parameter space between generic or task-specific sub-parts, and their allocation to each task is hard coded in the network, irrespective of the existing commonalties and differences in the data space.
 
-In this paper, we try to develop techniques that can take the similarity between tasks into account in a more effective way, by learning the network organization from the data.
+In this paper, we develop techniques that can take the similarity between tasks into account in a more effective way, by learning the network organization from the data.
 To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the parameter space, and show how training can still be performed end-to-end using a variational surrogate of the standard loss function.
-Our experiments with multilingual and multidomain confirm that this method can automatically detect similarities in the data, and yield related tasks to share more parameters.
+Our experiments with multilingual and multidomain machine translation confirm that this method can automatically detect similarities in the data, and yield related tasks to share more parameters.
 They also show that this method is comparable to using adapter layers in a number of empirical comparisons; contrarily to adapters, these performance are obtained without any increase of the model size.
-Our main contribution are summarized below:
+Our main contributions are summarized below:
 \begin{itemize}
@@
   \setlength{\parsep}{0pt}
-\item We provide a rigorous mathematical formulattion to the problem of jointly learning routing task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling;
+\item We provide a sound mathematical formulation to the problem of jointly learning routing task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling;
 \item We introduce algorithms to train this model end to end with very little extra computational cost.
-\item We report, using an extensive experimental setting, significant gains of this method with respect to strong baselines.
-\item We study how this method can actually exploit the similarities between tasks to learn consistent sub-networks.
+\item We report, using an extensive experimental setting, significant gains of this method with respect to strong baselines;
+\item We study how this method can actually exploit the similarities between tasks to learn interpretable sub-networks.
 \end{itemize}
@@
 Most architectures for multitask learning are based on a matching of subset of model parameters with tasks.
-While our model also needs to know the task $d \in [0 \dots n_d-1$ for each training and test instance, the structure of our Transformer networks \citep{Vaswani17attention} is emphbased on the notion of \emph{groups of parameters}.
-In each Transformer layer $l \in [1 \dots L]$, we identify $n_p$ groups of parameters.\fyTodo{In the attention heads only?
-In the FF?} The assignment of tasks to groups will be learned from the data, under the constraint that each task has only access to exactly $k$ groups of \emph{active parameters}, while the all the other parameter values are deemed to be \emph{dropped} (see Figure~\ref{fig:group_dropout}).
+While our model also needs to know the task $d \in [0 \dots n_d-1]$ for each training and test instance, the structure of our Transformer networks \citep{Vaswani17attention} is based on the notion of \emph{groups of nodes in the computation graph}.
+At the input of each Transformer layer $l \in [1 \dots L]$, we partition all input state vectors into $n_p$ groups of nodes, and zero-out a task-dependent subset of these groups.\fyDone{In the attention heads only?
+In the FF?} The assignment of tasks to groups will be learned from the data, under the constraint that each task only activates exactly $k$ groups of \emph{active nodes}, while the all the other values are nullified akin to a dropout process (see Figure~\ref{fig:group_dropout}).
 \begin{figure}
@@
 \caption{Latent group dropout.
-The set of nodes\fyTodo{parameters ?} in each layer is divided into equal-sized groups.
-For a task, we only keep a fixed number of groups.
+The set of nodes\fyDone{parameters ?} in each layer is divided into equal-sized groups.
+For each task, we only keep a fixed number of groups and nullify all the other nodes.
 Our method learns which groups will be kept / dropped for each task.}
@@
 \end{figure}
-Formally, we define the \emph{group dropout mask} $m_d^l$ as a $n_p$-dimensional binary vector containing exactly $k$ ones: the group $p$ ($\in  [0,\dots,n_p\text{-}1]$) is retained for task $d$ if $m_l^d(p)\text{\small =}1$ and is dropped if $m_l^d(p)\text{\small =} 0$.
-We denote $\Delta^{n_p}_k = \{ m \in \{0,1\}^{n_p} \text{such that} \mid{} m \mid_{L_1} = k \}$, where $\mid{} m \mid_{L_1}$ is the $L_1$ norm, the set of all admissible dropping masks.
+Formally, a \emph{group dropout mask} $m_l^d$ is a $n_p$-dimensional binary vector containing exactly $k$ ones: the group $p$ ($\in  [0,\dots,n_p\text{-}1]$) is retained for task $d$ if $m_l^d(p)\text{\small =}1$ and is dropped if $m_l^d(p)\text{\small =} 0$.
+We denote $\Delta^{n_p}_k = \{ m \in \{0,1\}^{n_p} \text{ such that }  \mid{} m \mid_{L_1} = k \}$ the set of all admissible dropout masks, with $\mid{} m \mid_{L_1}$ the $L_1$ norm of vector $m$.
 
-Given $m_l^d$, the parameter mask $r_l^d$ for task $d$ in layer $l$ is then derived as follows:
+Given $m_l^d$, the mask $r_l^d$ for task $d$ in layer $l$ is then derived as follows:
 \begin{align*}
@@
 \end{align*}
-where $d_k$ is the dimension of the hidden state.\fyTodo{Check this} The propagation of information within the network then depends on the current task value and can be formalized as:
+where $d_k$ is the dimension of the hidden state.\fyDone{Check this} The propagation of information within the network then depends on the current task value and can be formalized as:
 \begin{align*}
-  r_l^d &\in \{ 0,1 \}^{d_k}, \\
-  \tilde{h}^l &= h^l \odot r_l^d ,\\
-  h^{l+1} &= \text{LAYER}^{l+1}(h^l) ,\\
-  l & \in [0,\cdots, L-1] ,
+  \forall l \in [0,\cdots, L-1]: \tilde{h}^l &= h^l \odot r_l^d ,\\
+  h^{l+1} &= \operatorname{LAYER}^{l+1}(h^l) ,\\
 \end{align*}
-where $\text{LAYER}^l()$ represents all the computations in Transformer layer $l$.
+where $\operatorname{LAYER}^l()$ represents all the computations in Transformer layer $l$.
 % The dropping mask $r_l^d$ is computed as follows
 % where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
+
 \subsection{Training with latent dropout masks}
-We incorporate the latent variables  $m_{l}^d, l \in[0,\dots,L]$ into the log-likelihood as follows.
-For the sake of brevity, we denote $m_l^d$ the set of latent variables, \fyTodo{what do you mean here - this is all there is?} and use $x$, $y$, and $\theta$ to refer to the input, output, and parameter vectors, respectively; finally $\Phi^d_i \in \mathbb{R}^{n_p}$ denotes the set of variational parameters in the model:\fyTodo{Again this is complicated.}
+Assuming standard notation for our translation model  $P(y|x,d;\theta)$ where $x$, $y$ and $\theta$ respectively refer to the input, output, and parameter vector, the latent variables $m_{l}^d, l \in[0,\dots,L], d \in [0, \dots, n_d-1]$ are introduced as follows.
+We also assume a prior distribution for $m_{l}^d$ chosen to be the uniform distribution over $\Delta^{n_p}_k$: $P(m_l^d | x, d; \theta) = \operatorname{Unif}(\Delta^{n_p}_k)$; variables for each layer are independent and collectively refered to as $m^d$.
+For any (variational) distribution $Q(m^1 \dots m^{n_d}; \Phi)$ with parameters $\Phi$, it is well-known that the log-likelihood is lower-bounded by the so-called $\operatorname{ELBO}$ function, made of a summation of two terms: the \emph{distortion} $D$ and the \emph{rate} $R$ defined as follows:\footnote{We drop the dependency in $x$ to avoid notational clutter.}
 \begin{align}
-\hspace{-3.em}
-\text{log} P(y|x,\theta,d) \text{=} &\text{log} P(y, m_l^d |x,d;\theta) \text{-} \text{log} P(m_l^d|y, x, d ;\theta) \nonumber \\
-\text{=} & \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,d;\theta) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
-& \quad + KL(P(m_l^d | \Phi_l^d) \parallel P(m_l^d|y, x,\theta,d)) \nonumber \\
-\geqslant & \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
- = & \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y | x,\theta,m_l^d, d) \big] \label{eq:lower-bound} \\
-& \quad - KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\
- = & \mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y). \nonumber
+  \log P(y|d;\theta) \ge & \operatorname{ELBO}(y,d; \theta, \Phi) \nonumber \\
+  \operatorname{ELBO}(y,d; \theta, \Phi) =&  D(y,d; \theta, \Phi) - R(y,d; \theta, \Phi) \label{eq:lower-bound}\\
+  D(y,d; \theta, \phi) = & \mathbb{E}_{m^d \sim Q(m^d |d,\Phi)} \log P(y | m^d, d; \theta) \nonumber\\
+  R(y,d; \theta, \phi) = & \operatorname{KL}(Q(m^d | d, \Phi)||P(m^d | d; \theta)), \nonumber
 \end{align}
-$\mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y)$ is a lowerbound of the log-likelihood, known as the evidence lower bound (ELBO) in variational statistics \citep{Diederick14auto}, of our log-likelihood.
-Our training loss will be $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$.
+where $\operatorname{KL}$ is the Kullback-Leibler divergence.
+We use $-\operatorname{ELBO}(x, y\theta,\Phi_l)$ as our surrogate training objective, as a tractable approximation of the untractable likelihood, and try to minimize this function in $\theta$ and $\Phi$.
 
-The variational distribution of $m_l^d$, $P(m_l^d|\Phi_l^d)$, is defined as follows:
+The variational distribution $Q$ of $m^d$ is defined on layerwise basis, each layer $l$ involving only a subset $\Phi_l^d$ of parameters:
+\begin{align}
+  \operatorname{Ind}^d = \{ i_1, \cdots, i_k \} & \sim \text{SRS}(\operatorname{softmax}(\Phi_l^d), k) \nonumber \\
+  m_l^d(i) & =  \text{$\mathbb{I}$}(i \in \operatorname{Ind}^d), \nonumber
+\end{align}
+where $\text{SRS}(\pi,k)$ denotes the process of sampling $k$ times without replacement from the distribution $\pi$, and $\mathbb{I}$ is the indicator function.
+This particular choice the modeling of the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick of \citet{Kool19stochastic} that we use below.
+Given our assumptions for the prior distribution and our choice for the variational distribution, the two terms in Eq.~\eqref{eq:lower-bound} can be computed as:
 \begin{align}
 \hspace{-2.em}
-% &\Phi_d \in \mathbb{R}^{n_p} , \nonumber \\
-  Ind^d =& \{ i_1, \cdots, i_k \} \sim \text{SRS}(\operatorname{softmax}(\Phi_l^d)) , \nonumber \\
-  m_l^d(i) =& \mathbb{I}(i \in Ind^d) ,\nonumber
+D &= \mathbb{E}_{m^d \sim Q(m^d | d | \Phi)} \text{log} P(y | m^d, d, \theta) \nonumber \\
+&= \mathbb{E}_{g^d \sim^{\text{i.i.d}} G(0,1)} \big[ \log P(y |\tilde{m}^d,d, \theta,) \big] \nonumber
 \end{align}
-with $\text{SRS}(\pi)$ denoting the process of sampling without replacement from the distribution $\pi$, and $\mathbb{I}$ the indicator function.
-This choice for the modeling of the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick of \citet{Kool19stochastic}.
-In addition, we assume that the prior distribution of $m_l^d$ is uniform over the admissible set $\Delta^{n_p}_k$, $P(m_l^d | x, \theta, d) \ \text{=} \ U(\Delta^{n_p}_k)$.
-
-Given these assumptions for the prior distribution and our choice for the variational distribution, the two terms in Eq.~\eqref{eq:lower-bound} can be reformulated as:
-
+where the generation process $G(0,1)$ is a product of independent Gumbel distributions, yielding:
 \begin{align}
   \hspace{-2.em}
-T_1 &= \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y | x,\theta,m_l^d, d)  \big] \nonumber \\
-&= \displaystyle{\mathop{\mathbb{E}}_{g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} Gumbel(0,1) \ }} \big[ \text{log} P(y | x,\theta,\tilde{m}_l^d,d) \big] \nonumber
+  \forall d,  g^d & =  [g_0^d, \dots, g_L^d],  \text{ with } g_l^d \in \mathbb{R}^{n_p} \nonumber \\
+  \forall p,  g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1) \nonumber \\
+  \{ i_1, \cdots, i_k \} &= \operatorname{Top-k} \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \label{eq:top-k} \\ 
+       & \quad \quad g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \} \nonumber \\
+\tilde{m}_l^d(p) &= \mathbb{I}(p \in Ind^d). \nonumber 
 \end{align}
-where
+For the second term, the derivation is the following:\fyTodo{Check this: we have a sum over layers ?Also: R or -R ?} 
 \begin{align}
 \hspace{-2.em}
-g_l^d & \in \mathbb{R}^{n_p} \nonumber \\
-g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} Gumbel(0,1) \nonumber \\
-& \forall p \in \{ 0,\cdots,n_p\text{-}1 \} \nonumber \\
-i_1, \cdots, i_k &= Top\_k \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \label{eq:top-k} \\ 
-&g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \} \nonumber \\
-\tilde{m}_l^d(p) &= \mathbb{I}(p \in Ind^d). \nonumber 
+  R    &= \operatorname{KL}(Q(m^d | d, \Phi)||P(m^d | d; \theta)), \nonumber \\ 
+	&= \mathbb{H} \big[ Q(m_l^d | d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\ 
+	&= \mathbb{H} \big[ Q(i_1,\cdots,i_k | d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\
+	&  \geqslant \mathbb{H} \big[ Q(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \label{eq:entropy}
 \end{align}
-and
-\begin{align}
-\hspace{-2.em}
-T_2 &= - KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\ 
-	&= \mathbb{H} \big[ P(m_l^d | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\ 
-	&= \mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\
-	&\geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \label{eq:entropy}
-\end{align}
+We prove inequality~\eqref{eq:entropy} in Appendix~\ref{appendix:b}.
 
-we will prove Inequality \eqref{eq:entropy} in Appendix \ref{appendix:b}.
+Inequality~\eqref{eq:entropy} shows that an upperbound of $R$ is $\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d))$.
+During training, we maximize the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$\fyTodo{Check sum} as a regularization over the parameters $\Phi_l^d$ of the variational distribution.
 
-Now we go back to our loss function, which is the negative of the lower-bound \eqref{eq:lower-bound}, consisting of 2 terms $-T_1$ and $-T_2$.
-The inequality \eqref{eq:entropy} shows that an upperbound of $-T_2$ is $\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d))$.
-During training, we maximize the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ as a regularization over the parameters $\Phi_l^d$ of the variational distribution.
-
-Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi_l^d$ into log-likelihood function and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}.
-However, the operator Top-K, which defines $\tilde{m}_l^d$ in Equation \eqref{eq:top-k}, is not differentiable.
-Therefore, we approximate this function by the Soft-Top-K function defined as follows
-
+Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi$ into log-likelihood function\fyTodo{into the objective function?} and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}.
+However, the operator $\operatorname{Top-k}$, which serves to define $\tilde{m}_l^d$ in Equation~\eqref{eq:top-k}, is not differentiable.
+Therefore, we approximate this function by the $\operatorname{Soft-Top-K}$ function defined as follows:
 \begin{align}
@@
 During training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyperparameter $\tau$ to $0$.
-The gradient of the term $T_1$ w.r.t $\Phi_l^d$ is computed via the chain rule as follows:
+The gradient of the term $D$ w.r.t $\Phi_l^d$ is computed via the chain rule as follows:
 \begin{equation}
-\frac{\partial T_1}{\Phi_l^d} = \frac{\partial T_1}{\hat{m}_l^d(\tau)} \times \frac{\hat{m}_l^d(\tau)}{\Phi_l^d}
+\frac{\partial D}{\partial \Phi_l^d} = \frac{\partial T_1}{\partial \hat{m}_l^d(\tau)} \times \frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}
 \end{equation}
-The gradient $\frac{\partial T_1}{\hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\hat{m}_l^d(\tau)}{\Phi_l^d}$ is computed via implicit differentiation, as explained in Appendix~\ref{appendix:a}.
+The gradient $\frac{\partial T_1}{\partial \hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$ is computed via implicit differentiation, as explained in Appendix~\ref{appendix:a}.
 
-We jointly train the Transformer parameters $\theta$ and the parameters of the latent dropping masks $\Phi_l^d$ using the following multi-task loss.\fyTodo{This is repeated}
+We jointly train the Transformer parameters $\theta$ and the parameters of the variational distribution $\Phi$ using the following multi-task loss.\fyDone{This is repeated}
 \begin{align}
 \hspace{-0.em}
-\mathcal{L}(\theta,\Phi_l^1,\cdots,\Phi_l^k) = \displaystyle{\mathop{\sum}_{d=1}^{k}} \displaystyle{\mathop{\mathbb{E}}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)}} \big[-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)\big]
+\ell(\theta,\Phi_l^1,\cdots,\Phi_l^k) = \displaystyle{\mathop{\sum}_{d=1}^{k}} \displaystyle{\mathop{\mathbb{E}}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)}} \big[-\operatorname{ELBO}(\theta,\Phi_l^d,x,y)\big]
 \end{align}
 
-in which $\mathcal{D}_s^d$ is distribution of the task $d$ over the input space $\Omega^d_s$; $g^d: \Omega^d_s \rightarrow \Omega^d_t$ is the mapping function, which our multi-task modern needs to learn, of the task $d$; $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$ is the ELBO loss, defined in Equation~\ref{eq:lower-bound}.
+in which $\mathcal{D}_s^d$ is distribution of task $d$ over the input space $\Omega^d_s$; $g^d: \Omega^d_s \rightarrow \Omega^d_t$\fyTodo{allready defined} is the mapping function for the task $d$, which our multi-task model needs to learn; $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$ is the ELBO loss, defined in Equation~\ref{eq:lower-bound}.
 
-Finally, during inference, we define the dropping mask for $l^{th}$ layer and task $d$ as follows:
+Finally, during inference, we define the dropout mask for layer $l$ and task $d$ as follows:
 \begin{align*}
-  Ind_l^d &= Top_k(\Phi_l^d) \\
+  Ind_l^d &= \operatorname{Top-k}(\Phi_l^d) \\
   m_l^d &= \mathbb{I}(i\in Ind_l^d),\\

### 2021-11-12 2022/acl/acl_latex-fy.tex
math is done

@@
   \forall l \in [0,\cdots, L-1]: \tilde{h}^l &= h^l \odot r_l^d ,\\
-  h^{l+1} &= \operatorname{LAYER}^{l+1}(h^l) ,\\
+  h^{l+1} &= \operatorname{LAYER}^{l+1}(\tilde{h}^l) ,
 \end{align*}
@@
 We also assume a prior distribution for $m_{l}^d$ chosen to be the uniform distribution over $\Delta^{n_p}_k$: $P(m_l^d | x, d; \theta) = \operatorname{Unif}(\Delta^{n_p}_k)$; variables for each layer are independent and collectively refered to as $m^d$.
-For any (variational) distribution $Q(m^1 \dots m^{n_d}; \Phi)$ with parameters $\Phi$, it is well-known that the log-likelihood is lower-bounded by the so-called $\operatorname{ELBO}$ function, made of a summation of two terms: the \emph{distortion} $D$ and the \emph{rate} $R$ defined as follows:\footnote{We drop the dependency in $x$ to avoid notational clutter.}
+For any (variational) distribution $Q(m^1 \dots m^{n_d}; \Phi)$ with parameters $\Phi=\{\phi_l^1, ..., \phi_L^{n_d}\}$, it is well-known that the log-likelihood is lower-bounded by the so-called $\operatorname{ELBO}$ function, made of a summation of two terms: the \emph{distortion} $D$ and the \emph{rate} $R$ defined as follows:%\footnote{We drop the dependency in $x$ to avoid notational clutter.}
 \begin{align}
-  \log P(y|d;\theta) \ge & \operatorname{ELBO}(y,d; \theta, \Phi) \nonumber \\
-  \operatorname{ELBO}(y,d; \theta, \Phi) =&  D(y,d; \theta, \Phi) - R(y,d; \theta, \Phi) \label{eq:lower-bound}\\
-  D(y,d; \theta, \phi) = & \mathbb{E}_{m^d \sim Q(m^d |d,\Phi)} \log P(y | m^d, d; \theta) \nonumber\\
-  R(y,d; \theta, \phi) = & \operatorname{KL}(Q(m^d | d, \Phi)||P(m^d | d; \theta)), \nonumber
+  \log P(y|x,d;\theta) \ge & \operatorname{ELBO}(x,y,d; \theta, \Phi) \nonumber \\
+  \operatorname{ELBO}(x,y,d; \theta, \Phi) =&  D(x,y,d; \theta, \Phi) - R(x,y,d; \theta, \Phi) \label{eq:lower-bound}\\
+  D(x,y,d; \theta, \phi) = & \mathbb{E}_{m^d \sim Q(m^d |d,\Phi)} \log P(y | m^d, x, d; \theta) \nonumber\\
+  R(x,y,d; \theta, \phi) = & \operatorname{KL}(Q(m^d | x,d, \Phi)||P(m^d | x,d; \theta)), \nonumber
 \end{align}
 where $\operatorname{KL}$ is the Kullback-Leibler divergence.
-We use $-\operatorname{ELBO}(x, y\theta,\Phi_l)$ as our surrogate training objective, as a tractable approximation of the untractable likelihood, and try to minimize this function in $\theta$ and $\Phi$.
+We use $-\operatorname{ELBO}(x,y,d; \theta,\Phi)$ as our surrogate training objective, as a tractable approximation of the untractable likelihood, and try to minimize this function in $\theta$ and $\Phi$.
 
-The variational distribution $Q$ of $m^d$ is defined on layerwise basis, each layer $l$ involving only a subset $\Phi_l^d$ of parameters:
+The variational distribution $Q$ of $m^d$ is defined independently on a layerwise basis, each layer $l$ involving only a subset $\Phi_l^d$ of parameters:
 \begin{align}
@@
 \hspace{-2.em}
-D &= \mathbb{E}_{m^d \sim Q(m^d | d | \Phi)} \text{log} P(y | m^d, d, \theta) \nonumber \\
-&= \mathbb{E}_{g^d \sim^{\text{i.i.d}} G(0,1)} \big[ \log P(y |\tilde{m}^d,d, \theta,) \big] \nonumber
+D(\dots) &= \mathbb{E}_{m^d \sim Q(m^d | d; \Phi)} \text{log} P(y | m^d, x, d, \theta) \nonumber \\
+&= \mathbb{E}_{g^d \sim^{\text{i.i.d}} G(0,1)} \big[ \log P(y |\tilde{m}^d, x, d, \theta,) \big] \nonumber
 \end{align}
@@
   \hspace{-2.em}
-  \forall d,  g^d & =  [g_0^d, \dots, g_L^d],  \text{ with } g_l^d \in \mathbb{R}^{n_p} \nonumber \\
+  \forall d,  g^d & =  [g_1^d, \dots, g_L^d],  \text{ with } g_l^d \in \mathbb{R}^{n_p} \nonumber \\
   \forall p,  g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1) \nonumber \\
-  \{ i_1, \cdots, i_k \} &= \operatorname{Top-k} \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \label{eq:top-k} \\ 
-       & \quad \quad g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \} \nonumber \\
+  \operatorname{Ind}^d = \{ i_1, \cdots, i_k \} &= \operatorname{Top-k} \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \nonumber \\ 
+       & \quad \quad g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \}\label{eq:top-k} \\
 \tilde{m}_l^d(p) &= \mathbb{I}(p \in Ind^d). \nonumber 
 \end{align}
-For the second term, the derivation is the following:\fyTodo{Check this: we have a sum over layers ?Also: R or -R ?} 
+For the second term, the derivation is the following:\fyDone{Check this: we have a sum over layers ?Also: R or -R ?} 
 \begin{align}
@@
   R    &= \operatorname{KL}(Q(m^d | d, \Phi)||P(m^d | d; \theta)), \nonumber \\ 
-	&= \mathbb{H} \big[ Q(m_l^d | d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\ 
-	&= \mathbb{H} \big[ Q(i_1,\cdots,i_k | d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\
-	&  \geqslant \mathbb{H} \big[ Q(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \label{eq:entropy}
+	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(m_l^d | d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big) \nonumber \\ 
+	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1,\cdots,i_k | d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big)\nonumber \\
+	&\leqslant  -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \big)\label{eq:entropy}
 \end{align}
@@
 
-Inequality~\eqref{eq:entropy} shows that an upperbound of $R$ is $\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d))$.
-During training, we maximize the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$\fyTodo{Check sum} as a regularization over the parameters $\Phi_l^d$ of the variational distribution.
+Inequality~\eqref{eq:entropy} shows that an upperbound of $R$ is $\sum_{l=1}(\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d)))$.
+During training, we maximize a sum over layers of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$\fyDone{Check sum} which performs a regularization over the parameters $\Phi^d$ of the variational distribution.
 
-Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi$ into log-likelihood function\fyTodo{into the objective function?} and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}.
+Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi$ into the objective function and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}.
 However, the operator $\operatorname{Top-k}$, which serves to define $\tilde{m}_l^d$ in Equation~\eqref{eq:top-k}, is not differentiable.
@@
 \hat{m}_l^d(\tau) &= \displaystyle{\mathop{argmin}_{\substack{
-       0 \leqslant m_i \leqslant 1 \ \forall 0 \leqslant i \leqslant n_d\text{-}1 \label{eq:soft-top-k}\\
+                    0 \leqslant m_i \leqslant 1 \\
+  \forall 0 \leqslant i \leqslant n_d\text{-}1 \label{eq:soft-top-k}\\
         1^{T}.m = k
-      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \\
+      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \\ \nonumber
 & \text{in which} \nonumber \\
@@
 In Appendix \ref{appendix:a}, we prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
-Furthermore, we also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and that its differentiation can be computed using the implicit differentiation theorem.
+Furthermore, we also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and that its gradients can be computed using the implicit differentiation theorem.
 During training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyperparameter $\tau$ to $0$.
-The gradient of the term $D$ w.r.t $\Phi_l^d$ is computed via the chain rule as follows:
+The gradient of $D$ w.r.t $\Phi_l^d$ is computed using the chain rule as follows:
 \begin{equation}
-\frac{\partial D}{\partial \Phi_l^d} = \frac{\partial T_1}{\partial \hat{m}_l^d(\tau)} \times \frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}
+\frac{\partial D}{\partial \Phi_l^d} = \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}
 \end{equation}
-The gradient $\frac{\partial T_1}{\partial \hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$ is computed via implicit differentiation, as explained in Appendix~\ref{appendix:a}.
+The gradient $\frac{\partial D}{\partial \hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$ is computed via implicit differentiation, as explained in Appendix~\ref{appendix:a}.
 
@@
 \hspace{-0.em}
-\ell(\theta,\Phi_l^1,\cdots,\Phi_l^k) = \displaystyle{\mathop{\sum}_{d=1}^{k}} \displaystyle{\mathop{\mathbb{E}}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)}} \big[-\operatorname{ELBO}(\theta,\Phi_l^d,x,y)\big]
+\ell(\theta,\Phi) = \sum_{d=1}^{n_d} \mathbb{E}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)} \big[-\operatorname{ELBO}(x,y,d; \theta,\Phi)\big]
 \end{align}
 
-in which $\mathcal{D}_s^d$ is distribution of task $d$ over the input space $\Omega^d_s$; $g^d: \Omega^d_s \rightarrow \Omega^d_t$\fyTodo{allready defined} is the mapping function for the task $d$, which our multi-task model needs to learn; $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$ is the ELBO loss, defined in Equation~\ref{eq:lower-bound}.
+in which $\mathcal{D}_s^d$ is distribution of task $d$ over the input space $\Omega^d_s$; $\operatorname{MT}^d: \Omega^d_s \rightarrow \Omega^d_t$\fyDone{already defined} is the translation function for task $d$, which our multi-task model needs to learn; $-\mathcal{L}_{\text{ELBO}}(x,y, d, \theta,\Phi)$ is the ELBO loss, defined in Equation~\ref{eq:lower-bound}.
 
@@
   Ind_l^d &= \operatorname{Top-k}(\Phi_l^d) \\
-  m_l^d &= \mathbb{I}(i\in Ind_l^d),\\
+  m_l^d &= \mathbb{I}(i\in Ind_l^d)
 \end{align*}

### 2021-11-12 2022/acl/acl_latex.tex
figures

@@
 
+%subfigure
+\usepackage{caption}
+\usepackage{subcaption}
+
 \title{Multi-domain, multilingual translation with latent multi-task group dropout}
@@
 
-Multidomain and multilingual machine translation aim to develop one single model to perform translations for multiple domains and multiple language pairs, respectively.\footnote{We will refer to these two situations as 'multi-task MT' (or \mtsquare{} for short), and refer to individual domains and languages as 'tasks'.} These paradigms are motivated by the compactness of the resulting translation system \citep{dabre20survey,Chu18multilingual}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
+Multidomain and multilingual machine translation aim to develop one single model to perform translations for multiple domains and multiple language pairs, respectively.\footnote{We will refer to these two situations as 'multi-task MT' (or \mtsquare{} for short), and refer to individual domains and languages as 'tasks'.} These paradigms are motivated by the compactness of the resulting translation system \citep{Chu18multilingual,dabre20survey}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
 However, having all the tasks use exactly the same model parameters can cause negative interference between unrelated tasks \citep{conneau20unsupervised,wang20negative}.
-Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using adapter layers as studied in \citet{houlsby19parameter,Bapna19simple,Pham20Study}.
-If these techniques have proven effective to achieve strong baselines, such fail to fully take advantage of the similarities that exist between domains and tasks, as documented eg.\ in \citep{Pham21revisiting}.
-This is because the partition of the parameter space between generic or task-specific sub-parts, and their allocation to each task is hard coded in the network, irrespective of the existing commonalties and differences in the data space.
+Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using adapter layers as studied in \citep{houlsby19parameter,Bapna19simple,Pham20Study}.
+If these techniques have proven effective for building strong baselines, such fail to fully take advantage of the similarities that exist between domains and tasks, as documented eg.\ in \citet{Pham21revisiting}.
+This is because the partition of the parameter space between generic or task-specific sub-parts, and their allocation to each task is hard coded in the network, irrespective of the existing commonalities and differences in the data space.
 
-In this paper, we try to develop techniques that can take the similarity between tasks into account in a more effective way, by learning the network organization from the data.
+In this paper, we develop techniques that can take the similarity between tasks into account in a more effective way, by learning the network organization from the data.
 To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the parameter space, and show how training can still be performed end-to-end using a variational surrogate of the standard loss function.
-Our experiments with multilingual and multidomain confirm that this method can automatically detect similarities in the data, and yield related tasks to share more parameters.
+Our experiments with multilingual and multidomain machine translation confirm that this method can automatically detect similarities in the data, and yield related tasks to share more parameters.
 They also show that this method is comparable to using adapter layers in a number of empirical comparisons; contrarily to adapters, these performance are obtained without any increase of the model size.
-Our main contribution are summarized below:
+Our main contributions are summarized below:
 \begin{itemize}
@@
   \setlength{\parsep}{0pt}
-\item We provide a rigorous mathematical formulattion to the problem of jointly learning routing task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling;
+\item We provide a rigorous mathematical formulation to the problem of jointly learning routing task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling;
 \item We introduce algorithms to train this model end to end with very little extra computational cost.
 \item We report, using an extensive experimental setting, significant gains of this method with respect to strong baselines.
-\item We study how this method can actually exploit the similarities between tasks to learn consistent sub-networks.
+\item We study how this method can actually exploit the similarities between tasks to learn interpretable sub-networks.
 \end{itemize}
 
-\section{Latent multi-task group dropout}
-We define a group dropout for a domain $d$ as follows
-\begin{align*}
-\tilde{h}^l &= h^l \odot r_l^d ,\\
-r_l^d &\in \{ 0,1 \}^{d_k}, \\
-h^{l+1} &= \text{LAYER}^{l+1}(h^l) ,\\
-l & \in [0,\cdots,L] ,\\
-\end{align*}
-where $L$ is number of Transformer layers of our NMT model, the binary vector $r_l^d$ is a dropping mask. We illustrate the group dropout in Figure \ref{fig:group_dropout}
-\begin{figure}
-\includegraphics[width=0.4\textwidth]{group_dropout}
-\caption{Latent group dropout.
+%\section{Latent multi-task group dropout}
+%We define a group dropout for a domain $d$ as follows
+%\begin{align*}
+%\tilde{h}^l &= h^l \odot r_l^d ,\\
+%r_l^d &\in \{ 0,1 \}^{d_k}, \\
+%h^{l+1} &= \text{LAYER}^{l+1}(h^l) ,\\
+%l & \in [0,\cdots,L] ,\\
+%\end{align*}
+%where $L$ is number of Transformer layers of our NMT model, the binary vector $r_l^d$ is a dropping mask. We illustrate the group dropout in Figure \ref{fig:group_dropout}
+%\begin{figure}
+%\includegraphics[width=0.4\textwidth]{group_dropout}
+%\caption{Latent group dropout.
 The set of nodes in one layer is divided into equal-sized groups.
@@
 Our method learns which groups will be dropped for a domain.}
+%\label{fig:group_dropout}
+%\end{figure}
+% The dropping mask $r_l^d$ of $l^{th}$ layer and domain $d$ is defined as follows
+%\begin{align*}
+%r_l^d(i) &= \begin{cases}
+%      1, & \text{if}\ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p} \\
+%      & \text{AND} \  m_l^d(p) \text{\small ==} 1 \\
+%      0, & \text{otherwise},
+%    \end{cases} \\
+%p & \in  \{0,\dots,n_p\text{-}1 \}, \\
+%d & \in  \{0,\dots,n_d\text{-}1 \}, \\
+%i & \in  \{0,\dots,d_k\text{-}1 \}, \\
+%m_l^d & \in  \{0,1\}^{n_p} \ \forall d,l \, \\
+%\mid m_l^d \mid_{L_1} & =  k , \\
+%\Delta^{n_p}_k & =  \{ m \in \{0,1\}^{n_p} | \mid m \mid_{L_1} = k \},
+%\end{align*}
+%where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
+The group $p$ is retained for domain $d$ if $m_l^d(p)\text{\small ==}1$ and is dropped if $m_l^d(p)\text{\small ==}0$.
+We also denote $\Delta^{n_p}_k$ the set of all admissible dropping masks.
+%
+%We incorporate the latent variables $m_{l\in[0,\cdots,L]}^d$ into the log-likelihood of our model as below.
+For the sake of brevity, we denote $m_l^d$ the set of latent variables $m_{l\in[0,\cdots,L]}^d$
+%\begin{align}
+%\hspace{-3.em}
+%\text{log} P(y|x,\theta,d) & \text{=} \text{log} P(y, m_l^d |x,\theta,d) \text{-} \text{log} P(m_l^d|y, x,\theta,d) \nonumber \\
+%& \text{=} \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
+%& \text{+} KL(P(m_l^d | \Phi_l^d) \parallel P(m_l^d|y, x,\theta,d)) \nonumber \\
+%& \geqslant \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
+%& \text{=} \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y | x,\theta,m_l^d, d) \big] \label{eq:lower-bound} \\
+%& \text{-} KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\
+%& \text{=} \mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y). \nonumber
+%\end{align}
+%$\mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y)$ is the lowerbound, which is called the evidence lower bound (ELBO) in variational statistic \citep{Diederick14auto}, of our log-likelihood.
+Our training loss will be $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$.
+%
+%The variational distribution of $m_l^d$, $P(m_l^d|\Phi_l^d)$, is defined as follows
+%
+%\begin{align}
+%\hspace{-2.em}
+%&\Phi_d \in \mathbb{R}^{n_p} , \nonumber \\
+%&Ind^d \text{=}  \{ i_1, \cdots, i_k \} \nonumber \\
+%&\sim \text{\small Sampling\_without\_replacement} \big(\text{softmax}(\Phi_l^d) \big) , \nonumber \\
+%&m_l^d(i) \text{=} \mathbb{I}(i \in Ind^d) .\nonumber
+%\end{align}
+%
+%The choice of the modeling of the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick \citep{Kool19stochastic}.
+In addition, we suppose the prior distribution of $m_l^d$ is uniform over the admissible set $\Delta^{n_p}_k$, $P(m_l^d | x, \theta, d) \ \text{=} \ U(\Delta^{n_p}_k)$.
+%
+%With our assumption over the prior distribution and our choice of the variational distribution, we express the two terms of the lower-bound \eqref{eq:lower-bound} as follows
+%
+%\begin{align}
+%\hspace{-2.em}
+%T_1 &= \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y | x,\theta,m_l^d, d)  \big] \nonumber \\
+%&= \displaystyle{\mathop{\mathbb{E}}_{g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} Gumbel(0,1) \ }} \big[ \text{log} P(y | x,\theta,\tilde{m}_l^d,d) \big] \nonumber
+%\end{align}
+%where
+%\begin{align}
+%\hspace{-2.em}
+%g_l^d & \in \mathbb{R}^{n_p} \nonumber \\
+%g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} Gumbel(0,1) \nonumber \\
+%& \forall p \in \{ 0,\cdots,n_p\text{-}1 \} \nonumber \\
+%i_1, \cdots, i_k &= Top\_k \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \label{eq:top-k} \\ 
+%&g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \} \nonumber \\
+%\tilde{m}_l^d(p) &= \mathbb{I}(p \in Ind^d). \nonumber 
+%\end{align}
+%and
+%\begin{align}
+%\hspace{-2.em}
+%T_2 &= - KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\ 
+%	&= \mathbb{H} \big[ P(m_l^d | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\ 
+%	&= \mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\
+%	&\geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \label{eq:entropy}
+%\end{align}
+%
+%we will prove Inequality \eqref{eq:entropy} in Appendix \ref{appendix:b}.
+%
+%Now we go back to our loss function, which is the negative of the lower-bound \eqref{eq:lower-bound}, consisting of 2 terms $-T_1$ and $-T_2$.
+The inequality \eqref{eq:entropy} shows that the upperbound of $-T_2$ is $\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d))$.
+In the training, we will maximize the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ as a regularization over the parameters $\Phi_l^d$ of the variational distribution.
+%
+%Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi_l^d$ into log-likelihood function and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}.
+However, the operator Top-K, which defines $\tilde{m}_l^d$ in Equation \eqref{eq:top-k}, is not differentiable.
+Therefore, we approximate this function by the Soft-Top-K function defined as follows
+%
+%\begin{align}
+%\hat{m}_l^d(\tau) &= \displaystyle{\mathop{argmin}_{\substack{
+%       0 \leqslant m_i \leqslant 1 \ \forall 0 \leqslant i \leqslant n_d\text{-}1 \label{eq:soft-top-k}\\
+%        1^{T}.m = k
+%      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \\
+%& \text{in which} \nonumber \\
+%H_b(m) &= - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i) \nonumber 
+%\end{align}
+%
+%In Appendix \ref{appendix:a}, we will prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
+Furthermore, we will also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and its differentiation is computed via implicit differentiation theorem.
+During the training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyperparameter $\tau$ to $0$.
+The gradient of the term $T_1$ w.r.t $\Phi_l^d$ is computed via the chain rule as follows
+%\begin{equation}
+%\frac{\partial T_1}{\Phi_l^d} = \frac{\partial T_1}{\hat{m}_l^d(\tau)} \times \frac{\hat{m}_l^d(\tau)}{\Phi_l^d}
+%\end{equation}
+%The gradient $\frac{\partial T_1}{\hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\hat{m}_l^d(\tau)}{\Phi_l^d}$ is computed via implicit differentiation, that we will explain carefully in Appendix \ref{appendix:a}.
+%
+%We jointly train the Transformer parameters $\theta$ and the parameters of the latent dropping masks $\Phi_l^d$ using the following multi-task loss.
+%
+%\begin{align}
+%\hspace{-0.em}
+%\mathcal{L}(\theta,\Phi_l^1,\cdots,\Phi_l^k) = \displaystyle{\mathop{\sum}_{d=1}^{k}} \displaystyle{\mathop{\mathbb{E}}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)}} \big[-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)\big]
+%\end{align}
+%
+%in which $\mathcal{D}_s^d$ is distribution of the task $d$ over the input space $\Omega^d_s$; $g^d: \Omega^d_s \rightarrow \Omega^d_t$ is the mapping function, which our multi-task modern needs to learn, of the task $d$; $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$ is the ELBO loss, which we defined in Equation \ref{eq:lower-bound}.
+%
+%Finally, in the inference, we define the dropping mask for $l^{th}$ layer and domain $d$ as follows
+%
+%\begin{align*}
+%Ind_l^d &= Top_k(\Phi_l^d) \\
+%m_l^d &= \mathbb{I}(i\in Ind_l^d)\\
+%\end{align*}
+
+\section{Multi-task group dropout \label{sec:architecture}}
+\subsection{Network architecture, groups and layers}
+Most architectures for multitask learning are based on a matching of subset of model parameters with tasks.
+While our model also needs to know the task $d \in [0 \dots n_d-1]$ for each training and test instance, the structure of our Transformer networks \citep{Vaswani17attention} is based on the notion of \emph{groups of nodes in the computation graph}.
+At the input of each Transformer layer $l \in [1 \dots L]$, we partition all input state vectors into $n_p$ groups of nodes, and zero-out a task-dependent subset of these groups.\fyDone{In the attention heads only?
+In the FF?} The assignment of tasks to groups will be learned from the data, under the constraint that each task only activates exactly $k$ groups of \emph{active nodes}, while the all the other values are nullified akin to a dropout process (see Figure~\ref{fig:group_dropout}).
+\begin{figure}
+  \centering
+\includegraphics[width=0.4\textwidth]{group_dropout}
+\caption{Latent group dropout.
+The set of nodes\fyDone{parameters ?} in each layer is divided into equal-sized groups.
+For each task, we only keep a fixed number of groups and nullify all the other nodes.
+Our method learns which groups will be kept / dropped for each task.}
 \label{fig:group_dropout}
 \end{figure}
- The dropping mask $r_l^d$ of $l^{th}$ layer and domain $d$ is defined as follows
+Formally, a \emph{group dropout mask} $m_l^d$ is a $n_p$-dimensional binary vector containing exactly $k$ ones: the group $p$ ($\in  [0,\dots,n_p\text{-}1]$) is retained for task $d$ if $m_l^d(p)\text{\small =}1$ and is dropped if $m_l^d(p)\text{\small =} 0$.
+We denote $\Delta^{n_p}_k = \{ m \in \{0,1\}^{n_p} \text{ such that }  \mid{} m \mid_{L_1} = k \}$ the set of all admissible dropout masks, with $\mid{} m \mid_{L_1}$ the $L_1$ norm of vector $m$.
+
+Given $m_l^d$, the mask $r_l^d$ for task $d$ in layer $l$ is then derived as follows:
 \begin{align*}
@@
     1, & \text{if}\ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p} \\
-      & \text{AND} \  m_l^d(p) \text{\small ==} 1 \\
+    & \text{AND} \  m_l^d(p) \text{\small =} 1 \\
     0, & \text{otherwise},
-    \end{cases} \\
-p & \in  \{0,\dots,n_p\text{-}1 \}, \\
-d & \in  \{0,\dots,n_d\text{-}1 \}, \\
-i & \in  \{0,\dots,d_k\text{-}1 \}, \\
-m_l^d & \in  \{0,1\}^{n_p} \ \forall d,l \, \\
-\mid m_l^d \mid_{L_1} & =  k , \\
-\Delta^{n_p}_k & =  \{ m \in \{0,1\}^{n_p} | \mid m \mid_{L_1} = k \},
+  \end{cases} 
+  % p & \in  \{0,\dots,n_p\text{-}1 \}, \\
+  % d & \in  \{0,\dots,n_d\text{-}1 \}, \\
+  % i & \in  \{0,\dots,d_k\text{-}1 \}, \\
+  % m_l^d & \in  \{0,1\}^{n_p} \ \forall d,l \, \\
+  % \mid m_l^d \mid_{L_1} & =  k , \\
+  % \Delta^{n_p}_k & =  \{ m \in \{0,1\}^{n_p} | \mid m \mid_{L_1} = k \},
 \end{align*}
-where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
-The group $p$ is retained for domain $d$ if $m_l^d(p)\text{\small ==}1$ and is dropped if $m_l^d(p)\text{\small ==}0$.
-We also denote $\Delta^{n_p}_k$ the set of all admissible dropping masks.
+where $d_k$ is the dimension of the hidden state; $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
+The propagation of information within the network then depends on the current task value and can be formalized as:
+\begin{align*}
+  \forall l \in [0,\cdots, L-1]: \tilde{h}^l &= h^l \odot r_l^d ,\\
+  h^{l+1} &= \operatorname{LAYER}^{l+1}(\tilde{h}^l) ,\\
+\end{align*}
+where $\operatorname{LAYER}^l()$ represents all the computations in Transformer layer $l$.
 
-We incorporate the latent variables $m_{l\in[0,\cdots,L]}^d$ into the log-likelihood of our model as below.
-For the sake of brevity, we denote $m_l^d$ the set of latent variables $m_{l\in[0,\cdots,L]}^d$
+\subsection{Training with latent dropout masks}
+Assuming standard notation for our translation model  $P(y|x,d;\theta)$ where $x$, $y$ and $\theta$ respectively refer to the input, output, and parameter vector, the latent variables $m_{l}^d, l \in[0,\dots,L], d \in [0, \dots, n_d-1]$ are introduced as follows.
+We also assume a prior distribution for $m_{l}^d$ chosen to be the uniform distribution over $\Delta^{n_p}_k$: $P(m_l^d | x, d; \theta) = \operatorname{Unif}(\Delta^{n_p}_k)$; variables for each layer are independent and collectively referred to as $m^d$.
+For any (variational) distribution $Q(m^1 \dots m^{n_d}; \Phi)$ with parameters $\Phi$, it is well-known that the log-likelihood is lower-bounded by the so-called $\operatorname{ELBO}$ function, made of a summation of two terms: the \emph{distortion} $D$ and the \emph{rate} $R$ defined as follows:\footnote{We drop the dependency in $x$ to avoid notational clutter.}
 \begin{align}
-\hspace{-3.em}
-\text{log} P(y|x,\theta,d) & \text{=} \text{log} P(y, m_l^d |x,\theta,d) \text{-} \text{log} P(m_l^d|y, x,\theta,d) \nonumber \\
-& \text{=} \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
-& \text{+} KL(P(m_l^d | \Phi_l^d) \parallel P(m_l^d|y, x,\theta,d)) \nonumber \\
-& \geqslant \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
-& \text{=} \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y | x,\theta,m_l^d, d) \big] \label{eq:lower-bound} \\
-& \text{-} KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\
-& \text{=} \mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y). \nonumber
+  \log P(y|d;\theta) \ge & \operatorname{ELBO}(y,d; \theta, \Phi) \nonumber \\
+  \operatorname{ELBO}(y,d; \theta, \Phi) =&  D(y,d; \theta, \Phi) - R(y,d; \theta, \Phi) \label{eq:lower-bound}\\
+  D(y,d; \theta, \phi) = & \mathbb{E}_{m^d \sim Q(m^d |d,\Phi)} \log P(y | m^d, d; \theta) \nonumber\\
+  R(y,d; \theta, \phi) = & \operatorname{KL}(Q(m^d | d, \Phi)||P(m^d | d; \theta)), \nonumber
 \end{align}
-$\mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y)$ is the lowerbound, which is called the evidence lower bound (ELBO) in variational statistic \citep{Diederick14auto}, of our log-likelihood.
-Our training loss will be $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$.
+where $\operatorname{KL}$ is the Kullback-Leibler divergence.
+We use $-\operatorname{ELBO}(x, y\theta,\Phi_l)$ as our surrogate training objective, as a tractable approximation of the intractable likelihood, and try to minimize this function in $\theta$ and $\Phi$.
 
-The variational distribution of $m_l^d$, $P(m_l^d|\Phi_l^d)$, is defined as follows
-
+The variational distribution $Q$ of $m^d$ is defined on layerwise basis, each layer $l$ involving only a subset $\Phi_l^d$ of parameters:
+\begin{align}
+  \operatorname{Ind}^d = \{ i_1, \cdots, i_k \} & \sim \text{SRS}(\operatorname{softmax}(\Phi_l^d), k) \nonumber \\
+  m_l^d(i) & =  \text{$\mathbb{I}$}(i \in \operatorname{Ind}^d), \nonumber
+\end{align}
+where $\text{SRS}(\pi,k)$ denotes the process of sampling $k$ times without replacement from the distribution $\pi$, and $\mathbb{I}$ is the indicator function.
+This particular choice the modeling of the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick of \citet{Kool19stochastic} that we use below.
+Given our assumptions for the prior distribution and our choice for the variational distribution, the two terms in Eq.~\eqref{eq:lower-bound} can be computed as:
 \begin{align}
 \hspace{-2.em}
-&\Phi_d \in \mathbb{R}^{n_p} , \nonumber \\
-&Ind^d \text{=}  \{ i_1, \cdots, i_k \} \nonumber \\
-&\sim \text{\small Sampling\_without\_replacement} \big(\text{softmax}(\Phi_l^d) \big) , \nonumber \\
-&m_l^d(i) \text{=} \mathbb{I}(i \in Ind^d) .\nonumber
+D &= \mathbb{E}_{m^d \sim Q(m^d | d | \Phi)} \text{log} P(y | m^d, d, \theta) \nonumber \\
+&= \mathbb{E}_{g^d \sim^{\text{i.i.d}} G(0,1)} \big[ \log P(y |\tilde{m}^d,d, \theta,) \big] \nonumber
 \end{align}
-
-The choice of the modeling of the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick \citep{Kool19stochastic}.
-In addition, we suppose the prior distribution of $m_l^d$ is uniform over the admissible set $\Delta^{n_p}_k$, $P(m_l^d | x, \theta, d) \ \text{=} \ U(\Delta^{n_p}_k)$.
-
-With our assumption over the prior distribution and our choice of the variational distribution, we express the two terms of the lower-bound \eqref{eq:lower-bound} as follows
-
+where the generation process $G(0,1)$ is a product of independent Gumbel distributions, yielding:
 \begin{align}
   \hspace{-2.em}
-T_1 &= \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y | x,\theta,m_l^d, d)  \big] \nonumber \\
-&= \displaystyle{\mathop{\mathbb{E}}_{g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} Gumbel(0,1) \ }} \big[ \text{log} P(y | x,\theta,\tilde{m}_l^d,d) \big] \nonumber
+  \forall d,  g^d & =  [g_0^d, \dots, g_L^d],  \text{ with } g_l^d \in \mathbb{R}^{n_p} \nonumber \\
+  \forall p,  g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1) \nonumber \\
+  \{ i_1, \cdots, i_k \} &= \operatorname{Top-k} \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \label{eq:top-k} \\ 
+       & \quad \quad g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \} \nonumber \\
+\tilde{m}_l^d(p) &= \mathbb{I}(p \in Ind^d). \nonumber 
 \end{align}
-where
+For the second term, the derivation is the following:
 \begin{align}
 \hspace{-2.em}
-g_l^d & \in \mathbb{R}^{n_p} \nonumber \\
-g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} Gumbel(0,1) \nonumber \\
-& \forall p \in \{ 0,\cdots,n_p\text{-}1 \} \nonumber \\
-i_1, \cdots, i_k &= Top\_k \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \label{eq:top-k} \\ 
-&g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \} \nonumber \\
-\tilde{m}_l^d(p) &= \mathbb{I}(p \in Ind^d). \nonumber 
+  R    &= \operatorname{KL}(Q(m^d | d, \Phi)||P(m^d | d; \theta)), \nonumber \\
+  &= \displaystyle{\mathop{\sum}_{l=1}^L} \mathbb{H} \big[ Q(m_l^d | d, \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\ 
+  &= \displaystyle{\mathop{\sum}_{l=1}^L} \mathbb{H}  \big[ Q(i_1,\cdots,i_k | d, \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k)  \nonumber  \\ 
+	&  \geqslant \displaystyle{\mathop{\sum}_{l=1}^L} \mathbb{H} \big[ Q(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \label{eq:entropy}
 \end{align}
-and
-\begin{align}
-\hspace{-2.em}
-T_2 &= - KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\ 
-	&= \mathbb{H} \big[ P(m_l^d | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\ 
-	&= \mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\
-	&\geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \label{eq:entropy}
-\end{align}
+We prove inequality~\eqref{eq:entropy} in Appendix~\ref{appendix:b}.
 
-we will prove Inequality \eqref{eq:entropy} in Appendix \ref{appendix:b}.
+Inequality~\eqref{eq:entropy} shows that a lowerbound of $R$ is $\displaystyle{\mathop{\sum}_{l=1}^L} -\text{log}(\#\Delta^{n_p}_k) + \mathbb{H}(\text{softmax}(\Phi_l^d))$.
+During training, we maximize the entropy $\displaystyle{\mathop{\sum}_{l}^L} \mathbb{H}(\text{softmax}(\Phi_l^d))$ as a regularization over the parameters $\Phi_l^d$ of the variational distribution.
 
-Now we go back to our loss function, which is the negative of the lower-bound \eqref{eq:lower-bound}, consisting of 2 terms $-T_1$ and $-T_2$.
-The inequality \eqref{eq:entropy} shows that the upperbound of $-T_2$ is $\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d))$.
-In the training, we will maximize the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ as a regularization over the parameters $\Phi_l^d$ of the variational distribution.
-
-Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi_l^d$ into log-likelihood function and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}.
-However, the operator Top-K, which defines $\tilde{m}_l^d$ in Equation \eqref{eq:top-k}, is not differentiable.
-Therefore, we approximate this function by the Soft-Top-K function defined as follows
-
+Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi$ into the objective function and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}.
+However, the operator $\operatorname{Top-k}$, which serves to define $\tilde{m}_l^d$ in Equation~\eqref{eq:top-k}, is not differentiable.
+Therefore, we approximate this function by the $\operatorname{Soft-Top-K}$ function defined as follows:
 \begin{align}
@@
 
-In Appendix \ref{appendix:a}, we will prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
-Furthermore, we will also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and its differentiation is computed via implicit differentiation theorem.
-During the training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyperparameter $\tau$ to $0$.
-The gradient of the term $T_1$ w.r.t $\Phi_l^d$ is computed via the chain rule as follows
+In Appendix \ref{appendix:a}, we prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
+Furthermore, we also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and that its differentiation can be computed using the implicit differentiation theorem.
+During training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyperparameter $\tau$ to $0$.
+The gradient of the term $D$ w.r.t $\Phi_l^d$ is computed via the chain rule as follows:
 \begin{equation}
-\frac{\partial T_1}{\Phi_l^d} = \frac{\partial T_1}{\hat{m}_l^d(\tau)} \times \frac{\hat{m}_l^d(\tau)}{\Phi_l^d}
+\frac{\partial D}{\partial \Phi_l^d} = \frac{\partial T_1}{\partial \hat{m}_l^d(\tau)} \times \frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}
 \end{equation}
-The gradient $\frac{\partial T_1}{\hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\hat{m}_l^d(\tau)}{\Phi_l^d}$ is computed via implicit differentiation, that we will explain carefully in Appendix \ref{appendix:a}.
+The gradient $\frac{\partial T_1}{\partial \hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$ is computed via implicit differentiation, as explained in Appendix~\ref{appendix:a}.
 
-We jointly train the Transformer parameters $\theta$ and the parameters of the latent dropping masks $\Phi_l^d$ using the following multi-task loss.
-
+We jointly train the Transformer parameters $\theta$ and the parameters of the variational distribution $\Phi$ using the following multi-task loss.\fyDone{This is repeated}
 \begin{align}
 \hspace{-0.em}
-\mathcal{L}(\theta,\Phi_l^1,\cdots,\Phi_l^k) = \displaystyle{\mathop{\sum}_{d=1}^{k}} \displaystyle{\mathop{\mathbb{E}}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)}} \big[-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)\big]
+\ell(\theta,\Phi_l^1,\cdots,\Phi_l^k) = \displaystyle{\mathop{\sum}_{d=1}^{k}} \displaystyle{\mathop{\mathbb{E}}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)}} \big[-\operatorname{ELBO}(\theta,\Phi^d,x,y)\big]
 \end{align}
 
-in which $\mathcal{D}_s^d$ is distribution of the task $d$ over the input space $\Omega^d_s$; $g^d: \Omega^d_s \rightarrow \Omega^d_t$ is the mapping function, which our multi-task modern needs to learn, of the task $d$; $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$ is the ELBO loss, which we defined in Equation \ref{eq:lower-bound}.
+in which $\mathcal{D}_s^d$ is distribution of task $d$ over the input space $\Omega^d_s$; $g^d: \Omega^d_s \rightarrow \Omega^d_t$\fyDone{allready defined} is the mapping function for the task $d$, which our multi-task model needs to learn; $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$ is the ELBO loss, defined in Equation~\ref{eq:lower-bound}.
 
-Finally, in the inference, we define the dropping mask for $l^{th}$ layer and domain $d$ as follows
-
+Finally, during inference, we define the dropout mask for layer $l$ and task $d$ as follows:
 \begin{align*}
-Ind_l^d &= Top_k(\Phi_l^d) \\
-m_l^d &= \mathbb{I}(i\in Ind_l^d)\\
+  Ind_l^d &= \operatorname{Top-k}(\Phi_l^d) \\
+  m_l^d &= \mathbb{I}(i\in Ind_l^d),\\
 \end{align*}
+meaning that we simply pick the $k$ most likely parameter groups for the task at hand, and define the dropout mask accordingly.
 
@@
 
+We also plot the languages based on their dropouts in \ref{fig:visual}
+\begin{figure*}[htbp]
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include first image
+  \includegraphics[width=0.97\textwidth]{ted_related}  
+  \caption{TED-Related}
+  \label{fig:visual-related}
+\end{subfigure}
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=0.97\textwidth]{ted_diverse}  
+  \caption{TED-Diverse}
+  \label{fig:visual-diverse}
+\end{subfigure}
+\caption{Visualization of languages according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
+\label{fig:visual}
+\end{figure*}
+
 \section{Related work}
@@
 
-\citet{xian20deep} proposed using different depths for different tasks.
+\textbf{Sparse Transformer}\citet{xian20deep} proposed using different depths for different tasks.
 The authors aimed to match the depth of the sub-network to the complexity of the task.
@@
 
-%Adapters
-%
-%With the similar idea of language proximity based parameter sharing, recent works propose multi-decoder model.
-Sharing the same encoder among languages, they route languages in different families to different decoders
+\textbf{Adapters} have proven to be very efficient for NLP multi-task modeling \citep{houlsby19parameter,Bapna19simple,Pham20Study,pfeiffer20adapterhub}.
+The method consists of plugging several adapters to the intermediate layers of a pretrained Transformer and finetuning these adapters on the downstream tasks.
+Adapters can also be trained unsurpervisedly in multilingual translation \citep{Philip20monolingual}.
+However, hard-coded separation between the domains of different tasks leads to a catastrophic forgetting \citep{pfeiffer21adapterfusion}, which is a common problem in multi-task modeling using neural networks \cite{Michael89catastrophic}.
+In multi-domain translation, recently \citet{Pham21revisiting} demonstrated the brittleness of domain adapters against fuzzy domain separations, out-of-domain distributions and erroneous domain tags.
+Recently, several works have been dedicated to mitigate this weakness through mixture of expert mechanisms, for example,\cite{pfeiffer21adapterfusion}.
+
+Recently, \citet{biao21share} proposed jointly learning how to route between shared representation and language-specific representation via a conditional language-specific routing and the parameters of the underlying Transformer.
+This method is very related to the FusionAdapters of \cite{pfeiffer21adapterfusion}.
+Both approaches assume the final representation is a combination or a choice between shared representation and task-specific representations.
+The proximity between task is not taken into account in the routing mechanism.
+Our method proposed a different point of view on the problem of multi-task routing in the underlying network.
 
@@
 	\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048, the multi-head attentions contain with 8 heads.
-	\item Adapter based Transformer: The intermediate feeforward dimension is set as 2048 as in \citet{Pham21Revisiting}.
+	\item Adapter based Transformer: The intermediate feeforward dimension is set as 2048 as in \citet{Pham21revisiting}.
 	\item \system{LaMGD} Tranformer: There is no change in the architecture.

### 2021-11-14 2022/acl/acl_latex-fy.tex
new title
relecture sections 1 / 2

@@
 \title{Multi-domain, multilingual translation with latent multi-task group dropout}
-\title{Latent group dropout for multilingual and multidomain machine translation}
+\title{Latent Group Dropout for Multilingual and Multidomain Machine Translation}
 % Author information can be set in various styles:
@@
 \begin{abstract}
-  [FY] Multidomain and multilingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalties of the task at hand, while smaller parts are reserved to model the peculariaties of a language or a domain.
-In adapter-based approaches, these strategies are hard coded in the network architecture.
-In this work, we propose a new method aimed at learning the parameter sharing pattern, using a latent-variable model.
-We show how this model can be trained end to end, and report experimental results which show that the learned pattern are both meaningful, and yield improved translation performance without any increase of the model size.
+  Multidomain and multilingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalties of the tasks at hand, while smaller parts are reserved to model the peculariaties of a language or a domain.
+In adapter-based approaches, these strategies are hard coded in the network architecture, independent of the similarities that exist between tasks.
+In this work, we propose a new method aimed at better taking advantage of these similarities, using a latent-variable model.
+We also develop new techniques to train this model end to end and report multiple experimental results that show the the learned patterns are both meaningful and yield improved translation performance, without any increase of the model size.
 \end{abstract}
@@
 Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using adapter layers as studied in \citep{houlsby19parameter,Bapna19simple,Pham20Study}.
-If these techniques have proven effective for building strong baselines, such fail to fully take advantage of the similarities that exist between domains and tasks, as documented eg.\ in \citep{Pham21revisiting}.
-This is because the partition of the parameter space between generic or task-specific sub-parts, and their allocation to each task is hard coded in the network, irrespective of the existing commonalties and differences in the data space.
+If these techniques have proven effective for building strong baselines, they fail to fully take advantage of the similarities that exist between domains and tasks, as documented eg.\ in \citep{Pham21revisiting}.
+This is because the partition of the parameter space between generic or task-specific subparts, and their allocation to each task is hard-coded in the network, irrespective of the actual commonalties and differences in the data space.
 
-In this paper, we develop techniques that can take the similarity between tasks into account in a more effective way, by learning the network organization from the data.
+In this work, we develop techniques that can take the similarity between tasks into account in a more effective way, by learning the network organization from the data.
 To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the parameter space, and show how training can still be performed end-to-end using a variational surrogate of the standard loss function.
-Our experiments with multilingual and multidomain machine translation confirm that this method can automatically detect similarities in the data, and yield related tasks to share more parameters.
-They also show that this method is comparable to using adapter layers in a number of empirical comparisons; contrarily to adapters, these performance are obtained without any increase of the model size.
+Our experiments with multilingual and multidomain machine translation confirm that this method can automatically detect similarities in the data, meaning that related tasks use the same subparts of the network.
+Our results also show that this method is comparable to using adapter layers in a number of empirical comparisons; however, contrarily to adapters, these performance are obtained without any increase of the model size.
 Our main contributions are summarized below:
@@
   \setlength{\parsep}{0pt}
-\item We provide a sound mathematical formulation to the problem of jointly learning routing task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling;
-\item We introduce algorithms to train this model end to end with very little extra computational cost.
+\item We provide a sound mathematical formulation to the problem of jointly learning task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling approach;
+\item We present algorithms to train this model end to end with very little extra computational cost.
 \item We report, using an extensive experimental setting, significant gains of this method with respect to strong baselines;
@@
 \end{itemize}
-
-\fyTodo{Can you do distillation ? That is only keep the parameter for one domain and still get good results (better than DA ?)}
+\fyDone{Can you do distillation ? That is only keep the parameter for one domain and still get good results (better than DA ?)}
 
@@
 \subsection{Network architecture, groups and layers}
-Most architectures for multitask learning are based on a matching of subset of model parameters with tasks.
-While our model also needs to know the task $d \in [0 \dots n_d-1]$ for each training and test instance, the structure of our Transformer networks \citep{Vaswani17attention} is based on the notion of \emph{groups of nodes in the computation graph}.
+Many architectures for multitask learning are based on a matching of subset of model parameters with tasks.
+Given the task and the input instance, only a subpart of the network will be involved in the computation of the output value, based on a predefined association between subnetworks and tasks.
+The adapter architecture of \citep{Bapna19simple} illustrates this strategy, where a task-dependent set of layers is activated for each task.
+
+In our approach, we also require to know the task $d \in [0 \dots n_d-1]$ for each training and test instance.
+The structure of our Transformer networks \citep{Vaswani17attention} is however based on the notion of \emph{groups of nodes in the computation graph}.
 At the input of each Transformer layer $l \in [1 \dots L]$, we partition all input state vectors into $n_p$ groups of nodes, and zero-out a task-dependent subset of these groups.\fyDone{In the attention heads only?
-In the FF?} The assignment of tasks to groups will be learned from the data, under the constraint that each task only activates exactly $k$ groups of \emph{active nodes}, while the all the other values are nullified akin to a dropout process (see Figure~\ref{fig:group_dropout}).
+In the FF?} The assignment of tasks to groups will be learned from the data, under the constraint that each task only activates exactly $k$ groups of \emph{active nodes}, while the all the other values are nullified, akin to a dropout process (see Figure~\ref{fig:group_dropout}).
 \begin{figure}
@@
 The set of nodes\fyDone{parameters ?} in each layer is divided into equal-sized groups.
-For each task, we only keep a fixed number of groups and nullify all the other nodes.
+For each task, we only keep a fixed number of active groups of nodes and nullify all the other nodes.
 Our method learns which groups will be kept / dropped for each task.}
@@
 \end{figure}
-Formally, a \emph{group dropout mask} $m_l^d$ is a $n_p$-dimensional binary vector containing exactly $k$ ones: the group $p$ ($\in  [0,\dots,n_p\text{-}1]$) is retained for task $d$ if $m_l^d(p)\text{\small =}1$ and is dropped if $m_l^d(p)\text{\small =} 0$.
-We denote $\Delta^{n_p}_k = \{ m \in \{0,1\}^{n_p} \text{ such that }  \mid{} m \mid_{L_1} = k \}$ the set of all admissible dropout masks, with $\mid{} m \mid_{L_1}$ the $L_1$ norm of vector $m$.
+Formally, a \emph{group dropout mask} $m_l^d$ is a $n_p$-dimensional binary vector containing exactly $k$ ones: the group $p$ ($\in  [0,\dots,n_p\text{-}1]$) is retained for task $d$ if $m_l^d(p) =1$ and is dropped if $m_l^d(p) = 0$.
+We denote $\Delta^{n_p}_k = \{ m \in \{0,1\}^{n_p} \text{ such that }  \mid{} m \mid_{L_1} = k \}$ the set of all admissible dropout masks, with $\mid{} m \mid_{L_1}$ the $L_1$ norm of vector $m$; $\#\Delta^{n_p}_k$ is the cardinal of $\Delta^{n_p}$.
 
-Given $m_l^d$, the mask $r_l^d$ for task $d$ in layer $l$ is then derived as follows:
+Given $m_l^d$, the mask $r_l^d$ for task $d$ in layer $l$ is then derived as:
 \begin{align*}
@@
 \end{align*}
-where $\operatorname{LAYER}^l()$ represents all the computations in Transformer layer $l$.
+where $\operatorname{LAYER}^l()$ represents all the computations in Transformer layer $l$. It is applied at all positions of each layer in the encoder and in the decoder.  
 % The dropping mask $r_l^d$ is computed as follows
@@
 Assuming standard notation for our translation model  $P(y|x,d;\theta)$ where $x$, $y$ and $\theta$ respectively refer to the input, output, and parameter vector, the latent variables $m_{l}^d, l \in[0,\dots,L], d \in [0, \dots, n_d-1]$ are introduced as follows.
-We also assume a prior distribution for $m_{l}^d$ chosen to be the uniform distribution over $\Delta^{n_p}_k$: $P(m_l^d | x, d; \theta) = \operatorname{Unif}(\Delta^{n_p}_k)$; variables for each layer are independent and collectively refered to as $m^d$.
-For any (variational) distribution $Q(m^1 \dots m^{n_d}; \Phi)$ with parameters $\Phi=\{\phi_l^1, ..., \phi_L^{n_d}\}$, it is well-known that the log-likelihood is lower-bounded by the so-called $\operatorname{ELBO}$ function, made of a summation of two terms: the \emph{distortion} $D$ and the \emph{rate} $R$ defined as follows:%\footnote{We drop the dependency in $x$ to avoid notational clutter.}
+We chose the prior distribution for $m_{l}^d$ as the uniform distribution over $\Delta^{n_p}_k$: $P(m_l^d | x, d; \theta) = \operatorname{Unif}(\Delta^{n_p}_k)$; variables for each layer are independent and collectively refered to as $m^d$.
+For any (variational) distribution $Q(m^1 \dots m^{n_d}; \Phi)$ with parameters $\Phi=\{\phi_l^1, ..., \phi_L^{n_d}\}$, it is well-known that the log-likelihood is lower-bounded by the so-called $\operatorname{ELBO}$ function (hereafter denoted $\ell$), made of a summation of two terms: the \emph{distortion} $D$ and the \emph{rate} $R$ defined as follows:%\footnote{We drop the dependency in $x$ to avoid notational clutter.}
 \begin{align}
-  \log P(y|x,d;\theta) \ge & \operatorname{ELBO}(x,y,d; \theta, \Phi) \nonumber \\
-  \operatorname{ELBO}(x,y,d; \theta, \Phi) =&  D(x,y,d; \theta, \Phi) - R(x,y,d; \theta, \Phi) \label{eq:lower-bound}\\
+  \log P(y|x,d;\theta) \ge & \ell(x,y,d; \theta, \Phi) \nonumber \\
+  \ell(x,y,d; \theta, \Phi) =&  D(x,y,d; \theta, \Phi) - R(x,y,d; \theta, \Phi) \label{eq:lower-bound}\\
   D(x,y,d; \theta, \phi) = & \mathbb{E}_{m^d \sim Q(m^d |d,\Phi)} \log P(y | m^d, x, d; \theta) \nonumber\\
@@
 where $\operatorname{KL}$ is the Kullback-Leibler divergence.
-We use $-\operatorname{ELBO}(x,y,d; \theta,\Phi)$ as our surrogate training objective, as a tractable approximation of the untractable likelihood, and try to minimize this function in $\theta$ and $\Phi$.
+We use $-\ell{ELBO}(x,y,d; \theta,\Phi)$ as our surrogate training objective, as a tractable approximation of the untractable likelihood, and try to minimize this function in $\theta$ and $\Phi$.
 
-The variational distribution $Q$ of $m^d$ is defined independently on a layerwise basis, each layer $l$ involving only a subset $\Phi_l^d$ of parameters:
+The variational distribution $Q$ of $m^d$ is defined independently on a layerwise basis; this means that each layer only involves a subset $\Phi_l^d$ of variational parameters.
+$Q$ is computed as follows:
 \begin{align}
@@
 where $\text{SRS}(\pi,k)$ denotes the process of sampling $k$ times without replacement from the distribution $\pi$, and $\mathbb{I}$ is the indicator function.
-This particular choice the modeling of the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick of \citet{Kool19stochastic} that we use below.
+This modeling choice for the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick of \citet{Kool19stochastic} that we use below.
 Given our assumptions for the prior distribution and our choice for the variational distribution, the two terms in Eq.~\eqref{eq:lower-bound} can be computed as:
@@
 \hspace{-2.em}
-  R    &= \operatorname{KL}(Q(m^d | d, \Phi)||P(m^d | d; \theta)), \nonumber \\ 
-	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(m_l^d | d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big) \nonumber \\ 
-	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1,\cdots,i_k | d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big)\nonumber \\
-	&\leqslant  -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \big)\label{eq:entropy}
+  R    &= \operatorname{KL}(Q(m^d | x, d, \Phi)||P(m^d | x, d; \theta)), \nonumber \\ 
+	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(m_l^d | x, d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big) \nonumber \\ 
+	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1,\cdots,i_k | x, d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big)\nonumber \\
+	&\leqslant  -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \big). \label{eq:entropy}
 \end{align}
 We prove inequality~\eqref{eq:entropy} in Appendix~\ref{appendix:b}.
-
-Inequality~\eqref{eq:entropy} shows that an upperbound of $R$ is $\sum_{l=1}(\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d)))$.
-During training, we maximize a sum over layers of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$\fyDone{Check sum} which performs a regularization over the parameters $\Phi^d$ of the variational distribution.
+This inequality shows that an upperbound of $R$ is $\sum_{l=1}^L(\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d)))$.
+During training, we thus maximize a sum over layers of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$\fyDone{Check sum} which performs a regularization over the parameters $\Phi^d$ of the variational distribution.
 
@@
 & \text{in which} \nonumber \\
-H_b(m) &= - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i) \nonumber 
+H_b(m) &= - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i). \nonumber 
 \end{align}
@@
 \hspace{-0.em}
-\ell(\theta,\Phi) = \sum_{d=1}^{n_d} \mathbb{E}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)} \big[-\operatorname{ELBO}(x,y,d; \theta,\Phi)\big]
+\mathcal{L}(\theta,\Phi) = \sum_{d=1}^{n_d} \mathbb{E}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)} \big[-\ell(x,y,d; \theta,\Phi)\big]
 \end{align}
 
-in which $\mathcal{D}_s^d$ is distribution of task $d$ over the input space $\Omega^d_s$; $\operatorname{MT}^d: \Omega^d_s \rightarrow \Omega^d_t$\fyDone{already defined} is the translation function for task $d$, which our multi-task model needs to learn; $-\mathcal{L}_{\text{ELBO}}(x,y, d, \theta,\Phi)$ is the ELBO loss, defined in Equation~\ref{eq:lower-bound}.
+in which $\mathcal{D}_s^d$ is distribution of task $d$ over the input space $\Omega^d_s$; $\operatorname{MT}^d: \Omega^d_s \rightarrow \Omega^d_t$\fyDone{already defined} is the translation function for task $d$, which our multi-task model needs to learn; $-\ell(x,y, d, \theta,\Phi)$ is the $\operatorname{ELBO}$ loss, defined in Equation~\ref{eq:lower-bound}.
 
@@
 \end{align*}
-meaning that we simply pick the $k$ most likely parameter groups for the task at hand, and define the dropout mask accordingly.
+meaning that we simply pick the $k$ most likely parameter groups for the task at hand, and define the state dropout mask accordingly.
 
-\section{Experimental settings}
+\section{Experimental settings \label{sec:experiments}}
 \subsection{Baselines}

### 2021-11-15 2022/acl/acl_latex.tex
fin de la discussion
modifies figures, related works

@@
 % Remove the "review" option to generate the final version.
+%\usepackage[review,nohyperref]{acl}
 \usepackage[review]{acl}
@@
 
+\makeatletter
+\newenvironment{breakablealgorithm}
+  {% \begin{breakablealgorithm}
+   \begin{center}
+     \refstepcounter{algorithm}% New algorithm
+     \hrule height.8pt depth0pt \kern2pt% \@fs@pre for \@fs@ruled
+     \renewcommand{\caption}[2][\relax]{% Make a new \caption
+       {\raggedright\textbf{\fname@algorithm~\thealgorithm} ##2\par}%
+       \ifx\relax##1\relax % #1 is \relax
+         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##2}%
+       \else % #1 is not \relax
+         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##1}%
+       \fi
+       \kern2pt\hrule\kern2pt
+     }
+  }{% \end{breakablealgorithm}
+     \kern2pt\hrule\relax% \@fs@post for \@fs@ruled
+   \end{center}
+  }
+\makeatother
+
 %subfigure
@@
 
-\title{Multi-domain, multilingual translation with latent multi-task group dropout}
+\title{Latent Group Dropout for Multilingual and Multidomain Machine Translation}
 
@@
 \end{itemize}
+We clarify the notation of the languages in multilingual experiments as follows
+\begin{itemize}
+	\item Diverse: bos (bosnia), mar(marathi),
+         hin(hindi), mkd(macedonia), ell(greek),
+         bulgaria(bul), french(fra), korean(kor) 
+	\item Related: azerbajiani (aze), belarusian(bel),
+          glacian(glg), slovak (slk), turkisk (tur),
+		  russian(rus), portugese(por),czech(ces)
+\end{itemize}
 The detailed description of the datasets is provided in Appendix \ref{appendix:c}.
@@
 \label{ssec:fuzzy}
-In this experiment, we reuse an experiment proposed by \citet{Pham21revisiting} which measures the efficiency of a multi-domain NMT system exploit the proximity between domains.
-The experiment use the same data as the previous experiment.
-However, the domain \domain{LAW} is now randomly split into two pseudo domains \domain{law$_1$} and \domain{law$_2$}.
+In this experiment, we reuse an experiment proposed by \citet{Pham21revisiting} which measures the efficiency of a multi-domain NMT system exploiting the proximity between domains.
+The experiment uses the same data as the previous experiment.
+However, the domain \domain{law} is now randomly split into two pseudo domains \domain{law$_1$} and \domain{law$_2$}.
 If the multi-domain system made full use of the proximity between \domain{law$_1$} and \domain{law$_2$}, there should be no significant difference between the performance of the system trained with six original domains and with seven domains replacing \domain{law} by \domain{law$_1$} and \domain{law$_2$}.
@@
 We report below the gap between two settings when using \system{LaMGD} Tranformer.
+
 \begin{table*}[h!]
@@
 
-Table \ref{tab:fuzzy} demonstrates a significant loss when training with two pseudo domain \domain{law$_1$} and \domain{law$_2$} compared to training with the real domain \domain{law}.
-Our system does not have this loss between two situation.
-In section \ref{ssec:abalation}, we show that our algorithm outputs the same sub-network for \domain{law$_1$} and \domain{law$_2$}, that allows a full transfer between two domain.
+Table \ref{tab:fuzzy} demonstrates a significant loss when training with two pseudo domains \domain{law$_1$} and \domain{law$_2$} compared to training with the real domain \domain{law}.
+Our system does not have this loss between two situations.
+In Section \ref{ssec:abalation}, we show that our algorithm outputs the same sub-network for \domain{law$_1$} and \domain{law$_2$}, that allows a full transfer between two domains.
 
@@
 Significant gains are observed in \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resourced languages in the list.
-However, \system{LaMGD} Tranformer lags behind multilingual \system{Transformer} in O2M-diverse.
+However, \system{LaMGD} Tranformer lags behind multilingual \system{Transformer} and Language Adapters in O2M-diverse.
 \begin{table*}[h!]
@@
 \label{ssec:abalation}
-In this section, we compare the sub-networks of the domains or of the language pairs by computing the average of the similarities between the corresponding dropping mask in each layer.
+This section compares the sub-networks of the domains or the language pairs by computing the average of the similarities between the corresponding dropping masks concatenated through all the layers of the underlying model.
 
@@
 Table \ref{tab:fuzzy-sim} shows that the sub-networks of \domain{law$_1$} and \domain{law$_2$} are identical, which is one of the goals of our method.
+Furthermore, we remark a large distance between \domain{rel} and the rest of the domains, which correlates to the fact that \domain{rel} has the most distant topic to other domains.
+In Figure ~\ref{fig:domain}, we visualize the domains using their dropping masks concatenated and mapped to 2D space by PCA.
+
 \begin{table*}[h!]
@@
 
-We also plot the languages based on their dropouts in \ref{fig:visual}
+We also plot the languages based on their dropouts in Figure\ref{fig:visual}.
 \begin{figure*}[htbp]
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{\textwidth}
   \centering
   % include first image
-  \includegraphics[width=0.97\textwidth]{ted_related}  
+  \includegraphics[width=0.7\textwidth]{ted_related}  
   \caption{TED-Related}
@@
 \end{subfigure}
-\begin{subfigure}{.5\textwidth}
+\\
+\begin{subfigure}{\textwidth}
   \centering
   % include second image
-  \includegraphics[width=0.97\textwidth]{ted_diverse}  
+  \includegraphics[width=0.7\textwidth]{ted_diverse}  
   \caption{TED-Diverse}
@@
 
+\begin{figure*}[htbp]
+\includegraphics[width=0.7\textwidth]{multi_domain}
+\caption{Visualization of domains according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
+\label{fig:domain}
+\end{figure*}
+
 \section{Related work}
-The most adopted paradigm is to have two groups of parameters including task-agnostic parameters and task-specific parameters.
-Domain adapters and language adapters are strong baselines of this paradigm.
-However, adapters does not take into account the proximity of the task in their designed architectures.
+Multidomain, multilingual translation
 
-\citep{sen19multilingual,kong21multilingual} used the similar idea of language proximity based parameter sharing.
+\citep{sen19multilingual,kong21multilingual} used the similar idea of language proximity-based parameter sharing.
 The authors proposed a multi-decoder model sharing the same encoder among languages and routed languages in different families to different decoders.
@@
 The authors aimed to match the depth of the sub-network to the complexity of the task.
-\citet{Gong21pay,Gong21adaptive} also take interest in the adaptive sparse Transformers which differ the selection of heads in multi-head attention, of layers and of blocks in feed-forward matrices between different tasks.
-Mixture-of-experts (MoE) is another effective approach of sparse modeling.
-Within the transformer model, GShard model replaces a single feedforward (FFN) sub-layer with an MoE module consisting of multiple FFN sub-layers \citep{lepikhin21gshard}.
+\citet{Gong21pay,Gong21adaptive} also take an interest in the adaptive sparse Transformers, which differ the selection of heads in multi-head attention, layers, and blocks in feedforward matrices between different tasks.
+Mixture-of-experts (MoE) is another effective approach to sparse modeling.
+Within the transformer model, the GShard model replaces a single feedforward (FFN) sub-layer with an MoE module consisting of multiple FFN sub-layers \citep{lepikhin21gshard}.
 \citet{william21switch}
@@
 The method consists of plugging several adapters to the intermediate layers of a pretrained Transformer and finetuning these adapters on the downstream tasks.
-Adapters can also be trained unsurpervisedly in multilingual translation \citep{Philip20monolingual}.
-However, hard-coded separation between the domains of different tasks leads to a catastrophic forgetting \citep{pfeiffer21adapterfusion}, which is a common problem in multi-task modeling using neural networks \cite{Michael89catastrophic}.
-In multi-domain translation, recently \citet{Pham21revisiting} demonstrated the brittleness of domain adapters against fuzzy domain separations, out-of-domain distributions and erroneous domain tags.
-Recently, several works have been dedicated to mitigate this weakness through mixture of expert mechanisms, for example,\cite{pfeiffer21adapterfusion}.
+Adapters can also be trained unsupervised in multilingual translation \citep{Philip20monolingual}.
+However, the hard-coded separation between the domains of different tasks leads to a catastrophic forgetting \citep{pfeiffer21adapterfusion}, which is a common problem in multi-task modeling using neural networks \cite{Michael89catastrophic}.
+In multi-domain translation, recently \citet{Pham21revisiting} demonstrated the brittleness of domain adapters against fuzzy domain separations, out-of-domain distributions, and erroneous domain tags.
+Recently, several works have aimed to mitigate this weakness through a mixture of expert mechanisms, for example,\cite{pfeiffer21adapterfusion}.
 
@@
 Both approaches assume the final representation is a combination or a choice between shared representation and task-specific representations.
-The proximity between task is not taken into account in the routing mechanism.
+The proximity between tasks is not taken into account in the routing mechanism.
 Our method proposed a different point of view on the problem of multi-task routing in the underlying network.
@@
 Secondly, to get rid of the requirement of domain information in the inference, we could condition the dropping selection directly on the input $x$ instead of the domain $d$.
-These two open question are both very interesting and promising.
+These two open questions are both very interesting and promising.
 \bibliography{bibliography}
@@
 \subsection{Multilingual datasets}
-We clarify the notation of the languages in multilingual experiments as follows
-\begin{itemize}
-	\item Diverse: bos (bosnia), mar(marathi),
-         hin(hindi), mkd(macedonia), ell(greek),
-         bulgaria(bul), french(fra), korean(kor) 
-	\item Related: azerbajiani (aze), belarusian(bel),
-          glacian(glg), slovak (slk), turkisk (tur),
-		  russian(rus), portugese(por),czech(ces)
-\end{itemize}
 Table \ref{tab:Corpora-multilingual} describes the subset of TED8 that we used in our experiments.
@@
 
+\section{\system{LaMGD}'s training algorithm}
+
+\begin{breakablealgorithm}
+\caption{Training \system{LaMGD}} \label{alg:mdmt}
+\begin{algorithmic}[1]
+\REQUIRE {
+  \begin{itemize}
+    \setlength{\itemsep}{-1pt}    \setlength{\parsep}{0pt}
+  \item[] 
+  \item $n_d$ corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equipped by an empirical distribution $D_d(x)$
+  \item number of retained groups: k, number of groups: $n_p$
+ \end{itemize}}
+\REPEAT 
+\STATE{Pick a batch from domain $d$}
+\STATE{Sampling $\forall p,  g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
+\IF{$Early\_stopping(Eval\_scores)$}
+	\STATE{break}
+\ENDIF
+\STATE{$i=i+1$}
+\UNTIL{$i> iter\_num$}
+\end{algorithmic}
+\end{breakablealgorithm}
+
 \end{document}

### 2021-11-15 2022/acl/acl_latex.tex
merged two versions

@@
 
+\title{Multi-domain, multilingual translation with latent multi-task group dropout}
 \title{Latent Group Dropout for Multilingual and Multidomain Machine Translation}
-
 % Author information can be set in various styles:
@@
 \begin{document}
-%\setlength{\abovedisplayskip}{3pt}
-%\setlength{\belowdisplayskip}{3pt}
-%\setlength{\parskip}{0pt}
-%\setlength{\abovedisplayskip}{0pt}
-%\setlength{\belowdisplayskip}{0pt}
-%\setlength{\abovedisplayshortskip}{0pt}
-%\setlength{\belowdisplayshortskip}{0pt}
+\setlength{\abovedisplayskip}{3pt}
+\setlength{\belowdisplayskip}{3pt}
 \maketitle
-%\begin{abstract}
-%Multi-domain and multilingual translation can be cast under multi-task learning in which each domain or language pair is considered a task.
-Sharing most of the parameters of the model causes a catastrophic interference between unrelated tasks.
-Most of the proposed multi-task architectures propose sharing the parameters of an underlying generic model and using different output layers to predict different tasks.
-However, these methods do not explicitly exploit the proximity between tasks.
-In this work, we aim to express the task proximity in the multi-task network.
-Instead of using a predefined multi-task network, we jointly learn routing task-dependent sub-networks and the parameters of the whole network.
-More precisely, we propose clustering the nodes of each layer into equal-sized groups.
-For each layer and each domain, a number of groups will be dropped.
-By consequence, we assign different sparse sub-networks to different tasks.
-Our method learns the task-dependent dropping selections and the parameters of the underlying network.
-Our approach allows the proximity between tasks to express in their dropping selections.
-Through an extensive set of experiments, we demonstrate the advantages of our method over strong baselines in multi-domain and multilingual translation.
-Furthermore, through an ablation study, we illustrate a significant correlation between the similarity between learned task-dependent dropping selections and the proximity between tasks.
-%\end{abstract}
 
 \begin{abstract}
-  Multidomain and multilingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalties of the task at hand, while smaller parts are reserved to model the peculariaties of a language or a domain.
-In adapter-based approaches, these strategies are hard coded in the network architecture.
-In this work, we propose a new method aimed at learning the parameter sharing pattern, using a latent-variable model.
-We show how this model can be trained end to end, and report experimental results which show that the learned pattern are both meaningful, and yield improved translation performance without any increase of the model size.
+  Multidomain and multilingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalties of the tasks at hand, while smaller parts are reserved to model the peculariaties of a language or a domain.
+In adapter-based approaches, these strategies are hard coded in the network architecture, independent of the similarities that exist between tasks.
+In this work, we propose a new method aimed at better taking advantage of these similarities, using a latent-variable model.
+We also develop new techniques to train this model end to end and report multiple experimental results that show the the learned patterns are both meaningful and yield improved translation performance, without any increase of the model size.
 \end{abstract}
@@
 \section{Introduction}
-
-%Multi-domain translation and multilingual translation develop one model for translations in multi-domain and multiple language pairs, respectively.
-These paradigms are motivated by the compactness of the translation model \cite{dabre20survey,Chu18multilingual} and the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
-However, sharing parameters of the model can cause the catastrophic interference between unrelated tasks \citep{conneau20unsupervised,wang20negative}.
-%
-%In multi-domain setting, \citet{Pham21revisiting} pointed out a challenge in how multi-domain systems fully exploit the similarity between domains.
-The authors demonstrated the weakness of these systems in handling the proximity between domains through a simple setting using a pseudo domain separation.
-Domain Adapters, which are strong baselines for multi-domain translation \cite{Pham20Study}, failed this test.
-Such predefined allocations of parameters to domains do not take into account the proximity of domains in their choice and that motivates us to learn both the .
-%
-%The method we developed here mitigates the catastrophic interference in training a neural machine translation (NMT) model for multiple domains or multiple language pairs.
-Our main contribution are summarized below:
-%\begin{itemize}
-%\setlength{\itemsep}{1pt}
-%  \setlength{\parskip}{0pt}
-%  \setlength{\parsep}{0pt}
-%	\item We provide a rigorous mathematical modeling the problem of jointly learning routing task-dependent sub-network and the parameters of the underlying model using variational probabilistic modeling.
-%	\item Our method propose end to end training with very little extra computational cost.
-%	\item With an extensive experimental setting, we demonstrate a significant gain of our method against strong baselines.
-%	\item We illustrate a significant correlation between the similarity between learned sub-networks and the proximity of their corresponding tasks.
-%\end{itemize}
-
 Multidomain and multilingual machine translation aim to develop one single model to perform translations for multiple domains and multiple language pairs, respectively.\footnote{We will refer to these two situations as 'multi-task MT' (or \mtsquare{} for short), and refer to individual domains and languages as 'tasks'.} These paradigms are motivated by the compactness of the resulting translation system \citep{Chu18multilingual,dabre20survey}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
@@
 Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using adapter layers as studied in \citep{houlsby19parameter,Bapna19simple,Pham20Study}.
-If these techniques have proven effective for building strong baselines, such fail to fully take advantage of the similarities that exist between domains and tasks, as documented eg.\ in \citet{Pham21revisiting}.
-This is because the partition of the parameter space between generic or task-specific sub-parts, and their allocation to each task is hard coded in the network, irrespective of the existing commonalities and differences in the data space.
+If these techniques have proven effective for building strong baselines, they fail to fully take advantage of the similarities that exist between domains and tasks, as documented eg.\ in \citep{Pham21revisiting}.
+This is because the partition of the parameter space between generic or task-specific subparts, and their allocation to each task is hard-coded in the network, irrespective of the actual commonalties and differences in the data space.
 
-In this paper, we develop techniques that can take the similarity between tasks into account in a more effective way, by learning the network organization from the data.
+In this work, we develop techniques that can take the similarity between tasks into account in a more effective way, by learning the network organization from the data.
 To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the parameter space, and show how training can still be performed end-to-end using a variational surrogate of the standard loss function.
-Our experiments with multilingual and multidomain machine translation confirm that this method can automatically detect similarities in the data, and yield related tasks to share more parameters.
-They also show that this method is comparable to using adapter layers in a number of empirical comparisons; contrarily to adapters, these performance are obtained without any increase of the model size.
+Our experiments with multilingual and multidomain machine translation confirm that this method can automatically detect similarities in the data, meaning that related tasks use the same subparts of the network.
+Our results also show that this method is comparable to using adapter layers in a number of empirical comparisons; however, contrarily to adapters, these performance are obtained without any increase of the model size.
 Our main contributions are summarized below:
@@
   \setlength{\parsep}{0pt}
-\item We provide a rigorous mathematical formulation to the problem of jointly learning routing task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling;
-\item We introduce algorithms to train this model end to end with very little extra computational cost.
-\item We report, using an extensive experimental setting, significant gains of this method with respect to strong baselines.
+\item We provide a sound mathematical formulation to the problem of jointly learning task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling approach;
+\item We present algorithms to train this model end to end with very little extra computational cost.
+\item We report, using an extensive experimental setting, significant gains of this method with respect to strong baselines;
 \item We study how this method can actually exploit the similarities between tasks to learn interpretable sub-networks.
 \end{itemize}
-
-%\section{Latent multi-task group dropout}
-%We define a group dropout for a domain $d$ as follows
-%\begin{align*}
-%\tilde{h}^l &= h^l \odot r_l^d ,\\
-%r_l^d &\in \{ 0,1 \}^{d_k}, \\
-%h^{l+1} &= \text{LAYER}^{l+1}(h^l) ,\\
-%l & \in [0,\cdots,L] ,\\
-%\end{align*}
-%where $L$ is number of Transformer layers of our NMT model, the binary vector $r_l^d$ is a dropping mask. We illustrate the group dropout in Figure \ref{fig:group_dropout}
-%\begin{figure}
-%\includegraphics[width=0.4\textwidth]{group_dropout}
-%\caption{Latent group dropout.
-The set of nodes in one layer is divided into equal-sized groups.
-For a domain, we drop a fixed number of groups.
-Our method learns which groups will be dropped for a domain.}
-%\label{fig:group_dropout}
-%\end{figure}
-% The dropping mask $r_l^d$ of $l^{th}$ layer and domain $d$ is defined as follows
-%\begin{align*}
-%r_l^d(i) &= \begin{cases}
-%      1, & \text{if}\ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p} \\
-%      & \text{AND} \  m_l^d(p) \text{\small ==} 1 \\
-%      0, & \text{otherwise},
-%    \end{cases} \\
-%p & \in  \{0,\dots,n_p\text{-}1 \}, \\
-%d & \in  \{0,\dots,n_d\text{-}1 \}, \\
-%i & \in  \{0,\dots,d_k\text{-}1 \}, \\
-%m_l^d & \in  \{0,1\}^{n_p} \ \forall d,l \, \\
-%\mid m_l^d \mid_{L_1} & =  k , \\
-%\Delta^{n_p}_k & =  \{ m \in \{0,1\}^{n_p} | \mid m \mid_{L_1} = k \},
-%\end{align*}
-%where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
-The group $p$ is retained for domain $d$ if $m_l^d(p)\text{\small ==}1$ and is dropped if $m_l^d(p)\text{\small ==}0$.
-We also denote $\Delta^{n_p}_k$ the set of all admissible dropping masks.
-%
-%We incorporate the latent variables $m_{l\in[0,\cdots,L]}^d$ into the log-likelihood of our model as below.
-For the sake of brevity, we denote $m_l^d$ the set of latent variables $m_{l\in[0,\cdots,L]}^d$
-%\begin{align}
-%\hspace{-3.em}
-%\text{log} P(y|x,\theta,d) & \text{=} \text{log} P(y, m_l^d |x,\theta,d) \text{-} \text{log} P(m_l^d|y, x,\theta,d) \nonumber \\
-%& \text{=} \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
-%& \text{+} KL(P(m_l^d | \Phi_l^d) \parallel P(m_l^d|y, x,\theta,d)) \nonumber \\
-%& \geqslant \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y, m_l^d | x,\theta,d) \text{-} \text{log} P(m_l^d | \Phi_l^d) \big] \nonumber \\
-%& \text{=} \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y | x,\theta,m_l^d, d) \big] \label{eq:lower-bound} \\
-%& \text{-} KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\
-%& \text{=} \mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y). \nonumber
-%\end{align}
-%$\mathcal{L}_{\text{ELBO}}(\theta, \Phi_l^d, x , y)$ is the lowerbound, which is called the evidence lower bound (ELBO) in variational statistic \citep{Diederick14auto}, of our log-likelihood.
-Our training loss will be $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$.
-%
-%The variational distribution of $m_l^d$, $P(m_l^d|\Phi_l^d)$, is defined as follows
-%
-%\begin{align}
-%\hspace{-2.em}
-%&\Phi_d \in \mathbb{R}^{n_p} , \nonumber \\
-%&Ind^d \text{=}  \{ i_1, \cdots, i_k \} \nonumber \\
-%&\sim \text{\small Sampling\_without\_replacement} \big(\text{softmax}(\Phi_l^d) \big) , \nonumber \\
-%&m_l^d(i) \text{=} \mathbb{I}(i \in Ind^d) .\nonumber
-%\end{align}
-%
-%The choice of the modeling of the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick \citep{Kool19stochastic}.
-In addition, we suppose the prior distribution of $m_l^d$ is uniform over the admissible set $\Delta^{n_p}_k$, $P(m_l^d | x, \theta, d) \ \text{=} \ U(\Delta^{n_p}_k)$.
-%
-%With our assumption over the prior distribution and our choice of the variational distribution, we express the two terms of the lower-bound \eqref{eq:lower-bound} as follows
-%
-%\begin{align}
-%\hspace{-2.em}
-%T_1 &= \displaystyle{\mathop{\mathbb{E}}_{P(m_l^d | \Phi_l^d)}} \big[ \text{log} P(y | x,\theta,m_l^d, d)  \big] \nonumber \\
-%&= \displaystyle{\mathop{\mathbb{E}}_{g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} Gumbel(0,1) \ }} \big[ \text{log} P(y | x,\theta,\tilde{m}_l^d,d) \big] \nonumber
-%\end{align}
-%where
-%\begin{align}
-%\hspace{-2.em}
-%g_l^d & \in \mathbb{R}^{n_p} \nonumber \\
-%g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} Gumbel(0,1) \nonumber \\
-%& \forall p \in \{ 0,\cdots,n_p\text{-}1 \} \nonumber \\
-%i_1, \cdots, i_k &= Top\_k \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \label{eq:top-k} \\ 
-%&g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \} \nonumber \\
-%\tilde{m}_l^d(p) &= \mathbb{I}(p \in Ind^d). \nonumber 
-%\end{align}
-%and
-%\begin{align}
-%\hspace{-2.em}
-%T_2 &= - KL \big( P(m_l^d | \Phi_l^d) \parallel P(m_l^d | x,\theta,d) \big) \nonumber \\ 
-%	&= \mathbb{H} \big[ P(m_l^d | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\ 
-%	&= \mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\
-%	&\geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \label{eq:entropy}
-%\end{align}
-%
-%we will prove Inequality \eqref{eq:entropy} in Appendix \ref{appendix:b}.
-%
-%Now we go back to our loss function, which is the negative of the lower-bound \eqref{eq:lower-bound}, consisting of 2 terms $-T_1$ and $-T_2$.
-The inequality \eqref{eq:entropy} shows that the upperbound of $-T_2$ is $\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d))$.
-In the training, we will maximize the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ as a regularization over the parameters $\Phi_l^d$ of the variational distribution.
-%
-%Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi_l^d$ into log-likelihood function and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}.
-However, the operator Top-K, which defines $\tilde{m}_l^d$ in Equation \eqref{eq:top-k}, is not differentiable.
-Therefore, we approximate this function by the Soft-Top-K function defined as follows
-%
-%\begin{align}
-%\hat{m}_l^d(\tau) &= \displaystyle{\mathop{argmin}_{\substack{
-%       0 \leqslant m_i \leqslant 1 \ \forall 0 \leqslant i \leqslant n_d\text{-}1 \label{eq:soft-top-k}\\
-%        1^{T}.m = k
-%      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \\
-%& \text{in which} \nonumber \\
-%H_b(m) &= - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i) \nonumber 
-%\end{align}
-%
-%In Appendix \ref{appendix:a}, we will prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
-Furthermore, we will also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and its differentiation is computed via implicit differentiation theorem.
-During the training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyperparameter $\tau$ to $0$.
-The gradient of the term $T_1$ w.r.t $\Phi_l^d$ is computed via the chain rule as follows
-%\begin{equation}
-%\frac{\partial T_1}{\Phi_l^d} = \frac{\partial T_1}{\hat{m}_l^d(\tau)} \times \frac{\hat{m}_l^d(\tau)}{\Phi_l^d}
-%\end{equation}
-%The gradient $\frac{\partial T_1}{\hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\hat{m}_l^d(\tau)}{\Phi_l^d}$ is computed via implicit differentiation, that we will explain carefully in Appendix \ref{appendix:a}.
-%
-%We jointly train the Transformer parameters $\theta$ and the parameters of the latent dropping masks $\Phi_l^d$ using the following multi-task loss.
-%
-%\begin{align}
-%\hspace{-0.em}
-%\mathcal{L}(\theta,\Phi_l^1,\cdots,\Phi_l^k) = \displaystyle{\mathop{\sum}_{d=1}^{k}} \displaystyle{\mathop{\mathbb{E}}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)}} \big[-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)\big]
-%\end{align}
-%
-%in which $\mathcal{D}_s^d$ is distribution of the task $d$ over the input space $\Omega^d_s$; $g^d: \Omega^d_s \rightarrow \Omega^d_t$ is the mapping function, which our multi-task modern needs to learn, of the task $d$; $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$ is the ELBO loss, which we defined in Equation \ref{eq:lower-bound}.
-%
-%Finally, in the inference, we define the dropping mask for $l^{th}$ layer and domain $d$ as follows
-%
-%\begin{align*}
-%Ind_l^d &= Top_k(\Phi_l^d) \\
-%m_l^d &= \mathbb{I}(i\in Ind_l^d)\\
-%\end{align*}
+\fyDone{Can you do distillation ? That is only keep the parameter for one domain and still get good results (better than DA ?)}
 
@@
 \end{align*}
-where $d_k$ is the dimension of the hidden state; $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
-The propagation of information within the network then depends on the current task value and can be formalized as:
+where $d_k$ is the dimension of the hidden state.\fyDone{Check this} The propagation of information within the network then depends on the current task value and can be formalized as:
 \begin{align*}
   \forall l \in [0,\cdots, L-1]: \tilde{h}^l &= h^l \odot r_l^d ,\\
-  h^{l+1} &= \operatorname{LAYER}^{l+1}(\tilde{h}^l) ,\\
+  h^{l+1} &= \operatorname{LAYER}^{l+1}(\tilde{h}^l) ,
 \end{align*}
-where $\operatorname{LAYER}^l()$ represents all the computations in Transformer layer $l$.
+where $\operatorname{LAYER}^l()$ represents all the computations in Transformer layer $l$. It is applied at all positions of each layer in the encoder and in the decoder.  
+% The dropping mask $r_l^d$ is computed as follows
+% where $n_p$ is the number of dropout groups; $n_d$ is the number of domains; $d_k$ is the size of Transformer layers; $k$ is the number of retained groups at each layer; $m_l^d$ is a binary vector indicating which group of the nodes of the $l^{th}$ layer is dropped in the sub-network modeling domain $d$.
 
@@
 Assuming standard notation for our translation model  $P(y|x,d;\theta)$ where $x$, $y$ and $\theta$ respectively refer to the input, output, and parameter vector, the latent variables $m_{l}^d, l \in[0,\dots,L], d \in [0, \dots, n_d-1]$ are introduced as follows.
-We also assume a prior distribution for $m_{l}^d$ chosen to be the uniform distribution over $\Delta^{n_p}_k$: $P(m_l^d | x, d; \theta) = \operatorname{Unif}(\Delta^{n_p}_k)$; variables for each layer are independent and collectively referred to as $m^d$.
-For any (variational) distribution $Q(m^1 \dots m^{n_d}; \Phi)$ with parameters $\Phi$, it is well-known that the log-likelihood is lower-bounded by the so-called $\operatorname{ELBO}$ function, made of a summation of two terms: the \emph{distortion} $D$ and the \emph{rate} $R$ defined as follows:\footnote{We drop the dependency in $x$ to avoid notational clutter.}
+We chose the prior distribution for $m_{l}^d$ as the uniform distribution over $\Delta^{n_p}_k$: $P(m_l^d | x, d; \theta) = \operatorname{Unif}(\Delta^{n_p}_k)$; variables for each layer are independent and collectively refered to as $m^d$.
+For any (variational) distribution $Q(m^1 \dots m^{n_d}; \Phi)$ with parameters $\Phi=\{\phi_l^1, ..., \phi_L^{n_d}\}$, it is well-known that the log-likelihood is lower-bounded by the so-called $\operatorname{ELBO}$ function (hereafter denoted $\ell$), made of a summation of two terms: the \emph{distortion} $D$ and the \emph{rate} $R$ defined as follows:%\footnote{We drop the dependency in $x$ to avoid notational clutter.}
 \begin{align}
-  \log P(y|d;\theta) \ge & \operatorname{ELBO}(y,d; \theta, \Phi) \nonumber \\
-  \operatorname{ELBO}(y,d; \theta, \Phi) =&  D(y,d; \theta, \Phi) - R(y,d; \theta, \Phi) \label{eq:lower-bound}\\
-  D(y,d; \theta, \phi) = & \mathbb{E}_{m^d \sim Q(m^d |d,\Phi)} \log P(y | m^d, d; \theta) \nonumber\\
-  R(y,d; \theta, \phi) = & \operatorname{KL}(Q(m^d | d, \Phi)||P(m^d | d; \theta)), \nonumber
+  \log P(y|x,d;\theta) \ge & \ell(x,y,d; \theta, \Phi) \nonumber \\
+  \ell(x,y,d; \theta, \Phi) =&  D(x,y,d; \theta, \Phi) - R(x,y,d; \theta, \Phi) \label{eq:lower-bound}\\
+  D(x,y,d; \theta, \phi) = & \mathbb{E}_{m^d \sim Q(m^d |d,\Phi)} \log P(y | m^d, x, d; \theta) \nonumber\\
+  R(x,y,d; \theta, \phi) = & \operatorname{KL}(Q(m^d | x,d, \Phi)||P(m^d | x,d; \theta)), \nonumber
 \end{align}
 where $\operatorname{KL}$ is the Kullback-Leibler divergence.
-We use $-\operatorname{ELBO}(x, y\theta,\Phi_l)$ as our surrogate training objective, as a tractable approximation of the intractable likelihood, and try to minimize this function in $\theta$ and $\Phi$.
+We use $-\ell{ELBO}(x,y,d; \theta,\Phi)$ as our surrogate training objective, as a tractable approximation of the untractable likelihood, and try to minimize this function in $\theta$ and $\Phi$.
 
-The variational distribution $Q$ of $m^d$ is defined on layerwise basis, each layer $l$ involving only a subset $\Phi_l^d$ of parameters:
+The variational distribution $Q$ of $m^d$ is defined independently on a layerwise basis; this means that each layer only involves a subset $\Phi_l^d$ of variational parameters.
+$Q$ is computed as follows:
 \begin{align}
@@
   \hspace{-2.em}
-  \forall d,  g^d & =  [g_0^d, \dots, g_L^d],  \text{ with } g_l^d \in \mathbb{R}^{n_p} \nonumber \\
+  \forall d,  g^d & =  [g_1^d, \dots, g_L^d],  \text{ with } g_l^d \in \mathbb{R}^{n_p} \nonumber \\
   \forall p,  g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1) \nonumber \\
-  \{ i_1, \cdots, i_k \} &= \operatorname{Top-k} \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \label{eq:top-k} \\ 
-       & \quad \quad g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \} \nonumber \\
+  \operatorname{Ind}^d = \{ i_1, \cdots, i_k \} &= \operatorname{Top-k} \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \nonumber \\ 
+       & \quad \quad g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \}\label{eq:top-k} \\
 \tilde{m}_l^d(p) &= \mathbb{I}(p \in Ind^d). \nonumber 
 \end{align}
-For the second term, the derivation is the following:
+For the second term, the derivation is the following:\fyDone{Check this: we have a sum over layers ?Also: R or -R ?} 
 \begin{align}
 \hspace{-2.em}
-  R    &= \operatorname{KL}(Q(m^d | d, \Phi)||P(m^d | d; \theta)), \nonumber \\
-  &= \displaystyle{\mathop{\sum}_{l=1}^L} \mathbb{H} \big[ Q(m_l^d | d, \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \nonumber \\ 
-  &= \displaystyle{\mathop{\sum}_{l=1}^L} \mathbb{H}  \big[ Q(i_1,\cdots,i_k | d, \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k)  \nonumber  \\ 
-	&  \geqslant \displaystyle{\mathop{\sum}_{l=1}^L} \mathbb{H} \big[ Q(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \label{eq:entropy}
+  R    &= \operatorname{KL}(Q(m^d | x, d, \Phi)||P(m^d | x, d; \theta)), \nonumber \\ 
+	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(m_l^d | x, d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big) \nonumber \\ 
+	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1,\cdots,i_k | x, d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big)\nonumber \\
+	&\leqslant  -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \big). \label{eq:entropy}
 \end{align}
 We prove inequality~\eqref{eq:entropy} in Appendix~\ref{appendix:b}.
-
-Inequality~\eqref{eq:entropy} shows that a lowerbound of $R$ is $\displaystyle{\mathop{\sum}_{l=1}^L} -\text{log}(\#\Delta^{n_p}_k) + \mathbb{H}(\text{softmax}(\Phi_l^d))$.
-During training, we maximize the entropy $\displaystyle{\mathop{\sum}_{l}^L} \mathbb{H}(\text{softmax}(\Phi_l^d))$ as a regularization over the parameters $\Phi_l^d$ of the variational distribution.
+This inequality shows that an upperbound of $R$ is $\sum_{l=1}^L(\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d)))$.
+During training, we thus maximize a sum over layers of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$\fyDone{Check sum} which performs a regularization over the parameters $\Phi^d$ of the variational distribution.
 
@@
 \hat{m}_l^d(\tau) &= \displaystyle{\mathop{argmin}_{\substack{
-       0 \leqslant m_i \leqslant 1 \ \forall 0 \leqslant i \leqslant n_d\text{-}1 \label{eq:soft-top-k}\\
+                    0 \leqslant m_i \leqslant 1 \\
+  \forall 0 \leqslant i \leqslant n_d\text{-}1 \label{eq:soft-top-k}\\
         1^{T}.m = k
-      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \\
+      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \\ \nonumber
 & \text{in which} \nonumber \\
-H_b(m) &= - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i) \nonumber 
+H_b(m) &= - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i). \nonumber 
 \end{align}
@@
 \hspace{-0.em}
-\ell(\theta,\Phi_l^1,\cdots,\Phi_l^k) = \displaystyle{\mathop{\sum}_{d=1}^{k}} \displaystyle{\mathop{\mathbb{E}}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)}} \big[-\operatorname{ELBO}(\theta,\Phi^d,x,y)\big]
+\mathcal{L}(\theta,\Phi) = \sum_{d=1}^{n_d} \mathbb{E}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)} \big[-\ell(x,y,d; \theta,\Phi)\big]
 \end{align}
 
-in which $\mathcal{D}_s^d$ is distribution of task $d$ over the input space $\Omega^d_s$; $g^d: \Omega^d_s \rightarrow \Omega^d_t$\fyDone{allready defined} is the mapping function for the task $d$, which our multi-task model needs to learn; $-\mathcal{L}_{\text{ELBO}}(\theta,\Phi_l^d,x,y)$ is the ELBO loss, defined in Equation~\ref{eq:lower-bound}.
+in which $\mathcal{D}_s^d$ is distribution of task $d$ over the input space $\Omega^d_s$; $\operatorname{MT}^d: \Omega^d_s \rightarrow \Omega^d_t$\fyDone{already defined} is the translation function for task $d$, which our multi-task model needs to learn; $-\ell(x,y, d, \theta,\Phi)$ is the $\operatorname{ELBO}$ loss, defined in Equation~\ref{eq:lower-bound}.
 
@@
   Ind_l^d &= \operatorname{Top-k}(\Phi_l^d) \\
-  m_l^d &= \mathbb{I}(i\in Ind_l^d),\\
+  m_l^d &= \mathbb{I}(i\in Ind_l^d)
 \end{align*}
-meaning that we simply pick the $k$ most likely parameter groups for the task at hand, and define the dropout mask accordingly.
+meaning that we simply pick the $k$ most likely parameter groups for the task at hand, and define the state dropout mask accordingly.
 
-\section{Experimental settings}
+\section{Experimental settings \label{sec:experiments}}
 \subsection{Baselines}
@@
 \STATE{Pick a batch from domain $d$}
-\STATE{Sampling $\forall p,  g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
+\STATE{Sampling $\forall p,  g_l^d(p) \AND \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
 \IF{$Early\_stopping(Eval\_scores)$}

### 2021-11-15 2022/acl/acl_latex.tex
local changes
revised related work and conclusion

@@
 \section{Introduction}
-Multidomain and multilingual machine translation aim to develop one single model to perform translations for multiple domains and multiple language pairs, respectively.\footnote{We will refer to these two situations as 'multi-task MT' (or \mtsquare{} for short), and refer to individual domains and languages as 'tasks'.} These paradigms are motivated by the compactness of the resulting translation system \citep{Chu18multilingual,dabre20survey}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
+Multidomain and multilingual machine translation aim to develop one single model to perform translations for multiple domains and multiple language pairs, respectively.\footnote{We will refer to these two situations as 'multi-task MT' and refer to individual domains and languages as 'tasks'.} These paradigms are motivated by the compactness of the resulting translation system \citep{Chu18multilingual,dabre20survey}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
 However, having all the tasks use exactly the same model parameters can cause negative interference between unrelated tasks \citep{conneau20unsupervised,wang20negative}.
-Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using adapter layers as studied in \citep{houlsby19parameter,Bapna19simple,Pham20Study}.
+Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using adapter layers as studied in \citep{houlsby19parameter,Bapna19simple,Pham20Study,Philip20monolingual}.
 If these techniques have proven effective for building strong baselines, they fail to fully take advantage of the similarities that exist between domains and tasks, as documented eg.\ in \citep{Pham21revisiting}.
@@
   \setlength{\parsep}{0pt}
-\item We provide a sound mathematical formulation to the problem of jointly learning task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling approach;
+\item We provide a sound mathematical formulation to the problem of jointly learning task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling techniques;
 \item We present algorithms to train this model end to end with very little extra computational cost.
@@
 
-For multilingual case TED-related, we have four language families including (1) Turkic family with Azerbaijani and Turkish(\domain{aze},\domain{tur}); (2) Slavic family with Belarusian and Russian (\domain{bel},\domain{rus}); (3) Romance family with Glacian and Portuguese (\domain{glg},\domain{por}); and (4) Czech-Slovak family with Slovak and Czech (\domain{ces},\domain{slk}).
-The similarities between the sub-networks of each pair are 0.95, 0.94, 0.99 and 0.96 which are the maximum in each row except the diagonal, which stands for the self-similarity.
+For multilingual (TED-related) case, the data contain four language families including (1) Turkic family with Azerbaijani and Turkish(\domain{aze},\domain{tur}); (2) Slavic family with Belarusian and Russian (\domain{bel},\domain{rus}); (3) Romance family with Galician and Portuguese (\domain{glg}, \domain{por}); and (4) Czech-Slovak family with Slovak and Czech (\domain{ces}, \domain{slk}).
+The similarities between the sub-networks for each pair are respectively 0.95, 0.94, 0.99 and 0.96: for each language, the maximally similar one (excluding the diagonoal) is always from the same language family.
 \begin{table*}[h!]
@@
 
-We also plot the languages based on their dropouts in Figure\ref{fig:visual}.
+We also plot the languages based on their dropout masks in Figure~\ref{fig:visual} using a 2d projection computed by Principal Component Analysis (PCA).
 \begin{figure*}[htbp]
@@
 \end{subfigure}
-\caption{Visualization of languages according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
+\caption{Visualization of languages according to their dropout masks (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
 \label{fig:visual}
@@
 \includegraphics[width=0.7\textwidth]{multi_domain}
-\caption{Visualization of domains according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
+\caption{Visualization of domains according to their dropout masks (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
 \label{fig:domain}
@@
 
-\section{Related work}
-Multidomain, multilingual translation
+\section{Related work}\label{sec:related}
+Multidomain and multilingual translation systems have received considerable attention in the recent years, and a exhaustive survey is beyond the goal of this paper.
+Domain adaptation for neural MT is surveyed in \citep{Chu17comparison}, while multidomain MT systems are notably studied in \cite{Saunders21Asurvey,Pham21revisiting}; for multilingual MT, the reader is refered eg.\ to \citep{Chu18multilingual,dabre20survey}.
+We focus on the most relevant subset of this literature below.
 
-\citep{sen19multilingual,kong21multilingual} used the similar idea of language proximity-based parameter sharing.
-The authors proposed a multi-decoder model sharing the same encoder among languages and routed languages in different families to different decoders.
+\textbf{Language similarity} The methods developed by \citep{sen19multilingual,kong21multilingual} use language proximity to design parameter sharing strategies.
+The authors propose a multi-decoder model sharing the same encoder among languages and routing languages in different families to different decoders.
 
-\textbf{Sparse Transformer}\citet{xian20deep} proposed using different depths for different tasks.
+\textbf{Sparse Transformer} The idea of adaptive sparsity is studied in several works.
+For instance, \citet{xian20deep} propose to use a variable depth for different tasks.
 The authors aimed to match the depth of the sub-network to the complexity of the task.
-\citet{Gong21pay,Gong21adaptive} also take an interest in the adaptive sparse Transformers, which differ the selection of heads in multi-head attention, layers, and blocks in feedforward matrices between different tasks.
-Mixture-of-experts (MoE) is another effective approach to sparse modeling.
-Within the transformer model, the GShard model replaces a single feedforward (FFN) sub-layer with an MoE module consisting of multiple FFN sub-layers \citep{lepikhin21gshard}.
-\citet{william21switch}
+\citet{Gong21pay,Gong21adaptive} also take an interest in the adaptive sparse Transformers, in which differ each task triggers the selection of specific heads in multi-head attention, layers, and blocks in feedforward matrices.
+Mixture-of-experts (MoE) constitute another effective approach to achieve sparsity.
+Using the Transformer architecture, the GShard model replaces a single feedforward (FFN) sub-layer with an MoE module consisting of multiple FFN sub-layers \citep{lepikhin21gshard,william21switch} .
 
-\textbf{Adapters} have proven to be very efficient for NLP multi-task modeling \citep{houlsby19parameter,Bapna19simple,Pham20Study,pfeiffer20adapterhub}.
+\textbf{Adapter modules} Adapters have proven to be very efficient for NLP multi-task modeling \citep{houlsby19parameter,Bapna19simple,Pham20Study,pfeiffer20adapterhub}.
 The method consists of plugging several adapters to the intermediate layers of a pretrained Transformer and finetuning these adapters on the downstream tasks.
 Adapters can also be trained unsupervised in multilingual translation \citep{Philip20monolingual}.
-However, the hard-coded separation between the domains of different tasks leads to a catastrophic forgetting \citep{pfeiffer21adapterfusion}, which is a common problem in multi-task modeling using neural networks \cite{Michael89catastrophic}.
-In multi-domain translation, recently \citet{Pham21revisiting} demonstrated the brittleness of domain adapters against fuzzy domain separations, out-of-domain distributions, and erroneous domain tags.
-Recently, several works have aimed to mitigate this weakness through a mixture of expert mechanisms, for example,\cite{pfeiffer21adapterfusion}.
+However, the hard-coded separation between the domains of different tasks may lead to a catastrophic forgetting effect \citep{pfeiffer21adapterfusion}, which is a common problem in multi-task modeling using neural networks \cite{Michael89catastrophic}.
+In multi-domain translation, \citet{Pham21revisiting} recently demonstrated the brittleness of domain adapters against fuzzy domain separations, out-of-domain distributions, and erroneous domain tags.
+Several subsequent studies have aimed to mitigate this weakness through a mixture of expert mechanisms, for example,\cite{pfeiffer21adapterfusion}.
 
@@
 
-\section{Conclusions and outlook}
-In this work, we present a novel method allows jointly searching optimal sub-networks for multiple tasks and the parameters of the underlying network.
-Our method proposed a rigorous mathematical modeling and an end-to-end optimization with very little extra computational cost.
-We demonstrate an improvement of our model with large margin over multi-domain Transformer without zero additional parameters and an equivalent performance compared to a strong baseline such as Domain Adapters consisting of large number of extra parameters.
-In multilingual translation, our model outperforms multilingual Transformer and Language Adapters in 3/4 cases.
-Besides, we provided an thorough ablation study on the similarities between learned sub-networks and demonstrated a strong correlation between the sub-networks' similarity and the proximity of their tasks (domain or language).
+\section{Conclusions and outlook}\label{sec:conclusion}
+In this work, we present a novel method for multdomain and multilingual translation.
+It allows us to jointly search for an optimal assignment of sub-networks to tasks and learn the parameters of the underlying network.
+Our method relies on a sound mathematical model and an end-to-end optimization with very little extra computational cost.
+Experimentally, we achieve a large improvement over Transformer baseline without zero additional parameters; our performance are also comparable to that of a strong a multi-task baseline using residual adapter modules which however includes a large number of extra parameters.
+For multilingual translation, our model outperforms multilingual Transformer and Language Adapters in 3 our of 4 cases.
+Besides, we provided an thorough analysis of the similarities between learned sub-networks and demonstrated a strong correlation between the sub-networks' similarity and the proximity of the corresponding tasks (domain or language).
 

### 2021-11-15 2022/acl/acl_latex.tex
revised the conclusion

@@
 
-\textbf{Adapter modules} Adapters have proven to be very efficient for NLP multi-task modeling \citep{houlsby19parameter,Bapna19simple,Pham20Study,pfeiffer20adapterhub}.
-The method consists of plugging several adapters to the intermediate layers of a pretrained Transformer and finetuning these adapters on the downstream tasks.
-Adapters can also be trained unsupervised in multilingual translation \citep{Philip20monolingual}.
-However, the hard-coded separation between the domains of different tasks may lead to a catastrophic forgetting effect \citep{pfeiffer21adapterfusion}, which is a common problem in multi-task modeling using neural networks \cite{Michael89catastrophic}.
+\textbf{Adapter modules} Adapters have proven to be very efficient for multi-task NLP \citep{houlsby19parameter,Bapna19simple,Pham20Study,pfeiffer20adapterhub}.
+In a nutshell, this technique consists in plugging several so-called adapter modules to the intermediate layers of a pretrained Transformer and finetuning these adapters on the downstream tasks.
+Adapters can also be trained without supervision for multilingual translation \citep{Philip20monolingual}.
+However, the hard-coded separation between the domains of different tasks may lead to a catastrophic forgetting effect \citep{pfeiffer21adapterfusion}, which is a common problem in multi-task modeling using neural networks \citep{Michael89catastrophic}.
 In multi-domain translation, \citet{Pham21revisiting} recently demonstrated the brittleness of domain adapters against fuzzy domain separations, out-of-domain distributions, and erroneous domain tags.
-Several subsequent studies have aimed to mitigate this weakness through a mixture of expert mechanisms, for example,\cite{pfeiffer21adapterfusion}.
+Several subsequent studies have aimed to mitigate this weakness through a mixture of expert mechanism, for example,\cite{pfeiffer21adapterfusion}.
 
 Recently, \citet{biao21share} proposed jointly learning how to route between shared representation and language-specific representation via a conditional language-specific routing and the parameters of the underlying Transformer.
-This method is very related to the FusionAdapters of \cite{pfeiffer21adapterfusion}.
-Both approaches assume the final representation is a combination or a choice between shared representation and task-specific representations.
+This method is closely related to the FusionAdapters of \citet{pfeiffer21adapterfusion}.
+Both approaches assume that the final representation\fyTodo{??} is a combination or a choice between shared representation and task-specific representations.
 The proximity between tasks is not taken into account in the routing mechanism.
-Our method proposed a different point of view on the problem of multi-task routing in the underlying network.
+Our method proposed a different approach to the problem of multi-task routing in the underlying network.
 
@@
 
-Despite several advantages, our method has several weaknesses including the heuristically predefined number of retained groups, $k$ , and the need of domain label in the inference.
-In future work, we would like to provide an extensive variational framework to model both the number of retained groups, $k$ and the dropping selections.
-Secondly, to get rid of the requirement of domain information in the inference, we could condition the dropping selection directly on the input $x$ instead of the domain $d$.
-These two open questions are both very interesting and promising.
+There are several ways in which our methodology could be improved and generalized.
+In future work, we would first like to provide an complete variational framework to model both the number of groups, $k$ and the selection of the dropout masks.
+Second, we also intend to get rid of the requirement of knowing the domain information during inference: this would mean replacing the dependency on $d$ in the variational distribution by a dependency in the input $x$.
+These two questions seem both very interesting and promising; they would allow to us to dispense with heuristic choices in the architecture, and replace them with an increase dependency on the training data.
+
 \bibliography{bibliography}

### 2021-11-15 2022/acl/acl_latex.tex

--- 2021-11-15 2022/acl/acl_latex.tex
@@
 We prove inequality~\eqref{eq:entropy} in Appendix~\ref{appendix:b}.
-This inequality shows that an upperbound of $R$ is $\sum_{l=1}^L(\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d)))$.
+This inequality shows that an upperbound of $R$ is $\sum_{l=1}^L(\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d)))$ since $i_1 | \Phi_l^d \sim \operatorname{softmax}(\Phi_l^d)$.
 During training, we thus maximize a sum over layers of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$\fyDone{Check sum} which performs a regularization over the parameters $\Phi^d$ of the variational distribution.
@@
                     0 \leqslant m_i \leqslant 1 \\
-  \forall 0 \leqslant i \leqslant n_d\text{-}1 \label{eq:soft-top-k}\\
+  \forall 0 \leqslant i \leqslant n_d\text{-}1 \\
         1^{T}.m = k
-      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \\ \nonumber
-& \text{in which} \nonumber \\
+      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \label{eq:soft-top-k}
+\end{align}
+in which
+\begin{align}
 H_b(m) &= - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i). \nonumber 
@@
 \subsection{Baselines}
-\subsubsection{Multi-domain, multilingual translation}
-In both cases, we examine the performance of three following MT systems.
+\subsubsection{Multi-domain translation}
+We describe the choice of our systems in multi-domain experiments as follows
 \begin{itemize}
-	\item Transformer
-	\item Adapter based Transformer
-	\item \system{LaMGD} Tranformer
+	\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048, the multi-head attentions contain with 8 heads.
+	\item Adapter based Transformer: The intermediate feeforward dimension is set as 2048 as in \citet{Pham21revisiting}.
+	\item \system{LaMGD} Tranformer: There is no change in the architecture.
+We group 512 nodes in each layer into 16 groups of 32 consecutive nodes.
+For each domain, only 12 groups are selected.
+The number of parameters of variational distributions are $k \times L \times n_d$ which is tiny compared to the size of Transformer.
 \end{itemize}
-For the sake of brevity, we put the detailed descriptions of these systems in Appendix \ref{appendix:c}.
+
+We set the dropout probability as 0.1.
+We train the multi-domain Transformer model in 200~K iterations with a batchsize of 12~K tokens.
+We train \system{LaMGD} Tranformer in 300~K iterations with a batchsize of 12~K tokens since \system{LaMGD} Tranformer has to learn both the sub-networks and the parameters of the underlying Transformer.
+Finally, we plug Domain Adapters to the multi-domain Transformer model and finetune them with 25~K iterations using the same batchsize as the former.
+\subsubsection{Multilingual translation}
+We describe the choice of our systems in multilingual experiments as follows
+\begin{itemize}
+	\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 1024, the multi-head attentions contain with 8 heads as in \citet{Wang20balancing}.
+	\item Adapter based Transformer: The intermediate feeforward dimension is set as 128 as in \citet{Gong21adaptive}.
+	\item \system{LaMGD} Tranformer: There is no change in the architecture.
+We group 512 nodes in each layer into 16 groups We group 512 nodes in each layer into 16 groups of 32 consecutive nodes.
+For each domain, only 12 groups are selectedof 32 consecutive nodes.
+For each domain, only 12 groups are selected.
+\end{itemize}
+
+We set the dropout probability as 0.3.
+We train the multilingual Transformer model in 40~K iterations with a batchsize of 9600 tokens on 16 GPUs as in \citet{Gong21adaptive}.
+We train \system{LaMGD} Tranformer in 50~K iterations with the same batchsize.
+Finally, we finetune the Language Adapters in 5~K iterations.
+\subsubsection{Hyperparameters}
+\label{ssec:hyperparams}
+We choose $n_d = 16$ so that the dropout groups' size is not too small or too large.
+In our future works, we will do more experiments by changing $n_d$ to examine how this hyperparameters affect the performance of the model.
+The second hyperparameter in \system{LaMGD} is the number of retained groups in each layer, $k$, which we predetermined as 12 in every experiments.
+By retaining $12/16$ groups, we share $75 \%$ active groups between two domains or languages in average.
+
+Furthermore, the temperature $\tau$ of our $\operatorname{Soft-Top-K}$ operator will be gradually decreased from $0.5$ to $0.2$ as follows
+\begin{align*}
+\tau = \operatorname{min}\{ 0.2, 0.5 * \exp^{-r*step} \}
+\end{align*}
+in which $r=0.0001$.
+
+Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments.
+The choice of the values for our hyperparameters was not optimized and selected by suggestions from previous papers \citep{Jang17categorical,Gong21pay,Gong21adaptive}.
 \subsection{Datasets}
 \subsubsection{Multi-domain translation}
-We reuse the same data as the recent work of \citet{Pham21revisiting} in multi-domain translation. For the sake of brevity, we leave most of the details of data in Appendix \ref{appendix:c}.
+We reuse the same data as the recent work of \citet{Pham21revisiting} in multi-domain translation. We describe our datasets for multi-domain translation experiment in Table \ref{tab:Corpora-chap4}
+\begin{table*}[h!]
+  \centering
+  \begin{tabular}{|l|cccccc|} %*{4}{|r|}}
+    \cline{2-7} 
+    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} \\
+    \hline 
+    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) \\
+    \# \revisiondone{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 \\
+    \# \revisiondone{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 \\
+    \# \revisiondone{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 \\
+    \hline
+  \end{tabular}
+  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$).
+\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
+  }
+\label{tab:Corpora-chap4}
+\end{table*}
 \subsubsection{Multilingual translation}
@@
 \end{itemize}
-The detailed description of the datasets is provided in Appendix \ref{appendix:c}.
+Table \ref{tab:Corpora-multilingual} describes the subset of TED8 that we used in our experiments.
+\begin{table*}[h!]
+  \centering
+  \begin{tabular}{l|cccc} 
+  \hline
+    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c}{\domain{test}} \\
+    \hline 
+    \multirow{8}{*}{Related} & Azerbaijani & 5.94k & 671 & 903 \\
+							 & Belarusian & 4.51k & 248 & 664 \\
+							 & Glacian & 10.0k & 682 & 1007 \\
+							 & Slovak & 61.5k & 2271 & 2445 \\
+							 & Turkish & 182k & 4045 & 5029 \\
+                             & Russian & 208k & 4814 & 5483 \\
+                             & Portuguese & 185k & 4035 & 4855 \\
+                             & Czech & 103k & 3462 & 3831 \\
+    \hline
+    \multirow{8}{*}{Related} & Bosnian & 5.64k & 474 & 463 \\
+                             & Marathi & 9.84k & 767 & 1090 \\
+                             & Hindi & 18.79k & 854 & 1243 \\
+                             & Macedonian & 25.33k & 640 & 438 \\
+                             & Greek & 134k & 3344 & 4433 \\
+                             & Bulgarian & 174k & 4082 & 5060 \\
+                             & French & 192k & 4320 & 4866 \\
+                             & Korean & 205k & 4441 & 5637 \\
+  \end{tabular}
+  \caption{Data Statistics of TED8 Datasets}
+\label{tab:Corpora-multilingual}
+\end{table*}
 \section{Results and ablation studies}
@@
     O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.80 &7.30&20.80&21.10&39.70&19.80&22.60&15.20&18.91 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &4.30&6.80&21.10&22.00&39.70&20.00&23.00&15.20&19.01 \\ 
-    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]}  & 5.20&\SB{9.40}&20.60&\SB{22.80}&39.60&19.60&22.40&15.00&19.33 \\ 
+    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.8 &7.3&20.8&21.1&39.7&19.8&22.6&15.2&18.9 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &4.3&6.8&21.1&22&39.7&20&23&15.2&19 \\ 
+    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.33 \\ 
 	\hline
@@
     M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.40&16.60&28.50&	27.10&43.70&24.60&30.30&25.60&25.98 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &10.10&15.80&28.40&26.80&43.70&24.50&30.20&25.60&25.64\\ 
-    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  &11.30&\SB{17.40}&28.60&\SB{28.70}&43.70&24.50&30.70&25.60&26.31 \\ 
+    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.4&16.6&28.5&	27.1&43.7&24.6&30.3&25.6&25.98 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.64\\ 
+    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.31 \\ 
     \hline
@@
     O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.20&4.0&12.70&22.20&31.80&\SB{34.00}&38.30&8.30&20.19 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &10.20&4.00&\SB{13.30}&21.90&32.20&\SB{34.10}&38.50&8.30&20.31 \\ 
-    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}&10.10&3.80&12.60&\SB{22.80}&31.80&33.40&38.10&8.10&20.09\\
+    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&\SB{34}&38.3&8.3&20.19 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &10.2&4&\SB{13.3}&21.9&32.2&\SB{34.1}&38.5&8.3&20.31 \\ 
+    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
     \hline 
@@
     M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.40&9.70&20.50&31.80&37.50&38.70&39.80&19.00&27.43 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.50&9.40&20.00&30.60&37.20&38.20&39.30&19.00&27.03\\ 
-    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]} &\SB{23.50}&9.60&\SB{21.50}&32.20&37.70&38.60&40.00&18.90&27.75 \\
+    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19&27.43 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.5&9.4&20&30.6&37.2&38.2&39.3&19&27.03\\ 
+    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40&18.9&27.75 \\
     \hline
@@
 Furthermore, we remark a large distance between \domain{rel} and the rest of the domains, which correlates to the fact that \domain{rel} has the most distant topic to other domains.
+\domain{rel} only share around $75 \%$ its active groups with other domains whereas $75 \%$ is the average of sharing proportions by default (see Section \ref{ssec:hyperparams}).
 In Figure ~\ref{fig:domain}, we visualize the domains using their dropping masks concatenated and mapped to 2D space by PCA.
-
 \begin{table*}[h!]
@@
 We also plot the languages based on their dropouts in Figure\ref{fig:visual}.
-\begin{figure*}[htbp]
-\begin{subfigure}{\textwidth}
+\begin{figure*}[h!]
+\begin{subfigure}{0.5\textwidth}
   \centering
   % include first image
-  \includegraphics[width=0.7\textwidth]{ted_related}  
+  \includegraphics[width=0.9\textwidth]{ted_related}  
   \caption{TED-Related}
@@
 \end{subfigure}
-\\
-\begin{subfigure}{\textwidth}
+\begin{subfigure}{0.5\textwidth}
   \centering
   % include second image
-  \includegraphics[width=0.7\textwidth]{ted_diverse}  
+  \includegraphics[width=0.85\textwidth]{ted_diverse}  
   \caption{TED-Diverse}
@@
 \end{figure*}
-
-\begin{figure*}[htbp]
-\includegraphics[width=0.7\textwidth]{multi_domain}
-\caption{Visualization of domains according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
-\label{fig:domain}
+\begin{figure*}[h!]
+\begin{subfigure}{0.5\textwidth}
+  \centering
+  % include first image
+  \includegraphics[width=0.9\textwidth]{phi_ted_related_o2m}  
+  \caption{TED-Related}
+  \label{fig:phi-related}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=0.85\textwidth]{phi_ted_diverse_o2m}  
+  \caption{TED-Diverse}
+  \label{fig:phi-diverse}
+\end{subfigure}
+\caption{$\Phi_L^{0}$ in O2M experiments}
+\label{fig:phi}
 \end{figure*}
 
+\subsection{$\Phi_l^d$'s learning curve}
+In this section, we analyze the convergence of $\Phi_l^d$ during the training.
+For sake of brevity, we only plot in Figure ~\ref{fig:phi} the learning curve of $\Phi_{L}^0 \in \mathbb{R}^{12}$ (as $n_p=12$ in our setting) of the last layer (in decoder) and the first language \domain{aze}/\domain{bos} in the TED-related/diverse O2M experiment.
+We remark a fast convergence of 4 groups to "dropping" zone  which is a low region of negative half plane.
+There are three groups having changing movement when dropping to the negative half plane in the beginning but eventually approaching the positive half plane.
+
 \section{Related work}
-Multidomain, multilingual translation
+In \textbf{Multilingual translation},\citep{sen19multilingual,kong21multilingual} used the idea of language proximity-based parameter sharing.
+The authors proposed a multi-decoder model sharing the same encoder among languages and routed languages in different families to different decoders.
+These approaches share the same interest in expressing the proximity between tasks in the selection of task-specific parameters as our approach.
+However, our method learn the selection from a latent commonality in data instead of using a predefined selection such as "One language family per decoder" in \cite{kong21multilingual}.
 
-\citep{sen19multilingual,kong21multilingual} used the similar idea of language proximity-based parameter sharing.
-The authors proposed a multi-decoder model sharing the same encoder among languages and routed languages in different families to different decoders.
-
-\textbf{Sparse Transformer}\citet{xian20deep} proposed using different depths for different tasks.
+\textbf{Sparse Transformers} are another approaches for task-specific parameters selections.\citet{xian20deep} proposed using different depths for different tasks.
 The authors aimed to match the depth of the sub-network to the complexity of the task.
@@
 Within the transformer model, the GShard model replaces a single feedforward (FFN) sub-layer with an MoE module consisting of multiple FFN sub-layers \citep{lepikhin21gshard}.
-\citet{william21switch}
+\citet{william21switch} proposed a more efficient gating function to combine these FFN sub-layers.
 
@@
 In multi-domain translation, recently \citet{Pham21revisiting} demonstrated the brittleness of domain adapters against fuzzy domain separations, out-of-domain distributions, and erroneous domain tags.
-Recently, several works have aimed to mitigate this weakness through a mixture of expert mechanisms, for example,\cite{pfeiffer21adapterfusion}.
+Recently, several works have aimed to mitigate this weakness through a mixture of expert mechanisms, for example  \cite{pfeiffer21adapterfusion}.
 
@@
 \section{Appendix C}
-\label{appendix:c}
-\subsection{Multi-domain datasets}
-We describe our datasets for multi-domain translation experiment in Table \ref{tab:Corpora-chap4}
-\begin{table*}
-  \centering
-  \begin{tabular}{|l|cccccc|} %*{4}{|r|}}
-    \cline{2-7} 
-    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} \\
-    \hline 
-    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) \\
-    \# \revisiondone{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 \\
-    \# \revisiondone{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 \\
-    \# \revisiondone{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 \\
-    \hline
-  \end{tabular}
-  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$).
-\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
-  }
-\label{tab:Corpora-chap4}
-\end{table*}
-\subsection{Multilingual datasets}
-Table \ref{tab:Corpora-multilingual} describes the subset of TED8 that we used in our experiments.
-\begin{table*}
-  \centering
-  \begin{tabular}{l|cccc} 
-  \hline
-    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c}{\domain{test}} \\
-    \hline 
-    \multirow{8}{*}{Related} & Azerbaijani & 5.94k & 671 & 903 \\
-							 & Belarusian & 4.51k & 248 & 664 \\
-							 & Glacian & 10.0k & 682 & 1007 \\
-							 & Slovak & 61.5k & 2271 & 2445 \\
-							 & Turkish & 182k & 4045 & 5029 \\
-                             & Russian & 208k & 4814 & 5483 \\
-                             & Portuguese & 185k & 4035 & 4855 \\
-                             & Czech & 103k & 3462 & 3831 \\
-    \hline
-    \multirow{8}{*}{Related} & Bosnian & 5.64k & 474 & 463 \\
-                             & Marathi & 9.84k & 767 & 1090 \\
-                             & Hindi & 18.79k & 854 & 1243 \\
-                             & Macedonian & 25.33k & 640 & 438 \\
-                             & Greek & 134k & 3344 & 4433 \\
-                             & Bulgarian & 174k & 4082 & 5060 \\
-                             & French & 192k & 4320 & 4866 \\
-                             & Korean & 205k & 4441 & 5637 \\
-  \end{tabular}
-  \caption{Data Statistics of TED8 Datasets}
-\label{tab:Corpora-multilingual}
-\end{table*}
+\begin{figure*}
+\includegraphics[width=0.7\textwidth]{multi_domain}
+\caption{Visualization of domains according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
+\label{fig:domain}
+\end{figure*}
 
-\subsection{Multi-domain systems}
-We describe the choice of our systems in multi-domain experiments as follows
-\begin{itemize}
-	\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048, the multi-head attentions contain with 8 heads.
-	\item Adapter based Transformer: The intermediate feeforward dimension is set as 2048 as in \citet{Pham21revisiting}.
-	\item \system{LaMGD} Tranformer: There is no change in the architecture.
-We group 512 nodes in each layer into 16 groups of 32 consecutive nodes.
-For each domain, only 12 groups are selected.
-The number of parameters of variational distributions are $k \times L \times n_d$ which is tiny compared to the size of Transformer.
-\end{itemize}
-
-We set the dropout probability as 0.1.
-We train the multi-domain Transformer model in 200~K iterations with a batchsize of 12~K tokens.
-We train \system{LaMGD} Tranformer in 300~K iterations with a batchsize of 12~K tokens since \system{LaMGD} Tranformer has to learn both the sub-networks and the parameters of the underlying Transformer.
-Finally, we plug Domain Adapters to the multi-domain Transformer model and finetune them with 25~K iterations using the same batchsize as the former.
-\subsection{Multilingual systems}
-We describe the choice of our systems in multilingual experiments as follows
-\begin{itemize}
-	\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 1024, the multi-head attentions contain with 8 heads as in \citet{Wang20balancing}.
-	\item Adapter based Transformer: The intermediate feeforward dimension is set as 128 as in \citet{Gong21adaptive}.
-	\item \system{LaMGD} Tranformer: There is no change in the architecture.
-We group 512 nodes in each layer into 16 groups We group 512 nodes in each layer into 16 groups of 32 consecutive nodes.
-For each domain, only 12 groups are selectedof 32 consecutive nodes.
-For each domain, only 12 groups are selected.
-\end{itemize}
-
-We set the dropout probability as 0.3.
-We train the multilingual Transformer model in 40~K iterations with a batchsize of 9600 tokens on 16 GPUs as in \citet{Gong21adaptive}.
-We train \system{LaMGD} Tranformer in 50~K iterations with the same batchsize.
-Finally, we finetune the Language Adapters in 5~K iterations.
-
-\section{\system{LaMGD}'s training algorithm}
-
+\section{Appendix D}
 \begin{breakablealgorithm}
@@
 \STATE{Pick a batch from domain $d$}
-\STATE{Sampling $\forall p,  g_l^d(p) \AND \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
+\STATE{Sampling $\forall p,  g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
 \IF{$Early\_stopping(Eval\_scores)$}

### 2021-11-15 2022/acl/acl_latex.tex

--- 2021-11-15 2022/acl/acl_latex.tex
@@
 \begin{figure*}[htbp]
-\begin{subfigure}{\textwidth}
+\begin{subfigure}{0.5\textwidth}
   \centering
   % include first image
-  \includegraphics[width=0.7\textwidth]{ted_related}  
+  \includegraphics[width=0.85\textwidth]{ted_related}  
   \caption{TED-Related}
@@
 
-\begin{figure*}[htbp]
-\includegraphics[width=0.7\textwidth]{multi_domain}
-\caption{Visualization of domains according to their dropout masks (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
-\label{fig:domain}
+\begin{figure*}[h!]
+\begin{subfigure}{0.5\textwidth}
+  \centering
+  % include first image
+  \includegraphics[width=0.9\textwidth]{phi_ted_related_o2m}  
+  \caption{TED-Related}
+  \label{fig:phi-related}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=0.85\textwidth]{phi_ted_diverse_o2m}  
+  \caption{TED-Diverse}
+  \label{fig:phi-diverse}
+\end{subfigure}
+\caption{$\Phi_L^{0}$ in O2M experiments}
+\label{fig:phi}
 \end{figure*}
@@
 The authors propose a multi-decoder model sharing the same encoder among languages and routing languages in different families to different decoders.
+These approaches share the same interest in expressing the proximity between tasks in the selection of task-specific parameters as our approach.
+However, our method learn the selection from a latent commonality in data instead of using a predefined selection such as "One language family per decoder" in \cite{kong21multilingual}.
 
@@
 Mixture-of-experts (MoE) constitute another effective approach to achieve sparsity.
-Using the Transformer architecture, the GShard model replaces a single feedforward (FFN) sub-layer with an MoE module consisting of multiple FFN sub-layers \citep{lepikhin21gshard,william21switch} .
+Using the Transformer architecture, the GShard model replaces a single feedforward (FFN) sub-layer with an MoE module consisting of multiple FFN sub-layers \citep{lepikhin21gshard,william21switch}.
 

### 2021-11-15 2022/acl/acl_latex.tex
read section 3, fix the duplicate in bib

@@
 \section{Experimental settings \label{sec:experiments}}
-\subsection{Baselines}
-\subsubsection{Multi-domain translation}
-We describe the choice of our systems in multi-domain experiments as follows
+
+\subsection{System design and configuration}
+\subsubsection{Multidomain translation systems}
+Our systems for the multidomain experiments are designed as follows:
 \begin{itemize}
-	\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048, the multi-head attentions contain with 8 heads.
-	\item Adapter based Transformer: The intermediate feeforward dimension is set as 2048 as in \citet{Pham21revisiting}.
-	\item \system{LaMGD} Tranformer: There is no change in the architecture.
-We group 512 nodes in each layer into 16 groups of 32 consecutive nodes.
-For each domain, only 12 groups are selected.
-The number of parameters of variational distributions are $k \times L \times n_d$ which is tiny compared to the size of Transformer.
+\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048; the multi-head attention mechanism contains 8 heads;
+\item Adapter-based Transformer: The intermediate feedforward dimension is set to 2048;% as in \citet{Pham21revisiting}.
+\item \system{LaMGD} Transformer:\fyTodo{Explain acronym} There is no change in the architecture.
+We group the 512 nodes in each layer into 16 groups of 32 consecutive nodes.
+For each domain, only 12 out of the 16 groups are selected.
+The number of parameters of the variational distributions is $k \times L \times n_d$, which is negligible when compared to the size of the Transformer model;
 \end{itemize}
@@
 We set the dropout probability as 0.1.
-We train the multi-domain Transformer model in 200~K iterations with a batchsize of 12~K tokens.
-We train \system{LaMGD} Tranformer in 300~K iterations with a batchsize of 12~K tokens since \system{LaMGD} Tranformer has to learn both the sub-networks and the parameters of the underlying Transformer.
-Finally, we plug Domain Adapters to the multi-domain Transformer model and finetune them with 25~K iterations using the same batchsize as the former.
-\subsubsection{Multilingual translation}
-We describe the choice of our systems in multilingual experiments as follows
+We train the multi-domain Transformer model for 20k iterations with a batch size of 12k tokens.
+We train \system{LaMGD} Tranformer for 300k iterations with a batch size of 12k tokens: these extra iterations are justified by the need for the \system{LaMGD} Tranformer to learn both the sub-networks and the parameters of the underlying Transformer.
+Finally, we plug domain Adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline.
+
+\subsubsection{Multilingual translation systems}
+The systems used in our multilingual experiments are implemented as follows:
 \begin{itemize}
-	\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 1024, the multi-head attentions contain with 8 heads as in \citet{Wang20balancing}.
-	\item Adapter based Transformer: The intermediate feeforward dimension is set as 128 as in \citet{Gong21adaptive}.
-	\item \system{LaMGD} Tranformer: There is no change in the architecture.
-We group 512 nodes in each layer into 16 groups We group 512 nodes in each layer into 16 groups of 32 consecutive nodes.
-For each domain, only 12 groups are selectedof 32 consecutive nodes.
-For each domain, only 12 groups are selected.
+\item Transformer models: the embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 1024, each multi-head attentions contains 8~heads as in \citep{Wang20balancing}.
+\item Adapter based Transformer: the intermediate feedforward dimension is set as 128 as in \citep{Gong21adaptive}.
+\item \system{LaMGD} Tranformer: There is no change in the architecture. We group 512 nodes in each layer into 16 groups of 32 consecutive nodes. For each language, only 12 groups are selected. 
 \end{itemize}
@@
 We set the dropout probability as 0.3.
-We train the multilingual Transformer model in 40~K iterations with a batchsize of 9600 tokens on 16 GPUs as in \citet{Gong21adaptive}.
-We train \system{LaMGD} Tranformer in 50~K iterations with the same batchsize.
-Finally, we finetune the Language Adapters in 5~K iterations.
+We train the multilingual Transformer model in 40~K iterations with a batch size of 9600 tokens on 16 GPUs as in \citet{Gong21adaptive}.
+We train \system{LaMGD} Tranformer for 50k iterations with the same batch size.
+Finally, we finetune the language-specific Adapters for 5k iterations.
+
+All the translation systems are implemented OpenNMT\fyTodo{Fix this, add reference to framework if needed}.
+
 \subsubsection{Hyperparameters}
 \label{ssec:hyperparams}
-We choose $n_d = 16$ so that the dropout groups' size is not too small or too large.
-In our future works, we will do more experiments by changing $n_d$ to examine how this hyperparameters affect the performance of the model.
-The second hyperparameter in \system{LaMGD} is the number of retained groups in each layer, $k$, which we predetermined as 12 in every experiments.
-By retaining $12/16$ groups, we share $75 \%$ active groups between two domains or languages in average.
+We choose $n_d = 16$ so that the size of the dropout group is neither too small nor too large.
+The second important hyperparameter in \system{LaMGD} is the number of selected groups in each layer, $k$, which we set to 12 in every experiments.
+By retaining $12/16$ groups, we share on average $75 \%$ active groups between two domains or languages.
+This design ensures that the percentage of sharing is in the same order of magnitude as what we obtain with adapter modules.
+In our future work, we intend to analyze how these choice affact affect the final performance of the model.
 
-Furthermore, the temperature $\tau$ of our $\operatorname{Soft-Top-K}$ operator will be gradually decreased from $0.5$ to $0.2$ as follows
+The temperature parameter $\tau$ for the $\operatorname{Soft-Top-K}$ operator is gradually decreased from $0.5$ to $0.2$ according to the following policy:
 \begin{align*}
-\tau = \operatorname{min}\{ 0.2, 0.5 * \exp^{-r*step} \}
+\tau = \operatorname{min}\{ 0.2, 0.5 * \exp^{-r*step} \},
 \end{align*}
@@
 The choice of the values for our hyperparameters was not optimized and selected by suggestions from previous papers \citep{Jang17categorical,Gong21pay,Gong21adaptive}.
+
 \subsection{Datasets}
-\subsubsection{Multi-domain translation}
-We reuse the same data as the recent work of \citet{Pham21revisiting} in multi-domain translation. We describe our datasets for multi-domain translation experiment in Table \ref{tab:Corpora-chap4}
+\subsubsection{Multidomain translation}
+We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyTodo{Add public to the data link} We describe the datasets for multidomain translation experiments in Table~\ref{tab:Corpora-chap4}.
 \begin{table*}[h!]
@@
 We evaluate our model on both one-to-many (O2M) and many-to-one (M2O)
-translation tasks. We evaluate models on widely used multilingual translation datasets.
+translation tasks borrowing the multilingual translation datasets from past studies. More precisely, we used: 
 \begin{itemize}
-	\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations of \citet{qi18when} between English and eight related languages.
-	\item TED8-Diverse. The dataset consists of parallel sentences between English and eight diverse languages as in \citet{Wang20balancing}.
+\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations from \citet{qi18when} between English and eight \emph{related languages}.
+\item TED8-Diverse. The dataset consists of parallel sentences between English and eight \emph{diverse languages} as in \citet{Wang20balancing}.
 \end{itemize}
-We clarify the notation of the languages in multilingual experiments as follows
+
+The languages used in multilingual experiments are refered to by their ISO 639-1 code that we reproduce below for convenience:
 \begin{itemize}
-	\item Diverse: bos (bosnia), mar(marathi),
-         hin(hindi), mkd(macedonia), ell(greek),
-         bulgaria(bul), french(fra), korean(kor) 
-	\item Related: azerbajiani (aze), belarusian(bel),
-          glacian(glg), slovak (slk), turkisk (tur),
-		  russian(rus), portugese(por),czech(ces)
+\item \textbf{Diverse set}: bos (Bosnian), Bulgarian (bul), French (fra), ell (Greek),
+  hin (Hindi), Korean (kor) mkd (Macedonian), mar (Marathi);
+\item \textbf{Related set}: Azerbajiani (aze), Belarusian (bel),
+  Czech (ces), Galician (glg), Portuguese (por), Russian (rus), Slovak (slk), Turkish (tur).
 \end{itemize}
-Table \ref{tab:Corpora-multilingual} describes the subset of TED8 that we used in our experiments.
+
+Table~\ref{tab:Corpora-multilingual} describes the subsets of TED8 that were used in our experiments.
 \begin{table*}[h!]
@@
 \end{table*}
-\section{Results and ablation studies}
-\subsection{Multi-domain translation}
-According to Table \ref{tab:mdmt}, \system{LaMGD} Tranformer shows a significant improvement (+2.78) over the generic system \system{Transformer} with zero extra parameters.
-Moreover, \system{LaMGD} Tranformer achieves equivalent performance in average compared to \system{Adapter}, which is carrefully finetuned with a approximately 12.4M additional parameters.
+
+\section{Results and analyses}
+
+\subsection{Multidomain translation}
+For these experiments, our main results are in Table ~\ref{tab:mdmt}, where we observe that the \system{LaMGD} Tranformer achieves a significant improvement (+2.78) over the generic \system{Transformer} system with zero extra parameters.
+Moreover, \system{LaMGD} Tranformer achieves performance that are equivalent on average to that of the \system{Adapter} sytems, which is carrefully finetuned and contains approximately 12.4M additional parameters.
 \begin{table*}[h!]

### 2021-11-15 2022/acl/acl_latex.tex
relecture complete

@@
 For each domain, only 12 out of the 16 groups are selected.
-The number of parameters of the variational distributions is $k \times L \times n_d$, which is negligible when compared to the size of the Transformer model;
+The number of parameters of the variational distributions is $k \times L \times n_d$, which is negligible in comparison to the size of the Transformer model.
 \end{itemize}
 
-We set the dropout probability as 0.1.
+We set the dropout probability to 0.1.
 We train the multi-domain Transformer model for 20k iterations with a batch size of 12k tokens.
@@
 
-We set the dropout probability as 0.3.
-We train the multilingual Transformer model in 40~K iterations with a batch size of 9600 tokens on 16 GPUs as in \citet{Gong21adaptive}.
+We set the dropout probability to 0.3.
+We train the multilingual Transformer model for 40k iterations with a batch size of 9600 tokens on 16 V100 GPUs as in \citet{Gong21adaptive}.
 We train \system{LaMGD} Tranformer for 50k iterations with the same batch size.
@@
 
+For all experiments, we report the BLEU score of \citet{Papineni02bleu} computed with SacreBleu \cite{Post18call}.\fyTodo{Check this}
+
 \subsection{Datasets}
 \subsubsection{Multidomain translation}
-We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyTodo{Add public to the data link} We describe the datasets for multidomain translation experiments in Table~\ref{tab:Corpora-chap4}.
+We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyTodo{Add public to the data link} The datasets for the multidomain translation experiments are detailled in Table~\ref{tab:Corpora-chap4}.
 \begin{table*}[h!]
@@
     \system{Transformer}  \hfill{\footnotesize[65m]} & 40.3 & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[+12.4m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.28 \\ 
-    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & 39.0 & 52.4 & 87.5 & 55.33 \\ 
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+12.4m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.3 \\ 
+    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & 39.0 & 52.4 & 87.5 & 55.3 \\ 
     \hline
@@
 \end{table*}
+
 \subsubsection{Fuzzy domain separation}
 \label{ssec:fuzzy}
-In this experiment, we reuse an experiment proposed by \citet{Pham21revisiting} which measures the efficiency of a multi-domain NMT system exploiting the proximity between domains.
-The experiment uses the same data as the previous experiment.
-However, the domain \domain{law} is now randomly split into two pseudo domains \domain{law$_1$} and \domain{law$_2$}.
-If the multi-domain system made full use of the proximity between \domain{law$_1$} and \domain{law$_2$}, there should be no significant difference between the performance of the system trained with six original domains and with seven domains replacing \domain{law} by \domain{law$_1$} and \domain{law$_2$}.
-\citet{Pham21revisiting} showed a large gap between two training settings when using residual adapters.
-We report below the gap between two settings when using \system{LaMGD} Tranformer.
+For this experiment, we reuse proposal of \citet{Pham21revisiting}, who measure the efficiency of a multidomain NMT system exploiting the proximity between domains.
+It uses the same data as in the previous experiment; however, the domain \domain{law} is now randomly split into two pseudo-domains \domain{law$_1$} and \domain{law$_2$} of equal size.
+A truly multidomain system should be able to automatically detect the proximity between \domain{law$_1$} and \domain{law$_2$}, and there should be no significant difference between the performance of a system trained with the six original domains (including \domain{law}) or with the seven domains (including \domain{law} by \domain{law$_1$} and \domain{law$_2$}).
+The experiments in \citep{Pham21revisiting} reported a large gap between the two settings when using residual adapters.
+We replicated this setting and also report the results obtained with the \system{LaMGD} Tranformer system in Table~\ref{tab:fuzzy}.
 
@@
   \centering
-  \begin{tabular}{|p{4cm}|*{2}{r|}} \hline
-    Model / Domain & \multicolumn{1}{c|}{\domain{law$_1$}} & \multicolumn{1}{c|}{\domain{law$_2$}} \\ \hline 
-    \system{Adapter}   \hfill{\footnotesize[+25m]}  & -0.6 & -0.8 \\ 
-    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 0.0 & 0.0 \\ 
+  \begin{tabular}{|p{4cm}|*{3}{r|}} \hline
+    Model / Domain & \multicolumn{1}{c|}{\domain{law}} & \multicolumn{1}{c|}{\domain{law$_1$}} & \multicolumn{1}{c|}{\domain{law$_2$}} \\ \hline 
+    \system{Adapter}   \hfill{\footnotesize[+25m]}  & 61.0 & 59.4 (-0.6) &  50.2 (-0.8) \\ 
+    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 61.4 & 61.4 (=) & 61.4 (=) \\ 
     \hline
   \end{tabular}
-  \caption{Fuzzy domain separation}
+  \caption{Experiments with two similar pseudo-domains }
   \label{tab:fuzzy}
@@
 
-Table \ref{tab:fuzzy} demonstrates a significant loss when training with two pseudo domains \domain{law$_1$} and \domain{law$_2$} compared to training with the real domain \domain{law}.
-Our system does not have this loss between two situations.
-In Section \ref{ssec:abalation}, we show that our algorithm outputs the same sub-network for \domain{law$_1$} and \domain{law$_2$}, that allows a full transfer between two domains.
+The results in Table~\ref{tab:fuzzy} show a performance decrease for the adapter-based system when training with two pseudo-domains \domain{law$_1$} and \domain{law$_2$}.
+In contrast, the \system{LaMGD} model obtains very stable results.
+In Section~\ref{ssec:abalation}, we show that our algorithm in fact computes the same sub-network for \domain{law$_1$} and \domain{law$_2$}, that allows a full sharing of information between these two pseudo-domains.
 
 \subsection{Multilingual translation}
-According to Table \ref{tab:multilingual}, \system{LaMGD} Tranformer shows a boost of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, M2O-diverse respectively.
-Significant gains are observed in \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resourced languages in the list.
-However, \system{LaMGD} Tranformer lags behind multilingual \system{Transformer} and Language Adapters in O2M-diverse.
+Results for the multilingual experiments are in Table~\ref{tab:multilingual}.
+The \system{LaMGD} Tranformer achieves an improvement of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in the O2M-related, M2O-related, M2O-diverse conditions, respectively.
+Significant gains are observed for languages \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resource languages in our sets.
+However, \system{LaMGD} Tranformer is outperformed by the multilingual \system{Transformer} and Language Adapters for the O2M-diverse condition.
 \begin{table*}[h!]
@@
     O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&\SB{34}&38.3&8.3&20.19 \\
+    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&\SB{34.0}&38.3&8.3&20.19 \\
     \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &10.2&4&\SB{13.3}&21.9&32.2&\SB{34.1}&38.5&8.3&20.31 \\ 
@@
     M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19&27.43 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.5&9.4&20&30.6&37.2&38.2&39.3&19&27.03\\ 
-    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40&18.9&27.75 \\
+    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19.0&27.43 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.5&9.4&20.0&30.6&37.2&38.2&39.3&19.0&27.03\\ 
+    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40.0&18.9&27.75 \\
     \hline
   \end{tabular}
-  \caption{Multilingual translation}
+  \caption{Experiments with multilingual translation.\fyTodo{What is the bold for ?}}
   \label{tab:multilingual}
@@
 \label{ssec:abalation}
-This section compares the sub-networks of the domains or the language pairs by computing the average of the similarities between the corresponding dropping masks concatenated through all the layers of the underlying model.
+This section compares the sub-networks learnt for each domain or language pair by computing the average similarity between the corresponding dropout masks concatenated for all the layers of the underlying model.
+For the multidomain experiment, we analyze the case of pseud-domain separation reported in Section~\ref{ssec:fuzzy} in Table~\ref{tab:fuzzy-sim}.
+We see that the sub-networks for \domain{law$_1$} and \domain{law$_2$} are identical, yielding a full sharing between the corresponding training sets.
+Furthermore, we observe a large distance between \domain{rel} and the other domains, which is expected given that \domain{rel} is quite distinct from the other domains.
+\domain{rel} only share around $75 \%$ its active groups with other domains, as would be obtained by chance in our setting (see Section~\ref{ssec:hyperparams}).
+In Figure ~\ref{fig:domain}, we visualize the domains using their dropping masks concatenated and mapped to a 2d space using Principal Component Analysis (PCA).
 
-For the multi-domain experiment, we analyze the case of fuzzy domain separation in Section \ref{ssec:fuzzy}.
-Table \ref{tab:fuzzy-sim} shows that the sub-networks of \domain{law$_1$} and \domain{law$_2$} are identical, which is one of the goals of our method.
-Furthermore, we remark a large distance between \domain{rel} and the rest of the domains, which correlates to the fact that \domain{rel} has the most distant topic to other domains.
-\domain{rel} only share around $75 \%$ its active groups with other domains whereas $75 \%$ is the average of sharing proportions by default (see Section \ref{ssec:hyperparams}).
-In Figure ~\ref{fig:domain}, we visualize the domains using their dropping masks concatenated and mapped to 2D space by PCA.
 \begin{table*}[h!]
@@
   \end{tabular}
-  \caption{Multi-domain sub-networks' similarities}
+  \caption{Multidomain experiments: sub-network similarities}
   \label{tab:fuzzy-sim}
@@
 
-For multilingual (TED-related) case, the data contain four language families including (1) Turkic family with Azerbaijani and Turkish(\domain{aze},\domain{tur}); (2) Slavic family with Belarusian and Russian (\domain{bel},\domain{rus}); (3) Romance family with Galician and Portuguese (\domain{glg}, \domain{por}); and (4) Czech-Slovak family with Slovak and Czech (\domain{ces}, \domain{slk}).
-The similarities between the sub-networks for each pair are respectively 0.95, 0.94, 0.99 and 0.96: for each language, the maximally similar one (excluding the diagonoal) is always from the same language family.
+For multilingual (TED-related) experiments, the training data contains four language families: (1) Turkic, with Azerbaijani and Turkish(\domain{aze},\domain{tur}); (2) Slavic, with Belarusian and Russian (\domain{bel},\domain{rus}); (3) Romance, with Galician and Portuguese (\domain{glg}, \domain{por}); and (4) Czech-Slovak, with Slovak and Czech (\domain{ces}, \domain{slk}).
+The similarities between the sub-networks for each pair are respectively 0.95, 0.94, 0.99 and 0.96: for each language, the most similar language is always from the same language family.
 \begin{table*}[h!]
@@
 
-We also plot the languages based on their dropout masks in Figure~\ref{fig:visual} using a 2d projection computed by Principal Component Analysis (PCA).
+We also plot the languages based on their dropout masks in Figure~\ref{fig:visual} using a 2d projection computed by PCA.
 \begin{figure*}[htbp]
@@
 
-\begin{figure*}[h!]
-\begin{subfigure}{0.5\textwidth}
-  \centering
-  % include first image
-  \includegraphics[width=0.9\textwidth]{phi_ted_related_o2m}  
-  \caption{TED-Related}
-  \label{fig:phi-related}
-\end{subfigure}
-\begin{subfigure}{0.5\textwidth}
-  \centering
-  % include second image
-  \includegraphics[width=0.85\textwidth]{phi_ted_diverse_o2m}  
-  \caption{TED-Diverse}
-  \label{fig:phi-diverse}
-\end{subfigure}
-\caption{$\Phi_L^{0}$ in O2M experiments}
-\label{fig:phi}
-\end{figure*}
+% \begin{figure*}[h!]
+% \begin{subfigure}{0.5\textwidth}
+%   \centering
+%   % include first image
+%   \includegraphics[width=0.9\textwidth]{phi_ted_related_o2m}  
+%   \caption{TED-Related}
+%   \label{fig:phi-related}
+% \end{subfigure}
+% \begin{subfigure}{0.5\textwidth}
+%   \centering
+%   % include second image
+%   \includegraphics[width=0.85\textwidth]{phi_ted_diverse_o2m}  
+%   \caption{TED-Diverse}
+%   \label{fig:phi-diverse}
+% \end{subfigure}
+% \caption{$\Phi_L^{0}$ in O2M experiments}
+% \label{fig:phi}
+% \end{figure*}
 
@@
 However, the hard-coded separation between the domains of different tasks may lead to a catastrophic forgetting effect \citep{pfeiffer21adapterfusion}, which is a common problem in multi-task modeling using neural networks \citep{Michael89catastrophic}.
-In multi-domain translation, \citet{Pham21revisiting} recently demonstrated the brittleness of domain adapters against fuzzy domain separations, out-of-domain distributions, and erroneous domain tags.
-Several subsequent studies have aimed to mitigate this weakness through a mixture of expert mechanism, for example,\cite{pfeiffer21adapterfusion}.
+In multidomain translation, \citet{Pham21revisiting} recently demonstrated the brittleness of adapters against fuzzy domain separations, out-of-domain distributions, and erroneous domain tags.
+Several subsequent studies have aimed to mitigate this weakness through a mixture of expert mechanism (e.g.\ \cite{pfeiffer21adapterfusion}).
 
@@
 \section{Conclusions and outlook}\label{sec:conclusion}
-In this work, we present a novel method for multdomain and multilingual translation.
-It allows us to jointly search for an optimal assignment of sub-networks to tasks and learn the parameters of the underlying network.
-Our method relies on a sound mathematical model and an end-to-end optimization with very little extra computational cost.
-Experimentally, we achieve a large improvement over Transformer baseline without zero additional parameters; our performance are also comparable to that of a strong a multi-task baseline using residual adapter modules which however includes a large number of extra parameters.
+In this work, we have presented a novel method for multdomain and multilingual translation.
+It allows us to jointly search for an optimal assignment of sub-networks to tasks and to learn the parameters of the underlying network.
+Our method relies on a sound mathematical framework and an end-to-end optimization procedure which implies very little extra parameters.\fyTodo{There is a computational cost = 100k iterations} Experimentally, we achieve a large improvement over Transformer baseline without zero additional parameters; our performance are also comparable to that of a strong a multi-task baseline using residual adapter modules which however includes a large number of extra parameters.
 For multilingual translation, our model outperforms multilingual Transformer and Language Adapters in 3 our of 4 cases.

### 2021-11-15 2022/acl/acl_latex-mq.tex
my local

--- 2021-11-15 2022/acl/acl_latex.tex
@@
   \setlength{\parsep}{0pt}
-\item We provide a sound mathematical formulation to the problem of jointly learning task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling techniques;
+\item We provide a rigorous mathematical formulation to the problem of jointly learning task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling techniques;
 \item We present algorithms to train this model end to end with very little extra computational cost.
@@
   \centering
-  \begin{tabular}{l|cccc} 
+  \begin{tabular}{l|cccc|l|cccc} 
   \hline
-    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c}{\domain{test}} \\
+    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c}{\domain{test}} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c}{\domain{test}} \\
     \hline 
-    \multirow{8}{*}{Related} & Azerbaijani & 5.94k & 671 & 903 \\
-							 & Belarusian & 4.51k & 248 & 664 \\
-							 & Glacian & 10.0k & 682 & 1007 \\
-							 & Slovak & 61.5k & 2271 & 2445 \\
-							 & Turkish & 182k & 4045 & 5029 \\
-                             & Russian & 208k & 4814 & 5483 \\
-                             & Portuguese & 185k & 4035 & 4855 \\
-                             & Czech & 103k & 3462 & 3831 \\
+    \multirow{2}{*}{Related} & Azerbaijani & 5.94k & 671 & 903 & \multirow{2}{*}{Diverse} & Bosnian & 5.64k & 474 & 463 \\
+							 & Belarusian & 4.51k & 248 & 664 & & Marathi & 9.84k & 767 & 1090\\
+							 & Glacian & 10.0k & 682 & 1007 & & Hindi & 18.79k & 854 & 1243\\
+							 & Slovak & 61.5k & 2271 & 2445 & & Macedonian & 25.33k & 640 & 438 \\
+							 & Turkish & 182k & 4045 & 5029 & & Greek & 134k & 3344 & 4433\\
+                             & Russian & 208k & 4814 & 5483 & & Bulgarian & 174k & 4082 & 5060 \\
+                             & Portuguese & 185k & 4035 & 4855 & & French & 192k & 4320 & 4866\\
+                             & Czech & 103k & 3462 & 3831 & & Korean & 205k & 4441 & 5637 \\
     \hline
-    \multirow{8}{*}{Related} & Bosnian & 5.64k & 474 & 463 \\
-                             & Marathi & 9.84k & 767 & 1090 \\
-                             & Hindi & 18.79k & 854 & 1243 \\
-                             & Macedonian & 25.33k & 640 & 438 \\
-                             & Greek & 134k & 3344 & 4433 \\
-                             & Bulgarian & 174k & 4082 & 5060 \\
-                             & French & 192k & 4320 & 4866 \\
-                             & Korean & 205k & 4441 & 5637 \\
   \end{tabular}

### 2021-11-15 2022/acl/acl_latex.tex
answer fyTodos, correct spelling errors in abstract

@@
 \begin{abstract}
-  Multidomain and multilingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalties of the tasks at hand, while smaller parts are reserved to model the peculariaties of a language or a domain.
-In adapter-based approaches, these strategies are hard coded in the network architecture, independent of the similarities that exist between tasks.
-In this work, we propose a new method aimed at better taking advantage of these similarities, using a latent-variable model.
-We also develop new techniques to train this model end to end and report multiple experimental results that show the the learned patterns are both meaningful and yield improved translation performance, without any increase of the model size.
+  Multidomain and multilingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalities of the tasks at hand, while smaller parts are reserved to model the peculiarities of a language or a domain.
+In adapter-based approaches, these strategies are hardcoded in the network architecture, independent of the similarities between tasks.
+In this work, we propose a new method to better take advantage of these similarities, using a latent-variable model.
+We also develop new techniques to train this model end to end and report multiple experimental results that show the learned patterns are both meaningful and yield improved translation performance without any increase of the model size.
 \end{abstract}
@@
 \item Adapter-based Transformer: The intermediate feedforward dimension is set to 2048;% as in \citet{Pham21revisiting}.
-\item \system{LaMGD} Transformer:\fyTodo{Explain acronym} There is no change in the architecture.
+\item Transformer using Latent multi-task group dropout (denoted \system{LaMGD} Transformer):\fyDone{Explain acronym} There is no change in the architecture.
 We group the 512 nodes in each layer into 16 groups of 32 consecutive nodes.
@@
 
-All the translation systems are implemented OpenNMT\fyTodo{Fix this, add reference to framework if needed}.
+All the translation systems are implemented with OpenNMT-tf \footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}\fyDone{Fix this, add reference to framework if needed}.
 
@@
 
-For all experiments, we report the BLEU score of \citet{Papineni02bleu} computed with SacreBleu \cite{Post18call}.\fyTodo{Check this}
+For all experiments, we report the BLEU score of \citet{Papineni02bleu} computed with SacreBleu \cite{Post18call}.\fyDone{Check this}
 
@@
 \subsubsection{Multidomain translation}
-We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyTodo{Add public to the data link} The datasets for the multidomain translation experiments are detailled in Table~\ref{tab:Corpora-chap4}.
+We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyDone{Add public to the data link} The datasets \footnote{The raw data can be downloaded from \url{https://github.com/qmpham/experiments/tree/main/tacl20}} for the multidomain translation experiments are detailed in Table~\ref{tab:Corpora-chap4}.
 \begin{table*}[h!]
@@
 
-The languages used in multilingual experiments are refered to by their ISO 639-1 code that we reproduce below for convenience:
+The languages used in multilingual experiments are referred to by their ISO 639-1 code that we reproduce below for convenience:
 \begin{itemize}
@@
 \end{table*}
-
 \section{Results and analyses}
@@
   \end{tabular}
-  \caption{Experiments with multilingual translation.\fyTodo{What is the bold for ?}}
+  \caption{Experiments with multilingual translation. Boldface denotes significant gains over \system{Transformer}\fyDone{What is the bold for ?}}
   \label{tab:multilingual}
@@
   % include first image
-  \includegraphics[width=0.85\textwidth]{ted_related}  
+  \includegraphics[width=0.9\textwidth]{ted_related}  
   \caption{TED-Related}
@@
 This method is closely related to the FusionAdapters of \citet{pfeiffer21adapterfusion}.
-Both approaches assume that the final representation\fyTodo{??} is a combination or a choice between shared representation and task-specific representations.
+Both approaches aimed to select between shared representation and task-specific representations.
 The proximity between tasks is not taken into account in the routing mechanism.
@@
 It allows us to jointly search for an optimal assignment of sub-networks to tasks and to learn the parameters of the underlying network.
-Our method relies on a sound mathematical framework and an end-to-end optimization procedure which implies very little extra parameters.\fyTodo{There is a computational cost = 100k iterations} Experimentally, we achieve a large improvement over Transformer baseline without zero additional parameters; our performance are also comparable to that of a strong a multi-task baseline using residual adapter modules which however includes a large number of extra parameters.
+Our method relies on a sound mathematical framework and an end-to-end optimization procedure which implies very little extra parameters.
+The training course is reasonably longer (100k iterations in multidomain translation) than training standard models while considering the gain in performance.\fyDone{There is a computational cost = 100k iterations} Experimentally, we achieve a large improvement over Transformer baseline without zero additional parameters; our performance are also comparable to that of a strong a multi-task baseline using residual adapter modules which however includes a large number of extra parameters.
 For multilingual translation, our model outperforms multilingual Transformer and Language Adapters in 3 our of 4 cases.
@@
 \begin{align*}
-&\frac{\partial g}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d} + \frac{\partial g}{\partial \Phi_l^d} = 0 \\
-& \Rightarrow \frac{\partial \nu}{\partial \Phi_l^d} = - \big(\frac{\partial g}{\partial \nu}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}
+&\frac{\partial g}{\partial \bar{\nu}} \times \frac{\partial \bar{\nu}}{\partial \Phi_l^d} + \frac{\partial g}{\partial \Phi_l^d} = 0 \\
+& \Rightarrow \frac{\partial \bar{\nu}}{\partial \Phi_l^d} = - \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}
 \end{align*}
@@
 \section{Appendix C}
-\begin{figure*}
-\includegraphics[width=0.7\textwidth]{multi_domain}
-\caption{Visualization of domains according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
-\label{fig:domain}
-\end{figure*}
-
-\section{Appendix D}
 \begin{breakablealgorithm}
@@
     \setlength{\itemsep}{-1pt}    \setlength{\parsep}{0pt}
-  \item[] 
   \item $n_d$ corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equipped by an empirical distribution $D_d(x)$
   \item number of retained groups: k, number of groups: $n_p$
+  \item $i=0$
  \end{itemize}}
@@
 \STATE{Sampling $\forall p,  g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
-\IF{$Early\_stopping(Eval\_scores)$}
-	\STATE{break}
-\ENDIF
+\STATE{Solve the equation
+\begin{align*}
+\displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = k
+\end{align*}
+using binary search}
+\STATE{Computing mask of each layer
+\begin{align*}
+\hat{m}_l^d(\tau) &= \sigma(\frac{g_l^d + \Phi_l^d + \bar{\nu}}{\tau})
+\end{align*}}
+\STATE{Applying computed masks to their corresponding layer
+\begin{align*}
+  \forall l \in [0,\cdots, L-1]: \tilde{h}^l &= h^l \odot r_l^d ,\\
+  h^{l+1} &= \operatorname{LAYER}^{l+1}(\tilde{h}^l) ,
+\end{align*}
+}
+\STATE{Compute gradient of training loss over the underlying Transformer
+$$\Delta_{\theta} = \frac{\partial L}{\partial \theta}$$}
+\STATE{Compute gradient over the $\operatorname{Soft-Top-K}$ masks
+$$\frac{\partial D}{\partial \hat{m}_l^d(\tau)}$$}
+\STATE{Compute implicit gradient of the $\operatorname{Soft-Top-K}$ masks over $\Phi^d$
+$$- \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}$$}
+\STATE{Compute the gradient the training over $\Phi$
+$$ - \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d} + \frac{\partial \mathbb{H} \big[ \operatorname{softmax}(\Phi_l^d)\big] }{\partial \Phi_l^d} $$}
+\STATE{Update $\theta$ and $\Phi^d$ with their gradients}
 \STATE{$i=i+1$}
@@
 \end{breakablealgorithm}
-
+\section{Appendix D}
+\begin{figure*}[h!]
+\includegraphics[width=0.7\textwidth]{multi_domain}
+\caption{Visualization of domains according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
+\label{fig:domain}
+\end{figure*}
 \end{document}

### 2021-11-15 2022/acl/acl_latex.tex
correct g^d -> MT^d

@@
 where $\operatorname{KL}$ is the Kullback-Leibler divergence.
-We use $-\ell{ELBO}(x,y,d; \theta,\Phi)$ as our surrogate training objective, as a tractable approximation of the untractable likelihood, and try to minimize this function in $\theta$ and $\Phi$.
+We use $-\ell(x,y,d; \theta,\Phi)$ as our surrogate training objective, as a tractable approximation of the untractable likelihood, and try to minimize this function in $\theta$ and $\Phi$.
 
@@
 \hspace{-0.em}
-\mathcal{L}(\theta,\Phi) = \sum_{d=1}^{n_d} \mathbb{E}_{x \sim \mathcal{D}_s^d,y \sim g^d(x)} \big[-\ell(x,y,d; \theta,\Phi)\big]
+\mathcal{L}(\theta,\Phi) = \sum_{d=1}^{n_d} \mathbb{E}_{x \sim \mathcal{D}_s^d,y \sim MT^d(x)} \big[-\ell(x,y,d; \theta,\Phi)\big]
 \end{align}

### 2021-11-16 2022/acl/acl_latex.tex
Fixed small glitches, improved table 3, more TODOs

@@
 \newcommand{\mtsquare}[0]{MT$^2$}
+\newcommand{\lagd}[0]{LaGD$^2$}
 
@@
 \section{Introduction}
-Multidomain and multilingual machine translation aim to develop one single model to perform translations for multiple domains and multiple language pairs, respectively.\footnote{We will refer to these two situations as 'multi-task MT' and refer to individual domains and languages as 'tasks'.} These paradigms are motivated by the compactness of the resulting translation system \citep{Chu18multilingual,dabre20survey}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
+Multidomain and multilingual machine translation aim to develop one single model to perform translation for multiple domains and multiple language pairs, respectively.\footnote{We will refer to these two situations as 'multi-task MT' and refer to individual domains and languages as 'tasks'.} These paradigms are motivated by the compactness of the resulting translation system \citep{Chu18multilingual,dabre20survey}, the hypothetical positive knowledge transfer between similar domains \citep{Pham21revisiting} or between languages in the same family \citep{Tan19multilingual}.
 However, having all the tasks use exactly the same model parameters can cause negative interference between unrelated tasks \citep{conneau20unsupervised,wang20negative}.
-Hence the recent development of approaches relying on a partial sharing of the parameters, eg.\ using adapter layers as studied in \citep{houlsby19parameter,Bapna19simple,Pham20Study,Philip20monolingual}.
+Hence, the recent development of approaches relying on a partial sharing of the parameters, eg.\ using adapter layers as studied in \citep{houlsby19parameter,Bapna19simple,Pham20Study,Philip20monolingual}.
 If these techniques have proven effective for building strong baselines, they fail to fully take advantage of the similarities that exist between domains and tasks, as documented eg.\ in \citep{Pham21revisiting}.
-This is because the partition of the parameter space between generic or task-specific subparts, and their allocation to each task is hard-coded in the network, irrespective of the actual commonalties and differences in the data space.
+This is because the partition of the parameter space between generic or task-specific subparts, and their allocation to each task, is hard-coded in the network, irrespective of the actual commonalties and differences in the data space.
 
-In this work, we develop techniques that can take the similarity between tasks into account in a more effective way, by learning the network organization from the data.
-To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the parameter space, and show how training can still be performed end-to-end using a variational surrogate of the standard loss function.
+In this work, we develop techniques that can take the similarity between tasks into account in a more effective way, by \emph{learning the network organization from the data}.
+To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the parameter space, and show how training can still be performed end-to-end using a variational surrogate of the log-likelihood loss function.
 Our experiments with multilingual and multidomain machine translation confirm that this method can automatically detect similarities in the data, meaning that related tasks use the same subparts of the network.
@@
 \begin{itemize}
-\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048; the multi-head attention mechanism contains 8 heads;
+\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048; the multi-head attention mechanism contains 8 heads;\fyTodo{How many layers}
 \item Adapter-based Transformer: The intermediate feedforward dimension is set to 2048;% as in \citet{Pham21revisiting}.
-\item Transformer using Latent multi-task group dropout (denoted \system{LaMGD} Transformer):\fyDone{Explain acronym} There is no change in the architecture.
+\item Transformer using Latent multi-task group dropout (\system{LaMGD} Transformers):\fyDone{Explain acronym} There is no change in the architecture.
 We group the 512 nodes in each layer into 16 groups of 32 consecutive nodes.
 For each domain, only 12 out of the 16 groups are selected.
-The number of parameters of the variational distributions is $k \times L \times n_d$, which is negligible in comparison to the size of the Transformer model.
+The number of parameters of the variational distribution is $L\times k \times L \times n_d$, which is negligible in comparison to the size of the Transformer model.
 \end{itemize}
@@
 We train the multi-domain Transformer model for 20k iterations with a batch size of 12k tokens.
-We train \system{LaMGD} Tranformer for 300k iterations with a batch size of 12k tokens: these extra iterations are justified by the need for the \system{LaMGD} Tranformer to learn both the sub-networks and the parameters of the underlying Transformer.
-Finally, we plug domain Adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline.
+We train \system{LaMGD} Tranformer for 300k iterations with a batch size of 12k tokens: these extra iterations are justified by the need for the \system{LaMGD} Tranformer to learn both the sub-networks and the parameters of the underlying Transformer.\fyTodo{Have we converged?} Finally, we plug domain adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline.
 
@@
 
-All the translation systems are implemented with OpenNMT-tf \footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}\fyDone{Fix this, add reference to framework if needed}.
+All the translation systems are implemented with OpenNMT-tf \footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.\fyDone{Fix this, add reference to framework if needed}
 
@@
 By retaining $12/16$ groups, we share on average $75 \%$ active groups between two domains or languages.
-This design ensures that the percentage of sharing is in the same order of magnitude as what we obtain with adapter modules.
-In our future work, we intend to analyze how these choice affact affect the final performance of the model.
+This design ensures that the percentage of sharing is in the same ballpark as what we obtain with adapter modules.
+In our future work, we intend to analyze how these choices affect the final performance of the model.
 
@@
 Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments.
-The choice of the values for our hyperparameters was not optimized and selected by suggestions from previous papers \citep{Jang17categorical,Gong21pay,Gong21adaptive}.
+The choice of the values for our hyperparameters was not optimized and selected by suggestions from previous studies \citep{Jang17categorical,Gong21pay,Gong21adaptive}.
 
-For all experiments, we report the BLEU score of \citet{Papineni02bleu} computed with SacreBleu \cite{Post18call}.\fyDone{Check this}
+For all experiments, we report the BLEU score of \citet{Papineni02bleu} computed with SacreBleu \citep{Post18call}.\fyDone{Check this}
 
@@
 Table~\ref{tab:Corpora-multilingual} describes the subsets of TED8 that were used in our experiments.
-\begin{table*}[h!]
+\begin{table*}[h]
   \centering
-  \begin{tabular}{l|cccc|l|cccc} 
+  \begin{tabular}{|cccc||cccc|} 
   \hline
-    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c}{\domain{test}} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c}{\domain{test}} \\
+    \multicolumn{4}{|c||}{Related} & \multicolumn{4}{c|}{Diverse} \\
+                            \multicolumn{1}{|c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c||}{\domain{test}} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c|}{\domain{test}} \\
     \hline 
-    \multirow{2}{*}{Related} & Azerbaijani & 5.94k & 671 & 903 & \multirow{2}{*}{Diverse} & Bosnian & 5.64k & 474 & 463 \\
-							 & Belarusian & 4.51k & 248 & 664 & & Marathi & 9.84k & 767 & 1090\\
-							 & Glacian & 10.0k & 682 & 1007 & & Hindi & 18.79k & 854 & 1243\\
-							 & Slovak & 61.5k & 2271 & 2445 & & Macedonian & 25.33k & 640 & 438 \\
-							 & Turkish & 182k & 4045 & 5029 & & Greek & 134k & 3344 & 4433\\
-                             & Russian & 208k & 4814 & 5483 & & Bulgarian & 174k & 4082 & 5060 \\
-                             & Portuguese & 185k & 4035 & 4855 & & French & 192k & 4320 & 4866\\
-                             & Czech & 103k & 3462 & 3831 & & Korean & 205k & 4441 & 5637 \\
+    Azerbaijani & 5.94k & 671 & 903 & Bosnian & 5.64k & 474 & 463 \\
+    Belarusian & 4.51k & 248 & 664 & Marathi & 9.84k & 767 & 1090\\
+    Galician & 10.0k & 682 & 1007 & Hindi & 18.79k & 854 & 1243\\
+    Slovak & 61.5k & 2271 & 2445 & Macedonian & 25.33k & 640 & 438 \\
+    Turkish & 182k & 4045 & 5029 & Greek & 134k & 3344 & 4433\\
+    Russian & 208k & 4814 & 5483 &  Bulgarian & 174k & 4082 & 5060 \\
+    Portuguese & 185k & 4035 & 4855 & French & 192k & 4320 & 4866\\
+    Czech & 103k & 3462 & 3831 & Korean & 205k & 4441 & 5637 \\
     \hline
@@
 The experiments in \citep{Pham21revisiting} reported a large gap between the two settings when using residual adapters.
-We replicated this setting and also report the results obtained with the \system{LaMGD} Tranformer system in Table~\ref{tab:fuzzy}.
+We replicated this setting and also report the results obtained with the \system{LaMGD} Transformer system in Table~\ref{tab:fuzzy}.
 
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{law}} & \multicolumn{1}{c|}{\domain{law$_1$}} & \multicolumn{1}{c|}{\domain{law$_2$}} \\ \hline 
-    \system{Adapter}   \hfill{\footnotesize[+25m]}  & 61.0 & 59.4 (-0.6) &  50.2 (-0.8) \\ 
-    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 61.4 & 61.4 (=) & 61.4 (=) \\ 
+    \system{Adapter}   \hfill{\footnotesize[+25m]}  & 61.0 & 59.4 (-0.6) &  59.2 (-0.8) \\ 
+    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  & 61.4 & 61.4 (=) & 61.4 (=) \\ 
     \hline
@@
 Results for the multilingual experiments are in Table~\ref{tab:multilingual}.
-The \system{LaMGD} Tranformer achieves an improvement of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in the O2M-related, M2O-related, M2O-diverse conditions, respectively.
+The \system{LaMGD} Transformer achieves an improvement of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in the O2M-related, M2O-related, M2O-diverse conditions, respectively.
 Significant gains are observed for languages \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resource languages in our sets.
-However, \system{LaMGD} Tranformer is outperformed by the multilingual \system{Transformer} and Language Adapters for the O2M-diverse condition.
+However, \system{LaMGD} Transformer is outperformed by the multilingual \system{Transformer} and language Adapters for the O2M-diverse condition.
 \begin{table*}[h!]
@@
     \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &4.3&6.8&21.1&22&39.7&20&23&15.2&19 \\ 
-    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.33 \\ 
+    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.33 \\ 
 	\hline
@@
     \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.64\\ 
-    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.31 \\ 
+    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.31 \\ 
     \hline
@@
     \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &10.2&4&\SB{13.3}&21.9&32.2&\SB{34.1}&38.5&8.3&20.31 \\ 
-    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
+    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
     \hline 
@@
     \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.5&9.4&20.0&30.6&37.2&38.2&39.3&19.0&27.03\\ 
-    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40.0&18.9&27.75 \\
+    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40.0&18.9&27.75 \\
     \hline
   \end{tabular}
-  \caption{Experiments with multilingual translation. Boldface denotes significant gains over \system{Transformer}\fyDone{What is the bold for ?}}
+  \caption{Experiments with multilingual translation. Boldface denotes significant gains over \system{Transformer}.\fyDone{What is the bold for ?}}
   \label{tab:multilingual}

### 2021-11-16 2022/acl/acl_latex.tex
quick check of the appendix

@@
 \label{appendix:a}
-This section explains how we calculate $\hat{m}_l^d(\tau)$ by solving the optimization problem \eqref{eq:soft-top-k} and then how to compute the gradient $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$.
+This section explains how to compute $\hat{m}_l^d(\tau)$ by solving the optimization problem \eqref{eq:soft-top-k} and then how to compute the gradients $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$.
 
-First, to solve \eqref{eq:soft-top-k} we follow the same approach as in \cite{amos19lml,amos20differential} by applying the Karush–Kuhn–Tucker (KKT) conditions to \eqref{eq:soft-top-k}.
-The solution of \eqref{eq:soft-top-k} will have the following form
-
+First, to solve \eqref{eq:soft-top-k} we follow the same approach as in \cite{amos19lml,amos20differential} by applying the Karush–Kuhn–Tucker (KKT) conditions to~\eqref{eq:soft-top-k}.
+The solution of \eqref{eq:soft-top-k} will have the following form:
 \begin{align}
@@
 \end{align}
-in which $\sigma(.)$ is sigmoid function and $\bar{\nu}$ is the solution of the following equation
+in which $\sigma(.)$ is the sigmoid function and $\bar{\nu}$ is the solution of the following equation:
 \begin{align}
@@
 
-Because sigmoid is monotonic, the equation \eqref{eq:nu} has unique solution.
-Furthermore,  because of the smoothness of $g(\nu,\Phi_l^d) = \displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau})$ w.r.t $\nu$ and $\Phi_l^d$, we could deduce the implicit differentiation of its solution $\bar{\nu}$ w.r.t $\Phi_l^d$ as below although the solution of \eqref{eq:nu} does not have explicit form.
+Because sigmoid is monotonous, equation~\eqref{eq:nu} has a unique solution.
+Furthermore,  because of the smoothness of $g(\nu,\Phi_l^d) = \displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau})$ w.r.t $\nu$ and $\Phi_l^d$, we can perform the implicit differentiation of its solution $\bar{\nu}$ w.r.t $\Phi_l^d$ as below, even though the solution of \eqref{eq:nu} does not have an explicit form.
 
@@
 Therefore, we do not need autograd to compute the implicit gradient $\frac{\partial \nu}{\partial \Phi_l^d}$.
-The gradient of $\hat{m}_l^d(\tau)$ w.r.t $\Phi_l^d$ is deduced as follows
+The gradient of $\hat{m}_l^d(\tau)$ w.r.t $\Phi_l^d$ is computed as follows:
 \begin{align}
@@
 The convergence of binary search is extremely fast and assured by the monotonicity of $g(\nu,\Phi_l^d)$.
-In our experiments, we set our search range to $[-100,100]$.
+In our experiments, we set the search range to $[-100,100]$.
 
-Finally, we need prove the limit $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
-We suppose $g_l^d(i_1) + \Phi_l^d(i_1) > g_l^d(i_2) + \Phi_l^d(i_2) > \cdots > g_l^d(i_{n_p}) + \Phi_l^d(i_{n_p})$.
+Finally, we need prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
+We assume $g_l^d(i_1) + \Phi_l^d(i_1) > g_l^d(i_2) + \Phi_l^d(i_2) > \cdots > g_l^d(i_{n_p}) + \Phi_l^d(i_{n_p})$.
 
-Because
-
+Because:
 \begin{align*}
@@
 \end{align*}
-
-there exist $\epsilon$ such that $\forall \tau < \epsilon$, the solution $\bar{\nu}$ of \eqref{eq:nu} is between $- (g_l^d(i_{k+1}) + \Phi_l^d(i_{k+1})) > \bar{\nu} > - (g_l^d(i_k) + \Phi_l^d(i_k))$.
-Furthermore, because sigmoid is monotonic,
+there exist $\epsilon$ such that $\forall \tau < \epsilon$, the solution $\bar{\nu}$ of \eqref{eq:nu} satisfies $- (g_l^d(i_{k+1}) + \Phi_l^d(i_{k+1})) > \bar{\nu} > - (g_l^d(i_k) + \Phi_l^d(i_k))$.
+Furthermore, because sigmoid is monotonous,
 
@@
 
-By taking the limit of both sides, we have following limits
-
+By taking the limit on both sides, we get the following results:
 \begin{align*}
@@
 
-And, because $ \displaystyle{\mathop{\sum}_{u=1}^{n_p}}\hat{m}_l^d(\tau)(i_u) = k$, by taking the limit at both side, we will have $\lim_{\tau \rightarrow 0} \hat{m}_l^d(\tau)(i_k) = 1$.
+And, because $ \displaystyle{\mathop{\sum}_{u=1}^{n_p}}\hat{m}_l^d(\tau)(i_u) = k$, by taking the limit on both sides, we will have $\lim_{\tau \rightarrow 0} \hat{m}_l^d(\tau)(i_k) = 1$.
 Finally, we have
@@
 \label{appendix:b}
-In this section, we give a simple proof of Inequality \ref{eq:entropy}.
+In this section, we give a simple proof of inequality~\eqref{eq:entropy}.
 In fact, we only need to prove $\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] \geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]$.
-The proof is as follows
+The proof is as follows:
 \begin{align*}
@@
 \begin{algorithmic}[1]
-\REQUIRE {
+  \REQUIRE {%    
   \begin{itemize}
     \setlength{\itemsep}{-1pt}    \setlength{\parsep}{0pt}
-  \item $n_d$ corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equipped by an empirical distribution $D_d(x)$
-  \item number of retained groups: k, number of groups: $n_p$
+  \item[] 
+  \item $n_d$ corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equiped by an empirical distribution $D_d(x)$
+  \item number of groups: $n_p$; number of retained groups: $k$ 
   \item $i=0$
@@
 \STATE{Pick a batch from domain $d$}
-\STATE{Sampling $\forall p,  g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
-\STATE{Solve the equation
+\STATE{Sample $\forall l, \forall p:  g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
+\STATE{Solve the equation $\forall l$
 \begin{align*}
@@
 using binary search}
-\STATE{Computing mask of each layer
+\STATE{Compute mask of each layer
 \begin{align*}
-\hat{m}_l^d(\tau) &= \sigma(\frac{g_l^d + \Phi_l^d + \bar{\nu}}{\tau})
+\forall l, \hat{m}_l^d(\tau) &= \sigma(\frac{g_l^d + \Phi_l^d + \bar{\nu}}{\tau})
 \end{align*}}
-\STATE{Applying computed masks to their corresponding layer
+\STATE{Apply masks to their corresponding layer
 \begin{align*}
@@
 \end{breakablealgorithm}
-\section{Appendix D}
+\section{Appendix D}\fyTodo{fix this}
 \begin{figure*}[h!]

### 2021-11-16 2022/acl/acl_latex.tex
final version, correct figures, monotonous -> monotonically increasing, check convergence, refit the figure of domains

@@
 \begin{itemize}
-\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048; the multi-head attention mechanism contains 8 heads;\fyTodo{How many layers}
+\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048; the multi-head attention mechanism contains 8 heads; 6 layers in the encoder; 6 layers in the decoder.\fyDone{How many layers}
 \item Adapter-based Transformer: The intermediate feedforward dimension is set to 2048;% as in \citet{Pham21revisiting}.
@@
 We train the multi-domain Transformer model for 20k iterations with a batch size of 12k tokens.
-We train \system{LaMGD} Tranformer for 300k iterations with a batch size of 12k tokens: these extra iterations are justified by the need for the \system{LaMGD} Tranformer to learn both the sub-networks and the parameters of the underlying Transformer.\fyTodo{Have we converged?} Finally, we plug domain adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline.
+We train \system{LaMGD} Tranformer for 300k iterations with a batch size of 12k tokens: these extra iterations are justified by the need for the \system{LaMGD} Tranformer to learn both the sub-networks and the parameters of the underlying Transformer.\fyDone{Have we converged?} Finally, we plug domain adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline.
 
@@
 \begin{itemize}
-\item Transformer models: the embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 1024, each multi-head attentions contains 8~heads as in \citep{Wang20balancing}.
+\item Multilingual Transformer: the embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 1024, each multi-head attentions contains 8~heads as in \citep{Wang20balancing}.
 \item Adapter based Transformer: the intermediate feedforward dimension is set as 128 as in \citep{Gong21adaptive}.
@@
 
-Because sigmoid is monotonous, equation~\eqref{eq:nu} has a unique solution.
+Because sigmoid is monotonically increasing, equation~\eqref{eq:nu} has a unique solution.
 Furthermore,  because of the smoothness of $g(\nu,\Phi_l^d) = \displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau})$ w.r.t $\nu$ and $\Phi_l^d$, we can perform the implicit differentiation of its solution $\bar{\nu}$ w.r.t $\Phi_l^d$ as below, even though the solution of \eqref{eq:nu} does not have an explicit form.
@@
 there exist $\epsilon$ such that $\forall \tau < \epsilon$, the solution $\bar{\nu}$ of \eqref{eq:nu} satisfies $- (g_l^d(i_{k+1}) + \Phi_l^d(i_{k+1})) > \bar{\nu} > - (g_l^d(i_k) + \Phi_l^d(i_k))$.
-Furthermore, because sigmoid is monotonous,
+Furthermore, because sigmoid is monotonically increasing,
 
@@
 \section{Appendix C}
+\begin{figure}[H]
+\centering{
+\includegraphics[width=0.5\textwidth]{multi_domain}}
+\caption{Visualization of domains according to their dropout masks (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
+\label{fig:domain}
+\end{figure}
+\section{Appendix D}
 \begin{breakablealgorithm}
@@
   \item number of groups: $n_p$; number of retained groups: $k$ 
-  \item $i=0$
+  \item $i=0$; $iter\_num$
  \end{itemize}}
@@
 $$\frac{\partial D}{\partial \hat{m}_l^d(\tau)}$$}
-\STATE{Compute implicit gradient of the $\operatorname{Soft-Top-K}$ masks over $\Phi^d$
+\STATE{Compute implicit gradient of the $\operatorname{Soft-Top-K}$ masks over $\Phi_l^d$
 $$- \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}$$}
-\STATE{Compute the gradient the training over $\Phi$
-$$ - \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d} + \frac{\partial \mathbb{H} \big[ \operatorname{softmax}(\Phi_l^d)\big] }{\partial \Phi_l^d} $$}
-\STATE{Update $\theta$ and $\Phi^d$ with their gradients}
+\STATE{Compute the gradient the training over $\Phi_l^d$
+$$\Delta_{\Phi_l^d} = - \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d} + \frac{\partial \mathbb{H} \big[ \operatorname{softmax}(\Phi_l^d)\big] }{\partial \Phi_l^d} $$}
+\STATE{Update $\theta$ and $\Phi_l^d$ with their gradients}
 \STATE{$i=i+1$}
@@
 \end{breakablealgorithm}
-\section{Appendix D}\fyTodo{fix this}
-\begin{figure*}[h!]
-\includegraphics[width=0.7\textwidth]{multi_domain}
-\caption{Visualization of domains according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
-\label{fig:domain}
-\end{figure*}
+
 \end{document}
+
+%\en

### 2021-11-25 2022/acl/acl_latex.tex
remove x from Q(m | x,d,\Phi)

@@
   D(x,y,d; \theta, \phi) = & \mathbb{E}_{m^d \sim Q(m^d |d,\Phi)} \log P(y | m^d, x, d; \theta) \nonumber\\
-  R(x,y,d; \theta, \phi) = & \operatorname{KL}(Q(m^d | x,d, \Phi)||P(m^d | x,d; \theta)), \nonumber
+  R(x,y,d; \theta, \phi) = & \operatorname{KL}(Q(m^d | d, \Phi)||P(m^d | x,d; \theta)), \nonumber
 \end{align}
@@
 \hspace{-2.em}
-  R    &= \operatorname{KL}(Q(m^d | x, d, \Phi)||P(m^d | x, d; \theta)), \nonumber \\ 
-	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(m_l^d | x, d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big) \nonumber \\ 
-	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1,\cdots,i_k | x, d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big)\nonumber \\
-	&\leqslant  -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \big). \label{eq:entropy}
+  R    &= \operatorname{KL}(Q(m^d | d, \Phi)||P(m^d | x, d; \theta)), \nonumber \\ 
+	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(m_l^d | d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big) \nonumber \\ 
+	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1,\cdots,i_k | d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big)\nonumber \\
+	&\leqslant  -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1 | d, \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \big). \label{eq:entropy}
 \end{align}
@@
 For these experiments, our main results are in Table ~\ref{tab:mdmt}, where we observe that the \system{LaMGD} Tranformer achieves a significant improvement (+2.78) over the generic \system{Transformer} system with zero extra parameters.
-Moreover, \system{LaMGD} Tranformer achieves performance that are equivalent on average to that of the \system{Adapter} sytems, which is carrefully finetuned and contains approximately 12.4M additional parameters.
+Moreover, \system{LaMGD} Tranformer achieves performance that are equivalent on average to that of the \system{Adapter} sytems, which is carrefully finetuned and contains approximately 25M additional parameters per domain.
 \begin{table*}[h!]
@@
     \system{Transformer}  \hfill{\footnotesize[65m]} & 40.3 & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[+12.4m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.3 \\ 
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.3 \\ 
     \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & 39.0 & 52.4 & 87.5 & 55.3 \\ 
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{law}} & \multicolumn{1}{c|}{\domain{law$_1$}} & \multicolumn{1}{c|}{\domain{law$_2$}} \\ \hline 
-    \system{Adapter}   \hfill{\footnotesize[+25m]}  & 61.0 & 59.4 (-0.6) &  59.2 (-0.8) \\ 
-    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  & 61.4 & 61.4 (=) & 61.4 (=) \\ 
+    \system{Adapter}   \hfill{\footnotesize[+151m]}  & 61.0 & 60.4 (-0.6) &  60.2 (-0.8) \\ 
+    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  & 60.4 & 60.4 (=) & 60.4 (=) \\ 
     \hline
@@
     \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.8 &7.3&20.8&21.1&39.7&19.8&22.6&15.2&18.9 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &4.3&6.8&21.1&22&39.7&20&23&15.2&19 \\ 
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &4.3&6.8&21.1&22&39.7&20&23&15.2&19 \\ 
     \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.33 \\ 
@@
     \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.4&16.6&28.5&	27.1&43.7&24.6&30.3&25.6&25.98 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.64\\ 
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.64\\ 
     \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.31 \\ 
@@
     \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&\SB{34.0}&38.3&8.3&20.19 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &10.2&4&\SB{13.3}&21.9&32.2&\SB{34.1}&38.5&8.3&20.31 \\ 
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &10.2&4&\SB{13.3}&21.9&32.2&\SB{34.1}&38.5&8.3&20.31 \\ 
     \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
@@
     \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19.0&27.43 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.5&9.4&20.0&30.6&37.2&38.2&39.3&19.0&27.03\\ 
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &22.5&9.4&20.0&30.6&37.2&38.2&39.3&19.0&27.03\\ 
     \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40.0&18.9&27.75 \\
@@
   \end{tabular}
-  \caption{Multilingual sub-networks' similarities (M2O Ted-related)}
+  \caption{Multilingual sub-networks' similarities (M2O Ted-related). Boldface denotes the similarity of languages in a same family.}
   \label{tab:related-sim}

### 2022-01-12 2022/acl/acl_latex.tex
replace table of similarities by heatmaps, add results on the effect of K, add rebutal

@@
 \begin{align*}
-  r_l^d(i) &= \begin{cases}
-    1, & \text{if}\ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p} \\
-    & \text{AND} \  m_l^d(p) \text{\small =} 1 \\
-    0, & \text{otherwise},
-  \end{cases} 
+  r_l^d(i) = m_l^d(p) \ \ \ \text{if}\ \ \ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p}
+%  \begin{cases}
+%    1, & \text{if}\ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p} \\
+%    & \text{AND} \  m_l^d(p) \text{\small =} 1 \\
+%    0, & \text{otherwise},
+%  \end{cases} 
   % p & \in  \{0,\dots,n_p\text{-}1 \}, \\
@@
 \end{align*}
-where $\operatorname{LAYER}^l()$ represents all the computations in Transformer layer $l$. It is applied at all positions of each layer in the encoder and in the decoder.  
+where $\operatorname{LAYER}^l()$ represents all the computations in Transformer layer $l$, $\odot$ is element-wise product.
+It is applied at all positions of each layer in the encoder and in the decoder.
 % The dropping mask $r_l^d$ is computed as follows
@@
 This section compares the sub-networks learnt for each domain or language pair by computing the average similarity between the corresponding dropout masks concatenated for all the layers of the underlying model.
-For the multidomain experiment, we analyze the case of pseud-domain separation reported in Section~\ref{ssec:fuzzy} in Table~\ref{tab:fuzzy-sim}.
+For the multidomain experiment, we analyze the case of pseud-domain separation reported in Section~\ref{ssec:fuzzy} in Figure ~\ref{fig:md-heatmap}.
 We see that the sub-networks for \domain{law$_1$} and \domain{law$_2$} are identical, yielding a full sharing between the corresponding training sets.
@@
 
-\begin{table*}[h!]
+\begin{figure}[h!]
+\begin{subfigure}{0.5\textwidth}
   \centering
-  \begin{tabular}{|p{1cm}|*{7}{r|}} \hline
-    & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{bank}}& \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{law$_1$}} & \multicolumn{1}{c|}{\domain{law$_2$}} & \multicolumn{1}{c|}{\domain{rel}} &\multicolumn{1}{c|}{\domain{talk}} \\ \hline
-    \domain{med} &1.0&0.80&0.85&0.83&0.83&0.78&0.89\\
-    \domain{bank} &0.80&1.0&0.80&0.90&0.90&0.73&0.82\\
-    \domain{it} &0.85&0.80&1.0&0.80&0.80&0.78&0.85\\
-    \domain{law$_1$} &0.83&0.90&0.80&1.0&\SB{1.0}&0.75&0.82\\ 
-    \domain{law$_2$} &0.83&0.90&0.80&\SB{1.0}&1.0&0.75&0.82\\ 
-    \domain{rel} &0.78&0.73&0.78&0.75&0.75&1.0&0.77\\
-    \domain{talk} &0.89&0.82&0.85&0.82&0.82&0.77&1.0\\
-    \hline
-  \end{tabular}
-  \caption{Multidomain experiments: sub-network similarities}
-  \label{tab:fuzzy-sim}
-\end{table*}
+  % include first image
+  \includegraphics[width=0.7\textwidth]{multi_domain_heatmap.png}  
+  \caption{Multidomain}
+  \label{fig:md-heatmap}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=0.7\textwidth]{multilingual_heatmap.png}  
+  \caption{Multilingual}
+  \label{fig:ml-heatmap}
+\end{subfigure}
+\caption{Heatmap visualization of the similarities between dropout masks of domains(languages)}
+\label{fig:heatmap}
+\end{figure}
+
+%\begin{table*}[h!]
+%  \centering
+%  \begin{tabular}{|p{1cm}|*{7}{r|}} \hline
+%    & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{bank}}& \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{law$_1$}} & \multicolumn{1}{c|}{\domain{law$_2$}} & \multicolumn{1}{c|}{\domain{rel}} &\multicolumn{1}{c|}{\domain{talk}} \\ \hline
+%    \domain{med} &1.0&0.80&0.85&0.83&0.83&0.78&0.89\\
+%    \domain{bank} &0.80&1.0&0.80&0.90&0.90&0.73&0.82\\
+%    \domain{it} &0.85&0.80&1.0&0.80&0.80&0.78&0.85\\
+%    \domain{law$_1$} &0.83&0.90&0.80&1.0&\SB{1.0}&0.75&0.82\\ 
+%    \domain{law$_2$} &0.83&0.90&0.80&\SB{1.0}&1.0&0.75&0.82\\ 
+%    \domain{rel} &0.78&0.73&0.78&0.75&0.75&1.0&0.77\\
+%    \domain{talk} &0.89&0.82&0.85&0.82&0.82&0.77&1.0\\
+%    \hline
+%  \end{tabular}
+%  \caption{Multidomain experiments: sub-network similarities}
+%  \label{tab:fuzzy-sim}
+%\end{table*}
 
@@
 The similarities between the sub-networks for each pair are respectively 0.95, 0.94, 0.99 and 0.96: for each language, the most similar language is always from the same language family.
-\begin{table*}[h!]
-  \centering
-  \begin{tabular}{|p{1cm}|*{8}{r|}} \hline
-    & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{bel}}& \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} &\multicolumn{1}{c|}{\domain{slk}}&\multicolumn{1}{c|}{\domain{tur}} \\ \hline
-    \domain{aze} &1.0&0.90&0.89&0.86&0.85&0.91&0.89&\SB{0.95}\\
-    \domain{bel} &0.90&1.0&0.94&0.91&0.91&\SB{0.94}&0.94&0.88\\
-    \domain{ces} &0.89&0.94&1.0&0.92&0.94&0.98&\SB{0.99}&0.91\\
-    \domain{glg} &0.86&0.91&0.92&1.0&\SB{0.96}&0.92&0.93&0.86\\ 
-    \domain{por} &0.85&0.91&0.94&\SB{0.96}&1.0&0.93&0.94&0.87\\ 
-    \domain{rus} &0.91&\SB{0.94}&0.98&0.92&0.93&1.0&0.98&0.93\\
-    \domain{slk} &0.89&0.94&\SB{0.99}&0.93&0.94&0.98&1.0&0.91\\
-    \domain{tur} &\SB{0.95}&0.88&0.91&0.86&0.87&0.93&0.91&1.0\\
-    \hline
-  \end{tabular}
-  \caption{Multilingual sub-networks' similarities (M2O Ted-related). Boldface denotes the similarity of languages in a same family.}
-  \label{tab:related-sim}
-\end{table*}
+We provide in Figure~\ref{fig:ml-heatmap} the heatmap of the similarities between the dropout masks of our objective languages.
+%\begin{table*}[h!]
+%  \centering
+%  \begin{tabular}{|p{1cm}|*{8}{r|}} \hline
+%    & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{bel}}& \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} &\multicolumn{1}{c|}{\domain{slk}}&\multicolumn{1}{c|}{\domain{tur}} \\ \hline
+%    \domain{aze} &1.0&0.90&0.89&0.86&0.85&0.91&0.89&\SB{0.95}\\
+%    \domain{bel} &0.90&1.0&0.94&0.91&0.91&\SB{0.94}&0.94&0.88\\
+%    \domain{ces} &0.89&0.94&1.0&0.92&0.94&0.98&\SB{0.99}&0.91\\
+%    \domain{glg} &0.86&0.91&0.92&1.0&\SB{0.96}&0.92&0.93&0.86\\ 
+%    \domain{por} &0.85&0.91&0.94&\SB{0.96}&1.0&0.93&0.94&0.87\\ 
+%    \domain{rus} &0.91&\SB{0.94}&0.98&0.92&0.93&1.0&0.98&0.93\\
+%    \domain{slk} &0.89&0.94&\SB{0.99}&0.93&0.94&0.98&1.0&0.91\\
+%    \domain{tur} &\SB{0.95}&0.88&0.91&0.86&0.87&0.93&0.91&1.0\\
+%    \hline
+%  \end{tabular}
+%  \caption{Multilingual sub-networks' similarities (M2O Ted-related). Boldface denotes the similarity of languages in a same family.}
+%  \label{tab:related-sim}
+%\end{table*}
 
@@
 % \end{figure*}
-
+\subsection{Ablation study}
+We provide more analysis over the choice of hyperparameters $k$ which is the number of activated nodes in each layer, i.e, controls the sharing level between the tasks.
+Table ~\ref{tab:K} shows the degradation of the performance when the number of activated nodes is reduced, and the sharing level between tasks decreases in consequence.
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{|c|c|} \hline
+    K & \domain{avg}\\ \hline 
+    8 & 18.1 \\
+    10 & 19.15 \\
+    12 & 19.33 \\
+    14 & 19.44 \\
+    \hline
+  \end{tabular}
+  \caption{Variation of the performance w.r.t K while we fix $n_p=16$}
+  \label{tab:K}
+\end{table}
 \section{Related work}\label{sec:related}
 Multidomain and multilingual translation systems have received considerable attention in the recent years, and a exhaustive survey is beyond the goal of this paper.
-Domain adaptation for neural MT is surveyed in \citep{Chu17comparison}, while multidomain MT systems are notably studied in \cite{Saunders21Asurvey,Pham21revisiting}; for multilingual MT, the reader is refered eg.\ to \citep{Chu18multilingual,dabre20survey}.
+Domain adaptation for neural MT is surveyed in \citep{Chu17comparison}, while multidomain MT systems are notably studied in \cite{Saunders21Asurvey,Pham21revisiting}; for multilingual MT, the reader is referred eg.\ to \citep{Chu18multilingual,dabre20survey}.
 We focus on the most relevant subset of this literature below.
@@
 However, our method learn the selection from a latent commonality in data instead of using a predefined selection such as "One language family per decoder" in \cite{kong21multilingual}.
+
+\textbf{Language-specific sub-networks}.
+\citet{Frankle19Lottery,liu18rethinking} proposed different techniques to locate most important parameters for the task so that when training the model while masking other parameters the performance of the model does not degrade.\citet{lin21learning} hypothesized that among parameters of a multilingual NMT model, there exists a subset of parameters more biased to a specific language.
+The authors proposed to locate these subsets by pruning the parameters of each fine-tuned model.
+Our approach also aims to assign a sub-network to each task by masking the output of each layer rather than masking the parameters.
+Furthermore, \citet{lin21learning} computes the masks via a heuristic selection while our approach learn the masks via a variational technique.
 

### 2022-01-12 2022/acl/responses.tex
replace table of similarities by heatmaps, add results on the effect of K, add rebutal

\documentclass[12pt,times,a4paper,twoside]{article}

\usepackage[utf8]{inputenc}	% Para caracteres en español
\usepackage[left=3cm,right=3cm,top=2cm,bottom=3cm]{geometry}
\usepackage{pagenote}

\usepackage{algorithm}
\usepackage{algorithmic}

\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{mathtools,xparse}
\usepackage{mathrsfs}

\usepackage{multirow,booktabs}
\usepackage[table]{xcolor}
\usepackage{fullpage}
\usepackage{lastpage}
\usepackage{enumitem}
\usepackage{fancyhdr}

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{calc}
% \usepackage{color,soul}
\usepackage{multicol}
\usepackage{cancel}
\usepackage[retainorgcmds]{IEEEtrantools}
% \usepackage{xcolor} -- implied by xcolor

\usepackage{listings}
\usepackage{spverbatim}
\usepackage{fancyvrb}
\usepackage{hyperref}
\usepackage{float}
\usepackage{natbib}

\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}

\usepackage{soulutf8}
\usepackage{tabularx}

\usepackage[draft]{todo}
\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}

\colorlet{shadecolor}{orange!15}
\parindent 0in
\parskip 12pt
\geometry{margin=1in, headsep=0.25in}
\setlength{\belowdisplayskip}{8pt} \setlength{\belowdisplayshortskip}{8pt}
\setlength{\abovedisplayskip}{8pt} \setlength{\abovedisplayshortskip}{8pt}
\setlist{nosep}
\setlength{\parskip}{0.1cm}
\setlength{\parindent}{1em}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{definition}{Definition}[section]
\DeclareMathOperator*{\argmax}{argmax}
\newcommand{\R}{\ensuremath{\mathbb{R}}}

\pgfkeys{
    /tr/rowfilter/.style 2 args={
        /pgfplots/x filter/.append code={
            \edef\arga{\thisrow{#1}}
            \edef\argb{#2}
            \ifx\arga\argb
            \else
                \def\pgfmathresult{}
            \fi
        }
    }
}

\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\NewDocumentCommand{\normL}{ s O{} m }{%
  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{L_1}%
}

\title{Responses to the reviews of ACL rolling review paper \#162:
  ``Latent Group Dropout for Multilingual and Multidomain Machine Translation''
}
\date{}

\begin{document}
\maketitle

\textcolor{blue}{We thank the reviewers for sharing their suggestions which could improve the clarity of our paper and strengthen our claim over the efficacy of our proposed method}

\section*{Reviewer iMf4}
\subsubsection*{Paper Summary:}

This paper considers multidomain and multilingual NMT as multi-task learning and proposed a method to automatically determine task-specific and task-agnostic parameters in the neural networks without explicitly defining or adding any additional layers. learning task dependent sub-networks. They formulate such problem as latent dropout mask learning in which the parameters in each transformer layer are divided into multiple groups and learn to select the k active groups for each task so that the multitask loss is minimized. Experiments in multidomain NMT and multilingual NMT show that their proposed methods achieve better average BLEU scores than the vanilla transformer and adapter methods.

\subsubsection*{Summary of Strengths:}

\begin{itemize}
\item The problem setting is interesting and well-formulated.
\end{itemize}

\subsubsection*{Summary of Weaknesses:}

\begin{itemize}
\item Missing citation and baseline: Learning task-specific and task-agnostic parameters in multitask learning for NMT has been proposed in Lin et al.2021. In this work, they learn a mask for each neuron instead of group of neurons. You should compare the proposed LaMGD method with this baseline.
\item Lack of ablation study on the dropout group selection. What if the groups are selected randomly or using some predefined heuristics, e.g. first 2 groups are shared across tasks, some other groups for language family, or a single language.

\item You should have an analysis on the effect of k which potentially controls the level of sharing between tasks.

\item In table 5, although the average BLEU scores of LaMGD is higher, most of the results are not statistically significant.
  \end{itemize}

\subsubsection*{Comments, Suggestions And Typos:}

I Table 7 can be improved in presentation by using heatmap.

\textbf{Overall Assessment:} 2

\textbf{Confidence:} 3 =  Pretty sure, but there's a chance I missed something. Although I have a good feel for this area in general, I did not carefully check the paper's details, e.g., the math or experimental design.

\textbf{Best Paper:} No

\textbf{Replicability:} 2 = They would be hard pressed to reproduce the results: The contribution depends on data that are simply not available outside the author's institution or consortium and/or not enough details are provided.

\textbf{Datasets:} 1 = No usable datasets submitted.

\textbf{Software:} 2 = Documentary: The new software will be useful to study or replicate the reported research, although for other purposes it may have limited interest or limited usability. (Still a positive rating)

\textbf{Author Identity Guess:} 1 = I do not have even an educated guess about author identity.

\subsubsection*{Rebuttal}
{\color{blue}%
We thank the reviewer for his suggestion on a citation that we accidentally missed. For the sake of the clarification of our approach, we would like to claim several misunderstandings in the summary made by Reviewer iMf4:
\begin{itemize}
\item Our approach does not aim to find task-specific and task-agnostic parameters in the neural network. The objective of our algorithm is to optimize the assignment of a sub-group of "nodes", which are the dimensions of the output of each intermediate layer, to each task.
\item Our approach does not mask the sub-parts of each weight matrix but masks a fixed number of sub-groups of elements of the output of each intermediate layer.
\end{itemize}
We would like to answer the concerns of the reviews as follows.
\begin{itemize}
\item We admit the weakness in our choice of baseline according to the reviewer. However, our claim is that embedding the similarity between tasks in the construction of the multi-task network is beneficial, and the fine-tuning-based approaches ignore these similarities. We prove our claim by proposing a method that automatically optimizes the structure of a multi-task network by learning from data, i.e., implicitly learning the similarity between tasks and embedding it to the organization of the network. Besides, the Adapters are a robust method for multi-task learning, and a comparison with this baseline strongly supports our claim over the efficiency of our proposed method.
\item We are happy to provide more comparisons with heuristic strategies of the dropout group selection in the final version. The flaws of these heuristic strategies motivate us to develop this technique.
\item We have finished the analysis on the effect of $k$ on the performance and will provide them in the final version.
\item The multilingual experiments use two groups of 8 languages, which are all low-resourced; this might lower the significance of the improvement as observed in the paper of Lin et al. 21 (only 0.7 for low). Our method significantly improves the minor languages among these languages.
\end{itemize}

Finally we hope the reviewer change the view on the objective of our paper, reconsider the solid mathematical foundation of our method supported by an explicit implementation explained in the Appendix, the improvement with high margin in multi-domain, the reduced computational cost in the inference compared to the Adapters, and the novelty in the use of node-masking instead of weight-masking as in previous work such as in Lin et al.}
\section*{Reviewer 681g}

\subsubsection*{Paper Summary:}

This paper proposes a routing strategy for multitask machine translation. For a given language or domain, only a fixed amount of nodes is activated at each layer and the model learns which nodes should be activated. They experiment on both multilingual and multidomain settings, and the experimental results demonstrate that they can achieve comparable and sometimes marginally better performance than adapter-based methods with fewer parameters.

\subsubsection*{Summary of Strengths:}
\begin{itemize}
\item The general idea is intuitive and makes sense, and the paper proposes a technically sound way of implementing it.
\item The empirical results are good as they can achieve comparable performance with an adapter-based method with fewer parameters.
\item The paper is well-organized.
\end{itemize}

\subsubsection*{Summary Of Weaknesses:}

\begin{itemize}
\item There is little analysis of their methods. Sensitivity analysis of their hyper-parameters $(n_d,k,\tau)$, some qualitative analysis are required.
\item Requiring partitioning nodes into groups and a fixed number of nodes to be activated at each layer makes their method less flexible.
\item More baselines should be included (e.g. Li et al., (2020), Gong et al., (2021a, b) as mentioned in the paper).
\end{itemize}

\subsubsection*{Comments, Suggestions And Typos:}
It'd be good to also include a mixture-of-experts baseline.

\textbf{Overall Assessment:} 3 = Good: This paper is of interest to the *ACL audience and could be published, but might not be appropriate for a top-tier publication venue. It would likely be a strong paper in a suitable workshop.

\textbf{Confidence:} 4 = Quite sure. I tried to check the important points carefully. It's unlikely, though conceivable, that I missed something that should affect my ratings.

Best Paper: No

\textbf{Replicability:} 3 = They could reproduce the results with some difficulty. The settings of parameters are underspecified or subjectively determined, and/or the training/evaluation data are not widely available.

\textbf{Datasets:} 1 = No usable datasets submitted.

\textbf{Software:} 1 = No usable software released.

\textbf{Author Identity Guess:} 1 = I do not have even an educated guess about author identity.

\subsubsection*{Rebuttal}

{\color{blue}
We thank the reviewer for appreciating our mathematical work to support our implementation and the suggestions for a better understanding of our proposed approach. We would like to answer the reviewer's concerns as follows.

\begin{itemize}
\item We did admit our lack of analysis on the hyperparameters in the conclusion of our paper. We have collected more data for better analyzing the effect of hyper-parameters on the performance of our method since the submission. We will provide these results in our final version.
\item We agree with the reviewer. However, we could reduce the group size to 1 node only and use hierarchical modeling to optimize the number of nodes to be activated. These changes would improve the method's flexibility in exchange for computational cost in training.
\item We are aware of these works. However, we do not attempt to beat a race in the translation performance in multilingual translation or multi-domain translation. In this work, our claim is that embedding the similarity between tasks in the construction of the multi-task network is beneficial, and the fine-tuning-based approaches ignore these similarities. And, we prove our claim by proposing a method that automatically optimizes the structure of a multi-task network by learning from data, i.e., implicitly learning the similarity between tasks and embedding it to the organization of the network
\end{itemize}

We agree that a mixture of experts like Gshard is an excellent candidate for the baseline. We have been working on their implementations but lack the computational resource for training. Finally we hope the reviewer change the view on the objective of our paper and reconsider our solid mathematical modeling supported by supported by an explicit implementation explained in the Appendix.
}
\section*{Reviewer 3koi}

\subsubsection*{Paper Summary:}

This paper proposes a latent-variable model based on a variational probabilistic modeling framework for conditional compute in multi-domain and multilingual machine translation. They introduce latent variables which are learned during model training. These latent variables decide which sub-network of the model is selected for each task / language pair. The results on multi-domain machine translation show promise compared to vanilla Transformer and Adapter baselines. The results on multi-lingual machine translation are somewhat mixed with modest improvements in 3 settings and worse accuracy in 1 setting. The paper includes some interesting analysis of the models as well. On the other hand, some important ablations/hyper-parameter tuning experiments are missing and it is unclear how statistically significant the results on multilingual translation are.

\subsubsection*{Summary of Strengths:}

\begin{itemize}
\item This work proposes a technique to learn which subnetworks to use for which task / language pair based on learning from the data.
\item The proposed probabilistic framework is mathematically sound and explained clearly and completely.
\item There are experiments compared with decent baselines and multiple benchmarks and there has been a good attempt to evaluate the new technique well.
\item The analysis showing that accuracy of Adapter baselines declines on a domain when that domain is divided into two pseudo-domains, where as the accuracy on this proposed framework remains the same motivates the reason of using the proposed technique very well
\item A nice advantage of this approach is that it doesn’t need additional parameters to the added to the baseline Transformer model. There are some additional hyper-parameters to tune correctly per setup though.
\item There is also analysis showing whether the learned subnetworks for related domains/languages overlap significantly.

\end{itemize}

\subsubsection*{Summary of Weaknesses:}

\begin{itemize}
\item It is unclear how the latent variables that decide the subnetworks to be used per task/domain are initialized at the start of model training.
\item While the authors mention leaving this to future work, the study of how many groups to divide each layer into and how many groups to dropout per layer is missing.
\item The ablation about the effect of tuning temperature parameter $\tau$ is missing.
\item It is unclear whether the baselines have been allowed to train all the way till convergence. The multi-domain baseline model is trained for 20K updates, while the proposed model is trained for 300K updates. This looks weird.
\item Similarly, but to a lesser extent, the multilingual MT baselines are trained for 40K/(40K+5K) updates. The proposed model is trained for 50K updates. So it is unclear if the baselines have been given enough chance to convergence well.
\item It is unclear how statistically significant the improvements or decline in accuracy in the 4 setups for multilingual MT are.
\end{itemize}

\subsubsection*{Comments, Suggestions and Typos:}
[nice to have] comparison to Adapter baseline with the same number of parameters and approximate FLOPs per update as the Transformer baseline and the proposed approach would be useful.
While the proposed approach and results are quite interesting, there is some need to qualify some of the claims in the Introduction and Conclusion such as “very little extra computational cost” (it looks like training cost increases by 33$\%$); “significant gains of this method with respect to baselines” (the gains are great in the multi-domain benchmark, but modest in the multilingual translation benchmark).

\textbf{Overall Assessment:} 3 = Good: This paper is of interest to the *ACL audience and could be published, but might not be appropriate for a top-tier publication venue. It would likely be a strong paper in a suitable workshop.

\textbf{Confidence:} 3 =  Pretty sure, but there's a chance I missed something. Although I have a good feel for this area in general, I did not carefully check the paper's details, e.g., the math or experimental design.

\textbf{Best Paper:} No

\textbf{Replicability:} 2 = They would be hard pressed to reproduce the results: The contribution depends on data that are simply not available outside the author's institution or consortium and/or not enough details are provided.

\textbf{Datasets:} 1 = No usable datasets submitted.

\textbf{Software:} 1 = No usable software released.

\textbf{Author Identity Guess:} 1 = I do not have even an educated guess about author identity.

\subsubsection*{Rebuttal}
{\color{blue}%
We thank the reviewer for appreciating our mathematical modeling and the suggestions for a better understanding of our proposed technique. We would like to answer the concerns of the reviewer as follows.

\begin{itemize}
\item We initialize the probabilistic distribution of the latent variables as uniform distribution.
\item We will provide more data on how $k$ affect the performance of our approach in the final version.
\item As reported from many works using the Gumbel reparameterization trick, the effect of $\tau$ is relatively insignificant. Therefore, we focused more on selecting $k$.
\item Because we would like to have the same total amount of training iterations between Adapter-based and LaMGD, we extend the training of LaMGD to 300K. The convergence of the standard Transformer is before 200K as its validation curve became flat near the 200K-th iteration.
\item The previous reason is applied for the reviewer's concern over multilingual experiments.
\item As we pointed out, the significant gains are observed in several very low-resourced languages, not globally. This might be explained by the fact that low-resourced languages benefit from the sharing nodes with their related well-resourced languages while less suffering from the inference of unrelated languages.
\end{itemize}
Our claim on the advantage in computational cost repose on the inference stage while there are no extra parameters added to the standard Transformer compared to Adapter-based methods. However, the training time will cost more than usual because of the joint optimization of the network's organization and the parameters. Finally, we hope the reviewer change the view on the objective of our paper and reconsider our solid mathematical modeling supported by the explicit implementation explained in Appendix.
}%

\end{document}

### 2022-01-12 2022/acl/acl_latex.tex
correct the heatmap

@@
 \end{subfigure}
-\caption{Heatmap visualization of the similarities between dropout masks of domains(languages)}
+\caption{Heatmap visualization of the similarities between dropout masks of domains(languages). Brighter colors correspond to higher similarity scores.}
 \label{fig:heatmap}
@@
 For multilingual (TED-related) experiments, the training data contains four language families: (1) Turkic, with Azerbaijani and Turkish(\domain{aze},\domain{tur}); (2) Slavic, with Belarusian and Russian (\domain{bel},\domain{rus}); (3) Romance, with Galician and Portuguese (\domain{glg}, \domain{por}); and (4) Czech-Slovak, with Slovak and Czech (\domain{ces}, \domain{slk}).
-The similarities between the sub-networks for each pair are respectively 0.95, 0.94, 0.99 and 0.96: for each language, the most similar language is always from the same language family.
 We provide in Figure~\ref{fig:ml-heatmap} the heatmap of the similarities between the dropout masks of our objective languages.
+We observe that each pair of languages in the same family correspond to brightest color except the diagonal in every column or every row.
 %\begin{table*}[h!]

### 2022-01-13 2022/acl/acl_latex.tex
add analysis

@@
 Furthermore, we also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and that its gradients can be computed using the implicit differentiation theorem.
-During training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyperparameter $\tau$ to $0$.
+During training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyper-parameter $\tau$ to $0$.
 The gradient of $D$ w.r.t $\Phi_l^d$ is computed using the chain rule as follows:
@@
 
-\subsubsection{Hyperparameters}
+\subsubsection{Hyper-parameters}
 \label{ssec:hyperparams}
 We choose $n_d = 16$ so that the size of the dropout group is neither too small nor too large.
-The second important hyperparameter in \system{LaMGD} is the number of selected groups in each layer, $k$, which we set to 12 in every experiments.
+The second important hyper-parameter in \system{LaMGD} is the number of selected groups in each layer, $k$, which we set to 12 in every experiments.
 By retaining $12/16$ groups, we share on average $75 \%$ active groups between two domains or languages.
@@
 Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments.
-The choice of the values for our hyperparameters was not optimized and selected by suggestions from previous studies \citep{Jang17categorical,Gong21pay,Gong21adaptive}.
+The choice of the values for our hyper-parameters was not optimized and selected by suggestions from previous studies \citep{Jang17categorical,Gong21pay,Gong21adaptive}.
 
@@
 \subsection{Ablation study}
-We provide more analysis over the choice of hyperparameters $k$ which is the number of activated nodes in each layer, i.e, controls the sharing level between the tasks.
-Table ~\ref{tab:K} shows the degradation of the performance when the number of activated nodes is reduced, and the sharing level between tasks decreases in consequence.
+We discuss here the choice of the hyper-parameters $k$, the number of activated nodes in each layer, and its impact on the sharing level between the tasks.
+Table ~\ref{tab:K} shows the variance of performance when the number of activated nodes is changed, and the sharing level between tasks decreases in consequence.\Todo{add sharing results.
+}\Todo{check caption consistency}
 \begin{table}[h!]
@@
   \begin{tabular}{|c|c|} \hline
-    K & \domain{avg}\\ \hline 
+    $k$ & \domain{avg}\\ \hline 
     8 & 18.1 \\
@@
   \end{tabular}
-  \caption{Variation of the performance w.r.t K while we fix $n_p=16$}
+  \caption{Variation of the performance w.r.t $k$, while we fix $n_p=16$. These results correspond to the multilingual setting (o2m-diverse).}
   \label{tab:K}
 \end{table}
+
 \section{Related work}\label{sec:related}

### 2022-01-13 2022/acl/responses.tex
started work on responses

@@
 
-\textcolor{blue}{We thank the reviewers for sharing their suggestions which could improve the clarity of our paper and strengthen our claim over the efficacy of our proposed method}
+\textcolor{blue}{We thank the reviewers for sharing their suggestions which will help us improve the clarity of our paper and strengthen the claim regarding the efficacy of our method.}
 
@@
 
-\item You should have an analysis on the effect of k which potentially controls the level of sharing between tasks.
+\item You should have an analysis on the effect of $k$ which potentially controls the level of sharing between tasks.
 
@@
 {\color{blue}%
-We thank the reviewer for his suggestion on a citation that we accidentally missed.
-For the sake of the clarification of our approach, we would like to claim several misunderstandings in the summary made by Reviewer iMf4:
+  We thank the reviewer for pointing additional relevant work. We cite and discuss it in the revised version.
+
+For the sake of the clarification of our approach, we would like to point out several misunderstandings in the summary made by Reviewer iMf4:
 \begin{itemize}
-\item Our approach does not aim to find task-specific and task-agnostic parameters in the neural network.
+\item Our approach does not aim to find task-specific and task-agnostic parameters in the neural network: there is no notion of domain specific or domain agnostic parameters.
 The objective of our algorithm is to optimize the assignment of a sub-group of "nodes", which are the dimensions of the output of each intermediate layer, to each task.
@@
 \end{itemize}
-We would like to answer the concerns of the reviews as follows.
-\begin{itemize}
-\item We admit the weakness in our choice of baseline according to the reviewer.
-However, our claim is that embedding the similarity between tasks in the construction of the multi-task network is beneficial, and the fine-tuning-based approaches ignore these similarities.
+We have tried to improve our presentation of our method in the revised version of the paper.\todo{Change presention, give line numbers}
+
+We would like to answer the other concerns of the reviewier as follows.
+\begin{itemize} CLAIM : baseline + parameter sharing + resultats CDR
+\item Regarding baselines, we insist ur claim is that embedding the similarity between tasks in the construction of the multi-task network is beneficial, and the fine-tuning-based approaches ignore these similarities.
 We prove our claim by proposing a method that automatically optimizes the structure of a multi-task network by learning from data, i.e., implicitly learning the similarity between tasks and embedding it to the organization of the network.
@@
 The flaws of these heuristic strategies motivate us to develop this technique.
-\item We have finished the analysis on the effect of $k$ on the performance and will provide them in the final version.
-\item The multilingual experiments use two groups of 8 languages, which are all low-resourced; this might lower the significance of the improvement as observed in the paper of Lin et al.
-21 (only 0.7 for low).
-Our method significantly improves the minor languages among these languages.
+\item We have included the analysis on the effect of $k$ on the performance in the revised version (see new section 4.4)
+\item The multilingual experiments use two groups of $8$ languages, which are all low-resourced; this might lower the significance of the improvement as observed in the paper of Lin et al.
+21 (only +0.7 for low resource).
+Our method significantly improves the less resource languages among these groups.
 \end{itemize}
 
-Finally we hope the reviewer change the view on the objective of our paper, reconsider the solid mathematical foundation of our method supported by an explicit implementation explained in the Appendix, the improvement with high margin in multi-domain, the reduced computational cost in the inference compared to the Adapters, and the novelty in the use of node-masking instead of weight-masking as in previous work such as in Lin et al.}
+\todo{Make this contribution} Finally, we hope the reviewer change the view on the objective of our paper, reconsider the solid mathematical foundation of our method supported by an explicit implementation explained in the Appendix, the improvement with high margin in multi-domain, the reduced computational cost in the inference compared to the Adapters, and the novelty in the use of node-masking instead of weight-masking as in previous work such as in Lin et al.}
 \section*{Reviewer 681g}
@@
 {\color{blue}
-We thank the reviewer for appreciating our mathematical work to support our implementation and the suggestions for a better understanding of our proposed approach.
+\tofo{Keep this ?}We thank the reviewer for appreciating our mathematical work to support our implementation and the suggestions for a better understanding of our proposed approach.
 We would like to answer the reviewer's concerns as follows.
@@
 \begin{itemize}
-\item We did admit our lack of analysis on the hyperparameters in the conclusion of our paper.
-We have collected more data for better analyzing the effect of hyper-parameters on the performance of our method since the submission.
-We will provide these results in our final version.
+\item % We did admit our lack of analysis on the hyperparameters in the conclusion of our paper.
+  We have now collected more data for better analyzing the effect of hyper-parameters on the performance of our method since the submission. These results in our revised version.
 \item We agree with the reviewer.
@@
 These changes would improve the method's flexibility in exchange for computational cost in training.
-\item We are aware of these works.
+\item We are aware of these works and they are discussed in related work section.
 However, we do not attempt to beat a race in the translation performance in multilingual translation or multi-domain translation.
 In this work, our claim is that embedding the similarity between tasks in the construction of the multi-task network is beneficial, and the fine-tuning-based approaches ignore these similarities.
-And, we prove our claim by proposing a method that automatically optimizes the structure of a multi-task network by learning from data, i.e., implicitly learning the similarity between tasks and embedding it to the organization of the network
+And, we prove our claim by proposing a method that automatically optimizes the structure of a multi-task network by learning from data, i.e., implicitly learning the similarity between tasks and embedding it to the organization of the network \Todo{Reponse à tous les reviewers}
 \end{itemize}
 
-We agree that a mixture of experts like Gshard is an excellent candidate for the baseline.
-We have been working on their implementations but lack the computational resource for training.
+We agree that a mixture of experts is an relevant alternative baseline.
+Again resource limitation in training --  We have been working on their implementations but lack the computational resource for training.
 Finally we hope the reviewer change the view on the objective of our paper and reconsider our solid mathematical modeling supported by supported by an explicit implementation explained in the Appendix.
@@
 \begin{itemize}
-\item We initialize the probabilistic distribution of the latent variables as uniform distribution.
-\item We will provide more data on how $k$ affect the performance of our approach in the final version.
-\item As reported from many works using the Gumbel reparameterization trick, the effect of $\tau$ is relatively insignificant. Therefore, we focused more on selecting $k$.
+\item We initialize the probabilistic distribution of the latent variables as uniform distribution. This is clarified in the revised text.\Todo{Check clarified which line}
+\item The revised version includes a discussion  regarding $k$ and how it affects the performance of our approach (section 4.4).
+\item As reported in many works using the Gumbel reparameterization trick, the effect of $\tau$ is relatively insignificant.
+This is discussed in the revised version in section.
+\Todo{which section}Therefore, we focused more on selecting $k$.
+\Todo{Add references}
 \item Because we would like to have the same total amount of training iterations between Adapter-based and LaMGD, we extend the training of LaMGD to 300K.
 The convergence of the standard Transformer is before 200K as its validation curve became flat near the 200K-th iteration.
+These has been clarified in the paper.
+\Todo{Which section ?}
 \item The previous reason is applied for the reviewer's concern over multilingual experiments.
@@
 Finally, we hope the reviewer change the view on the objective of our paper and reconsider our solid mathematical modeling supported by the explicit implementation explained in Appendix.
+\todo{Not so great.
+We cannot be sure the reviewers stay the same.}
 }%

### 2022-01-13 2022/acl/acl_latex.tex
add answers for todos, local changes

@@
 The number of parameters of the variational distribution is $L\times k \times L \times n_d$, which is negligible in comparison to the size of the Transformer model.
+\item \item Transformer using heuristic multi-task group dropout (\system{HMGD} Transformer): we share 320 nodes for every task, and reserve 32 nodes for each task (making $320 + 32*6 = 512$ nodes in total).
 \end{itemize}
@@
 We set the dropout probability to 0.1.
-We train the multi-domain Transformer model for 20k iterations with a batch size of 12k tokens.
-We train \system{LaMGD} Tranformer for 300k iterations with a batch size of 12k tokens: these extra iterations are justified by the need for the \system{LaMGD} Tranformer to learn both the sub-networks and the parameters of the underlying Transformer.\fyDone{Have we converged?} Finally, we plug domain adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline.
+We train the multi-domain Transformer model for 200k iterations with a batch size of 12k tokens.
+The convergence of the standard Transformer is before 200K as its validation curve became flat near the 200K-th iteration.
+The \system{LaMGD} Tranformer converged after 300k iterations with a batch size of 12k tokens.
+The convergence of \system{LaMGD} is controlled by its validation curve.
+Finally, we plug domain adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline.
+With approximately the same amount of training iterations, \system{LaMGD} and Adapter-based Transformer are trained.
 
@@
 We train the multilingual Transformer model for 40k iterations with a batch size of 9600 tokens on 16 V100 GPUs as in \citet{Gong21adaptive}.
-We train \system{LaMGD} Tranformer for 50k iterations with the same batch size.
+We train \system{LaMGD} Transformer for 50k iterations with the same batch size.
+The convergence of the models are controlled via their validation curves.
 Finally, we finetune the language-specific Adapters for 5k iterations.
@@
 in which $r=0.0001$.
-
-Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments.
-The choice of the values for our hyper-parameters was not optimized and selected by suggestions from previous studies \citep{Jang17categorical,Gong21pay,Gong21adaptive}.
+While \citet{Gong21pay,Gong21adaptive} fixed $\tau$ to be 0.2, we select an anneal policy for $\tau$ proposed by previous studies \citep{Jang17categorical}.
+Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments
 
 For all experiments, we report the BLEU score of \citet{Papineni02bleu} computed with SacreBleu \citep{Post18call}.\fyDone{Check this}
-
+\subsubsection{Latent variables' initialization}
+We initialize the probabilistic distribution of the latent variables by uniform distribution.
+More precisely, we set $\Phi_l^d$, which generates the probability of the masks via the softmax activation function, to $0^{n_d}$.
 \subsection{Datasets}
@@
 Moreover, \system{LaMGD} Tranformer achieves performance that are equivalent on average to that of the \system{Adapter} sytems, which is carrefully finetuned and contains approximately 25M additional parameters per domain.
+Variational mask learned from data by \system{LaMGD} also outperforms heuristic dropout mask \system{HMGD} by 0.5 in average.
 \begin{table*}[h!]
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[65m]} & 40.3 & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.3 \\ 
-    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & 39.0 & 52.4 & 87.5 & 55.3 \\ 
+    \system{Transformer}  \hfill{\footnotesize[65m]} & \SB{40.3} & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & \SB{61.0} & \SB{53.1} & 37.5 & 49.6 & \SB{91.0} & \SB{55.3} \\ 
+    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & \SB{40.3} & 60.4 & 52.4 & \SB{39.0} & \SB{52.4} & 87.5 & \SB{55.3} \\ 
+    \system{HMGD} Tranformer   \hfill{\footnotesize[+0m]}  & \SB{40.4} & 60.4 & 51.9 & 38.7 &	50.80 &	86.80 & 54.8 \\ 
     \hline
@@
   \centering
-  \begin{tabular}{|c|c|} \hline
-    $k$ & \domain{avg}\\ \hline 
-    8 & 18.1 \\
-    10 & 19.15 \\
-    12 & 19.33 \\
-    14 & 19.44 \\
+  \begin{tabular}{|c|c|c|} \hline
+    $k$ & \domain{avg} & sharing rate \\ \hline 
+    8 & 18.1 & 0.63\\
+    10 & 19.15 & 0.73\\
+    12 & 19.33 & 0.78\\
+    14 & 19.44 & 0.88\\
     \hline
   \end{tabular}
-  \caption{Variation of the performance w.r.t $k$, while we fix $n_p=16$. These results correspond to the multilingual setting (o2m-diverse).}
+  \caption{Variation of the performance w.r.t $k$, while we fix $n_p=16$. These results correspond to the multilingual setting (o2m-related).}
   \label{tab:K}

### 2022-01-13 2022/acl/responses.tex

@@
 \usepackage{pgfplotstable}
-
 \usepackage{soulutf8}
 \usepackage{tabularx}
+\usepackage{url}
 
@@
 }
+\author{}
 \date{}
@@
 
-\textcolor{blue}{We thank the reviewers for sharing their suggestions which will help us improve the clarity of our paper and strengthen the claim regarding the efficacy of our method.}
+\textcolor{blue}{We thank the meta-reviewer and reviewers for sharing their suggestions which will help us improve the clarity of our paper and strengthen the claim regarding the efficacy of our method.}
+
+\section*{Meta Review Tdeo}
+
+\subsubsection*{Meta-review:}
+
+This paper proposes a method for parameter sharing in multitask machine translation.
+For a given language or domain, the model learns which groups of nodes to activate in order to optimize the multitask loss.
+
+\subsubsection*{Summary Of Reasons To Publish:}
+
+The reviewers mostly agree that this is an interesting, well-presented method.
+Multitask training is very relevant for MT, and this method achieves comparable performance to an adapter-based method with fewer parameters.
+
+\subsubsection*{Summary Of Suggested Revisions:}
+
+From the reviews and the ensuing discussion, it seems that the paper would benefit from comparison to more baselines.
+In particular, comparisons to the paper highlighted by iMF4, or to other methods already mentioned in the paper: (Li et al., (2020), Gong et al., (2021a, b)).
+Some reviewers mention the need for more ablations and a better understanding of hyper-parameter sensitivity.
+
+\paragraph{}\textbf{Overall Assessment}: 3 = There are major points that may be revised
+\paragraph{}\textbf{Suggested Venues:} WMT.
+\paragraph{}\textbf{Ethical Concerns:} None.
+
+\subsubsection*{Rebuttal}
+{
+  \color{blue}
+
+As a follow up to these comments, the paper has been revised to clarify the main claims, the description of the method and presentation of experimental results with new graphs.
+We have also included a comparison with a new baseline suggested by Reviewer iMf4, and an analysis of the impact of meta parameters.
+
+However, we would also like to outline the fact that our contribution, in this paper, is *not* to outperform the best published state-of-the-art results for multilingual / multidomain MT.
+Instead, our goal is to present a novel, mathematically sound, methodology to automatically learn shared nodes in multi-task (multi-domain, multi-lingual) systems and to show (a) that these new patterns yield performance that are better that alternative approaches for sharing / unsharing subpart of the computation graph accross tasks and also compare favorably in terms of computational cost; (b) that these patterns actually capture similarities between tasks.
+We have therefore planned our experiments, comparisons and analyses insofar as they were relevant to illustrate these aspects.
+This has notably motivate our choice the adaptor model, which has been used in many prior work, as a reasonable point of comparison.
+We however respectfully disagree on the fact that unpublished papers from mid 2021 should be the ``baseline'' for reporting experiments in november 2021.
+
+We discuss each review below, and explain the changes that have been made in response to each reviewers' suggestion.
+}
+
+% \todo{Make this contribution} Finally, we hope the reviewer change the view on the objective of our paper, reconsider the solid mathematical foundation of our method supported by an explicit implementation explained in the Appendix, the improvement with high margin in multi-domain, the reduced computational cost in the inference compared to the Adapters, and the novelty in the use of node-masking instead of weight-masking as in previous work such as in Lin et al.
+% }
 
@@
 
-I Table 7 can be improved in presentation by using heatmap.
+Table 7 can be improved in presentation by using heatmaps.
 
@@
 {\color{blue}%
-  We thank the reviewer for pointing additional relevant work. We cite and discuss it in the revised version.
+We thank the reviewer for pointing additional relevant work. We cite and discuss it in the revised version.\Todo{Check this}
 
@@
 \begin{itemize}
-\item Our approach does not aim to find task-specific and task-agnostic parameters in the neural network: there is no notion of domain specific or domain agnostic parameters.
-The objective of our algorithm is to optimize the assignment of a sub-group of "nodes", which are the dimensions of the output of each intermediate layer, to each task.
-\item Our approach does not mask the sub-parts of each weight matrix but masks a fixed number of sub-groups of elements of the output of each intermediate layer.
+\item Our approach does not aim to find task-specific and task-agnostic \textbf{parameters} in the neural network: there is no notion of domain-specific or domain-agnostic parameters in the approach we present.
+The objective of our algorithm is to optimize the assignment \emph{of a sub-group of "nodes"}, at the output of each intermediate layer, to each task.
+\item Our approach does not mask the sub-parts of each weight matrix, but instead it masks a fixed number of sub-groups of elements of the output of each intermediate layer.
 \end{itemize}
-We have tried to improve our presentation of our method in the revised version of the paper.\todo{Change presention, give line numbers}
+We have tried to improve our presentation of our method in the revised version of the paper to avoid further misunderstandings.\todo{Change presention, give line numbers}
 
-We would like to answer the other concerns of the reviewier as follows.
-\begin{itemize} CLAIM : baseline + parameter sharing + resultats CDR
-\item Regarding baselines, we insist ur claim is that embedding the similarity between tasks in the construction of the multi-task network is beneficial, and the fine-tuning-based approaches ignore these similarities.
-We prove our claim by proposing a method that automatically optimizes the structure of a multi-task network by learning from data, i.e., implicitly learning the similarity between tasks and embedding it to the organization of the network.
-Besides, the Adapters are a robust method for multi-task learning, and a comparison with this baseline strongly supports our claim over the efficiency of our proposed method.
-\item We are happy to provide more comparisons with heuristic strategies of the dropout group selection in the final version.
-The flaws of these heuristic strategies motivate us to develop this technique.
-\item We have included the analysis on the effect of $k$ on the performance in the revised version (see new section 4.4)
+We would like to answer the other concerns of the reviewier as follows. \todo{CLAIM : baseline + parameter sharing + resultats CDR}
+\begin{itemize} 
+\item Regarding baselines, we generally refer to the rebuttal to the meta review.
+Note that according to your suggestion, we have added new multidomain experiments with a heuristic definition of blocks.
+See the updated table~XXX.\Todo{Add table}
+\item We have now included an analysis on the effect of $k$ on the performance in the revised version (see new section 4.4)
 \item The multilingual experiments use two groups of $8$ languages, which are all low-resourced; this might lower the significance of the improvement as observed in the paper of Lin et al.
-21 (only +0.7 for low resource).
-Our method significantly improves the less resource languages among these groups.
+21 (only +0.7 for low resource languages).
+Our method significantly improves the less resources languages among these groups.
+\item As suggested, we are now using heatmaps instead of Table 7.
+\item Statistical confidence scores are also reported. \todo{Add significancy tests}
 \end{itemize}
+}
 
-\todo{Make this contribution} Finally, we hope the reviewer change the view on the objective of our paper, reconsider the solid mathematical foundation of our method supported by an explicit implementation explained in the Appendix, the improvement with high margin in multi-domain, the reduced computational cost in the inference compared to the Adapters, and the novelty in the use of node-masking instead of weight-masking as in previous work such as in Lin et al.}
 \section*{Reviewer 681g}
@@
 
-Best Paper: No
+\textbf{Best Paper}: No
 
@@
 {\color{blue}
-\tofo{Keep this ?}We thank the reviewer for appreciating our mathematical work to support our implementation and the suggestions for a better understanding of our proposed approach.
-We would like to answer the reviewer's concerns as follows.
+We thank the reviewer for their suggestest improvements to this work. We would like to answer the reviewer's concerns as follows.
 
@@
 \item % We did admit our lack of analysis on the hyperparameters in the conclusion of our paper.
-  We have now collected more data for better analyzing the effect of hyper-parameters on the performance of our method since the submission. These results in our revised version.
-\item We agree with the reviewer.
-However, we could reduce the group size to 1 node only and use hierarchical modeling to optimize the number of nodes to be activated.
-These changes would improve the method's flexibility in exchange for computational cost in training.
-\item We are aware of these works and they are discussed in related work section.
-However, we do not attempt to beat a race in the translation performance in multilingual translation or multi-domain translation.
-In this work, our claim is that embedding the similarity between tasks in the construction of the multi-task network is beneficial, and the fine-tuning-based approaches ignore these similarities.
-And, we prove our claim by proposing a method that automatically optimizes the structure of a multi-task network by learning from data, i.e., implicitly learning the similarity between tasks and embedding it to the organization of the network \Todo{Reponse à tous les reviewers}
+  We have now collected more data for better analyzing the effect of hyper-parameters on the performance of our method. These results in our revised version (section 4.4).
+
+\item We agree with the reviewer that our method requires to set the number of blocks a priori: for us, this is just one more hyper-parameter on top of the many parameters in the Transformer (depth, number of heads, representation side) and should not be regarded as a lack of flexibility.
+Our analysis show that in our experimental conditions, the number of groups impacts the performance, which seem to quicly plateau as the number of blocks increase.
+
+\item Regarding alternative approaches that have been recently proposed, we are aware of these works and they are duely discussed in related work section.
+However, as explain in our general rebutal above, we do not consider that comparing with these would change any of our claim in any way.\done\Todo{Reponse à tous les reviewers}
 \end{itemize}
 
-We agree that a mixture of experts is an relevant alternative baseline.
-Again resource limitation in training --  We have been working on their implementations but lack the computational resource for training.
-Finally we hope the reviewer change the view on the objective of our paper and reconsider our solid mathematical modeling supported by supported by an explicit implementation explained in the Appendix.
+We agree that a mixture of experts model would also provide yet another possible alternative baseline.
+Note however that mixture of experts are much costly to train, much more than our method, as we would need to fully train one model for each expert; furthermore, a mixture of expert would not bring us any insight regarding the benefits of sharing / unsharing patterns for related domains.
+Therefore, we do not see that the outcome of such comparison would allow us to revise or amend or main claims.
 }
@@
 {\color{blue}%
-We thank the reviewer for appreciating our mathematical modeling and the suggestions for a better understanding of our proposed technique.
-We would like to answer the concerns of the reviewer as follows.
+We thank the reviewer for their suggestest improvements to this work. We would like to answer the concerns of the reviewer as follows.
 
@@
 \item As reported in many works using the Gumbel reparameterization trick, the effect of $\tau$ is relatively insignificant.
-This is discussed in the revised version in section.
+This is discussed in the revised version in section XXXX.
 \Todo{which section}Therefore, we focused more on selecting $k$.
 \Todo{Add references}
-\item Because we would like to have the same total amount of training iterations between Adapter-based and LaMGD, we extend the training of LaMGD to 300K.
-The convergence of the standard Transformer is before 200K as its validation curve became flat near the 200K-th iteration.
+\item Because we would like to have the same \emph{total amount of training iterations between Adapter-based and LaMGD}, we extend the training of LaMGD to 300K.
+The convergence of the standard Transformer is achieved before 200K as its validation curve became flat near the 200K-th iteration.
 These has been clarified in the paper.
 \Todo{Which section ?}
-\item The previous reason is applied for the reviewer's concern over multilingual experiments.
-\item As we pointed out, the significant gains are observed in several very low-resourced languages, not globally.
-This might be explained by the fact that low-resourced languages benefit from the sharing nodes with their related well-resourced languages while less suffering from the inference of unrelated languages.
+\item The same answer applies to the reviewer's concern regarding the multilingual experiments.
+\item As pointed out\todo{Where ?}, significant gains are observed in several very low-resourced languages, not globally.
+This might be explained by the fact that low-resourced languages benefit most from the sharing nodes with a related well-resourced languages, and are also less affected by the inference of unrelated languages.
 \end{itemize}
-Our claim on the advantage in computational cost repose on the inference stage while there are no extra parameters added to the standard Transformer compared to Adapter-based methods.
-However, the training time will cost more than usual because of the joint optimization of the network's organization and the parameters.
-Finally, we hope the reviewer change the view on the objective of our paper and reconsider our solid mathematical modeling supported by the explicit implementation explained in Appendix.
+
+Our claim on the advantage in computational cost wrt.
+Adapters relies on the inference stage while there are no extra parameters added to the standard Transformer compared to Adapter-based methods.
+At the same time, the training time will  more than usual because of the joint optimization of the network's organization and the parameters.
 \todo{Not so great.
-We cannot be sure the reviewers stay the same.}
+Can we say more ?
+There is a nice blog post on FLOPS here: \url{%
+https://medium.com/@dzmitrybahdanau/the-flops-calculus-of-language-model-training-3b19c1f025e4
+}%
+}
 }%

### 2022-01-13 2022/acl/responses.tex
merge remote changes from FY
answers for todos in responses.text

@@
 \end{itemize}
-We have tried to improve our presentation of our method in the revised version of the paper to avoid further misunderstandings.\todo{Change presention, give line numbers}
+We have tried to improve our presentation of our method in the revised version of the paper to avoid further misunderstandings.\todo{Change presentation, give line numbers}
 
@@
 Note that according to your suggestion, we have added new multidomain experiments with a heuristic definition of blocks.
-See the updated table~XXX.\Todo{Add table}
+See the updated table $3$.\Todo{Add table}
 \item We have now included an analysis on the effect of $k$ on the performance in the revised version (see new section 4.4)
@@
 \begin{itemize}
-\item We initialize the probabilistic distribution of the latent variables as uniform distribution. This is clarified in the revised text.\Todo{Check clarified which line}
+\item We initialize the probabilistic distribution of the latent variables as uniform distribution. This is clarified in the revised text in Section$3.1.4$.\Todo{Check clarified which line}
 \item The revised version includes a discussion regarding $k$ and how it affects the performance of our approach (section 4.4).
 \item As reported in many works using the Gumbel reparameterization trick, the effect of $\tau$ is relatively insignificant.
-This is discussed in the revised version in section XXXX.
+This is discussed in the revised version in section $3.1.3$.
 \Todo{which section}Therefore, we focused more on selecting $k$.
@@
 These has been clarified in the paper.
-\Todo{Which section ?}
+\Todo{Which section ?} Section $3.1.1$
 \item The same answer applies to the reviewer's concern regarding the multilingual experiments.
-\item As pointed out\todo{Where ?}, significant gains are observed in several very low-resourced languages, not globally.
+\item As pointed out\todo{Where ?} in Table 5, significant gains are observed in several very low-resourced languages, not globally.
 This might be explained by the fact that low-resourced languages benefit most from the sharing nodes with a related well-resourced languages, and are also less affected by the inference of unrelated languages.
@@
 Adapters relies on the inference stage while there are no extra parameters added to the standard Transformer compared to Adapter-based methods.
-At the same time, the training time will  more than usual because of the joint optimization of the network's organization and the parameters.
+At the same time, the training time will be more than usual because of the joint optimization of the network's organization and the parameters.
 \todo{Not so great.

### 2022-01-13 2022/acl/responses.tex
small fixes in responses

@@
 We have therefore planned our experiments, comparisons and analyses insofar as they were relevant to illustrate these aspects.
-This has notably motivate our choice the adaptor model, which has been used in many prior work, as a reasonable point of comparison.
-We however respectfully disagree on the fact that unpublished papers from mid 2021 should be the ``baseline'' for reporting experiments in november 2021.
+This has notably motivated our choice the adapter model, which has been used in many prior studies, as a reasonable point of comparison.
+We however respectfully disagree on the fact that unpublished papers from mid 2021 cited in the reviews should be the ``baseline'' for reporting experiments in november 2021.
 
@@
 \begin{itemize}
-\item Our approach does not aim to find task-specific and task-agnostic \textbf{parameters} in the neural network: there is no notion of domain-specific or domain-agnostic parameters in the approach we present.
+\item Our approach does not aim to find task-specific and task-agnostic \emph{parameters} in the neural network: there is no notion of domain-specific or domain-agnostic parameters in the approach we present.
 The objective of our algorithm is to optimize the assignment \emph{of a sub-group of "nodes"}, at the output of each intermediate layer, to each task.
@@
 Note that according to your suggestion, we have added new multidomain experiments with a heuristic definition of blocks.
-See the updated table $3$.\Todo{Add table}
+See the updated Table $3$.\Todo{Add table}
 \item We have now included an analysis on the effect of $k$ on the performance in the revised version (see new section 4.4)
@@
 
-\item We agree with the reviewer that our method requires to set the number of blocks a priori: for us, this is just one more hyper-parameter on top of the many parameters in the Transformer (depth, number of heads, representation side) and should not be regarded as a lack of flexibility.
-Our analysis show that in our experimental conditions, the number of groups impacts the performance, which seem to quicly plateau as the number of blocks increase.
+\item We agree with the reviewer that our method requires to set the number of blocks a priori: for us, this is just one more hyper-parameter on top of the many parameters in the Transformer (depth, number of heads, size of representations, etc) and should not be regarded as a lack of flexibility.
+Our analysis show that in our experimental conditions, the number of groups impacts the performance, which seem to quickly plateau as the number of blocks increases.
 
-\item Regarding alternative approaches that have been recently proposed, we are aware of these works and they are duely discussed in related work section.
-However, as explain in our general rebutal above, we do not consider that comparing with these would change any of our claim in any way.\done\Todo{Reponse à tous les reviewers}
+\item Regarding alternative approaches that have been recently proposed, we are aware of these works and they are duly discussed in the related work section.
+However, as explained in our general rebuttal above, we do not consider that comparing what we try to do with these would change any of our main claims in any way.\done\Todo{Reponse à tous les reviewers}
 \end{itemize}
@@
 \begin{itemize}
-\item We initialize the probabilistic distribution of the latent variables as uniform distribution. This is clarified in the revised text in Section$3.1.4$.\Todo{Check clarified which line}
+\item We initialize the probabilistic distribution of the latent variables as uniform distribution. This is clarified in the revised text in Section$3.1.4$.\done\Todo{Check clarified which line}
 \item The revised version includes a discussion regarding $k$ and how it affects the performance of our approach (section 4.4).
 \item As reported in many works using the Gumbel reparameterization trick, the effect of $\tau$ is relatively insignificant.
-This is discussed in the revised version in section $3.1.3$.
-\Todo{which section}Therefore, we focused more on selecting $k$.
+This is discussed in the revised version in section 3.1.3.
+\done\Todo{which section}Therefore, we focused more on selecting $k$.
 \Todo{Add references}
@@
 The convergence of the standard Transformer is achieved before 200K as its validation curve became flat near the 200K-th iteration.
-These has been clarified in the paper.
-\Todo{Which section ?} Section $3.1.1$
+These has been clarified in the paper in section 3.1.1.
+\done\Todo{Which section ?}
 \item The same answer applies to the reviewer's concern regarding the multilingual experiments.
-\item As pointed out\todo{Where ?} in Table 5, significant gains are observed in several very low-resourced languages, not globally.
+\item As pointed out\done\todo{Where ?} in Table~5, significant gains are observed in several very low-resourced languages, not globally.
 This might be explained by the fact that low-resourced languages benefit most from the sharing nodes with a related well-resourced languages, and are also less affected by the inference of unrelated languages.

### 2022-01-14 2022/acl/acl_latex.tex
answers for todos in responses.text
adjust length

@@
 \item We provide a sound mathematical formulation to the problem of jointly learning task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling techniques;
-\item We present algorithms to train this model end to end with very little extra computational cost.
-\item We report, using an extensive experimental setting, significant gains of this method with respect to strong baselines;
+\item We present algorithms to train this model end-to-end with very little extra parameters;
+\item We report, using an extensive experimental setting, significant gains of this method with respect to strong baselines for low-resource multilingual MT;
 \item We study how this method can actually exploit the similarities between tasks to learn interpretable sub-networks.
@@
 For each task, we only keep a fixed number of active groups of nodes and nullify all the other nodes.
-Our method learns which groups will be kept / dropped for each task.}
+%  Our method learns which groups will be kept / dropped for each task.
+}
 \label{fig:group_dropout}
 \end{figure}
-Formally, a \emph{group dropout mask} $m_l^d$ is a $n_p$-dimensional binary vector containing exactly $k$ ones: the group $p$ ($\in  [0,\dots,n_p\text{-}1]$) is retained for task $d$ if $m_l^d(p) =1$ and is dropped if $m_l^d(p) = 0$.
-We denote $\Delta^{n_p}_k = \{ m \in \{0,1\}^{n_p} \text{ such that }  \mid{} m \mid_{L_1} = k \}$ the set of all admissible dropout masks, with $\mid{} m \mid_{L_1}$ the $L_1$ norm of vector $m$; $\#\Delta^{n_p}_k$ is the cardinal of $\Delta^{n_p}$.
+Formally, a \emph{group dropout mask} $m_l^d$ is a $n_p$-dimensional binary vector containing exactly $k$ ones: group $p$ ($\in  [0,\dots,n_p\text{-}1]$) is retained for task $d$ if $m_l^d(p) =1$ and is dropped if $m_l^d(p) = 0$.
+We denote $\Delta^{n_p}_k = \{ m \in \{0,1\}^{n_p} \text{ such that }  \mid{} m \mid_{L_1} = k \}$ the set of all admissible masks, with $\mid{} m \mid_{L_1}$ the $L_1$ norm of vector $m$; $\#\Delta^{n_p}_k$ is the cardinal of $\Delta^{n_p}$.
 
@@
 \begin{align*}
-  r_l^d(i) = m_l^d(p) \ \ \ \text{if}\ \ \ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p}
+  r_l^d(i) = m_l^d(p) \ \ \ \text{if}\ \ \ p \times \frac{d_k}{n_p} \leqslant i < (p+1) \times \frac{d_k}{n_p},
 %  \begin{cases}
@@
 \end{align*}
-where $d_k$ is the dimension of the hidden state.\fyDone{Check this} The propagation of information within the network then depends on the current task value and can be formalized as:
+where $d_k$ is the dimension of the hidden state.\fyDone{Check this} The propagation of information within the network then depends on the current task value as follows:
 \begin{align*}
@@
 where $\operatorname{KL}$ is the Kullback-Leibler divergence.
-We use $-\ell(x,y,d; \theta,\Phi)$ as our surrogate training objective, as a tractable approximation of the untractable likelihood, and try to minimize this function in $\theta$ and $\Phi$.
+We use $-\ell(x,y,d; \theta,\Phi)$ as our surrogate training loss, as a tractable approximation of the likelihood, and try to minimize this function in $\theta$ and $\Phi$.
 
@@
 This modeling choice for the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick of \citet{Kool19stochastic} that we use below.
-Given our assumptions for the prior distribution and our choice for the variational distribution, the two terms in Eq.~\eqref{eq:lower-bound} can be computed as:
+Given our choices for the prior and the variational distributions, the two terms in Eq.~\eqref{eq:lower-bound} can be computed as:
 \begin{align}
@@
 The number of parameters of the variational distribution is $L\times k \times L \times n_d$, which is negligible in comparison to the size of the Transformer model.
-\item \item Transformer using heuristic multi-task group dropout (\system{HMGD} Transformer): we share 320 nodes for every task, and reserve 32 nodes for each task (making $320 + 32*6 = 512$ nodes in total).
+\item Transformer using heuristic multi-task group dropout (\system{HMGD} Transformer): we share 320 nodes for every task, and reserve 32 nodes for each task (totalling $320 + 32*6 = 512$ nodes).
 \end{itemize}
@@
 The convergence of the standard Transformer is before 200K as its validation curve became flat near the 200K-th iteration.
-The \system{LaMGD} Tranformer converged after 300k iterations with a batch size of 12k tokens.
+The \system{LaMGD} Transformer converged after 300k iterations with a batch size of 12k tokens.
 The convergence of \system{LaMGD} is controlled by its validation curve.
-Finally, we plug domain adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline.
-With approximately the same amount of training iterations, \system{LaMGD} and Adapter-based Transformer are trained.
+Finally, we plug domain adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline: this ensures that \system{LaMGD} and Adapter-based Transformer have the same amount of training iterations.
 
@@
 \item Adapter based Transformer: the intermediate feedforward dimension is set as 128 as in \citep{Gong21adaptive}.
-\item \system{LaMGD} Tranformer: There is no change in the architecture. We group 512 nodes in each layer into 16 groups of 32 consecutive nodes. For each language, only 12 groups are selected. 
+\item \system{LaMGD} Transformer: There is no change in the architecture. We group 512 nodes in each layer into 16 groups of 32 consecutive nodes. For each language, we select 12 groups.
 \end{itemize}
@@
 Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments
-
-For all experiments, we report the BLEU score of \citet{Papineni02bleu} computed with SacreBleu \citep{Post18call}.\fyDone{Check this}
-\subsubsection{Latent variables' initialization}
-We initialize the probabilistic distribution of the latent variables by uniform distribution.
-More precisely, we set $\Phi_l^d$, which generates the probability of the masks via the softmax activation function, to $0^{n_d}$.
-\subsection{Datasets}
+\subsubsection{Latent variables initialization}
+We initialize the distribution of the latent variables uniformly. More precisely, we set $\Phi_l^d$, which generates the probability of the masks via the softmax activation function, to $0^{n_d}$.
+\subsection{Datasets and metrics}
 \subsubsection{Multidomain translation}
-We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyDone{Add public to the data link} The datasets \footnote{The raw data can be downloaded from \url{https://github.com/qmpham/experiments/tree/main/tacl20}} for the multidomain translation experiments are detailed in Table~\ref{tab:Corpora-chap4}.
+We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyDone{Add public to the data link} The datasets \footnote{See \url{https://github.com/qmpham/experiments/tree/main/tacl20}} for the multidomain translation experiments are detailed in Table~\ref{tab:Corpora-chap4}.
 \begin{table*}[h!]
@@
 
-The languages used in multilingual experiments are referred to by their ISO 639-1 code that we reproduce below for convenience:
+The languages used in the multilingual experiments are as follows (see statistics in Table~\ref{tab:Corpora-multilingual}):
 \begin{itemize}
@@
 \item \textbf{Related set}: Azerbajiani (aze), Belarusian (bel),
-  Czech (ces), Galician (glg), Portuguese (por), Russian (rus), Slovak (slk), Turkish (tur).
+  Czech (ces), Galician (glg), Portuguese (por), Russian (rus), Slovak (slk), Turkish~(tur).
 \end{itemize}
 
-Table~\ref{tab:Corpora-multilingual} describes the subsets of TED8 that were used in our experiments.
 \begin{table*}[h]
@@
 \end{table*}
+
+For all experiments, we report the BLEU score of \citet{Papineni02bleu} computed with SacreBleu \citep{Post18call}.
+Statistical significance is computed with compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+We report significant differences at the level of $p=0.05$.\fyDone{Check this}
+
 \section{Results and analyses}
@@
 \subsection{Multidomain translation}
-For these experiments, our main results are in Table ~\ref{tab:mdmt}, where we observe that the \system{LaMGD} Tranformer achieves a significant improvement (+2.78) over the generic \system{Transformer} system with zero extra parameters.
-Moreover, \system{LaMGD} Tranformer achieves performance that are equivalent on average to that of the \system{Adapter} sytems, which is carrefully finetuned and contains approximately 25M additional parameters per domain.
+For these experiments, our main results are in Table~\ref{tab:mdmt}, where we observe that the \system{LaMGD} Transformer achieves a significant improvement (+2.78) over the generic \system{Transformer} system with zero extra parameters.
+Moreover, \system{LaMGD} Transformer achieves performance that are equivalent on average to that of the \system{Adapter} sytems, which is finetuned and contains approximately 25M additional parameters per domain.
 Variational mask learned from data by \system{LaMGD} also outperforms heuristic dropout mask \system{HMGD} by 0.5 in average.
@@
     \system{Transformer}  \hfill{\footnotesize[65m]} & \SB{40.3} & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
+    \system{HMGD} Transformer   \hfill{\footnotesize[+0m]}  & \SB{40.4} & 60.4 & 51.9 & 38.7 &	50.80 &	86.80 & 54.8 \\ 
     \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & \SB{61.0} & \SB{53.1} & 37.5 & 49.6 & \SB{91.0} & \SB{55.3} \\ 
-    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & \SB{40.3} & 60.4 & 52.4 & \SB{39.0} & \SB{52.4} & 87.5 & \SB{55.3} \\ 
-    \system{HMGD} Tranformer   \hfill{\footnotesize[+0m]}  & \SB{40.4} & 60.4 & 51.9 & 38.7 &	50.80 &	86.80 & 54.8 \\ 
+    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  & \SB{40.3} & 60.4 & 52.4 & \SB{39.0} & \SB{52.4} & 87.5 & \SB{55.3} \\ 
     \hline
@@
   \end{tabular}
-  \caption{Experiments with multilingual translation. Boldface denotes significant gains over \system{Transformer}.\fyDone{What is the bold for ?}}
+  \caption{Multilingual Translation experiments. Boldface denotes significant gains over \system{Transformer} ($p=0.05$).\fyDone{What is the bold for ?}}
   \label{tab:multilingual}
@@
 \textbf{Language-specific sub-networks}.
-\citet{Frankle19Lottery,liu18rethinking} proposed different techniques to locate most important parameters for the task so that when training the model while masking other parameters the performance of the model does not degrade.\citet{lin21learning} hypothesized that among parameters of a multilingual NMT model, there exists a subset of parameters more biased to a specific language.
+\todo{}\citet{Frankle19Lottery,liu18rethinking} proposed different techniques to identify most important parameters for the current task so that when training the model while masking other parameters the performance of the model does not degrade.
+\citet{lin21learning} hypothesized that among parameters of a multilingual NMT model, there exists a subset of parameters more biased to a specific language.
 The authors proposed to locate these subsets by pruning the parameters of each fine-tuned model.

### 2022-01-14 2022/acl/acl_latex.tex
local changes

--- 2022-01-13 2022/acl/acl_latex.tex
@@
 The number of parameters of the variational distribution is $L\times k \times L \times n_d$, which is negligible in comparison to the size of the Transformer model.
-\item \item Transformer using heuristic multi-task group dropout (\system{HMGD} Transformer): we share 320 nodes for every task, and reserve 32 nodes for each task (making $320 + 32*6 = 512$ nodes in total).
+\item Transformer using heuristic multi-task group dropout (\system{HMGD} Transformer): we share 320 nodes for every task, and reserve 32 nodes for each task (making $320 + 32*6 = 512$ nodes in total).
 \end{itemize}
@@
 in which $r=0.0001$.
-While \citet{Gong21pay,Gong21adaptive} fixed $\tau$ to be 0.2, we select an anneal policy for $\tau$ proposed by previous studies \citep{Jang17categorical}.
-Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments
+While \citet{Gong21pay,Gong21adaptive} fixed $\tau$ to be 0.2, we combine it with an anneal policy for $\tau$ proposed by previous studies \citep{Jang17categorical} to encourage the diversity in the beginning of the training.
+Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments.
 
@@
   \end{tabular}
-  \caption{Experiments with multilingual translation. Boldface denotes significant gains over \system{Transformer}.\fyDone{What is the bold for ?}}
+  \caption{Experiments with multilingual translation. Boldface denotes significant gains over \system{Transformer} ($p=0.05$).\fyDone{What is the bold for ?}}
   \label{tab:multilingual}

### 2022-01-14 2022/acl/acl_latex.tex
revise related work

--- 2022-01-14 2022/acl/acl_latex.tex
@@
 These approaches share the same interest in expressing the proximity between tasks in the selection of task-specific parameters as our approach.
-However, our method learn the selection from a latent commonality in data instead of using a predefined selection such as "One language family per decoder" in \cite{kong21multilingual}.
+However, our method learn the selection from a latent commonality in data instead of using a predefined selection such as "One language family per decoder" in \citep{kong21multilingual}.
 
 \textbf{Language-specific sub-networks}.
-\todo{}\citet{Frankle19Lottery,liu18rethinking} proposed different techniques to identify most important parameters for the current task so that when training the model while masking other parameters the performance of the model does not degrade.
-\citet{lin21learning} hypothesized that among parameters of a multilingual NMT model, there exists a subset of parameters more biased to a specific language.
-The authors proposed to locate these subsets by pruning the parameters of each fine-tuned model.
-Our approach also aims to assign a sub-network to each task by masking the output of each layer rather than masking the parameters.
-Furthermore, \citet{lin21learning} computes the masks via a heuristic selection while our approach learn the masks via a variational technique.
+\done\todo{}\citet{Frankle19Lottery,liu18rethinking} study techniques to identify the most important parameters for the current task, so that masking the less important parameters during training does not hurt performance.
+\citet{lin21learning} adapts this idea for multilingual NMT, trying to identify language dependent subsets of parameters by pruning a fine-tuned model.
+Our approach also aims to map sub-networks to tasks: we do so by masking the output of each layer, rather than masking parameters.
+Furthermore, \citet{lin21learning} computes the masks via a heuristic selection; while our approach learns the masks with variational techniques.
 

### 2022-01-14 2022/acl/acl_latex.tex
revise related work

@@
 In this work, we propose a new method to better take advantage of these similarities, using a latent-variable model.
-We also develop new techniques to train this model end to end and report multiple experimental results that show the learned patterns are both meaningful and yield improved translation performance without any increase of the model size.
+We also develop new techniques to train this model end-to-end and report experimental results showing that the learned patterns are both meaningful and yield improved translation performance without any increase of the model size.
 \end{abstract}

### 2022-01-14 2022/acl/acl_latex.tex

@@
 Therefore, we approximate this function by the $\operatorname{Soft-Top-K}$ function defined as follows:
-\begin{align}
+\begin{align}\label{eq:soft-top-k}
 \hat{m}_l^d(\tau) &= \displaystyle{\mathop{argmin}_{\substack{
@@
         1^{T}.m = k
-      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \label{eq:soft-top-k}
+      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) 
 \end{align}
@@
 Furthermore, we also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and that its gradients can be computed using the implicit differentiation theorem.
-During training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyper-parameter $\tau$ to $0$.
+During training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyper-parameter $\tau$ to $0.2$.
 The gradient of $D$ w.r.t $\Phi_l^d$ is computed using the chain rule as follows:
-\begin{equation}
+\begin{align*}
 \frac{\partial D}{\partial \Phi_l^d} = \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} 
-\end{equation}
+\end{align*}
 The gradient $\frac{\partial D}{\partial \hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$ is computed via implicit differentiation, as explained in Appendix~\ref{appendix:a}.
@@
 We jointly train the Transformer parameters $\theta$ and the parameters of the variational distribution $\Phi$ using the following multi-task loss.\fyDone{This is repeated}
-\begin{align}
+\begin{align*}
 \hspace{-0.em}
 \mathcal{L}(\theta,\Phi) = \sum_{d=1}^{n_d} \mathbb{E}_{x \sim \mathcal{D}_s^d,y \sim MT^d(x)} \big[-\ell(x,y,d; \theta,\Phi)\big]
-\end{align}
+\end{align*}
 
@@
 We set the dropout probability to 0.1.
-We train the multi-domain Transformer model for 200k iterations with a batch size of 12k tokens.
+We train the multidomain Transformer model for 200k iterations with a batch size of 12k tokens.
 The convergence of the standard Transformer is before 200K as its validation curve became flat near the 200K-th iteration.
-The \system{LaMGD} Transformer converged after 300k iterations with a batch size of 12k tokens.
+The \system{LaMGD} Transformer converged after 300k iterations with the same batch size.
 The convergence of \system{LaMGD} is controlled by its validation curve.
-Finally, we plug domain adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline: this ensures that \system{LaMGD} and Adapter-based Transformer have the same amount of training iterations.
+Finally, we plug adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline.
+%: this ensures that \system{LaMGD} and Adapter-based Transformer have the same amount of training iterations.
 
@@
 While \citet{Gong21pay,Gong21adaptive} fixed $\tau$ to be 0.2, we select an anneal policy for $\tau$ proposed by previous studies \citep{Jang17categorical}.
-Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments
+Finally, we set the weight of the entropy term to $0.0001$ in the training loss in every experiments.
+%$\mathbb{H}(\text{softmax}(\Phi_l^d))$
 \subsubsection{Latent variables initialization}
@@
   \end{tabular}
-  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$).
-\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
-  }
+  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$).}
+%\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
 \label{tab:Corpora-chap4}
@@
     Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[65m]} & \SB{40.3} & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
-    \system{HMGD} Transformer   \hfill{\footnotesize[+0m]}  & \SB{40.4} & 60.4 & 51.9 & 38.7 &	50.80 &	86.80 & 54.8 \\ 
-    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & \SB{61.0} & \SB{53.1} & 37.5 & 49.6 & \SB{91.0} & \SB{55.3} \\ 
-    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  & \SB{40.3} & 60.4 & 52.4 & \SB{39.0} & \SB{52.4} & 87.5 & \SB{55.3} \\ 
+    \system{Transformer}  \hfill{\footnotesize[65m]} & 40.3 & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
+    \system{HMGD} Transformer   \hfill{\footnotesize[+0m]}  & 40.4 & 60.4 & 51.9 & 38.7 &	50.80 &	86.80 & 54.8 \\ 
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & \SB{91.0} & \SB{55.3} \\ 
+    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & \SB{39.0} & \SB{52.4} & 87.5 & \SB{55.3} \\ 
     \hline
@@
 A truly multidomain system should be able to automatically detect the proximity between \domain{law$_1$} and \domain{law$_2$}, and there should be no significant difference between the performance of a system trained with the six original domains (including \domain{law}) or with the seven domains (including \domain{law} by \domain{law$_1$} and \domain{law$_2$}).
-The experiments in \citep{Pham21revisiting} reported a large gap between the two settings when using residual adapters.
-We replicated this setting and also report the results obtained with the \system{LaMGD} Transformer system in Table~\ref{tab:fuzzy}.
+\citet{Pham21revisiting} reported a large gap between the two settings when using residual adapters.
+We replicated this setting and report the results obtained with the \system{LaMGD} Transformer system in Table~\ref{tab:fuzzy}.
 
@@
     O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&\SB{34.0}&38.3&8.3&20.19 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &10.2&4&\SB{13.3}&21.9&32.2&\SB{34.1}&38.5&8.3&20.31 \\ 
+    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&34.0&38.3&8.3&20.19 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &10.2&4&\SB{13.3}&21.9&32.2&34.1&38.5&8.3&20.31 \\ 
     \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
@@
 \end{subfigure}
-\caption{Heatmap visualization of the similarities between dropout masks of domains(languages). Brighter colors correspond to higher similarity scores.}
+\caption{Heatmap visualization of the similarities between dropout masks of domains(languages).}
 \label{fig:heatmap}
@@
 
-We also plot the languages based on their dropout masks in Figure~\ref{fig:visual} using a 2d projection computed by PCA.
+We also plot the languages based on their dropout masks in Figure~\ref{fig:visual} using a 2d PCA projection.
 \begin{figure*}[htbp]
@@
   \end{tabular}
-  \caption{Variation of the performance w.r.t $k$, while we fix $n_p=16$. These results correspond to the multilingual setting (o2m-related).}
+  \caption{Variation of the performance w.r.t $k$, while we fix $n_p=16$ (o2m-related experiment).}
   \label{tab:K}

### 2022-01-14 2022/acl/acl_latex.tex

@@
 In this work, we propose a new method to better take advantage of these similarities, using a latent-variable model.
-We also develop new techniques to train this model end-to-end and report experimental results showing that the learned patterns are both meaningful and yield improved translation performance without any increase of the model size.
+We also develop new techniques to train this model end to end and report multiple experimental results that show the learned patterns are both meaningful and yield improved translation performance without any increase of the model size.
 \end{abstract}
@@
 These approaches share the same interest in expressing the proximity between tasks in the selection of task-specific parameters as our approach.
-However, our method learn the selection from a latent commonality in data instead of using a predefined selection such as "One language family per decoder" in \citep{kong21multilingual}.
+However, our method learn the selection from a latent commonality in data instead of using a predefined selection such as "One language family per decoder" in \cite{kong21multilingual}.
 
 \textbf{Language-specific sub-networks}.
-\done\todo{}\citet{Frankle19Lottery,liu18rethinking} study techniques to identify the most important parameters for the current task, so that masking the less important parameters during training does not hurt performance.
-\citet{lin21learning} adapts this idea for multilingual NMT, trying to identify language dependent subsets of parameters by pruning a fine-tuned model.
-Our approach also aims to map sub-networks to tasks: we do so by masking the output of each layer, rather than masking parameters.
-Furthermore, \citet{lin21learning} computes the masks via a heuristic selection; while our approach learns the masks with variational techniques.
+\todo{}\citet{Frankle19Lottery,liu18rethinking} proposed different techniques to identify most important parameters for the current task so that when training the model while masking other parameters the performance of the model does not degrade.
+\citet{lin21learning} hypothesized that among parameters of a multilingual NMT model, there exists a subset of parameters more biased to a specific language.
+The authors proposed to locate these subsets by pruning the parameters of each fine-tuned model.
+Our approach also aims to assign a sub-network to each task by masking the output of each layer rather than masking the parameters.
+Furthermore, \citet{lin21learning} computes the masks via a heuristic selection while our approach learn the masks via a variational technique.
 

### 2022-01-14 2022/acl/responses.tex

@@
 
-However, we would also like to outline the fact that our contribution, in this paper, is *not* to outperform the best published state-of-the-art results for multilingual / multidomain MT.
+However, we would also like to outline the fact that our contribution, in this paper, is not to outperform the best published state-of-the-art results for multilingual / multidomain MT.
 Instead, our goal is to present a novel, mathematically sound, methodology to automatically learn shared nodes in multi-task (multi-domain, multi-lingual) systems and to show (a) that these new patterns yield performance that are better that alternative approaches for sharing / unsharing subpart of the computation graph accross tasks and also compare favorably in terms of computational cost; (b) that these patterns actually capture similarities between tasks.
@@
 {\color{blue}%
-We thank the reviewer for pointing additional relevant work. We cite and discuss it in the revised version.\Todo{Check this}
+We thank the reviewer for suggesting additional relevant work. We cite and discuss it in the revised version in section 5.\done\Todo{Check this}
 
@@
 \end{itemize}
-We have tried to improve our presentation of our method in the revised version of the paper to avoid further misunderstandings.\todo{Change presentation, give line numbers}
+We have tried to improve our presentation of our method in the revised version of the paper to avoid further misunderstandings (see lines 92-95, lines 115-119).\done\todo{Change presentation, give line numbers}
 
-We would like to answer the other concerns of the reviewier as follows. \todo{CLAIM : baseline + parameter sharing + resultats CDR}
+We would like to answer the other concerns of the reviewer as follows. \done\todo{CLAIM : baseline + parameter sharing + resultats CDR}
 \begin{itemize} 
@@
 Note that according to your suggestion, we have added new multidomain experiments with a heuristic definition of blocks.
-See the updated Table $3$.\Todo{Add table}
+See the updated Table $3$.\done\Todo{Add table}
 \item We have now included an analysis on the effect of $k$ on the performance in the revised version (see new section 4.4)
@@
 \item As suggested, we are now using heatmaps instead of Table 7.
-\item Statistical confidence scores are also reported. \todo{Add significancy tests}
+\item Statistical confidence scores are also reported in Section 3.2. \done\todo{Add significancy tests}
 \end{itemize}
@@
 We agree that a mixture of experts model would also provide yet another possible alternative baseline.
-Note however that mixture of experts are much costly to train, much more than our method, as we would need to fully train one model for each expert; furthermore, a mixture of expert would not bring us any insight regarding the benefits of sharing / unsharing patterns for related domains.
-Therefore, we do not see that the outcome of such comparison would allow us to revise or amend or main claims.
+Note however that mixture of experts are much costly to train, much more than our method, as we would need add new sets of parameters for each expert.
+Furthermore, we do not see that the outcome of such comparison would allow us to revise or amend or main claims.
 }
@@
 \begin{itemize}
-\item We initialize the probabilistic distribution of the latent variables as uniform distribution. This is clarified in the revised text in Section$3.1.4$.\done\Todo{Check clarified which line}
+\item We initialize the probabilistic distribution of the latent variables as a uniform distribution. This is clarified in the revised text in Section~3.1.4.\done\Todo{Check clarified which line}
 \item The revised version includes a discussion regarding $k$ and how it affects the performance of our approach (section 4.4).
 \item As reported in many works using the Gumbel reparameterization trick, the effect of $\tau$ is relatively insignificant.
-This is discussed in the revised version in section 3.1.3.
+This is discussed in the revised version in Section 3.1.3.
 \done\Todo{which section}Therefore, we focused more on selecting $k$.
-\Todo{Add references}
+\done\Todo{Add references}
 \item Because we would like to have the same \emph{total amount of training iterations between Adapter-based and LaMGD}, we extend the training of LaMGD to 300K.
 The convergence of the standard Transformer is achieved before 200K as its validation curve became flat near the 200K-th iteration.
-These has been clarified in the paper in section 3.1.1.
+These has been clarified in the paper in Section~3.1.1.
 \done\Todo{Which section ?}
@@
 \item As pointed out\done\todo{Where ?} in Table~5, significant gains are observed in several very low-resourced languages, not globally.
+We have checked that these differences were significant.
 This might be explained by the fact that low-resourced languages benefit most from the sharing nodes with a related well-resourced languages, and are also less affected by the inference of unrelated languages.
@@
 At the same time, the training time will be more than usual because of the joint optimization of the network's organization and the parameters.
-\todo{Not so great.
+\done\todo{Not so great.
 Can we say more ?

### 2022-01-14 2022/acl/acl_latex.tex
shorter

--- 2022-01-14 2022/acl/acl_latex.tex
@@
 
-Recently, \citet{biao21share} proposed jointly learning how to route between shared representation and language-specific representation via a conditional language-specific routing and the parameters of the underlying Transformer.
-This method is closely related to the FusionAdapters of \citet{pfeiffer21adapterfusion}.
-Both approaches aimed to select between shared representation and task-specific representations.
+\citet{biao21share} propose to learn to route between shared and language-specific representations with a conditional language-specific routing while training the parameters of the underlying Transformer.
+This method is related to the FusionAdapters of \citet{pfeiffer21adapterfusion}.
+Both approaches aim to select between shared and task-specific representations.
 The proximity between tasks is not taken into account in the routing mechanism.
-Our method proposed a different approach to the problem of multi-task routing in the underlying network.
+We propose a different approach to the problem of multi-task routing in the underlying network.
 

### 2022-01-14 2022/acl/acl_latex.tex
8 pages version

@@
 It allows us to jointly search for an optimal assignment of sub-networks to tasks and to learn the parameters of the underlying network.
-Our method relies on a sound mathematical framework and an end-to-end optimization procedure which implies very little extra parameters.
-The training course is reasonably longer (100k iterations in multidomain translation) than training standard models while considering the gain in performance.\fyDone{There is a computational cost = 100k iterations} Experimentally, we achieve a large improvement over Transformer baseline without zero additional parameters; our performance are also comparable to that of a strong a multi-task baseline using residual adapter modules which however includes a large number of extra parameters.
-For multilingual translation, our model outperforms multilingual Transformer and Language Adapters in 3 our of 4 cases.
-Besides, we provided an thorough analysis of the similarities between learned sub-networks and demonstrated a strong correlation between the sub-networks' similarity and the proximity of the corresponding tasks (domain or language).
+Our method relies on a sound mathematical framework and an end-to-end optimization procedure; it only adds a small number of extra parameters.
+The additional training cost is also reasonable, amounting to 100k iterations in the multidomain setting, given the observed gains in performance.\fyDone{There is a computational cost = 100k iterations} Experimentally, we achieve a large improvement over a Transformer baseline; our performance are also comparable to that of a strong a multi-task baseline using residual adapter modules which rely on a large number of extra parameters.
+For multilingual translation, our model outperforms multilingual Transformer and Language Adapters in 3 our of 4 settings.
+Besides, we provided an thorough analysis of the similarities between learned sub-networks and demonstrate a strong correlation between the learned similarities and the proximity of the corresponding tasks (domain or language).
 
-There are several ways in which our methodology could be improved and generalized.
+There are several ways in which our methodology can be improved.
 In future work, we would first like to provide an complete variational framework to model both the number of groups, $k$ and the selection of the dropout masks.
-Second, we also intend to get rid of the requirement of knowing the domain information during inference: this would mean replacing the dependency on $d$ in the variational distribution by a dependency in the input $x$.
-These two questions seem both very interesting and promising; they would allow to us to dispense with heuristic choices in the architecture, and replace them with an increase dependency on the training data.
+Second, we also intend to dispense with the domain information during inference: this would mean replacing the dependency on $d$ in the variational distribution by a dependency on the input $x$.
+Adressing these two questions will allow to us replace heuristic choices in the architecture design with an increased dependency on the training data.
 

### 2022-01-14 2022/acl/acl_latex.tex
8 pages version

@@
 
-In this work, we develop techniques that can take the similarity between tasks into account in a more effective way, by \emph{learning the network organization from the data}.
-To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the parameter space, and show how training can still be performed end-to-end using a variational surrogate of the log-likelihood loss function.
+In this work, we study and develop new methods aimed to take into account the similarity between tasks in a more effective way, by \emph{learning the network organization from the data}.
+To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the representation space and show how training can still be performed end-to-end using a variational surrogate of the log-likelihood loss function.
 Our experiments with multilingual and multidomain machine translation confirm that this method can automatically detect similarities in the data, meaning that related tasks use the same subparts of the network.

### 2022-01-14 2022/acl/acl_latex.tex
revised introduction

@@
 
-In this work, we study and develop new methods aimed to take into account the similarity between tasks in a more effective way, by \emph{learning the network organization from the data}.
+In this work, we study and develop a new method, \emph{multi-task group dropout}, aimed to take into account the similarity between tasks in a more effective way, by \emph{learning the network organization from the data}.
 To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the representation space and show how training can still be performed end-to-end using a variational surrogate of the log-likelihood loss function.
@@
 Our results also show that this method is comparable to using adapter layers in a number of empirical comparisons; however, contrarily to adapters, these performance are obtained without any increase of the model size.
-Our main contributions are summarized below:
-\begin{itemize}
+Our contributions are primarily methodological and can be summarized as follows:
+\begin{enumerate}
   \setlength{\itemsep}{1pt}
@@
   \setlength{\parsep}{0pt}
-\item We provide a sound mathematical formulation to the problem of jointly learning task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling techniques;
+\item We introduce a novel, sound mathematical formulation to the problem of jointly learning task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling techniques;
 \item We present algorithms to train this model end-to-end with very little extra parameters;
-\item We report, using an extensive experimental setting, significant gains of this method with respect to strong baselines for low-resource multilingual MT;
+\item We report, using an extensive set of experiments, significant gains with respect to strong baselines for low-resource multilingual MT;
 \item We study how this method can actually exploit the similarities between tasks to learn interpretable sub-networks.
-\end{itemize}
+\end{enumerate}
 \fyDone{Can you do distillation ? That is only keep the parameter for one domain and still get good results (better than DA ?)}

### 2022-02-19 2022/eamt/eamt-latex (1)/eamt22.tex
add eamt

%
% File eamt20.tex
%
% Contact: cettolo@fbk.eu, mlf@dlsi.ua.es

%%% To ease future customizations, various replaceables have been paramaterized
%%% as listed in the newcommands section

\documentclass[11pt]{article}
\usepackage{eamt22}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage[small,bf]{caption} % added MLF 20171211
\setlength\titlebox{6.5cm}    % Expanding the titlebox
%%% YOUR PACKAGES BELOW THIS LINE %%%

\newcommand{\confname}{EAMT 2022}
\newcommand{\website}{https://eamt2022.com/}
\newcommand{\urlwebsite}{https://eamt2022.com/}
\newcommand{\contactname}{the guest organizers, LT3,}
\newcommand{\contactemail}{\texttt{eamt2022@ugent.be}}
\newcommand{\conffilename}{eamt22}
\newcommand{\downloadsite}{https://eamt2022.com/templates}

\newcommand{\citet}[1]{\cite{#1}}

\newcommand{\paperlength}{$10$ (ten)}
\newcommand{\shortpaperlength}{$10$ (ten)}
\newcommand{\projectlength}{$2$ (two)}
\newcommand{\translatorlength}{$10$ (ten)}

\title{Instructions for \confname\ Proceedings}

\author{First Author\\
  Affiliation / Address line 1\\
  Affiliation / Address line 2\\
  Affiliation / Address line 3\\
  {\tt email@domain}  \And
  Second Author\\
  Affiliation / Address line 1\\
  Affiliation / Address line 2\\
  Affiliation / Address line 3\\
  {\tt email@domain}}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  This document contains the instructions for preparing a camera-ready
  manuscript for the papers in the
  proceedings of \confname . The document itself
  conforms to its own specifications, and is therefore an example of
  what your manuscript should look like. Authors are asked to conform
  to all the directions reported in this document.
\end{abstract}

\section{Credits}

This document is derived from the instructions for EAMT 2020, % update this list at next change
which was the latest in a long line of adaptations: EAMT 2009, EAMT 2011, EAMT 2012, EAMT 2014, EAMT 2015, EAMT 2016, EAMT 2017, EAMT 2018, Coling 2008, ACL-07, Coling/ACL-06, EACL-06, ACL-05 and EACL-03.  All these were based on the formats of earlier ACL and EACL Conference proceedings.  Those versions were written by several people, including John
Chen, Henry S. Thompson and Donald Walker.

\section{Introduction}

The following formatting instructions are directed to authors of
papers accepted for publication in \confname{} proceedings.

See also the
conference website\footnote{\urlwebsite} for additional
advice and information regarding submission.  All authors are required to adhere to these
specifications. Authors are required to submit their papers in PDF  (Portable Document
Format). The proceedings will be printed on
{\bf A4 paper}. Authors from countries in which access to
word processing systems is limited should contact \contactname~at \contactemail~ as soon as possible.

\section{General Instructions}

Manuscripts must be in two-column format.  Exceptions to the
two-column format include the {\nobreak{title}}, authors' names and complete
addresses, which must be centered at the top of the first page, and
any full-width figures or tables (see the guidelines in
Subsection~\ref{ssec:first}). {\bf Type single-spaced.}  Start all
pages directly under the top margin. See the guidelines later
regarding formatting the first page.

See Section~\ref{sec:length} for information on the
maximum number of pages.

If you are using Overleaf, you can use this document as a template for your submission. Simply create your own copy 
of this read-only document by clicking on the main Overleaf menu (see Figure~\ref{fig:clone})  at the top left of the document editing window and then click on ``Copy Project''. 
You can then modify the newly created document.

\begin{figure*}
 \centering
 \includegraphics[scale=0.5]{clone1.png}
 \includegraphics[scale=0.5]{clone2.png}   
 \caption{How to clone the LaTeX template in Overleaf}\label{fig:clone}
\end{figure*}

\subsection{Electronically-available resources}

This description is provided in \LaTeX2e (\nobreak{\conffilename.tex}) along
with the \LaTeX2e style file used to format it (\nobreak{\conffilename.sty})
and an EAMT bibliography style (\nobreak{\conffilename.bst}); and in PDF format
(\nobreak{\conffilename.pdf}). These files are all available at
\downloadsite. There is also a Microsoft Word (\nobreak{\conffilename.dot}) and LibreOffice.org/OpenOffice (\nobreak{\conffilename.odt}) document template available at the same URL. We
strongly re\-commend the use of these style files, which have been
appropriately tailored for the \confname~proceedings.

\subsection{Format of Electronic Manuscript}
\label{sect:pdf}

For the production of the electronic manuscript you must use Adobe's
Portable Document Format (PDF). This format can be generated from
postscript files. On Unix systems, you can use {\tt ps2pdf} for
this purpose. Under Microsoft Windows, you can use Adobe's Distiller
or GSview (File$>$Convert$>$pdfwrite); if you have \textit{cygwin}
installed, you can use \textit{ps2pdf}. Note that
some word processing programs generate PDF which may not include all
the necessary fonts (esp. tree diagrams, symbols). When you print or
create the PDF file, there is usually an option in your printer setup
to include none, all or just non-standard fonts.  Please make sure
that you select the option of including ALL the fonts. {\em Before
sending it, test your PDF by printing it from a computer different
from the one where it was created.} Moreover, some word processors may
generate very large postscript / PDF files, where each page is rendered
as an image. Such images may reproduce poorly. In this case, try
alternative ways to obtain the postscript and / or PDF. One way on some
systems is to install a driver for a postscript printer, send your
document to the printer specifying ``Output to a file'', then convert
the file to PDF.

% US Letter
%It is of utmost importance to specify the \textbf{US Letter format} (21.6 cm x
%27.9 cm) / (8.5 in x 11 in) when formatting the paper. When working with {\tt
%dvips}, for instance, one should specify {\tt -t letter}.

% A4
It is of utmost importance to specify the \textbf{A4 format} (21.0 cm
x 29.7 cm) / (8.3 in x 11.7 in) when formatting the paper. When working with
{\tt dvips}, for instance, one should specify {\tt -t a4}.

Print-outs of the PDF file on A4 paper should look like the present document,
which conforms to the formatting requirements. {\em Note that in order for
your paper to print correctly, you should disable centering and scale-to-fit
options on your printer.} If you cannot meet the above requirements about the
production of your camera-ready paper, please contact the programme chairs
as soon as possible.

\subsection{Layout}
\label{ssec:layout}

Format manuscripts two columns to a page, in the manner these
instructions are formatted. The exact dimensions for a page on
A4 paper are:

% US letter
% \begin{itemize}
% \item Left and right margins: 2.5 cm (1 in)
% \item Top margin: 2.5 cm (1 in)
% \item Bottom margin: 2.5 cm (1 in)
% \item Column width: 8 cm (3.15 in)
% \item Column height: 22.9 cm (9 in)
% \item Gap between columns: 0.5 cm (0.2 in)
% \end{itemize}

% A4
\begin{itemize}
\itemsep 0mm
\item Left and right margins: 2.5 cm (1 in)
\item Top margin: 2.5 cm (1 in)
\item Bottom margin: 2.5 cm (1 in)
\item Column width: 7.7 cm (3.1 in)
\item Column height: 23.7 cm (9.4 in)
\item Gap between columns: 0.5 cm (0.2 in)
\end{itemize}

\subsection{Fonts}

For uniformity, Adobe's {\bf Times Roman} font should be
used. In \LaTeX2e{} this is accomplished by putting

\begin{quote}
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\end{quote}
in the preamble. If Times Roman is unavailable, use {\bf Computer
  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
  10\% less dense than Adobe's Times Roman font.

\subsection{The First Page}
\label{ssec:first}

Center the title, authors' names and affiliations across both
columns. Do not use footnotes for affiliations. Do not include the
paper ID number assigned during the submission process. Use the
two-column format only when you begin the abstract.

\paragraph{Title:} Place the title centered at the top of the first page, in
a 15-point bold font. Long titles should be typed on two lines without
a blank line intervening. Approximately, put the title at 2.5 cm (1 in) from
the top of the page, followed by a blank line, then the authors'
names, and the affiliations on the following line. Do not use only
initials for given names (middle initials are allowed). Avoid
capitalizing last names. The affiliation should contain the author's
complete address, and if possible an electronic mail address. Leave
about 2 cm (0.75 in) between the affiliation and the body of the first page.

\paragraph{Abstract:} Type the abstract at the beginning of the first
column. The width of the abstract text should be smaller than the
width of the columns for the text in the body of the paper by about
0.6 cm (0.25 in) on each side. Center the word {\bf Abstract} in a 12 point bold
font above the body of the abstract. The abstract should be a concise
summary of the general thesis and conclusions of the paper. It should
be no longer than 200 words.

\paragraph{Text:} Begin typing the main body of the text immediately after the
abstract, observing the two-column format as shown in the present document.
Use 11 point font for text. {\bf Indent} when starting a new paragraph, except
for the first paragraph following each heading.

\subsection{Sections}

\paragraph{Headings:} Type and label section and subsection headings in the
style shown on the present document.  Use numbered sections (Arabic
numerals) in order to facilitate cross references. Number subsections
with the section number and the subsection number separated by a dot,
in Arabic numerals. Do not number subsubsections. Use 11 point font for
subsection headings and 12 point font for section headings.

\paragraph{Citations:} Citations within the text appear in parentheses
as~\cite{Gusfield:97} or, if the author's name appears in the text itself, as
Gusfield~\shortcite{Gusfield:97}. Citations in parentheses should not be used
as linguistic phrases; for example, instead of ``\cite{Gusfield:97} \nobreak{argues}
that \ldots'' say ``Gusfield~\shortcite{Gusfield:97} \nobreak{argues} that \ldots''.
Treat double authors as in~\cite{Aho:72}, but write as in~\cite{Chandra:81} when more than
two authors are involved. 
\nobreak{Append} lowercase letters to the year in cases of ambiguity as in \cite{Gledson:08homog}. 
Collapse multiple citations in parenthesis as
in~\cite{Gusfield:97,Aho:72} and like this for multiple citations with the same-named \nobreak{author:}
% \cite{Tam:06,Tam:07,Gledson:08homog,Gledson:08websearch}
(Tam and Schultz, 2006, 2007; Gledson and Keane, 2008a,b).

\paragraph{References:} Gather the full set of references together under the
heading {\bf References}; place the section before any Appendices, unless they
contain references. Arrange the references alphabetically by the first
author's last-name, rather than by order of occurrence in the text, and invert
the first-name and last-name of the first author (only). Provide as complete
a citation as possible, using a consistent format, such as the one for {\em
Computational Linguistics\/} or the one in the {\em Publication Manual of the
American Psychological Association\/}~\cite{APA:83}.  Use of full names for
authors rather than initials is preferred.  Use full names for journals and
conferences, not abbreviations (for example ``45th Meeting of the Association
for Computational Linguistics'', not ``ACL07'').

The \LaTeX2e{} and Bib\TeX{} style files provided roughly fit the
American Psychological Association format, allowing regular citations,
short citations and multiple citations as described above.

\paragraph{Appendices:} Appendices, if any, directly follow the text and the
references (but see above).  Letter them in sequence and provide an
informative title: {\bf Appendix A. Title of Appendix}.

\paragraph{Acknowledgements:} the acknowledgement section should go as a last section immediately
\textit{before the references}.  Do not number the acknowledgement section.

\subsection{Footnotes}

Put footnotes at the bottom of the page and use
9-point font. They may be numbered or referred to by asterisks or
other symbols.\footnote{This is how a footnote should appear. Footnote numbers in the main text should follow punctuation}
Footnotes should be separated from the main text by a
line.\footnote{Note the line separating the footnotes from the text.}

\subsection{Other formatting hints}
Camera-ready papers should follow commonly accepted typographical conventions. Here are some of them:
\begin{itemize}
\item Use an en-dash (\LaTeX{} ``--'') for language pairs: ``English--Swahili'', not ``English-Swahili''.
\item Use an en-dash (\LaTeX{} ``--'') for number ranges: ``3--4'', not ``3-4''.
\item When typesetting math, function or variable names with more than one letter should be typeset in Roman (using \verb|\mathrm|), not italics: ``\(\mathrm{distance}(a,b)\)'', not ``\(distance(a,b)\)''.
\item When typesetting subscripts which are words or word abbreviations, make sure you typeset them in Roman (using \verb|\mathrm|): ``\(A_\mathrm{max}\)'', not ``\(A_{max}\)''
\item Avoid informal abbreviations such as ``10k'' for ``10,000'' or ``4M'' for ``4,000,000'' or ``4 million''.
\item Separate quantities from units: ``4~GB'', not ``4GB''. Use an unbreakable space (\LaTeX{} ``\verb|~|'')
\end{itemize}

\subsection{Copyright}

\confname{} requires a copyright license statement. This should be inserted as
an unnumbered footnote on the first column of the first page. The \LaTeX{} style file
(and the Word template) adds the copyright statement automatically. Refer to the style file for instructions to
change this if required.

\subsection{Graphics}

\paragraph{Illustrations:} Place figures, tables, and photo\-graphs in the
paper near where they are first discussed, rather than at the end, if
possible.  Wide illustrations may run across both columns. Do not use
colour illustrations as they may reproduce poorly.

\paragraph{Captions:} Provide a caption for every illustration; number
each one sequentially in the form:  ``Figure 1. Caption of the
Figure.'', ``Table 1. Caption of the Table.''  Type the captions of
the figures and tables below the body, using 11-point text.

\section{Length of Submission}
\label{sec:length}

Unless otherwise specified, the maximum length is \paperlength~pages for research papers, \shortpaperlength~pages for user papers, \projectlength~pages for projects/products and
\translatorlength~pages for the translators' track papers. 
The page limit should be
observed strictly. All illustrations, references, and appendices must be
accommodated within these page limits, following the formatting instructions
given in the present document.

% \bibliography{\confname}

\begin{thebibliography}{}

\bibitem[\protect\citename{Aho and Ullman}1972]{Aho:72}
Aho, Alfred~V. and Jeffrey~D. Ullman.
\newblock 1972.
\newblock {\em The Theory of Parsing, Translation and Compiling}, volume~1.
\newblock Prentice-{Hall}, Englewood Cliffs, NJ.

\bibitem[\protect\citename{{American Psychological Association}}1983]{APA:83}
{American Psychological Association}.
\newblock 1983.
\newblock {\em Publications Manual}.
\newblock American Psychological Association, Washington, DC.

\bibitem[\protect\citename{{Association for Computing Machinery}}1983]{ACM:83}
{Association for Computing Machinery}.
\newblock 1983.
\newblock {\em Computing Reviews}, 24(11):503--512.

\bibitem[\protect\citename{Chandra \bgroup et al.\egroup }1981]{Chandra:81}
Chandra, Ashok~K., Dexter~C. Kozen, and Larry~J. Stockmeyer.
\newblock 1981.
\newblock Alternation.
\newblock {\em Journal of the Asso\-ciation for Computing Machinery},
  28(1):114--133.

\bibitem[\protect\citename{Gledson and Keane}2008a]{Gledson:08homog}
Gledson, Anne, and John Keane. 
\newblock 2008a. 
\newblock Measuring Topic Homogeneity and its Application to Dictionary-Based Word-Sense Disambiguation. 
\newblock {\em Coling 2008, 22nd International Conference on Computational Linguistics}, Manchester, UK.
\newblock 273--280.
\bibitem[\protect\citename{Gledson and Keane}2008b]{Gledson:08websearch}
Gledson, Anne, and John Keane. 
\newblock 2008b. 
\newblock Using Web-Search Results to Measure Word-group Similarity. \newblock {\em Coling 2008, 22nd International Conference on Computational Linguistics}, Manchester, UK.
\newblock 281--288.
\bibitem[\protect\citename{Gusfield}1997]{Gusfield:97}
Gusfield, Dan.
\newblock 1997.
\newblock {\em Algorithms on Strings, Trees and Sequences}.
\newblock Cambridge University Press, Cambridge, UK.

\bibitem[\protect\citename{Tam and Schultz}2006]{Tam:06}
Tam, Yik-Cheung and Tanja Schultz.
\newblock 2006. 
\newblock Unsupervised Language Model Adaptation Using
  \nobreak{Latent} Semantic Marginals.
\newblock {\em Interspeech 2006 -- ICSLP, Ninth International Conference on Spoken Language Processing}, 
Pittsburgh, Pennsylvania, paper 1705-Thu1A2O.2.

\bibitem[\protect\citename{Tam and Schultz}2007]{Tam:07}
Tam, Yik-Cheung and Tanja Schultz.
\newblock 2007.
\newblock Correlated \nobreak{Latent} Semantic Model for
  Unsupervised Language Model Adaptation.
\newblock {\em Proceedings of ICASSP 2007, International Conference on 
  Acoustics, Speech, and Signal Processing}, Honolulu, Hawaii, Vol. IV, 41--44.

\end{thebibliography}

\end{document}

### 2022-02-19 2022/eamt/eamt-latex/appendix.tex
add eamt

--- 2021-05-17 2021/emnlp/supplementary_material.tex
@@
-% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
-\pdfoutput=1
-% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.
-
-\documentclass[11pt]{article}
-
-% Remove the "review" option to generate the final version.
-\usepackage[review]{emnlp2021}
-
-% Standard package includes
-\usepackage{times}
-\usepackage{latexsym}
-
-% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
-\usepackage[T1]{fontenc}
-% For Vietnamese characters
-% \usepackage[T5]{fontenc}
-% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets
-
-% This assumes your files are encoded as UTF8
-\usepackage[utf8]{inputenc}
-\renewcommand{\UrlFont}{\ttfamily\small}
-\usepackage{graphicx}
-\usepackage{grffile}
-\usepackage{multirow}
-\usepackage{xcolor,colortbl}
-\usepackage{amsmath}
-%\usepackage{fdsymbol}
-\usepackage{amssymb}
-\usepackage{makecell}
-\usepackage[super]{nth}
-\usepackage{arydshln}
-\usepackage{algorithm,algorithmic}
-%\usepackage[noend]{algpseudocode}
-\usepackage{regexpatch}
-\usepackage{subcaption}
-% This is not strictly necessary, and may be commented out,
-% but it will improve the layout of the manuscript,
-% and will typically save some space.
-\usepackage{microtype}
-
-\usepackage{url}
-\usepackage{longtable}
-\usepackage{tikz}
-\usetikzlibrary{calc}
-\usepackage[draft]{todo}
-\usepackage[normalem]{ulem}
-\usepackage{xspace}
-\usepackage{float}
-
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-
-\let\oldbibitem\bibitem
-\def\bibitem{\vfill\oldbibitem}
-
-\usepackage{soulutf8}
-\usepackage{tabularx}
-
-\usepackage{multirow}
-\usepackage{adjustbox}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
-\newcommand{\fyFuture}[1]{\done[FY]\Todo[FY:]{\textcolor{red}{#1}}}
-\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
-
-\newcommand{\revision}[1]{\textcolor{red}{#1}}
-\newcommand{\revisiondel}[1]{}
-\newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
-\newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
-\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
-\newcommand{\system}[1]{\texttt{{#1}}}
-
-\newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
-\newcommand{\vtheta}{\ensuremath{\boldsymbol\theta}\xspace} % parameters vector for a distribution
-\newcommand{\vpsi}{\ensuremath{\boldsymbol\psi}\xspace} % parameters vector for a distribution
-\newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}
-% \newcommand{\SB}[1]{\textcolor{green}{#1}}
-% \newcommand{\SW}[1]{\textcolor{red}{#1}}
-\newcommand{\SB}[1]{\textbf{#1}}
-\newcommand{\SW}[1]{\underline{#1}}
-% limits underneath
-\DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\renewcommand\textfraction{.1}
-\renewcommand\floatpagefraction{.95}
-\newcommand{\sbcl}[2]{{\scriptsize #1 \hfill $|$ \hfill  #2}}
-\usepackage{multirow}
-\usepackage{adjustbox}
-
-\makeatletter
-\newenvironment{breakablealgorithm}
-  {% \begin{breakablealgorithm}
-   \begin{center}
-     \refstepcounter{algorithm}% New algorithm
-     \hrule height.8pt depth0pt \kern2pt% \@fs@pre for \@fs@ruled
-     \renewcommand{\caption}[2][\relax]{% Make a new \caption
-       {\raggedright\textbf{\fname@algorithm~\thealgorithm} ##2\par}%
-       \ifx\relax##1\relax % #1 is \relax
-         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##2}%
-       \else % #1 is not \relax
-         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##1}%
-       \fi
-       \kern2pt\hrule\kern2pt
-     }
-  }{% \end{breakablealgorithm}
-     \kern2pt\hrule\relax% \@fs@post for \@fs@ruled
-   \end{center}
-  }
-\makeatother
-
-\title{Supplementary material for paper ``Multi-Domain Adaptation in Neural Machine Translation with Dynamic Sampling Strategies''}
-
-\date{}
-
-\begin{document}
-\maketitle
-\setlength{\abovedisplayskip}{2pt}
-\setlength{\belowdisplayskip}{2pt}
-
 \section*{A - Generalized Multi-Domain Dynamic Adaptation Curriculum Algorithm}
 The pseudo-code for the generalized Multi-Domain Adaptation Dynamic Sampling Algorithm  is in algorithm~\ref{alg:mdmt}.
+\begin{breakablealgorithm}
+\caption{Multi-Domain Adaptation Dynamic Sampling} \label{alg:mdmt}
+\begin{algorithmic}[1]
+\REQUIRE {
+  \begin{itemize}
+    \setlength{\itemsep}{-1pt}    \setlength{\parsep}{0pt}
+  \item[] 
+  \item $n_d$ corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equipped by an empirical distribution $D_d(x)$
+  \item $n_d$  dev sets $Dev^d, d\in [1, \dots,n_d]$ for $n_d$ domains.
+  \item Domain testing distribution $\vlambda^t \in \mathbb{R}^{n_d}$
+  \item Batch size $B$
+  \item Domain Dynamic Sampling Distribution $\lambda^l_i \in \mathbb{R}^{n_d}$ where $i$ indexes iterations.
+  \item $Eval\_scores = []$
+  \item $Early\_stopping$ criterion
+  \item Total training iterations $iter\_num$
+  \item Update rule for sampling distribution $\vlambda^l_i$
+ \end{itemize}}
+\REPEAT 
+\STATE{// Start of iteration i}
+\STATE{Randomly pick $d \in [1,\dots,n_d]$ from sampling distribution $\lambda^l_i$}
+\STATE{Sample $B$ sentences from $C^d$ with empirical distribution $D_d(x)$}
+\STATE{Update model by applying SGD computed from $B$ sampled sentences}
+\IF{$i \equiv 0 \mod{eval\_step}$}
+	\STATE{Evaluate current model with $n_d$ dev sets. $S^d_i$ is the performance at iteration $i^{th}$ on domain $d$}
+	\STATE{Report weighted score using test distribution $\vlambda^t$. $$eval(i) = \displaystyle{\mathop{\sum}_d^{n_d} \lambda^t(d) S^d_i}$$}
+	\STATE{$Eval\_scores.append(eval(i))$}
+\ENDIF
+\IF{$i \equiv 0 \mod{sampler\_updating\_step}$}
+	\STATE Update $\vlambda^l_i$
+\ENDIF
+\IF{$Early\_stopping(Eval\_scores)$}
+	\STATE{break}
+\ENDIF
+\STATE{$i=i+1$}
+\UNTIL{$i> iter\_num$}
+\end{algorithmic}
+\end{breakablealgorithm}
 \section*{B - Experiments with automatic domains}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{breakablealgorithm}
-\caption{Multi-Domain Adaptation Dynamic Sampling} \label{alg:mdmt}
-\begin{algorithmic}[1]
-\REQUIRE {
-  \begin{itemize}
-    \setlength{\itemsep}{-1pt}    \setlength{\parsep}{0pt}
-  \item[] 
-  \item $n_d$ corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equipped by an empirical distribution $D_d(x)$
-  \item $n_d$  dev sets $Dev^d, d\in [1, \dots,n_d]$ for $n_d$ domains.
-  \item Domain testing distribution $\vlambda^t \in \mathbb{R}^{n_d}$
-  \item Batch size $B$
-  \item Domain Dynamic Sampling Distribution $\lambda^l_i \in \mathbb{R}^{n_d}$ where $i$ indexes iterations.
-  \item $Eval\_scores = []$
-  \item $Early\_stopping$ criterion
-  \item Total training iterations $iter\_num$
-  \item Update rule for sampling distribution $\vlambda^l_i$
- \end{itemize}}
-\REPEAT 
-\STATE{// Start of iteration i}
-\STATE{Randomly pick $d \in [1,\dots,n_d]$ from sampling distribution $\lambda^l_i$}
-\STATE{Sample $B$ sentences from $C^d$ with empirical distribution $D_d(x)$}
-\STATE{Update model by applying SGD computed from $B$ sampled sentences}
-\IF{$i \equiv 0 \mod{eval\_step}$}
-	\STATE{Evaluate current model with $K$ dev sets. $S^d_i$ is the performance at iteration $i^{th}$ on domain $d$}
-	\STATE{Report weighted score using test distribution $\vlambda^t$. $$eval(i) = \displaystyle{\mathop{\sum}_d^{n_d} \lambda^t(d) S^d_i}$$}
-	\STATE{$Eval\_scores.append(eval(i))$}
-\ENDIF
-\IF{$i \equiv 0 \mod{sampler\_updating\_step}$}
-	\STATE Update $\vlambda^l_i$
-\ENDIF
-\IF{$Early\_stopping(Eval\_scores)$}
-	\STATE{break}
-\ENDIF
-\UNTIL{$i> iter\_num$}
-\end{algorithmic}
-\end{breakablealgorithm}
-
-\end{document}

### 2022-02-19 2022/eamt/eamt-latex/eamt22.tex
add eamt

--- 2021-08-04 2021/wmt/auto_cl_mdmt.tex
@@
-% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
-\pdfoutput=1
-% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.
+%
+% File eamt20.tex
+%
+% Contact: cettolo@fbk.eu, mlf@dlsi.ua.es
+
+%%% To ease future customizations, various replaceables have been paramaterized
+%%% as listed in the newcommands section
 
 \documentclass[11pt]{article}
+\usepackage{eamt22}
+\usepackage{times}
+\usepackage{url}
+\usepackage{latexsym}
+\usepackage[small,bf]{caption} % added MLF 20171211
+\setlength\titlebox{6.5cm}    % Expanding the titlebox
+%%% YOUR PACKAGES BELOW THIS LINE %%%
 
-% Remove the "review" option to generate the final version.
-\usepackage[review]{emnlp2021}
+\newcommand{\confname}{EAMT 2022}
+\newcommand{\website}{https://eamt2022.com/}
+\newcommand{\urlwebsite}{https://eamt2022.com/}
+\newcommand{\contactname}{the guest organizers, LT3,}
+\newcommand{\contactemail}{\texttt{eamt2022@ugent.be}}
+\newcommand{\conffilename}{eamt22}
+\newcommand{\downloadsite}{https://eamt2022.com/templates}
 
-% Standard package includes
-\usepackage{times}
-\usepackage{latexsym}
+\newcommand{\paperlength}{$10$ (ten)}
+\newcommand{\shortpaperlength}{$10$ (ten)}
+\newcommand{\projectlength}{$2$ (two)}
+\newcommand{\translatorlength}{$10$ (ten)}
 
-% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
-\usepackage[T1]{fontenc}
-% For Vietnamese characters
-% \usepackage[T5]{fontenc}
-% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets
+\title{Multi-Domain Adaptation in Neural Machine Translation with Dynamic Sampling Strategies}
 
+\author{First Author\\
+  Affiliation / Address line 1\\
+  Affiliation / Address line 2\\
+  Affiliation / Address line 3\\
+  {\tt email@domain}  \And
+  Second Author\\
+  Affiliation / Address line 1\\
+  Affiliation / Address line 2\\
+  Affiliation / Address line 3\\
+  {\tt email@domain}}
+
+\date{}
 % This assumes your files are encoded as UTF8
@@
 \usepackage{arydshln}
+\usepackage{algorithm,algorithmic}
 
-\usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
+%\usepackage[noend]{algpseudocode}
 \usepackage{regexpatch}
@@
 
-\title{Multi-Domain Adaptation in Neural Machine Translation with Dynamic Sampling Strategies}
-
-\author{First Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{email@domain} \\\And
-  Second Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{email@domain} \\}
-
-\date{}
+\makeatletter
+\newenvironment{breakablealgorithm}
+  {% \begin{breakablealgorithm}
+   \begin{center}
+     \refstepcounter{algorithm}% New algorithm
+     \hrule height.8pt depth0pt \kern2pt% \@fs@pre for \@fs@ruled
+     \renewcommand{\caption}[2][\relax]{% Make a new \caption
+       {\raggedright\textbf{\fname@algorithm~\thealgorithm} ##2\par}%
+       \ifx\relax##1\relax % #1 is \relax
+         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##2}%
+       \else % #1 is not \relax
+         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##1}%
+       \fi
+       \kern2pt\hrule\kern2pt
+     }
+  }{% \end{breakablealgorithm}
+     \kern2pt\hrule\relax% \@fs@post for \@fs@ruled
+   \end{center}
+  }
+\makeatother
 
@@
 In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key component in training is then to mitigate the detrimental effects of a possible mismatch between these distributions.
-Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from and should not be confused with multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison,Saunders21asurvey} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
-A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Aharoni19massively},\fyDone{Fix ref}
-% \revision{cite also \citep{Shen21source}} -> I think this is not so appropriate here
+Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from and should not be confused with multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \newcite{Chu2017comparison,Saunders21asurvey} for a review), and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+A related situation is multilingual machine translation \cite{Firat16multiway,Ha16towards,Johnson17google,Aharoni19massively},\fyDone{Fix ref}
+% \revision{cite also \cite{Shen21source}} -> I think this is not so appropriate here
 where the heterogeneity of training data not only corresponds to variations in the topic, genre, or register but also in language.
 
-This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the test conditions \citep{Moore10selection,Axelrod11domain}.
-Finding the optimal balance of training data is, however, a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data \citep{Miceli-barone17regularization}.
+This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the test conditions \cite{Moore10selection,Axelrod11domain}.
+Finding the optimal balance of training data is, however, a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data \cite{Miceli-barone17regularization}.
 It may also be suboptimal, as some target domains or languages might be easier to train than others.
-Finally, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not achieve optimally for any particular domain.
+Finally, improving the performance of the MT system in one domain will often hurt that of another \cite{Vanderwees17dynamic,Britz17mixing}, and improving model generalization across all domains \cite{koehn18findings} may not achieve optimally for any particular domain.
 
 Several recent proposals
-% \citep{Vanderwees17dynamic,Graves17automated,platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Wang20balancing,Wang20learning-multi}
+% \cite{Vanderwees17dynamic,Graves17automated,platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Wang20balancing,Wang20learning-multi}
 have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies.
-Notably, \citet{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while \citet{Graves17automated,Platanios19competence,Kumar19reinforcement,Wang20learning-multi,Wang20balancing} build curricula that automatically adapt to the training data.
+Notably, \newcite{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while \newcite{Graves17automated,Platanios19competence,Kumar19reinforcement,Wang20learning-multi,Wang20balancing} build curricula that automatically adapt to the training data.
 
 In this paper, we contribute to this line of research in several ways. 
-First, we propose a novel framework (\emph{Multi-Domain Automated Curriculum}, MDAC for short) that simultaneously accounts for the domain adaptation and the multi-domain adaptation problems.
-Second, we propose a novel method of building an automated curriculum for training Multi-Domain NMT models.
-We evaluate our method in various single-domain and multi-domain adaptation settings.
-We compare our method with several contrasts, including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}, which was previously only applied to multilingual MT.
-% From now on, we name our method MDAC which is brief for .
+\begin{itemize}
+	\item First, we propose a novel framework (\emph{Multi-Domain Automated Curriculum}, MDAC for short), which is a variant of Differentiable Data Selection (DDS) of \newcite{Wang20balancing}, that simultaneously accounts for the domain adaptation and the multi-domain adaptation problems.
+	\item We demonstrate MDAC's equivalent performance compared to fine-tuning for domain adaptation; and better performance compared to classical choices of static data sampling strategies.
+	\item We show that our variant MDAC mitigate some failures of DDS in handling multi-domain training.
+	\item We demonstrate various exotic use cases of differentiable data selection frameworks (MDAC and original DDS), including, domain adaptation using unsupervised clustering instead of using domain label a priori which is barely available in real context; domain adaptation using out-of-domain training data and small in-domain validation data; and bi-domain adaptation in which we are only interested in two domain of the training mixture.
+\end{itemize}
+
 Based on these experimental results, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard fine-tuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test situations, from targeting one single domain to the full multi-domain scenario.
-% distributions
 
@@
 
-We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \citep{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \citep{Chen16guided,Zhang16topicinformed}, register \citep{Sennrich16politeness}, style \citep{Niu18multitask}, etc.
+We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
 Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
@@
 We consider just three domains, and represent vectors of mixture weights $\vlambda^{s}$ and $\vlambda^{t}$ in the 3-dimensional simplex.
-Training with weights in (a) and testing with weights in (c) is supervised multi-source domain adaptation to domain~2 ($d_2$), while (b)-(c) is the unsupervised\fyTodo{Unseen ?} version, with no training data from $d_2$; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with configurations (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
+Training with weights in (a) and testing with weights in (c) is supervised multi-source domain adaptation to domain~2 ($d_2$), while (b)-(c) is the unsupervised\fyDone{Unseen ?} version, with no training data from $d_2$; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with configurations (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
 \end{figure}
 
-These situations have been amply documented from a theoretical perspective (eg.\ \citet{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}).
+These situations have been amply documented from a theoretical perspective (eg.\ \newcite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}).
 A general recommendation in the DA setting is to adjust the sampling distribution used to optimize the system so as to compensate for the mismatch between $\mathcal{D}^s(x)$ and $\mathcal{D}^t(x)$.
@@
 
-A widely used approach to supervised DA is \emph{fine-tuning} \citep{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
+A widely used approach to supervised DA is \emph{fine-tuning} \newcite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
 With our notations, this approach amounts to first learning an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$) with $\indic{A}$ the indicator function for predicate $A$.
@@
 We divide the training into many short sessions; in each session $t$, the model is trained with a static data distribution $\vlambda^{l}(\psi_t)$.
-After one learning session, we update the data distribution using the REINFORCE algorithm of \citet{Williams92simple}.
+After one learning session, we update the data distribution using the REINFORCE algorithm of \newcite{Williams92simple}.
 The evolution of $\psi$ is thus defined by:
@@
 To compute the reward $R(d)$ of using data from domain $d$ to train the model, we simulate $k$ training steps from the current checkpoint by training the model with $k$ batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
-This computation is inspired by the target prediction gain in \citet{Graves17automated}.
-However, where \citet{Graves17automated} used accumulated gains from the past as rewards, we instead predict the usefulness of each domain for improving the future performance of the system given its current state.
+This computation is inspired by the target prediction gain in \newcite{Graves17automated}.
+However, where \newcite{Graves17automated} used accumulated gains from the past as rewards, we instead predict the usefulness of each domain for improving the future performance of the system given its current state.
 This is achieved by simulating a round of training with only the data from one domain.
-Furthermore, the choice of the parameterization of the sampling distribution in our method differs from that of \citet{Graves17automated}, although this choice is empirical.
+Furthermore, the choice of the parameterization of the sampling distribution in our method differs from that of \newcite{Graves17automated}, although this choice is empirical.
 
-The work of \citet{Wang20balancing} is also related: it is based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\vlambda^{l}$ that will result in the best NMT model with respect to a given target dev set at the end of training.
+The work of \newcite{Wang20balancing} is also related: it is based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\vlambda^{l}$ that will result in the best NMT model with respect to a given target dev set at the end of training.
 These authors also derive a similar form of update for $\psi$.
@@
 We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.
-We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}) \citep{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \citep{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \citep{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \citep{Cettolo12wit}, and the Koran (\domain{rel}).
+We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \cite{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
 Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
 Most corpora are available from the Opus website.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
-To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \citep{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
+To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
 %
 We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \citep{Papineni02bleu}.\footnote{We use truecasing and \texttt{sacrebleu} \citep{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \citep{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \citep{Neubig19compare-mt}.
+Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and \texttt{sacrebleu} \cite{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
 We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
@@
 \subsection{Baseline systems \label{ssec:baseline}}
-Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \citep{Britz17mixing}.} Using Transformers \citep{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \citep{Klein17opennmt}, we build the following systems:
+Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
 
@@
 with $\alpha \in \{0,0.25,0.5,0.75,1.0\}$. We denote these as \system{Mixed-$\alpha$} below. \system{Mixed-$0$} uses a uniform distribution, \system{Mixed-$1.0$} the empirical distribution of domains.
-\item fine-tuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ??
+\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ??
 iterations} The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems, each adapted for one domain, with no parameter sharing across domains.
 \item systems trained with fixed data mixtures corresponding to $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$; these are used in the multi-domain experiments of Section~\ref{ssec:mda};
-\item  our own implementations of recent dynamic sampling proposals from the literature: Curriculum Learning (CL) of \citet{Zhang19curriculum} and Differential Data Selection (DDS) of \citet{Wang20balancing} (see details below);
+\item  our own implementations of recent dynamic sampling proposals from the literature: Curriculum Learning (CL) of \newcite{Zhang19curriculum} and Differential Data Selection (DDS) of \newcite{Wang20balancing} (see details below);
 \end{itemize}
@@
 \subsection{CL and DDS's re-implementation}\fyDone{Right place for this ?}
-We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to the original code of \citet{Wang20balancing}.\footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}
+We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to the original code of \newcite{Wang20balancing}.\footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}
 % 
-We also re-implement the approach of \citet{Zhang19curriculum} according to the authors' description.
+We also re-implement the approach of \newcite{Zhang19curriculum} according to the authors' description.
 For each domain adaptation experiment, we combine the training data of all other domains into one corpus then compute the cross-entropy difference score of each source sentence of this combined dataset.
@@
 
-Finally, in the \emph{unsupervised domain adaptation task}, we cluster all available training data into 30~clusters using the KNN algorithm in the same way as in \citep{Tars18multidomain}, then adapt these clusters to one of 6 domains using the corresponding in-domain dev set.
+Finally, in the \emph{unsupervised domain adaptation task}, we cluster all available training data into 30~clusters using the KNN algorithm in the same way as in \cite{Tars18multidomain}, then adapt these clusters to one of 6 domains using the corresponding in-domain dev set.
 We compare MDAC to DDS for each of our 6 test sets.
@@
 This setting targets a uniform test distribution $\vlambda^t = \vlambda_0$.
-In this situation, CL \citep{Zhang19curriculum} does not apply.
+In this situation, CL \cite{Zhang19curriculum} does not apply.
 We therefore only contrast the performance of MDAC, DDS and several fixed training data distribution $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$, where $\vlambda_{\alpha}$ is defined according to equation~\eqref{mixture:trn}.
@@
 Domain adaptation is an old problem that has been studied from many angles, both for SMT and NMT.
-A survey of supervised and unsupervised DA for NMT is in \citet{Chu18asurvey}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the more recent survey of \citet{Saunders21asurvey}.
+A survey of supervised and unsupervised DA for NMT is in \newcite{Chu18asurvey}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the more recent survey of \newcite{Saunders21asurvey}.
 Our approach to DA in this paper falls under the former category.
@@
 Multi-domain adaptation (MDA) aims to develop systems that simultaneously bode well for several domains.
-Like for DA, techniques for supervised MDA combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \citep{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citet{Jiang19multidomain}.
-Unsupervised MDA is studied in \citet{Farajian17multidomain}, as an instance of unsupervised DA.
+Like for DA, techniques for supervised MDA combine one or several ingredients: (a) the specialization of data representations \cite{Kobus17domaincontrol} or of sub-networks \cite{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \newcite{Jiang19multidomain}.
+Unsupervised MDA is studied in \newcite{Farajian17multidomain}, as an instance of unsupervised DA.
 
-Most approaches to adaptive/dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of curriculum learning is introduced.
+Most approaches to adaptive/dynamic data selection take inspiration from \newcite{Bengio09curriculum}, where the notion of curriculum learning is introduced.
 CL relies on the notion of the ``easiness'' of a sample to schedule the presentation of training data so that the easiest examples are presented first and the hardest last.
-Various ways to automate CL using the framework of multi-armed bandits are explored in \citep{Graves17automated}, which has been an inspiration for our implementation.
+Various ways to automate CL using the framework of multi-armed bandits are explored in \cite{Graves17automated}, which has been an inspiration for our implementation.
 While the initial aim was primarily to improve and speed up training, CL has also proven useful for domain adaptive / multi-domain / multilingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
-For instance, \citet{Zhang19curriculum} study supervised DA and propose a curriculum approach which progressively augments the training data: in the early stages, only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
-This is somehow opposite to the recommendations of \citet{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyDone{These have not been compared?
+For instance, \newcite{Zhang19curriculum} study supervised DA and propose a curriculum approach which progressively augments the training data: in the early stages, only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
+This is somehow opposite to the recommendations of \newcite{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyDone{These have not been compared?
 and also to what we do ?}
 
-\citet{Kumar19reinforcement} use reinforcement techniques (deep Q-learning) to learn the curriculum strategy: in this work, complexity corresponds to difficulty levels which are binned using contrastive data selection.
-The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyDone{Alert: what do we do during warm-up ?} The same technique is recently applied to multilingual NMT in \citep{Kumar21learning}.
-\citet{Zhou20uncertainty} propose another curriculum-based approach which instead relies on \emph{instance uncertainty} as a measure of their difficulty and presents the data sample starting with the easiest (more predictable) first.
+\newcite{Kumar19reinforcement} use reinforcement techniques (deep Q-learning) to learn the curriculum strategy: in this work, complexity corresponds to difficulty levels which are binned using contrastive data selection.
+The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyDone{Alert: what do we do during warm-up ?} The same technique is recently applied to multilingual NMT in \cite{Kumar21learning}.
+\newcite{Zhou20uncertainty} propose another curriculum-based approach which instead relies on \emph{instance uncertainty} as a measure of their difficulty and presents the data sample starting with the easiest (more predictable) first.
 Another contribution of this paper is an alternative stopping criterium.
-Closest to our problems, \citet{Wang20learning-multi} adapt CL for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
+Closest to our problems, \newcite{Wang20learning-multi} adapt CL for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
 Each step consists of (a) weighting instances based on relevance features, (b) fine-tuning a pretrained model using the weighted training set, and is applied iteratively to train a sequence of models.
@@
 In the future, we intend to continue developing this approach and improve its effectiveness.
-One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training \citep{Kumar19reinforcement}.
+One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training \cite{Kumar19reinforcement}.
 Another area where we need to progress is the unsupervised learning setting of Section~\ref{ssec:clda}, where our results still lag way behind that of supervised DA - this might be due to the inability of our simplistic optimization strategy to handle situations where the number of clusters is too large.
+%\clearpage{}
 % slow down learning during warm up or normalize reward; handle large number of classes;
@@
 % \section*{Acknowledgments}
-\bibliographystyle{acl_natbib}
+%The work presented in this paper was partially sup-ported by the European Commission under contractH2020-787061 ANITA
 \bibliography{multidomain}
+\bibliographystyle{eamt22}
+%\clearpage{}
 %\appendix
+%\input{appendix.tex}
 \end{document}

### 2022-03-01 2022/eamt/eamt-latex/eamt22.tex

@@
 \begin{itemize}
-	\item First, we propose a novel framework (\emph{Multi-Domain Automated Curriculum}, MDAC for short), which is a variant of Differentiable Data Selection (DDS) of \newcite{Wang20balancing}, that simultaneously accounts for the domain adaptation and the multi-domain adaptation problems.
-	\item We demonstrate MDAC's equivalent performance compared to fine-tuning for domain adaptation; and better performance compared to classical choices of static data sampling strategies.
-	\item We show that our variant MDAC mitigate some failures of DDS in handling multi-domain training.
-	\item We demonstrate various exotic use cases of differentiable data selection frameworks (MDAC and original DDS), including, domain adaptation using unsupervised clustering instead of using domain label a priori which is barely available in real context; domain adaptation using out-of-domain training data and small in-domain validation data; and bi-domain adaptation in which we are only interested in two domain of the training mixture.
+	\item First, we propose a novel framework (\emph{Multi-Domain Automated Curriculum}, MDAC for short), a variant of Differentiable Data Selection (DDS) of \newcite{Wang20balancing} initially applied to multilingual NMT, that simultaneously accounts for the domain adaptation and the multi-domain adaptation problems.
+	\item We demonstrate MDAC's equivalent performance compared to fine-tuning for domain adaptation \ref{ssec:da} and better performance compared to classical choices of static data sampling strategies for multi-domain settings \ref{ssec:mda}.
+	\item We show that our variant MDAC mitigates some failures of DDS in handling multi-domain training.
+	\item We demonstrate various exotic use cases of differentiable data selection frameworks (MDAC and original DDS), including domain adaptation using unsupervised clustering instead of using domain label a priori \ref{ssec:clda}, which is barely available in real context; domain adaptation using out-of-domain training data and small in-domain validation data \ref{ssec:uda}; and bi-domain adaptation in which we are only interested in two domain of the training mixture \ref{ssec:bida}.
 \end{itemize}
-
-Based on these experimental results, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard fine-tuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test situations, from targeting one single domain to the full multi-domain scenario.
 
@@
 
-\subsection{Multi-domain adaptation}\label{ssec:mda}
+\subsection{Multi-domain NMT}\label{ssec:mda}
 
-We now turn to a more realistic scenario and consider multi-domain adaptation, which aims to train one single system with optimal performance averaged over 6 domains.
+We now turn to a more realistic scenario and consider multi-domain NMT, which aims to train one single system with optimal performance averaged over 6 domains.
 This setting targets a uniform test distribution $\vlambda^t = \vlambda_0$.
@@
 
-Multi-domain adaptation (MDA) aims to develop systems that simultaneously bode well for several domains.
-Like for DA, techniques for supervised MDA combine one or several ingredients: (a) the specialization of data representations \cite{Kobus17domaincontrol} or of sub-networks \cite{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \newcite{Jiang19multidomain}.
-Unsupervised MDA is studied in \newcite{Farajian17multidomain}, as an instance of unsupervised DA.
+Multi-domain NMT (MDMT) aims to develop systems that simultaneously bode well for several domains.
+Like for DA, techniques for supervised MDMT combine one or several ingredients: (a) the specialization of data representations \cite{Kobus17domaincontrol} or of sub-networks \cite{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \newcite{Jiang19multidomain}.
+Unsupervised MDMT is studied in \newcite{Farajian17multidomain}, as an instance of unsupervised DA.
 
@@
 Another contribution of this paper is an alternative stopping criterium.
-Closest to our problems, \newcite{Wang20learning-multi} adapt CL for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
+Closest to our problems, \newcite{Wang20learning-multi} adapt CL for multi-domain NMT, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
 Each step consists of (a) weighting instances based on relevance features, (b) fine-tuning a pretrained model using the weighted training set, and is applied iteratively to train a sequence of models.
@@
 \fyDone{Write conclusion}
-In this study, we have presented a generic framework to perform a variety of adaptation tasks for machine translation, ranging from conventional supervised domain adaptation to multi-domain adaptation and unseen domain adaptation.
+In this study, we have presented a generic framework to perform a variety of adaptation tasks for machine translation, ranging from conventional supervised domain adaptation to multi-domain NMT and unseen domain adaptation.
 By experimenting with all these settings, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during the course of training, could be used in all these situations.

### 2022-03-12 2022/acl/acl_latex.tex
Few fixes after the ARR comments
local changes

@@
 
-\title{Multi-domain, multilingual translation with latent multi-task group dropout}
+%\title{Multi-domain, multilingual translation with latent multi-task group dropout}
 \title{Latent Group Dropout for Multilingual and Multidomain Machine Translation}
@@
 \item We present algorithms to train this model end-to-end with very little extra parameters;
-\item We report, using an extensive set of experiments, significant gains with respect to strong baselines for low-resource multilingual MT;
+\item We report, using an extensive set of experiments, gains for multidomain MT and very low-resourced languages in multilingual MT;
 \item We study how this method can actually exploit the similarities between tasks to learn interpretable sub-networks.
@@
 We set the dropout probability to 0.1.
-We train the multidomain Transformer model for 200k iterations with a batch size of 12k tokens.
+We train the multidomain Transformer model for 200k iterations with a batch size of 12k tokens using 4 V100 GPUs.
 The convergence of the standard Transformer is before 200K as its validation curve became flat near the 200K-th iteration.
@@
 We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyDone{Add public to the data link} The datasets \footnote{See \url{https://github.com/qmpham/experiments/tree/main/tacl20}} for the multidomain translation experiments are detailed in Table~\ref{tab:Corpora-chap4}.
+For each domain, the size of the dev set and the test set is 1~K.
 \begin{table*}[h!]
@@
     \system{HMGD} Transformer   \hfill{\footnotesize[+0m]}  & 40.4 & 60.4 & 51.9 & 38.7 &	50.80 &	86.80 & 54.8 \\ 
-    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & \SB{91.0} & \SB{55.3} \\ 
-    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & \SB{39.0} & \SB{52.4} & 87.5 & \SB{55.3} \\ 
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.3 \\ 
+    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  & 40.3 & \SB{60.4} & \SB{52.4} & \SB{39.0} & \SB{52.4} & \SB{87.5} & \SB{55.3} \\ 
     \hline
   \end{tabular}
-  \caption{Multi-domain translation}
+  \caption{Multidomain translation experiment. Boldface denotes significant gains over \system{Transformer} ($p=0.05$)}
   \label{tab:mdmt}
@@
   \end{tabular}
-  \caption{Experiments with two similar pseudo-domains }
+  \caption{Experiments with two similar pseudo-domains}
   \label{tab:fuzzy}
@@
     \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&34.0&38.3&8.3&20.19 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &10.2&4&\SB{13.3}&21.9&32.2&34.1&38.5&8.3&20.31 \\ 
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &10.2&4&13.3&21.9&32.2&34.1&38.5&8.3&20.31 \\ 
     \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
@@
   \includegraphics[width=0.7\textwidth]{multilingual_heatmap.png}  
-  \caption{Multilingual}
+  \caption{Multilingual (Related)}
   \label{fig:ml-heatmap}
@@
 We discuss here the choice of the hyper-parameters $k$, the number of activated nodes in each layer, and its impact on the sharing level between the tasks.
-Table ~\ref{tab:K} shows the variance of performance when the number of activated nodes is changed, and the sharing level between tasks decreases in consequence.\Todo{add sharing results.
-}\Todo{check caption consistency}
+Table ~\ref{tab:K} shows the variance of performance when the number of activated nodes is changed, and the sharing level between tasks decreases in consequence.
 \begin{table}[h!]
@@
 \begin{align}
-\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} = \frac{\partial \hat{m}_l^d(\tau)}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d}
+\hspace{-3.em}
+\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} = \frac{\partial \hat{m}_l^d(\tau)}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d} + \frac{1}{\tau} \frac{exp(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau})}{(1+exp(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}))^2}
 \end{align}
@@
 \STATE{Compute implicit gradient of the $\operatorname{Soft-Top-K}$ masks over $\Phi_l^d$
-$$- \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}$$}
+$$\frac{\partial \bar{\nu}}{\partial \Phi_l^d} = - \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}$$
+$$\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} = \frac{\partial \hat{m}_l^d(\tau)}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d} + \frac{1}{\tau} \frac{exp(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau})}{(1+exp(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}))^2}$$}
 \STATE{Compute the gradient the training over $\Phi_l^d$
-$$\Delta_{\Phi_l^d} = - \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d} + \frac{\partial \mathbb{H} \big[ \operatorname{softmax}(\Phi_l^d)\big] }{\partial \Phi_l^d} $$}
+$$\Delta_{\Phi_l^d} = \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} + \frac{\partial \mathbb{H} \big[ \operatorname{softmax}(\Phi_l^d)\big] }{\partial \Phi_l^d} $$}
 \STATE{Update $\theta$ and $\Phi_l^d$ with their gradients}

### 2022-03-28 2022/eamt/eamt-latex/eamt22.tex
start shortening the paper

@@
   Affiliation / Address line 2\\
-  Affiliation / Address line 3\\
+%  Affiliation / Address line 3\\
   {\tt email@domain}  \And
@@
   Affiliation / Address line 2\\
-  Affiliation / Address line 3\\
+%  Affiliation / Address line 3\\
   {\tt email@domain}}
@@
 Such multi-source / multi-domain adaptation problems are typically approached through instance selection or reweighting strategies, based on a static assessment of the relevance of training instances with respect to the task at hand.
-In this paper, we study dynamic data selection strategies that are able to automatically re-evaluate the usefulness of data samples and to evolve a data selection policy in the course of training.
-Based on the results of multiple experiments, we show that such methods constitute a generic framework to automatically and effectively handle a variety of real-world situations, from multi-source domain adaptation to multi-domain learning and unsupervised domain adaptation.
+In this paper, we study dynamic data selection strategies that are able to automatically re-evaluate the usefulness of data samples
+  % and to evolve a data selection policy
+  in the course of training. Based on the results of multiple experiments, we show that our method offer a generic framework to automatically
+  % and effectively
+  handle several real-world situations, from multi-source or unsupervised domain adaptation to multidomain learning.
 \end{abstract}
@@
 A typical setting in machine translation (MT) is to collect the largest possible collection of parallel data for the chosen language pair, with the intent to achieve optimal performance for the task of interest.
-In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key component in training is then to mitigate the detrimental effects of a possible mismatch between these distributions.
-Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from and should not be confused with multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \newcite{Chu2017comparison,Saunders21asurvey} for a review), and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
-A related situation is multilingual machine translation \cite{Firat16multiway,Ha16towards,Johnson17google,Aharoni19massively},\fyDone{Fix ref}
+In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key aspect of training is then to mitigate the detrimental effects of a mismatch between these distributions.
+Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from; this setting differs from multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \newcite{Chu2017comparison,Saunders21asurvey} for a review), and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+A related situation is multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Aharoni19massively},\fyDone{Fix ref}
 % \revision{cite also \cite{Shen21source}} -> I think this is not so appropriate here
-where the heterogeneity of training data not only corresponds to variations in the topic, genre, or register but also in language.
+where the diversity of training data not only corresponds to variations in the topic, genre, or register but also in language.
 
-This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the test conditions \cite{Moore10selection,Axelrod11domain}.
-Finding the optimal balance of training data is, however, a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data \cite{Miceli-barone17regularization}.
-It may also be suboptimal, as some target domains or languages might be easier to train than others.
-Finally, improving the performance of the MT system in one domain will often hurt that of another \cite{Vanderwees17dynamic,Britz17mixing}, and improving model generalization across all domains \cite{koehn18findings} may not achieve optimally for any particular domain.
+This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the testing conditions \cite{Moore10selection,Axelrod11domain}.
+Finding the optimal balance of training data is however, a challenging task due, for instance, to the similarity between domains/languages, or to the regularization effects of out-of-domain data \cite{Miceli-barone17regularization}.
+A static policy may also be suboptimal when some target domains or languages are easier to train than others.
+Finally, improving the performance of the MT system in one domain will often hurt that of another \cite{Vanderwees17dynamic,Britz17mixing} and improving model generalization across all domains \cite{koehn18findings} may not achieve optimally for any particular domain.
 
@@
 % \cite{Vanderwees17dynamic,Graves17automated,platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Wang20balancing,Wang20learning-multi}
-have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies.
-Notably, \newcite{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while \newcite{Graves17automated,Platanios19competence,Kumar19reinforcement,Wang20learning-multi,Wang20balancing} build curricula that automatically adapt to the training data.
-
+explore ways to instead consider \emph{dynamic} data selection and sampling strategies: \newcite{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while \newcite{Graves17automated,Platanios19competence,Kumar19reinforcement,Wang20learning-multi,Wang20balancing} build curricula that automatically adapt to the training data.
 In this paper, we contribute to this line of research in several ways.
 \begin{itemize}
-	\item First, we propose a novel framework (\emph{Multi-Domain Automated Curriculum}, MDAC for short), a variant of Differentiable Data Selection (DDS) of \newcite{Wang20balancing} initially applied to multilingual NMT, that simultaneously accounts for the domain adaptation and the multi-domain adaptation problems.
-	\item We demonstrate MDAC's equivalent performance compared to fine-tuning for domain adaptation \ref{ssec:da} and better performance compared to classical choices of static data sampling strategies for multi-domain settings \ref{ssec:mda}.
-	\item We show that our variant MDAC mitigates some failures of DDS in handling multi-domain training.
-	\item We demonstrate various exotic use cases of differentiable data selection frameworks (MDAC and original DDS), including domain adaptation using unsupervised clustering instead of using domain label a priori \ref{ssec:clda}, which is barely available in real context; domain adaptation using out-of-domain training data and small in-domain validation data \ref{ssec:uda}; and bi-domain adaptation in which we are only interested in two domain of the training mixture \ref{ssec:bida}.
+	\item First, we propose a novel framework (\emph{Multi-Domain Automated Curriculum}, MDAC for short), a variant of Differentiable Data Selection (DDS) of \newcite{Wang20balancing}, initially applied to multilingual NMT, that simultaneously accounts for the domain adaptation and the multidomain adaptation problems.
+	\item We show that MDAC achieves performance that compare to fine-tuning strategies for DA (\textsection~\ref{ssec:da}) and outperform some static data sampling strategies for multidomain settings (\ref{ssec:mda}).
+	\item We show that our variant MDAC mitigates some failures of DDS in multidomain training.
+	\item We illustrate the generality of differentiable data selection frameworks (both MDAC and DDS) on less common situations such as DA using unsupervised clustering (\textsection~\ref{ssec:clda}); DA using out-of-domain training data and small in-domain validation data (\textsection~\ref{ssec:uda}); and two-domain adaptation where the test distribution only mixes two of the training domain (\textsection~\ref{ssec:bida}).
 \end{itemize}

### 2022-03-28 2022/eamt/eamt-latex/eamt22.tex
reviewed till section 5

@@
 
-We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
+We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
 Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
-$h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$), but might differ from one domain to the other.
+$h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$) but may differ across domains.
 
-It is usual in MT to opportunistically collect training samples from several domains, which means that training instances are distributed according to a mixture $\mathcal{D}^s$ such that $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}(d) \mathcal{D}_d(x)$, with $\{\lambda^{s}(d), d=1 \dots n_d\}$ the mixture weights satisfying $\sum_d \lambda^{s}(d)=1$.
-In the sequel, we also use boldface to denote the vector $\vlambda$, and $\lambda(d)$ is used to denote a scalar corresponding to the $d^{th}$ element of the vector $\vlambda$.
+It is usual in MT to opportunistically collect corpora from several domains, which means that training instances are distributed according to a mixture $\mathcal{D}^s$ such that $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}(d) \mathcal{D}_d(x)$, with $\{\lambda^{s}(d), d=1 \dots n_d\}$ the mixture weights satisfying $\sum_d \lambda^{s}(d)=1$.
+In the sequel, boldface $\vlambda$ denotes a vector with $\lambda(d)$ the $d^{th}$ component of $\vlambda$.
 
-The main challenge in this situation is to make the best of these heterogeneous data, with the aim to achieve the optimal performance for the intended test conditions.
-These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unseen domain adaptation); in multi-domain adaptation finally, the test distribution is itself a mixture of domains, some of which may also be observed in training.
-Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of pure domain adaptation (see illustration on Figure~\ref{fig:mdmt-lambdas}).
+The main challenge in this situation is to make the best of heterogeneous data, with the aim to achieve optimal performance for the target test conditions.
+These might correspond to data from just one of the training domains, as in standard supervised domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unseen domain adaptation); in multidomain adaptation finally, the test distribution is itself a mixture of domains, some of which may also be observed in training.
+We thus assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of domain adaptation (see Figure~\ref{fig:mdmt-lambdas}).
 \begin{figure}[h]
   \centering
+  \vspace{-\baselineskip}
   \includegraphics[width=0.48\textwidth]{mdmt-lambdas}
   \caption{Training and testing with distribution mismatch.
-We consider just three domains, and represent vectors of mixture weights $\vlambda^{s}$ and $\vlambda^{t}$ in the 3-dimensional simplex.
-Training with weights in (a) and testing with weights in (c) is supervised multi-source domain adaptation to domain~2 ($d_2$), while (b)-(c) is the unsupervised\fyDone{Unseen ?} version, with no training data from $d_2$; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with configurations (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
+We consider three domains and represent $\vlambda^{s}$ and $\vlambda^{t}$ in the 3-dimensional simplex.
+Training with weights in (a) and testing with weights in (c) is supervised multi-source domain adaptation to domain~2 ($d_2$), while (b)-(c) is the unsupervised\fyDone{Unseen ?} version, with no training data from $d_2$; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with settings (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
 \end{figure}
@@
 
-A widely used approach to supervised DA is \emph{fine-tuning} \newcite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning.
-With our notations, this approach amounts to first learning an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$) with $\indic{A}$ the indicator function for predicate $A$.
-Note that this strategy is potentially suboptimal, as some out-of-domain samples may contribute to the final performance due to eg.\ domain similarity.
-Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains might be easier to learn than others.\fyDone{How to measure this?}
+A widely-used approach to supervised DA is \emph{fine-tuning} \cite{Luong15stanford}, where $\vlambda^{l}$ varies during learning.
+With our notations, this approach first learns an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then continues training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$), with $\indic{A}$ the indicator function for predicate $A$.
+This strategy is potentially suboptimal as some out-of-domain samples may contribute to the final performance due to e.g.\ domain similarity.
+Optimizing the learning distribution in multidomain settings is even more challenging as the learner needs to take advantage of possible domains overlaps and also of the fact that some domains might be easier to learn than others.\fyDone{How to measure this?}
 % \revision{Answer: The graphic of \system{Mixed-0} shows that with the same amount of data(batches) the convergences of the domains vary significantly}
@@
 \subsection{Basic principles}
-Assuming training data in each of the $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
-We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t_d}$ to denote the empirical train and test distributions for domain $d$, and $\widehat{\mathcal{D}^{u}}(x;\vlambda^{u}) = \sum_{d} \lambda^{u}(d) \widehat{\mathcal{D}^{u}_d}(x)$ for $u\in\{l,t\}$.
-In our setting,  $\vlambda^t$, and hence $\widehat{\mathcal{D}^t}(x;\vlambda^t)$ are fixed and predefined, approximated with an equivalent number of development corpora.
+Assuming training data in each of the $n_d$ domains $d_1 \dots d_{n_d}$, the size of the training corpus in domain $d$ is denoted $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples.
+$\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t_d}$ denote the empirical train and test distributions for domain $d$ and $\widehat{\mathcal{D}^{u}}(x;\vlambda^{u}) = \sum_{d} \lambda^{u}(d) \widehat{\mathcal{D}^{u}_d}(x)$ for $u\in\{l,t\}$.
+In our setting,  $\vlambda^{t}$ and hence $\widehat{\mathcal{D}^t}(x;\vlambda^t)$ are fixed and predefined, approximated with an equivalent number of development corpora.
 
 % Our method
-MDAC constructs an adaptative training distribution $\vlambda^{l}$ that optimizes the data selection policy along with the training of the NMT model.
-We parameterize $\vlambda^{l}$ by a differentiable function $\vlambda^l(\psi)$, which will be described in detail in Section \ref{ssec:dds-sys}.
+MDAC builds an adaptative training distribution $\vlambda^{l}$ that optimizes the data selection policy along with the training of the model.
+We parameterize $\vlambda^{l}$ by a differentiable function $\vlambda^l(\psi)$, which is described in \textsection~\ref{ssec:dds-sys}.
 We divide the training into many short sessions; in each session $t$, the model is trained with a static data distribution $\vlambda^{l}(\psi_t)$.
 After one learning session, we update the data distribution using the REINFORCE algorithm of \newcite{Williams92simple}.
-The evolution of $\psi$ is thus defined by:
+The evolution of $\psi$ is thus defined by:\fyTodo{$t$ is both for time, and for test - change into $\tau$?}
 \begin{align*}
@@
 where the reward $R(d)$ is computed as:
-\begin{align*}
-  R(d) = J^t(\theta_{t+k},\vlambda^t) - J^t(\theta_t,\vlambda^t),
-\end{align*}
+\begin{align}
+  R(d) = J^t(\theta_{t+k},\vlambda^t) - J^t(\theta_t,\vlambda^t), \label{eq:reward}
+\end{align}
 and where we also define:
@@
 
-To compute the reward $R(d)$ of using data from domain $d$ to train the model, we simulate $k$ training steps from the current checkpoint by training the model with $k$ batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss.
-This computation is inspired by the target prediction gain in \newcite{Graves17automated}.
+To compute the reward $R(d)$ associated to training the model with data from domain $d$, we simulate $k$ training steps from the current checkpoint, using $k$ batches sampled from $D^l(d)$ and computing the gain of the weighted dev-loss.
+This computation is inspired by the target prediction gain of \newcite{Graves17automated}.
 However, where \newcite{Graves17automated} used accumulated gains from the past as rewards, we instead predict the usefulness of each domain for improving the future performance of the system given its current state.
 This is achieved by simulating a round of training with only the data from one domain.
-Furthermore, the choice of the parameterization of the sampling distribution in our method differs from that of \newcite{Graves17automated}, although this choice is empirical.
+We also differ from this authors in the parameterization of the sampling distribution.
+% Furthermore, the choice of the parameterization of the sampling distribution in our method differs from that of \newcite{Graves17automated}, although this choice is empirical.
 
-The work of \newcite{Wang20balancing} is also related: it is based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\vlambda^{l}$ that will result in the best NMT model with respect to a given target dev set at the end of training.
+The work of \newcite{Wang20balancing} is also related: it is based on the bi-level optimization framework, which aims to find an optimal static distribution $\vlambda^{l}$ that will result in the best model with respect to a given target dev set at the end of training.
 These authors also derive a similar form of update for $\psi$.
@@
 \subsection{MDAC for (multi) domain adaptation}
-The setting developed in previous sections is quite general and can, in principle, accommodate the variety of situations mentioned above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training.
-In our experiments, we would like to better assess the true potential of MDAC in these settings and seek to experimentally answer the following questions:
+The setting developed in previous sections is quite general and can, in principle, accommodate the variety of situations mentioned above, and many more: basic DA, multidomain adaptation with various target distributions, possibly including domains unseen in training.
+In our experiments, we would like to better assess the potential of MDAC in these settings and seek to study the following questions:
 \begin{itemize}
-\item is MDAC a viable alternative to conventional fine-tuning? In particular, does it enable to better take advantage of relevant data from other domains?
-\item is MDAC also a viable option in multi-domain adaptation scenarios?
-\item does MDAC also enable to perform \emph{unsupervised} (multi-)domain adaptation? \fyDone{TBContinued}
+\item is MDAC a viable alternative to fine-tuning? In particular, does it enable to better take advantage of relevant data from other domains?
+\item is MDAC a viable option in multidomain adaptation scenarios?
+\item does MDAC enable to perform \emph{unsupervised} (multi-)domain adaptation? \fyDone{TBContinued}
 \end{itemize}
@@
 \subsection{Data and metrics \label{ssec:corpora}}
-We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.
+We experiment with translation from English into French in 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.
 We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \cite{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
-Complementary experiments also use v12 of the News Commentary corpus (\domain{news}).
+Additional experiments use the News Commentary corpus (\domain{news}).
 Most corpora are available from the Opus website.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
-To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
+To reduce the number of types, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
 %
@@
 \subsection{Baseline systems \label{ssec:baseline}}
-Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
-
+Our baselines are standard for multidomain settings.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:
 \begin{itemize}
@@
 \begin{align} \label{mixture:trn}
-\lambda_{\alpha}(d) = \frac{q_d^{\alpha}}{\displaystyle{\mathop{\sum}_{d=1}^{n_d}q_d^{\alpha}}} &&
+\lambda_{\alpha}(d) = (\sum_{d=1}^{n_d}q_d^{\alpha})^{-1} (q_d^{\alpha}) &&
 q_d = \frac{\mid N^{s}_d \mid}{\displaystyle{N^{s}}} % \mathop{\sum}_{i=1}^K\mid D_i \mid}}
@@
 with $\alpha \in \{0,0.25,0.5,0.75,1.0\}$. We denote these as \system{Mixed-$\alpha$} below. \system{Mixed-$0$} uses a uniform distribution, \system{Mixed-$1.0$} the empirical distribution of domains.
-\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ??
-iterations} The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems, each adapted for one domain, with no parameter sharing across domains.
-\item systems trained with fixed data mixtures corresponding to $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$; these are used in the multi-domain experiments of Section~\ref{ssec:mda};
-\item  our own implementations of recent dynamic sampling proposals from the literature: Curriculum Learning (CL) of \newcite{Zhang19curriculum} and Differential Data Selection (DDS) of \newcite{Wang20balancing} (see details below);
+\item fine-tuned models based on \system{Mixed-$1.0$}, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ??
+iterations} The fine-tuning (\system{FT-Full}) procedure updates all the parameters of the initial model, resulting in six systems, one per domain, with no parameter sharing across domains.
+\item systems trained with fixed mixtures with $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$; these are used in the multidomain experiments of \textsection~\ref{ssec:mda};
+\item  our implementations of dynamic sampling proposals from the literature: Curriculum Learning (CL) of \newcite{Zhang19curriculum} and Differential Data Selection (DDS) of \newcite{Wang20balancing} (see below);
 \end{itemize}
@@
 
-\subsection{CL and DDS's re-implementation}\fyDone{Right place for this ?}
+\subsection{CL and DDS re-implementations}\fyDone{Right place for this ?}
 We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to the original code of \newcite{Wang20balancing}.\footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}
@@
 We also re-implement the approach of \newcite{Zhang19curriculum} according to the authors' description.
-For each domain adaptation experiment, we combine the training data of all other domains into one corpus then compute the cross-entropy difference score of each source sentence of this combined dataset.
-We then sort and split the corpus into 9 shards and execute curriculum learning with 10 shards, using the in-domain corpus as the first shard.
+For each DA experiment, we combine the training data of all other domains into one corpus then compute the cross-entropy difference score of each source sentence of this combined dataset.
+We then sort and split the corpus into 9 shards and execute curriculum learning with 10 shards, using the in-domain data as the first shard.
 
 \subsection{MDAC systems} \label{ssec:dds-sys}
-The behavior of MDAC only depends on (a) the initial domain distribution at the start of training $\vlambda^{l}_{t=0}$, and (b) the targeted (dev/test) distribution $\vlambda^{t}$.
+The behavior of MDAC only depends on (a) the initial domain distribution at the start of training $\vlambda^{l}_{t=0}$, and (b) the target (dev/test) distribution $\vlambda^{t}$.
 We thus report these systems as \system{MDAC($\vlambda^{l}_{t=0}$, $\vlambda^{t}$)} and compare with DDS using the same settings.
 
-In our work, we parameterize the distribution $\vlambda^l$ as follows (with $\beta=2$ in all experiments):\footnote{This setting corresponds to the \emph{spherical softmax} of \cite{Brebisson16anexploration}.}\fyDone{Explain why}
+In our work, we parameterize the distribution $\vlambda^l$ as follows (with $\beta=2$ in all experiments):\footnote{The \emph{spherical softmax} in \cite{Brebisson16anexploration}.}\fyDone{Explain why}
 \begin{equation}
@@
 \end{equation}
-This parameterization avoids the ``rich-get-richer'' effect that we observe when using $\vlambda(\vpsi) = \operatorname{softmax}(\vpsi)$, which yields gradients wrt.
+This parameterization avoids the ``rich-get-richer'' effect that we observe with $\vlambda(\vpsi) = \operatorname{softmax}(\vpsi)$, which yields gradients wrt.
 $\psi[d]$ that are proportional to $\exp(\psi[d])$ (see also Figure~\ref{fig:sampling}).
 Additional settings for the hyper-parameters of our method include the number of simulation steps $k=10$ and the learning rate $\mathbf{lr}_{data}=0.001$.
-We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section~\ref{ssec:clda}), where we use 20~gradient descent iterations to avoid converging to degenerate distributions.\fyFuture{Why?
+We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (\textsection~\ref{ssec:clda}), where we use 20~gradient descent iterations to avoid converging to degenerate distributions.\fyFuture{Why?
 computation?} We split the training into 100 short sessions that last 2000 training steps each.
@@
 
-The computational cost of our approach is due to the simulation step, which is conducted after every 2000 iterations to compute the reward of each domain.
-During the simulation step, we update the temporary checkpoint with $k$ updates for each domain, which cost as much as $k$ training updates.
-Therefore, we execute $k \times n_d$ updates in total after every 2000 iterations.
-Our algorithm approximately costs $1 + \frac{k \times n_d}{2000}$ times as much as a standard training.
+The computational cost of our approach is due to the simulation step, which is conducted after every 2,000 iterations to compute the reward of each domain (eq.~\eqref{eq:reward}).
+During this step, we update the temporary checkpoint with $k$ updates for each domain, which costs as much as $k$ training updates.
+Therefore, we execute $k \times n_d$ updates after every 2,000 iterations.
+Our algorithm approximately costs $1+\frac{k \times n_d}{2000}$ times as much as a standard training.
 
 \subsection{Experimental tasks}
-We evaluate and compare our method with multiple baselines in the 5 following conditions.
+We evaluate our method in the 5 following conditions.
 %including: domain adaptation; multi-domain adaptation, bi-domain adaptation, unseen domain adaptation, and unsupervised domain adaptation.
-In the \emph{supervised domain adaptation task}, given the data from 6 domains (\domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}), we aim to build distinct expert NMT models for each domain.
-To challenge the flexibility of the method, we also consider a \emph{bi-domain adaptation task}, where given the same 6 domains, we only focus on adapting to a mixture of 2~domains.
+In the \emph{supervised domain adaptation task}, given the data from 6 domains (\domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}), we aim to build expert NMT models for each domain.
+To challenge the flexibility of the method, we also consider a \emph{two-domain adaptation task}, where given the same 6~domains, we focus on adapting to a mixture of 2~domains.
 
-In the \emph{multi-domain adaptation task}, given the same 6 domains, we aim to build one single NMT model that would perform optimally, assuming a uniform distribution of domains during the test.
+In the \emph{multidomain adaptation task}, we use the same 6 domains to build one single NMT model that should perform optimally, assuming a uniform distribution of domains during the test.
 
-In a fourth experiment (\emph{unsupervised domain adaptation}), given training data in 6 domains and a small development set of a new domain (\domain{News} in our case), we aim to build an NMT model which performs well for the unseen domain.
+In a fourth experiment (\emph{unseen domain adaptation}), given training data in 6 domains and a small dev set of a new domain (\domain{News}), we build a model which performs well for the unseen domain.
 

### 2022-03-29 2022/eamt/eamt-latex/eamt22.tex
full review

@@
 \usepackage{adjustbox}
-\renewcommand{\floatpagefraction}{.9}
 
@@
 % \cite{Vanderwees17dynamic,Graves17automated,platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Wang20balancing,Wang20learning-multi}
-explore ways to instead consider \emph{dynamic} data selection and sampling strategies: \newcite{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while \newcite{Graves17automated,Platanios19competence,Kumar19reinforcement,Wang20learning-multi,Wang20balancing} build curricula that automatically adapt to the training data.
+explore ways to instead consider \emph{dynamic} data selection and sampling strategies: \newcite{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while \newcite{Wang20learning-multi,Wang20balancing} build curricula that automatically adapt to the training data.
 In this paper, we contribute to this line of research in several ways.
@@
 \subsection{Data and metrics \label{ssec:corpora}}
-We experiment with translation from English into French in 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.
-We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \cite{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
-Additional experiments use the News Commentary corpus (\domain{news}).
-Most corpora are available from the Opus website.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
-To reduce the number of types, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
-%
-We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and \texttt{sacrebleu} \cite{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
-We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
-
 \begin{table*}[htbp]
-  \centering
+  \vspace{-\baselineskip}
+  \centering \footnotesize
   \begin{tabular}{|l|ccccccc|} %*{4}{|r|}}
@@
   \end{tabular}
-  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the training domain mixture (which does not contain \domain{news}), number English and French tokens ($\times 10^6$), number English and French types ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$).
-\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
+  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the training domain mixture (exluding \domain{news}), number English and French tokens ($\times 10^6$), types and uniq types ($\times 10^3$): the latter are types that only appear in a given domain.
+    % \domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
   }
+  \vspace{-\baselineskip}
 \label{tab:Corpora}
 \end{table*}
+
+We experiment with translation from English into French in 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.
+We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \cite{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
+Additional experiments use the News Commentary corpus (\domain{news}).
+Most corpora are available from the Opus website.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
+To reduce the number of types, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
+%
+We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
+Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and \texttt{sacrebleu} \cite{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
 
@@
 To challenge the flexibility of the method, we also consider a \emph{two-domain adaptation task}, where given the same 6~domains, we focus on adapting to a mixture of 2~domains.
-
+%
 In the \emph{multidomain adaptation task}, we use the same 6 domains to build one single NMT model that should perform optimally, assuming a uniform distribution of domains during the test.
-
-In a fourth experiment (\emph{unseen domain adaptation}), given training data in 6 domains and a small dev set of a new domain (\domain{News}), we build a model which performs well for the unseen domain.
-
-Finally, in the \emph{unsupervised domain adaptation task}, we cluster all available training data into 30~clusters using the KNN algorithm in the same way as in \cite{Tars18multidomain}, then adapt these clusters to one of 6 domains using the corresponding in-domain dev set.
+%
+A fourth experiment (\emph{unseen domain adaptation}), adds to the training data for 6~domains a small dev set in a new domain (\domain{News}): our target is a model which performs well for the unseen domain.
+%
+Finally, in the \emph{unsupervised domain adaptation task}, we cluster all available training data into 30~clusters using the KNN algorithm as in \cite{Tars18multidomain}, then learn mixture weights these clusters to one of 6 domains using the corresponding dev set.
 We compare MDAC to DDS for each of our 6 test sets.
@@
 \section{Results and discussion \label{sec:results}}
-
-\subsection{Domain Adaptation}\label{ssec:da}
-In this setting, we aim to build an NMT model for one single domain: we accordingly set $\vlambda^t$ to a deterministic distribution $\vlambda_d$, where the target domain $d$ has probability~1.
-
-We consider three initializations for MDAC and DDS, using $\vlambda_0$, $\vlambda_1$ and $\vlambda_d$.
-According to Table~\ref{tab:da}, MDAC achieves the overall best performance when $\vlambda_{t=0} = \vlambda_0$.
-Doing so proves much better than initializing with $\vlambda_d$ for small domains: \domain{talk}, \domain{bank} and\domain{it}.
-Conversely, initializing with $\vlambda_d$ is beneficial when targeting large domains such as \domain{med} and \domain{law}.
-The same conclusion holds for DDS.
-
-We now compare the best MDAC system (using $\vlambda_{t=0} = \vlambda_0$) to full fine-tuning.
-According to Table~\ref{tab:da}, fine-tuning is better for large domains such as \domain{med} and \domain{law}, while MDAC outperforms fine-tuning by approximately 1.2 BLEU for \domain{bank} and 1.0 BLEU for \domain{rel}.
-This indicates that for small domains, out-of-domain data helps improve the generalization and that MDAC is able to exploit both the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
-Results for DDS display similar trends but are always outperformed by MDAC.
-The behavior of CL, which does only well the large domain \domain{med} lag somewhat behind.
 
@@
   \end{tabular}
-  \caption{Domain adaptation experiments.
+  \caption{Single domain adaptation.
 We report BLEU scores of each method for 6 target domains and their average: each column corresponds to a distinct system.
@@
 ($^{**}$) MDAC is significantly better than CL and DDS with $p<0.05$.
-($^{***}$) MDAC is significantly better than CL, fine-tuning with $p<0.05$.}
+($^{***}$) MDAC is significantly better than CL, fine-tuning with $p<0.05$.} \vspace{-\baselineskip}
   \label{tab:da}
 \end{table*}
+\subsection{Domain Adaptation}\label{ssec:da}
+In this setting, we aim to build an NMT model for one single domain: we accordingly set $\vlambda^t$ to a deterministic distribution $\vlambda_d$, where the target domain $d$ has probability~1.
 
-\subsection{Bi-domain adaptation}\label{ssec:bida}
+We consider three initializations for MDAC and DDS, using $\vlambda_0$, $\vlambda_1$ and $\vlambda_d$.
+According to Table~\ref{tab:da}, MDAC achieves the overall best performance when $\vlambda_{t=0} = \vlambda_0$.
+Doing so proves much better than initializing with $\vlambda_d$ for small domains: \domain{talk}, \domain{bank} and\domain{it}.
+Conversely, initializing with $\vlambda_d$ is beneficial when targeting large domains such as \domain{med} and \domain{law}.
+The same conclusion holds for DDS.
 
-In these control experiments, we showcase the flexibility of dynamic sampling and try to adapt to (arbitrary) pairs of target domains with equal weight and contrast our results with those of DDS.
-Results are in Table~\ref{tab:bi-da}.
-Here, MDAC significantly outperforms DDS in two settings (\domain{med+it} and \domain{law+bank}) while being surpassed in \domain{talk+rel}.
+We now compare the best MDAC system (using $\vlambda_{t=0} = \vlambda_0$) to full fine-tuning.
+According to Table~\ref{tab:da}, fine-tuning is better for large domains such as \domain{med} and \domain{law}, while MDAC outperforms fine-tuning by approximately 1.2 BLEU for \domain{bank} and 1.0 BLEU for \domain{rel}.
+This suggests that for small domains, out-of-domain data helps improve the generalization and that MDAC is able to exploit both the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.
+Results for DDS display similar trends but are always outperformed by MDAC.
+Results for CL, which does only well the large domain \domain{med}, lag somewhat behind.
+
+\subsection{Two-domain adaptation}\label{ssec:bida}
+
+In these control experiments, we showcase the flexibility of dynamic sampling and adapt to (arbitrary) pairs of target domains with equal weight, contrasting MDAC with DDS in Table~\ref{tab:bi-da}.
+Here, MDAC significantly outperforms DDS in two settings (\domain{med+it} and \domain{law+bank}) out of three.
+%while being surpassed in \domain{talk+rel}.
 % Allthough we showing that MDAC is capable to adapt to a variable set of target domains.
@@
 
-We now turn to a more realistic scenario and consider multi-domain NMT, which aims to train one single system with optimal performance averaged over 6 domains.
-This setting targets a uniform test distribution $\vlambda^t = \vlambda_0$.
-In this situation, CL \cite{Zhang19curriculum} does not apply.
-We therefore only contrast the performance of MDAC, DDS and several fixed training data distribution $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$, where $\vlambda_{\alpha}$ is defined according to equation~\eqref{mixture:trn}.
+We now turn to a more realistic scenario and consider multidomain NMT, which aims to train one single system with optimal performance averaged over 6 domains and targets a uniform test distribution $\vlambda^t = \vlambda_0$.
+In this situation, CL \cite{Zhang19curriculum} does not apply: we only contrast the performance of MDAC, DDS and several fixed training data distribution $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$, where $\vlambda_{\alpha}$ is defined according to equation~\eqref{mixture:trn}.
 
@@
 For this configuration, MDAC outperforms static training distributions including $\big[ \vlambda_0, \vlambda_{0.75}, \vlambda_{1.0}\big]$ by a significant margin, and performs slightly better than $\big[ \vlambda_{0.25}, \vlambda_{0.5} \big]$.
-This indicates that MDAC allows us to skip the somewhat heuristic choice of the optimal training mixture.
+Using MDAC thus dispenses with the empirical search of an optimal training mixture.
 
 A second observation is that MDAC again outperforms DDS by a wide margin (+1.5 BLEU on average); the only domain where DDS does better is \domain{med}.
-Figure~\ref{fig:sampling}, which displays the evolution of the training mixture in the course of training, helps to understand the difference between the two methods.
-For DDS (Figure~\ref{fig:DDS}), the sampling distribution quickly reaches a bi-domain mode in which only \domain{med} and \domain{rel} have significant probability -- hence the good performance on the former domain.
-In contrast, the distribution computed by MDAC evolves more smoothly; small domains such as \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} receive a larger proportion of training data in the early stages; their weights then slowly decrease as larger domains such as \domain{med} and \domain{law} increase their share.
-This, however, only happens at the end of training, when some NMT models might already be close to their best performance for the small domains.
+Figure~\ref{fig:sampling}, which plots the evolution of the mixture weights during training, helps to understand the difference between the two methods.
+For DDS (Figure~\ref{fig:DDS}), the sampling distribution quickly reaches a bi-modal regime in which only \domain{med} and \domain{rel} have significant probability -- hence the good performance on the former domain.
+In contrast, the distribution computed by MDAC evolves more smoothly; small domains such as \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} receive a larger part of training data in the early stages; their weights then slowly decrease as larger domains such as \domain{med} and \domain{law} increase their share.
+This only happens at the end of training, when some NMT models might already be close to their peak performance for the small domains.
 
 \begin{table*}[htbp]
+  \vspace{\baselineskip}
   \centering \small
@@
 \begin{table*}[htbp]
+  \vspace{-2\baselineskip}
   \centering \small
@@
   \end{tabular}
-  \caption{Multi domain adaptation.
+  \caption{Multidomain adaptation.
 For a given line, all the columns \emph{correspond to the same multi-domain system}.
@@
 \begin{figure*}[htbp]
+  \vspace{-2\baselineskip}
   \begin{subfigure}{.5\textwidth}
@@
 \subsection{Unseen domain}\label{ssec:uda}
-The left part of Table~\ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$ and with dynamic data selection (MDAC and DDS).
-These systems have insignificant differences in BLEU, showing that dynamic mixture does not improve the robustness of the NMT system against unseen domains.
-However, the fact that the performance of MDAC and DDS is close to the best performance suggests that they also apply in such challenging situations.
+The left part of Table~\ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for systems trained with mixtures $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$ and with dynamic data selection (MDAC and DDS).
+These systems have insignificant differences in BLEU, suggesting that dynamic mixtures do not improve the robustness of NMT systems against unseen domains.
+However, the performance of MDAC and DDS remains close to the best performance, showing that they also apply in such settings.
 % the robustness of the NMT model.
@@
 \subsection{Automatic clustering}\label{ssec:clda}
-The right part of Table~\ref{tab:unsupervised-da} reports the performance of NMT systems adapted to each domain.\footnote{A more detailed analysis is in the supplementary material.} In comparison to Section~\ref{ssec:da}, the training data is distributed in 30 automatic clusters instead of the 6 original domains.
-Splitting the train data into small groups provides the learner with extra degrees of freedom when selecting the best distribution.
+The right part of Table~\ref{tab:unsupervised-da} reports the performance of NMT systems adapted to each domain.
+% \footnote{A more detailed analysis is in the supplementary material.}
+In comparison to Section~\ref{ssec:da}, the training data is distributed in 30 automatic clusters instead of the 6 original domains.
+Splitting the train data into small groups gives the learner extra degrees of freedom when selecting the best distribution.
 However, as these clusters are built automatically, they are noisier in nature.
-According to the results in Table~\ref{tab:unsupervised-da}, this scenario is hard both for DDS and MDAC, which performs much worse than for the supervised DA setting.
+According to results in Table~\ref{tab:unsupervised-da}, this scenario is hard both for DDS and MDAC, which performs much worse than for the supervised DA setting.
 This again signals the importance of initialization: analyzing the clustering, we find that the data for \domain{rel} mostly correspond to one single cluster.
-When using a uniform initialization, this cluster starts with a very small weight and never succeeds in matching the (good) performance observed in the DA setting.
+With a uniform initialization, this cluster starts with a small weight and never succeeds in matching the good performance observed in the DA setting.
 
@@
 Domain adaptation is an old problem that has been studied from many angles, both for SMT and NMT.
-A survey of supervised and unsupervised DA for NMT is in \newcite{Chu18asurvey}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the more recent survey of \newcite{Saunders21asurvey}.
+A survey of supervised and unsupervised DA for NMT is in \cite{Chu18asurvey}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the recent survey of \newcite{Saunders21asurvey}.
 Our approach to DA in this paper falls under the former category.
-We refer interested readers to these references and do not discuss DA any further.
+We refer readers interested in DA to these papers.
 
-Multi-domain NMT (MDMT) aims to develop systems that simultaneously bode well for several domains.
+Multidomain NMT (MDMT) aims to develop systems that simultaneously bode well for several domains.
 Like for DA, techniques for supervised MDMT combine one or several ingredients: (a) the specialization of data representations \cite{Kobus17domaincontrol} or of sub-networks \cite{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \newcite{Jiang19multidomain}.
@@
 Various ways to automate CL using the framework of multi-armed bandits are explored in \cite{Graves17automated}, which has been an inspiration for our implementation.
-While the initial aim was primarily to improve and speed up training, CL has also proven useful for domain adaptive / multi-domain / multilingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
-For instance, \newcite{Zhang19curriculum} study supervised DA and propose a curriculum approach which progressively augments the training data: in the early stages, only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
-This is somehow opposite to the recommendations of \newcite{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyDone{These have not been compared?
+While the initial aim was primarily to improve and speed up training, CL has also proven useful for domain adaptive / multidomain / multilingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
+For instance, \newcite{Zhang19curriculum} study supervised DA and propose a curriculum approach which progressively augments the training data: early stages only use in-data, while less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
+This is opposite to the policy of \newcite{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively increases the focus on in-domain data.\fyDone{These have not been compared?
 and also to what we do ?}
@@
 \newcite{Kumar19reinforcement} use reinforcement techniques (deep Q-learning) to learn the curriculum strategy: in this work, complexity corresponds to difficulty levels which are binned using contrastive data selection.
-The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyDone{Alert: what do we do during warm-up ?} The same technique is recently applied to multilingual NMT in \cite{Kumar21learning}.
-\newcite{Zhou20uncertainty} propose another curriculum-based approach which instead relies on \emph{instance uncertainty} as a measure of their difficulty and presents the data sample starting with the easiest (more predictable) first.
+The reward is based on the increase of the development set loss that results from the current data selection strategy.\fyDone{Alert: what do we do during warm-up ?} This technique is applied to multilingual NMT in \cite{Kumar21learning}.
+\newcite{Zhou20uncertainty} propose another curriculum-based approach which relies on \emph{instance uncertainty} as a measure of their difficulty and presents data samples starting with the easiest (more predictable) first.
 Another contribution of this paper is an alternative stopping criterium.
-Closest to our problems, \newcite{Wang20learning-multi} adapt CL for multi-domain NMT, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
+Closest to our problems, \newcite{Wang20learning-multi} adapt CL for multidomain NMT, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
 Each step consists of (a) weighting instances based on relevance features, (b) fine-tuning a pretrained model using the weighted training set, and is applied iteratively to train a sequence of models.
-The one that maximizes the development set performance is finally retained.
+The one that maximizes the dev set performance is finally retained.
 
@@
 \fyDone{Write conclusion}
-In this study, we have presented a generic framework to perform a variety of adaptation tasks for machine translation, ranging from conventional supervised domain adaptation to multi-domain NMT and unseen domain adaptation.
-By experimenting with all these settings, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during the course of training, could be used in all these situations.
+In this study, we have presented a generic framework to perform a variety of adaptation tasks for machine translation, ranging from conventional supervised domain adaptation to multidomain NMT and unseen domain adaptation.
+By experimenting with all these settings, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during the course of training, can be used in all these situations.
 This algorithm, we believe, provides us with a more sound approach to (multi-domain) DA than existing heuristics and dispenses with the costly search of optimal meta-parameters.
@@
 
-In the future, we intend to continue developing this approach and improve its effectiveness.
+Our future work will continue developing this approach and improve its effectiveness.
 One issue that we have left unaddressed is reward normalization, which is especially important in the early stages of training \cite{Kumar19reinforcement}.
-Another area where we need to progress is the unsupervised learning setting of Section~\ref{ssec:clda}, where our results still lag way behind that of supervised DA - this might be due to the inability of our simplistic optimization strategy to handle situations where the number of clusters is too large.
+Another area where we need to progress is the unsupervised learning setting of \textsection~\ref{ssec:clda}, where our results lag behind supervised DA.
+This might be due to the inability of our simplistic optimization strategy to handle situations where the number of clusters is large.
 %\clearpage{}

### 2022-03-31 2022/eamt/eamt-latex/eamt22.tex
10 pages - sharp

@@
 
-We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc.
+We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
 Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
@@
 We experiment with translation from English into French in 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}.
-We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \cite{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
+We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}); the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}).
 Additional experiments use the News Commentary corpus (\domain{news}).
-Most corpora are available from the Opus website.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
+Most corpora are available from the Opus website \cite{Tiedeman12parallel}.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
 To reduce the number of types, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%
@@
 Domain adaptation is an old problem that has been studied from many angles, both for SMT and NMT.
-A survey of supervised and unsupervised DA for NMT is in \cite{Chu18asurvey}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the recent survey of \newcite{Saunders21asurvey}.
+A survey of supervised and unsupervised DA for NMT is in \cite{Chu2017comparison}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the recent survey of \newcite{Saunders21asurvey}.
 Our approach to DA in this paper falls under the former category.

### 2022-04-13 2022/acl/acl_latex.tex
started the final version

@@
 %\usepackage[review,nohyperref]{acl}
-\usepackage[review]{acl}
+\usepackage{acl}
 
@@
 
-\author{First Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{email@domain} \\\And
-  Second Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{email@domain} \\}
+\author{
+  MinhQuang Pham$^{\dag\ddag}$, \ \ Josep Crego$^\dag$,\ \ Fran\c cois Yvon$^\ddag$ \\
+  $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France\\
+  {\tt firstname.lastname@systrangroup.com}\\
+  $^\ddag$Universit\'e Paris-Saclay, CNRS, LISN,  91405 Orsay, France\\
+  {\tt firstname.lastname@limsi.fr}
+}
 
@@
 \item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048; the multi-head attention mechanism contains 8 heads; 6 layers in the encoder; 6 layers in the decoder.\fyDone{How many layers}
-\item Adapter-based Transformer: The intermediate feedforward dimension is set to 2048;% as in \citet{Pham21revisiting}.
+\item Adapter-based Transformer: the intermediate feedforward dimension is set to 2048, as in \citet{Pham21revisiting}.
 \item Transformer using Latent multi-task group dropout (\system{LaMGD} Transformers):\fyDone{Explain acronym} There is no change in the architecture.
@@
 \item Multilingual Transformer: the embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 1024, each multi-head attentions contains 8~heads as in \citep{Wang20balancing}.
-\item Adapter based Transformer: the intermediate feedforward dimension is set as 128 as in \citep{Gong21adaptive}.
+\item Adapter based Transformer: the intermediate feedforward dimension is set as 128. We follow here the parameter setting of \citep{Gong21adaptive}.
 \item \system{LaMGD} Transformer: There is no change in the architecture. We group 512 nodes in each layer into 16 groups of 32 consecutive nodes. For each language, we select 12 groups.
@@
     \system{Transformer}  \hfill{\footnotesize[65m]} & 40.3 & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
-    \system{HMGD} Transformer   \hfill{\footnotesize[+0m]}  & 40.4 & 60.4 & 51.9 & 38.7 &	50.80 &	86.80 & 54.8 \\ 
-    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.3 \\ 
-    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  & 40.3 & \SB{60.4} & \SB{52.4} & \SB{39.0} & \SB{52.4} & \SB{87.5} & \SB{55.3} \\ 
+    \system{HMGD} Transformer   \hfill{\footnotesize[+0m]}  & \SB{40.4} & 60.4 & 51.9 & 38.7 &	50.8 &	86.80 & 54.8 \\ 
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & \SB{61.0} & \SB{53.1} & 37.5 & 49.6 & \SB{91.0} & \SB{55.3} \\ 
+    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & \SB{39.0} & \SB{52.4} & 87.5 & \SB{55.3} \\ 
     \hline
   \end{tabular}
-  \caption{Multidomain translation experiment. Boldface denotes significant gains over \system{Transformer} ($p=0.05$)}
+  \caption{Multi-domain translation. Boldface identifies best system for each domain.}
   \label{tab:mdmt}
@@
   \begin{tabular}{|p{4cm}|*{9}{r|}} \hline
-    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
+    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
     \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.8 &7.3&20.8&21.1&39.7&19.8&22.6&15.2&18.9 \\
@@
     \hline
-    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
+    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
     \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.4&16.6&28.5&	27.1&43.7&24.6&30.3&25.6&25.98 \\
@@
     O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&34.0&38.3&8.3&20.19 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &10.2&4&13.3&21.9&32.2&34.1&38.5&8.3&20.31 \\ 
-    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
+    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&34.0&38.3&8.3&20.2 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &10.2&4&\SB{13.3}&21.9&32.2&34.1&38.5&8.3&20.3 \\ 
+    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.1\\
     \hline 
@@
     M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19.0&27.43 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &22.5&9.4&20.0&30.6&37.2&38.2&39.3&19.0&27.03\\ 
-    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40.0&18.9&27.75 \\
+    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19.0&27.4 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &22.5&9.4&20.0&30.6&37.2&38.2&39.3&19.0&27.0\\ 
+    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40.0&18.9&27.7 \\
     \hline
@@
 For multilingual translation, our model outperforms multilingual Transformer and Language Adapters in 3 our of 4 settings.
+LAMGD seems specially beneficial for training languages with little parallel data, which can take advantage of the resources that are available for related languages.
 Besides, we provided an thorough analysis of the similarities between learned sub-networks and demonstrate a strong correlation between the learned similarities and the proximity of the corresponding tasks (domain or language).
@@
 Second, we also intend to dispense with the domain information during inference: this would mean replacing the dependency on $d$ in the variational distribution by a dependency on the input $x$.
-Adressing these two questions will allow to us replace heuristic choices in the architecture design with an increased dependency on the training data.
+Another interesting direction will be to consider adapting the size and capacity allocated to each domain / language, depending on the difficulty of the associated translation task.
+Addressing these questions will allow to us replace heuristic choices in the architecture design with an increased dependency on the training data.
+
+\section{Acknowledgements}
+This work was granted access to the Jean Zay HPC resources of [TGCC/CINES/IDRIS] under the allocation 2020-[AD011011270] made by GENCI (Grand Equipement National de Calcul Intensif).
+
+\section*{Ethic Statement}
 

### 2022-04-13 2022/acl/acl_latex.tex
Added Ethical statement

@@
 
+\section{Ethical Considerations}
+MT technologies are generally intended to facilitate cross-lingual as well as cross-cultural communications.
+The methods presented here are notably interesting in the view to improve MT from and into English for low-resource languages, subject to the availability of data for a related language.
+We acknowledge that (a) our results should ultimately be backep-up large scale experiments involving much more languages -- even though this goes against the idea of limitating the computing cost our experiments; (b) better architectures and training regimes can improve the translation quality for low-resource languages, yet will not solve the problem entirely.
+This means that additional work focusing speciifically on developing resources for these languages should remain an important objective for future work.
+
 \section{Acknowledgements}
 This work was granted access to the Jean Zay HPC resources of [TGCC/CINES/IDRIS] under the allocation 2020-[AD011011270] made by GENCI (Grand Equipement National de Calcul Intensif).
-
-\section*{Ethic Statement}
 

### 2022-04-22 2022/acl/acl_latex.tex
local

--- 2022-04-13 2022/acl/acl_latex.tex
@@
     \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &4.3&6.8&21.1&22&39.7&20&23&15.2&19 \\ 
-    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.33 \\ 
+    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.3 \\ 
 	\hline
     \hline
-    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
-    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.4&16.6&28.5&	27.1&43.7&24.6&30.3&25.6&25.98 \\
-    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.64\\ 
-    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.31 \\ 
+    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline
+    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.4&16.6&28.5&	27.1&43.7&24.6&30.3&25.6&26.0 \\
+    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.6\\ 
+    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.3 \\ 
     \hline
@@
 \end{table}
-
+In addition, we also report in this section the effect of not choosing the number of groups $n_p$, which is assigned to 16 in the comparison of \system{LaMGD} and the contrasting methods.
+We show that setting $n_p$ to the layer's size, which means the group size is 1, has a very similar performance as choosing $n_p$ heuristically.
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{|c|c|c|} \hline
+    $k$/$n_p$ & \domain{avg} \\ \hline 
+    12 / 16 & 19.33\\
+    384 / 512 & 19.26\\
+    \hline
+  \end{tabular}
+  \caption{Setting the size of group to 1 (o2m-related experiment). The quota of activated nodes is keep unchanged to $75$\%}
+  \label{tab:K}
+\end{table}
 \section{Related work}\label{sec:related}
@@
 
+\section{Ethical Considerations}
+MT technologies are generally intended to facilitate cross-lingual as well as cross-cultural communications.
+The methods presented here are notably interesting in the view to improve MT from and into English for low-resource languages, subject to the availability of data for a related language.
+We acknowledge that (a) our results should ultimately be backed-up large scale experiments involving much more languages -- even though this goes against the idea of limiting the computing cost our experiments; (b) better architectures and training regimes can improve the translation quality for low-resource languages, yet will not solve the problem entirely.
+This means that additional work focusing specifically on developing resources for these languages should remain an important objective for future work.
+
 \section{Acknowledgements}
 This work was granted access to the Jean Zay HPC resources of [TGCC/CINES/IDRIS] under the allocation 2020-[AD011011270] made by GENCI (Grand Equipement National de Calcul Intensif).
-
-\section*{Ethic Statement}
 

### 2022-05-09 2022/eamt/eamt-latex/eamt22.tex
local changes
correct authors, citations

@@
 
-\author{First Author\\
-  Affiliation / Address line 1\\
-  Affiliation / Address line 2\\
-%  Affiliation / Address line 3\\
-  {\tt email@domain}  \And
-  Second Author\\
-  Affiliation / Address line 1\\
-  Affiliation / Address line 2\\
-%  Affiliation / Address line 3\\
-  {\tt email@domain}}
+%\author{First Author\\
+%  Affiliation / Address line 1\\
+%  Affiliation / Address line 2\\
+%%  Affiliation / Address line 3\\
+%  {\tt email@domain}  \And
+%  Second Author\\
+%  Affiliation / Address line 1\\
+%  Affiliation / Address line 2\\
+%%  Affiliation / Address line 3\\
+%  {\tt email@domain}}
+
+\author{Minh-Quang Pham\\
+  Universit\'e Paris-Saclay, \\
+  CNRS, LISN,  91405 Orsay, France \\
+  {\tt minhquang.pham@limsi.fr}  \\ \\
+  \textbf{Josep Crego}\\
+  SYSTRAN ,\\
+  5 rue Feydeau, 75002 Paris, France\\
+  {\tt josep.crego@systrangroup.com} \And
+  Fran\c cois Yvon\\
+  Universit\'e Paris-Saclay, \\
+  CNRS, LISN,  91405 Orsay, France\\
+  {\tt francois.yvon@limsi.fr}}
 
@@
 \usepackage{adjustbox}
-
+\usepackage[hidelinks]{hyperref}
 \makeatletter
@@
 In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key aspect of training is then to mitigate the detrimental effects of a mismatch between these distributions.
-Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from; this setting differs from multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \newcite{Chu2017comparison,Saunders21asurvey} for a review), and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from; this setting differs from multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \newcite{Chu2017comparison}; \newcite{Saunders21asurvey} for a review), and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
 A related situation is multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Aharoni19massively},\fyDone{Fix ref}
@@
 % \cite{Vanderwees17dynamic,Graves17automated,platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Wang20balancing,Wang20learning-multi}
-explore ways to instead consider \emph{dynamic} data selection and sampling strategies: \newcite{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while \newcite{Wang20learning-multi,Wang20balancing} build curricula that automatically adapt to the training data.
+explore ways to instead consider \emph{dynamic} data selection and sampling strategies: \newcite{Vanderwees17dynamic}; \newcite{Zhang19curriculum} construct a static curriculum, while \newcite{Wang20learning-multi}; \newcite{Wang20balancing} build curricula that automatically adapt to the training data.
 In this paper, we contribute to this line of research in several ways.
@@
 We consider three domains and represent $\vlambda^{s}$ and $\vlambda^{t}$ in the 3-dimensional simplex.
-Training with weights in (a) and testing with weights in (c) is supervised multi-source domain adaptation to domain~2 ($d_2$), while (b)-(c) is the unsupervised\fyDone{Unseen ?} version, with no training data from $d_2$; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with settings (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
+Training with weights in (a) and testing with weights in (c) is supervised multi-source domain adaptation to domain~2 ($d_2$), while (b)-(c) is the unsupervised version, with no training data from $d_2$; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with settings (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
 \end{figure}
 
-These situations have been amply documented from a theoretical perspective (eg.\ \newcite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}).
+These situations have been amply documented from a theoretical perspective (eg.\ \newcite{Mansour09multiple}; \newcite{Mansour09domainadaptation}; \newcite{Hoffman18algorithms}).
 A general recommendation in the DA setting is to adjust the sampling distribution used to optimize the system so as to compensate for the mismatch between $\mathcal{D}^s(x)$ and $\mathcal{D}^t(x)$.
@@
 After one learning session, we update the data distribution using the REINFORCE algorithm of \newcite{Williams92simple}.
-The evolution of $\psi$ is thus defined by:\fyTodo{$t$ is both for time, and for test - change into $\tau$?}
+The evolution of $\psi$ is thus defined by:\fyDone{$t$ is both for time, and for test - change into $\tau$?}
 \begin{align*}
@@
 This is achieved by simulating a round of training with only the data from one domain.
-We also differ from this authors in the parameterization of the sampling distribution.
+We also differ from these authors in the parameterization of the sampling distribution.
 % Furthermore, the choice of the parameterization of the sampling distribution in our method differs from that of \newcite{Graves17automated}, although this choice is empirical.
@@
 Additional experiments use the News Commentary corpus (\domain{news}).
-Most corpora are available from the Opus website \cite{Tiedeman12parallel}.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
+Most corpora are available from the Opus website\footnote{\url{http://opus.nlpl.eu}}.
+These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}.
 To reduce the number of types, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%

### 2022-05-09 2022/eamt/eamt-latex/eamt22.tex
try to fix style

@@
 \author{Minh-Quang Pham\\
-  Universit\'e Paris-Saclay, \\
-  CNRS, LISN,  91405 Orsay, France \\
-  {\tt minhquang.pham@limsi.fr}  \\ \\
+  Uni. Paris-Saclay,  CNRS, LISN \\
+  F-91405 Orsay \\
+  {\tt pham@limsi.fr}  \And
   \textbf{Josep Crego}\\
   SYSTRAN ,\\
-  5 rue Feydeau, 75002 Paris, France\\
-  {\tt josep.crego@systrangroup.com} \And
+  5 rue Feydeau, F-75002 Paris\\
+  {\tt crego@systrangroup.com} \And
   Fran\c cois Yvon\\
-  Universit\'e Paris-Saclay, \\
-  CNRS, LISN,  91405 Orsay, France\\
-  {\tt francois.yvon@limsi.fr}}
+  Uni. Paris-Saclay, CNRS, LISN\\
+  F-91405 Orsay, France\\
+  {\tt yvon@limsi.fr}}
 
@@
 In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key aspect of training is then to mitigate the detrimental effects of a mismatch between these distributions.
-Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from; this setting differs from multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \newcite{Chu2017comparison}; \newcite{Saunders21asurvey} for a review), and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from; this setting differs from multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \cite{Chu2017comparison,Saunders21asurvey} for a review), and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
 A related situation is multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Aharoni19massively},\fyDone{Fix ref}
@@
 % \cite{Vanderwees17dynamic,Graves17automated,platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Wang20balancing,Wang20learning-multi}
-explore ways to instead consider \emph{dynamic} data selection and sampling strategies: \newcite{Vanderwees17dynamic}; \newcite{Zhang19curriculum} construct a static curriculum, while \newcite{Wang20learning-multi}; \newcite{Wang20balancing} build curricula that automatically adapt to the training data.
+explore ways to instead consider \emph{dynamic} data selection and sampling strategies: \newcite{Vanderwees17dynamic} and \newcite{Zhang19curriculum} construct a static curriculum, while \newcite{Wang20learning-multi} and \newcite{Wang20balancing} build curricula that automatically adapt to the training data.
 In this paper, we contribute to this line of research in several ways.
@@
 
-These situations have been amply documented from a theoretical perspective (eg.\ \newcite{Mansour09multiple}; \newcite{Mansour09domainadaptation}; \newcite{Hoffman18algorithms}).
+These situations have been amply documented from a theoretical perspective (eg.\ \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}).
 A general recommendation in the DA setting is to adjust the sampling distribution used to optimize the system so as to compensate for the mismatch between $\mathcal{D}^s(x)$ and $\mathcal{D}^t(x)$.
@@
 Multidomain NMT (MDMT) aims to develop systems that simultaneously bode well for several domains.
-Like for DA, techniques for supervised MDMT combine one or several ingredients: (a) the specialization of data representations \cite{Kobus17domaincontrol} or of sub-networks \cite{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \newcite{Jiang19multidomain}.
+Like for DA, techniques for supervised MDMT combine one or several ingredients: (a) the specialization of data representations \cite{Kobus17domaincontrol} or of sub-networks \cite{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \cite{Jiang19multidomain}.
 Unsupervised MDMT is studied in \newcite{Farajian17multidomain}, as an instance of unsupervised DA.
@@
 Most approaches to adaptive/dynamic data selection take inspiration from \newcite{Bengio09curriculum}, where the notion of curriculum learning is introduced.
-CL relies on the notion of the ``easiness'' of a sample to schedule the presentation of training data so that the easiest examples are presented first and the hardest last.
-Various ways to automate CL using the framework of multi-armed bandits are explored in \cite{Graves17automated}, which has been an inspiration for our implementation.
-While the initial aim was primarily to improve and speed up training, CL has also proven useful for domain adaptive / multidomain / multilingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
+CL relies on the notion of the ``easiness'' of a sample to schedule the presentation of training data so as to start with the easiest examples and end with the hardest.
+Various ways to automate CL in the framework of multi-armed bandits are explored in \cite{Graves17automated}, which has been an inspiration for our implementation.
+While the initial aim was primarily to improve and speed up training, CL has also proven useful for multidomain and multilingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
 For instance, \newcite{Zhang19curriculum} study supervised DA and propose a curriculum approach which progressively augments the training data: early stages only use in-data, while less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages.
@@
 
-\newcite{Kumar19reinforcement} use reinforcement techniques (deep Q-learning) to learn the curriculum strategy: in this work, complexity corresponds to difficulty levels which are binned using contrastive data selection.
-The reward is based on the increase of the development set loss that results from the current data selection strategy.\fyDone{Alert: what do we do during warm-up ?} This technique is applied to multilingual NMT in \cite{Kumar21learning}.
-\newcite{Zhou20uncertainty} propose another curriculum-based approach which relies on \emph{instance uncertainty} as a measure of their difficulty and presents data samples starting with the easiest (more predictable) first.
-Another contribution of this paper is an alternative stopping criterium.
+\newcite{Kumar19reinforcement} use reinforcement learning to learn the curriculum strategy: in this work, complexity corresponds to difficulty levels which are binned using contrastive data selection.
+The reward is based on the increase of the devset loss that results from the current data selection strategy.\fyDone{Alert: what do we do during warm-up ?} This technique is applied to multilingual NMT in \cite{Kumar21learning}.
+\newcite{Zhou20uncertainty} propose another CL-based approach which relies on \emph{instance uncertainty} as a measure of their difficulty and presents data samples starting with the easiest.
+Another contribution of this work is a new stopping criterium.
 Closest to our problems, \newcite{Wang20learning-multi} adapt CL for multidomain NMT, where an optimal instance weighting scheme is found using Bayesian optimization techniques.
-Each step consists of (a) weighting instances based on relevance features, (b) fine-tuning a pretrained model using the weighted training set, and is applied iteratively to train a sequence of models.
-The one that maximizes the dev set performance is finally retained.
+Each step consists of (a) weighting instances based on relevance features, (b) fine-tuning a pretrained model using the weighted training set, and is applied to train a sequence of models.
+The one that maximizes the devset performance is finally retained.
 
@@
 \fyDone{Write conclusion}
-In this study, we have presented a generic framework to perform a variety of adaptation tasks for machine translation, ranging from conventional supervised domain adaptation to multidomain NMT and unseen domain adaptation.
-By experimenting with all these settings, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during the course of training, can be used in all these situations.
+In this study, we have presented a generic framework to perform multiple adaptation tasks for machine translation, ranging from supervised domain adaptation to multidomain NMT and unseen domain adaptation.
+In our experiments, we have shown that the same algorithm, aimed at automatically finding an effective data sampling scheme during the course of training, can be used in all these situations.
 This algorithm, we believe, provides us with a more sound approach to (multi-domain) DA than existing heuristics and dispenses with the costly search of optimal meta-parameters.
@@
 
-% \section*{Acknowledgments}
-%The work presented in this paper was partially sup-ported by the European Commission under contractH2020-787061 ANITA
 \bibliography{multidomain}
 \bibliographystyle{eamt22}
-%\clearpage{}
-%\appendix
+% \clearpage{}
+
+% \appendix
 %\input{appendix.tex}

### 2022-05-16 2022/eamt/eamt-latex/eamt22.tex
fixed length and ref issue

@@
 \usepackage[small,bf]{caption} % added MLF 20171211
-\setlength\titlebox{6.5cm}    % Expanding the titlebox
+\setlength\titlebox{6.8cm}    % Expanding the titlebox
 %%% YOUR PACKAGES BELOW THIS LINE %%%
@@
   {\tt pham@limsi.fr}  \And
-  \textbf{Josep Crego}\\
+  Josep Crego\\
   SYSTRAN ,\\
@@
 %\usepackage[noend]{algpseudocode}
-\usepackage{regexpatch}
-\usepackage{subcaption}
+% \usepackage{regexpatch}
+% \usepackage{subcaption}
 % This is not strictly necessary, and may be commented out,
@@
 
-\usepackage{url}
 \usepackage{longtable}
@@
 
-\let\oldbibitem\bibitem
-\def\bibitem{\vfill\oldbibitem}
+%\let\oldbibitem\bibitem
+% \def\bibitem{\vfill\oldbibitem}
 
@@
 \usepackage{tabularx}
-
-\usepackage{multirow}
 \usepackage{adjustbox}
-\usepackage{caption}
 \usepackage{subcaption}
+% \usepackage{hyperref}
 
@@
 \newcommand{\sbcl}[2]{{\scriptsize #1 \hfill $|$ \hfill  #2}}
-\usepackage{multirow}
-\usepackage{adjustbox}
-\usepackage[hidelinks]{hyperref}
+
 \makeatletter
@@
 
-These situations have been amply documented from a theoretical perspective (eg.\ \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}).
+These situations have been amply documented from a theoretical perspective \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}.
 A general recommendation in the DA setting is to adjust the sampling distribution used to optimize the system so as to compensate for the mismatch between $\mathcal{D}^s(x)$ and $\mathcal{D}^t(x)$.
@@
 Like for DA, techniques for supervised MDMT combine one or several ingredients: (a) the specialization of data representations \cite{Kobus17domaincontrol} or of sub-networks \cite{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \cite{Jiang19multidomain}.
-Unsupervised MDMT is studied in \newcite{Farajian17multidomain}, as an instance of unsupervised DA.
+Unsupervised MDMT is studied in \cite{Farajian17multidomain}, as an instance of unsupervised DA.
 
@@
 \bibliographystyle{eamt22}
+
 % \clearpage{}

### 2022-05-16 2022/acl/acl_latex.tex
fixed formatting errors

@@
 %graphic packages
+\usepackage[export]{adjustbox}
 \usepackage{graphicx}
@@
 We discuss here the choice of the hyper-parameters $k$, the number of activated nodes in each layer, and its impact on the sharing level between the tasks.
-Table ~\ref{tab:K} shows the variance of performance when the number of activated nodes is changed, and the sharing level between tasks decreases in consequence.
+Table~\ref{tab:K} reports the variance of performance when the number of activated nodes is changed and the sharing level between tasks decreases accordingly.
 \begin{table}[h!]
@@
 In addition, we also report in this section the effect of not choosing the number of groups $n_p$, which is assigned to 16 in the comparison of \system{LaMGD} and the contrasting methods.
-We show that setting $n_p$ to the layer's size, which means the group size is 1, has a very similar performance as choosing $n_p$ heuristically.
+We show that setting $n_p$ to the layer's size, which means the group size is 1, has a very similar performance as choosing $n_p$ heuristically (Table~\ref{tab:K2}).
 \begin{table}[h!]
@@
   \caption{Setting the size of group to 1 (o2m-related experiment). The quota of activated nodes is keep unchanged to $75$\%}
-  \label{tab:K}
+  \label{tab:K2}
 \end{table}
+
 \section{Related work}\label{sec:related}
@@
 \hspace{-3.em}
-\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} = \frac{\partial \hat{m}_l^d(\tau)}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d} + \frac{1}{\tau} \frac{exp(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau})}{(1+exp(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}))^2}
+  \frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} = & \frac{\partial \hat{m}_l^d(\tau)}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d} +  \nonumber \\
+  & \quad \frac{1}{\tau} \frac{exp(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau})}{(1+exp(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}))^2}
 \end{align}
@@
 
-Finally, we need prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
+Finally, we need to prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$.
 We assume $g_l^d(i_1) + \Phi_l^d(i_1) > g_l^d(i_2) + \Phi_l^d(i_2) > \cdots > g_l^d(i_{n_p}) + \Phi_l^d(i_{n_p})$.
@@
 \begin{align*}
-\hspace{-3.em}
-\lim_{\tau \rightarrow 0} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = \begin{cases}
+% \hspace{-3.em}
+  \lim_{\tau \rightarrow 0} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) & = \\
+  \quad \begin{cases}
     1, & \text{if}\ \tau > - (g_l^d(i) + \Phi_l^d(i)), \\
@@
 \begin{align*}
-&\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] = - \displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \text{log}P(i_1,\cdots,i_k | \Phi_l^d) \big] \\
-&= -\displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \displaystyle{\sum_{j=2}^k}\text{log}P(i_j | i_1,\cdots,j_{j-1},\Phi_l^d) +  \text{log} P(i_1 | \Phi_l^d) \big] \\
-&\geqslant -\displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] \\
+&\mathbb{H} \big[ P(i_1\cdots i_k | \Phi_l^d) \big] = - \displaystyle{\mathop{\mathbb{E}}_{i_1\cdots i_k | \Phi_l^d}} \big[ \text{log}P(i_1 \cdots i_k | \Phi_l^d) \big] \\
+&= -\!\! \!\!\!\! \displaystyle{\mathop{\mathbb{E}}_{i_1\cdots i_k | \Phi_l^d}} \big[ \displaystyle{\sum_{j=2}^k}\text{log}P(i_j | i_1 \cdots j_{j-1},\Phi_l^d) +  \text{log} P(i_1 | \Phi_l^d) \big] \\
+&\geqslant -\!\! \!\! \!\!\displaystyle{\mathop{\mathbb{E}}_{i_1 \cdots i_k | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] \\
 &= -\displaystyle{\mathop{\mathbb{E}}_{i_1 | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] = \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]
@@
 \centering{
-\includegraphics[width=0.5\textwidth]{multi_domain}}
+\includegraphics[width=0.5\textwidth,frame]{multi_domain}}
 \caption{Visualization of domains according to their dropout masks (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
@@
 using binary search}
-\STATE{Compute mask of each layer
+\STATE{Compute mask of each layer:
 \begin{align*}
@@
 \end{align*}}
-\STATE{Apply masks to their corresponding layer
+\STATE{Apply masks to their corresponding layer:
 \begin{align*}
@@
 }
-\STATE{Compute gradient of training loss over the underlying Transformer
+\STATE{Compute gradient of training loss over the underlying Transformer:
 $$\Delta_{\theta} = \frac{\partial L}{\partial \theta}$$}
-\STATE{Compute gradient over the $\operatorname{Soft-Top-K}$ masks
+\STATE{Compute gradient over the $\operatorname{Soft-Top-K}$ masks:
 $$\frac{\partial D}{\partial \hat{m}_l^d(\tau)}$$}
-\STATE{Compute implicit gradient of the $\operatorname{Soft-Top-K}$ masks over $\Phi_l^d$
+\STATE{Compute implicit gradient of the $\operatorname{Soft-Top-K}$ masks over $\Phi_l^d$:
 $$\frac{\partial \bar{\nu}}{\partial \Phi_l^d} = - \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}$$
-$$\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} = \frac{\partial \hat{m}_l^d(\tau)}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d} + \frac{1}{\tau} \frac{exp(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau})}{(1+exp(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}))^2}$$}
-\STATE{Compute the gradient the training over $\Phi_l^d$
-$$\Delta_{\Phi_l^d} = \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} + \frac{\partial \mathbb{H} \big[ \operatorname{softmax}(\Phi_l^d)\big] }{\partial \Phi_l^d} $$}
+$$\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} = \frac{\partial \hat{m}_l^d(\tau)}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d}  + $$
+$$ \quad\quad \frac{1}{\tau} \frac{exp(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau})}{(1+exp(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}))^2}$$}
+\STATE{Compute the gradient the training over $\Phi_l^d$:
+$$ \Delta_{\Phi_l^d} = \frac{\partial D}{\partial \hat{m}_l^d(\tau)}  \frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} + \frac{\partial \mathbb{H} \big[ \operatorname{softmax}(\Phi_l^d)\big] }{\partial \Phi_l^d} $$}
 \STATE{Update $\theta$ and $\Phi_l^d$ with their gradients}

### 2022-05-16 2022/acl/acl_latex.tex
local changes

--- 2022-04-22 2022/acl/acl_latex.tex
@@
 
-\author{
-  MinhQuang Pham$^{\dag\ddag}$, \ \ Josep Crego$^\dag$,\ \ Fran\c cois Yvon$^\ddag$ \\
+\author{MinhQuang Pham$^{\dag\ddag}$\thanks{$\ \ \ \ $Now Research Scientist at Zoom Video Communications}, \ \ Josep Crego$^\dag$,\ \ Fran\c cois Yvon$^\ddag$ \\
   $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France\\
   {\tt firstname.lastname@systrangroup.com}\\
-  $^\ddag$Universit\'e Paris-Saclay, CNRS, LISN,  91405 Orsay, France\\
-  {\tt firstname.lastname@limsi.fr}
-}
+  $^\ddag$Universit\'e Paris-Saclay, CNRS, LIMSI,  91405 Orsay, France\\
+  {\tt firstname.lastname@limsi.fr}}
 
@@
 \begin{align*}
-\hspace{-3.em}
-\lim_{\tau \rightarrow 0} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = \begin{cases}
+\lim_{\tau \rightarrow 0} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = \\
+= \begin{cases}
       1, & \text{if}\ \tau > - (g_l^d(i) + \Phi_l^d(i)), \\

### 2022-05-16 2022/eamt/eamt-latex/eamt22.tex
local changes

--- 2022-05-09 2022/eamt/eamt-latex/eamt22.tex
@@
 In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key aspect of training is then to mitigate the detrimental effects of a mismatch between these distributions.
-Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from; this setting differs from multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \cite{Chu2017comparison,Saunders21asurvey} for a review), and so is multi-domain (MD) learning \cite{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}.
+Single-source and multi-source\footnote{In this paper, multi-source DA means having multiple domains to adapt from; this setting differs from multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \newcite{Chu2017comparison} and \newcite{Saunders21asurvey} for a review), and so is multi-domain (MD) learning \cite{Chu18multilingual,Jiang19multidomain,Pham21revisiting}.
 A related situation is multilingual MT \cite{Firat16multiway,Ha16towards,Johnson17google,Aharoni19massively},\fyDone{Fix ref}
@@
 
-This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the testing conditions \cite{Moore10selection,Axelrod11domain}.
+This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the testing conditions \cite{Axelrod11domain}.
 Finding the optimal balance of training data is however, a challenging task due, for instance, to the similarity between domains/languages, or to the regularization effects of out-of-domain data \cite{Miceli-barone17regularization}.
@@
 
-We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc.
+We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from other data sources in terms of textual genre, thematic content \cite{Chen16guided}, register, style \cite{Niu18multitask}, etc.
 Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$.
@@
 
-These situations have been amply documented from a theoretical perspective (eg.\ \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}).
+These situations have been amply documented from a theoretical perspective \ \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}.
 A general recommendation in the DA setting is to adjust the sampling distribution used to optimize the system so as to compensate for the mismatch between $\mathcal{D}^s(x)$ and $\mathcal{D}^t(x)$.
@@
 We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.
-Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and \texttt{sacrebleu} \cite{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
+Validation sets are used to chose the best model according to the average BLEU score \cite{Post18call}.\footnote{We use truecasing and \texttt{sacrebleu} \cite{Post18call}.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}.
 We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}
@@
 Domain adaptation is an old problem that has been studied from many angles, both for SMT and NMT.
-A survey of supervised and unsupervised DA for NMT is in \cite{Chu2017comparison}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the recent survey of \newcite{Saunders21asurvey}.
+A survey of supervised and unsupervised DA for NMT is in \newcite{Chu2017comparison}, where the authors distinguish between data-centric and model-centric DA, a view also adopted in the recent survey of \newcite{Saunders21asurvey}.
 Our approach to DA in this paper falls under the former category.
@@
 Multidomain NMT (MDMT) aims to develop systems that simultaneously bode well for several domains.
-Like for DA, techniques for supervised MDMT combine one or several ingredients: (a) the specialization of data representations \cite{Kobus17domaincontrol} or of sub-networks \cite{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \cite{Jiang19multidomain}.
+Like for DA, techniques for supervised MDMT combine one or several ingredients: (a) the specialization of data representations \cite{Kobus17domaincontrol} or of sub-networks \cite{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing}; (c) the use of automatic domain identification \cite{Jiang19multidomain}.
 Unsupervised MDMT is studied in \newcite{Farajian17multidomain}, as an instance of unsupervised DA.
@@
 CL relies on the notion of the ``easiness'' of a sample to schedule the presentation of training data so as to start with the easiest examples and end with the hardest.
-Various ways to automate CL in the framework of multi-armed bandits are explored in \cite{Graves17automated}, which has been an inspiration for our implementation.
+Various ways to automate CL in the framework of multi-armed bandits are explored in \newcite{Graves17automated}, which has been an inspiration for our implementation.
 While the initial aim was primarily to improve and speed up training, CL has also proven useful for multidomain and multilingual MT, based on alternative definitions of ``easiness''.\fyDone{Ajouter Grave}
@@
 \newcite{Kumar19reinforcement} use reinforcement learning to learn the curriculum strategy: in this work, complexity corresponds to difficulty levels which are binned using contrastive data selection.
-The reward is based on the increase of the devset loss that results from the current data selection strategy.\fyDone{Alert: what do we do during warm-up ?} This technique is applied to multilingual NMT in \cite{Kumar21learning}.
+The reward is based on the increase of the devset loss that results from the current data selection strategy.\fyDone{Alert: what do we do during warm-up ?} This technique is applied to multilingual NMT in \newcite{Kumar21learning}.
 \newcite{Zhou20uncertainty} propose another CL-based approach which relies on \emph{instance uncertainty} as a measure of their difficulty and presents data samples starting with the easiest.

### 2022-05-16 2022/acl/acl_latex.tex

--- 2022-05-16 2022/acl/acl_latex.tex
@@
 \begin{align*}
-% \hspace{-3.em}
-  \lim_{\tau \rightarrow 0} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) & = \\
-  \quad \begin{cases}
+\lim_{\tau \rightarrow 0} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = \\
+= \begin{cases}
       1, & \text{if}\ \tau > - (g_l^d(i) + \Phi_l^d(i)), \\
@@
     \end{cases}
+%% \hspace{-3.em}
+%  \lim_{\tau \rightarrow 0} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) & = \\
+%  \quad \begin{cases}
+%    1, & \text{if}\ \tau > - (g_l^d(i) + \Phi_l^d(i)), \\
+%    0, & \text{if}\ \tau < - (g_l^d(i) + \Phi_l^d(i)), \\
+%    \frac{1}{2} &\text{otherwise}\\
+%  \end{cases}
 \end{align*}

### 2022-05-16 2022/eamt/eamt-latex/eamt22.tex

@@
 
-These situations have been amply documented from a theoretical perspective \ \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}.
+These situations have been amply documented from a theoretical perspective \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}.
 A general recommendation in the DA setting is to adjust the sampling distribution used to optimize the system so as to compensate for the mismatch between $\mathcal{D}^s(x)$ and $\mathcal{D}^t(x)$.
@@
 Multidomain NMT (MDMT) aims to develop systems that simultaneously bode well for several domains.
-Like for DA, techniques for supervised MDMT combine one or several ingredients: (a) the specialization of data representations \cite{Kobus17domaincontrol} or of sub-networks \cite{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing}; (c) the use of automatic domain identification \cite{Jiang19multidomain}.
-Unsupervised MDMT is studied in \newcite{Farajian17multidomain}, as an instance of unsupervised DA.
+Like for DA, techniques for supervised MDMT combine one or several ingredients: (a) the specialization of data representations \cite{Kobus17domaincontrol} or of sub-networks \cite{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \cite{Jiang19multidomain}.
+Unsupervised MDMT is studied in \cite{Farajian17multidomain}, as an instance of unsupervised DA.
 
